### 2014-04-25 sequential/latex/DAC2013/mainframe.tex
upload files from laptop linux

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

\begin{document}

\title{Verification of Sequential Galois Field Circuits using Algebraic Geometry
\titlenote{First version, only contains abstract and introduction part.}}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{2} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Xiaojun Sun\\
       \affaddr{University of Utah}\\
       \affaddr{Department of Electrical \& Computer Engineering}\\
       \affaddr{Salt Lake City, USA}\\
       \email{xiaojun.sun@utah.edu}
% 2nd. author
\alignauthor
Priyank Kalla\\
       \affaddr{University of Utah}\\
       \affaddr{Department of Electrical \& Computer Engineering}\\
       \affaddr{Salt Lake City, USA}\\
       \email{kalla@ece.utah.edu}
% 3rd. author
%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
%one who did all the really hard work.}\\
%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
%       \affaddr{Hekla, Iceland}\\
%       \email{larst@affiliation.org}
%\and  % use '\and' if you need 'another row' of author names
% 4th. author
%\alignauthor Lawrence P. Leipuner\\
%       \affaddr{Brookhaven Laboratories}\\
%       \affaddr{Brookhaven National Lab}\\
%       \affaddr{P.O. Box 5000}\\
%       \email{lleipuner@researchlabs.org}
% 5th. author
%\alignauthor Sean Fogarty\\
%       \affaddr{NASA Ames Research Center}\\
%       \affaddr{Moffett Field}\\
%       \affaddr{California 94035}\\
%       \email{fogartys@amesres.org}
% 6th. author
%\alignauthor Charles Palmer\\
%       \affaddr{Palmer Research Laboratories}\\
%      \affaddr{8600 Datapoint Drive}\\
%%      \affaddr{San Antonio, Texas 78229}\\
%       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{07 Nov 2013}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Circuits working in Galois fields are increasingly employed in designs like Ellptic Curve Cryptography (ECC). 
This work proposes a new method to effectively verify sequential circuits in Galois fields. Algebraic geometry
is intriduced to describe circuits behavior and redefine implicit state space traversal. Moreover, Gr\"obner basis
representation is adopted for word-level abstraction on circuit variables to address state explosion problem
with BDDs. Experiments are run with F4-style reduction engine to get more competitive results.
\end{abstract}

% A category with the (minimum) three required fields
\category{EDA5.1}{Verification}{Functional, transaction-level, RTL, and gate-level modeling and verification of hardware design}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Verification, Algorithms}

\keywords{Verification, Sequential, Galois Fields, Algebraic Geometry} % NOT required for Proceedings

\section{Introduction}
Verification for sequential circuits has been extensively discussed for decade. Binary Decision Diagram (BDD) is the first
and most widely used canonical technique among those for sequential verification \cite{touati1990implicit}. It is easy to
manipulate any Boolean expressions with BDD and use them to enumerate implicit states. However, an issue that BDDs have to
face is the size explosion when Boolean variables increase significantly. Several modifications have been made to improve
this problem by optimizing original BDD \cite{gunther2001application} \cite{narayan1997reachability}, or by reducing Boolean
variables from circuits\cite{burch1991representing} \cite{jiang2003verification}. Some variants of BDD representation
have also been developed such as Linear Taylor Expansion Diagram (LTED) \cite{alizadeh2009sequential}.

Another sort of approaches are based on satisfiability theory (SAT). SAT-solver is applied to state space traversal and 
work through eliminations on Boolean constrains \cite{coudert1990verification}. By exploiting SAT-based algorithm, direct
reachability induction is proposed \cite{bjesse2000sat}. Still SAT-based approaches need to struggle with size explosion problem.

\section{Related Works}
G. Avrunin \cite{avrunin1996symbolic} introduced algebraic geometry into formal verification. In his paper, the corresponding
relation between varieties of ideals and circuit variables is discussed, and concepts such as intersect of varieties, ideal and
its generators and algebraic closure are explored. Yet, a self-contained system of algebraic geometry representation theory has
not been well-defined.

T. Pruss, el(which one should I cite?) developed a word-level abstraction method based on  Gr\"obner basis theory...

J. Lv, el (should I cite this part) gave out a F4-style reduction algorithm which can help speed up Gr\"obner basis calculation...

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{seq_verif}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
\balancecolumns
% That's all folks!
\end{document}

### 2014-04-25 sequential/latex/DAC2013/theory.tex
upload files from laptop linux

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

%packages
\usepackage[linesnumbered, ruled]{algorithm2e}
\RequirePackage{amssymb, mathptm}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{cases}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{comment}

%new format
\newtheorem{Algorithm}{Algorithm}
\newdef{Definition}{Definition}
\newtheorem{Theorem}{Theorem}
\newtheorem{Lemma}{Lemma}
\newdef{Example}{Example}

\begin{document}

\title{Verification of Sequential Galois Field Circuits using Algebraic Geometry
\titlenote{Version 0.3, refined theory part with normal basis, proposed details for designing 2 examples.}}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{2} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Xiaojun Sun\\
       \affaddr{University of Utah}\\
       \affaddr{Department of Electrical \& Computer Engineering}\\
       \affaddr{Salt Lake City, USA}\\
       \email{xiaojun.sun@utah.edu}
% 2nd. author
\alignauthor
Priyank Kalla\\
       \affaddr{University of Utah}\\
       \affaddr{Department of Electrical \& Computer Engineering}\\
       \affaddr{Salt Lake City, USA}\\
       \email{kalla@ece.utah.edu}
% 3rd. author
%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
%one who did all the really hard work.}\\
%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
%       \affaddr{Hekla, Iceland}\\
%       \email{larst@affiliation.org}
%\and  % use '\and' if you need 'another row' of author names
% 4th. author
%\alignauthor Lawrence P. Leipuner\\
%       \affaddr{Brookhaven Laboratories}\\
%       \affaddr{Brookhaven National Lab}\\
%       \affaddr{P.O. Box 5000}\\
%       \email{lleipuner@researchlabs.org}
% 5th. author
%\alignauthor Sean Fogarty\\
%       \affaddr{NASA Ames Research Center}\\
%       \affaddr{Moffett Field}\\
%       \affaddr{California 94035}\\
%       \email{fogartys@amesres.org}
% 6th. author
%\alignauthor Charles Palmer\\
%       \affaddr{Palmer Research Laboratories}\\
%      \affaddr{8600 Datapoint Drive}\\
%%      \affaddr{San Antonio, Texas 78229}\\
%       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{21 Jan 2014}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Circuits working in Galois fields are increasingly employed in designs like arithmetic component for Elliptic Curve Cryptography (ECC). 
This work proposes a new method to effectively verify sequential circuits in Galois fields. Algebraic geometry
is introduced to describe circuits behavior and expand the definition of implicit state space traversal. Moreover, Gr\"obner basis
representation is adopted for word-level abstraction on circuit variables to address state explosion problem
with BDDs. Experiments are run with fast Gr\"obner basis computation engine to get competitive results.
\end{abstract}

% A category with the (minimum) three required fields
\category{EDA5.1}{Verification}{Functional, transaction-level, RTL, and gate-level modeling and verification of hardware design}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Verification, Algorithms}

\keywords{Verification, Sequential, Galois Fields, Algebraic Geometry} % NOT required for Proceedings

\section{Essential Theory}
\subsection{Polynomial Representation}
A typical sequential circuit is composed as figure \ref{fig:sequential}. The combinational logic has primary input $x$, state
inputs $\{s_0, s_1, \dots, s_{k-1}\}$, primary output $z$ and state outputs $t_0, t_1, \dots, t_{k-1}$. State outputs
are latched into registers, which will be fed back to state inputs in next clock cycle.

\begin{figure}
\centering
\psfig{file=./sequential_fig.ps, height=2in, width=3.5in,}
\caption{A typical sequential circuit block view}
\label{fig:sequential}
\end{figure}

A transition function $\Delta$ is a Boolean function about state outputs and all inputs, i.e.
\begin{displaymath}
t_i = \Delta_i(x, s_0, s_1, \dots, s_{k-1})
\end{displaymath}
This constrain could also be written in the form of equation $polynomial = 0$:
\begin{displaymath}
t_i - \Delta_i(x, s_0, s_1, \dots, s_{k-1}) = 0
\end{displaymath}
Solution to this equation is limited within $\{0, 1\}$, which is set of all elements in Galois field $\mathbb{F}_2$.
Furthermore, there exists a one-to-one corresponding relation from Boolean space to Galois field on higher dimension: 
$\mathbb{B}^k \to \mathbb{F}_{2^k}$, which means it is possible to find a set of elements in $\mathbb{F}_{2^k}$ representing
all Boolean values every bit of a bit vector can take. In Galois field, a as \ref{table:booltogalois} shows.
\begin{table}[h]
\centering
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
\pmod{x^4+x^3+1}$}\label{table:booltogalois}  
\begin{tabular}{|c|c||c|c|} 
\hline
$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
\hline
$0000$        & $0$           & $1000$  &$\alpha^3$\\
\hline
$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
\hline
$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
\hline
$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
\hline
$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
\end{table}
A word-level representation on state inputs/outputs can be attained through above correspondence. If the set of 
state inputs $\{s_0, s_1, \dots, s_{k-1}\}$ is denoted by $S$ taking values from $\mathbb{F}_{2^k}$, and
set of outputs $\{t_0, t_1, \dots, t_{k-1}\}$ is denoted by $T$, a word-level transition function can be written as
\begin{displaymath}
T - \Delta(x, S) = 0
\end{displaymath}
If $T - \Delta(x, S)$ is taken as a polynomial $f$, then $f \in \mathbb{F}_{2^k}[x]\ mod\ P(\alpha)$.
\subsection{Gr\"obner Basis Theory}
Gr\"obner basis is used to calculate desired results out of a set of polynomials.
If there exists multiple polynomials $\{f_1, f_2, \dots, f_r\}$ representing constrains on $T, x$ and $S$,
then all polynomials should simultaneously equal to $0$.
\begin{displaymath}
\left\{
  \begin{array}{lc}
  f_1 & = 0\\
  f_2 & = 0\\
  \dots & \ \\
  f_r & = 0
  \end{array} \right.
\end{displaymath}
Solution to these equations is set of values of $(T, x, S)$, which is called \textbf{Variety} of \textbf{ideal}
$I = \langle f_1, f_2, \dots, f_r\rangle $.

Calculate Reduced Gr\"obner Basis (RGB) from ideal $I$ with \textbf{elimination term order}, there exists a polynomial contains
only $T$, if the values of states input $S$ and primary input $x$ are given in the form of polynomials and included by $I$,
the values of $T$ can be computed.

Otherwise, if RGB calculation is manipulated under \textbf{abstraction term order}, there exists a polynomial in
the form of $T - \mathcal{F}(x, S)$ representing transition function. In next clock cycle, assign new state input $S'$ with $T$, the next
state can be computed again, result is polynomial $T' - \mathcal{F}(x', S')$.

\subsection{Normal Basis Representation}

Let $\beta$ be an element in the Galois field $F_{2^n}$ constructed by primitive element $\alpha$ and minimal polynomial
$P(\alpha)$. Then a basis in the form $\{\beta, \beta^2, \beta^4, \beta^8, ... ,\beta^{2^{n-1}}\}$ is a
\emph{Normal Basis}; here $\beta$ is called \emph{Normal Element}.

For arithmetic operations in Galois fields, squaring and multiplication (with modulus) can be greatly simplified if 
normal basis representation is adopted to represent operands.
\begin{Example}
\label{ex:nb_sq}
Element squaring: In $F_{2^n}$, all coefficients which can be divided by 2 are reduced, so 
following equality holds for all field elements $a$ and $b$:
$$(a+b)^2 = a^2 + b^2$$ 
Apply this rule on element squaring of standard/polynomial basis:
\begin{align}
& (b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{n-1}\beta^{2^{n-1}})^2 \nonumber\\
&= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{n-1}^2\beta^{2^n} \nonumber\\
&= b_{n-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{n-2}\beta^{2^{n-1}} \nonumber
\end{align}
using Fermat's little theorem $\beta^{2^n} = \beta$. This shows that squaring of field elements
represented by normal bases can be easily completed by operating a simple right-cyclic rotation.
\end{Example}

\begin{Example}
\label{ex:nb_multi}
Normal basis multiplication: There are 2 binary vectors representing operands of multiplication in normal
basis: $$A = (a_0, a_1, \dots, a_{n-1}),\ B = (b_0, b_1, \dots, b_{n-1})$$ 
the product is also written in normal basis representation: $$C = A*B = (c_0, c_1, \dots, c_{n-1})$$
Describe calculation procedure for the most significant digit of product with a function: 
$$c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})$$
Square both side: $C^2 = A^2*B^2$, i.e. the second significant digit 
$$c_{n-2} = f(a_{n-1}, a_0, a_1, \dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2})$$ 
By this method it is easy to get all digits of product $C$.
\end{Example}

\section{Algebraic Geometry Approach}
\subsection{State Space Traversal}

\begin{figure}
\centering
\psfig{file=./fsm_fig.ps, height=2in, width=3.5in,}
\caption{Gate-level circuit of combinational logic block of sample FSM}
\label{fig:fsm}
\end{figure}

\begin{figure}
\centering
\psfig{file=./stg_fig.ps, height=2in, width=3.5in,}
\caption{State transition graph of sample FSM}
\label{fig:stg}
\end{figure}

The first example is a mealy finite state machine (FSM). Gate-level circuits of combinational logic
is figure \ref{fig:fsm}, $x$ is primary input,
$\{s_0, s_1\}$ are state inputs, $Z$ is primary output and $\{t_0, t_1\}$ are state outputs.
Word-level state input is $S = s_0 + s_1\cdot\alpha$, output is $T = t_0 + t_1\cdot\alpha$. In Galois field
$\mathbb{F}_{2^2}$, bit-level variables $x, s_0, s_1, Z, t_0, t_1$ can take values from $\{0, 1\}$, while
word-level variables $S, T$ may take values from $\{0, 1, \alpha, 1 + \alpha\}$. State transition graph (STG)
showed in figure \ref{fig:stg} uses 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2, S_3\}$, which
could be converted to elements in $\mathbb{F}_{2^2}$ similarly as table \ref{table:booltogalois} shows.

One important technique to check sequential equivalence is \textbf{state space traversal}. In this example,
Breath-First-Search (BFS) is employed for traversal, which can be described with following algorithm:

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}\label{alg:BFS}
\end{algorithm}

Our approach is implementing state space traversal by refined BFS algorithm combined with polynomial representation
and Gr\"obner basis theory. In practice, this example will show how to apply concepts and techniques from algebraic
geometry on each line of above BFS algorithm.

\subsection{States and Varieties of Ideal}

\begin{Theorem}
State variables $S, T$ and sets of states such as $from^i, to^i$ can always be represented as varieties of ideals.
\end{Theorem}
For example in Line 1 of BFS algorithm, assume initial state is $S_3$ in figure \ref{fig:stg}, then corresponding
polynomial $f = \mathcal{F}(S^0) = S^0 - 1 - \alpha$. Consider an ideal $I$ with only one generator $f$, its variety
$V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma - 1 - \alpha = 0\}$, which equals to $\{1 + \alpha\}$, the only
valid value $S^0$ can take.

If an ideal contains multiple polynomial specifications, it is necessary to compute Gr\"obner basis with elimination term
order to get one polynomial only on desired variable. In first iteration of Line 4, $to^i$ has multiple specifications,
some of them are transition functions on bit-level:
\begin{displaymath}
\begin{array}{ll}
f_1:& t_0 - (\overline{x}\ and\ \overline{s_0}\ and\ \overline{s_1})\ or\ (s_0\ and\ s_1) \\
f_2:& t_1 - (\overline{s_0}\ and\ x)\ or\ (p\ and\ \overline{s_1})\
\end{array}
\end{displaymath}
And some are bits-to-word definitions fitting polynomial representation of elements in $\mathbb{F}_{2^2}$:
\begin{displaymath}
\begin{array}{ll}
f_3:& S - s_0 - s_1\alpha \\
f_4:& T - t_0 - t_1\alpha
\end{array}
\end{displaymath}
And an polynomial about initial state as mentioned above:
\begin{displaymath}
f_5:\  S - 1 - \alpha
\end{displaymath}
And the rests are vanishing polynomials for every variable, bit-level and word level:
$f_6: x^2 - x; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: s_0^2 - s_0;
f_{11}: s_1^2 - s_1; f_{12}: T^4 - T$

Transition equations here contains some Boolean operators, they can be re-written in terms of operations in Galois fields. 
In $\mathbb{F}_{2}$, let $TRUE = 1, FALSE = 0$, for either element $a$ in this field, considering $0 + 0 = 1 + 1 \equiv 0\ (mod\ 2)$
and $0 + 1 = 1 + 0 \equiv 1\ (mod\ 2)$, the inverse of $a$ is: $\overline{a} = 1 + a$. Similarly all Boolean operators
can be converted within $\mathbb{F}_{2}$, table \ref{table:booltogalois_op} gives part of them and their corresponding
operations in $\mathbb{F}_{2}$:
Also note that there is no specification on initial primary input $x$, in Gr\"obner basis approach this means $x$ is smoothed by
reversely using Shannon's expansion.
\begin{table}
\centering
\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
\label{table:booltogalois_op}
\begin{tabular}{|c|c|} \hline
Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
$\overline{a}$ & $1 + a$\\ \hline
$a\ and\ b$ & $ab$\\ \hline
$a\ or\ b$ & $a + b + ab$\\ \hline
$a \oplus b$ & $a + b$\\
\hline\end{tabular}
\end{table}
Using elimination term order: $others\ >\ bit$-$level\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
$J = \langle f_1, f_2, \dots, f_{12}\rangle $, the result will include one polynomial $f_T$ contains only variable $T$. In this example,
$f_T = T + 1$. This polynomial equals to $T - 1$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
limited within $\mathbb{F}_{2}$, where $1 \equiv -1\ (mod\ 2)$. Solution to $f_T = 0$ is $T = 1$, which shows that next
state the machine just reached is $\{S_1(01)\}$.

\subsection{Application of Algebraic Geometry}

There are some difficulties with polynomial representation when executing Line 5 and Line 6, it is necessary to explore
how \textbf{Union}, \textbf{Intersection} and \textbf{Complement Set} works in Galois field $\mathbb{F}_{2^k}$. Since
state set variables $from^i, to^i, reached$ all have single polynomial, consider an ideal $I$ with single generator $I = \langle f\rangle $.
Example: assume $I = \langle f\rangle  = \langle T^2 + (1+\alpha)\cdot T+\alpha\rangle $, its variety $V(I) = \{a\ |\ a \in \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$.
So it is reasonable to specify: union, intersection and complement set mentioned in this paper are all functions on \textbf{varieties}.
To better discuss this problem, introduce following concepts from algebraic geometry:
\begin{Definition}
\label{def:sum}
({\bf Sum of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the 
{\bf sum} of $I$ and $J$, denoted by $I + J$, is the set
  \begin{equation}
  I + J = \{f + g\ |\ f \in I \ and\  g \in J\}.
  \end{equation}
Furthermore, if $I = \langle f_1, \dots, f_r\rangle$ and 
$J = \langle g_1, \dots, g_s\rangle$, then 
$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$.
\end{Definition}
\begin{Definition}
\label{def:prod}
({\bf Product of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the
{\bf product} of $I$ and $J$, denoted by $I \cdot J$, is defined to be the ideal generated 
by all polynomials $f \cdot g$ where $f \in I$ and $g \in J$. Furthermore, let
$I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$, then
  \begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle .
  \end{equation}
\end{Definition}
\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then $I:J$
is the set
  \begin{equation}
  \{f \in k[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}
These concepts can lead the way to solution by adopting following theorems:
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
\bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) = {\bf V}(I) \bigcup {\bf V}(J)$.
\end{Theorem}
\begin{Theorem}
\label{thm:quotient}
If $I, J$ are ideals with only one generator, then ${\bf V}(I:J) ={\bf V}(I) - {\bf V}(J)$.
\end{Theorem}
Proof to these theorems are referred to (that yellow book?) and (my write-up?).

For varieties intersection $\{1\}\bigcap\{1, \alpha\}$, calculate ideal sum $\langle T+1, T^2 + (1+\alpha)\cdot T+\alpha\rangle  = \langle T+1\rangle $,
its variety is $\{1\}$; for varieties union $\{1,1+\alpha\}\bigcup\{1+\alpha\}$, calculate
ideal product $\langle (T+1+\alpha)\cdot(T^2 + (1+\alpha)\cdot T+\alpha)\rangle  = \langle T^3 + 1\rangle $, its variety
is $\{1, \alpha, 1+\alpha\}$; for complement set of variety $\{1, \alpha\}$, the universal set is
the variety of ideal of vanishing polynomial $V(\langle T^4-T\rangle ) = \{0,1,\alpha,1+\alpha\}$,
so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
the result is $\{0,1+\alpha\}$.

From discussion above, the BFS traversal is completely implemented in Galois field.
If the initial input is $S_3$: $S + 1 + \alpha$, the final return value will be set of
reachable states: $T^4 + T$, i.e. the universal set, $\{S_0, S_1, S_2, S_3\}$.

\subsection{Discussion on Ideal quotient and its generator}
In nature, we are looking varieties as the solutions to specific polynomial $f=0$ WITHIN $\mathbb{F}_{2^n}$, because (affine) varieties are defined
within algebraic closure $k^n$ but we only care about those fall in $\mathbb{F}_{2^n}$.

Theorem \ref{thm:unionintersect} related to DEF. \ref{def:sum} and \ref{def:prod} counts on ideal generators, so there will be no ambiguity on 
new ideals we get from this theorem. It also appeared in Avrunin's CAV paper. The only problem is what ideal quotient operation will give us.
This should be further discussed because the completeness of our algebraic geometry approach will benefit from it. \emph{Singular} has the function of
ideal quotient, so ask from the author maybe another choice.

If we put this conception aside, it is also feasible to define complement set from "solution to 
equation" sense. Assume $f(x) = x + 1 + \alpha = 0$, its solution within $\mathbb{F}_{2^2}$ 
is $\{1+\alpha\}$; meanwhile solutions to vanishing polynomial $v(x) = x^2 + x = 0$ are
$\{0,1,\alpha,1+\alpha\}$. Now compute solutions to $$g(x) = \frac{v(x)}{f(x)} = 0$$,
this requires $v(x) = 0$ and $f(x) \neq 0$, which means the complement set of solution 
to $f(x) = 0$.

Issues to this interpretation: (a) Is $v(x)$ always divisible by $f(x)$? (b) Is $v(x)=0$ and
$f(x) \neq 0$ equivalent with $g(x)=0$ on their solutions? Could $f(x)$ equal to $0$?

\section{Functional Verification}
\subsection{Sequential Galois Arithmetic Multiplier}
Our experiment is to verify the function of a Galois field multiplier after certain clock cycles.
The following design is Sequential Multiplier with Parallel Output (SMPO),
a Normal Basis multiplier based on Massey-Omura algorithm. Inputs and outputs
are all 5-bit word taking value from ${\mathbb F}_{2^5}$. After loading operands
 A and B, setting all output latches to 0 and running for 5 iterations, the 
output should be $R = A\cdot B$ (mod $x^5 + x^2 + 1$).
\begin{figure}
\centering
\epsfig{file=./mySMPO.eps, height=2in, width=3.5in}
\caption{5-bit SMPO}
\label{fig:SMPO}
\end{figure}
Similarly, build elimination ideal for all gates/operations and induce
initial states of latches. However, instead of eliminating all variables
to one, this example adopts abstraction term order
and keeps the polynomial contains function between output and inputs.
Here the lex ordering is $\{others\} > R > \{A, B\}$, and objective
polynomial is $R + \mathcal{F}(A, B)$.

Apply this approach to calculate image in BFS traversal, but modify algorithm
\ref{alg:BFS} to make it adapt 5 steps reachable states enumeration.
The result is $R + AB$, which validates the function of this circuit.

\subsection{Gr\"obner basis based Approach}

For 5-bit normal basis multiplication, the $i$-th digit of product can be written as
\begin{align}
R_i &= b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
\end{align}
It is possible to calculate every conjunctive term in the product simultaneously within one clock cycle,
on distinct bits. 
Use cyclic similarity of above function, following executions are operated: 
\begin{Example}
Sequential Multiplier Protocol:
\begin{itemize}
\item \textbf{Initial}\ \ $R_0 = R_1 = R_2 = R_3 = R_4 = 0$
\item \textbf{Clock 1}\ \ $R_0 = a_1b_0, R_1 = b_2(a_1 + a_4), R_2 = b_4(a_0 + a_1), R_3 = b_1(a_4 + a_0), 
			R_4 = b_3(a_1 + a_3)$
\item \textbf{Clock 2}\ \ $R_0 = b_3(a_1 + a_3) + a_0b_4, R_1 = a_1b_0 + b_1(a_0 + a_3), R_2 = b_2(a_1 + a_4)
			+ b_3(a_4 + a_0), R_3 = b_4(a_0 + a_1) + b_0(a_3 + a_4), R_4 = b_1(a_4 + a_0) + b_2(a_0 + a_2)$
\item \textbf{$\dots$}
\item \textbf{Clock 5}\ \ $R_0 = c_0, R_1 = c_1, R_2 = c_2, R_3 = c_3, R_4 = c_4$, i.e. $R = A\cdot B$.
\end{itemize}
\end{Example}

In BFS algorithm, the \textbf{Image} function is implemented by constructing an elimination ideal then 
compute its Gr\"obner basis. 
\begin{Example}
\label{ex:SMPO}
For 5-bit SMPO, the ideal consists of (for the first clock cycle):

(a) {\bf Gate descriptions:}
$a_1+a_4+c_1, a_1+a_0+c_2, a_0+a_4+c_3, a_1+a_3+c_4,
		  a_1b_0+r_4+R_0, c_1b_2+r_0+R_1, c_2b_4+r_1+R_2, c_3b_1+r_2+R_3, c_4b_3+r_3+R_4;$

(b) {\bf Word-level variables:}
$A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18},
		  r+r_0\alpha^5+r_1\alpha^{10}+r_2\alpha^{20}+r_3\alpha^9+r_4\alpha^{18},
		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18};$

(c) {\bf Vanishing polynomials:}
		 $ a_0^2+a_0, a_1^2+a_1, a_2^2+a_2, a_3^2+a_3, a_4^2+a_4,
		  b_0^2+b_0, b_1^2+b_1, b_2^2+b_2, b_3^2+b_3, b_4^2+b_4,
		  r_0^2+r_0, r_1^2+r_1, r_2^2+r_2, r_3^2+r_3, r_4^2+r_4,
		  R_0^2+R_0, R_1^2+R_1, R_2^2+R_2, R_3^2+R_3, R_4^2+R_4,
		  c_1^2+c_1, c_2^2+c_2, c_3^2+c_3, c_4^2+c_4,
		  A^{32}+A, B^{32}+B, r^{32}+r, R^{32}+R;$

(d)	{\bf Feedback input:}	  $r_{in}$.

Polynomial $r_{in}$ equals to $from^{i-1}$ in Line 4, BFS algorithm. Under abstraction term ordering,
polynomial $to^i$ is assigned with a polynomial in result Gr\"obner basis which has the form $R + \mathcal{F}(A,B)$. 
Since all outputs are connected to feedback inputs
in SMPO, Line 5 and 6 will be neglected. Line 7 is finished by replace current output $R$ with previous 
output $r$. Initially $r_{in} = 0$.

In next clock cycle, $r_{in} = r + \mathcal{F}(A,B)$ updated from latest result; simultaneously operands
$A$ and $B$ have been cyclically shifted, so gate descriptions in (a) should be modified accordingly. The 
loop is operated for 5 times, result from the last step's Gr\"obner basis should be polynomial $R + AB$.
\end{Example}

This experiment can be repeated on $n$-bits SMPO after running for $n$ clock cycles.

\section{Fast Gr\"obner basis computation}
\subsection{Refined abstraction term order (RATO)}
A lexicographic order constrained by following relation $>_{r}$: "circuit variables ordered reverse topologically" $>$ 
"designated word-level output" $>$ "word-level inputs" is called the \emph{Refined Abstraction Term Order (RATO)}.

In Buchberger's algorithm, the specification polynomial (Spoly) for each pair is calculated. In RATO, most polynomials
have relatively prime leading terms/monomials (which means $Spoly \xrightarrow{J+J_0}_{+} 0$) except one pair:
word-level polynomial corresponding to outputs and its leading bit-level variable's gate description polynomial.
Its remainder $r$ lets following lemma hold:

\begin{Lemma}
\label{lem:1}
$r$ will only contain primary inputs (bit-level and word-level) and word-level output; furthermore, there will be one and
only one term containing word-level output whose monomial is word-level output itself rather than higher order form.
\end{Lemma}

\begin{proof}
First proposition is easy to prove by contradiction. Second part, the candidate pair of polynomials only have one term of
single word-level output variable (say it is $R$) and this term is the last term under RATO, which means there is only one term with
$R$ in Spoly. Meanwhile in other polynomials from $J+J_0$ there is no such term containing $R$, so this term will be
kept to remainder $r$, in first degree.
\end{proof}

\begin{Example}
The elimination ideal for 5-bit SMPO (Ex.\ref{ex:SMPO}) could be rewritten under RATO:
\begin{align}
&(R_0,R_1,R_2,R_3,R_4)>(r_0,r_1,r_2,r_3,r_4)\nonumber\\&>(c_1,c_2,c_3,c_4,b_0,b_1,b_2,b_3,b_4)\nonumber\\
&>(a_0,a_1,a_2,a_3,a_4)>R>r>(A,B)\nonumber
\end{align}
Thus the candidate pair is
$(f_w,f_g), f_w = R_0+r_4+b_0\cdot a_1, f_g =R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18} + R$.
Result after reduction is:
\begin{align}
&Spoly(f_w,f_g) \xrightarrow{J+J_0}_{+}\nonumber\\
&r_1+(\alpha)r_2+(\alpha^4+\alpha^2)r_3+(\alpha^3+\alpha^2)r_4+(\alpha^3)b_1a_1+(\alpha^4+\alpha^2)b_1a_2\nonumber\\
+&(\alpha^3+\alpha+1)b_1a_3+(\alpha^3+\alpha)b_1a_4+(\alpha+1)b_1A+(\alpha^4+\alpha^2+\alpha)b_2a_1\nonumber\\
+&(\alpha^4+\alpha^3+\alpha^2+\alpha)b_2a_4+(\alpha^3+\alpha^2+1)b_3a_1+(\alpha)b_3a_3\nonumber\\
+&(\alpha^2+\alpha+1)b_4a_1+(\alpha+1)b_4a_2+(\alpha^4+\alpha^2)b_4a_3+(\alpha^4+\alpha^3+\alpha+1)b_4a_4\nonumber\\
+&(\alpha^3+1)b_4A+(\alpha^4+\alpha^3+\alpha^2+1)a_1B+(\alpha^4+\alpha^3+\alpha^2+1)R\nonumber
\end{align}
The remainder satisfies Lemma \ref{lem:1}.
\end{Example}

\subsection{Bit-level Variable Substitution (BLVS)}
The remainder from \emph{Spoly} contains some bit-level variables, and our objective is to get a polynomial contains only word-level variables
(such as $R+\mathcal{F}(A,B)$). One possible method is to rewrite bit-level variables in term of function on word-level
variables, i.e. $a_i = \mathcal{G}(A)$, then do substitution. A Gaussian-elimination-fashion approach could be applied to
compute corresponding $\mathcal{G}(A)$ efficiently.

\begin{Example}
{\bf Objective}:\ Abstract polynomial $a_i + \mathcal{G}_i(A)$ from $f_0: a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A$.

First, compute $f_0^2: a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2$. Apparently variable $a_0$ can be
eliminated by operation 
\begin{align}
f_1 =& f_0\times \alpha^5 + f_0^2: \nonumber\\
&a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
+&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
\end{align}
Recursively eliminate $a_1$ from $f_1$, $a_2$ from $f_2$, etc. The final polynomial $f_4$ has the form 
$a_4 + \mathcal{G}_4(A) = a_4+(\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4+(\alpha^2+1)A^2+(\alpha+1)A$. Recursively substitute
$\mathcal{G}_4(A)$ back to $f_3$, etc. The result is a set of polynomials
\begin{displaymath}
\{g_i\ | \ g_i: a_i + \mathcal{G}_i(A)\}
\end{displaymath}
\end{Example}
In this approach it is easy to get word-level variable representation for each bit-level primary inputs. By substitution, a new polynomial in the form $R+\mathcal{F}(A,B)$
could be attained.

\textit{Discussion}\ \ (a) This approach won't conduct to a result reducing to 0, because \emph{Spoly}'s remainder contains information from the whole system,
the substitution will result something new, i.e. an abstraction of the system; (b) The Gaussian-like approach is a pre-processing of a variable, and could be immediately reused
to other words with the same length; (c) For $n$-bits SMPO, the fast GB is very efficient: for each cycle, first compute a \emph{Spoly} in $O(1)$, then do multi-division by $O(n)$ polynomials
(unknown complexity, maybe lower if using F4-style reduction?), the substitution cost at most $O(n^3)$ time (assuming $O(1)$ for each term). In total repeat for $n$ cycles.
The complexity should be much lower compared to Buchberger's algorithm.
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
%\bibliographystyle{abbrv}
%\bibliography{seq_verif}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
\appendix
\section{Normal Basis Multiplication}
\subsection{$\lambda$-Matrix}
$\lambda$-Matrix is defined with cross-product terms from multiplication. That is 
\begin{displaymath}
Product\ C = (\sum_{i=0}^{n-1}a_i\beta^{2^i})(\sum_{j=0}^{n-1}b_j\beta^{2^j}) = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}a_ib_j\beta^{2^i}\beta^{2^j}
\end{displaymath}
The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
normal basis, i.e.
\begin{displaymath}
\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1}\lambda_{ij}^{(k)}\beta^{2^k}, \ \ \lambda_{ij}^{(k)} \in F_2.
\end{displaymath}
Substitution yields, get the expression for $k$-th digit of product:
\begin{displaymath}
c_k = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}\lambda_{ij}^{(k)}a_ib_j
\end{displaymath}
$\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ in $k$-th $\lambda$-Matrix.

\begin{Example}
$\lambda$-Matrix: A binary $n\times n$ matrix $M$ could be employed to describe the "function"
 mentioned in Ex.\ref{ex:nb_multi}: $c_{n-1} = f(A, B) = A \cdot M \cdot B^T$, $B^T$ denotes vector transposition. 
More specifically, denote the matrix by \emph{$k$-th $\lambda$-Matrix}: $c_k = A \cdot M^{(k)} \cdot B^T$.
Then $c_{k-1} = A \cdot M^{(k-1)} \cdot B^T = rotate(A) \cdot M^{(k)} \cdot rotate(B)^T$, which means $M^{(k)}$ is generated
by right and down cyclic shifting $M^{(k-1)}$.

In $\mathbb{F}_{2^3}$ constructed by $\alpha^3 + \alpha + 1$, let $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
is a normal basis. $0$-th $\lambda$-Matrix
\begin{displaymath}
M^{(0)} = \left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right).
\end{displaymath}
i.e.,
\begin{displaymath}
c_0 = (a_0\  a_1\  a_2)\left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right)\left(
\begin{array} {lcr}
b_0\\
b_1\\
b_2
\end{array} \right).
\end{displaymath}
\end{Example}
\subsection{Optimal Normal Basis}
\begin{Definition}
The number of non-zero entries in $\lambda$-Matrix is known as  \emph{Complexity$(C_N)$}.
\end{Definition}
\begin{Theorem}
If N is a normal basis for $\mathbb{F}_{p^n}$ with $\lambda$-matrix $M^{(k)}$, then non-zero entries in 
matrix $C_N\geq 2n-1$.
\end{Theorem}
Proof omitted.
\begin{Definition}
If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as
\emph{Optimal Normal Basis (ONB)}.
\end{Definition}
There are 2 types of optimal normal basis existing.
Rules for creating Type-I ONB over $\mathbb{F}_{2^n}$ are:
\begin{itemize}
\item n+1 must be prime.
\item 2 must be primitive in $\mathbb{Z}_{n+1}$.
\end{itemize}
Rules for creating Type-II ONB over $\mathbb{F}_{2^n}$ are:
\begin{itemize}
\item 2n+1 must be prime. And either
\item 2 must be primitive in $\mathbb{Z}_{2n+1}$, OR
\item $2n+1 \equiv 3 \ \ mod\ \  4$ AND 2 generates the quadratic residues in $\mathbb{Z}_{2n+1}$
\end{itemize}
There are corresponding criteria for simply creating $\lambda$-Matrix for either type of ONB:
\begin{Lemma}
Type-I ONB implies a $\lambda$-Matrix that each nonzero entry $M_{i,j}$ satisfies
\begin{numcases}{}
2^i + 2^j = 1 \bmod  n+1\notag\\
2^i + 2^j = 0 \bmod n+1\notag
\end{numcases}
Type-II ONB implies a $\lambda$-Matrix that each nonzero entry $M_{i,j}$ satisfies
\begin{numcases}{}
2^i + 2^j = 1\bmod\  2n+1\notag\\
2^i + 2^j = -1 \bmod  2n+1\notag\\
2^i - 2^j = 1 \bmod  2n+1\notag\\
2^i - 2^j = -1 \bmod 2n+1\notag
\end{numcases}
\end{Lemma}
Proof omitted. (Issues here: We knew type-I \& II definition can deduce corresponding $\lambda$-Matrix, how about
reverse implication? i.e. can we prove the equivalence?)

\subsection{Design a Normal Basis Multiplier using $\lambda$-Matrix}
Imitating the structure of SMPO, just specify the gates' connections for the first clock cycle, then following cycles are
automatically completed by cyclic shifting operands $A$ and $B$. The whole design procedure is based on a $n \times n$ $k$-th $\lambda$-Matrix.

First figure out the first row (\emph{row} 0) in $\lambda$-Matrix, for any nonzero entry $M_{0,j}$ (there will be only 1 if taking $0$-th $\lambda$-Matrix), place an AND gate
$a_j\land b_0$. Connect different $a_j$ with a XOR gate before place the AND gate if there are 2 or more nonzero entries in this row;

Secondly, repeat this for each row. Note for nonzero entry $M_{i,j}$, corresponding index of $a_u$ and $b_v$ is incremented by $i$, i.e. $u = j + i \pmod n$, $v = i + i \pmod n$;

Finally, connect the output of AND gate for row $i$ to one input of a XOR gate, then connect the output of XOR gate to a register/flip-flop. The output of register/flip-flop is connected
to the other input of XOR gate, but at next row. All these gates and connections for row $i$ is called unit $R_i$. (Fig.\ref{fig:SMPO})

\section{Find Optimal Normal Basis}
\subsection{Transforming to SAT}
From last section we learned it is easy to compute $\lambda$-Matrix when given certain type of optimal normal basis. However in our approach, the mapping between 
optimal normal basis and standard basis should be explicitly specified. In other words, let \emph{Optimal Normal Element} $\beta = \alpha^k$, the exponential $k$ on
primitive element $\alpha$ needs to be calculated. Otherwise word-level variable interpreting polynomial $a_0\beta + a_1\beta^2 + \cdots + a_{n-1}\beta^{2^{n-1}}$
will be an unknown polynomial and GB computation is impossible.

Currently finding the optimal normal element $\beta$ in $2^n$ (over field $\mathbb{F}_{2^n}$) elements is NP-hard. However, it is possible to improve the performance 
by transforming this problem to SAT problem which can be solved by efficient SAT solver. This is what our tool is managing to do.

Given 0-th $\lambda$-Matrix $M^{(0)}$, other $\lambda$-Matrices can be obtained by cyclic right-down shifting, i.e. entry $M_{i,j}^{(k)} = M_{i-k,j-k}^{(0)}$. Note all
index operation results should modulus $n$. $\beta^3$ is an element which can be represented as product of normal basis multiplication: $\beta^3 = \beta\cdot\beta^2$.
Consider the following equation
\begin{displaymath}
c_k = (1\  0\ 0\ \cdots\ 0)\  M^{(k)} \left(
\begin{array} {lcr}
0\\
1\\
0\\
\vdots\\
0
\end{array} \right)
\end{displaymath}
$c_k$ is the $k$-th coefficient in normal basis expression of $\beta^3$: $\beta^3 = c_0\beta + c_1\beta^2 + \cdots + c_{n-1}\beta^{2^{n-1}}$.
Thus, $\beta^3$ could be written as
\begin{displaymath}
\beta^3 = \sum_{k=0}^{n-1} M_{0,1}^{(k)}\beta^{2^k} = \sum_{k=0}^{n-1} M_{-k,1-k}^{(0)}\beta^{2^k}
\end{displaymath}
Meanwhile, the standard basis representation of optimal normal element can be assumed as
\begin{displaymath}
\beta = a_0 + a_1\alpha + \cdots + a_{n-1}\alpha^{n-1}
\end{displaymath}
Do a substitution, vanish every coefficient. Finally we can get a linear system of $n$ boolean equations (composed by XOR and AND), with $n$ unknowns $a_0,a_1,\dots,a_{n-1}$. 
Examples are in the write-up by Florian's student.

\subsection{Solving XOR-SAT}
Solving the linear system of boolean equations is a tricky problem, because ordinary SAT solver only support Conjunctive Normal Form (CNF) inputs; while a XOR-SAT based solver 
\emph{CryptoMiniSAT} only support clauses composed by single literals and XOR connectives ($a\oplus b \oplus (c\land d)$ is illegal input), I call it "XOR-CNF". Basically there are 2 approaches 
to think about our tool flow:

(a) Use another engine (Commercial Symbolic Computing System such as \emph{Wolfram Mathematica}, or Circuit Based Synthesizer such as \emph{ABC}) to transform all equations to CNF clauses.
Advantage: Convenient, even no need to write our own tool in scripts; Disadvantage: cannot guarantee the efficiency, CNF may explode if we have hundreds of equations with hundreds of 
XOR terms.

(b) Imitate clause learning technique in CDCL, create new clauses to satisfy the input format of \emph{CryptoMiniSAT}. Advantage: \emph{CryptoMiniSAT} is a tool with performance enhancing
on XOR clauses, and since the number of XOR clauses will not explode (I will explain later), maybe the efficiency can be guaranteed (somehow); Disadvantage: need to build our own tool for experiments.

Let me have a tiny statement on the second choice. First is the approach, maybe better illustrated by an example:
\begin{Example}
Change boolean equation $a\oplus b\oplus c \oplus bc = 0$ to XOR-CNF. First, a SAT-solver always want whole CNF to be TRUE. Fortunately, we can negate a XOR clause by just negate one literal:
$\overline{a}\oplus b\oplus c \oplus bc = 1$. Then we need to solve the non-single-literal term. Replace term $bc$ with new literal $d$, and create a new clause to make $d=bc$, which is
$(\overline{b}\lor\overline{c}\lor d)\land(b\lor\overline{d})\land(c\lor\overline{d})$, adding 3 short clauses.
\end{Example}
You may ask why the number and size of clauses won't explode, the answer is easy: we only have 2-literal terms in worst case! Note all equations come from 
$$\beta^3 + c_0\beta + c_1\beta + \cdots + c_{n-1}\beta^{2^{n-1}} = 0$$
Even exponentials will never generate higher order coefficients in $\mathbb{F}_{2^n}$, so the only term generates higher-order coefficients is
$$\beta^3 = (a_0+a_1\alpha + \cdots a_{n-1}\alpha^{n-1})(a_0+a_1\alpha^2+\cdots + a_{n-1}\alpha^{2(n-1)})$$
Apparently cross-product terms here have at most 2 literals (degree 2).

Acknowledging this fact, for $n$ unknowns there will be at most $\dbinom{n}{2} = O(n^2)$ 2-literal terms, so the number of newly created clause will also be $O(n^2)$, with fixed tiny size.

By either approach, we finally get one legal assignment for bit vector $(a_0,a_1,\dots,c_{n-1})$ from the SAT-solver, by that we can compute $\beta$. We can directly use it as ONB, or setup a loop to keep squaring 
$\beta$ to find the minimum $k$ for $\alpha^k$, and adopt that one as ONB.

\textit{Discussion}\ \ The tool flow is feasible enough, so only issue is the soundness from the ONB theory to linear system of boolean equations. I asked following questions to Florian's student:
1. Solutions for linear system of equations in $\mathbb{F}_2$ (or say $\mathbb{Z}_2$). For example, in last page of ONB.pdf194 KB, there are 5 variables and 5 equations, provide 5 conjugate solutions forming a set of (optimal) normal basis.
Is it possible to dig into this, say to prove all solutions are conjugates $(x, x^2, x^4, x^8, \dots)$?
2. You chose $\beta^3$ to deduce those equations. Does solution to $\beta^3$ necessarily cover (or equal to) solutions for given lambda-Matrix M0? In other words, if I choose $\beta^5$ or $\beta^6$ to construct those equations, can I get the same answer?

If above concerns are resolved, we can come up with a complete proof.
\balancecolumns
% That's all folks!
\end{document}

### 2014-04-25 sequential/latex/DAC2013/xiaojun.tex
upload files from laptop linux

\section{Further discussion on Computer Algebra}
\label{sec:moretheory}

Bit level circuit variables including primary inputs/outputs can only take value
within $\Z_2$. They can also be written in functions of other variables and constants.
In computer algebra, circuit variable $a$ can be described as a polynomial, that is
$a = f(x_0, x_1, \dots, x_{n-1}) \in {\mathbb{F}}[x]$. Moreover, circuit variable can
have multiple specifications, they can be written in polynomials and included in an
ideal $I = \{f_0, f_1, \dots, f_k\}$. The varieties of this ideal are possible 
assignments of circuit variables. Similarly for $n$ bits world-level circuit 
variables, the varieties are elements from $\Fkk$. Thus it's convenient to use
ideals and varieties to describle circuit variables. \\

\begin{Definition}
({\bf Sum of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the 
{\bf sum} of $I$ and $J$, denoted by $I + J$, is the set
  \begin{equation}
  I + J = \{f + g : f \in I \ and\  g \in J\}.
  \end{equation}
Furthermore, if $I = \langle f_1, \dots, f_r\rangle$ and 
$J = \langle g_1, \dots, g_s\rangle$, then 
$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$.
\end{Definition}

\begin{Definition}
({\bf Product of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the
{\bf product} of $I$ and $J$, denoted by $I \cdot J$, is defined to be the ideal generated 
by all polynomials $f \cdot g$ where $f \in I$ and $g \in J$. Furthermore, let
$I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$, then
  \begin{equation}
  I \cdot J = \langle f_ig_j : 1 \leq i \leq r, 1 \leq j \leq s\rangle .
  \end{equation}
\end{Definition}

\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then $I:J$
is the set
  \begin{equation}
  \{f \in k[x_1, \dots, x_n] : fg \in I, \forall g \in J\}
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}

According to definitions above, it is possible to calculate ideal sum, product and 
quotient based on ideal generators. Additional conclusion about union, 
intersection and complement of varieties can be attained below: \\

\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
\bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) = {\bf V}(I) \bigcup {\bf V}(J)$.
\end{Theorem}

\begin{Theorem}
\label{thm:quotient}
Let $V$ and $W$ be varieties in $\Fkk$. Then ${\bf I}(V) : {\bf I}(W) = 
{\bf I}(V - W)$.
\end{Theorem}

For \ref{thm:quotient}, let $V = \Fkk = {\bf V}(\langle vanishing \  polynomials
\rangle)$, then $V - W$ is the \textbf{complementary set} of variety $W$.\\

Based on these conclusions it's easy to get a circuit variable's specifications
or possible assignments from the other known information.

\section{Sequential circuits and Finite State Machine}
\label{sec:fsm}
\subsection{Introduction to Finite State Machine:}
Borrow from somewhere.
\subsection{Symbolic Image Computation:}
Borrow from somewhere.
\subsection{Implicit State Enumeration:}
BFS algorithm below:
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cdot \overline{reached}$\;
  	$reached \gets reached + new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}\label{alg:gb}
\end{algorithm}

\section{Experiments using new approach}
\label{sec:exp}

First example is reachable state enumeration of a 2-bit finite state machine.
Breadth-First-Search traversal (mentioned in \ref{sec:fsm}) is adopted.

\begin{Example}
\label{ex:traversal}
Sample circuit is described below with gate-level implement and state transition
graph.
  \begin{figure}[hbt]
    \centerline{\includegraphics[scale=0.3]{./fsm_ckt.png}}
    \caption{Sample circuit}
  \label{fig:fsmckt}
  \end{figure}

  \begin{figure}[hbt]
    \centerline{\includegraphics[scale=0.3]{./fsm_stg.png}}
    \caption{State Transition Graph}
  \label{fig:fsmstg}
  \end{figure}

The following mapping from Boolean operations to Galois field $\mathbb{F}_2$ 
operations is straightforward. Assuming $a$ and $b$ are both bit-level variables:
\begin{itemize}
  \item $\overline{a} \rightarrow 1 + a$
  \item $a\ and\ b \rightarrow ab$
  \item $a\ or\ b \rightarrow a + b + ab$
\end{itemize}
Comparing to the quantification approach in \ref{sec:fsm}, using Gr\"obner bases
approach from \ref{sec:theory}. Construct an elimination ideal with following 
polynomials: $f_1: 1 + XNOR(t_0, OR(p\cdot q, (1 + x_i)\cdot(1 + p)\cdot(1 + q))); 
f_2: 1 + XNOR(t1, OR(xi\cdot(1+p),p\cdot(1+q))); f_3: S - p - q\cdot\alpha; 
f_4: T - t_0 - t_1\cdot \alpha$ and vanishing polynomials $f_5: p^2 - p; 
f_6: q^2 - q; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: x_i^2 - x_i; 
f_{11}: T^4 - T$, where $S$ and $T$ are 2-bit word representing current state
and next state as defined in $f_3$ and $f_4$, $f_1$ and $f_2$ describe the 
transition relations. At last the initial state should also be included in 
this elimination ideal. For example initial state is $S_3$, then last polynomial
is $f_{12}: S + 1 + \alpha$. According to elimination theorem \ref{thm:elim},
let the lex ordering be $\{all\ others\} > T$, calculate Gr\"obner basis, the 
result must contain a polynomial with only one variable $T$, which can indicate
the possible assignments of next state. \par

Apply this approach to calculate image in BFS traversal \ref{alg:BFS}.
In this example, $to^i$ and $reached$ are both ideals containing only one
generator, i.e. the polynomial with single variable $T$; universal set is
varieties from ideal $\langle T^4 - T\rangle$. Use \ref{thm:unionintersect} and
\ref{thm:quotient} to complete \ref{alg:BFS}, the final return value is 
polynomial $to^3 = T^4 + T$, which means all states are rechable from state $S_3$.
\end{Example}

The second example shows the application of the new approach on sequential 
arithematic circuits function verification.

\begin{Example}
\label{exp:multiplier}

The following design is Sequential Multiplier with Parallel Output (SMPO),
a Normal Basis multiplier based on Massey-Omura algorithm. Inputs and outputs
are all 5-bit word taking value from ${\mathbb F}_{2^5}$. After loading operands
 A and B, setting all output latches to 0 and running for 6 iterations, the 
output should be $R = A\cdot B$ (mod $x^5 + x^2 + 1$).
  \begin{figure}[hbt]
    \centerline{\includegraphics[scale=0.3]{./SMPO.png}}
    \caption{5-bit SMPO}
  \label{fig:SMPO}
  \end{figure}

Similarly, build elimination ideal for all gates/operations and induce
initial states of latches. However, instead of eliminating all variables
to one, this example adopts abstraction term order from \ref{thm:abs}
and keeps the polynomial contains function between output and inputs.
Here the lex ordering is $\{others\} > R > \{A, B\}$, and objective
polynimial is $R + \F(A, B)$. \par

Apply this approach to calculate image in BFS traversal, but modify 
\ref{alg:BFS} to make it adapt 6 steps reachable states enumeration.
The result is $R + AB$, which validates the function of this circuit.

\end{Example}

### 2014-04-25 sequential/latex/ICCAD2014/abstract_fmcad.tex
upload files from laptop linux

\begin{abstract}
Sequential circuits that implement Galois field arithmetic over
$\Fkk$ are prevalent in cryptography and error control coding, where
the datapath size $k$ is very large. Formal verification of sequential
arithmetic circuits with large datapath size is beyond the
capabilities of contemporary verification techniques. To address this
problem, this paper introduces a new concept of implicit state
enumeration of finite state machines (FSMs) performed at the
word-level. Using algebraic geometry, we show that the state-space of
a sequential circuit can be encoded, canonically, as solutions to a
multivariate word-level polynomial $\F(A)$ in $\Fkk$, where $A = {a_0,
\dots, a_{k-1}}$ represents the $k$-bit state-register (word). Subsequently,
the Groebner basis algorithm, along with elimination ideals and quotients 
of ideals can be employed for FSM traversal. We apply our approach 
to verify sequential Galois field multipliers with large datapath sizes, 
where contemporary techniques prove to be infeasible.
\end{abstract}

### 2014-04-25 sequential/latex/ICCAD2014/appendix_fmcad.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/DAC2013/theory.tex
@@
-% THIS IS SIGPROC-SP.TEX - VERSION 3.1
-% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
-% APRIL 2009
-%
-% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
-% LaTeX2e document class file for Conference Proceedings submissions.
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) Page numbering
-% ---------------------------------------------------------------------------------------------------------------
-% It is an example which *does* use the .bib file (from which the .bbl file
-% is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission,
-% you need to 'insert'  your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% Questions regarding SIGS should be sent to
-% Adrienne Griscti ---> griscti@acm.org
-%
-% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
-% Gerald Murray ---> murray@hq.acm.org
-%
-% For tracking purposes - this is V3.1SP - APRIL 2009
-
-\documentclass{acm_proc_article-sp}
-
-%packages
-\usepackage[linesnumbered, ruled]{algorithm2e}
-\RequirePackage{amssymb, mathptm}
-\usepackage{amsbsy}
-\usepackage{graphicx}
-\usepackage{helvet}
-\usepackage{enumerate}
-\usepackage{amsmath}
-\usepackage{cases}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{subfig}
-\usepackage{comment}
-
-%new format
-\newtheorem{Algorithm}{Algorithm}
-\newdef{Definition}{Definition}
-\newtheorem{Theorem}{Theorem}
-\newtheorem{Lemma}{Lemma}
-\newdef{Example}{Example}
-
-\begin{document}
-
-\title{Verification of Sequential Galois Field Circuits using Algebraic Geometry
-\titlenote{Version 0.3, refined theory part with normal basis, proposed details for designing 2 examples.}}
-%\subtitle{[Extended Abstract]
-%\titlenote{A full version of this paper is available as
-%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-%\LaTeX$2_\epsilon$\ and BibTeX} at
-%\texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Xiaojun Sun\\
-       \affaddr{University of Utah}\\
-       \affaddr{Department of Electrical \& Computer Engineering}\\
-       \affaddr{Salt Lake City, USA}\\
-       \email{xiaojun.sun@utah.edu}
-% 2nd. author
-\alignauthor
-Priyank Kalla\\
-       \affaddr{University of Utah}\\
-       \affaddr{Department of Electrical \& Computer Engineering}\\
-       \affaddr{Salt Lake City, USA}\\
-       \email{kalla@ece.utah.edu}
-% 3rd. author
-%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-%one who did all the really hard work.}\\
-%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-%       \affaddr{Hekla, Iceland}\\
-%       \email{larst@affiliation.org}
-%\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-%\alignauthor Lawrence P. Leipuner\\
-%       \affaddr{Brookhaven Laboratories}\\
-%       \affaddr{Brookhaven National Lab}\\
-%       \affaddr{P.O. Box 5000}\\
-%       \email{lleipuner@researchlabs.org}
-% 5th. author
-%\alignauthor Sean Fogarty\\
-%       \affaddr{NASA Ames Research Center}\\
-%       \affaddr{Moffett Field}\\
-%       \affaddr{California 94035}\\
-%       \email{fogartys@amesres.org}
-% 6th. author
-%\alignauthor Charles Palmer\\
-%       \affaddr{Palmer Research Laboratories}\\
-%      \affaddr{8600 Datapoint Drive}\\
-%%      \affaddr{San Antonio, Texas 78229}\\
-%       \email{cpalmer@prl.com}
-}
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{21 Jan 2014}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-\begin{abstract}
-Circuits working in Galois fields are increasingly employed in designs like arithmetic component for Elliptic Curve Cryptography (ECC). 
-This work proposes a new method to effectively verify sequential circuits in Galois fields. Algebraic geometry
-is introduced to describe circuits behavior and expand the definition of implicit state space traversal. Moreover, Gr\"obner basis
-representation is adopted for word-level abstraction on circuit variables to address state explosion problem
-with BDDs. Experiments are run with fast Gr\"obner basis computation engine to get competitive results.
-\end{abstract}
-
-% A category with the (minimum) three required fields
-\category{EDA5.1}{Verification}{Functional, transaction-level, RTL, and gate-level modeling and verification of hardware design}
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-\terms{Verification, Algorithms}
-
-\keywords{Verification, Sequential, Galois Fields, Algebraic Geometry} % NOT required for Proceedings
-
-\section{Essential Theory}
-\subsection{Polynomial Representation}
-A typical sequential circuit is composed as figure \ref{fig:sequential}. The combinational logic has primary input $x$, state
-inputs $\{s_0, s_1, \dots, s_{k-1}\}$, primary output $z$ and state outputs $t_0, t_1, \dots, t_{k-1}$. State outputs
-are latched into registers, which will be fed back to state inputs in next clock cycle.
-
-\begin{figure}
-\centering
-\psfig{file=./sequential_fig.ps, height=2in, width=3.5in,}
-\caption{A typical sequential circuit block view}
-\label{fig:sequential}
-\end{figure}
-
-A transition function $\Delta$ is a Boolean function about state outputs and all inputs, i.e.
-\begin{displaymath}
-t_i = \Delta_i(x, s_0, s_1, \dots, s_{k-1})
-\end{displaymath}
-This constrain could also be written in the form of equation $polynomial = 0$:
-\begin{displaymath}
-t_i - \Delta_i(x, s_0, s_1, \dots, s_{k-1}) = 0
-\end{displaymath}
-Solution to this equation is limited within $\{0, 1\}$, which is set of all elements in Galois field $\mathbb{F}_2$.
-Furthermore, there exists a one-to-one corresponding relation from Boolean space to Galois field on higher dimension: 
-$\mathbb{B}^k \to \mathbb{F}_{2^k}$, which means it is possible to find a set of elements in $\mathbb{F}_{2^k}$ representing
-all Boolean values every bit of a bit vector can take. In Galois field, a as \ref{table:booltogalois} shows.
-\begin{table}[h]
-\centering
-\caption{Bit-vector, Exponential and Polynomial representation of
-elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
-\pmod{x^4+x^3+1}$}\label{table:booltogalois}  
-\begin{tabular}{|c|c||c|c|} 
-\hline
-$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
-\hline
-$0000$        & $0$           & $1000$  &$\alpha^3$\\
-\hline
-$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
-\hline
-$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
-\hline
-$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
-\hline
-$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
-\hline
-$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
-\hline
-$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
-\hline
-$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
-\hline
-\end{tabular}
-\end{table}
-A word-level representation on state inputs/outputs can be attained through above correspondence. If the set of 
-state inputs $\{s_0, s_1, \dots, s_{k-1}\}$ is denoted by $S$ taking values from $\mathbb{F}_{2^k}$, and
-set of outputs $\{t_0, t_1, \dots, t_{k-1}\}$ is denoted by $T$, a word-level transition function can be written as
-\begin{displaymath}
-T - \Delta(x, S) = 0
-\end{displaymath}
-If $T - \Delta(x, S)$ is taken as a polynomial $f$, then $f \in \mathbb{F}_{2^k}[x]\ mod\ P(\alpha)$.
-\subsection{Gr\"obner Basis Theory}
-Gr\"obner basis is used to calculate desired results out of a set of polynomials.
-If there exists multiple polynomials $\{f_1, f_2, \dots, f_r\}$ representing constrains on $T, x$ and $S$,
-then all polynomials should simultaneously equal to $0$.
-\begin{displaymath}
-\left\{
-  \begin{array}{lc}
-  f_1 & = 0\\
-  f_2 & = 0\\
-  \dots & \ \\
-  f_r & = 0
-  \end{array} \right.
-\end{displaymath}
-Solution to these equations is set of values of $(T, x, S)$, which is called \textbf{Variety} of \textbf{ideal}
-$I = \langle f_1, f_2, \dots, f_r\rangle $.
-
-Calculate Reduced Gr\"obner Basis (RGB) from ideal $I$ with \textbf{elimination term order}, there exists a polynomial contains
-only $T$, if the values of states input $S$ and primary input $x$ are given in the form of polynomials and included by $I$,
-the values of $T$ can be computed.
-
-Otherwise, if RGB calculation is manipulated under \textbf{abstraction term order}, there exists a polynomial in
-the form of $T - \mathcal{F}(x, S)$ representing transition function. In next clock cycle, assign new state input $S'$ with $T$, the next
-state can be computed again, result is polynomial $T' - \mathcal{F}(x', S')$.
-
-\subsection{Normal Basis Representation}
-
-Let $\beta$ be an element in the Galois field $F_{2^n}$ constructed by primitive element $\alpha$ and minimal polynomial
-$P(\alpha)$. Then a basis in the form $\{\beta, \beta^2, \beta^4, \beta^8, ... ,\beta^{2^{n-1}}\}$ is a
-\emph{Normal Basis}; here $\beta$ is called \emph{Normal Element}.
-
-For arithmetic operations in Galois fields, squaring and multiplication (with modulus) can be greatly simplified if 
-normal basis representation is adopted to represent operands.
-\begin{Example}
-\label{ex:nb_sq}
-Element squaring: In $F_{2^n}$, all coefficients which can be divided by 2 are reduced, so 
-following equality holds for all field elements $a$ and $b$:
-$$(a+b)^2 = a^2 + b^2$$ 
-Apply this rule on element squaring of standard/polynomial basis:
-\begin{align}
-& (b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{n-1}\beta^{2^{n-1}})^2 \nonumber\\
-&= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{n-1}^2\beta^{2^n} \nonumber\\
-&= b_{n-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{n-2}\beta^{2^{n-1}} \nonumber
-\end{align}
-using Fermat's little theorem $\beta^{2^n} = \beta$. This shows that squaring of field elements
-represented by normal bases can be easily completed by operating a simple right-cyclic rotation.
-\end{Example}
-
-\begin{Example}
-\label{ex:nb_multi}
-Normal basis multiplication: There are 2 binary vectors representing operands of multiplication in normal
-basis: $$A = (a_0, a_1, \dots, a_{n-1}),\ B = (b_0, b_1, \dots, b_{n-1})$$ 
-the product is also written in normal basis representation: $$C = A*B = (c_0, c_1, \dots, c_{n-1})$$
-Describe calculation procedure for the most significant digit of product with a function: 
-$$c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})$$
-Square both side: $C^2 = A^2*B^2$, i.e. the second significant digit 
-$$c_{n-2} = f(a_{n-1}, a_0, a_1, \dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2})$$ 
-By this method it is easy to get all digits of product $C$.
-\end{Example}
-
-\section{Algebraic Geometry Approach}
-\subsection{State Space Traversal}
-
-\begin{figure}
-\centering
-\psfig{file=./fsm_fig.ps, height=2in, width=3.5in,}
-\caption{Gate-level circuit of combinational logic block of sample FSM}
-\label{fig:fsm}
-\end{figure}
-
-\begin{figure}
-\centering
-\psfig{file=./stg_fig.ps, height=2in, width=3.5in,}
-\caption{State transition graph of sample FSM}
-\label{fig:stg}
-\end{figure}
-
-The first example is a mealy finite state machine (FSM). Gate-level circuits of combinational logic
-is figure \ref{fig:fsm}, $x$ is primary input,
-$\{s_0, s_1\}$ are state inputs, $Z$ is primary output and $\{t_0, t_1\}$ are state outputs.
-Word-level state input is $S = s_0 + s_1\cdot\alpha$, output is $T = t_0 + t_1\cdot\alpha$. In Galois field
-$\mathbb{F}_{2^2}$, bit-level variables $x, s_0, s_1, Z, t_0, t_1$ can take values from $\{0, 1\}$, while
-word-level variables $S, T$ may take values from $\{0, 1, \alpha, 1 + \alpha\}$. State transition graph (STG)
-showed in figure \ref{fig:stg} uses 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2, S_3\}$, which
-could be converted to elements in $\mathbb{F}_{2^2}$ similarly as table \ref{table:booltogalois} shows.
-
-One important technique to check sequential equivalence is \textbf{state space traversal}. In this example,
-Breath-First-Search (BFS) is employed for traversal, which can be described with following algorithm:
-
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
- \KwIn{Transition functions $\Delta$, initial state $S^0$}
-
-  $from^0 = reached = S^0$\;
-  \Repeat{$new^i == 0$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$Img$(\Delta, from^{i-1})$\;
-	$new^i \gets to^i \cap \overline{reached}$\;
-  	$reached \gets reached \cup new^i$\;
-	$from^i \gets new^i$\;
-  }
-\Return{$reached$}
-\caption {Breadth-first Traversal Algorithm}\label{alg:BFS}
-\end{algorithm}
-
-Our approach is implementing state space traversal by refined BFS algorithm combined with polynomial representation
-and Gr\"obner basis theory. In practice, this example will show how to apply concepts and techniques from algebraic
-geometry on each line of above BFS algorithm.
-
-\subsection{States and Varieties of Ideal}
-
-\begin{Theorem}
-State variables $S, T$ and sets of states such as $from^i, to^i$ can always be represented as varieties of ideals.
-\end{Theorem}
-For example in Line 1 of BFS algorithm, assume initial state is $S_3$ in figure \ref{fig:stg}, then corresponding
-polynomial $f = \mathcal{F}(S^0) = S^0 - 1 - \alpha$. Consider an ideal $I$ with only one generator $f$, its variety
-$V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma - 1 - \alpha = 0\}$, which equals to $\{1 + \alpha\}$, the only
-valid value $S^0$ can take.
-
-If an ideal contains multiple polynomial specifications, it is necessary to compute Gr\"obner basis with elimination term
-order to get one polynomial only on desired variable. In first iteration of Line 4, $to^i$ has multiple specifications,
-some of them are transition functions on bit-level:
-\begin{displaymath}
-\begin{array}{ll}
-f_1:& t_0 - (\overline{x}\ and\ \overline{s_0}\ and\ \overline{s_1})\ or\ (s_0\ and\ s_1) \\
-f_2:& t_1 - (\overline{s_0}\ and\ x)\ or\ (p\ and\ \overline{s_1})\
-\end{array}
-\end{displaymath}
-And some are bits-to-word definitions fitting polynomial representation of elements in $\mathbb{F}_{2^2}$:
-\begin{displaymath}
-\begin{array}{ll}
-f_3:& S - s_0 - s_1\alpha \\
-f_4:& T - t_0 - t_1\alpha
-\end{array}
-\end{displaymath}
-And an polynomial about initial state as mentioned above:
-\begin{displaymath}
-f_5:\  S - 1 - \alpha
-\end{displaymath}
-And the rests are vanishing polynomials for every variable, bit-level and word level:
-$f_6: x^2 - x; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: s_0^2 - s_0;
-f_{11}: s_1^2 - s_1; f_{12}: T^4 - T$
-
-Transition equations here contains some Boolean operators, they can be re-written in terms of operations in Galois fields. 
-In $\mathbb{F}_{2}$, let $TRUE = 1, FALSE = 0$, for either element $a$ in this field, considering $0 + 0 = 1 + 1 \equiv 0\ (mod\ 2)$
-and $0 + 1 = 1 + 0 \equiv 1\ (mod\ 2)$, the inverse of $a$ is: $\overline{a} = 1 + a$. Similarly all Boolean operators
-can be converted within $\mathbb{F}_{2}$, table \ref{table:booltogalois_op} gives part of them and their corresponding
-operations in $\mathbb{F}_{2}$:
-Also note that there is no specification on initial primary input $x$, in Gr\"obner basis approach this means $x$ is smoothed by
-reversely using Shannon's expansion.
-\begin{table}
-\centering
-\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
-\label{table:booltogalois_op}
-\begin{tabular}{|c|c|} \hline
-Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
-$\overline{a}$ & $1 + a$\\ \hline
-$a\ and\ b$ & $ab$\\ \hline
-$a\ or\ b$ & $a + b + ab$\\ \hline
-$a \oplus b$ & $a + b$\\
-\hline\end{tabular}
-\end{table}
-Using elimination term order: $others\ >\ bit$-$level\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
-$J = \langle f_1, f_2, \dots, f_{12}\rangle $, the result will include one polynomial $f_T$ contains only variable $T$. In this example,
-$f_T = T + 1$. This polynomial equals to $T - 1$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
-limited within $\mathbb{F}_{2}$, where $1 \equiv -1\ (mod\ 2)$. Solution to $f_T = 0$ is $T = 1$, which shows that next
-state the machine just reached is $\{S_1(01)\}$.
-
-\subsection{Application of Algebraic Geometry}
-
-There are some difficulties with polynomial representation when executing Line 5 and Line 6, it is necessary to explore
-how \textbf{Union}, \textbf{Intersection} and \textbf{Complement Set} works in Galois field $\mathbb{F}_{2^k}$. Since
-state set variables $from^i, to^i, reached$ all have single polynomial, consider an ideal $I$ with single generator $I = \langle f\rangle $.
-Example: assume $I = \langle f\rangle  = \langle T^2 + (1+\alpha)\cdot T+\alpha\rangle $, its variety $V(I) = \{a\ |\ a \in \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$.
-So it is reasonable to specify: union, intersection and complement set mentioned in this paper are all functions on \textbf{varieties}.
-To better discuss this problem, introduce following concepts from algebraic geometry:
-\begin{Definition}
-\label{def:sum}
-({\bf Sum of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the 
-{\bf sum} of $I$ and $J$, denoted by $I + J$, is the set
-  \begin{equation}
-  I + J = \{f + g\ |\ f \in I \ and\  g \in J\}.
-  \end{equation}
-Furthermore, if $I = \langle f_1, \dots, f_r\rangle$ and 
-$J = \langle g_1, \dots, g_s\rangle$, then 
-$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$.
-\end{Definition}
-\begin{Definition}
-\label{def:prod}
-({\bf Product of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the
-{\bf product} of $I$ and $J$, denoted by $I \cdot J$, is defined to be the ideal generated 
-by all polynomials $f \cdot g$ where $f \in I$ and $g \in J$. Furthermore, let
-$I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$, then
-  \begin{equation}
-  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle .
-  \end{equation}
-\end{Definition}
-\begin{Definition}
-({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then $I:J$
-is the set
-  \begin{equation}
-  \{f \in k[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}
-  \end{equation}
-and is called the {\bf ideal quotient} of $I$ by $J$.
-\end{Definition}
-These concepts can lead the way to solution by adopting following theorems:
-\begin{Theorem}
-\label{thm:unionintersect}
-If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
-\bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) = {\bf V}(I) \bigcup {\bf V}(J)$.
-\end{Theorem}
-\begin{Theorem}
-\label{thm:quotient}
-If $I, J$ are ideals with only one generator, then ${\bf V}(I:J) ={\bf V}(I) - {\bf V}(J)$.
-\end{Theorem}
-Proof to these theorems are referred to (that yellow book?) and (my write-up?).
-
-For varieties intersection $\{1\}\bigcap\{1, \alpha\}$, calculate ideal sum $\langle T+1, T^2 + (1+\alpha)\cdot T+\alpha\rangle  = \langle T+1\rangle $,
-its variety is $\{1\}$; for varieties union $\{1,1+\alpha\}\bigcup\{1+\alpha\}$, calculate
-ideal product $\langle (T+1+\alpha)\cdot(T^2 + (1+\alpha)\cdot T+\alpha)\rangle  = \langle T^3 + 1\rangle $, its variety
-is $\{1, \alpha, 1+\alpha\}$; for complement set of variety $\{1, \alpha\}$, the universal set is
-the variety of ideal of vanishing polynomial $V(\langle T^4-T\rangle ) = \{0,1,\alpha,1+\alpha\}$,
-so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
-which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
-the result is $\{0,1+\alpha\}$.
-
-From discussion above, the BFS traversal is completely implemented in Galois field.
-If the initial input is $S_3$: $S + 1 + \alpha$, the final return value will be set of
-reachable states: $T^4 + T$, i.e. the universal set, $\{S_0, S_1, S_2, S_3\}$.
-
-\subsection{Discussion on Ideal quotient and its generator}
-In nature, we are looking varieties as the solutions to specific polynomial $f=0$ WITHIN $\mathbb{F}_{2^n}$, because (affine) varieties are defined
-within algebraic closure $k^n$ but we only care about those fall in $\mathbb{F}_{2^n}$.
-
-Theorem \ref{thm:unionintersect} related to DEF. \ref{def:sum} and \ref{def:prod} counts on ideal generators, so there will be no ambiguity on 
-new ideals we get from this theorem. It also appeared in Avrunin's CAV paper. The only problem is what ideal quotient operation will give us.
-This should be further discussed because the completeness of our algebraic geometry approach will benefit from it. \emph{Singular} has the function of
-ideal quotient, so ask from the author maybe another choice.
-
-If we put this conception aside, it is also feasible to define complement set from "solution to 
-equation" sense. Assume $f(x) = x + 1 + \alpha = 0$, its solution within $\mathbb{F}_{2^2}$ 
-is $\{1+\alpha\}$; meanwhile solutions to vanishing polynomial $v(x) = x^2 + x = 0$ are
-$\{0,1,\alpha,1+\alpha\}$. Now compute solutions to $$g(x) = \frac{v(x)}{f(x)} = 0$$,
-this requires $v(x) = 0$ and $f(x) \neq 0$, which means the complement set of solution 
-to $f(x) = 0$.
-
-Issues to this interpretation: (a) Is $v(x)$ always divisible by $f(x)$? (b) Is $v(x)=0$ and
-$f(x) \neq 0$ equivalent with $g(x)=0$ on their solutions? Could $f(x)$ equal to $0$?
-
-\section{Functional Verification}
-\subsection{Sequential Galois Arithmetic Multiplier}
-Our experiment is to verify the function of a Galois field multiplier after certain clock cycles.
-The following design is Sequential Multiplier with Parallel Output (SMPO),
-a Normal Basis multiplier based on Massey-Omura algorithm. Inputs and outputs
-are all 5-bit word taking value from ${\mathbb F}_{2^5}$. After loading operands
- A and B, setting all output latches to 0 and running for 5 iterations, the 
-output should be $R = A\cdot B$ (mod $x^5 + x^2 + 1$).
-\begin{figure}
-\centering
-\epsfig{file=./mySMPO.eps, height=2in, width=3.5in}
-\caption{5-bit SMPO}
-\label{fig:SMPO}
-\end{figure}
-Similarly, build elimination ideal for all gates/operations and induce
-initial states of latches. However, instead of eliminating all variables
-to one, this example adopts abstraction term order
-and keeps the polynomial contains function between output and inputs.
-Here the lex ordering is $\{others\} > R > \{A, B\}$, and objective
-polynomial is $R + \mathcal{F}(A, B)$.
-
-Apply this approach to calculate image in BFS traversal, but modify algorithm
-\ref{alg:BFS} to make it adapt 5 steps reachable states enumeration.
-The result is $R + AB$, which validates the function of this circuit.
-
-\subsection{Gr\"obner basis based Approach}
-
-For 5-bit normal basis multiplication, the $i$-th digit of product can be written as
-\begin{align}
-R_i &= b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
-&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
-(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
-\end{align}
-It is possible to calculate every conjunctive term in the product simultaneously within one clock cycle,
-on distinct bits. 
-Use cyclic similarity of above function, following executions are operated: 
-\begin{Example}
-Sequential Multiplier Protocol:
-\begin{itemize}
-\item \textbf{Initial}\ \ $R_0 = R_1 = R_2 = R_3 = R_4 = 0$
-\item \textbf{Clock 1}\ \ $R_0 = a_1b_0, R_1 = b_2(a_1 + a_4), R_2 = b_4(a_0 + a_1), R_3 = b_1(a_4 + a_0), 
-			R_4 = b_3(a_1 + a_3)$
-\item \textbf{Clock 2}\ \ $R_0 = b_3(a_1 + a_3) + a_0b_4, R_1 = a_1b_0 + b_1(a_0 + a_3), R_2 = b_2(a_1 + a_4)
-			+ b_3(a_4 + a_0), R_3 = b_4(a_0 + a_1) + b_0(a_3 + a_4), R_4 = b_1(a_4 + a_0) + b_2(a_0 + a_2)$
-\item \textbf{$\dots$}
-\item \textbf{Clock 5}\ \ $R_0 = c_0, R_1 = c_1, R_2 = c_2, R_3 = c_3, R_4 = c_4$, i.e. $R = A\cdot B$.
-\end{itemize}
-\end{Example}
-
-In BFS algorithm, the \textbf{Image} function is implemented by constructing an elimination ideal then 
-compute its Gr\"obner basis. 
-\begin{Example}
-\label{ex:SMPO}
-For 5-bit SMPO, the ideal consists of (for the first clock cycle):
-
-(a) {\bf Gate descriptions:}
-$a_1+a_4+c_1, a_1+a_0+c_2, a_0+a_4+c_3, a_1+a_3+c_4,
-		  a_1b_0+r_4+R_0, c_1b_2+r_0+R_1, c_2b_4+r_1+R_2, c_3b_1+r_2+R_3, c_4b_3+r_3+R_4;$
-
-(b) {\bf Word-level variables:}
-$A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
-		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18},
-		  r+r_0\alpha^5+r_1\alpha^{10}+r_2\alpha^{20}+r_3\alpha^9+r_4\alpha^{18},
-		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18};$
-
-(c) {\bf Vanishing polynomials:}
-		 $ a_0^2+a_0, a_1^2+a_1, a_2^2+a_2, a_3^2+a_3, a_4^2+a_4,
-		  b_0^2+b_0, b_1^2+b_1, b_2^2+b_2, b_3^2+b_3, b_4^2+b_4,
-		  r_0^2+r_0, r_1^2+r_1, r_2^2+r_2, r_3^2+r_3, r_4^2+r_4,
-		  R_0^2+R_0, R_1^2+R_1, R_2^2+R_2, R_3^2+R_3, R_4^2+R_4,
-		  c_1^2+c_1, c_2^2+c_2, c_3^2+c_3, c_4^2+c_4,
-		  A^{32}+A, B^{32}+B, r^{32}+r, R^{32}+R;$
-
-(d)	{\bf Feedback input:}	  $r_{in}$.
-
-Polynomial $r_{in}$ equals to $from^{i-1}$ in Line 4, BFS algorithm. Under abstraction term ordering,
-polynomial $to^i$ is assigned with a polynomial in result Gr\"obner basis which has the form $R + \mathcal{F}(A,B)$. 
-Since all outputs are connected to feedback inputs
-in SMPO, Line 5 and 6 will be neglected. Line 7 is finished by replace current output $R$ with previous 
-output $r$. Initially $r_{in} = 0$.
-
-In next clock cycle, $r_{in} = r + \mathcal{F}(A,B)$ updated from latest result; simultaneously operands
-$A$ and $B$ have been cyclically shifted, so gate descriptions in (a) should be modified accordingly. The 
-loop is operated for 5 times, result from the last step's Gr\"obner basis should be polynomial $R + AB$.
-\end{Example}
-
-This experiment can be repeated on $n$-bits SMPO after running for $n$ clock cycles.
-
-\section{Fast Gr\"obner basis computation}
-\subsection{Refined abstraction term order (RATO)}
-A lexicographic order constrained by following relation $>_{r}$: "circuit variables ordered reverse topologically" $>$ 
-"designated word-level output" $>$ "word-level inputs" is called the \emph{Refined Abstraction Term Order (RATO)}.
-
-In Buchberger's algorithm, the specification polynomial (Spoly) for each pair is calculated. In RATO, most polynomials
-have relatively prime leading terms/monomials (which means $Spoly \xrightarrow{J+J_0}_{+} 0$) except one pair:
-word-level polynomial corresponding to outputs and its leading bit-level variable's gate description polynomial.
-Its remainder $r$ lets following lemma hold:
-
-\begin{Lemma}
-\label{lem:1}
-$r$ will only contain primary inputs (bit-level and word-level) and word-level output; furthermore, there will be one and
-only one term containing word-level output whose monomial is word-level output itself rather than higher order form.
-\end{Lemma}
-
-\begin{proof}
-First proposition is easy to prove by contradiction. Second part, the candidate pair of polynomials only have one term of
-single word-level output variable (say it is $R$) and this term is the last term under RATO, which means there is only one term with
-$R$ in Spoly. Meanwhile in other polynomials from $J+J_0$ there is no such term containing $R$, so this term will be
-kept to remainder $r$, in first degree.
-\end{proof}
-
-\begin{Example}
-The elimination ideal for 5-bit SMPO (Ex.\ref{ex:SMPO}) could be rewritten under RATO:
-\begin{align}
-&(R_0,R_1,R_2,R_3,R_4)>(r_0,r_1,r_2,r_3,r_4)\nonumber\\&>(c_1,c_2,c_3,c_4,b_0,b_1,b_2,b_3,b_4)\nonumber\\
-&>(a_0,a_1,a_2,a_3,a_4)>R>r>(A,B)\nonumber
-\end{align}
-Thus the candidate pair is
-$(f_w,f_g), f_w = R_0+r_4+b_0\cdot a_1, f_g =R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18} + R$.
-Result after reduction is:
-\begin{align}
-&Spoly(f_w,f_g) \xrightarrow{J+J_0}_{+}\nonumber\\
-&r_1+(\alpha)r_2+(\alpha^4+\alpha^2)r_3+(\alpha^3+\alpha^2)r_4+(\alpha^3)b_1a_1+(\alpha^4+\alpha^2)b_1a_2\nonumber\\
-+&(\alpha^3+\alpha+1)b_1a_3+(\alpha^3+\alpha)b_1a_4+(\alpha+1)b_1A+(\alpha^4+\alpha^2+\alpha)b_2a_1\nonumber\\
-+&(\alpha^4+\alpha^3+\alpha^2+\alpha)b_2a_4+(\alpha^3+\alpha^2+1)b_3a_1+(\alpha)b_3a_3\nonumber\\
-+&(\alpha^2+\alpha+1)b_4a_1+(\alpha+1)b_4a_2+(\alpha^4+\alpha^2)b_4a_3+(\alpha^4+\alpha^3+\alpha+1)b_4a_4\nonumber\\
-+&(\alpha^3+1)b_4A+(\alpha^4+\alpha^3+\alpha^2+1)a_1B+(\alpha^4+\alpha^3+\alpha^2+1)R\nonumber
-\end{align}
-The remainder satisfies Lemma \ref{lem:1}.
-\end{Example}
-
-\subsection{Bit-level Variable Substitution (BLVS)}
-The remainder from \emph{Spoly} contains some bit-level variables, and our objective is to get a polynomial contains only word-level variables
-(such as $R+\mathcal{F}(A,B)$). One possible method is to rewrite bit-level variables in term of function on word-level
-variables, i.e. $a_i = \mathcal{G}(A)$, then do substitution. A Gaussian-elimination-fashion approach could be applied to
-compute corresponding $\mathcal{G}(A)$ efficiently.
-
-\begin{Example}
-{\bf Objective}:\ Abstract polynomial $a_i + \mathcal{G}_i(A)$ from $f_0: a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A$.
-
-First, compute $f_0^2: a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2$. Apparently variable $a_0$ can be
-eliminated by operation 
-\begin{align}
-f_1 =& f_0\times \alpha^5 + f_0^2: \nonumber\\
-&a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
-+&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
-\end{align}
-Recursively eliminate $a_1$ from $f_1$, $a_2$ from $f_2$, etc. The final polynomial $f_4$ has the form 
-$a_4 + \mathcal{G}_4(A) = a_4+(\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4+(\alpha^2+1)A^2+(\alpha+1)A$. Recursively substitute
-$\mathcal{G}_4(A)$ back to $f_3$, etc. The result is a set of polynomials
-\begin{displaymath}
-\{g_i\ | \ g_i: a_i + \mathcal{G}_i(A)\}
-\end{displaymath}
-\end{Example}
-In this approach it is easy to get word-level variable representation for each bit-level primary inputs. By substitution, a new polynomial in the form $R+\mathcal{F}(A,B)$
-could be attained.
-
-\textit{Discussion}\ \ (a) This approach won't conduct to a result reducing to 0, because \emph{Spoly}'s remainder contains information from the whole system,
-the substitution will result something new, i.e. an abstraction of the system; (b) The Gaussian-like approach is a pre-processing of a variable, and could be immediately reused
-to other words with the same length; (c) For $n$-bits SMPO, the fast GB is very efficient: for each cycle, first compute a \emph{Spoly} in $O(1)$, then do multi-division by $O(n)$ polynomials
-(unknown complexity, maybe lower if using F4-style reduction?), the substitution cost at most $O(n^3)$ time (assuming $O(1)$ for each term). In total repeat for $n$ cycles.
-The complexity should be much lower compared to Buchberger's algorithm.
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-%\bibliographystyle{abbrv}
-%\bibliography{seq_verif}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
 \appendix
-\section{Normal Basis Multiplication}
-\subsection{$\lambda$-Matrix}
+\subsection{Normal Basis Multiplication and $\lambda$-Matrix}
 $\lambda$-Matrix is defined with cross-product terms from multiplication. That is 
@@
 
-\section{Find Optimal Normal Basis}
-\subsection{Transforming to SAT}
-From last section we learned it is easy to compute $\lambda$-Matrix when given certain type of optimal normal basis. However in our approach, the mapping between 
-optimal normal basis and standard basis should be explicitly specified. In other words, let \emph{Optimal Normal Element} $\beta = \alpha^k$, the exponential $k$ on
-primitive element $\alpha$ needs to be calculated. Otherwise word-level variable interpreting polynomial $a_0\beta + a_1\beta^2 + \cdots + a_{n-1}\beta^{2^{n-1}}$
-will be an unknown polynomial and GB computation is impossible.
+\subsection{Find Optimal Normal Basis}
+This step is done by directly constructing the field using special primitive polynomial $P(x)$. Primitive element $\alpha$ which is root of that primitive polynomial( $P(\alpha)=0$),
+is also optimal normal element. i.e. $\beta = \alpha$.
 
-Currently finding the optimal normal element $\beta$ in $2^n$ (over field $\mathbb{F}_{2^n}$) elements is NP-hard. However, it is possible to improve the performance 
-by transforming this problem to SAT problem which can be solved by efficient SAT solver. This is what our tool is managing to do.
-
-Given 0-th $\lambda$-Matrix $M^{(0)}$, other $\lambda$-Matrices can be obtained by cyclic right-down shifting, i.e. entry $M_{i,j}^{(k)} = M_{i-k,j-k}^{(0)}$. Note all
-index operation results should modulus $n$. $\beta^3$ is an element which can be represented as product of normal basis multiplication: $\beta^3 = \beta\cdot\beta^2$.
-Consider the following equation
-\begin{displaymath}
-c_k = (1\  0\ 0\ \cdots\ 0)\  M^{(k)} \left(
-\begin{array} {lcr}
-0\\
-1\\
-0\\
-\vdots\\
-0
-\end{array} \right)
-\end{displaymath}
-$c_k$ is the $k$-th coefficient in normal basis expression of $\beta^3$: $\beta^3 = c_0\beta + c_1\beta^2 + \cdots + c_{n-1}\beta^{2^{n-1}}$.
-Thus, $\beta^3$ could be written as
-\begin{displaymath}
-\beta^3 = \sum_{k=0}^{n-1} M_{0,1}^{(k)}\beta^{2^k} = \sum_{k=0}^{n-1} M_{-k,1-k}^{(0)}\beta^{2^k}
-\end{displaymath}
-Meanwhile, the standard basis representation of optimal normal element can be assumed as
-\begin{displaymath}
-\beta = a_0 + a_1\alpha + \cdots + a_{n-1}\alpha^{n-1}
-\end{displaymath}
-Do a substitution, vanish every coefficient. Finally we can get a linear system of $n$ boolean equations (composed by XOR and AND), with $n$ unknowns $a_0,a_1,\dots,a_{n-1}$. 
-Examples are in the write-up by Florian's student.
-
-\subsection{Solving XOR-SAT}
-Solving the linear system of boolean equations is a tricky problem, because ordinary SAT solver only support Conjunctive Normal Form (CNF) inputs; while a XOR-SAT based solver 
-\emph{CryptoMiniSAT} only support clauses composed by single literals and XOR connectives ($a\oplus b \oplus (c\land d)$ is illegal input), I call it "XOR-CNF". Basically there are 2 approaches 
-to think about our tool flow:
-
-(a) Use another engine (Commercial Symbolic Computing System such as \emph{Wolfram Mathematica}, or Circuit Based Synthesizer such as \emph{ABC}) to transform all equations to CNF clauses.
-Advantage: Convenient, even no need to write our own tool in scripts; Disadvantage: cannot guarantee the efficiency, CNF may explode if we have hundreds of equations with hundreds of 
-XOR terms.
-
-(b) Imitate clause learning technique in CDCL, create new clauses to satisfy the input format of \emph{CryptoMiniSAT}. Advantage: \emph{CryptoMiniSAT} is a tool with performance enhancing
-on XOR clauses, and since the number of XOR clauses will not explode (I will explain later), maybe the efficiency can be guaranteed (somehow); Disadvantage: need to build our own tool for experiments.
-
-Let me have a tiny statement on the second choice. First is the approach, maybe better illustrated by an example:
-\begin{Example}
-Change boolean equation $a\oplus b\oplus c \oplus bc = 0$ to XOR-CNF. First, a SAT-solver always want whole CNF to be TRUE. Fortunately, we can negate a XOR clause by just negate one literal:
-$\overline{a}\oplus b\oplus c \oplus bc = 1$. Then we need to solve the non-single-literal term. Replace term $bc$ with new literal $d$, and create a new clause to make $d=bc$, which is
-$(\overline{b}\lor\overline{c}\lor d)\land(b\lor\overline{d})\land(c\lor\overline{d})$, adding 3 short clauses.
-\end{Example}
-You may ask why the number and size of clauses won't explode, the answer is easy: we only have 2-literal terms in worst case! Note all equations come from 
-$$\beta^3 + c_0\beta + c_1\beta + \cdots + c_{n-1}\beta^{2^{n-1}} = 0$$
-Even exponentials will never generate higher order coefficients in $\mathbb{F}_{2^n}$, so the only term generates higher-order coefficients is
-$$\beta^3 = (a_0+a_1\alpha + \cdots a_{n-1}\alpha^{n-1})(a_0+a_1\alpha^2+\cdots + a_{n-1}\alpha^{2(n-1)})$$
-Apparently cross-product terms here have at most 2 literals (degree 2).
-
-Acknowledging this fact, for $n$ unknowns there will be at most $\dbinom{n}{2} = O(n^2)$ 2-literal terms, so the number of newly created clause will also be $O(n^2)$, with fixed tiny size.
-
-By either approach, we finally get one legal assignment for bit vector $(a_0,a_1,\dots,c_{n-1})$ from the SAT-solver, by that we can compute $\beta$.
-We can directly use it as ONB, or setup a loop to keep squaring
-$\beta$ to find the minimum $k$ for $\alpha^k$, and adopt that one as ONB.
-
-\textit{Discussion}\ \ The tool flow is feasible enough, so only issue is the soundness from the ONB theory to linear system of boolean equations. I asked following questions to Florian's student:
-1.
-Solutions for linear system of equations in $\mathbb{F}_2$ (or say $\mathbb{Z}_2$).
-For example, in last page of ONB.pdf194 KB, there are 5 variables and 5 equations, provide 5 conjugate solutions forming a set of (optimal) normal basis.
-Is it possible to dig into this, say to prove all solutions are conjugates $(x, x^2, x^4, x^8, \dots)$?
-2.
-You chose $\beta^3$ to deduce those equations.
-Does solution to $\beta^3$ necessarily cover (or equal to) solutions for given lambda-Matrix M0?
-In other words, if I choose $\beta^5$ or $\beta^6$ to construct those equations, can I get the same answer?
-
-If above concerns are resolved, we can come up with a complete proof.
-\balancecolumns
-% That's all folks!
-\end{document}
+The special polynomial for type-I and type-II optimal normal basis can be computed from algorithms showed in page 108, IEEE standard 1363-2000 document.

### 2014-04-25 sequential/latex/ICCAD2014/experiment_fmcad.tex
upload files from laptop linux

\section{Experiment Results}
Our experiment on different size of SMPO designs is performed with Singular symbolic algebra computation system.
The SMPO designs are given as gate-level netlists with registers, then translated to polynomials to compose
elimination ideal in Singular for Gr\"obner basis calculation. The experiment is conducted on desktop with
3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and running 64-bit Linux OS.

\subsection{Computing Gr\"obner basis in Singular}
The Singular tool can read in scripts written in its own format similar to ANSI-C. For SMPO experiment, the main 
loop of our script file performs the same function as algorithm \ref{alg:modified} describes, while Gr\"obner basis
computation in main loop can be divided into 4 different function parts:

(i) Pre-process:

This step is executed only once before main loop starts. The function of pre-process is to compute following system
of equations for bit-level inputs $a_0 \sim a_{k-1}$:
\begin{displaymath}
\left\{
  \begin{array}{ll}
  a_0 & = f_0(A)\\
  a_1 & = f_1(A)\\
  \vdots & \ \\
  a_{k-1} & = f_{k-1}(A)
  \end{array} \right.
\end{displaymath}
The methodology has been discussed in section \ref{sec:blvs}. For 5-bit SMPO example, we start from word-level
expression polynomial
\begin{displaymath}
A + a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}
\end{displaymath}
and the result is
\begin{displaymath}
\left\{
  \begin{array}{lcl}
  a_0 & = & (\alpha+1)A^{16}+(\alpha^4+\alpha^3+\alpha)A^8+(\alpha^3+\alpha^2)A^4\\&&+(\alpha^4+1)A^2+(\alpha^2+1)A\\
  a_1 & = & (\alpha^2+1)A^{16}+(\alpha+1)A^8+(\alpha^4+\alpha^3+\alpha)A^4\\&&+(\alpha^3+\alpha^2)A^2+(\alpha^4+1)A\\
  a_2 & = & (\alpha^4+1)A^{16}+(\alpha^2+1)A^8+(\alpha+1)A^4\\&&+(\alpha^4+\alpha^3+\alpha)A^2+(\alpha^3+\alpha^2)A\\
  a_3 & = & (\alpha^3+\alpha^2)A^{16}+(\alpha^4+1)A^8+(\alpha^2+1)A^4\\&&+(\alpha+1)A^2+(\alpha^4+\alpha^3+\alpha)A\\
  a_4 & = & (\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4\\&&+(\alpha^2+1)A^2+(\alpha+1)A
  \end{array} \right.
\end{displaymath}
By replacing bit-level variable $a_i$ with $b_i, r_i$ or $R_i$, and word-level variable $A$ with $B, r, R$ respectively,
we can directly get bit-word relation functions for another operand input, pseudo input and pseudo output.

One limitation to Singular tool is the exponential cannot exceed $2^{63}$, so when doing experiments for SMPO larger than
62 bits, we use a little trick (the feasibility of this trick can also be verified in following steps). Since the BLVS method
only requires squaring of equations each time, the exponential of word $A$ can only be in the form $2^{i-1}$, i.e. 
$A^{2^0},A^{2^1},\dots,A^{2^{k-1}}$. To minimize the exponential presenting in Singular tool, we rewrite $2^{i-1}$ to $i$, i.e.
$(A^{2^0},A^{2^1},\dots,A^{2^{k-1}}) \to (A, A^2, \dots, A^k)$. In this way result is rewritten to be
$$a_0 = (\alpha+1)A^5+(\alpha^4+\alpha^3+\alpha)A^4+(\alpha^3+\alpha^2)A^3+(\alpha^4+1)A^2+(\alpha^2+1)A$$
Thus the exponential will not exceed the Singular data size limit.

This step requires limited substitution operations in Singular, so although we use the naive Gaussian elimination
method (whose time complexity is $O(k^3)$), the time cost is trivial comparing to following steps. For 33 bits experiment,
pre-process execution time is 2.7 sec; while for 100 bits experiment time cost is 36 sec.

(ii) Spoly reduction:

First, Spoly is calculated based on RATO, then reduced with the ideal composed by circuit description polynomials ($J$).
For already finished experiments, naive reduction (multi-division) is adopted, and this step takes largest portion of total 
time consumption.

For SMPO experiments, reduced Spoly has following generic form (all coefficients are omitted):
$$redSpoly = \sum r_i + \sum a_ib_i + \sum a_iB + \sum b_iA + R + r$$
Note there is no cross-term for bit-level or word-level variables from same side such as $a_ia_j, a_iA$, etc. Consider the necessary
condition of our trick, this property of reduced Spoly guarantees the word level variable can only exist in the form $A^{2^{i-1}}$,
after substituting bit-level variables with corresponding word-level variable.

(iii) Substitute bit-level variables in reduced Spoly:

Use the result from pre-process, get rid of $r_i, a_i$ and $b_i$ through substitution. This step yields following polynomial (consider the trick we used):
$$R + \sum r^i + \sum A^iB^j$$
all coefficients omitted.

(iv) Substitute present state word-level variable $r$ with inputs $A$ and $B$:

According to section \ref{sec:SMPOexperiment}, there is still a polynomial $r_{in}$ in the ideal we want to compute Gr\"obner basis.
This polynomial has form $r+f'(A,B)$, which is last clock cycle's output ($R+f'(A,B)$) with only leading term replaced in step "$from^i\gets to^i$" in 
algorithm \ref{alg:modified}. Basically this step has nothing different from last one, however, it must be taken good care of when using our
trick. There is power of $r$, $r^m$ is originally $r^{2^{m-1}}$; so if $r+f'(A,B)$ contains terms $A^iB^j$, the correct result after doing power
is $$(A^{2^{i-1}}B^{2^{j-1}})^{2^{m-1}} = A^{2^{((i+m-2)\bmod k)+1}}B^{2^{((j+m-2)\bmod k)+1}}$$
So the correct exponential for $A$ and $B$ in $(A^iB^j)^m$ should be $((i+m-2)\bmod k)+1$ and $((j+m-2)\bmod k)+1$, respectively.

Within one main loop, after finishing steps (ii) to (iv), the output should be intermediate multiplication result $R+f(A,B)$. After $k$ loops,
the output is $R+A\cdot B$ when SMPO is bug-free.

\subsection{Discussion}

I have concerns on this SMPO experiment because we use abstraction polynomial $R+f(A,B)$ instead of eliminating to univariate polynomial $g(R)$.
The only point of inconsistency to the FSM traversal we proposed is, currently the ideal quotient theory's correctness is only guaranteed under
univariate assumption. I think we have some options here: 1, expand the ideal quotient theory to multivariate; 2, argue that we use abstraction
only to make it easier to check the correctness of our experiment and let the output make sense to people; 3, abandon SMPO experiment in this paper
focusing on FSM traversal, try to do some new experiments on reachability checking, leave this SMPO experiment for some new paper concerning
functional verification.

### 2014-04-25 sequential/latex/ICCAD2014/intro_fmcad.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/DAC2013/mainframe.tex
@@
-% THIS IS SIGPROC-SP.TEX - VERSION 3.1
-% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
-% APRIL 2009
-%
-% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
-% LaTeX2e document class file for Conference Proceedings submissions.
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) Page numbering
-% ---------------------------------------------------------------------------------------------------------------
-% It is an example which *does* use the .bib file (from which the .bbl file
-% is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission,
-% you need to 'insert'  your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% Questions regarding SIGS should be sent to
-% Adrienne Griscti ---> griscti@acm.org
-%
-% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
-% Gerald Murray ---> murray@hq.acm.org
-%
-% For tracking purposes - this is V3.1SP - APRIL 2009
-
-\documentclass{acm_proc_article-sp}
-
-\begin{document}
-
-\title{Verification of Sequential Galois Field Circuits using Algebraic Geometry
-\titlenote{First version, only contains abstract and introduction part.}}
-%\subtitle{[Extended Abstract]
-%\titlenote{A full version of this paper is available as
-%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-%\LaTeX$2_\epsilon$\ and BibTeX} at
-%\texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Xiaojun Sun\\
-       \affaddr{University of Utah}\\
-       \affaddr{Department of Electrical \& Computer Engineering}\\
-       \affaddr{Salt Lake City, USA}\\
-       \email{xiaojun.sun@utah.edu}
-% 2nd. author
-\alignauthor
-Priyank Kalla\\
-       \affaddr{University of Utah}\\
-       \affaddr{Department of Electrical \& Computer Engineering}\\
-       \affaddr{Salt Lake City, USA}\\
-       \email{kalla@ece.utah.edu}
-% 3rd. author
-%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-%one who did all the really hard work.}\\
-%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-%       \affaddr{Hekla, Iceland}\\
-%       \email{larst@affiliation.org}
-%\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-%\alignauthor Lawrence P. Leipuner\\
-%       \affaddr{Brookhaven Laboratories}\\
-%       \affaddr{Brookhaven National Lab}\\
-%       \affaddr{P.O. Box 5000}\\
-%       \email{lleipuner@researchlabs.org}
-% 5th. author
-%\alignauthor Sean Fogarty\\
-%       \affaddr{NASA Ames Research Center}\\
-%       \affaddr{Moffett Field}\\
-%       \affaddr{California 94035}\\
-%       \email{fogartys@amesres.org}
-% 6th. author
-%\alignauthor Charles Palmer\\
-%       \affaddr{Palmer Research Laboratories}\\
-%      \affaddr{8600 Datapoint Drive}\\
-%%      \affaddr{San Antonio, Texas 78229}\\
-%       \email{cpalmer@prl.com}
-}
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{07 Nov 2013}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-\begin{abstract}
-Circuits working in Galois fields are increasingly employed in designs like Ellptic Curve Cryptography (ECC). 
-This work proposes a new method to effectively verify sequential circuits in Galois fields. Algebraic geometry
-is intriduced to describe circuits behavior and redefine implicit state space traversal. Moreover, Gr\"obner basis
-representation is adopted for word-level abstraction on circuit variables to address state explosion problem
-with BDDs. Experiments are run with F4-style reduction engine to get more competitive results.
-\end{abstract}
-
-% A category with the (minimum) three required fields
-\category{EDA5.1}{Verification}{Functional, transaction-level, RTL, and gate-level modeling and verification of hardware design}
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-\terms{Verification, Algorithms}
-
-\keywords{Verification, Sequential, Galois Fields, Algebraic Geometry} % NOT required for Proceedings
-
 \section{Introduction}
@@
 J. Lv, el (should I cite this part) gave out a F4-style reduction algorithm which can help speed up Gr\"obner basis calculation...
-
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{seq_verif}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-\balancecolumns
-% That's all folks!
-\end{document}

### 2014-04-25 sequential/latex/ICCAD2014/theory.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/DAC2013/theory.tex
@@
-% THIS IS SIGPROC-SP.TEX - VERSION 3.1
-% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
-% APRIL 2009
-%
-% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
-% LaTeX2e document class file for Conference Proceedings submissions.
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
-%       1) The Permission Statement
-%       2) The Conference (location) Info information
-%       3) The Copyright Line with ACM data
-%       4) Page numbering
-% ---------------------------------------------------------------------------------------------------------------
-% It is an example which *does* use the .bib file (from which the .bbl file
-% is produced).
-% REMEMBER HOWEVER: After having produced the .bbl file,
-% and prior to final submission,
-% you need to 'insert'  your .bbl file into your source .tex file so as to provide
-% ONE 'self-contained' source file.
-%
-% Questions regarding SIGS should be sent to
-% Adrienne Griscti ---> griscti@acm.org
-%
-% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
-% Gerald Murray ---> murray@hq.acm.org
-%
-% For tracking purposes - this is V3.1SP - APRIL 2009
-
-\documentclass{acm_proc_article-sp}
-
-%packages
-\usepackage[linesnumbered, ruled]{algorithm2e}
-\RequirePackage{amssymb, mathptm}
-\usepackage{amsbsy}
-\usepackage{graphicx}
-\usepackage{helvet}
-\usepackage{enumerate}
-\usepackage{amsmath}
-\usepackage{cases}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{subfig}
-\usepackage{comment}
-
-%new format
-\newtheorem{Algorithm}{Algorithm}
-\newdef{Definition}{Definition}
-\newtheorem{Theorem}{Theorem}
-\newtheorem{Lemma}{Lemma}
-\newdef{Example}{Example}
-
-\begin{document}
-
-\title{Verification of Sequential Galois Field Circuits using Algebraic Geometry
-\titlenote{Version 0.3, refined theory part with normal basis, proposed details for designing 2 examples.}}
-%\subtitle{[Extended Abstract]
-%\titlenote{A full version of this paper is available as
-%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-%\LaTeX$2_\epsilon$\ and BibTeX} at
-%\texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement
-% and alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time'
-% i.e. three 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of
-% "name/affiliations" may appear. We just ask that you restrict
-% the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate'
-% we ask you to refrain from putting more than six authors
-% (two rows with three columns) beneath the article title.
-% More than six makes the first-page appear very cluttered indeed.
-%
-% Use the \alignauthor commands to handle the names
-% and affiliations for an 'aesthetic maximum' of six authors.
-% Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the
-% \additionalauthors command.
-% These 'additional authors' will be output/set for you
-% without further effort on your part as the last section in
-% the body of your article BEFORE References or any Appendices.
-
-\numberofauthors{2} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting
-% reasons) and the remaining two appear in the \additionalauthors section.
-%
-\author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
-\alignauthor
-Xiaojun Sun\\
-       \affaddr{University of Utah}\\
-       \affaddr{Department of Electrical \& Computer Engineering}\\
-       \affaddr{Salt Lake City, USA}\\
-       \email{xiaojun.sun@utah.edu}
-% 2nd. author
-\alignauthor
-Priyank Kalla\\
-       \affaddr{University of Utah}\\
-       \affaddr{Department of Electrical \& Computer Engineering}\\
-       \affaddr{Salt Lake City, USA}\\
-       \email{kalla@ece.utah.edu}
-% 3rd. author
-%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
-%one who did all the really hard work.}\\
-%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
-%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
-%       \affaddr{Hekla, Iceland}\\
-%       \email{larst@affiliation.org}
-%\and  % use '\and' if you need 'another row' of author names
-% 4th. author
-%\alignauthor Lawrence P. Leipuner\\
-%       \affaddr{Brookhaven Laboratories}\\
-%       \affaddr{Brookhaven National Lab}\\
-%       \affaddr{P.O. Box 5000}\\
-%       \email{lleipuner@researchlabs.org}
-% 5th. author
-%\alignauthor Sean Fogarty\\
-%       \affaddr{NASA Ames Research Center}\\
-%       \affaddr{Moffett Field}\\
-%       \affaddr{California 94035}\\
-%       \email{fogartys@amesres.org}
-% 6th. author
-%\alignauthor Charles Palmer\\
-%       \affaddr{Palmer Research Laboratories}\\
-%      \affaddr{8600 Datapoint Drive}\\
-%%      \affaddr{San Antonio, Texas 78229}\\
-%       \email{cpalmer@prl.com}
-}
-% There's nothing stopping you putting the seventh, eighth, etc.
-% author on the opening page (as the 'third row') but we ask,
-% for aesthetic reasons that you place these 'additional authors'
-% in the \additional authors block, viz.
-%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
-%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
-%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
-\date{21 Jan 2014}
-% Just remember to make sure that the TOTAL number of authors
-% is the number that will appear on the first page PLUS the
-% number that will appear in the \additionalauthors section.
-
-\maketitle
-\begin{abstract}
-Circuits working in Galois fields are increasingly employed in designs like arithmetic component for Elliptic Curve Cryptography (ECC). 
-This work proposes a new method to effectively verify sequential circuits in Galois fields. Algebraic geometry
-is introduced to describe circuits behavior and expand the definition of implicit state space traversal. Moreover, Gr\"obner basis
-representation is adopted for word-level abstraction on circuit variables to address state explosion problem
-with BDDs. Experiments are run with fast Gr\"obner basis computation engine to get competitive results.
-\end{abstract}
-
-% A category with the (minimum) three required fields
-\category{EDA5.1}{Verification}{Functional, transaction-level, RTL, and gate-level modeling and verification of hardware design}
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-\terms{Verification, Algorithms}
-
-\keywords{Verification, Sequential, Galois Fields, Algebraic Geometry} % NOT required for Proceedings
-
 \section{Essential Theory}
@@
 
-\begin{figure}
-\centering
-\psfig{file=./sequential_fig.ps, height=2in, width=3.5in,}
+\begin{figure}[hbt]
+\begin{center}
+\includegraphics[width=3.5in]{./sequential_fig.eps}
+\end{center}
 \caption{A typical sequential circuit block view}
@@
 \centering
-\caption{Bit-vector, Exponential and Polynomial representation of
-elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
-\pmod{x^4+x^3+1}$}\label{table:booltogalois}  
 \begin{tabular}{|c|c||c|c|} 
@@
 \end{tabular}
+\caption{Bit-vector, Exponential and Polynomial representation of
+elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
+\pmod{x^4+x^3+1}$}\label{table:booltogalois}  
 \end{table}
@@
 
-\begin{figure}
-\centering
-\psfig{file=./fsm_fig.ps, height=2in, width=3.5in,}
+\begin{figure}[hbt]
+\begin{center}
+\includegraphics[scale=0.3]{./stg_fig.ps}
+\end{center}
+\caption{State transition graph of sample FSM}
+\label{fig:stg}
+\end{figure}
+
+\begin{figure}[hbt]
+\begin{center}
+\includegraphics[width=3.5in]{./fsm_fig.eps}
+\end{center}
 \caption{Gate-level circuit of combinational logic block of sample FSM}
 \label{fig:fsm}
-\end{figure}
-
-\begin{figure}
-\centering
-\psfig{file=./stg_fig.ps, height=2in, width=3.5in,}
-\caption{State transition graph of sample FSM}
-\label{fig:stg}
 \end{figure}
@@
 \end{Theorem}
-For example in Line 1 of BFS algorithm, assume initial state is $S_3$ in figure \ref{fig:stg}, then corresponding
-polynomial $f = \mathcal{F}(S^0) = S^0 - 1 - \alpha$. Consider an ideal $I$ with only one generator $f$, its variety
-$V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma - 1 - \alpha = 0\}$, which equals to $\{1 + \alpha\}$, the only
+For example in Line 1 of BFS algorithm, assume initial state is $S_0$ in figure \ref{fig:stg}, then corresponding
+polynomial $f = \mathcal{F}(S^0) = S^0 - 0$. Consider an ideal $I$ with only one generator $f$, its variety
+$V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 0\}$, which equals to $\{0\}$, the only
 valid value $S^0$ can take.
@@
 f_1:& t_0 - (\overline{x}\ and\ \overline{s_0}\ and\ \overline{s_1})\ or\ (s_0\ and\ s_1) \\
-f_2:& t_1 - (\overline{s_0}\ and\ x)\ or\ (p\ and\ \overline{s_1})\
+f_2:& t_1 - (\overline{s_0}\ and\ x)\ or\ (s_0\ and\ \overline{s_1})\
 \end{array}
@@
 \begin{displaymath}
-f_5:\  S - 1 - \alpha
+f_5:\  S
 \end{displaymath}
@@
 \centering
-\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
-\label{table:booltogalois_op}
 \begin{tabular}{|c|c|} \hline
@@
 \hline\end{tabular}
+\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
+\label{table:booltogalois_op}
 \end{table}
-Using elimination term order: $others\ >\ bit$-$level\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
+Using elimination term order: $intermediate\ bit$-$level\ signals\ >\ bit$-$level\ primary\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
 $J = \langle f_1, f_2, \dots, f_{12}\rangle $, the result will include one polynomial $f_T$ contains only variable $T$. In this example,
-$f_T = T + 1$. This polynomial equals to $T - 1$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
-limited within $\mathbb{F}_{2}$, where $1 \equiv -1\ (mod\ 2)$. Solution to $f_T = 0$ is $T = 1$, which shows that next
-state the machine just reached is $\{S_1(01)\}$.
+$f_T = T^2+(\alpha+1)T+\alpha$. This polynomial equals to $T^2-(\alpha+1)T+\alpha$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
+limited within $\mathbb{F}_{2}$, where $x \equiv -x\ (mod\ 2)$ for any element $x$ in the field. Solution to $f_T = 0$ is $T = 1\ or \ T = \alpha$, which shows that next
+state the machine just reached is $\{S_1(01), S_2(10)\}$.
+
+Continue to run this traversal algorithm, it will terminates with final reachable states $f_T = T^3+(\alpha+1)T^2+\alpha T$, its solution is $T=0\ or\ T=1\ or\ T=\alpha$,
+which shows within total 4 states, state $S_3$ is unreachable from initial state $S_0$.
 
@@
 Example: assume $I = \langle f\rangle  = \langle T^2 + (1+\alpha)\cdot T+\alpha\rangle $, its variety $V(I) = \{a\ |\ a \in \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$.
+
+\begin{figure*}[tb]
+\begin{center}
+\includegraphics[width=\textwidth]{./mySMPO.eps}
+\end{center}
+\caption{5-bit SMPO}
+\label{fig:SMPO}
+\end{figure*}
+
 So it is reasonable to specify: union, intersection and complement set mentioned in this paper are all functions on \textbf{varieties}.
@@
 For varieties intersection $\{1\}\bigcap\{1, \alpha\}$, calculate ideal sum $\langle T+1, T^2 + (1+\alpha)\cdot T+\alpha\rangle  = \langle T+1\rangle $,
-its variety is $\{1\}$; for varieties union $\{1,1+\alpha\}\bigcup\{1+\alpha\}$, calculate
+its variety is $\{1\}$; for varieties union $\{1,\alpha\}\bigcup\{1+\alpha\}$, calculate
 ideal product $\langle (T+1+\alpha)\cdot(T^2 + (1+\alpha)\cdot T+\alpha)\rangle  = \langle T^3 + 1\rangle $, its variety
@@
 If the initial input is $S_3$: $S + 1 + \alpha$, the final return value will be set of
-reachable states: $T^4 + T$, i.e. the universal set, $\{S_0, S_1, S_2, S_3\}$.
+reachable states: $T^4 + T$, i.e. the universal set, $\{S_0, S_1, S_2, S_3\}$. Modified algorithm could be
+written in following interpretation:
 
-\subsection{Discussion on Ideal quotient and its generator}
-In nature, we are looking varieties as the solutions to specific polynomial $f=0$ WITHIN $\mathbb{F}_{2^n}$, because (affine) varieties are defined
-within algebraic closure $k^n$ but we only care about those fall in $\mathbb{F}_{2^n}$.
+\begin{algorithm}[hbt]
+\SetAlgoNoLine
+ \KwIn{Input-output circuit characteristic polynomial ideal $I_{ckt}$, initial state polynomial $\F(S)$}
 
-Theorem \ref{thm:unionintersect} related to DEF. \ref{def:sum} and \ref{def:prod} counts on ideal generators, so there will be no ambiguity on 
-new ideals we get from this theorem. It also appeared in Avrunin's CAV paper. The only problem is what ideal quotient operation will give us.
-This should be further discussed because the completeness of our algebraic geometry approach will benefit from it. \emph{Singular} has the function of
-ideal quotient, so ask from the author maybe another choice.
+  $from^0 = reached = \F(S)$\;
+  \Repeat{$new^i == 1$}
+  {
+  	$i \gets i + 1$\;
+	$to^i \gets$GB w/ elimination term order$<I_{ckt}, from^{i-1}>$\;
+	$new^i \gets $generator of $<to^i> + (<T^4-T>:<reached>)$\;
+  	$reached \gets $generator of $<reached> \cdot <new^i>$\;
+	$from^i \gets new^i(S\setminus T)$\;
+  }
+\Return{$reached$}
+\caption {Algebraic Geometry based Traversal Algorithm}\label{alg:modified}
+\end{algorithm}
 
-If we put this conception aside, it is also feasible to define complement set from "solution to 
-equation" sense. Assume $f(x) = x + 1 + \alpha = 0$, its solution within $\mathbb{F}_{2^2}$ 
-is $\{1+\alpha\}$; meanwhile solutions to vanishing polynomial $v(x) = x^2 + x = 0$ are
-$\{0,1,\alpha,1+\alpha\}$. Now compute solutions to $$g(x) = \frac{v(x)}{f(x)} = 0$$,
-this requires $v(x) = 0$ and $f(x) \neq 0$, which means the complement set of solution 
-to $f(x) = 0$.
-
-Issues to this interpretation: (a) Is $v(x)$ always divisible by $f(x)$? (b) Is $v(x)=0$ and
-$f(x) \neq 0$ equivalent with $g(x)=0$ on their solutions? Could $f(x)$ equal to $0$?
+Here $from^i, to^i, new^i$ are all univariate polynomial about $S$ or $T$
 
@@
 output should be $R = A\cdot B$ (mod $x^5 + x^2 + 1$).
-\begin{figure}
-\centering
-\epsfig{file=./mySMPO.eps, height=2in, width=3.5in}
-\caption{5-bit SMPO}
-\label{fig:SMPO}
-\end{figure}
+
 Similarly, build elimination ideal for all gates/operations and induce
@@
 and keeps the polynomial contains function between output and inputs.
-Here the lex ordering is $\{others\} > R > \{A, B\}$, and objective
+Here the lex ordering is $\{all\ bit$-$level\ variables\} > R > \{A, B\}$, and objective
 polynomial is $R + \mathcal{F}(A, B)$.
@@
 \subsection{Gr\"obner basis based Approach}
+\label{sec:SMPOexperiment}
 
@@
 &Spoly(f_w,f_g) \xrightarrow{J+J_0}_{+}\nonumber\\
-&r_1+(\alpha)r_2+(\alpha^4+\alpha^2)r_3+(\alpha^3+\alpha^2)r_4+(\alpha^3)b_1a_1+(\alpha^4+\alpha^2)b_1a_2\nonumber\\
-+&(\alpha^3+\alpha+1)b_1a_3+(\alpha^3+\alpha)b_1a_4+(\alpha+1)b_1A+(\alpha^4+\alpha^2+\alpha)b_2a_1\nonumber\\
-+&(\alpha^4+\alpha^3+\alpha^2+\alpha)b_2a_4+(\alpha^3+\alpha^2+1)b_3a_1+(\alpha)b_3a_3\nonumber\\
-+&(\alpha^2+\alpha+1)b_4a_1+(\alpha+1)b_4a_2+(\alpha^4+\alpha^2)b_4a_3+(\alpha^4+\alpha^3+\alpha+1)b_4a_4\nonumber\\
-+&(\alpha^3+1)b_4A+(\alpha^4+\alpha^3+\alpha^2+1)a_1B+(\alpha^4+\alpha^3+\alpha^2+1)R\nonumber
+&r_1+(\alpha)r_2+(\alpha^4+\alpha^2)r_3+(\alpha^3+\alpha^2)r_4+(\alpha^3)b_1a_1\nonumber\\
++&(\alpha^4+\alpha^2)b_1a_2+(\alpha^3+\alpha+1)b_1a_3+(\alpha^3+\alpha)b_1a_4+(\alpha+1)b_1A\nonumber\\
++&(\alpha^4+\alpha^2+\alpha)b_2a_1+(\alpha^4+\alpha^3+\alpha^2+\alpha)b_2a_4+(\alpha^3+\alpha^2+1)b_3a_1\nonumber\\
++&(\alpha)b_3a_3+(\alpha^2+\alpha+1)b_4a_1+(\alpha+1)b_4a_2+(\alpha^4+\alpha^2)b_4a_3\nonumber\\
++&(\alpha^4+\alpha^3+\alpha+1)b_4a_4+(\alpha^3+1)b_4A+(\alpha^4+\alpha^3+\alpha^2+1)a_1B\nonumber\\
++&(\alpha^4+\alpha^3+\alpha^2+1)R\nonumber
 \end{align}
@@
 \subsection{Bit-level Variable Substitution (BLVS)}
+\label{sec:blvs}
 The remainder from \emph{Spoly} contains some bit-level variables, and our objective is to get a polynomial contains only word-level variables
@@
 %
-\appendix
-\section{Normal Basis Multiplication}
-\subsection{$\lambda$-Matrix}
-$\lambda$-Matrix is defined with cross-product terms from multiplication. That is 
-\begin{displaymath}
-Product\ C = (\sum_{i=0}^{n-1}a_i\beta^{2^i})(\sum_{j=0}^{n-1}b_j\beta^{2^j}) = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}a_ib_j\beta^{2^i}\beta^{2^j}
-\end{displaymath}
-The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
-normal basis, i.e.
-\begin{displaymath}
-\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1}\lambda_{ij}^{(k)}\beta^{2^k}, \ \ \lambda_{ij}^{(k)} \in F_2.
-\end{displaymath}
-Substitution yields, get the expression for $k$-th digit of product:
-\begin{displaymath}
-c_k = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}\lambda_{ij}^{(k)}a_ib_j
-\end{displaymath}
-$\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ in $k$-th $\lambda$-Matrix.
-
-\begin{Example}
-$\lambda$-Matrix: A binary $n\times n$ matrix $M$ could be employed to describe the "function"
- mentioned in Ex.\ref{ex:nb_multi}: $c_{n-1} = f(A, B) = A \cdot M \cdot B^T$, $B^T$ denotes vector transposition. 
-More specifically, denote the matrix by \emph{$k$-th $\lambda$-Matrix}: $c_k = A \cdot M^{(k)} \cdot B^T$.
-Then $c_{k-1} = A \cdot M^{(k-1)} \cdot B^T = rotate(A) \cdot M^{(k)} \cdot rotate(B)^T$, which means $M^{(k)}$ is generated
-by right and down cyclic shifting $M^{(k-1)}$.
-
-In $\mathbb{F}_{2^3}$ constructed by $\alpha^3 + \alpha + 1$, let $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
-is a normal basis. $0$-th $\lambda$-Matrix
-\begin{displaymath}
-M^{(0)} = \left(
-\begin{array} {lcr}
-0 & 1 & 0\\
-1 & 0 & 1\\
-0 & 1 & 1
-\end{array} \right).
-\end{displaymath}
-i.e.,
-\begin{displaymath}
-c_0 = (a_0\  a_1\  a_2)\left(
-\begin{array} {lcr}
-0 & 1 & 0\\
-1 & 0 & 1\\
-0 & 1 & 1
-\end{array} \right)\left(
-\begin{array} {lcr}
-b_0\\
-b_1\\
-b_2
-\end{array} \right).
-\end{displaymath}
-\end{Example}
-\subsection{Optimal Normal Basis}
-\begin{Definition}
-The number of non-zero entries in $\lambda$-Matrix is known as  \emph{Complexity$(C_N)$}.
-\end{Definition}
-\begin{Theorem}
-If N is a normal basis for $\mathbb{F}_{p^n}$ with $\lambda$-matrix $M^{(k)}$, then non-zero entries in 
-matrix $C_N\geq 2n-1$.
-\end{Theorem}
-Proof omitted.
-\begin{Definition}
-If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as
-\emph{Optimal Normal Basis (ONB)}.
-\end{Definition}
-There are 2 types of optimal normal basis existing.
-Rules for creating Type-I ONB over $\mathbb{F}_{2^n}$ are:
-\begin{itemize}
-\item n+1 must be prime.
-\item 2 must be primitive in $\mathbb{Z}_{n+1}$.
-\end{itemize}
-Rules for creating Type-II ONB over $\mathbb{F}_{2^n}$ are:
-\begin{itemize}
-\item 2n+1 must be prime. And either
-\item 2 must be primitive in $\mathbb{Z}_{2n+1}$, OR
-\item $2n+1 \equiv 3 \ \ mod\ \  4$ AND 2 generates the quadratic residues in $\mathbb{Z}_{2n+1}$
-\end{itemize}
-There are corresponding criteria for simply creating $\lambda$-Matrix for either type of ONB:
-\begin{Lemma}
-Type-I ONB implies a $\lambda$-Matrix that each nonzero entry $M_{i,j}$ satisfies
-\begin{numcases}{}
-2^i + 2^j = 1 \bmod  n+1\notag\\
-2^i + 2^j = 0 \bmod n+1\notag
-\end{numcases}
-Type-II ONB implies a $\lambda$-Matrix that each nonzero entry $M_{i,j}$ satisfies
-\begin{numcases}{}
-2^i + 2^j = 1\bmod\  2n+1\notag\\
-2^i + 2^j = -1 \bmod  2n+1\notag\\
-2^i - 2^j = 1 \bmod  2n+1\notag\\
-2^i - 2^j = -1 \bmod 2n+1\notag
-\end{numcases}
-\end{Lemma}
-Proof omitted. (Issues here: We knew type-I \& II definition can deduce corresponding $\lambda$-Matrix, how about
-reverse implication? i.e. can we prove the equivalence?)
-
-\subsection{Design a Normal Basis Multiplier using $\lambda$-Matrix}
-Imitating the structure of SMPO, just specify the gates' connections for the first clock cycle, then following cycles are
-automatically completed by cyclic shifting operands $A$ and $B$. The whole design procedure is based on a $n \times n$ $k$-th $\lambda$-Matrix.
-
-First figure out the first row (\emph{row} 0) in $\lambda$-Matrix, for any nonzero entry $M_{0,j}$ (there will be only 1 if taking $0$-th $\lambda$-Matrix), place an AND gate
-$a_j\land b_0$. Connect different $a_j$ with a XOR gate before place the AND gate if there are 2 or more nonzero entries in this row;
-
-Secondly, repeat this for each row. Note for nonzero entry $M_{i,j}$, corresponding index of $a_u$ and $b_v$ is incremented by $i$, i.e. $u = j + i \pmod n$, $v = i + i \pmod n$;
-
-Finally, connect the output of AND gate for row $i$ to one input of a XOR gate, then connect the output of XOR gate to a register/flip-flop. The output of register/flip-flop is connected
-to the other input of XOR gate, but at next row. All these gates and connections for row $i$ is called unit $R_i$. (Fig.\ref{fig:SMPO})
-
-\section{Find Optimal Normal Basis}
-\subsection{Transforming to SAT}
-From last section we learned it is easy to compute $\lambda$-Matrix when given certain type of optimal normal basis. However in our approach, the mapping between 
-optimal normal basis and standard basis should be explicitly specified. In other words, let \emph{Optimal Normal Element} $\beta = \alpha^k$, the exponential $k$ on
-primitive element $\alpha$ needs to be calculated. Otherwise word-level variable interpreting polynomial $a_0\beta + a_1\beta^2 + \cdots + a_{n-1}\beta^{2^{n-1}}$
-will be an unknown polynomial and GB computation is impossible.
-
-Currently finding the optimal normal element $\beta$ in $2^n$ (over field $\mathbb{F}_{2^n}$) elements is NP-hard. However, it is possible to improve the performance 
-by transforming this problem to SAT problem which can be solved by efficient SAT solver. This is what our tool is managing to do.
-
-Given 0-th $\lambda$-Matrix $M^{(0)}$, other $\lambda$-Matrices can be obtained by cyclic right-down shifting, i.e. entry $M_{i,j}^{(k)} = M_{i-k,j-k}^{(0)}$. Note all
-index operation results should modulus $n$. $\beta^3$ is an element which can be represented as product of normal basis multiplication: $\beta^3 = \beta\cdot\beta^2$.
-Consider the following equation
-\begin{displaymath}
-c_k = (1\  0\ 0\ \cdots\ 0)\  M^{(k)} \left(
-\begin{array} {lcr}
-0\\
-1\\
-0\\
-\vdots\\
-0
-\end{array} \right)
-\end{displaymath}
-$c_k$ is the $k$-th coefficient in normal basis expression of $\beta^3$: $\beta^3 = c_0\beta + c_1\beta^2 + \cdots + c_{n-1}\beta^{2^{n-1}}$.
-Thus, $\beta^3$ could be written as
-\begin{displaymath}
-\beta^3 = \sum_{k=0}^{n-1} M_{0,1}^{(k)}\beta^{2^k} = \sum_{k=0}^{n-1} M_{-k,1-k}^{(0)}\beta^{2^k}
-\end{displaymath}
-Meanwhile, the standard basis representation of optimal normal element can be assumed as
-\begin{displaymath}
-\beta = a_0 + a_1\alpha + \cdots + a_{n-1}\alpha^{n-1}
-\end{displaymath}
-Do a substitution, vanish every coefficient. Finally we can get a linear system of $n$ boolean equations (composed by XOR and AND), with $n$ unknowns $a_0,a_1,\dots,a_{n-1}$. 
-Examples are in the write-up by Florian's student.
-
-\subsection{Solving XOR-SAT}
-Solving the linear system of boolean equations is a tricky problem, because ordinary SAT solver only support Conjunctive Normal Form (CNF) inputs; while a XOR-SAT based solver 
-\emph{CryptoMiniSAT} only support clauses composed by single literals and XOR connectives ($a\oplus b \oplus (c\land d)$ is illegal input), I call it "XOR-CNF". Basically there are 2 approaches 
-to think about our tool flow:
-
-(a) Use another engine (Commercial Symbolic Computing System such as \emph{Wolfram Mathematica}, or Circuit Based Synthesizer such as \emph{ABC}) to transform all equations to CNF clauses.
-Advantage: Convenient, even no need to write our own tool in scripts; Disadvantage: cannot guarantee the efficiency, CNF may explode if we have hundreds of equations with hundreds of 
-XOR terms.
-
-(b) Imitate clause learning technique in CDCL, create new clauses to satisfy the input format of \emph{CryptoMiniSAT}. Advantage: \emph{CryptoMiniSAT} is a tool with performance enhancing
-on XOR clauses, and since the number of XOR clauses will not explode (I will explain later), maybe the efficiency can be guaranteed (somehow); Disadvantage: need to build our own tool for experiments.
-
-Let me have a tiny statement on the second choice. First is the approach, maybe better illustrated by an example:
-\begin{Example}
-Change boolean equation $a\oplus b\oplus c \oplus bc = 0$ to XOR-CNF. First, a SAT-solver always want whole CNF to be TRUE. Fortunately, we can negate a XOR clause by just negate one literal:
-$\overline{a}\oplus b\oplus c \oplus bc = 1$. Then we need to solve the non-single-literal term. Replace term $bc$ with new literal $d$, and create a new clause to make $d=bc$, which is
-$(\overline{b}\lor\overline{c}\lor d)\land(b\lor\overline{d})\land(c\lor\overline{d})$, adding 3 short clauses.
-\end{Example}
-You may ask why the number and size of clauses won't explode, the answer is easy: we only have 2-literal terms in worst case! Note all equations come from 
-$$\beta^3 + c_0\beta + c_1\beta + \cdots + c_{n-1}\beta^{2^{n-1}} = 0$$
-Even exponentials will never generate higher order coefficients in $\mathbb{F}_{2^n}$, so the only term generates higher-order coefficients is
-$$\beta^3 = (a_0+a_1\alpha + \cdots a_{n-1}\alpha^{n-1})(a_0+a_1\alpha^2+\cdots + a_{n-1}\alpha^{2(n-1)})$$
-Apparently cross-product terms here have at most 2 literals (degree 2).
-
-Acknowledging this fact, for $n$ unknowns there will be at most $\dbinom{n}{2} = O(n^2)$ 2-literal terms, so the number of newly created clause will also be $O(n^2)$, with fixed tiny size.
-
-By either approach, we finally get one legal assignment for bit vector $(a_0,a_1,\dots,c_{n-1})$ from the SAT-solver, by that we can compute $\beta$.
-We can directly use it as ONB, or setup a loop to keep squaring
-$\beta$ to find the minimum $k$ for $\alpha^k$, and adopt that one as ONB.
-
-\textit{Discussion}\ \ The tool flow is feasible enough, so only issue is the soundness from the ONB theory to linear system of boolean equations. I asked following questions to Florian's student:
-1.
-Solutions for linear system of equations in $\mathbb{F}_2$ (or say $\mathbb{Z}_2$).
-For example, in last page of ONB.pdf194 KB, there are 5 variables and 5 equations, provide 5 conjugate solutions forming a set of (optimal) normal basis.
-Is it possible to dig into this, say to prove all solutions are conjugates $(x, x^2, x^4, x^8, \dots)$?
-2.
-You chose $\beta^3$ to deduce those equations.
-Does solution to $\beta^3$ necessarily cover (or equal to) solutions for given lambda-Matrix M0?
-In other words, if I choose $\beta^5$ or $\beta^6$ to construct those equations, can I get the same answer?
-
-If above concerns are resolved, we can come up with a complete proof.
-\balancecolumns
-% That's all folks!
-\end{document}

### 2014-04-25 sequential/latex/Word-Level RTL.tex
upload files from laptop linux

%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference,11pt]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

\usepackage{graphicx}

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Word-Level RTL Equivalence Checking Using Word-Level Constraint-Models}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Xiaojun Sun}
\IEEEauthorblockA{Electrical and Computer Engineering\\
University of Utah\\
Salt Lake City, Utah\\
Email: jason.sj.sun@gmail.com}
\and
\IEEEauthorblockN{Shaofan Qi}
\IEEEauthorblockA{Electrical and Computer Engineering\\
University of Utah\\
Salt Lake City, Utah\\
Email: chi.cedric@gmail.com}
\and
\IEEEauthorblockN{Yi Guo}
\IEEEauthorblockA{School of Computing\\
University of Utah\\
Salt Lake City, Utah\\
Email: broroguoyi@hotmail.com}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

%\begin{abstract}
%%\boldmath
%The abstract goes here.
%\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\begin{abstract}
%\boldmath
To check RTL equivalence using word-level constraints,\emph{ LPSAT} is introduced to solve the satisfiability (SAT) problem for complex RTL-level
designs, which contain word-level arithmetic operators and Boolean logics. Our project uses linear-program solvers for constraint satisfaction.
The word-level SAT solver, called LPSAT, based on linear programming. In this approach, both word-level arithmetic constraints and control logic constraints are put together into a common linear programming domain.
\end{abstract}

\section{Introduction}
% no \IEEEPARstart
The Boolean satisfiability problem (SAT) belongs to the class of NP-complete problems, with algorithmic solutions having exponential worst-case complexity. Word-Level RTL equivalence checking using linear constraint-models. Linear Programming (LP) is a special case of a constrained optimization problem in which both the objective function and constraints are linear equations. In addition, variables in LP are assumed to be non-negative.

In this project, our major target is to implement the equivalence checking with word-level RTL and linear constraint. Register-transfer level (RTL) is referred as a design abstraction which models a synchronous digital circuit in terms of the flow of digital signals (data) between hardware registers, and the logical operations performed on those signals. Word-level abstraction or constraint-models are usually used to check world-level RTL equivalence. This project aims at proving the equivalence between two given RTLs, using integer linear-programming solvers for constraint satisfaction.

Satisfiability of complex word-level formulas described at the RTL is usually solved in the Boolean domain, with Boolean solvers. However, these engines often show a poor performance for data path verification. Instead of solving the problem at the bit-level, a method is proposed to transform conjunctions of bit-vector equalities and inequalities into sets of integer linear arithmetic constraints. Integer linear constraint solvers are used as a decision procedure for bit-vector arithmetic. Linear Programming (LP) is a special case of a constrained optimization problem in which both the objective function and constraints are linear equations. In addition, variables in LP are assumed to be non-negative.

 Bit-vectors are used to represent variables and their values by RTL-HDLs.  Important arithmetic bit-vector operators are for instance: \emph {negation, addition, multiplication, extension, concatenation, and shifting}. These operators are often used to model data paths of hardware designs.  Many design automation tools operating on HDLs, including logic verification, test generation, and synthesis, have to solve the satisfiability problem for word-level formulas.  Solving the SAT problem directly on the word-level, as opposed to ﬁrst translating it into the Boolean domain, bears a high potential for optimization for those tools.  In the RTL, we should build the linear model of the operators shown in table 1.

In this project, we first find all these operators’ linear constraint model, and the way to linearize arithmetic operators and Boolean gates. Then build a program to cover the circuit design, which is RTL-level (verilog file). Then use the Linear Programming (LP) technique (AMPL, IBM CPLEX) to model and solve this SAT problem.  ILOG CPLEX is a tool for solving linear optimization problems, commonly referred to as Linear Programming (LP) problems.[1]

\begin{table}[!t]
% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Operators}
\label{work}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|c|c|c|}
\hline
Arithmetic Operators& Bitwise Operators & Logical Operators \\
\hline

$+, -, *,  \% $ &   \& , $\vert$ , \^{} , \~{} , \^{}\~{} , \~{}\^{}   & ! , \&\& , $\vert$$\vert$  \\
\hline
\hline
Relational Operators& Equality Operators & Shift Operators \\
\hline
$>, < , >= , <= $ &   $ == , != $   & $ >> , << $  \\
\hline
\hline
Conditional Operators&& \\
\hline
$?: $ &&\\
\hline
\end{tabular}
\end{table}
%\hfill mds
%\hfill January 11, 2007

\section{Linear Programming}
Linear Programming (\emph{LP} or \emph{linear optimization}) is a mathematical method for determining a way to achieve the best outcome in which both the objective function and constraints are linear equations.

Linear programs are problems that can be expressed in canonical form:

\begin{eqnarray*}
\emph{Minimize} & : & c^T X\\
\emph{Subject  to} & : & AX <= b\\
 & & X>=0\\
\end{eqnarray*}
% this part is trivial
where $X$ represents the vector of variables (to be determined), $c$ and $b$ are vectors of (known) coefficients, $A$ is a (known) matrix of coefficients, and is the matrix transpose. The expression to be maximized or minimized is called the objective function ($c^T X$ in this case). The inequalities $Ax <= b$ are the constraints which specify a convex polytopes over which the objective function is to be optimized. In this context, two vectors are comparable when they have the same dimensions. If every entry in the first is less-than or equal-to the corresponding entry in the second then we can say the first vector is less-than or equal-to the second vector.[2]

\section{Modeling}
This part is using Linear Programming (LP) technique to model the SAT problem for a complex design composed of the arithmetic and Boolean logic blocks. Basic arithmetic word-level operators include: adders, subtractors, comparators, , shifters, and multipliers.  Any other word-level operators can be represented in
terms of these operators using Boolean connectors (AND, OR, NOT).

	\subsection{ Adder/Subtractor}
	Addition and subtraction are both linear operators and can be trivially represented as equality constraints:
	\begin{eqnarray*}
	& &C=A \pm B \Rightarrow C - A  \mp B = 0
	\end{eqnarray*}

	\subsection{ Comparison operators}
	There are six types of comparison operators:$ \leq , < , \geq, > , = , \neq$. For a comparator, $A = (B == C) \emph{ or } A = (B < C) \emph{ or }  A = (B > C)$, variables
$x , y$ are introduced to indicate that $(B==C) \Leftrightarrow (x y = 00)$ , $(B>C) \Leftrightarrow (x y = 10)$ and $(B<C) \Leftrightarrow (x y = 01)$
$$\left\{
\begin{array}{lcl}
{A = 1- x -y} \\
{B - C+ L \cdot (1-x) \geq 1} \\
 {B-C+L\cdot x \leq 0}\\
 {C -B +L\cdot (1-y) \geq 1}\\
  {C -B +L\cdot y \leq 0}\\
  {x +y \leq 1}
\end{array} \right. $$
	Here $L$ is a constant and satisfying the constraint $L \geq max(B,C)$. We assume the $max(B,C)$ is used to denote a constant,  it means the maximum value of the upper bound of A or B $(\emph {in our project, we made } L = 2^n -1)$.

	\subsection{Multiplier}
	Multiplication is a non-linear operator, like $A = B \times C$. So one of $B$ or $C$ must be expanded into binary variables. The best candidate for the expansion is the one which is driven
by the Boolean part. Let \emph{B} be an n-bit variable to be expanded, and let $P_i = B_i \times C$ be a partial product, and $B_i$ has to be declared as a binary variable. And the $A = \sum_{i=0} ^{n-1} 2^i \cdot P_i$, and for each $i \in [0,1,\ldots,n-1] $,
$$\left\{ \begin{array}{l}
{P_i -L \cdot B_i \leq 0} \\
{P_i - C+ L \cdot (1-B_i) \geq 0} \\
{0 \leq P_i \leq C}
\end{array} \right. $$
[3]

\section{Implementations}
For our implementation, we chose \emph{Perl} to implement converter, which will translate the RTL-level verilog to our constraints file.
The algorithm is: First, open the RTL verilog file as input, read in one line, find if it's an assertion, record signal name and its width;
Second, go to signal assignment part (usually right of first``=''), split whole sentence into words by space(s);
Then, record the number of operators as counter for iteration, each loop find a arithmetic unit where left operand does not contain ``)'' and right operand does not contain ``('',
calculate this unit with given operands and operator and their bits width, print out the constrains into result file;
Finally, judge if we need to generate a temporary variable to record the result, then replace the original expression ``(A op B)'' with new variable C in the string, continue iteration until all operations are done.
This is a \emph {pseudo-code}:

\begin{verbatim}
$readin = <>; 
\end{verbatim}
 \emph {read in a line}
\begin{verbatim}
@items = split /\s+/, $readin;  
\end{verbatim} 
\emph {split with spaces to get words}
\begin{verbatim}
if (Not an assertion) {next;}  
\end{verbatim} 
\emph {an assertion requires ``=" behind signal name}
\begin{verbatim}
$def_var = $items [1]; 
$wid_signal = /\[(\d+)\:/; \end{verbatim}  \emph {extract width of signal} \begin{verbatim} 
$var_list {$def_var} = $wid_signal;  \end{verbatim}  \emph {store this info into a hash table} \begin{verbatim} 
While (find basic operators)
{
    Operator count ++;
\end{verbatim}
\emph{find out the smallest unit, which just contains one operator}
\begin{verbatim}
	if (judge unit == FALSE) {next;}
\end{verbatim}
	\emph{Write linear constrains to output file}
\begin{verbatim}
    &extract($picked string) 
  $op_left, $op_right, $wid_left, $wid_right;  \end{verbatim}  \emph {extract 2 operands and their width (bits No.)} \begin{verbatim} 
if ($op_left || $op_right contain [i:j] or -) 
  {&unary_op($oprand);}  \end{verbatim}  \emph { judge if we need to process unary operand first } \begin{verbatim}
if (exsit $var_list {C} == FALSE) 
  {$obj = "my_var_$tmp_count";} 
&write_constrains 
  ($op_left, $op_right, $wid_left, 
	$wid_right, $obj) 
\end{verbatim}
\emph{remove excess brackets, store into new element. And replace the unit with an temporary variable}
\begin{verbatim}
    Find immediate (){},Remove(){};  \end{verbatim}  \emph {if operand is ``," for catenation then apply "\{\}" } \begin{verbatim} 
	replace substr $items[i-1][i][i+1] with $obj;  \end{verbatim}  \emph { replace for binary operation } \begin{verbatim}
}
\end{verbatim}

\section{Experimentation and Results}
The experimentation of our prototype required a little revise of the output constraints manually. Also the constraint of multiplier is using some logical function such as \emph{floor()}. The way we try to linearize these logical functions by introducing new temporary \textbf {Boolean} variables to regulate constraints. We conducted two tests (for translating a same verilog file but using different solvers). The first one with of standard AMPL/CPLEX:
\begin{verbatim}
ampl: model P0.mod;
ampl: option solver cplex;
ampl: solve;
CPLEX 11.2.0:
Objective = find a feasible point.
\end{verbatim}

 Second test was using MINOS, which is AMPL default solver :
\begin{verbatim}
ampl: option solver minos;
ampl: solve;
MINOS 5.51:
6 iterations
Nonlin evals: constrs = 6, Jac = 6.
Objective = find a feasible point.
\end{verbatim}

\section{Conclusion}
This project offered a successful demonstration of Word-Level RTL Equivalence Checking using Perl and AMPL/CPLEX. It red verilog file and converted the RTL-level descriptions into Linear Programming constraints. It also successfully verified the assignment with  AMPL/CPLEX.

\section{Future work}
While our prototype offers a working solution, it is far from its full potential. In the future work, we can add more LPATPG function into our project. We can also do more tests on boolean logical test. Designing a fully automatic converter is also in our a further plans. And build a user friendly GUI can be scheduled at same time.

\begin{thebibliography}{1}

\bibitem{ref1}
Z. Zeng, P. Kalla, M. J. Ciesielski, \emph{``LPSAT: A Uniﬁed Approach to RTL Satisﬁability,” } ICCAD'08, 2008.

\bibitem{ref2}
\emph{Linear Programming} http://en.wikipedia.org/wiki/Linear\_programming

\bibitem{ref3}
Z. Zeng \emph{``FUNCTIONAL TEST GENERATION BASED ON WORD-LEVEL SATISFIABILITY ,” } Dissertation of Zhihong Zeng 2002.

\end{thebibliography}

% that's all folks
\end{document}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

%\section{Conclusion}
%The conclusion goes here.

% conference papers do not normally have an appendix

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

% that's all folks
\end{document}

### 2014-04-25 sequential/latex/abstraction/aaproject.tex
upload files from laptop linux

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  IEEEsample.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%    More information: see the header of IEEEtran.sty
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%

\documentclass[11pt,twoside, onecolumn]{IEEEtran}
%\documentclass[conference]{IEEEtran}

%%%\IEEEoverridecommandlockouts

\usepackage[ruled]{./algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%% \RequirePackage{times}
%% \RequirePackage{algorithmic}
%% \PassOptionsToPackage{boxed}{algorithm}
%% \RequirePackage{algorithm}
%% \RequirePackage{multicol}
%\renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
%\renewcommand{\algorithmicensure}{\textbf{Outputs:}}
%\DeclareMathAlphabet{\mathtsl}{OT1}{ptm}{m}{sl}

%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em1
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{example}{Example}
%\newtheorem{corollary}{Corollary}

\RequirePackage{amssymb, mathptm}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{comment}

%%indent in algorithm

%\setcounter{page}{1}

% New command for the table notes.
\def\tabnote#1{{\small{#1}}}

% New command for the line spacing.
\newcommand{\ls}[1]
    {\dimen0=\fontdimen6\the\font
     \lineskip=#1\dimen0
     \advance\lineskip.5\fontdimen5\the\font
     \advance\lineskip-\dimen0
     \lineskiplimit=.9\lineskip
     \baselineskip=\lineskip
     \advance\baselineskip\dimen0
     \normallineskip\lineskip
     \normallineskiplimit\lineskiplimit
     \normalbaselineskip\baselineskip
     \ignorespaces
    }
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqarr}{\begin{eqnarray}}
\newcommand{\eeqarr}{\end{eqnarray}}
%\newcommand{\ov}{\overline}
\newcommand{\ov}{\bar}
\newcommand{\xor}{\bigoplus}
\newcommand{\Fm}{{\mathbb{F}}}

%the following is for space before and after align or other equation environment.

%%
\newtheorem{Algorithm}{Algorithm}[section]
\newtheorem{Definition}{Definition}[section]
\newtheorem{Example}{Example}[section]
\newtheorem{Proposition}{Proposition}[section]
\newtheorem{Lemma}{Lemma}[section]
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}{Corollary}[section]
\newtheorem{Conjecture}{Conjecture}[section]
\newtheorem{Problem}{Problem}[section]
\newtheorem{Notation}{Notation}[section]
\newtheorem{Setup}{Problem Setup}[section]
%%%

%%set spacing between table columns
\setlength{\tabcolsep}{3pt}

\begin{document}

%\thispagestyle{empty}
%\pagestyle{empty}

\ls{1.1}

\title{\large{\textsc{Report  on Algorithmic Tight Bound Analysis of
      Tree Metrics Approximation}}}
\author{Xiaojun Sun\\
A Final Project Report\\
as a satisfaction of requirement by Advanced Algorithms\\
Fall Semester 2013
}

%%\author{\IEEEauthorblockN{Jinpeng Lv and Priyank Kalla}\thanks{This work is sponsored in part by a grant from NSF \#CCF-546859.}
%\IEEEauthorblockA{Department of  Electrical and Computer Eng.\\
% University of Utah, Salt Lake City, UT-84112 \vspace{-0.3in}
 %\{lv, kalla\}@eng.utah.edu
% }
%\and
%\IEEEauthorblockN{Florian Enescu} \thanks{\normalsize  978-3-9810801-8-6/DATE12/$\copyright 2012$ EDAA}
%\IEEEauthorblockA{Department of Mathematics and Statistics\\
% Georgia State University,  Atlanta, GA 30302-4038 \vspace{-0.3in}
% fenescu@mathstat.gsu.edu
%} 
%

\maketitle

%\markboth{MS Proposal by Tim Pruss}{}
\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
\newcommand{\B}{{\mathbb{B}}}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\F}{{\mathcal{F}}}
\newcommand{\G}{{\mathcal{G}}}
\newcommand{\R}{\mathbb{R}}
%%%

\newcommand{\debug}[1]{\textcolor{gray}{[ #1 ]}}

%\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%% Include your files here %%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
This report is discussing the key techniques of J. Fakcharoenphol el.'s journal paper\cite{thispaper}.
First, we learn about the importance of adopting tree metric approximating algorithms. Then two key
techniques are found: the CKR like algorithm to decompose a graph, as well as a region-growing like
algorithm to derandomize the probabilistic approximation. Former algorithm guarantees the expected
distortion among a set of trees has an upper bound $O(logn)$, latter one tells the existence of a single
tree whose average distortion also satisfies upper bound $O(logn)$. At last this paper concludes that
the $O(logn)$ distortion is a tight bound, thus can improve the performance of all tree metric approximation
applications. 
\end{abstract}

\section{Problem Description}
\label{sec:desc}
The problem this paper addressing is as its title states: to find a tight bound for the algorithm 
which approximate arbitrary metrics with tree metrics.
\subsection{Importance of Tree Metrics Approximation Problem}
For a lot of problems in both algorithmic research and real life, there exists a metric representation.
One example is the Traveling Salesman Problem (TSP). This problem can be represented by a weighted graph, as long as specifying a metric. If we want to solve this
problem, we will need to run over all possible tours and find out the optimal, because this is currently
proved to be a NP-hard problem. However, since metric space satisfies triangle inequality (will be
discussed later in section \ref{sec:basic}), a Spanning Tree Heuristic\cite{h_tsp} can be adopted,
the minimum spanning tree is taken, and bunch of polynomial-time algorithms can be applied on this spanning
tree to provide a n approximation to TSP.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{h_tsp.png}
	\end{center}
	\caption{A spanning tree heuristic for TSP (revising from \cite{Jeff})}
	\label{fig:tsp}
\end{figure}

In nature, the spanning tree heuristic is a deterministic $\alpha$-approximation of a graph (without
loss of generality, we call it an arbitrary metric space) using tree metric.
This problem shows the importance of developing tree metrics approximation algorithms. Furthermore, if we could
embed an arbitrary metric into a tree with distortion $D$, then the optimal solution to this TSP
on the tree would be within $D$ of the optimal TSP tour on original input metric. As the distortion
be well-bounded at a lower level, the approximation algorithm will be further improved. More compelling
algorithmic applications of low distortion embeddings will be listed in section \ref{sec:potential}.

\subsection{History}
Approximations of metric spaces by "simpler" metric spaces has been intensively studied, both in
mathematics and computer science. 
"Embedding" is an important concept in these research areas,
it stands for a map from one metric space to another one, 
also can be defined in topology with geometry, algebra and domain theory with different sets/domains/fields. Johnson el.(1984)\cite{hilbert} discussed embeddings in Hilbert spaces at Functional
theory point of view; Graham el.(1985)\cite{zd} did some work about embeddings in $\Z^d$ in graph theory; moreover,
Linial el.(1994)\cite{realnorm} gave a research on low-distortion embeddings in low-dimensional real normed spaces.

As to research about tree metrics approximation, it origins from the paper by Cai el.(1995)\cite{Cai}
who gave out a tree spanner approximation to solve K-server problem. Soon, Bartal(1996)\cite{hst} introduced
an important structure called Hierarchically well Separated Tree (HST), and proved an arbitrary metric 
space can be $\alpha$-probabilistically-approximated with distortion $\alpha=O(log^2n)$, and lower bound is
$\Omega(logn)$(this will be extensively discussed in section \ref{sec:hst}). Later, Bartal(1998)\cite{bartal98}
developed Hierarchical Partition Metrics (HPM) based on HST fashion, and improved the bound of distortion
to $O(lognloglogn)$.

Later, people began to focus on how to find a decomposition of graph which can 
benefit the tree metrics approximation. Calinescu el.(2001)\cite{CKR} gave an $O(logk)$ approximation
algorithm based on a linear programming relaxation for 0-extension problem (k is number of terminals), which is known as
CKR algorithm (will be introduced in section \ref{sec:CKR}). The partition algorithm of CKR
was improved by Fakcharoenphol el.(2003, actually authors of this paper)\cite{thispaper}, which is adopted finally
in this paper to provide a probabilistic graph decomposition algorithm which guarantees the expected distortion
have a upper bound $O(logn)$.

\subsection{Potential Applications}
\label{sec:potential}
Tree metrics approximation can help solve bunch of applications in a variety of fields. The following
problems already have reasonable approximation algorithm on trees, so if our goal (to approximate
arbitrary metrics with tree metrics) is attained, these problems can be easily solved for much broader
inputs.

{\bf The group Steiner tree} problem. Given a weighted graph and a collection of $k$ sets, find
a Steiner tree connecting at least one element from each set. Garg el.\cite{GKR98} give an $O(logklogn)$ approximation 
algorithm for trees.

{\bf Buy-at-bulk network design}. Given a weighted graph and a sub-linear function of cost
describing the edge cost as a function of the load on it. If given a sequence of pairs of nodes,
the objective is to find a path connecting each pair, and minimize the total load. Awerbuch el. \cite{AA97}
give an $O(1)$-approximation algorithm on trees.

{\bf The communication spanning tree} problem\cite{Hu74}. Given a metric space network with costs on edges, the
communication spanning tree problem is to find a spanning tree and minimize a weighted sum of tree distances
over all pairs of nodes. Graph decomposition fashion approach can be applied trivially to find
desired spanning tree.

Meantime, probabilistic approximation of metric spaces is of particular importance in the case of
on-line problems where randomization against oblivious adversaries is very powerful.

{\bf Metric task systems} consists of a set of states forming a metric space and a set of tasks 
associated with costs in the different states. The goal is to schedule state transitions in order to
minimize the total move and task costs. This problem has deterministic competitive ratio of $2n-1$, and
proved to have same upper and lower bound to the $(n-1)$-server problem on $n$ points.\cite{MMS88}

{\bf The K-server problem} consists of K-servers in a metric space, points are requested over time
and a server must be moved to the request location. The memoryless K-server algorithm for trees
developed by Chrobak\cite{CL91} is very efficient randomized algorithm comparing to traditional work function 
algorithm.

\section{Techniques from the Paper}
\subsection{Introduction to Basic Concepts}
\label{sec:basic}
There are some definitions to make, which is necessary to understand the argument of this paper.
\begin{Definition}
Given a set $X$ of points, a \emph{distance function} on $X$ is a map $d : X \times X \to \R_+$ that is symmetric,
and satisfies $d(i, i) = 0\ \forall i \in X$. The distance is said to be a $metric$ if the triangle inequality
holds, i.e.
$$d(i, j) \leq d(i, k) + d(k, j),
\ \forall\ i, j, k \in X.$$
\end{Definition}
Furthermore, a $metric\ space$ is a set
where metric for every element pair is defined. In graph theory, this “distance” is weight of a
path between arbitrary vertices. So any undirected weighted graph (UWG) could be interpreted
as a metric space. Given metric space $G = (V, d)$ as an UWG, $E$ is a set including all edges in $G$, $|
V| = n$, pick arbitrary vertices $u, v$ from V, denote the distance between them as $d(u, v)$, The property of metric space can also be written in 3 specifications:\par
(a) $d(u, v) \geq 0$, and $d(u, v) = 0$ if and only if $u = v$ ;\par
(b) $d(u, v) = d(v, u)$ (symmetry);\par
(c) $d(u, v) \leq d(u, w) + d(w, v)$ (triangle inequality).
\\
We also mentioned the "embedding" and "distortion" in section \ref{sec:desc}, they can be formalized as
following:
\begin{Definition}
Given metric spaces $(X, d)$ and $(X', d' )$, a map $f : X \to X'$ is called an \emph{embedding}.
\end{Definition}
\begin{Definition}
Given two metrics $(X, d)$ and $(X' , d' )$ and a map $f : X \to X'$ , the \emph{contraction} of $f$ is the
maximum factor by which distances are shrunk, i.e.,
$$\max_{x,y\in X}\frac{d(x,y)}{d'(f(x),f(y))}$$
the $expansion$ or $stretch$ of $f$ is the maximum factor by which distances are stretched:
$$\max_{x,y\in X}\frac{d'(f(x),f(y))}{d(x,y)}$$
and the $distortion$ of $f$ , denoted by $||f||_{dist}$ , is the product of the contraction and the expansion.
\end{Definition}
Easy to understand that "distortion" is the maximum difference between 2 metrics. 
 Next it is necessary to 
define the most important concept in the tree approximation algorithm: "$\alpha$-probabilistic approximation".
\begin{Definition}
A metric space $N$ over $V$, \emph{dominates} a metric space $M$ over $V$, if for
every $u,v\in V$, $d_N(u,v)\geq d_M(u,v)$.
\end{Definition}
\begin{Definition}
A metric space $N$ over $V$, $\alpha$-$approximates$ a metric space $M$ over $V$, if it dominates $M$
and for every $u,v\in V$, $d_N(u,v) \leq \alpha\cdot d_M(u,v)$.
\end{Definition}
\begin{Definition}
A set of metric spaces $\mathcal{S}$ over $V$, $\alpha$-$probabilistically$-$approximates$ a metric
space $M$ over $V$, if every metric space in $\mathcal{S}$ dominates $M$ and $\exists$ a probability 
distribution over metric space $N\in \mathcal{S}$ such that for every $u,v\in V$, 
$Ex(d_N(u,v))\leq \alpha\cdot d_M(u,v)$.
\end{Definition}
Here $Ex$ means expectations. Consider an optimization problem $\mathcal{P}$ defined on metric spaces,
the cost of solution for $\mathcal{P}$ can be expressed as a linear combination of distances between
vertices in metric space.\cite{bartal98} Use the linearity of expectations, if the probability distribution is known,
$\alpha$-probabilistic approximation is a good approximation when deterministic approximation is 
unavailable, just like the problem we concern: approximate an arbitrary metric with a single tree metric.

There is an example to explain how to calculate distortion $\alpha$, for an $\alpha$-probabilistic approximation for a tree to a graph.
\begin{Example}
\label{ex:ncycle}
Suppose we have a $n$-cycle graph, remove one edge randomly to get a tree. For each edge $(u,v)\in E$ on this 
tree, the probability of distant change is $\frac{1}{n}$, while probability for unchange is $\frac{n-1}{n}$.
So the expectation of distortion on one edge is (assume initial distance in n-cycle for neighbors are all 1,
the distortion is then equal to distance on tree metric space):
$$Ex(d_T(u,v)) = \frac{1}{n}\cdot (n-1) + \frac{n-1}{n} \cdot 1 = 2\frac{n-1}{n} \leq 2$$
We can assert that this n-cycle can be \emph{2-probabilistically-approximated} by a tree with uniform probability distribution.
\end{Example}
\subsection{Graph Decomposition and CKR Algorithm}
\label{sec:CKR}
Example \ref{ex:ncycle} is a good explanation on how to build a tree out of a graph by cutting some edges.
The graph decomposition techniques reply on a CKR-like procedure. The CKR algorithm is shown below:
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{A semi-metric space (where distinct vertices may have 0 distance) $G=(V,d)$, while every edge has corresponding distance greater than 1}
 \KwOut{A tree semi-metric space dominating original semi-metric space\\} %, a Gr\"{o}bner basis
  Choose a random permutation $\sigma$ of terminals(for 0-extension problem);\\
  Choose a real number $\alpha$ uniformly at random from $[1,2)$;\\
  \For{each vertex $v\in V$}
  {
  	$l \gets 1$\;
  	\While{$v$ has not yet been assigned a terminal} 
	{
		\If{$d(v,\sigma(l)) \leq \alpha R(v)$}
		{
			assign $v$ to terminal $\sigma(l)$;
		}
		$l\gets l + 1$;
	}
   }
\caption {The CKR Algorithm}\label{alg:CKR}
\end{algorithm}
In this algorithm $R(v)$ is the distance of {\it closest terminal of $v$} to {\it this terminal's closest terminal}. 
Terminal is a special set of vertices in multi-way cut problem which can be used as roots of tree.
More details such as 0-extension problem and semi-metric are reasoned in \cite{CKR}, they do not make any difference to the new algorithm. Here, the authors make some modifications to this CKR algorithm
to achieve graph decomposition.

The procedure can be briefly described like below: use balls of different sizes to cut this graph, the first
ball's radius is $D_i$ which can contain all vertices. Then recursively choose vertices as centers to draw
balls in half of previous radius $D_{i-1} = \frac{D_i}{2}$. Once the circle (bound of ball) cuts any edges,
delete them, repeat drawing until every vertex is sitting within a ball, at this time 
we only keep edges from higher layer's center to lower layer's centers. Now doing same thing recursively
for each smaller ball with smaller radius $\frac{D_{i-1}}{2}$ (or in other words, go into lower layer $i-2$).
 Repeat this until every vertex is a center of a ball, all remaining edges and all vertices form a tree.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{laminar.png}
	\end{center}
	\caption{Brief procedure of graph decomposition by ball cutting\cite{thispaper}}
	\label{fig:laminar}
\end{figure}
We call this algorithm "FRT algorithm" in the same fashion of "CKR algorithm"(C,K,R are initials of authors).
We can see one main difference is the probability to choose ball radiuss, this will be important
in the tight bound analysis later in section \ref{sec:cutbound}.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{A metric space $G=(V,d)$, while every edge has corresponding distance strictly greater than 1}
 \KwOut{A tree metric space dominating original metric space\\} %, a Gr\"{o}bner basis
  Choose a random permutation $\pi$ of all vertices;\\
  Choose a real number $\beta$ randomly from $[1,2]$ at distribution $p(x) = \frac{1}{xln2}$;\\
  $D_\delta \gets V$; $i\gets \delta-1$;\ \  /* $D$ means set of vertices in the ball, largest ball has radius $2^\delta$ */\\
  \While{$D_{i+1}$ has non-singleton clusters}
  {
  	$\beta_i \gets 2^{i-1}\beta$;\ \ /* Choose radius not exactly on powers of 2 */\\
  	\For{$l = 1,2,\dots,n$} 
	{
		\For{every cluster $S$ in $D_{i+1}$}
		{
			Create a new cluster consisting of all unassigned vertices in $S$ closer than $\beta_i$ to $\pi(l)$
		}

	}
	$i\gets i-1$;
   }
\caption {The FRT Algorithm}\label{alg:FRT}
\end{algorithm}
A graphical example of FRT decomposition is attached in APPENDIX section.
\subsection{k-Hierarchically well Separated Trees}
\label{sec:hst}
The definition of k-hierarchically well separated tree ($k$-HST) was first state in Bartal's paper\cite{hst}.
\begin{Definition}
A {\it k-hierarchically well separated tree} is defined as a rooted weighted tree with following properties:\par
(a) The edge weight from any node to each of its children is the same.\par
(b) The edge weights along any path from the root to a leaf are decreasing by a factor of at least $k$.
\end{Definition}
Obviously in FRT graph decomposition, the tree we construct is 2-HST, because: weights on tree edge in the
same layer are all $2^{i}$ (upper bound), and the radius shrinks by factor 2 when going down one layer.
Actually in \cite{bartal98}, 2-HST can be converted to any $k$-HST with distortion $O(k/logk)$, this
proves the FRT algorithm is able to be better utilized on more general applications 
even taking another factor rather than 2.

\section{Computational Analysis}
\subsection{Graph Decomposition Bound Analysis}
\label{sec:cutbound}
In section \ref{sec:CKR} we mentioned the brief procedure of graph partition method (FRT algorithm) used 
in this paper. In this section, we will prove that the expectation of tree metric distortion compared
to original graph has a $O(log n)$ bound.

The computation of expectation of distortion follows the method in example \ref{ex:ncycle}. As the proof
in this paper, we define to formalize:
\begin{Definition}
Center $w$ \emph{settles} the edge $(u,v)$ at level/layer $i$ if it is the first center to which at least one 
end of this edge ($u$ and $v$) get assigned.
\end{Definition}

From this definition, there is exactly one center settles any edge $(u,v)$ at any particular level.
Note that edge $(u,v)$ is in level $i$ means $u$ and $v$ first get separated to different clusters at level $i$.
Also a formal definition of "cut" can be made here:
\begin{Definition}
Center $w$ \emph{cuts} edge $(u,v)$ at level $i$ if it settles $(u,v)$ at this level, but exactly one
of $u,v$ is assigned to (included in ball) $w$ at level $i$.
\end{Definition}

One example of "cut" is in fig \ref{fig:cut}, $\pi(1)$ is the center of grey shaded ball, it has radius
$2^{i-1}\beta$. Assume this ball is the first ball we drew in this level, and there exists an edge
$(\pi(10),\pi(2))$. Since $\pi(10)$ is first time to be assigned to $\pi(1)$, we call $\pi(1)$ cuts
this edge at level $i-1$.

%%%%%%%%figure:cut here
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{cut.png}
	\end{center}
	\caption{Sample for hierarchically ball cutting (revised from \cite{thispaper})}
	\label{fig:cut}
\end{figure}

Then we can review the proof in this paper (part 2.3, p 490-492) step by step. First, it says 
\begin{Theorem}
If $w$ cuts
edge $(u,v)$, the tree length of this edge $(u,v)$ is about $2^{i+2}$.
\end{Theorem}

{\it Proof}\ \ 
Tree metric we need
always dominates original metric, so we need to estimate the maximum; current radius that cut $(u,v)$ is
$2^i$ maximum (please refer to FRT algorithm, $2^{i-1}\beta \leq 2^i$), so the distance between $u,v$ is
2 times of current diameter, i.e. $4\times 2^{i} = 2^{i+1}$. Fig \ref{fig:distance}(a) illustrates this
assert. 
In that figure, $u$ is assigned to center $w$, and there must exist a center $w'$ comes later (in the same
level or lower levels) to which $v$ is assigned. We know center $w$ cuts edge $(u,v)$ at level $i$,
then the radius of ball $w$ is $r$, the radius of ball $w'$ must satisfy $r'\leq r$. Note that
the distance between $w,w'$ should be less than $2r$, otherwise $(u,v)$ is located already in a higher level.
 So the largest possible distance of $u,v$ is $4r$.
$\Box$

Then the new distance counting on the tree metric space associated with center $w$ is denoted with $d_w^T(u,v)$.
Consider the center $w$ is arbitrarily taken from a set of vertices, it is necessary
to apply the upper bound on an arbitrary partition. Arrange a sequence of these centers $\{w_1, w_2, \dots, w_s,\dots, w_n\}$ in the order of $d_{w_1}^T(u,v) \leq d_{w_2}^T(u,v)\leq\cdots d_{w_s}^T(u,v)\leq \cdots d_{w_t}^T(u,v)$, $w_s$ is picked arbitrarily. Without loss of generality, assume $d(w_s,u)\leq d(w_s,v)$.
To satisfy $w_s$ cuts edge $(u,v)$, 2 conditions must be fulfilled:\\
(a) $d(w_s,u) \leq \beta_i < d(w_s,v)$ for some $i$;\\
(b) $w_s$ settles $(u,v)$ at level $i$.

The expectation of tree metric distance when condition (a) is fulfilled is:
\begin{align}
Ex\left(d_{w_s}^T(u,v)\ |\ d(w_s,u) \leq \beta_i < d(w_s,v)\right) &= \sum Pr(\beta_i\ lies\ in\ [\ d(w_s,u),d(w_s,v)\ ))\cdot d_{w_s}^T(u,v)\nonumber\\
&= \int_{d(w_s,u)}^{d(w_s,v)}\frac{1}{xln2}\cdot d_{w_s}^T(u,v)\cdot dx\nonumber
\end{align} 
note that the value of $d_{w_s}^T(u,v)$ has upper bound $2^{i+2} = 8\cdot 2^{i-1}\leq 8\cdot 2^{i-1}\beta = 8\beta_i$.

Condition (b) is fulfilled when $w_s$ is the first in sequence of $<w>$ which can settle $(u,v)$
at level $i$. Since $w_j$ defines a tree distance smaller than $w_s$ when $j<s$, with the same radius,
 $w_j$ can surely "settle" this edge at level $i$, an example is fig \ref{fig:distance}(b). In this
 example we assume tree distance is $d(w_s,u)+d(w_s,v)$, easy to observe that any center $w_j$
 configuring a shorter tree distance is sitting at a location satisfying the definition of
 "settle" edge $(u,v)$ when applying radius $\beta_i$.

 %figure distances
 \begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{fig_distance.pdf}
	\end{center}
	\caption{Examples for partition upper bound proof}
	\label{fig:distance}
\end{figure}

Concluding above arguments, in a consecutive sequence $\{w_1,w_2,\dots,w_s\}$ every element can
settle $(u,v)$ if it is the first center selected at level $i$ under condition (a). Considering
that $w_j$ may also settle $(u,v)$ even $j>s$ and $w_s$ is randomly chosen, probability
$$Pr\text{(condition\ (b) } | \text{ condition\ (a))} \leq \frac{1}{s}$$
this is a conditional probability. Following assertion $Pr(A\ and\ B) = Pr(A)\cdot Pr(B|A)$,
probability for center $w_s$ to cut (u,v) is 
$$\frac{1}{s}\int_{d(w_s,u)}^{d(w_s,v)}\frac{1}{xln2}\cdot d_{w_s}^T(u,v)\cdot dx$$
since $\frac{1}{s}$ is irrelevant with variable $x$, it can be directly multiplied into
expectation calculation, so the expected tree metric distance between $u,v$ through $w_s$ is
\begin{align}
Ex\left(d_{w_s}^T(u,v)\right)
&\leq \int_{d(w_s,u)}^{d(w_s,v)}\frac{1}{xln2}\cdot 8x\cdot\frac{1}{s} dx\nonumber\\
&=\frac{8}{sln2}(d(w_s,v)-d(w_s,u))\\
&\leq \frac{8d(u,v)}{sln2}\ \ \ \text{(triangle\ inequality)}
\end{align}

$w_s$ is randomly chosen from sequence containing $n$ vertices. Using linearity of expectation,
the expected tree metric distance is bounded by
$$Ex(d_T(u,v)) \leq \sum_{s=1}^{n}\frac{8d(u,v)}{sln2} = O(logn)\cdot d(u,v)$$

The conclusion is
\begin{Theorem}
The distribution over tree metrics resulting from FRT algorithm $O(logn)$-probabilistically-approximates the original metric $d$.
\end{Theorem}
\subsection{Derandomization Bound Analysis}
Derandomization means that FRT algorithm not only provide a set of tree metrics with desired expectation
on distortion, but also could actually give a single tree metric such that its distortion is bounded by $O(logn)$.

To generalize, assume for each edge $(u,v)$ there is weight $c_{uv}$ assigned on it. 
We want to prove that
\begin{Proposition}
The FRT algorithm can find a tree metric space $d^T$ such that $(V,d^T)$ dominates original metric space
$G = (V,d)$
and satisfies $$\sum_{u,v\in V}c_{uv}\cdot d^T(u,v)\leq O(logn)\sum_{u,v\in V}c_{uv}\cdot d(u,v)$$
\end{Proposition}

In order to formalize, we can define
\begin{Definition}
Given an edge $e = (u,v)$ of length $d_e$ and weight $c_e$, the \emph{volume} of the edge is $d_e\cdot c_e$,
the \emph{sum\ of volume} $W = \sum_ec_ed_e$ corresponding to a set of qualified edges.
\end{Definition}

Using the definition of volume, it is easy to further define the following concepts vital in our proof.

\begin{Definition}
\emph{Ball} is a set of vertices located in a ball with center $t$ and radius $r$, denoted by $B(t,r)$.
The \emph{volume of neighborhood} $W(t,r)$ is computed by following criteria:\par
(a) edge $e=(u,v)$ with both ends in $B(t,r)$ contributes $c_ed_e$ to $W(t,r)$;\par
(b) edge $e=(u,v)$ with exactly one end such as $u$ in $B(t,r)$ contributes $c_e(r-d(t,u))$ to $W(t,r)$;\par
(c) edge $e=(u,v)$ with no ends in $B(t,r)$ contributes 0 to $W(t,r)$;\par
\emph{Cut volume} is denoted as $c(t,r)=\sum_{u\in B(t,r),v\notin B(t,r)}c_{uv}$ representing total weight of
edges cut by $B(t,r)$.
\end{Definition}

Based on above definitions, the proof in this paper can be reproduced. First, this assertion is proved:
\begin{Theorem}
There are radii $r_i: 2^{i-1}\leq r_i < 2^i$ such that $\sum_i\frac{c(t,r_i)}{W(t,r_i)}\cdot 2^i \leq O(logn)$
if $W$ is polynomial function of size $n$.
\end{Theorem}

The proof is very detailed in this paper already (part 3.1, p491), using region growing lemma from Garg el.\cite{Garg}
and contradiction. Then the author relaxed pre-condition that $W$ has to be polynomial function of 
size $n$ by find the actual upper bound of $W$. The largest edge length is $O(n)$, and unit volume
also affects the total volume linearly ($O(n)$), so the unit neighborhood volume on unit edge length $W_0 = W(t,1)$ is at least $(O(n\cdot n))^{-1} = \Omega(1/n^2)$ times total volume $W$, ratio $W/W_0$ is bounded
by $O(n^2)$, then $ln(n^2) = O(logn)$, the results with assumption still stand.

Above theorem is used to calculate the deterministic upper bound distortion of FRT algorithm.
Let $t$ be the center that maximizes $W(t,2^i)$ ($2^i$ is highest level maximum radius as mentioned in 
section \ref{sec:cutbound}). The FRT algorithm cut out $B(t,r_i)$ ($r_i$ is chosen in this level w.r.t. $t$),
and the new tree metric edge has length upper bound $2^{i+2}$ as discussed.

Recall the inequality in our proposition, $c_{uv}\cdot d^T(u,v) \leq c(t,r_i)\cdot 2^{i+2}$ for one edge at level $i$, 
$\sum_{u,v\in V}c_{uv}\cdot d(u,v) = W(t,r_i)$ according to definition. Sum up for all possible reached levels:
$$\frac{\sum_{u,v\in V}c_{uv}\cdot d^T(u,v)}{\sum_{u,v\in V}c_{uv}\cdot d(u,v)}=\sum_i\frac{c(t,r_i)}{W(t,r_i)}\cdot 2^{i+2} \leq 4ln\frac{W(t,2^{i+2})}{W(t,2^i)} \leq 8ln\frac{W}{W(t,1)} = O(log(n)),$$
our proposition is proved. Note here $t$ is chosen to guarantee it can still cut some edge with unit(minimum)
length 1 to form unit volume $W(t,1)$.

In conclusion:
\begin{Theorem}
FRT algorithm returns a 2-hierarchically well separated tree metric $d_T$ such that\par
(a) $\forall u,v \in V, d_T(u,v)\geq d(u,v)$;\par
(b) $\sum_{u,v\in V}c_{uv}\cdot d^T(u,v)\leq O(logn)\sum_{u,v\in V}c_{uv}\cdot d(u,v)$.
\end{Theorem}
\subsection{Computational Complexity Analysis}
In algorithm \ref{alg:FRT}, there are 3 cycles exploiting every vertices, the time complexity is 
$O(n^3)$, a polynomial time complexity.

\section{Further Developments on this Paper}
\subsection{Improvements on Relevant Applications}
\label{sec:app}
All applications we mentioned in \ref{sec:potential} and listed in Part 4 in this paper (we are discussing)
get an improvement on their performance since the upper bound of distortion is optimized. Moreover,
its graph decomposition approach also inspires some now approximations like spatial approximation in
metric labeling problem in image processing\cite{metriclabel} and trip planing about big data\cite{plan}.

In graph theory the low-stretch spanning tree is also discussed as an important topic\cite{lowspan}.
\subsection{Formalization and Refinement on Theory}
Some refinements are made on the FRT algorithm, especially on the probability distribution to select 
ball radii, such as Cai\cite{RMOP}. However, since the tight bound can be applied generally in
similar approximation algorithms, there is no improvement on further constraining the distortion.

And Shalekamp relaxed the distribution assumption made in this FRT algorithm, and directly gave a proof
for the tight bound, this is a big progress to formalize the tree approximation algorithms\cite{FRT}.

\section{My Comments}
The tree approximation algorithm is a good approach to solve some NP-hard problems in graph theory,
thus to find a tight bound of the tree metric distortion is important to optimize lots of approximation 
algorithms. The derandomization process guarantees its correctness for on-line problems. In conclusion,
this paper is innovative and vital enough and worth 400+ citations.
%%%%%%%%%%%%%%%%%%%% The bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ieee}
\bibliography{advalg}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-04-25 sequential/latex/abstraction/abstract.tex
upload files from laptop linux

\begin{abstract}
A combinational circuit with $k$-inputs and $k$-outputs implements
Boolean functions $f: \B^k \rightarrow \B^k$, where $\B = \{0, 1\}$.
Such functions can also be construed as a mapping $f: \Fkk
\rightarrow \Fkk$,  where $\Fkk$ denotes the Galois field of $2^k$
elements. Every function over $\Fkk$ is a polynomial function ---
i.e. there exists a unique, minimal, canonical polynomial
$\F$ that describes $f$. In this master's thesis, we propose to
investigate computer algebra based methods to derive the canonical
(word-level) polynomial representation of the circuit as $Y = \F (A)$
over $\Fkk$, where $A$ and $Y$ denote, respectively, the input and 
output bit-vectors of the circuit. We show that this can be achieved
by computing a Gr\"obner basis of a set of polynomials derived from
the circuit, using a specific elimination term order. 
%This
%approach can be generalized to any circuit with arbitray number of
%inputs and outputs to abstract polynomial functions over $f:
%{\mathbb{F}}^{2^n} \rightarrow {\mathbb{F}}^{2^m}$.
Computing Gr\"obner bases using elimination orders is, however,
practically infeasible for large circuits. To overcome this
limitation, we propose to investigate the use of the FGLM algorithm 
--- which converts a given Gr\"obner basis w.r.t. one term order $>_1$
into another w.r.t. an elimination term order $>_2$ ---  to  
derive the required polynomial abstraction. The objective of this
research is to: i) design an efficient CAD tool based on our approach;
and ii) apply it to reverse-engineer hardware implementations of
Elliptic Curve Cryptography (ECC) primitives over Galois fields $\Fkk$
--- with applications to design verification and function
identification. 
\end{abstract}

### 2014-04-25 sequential/latex/abstraction/abstraction.tex
upload files from laptop linux

\section{Word-Level Abstraction using Gr\"obner basis}
\label{sec:theory}
We are given a circuit $C$ with $k$-bit inputs and outputs that
performs a polynomial computation $Y = \F(A)$ over $\Fq = \Fkk$. Let
$P(x)$ be the {\it given} irreducible or primitive polynomial used for
field construction, and let $\alpha$ be its root, i.e. $P(\alpha) = 0
$. Note that we do not know the polynomial representation
$\F(A)$ and our objective is to identify (the coefficients of)
$\F(A)$. Let $\{a_0, \dots, a_{k-1}\}$ denote the primary inputs and
let $\{y_0, \dots, y_{k-1}\}$ be the primary outputs of $C$. Then, the
word-level and bit-level correspondences are: 
\begin{equation}
\label{eqn:words}
 A = a_0 + a_1 \alpha + \dots + a_{k-1} \alpha^{k-1}; ~~ Y = y_0 +
y_1 \alpha + \dots + y_{k-1} \alpha^{k-1};
\end{equation}

We analyze the circuit and model all the gate-level Boolean operators
as polynomials in ${\mathbb{F}}_2 \subset \Fkk$. To this set of
Boolean polynomials, we append the polynomials of Eqn
(\ref{eqn:words}) that relate the word-level and bit-level
variables. We model this set of polynomials as $F = \{f_1, \dots,
f_s\}$ over the ring $R = \Fq[x_1, \dots, x_d, Y, A]$. Here $x_1,
\dots, x_d$ denote, collectively, all the bit-level variables of the
circuit --- i.e. primary inputs, primary outputs and the intermediate
circuit variables --- and $Y, A$ the word-level variables. Denote the
generated ideal as $J = \langle F \rangle \subset R$. As $Y = \F(A)$
is a polynomial representation of the circuit, represent this (unknown)
``specification'' as a polynomial $f: Y - \F(A)$, or as $f: Y + \F(A)$
and $-1 = +1$ over $\Fkk$.

As the circuit $C$ implements the function $f$, clearly $f$ {\it
  agrees with the solutions} to $f_1 = \dots = f_s = 0$. In computer
algebra terminology, this means that $f$ {\it vanishes on the variety} 
$V_{\Fq}(J)$. If $f$ vanishes on $V_{\Fq}(J)$, then $f$ is a member of
the ideal $I(V_{\Fq}(J))$. Strong Nullstellensatz over Galois fields
(Theorem \ref{thm:strong-nullsatz-fq}) tells us that $I(V_{\Fq}(J)) =
J + J_0$, where $J_0 = \langle x_1^q - x_1, \dots, x_d^q - x_d, Y^q -
Y, A^q - A \rangle$ is the ideal of vanishing polynomials in
$R$. Consolidating these results, we deduce that the specification
polynomial $f \in (J+J_0)$.

If the specification polynomial is known, then the verification
problem can be solved using membership testing of $f$ in the ideal $(J
+ J_0)$ (\cite{lv:date2012} used such a formulation). We will now show
that by computing a Gr\"obner basis of $(J + J_0)$, using a specific
elimination term order, we can also identify the polynomial $f$ which
represents the function implemented by the circuit.

{\bf Reverse-Engineering $f$ from $C$:} The variety $V_{\Fq}(J)$ is
the set of all consistent assignments to the nets (signals) in the
circuit $C$. If we {\it project this variety on the word-level input and
output variables of the circuit $C$, we essentially generate the
function $f$ implemented by the circuit.} Projection of varieties from
$d$-dimensional space $\Fq^d$ onto a lower dimensional subspace
$\Fq^{d-l}$ is equivalent to {\it eliminating $l$ variables} from the
corresponding ideal.

\begin{Definition}
    ({\bf Elimination Ideal}) From \cite{ideals:book}:  Given
  $J=\langle f_1,\dots,f_s\rangle \subset \Fq[x_1,\dots,x_d]$, the
  $l$th {\bf elimination ideal} $J_l$ is the ideal of
  $\Fq[x_{j+1},\dots,x_d]$ defined by 
    \begin{equation}
        J_l= J \cap \Fq[x_{l+1},\dots,x_d]
    \end{equation}
\end{Definition}

In other words, the $l$th elimination ideal does not contain variables
$x_1,\dots,x_l$, nor do the generators of it.  This can aid in solving
systems of polynomial equations by isolating variables in a set of
constraints, as is the purpose of techniques such as Gaussian
elimination. 
%The generators of a $k$th elimination ideal are such that
%$variables  are not present.
Moreover, Gr\"obner bases may be used to generate an elimination ideal
by using an ``elimination order.''  One such ordering
is a pure lexicographic ordering, which features into a theorem:
%The ideal $I_k$ is spanned by elements which eliminate variables
%$x_1,\dots,x_k$.  This leads to an {\bf elimination theorem} for \Grobner
%bases:
\begin{Theorem} \label{thm:elim}
({\bf Elimination Theorem}) From \cite{ideals:book}: Let $J
  \subset \Fq[x_1,\dots,x_d]$ be an     ideal and let $G$ be a
  Gr\"obner basis of $J$ with respect to a lex ordering where $x_1
  > x_2 > \dots > x_d$.  Then for every $0     \leq l \leq
  d$, the set 
    \begin{equation}
        G_l= G \cap \Fq[x_{l+1},\dots,x_d]
    \end{equation}
    is a Gr\"obner basis of the $l$th elimination ideal $J_l$.
\end{Theorem}

We describe the application of elimination ideals using the following
example, borrowed from \cite{ideals:book}.

\begin{Example}

{\it 
Consider polynomials $f_1: x^2 - y - z - 1; ~~f_2: x - y^2 - z -1; 
~~f_3: x - y - z^2 - 1$ and ideal $J = \langle f_1, f_2, f_3\rangle
\subset {\mathbb{C}}[x, y, z]$. Let us compute a Gr\"obner basis $G$
of $J$ w.r.t. lex term order with $x > y > z$. Then $G = \{g_1, \dots,
g_4\}$ is obtained as: $g_1: x - y - z^2 - 1; ~~g_2: y^2 - y - z^2 - z;
~~g_3: 2yz^2 - z^4 - z^2; ~~g_4: z^6 - 4z^4 - 4z^3 - z^2$. 
Notice that the polynomial $g_4$ contains only the variable $z$, and
it {\bf eliminates} variables $x, y$. Similarly, polynomials $g_2,
g_3, g_4$, contain variables $y, z$ and eliminate $x$. According to
Theorem \ref{thm:elim}, $G_1 = G \cap {\mathbb{C}}[y, z] = \{g_2, g_3,
g_4\}$ and $G_2 = G \cap {\mathbb{C}}[z] = \{g_4\}$ are the Gr\"obner
bases of the $1^{st}$ and $2^{nd}$ elimination ideals of $J$, respectively.
}
\end{Example}

In conclusion, Gr\"obner basis computations w.r.t. pure lexicographic
term orders can be used to eliminate variables from an ideal. The
above example motivates our approach: Since we want to derive a
polynomial representation from a circuit in variables $Y, A$, we can
compute a Gr\"obner basis of $J + J_0$ w.r.t. an elimination order
that eliminates all the ($d$) bit-level variables of the
circuit. Then, the Gr\"obner basis $G_d = G \cap \Fq[x_1, \dots, x_d,
  Y, A]$ of the $d^{th}$ elimination ideal of $(J + J_0)$ will contain
polynomials in only $Y, A$. We now prove that the 
required polynomial representation will be found in $G_d$.
%$Y = \F(A)$
%will be found in $G_d$ and it will be the polynomial representation of
%the function implemented by the circuit.
First, let us formally ``setup'' the abstraction problem:

\begin{Setup}\label{not:abs}
Given a circuit $C$ with $k$-bit inputs and outputs which computes a
polyfunction $f: \Fkk \rightarrow \Fkk$. Let $\{a_0, \dots, a_{k-1}\}$
be the bit-level primary inputs and $\{y_0, \dots, y_{k-1}\}$ be the
primary outputs. Let $A, Y$ denote the word-level input and output
variables of the circuit, respectively, such that $A = a_0 + a_1
\alpha + \dots + a_{k-1}\alpha^{k-1}$ and $Y = y_0 + \dots +
y_{k-1}\alpha^{k-1}$, where $\alpha$ is a primitive element of $\Fkk$. 
Let ${\mathcal{F}}(A)$ be the (unknown) polynomial representation of
the function implemented by the circuit such that $Y =
{\mathcal{F}}(A)$.

Denote by $x_i, i = 1, \dots, d$ all the Boolean variables of the
circuit -- i.e. the input, output and the intermediate
variables. Let $R = \Fkk[x_i, Y, A: i = 1, \dots d]$ denote the
ring to model the polynomials that describe 
the circuit functionality. Let ideal $J \subset \Fkk[x_i, Y, A: i = 1
  \dots d]$ be generated by the bit-level and word-level polynomials
of the circuit. Let $J_0 = \langle x_i^2-x_i, Y^{2^k} - Y, A^{2^k} - A: i =
1, \dots, d\rangle$ denote the ideal of vanishing polynomials in $R$. 
\hfill$\Box$
\end{Setup}

Now, we will impose the following elimination order used for
abstraction:

\begin{Definition}
{\bf Abstraction Term Order $>$:} 
%For the given circuit $C$,
%let $x_1, \dots, x_d$ denote all the bit-level variables, and let $Y,
%A$ denote, respectively the word-level output and input
%variables. 
Using the variable order $x_1 > x_2 > \dots > x_d > Y > A$,
impose a lex term order $>$ on the polynomial ring $R = \Fq[x_1,
  \dots, x_d, Y, A]$. This elimination term order $>$ is defined as
the {\bf Abstraction Term Order}. 
\end{Definition}

\begin{Theorem} \label{thm:abs}
{\bf Abstraction Theorem:} Using the setup and notations from Problem
Setup \ref{not:abs} above, compute a Gr\"obner basis $G$ of ideal $(J
+ J_0)$ using the abstraction term order $>$. Then: \\
(i) $G$ must contain the vanishing polynomial $A^q - A$ as the only
polynomial with only $A$ as the support variable;\\
(ii) $G$ must contain a polynomial of the form $Y + {\mathcal{G}}(A)$;
and\\ 
(iii) $Y + {\mathcal{G}}(A)$ is such that $\F(A) = {\mathcal{G}}(A),
\forall A \in \Fq$. In other words, ${\mathcal{G}}(A)$ and $\F(A)$ are
equal as polynomial functions over $\Fq$.
\end{Theorem}

\begin{proof}
(i) $A^q -A$ is a given generator of $J_0$. Variable $A$ is also the
  last variable in the abstraction term order. Moreover, $A$ is an
  input to the circuit, so $A$ is an independent variable which can
  take any and all values in $\Fkk$. As a   result, $G_{d+1} = G \cap
  \Fkk[A] = \{A^q - A\}$.

(ii) Since $f:Y + {\mathcal{F}}(A)$ is a polynomial representation of
  the function of the circuit, $Y + {\mathcal{F}}(A) \in J + J_0$, as
  described above. Therefore, according to the definition of a
  Gr\"obner basis (Definition \ref{def:gb}), the leading term of $Y +
  {\mathcal{F}}(A)$ (which is $Y$) should be divisible by the leading
  term of some polynomial $g_i \in G$. The only way $lt(g_i)$ can
  divide $Y$ is when $lt(g_i) = Y$ itself. Moreover, due to our
  abstraction (lex) term order, $Y > A$, so this polynomial must be
  of the form $Y + {\mathcal{G}}(A)$.

(iii) As $Y = \F(A)$ represents the function of the circuit, $Y +
  \F(A) \in J + J_0$. Moreover, $V(J + J_0) \subset V(Y + \F(A))$. 
  Project this variety $V(J + J_0)$ onto the co-ordinates
  corresponding to $(A, Y)$. What we obtain is the {\it graph of the
  function} $(A) \mapsto \F(A)$ from $\Fkk \rightarrow \Fkk$. Since $Y
  + {\mathcal{G}}(A)$ is an element of the Gr\"obner basis of $J +
  J_0$, $V(J + J_0) \subset V(Y + {\mathcal{G}}(A))$ too. Therefore,
  $Y = {\mathcal{G}}(A)$ gives the same function as $Y = \F(A)$, for
  all $A \in \Fkk$.
\end{proof}

As a consequence of Theorem \ref{thm:abs}, if we compute a Gr\"obner
basis $G$ of $J + J_0$ using the abstraction term order, we will find
a polynomial of the form $Y + \G(A)$ in the Gr\"obner basis, such that
$Y = \G(A)$ is a polynomial representation of the circuit. However, if
the Gr\"obner basis is not reduced, it is possible to obtain multiple
polynomials in $G$ of the form $Y + \G _1(A), Y + \G _2(A), \dots,$;
all of which correspond to the same function.

\begin{Corollary}
Computing a {\bf reduced} Gr\"obner basis $G_r$ of $J + J_0$, we
will obtain {\bf one and only one polynomial} in $G_r$ of the form $Y
+ \G(A)$, such that $Y = \G(A)$ is the {\bf unique, minimal,
  canonical} representation of the function $f$ implemented by the
circuit.  
\end{Corollary}

The above results trivially extend to circuits with multiple
word-level input variables $A^1, \dots, A^n$, and the canonical
polynomial representation obtained by computing a reduced Gr\"obner
basis $G_r$ of $J + J_0$ using $>$ is of the form $Y = \F(A^1, \dots,
A^n)$.

\section{Research to be conducted further....}

{\it Gr\"obner basis Complexity:} To compute a Gr\"obner basis for
$J + J_0$ over $\Fq$, the following result is known \cite{gao:gf-gb-ms}:

\begin{Theorem}
Let $J = \langle f_1, \dots, f_s, ~x_1^q - x_1, \dots, x_d^q -
x_d\rangle \subset \Fq [x_1, \dots, x_d]$ be an ideal. The time and
space complexity of Buchberger's algorithm to compute a Gr\"obner
basis of $J$ is bounded by $q^{O(d)}$.
 assuming that the length of input $f_1, \dots, f_s$ is dominated by
 $q^{O(d)}$.  
\end{Theorem}

In our case, $q = 2^k$, and when $k$ and $d$ are large, this complexity
may make  verification infeasible. Therefore, I wish to conduct
research to make this approach scalable. Talk about FGLM here, and
maybe show an example corresponding to the same 2-bit multiplier
circuit....... You take it up from here....

### 2014-04-25 sequential/latex/abstraction/intro.tex
upload files from laptop linux

\section{Introduction}

Abstraction of word-level functionality from bit-level descriptions of
digital circuits is an important fundamental problem in Electronic
Design Automation (EDA) and design verification. Word-level abstractions
find applications in high-level/RTL datapath synthesis
\cite{demicheli:iccad_98} \cite{demicheli:dac_99}
\cite{demicheli:tcad_03}, RTL verification \cite{WLS} \cite{arditi:bmd}
\cite{lpsat}, word-level SMT and constraint solving \cite{ms:research}
\cite{boolector} \cite{tew:iccad08}, and can also be
applied to detect malicious modifications to a circuit -- potentially
inserted as hardware Trojan horses -- by reverse-engineering the
function implemented by the circuit. It is further desirable that the
obtained word-level abstraction be a {\it canonical} representation of
the function, to facilitate equivalence verification between a
specification model and an implementation circuit. This master's
thesis proposes the investigation of the problem of deriving word-level,
canonical, polynomial representations from circuits over Galois fields
using concepts from commutative algebra and algebraic geometry ---
notably, Gr\"obner bases \cite{gb_book}, Buchberger's algorithm
\cite{buchberger_thesis}, elimination ideals \cite{ideals:book}, 
and the FGLM algorithm \cite{fglm}.

The main motivation for this work is to ``reverse engineer'' or
identify the function implemented by a given Galois field arithmetic 
circuit as used in Elliptic Curve Cryptography (ECC). The main
operations of encryption, decryption and authentication in ECC rely on
point-addition and point-doubling operations on elliptic curves
defined over Galois fields $\Fkk$. These primitive operations are, in
turn, implemented as circuits that perform polynomial computations
over $k$-bit vectors \cite{eccld}. 
%Moreover, the datapath size $k$ in these applications
%can be very large, {\it e.g.} $k = 163, 233$ and larger, as specified
%by the US National Institute for Standards and Technology (NIST) for
%ECC. 
The objective is to apply our approach to a given circuit and
{\it extract  the word-level polynomial function} implemented by
it for design verification. The mathematical problems addressed in
this proposal, and the approaches to solve them, are described below:

{\bf Problem Statement:} Given a combinational circuit $C$ with
$k$-bit inputs and $k$-bit outputs, such that the circuit implements 
Boolean functions that are mappings between $k$-dimensional Boolean
spaces: $f: \B^k \rightarrow \B^k$. Let $\{a_0, \dots, a_{k-1}\}$
denote the primary inputs of the circuit, and let $\{y_0, \dots,
y_{k-1}\}$ denote the bit-level primary outputs. Suppose that we do
not know the function implemented by the circuit. We wish to
derive a {\it word-level, symbolic representation} of the function as
$Y = \F(A)$, where $Y = \{y_0, \dots, y_{k-1}\}$, $A = \{a_0, \dots, 
a_{k-1}\}$ are, respectively, the output and input bit-vectors,
and $\F$ denotes a polynomial representation of the
function $f$. We wish to further generalize our approach to any
circuit with arbitrary number  of inputs ($n$) and outputs ($m$), as
polynomial functions $f: {\mathbb{F}}_{2^n} \rightarrow
{\mathbb{F}}_{2^m}$.

{\bf Approach:} At a Boolean-level, canonical symbolic DAG
representations of functions have been extensively investigated; such
as decisions diagrams (ROBDDs \cite{BRYA86}, FDDs \cite{okfdd}),
moment diagrams (BMDs \cite{bmd}, K*BMDs \cite{kbmd}) and their
word-level variants \cite{WLS}. However, the decomposition principle
behind these representations is based on (variants of) the Shannon's
expansion, which is a bit-wise, Boolean decomposition. Such
approaches cannot be efficiently applied to word-level abstraction of
modulo-arithmetic circuits over Galois extension fields
$\Fkk$. Therefore, we approach this problem from a symbolic commutative
algebra perspective.

The function $f: \B^k \rightarrow \B^k$ is a
mapping among $2^k$ elements. Therefore, $f$ can also be interpreted,
algebraically, in the following two ways: i) as a function $f:
\Z_{2^k} \rightarrow \Z_{2^k}$, i.e. as a function over finite integer
rings $\pmod{ 2^k}$; and ii) as a function $f: \Fkk \rightarrow \Fkk$,
where $\Fkk$ represents the Galois field of $2^k$
elements. In this work, we will analyze $f$ as a polynomial function
over $\Fkk$, for the following reasons.

First of all, {\it not every
  function} of the type $f: \Z_{2^k} \rightarrow \Z_{2^k}$, is a
polynomial function; i.e. every function cannot be represented by a
polynomial $\F \pmod{ 2^k}$.  In commutative algebra, identification of
such polynomial functions is an important problem: i.e.  given a function
$f: \Z_n \rightarrow \Z_n$, $n \in {\mathbb{N}}$, identify if $f$ has
a polynomial representation. If so, derive a unique,
minimal, canonical polynomial $\F$, such that $f \equiv  \F \pmod{
  n}$. 
%that represents $f$. 
%This
%requires to analyze $f$ at each of the $n$ points, and to setup a
%system of $n$ linear congruences to solve. If this system of
%congruences has a solution, then $f$ is a polynomial function; 
%otherwise, $f$ is not a polynomial function. And the feasible
%solutions to these linear congruences correspond to the coefficients
%of a polynomial $\F$ that represents $f$.  
The papers
\cite{singmaster} \cite{chen_95} \cite{chen_96} addressed such
problems in number theory and algebra. {\it
 Shekhar} \cite{shekhar:phd} also addressed RTL verification of
polynomial datapaths over $k$-bit vectors using the above
concepts. However, in her work, the RTL datapath was already known to
be a polynomial function $\pmod{ 2^k}$. Unfortunately, an arbitrary
$k$-input/$k$-output combinational circuit cannot always be modeled as
a polynomial function over $\Z_{2^k}$; therefore, the $\Z_{2^k}$ model
is not considered in this work.

On the other hand, there is a well-known ``textbook'' result
\cite{ff:1997} which states that over Galois fields ($\Fq$) of $q$
elements, {\it every function} $f: \Fq \rightarrow \Fq$ is a
polynomial function. 
%By analyzing $f$ over each of the $q$ points, one can apply
%Langrange's interpolation formula and interpolate a polynomial $\F(x)
%= \sum_{k=1} ^q  \frac{ \prod_{i \neq k}  (x -x_i)}{\prod_{i \neq k}
%  (x_k -x_i)} \cdot f(x_k)$, which is a polynomial of degree at 
%most $q-1$ in $x$. One can easily see that $\F(a)=f(a)$ for all $a \in
%\Fq$, and $\F(x)$ is therefore the polynomial function representation
%of $f$. 
Motivated by this fundamental result, we will devise an approach
to extract a word-level polynomial function (abstraction) from a
circuit. Given the circuit $C$ which represents a function $f:
\B^k \rightarrow \B^k$, we will interpret $f$ as a function $f: \Fkk
\rightarrow \Fkk$, and derive a unique, minimal, canonical polynomial
representation as $Y = \F(A)$ over $\Fkk$, where $Y, A$ are,
respectively, the output and input bit-vectors of the circuit
$C$. 
%Note, however, that we are not given a function (mapping);
%rather, we are given a circuit, which is an encoding of the
%function. Interpolating a polynomial $\F$ over $2^k$ points is
%also going to be infeasible for large values of $k$. Therefore, 
We will extract a set of polynomials (ideal) from the circuit and employ
{\it Gr\"obner bases} techniques to abstract the canonical polynomial
representation. Our approach can be generalized to a circuit with
arbitrary number of inputs and outputs, i.e. to analyze polynomial
functions $f: \Fq^d \rightarrow \Fq$, $q = 2^k$, and further to $f:
{\mathbb{F}}_{2^n} \rightarrow {\mathbb{F}}_{2^m}$.

\begin{Example}
{\bf Motivating Example:} {\it
Consider a 2-bit multiplier circuit over ${\mathbb{F}}_{2^2}$ given in
Fig. \ref{fig:mul2bit}, which implements a function $f: {\mathbb{F}}_4
\times {\mathbb{F}}_4 \rightarrow {\mathbb{F}}_4$. It implements a
{\bf polynomial function}: $Z = A \times B, ~Z, A, B \in
{\mathbb{F}}_4$. Variables $a_0, a_1, b_0, b_1$ are primary inputs,
$z_0, z_1$ are primary outputs, and $s_0, s_1, s_2, s_3, r_0$ are
intermediate variables of the circuit. 
These bit-level variables are in ${\mathbb{F}}_2 = \{0, 1\}$.
%variables $a_0, a_1, b_0, b_1, g_0, g_1, s_0, s_1, s_2, s_3, r_0$ are
%in ${\mathbb{F}}_{2}$. 
Then, we can consider $A = a_0 + a_1 \alpha, ~B = b_0 +
b_1 \alpha$ as the word-level inputs and $Z = z_0 + z_1 \alpha$ 
as the output in ${\mathbb{F}}_4$. Here $P(x) = x^2 + x + 1$ is the primitive
polynomial of ${\mathbb{F}}_4$ and $P(\alpha) = 0$. 
\begin{figure}[htb]
\centerline{
\includegraphics[scale=0.4]{2bitmult.eps}
}
\caption{\small A 2-bit Multiplier over ${\mathbb{F}}_{2^2}$. The gate
  $\otimes$ corresponds to AND-gate, i.e. bit-level multiplication
  modulo 2. The gate $\oplus$ corresponds to XOR-gate, i.e. addition
  modulo 2.} 
\label{fig:mul2bit}
\end{figure}

The functionality of the entire circuit can be described using the
following polynomials derived from the Boolean gate-level operators: 
$f_1: z_0+z_1\alpha +Z; ~~f_2: b_0+b_1\alpha +B; ~~f_3: a_0+a_1 \alpha
+A; ~~f_4: s_0+a_0 \cdot b_0; ~~f_5: s_1+a_0 \cdot b_1; ~~f_6:
s_2+a_1 \cdot b_0; ~~f_7: s_3+a_1 \cdot b_1; ~~f_8: r_0+s_1 + s_2;
~~f_9: z_0+s_0 + s_3; ~~f_{10}: z_1 + r_0 + s_3$. If we impose the
following {\bf elimination} term order: {\bf lex term order} with
``circuit variables'' $>$ ``Output Z'' $>$ ``Inputs, A, B''
%More precisely, we can use {\bf lex} term order with $s_0 > s_1 >  s_2 >
%s_3 >  r_0 > a_0 > a_1 > b_0 > b_1 > g_0 > g_1 > A > B > G$.
%The above term order is an {\bf elimination} order. If we use the above
and compute a Gr\"obner basis $G$ of $\{f_1, \dots, f_{10}\}$, we observe
the following polynomials in the basis:  $g_1: z_0+z_1\alpha +Z;
~~g_2: b_0+b_1\alpha +B; ~~g_3: a_0+a_1 \alpha +A; ~~g_4: s_3+r_0+z_1;
g_5: s_1+s_2+r_0; ~~g_6: s_0+s_3+z_0; ~~{\mathbf{g_7: Z +  AB}}; ~~g_8:
a_1b_1+a_1B+b_1A+z_1; ~~g_9: r_0+a_1b_1+z_1; ~~g_{10}:
s_2+a_1b_0$. Notice that the polynomial $g_7: Z + AB$ describes $Z =
AB$ as the (canonical) polynomial function implemented by the circuit;
and we were able to {\it extract} this representation using the
Gr\"obner basis computation. We wish to explore such an approach to
derive an efficient decision procedure for word-level abstraction. 
}
\end{Example}

%\subsection{Contributions of this Thesis}
{\bf Contributions of this Thesis:} Our approach to solve this
problem, and the contributions of this thesis, can be outlined as
follows:

\begin{enumerate}
\item Using polynomial abstractions, we analyze the circuit, and
  model the gate-level Boolean operators as elements of a multivariate
  polynomial ring with coefficients in $\Fkk$.

\item Based on the concepts of {\it Strong Nullstellensatz, Gr\"obner
  bases, elimination ideals and projections of varieties}
  \cite{ideals:book}, we deduce that the polynomial abstraction
  problem can be formulated as one of {\it computing a Gr\"obner
    basis} of the set of polynomials derived from the given circuit
  netlist, using a specific elimination term order $>$.

\item Computing Gr\"obner bases using elimination term orders is
  infeasible for large circuits. To overcome this limitation, we will
  investigate the use of the FGLM algorithm \cite{fglm} to derive the
  polynomial representation. The FGLM algorithm takes an {\it already
    computed} Gr\"obner basis ($G_1$) w.r.t. an arbitrary term order
  $>_1$, and transforms it into another Gr\"obner basis ($G$)
  w.r.t. an elimination term order $>$. In the PhD thesis of {\it
  Lv} \cite{lv:phd}, it was shown that for any given
  circuit, there exists a term order $>_1$ that renders the set of
  polynomials derived from the circuit itself a Gr\"obner
  basis. Moreover, $>_1$ can be easily derived by performing a reverse
  topological traversal of the circuit. Consequently, $G_1$ is readily
  available as a Gr\"obner basis directly by construction. Using FGLM,
  we can then translate $G_1$ into the Gr\"obner basis $G$
  w.r.t. the required elimination order $>$.

\item The worst-case complexity of computing a Gr\"obner basis is
  known to be doubly exponential in $n$, the number of variables, and
  polynomial in $d$, the degree of the ideal. Most hardware/software
  verification applications do exhibit such high complexity, as the
  number of intermediate computations in the Buchbeger's algorithm
  \cite{buchberger_thesis} 
  tend to explode. Therefore, we conjecture that using the FGLM
  algorithm --- which relies on sparse linear algebra concepts ---
  perhaps this complexity can be avoided. In this thesis, we will
  develop a CAD framework using the above ideas and investigate the
  scalability of our methods.

\item As an application, we will use our CAD approach to
  reverse-engineer the function implemented by a given Galois field
  arithmetic circuits used in ECC   implementations. This can be used
  for function identification from a given circuit for formal
  verification.

\end{enumerate}

{\bf Proposal Organization:} The next section reviews previous work in
the area of canonical representations functions, word-level
abstractions, formal design verification using computer algebra
techniques, and also polynomial interpolation literature in symbolic
computing. Section \ref{sec:prelimgf} reviews 
Galois field theory and hardware design. Section \ref{sec:ideals}
reviews preliminary computer-algebra concepts related to ideals,
varieties, Gr\"obner bases, elimination ideals and
Nullstellensatz. Section \ref{sec:theory} describes our results on
polynomial abstraction from circuits, and demonstrates the application
of our work using preliminary experiments. The application of the FGLM
algorithm is covered in Section \ref{sec:fglm}. Finally, Section
\ref{sec:obj} outlines the research plan and concludes the proposal.

### 2014-04-25 sequential/latex/abstraction/iwls/abstract_iwls.tex
upload files from laptop linux

\begin{abstract}
A combinational circuit with $k$-inputs and $k$-outputs implements 
Boolean functions $f: \B^k \rightarrow \B^k$, where $\B = \{0, 1\}$.
The function can also be construed as a mapping $f: \Fkk
\rightarrow \Fkk$,  where $\Fkk$ denotes a Galois field of $2^k$
elements. Every function over $\Fkk$ is a polynomial function ---
i.e. there exists a unique, minimal, canonical polynomial
$\F$ that describes $f$. This paper presents novel techniques based on 
computer-algebra and algebraic-geometry to derive the canonical
(word-level) polynomial representation of the circuit as $Y = \F (A)$
over $\Fkk$, where $A$ and $Y$ denote, respectively, the input and 
output bit-vectors of the circuit. We show that this can be achieved
by computing a Gr\"obner basis of a set of polynomials derived from
the circuit, using a specific elimination term order. Our approach can
be generalized to circuits with arbitrary number of inputs and
outputs, as polyfunctions $f: \Fkk^d \rightarrow \Fkk$ or $f:
{\mathbb{F}}_{2^n} \rightarrow {\mathbb{F}}_{2^m}$.

Computing Gr\"obner bases using elimination orders is, however,
practically infeasible for large circuits. To overcome this
limitation, we present improvements to the computation
based on: i) term orders derived from the topological
traversal of the given circuit; and ii) the use of FGLM algorithm from 
algebraic geometry. We apply our approach to ``reverse-engineer''
hardware implementations of Elliptic Curve Cryptography (ECC)
primitives over Galois fields $\Fkk$ --- with applications to design
verification and function identification. 
\end{abstract}

### 2014-04-25 sequential/latex/abstraction/iwls/abstraction.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/abstraction/abstraction.tex
@@
 circuit variables --- and $Y, A$ the word-level variables. Denote the
-generated ideal as $J = \langle F \rangle \subset R$. As $Y = \F(A)$
-is a polynomial representation of the circuit, represent this (unknown)
-``specification'' as a polynomial $f: Y - \F(A)$, or as $f: Y + \F(A)$
-and $-1 = +1$ over $\Fkk$.
+generated ideal as $J = \langle F \rangle \subset R$. Also, let us
+denote the (unknown) ``specification'' of the circuit as a polynomial 
+$Y - \F(A)$; or equivalently as $Y + \F(A)$ and -1 = +1 over $\Fkk$.
 
-As the circuit $C$ implements the function $f$, clearly $f$ {\it
-  agrees with the solutions} to $f_1 = \dots = f_s = 0$. In computer
-algebra terminology, this means that $f$ {\it vanishes on the variety} 
-$V_{\Fq}(J)$. If $f$ vanishes on $V_{\Fq}(J)$, then $f$ is a member of
+As $Y = \F(A)$, clearly $Y + \F(A)$ {\it  agrees with the solutions}
+to the circuit equations $f_1 = \dots = f_s = 0$. In computer 
+algebra terminology, this means that $Y + \F(A)$ {\it vanishes on the
+  variety}  $V_{\Fq}(J)$. If $Y + \F(A)$ vanishes on $V_{\Fq}(J)$,
+then due to Proposition \ref{prop:finIV}, $Y + \F(A)$ is a member of
 the ideal $I(V_{\Fq}(J))$. Strong Nullstellensatz over Galois fields
@@
 Y, A^q - A \rangle$ is the ideal of vanishing polynomials in
-$R$. Consolidating these results, we deduce that the specification
-polynomial $f \in (J+J_0)$.
+$R$. Consolidating these results, we deduce that:
+
+\begin{Proposition} 
+\label{prop:spec_in_IV}
+The specification polynomial $Y + \F(A) \in
+  (J+J_0)$. 
+\end{Proposition}
 
 If the specification polynomial is known, then the verification
-problem can be solved using membership testing of $f$ in the ideal $(J
-+ J_0)$ (\cite{lv:date2012} used such a formulation). We will now show
-that by computing a Gr\"obner basis of $(J + J_0)$, using a specific
-elimination term order, we can also identify the polynomial $f$ which
-represents the function implemented by the circuit.
+problem can be solved via membership testing of $Y + \F(A)$ in the
+ideal $(J + J_0)$ (\cite{lv:date2012} actually used such a verification
+formulation). We will now show further that by computing a Gr\"obner
+basis of $(J + J_0)$, using a specific elimination term order, we can
+also identify the polynomial $Y + \F(A)$ which represents the function
+implemented by the circuit.
 
-{\bf Reverse-Engineering $f$ from $C$:} The variety $V_{\Fq}(J)$ is
+{\bf Reverse-Engineering $\F$ from $C$:} The variety $V_{\Fq}(J)$ is
 the set of all consistent assignments to the nets (signals) in the
@@
   $l$th {\bf elimination ideal} $J_l$ is the ideal of
-  $\Fq[x_{j+1},\dots,x_d]$ defined by 
+  $\Fq[x_{l+1},\dots,x_d]$ defined by 
     \begin{equation}
@@
 Moreover, Gr\"obner bases may be used to generate an elimination ideal
-by using an ``elimination order.''  One such ordering
+by using an ``elimination term order.''  One such ordering
 is a pure lexicographic ordering, which features into a theorem:
@@
 {\it 
-Consider polynomials $f_1: x^2 - y - z - 1; ~~f_2: x - y^2 - z -1; 
+Consider polynomials $f_1: x^2 - y - z - 1, ~~f_2: x - y^2 - z -1, 
 ~~f_3: x - y - z^2 - 1$ and ideal $J = \langle f_1, f_2, f_3\rangle
@@
 that eliminates all the ($d$) bit-level variables of the
-circuit. Then, the Gr\"obner basis $G_d = G \cap \Fq[x_1, \dots, x_d,
-  Y, A]$ of the $d^{th}$ elimination ideal of $(J + J_0)$ will contain
+circuit. Then, the Gr\"obner basis $G_d = G \cap \Fq[Y, A]$ 
+of the $d^{th}$ elimination ideal of $(J + J_0)$ will contain
 polynomials in only $Y, A$. We now prove that the 
@@
 Given a circuit $C$ with $k$-bit inputs and outputs which computes a
-polyfunction $f: \Fkk \rightarrow \Fkk$. Let $\{a_0, \dots, a_{k-1}\}$
+polyfunction $\F: \Fkk \rightarrow \Fkk$. Let $\{a_0, \dots, a_{k-1}\}$
 be the bit-level primary inputs and $\{y_0, \dots, y_{k-1}\}$ be the
@@
   \dots, x_d, Y, A]$. This elimination term order $>$ is defined as
-the {\bf Abstraction Term Order}. 
+the {\bf Abstraction Term Order}. The relative ordering among $x_1,
+\dots, x_d$ is not important and can be chosen arbitrarily. 
 \end{Definition}
@@
 \begin{Theorem} \label{thm:abs}
-{\bf Abstraction Theorem:} Using the setup and notations from Problem
-Setup \ref{not:abs} above, compute a Gr\"obner basis $G$ of ideal $(J
-+ J_0)$ using the abstraction term order $>$. Then: \\
+{\bf Abstraction Theorem:} Using the setup and notations from {\it
+  Problem Setup} \ref{not:abs} above, compute a Gr\"obner basis $G$ of
+ideal $(J + J_0)$ using the abstraction term order $>$. Then: \\
 (i) $G$ must contain the vanishing polynomial $A^q - A$ as the only
@@
 (ii) Since $f:Y + {\mathcal{F}}(A)$ is a polynomial representation of
-  the function of the circuit, $Y + {\mathcal{F}}(A) \in J + J_0$, as
-  described above. Therefore, according to the definition of a
-  Gr\"obner basis (Definition \ref{def:gb}), the leading term of $Y +
-  {\mathcal{F}}(A)$ (which is $Y$) should be divisible by the leading
-  term of some polynomial $g_i \in G$. The only way $lt(g_i)$ can
-  divide $Y$ is when $lt(g_i) = Y$ itself. Moreover, due to our
-  abstraction (lex) term order, $Y > A$, so this polynomial must be
-  of the form $Y + {\mathcal{G}}(A)$.
+  the function of the circuit, $Y + {\mathcal{F}}(A) \in J + J_0$, due
+  to Proposition \ref{prop:spec_in_IV}. Therefore, according to the
+  definition of a Gr\"obner basis (Definition \ref{def:gb}), the
+  leading term of $Y + {\mathcal{F}}(A)$ (which is $Y$) should be
+  divisible by the leading term of some polynomial $g_i \in G$. The
+  only way $lt(g_i)$ can divide $Y$ is when $lt(g_i) = Y$
+  itself. Moreover, due to our abstraction (lex) term order, $Y > A$,
+  so this polynomial must be of the form $Y + {\mathcal{G}}(A)$.
 
@@
   corresponding to $(A, Y)$. What we obtain is the {\it graph of the
-  function} $(A) \mapsto \F(A)$ from $\Fkk \rightarrow \Fkk$. Since $Y
+  function} $A \mapsto \F(A)$ over $\Fkk$. Since $Y
   + {\mathcal{G}}(A)$ is an element of the Gr\"obner basis of $J +
@@
 basis $G_r$ of $J + J_0$ using $>$ is of the form $Y = \F(A^1, \dots,
-A^n)$.
+A^n)$. Our approach can also be applied to circuits with $n$ inputs
+and $m$ outputs, to interpolate polyfunctions $f: {\mathbb{F}}_{2^n}
+\rightarrow {\mathbb{F}}_{2^m}$. This is achieved by modeling the
+function over ${\mathbb{F}}_{2^k}$ where $k = LCM(n, m)$ --- i.e.,
+over the smallest single field $\Fkk$ that contains both
+${\mathbb{F}}_{2^n}$ and ${\mathbb{F}}_{2^m}$.
 
-\section{Research to be conducted further....}
+\begin{Example}
+\label{ex:twomult}
+{\bf Demonstration of our approach:} {\it
+Consider a 2-bit multiplier circuit over ${\mathbb{F}}_{2^2}$ given in
+Fig. \ref{fig:mul2bit}, which implements a function $f: {\mathbb{F}}_4
+\times {\mathbb{F}}_4 \rightarrow {\mathbb{F}}_4$. It is a
+{\bf polynomial function}: $Z = A \times B, ~Z, A, B \in
+{\mathbb{F}}_4$. Variables $a_0, a_1, b_0, b_1$ are primary inputs,
+$z_0, z_1$ are primary outputs, and $s_0, s_1, s_2, s_3, r_0$ are
+intermediate variables of the circuit. 
+%These bit-level variables are in ${\mathbb{F}}_2 = \{0, 1\}$.
+%variables $a_0, a_1, b_0, b_1, g_0, g_1, s_0, s_1, s_2, s_3, r_0$ are
+%in ${\mathbb{F}}_{2}$. 
+Then, we can consider $A = a_0 + a_1 \alpha, ~B = b_0 +
+b_1 \alpha$ as the word-level inputs and $Z = z_0 + z_1 \alpha$ 
+as the output in ${\mathbb{F}}_4$. Here $P(x) = x^2 + x + 1$ is the primitive
+polynomial used to construct the field ${\mathbb{F}}_4$ and $P(\alpha)
+= 0$, i.e. $\alpha$ is a root of $P(x)$. 
+\begin{figure}[htb]
+\centerline{
+\includegraphics[scale=0.3]{2bitmult.eps}
+}
+\caption{\small A 2-bit Multiplier over ${\mathbb{F}}_{2^2}$. The gate
+  $\otimes$ corresponds to AND-gate, i.e. bit-level multiplication
+  modulo 2. The gate $\oplus$ corresponds to XOR-gate, i.e. addition
+  modulo 2.} 
+\label{fig:mul2bit}
+\end{figure}
 
-{\it Gr\"obner basis Complexity:} To compute a Gr\"obner basis for
-$J + J_0$ over $\Fq$, the following result is known \cite{gao:gf-gb-ms}:
-
-\begin{Theorem}
-Let $J = \langle f_1, \dots, f_s, ~x_1^q - x_1, \dots, x_d^q -
-x_d\rangle \subset \Fq [x_1, \dots, x_d]$ be an ideal. The time and
-space complexity of Buchberger's algorithm to compute a Gr\"obner
-basis of $J$ is bounded by $q^{O(d)}$.
- assuming that the length of input $f_1, \dots, f_s$ is dominated by
- $q^{O(d)}$.  
-\end{Theorem}
-
-In our case, $q = 2^k$, and when $k$ and $d$ are large, this complexity
-may make  verification infeasible. Therefore, I wish to conduct
-research to make this approach scalable. Talk about FGLM here, and
-maybe show an example corresponding to the same 2-bit multiplier
-circuit....... You take it up from here....
+The functionality of the entire circuit can be described using the
+following polynomials derived from the Boolean gate-level operators: 
+$f_1: z_0+z_1\alpha +Z; ~~f_2: b_0+b_1\alpha +B; ~~f_3: a_0+a_1 \alpha
++A; ~~f_4: s_0+a_0 \cdot b_0; ~~f_5: s_1+a_0 \cdot b_1; ~~f_6:
+s_2+a_1 \cdot b_0; ~~f_7: s_3+a_1 \cdot b_1; ~~f_8: r_0+s_1 + s_2;
+~~f_9: z_0+s_0 + s_3; ~~f_{10}: z_1 + r_0 + s_3$. Ideal $J = \langle
+f_1, \dots, f_{10}\rangle$. Generate $J_0$ as the ideal of vanishing
+polynomials. Impose the
+following {\bf elimination} term order: {\bf lex term order} with
+``circuit variables'' $>$ ``Output Z'' $>$ ``Inputs, A, B''
+%More precisely, we can use {\bf lex} term order with $s_0 > s_1 >  s_2 >
+%s_3 >  r_0 > a_0 > a_1 > b_0 > b_1 > g_0 > g_1 > A > B > G$.
+%The above term order is an {\bf elimination} order. If we use the above
+and compute a Gr\"obner basis $G$ of $J + J_0$, we observe
+the following polynomials in the basis:  $g_1: z_0+z_1\alpha +Z;
+~~g_2: b_0+b_1\alpha +B; ~~g_3: a_0+a_1 \alpha +A; ~~g_4: s_3+r_0+z_1;
+g_5: s_1+s_2+r_0; ~~g_6: s_0+s_3+z_0; ~~{\mathbf{g_7: Z +  AB}}; ~~g_8:
+a_1b_1+a_1B+b_1A+z_1; ~~g_9: r_0+a_1b_1+z_1; ~~g_{10}:
+s_2+a_1b_0$, and the polynomials of $J_0$. Notice that the polynomial
+$g_7: Z + AB$ describes $Z = AB$ as the (canonical) polynomial
+function implemented by the circuit; and we were able to {\bf extract}
+this representation using the Gr\"obner basis computation. 
+%We wish to explore such an approach to
+%derive an efficient decision procedure for word-level abstraction. 
+}
+\end{Example}

### 2014-04-25 sequential/latex/abstraction/iwls/fglm.tex
upload files from laptop linux

\section{Preliminary experiments and FGLM}
\label{sec:fglm}

%{\it Gr\"obner basis Complexity:} To compute a Gr\"obner basis for
%$J + J_0$ over $\Fq$, the following result is known \cite{gao:gf-gb-ms}:

\begin{Theorem}
(From \cite{lv:date2012}:) Let $J + J_0 = \langle f_1, \dots, f_s, ~x_1^q -
  x_1, \dots, x_d^q - x_d\rangle \subset \Fq [x_1, \dots, x_d]$ be an
  ideal. The time and space complexity of Buchberger's algorithm to
  compute a Gr\"obner basis of $J + J_0$ is bounded by $q^{O(d)}$.
 assuming that the length of input $f_1, \dots, f_s$ is dominated by
 $q^{O(d)}$.  
\end{Theorem}

In our case $q = 2^k$, and when $k$ and $d$ are large, this complexity 
makes verification infeasible. The impact of this complexity is clearly
visible in the results of preliminary experiments (TABLE \ref{tab:sT}).
The experiments are conducted with Mastrovito \cite{mastro:1989}
multiplier circuits of various word sizes $k$. Each multiplier
computes the polynomial function  $Y = \F(A,B) = A \cdot B$ over
$\Fkk$, where $Y, A, B$ are $k$-bit vectors. We extract the 
Boolean gate-level operators $J$ and generate vanishing polynomials $J_0$.
Then we compute the Gr\"obner basis of $J + J_0$
with respect to abstraction term order $>$. The computation was performed
using the "slimgb" command in the {\sc Singular'} computer algebra
tool \cite{DGPS}. The resulting Gr\"obner bases contained a polynomial
$Y + A \times B$.  
These experiments were run on a 64-bit Ubuntu OS running on a 2.4GHz 
Intel QuadCore processor with 8Gb of memory. The approach is
infeasible beyond 40-bit datapath size, as the Gr\"obner basis engine
encounters a memory explosion. BDDs cannot be constructed for
any of these circuits.

%In Elliptic Curve Cryptography, the 
%minimum suggested word-size is 163 bits, as set by the US National Institute 
%for Standards and Technology. Therefore, I wish to conduct research to make 
%this approach scalable.

\begin{table}[h]
{\small
	\begin{center}
	    \caption{Gr\"obner Basis runtime for
              Mastrovito Multipliers in Singular using Abstraction
              Term Order $>$.}\label{tab:sT} 
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & \# Polynomials (\# gates) & Time (minutes)   \\
		\hline
	        $16$	&  $1,871$  & $2.4$ \\
		$24$	&  $3,135$  & $12$  \\
	        $32$	&  $5,549$  & $22.6$ \\
	        $40$	&  $8,587$  & $266$ \\
		$48$	& $12,327$  & NA (Out of Memory) \\
	        \hline
	    \end{tabular}
	\end{center} 
}
\end{table}

{\it Proposed FGLM Approach:} To make this approach scalable, we are
investigating the use of the FGLM algorithm \cite{fglm}. This
algorithm provides a method for converting a Gr\"obner basis in one
term ordering to a  Gr\"obner Basis in another term ordering.
We use this method in conjunction with following result which was
derived and proved in \cite{lv:date2012}:

\begin{Theorem} \label{thm:top-order}
Let $C$ be any arbitrary combinational circuit. Let $\{x_1, \dots,
x_d\}$ denote the set of all variables (signals) in the circuit,
i.e. the primary input, intermediate and primary output
variables. Perform a {\it reverse topological traversal} of the
circuit and order the variables such that $x_i > x_j$ if $x_i$ appears
earlier in the reverse topological order. Impose a lex term order to
represent each gate as a polynomial $f_i$,
s.t. $f_i = x_i + \text{tail}(f_i)$. Then the set of all polynomials 
$G_1 = \{f_1, \dots, f_s, x_i^q - x_i: i = 1, \dots, d\}$ constitutes
a Gr\"obner basis.
%, as $lt(f_i)$ and $ lt(f_j)$ for $i\neq j$ are relatively prime.
\end{Theorem}

Using the term ordering specified by \cite{lv:date2012}, which we
denote $>_1$,  we can obviate the need to compute a Gr\"obner
basis. Imposing the monomial ordering $>_1$ already makes $G_1$ a
Gr\"obner basis of $J + J_0$. Now we apply the FGLM algorithm to
transform the Gr\"obner basis into another basis w.r.t. the
abstraction ordering $>$.  This new basis will also definitely contain
a polynomial in the form of $Y + \F(A)$ as desired. By using FGLM to
convert the Gr\"obner basis from $>_1$ to $>$, we are hoping to avoid
the complexity presented by computing a Gr\"obner basis over the
abstraction ordering $>$.

\begin{Example}
\label{ex:myapproach}
{\it
Let us revisit the 2-bit multiplier circuit from Example \ref{ex:twomult}, 
shown in Fig. \ref{fig:mul2bit}.

%\begin{figure}[htb]
%\centerline{
%\includegraphics[scale=0.35]{2bitmult.eps}
%}
%\caption{\small A 2-bit Multiplier over ${\mathbb{F}}_{2^2}$. The gate
%  $\otimes$ corresponds to AND-gate, i.e. bit-level multiplication
%  modulo 2. The gate $\oplus$ corresponds to XOR-gate, i.e. addition
%  modulo 2.} 
%\label{fig:mul2bit2}
%\end{figure}
As before, we can describe the functionality of this circuit, $J$, using the
following polynomials: $f_1: Z+z_0+z_1\alpha; ~~f_2: B+b_0+b_1\alpha; ~~f_3: A+a_0+a_1 \alpha
; ~~f_4: s_0+a_0 \cdot b_0; ~~f_5: s_1+a_0 \cdot b_1; ~~f_6:
s_2+a_1 \cdot b_0; ~~f_7: s_3+a_1 \cdot b_1; ~~f_8: r_0+s_1 + s_2;
~~f_9: z_0+s_0 + s_3; ~~f_{10}: z_1 + r_0 + s_3$.
By imposing the monomial order $>_1$, $J+J_0$ is already a Gr\"obner basis, 
where $J_0$ is the ideal of vanishing polynomials of the circuit:
$~~f_{11}: a_0^2+a_0; ~~f_{12}: a_1^2+a_1; ~~f_{13}: b_0^2+b_0;
~~f_{14}: b_1^2+b_1; ~~f_{15}: s_0^2+s_0; ~~f_{16}: s_1^2+s_1;
~~f_{17}: s_2^2+s_2; ~~f_{18}: s_3^2+s_3; ~~f_{19}: r_0^2+r_0;
~~f_{20}: z_0^2+z_0; ~~f_{21}: z_1^2+z_1$. We denote this Gr\"obner
basis as $G_1 = \{f_1, \dots, f_{21}\}$. Using the FGLM algorithm 
("fglm" command in  Singular) we convert $G_1$
from ordering $>_1$ to $G$ with the abstraction term ordering $>$. 
The resulting Gr\"obner basis $G$ contains the 
polynomial $Z \cdot (\alpha+1)+A \cdot B \cdot (\alpha+1)$, 
which can be equivalently reduced to $Z+A \cdot B$.
}
\end{Example}

FGLM exploits concepts from algebraic geometry and sparse linear
algebra to transform the Gr\"obner basis. While a detailed exposition
of FGLM is beyond the scope of this paper, we demonstrate its
operation on the 2-bit multiplier shown in Example
\ref{ex:myapproach}.

\begin{Example}
\label{ex:approach}
{\it
FGLM begins by taking the least variable in the the abstraction term 
ordering, ($A$ in our case). Starting with $m=0$, it computes $A^m
\xrightarrow{G_1}_+r$. It checks whether the remainder $r$ can be
represented as a linear combination of any remainders calculated thus
far. If so, it adds this  representation to $G$ and moves on to the
next monomial; otherwise it  increments $m$ and repeats $A^m
\xrightarrow{G_1}_+r$ computation.

\begin{itemize}
\item $A^0 \xrightarrow{G_1}_+ 1$
\item $A^1 \xrightarrow{G_1}_+ a_0+ a_1 \alpha$
\item $A^2 \xrightarrow{G_1}_+ a_0+a_1 \cdot (\alpha+1) $
\item $A^3 \xrightarrow{G_1}_+ a_0 \cdot a_1+a_0+a_1$
\item $A^4 \xrightarrow{G_1}_+ a_0+ a_1 \alpha = A$
\end{itemize}

$A^4$ can be composed of $A$, so $A^4 - A$ is added to $G$. Similarly,
$B^4 - B$ is also added to $G$. Continue to the next variable $Z$ in
the order as we have ``circuit variables'' $> Z > A > B$. 
\begin{itemize}
\item $Z^1 \xrightarrow{G_1}_+ a_0 \cdot b_0+(\alpha) \cdot a_0 \cdot
  b_1+(\alpha) \cdot a_1 \cdot b_0+(\alpha^2) \cdot a_1 \cdot b_1 = 
a_0(b_0 + b_1 \alpha) + a_1\alpha (b_0 + b_1 \alpha) = (a_0 + a_1
\alpha)(b_0 + b_1 \alpha) = A   \cdot B$ 
\end{itemize}
$Z$ is found to be a function of $A$ and $B$, therefore $Z - AB$ is
added to $G$.
}
\end{Example}

FGLM continues to convert the rest of the monomials into the new ordering. However, 
since we are only looking for $Y=\F(A)$ polynomial representation, we
can make this approach even more efficient by restricting the
conversion to only the word-level variables of the circuit. This idea
is currently under development.

\section{Conclusions}
\label{sec:concl}

This paper has described an approach to derive a word-level, canonical
polynomial representation from combinational circuits using Gr\"obner
bases. Our approach interprets the function of the circuit $f: \B^k
\rightarrow \B^k$ as a polynomial function over Galois fields $f: \Fkk
\rightarrow \Fkk$. By deriving a set of polynomials corresponding to
the circuit (ideal $J + J_0$), and computing its Gr\"obner basis
w.r.t. a specific elimination (abstraction) term order, the canonical
representation of this circuit can be derived. As Gr\"obner basis
computation exhibits high complexity, we are currently investigating
the use of the FGLM algorithm to make our approach scalable.

### 2014-04-25 sequential/latex/abstraction/iwls/intro_iwls.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/abstraction/intro.tex
@@
 find applications in high-level/RTL datapath synthesis
-\cite{demicheli:iccad_98} \cite{demicheli:dac_99}
-\cite{demicheli:tcad_03}, RTL verification \cite{WLS} \cite{arditi:bmd}
+%\cite{demicheli:iccad_98} \cite{demicheli:dac_99}
+\cite{demicheli:tcad_03}, RTL verification \cite{WLS} %\cite{arditi:bmd}
 \cite{lpsat}, word-level SMT and constraint solving \cite{ms:research}
-\cite{boolector} \cite{tew:iccad08}, and can also be
-applied to detect malicious modifications to a circuit -- potentially
-inserted as hardware Trojan horses -- by reverse-engineering the
-function implemented by the circuit. It is further desirable that the
-obtained word-level abstraction be a {\it canonical} representation of
-the function, to facilitate equivalence verification between a
-specification model and an implementation circuit. This master's
-thesis proposes the investigation of the problem of deriving word-level,
-canonical, polynomial representations from circuits over Galois fields
-using concepts from commutative algebra and algebraic geometry ---
-notably, Gr\"obner bases \cite{gb_book}, Buchberger's algorithm
-\cite{buchberger_thesis}, elimination ideals \cite{ideals:book}, 
-and the FGLM algorithm \cite{fglm}.
+\cite{boolector}, %\cite{tew:iccad08}, 
+etc. Functional abstraction can
+also be helpful in detecting malicious modifications to a circuit --
+potentially inserted as hardware Trojan horses -- by
+{\it reverse-engineering} the function implemented by the circuit\footnote{A
+hardware Trojan modifies the functionality of the 
+circuit. Abstraction of the function at word-level can perhaps give
+some insights into the functional {\it intent} of the Trojan. This can
+also be helpful in Trojan classification; such efforts are
+currently underway within the EDA community \cite{Bhunia:Trojan}.}. 
+It is further desirable that the obtained word-level abstraction be a
+{\it canonical} representation of the function, to facilitate
+equivalence verification between a specification model and an
+implementation circuit.
 
-The main motivation for this work is to ``reverse engineer'' or
-identify the function implemented by a given Galois field arithmetic 
-circuit as used in Elliptic Curve Cryptography (ECC). The main
-operations of encryption, decryption and authentication in ECC rely on
-point-addition and point-doubling operations on elliptic curves
-defined over Galois fields $\Fkk$. These primitive operations are, in
-turn, implemented as circuits that perform polynomial computations
-over $k$-bit vectors \cite{eccld}. 
+This paper presents a symbolic method to derive the {\it word-level,
+  canonical, polynomial representation} from a given combinational
+circuit. Our approach models the problem over Galois fields, and employs
+concepts from commutative algebra and algebraic geometry ---
+notably, Gr\"obner bases \cite{ideals:book} theory and
+technology --- to derive the word-level abstraction.
+%Buchberger's algorithm
+%\cite{buchberger_thesis}, elimination ideals 
+%and the FGLM algorithm \cite{fglm}.
+
+The main motivation for this work is to ``reverse engineer'' (or
+to identify) the function implemented by a given Galois field
+arithmetic circuit as used in Elliptic Curve Cryptography (ECC). The
+main operations of encryption, decryption and authentication in ECC
+rely on point-addition and point-doubling operations on elliptic
+curves defined over Galois fields $\Fkk$. These primitive operations
+are, in turn, implemented as circuits that perform polynomial
+computations over $k$-bit vectors \cite{eccld}. 
 %Moreover, the datapath size $k$ in these applications
@@
 The objective is to apply our approach to a given circuit and
-{\it extract  the word-level polynomial function} implemented by
-it for design verification. The mathematical problems addressed in
-this proposal, and the approaches to solve them, are described below:
+{\it extract  the word-level polynomial function (polyfunction)}
+implemented by it for design verification. While our approach can be
+applied to any arbitrary combinational circuit, it is most beneficial
+for arithmetic circuits where canonical Boolean representations prove
+to be infeasible.
 
@@
 derive a {\it word-level, symbolic representation} of the function as
-$Y = \F(A)$, where $Y = \{y_0, \dots, y_{k-1}\}$, $A = \{a_0, \dots, 
-a_{k-1}\}$ are, respectively, the output and input bit-vectors,
-and $\F$ denotes a polynomial representation of the
-function $f$. We wish to further generalize our approach to any
+$Y = \F(A)$ over $\Fkk$, where $Y = \{y_0, \dots, y_{k-1}\}$, $A =
+\{a_0, \dots,  a_{k-1}\}$ are, respectively, the output and input
+bit-vectors (words), and $\F$ denotes a polynomial representation of
+the function $f$. We wish to further generalize our approach to any
 circuit with arbitrary number  of inputs ($n$) and outputs ($m$), as
-polynomial functions $f: {\mathbb{F}}_{2^n} \rightarrow
+polyfunctions $f: {\mathbb{F}}_{2^n} \rightarrow
 {\mathbb{F}}_{2^m}$.
 
-{\bf Approach:} At a Boolean-level, canonical symbolic DAG
-representations of functions have been extensively investigated; such
-as decisions diagrams (ROBDDs \cite{BRYA86}, FDDs \cite{okfdd}),
-moment diagrams (BMDs \cite{bmd}, K*BMDs \cite{kbmd}) and their
-word-level variants \cite{WLS}. However, the decomposition principle
-behind these representations is based on (variants of) the Shannon's
-expansion, which is a bit-wise, Boolean decomposition. Such
-approaches cannot be efficiently applied to word-level abstraction of
-modulo-arithmetic circuits over Galois extension fields
-$\Fkk$. Therefore, we approach this problem from a symbolic commutative
-algebra perspective.
+{\bf Approach:} Canonical symbolic DAG 
+representations of Boolean functions, such
+as ROBDDs \cite{BRYA86}, FDDs \cite{okfdd}, BMDs \cite{bmd}, etc., are
+based on (variants of) the Shannon's  expansion, which is a bit-wise,
+Boolean decomposition. These are inapplicable to 
+word-level abstraction of modulo-arithmetic circuits over Galois
+extension fields $\Fkk$. Therefore, we approach this problem from the
+perspective of {\it symbolic computer algebra and algebraic geometry}.
 
@@
 polynomial $\F \pmod{ 2^k}$.  In commutative algebra, identification of
-such polynomial functions is an important problem: i.e.  given a function
+such polynomial functions is an important problem: given a function
 $f: \Z_n \rightarrow \Z_n$, $n \in {\mathbb{N}}$, identify if $f$ has
@@
 %of a polynomial $\F$ that represents $f$.  
-The papers
-\cite{singmaster} \cite{chen_95} \cite{chen_96} addressed such
-problems in number theory and algebra. {\it
- Shekhar} \cite{shekhar:phd} also addressed RTL verification of
-polynomial datapaths over $k$-bit vectors using the above
-concepts. However, in her work, the RTL datapath was already known to
+The work of 
+\cite{singmaster} 
+%\cite{chen_95} \cite{chen_96} 
+addresses such problems in number theory and algebra. {\it
+ Shekhar et al.} \cite{shekhar:tcad07} address RTL verification
+of polynomial datapaths over $k$-bit vectors using the above
+concepts. However, in their work, the RTL datapath is already known to
 be a polynomial function $\pmod{ 2^k}$. Unfortunately, an arbitrary
@@
 %of $f$. 
-Motivated by this fundamental result, we will devise an approach
-to extract a word-level polynomial function (abstraction) from a
-circuit. Given the circuit $C$ which represents a function $f:
-\B^k \rightarrow \B^k$, we will interpret $f$ as a function $f: \Fkk
-\rightarrow \Fkk$, and derive a unique, minimal, canonical polynomial
-representation as $Y = \F(A)$ over $\Fkk$, where $Y, A$ are,
-respectively, the output and input bit-vectors of the circuit
-$C$. 
+Motivated by this fundamental result, we devise an approach
+to derive a word-level polynomial function abstraction from the
+circuit using the $f: \Fkk \rightarrow \Fkk$ polyfunction model. 
+%Given the circuit $C$ which represents a function $f:
+%\B^k \rightarrow \B^k$, we interpret $f$ as a function $f: \Fkk
+%\rightarrow \Fkk$, and derive a unique, minimal, canonical polynomial
+%representation as $Y = \F(A)$ over $\Fkk$, where $Y, A$ are,
+%respectively, the output and input bit-vectors of the circuit
+%$C$. 
 %Note, however, that we are not given a function (mapping);
@@
 %also going to be infeasible for large values of $k$. Therefore, 
-We will extract a set of polynomials (ideal) from the circuit and employ
-{\it Gr\"obner bases} techniques to abstract the canonical polynomial
+We analyze the circuit, derive a specific set of polynomials modeled
+over $\Fkk$, and compute a {\it Gr\"obner basis} of these polynomials
+with a specific term order to abstract the canonical polynomial
 representation. Our approach can be generalized to a circuit with
-arbitrary number of inputs and outputs, i.e. to analyze polynomial
-functions $f: \Fq^d \rightarrow \Fq$, $q = 2^k$, and further to $f:
+arbitrary number of inputs and outputs, i.e. to model polynomial
+functions $f: \Fkk^d \rightarrow \Fkk$, and further to $f:
 {\mathbb{F}}_{2^n} \rightarrow {\mathbb{F}}_{2^m}$.
 
-\begin{Example}
-{\bf Motivating Example:} {\it
-Consider a 2-bit multiplier circuit over ${\mathbb{F}}_{2^2}$ given in
-Fig. \ref{fig:mul2bit}, which implements a function $f: {\mathbb{F}}_4
-\times {\mathbb{F}}_4 \rightarrow {\mathbb{F}}_4$. It implements a
-{\bf polynomial function}: $Z = A \times B, ~Z, A, B \in
-{\mathbb{F}}_4$. Variables $a_0, a_1, b_0, b_1$ are primary inputs,
-$z_0, z_1$ are primary outputs, and $s_0, s_1, s_2, s_3, r_0$ are
-intermediate variables of the circuit. 
-These bit-level variables are in ${\mathbb{F}}_2 = \{0, 1\}$.
-%variables $a_0, a_1, b_0, b_1, g_0, g_1, s_0, s_1, s_2, s_3, r_0$ are
-%in ${\mathbb{F}}_{2}$. 
-Then, we can consider $A = a_0 + a_1 \alpha, ~B = b_0 +
-b_1 \alpha$ as the word-level inputs and $Z = z_0 + z_1 \alpha$ 
-as the output in ${\mathbb{F}}_4$. Here $P(x) = x^2 + x + 1$ is the primitive
-polynomial of ${\mathbb{F}}_4$ and $P(\alpha) = 0$. 
-\begin{figure}[htb]
-\centerline{
-\includegraphics[scale=0.4]{2bitmult.eps}
-}
-\caption{\small A 2-bit Multiplier over ${\mathbb{F}}_{2^2}$. The gate
-  $\otimes$ corresponds to AND-gate, i.e. bit-level multiplication
-  modulo 2. The gate $\oplus$ corresponds to XOR-gate, i.e. addition
-  modulo 2.} 
-\label{fig:mul2bit}
-\end{figure}
-
-The functionality of the entire circuit can be described using the
-following polynomials derived from the Boolean gate-level operators: 
-$f_1: z_0+z_1\alpha +Z; ~~f_2: b_0+b_1\alpha +B; ~~f_3: a_0+a_1 \alpha
-+A; ~~f_4: s_0+a_0 \cdot b_0; ~~f_5: s_1+a_0 \cdot b_1; ~~f_6:
-s_2+a_1 \cdot b_0; ~~f_7: s_3+a_1 \cdot b_1; ~~f_8: r_0+s_1 + s_2;
-~~f_9: z_0+s_0 + s_3; ~~f_{10}: z_1 + r_0 + s_3$. If we impose the
-following {\bf elimination} term order: {\bf lex term order} with
-``circuit variables'' $>$ ``Output Z'' $>$ ``Inputs, A, B''
-%More precisely, we can use {\bf lex} term order with $s_0 > s_1 >  s_2 >
-%s_3 >  r_0 > a_0 > a_1 > b_0 > b_1 > g_0 > g_1 > A > B > G$.
-%The above term order is an {\bf elimination} order. If we use the above
-and compute a Gr\"obner basis $G$ of $\{f_1, \dots, f_{10}\}$, we observe
-the following polynomials in the basis:  $g_1: z_0+z_1\alpha +Z;
-~~g_2: b_0+b_1\alpha +B; ~~g_3: a_0+a_1 \alpha +A; ~~g_4: s_3+r_0+z_1;
-g_5: s_1+s_2+r_0; ~~g_6: s_0+s_3+z_0; ~~{\mathbf{g_7: Z +  AB}}; ~~g_8:
-a_1b_1+a_1B+b_1A+z_1; ~~g_9: r_0+a_1b_1+z_1; ~~g_{10}:
-s_2+a_1b_0$. Notice that the polynomial $g_7: Z + AB$ describes $Z =
-AB$ as the (canonical) polynomial function implemented by the circuit;
-and we were able to {\it extract} this representation using the
-Gr\"obner basis computation. We wish to explore such an approach to
-derive an efficient decision procedure for word-level abstraction. 
-}
-\end{Example}
-
 %\subsection{Contributions of this Thesis}
-{\bf Contributions of this Thesis:} Our approach to solve this
-problem, and the contributions of this thesis, can be outlined as
-follows:
+{\bf Approach \& Contributions:}
 
@@
   problem can be formulated as one of {\it computing a Gr\"obner
-    basis} of the set of polynomials derived from the given circuit
+    basis} of a set of polynomials derived from the given circuit
   netlist, using a specific elimination term order $>$.
@@
 \item Computing Gr\"obner bases using elimination term orders is
-  infeasible for large circuits. To overcome this limitation, we will
+  infeasible for large circuits. To overcome this limitation, we 
   investigate the use of the FGLM algorithm \cite{fglm} to derive the
@@
   $>_1$, and transforms it into another Gr\"obner basis ($G$)
-  w.r.t. an elimination term order $>$. In the PhD thesis of {\it
-  Lv} \cite{lv:phd}, it was shown that for any given
-  circuit, there exists a term order $>_1$ that renders the set of
-  polynomials derived from the circuit itself a Gr\"obner
+  w.r.t. an elimination term order $>$. In the work of {\it
+  Lv} \cite{lv:date2012} it was shown that for any
+  given circuit, there exists a term order $>_1$ that renders the set
+  of polynomials derived from the circuit itself a Gr\"obner
   basis. Moreover, $>_1$ can be easily derived by performing a reverse
@@
   verification applications do exhibit such high complexity, as the
-  number of intermediate computations in the Buchbeger's algorithm
-  \cite{buchberger_thesis} 
+  number of intermediate computations in the Gr\"obner basis algorithm
+  (Buchbeger's algorithm) \cite{buchberger_thesis} 
   tend to explode. Therefore, we conjecture that using the FGLM
-  algorithm --- which relies on sparse linear algebra concepts ---
-  perhaps this complexity can be avoided. In this thesis, we will
-  develop a CAD framework using the above ideas and investigate the
-  scalability of our methods.
+  algorithm --- which relies on {\it sparse linear algebra concepts}
+  --- perhaps this complexity can be avoided. We are currently
+  implementing a domain specific FGLM algorithm --- specifically for
+  the abstraction problems related to Galois field
+  crypto-circuits. Experimental results with FGLM are not yet available.
 
-\item As an application, we will use our CAD approach to
-  reverse-engineer the function implemented by a given Galois field
-  arithmetic circuits used in ECC   implementations. This can be used
-  for function identification from a given circuit for formal
-  verification.
+%\item As an application, we will use our CAD approach to
+%  reverse-engineer the function implemented by a given Galois field
+%  arithmetic circuits used in ECC   implementations. This can be used
+%  for function identification from a given circuit for formal
+%  verification.
 
@@
 
-{\bf Proposal Organization:} The next section reviews previous work in
+{\bf Paper Organization:} The next section reviews previous work in
 the area of canonical representations functions, word-level
-abstractions, formal design verification using computer algebra
-techniques, and also polynomial interpolation literature in symbolic
+abstractions, 
+%formal design verification using computer algebra techniques, 
+and also polynomial interpolation literature in symbolic
 computing. Section \ref{sec:prelimgf} reviews 
-Galois field theory and hardware design. Section \ref{sec:ideals}
+Galois field theory and describes the arithmetic circuits that we wish
+to verify using our approach. Section \ref{sec:ideals}
 reviews preliminary computer-algebra concepts related to ideals,
@@
 algorithm is covered in Section \ref{sec:fglm}. Finally, Section
-\ref{sec:obj} outlines the research plan and concludes the proposal.
+\ref{sec:concl} concludes the proposal.

### 2014-04-25 sequential/latex/abstraction/iwls/prelim_gf_iwls.tex
upload files from laptop linux

\section{Galois Fields, Polynomial Functions \& Hardware Design}
\label{sec:prelimgf}
%\vspace{-0.035in}

%% We briefly describe the relevant concepts related to Galois fields
%% $\Fkk$; for more details, interested readers may refer to the textbook 
%% \cite{galois_field:mceliece}. 
%% %to put the complexity of the verification problem in perspective, 
%% %to shed some light on the complexity of our design verification, 
%% We also review some VLSI architectures used for Galois field
%% computations \cite{mastro:phd} \cite{acar:1998} \cite{wu:2002}
%% \cite{Knezevic:2008} \cite{eccld} \cite{ecc163}. In our experiments,
%% we have verified custom designs based on these architectures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% %\begin{Definition}
%% A finite field, also called a {\bf Galois field}, denoted by $F_q$, is
%% a set of $q$ elements, including two distinguished elements 0 and 1,
%% along with operations $'+'$ and $'\cdot'$ (addition and
%% multiplication) that satisfy the following properties: 
%% \begin{itemize}
%% \item ($F_q, +, 0$) forms an Abelian group.

%% % \begin{itemize}
%% % \item  Closure: $\forall a, b \in F_q, a + b \in F_q$.
%% % \item Associativity: $\forall a, b, c \in F_q, a+(b+c)=(a+b)+c$.
%% % \item Commutativity: $\forall a, b \in F_q, a+b=b+a$.
%% % \item Additive Identity: $\forall a \in F_q, a+0=a$. 
%% % \item Additive Inverse: $\forall a \in F_q$, there is an additive
%% %   inverse element $-a \in F_q$ such that $a+(-a)=0$. 
%% %\end{itemize}

%% \item ($F_q, +, \cdot, 0, 1$) forms a commutative ring with unity.
%% % \begin{itemize}
%% % \item Multiplication closure: $\forall a, b \in F_q, a \cdot b \in
%% %   F_q$.
%% % \item Multiplication Associativity: $\forall a, b, c \in F_q, a \cdot
%% %   (b \cdot c)=(a \cdot b)\cdot c$.  
%% % \item Multiplication Commutativity: $\forall a, b \in F_q, a \cdot b=b \cdot a$. 
%% % \item Multiplication Identity: $\forall a \in F_q, a\cdot 1=a$. 
%% % \item Distributivity: $\forall a, b, c \in F_q, a \cdot (b+c)=(a
%% % \cdot b)+(a \cdot c)$. 
%% % \end{itemize}

%% \item Every non-zero element in $F_q$ has a   multiplicative inverse:
%%   $\forall a \in F_q - \{0\}, \exists a^{-1} \in F_q$ such that $a \cdot
%%   a^{-1}=1$.

%% \item The number of elements $q$ of the finite field is a
%%   power of a prime integer, i.e. $q = p^k, k \geq 1, p=$ prime.

%% \end{itemize}
%% %\end{Definition}

A Galois field $\Fq$ is a field with a finite number of elements. The
number of elements $q$ of the field is a power of a prime integer ---
i.e. $q = p^k$, where $p$ is a prime integer, and $k \geq 1$
\cite{galois_field:mceliece}. 
%Galois fields are denoted as ${\mathbb{F}}_{q}$ and also $GF(q
%=p^k)$. 
We are interested in fields where $p = 2$ and $k >1$ --- i.e. {\it
  binary Galois extension   fields} $\mathbb{F}_{2^k}$ --- as they are
widely employed in hardware implementations of cryptography
primitives.

To construct ${\mathbb{F}}_{2^k}$, we take the polynomial ring
${\mathbb{F}}_2[x]$, where ${\mathbb{F}}_{2} = \{0, 1\}$, and an
irreducible  polynomial $P(x) \in {\mathbb{F}}_2[x]$ of degree $k$, and
construct ${\mathbb{F}}_{2^k}$ as ${\mathbb{F}}_2[x] \pmod{   P(x)}$. As a
result, all field operations are performed 
modulo the irreducible polynomial $P(x)$ and the coefficients are
reduced modulo $p=2$. %=2$; due to which $-1 = +1$ over $\Fkk$. 
Any element $A \in {\mathbb{F}}_{2^k}$ can be represented in polynomial
form as $A = a_0 +  a_1 \alpha + \dots + a_{k-1} \alpha^{k-1}$, where
$a_i \in {\mathbb{F}}_2, i = 0, \dots, k-1$, and $\alpha$ is the root of
the irreducible polynomial, {\it i.e.} $P(\alpha)=0$. 
%The field
%$\Fkk$ can therefore be construed as a $k$-dimensional vector space
%over ${\mathbb{F}}_2$.

%For
%example, $\mathbb{F}_{16} =  \mathbb{F}_2[x] \pmod{ x^4 + x + 1}$.

%% \debug{
%% The {\it characteristic} of any finite field with
%% unity element $1$ is the least integer $n$ such that $1 + \dots + 1$
%% {\it (n times)} $= 0$. The characteristic of fields of the type
%% ${\mathbb{F}}_{p^k}$ %{\mathbb{F}}_p[x] \pmod{ P(x)}$ 
%% is the prime integer $p$. Since in our
%% case $p = 2$, all fields of the type $\Fkk$, for any given $k$, have 
%% characteristic 2. As a result, all field operations are performed
%% modulo the irreducible polynomial $P(x)$ and the coefficients are
%% reduced modulo $p=2$; due to which $-1 = +1$ over $\Fkk$. 
%% }

%% Any element $A \in \mathbb{F}_{2^k}$ can be represented in polynomial
%% form as $A = a_0 +  a_1 \alpha + \dots + a_{k-1} \alpha^{k-1}$, where
%% $a_i \in \mathbb{F}_2, i = 0, \dots, k-1$, and $\alpha$ is the root of
%% the irreducible polynomial, {\it i.e.} $P(\alpha)=0$. The field
%% $\Fkk$ can therefore be construed as a $k$-dimensional vector space
%% over $\mathbb{F}_2$.

\begin{Example}\label{ex:1}
{\it
Let us construct ${\mathbb{F}}_{2^4}$ as ${\mathbb{F}}_2[x] \pmod{
  P(x)}$, where $P(x)=x^4+x^3+1 \in {\mathbb{F}}_2[x]$ is an
irreducible polynomial of degree $k =4$. Let $\alpha$ be the root of
$P(x)$, i.e. $P(\alpha)=0$.  Any element $A \in {\mathbb{F}}_2[x]
\pmod{ x^4 + x^3 + 1}$ has a representation of the type: $A = a_3 x^3
+ a_2 x^2 +  a_1 x + a_0$ where the coefficients $a_3, \dots, a_0$ are
in ${\mathbb{F}}_2 = \{0, 1\}$. Since there are only 16 such
polynomials, we obtain 16 elements in the field
${\mathbb{F}}_{16}$. Each element in can then be viewed as a 4-bit
vector over ${\mathbb{F}}_2$: ${\mathbb{F}}_{16}=\{(0000),(0001),
\dots (1110),(1111)\}$.  Each element also has an exponential
representation; all three representations are shown in Table
\ref{tab:gfelement}. For example, consider the element $\alpha^{12}$.
Computing $\alpha^{12} \pmod{ \alpha^4+\alpha^3+1} = \alpha + 1
= (0011)$; hence we have the three equivalent representations. 
}

\vspace{-0.1in}
\begin{table}[h]
\begin{center}
{\tiny
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
\pmod{x^4+x^3+1}$}\label{tab:gfelement}  
\begin{tabular}{|c|c|c||c|c|c|} 
\hline
$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
\hline
$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
\hline
$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
\hline
$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
\hline
$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
\hline
$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
}
\end{center}
\end{table}
\end{Example}

%%%%%%%%%
{\bf Polynomial Functions $f: \Fkk \rightarrow \Fkk$:} 
Arbitrary mappings among $k$-bit vectors can be constructed; each such
mapping generates a function $f: \B^k \rightarrow \B^k$. 
%. Since every
%$k$-bit vector can be construed as an element in $\Fkk$ (as shown in
%the above example), every such function corresponds to a function over
Every such function is also a polynomial function over Galois fields:
$f: \Fkk \rightarrow \Fkk$.

\begin{Theorem}
From \cite{ff:1997}: 
Any  function $f: \Fq \to \Fq$ is a polynomial function
over $\Fq$, that is there exists a polynomial $\F \in \Fq [x]$ such
  that $f(a) = \F(a)$, for all $a \in \Fq$. 
\end{Theorem}

By analyzing $f$ over each of the $q$ points, one can apply
{\bf Langrange's interpolation formula} and interpolate a polynomial
$\F(x) = \sum_{k=1} ^q  \frac{ \prod_{i \neq k}  (x -x_i)}{\prod_{i \neq k}
  (x_k -x_i)} \cdot f(x_k)$, which is a polynomial of degree at 
most $q-1$ in $x$. One can easily see that $\F(a)=f(a)$ for all $a \in
\Fq$, and $\F(x)$ is therefore the polynomial function representation
of $f$.

%% \begin{Example} {\it
%% Let $A = \{a_2, a_1, a_0\}$ and $Y = \{y_2, y_1,
%% y_0\}$ be 3-bit vectors.  Consider the function $Y[2:0] = A[2:0]
%% >> 1$, i.e. a {\bf bit-vector right shift} operation on $A$. 
%% The function maps as follows:

%% \begin{center}
%% {\small
%% \begin{tabular}{c|ccc|c||c|ccc|c}
%% $\{a_2a_1a_0\}$  & $A$ &$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$ &
%%   $\{a_2a_1a_0\}$  &$A$&$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$\\
%% \hline
%% 000  &0 &$\rightarrow$& 000 & 0 & 100  &$\alpha^2$ &$\rightarrow$& 010 &  $\alpha$ \\
%% 001  &1 &$\rightarrow$& 000 & 0 & 101  &$\alpha^2 + 1$ &$\rightarrow$&010 & $\alpha$ \\
%% 010  &$\alpha$ & $\rightarrow$ & 001& 1 & 110  &$\alpha^2 + \alpha$&$\rightarrow$& 011 &$\alpha + 1$ \\   
%% 011  &$\alpha + 1$ &$\rightarrow$& 001 &1 & 111& $\alpha^2 + \alpha + 1$ &$\rightarrow$& 011 &$\alpha + 1$\\
%% \hline
%% \end {tabular}
%% }
%% \end{center}

%% If we model this function as $f: {\mathbb{Z}}_{8} \rightarrow
%% {\mathbb{Z}}_{8}$, then using the results of \cite{singmaster}
%% \cite{chen_95} \cite{chen_96}, we deduce that this function is not a
%% polynomial function over ${\mathbb{Z}}_{8}$. However, applying
%% Largange's interpolation formula over ${\mathbb{F}}_{2^3}$, we obtain
%% the following polynomial function representation, $Y =
%% (\alpha^2+1)A^4+(\alpha^2+1)A^2$, where $P(\alpha) = \alpha^3 +
%% \alpha + 1 = 0$. 
%% }
%% \end{Example}

An important property of Galois fields is that for all elements $A \in
{\mathbb{F}}_q, A^q = A$, and hence $A^q - A =  0.$ Therefore, the
polynomial $x^q -x$ {\it vanishes} on all points in
$\Fq$. Consequently, any polynomial $\F(X)$ can be reduced $\pmod{ X^q
  - X}$ to obtain a canonical representation $\F(X) \pmod{X^q - X}$
with degree at most $q-1$. 
%Such {\it vanishing polynomials} will form
%an important part of our formulation.

\begin{Definition}
Any function $f: \Fq^d \to \Fq$ has a unique canonical representation
(UCR) as polynomial $\F \in \Fq[x_1,\dots, x_d]$ such that all its 
nonzero monomials are of the form $x_1^{i_1}\cdots x_d^{i_d}$ where $0
\leq i_j \leq q-1$, for all $j=1, \ldots d$.
\end{Definition}

\subsection{Hardware Implementations of Galois Field Arithmetic}
% Operations Over $\mathbb{F}_{2^k}$}

{\bf Point Addition over Elliptic Curves:} The main operations of
encryption, decryption and authentication in elliptic curve
cryptography (ECC) rely on {\it point additions} and {\it doubling}
operations on elliptic curves designed over Galois fields. 
%Point multiplication involves a series of addition and doubling of
%points on the elliptic curve. % -- this, in turn, requires finite
%field multiplications.  
%A drawback of traditional approaches is that they require
%point multiplication is that each point addition and doubling 
In general, this requires computation of multiplicative inverses over
the field - which is expensive. Modern approaches represent
the points in projective coordinate systems, {\it e.g.}, the
L$\acute{o}$pez-Dahab (LD) projective coordinate \cite{eccld},
%\cite{ecc:software},
which eliminates the need for multiplicative inverses and improves the
efficiency of these operations.

%%%

%% \begin{Example}
%% {\it 
%% Consider point addition in L$\acute{o}$pez-Dahab (LD) projective
%% coordinate. Given an elliptic curve: $Y^2 + XYZ = X^3Z + aX^2Z^2 +
%% bZ^4$ over $\mathbb{F}_{2^k}$,   where $X,Y,Z$ are $k$-bit vectors
%% that are elements in $\mathbb{F}_{2^k}$ and similarly, $a, b$ are
%% constants from the field.   Let ($X_3$, $Y_3$, $Z_3$) = ($X_1$, $Y_1$,
%% $Z_1$) + ($X_2$, $Y_2$, $1$)  represent point addition over the
%% elliptic curve.  Then $X_3$, $Y_3$, $Z_3$ can be computed as follows:

%% \begin{align*}
%% A &= Y_2 \cdot Z_1^2 + Y_1 \\
%% B &= X_2 \cdot Z_1 + X_1 \\
%% C &= Z_1 \cdot B \\
%% D &= B^2 \cdot(C + a Z_1^2) \\
%% Z_3 &= C^2 \\
%% E &= A \cdot C  \\
%% X_3 &= A^2 + D + E  \\
%% F &= X_3 + X_2 \cdot Z_3 \\
%% G &= X_3 + Y_2\cdot Z_3 \\
%% Y_3 &= E\cdot F + Z_3 \cdot G \\
%% \end{align*}
%% }
%% \end{Example}

\begin{Example}
{\it 
Consider point doubling in LD projective coordinate system. Given an
elliptic curve: $Y^2 + XYZ = X^3Z + aX^2Z^2 + bZ^4$.  
Let  ($X_3$, $Y_3$, $Z_3$) = 2($X_1$, $Y_1$, $Z_1$), then $X_3, Y_3,
Z_3$ can be computed as: $X_3 = X_1^4 + b \cdot Z_1^4; ~~Z_3 = X_1^2
\cdot Z_1^2; ~~Y_3 = b Z_1^4 \cdot Z_3 + X_3 \cdot (aZ_3 + Y_1^2 +
bZ_1^4 )$. 
}
\end{Example}

The multiplication and iterative squaring operations in the above
computation are usually implemented using custom-designed Galois field
multipliers, such as the Mastrovito \cite{mastro:1989}, Montgomery
\cite{acar:1998}, Barrett multipliers \cite{Knezevic:2008}, or
composite-field multipliers \cite{cfmulti:1996} --- which are, in
turn, hierarchically designed.  
For example, Montgomery reduction (MR)
computes: $MR(A,B)=A\cdot B \cdot R^{-1} \pmod {P(x)}$, 
where $A,B$ are $k$-bit inputs, $R$ is suitably chosen as
$R={\alpha}^k$, $R^{-1}$ is multiplicative inverse of $R$ in
${\mathbb{F}}_{2^k}$, and $P(x)$ is the irreducible polynomial.
Since Montgomery reduction cannot directly compute $A\cdot B \pmod
{P(x)}$, we need to pre-compute $A\cdot R$ and $B\cdot R$, as shown in
Figure \ref{fig:mm4}.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{../new_mmcircuit.eps}
	\end{center}
	\caption{{\it Montgomery} multiplication over $\mathbb{F}_{2^k}$
          using four Montgomery reductions.}
	\label{fig:mm4}
\end{figure}

Given a hierarchically designed Montgomery multiplier, we will first
extract polynomials $AR, BR, ABR$ from the sub-circuit blocks. By
analyzing the interconnection of these sub-circuits at word-level, we
can then apply our approach at a higher-level, to extract the function
of the entire circuit. Performing such operations hierarchically, we
will apply our approach to reverse-engineer point-addition circuits
designed using a variety of such Galois field adder and multiplier
circuits.

### 2014-04-25 sequential/latex/abstraction/iwls/prev_iwls.tex
upload files from laptop linux

\section{Review of Previous Work}
\label{sec:prev}

{\bf Canonical Decision Diagrams:} 
The Reduced Ordered Binary Decision Diagram
(ROBBD) \cite{BRYA86}, based on the Shannon's decomposition, was the
first significant contribution in this area. 
Motivated by the success of BDDs, 
variants of the Shannon's decomposition principle (Davio,
Reed-Muller, etc.) were explored to develop other functional decision
diagrams: FDDs \cite{okfdd}, ADDs \cite{add}, %MTBDDs\cite{mtbdd},  
HDDs \cite{hdd} and EVBDDs \cite{evbdd}. 
%While these are %{\it curiously} still 
%referred to as {\it Word-Level Decision Diagrams} \cite{WLS}, 
Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivative K*BMDs
\cite{kbmd}, % and *PHDDs \cite{phdd}, 
depart from the Boolean
decomposition and perform the decomposition of a {\it linear} function
based on its two moments. The decomposition is still point-wise,
binary, w.r.t. each Boolean variable; these do not
serve the purpose of word-level abstraction from bit-level
representations.

%% BMDs provide a compact representation for
%% integer arithmetic circuits such as multipliers and squarers. However,
%% these are inapplicable for modulo-arithmetic circuits over Galois fields.
%in our application, we encounter modulo-arithmetic circuits and that
%too over Galois fields. For such applications, these representations
%are inapplicable. Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp}
%generalize the linear moment decomposition of BMDs into a Taylor
%series expansion --- this allows the representation of RTL
%descriptions as polynomials over bit-vectors. However, 
Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp} are a word-level
canonical representation of a {\it polynomial expression}, but they do
not represent a {\it polynomial function} canonically. 
%For example, $f_1 = 0$ and $f_2 = 2x^2 - 2x
%\pmod{ 4}$ are two different polynomial representations of the zero
%function over $\Z_4$; but they are symbolically different
%polynomials and they have different (non-isomorphic) TED DAGs. 
While \cite{shekhar:tcad07}  and \cite{alizadeh:tcad2010} provide
canonical representations of polynomial functions, they do so over
$\Z_{2^k}$ and not over $\Fkk$.
%finite integer rings and not over Galois fields.

MODDs \cite{modd} \cite{modd_tcomp} are a DAG representation of the
characteristic function of a circuit over Galois fields $\Fkk$. MODDs
come very close to satisfying our requirements as a canonical
word-level representation that can be employed over Galois fields, as
it essentially interpolates a polynomial from the characteristic 
function. However, MODDs do not scale very well for large circuits ---
this is because every node in the DAG can have up to $k$ children and
the normalization operations are very complicated for MODDs. They 
suffer from the size explosion problem during intermediate
computations. They are known to be infeasible beyond
32-bit circuits.

%In the realm of theorem-proving and SMT-solving, many automated
%decision procedures have been devised that can decide
%satisfiability/validity of word-level formulas represented by a
%combination of theories, but they do not provide a canonical
%representation of the function implemented by a ``bit-blasted''
%circuit.

{\bf Verification of Galois field circuits:} 
%% Symbolic computer algebra techniques have been employed for formal
%% verification of circuits over $\Z_{2^k}$ and also over $\Fkk$. 
%% %% The work of \cite{gao:gf-gb-ms} shows how to use Gr\"obner basis
%% %% techniques to count the zeros of an ideal $J$ over ${\mathbb{F}}_q$
%% %% (i.e. count $V_{{\mathbb{F}}_q}(J)$). The authors then follow-up with
%% %% an approach for {\it quantifier elimination} over Galois fields
%% %% ${\mathbb{F}}_q$ \cite{gao:qe-gf-gb}. While the above works present
%% %% the theory and algorithms, efficiency/improvements in Gr\"obner basis
%% %% computation is not addressed. So is also the case with other general
%% %% verification techniques using Gr\"obner bases \cite{Avrunin:CAV}
%% %% \cite{gbverify:2007} \cite{manna:program}, etc. 
%% The paper \cite{wienand:cav08} addresses verification of
%% finite-precision integer datapath circuits using the concepts of
%% Gr\"obner bases over the ring ${\mathbb{Z}}_{2^k}$. 
%% %They model the
%% %circuit constraints by way of arithmetic-bit-level (ABL) polynomials
%% %($\{G\}$) and formulate the verification test as an equivalent
%% %{\it variety subset problem}. To solve this, first they derive a term
%% %order to represent the circuit polynomials $\{G\}$ with a Gr\"obner
%% %basis. Then they compute a normal form $f$ of the specification $g$
%% %w.r.t. $\{G\}$. They show that, under certain constraints that are
%% %satisfied by their term order, the circuit is correct if and only if
%% %$f$ is a vanishing polynomial over ${\mathbb{Z}}_{2^k}$
%% %\cite{shekhar:tcad07}. 
%% In \cite{wedler:date11}, the authors further
%% show an integration of their approach within an SMT-solver for
%% arithmetic instances.  
%% %show that  the vanishing polynomial test can be
%% %omitted by formulating the problem directly over $Q :=
%% %{\mathbb{Z}}_{2^k} [X]/\langle x^2-x : x \in X \rangle$. 
In %\cite{ibm:ecc-ita11} \cite{blueveri:ita} 
\cite{ibm:blueveri},
%from IBM needs special mention: T
the authors present the {\sc BLUEVERI} tool from IBM for verification
of Galois field circuits for error correcting codes against an
algorithmic spec. The implementation consists of a set of
(pre-designed and verified) circuit blocks that are interconnected to
form the error correcting system. The spec is given as a set of design
constraints on a ``check file''. Their objective is to prove the 
equivalence of the implementation against this check file,
%They model
%the verification instance as a data-flow graph, represent each
%sub-circuit block with its known (word-level) polynomial over $\Fq$,
%and formulate the verification problem using the {\it Weak
%  Nullstellensatz} --- i.e. to check if the {\it variety} of the
%algebraic system ``{\it spec} $\neq$ {\it implementation}'' is empty
for which they employ a Nullstellensatz formulation.
%for which they use a Gr\"obner basis engine. 
%Their main
%contributions are: i) a ``term re-writing'' to specify the algorithmic
%description using polynomials (ideal); and ii) integrating an
%AIG-style \cite{AIG:2002} Boolean solver with their word-level
%decision procedure, with lazy signal computations and Boolean
%reasoning. 
For final verification, the polynomial system is given
to a computer algebra tool ({\sc Singular} \cite{DGPS}) to {\it
  compute} a reduced Gr\"obner basis. However, improvements to the
core Gr\"obner basis computational engine are not the subject of their
work.

In %\cite{lv:phd} 
\cite{lv:date2012} 
%\cite{lv:hldvt2011} \cite{lv:vlsi2012}, 
{\it Lv et al.} present computer algebra
techniques for formal verification of Galois field arithmetic
circuits. Given a specification polynomial $f$, and a circuit $C$,
they formulate the verification problems as an ideal membership test
using the Strong Nullstellensatz and Gr\"obner bases. In \cite{lv:date2012}, the
authors show that for any combinational circuit, {\it there exists
a term order $>_1$ that renders the set of polynomials of the circuit
itself a Gr\"obner basis --- and this term order can be easily derived
by performing a topological traversal of the circuit}. By exploiting
this term order, verification can be significantly scaled
to $163$-bit (NIST-specified) cryptography circuits. 
In contrast to the work of \cite{lv:date2012}, we are not given a
specification polynomial. Given the circuit $C$, we want
to derive (extract) the word-level specification $f$. 
In this work, we build upon the results of
%\cite{gao:gf-gb-ms} \cite{gao:qe-gf-gb} %\cite{modd_tcomp} 
\cite{lv:date2012}.

{\bf Polynomial Interpolation in Symbolic Computation:}
The problem of polynomial interpolation is a fundamental problem in
symbolic and algebraic computing which finds application in modular
algorithms, such as the GCD computation and polynomial
factorization. The problem is stated as follows: Given $n$ distinct
data points $x_1, \dots, x_n$, and their evaluations at these points
$y_1, \dots, y_n$,  {\it interpolate} a polynomial $\F(X)$ of degree
$n-1$ (or less) such that $\F(x_i) = y_i$ for $1 \leq i \leq n$. 
Let $t$ be the number of non-zero terms in $\F$ and let $T$ be the
total number of possible terms. When ${{t} \over {T}} << 1$, the
polynomial $\F$ is {\it sparse}, otherwise it is {\it dense}. Much of
the work in polynomial interpolation addresses sparse 
interpolation using the ``black-box'' model (also called the
algebraic circuit model) as shown in Fig. \ref{fig:blackbox}.

\begin{figure}[hbt]
\centerline{
%\includegraphics[scale=0.3]{../figures/2bitmultiplier.eps}
\includegraphics[scale=0.35]{../blackbox.eps}
}
\caption{The black-box or the algebraic circuit representation.}
\label{fig:blackbox}
\end{figure}

Let $\F$ be a multivariate polynomial in $n$ variables $\{x_1, \dots,
x_n\}$, with $t$ non-zero terms ($0 < t < T$), represented with a
black-box $B$. On input $(x_1, \dots, x_n)$,
the black-box evaluates $y_i = \F(x_1, \dots, x_n)$. Given also a
degree bound $d$ on $\F$, the goal is to interpolate the polynomial
$\F$ with a minimum number of {\it probes} to the black-box. The early
work of Zippel \cite{zippel:interpolate} and
Ben-Or/Tiwari \cite{ben-or-tiwari:interpolate} require $O(ndt)$ and
$O(T \log n)$ probes, respectively, to the black-box. These bounds
have since been improved significantly; the recent algorithm of
\cite{monagan:interpolate} interpolates with $O(nt)$ probes.

Our problem falls into the category of {\it dense} interpolation, as
we require a polynomial that describes the function at each of the $q$
points of the field $\Fq$. Newton's dense interpolation technique,
with the black-box model, bounds the number of probes by $(d+1)^n$ ---
which  exhibits very high complexity. In the logic synthesis area, the
work of \cite{zilic:interpolate} investigates dense interpolation. Due
to the inherent high-complexity, their approach is feasible only for
applications over small fields, {\it e.g.} computing  Reed-Muller
forms for multi-valued logic.

We can also employ the black-box model by replacing the black-box
(algebraic circuit) by the given circuit $C$; then every {\it probe}
of the black-box would correspond to a {\it simulation of the
  circuit}. As we desire a polynomial representation of the entire
function, exhaustive simulation would be required, which is
infeasible. Therefore, we propose a {\it symbolic approach} to
polynomial interpolation from a circuit using the Gr\"obner basis
computation. 
%It can be shown that the computational complexity of
%computing a Gr\"obner basis for our problem, in the worst case, is
%$q^{O(n)}$, where $q = 2^k$, $k$ is the datapath width and $n$ is the
%number of variables in the circuit.

### 2014-04-25 sequential/latex/abstraction/iwls/xiaojun.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/DAC2013/xiaojun.tex
@@
 ideal $I = \{f_0, f_1, \dots, f_k\}$. The varieties of this ideal are possible 
-assignments of circuit variables. Similarly for $n$ bits world-level circuit 
+assignments of circuit variable. Similarly for $n$ bits world-level circuit 
 variables, the varieties are elements from $\Fkk$. Thus it's convenient to use
-ideals and varieties to describle circuit variables. \\
+ideals and varieties to describle circuit variables. \par
 
 \begin{Definition}
-({\bf Sum of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the 
-{\bf sum} of $I$ and $J$, denoted by $I + J$, is the set
+({\bf Sum of Ideals}) If I and J are ideals in $k[x_1, \dots, x_n]$, then the 
+{\bf sum} of I and J, denoted I + J, is the set
   \begin{equation}
-  I + J = \{f + g : f \in I \ and\  g \in J\}.
+  I + J = \{f + g : f \in I and g \in J\}.
   \end{equation}
@@
 \begin{Definition}
-({\bf Product of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the
-{\bf product} of $I$ and $J$, denoted by $I \cdot J$, is defined to be the ideal generated 
+({\bf Product of Ideals}) If I and J are ideals in $k[x_1, \dots, x_n]$, then the
+{\bf product} of I and J, denoted $I \cdot J$, is defined to be the ideal generated 
 by all polynomials $f \cdot g$ where $f \in I$ and $g \in J$. Furthermore, let
@@
 \begin{Definition}
-({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then $I:J$
+({\bf Quotient of Ideals}) If I and J are ideals in $k[x_1, \dots, x_n]$, then I:J
 is the set
   \begin{equation}
-  \{f \in k[x_1, \dots, x_n] : fg \in I, \forall g \in J\}
+  \{f \in k[x_1, \dots, x_n] : fg \in I \forall g \in J\}
   \end{equation}
-and is called the {\bf ideal quotient} of $I$ by $J$.
+and is called the {\bf ideal quotient} of I by J.
 \end{Definition}
 
-According to definitions above, it is possible to calculate ideal sum, product and 
-quotient based on ideal generators. Additional conclusion about union, 
-intersection and complement of varieties can be attained below: \\
+According to definitions above, it's possible to calculate ideal sum, product and 
+quotient based on ideal generators. Additionally conclusion about union, 
+intersection and complement of varieties can be attained below: \par
 
@@
 \label{thm:unionintersect}
-If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
+If I and J are ideals in $k[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
 \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) = {\bf V}(I) \bigcup {\bf V}(J)$.
@@
 \label{thm:quotient}
-Let $V$ and $W$ be varieties in $\Fkk$. Then ${\bf I}(V) : {\bf I}(W) = 
+Let V and W be varieties in $\Fkk$. Then ${\bf I}(V) : {\bf I}(W) = 
 {\bf I}(V - W)$.
@@
 
-For \ref{thm:quotient}, let $V = \Fkk = {\bf V}(\langle vanishing \  polynomials
-\rangle)$, then $V - W$ is the \textbf{complementary set} of variety $W$.\\
+For \ref{thm:quotient}, let $V = \Fkk = {\bf V}(\langle vanishing polynomials
+\rangle)$, then $V - W$ is the complementary set of variety $W$.\\
 
@@
 \label{sec:fsm}
-\subsection{Introduction to Finite State Machine:}
-Borrow from somewhere.
-\subsection{Symbolic Image Computation:}
-Borrow from somewhere.
-\subsection{Implicit State Enumeration:}
-BFS algorithm below:
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
- \KwIn{Transition functions $\Delta$, initial state $S^0$}
-
-  $from^0 = reached = S^0$\;
-  \Repeat{$new^i == 0$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$Img$(\Delta, from^{i-1})$\;
-	$new^i \gets to^i \cdot \overline{reached}$\;
-  	$reached \gets reached + new^i$\;
-	$from^i \gets new^i$\;
-  }
-\Return{$reached$}
-\caption {Breadth-first Traversal Algorithm}\label{alg:gb}
-\end{algorithm}
+{\bf Introduction to Finite State Machine:}\par
+{\bf Symbolic Image Computation:}\par
+{\bf Implicit State Enumeration:}\par
 
@@
 
-The following mapping from Boolean operations to Galois field $\mathbb{F}_2$ 
+The following mapping from Boolean operations to Gal\"ois field $\mathbb{F}_2$ 
 operations is straightforward. Assuming $a$ and $b$ are both bit-level variables:
 \begin{itemize}
-  \item $\overline{a} \rightarrow 1 + a$
-  \item $a\ and\ b \rightarrow ab$
-  \item $a\ or\ b \rightarrow a + b + ab$
+  \item $\bar{a} \rightarrow 1 + a$
+  \item $a and b \rightarrow ab$
+  \item $a or b \rightarrow a + b + ab$
 \end{itemize}
@@
 approach from \ref{sec:theory}. Construct an elimination ideal with following 
-polynomials: $f_1: 1 + XNOR(t_0, OR(p\cdot q, (1 + x_i)\cdot(1 + p)\cdot(1 + q))); 
+polynomials: $f_1: 1 + XNOR(t0, OR(p\cdot q, (1+xi)\cdot(1+p)\cdot(1+q))); 
 f_2: 1 + XNOR(t1, OR(xi\cdot(1+p),p\cdot(1+q))); f_3: S - p - q\cdot\alpha; 
-f_4: T - t_0 - t_1\cdot \alpha$ and vanishing polynomials $f_5: p^2 - p; 
-f_6: q^2 - q; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: x_i^2 - x_i; 
+f_4: T - t0 - t1\cdot\alpha$ and vanishing polynomials $f_5: p^2 - p; 
+f_6: q^2 - q; f_7: t0^2 - t0; f_8: t1^2 - t1; f_9: S^4 - S; f_{10}: xi^2 - xi; 
 f_{11}: T^4 - T$, where $S$ and $T$ are 2-bit word representing current state
@@
 is $f_{12}: S + 1 + \alpha$. According to elimination theorem \ref{thm:elim},
-let the lex ordering be $\{all\ others\} > T$, calculate Gr\"obner basis, the 
+let the lex ordering be $\{all others\} > T$, calculate Gr\"obner basis, the 
 result must contain a polynomial with only one variable $T$, which can indicate

### 2014-04-25 sequential/latex/abstraction/ken.tex
upload files from laptop linux

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  IEEEsample.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%    More information: see the header of IEEEtran.sty
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%

\documentclass[11pt,twoside, onecolumn]{IEEEtran}
%\documentclass[conference]{IEEEtran}

%%%\IEEEoverridecommandlockouts

\usepackage[ruled]{./algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%% \RequirePackage{times}
%% \RequirePackage{algorithmic}
%% \PassOptionsToPackage{boxed}{algorithm}
%% \RequirePackage{algorithm}
%% \RequirePackage{multicol}
%\renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
%\renewcommand{\algorithmicensure}{\textbf{Outputs:}}
%\DeclareMathAlphabet{\mathtsl}{OT1}{ptm}{m}{sl}

%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em1
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{example}{Example}
%\newtheorem{corollary}{Corollary}

\RequirePackage{amssymb, mathptm}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{comment}

%%indent in algorithm

%\setcounter{page}{1}

% New command for the table notes.
\def\tabnote#1{{\small{#1}}}

% New command for the line spacing.
\newcommand{\ls}[1]
    {\dimen0=\fontdimen6\the\font
     \lineskip=#1\dimen0
     \advance\lineskip.5\fontdimen5\the\font
     \advance\lineskip-\dimen0
     \lineskiplimit=.9\lineskip
     \baselineskip=\lineskip
     \advance\baselineskip\dimen0
     \normallineskip\lineskip
     \normallineskiplimit\lineskiplimit
     \normalbaselineskip\baselineskip
     \ignorespaces
    }
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqarr}{\begin{eqnarray}}
\newcommand{\eeqarr}{\end{eqnarray}}
%\newcommand{\ov}{\overline}
\newcommand{\ov}{\bar}
\newcommand{\xor}{\bigoplus}
\newcommand{\Fm}{{\mathbb{F}}}

%the following is for space before and after align or other equation environment.

%%
\newtheorem{Algorithm}{Algorithm}[section]
\newtheorem{Definition}{Definition}[section]
\newtheorem{Example}{Example}[section]
\newtheorem{Proposition}{Proposition}[section]
\newtheorem{Lemma}{Lemma}[section]
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}{Corollary}[section]
\newtheorem{Conjecture}{Conjecture}[section]
\newtheorem{Problem}{Problem}[section]
\newtheorem{Notation}{Notation}[section]
\newtheorem{Setup}{Problem Setup}[section]
%%%

%%set spacing between table columns
\setlength{\tabcolsep}{3pt}

\begin{document}

%\thispagestyle{empty}
%\pagestyle{empty}

\ls{1.1}

\title{\large{\textsc{Report on Arithmetic Circuits Verification}}}
\author{Xiaojun Sun\\
A Technical Report\\
in partial fulfillment of requirements for PhD qualification\\
of Computer Engineering Program, University of Utah\\
Fall Semester 2013
}

%%\author{\IEEEauthorblockN{Jinpeng Lv and Priyank Kalla}\thanks{This work is sponsored in part by a grant from NSF \#CCF-546859.}
%\IEEEauthorblockA{Department of  Electrical and Computer Eng.\\
% University of Utah, Salt Lake City, UT-84112 \vspace{-0.3in}
 %\{lv, kalla\}@eng.utah.edu
% }
%\and
%\IEEEauthorblockN{Florian Enescu} \thanks{\normalsize  978-3-9810801-8-6/DATE12/$\copyright 2012$ EDAA}
%\IEEEauthorblockA{Department of Mathematics and Statistics\\
% Georgia State University,  Atlanta, GA 30302-4038 \vspace{-0.3in}
% fenescu@mathstat.gsu.edu
%} 
%

\maketitle

%\markboth{MS Proposal by Tim Pruss}{}
\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
\newcommand{\B}{{\mathbb{B}}}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\F}{{\mathbb{F}}}
\newcommand{\G}{{\mathcal{G}}}
\newcommand{\R}{\mathbb{R}}
%%%

\newcommand{\debug}[1]{\textcolor{gray}{[ #1 ]}}

%\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%% Include your files here %%%%%%%%%%%%%%%%%%%%%
\section{Arithmetic Hardware Implementation}
\subsection{Speed Improvement of Arithmetic Hardware Implementation}
The hardware implementations of arithmetic computations can be classified to 2 types:
one is fully-customized ASIC chips (ASIC method), it requires designing on lower level such as 
layouts and full fabrication process in foundries. The other one is all sorts of programmable
logic devices (PLD), including programmable logic array (PLA), complex programmable logic device 
(CPLD) and field programmable gate array (FPGA). In general case, hardware implementation 
receives inputs in the form of signals directly from environment or from other block/devices,
and give out outputs also in the form of signals.

The software implementation is to describe this computation in some programming languages.
Codes are compiled and interpreted into instructions of machine language,
then fed into a general-purpose CPU for solutions.

When talking about advantages to implement arithmetic computations into hardware than in software,
the improvement on computing speed is the most salient. The following part reasons from 2 aspects
why hardware implementation usually has superiority on speed.

(1) Highly customized logic design in hardware implementations can reach a simpler structure
than CPU, which means for the same computation less instruction will be executed and less
gates will be used.

\begin{Example}
\label{ex:4adder}
Assume an addition in finite field $\F_2^4$ needs to be executed. Given boolean vector operands
$A = \{a_3a_2a_1a_0\}$, $B = \{b_3b_2b_1b_0\}$, problem is to calculate the output 
$Z = \{z_3z_2z_1z_0\}$. In $\F_2^4$, addition $1 + 1 = 0$ on each bit and there is no carry.
From this property, a hardware implementation is showed in fig.\ref{fig:4adder}.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{4adder.png}
	\end{center}
	\caption{A 4-bit adder gate level implementation over $\F_2^4$[1]}
	\label{fig:4adder}
\end{figure}

This is a very simple implementation. However, when executing this addition in CPU,
at least a 4-bit addition and a shift operation (for mod $2^4$) will be executed.
Under the same fabrication condition, the hardware implementation cost less time.
\end{Example}

Another advantage can be learnt from above example is the capability of parallel computing.
In this example, each bit during addition is independent from each other. Based on this 
knowledge, the hardware implementation is designed to compute each bit in parallel.
However, a general-purpose CPU must take ordinary sequential computing into consideration
for most bit-dependent additions, so the efficiency of CPU is usually inferior than customized
hardware implementations.

(2) The standard process for CPU to execute an instruction contains some redundant stages.

In example \ref{ex:4adder}, assume this addition is implemented in software. 
If an interpreter is very smart to deduce a 4-parallel-XOR algorithm for this addition 
and the CPU has 4 XOR gates in ALU to execute XOR operation in parallel, can this software 
implementation compete with hardware implementation? The answer is no, because in a combinational
circuit hardware implementation, the input can be read immediately; while in CPU,
there are at least an instruction fetch stage and a decoding stage lagging the whole process.

In conclusion, one advantage of adopting hardware implementations is the higher speed comparing to
software implementations.

\subsection{Other advantages on Hardware implementations}
Besides speed improvement, hardware implementations also have other advantages.
One is less energy consumption. Since there is less time for a customized hardware
design to execute 1 instruction, under the same work load its energy consumption
will be lower than CPU if they work with same power.

In practice, a hardware design may require lower voltage to work, so the power
also could be reduced.

Another advantage is the quality (accuracy) of results. For example, before floating-point unit
(FPU) was embedded into general-purpose CPU, floating-point operations are emulated by CPU 
using a series of simpler fixed-point arithmetic operations that run on integer ALU, the 
accuracy of floating-point computations are limited. Intel 80287 then is an arithmetic hardware
implementation customized for floating-point computations, which is efficient on IEEE 754 compliant
high accuracy floating-point computations. Hardware implementations can be customized to reach 
a higher accuracy which may exceed the CPU limit.

\subsection{Disadvantages on Hardware Implementation}
First disadvantage is loss of convenience. Describing the computation or algorithm in high-level
programming language usually cost relatively short time, while designing layouts and fabrication
cost half to several years in most cases.

Another disadvantage is money cost. A PC/server with general-purpose CPU is standard facility in most labs and 
companies, so there is no extra money cost in software implementations. However, PLDs
and equipments to program PLDs are a fixed expense for hardware implementations. If
the computation is implemented in fully customized design chips, the fabrication charges even
more. I used to send a batch (20 pcs) of 2-inch silicon wafers for simple ion implantations,
the vender charged me more than \$800 for a single process. A set of masks for photolithography
with low resolution ($> 1 \mu m$) pattern costs thousands of dollars. Money is an important 
factor that needs to be taken into consideration.

\subsection{Discussion}
Lots of factors affects the decision to choose hardware implementation or software implementation. 
A balance need to be achieved between disadvantages and advantages while considering their weights
on the decision problem.

From my point of view, situations to choose hardware implementations includes:\par
(1) Speed/energy consumption/accuracy is greatly improved and they are key factors.
Noticing this point, the FPU was integrated with CPU by Intel and other chip manufacturers.
In labs, when CPU performance is bottleneck in computing, it is also worth
to get a hardware implementation for certain computations/algorithms.\par
(2) If it has to be implemented in hardware. Many computations/algorithms are supposed to be
applied on circuits, such as a secure ID.

 \begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{ID.png}
	\end{center}
	\caption{A secure ID based on RSA}
	\label{fig:ID}
\end{figure}

\section{Verification of Arithmetic Circuits}
\subsection{Difficulties on Verification}
Control logic is usually implemented through synthesis result of some state graphs.
Each state graph has limited number of states, after state assignment the bit-width
of control signals is relatively small. On the other hand, arithmetic units deal with
numbers (integers, floating-point numbers or polynomials) at word level. Since the computing
logic is not complicated in most case (such as a multiplier), arithmetic unit is easy to
synthesize,i.e. mapping word-level arithmetic function to bit-level boolean functions (gate implementations).
So designer's job is to design an arithmetic unit has the same size required by application.
Unfortunately, the application's request usually covers large-size or high-accuracy input data;
this lead to a wide datapath connecting to arithmetic unit with lots of bits.
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.7]{FIR.png}
	\end{center}
	\caption{Comparison between control logic and datapath[2]}
	\label{fig:FIR}
\end{figure}
Fig.\ref{fig:FIR} shows block diagram of order-n FIR filter. It is clear to see there are only
1 bit input and 5 1-bit outputs for the control logic (FSM), but the datapath have 2 $n$-bits
input, $n$ could be a number much larger than 10.

For example, IEEE 754 compliant double precision floating point (round to 0) multiplier has
2 input operands, each of them has 64 bits:

 \begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{double.png}
	\end{center}
	\caption{Double precision floating-point number in IEEE 754[5]}
	\label{fig:double}
\end{figure}

including 1 sign bit, 11 exponent bits and 52 fraction bits. In total there are 128 bits input,
to exhaustively check all possible inputs, $2^{128}$ test vectors are needed. This is difficult 
to enumerate.

\subsection{Approaches for arithmetic hardware verification}
Verification techniques includes 2 basic types: equivalence checking (EC) and property checking
(PC). Equivalence checking does not suit arithmetic circuit verification very well, because it 
relies on the bit-level miter circuit output, so could not avoid the "bit blasting" issue;
Property checking fits arithmetic verification well because the computational function itself is 
the word-level property need to be checked. To check the property in word-level rather than
bit-level, word-level abstraction techniques are necessary. So in all kind of verification
techniques, those belong to property and can solve the abstraction problem are best techniques
for arithmetic hardware verification.

There are several typical techniques used to address arithmetic circuit verification.
First one is a BDD-like canonical graph-based diagram called Binary Moment Diagram (BMD). [8]
BMDs are used in verifying arithmetic designs with bit-level inputs and integer outputs.
BMDs use modified Shannon's expansion, where Boolean variable is treated as binary integer 
variable. For example, the complement of variable $x$ is modeled as $\bar{x} = 1-x$
and according to Shannon's expansion
$$f(x) = (1-x)\cdot f_{\bar{x}} + x\cdot f_x = f_{\bar{x}} + x\cdot(f_x - f_{\bar{x}}) = f_{\bar{x}} + x\cdot f_{\Delta x}$$
The above decomposition is called \emph{moment} decomposition where $f_{\bar{x}}$ is the 
\emph{constant moment}, $f_{\Delta x}$ is the \emph{linear moment}. In this way a Boolean
function $f$ is "transformed" to a linear function of $x$.
 \begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{BMD.png}
	\end{center}
	\caption{An example of BMD[8]}
	\label{fig:BMD}
\end{figure}
Fig.\ref{fig:BMD} is an example of BMD representation, it stands for an integer encoded
by 3 bits: $X = 4x_2 + 2x_1 + x_0$,
dashed edges means the constant moment and solid edges represent the first moment of the function 
w.r.t. the decomposing variable. As an application, a word-level BMD can efficiently represent an integer multiplication
because its size grows linearly with the size of multiplier's input, which is easy to observe
from fig.\ref{fig:multi} 
 \begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{multi.png}
	\end{center}
	\caption{BMD for $F = x \cdot y$, $x,y$ are 2-bit words, $F$ is 4-bit word[1]}
	\label{fig:multi}
\end{figure}

Another approach is \emph{Tylor Expansion Diagram} (TED)[1]. The abstraction method of TED is borrowed
from Tylor series expansion. Assume a Boolean function $f(x,y,\dots)$ be differentiable, it can 
be expanded w.r.t. variable $x$:
$$f(x,y,\dots) = f(x=0,y,\dots) + x\cdot f'(x=0,y,\dots) + \frac{1}{2}x^2\cdot f''(x=0,y,\dots) + \cdots$$
and it can be further expanded on remaining variables because taking $x=0$ already ensure any derivatives of $f$ are 
independent from $x$; recursive decomposition results a tree which is $TED$.

Moreover, some symbolic algebraic methods also provide effective abstraction. Dr. Kalla [3] proposed
an abstraction based on computing the Gr\"obner basis with abstraction term order on 
elimination ideal, the approach will directly given out the polynomial of system specification,
which can be compared to equations from arithmetic computations.

\subsection{Discussion}
Although I mentioned several approaches suit arithmetic verification well, till now there is 
no optimal method can deal with all arithmetic verification problems. There are many limitations
on the applications of those approaches.

One of the main limitations of BMD is that performing certain arithmetic operations cost badly,
such as power function and modular operations; TED has similar limitations as BMD. For Gr\"obner
basis approach, modeling method on complex arithmetic designs for integers has not been developed,
and in general case Gr\"obner basis computing costs exponential complexity time, symbolic algebraic 
tools have not yet been widely used.

In conclusion, the nature of difficulties is lack of computing capabilities so that it is 
not practical to exhaust $O(2^n)$ input search space. In order to get around this problem,
some approaches are proposed but they can only be effective/efficient to part of original 
problems.

\section{Prevention from design flaws}
\subsection{Intel FDIV bug}
The Intel FDIV bug occurs in 5 entries in the lookup table (LUT) of SRT division for floating-point
division algorithm in Pentium processor.

{\bf SRT Division} is a fast division algorithm widely adopted in CPU/FPU computing.
It is highly similar to the long division method we learned from elementary school,
the only difference is:
given dividend $p_0$ and divisor $d$, within each iteration $i$, $p_i$ is temporary 
dividend, the partial quotient
$q_i$ is chosen to satisfy
$$0\leq p_i - q_i\cdot d < d \ \ \ \ (\text{for long division})$$
$$|p_i - q_i\cdot d| \leq \frac{k}{r}\cdot d\ \ \ \ (\text{for $r$-radix SRT division})$$

Radix $r$ limits the range of quotient we can choose from, $k$ is a parameter
calculated according to $r$ to guarantee at every iteration only reasonable value of partial quotient
falls into interval $[-\frac{k}{r}\cdot d, \frac{k}{r}\cdot d]$.

\begin{Example}
Assume $r=10$, which means partial quotient could take value from $\{-5,-4,\dots,4,5\}$
corresponding $k/r = 5/9$.

Given dividend $p_0 = 1.000$, divisor $d = 8.000$, now compute 1/8.

Iteration 1: Choose first digit of quotient $q_0$ so that $|p_0-q_0\cdot d|\leq \frac{5}{9}\cdot 8.000$.
The only candidate is $q_0 = 0$. Then compute temporary dividend for next iteration:
\[
p_1 = 10\cdot(1.000 - 0\cdot 8.000) = 10.000
\]

Iteration 2: Choose $q_1$ so that $|10.000-q_1\cdot 8.000|\leq \frac{5}{9}\cdot 8.000$.
The only choice is $q_1 = 1$.
$$p_2 = 10\cdot(10.000 - 1\cdot 8.000) = 20.000$$

Iteration 3: Choose $q_2$ so that $|20.000-q_2\cdot 8.000|\leq \frac{5}{9}\cdot 8.000$.
Now there are two candidates, 2 and 3. Assume the algorithm chooses $q_2 = 3$ (can be either
deterministically or nondeterministically).
$$p_3 = 10\cdot(20.000 - 3\cdot 8.000) = -40.000$$

Iteration 4: Choose $q_3$ so that $|-40.000-q_3\cdot 8.000|\leq \frac{5}{9}\cdot 8.000$.
$q_3 = -5$ is the only choice.
$$p_4 = 10\cdot(-40.000 - (-5\cdot 8.000)) = 0.000$$
Algorithm terminates. The final quotient is $0+0.1+0.03+(-0.005) = 0.125$. [6]
\end{Example}

The SRT division adopted by FDIV has $r = 4$, which means quotient can be chosen
from $\{-2,-1,0,1,2\}$, and corresponding $k/r = 2/3$. If temporary dividend is given,
then the possible value(s) of quotient can be computed. In order to achieve higher efficiency,
FDIV uses LUT shown on last page (source: http://daviddeley.com/pentbug/pentbug3.htm).
As soon as FDIV algorithm get temporary dividend $p_i$, in LUT, unit $(d,p_i)$ stores
the value of corresponding quotient, reading from LUT cost $O(1)$ time.

This LUT is implemented in forms of PLA.[10] The bug occurs at 5 red units (also shown on last page),
in these units, pre-stored quotient value should be 2, but they are all equal to 0 in the buggy FPU.[4]
These 5 units are rarely used in most divisions, so the bug failed to be detected by test engineer
from Intel.

\subsection{Detection of similar flaws}
This LUT passed Intel's validation, because the prover only checked the lower half of
LUT (which is correct) and believed the upper half would be mirror to lower half.
After chips were fabricated, Intel also test it with millions of test vectors, unfortunately
no one hit this bug. This indicates that a formal method which can cover all possible
input logic is necessary.

I have 2 suggestions to detect this kind of bugs. One is to verify the synthesis result 
(values assigned to LUT) with a theorem prover such as HOL and PVS. In this example, for 
each $unit(d,p_i)$, check whether assertion 
$$|p_i - unit(d,p_i)\cdot d| \leq \frac{2}{3}\cdot d$$
is satisfied. Similar checking can be embedded into the synthesis tool, to eliminate bugs
inside synthesis tools (software verification).

The other one is to build a built-in module for test. This module could be designed based on 
the property we are aiming. For example, if we build an adder module to calculate 
the sum of all entries in LUT, a lot of bugs can be detected (including this FDIV bug)
except which does not change the sum. By designing reasonable built-in test modules,
time for test could be shorter while the coverage on bugs will be improved.

\subsection{Discussion}
The first paper[7] uses a lot of implementations in HOL theorem prover to argue that any theorems
adopted by floating-point algorithm synthesis tools can be verified. One of these implementations
is:

The following theorem can be implemented with HOL theorem prover statements in fig.\ref{fig:HOL}:
\begin{Theorem}
\label{alg:HOL}
If $q$ is a floating point number within 1 $ulp$ of the true quotient $a/b$ of two floating point
numbers, and $y$ is the correctly rounded-to-nearest approximation of the exact reciprocal $\frac{1}{b}$,
then the following two floating point operations:
$$r = a - bq$$
$$q' = q+ry$$
using round-to-nearest in each case, yield the correctly round-to-nearest quotient $q'$.
 \begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.6]{HOL.png}
	\end{center}
	\caption{HOL statements to prove theorem \ref{alg:HOL}[7]}
	\label{fig:HOL}
\end{figure}
\end{Theorem}

There is a whole argument on applying HOL prover to verify Intel Itanium floating point architecture,
which supports my suggestion on using theorem provers.

The second paper[9] talks about the built-in self speed test (BISST). BISST supports my ideal about
built-in test module, but works in slightly different way. Fig.\ref{fig:BISST} describes the implementation 
of BISST. First, only key function modules are tested; second, the essential idea of BISST is to 
build a bug-free module realizing the same function, then use a miter structure to check equivalence
of outputs; third, the BISST can be activated separately when testing.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{BISST.png}
	\end{center}
	\caption{A block diagram for BISST[9]}
	\label{fig:BISST}
\end{figure}

This paper also talked about an issue which my suggestion does not concern. To build a 
bug-free copy for every key module will occupy too much area, so this paper discussed how to
shrink the output signals so that the area cost could be lower down and keep a reasonable probability
of aliasing. In my suggestion, every module is designed as a property checker, so it will have much
smaller size than the module it tests.

%%%%%%%%%%%%%%%%%%%% The bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{1}

\bibitem{ref1}
Lv, Jinpeng. Scalable formal Verification of finite field Arithmetic circuits using
computer algebra techniques. Diss. University of Utah, 
Dept. of Electrical and Computer Engineering, 2012.

\bibitem{ref2}
Raudvere, Tarvo, et al. "System level verification of digital signal processing applications based on the polynomial abstraction technique." Computer-Aided Design, 2005. ICCAD-2005. IEEE/ACM International Conference on. IEEE, 2005.

\bibitem{ref3}
T. Pruss, P. Kalla and F. Enescu. Word-level Abstraction from Bit-level circuits using Gr\"obner Bases. IWLS 2013

\bibitem{ref4}
David W. Deley. "The Pentium Divison Flaw".

\bibitem{ref5}

Wikipedia: IEEE 754. http://en.wikipedia.org/wiki/IEEE\_floating\_point

\bibitem{ref6}

Clarke, Edmund M., Steven M. German, and Xudong Zhao. "Verifying the SRT division algorithm using theorem proving techniques." Computer Aided Verification. Springer Berlin Heidelberg, 1996.
\bibitem{ref7}

Harrison J. \emph{Floating-point verification using theorem proving[M]. Formal Methods for Hardware Verification}. Springer Berlin Heidelberg, 2006: 211-242.

\bibitem{ref8}

Chen Y A. \emph{Arithmetic circuit verification based on word-level decision diagrams}[D]. Intel, 1998.

\bibitem{ref9}

Hagihara, Yasuhiko, et al. "Floating-point datapaths with online built-in self speed test." Solid-State Circuits, IEEE Journal of 32.3 (1997): 444-449.

\bibitem{ref10}

Price D. \emph{Pentium FDIV flaw-lessons learned}[J]. Micro, IEEE, 1995, 15(2): 86-88.

\end{thebibliography}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-04-25 sequential/latex/abstraction/myers.tex
upload files from laptop linux

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  IEEEsample.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%    More information: see the header of IEEEtran.sty
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%

\documentclass[11pt,twoside, onecolumn]{IEEEtran}
%\documentclass[conference]{IEEEtran}

%%%\IEEEoverridecommandlockouts

\usepackage[ruled]{./algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%% \RequirePackage{times}
%% \RequirePackage{algorithmic}
%% \PassOptionsToPackage{boxed}{algorithm}
%% \RequirePackage{algorithm}
%% \RequirePackage{multicol}
%\renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
%\renewcommand{\algorithmicensure}{\textbf{Outputs:}}
%\DeclareMathAlphabet{\mathtsl}{OT1}{ptm}{m}{sl}

%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em1
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{example}{Example}
%\newtheorem{corollary}{Corollary}

\RequirePackage{amssymb, mathptm}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{comment}

%%indent in algorithm

%\setcounter{page}{1}

% New command for the table notes.
\def\tabnote#1{{\small{#1}}}

% New command for the line spacing.
\newcommand{\ls}[1]
    {\dimen0=\fontdimen6\the\font
     \lineskip=#1\dimen0
     \advance\lineskip.5\fontdimen5\the\font
     \advance\lineskip-\dimen0
     \lineskiplimit=.9\lineskip
     \baselineskip=\lineskip
     \advance\baselineskip\dimen0
     \normallineskip\lineskip
     \normallineskiplimit\lineskiplimit
     \normalbaselineskip\baselineskip
     \ignorespaces
    }
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqarr}{\begin{eqnarray}}
\newcommand{\eeqarr}{\end{eqnarray}}
%\newcommand{\ov}{\overline}
\newcommand{\ov}{\bar}
\newcommand{\xor}{\bigoplus}
\newcommand{\Fm}{{\mathbb{F}}}

%the following is for space before and after align or other equation environment.

%%
\newtheorem{Algorithm}{Algorithm}[section]
\newtheorem{Definition}{Definition}[section]
\newtheorem{Example}{Example}[section]
\newtheorem{Proposition}{Proposition}[section]
\newtheorem{Lemma}{Lemma}[section]
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}{Corollary}[section]
\newtheorem{Conjecture}{Conjecture}[section]
\newtheorem{Problem}{Problem}[section]
\newtheorem{Notation}{Notation}[section]
\newtheorem{Setup}{Problem Setup}[section]
%%%

%%set spacing between table columns
\setlength{\tabcolsep}{3pt}

\begin{document}

%\thispagestyle{empty}
%\pagestyle{empty}

\ls{1.1}

\title{\large{\textsc{Report  on Algorithmic Complexity Analysis in Design Automation and Formal Verification}}}
\author{Xiaojun Sun\\
A Technical Report\\
in partial fulfillment of requirements for PhD qualification\\
of Computer Engineering Program, University of Utah\\
Fall Semester 2013
}

%%\author{\IEEEauthorblockN{Jinpeng Lv and Priyank Kalla}\thanks{This work is sponsored in part by a grant from NSF \#CCF-546859.}
%\IEEEauthorblockA{Department of  Electrical and Computer Eng.\\
% University of Utah, Salt Lake City, UT-84112 \vspace{-0.3in}
 %\{lv, kalla\}@eng.utah.edu
% }
%\and
%\IEEEauthorblockN{Florian Enescu} \thanks{\normalsize  978-3-9810801-8-6/DATE12/$\copyright 2012$ EDAA}
%\IEEEauthorblockA{Department of Mathematics and Statistics\\
% Georgia State University,  Atlanta, GA 30302-4038 \vspace{-0.3in}
% fenescu@mathstat.gsu.edu
%} 
%

\maketitle

%\markboth{MS Proposal by Tim Pruss}{}
\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
\newcommand{\B}{{\mathbb{B}}}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\F}{{\mathcal{F}}}
\newcommand{\G}{{\mathcal{G}}}
\newcommand{\R}{\mathbb{R}}
%%%

\newcommand{\debug}[1]{\textcolor{gray}{[ #1 ]}}

%\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%% Include your files here %%%%%%%%%%%%%%%%%%%%%
\section{Linear Complexity}
\subsection{Definition}
Computational complexity notions strictly defined in [1] are preliminaries to this report.
\begin{Definition}
Computational complexity $\Theta(n)$ is defined as \emph{linear complexity}. The property of
linear complexity algorithm is the time or space cost of this algorithm is a linear function
of input data size $n$.
\end{Definition}

The following algorithm is a simple example of linear complexity algorithm on both time
and space consumption.
\begin{Example}
An array with $n$ integers is stored within a 1-dimension linked list. A traversal algorithm
starts from the head of list, transit through the pointer pointing to successor node, stops
when reaching the end node. The space complexity of this linked list is $\Theta(n)$, and
time complexity of traversal algorithm is $\Theta(n)$.
\end{Example}

\subsection{Design Automation Applications}
\label{sec:graph}
Algorithms from graph theory are a big portion of algorithms in design automation because
graph representation is widely adopted in different aspects of electrical engineering. For example,
graph representation covers finite state machine, labeled petri-net, delay network and abstraction
of circuits. If there is a loop in state transition sequence, or feedback loop in gate-level
implementation, there is also a cycle in corresponding graph representation. An efficient
algorithm can check whether there exists a cycle in a directed graph, which is
\emph{Topological Sort}.

In graph theory, a topological sort or topological ordering of a directed acyclic graph (DAG) is a linear
 ordering of its vertices, in which each vertex comes before all vertices to which it has outbound edges. 
 Every DAG has one or more topological sorts.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{A directed graph $G=(V,E)$}
 \KwOut{A list of topological sorted vertices\\} %, a Gr\"{o}bner basis
  Initialize $S$;\\
  $V_0 \gets$ all vertices with no incoming edges;\\
  \While{$V$ is non-empty}{
 	 remove a vertex $v$ from $V_0$;\\
 	 insert $v$ into $S$;\\
 	 \For{each vertex $u$ with an edge $e = <v,u>$}
 	 {
 	 	remove edge $e$ from $G$;\\
 	 	\If{u has no incoming edges}
 	 	{
 	 		insert $u$ into $V_0$;
 	 	}
 	 }
   }
   \If{$G$ still has edges}
   {
   		\Return{ "Graph has cycle(s)"};\\
   		}
   	\Else{
   		\Return $S$;
   		}
\caption{The Topological Sort Algorithm$^{[2]}$}\label{alg:top}
\end{algorithm}

\subsection{Computational Analysis}
According to algorithm \ref{alg:top}, before algorithm terminated every vertex and edge is scanned only once then directly
removed. So in the worst case, there are no edges left in original graph, all vertices and edges are
visited, the time complexity is $\Theta(|V|+|E|)$, linear to number of vertices and edges in graph.

\subsection{Discussion}
Linear complexity algorithm is low-complexity algorithm, the time/space cost grows along with the
size of input (usually given in a limited size). In general an even lower complexity usually means
$O(logn)$ or $O(n^{1/k})$ complexity which is hard to achieve. As to topological sort, when the vertices
and edges in original graph are given arbitrarily, there is no alternative algorithm with lower complexity.

\section{Polynomial Complexity}
\subsection{Definition}
\begin{Definition}
\emph{Polynomial complexity} is denoted by $O(n^d)$, where $d$ is a constant number. The property
of polynomial complexity algorithm is the time and space cost of this algorithm is polynomial function
of input size $n$.
\end{Definition}

Logarithmic complexity is
also polynomial complexity because it satisfies the upper bound when taking $d=1$. Linear complexity belongs
to polynomial complexity since $\Theta(n) = O(n^1)$. If $d=0$, it degenerates to constant complexity $O(1)$.

\begin{Example}
The insert sort of an array has polynomial complexity $O(n^2)$.
\end{Example}

\subsection{Design Automation Applications}
In physical design automation there is a special routing problem for interconnect wires. In PCB layout,
the interconnect wires are usually assigned to different copper layers to avoid crossover; wires within the 
same layer are not allowed for crossover. Since copper on the top layer is easy to be etched, industry
requires designers to put the most complicated pattern on top layer. So top layer of interconnection usually requires
the largest number of wires. This problem can be abstracted as:

%%%%%%%%%%%%%%%%%%%%%%% fig:pcb here!!
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{fig_pcb.png}
	\end{center}
	\caption{Example bipartite graph of PCB routing problem[10]}
	\label{fig:pcb}
\end{figure}

Given a bipartition graph, there exists an edge for a pair between 2 sets: $(i,\pi(i))$, $\pi(i)$
is a permutation of $\{1,2,\dots,n\}$. 
$\forall 1\leq i < j\leq n$, edge $(i,\pi(i))$ has crossover with edge $(j,\pi(j))$ if and only if
$\pi(i) > \pi(j)$.
Find the maximal subset of these edges with no crossover. In fig.\ref{fig:pcb}, the desired \emph{maximal no-crossover subset} (MNS) is $\{(4,2),(5,5),(7,9),(9,10)\}$.

This problem has a solution based on a polynomial complexity algorithm. The main idea of that algorithm
is \emph{dynamic programming} (DP).

Denote a part of wires by set $N(i,j) = \{t | (t, \pi(t)) \in Nets, t \leq i, \pi(t) \leq j \}$, its
MNS is denoted by $MNS(i,j)$, its size $|MNS(i,j)| = Size(i,j)$. Then it follows deduction equations:
\begin{equation}
 Size(1,j) = \left\{
\begin{array}{lcl}
0       &      & {j < \pi(1)}\\
1     &      & {j \geq \pi(1)}
\end{array} \right. 
\end{equation}
\begin{equation}
 Size(i,j)\ \ (i>1) = \left\{
\begin{array}{lcl}
Size(i-1,j)       &      & {j < \pi(i)}\\
max\{Size(i-1,j), Size(i-1,\pi(i)-1)+1\}     &      & {j \geq \pi(i)}
\end{array} \right. 
\end{equation}
Equation (1) is the initial value of $Size(i,j)$ when $i=1$, it is deduced with $j$ increasing, MNS size equals to 0 until $j$ reaches the position defining first edge. Equation (2) means every time the algorithm
sweeping $j$ to define a new edge, it is necessary to compare the new size including this edge or not.

Pseudo codes are easy to get from deduction equations for dynamic programming problems. 
\subsection{Computational Analysis}
From the deduction equations, this dynamic programming algorithm will fill up a matrix when getting
answer to the problem, and each entry in the matrix will be updated only once, so both time complexity
and space complexity for this DP algorithm is $O(n^2)$, a polynomial complexity.

\subsection{Discussion}
\label{sec:LIS}
An alternative algorithm for this problem lowers down the time complexity.
By observing optimal solution to MNS problem, the equivalent problem is to find a Longest Increasing
Subsequence (LIS) in permutation $\{\pi(1),\pi(2),\dots,\pi(n)\}$. Fortunately there is a $O(nlogk)$
algorithm to compute LIS.

The thought of algorithm is: use a stack to maintain a current longest sequence. Scan the whole array,
compare the top of stack with current element $temp$. If $temp$ is larger than stack top, push it in
stack so that current LIS is extended. Otherwise, binary-search the first element larger than $temp$
in the stack, then replace it with $temp$. At last the size of stack is the length of LIS.

For example, given sequence $1,5,8,3,6,7$. The stack is $1,5,8$, when reading 3, search and find 5 in
stack, replace with 3 to get 1,3,8; then read in 6, replace 8 in stack again; at last read in 7. The
final stack is 1,3,6,7.

This improved algorithm uses a loop to scan the array, then use binary search to find desired replacing
elements. Assume the length of LIS is $k$, then in the worst case every time a binary search is required
in $k$ length subsequence, the time complexity is $O(nlogk)$. The result given by this algorithm is
guaranteed to be optimum, so it is not heuristic.

\section{Exponential Complexity}
\subsection{Definition}
\begin{Definition}
\emph{Exponential complexity} is complexity $O(r^n)$, where $r$ is a constant number. Exponential complexity
algorithm has the property that time/space cost is exponential function of problem size $n$.
\end{Definition}

Exponential complexity is not a polynomial complexity. Usually it indicates the time/space cost explodes
with input size increasing; an algorithm with exponential complexity is common in brute-force techniques
and not considered as an efficient and optimal choice in general case.

\subsection{Formal Verification Applications}
\label{sec:3sat}
Circuit satisfiability (SAT) problem is a typical problem which has no polynomial complexity algorithm
to find optimal solution. For example, given a circuit with primary inputs $\{a,b,c\}$, the output
$Y = (a \land b) \lor \bar{c}$. SAT problem is: is there any boolean vector assignment to primary inputs
so that $Y = 1$?

%%%%%%%%%%%%% fig:SAT here !!!
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{fig_SAT.pdf}
	\end{center}
	\caption{Sample circuit for SAT problem}
	\label{fig:SAT}
\end{figure}

Following the boolean logic, any boolean description of SAT problem can be written in conjunctive
normal form (CNF). If the number of literals in every clause of CNF expression is limited to 3, 
the SAT problem degenerates to 3-SAT problem. The naive method to solve this problem is to scan
all possible assignments on inputs to check 3-SAT. In worst case, if a circuit has $n$ primary
inputs, every input could be assigned $\{0,1\}$, the exhaustive checking algorithm cost $2^n$ checks,
the time complexity is $O(2^n)$, this is an exponential complexity.

\subsection{Discussion}
According to literature search, currently no polynomial lower bound has been proved for 3-SAT problems,
which means there is no algorithm can provide a polynomial complexity in worst case.
However, some refinements are effective to lower down time cost in practice.
One algorithm improves time complexity to $O(r^n), r < 2$.
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{A CNF expression with at most 3 literals in a clause}
 \KwOut{An legal assignment of input satisfies given CNF or claim no available assignments\\} %, a Gr\"{o}bner basis
  Check whether current assignment satisfies, if yes return current assignment;\\
  \While{pick arbitrary clause $C = \alpha_1 \lor \alpha_2 \lor \alpha_3$ that still unsatisfied}{
 	 set $\alpha_1$ to "true", reduce original CNF and recursively check rest clauses;\\
 	 set $\alpha_1$ to "false" and $\alpha_2$ to "true", reduce original CNF and recursively check rest clauses;\\
 	 set both $\alpha_1$ and $\alpha_2$ to "false" and assign "true" to $\alpha_3$, reduce original CNF and recursively check rest clauses;\\
 }
 \If{Cannot find any assignment to satisfy original CNF}
 {
 \Return{"No satisfiable assignments"}
 }
 	 \caption{The Refined Exponential Time 3-SAT Algorithm$^{[2]}$}\label{alg:3sat}
\end{algorithm}
\begin{Theorem}
Algorithm \ref{alg:3sat} solves 3-SAT problem with time complexity $O(1.8393^n)$.
\end{Theorem}

\textit{Proof}[10]\ \ \ Correctness: assignment returned by the algorithm satisfies original CNF expression.
When the algorithm claims "no satisfiable assignments", for all clauses in form
$C = \alpha_1 \lor \alpha_2 \lor \alpha_3$, following possible assignments are checked:\par
(1) $\alpha_1 = TRUE$;\par
(2) $\alpha_1 = FALSE$ and $\alpha_2 = TRUE$;\par
(3) $\alpha_1 = \alpha_2 = FALSE$ and $\alpha_3 = TRUE$.\\
Any assignments must belong to one and only one kind of above 3 situations. Since the while loop
will not terminate unless all clauses are recursively checked, at this time there is no assignment
making original CNF expression be true.

Complexity: Assume $T(n)$ is the time algorithm operates on CNF containing $n$ literals in total.
Let $m$ be the number of clauses in this CNF expression, then $m = O(n^3)$ since every clause
has at most $3$ literals. In each situation of recursion, the size of original problem is
scaled down to $(n-1), (n-2), (n-3)$ after determining assignments for $\alpha_1,(\alpha_1,\alpha_2),(\alpha_1,\alpha_2,\alpha_3)$. Reducing procedure scans $m$ clauses and justifying satisfaction requires 
scan all $n$ literals to take their values. So deduction equation on runtime is
$$T(n) \leq T(n-1) + T(n-2) + T(n-3) + O(n+m)$$
Solution is $T(n) = O(1.8393^n)$, while $1.8393$ is the largest root of equation $x^3 = x^2 + x + 1$. \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ $\Box$

\section{NP-Complete}
\subsection{Definition}
\begin{Definition}
A decision problem belongs to $P$ when there is a polynomial time complexity solution.
\end{Definition}

Decision problem is a kind of problems can be answered by $"YES"$ or $"No"$. For example, the LIS problem
mentioned in \ref{sec:LIS} can be classified to $P$ if it is asked in this form: "Does there exist a $LIS$ of
length $k$?". The LIS algorithm also gives the subsequence when solving this decision problem.

\begin{Definition}
Nondeterministic polynomial problems ($NP$) is a class of decision problems, when given a testcase,
the problem can verify the testcase within polynomial complexity time.
\end{Definition}

A set of assignments can be verified in 3-SAT problem within polynomial time, and 3-SAT is a decision
problem. So 3-SAT$\in NP$.

\begin{Definition}
\label{def:nphard}
A problem $\Pi$ is \emph{NP-hard} if and only if, for any problem $\Pi' \in NP$, there is a polynomial-time Turing reduction
from $\Pi'$ to $\Pi$. A Turing reduction means a reduction that can be executed on a Turing machine. Polynomial-time Turing reductions are also called \emph{Cook reductions}.
\end{Definition}

Note that NP-hard problems are not necessarily decision problems. 
According to this definition, a NP-hard problem is not easier than any other NP problems. If a NP-hard
problem is solved in polynomial time, then all $NP$ problems have algorithms to find solutions in polynomial time.

\begin{Definition}
If a problem is NP-hard, and it belongs to NP, then this problem is \emph{NP-Complete}.
\end{Definition}

%%%%%%%%%%%%%%%%% fig:NP here !!!!!!!!!!!!
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{fig_NP.pdf}
	\end{center}
	\caption{A popular graph illustrating relations among P, NP, NP-hard and NP-Complete}
	\label{fig:NP}
\end{figure}

Fig.\ref{fig:NP} illustrates relations between these classes of problems. Note that $P!=NP$ and
$P$ disjoint with $NP$-$hard$ are just conjectures and need to be proved.

To directly prove that a problem is NP-hard or NP-Complete is difficult. Currently most proofs are based
on finding a polynomial time Turing reduction from SAT problem. SAT problem is proved to be NP-Complete in
[4]. If a polynomial time Turing reduction from SAT problem to given problem $\Pi$ is found, then $\Pi$
is not easier than SAT problem. So $\Pi$ is proved to be NP-hard or NP-Complete (if $\Pi \in NP$).

\subsection{Design Automation Applications}
Covering problem is important in logic synthesis. For a circuit with specification set $S = \{s_1, s_2, s_3, s_4\}$, there are 3 possible gate implementations, where $G_1$ satisfies $\{s_1, s_2\}$, $G_2$ satisfies
$\{s_2,s_3\}$ and $G_3$ satisfies $\{s_3,s_4\}$. The question is: is there a set of gate implementations
$\{G_i\}$ that every gate implementation is disjoint with each other on satisfiable specifications, 
and their union equals to $S$? In this example the answer is yes, the a desired solution is $\{G_1, G_3\}$.

This problem is called \emph{exact covering} problem. 
\begin{Theorem}
Exact covering problem (ECP) is NP-Complete.
\end{Theorem}

\textit{Proof} \ \ \ Given a set $\{G_i\}$, this problem can be verified by scanning every element in
$\{G_i\}$ can compare with $S$, this can be done within polynomial time, so $ECP \in NP$.

Following argument proves that there is a polynomial time reduction from ECP to SAT.

Construction: $n$ is the number of specifications, $m$ is the number of candidate gate implements. Assume a set of
variables $x_1,x_2,\dots,x_n$ and a CNF $F = C_1\land C_2\land \cdots \land C_m$ where clause 
$C_j = z_{j1}\lor z_{j2} \lor \cdots \lor z_{js_j}$, $j = 1,2,\dots,m$; $z$ is literal based on $x$.
Objective is to construct desired subset $W = \{G_i\}$ of finite set $S$ so that $F$ is satisfied
if and only if $W$ contains the exact cover of $S$.
Let $S = \{x_1,x_2,\dots, x_n, C_1, C_2,\dots, C_m\} \bigcup \{p_{jt}\ |\ 1\leq t \leq s_j, 1\leq j\leq m \}$,
where $p_{jt}$ represents literal $z_{jt}$ in $C_j$. For each variable $x_i$, there are 2 subset
$T_i^T$ and $T_i^F$. $T_i^T$ includes variable $x_i$ itself, and all corresponding literals
equal to $\neg x_i$; $T_i^F$ also includes variable $x_i$ itself, with all corresponding literals
equal to $x_i$, i.e.
$$T_i^T = \{x_i, p_{jt}\ |\ z_{jt} = \neg x_i, 1\leq t\leq s_j, 1\leq j\leq m\},$$
$$T_i^T = \{x_i, p_{jt}\ |\ z_{jt} =  x_i, 1\leq t\leq s_j, 1\leq j\leq m\}.$$
Since other subsets do not include $x_i$, any exact cover $U$ must include exactly one
of $T_i^T$ and $T_i^F$. Assign $x_i = 1$ if and only if $U$ contains $T_i^T$, thus assignment
$x_i = 0$ equals to $U$ containing $T_i^F$. Additionally, for each clause $C_j$, it has $s_j$ subsets:
$C_{jt} = \{C_j, p_{jt}\}, 1\leq t \leq s_j$. Besides these subsets, every $p_{jt}$ constructs a unit
subset by itself: $\{p_{jt}\}$.

To conclude, $W$ includes following subsets:\par
(1) $\{p_{jt}\}$ (number: $\sum_{j=1}^{m}s_j$)\par
(2) $T_i^T, T_i^F$ (number: $2\times n$)\par
(3) $C_{jt}$ (number: $\sum_{j=1}^{m}s_j$).\par

Reasoning: To prove $F$ is satisfied
if and only if $W$ contains the exact cover of $S$.

Assume $U \subseteq W$ is exact cover for $S$. Assign $1$ to $x_i$ if $T_i^T \in U$;
assign $0$ to $x_i$ if $T_i^F \in U$. For all $1\leq j\leq m$, there is exactly one $C_{jt} \in U$.
If $T_i^T \in U$, then $p_{jt}\notin T_i^T$, thus $z_{jt} = x_i$, and $x_i$ is already assigned "1", so
$C_j$ is satisfied. On the other hand if $T_i^F \in U$, then $p_{jt}\notin T_i^F$, thus $z_{jt} = \neg x_i$, 
and $x_i$ is already assigned "0", so $C_j$ is also satisfied. Consider every clause been satisfied,
the corresponding assignment satisfies $F$.

If $F$ is satisfiable, assume an assignment $t$ under which $F$ is true. For each $1\leq i \leq n$,
$t(x_i) = 1$ implies $U$ contains $T_i^T$; $t(x_i) = 0$ implies $U$ contains $T_i^F$. For each
$1\leq j\leq m$, since $t$ satisfies $C_j$, there exists a literal $z_{jt}$ so that $t(z_{jt}) = 1$
and $p_{jt}$ not included by $U$, thus $C_{jt}$ could be inserted into $U$. Until this time,
$U$ already covers all $x_i$ and $C_j$, as well as a part of $p_{jt}$. So after add the uncovered
$p_{jt}$ in forms of unit subset $\{p_{jt}\}$ into $U$, $U$ becomes the exact cover of $S$.

With above arguments, $W$ contains exact cover $U$ (existence of $U$) as long as $F$ is satisfied. If $W$ can be 
constructed within polynomial time, it means the ECP can be reduced to a SAT problem in
polynomial time. The maximal number of literals in $F$ is $mn$, so $|S|\leq n+m+mn$, and
number of subsets in $W$ will not exceed $2n+2mn$ while size of single subset is smaller than $n+1$.
Moreover, the construction method is straightforward and able to finish within polynomial time.
Thus, a polynomial time Turing reduction from ECP to SAT is proved.

In conclusion, $ECP \in NP$ and can be reduced to SAT problem, which is NP-Complete. So ECP is NP-Complete.$\Box$

\subsection{Computational Analysis and Discussion}
A naive algorithm for ECP is to enumerate all subsets from $\{G_i\}$. If there are $n$ gate implementations,
number to enumerate is $2^n$, so the time complexity for naive algorithm is $O(2^n)$, it is not a polynomial
complexity.

ECP is NP-Complete indicates there is currently no polynomial time complexity algorithm to find the
solution. However, there are several heuristics can be applied to minimize search space as much as possible.
One of them is DLX algorithm invented by D.E. Knuth[5].

Assume there is a binary matrix $M$, each row of $M$ can represent a gate implementation $G_i$.
Entry $M(i,j) = 0$ means $G_i$ does not satisfy specification $s_j$, $M(i,j) = 1$ means it does satisfy.
For example, following matrix
\begin{equation}
M = \left(
\begin{array}{lcccccr}
0 & 0 & 1 & 0 & 1 & 1 & 0\\
1 & 0 & 0 & 1 & 0 & 0 & 1\\
0 & 1 & 1 & 0 & 0 & 1 & 0\\
0 & 0 & 0 & 1 & 0 & 1 & 0\\
0 & 1 & 0 & 0 & 0 & 0 & 1\\
0 & 0 & 0 & 1 & 1 & 0 & 1
\end{array}
\right)\nonumber
\end{equation}
row 1,4,5 is an exact cover of specification set, because by merging these rows (do XOR for corresponding
entries) the result is a row with all "1" on every column.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Matrix to be covered $M$}
 \KwOut{A list of row numbers as solution\\} %, a Gr\"{o}bner basis
  \If{Matrix $M$ is empty}
  {
  		\Return{List of solution};
  	}
  Otherwise choose a column $c$; \ \ \ /* deterministically, usually choose the one with least number of "1" entries */\\
  Choose a row $r$ such that $M(r,c) = 1$; \ \ \ /* nondeterministically */ \\
  Add row $r$ to solution list;\\
  \For{each column $j$ such that $M(r,j) = 1$}{
 	 \For{each row $i$ such that $M(i,j) = 1$}
 	 {
 	 	delete row $i$ from $M$;
 	 }
 	 delete column $j$ from $M$;
   }
   Recursively repeat on reduced Matrix;
\caption{Knuth's DLX Algorithm for ECP$^{[5]}$}\label{alg:knuth}
\end{algorithm}

The upper bound of this algorithm is still exponential time complexity. However in experiment,
the search tree usually get very sparse after a few recursions. In general situation, this algorithm
is an efficient heuristic (nondeterministically branch-and-bound) algorithm.

\section{NP-hard}
\subsection{Definition}
The definition of NP-hard problem is stated in DEF. \ref{def:nphard}. In practice, most NP-hard problems
are problems proved no easier than NP-Complete problems but not decision problems.

\subsection{Design Automation Applications} 
A problem related to route among pins is important in design automation.
In fig.\ref{fig:MST}(a), there are 4 pins and routing is only allowed along the grids.
The problem is to find a path of routing to connect all pins with lowest cost on total length of
wires. This problem is called minimal Rectilinear Steiner tree problem (MRST), highly similar to 
minimal Steiner tree problem, which is proved to be NP-hard[7]. Since MRST problem is not a 
decision problem ($\notin NP$), it is NP-hard rather than NP-Complete.

\subsection{Computational Analysis}
There is no polynomial time complexity algorithm to find an optimal solution to this problem.
In order to lower down the time complexity, some heuristics are adopted to approximate the 
optimal solution. One algorithm is to use minimal spanning tree (MST) to approximate MRST,
and the distortion is bounded to a small value.

\begin{Theorem}
The relationship between MST and MRST is:
$$\frac{Cost(MST)}{Cost(MRST)} \leq \frac{3}{2}$$
\end{Theorem}

This assertion is proved by K. Hwang[8], it shows MST algorithm provides a $3/2$-approximation of
MRST problem. MST has deterministic algorithm such as Prim's Algorithm and Kruskal's Algorithm,
both have polynomial time complexity $O(|E|log|V|)$. Considering $|V|$ is the number of pins $n$,
and $|E|$ is the number of edges connect every pair of pins ($n(n-1)/2$), the complexity is
written as $O(n^2logn)$.

\subsection{Discussion}
It is hard to find even lower complexity algorithm than MST heuristic. However, the quality
of solution given by MST approximation is relatively low (compare fig.\ref{fig:MST}(b) to fig.\ref{fig:Steiner}(b)).

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{MST.png}
	\end{center}
	\caption{Example of MST heuristic[9]}
	\label{fig:MST}
\end{figure}

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{Steiner.png}
	\end{center}
	\caption{Example of optimal routing with Steiner points [9]}
	\label{fig:Steiner}
\end{figure}

The iterated 1-Steiner heuristic is applied to improve the quality. 
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{A set of $n$ pins $P$}
 \KwOut{A Steiner tree on $P$\\} %, a Gr\"{o}bner basis
  \While{There are places on grids to insert Steiner points}
  {
  	Find a 1-Steiner point which brings maximal improvement on cost by constructing MST with it;\\
  	Add this Steiner point into recording list $S$;\\
  	In refined MST, find existing Steiner points with degree $\leq 2$, remove them from $MST$ and $S$;
  }
  \Return{MST generated by $P \cup S$}
\caption{Iterated 1-Steiner Heuristic Algorithm$^{[6]}$}\label{alg:iter}
\end{algorithm}

In this heuristic, in preparation stage, an ordering of potential Steiner point candidates is calculated by checking the improvement on total
cost of MST generated by $P \cup \{x\}$, $x$ is a candidate. Then the best $n$ candidates are kept,
so the number of iterations are limited to $n$. During each iteration, the removal and MST re-generation
cost $O(n^2logn)$ time. So total time complexity is $O(n^3logn)$, higher than the MST heuristic.
Fig.\ref{fig:iter}(b)-(e) depicts all iterations of this algorithm ($n=4$).

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.4]{iter.png}
	\end{center}
	\caption{Example of iterated 1-Steiner heuristic algorithm[9]}
	\label{fig:iter}
\end{figure}

Although executed with a higher time complexity, the quality of the solution given by this algorithm is 
greatly improved comparing to MST heuristic. In experiment, this algorithm gives optimal solution for
$n\leq 4$; for $n=5$ which MST heuristic may give a $3/2$ approximation, the iterated 1-Steiner heuristic
gives a tighter $7/6$ approximation in worst case.

%%%%%%%%%%%%%%%%%%%% The bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{1}

\bibitem{ref1}
Web lecture from MIT, \emph{Big O Notation}. http://web.mit.edu/16.070/www/lecture/big\_o.pdf

\bibitem{ref2}
Kleinberg J, Tardos E. \emph{Algorithm design}[M]. Pearson Education India, 2006.

\bibitem{ref3}
Jeff Erickson, \emph{"Algorithms course materials"}.

\bibitem{ref4}
Stephen A. Cook (1971). \emph{The Complexity of Theorem-Proving Procedures}. Proc. 3rd Ann. Symp. on Theory of Computer. pp151-158.

\bibitem{ref5}
Knuth D E. \emph{Dancing links}[J]. arXiv preprint cs/0011047, 2000.

\bibitem{ref6}

Kahng, Andrew, and Gabriel Robins. \emph{A new class of Steiner tree heuristics with good performance: the iterated 1-Steiner approach.}  International Conference on Computer-Aided Design,. IEEE, 1990.

\bibitem{ref7}

Richard M. Karp. \emph{Reducibility Among Combinatorial Problems}. In R. E. Miller and J. W. Thatcher (editors). \emph{Complexity of Computer Computations}. New York: Plenum. pp. 85-103.

\bibitem{ref8}

Hwang, Frank K. \emph{On Steiner minimal trees with rectilinear distance.} SIAM journal on Applied Mathematics 30.1 (1976): 104-114.

\bibitem{ref9}

L. Wang, Y. Chang, K. Cheng , Electronic design automation: synthesis, verification, and test[M]. Elsevier, 2009.

\bibitem{ref10}

W. Qu, T. Liu et al., \emph{Algorithm Design and Analysis}. Tsinghua University, 2011

\end{thebibliography}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-04-25 sequential/latex/abstraction/prelim_gf.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/abstraction/iwls/prelim_gf_iwls.tex
@@
 
-A Galois field $\Fq$ is a field with a finite number of elements. The
-number of elements $q$ of the field is a power of a prime integer ---
-i.e. $q = p^k$, where $p$ is a prime integer, and $k \geq 1$
-\cite{galois_field:mceliece}. 
-%Galois fields are denoted as ${\mathbb{F}}_{q}$ and also $GF(q
-%=p^k)$. 
-We are interested in fields where $p = 2$ and $k >1$ --- i.e. {\it
-  binary Galois extension   fields} $\mathbb{F}_{2^k}$ --- as they are
-widely employed in hardware implementations of cryptography
-primitives.
+A Galois field is a field with a finite number of elements. The number
+of elements $q$ of the field is a power of a prime integer --- i.e. $q
+= p^k$, where $p$ is a prime integer, and $k \geq 1$ is a positive
+integer \cite{galois_field:mceliece}. Galois fields are denoted as
+${\mathbb{F}}_{q}$ and also $GF(q =p^k)$. We are interested in fields
+where $p = 2$ and $k >1$ --- i.e. {\it   binary Galois extension
+  fields} $\mathbb{F}_{2^k}$ --- as they are widely employed in
+hardware implementations of cryptography primitives.
 
@@
 $a_i \in {\mathbb{F}}_2, i = 0, \dots, k-1$, and $\alpha$ is the root of
-the irreducible polynomial, {\it i.e.} $P(\alpha)=0$. 
-%The field
-%$\Fkk$ can therefore be construed as a $k$-dimensional vector space
-%over ${\mathbb{F}}_2$.
+the irreducible polynomial, {\it i.e.} $P(\alpha)=0$. The field
+$\Fkk$ can therefore be construed as a $k$-dimensional vector space
+over ${\mathbb{F}}_2$.
 
@@
 \begin{Example}\label{ex:1}
-{\it
 Let us construct ${\mathbb{F}}_{2^4}$ as ${\mathbb{F}}_2[x] \pmod{
@@
 = (0011)$; hence we have the three equivalent representations.
-}
 
@@
 \begin{center}
-{\tiny
+{\small
 \caption{Bit-vector, Exponential and Polynomial representation of
@@
 Arbitrary mappings among $k$-bit vectors can be constructed; each such
-mapping generates a function $f: \B^k \rightarrow \B^k$. 
-%. Since every
-%$k$-bit vector can be construed as an element in $\Fkk$ (as shown in
-%the above example), every such function corresponds to a function over
-Every such function is also a polynomial function over Galois fields:
-$f: \Fkk \rightarrow \Fkk$.
+mapping generates a function $f: \B^k \rightarrow \B^k$. Since every
+$k$-bit vector can be construed as an element in $\Fkk$ (as shown in
+the above example), every such function corresponds to a function over
+Galois fields: $f: \Fkk \rightarrow \Fkk$. Every such function is also
+a polynomial function.
 
@@
 From \cite{ff:1997}: 
-Any  function $f: \Fq \to \Fq$ is a polynomial function
+Let $\Fq$ be a Galois field of $q$ elements where $q$ is a power of a
+prime integer. Any  function $f: \Fq \to \Fq$ is a polynomial function
 over $\Fq$, that is there exists a polynomial $\F \in \Fq [x]$ such
@@
 
-%% \begin{Example} {\it
-%% Let $A = \{a_2, a_1, a_0\}$ and $Y = \{y_2, y_1,
-%% y_0\}$ be 3-bit vectors.  Consider the function $Y[2:0] = A[2:0]
-%% >> 1$, i.e. a {\bf bit-vector right shift} operation on $A$. 
-%% The function maps as follows:
+\begin{Example} {\it
+Let $A = \{a_2, a_1, a_0\}$ and $Y = \{y_2, y_1,
+y_0\}$ be 3-bit vectors.  Consider the function $Y[2:0] = A[2:0]
+>> 1$, i.e. a {\bf bit-vector right shift} operation on $A$. 
+The function maps as follows:
 
-%% \begin{center}
-%% {\small
-%% \begin{tabular}{c|ccc|c||c|ccc|c}
-%% $\{a_2a_1a_0\}$  & $A$ &$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$ &
-%%   $\{a_2a_1a_0\}$  &$A$&$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$\\
-%% \hline
-%% 000  &0 &$\rightarrow$& 000 & 0 & 100  &$\alpha^2$ &$\rightarrow$& 010 &  $\alpha$ \\
-%% 001  &1 &$\rightarrow$& 000 & 0 & 101  &$\alpha^2 + 1$ &$\rightarrow$&010 & $\alpha$ \\
-%% 010  &$\alpha$ & $\rightarrow$ & 001& 1 & 110  &$\alpha^2 + \alpha$&$\rightarrow$& 011 &$\alpha + 1$ \\   
-%% 011  &$\alpha + 1$ &$\rightarrow$& 001 &1 & 111& $\alpha^2 + \alpha + 1$ &$\rightarrow$& 011 &$\alpha + 1$\\
-%% \hline
-%% \end {tabular}
-%% }
-%% \end{center}
+\begin{center}
+{\small
+\begin{tabular}{c|ccc|c||c|ccc|c}
+$\{a_2a_1a_0\}$  & $A$ &$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$ &
+  $\{a_2a_1a_0\}$  &$A$&$\rightarrow$& $\{y_2y_1y_0\}$ &$Y$\\
+\hline
+000  &0 &$\rightarrow$& 000 & 0 & 100  &$\alpha^2$ &$\rightarrow$& 010 &  $\alpha$ \\
+001  &1 &$\rightarrow$& 000 & 0 & 101  &$\alpha^2 + 1$ &$\rightarrow$&010 & $\alpha$ \\
+010  &$\alpha$ & $\rightarrow$ & 001& 1 & 110  &$\alpha^2 + \alpha$&$\rightarrow$& 011 &$\alpha + 1$ \\   
+011  &$\alpha + 1$ &$\rightarrow$& 001 &1 & 111& $\alpha^2 + \alpha + 1$ &$\rightarrow$& 011 &$\alpha + 1$\\
+\hline
+\end {tabular}
+}
+\end{center}
 
-%% If we model this function as $f: {\mathbb{Z}}_{8} \rightarrow
-%% {\mathbb{Z}}_{8}$, then using the results of \cite{singmaster}
-%% \cite{chen_95} \cite{chen_96}, we deduce that this function is not a
-%% polynomial function over ${\mathbb{Z}}_{8}$. However, applying
-%% Largange's interpolation formula over ${\mathbb{F}}_{2^3}$, we obtain
-%% the following polynomial function representation, $Y =
-%% (\alpha^2+1)A^4+(\alpha^2+1)A^2$, where $P(\alpha) = \alpha^3 +
-%% \alpha + 1 = 0$. 
-%% }
-%% \end{Example}
+If we model this function as $f: {\mathbb{Z}}_{8} \rightarrow
+{\mathbb{Z}}_{8}$, then using the results of \cite{singmaster}
+\cite{chen_95} \cite{chen_96}, we deduce that this function is not a
+polynomial function over ${\mathbb{Z}}_{8}$. However, applying
+Largange's interpolation formula over ${\mathbb{F}}_{2^3}$, we obtain
+the following polynomial function representation, $Y =
+(\alpha^2+1)A^4+(\alpha^2+1)A^2$, where $P(\alpha) = \alpha^3 +
+\alpha + 1 = 0$. 
+}
+\end{Example}
 
@@
   - X}$ to obtain a canonical representation $\F(X) \pmod{X^q - X}$
-with degree at most $q-1$. 
-%Such {\it vanishing polynomials} will form
-%an important part of our formulation.
+with degree at most $q-1$. Such {\it vanishing polynomials} will form
+an important part of our ideal membership formulation.
 
@@
 composite-field multipliers \cite{cfmulti:1996} --- which are, in
-turn, hierarchically designed.  
-For example, Montgomery reduction (MR)
+turn, hierarchically designed.  For example, Montgomery reduction (MR)
 computes: $MR(A,B)=A\cdot B \cdot R^{-1} \pmod {P(x)}$, 
@@
 	\begin{center}
-	\includegraphics[scale=0.4]{../new_mmcircuit.eps}
+	\includegraphics[scale=0.4]{new_mmcircuit.eps}
 	\end{center}

### 2014-04-25 sequential/latex/abstraction/prev.tex
upload files from laptop linux

--- 2014-04-25 sequential/latex/abstraction/iwls/prev_iwls.tex
@@
 
-{\bf Canonical Decision Diagrams:} 
-The Reduced Ordered Binary Decision Diagram
-(ROBBD) \cite{BRYA86}, based on the Shannon's decomposition, was the
-first significant contribution in this area. 
-Motivated by the success of BDDs, 
+\subsection{Canonical Decision Diagrams}
+The problem of canonical representations of Boolean functions has
+received a lot of attention. The Reduced Ordered Binary Decision Diagram
+(ROBBD) \cite{BRYA86} was the first significant contribution in this
+area. 
+%Efficient implementation of ROBDDs as a software package
+%\cite{brace} allowed for efficient formal verification of
+%combinational and sequential circuits. 
+ROBDDs represent a Boolean function as an implicit set of points on
+a canonical directed acyclic graph (DAG). 
+%Manipulation of Boolean functions can then be carried
+%out as composition operations on their respective DAGs. 
+The decomposition principle behind BDDs is one of Shannon's expansion,
+i.e. $f(x, y, \dots) = x f_x + x' f_{x'}$, where $f_x = f(x = 1)$ and
+$f_{x'} = f(x = 0)$ denote the positive and negative co-factors of
+$f$ w.r.t. $x$, respectively. Motivated by the success of BDDs, 
 variants of the Shannon's decomposition principle (Davio,
 Reed-Muller, etc.) were explored to develop other functional decision
-diagrams: FDDs \cite{okfdd}, ADDs \cite{add}, %MTBDDs\cite{mtbdd},  
-HDDs \cite{hdd} and EVBDDs \cite{evbdd}. 
-%While these are %{\it curiously} still 
-%referred to as {\it Word-Level Decision Diagrams} \cite{WLS}, 
-Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivative K*BMDs
-\cite{kbmd}, % and *PHDDs \cite{phdd}, 
-depart from the Boolean
-decomposition and perform the decomposition of a {\it linear} function
-based on its two moments. The decomposition is still point-wise,
-binary, w.r.t. each Boolean variable; these do not
+diagrams. These include FDDs \cite{okfdd}, ADDs \cite{add}, MTBDDs
+\cite{mtbdd}, their hybrid and edge-valued counterparts, HDDs
+\cite{hdd} and EVBDDs \cite{evbdd}. While these are
+%{\it curiously} still 
+referred to as {\it Word-Level Decision Diagrams} \cite{WLS}, the
+decomposition is still point-wise, binary, 
+w.r.t. each Boolean variable. These representations do not
 serve the purpose of word-level abstraction from bit-level
@@
 
-%% BMDs provide a compact representation for
-%% integer arithmetic circuits such as multipliers and squarers. However,
-%% these are inapplicable for modulo-arithmetic circuits over Galois fields.
+Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivatives K*BMDs
+\cite{kbmd} and *PHDDs \cite{phdd}, depart from the Boolean
+decomposition and perform the decomposition of a {\it linear} function
+based on its two moments. BMDs provide a compact representation for
+integer arithmetic circuits such as multipliers and squarers. However,
+these are inapplicable for modulo-arithmetic circuits over Galois fields.
 %in our application, we encounter modulo-arithmetic circuits and that
@@
 %descriptions as polynomials over bit-vectors. However, 
-Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp} are a word-level
+Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp} are a
 canonical representation of a {\it polynomial expression}, but they do
@@
 %polynomials and they have different (non-isomorphic) TED DAGs. 
-While \cite{shekhar:tcad07}  and \cite{alizadeh:tcad2010} provide
+While \cite{shekhar:phd}  and \cite{alizadeh:tcad2010} provide
 canonical representations of polynomial functions, they do so over
@@
 this is because every node in the DAG can have up to $k$ children and
-the normalization operations are very complicated for MODDs. They 
+the normalization operations are very complicated for MODDs. They also
 suffer from the size explosion problem during intermediate
@@
 
-%In the realm of theorem-proving and SMT-solving, many automated
-%decision procedures have been devised that can decide
-%satisfiability/validity of word-level formulas represented by a
-%combination of theories, but they do not provide a canonical
-%representation of the function implemented by a ``bit-blasted''
-%circuit.
+In the realm of theorem-proving and SMT-solving, many automated
+decision procedures have been devised that can decide
+satisfiability/validity of word-level formulas represented by a
+combination of theories, but they do not provide a canonical
+representation of the function implemented by a ``bit-blasted''
+circuit.
 
-{\bf Verification of Galois field circuits:} 
-%% Symbolic computer algebra techniques have been employed for formal
-%% verification of circuits over $\Z_{2^k}$ and also over $\Fkk$. 
-%% %% The work of \cite{gao:gf-gb-ms} shows how to use Gr\"obner basis
-%% %% techniques to count the zeros of an ideal $J$ over ${\mathbb{F}}_q$
-%% %% (i.e. count $V_{{\mathbb{F}}_q}(J)$). The authors then follow-up with
-%% %% an approach for {\it quantifier elimination} over Galois fields
-%% %% ${\mathbb{F}}_q$ \cite{gao:qe-gf-gb}. While the above works present
-%% %% the theory and algorithms, efficiency/improvements in Gr\"obner basis
-%% %% computation is not addressed. So is also the case with other general
-%% %% verification techniques using Gr\"obner bases \cite{Avrunin:CAV}
-%% %% \cite{gbverify:2007} \cite{manna:program}, etc. 
-%% The paper \cite{wienand:cav08} addresses verification of
-%% finite-precision integer datapath circuits using the concepts of
-%% Gr\"obner bases over the ring ${\mathbb{Z}}_{2^k}$. 
-%% %They model the
-%% %circuit constraints by way of arithmetic-bit-level (ABL) polynomials
-%% %($\{G\}$) and formulate the verification test as an equivalent
-%% %{\it variety subset problem}. To solve this, first they derive a term
-%% %order to represent the circuit polynomials $\{G\}$ with a Gr\"obner
-%% %basis. Then they compute a normal form $f$ of the specification $g$
-%% %w.r.t. $\{G\}$. They show that, under certain constraints that are
-%% %satisfied by their term order, the circuit is correct if and only if
-%% %$f$ is a vanishing polynomial over ${\mathbb{Z}}_{2^k}$
-%% %\cite{shekhar:tcad07}. 
-%% In \cite{wedler:date11}, the authors further
-%% show an integration of their approach within an SMT-solver for
-%% arithmetic instances.  
-%% %show that  the vanishing polynomial test can be
-%% %omitted by formulating the problem directly over $Q :=
-%% %{\mathbb{Z}}_{2^k} [X]/\langle x^2-x : x \in X \rangle$. 
-In %\cite{ibm:ecc-ita11} \cite{blueveri:ita} 
-\cite{ibm:blueveri},
+\subsection{Prior work on Verification of Galois field circuits}
+
+Symbolic computer algebra techniques have been employed for formal
+verification of circuits over $\Z_{2^k}$ and also over $\Fkk$. 
+The work of \cite{gao:gf-gb-ms} shows how to use Gr\"obner basis
+techniques to count the zeros of an ideal $J$ over ${\mathbb{F}}_q$
+(i.e. count $V_{{\mathbb{F}}_q}(J)$). The authors then follow-up with
+an approach for {\it quantifier elimination} over Galois fields
+${\mathbb{F}}_q$ \cite{gao:qe-gf-gb}. While the above works present
+the theory and algorithms, efficiency/improvements in Gr\"obner basis
+computation is not addressed. So is also the case with other general
+verification techniques using Gr\"obner bases \cite{Avrunin:CAV}
+\cite{gbverify:2007} \cite{manna:program}, etc. 
+The paper \cite{wienand:cav08} addresses verification of
+finite-precision integer datapath circuits using the concepts of
+Gr\"obner bases over the ring ${\mathbb{Z}}_{2^k}$. 
+%They model the
+%circuit constraints by way of arithmetic-bit-level (ABL) polynomials
+%($\{G\}$) and formulate the verification test as an equivalent
+%{\it variety subset problem}. To solve this, first they derive a term
+%order to represent the circuit polynomials $\{G\}$ with a Gr\"obner
+%basis. Then they compute a normal form $f$ of the specification $g$
+%w.r.t. $\{G\}$. They show that, under certain constraints that are
+%satisfied by their term order, the circuit is correct if and only if
+%$f$ is a vanishing polynomial over ${\mathbb{Z}}_{2^k}$
+%\cite{shekhar:tcad07}. 
+In \cite{wedler:date11}, the authors further
+show an integration of their approach within an SMT-solver for
+arithmetic instances.  
+%show that  the vanishing polynomial test can be
+%omitted by formulating the problem directly over $Q :=
+%{\mathbb{Z}}_{2^k} [X]/\langle x^2-x : x \in X \rangle$.
+
+In \cite{ibm:ecc-ita11}
+\cite{blueveri:ita} \cite{ibm:blueveri},
 %from IBM needs special mention: T
@@
 
-In %\cite{lv:phd} 
-\cite{lv:date2012} 
-%\cite{lv:hldvt2011} \cite{lv:vlsi2012}, 
-{\it Lv et al.} present computer algebra
+In \cite{lv:phd} \cite{lv:date2012} \cite{lv:hldvt2011}
+\cite{lv:vlsi2012}, {\it Lv et al.} present computer algebra
 techniques for formal verification of Galois field arithmetic
@@
 to $163$-bit (NIST-specified) cryptography circuits. 
-In contrast to the work of \cite{lv:date2012}, we are not given a
+In contrast to the work of \cite{lv:phd}, we are not given a
 specification polynomial. Given the circuit $C$, we want
-to derive (extract) the word-level specification $f$. 
-In this work, we build upon the results of
-%\cite{gao:gf-gb-ms} \cite{gao:qe-gf-gb} %\cite{modd_tcomp} 
-\cite{lv:date2012}.
+to derive (extract) the word-level specification $f$. In our work, we
+borrow and further build upon the results of \cite{gao:gf-gb-ms}
+\cite{gao:qe-gf-gb} %\cite{modd_tcomp} 
+\cite{lv:date2012} \cite{lv:phd}.
 
-{\bf Polynomial Interpolation in Symbolic Computation:}
+\subsection{Polynomial Interpolation in Symbolic Computation}
+
 The problem of polynomial interpolation is a fundamental problem in
@@
 %\includegraphics[scale=0.3]{../figures/2bitmultiplier.eps}
-\includegraphics[scale=0.35]{../blackbox.eps}
+\includegraphics[scale=0.45]{blackbox.eps}
 }
@@
 
-Our problem falls into the category of {\it dense} interpolation, as
-we require a polynomial that describes the function at each of the $q$
-points of the field $\Fq$. Newton's dense interpolation technique,
-with the black-box model, bounds the number of probes by $(d+1)^n$ ---
-which  exhibits very high complexity. In the logic synthesis area, the
-work of \cite{zilic:interpolate} investigates dense interpolation. Due
-to the inherent high-complexity, their approach is feasible only for
-applications over small fields, {\it e.g.} computing  Reed-Muller
-forms for multi-valued logic.
+Our problem falls into the category of dense interpolation, as we
+require a polynomial that describes the function at each of the $q$
+points of the field $\Fq$. Newton's interpolation technique, with the
+black-box model, bounds the number of probes by $(d+1)^n$ --- which 
+exhibits very high complexity. In the logic synthesis area, the work of
+\cite{zilic:interpolate} investigates dense interpolation. Due to this
+high-complexity, their approach is feasible only for applications over
+small fields, {\it e.g.} computing  Reed-Muller forms for multi-valued
+logic.
 
@@
 polynomial interpolation from a circuit using the Gr\"obner basis
-computation. 
-%It can be shown that the computational complexity of
-%computing a Gr\"obner basis for our problem, in the worst case, is
-%$q^{O(n)}$, where $q = 2^k$, $k$ is the datapath width and $n$ is the
-%number of variables in the circuit.
+computation. It can be shown that the computational complexity of
+computing a Gr\"obner basis for our problem, in the worst case, is
+$q^{O(n)}$, where $q = 2^k$, $k$ is the datapath width and $n$ is the
+number of variables in the circuit.

### 2014-04-25 sequential/latex/abstraction/project.tex
upload files from laptop linux

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  IEEEsample.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%    More information: see the header of IEEEtran.sty
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%

\documentclass[conference,10pt]{IEEEtran}
%\documentclass[conference]{IEEEtran}

%%%\IEEEoverridecommandlockouts

%\usepackage[ruled]{./algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
%\renewcommand{\algorithmcfname}{ALGORITHM}
%\SetAlFnt{\small}
%\SetAlCapFnt{\small}
%\SetAlCapNameFnt{\small}
%\SetAlCapHSkip{0pt}
%\IncMargin{-\parindent}

%% \RequirePackage{times}
%% \RequirePackage{algorithmic}
%% \PassOptionsToPackage{boxed}{algorithm}
%% \RequirePackage{algorithm}
%% \RequirePackage{multicol}
%\renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
%\renewcommand{\algorithmicensure}{\textbf{Outputs:}}
%\DeclareMathAlphabet{\mathtsl}{OT1}{ptm}{m}{sl}

%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em1
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{example}{Example}
%\newtheorem{corollary}{Corollary}

\RequirePackage{amssymb, mathptm}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{comment}
\usepackage{leqno}
\usepackage{cases}
\usepackage{multirow}

\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{listings}
\lstset{language=C}

%%indent in algorithm

%\setcounter{page}{1}

% New command for the table notes.
\def\tabnote#1{{\small{#1}}}

% New command for the line spacing.
\newcommand{\ls}[1]
    {\dimen0=\fontdimen6\the\font
     \lineskip=#1\dimen0
     \advance\lineskip.5\fontdimen5\the\font
     \advance\lineskip-\dimen0
     \lineskiplimit=.9\lineskip
     \baselineskip=\lineskip
     \advance\baselineskip\dimen0
     \normallineskip\lineskip
     \normallineskiplimit\lineskiplimit
     \normalbaselineskip\baselineskip
     \ignorespaces
    }
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqarr}{\begin{eqnarray}}
\newcommand{\eeqarr}{\end{eqnarray}}
%\newcommand{\ov}{\overline}
\newcommand{\ov}{\bar}
\newcommand{\xor}{\bigoplus}
\newcommand{\Fm}{{\mathbb{F}}}

%the following is for space before and after align or other equation environment.

%%
\newtheorem{Algorithm}{Algorithm}[section]
\newtheorem{Definition}{Definition}[section]
\newtheorem{Example}{Example}[section]
\newtheorem{Proposition}{Proposition}[section]
\newtheorem{Lemma}{Lemma}[section]
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}{Corollary}[section]
\newtheorem{Conjecture}{Conjecture}[section]
\newtheorem{Problem}{Problem}[section]
\newtheorem{Notation}{Notation}[section]
\newtheorem{Setup}{Problem Setup}[section]
%%%

%%set spacing between table columns
\setlength{\tabcolsep}{3pt}

\begin{document}

%\thispagestyle{empty}
%\pagestyle{empty}

\title{Project Proposal: Implementing Flow-based Balanced Bipartition Algorithm}

\author{
\IEEEauthorblockN{Xiaojun Sun}
\IEEEauthorblockA{Electrical and Computer Engineering\\
University of Utah\\
Salt Lake City, Utah 84112\\
Email: xiaojuns@ece.utah.edu}
}

%%\author{\IEEEauthorblockN{Jinpeng Lv and Priyank Kalla}\thanks{This work is sponsored in part by a grant from NSF \#CCF-546859.}
%\IEEEauthorblockA{Department of  Electrical and Computer Eng.\\
% University of Utah, Salt Lake City, UT-84112 \vspace{-0.3in}
 %\{lv, kalla\}@eng.utah.edu
% }
%\and
%\IEEEauthorblockN{Florian Enescu} \thanks{\normalsize  978-3-9810801-8-6/DATE12/$\copyright 2012$ EDAA}
%\IEEEauthorblockA{Department of Mathematics and Statistics\\
% Georgia State University,  Atlanta, GA 30302-4038 \vspace{-0.3in}
% fenescu@mathstat.gsu.edu
%} 
%

\maketitle

%\markboth{MS Proposal by Tim Pruss}{}
\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
\newcommand{\B}{{\mathbb{B}}}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\F}{{\mathcal{F}}}
\newcommand{\G}{{\mathcal{G}}}
\newcommand{\R}{\mathbb{R}}

\newcommand{\HMetis}{{\scriptsize{H}}M{\scriptsize{E}}T{\scriptsize{I}}S}
%%%

\newcommand{\debug}[1]{\textcolor{gray}{[ #1 ]}}

%\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%% Include your files here %%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Flow-based balanced bipartition (FBB) algorithm is a max-flow/min-cut based circuit partitioning
heuristic. The core algorithm is revised from max-flow algorithm on graphs/hypergraphs while taking
the balancing criteria into consideration. Additionally, an incrementing max-flow algorithm
is proposed to improve the complexity of whole approach. My project will focus on implementing
FBB algorithm on variant netlists denoting by graphs and hypergraphs, exploring better initial
assignments for attaining global optimal results and doing reasonable modification on original
algorithms.
\end{abstract}

\section{Flow-based Bipartition Algorithm}
\label{sec:intro}

\subsection{Modeling a Net in a Flow Network}
A net in circuit netlist could be represented as an edge or a hyperedge in graphs/hypergraphs.
When we talk about the cost of cut set, it is defined as the number of nets in cut set. For
general max-flow/min-cut problems, they deal with graphs with ordinary edges (2-ends edge).
In order to utilize these high efficient max-flow/min-cut algorithms in our implementation,
it is necessary to translate a net (hyperedge) into an ordinary edge, i.e. model the circuit in
graphs which general max-flow/min-cut algorithm can deal with.
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.5]{figckt.png}
	\end{center}
	\caption{Circuit nets example}
	\label{fig:ckt}
\end{figure}

Fig.\ref{fig:ckt} is an example netlist with 2 different cut sets. In this figure, net $a$ and $b$
could be represented with hyperedges. Fig.\ref{fig:net} explains a method to translate such nets into
edges that could be dealt with general min-cut algorithm. 2 dummy vertices are inserted as long as a "bridging edge"
with weight 1. Once this net is cut, the bridging edge must be included in the cut set from the output
of any general min-cut algorithm, and cost is 1.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.5]{fignet.png}
	\end{center}
	\caption{Transform a Net to subgraph with bridging edge}
	\label{fig:net}
\end{figure}

\begin{algorithm}[hbt]
\caption{Algorithm for constructing flow network from circuit netlists}
\label{alg:construct}
\begin{algorithmic}[1]
\Function{Construct}{Netlist allowing hyperedges $N = (V,E)$}
  \State $V'\gets V$   /*copy all vertices*/
  \For{each net $n\in E$}
  	\State Insert auxiliary vertices $n_1$ and $n_2$ into $V'$
  	\State Insert bridging edge $e = (n_1,n_2)$ with unit weight into $E'$
  	\For{each node $u$ incident to $n$}
  		\State Insert edge $(u,n_1)$ and $(n_2,u)$ with infinite weight into $E'$
  	\EndFor
  \EndFor
  \State \Return graph $N' = (V',E')$
\EndFunction
\end{algorithmic}
\end{algorithm}

It showes from algorithm \ref{alg:construct} that the size of newly constructed graph will not explode.
And the following theorem guarantees that the result of general min-cut algorithm can be adopted.

\begin{Theorem}
$N$ has a cut of net whose size is at most $C$ if and only if $N'$ has a cut of capacity at most $C$.
\end{Theorem}

\subsection{Min-cut Balanced Bipartition}
It is easy to see a minimum cut of nets in netlist can be computed from general min-cut algorithm, based on
our netlist-graph translation. However, the min-cut cannot guarantee a balanced bipartition. It is necessary
to build an algorithm iterating with optimal min-cut and improve the balance every loop, which is the FBB algorithm we want to propose.

\begin{algorithm}
\caption{FBB Algorithm}
\label{alg:FBB}
\begin{algorithmic}[1]
\Function{FBB}{graph $G = (V,E)$}
  \State Randomly pick a pair of vertices as source $s$ and sink $t$, delete all edges starting from $t$ or
  ending at $s$
  \State $C \gets $\Call{MinCut}{$G$}
  \State $X\gets$ vertices from source part $\setminus C$
  \State $X'\gets$ vertices from sink part $\setminus C$
  \While{$C$ violates balancing criteria}
  	\If{$X$ is too small}
  		\State Collapse $X$ to $s$
  		\State Collapse to $s$ a vertex $v\in X'$ adjacent to $C$
  	\Else
  		\State Collapse $X'$ to $t$
  		\State Collapse to $t$ a vertex $v\in X$ adjacent to $C$
  	\EndIf
    \State $C \gets $\Call{MinCut}{$G$}
  \State $X\gets$ vertices from source part $\setminus C$
  \State $X'\gets$ vertices from sink part $\setminus C$
  \EndWhile
  \State \Return $C$
\EndFunction
\end{algorithmic}
\end{algorithm}

\begin{figure*}[hbt]
	\begin{center}
	\includegraphics[width=5in]{figfbb.png}
	\end{center}
	\caption{An example on FBB algorithm execution}
	\label{fig:FBB}
\end{figure*}

Fig.\ref{fig:FBB} illustrates what algorithm \ref{alg:FBB} does. First step, $s$ and $t$ are randomly chosen
from all vertices, and a min-cut $(X1,\overline{X1})$ of cost 2 is computed. Note that a small dot denote
a pair of auxiliary nodes we added for each net. Apparently this is not a balanced cut 
partition, so we merge $X$ ($\varnothing$ in first iteration) into $s$, then pick vertex $a$ from $X'$, also
merge it into $s$. In the second iteration, we run min-cut algorithm again, and get a cut $(X2,\overline{X2})$.
The cost is also 2, but this is still not balanced in the other direction. This time we merge $X'$ 
(vertices $e$ and $i$) into $t$, as well as vertex $h$ chosen from $X$. In third iteration, a cut $(X3,\overline{X3})$ of cost 3 satisfies the balancing criteria, which is returned as final result.

\subsection{Efficient Min-cut Computation}
The algorithm's complexity is retarded by computing min-cut within each iteration. In order to improve this,
a new incremental flow computation based algorithm is proposed. The main idea is to keep track of the flow
information from last iteration, and start the max-flow computing from this initial flow instead of zero flow.

\begin{algorithm}
\caption{Incremental Flow Min-cut}
\label{alg:incflow}
\begin{algorithmic}[1]
\Function{IncFlowMinCut}{residue flow graph $G'$}
	\State copy flow info from last iteration
  \While{$\exists$ an augmenting path from $s$ to $t$}
  	\State saturate this path
  \EndWhile
  \Comment/* there is no more augmenting path from $s$ to $t$ */
  \For{all vertices $u$}
  	\If{$\exists$ an augmenting path from $s$ to $u$}
  		\State Insert $u$ to set $X$
  	\EndIf
  \EndFor
  \For{all vertices $u\in X$}
  	\If{$e=(u,v)$ is bridging edge and $v\notin X$}
  		\State Insert $e$ into $C'$
  	\EndIf
  \EndFor
  \State \Return $C$ whose nets contain $C'$ as min-cut
  \State \Return $X$, and $V'\setminus(X\cup C')$ as $X'$
\EndFunction
\end{algorithmic}
\end{algorithm}

The refined algorithm with $k$ iterations is proved to have same complexity as a zero-flow min-cut algorithm.

\section{Experiment}
I plan to execute experiment on ISCAS benchmarks, and do comparison with K-L based tools and \emph{\HMetis} tool.
This heuristic depends on the initial assignment of source and sink, so FBB should be executed for multiple
times to test average performance.

%%%%%%%%%%%%%%%%%%%% The bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{1}
\bibitem{ref1}
Yang H H, Wong D F. \emph{Efficient network flow based min-cut balanced partitioning}[M], The Best of ICCAD. Springer US, 2003: 521-534.
\end{thebibliography}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-04-25 sequential/latex/proposal.tex
upload files from laptop linux

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{fullpage}
\usepackage{fancyhdr}
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=0.5cm,bottom=1cm}
\pagestyle{fancyplain}
\headheight 15pt
%\lhead{Xiaojun Sun\ (u0741290)}
%\chead{\textbf{\Large Homework \asgnname}}
%\rhead{\today}
%\headsep 10pt

\begin{document}

\title{Proposal: Arithmetic circuits verification using Network Flow model}
\author{Xiaojun Sun}
\maketitle

\section{Theory}
This paper discussed a new method modeling each gate and fanout branch by arithmetic signature expressions.
First, the compatibility for each cut set is computed using weight propagation. Second, based on a valid 
assignment on input signal that makes the circuit compatible, total flow conducting into fanout blocks is
equal to the flow consuming by floating signals, i.e.\\
\textbf{Theorem 1}\ \ \ The circuit is functionally correct if and only if:
\\
(i) There exists a compatible assignment of weights consistent with the input signature $Sig_{in}$;
\\(ii) The amount of the flow introduced by fanouts $\Delta_{fn}$ is equal to the flow consumed by floating signals $\Sigma_{fi}$.

Both the consistency checking and fanouts flow conservation checking are equivalence checking, which is 
accomplished by TED method in original paper. My work is to find a substitution method such as word-level
compatible computer algebraic method to verify the equivalence, thus check the functional correctness of
the arithmetic circuits.

\section{Implementation}
The test was running on as large as $62\times 62$ multiplier. The netlist is transformed to network of
half-adders, full-adders and other logic gates remained, which is easy to deduce signature expressions
to primary inputs signals.

Several cuts are selected, and for each cut the signature (flow) equation is checked equality with its higher 
level. If any difference detected, the residual expression must contain affected signals by the potential bug(s).

After the compatibility/consistency checking is completed, the flow generated by fanouts and consumed by floating
signals are compared to check equivalence. Then the functional correctness is asserted based on the conservative
condition of network flow model.

\begin{thebibliography}{1}
\bibitem{ref1}
Ciesielski, Maciej, Walter Brown, and André Rossi. \emph{"Arithmetic Bit-Level Verification Using Network Flow Model."} Hardware and Software: Verification and Testing. Springer International Publishing, 2013. 327-343.
\end{thebibliography}

\end{document}

### 2014-04-25 sequential/latex/topic.tex
upload files from laptop linux

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{fullpage}
\usepackage{fancyhdr}
\usepackage{cases}
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=1cm,bottom=2cm}
\pagestyle{fancyplain}
\headheight 15pt
%\lhead{Xiaojun Sun\ (u0741290)}
%\chead{\textbf{\Large Homework \asgnname}}
%\rhead{\today}
%\headsep 10pt

\begin{document}

\title{Illustration on my write-up and some issues to address}
\author{Xiaojun Sun}
\maketitle

\section{About my write-up}
My write-up (attached in same tar) is the theory part of proposed ICCAD paper with examples in details. I think
the essence of our theory should focus on the state space traversal on general FSM part, because this part
introduces algebraic geometry concepts (sum, product and quotient of ideals). The quotient of ideal concept has never
been used to describe state transition (including complement) based on my survey, and its soundness has been
validated by Florian's proof.

A general example circuit/FSM, please refer to \textbf{Part 2.1, 2.2} and {\bf 2.3}, basically I had done a step-by-step comment on 
this example. Also I attach the \emph{Singular} file with full comments, please try to run this \emph{traversal.sing}.

There are limitations on my technique, one is that the polynomial (ideal generator) must be univariate, so using
Tim's abstraction is necessary. As to the normal basis multiplier, since this implementation is like a one-way
FSM, the state transition graph is a chain, there is no need to use techniques like ideal sum/product/quotient.
Thus the main contribution of this multiplier is providing an sequential application of Tim's abstraction technique,
but it can still fit our state-space-traversal framework (see last paragraph of \textbf{Part 3.1}, then \textbf{Part 3.2}).

Still, in this multiplier example, I can propose another contribution on fast Gr\"obner basis computing (see \textbf{Part 4.2}),
it is originally from Tim's idea, but not exactly the same, because the property of Normal Basis allows us
to square the polynomial conveniently.

\section{Soundness of Optimal Normal Basis Theory}
This is more likely related to mathematical stuff rather than computer engineering. My following questions are necessary to
prove the soundness of the SAT based technique to compute optimal normal element.

(a) We know how to get the $\lambda$-Matrix from a set of criteria $C$ such as
\begin{equation}
\begin{cases}
2^i + 2^j = 1 \ \mod\ \  2n+1\\
2^i + 2^j = -1 \ \ mod \ \ 2n+1\\
2^i - 2^j = 1 \ \ mod\ \  2n+1\\
2^i - 2^j = -1 \ \ mod \ \ 2n+1
\end{cases}
\end{equation}
but a formal proof of following equivalence is needed:
$$N\text{ is type-II optimal normal basis} \Leftrightarrow N\text{ is the only solution of }\lambda\text{-Matrix } M^0\text{generated by }C$$
Considering Rosing's book already proved the sufficiency, we want to confirm the necessity:
$$N\text{ is one solution of }\lambda\text{-Matrix } M^0\text{generated by }C \implies N\text{ is type-II optimal normal basis}$$

(b) This relates to the approach from Florian's student: to convert the $\lambda$-Matrix satisfiability problem to solving system
of boolean equations (then we can solve the system using SAT/BDD). I also include my questions in {\bf Part B.2} last paragraph of my write-up,
here I state more formally:

Prove: solutions to system of boolean equations are always conjugates;

Prove: choosing equation on $\beta^3$ is sufficient to cover all solutions to original $\lambda$-Matrix.\\
\vspace{5mm}\\
If we proved all above, then it is sufficient to write a paper on "methodology to compute optimal normal basis".

\section{Challenges on 32/64-bit multiplier experiment}
I think the main thing lagging my tool is the bad performance for \emph{Singular} to convert a high-degree power representation ring variable
to its standard basis representation. For example, if the minimal polynomial is
$$p(\alpha) = \alpha^{33} + \alpha^{27} + \alpha^{20} + \alpha^{19} + \alpha^{17} + \alpha^{11} + \alpha^5 + \alpha + 1$$
and the normal element (in std basis) is 
$$\beta = 1+\alpha^2+\alpha^4+\alpha^6+\alpha^7+\alpha^9+\alpha^{14}+\alpha^{15}+\alpha^{16}+\alpha^{19}+\alpha^{21}+\alpha^{22}+\alpha^{26}+\alpha^{27}+\alpha^{29}+\alpha^{30}+\alpha^{32}$$
Compute $\beta^{4294967296}$ (in std basis).

%\begin{thebibliography}{1}
%\bibitem{ref1}
%J. Lv, \emph{Scalable Formal Verification of Finite Field Arithmetic Circuits using Computer Algebra Techniques}, Ph.D. Dissertation, The University of Utah, 2012
%\end{thebibliography}

\end{document}

### 2014-04-25 sequential/csource/SAplacement.c

@@ -1,0 +1,35 @@
+# include "stdlib.h"
+#include "stdio.h"
+# include "time.h"
+
+#define dm 8
+int a[dm];
+void permutation()
+{
+ int b;                       //定义一个变量存取随机生成的一个整数
+ int j,i=0;          
+ srand((unsigned)time(NULL));
+ a[0]=rand()%dm+1;
+ for(i=1;i<dm;i++)           //循环生成随机数，并存取在数组a[]中。
+ {
+  srand((unsigned)time(NULL));
+  loop:   b=rand()%dm+1;
+     for(j=0;j<=i;j++)       //把后一个生成的数和前面的数一一对比，若是有相同的GOTO LOOP再产生。
+  {
+        if(b==a[j])
+        goto loop;
+  }
+     a[i]=b;
+  }
+/* for(i=0;i<dm;i++)
+ {printf("%d ",a[i]);}
+ printf("\n");*/
+}
+
+int
+
+int metropolis(int S, int T, int M)
+void main()
+{
+  permutation();
+}

### 2014-04-25 sequential/csource/bcp_source.cpp

@@ -1,0 +1,484 @@
+#include<iostream>
+#include<fstream>
+#include<stdlib.h>
+#include<string>
+#include "bcphead.h"
+
+using namespace std;
+
+/*******************************************************************************************************/
+/**************************                Class Functions                ******************************/
+/*******************************************************************************************************/
+
+Matrix::Matrix(int x, int y)  //Construct new object, allocate memory, set column tag all to -1(unknown)
+{
+	m = x;
+	n = y;
+	noSolu = false;
+	A = new int *[x];
+	int *tmp = new int [x*y];
+	for(int i = 0; i < x; i++)
+		A[i] = &(tmp[i*y]);
+	len = new int [x];
+	col = new int [y];
+	for(int j = 0; j < n; j++)
+		col[j] = -1;
+}
+
+void Matrix::update_len()  //Update length of row, call this when finish intializing or setting up column tag
+{
+	for (int i = 0; i < m ; i++){
+		if(len[i] == 0) continue; // Never update deleted rows
+		int cnt = 0;
+		for(int j = 0; j < n; j++){
+			if( (A[i][j] != -1) && (col[j] == -1) ) // Count only if entries are available
+				cnt++;
+		}
+		len[i] = cnt;
+	}
+}
+
+Matrix::Matrix(const Matrix &s) // Copy-constructor, used for creating formal arguments
+{
+	m = s.m;
+	n = s.n;
+	noSolu = s.noSolu;
+	int i,j;
+	A = new int *[m];
+	int *tmp = new int [m*n];
+	for(i = 0; i < m; i++)
+		A[i] = &(tmp[i*n]);
+	len = new int [m];
+	col = new int [n];
+	for (i = 0; i < m; i++){  // Copy everything
+		len[i] = s.rowLen(i);
+		for (j = 0; j < n; j++)
+			assign_A(i, j, s.read_A(i,j));
+	}
+	for(j = 0; j<n; j++)
+		col[j] = s.xValue(j);
+}
+
[…]

### 2014-04-25 sequential/csource/bcphead.h

@@ -1,0 +1,28 @@
+#ifndef BCPHEAD_H
+#define BCPHEAD_H
+
+class Matrix
+{
+private:
+	int **A;
+	int *len;
+	int *col;
+public:
+	int m,n;
+	bool noSolu;
+	Matrix(int,int);
+	Matrix(const Matrix &);
+	int read_A(int,int) const;
+	~Matrix();
+	int rowLen(int) const;
+	int xValue(int) const;
+	void assign_A(int,int,int);
+	void update_len();
+	void deleteRow(int);
+	void setColZero(int);
+	void setColOne(int);
+	bool dom_R(int,int);
+	bool dom_C(int,int);
+};
+
+#endif

### 2014-04-25 sequential/csource/binary.c

@@ -1,0 +1,49 @@
+/* binary search -- maintained by Xiaojun Sun   ********************************************
+ * Implemented with C
+ * Compiled and linked with gcc-4.6.3
+ * Test cases borrowed from http://www.cs.utsa.edu/~wagner/CS3343/recursion/binsearch.html
+ *******************************************************************************************/
+/* desired output for this test case:   ****************************************************
+ * 0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 
+ * -1 -1 -1 -1
+ *******************************************************************************************/
+
+#include <stdio.h>
+#include <stdbool.h>
+#define LEN 29
+
+int binarysearch(int MAX, int *A, int item) {
+   int low = 0;
+   int high = MAX;
+   int loc = -1;
+   bool found = false;
+   while ((low < high) && !found) {
+      int mid = (low + high - 1) / 2;
+      if (A[mid] == item) 
+	  {
+		found = true;
+		loc = mid;
+	  }
+      else if (A[mid] < item) low = mid + 1;
+      else high = mid;
+   }
+   return loc;
+}
+
+int main() {
+
+   int A[] =
+    { 2, 8,12,14,16,19,24,28,31,33,// 0-9
+     39,40,45,49,51,53,54,56,57,60,// 10-19
+     63,69,77,82,88,89,94,96,97};  // 20-28
+   int i;
+   for (i = 0; i < 29; i++)
+      printf("%i ", binarysearch(LEN, A, A[i]));
+   printf("\n");
+   printf("%i ", binarysearch(LEN, A, 1));
+   printf("%i ", binarysearch(LEN, A, 17));
+   printf("%i ", binarysearch(LEN, A, 42));
+   printf("%i ", binarysearch(LEN, A, 99));
+   printf("\n");
+   return 0;
+}

### 2014-04-25 sequential/csource/matrix.cpp

@@ -1,0 +1,106 @@
+#include <iostream>
+#include <cstring>
+#include <cmath>
+using namespace std;
+
+bool onb(double a, double b, int n)
+{
+	long sum, diff, power_a, power_b;
+	power_a = long(pow(2,a))%(2*n+1);
+	power_b = long(pow(2,b))%(2*n+1);
+	if (a>16) 
+		power_a = (long(pow(2,16))%(2*n+1)) * (long(pow(2,(a - 16)))%(2*n+1));
+	if (b>16) 
+		power_b = (long(pow(2,16))%(2*n+1)) * (long(pow(2,(b - 16)))%(2*n+1));
+	sum = power_a + power_b;
+	diff = power_a - power_b;
+	while(diff < 0)
+		diff += (2*n+1);
+	sum = sum%(2*n+1);
+	diff = diff%(2*n+1);
+	if (sum == 1 || sum == (2*n) || diff == 1 || diff == (2*n))
+		return 1;
+	else return 0;
+}
+
+void beta_cube(int **idx, int n)
+{
+	int i,d;
+	cout<<endl;
+	if(idx[0][0] == 1)
+		cout<<"0 ";
+	if(idx[n-1][0] == 0)
+		cout<<"1 ";
+	for(i=n-2;i>0;i--)
+	{
+		d = i+1;
+		if((d == idx[i][0]) || (d == idx[i][1]))
+			cout<<n-i<<" ";
+	}
+	cout<<endl;
+}
+
+int main()
+{
+	int n = 33; // You can change n, currently plz keep n<64
+	bool **l=new bool*[n];
+	int **idx = new int*[n];
+	for(int k=0;k<n;k++)
+	{
+		l[k] = new bool[n];
+		idx[k] = new int[2];
+	}
+//	bool l[11][11];
+//	int idx[11][2];
+	int cnt=0;
+	int x, ip, jp;
+	double i = 0, j = 0;
+	for(i=0; i<n; i++)
+		memset(l[int(i)],0,n);
+	for(i=0; i<n; i++)
+		memset(idx[int(i)],0,2);
+	for(i=0; i < n; i++){
+		cnt = 0;
+		for(j = 0; j<n; j++){
+			if(onb(i,j,n) == 1){
+			    ip = ( int(j) - int(i) + n )%n;
+				jp = ( n - int(i) )%n;
[…]

### 2014-04-25 sequential/csource/maxsubarray.cpp

@@ -1,0 +1,31 @@
+#include<iostream>
+using namespace std;
+
+int Maxsum_ultimate(int * arr, int size)
+{
+    int maxSum = -9999;
+    int sum = 0;
+    for(int i = 0; i < size; ++i)
+    {
+        if(sum < 0)
+        {
+            sum = arr[i];
+        }else
+        {
+            sum += arr[i];
+        }
+        if(sum > maxSum)
+        {
+            maxSum = sum;
+        }
+    }
+    return maxSum;
+}
+
+int main()
+{
+	int a[] = {-5,-4,-3,-1,-1,-2};
+	int max = Maxsum_ultimate(a,6);
+	cout<<"result is"<<max<<endl;
+	return 0;
+}

### 2014-04-25 sequential/csource/repeating.cpp

@@ -1,0 +1,17 @@
+#include <iostream>
+
+using namespace std;
+
+int main(){
+  int i,j;
+  char *word;
+  cout<<"Plz input prefix:";
+  cin>>word;
+  cout<<"Plz input end No.:";
+  cin>>j;
+  for(i = 0; i <= j; i++){
+    cout<<word<<i<<", ";
+  }
+  cout<<endl;
+  return 0;
+}

### 2014-04-25 sequential/csource/type1onb.cpp

@@ -1,0 +1,71 @@
+#include <iostream>
+#include <cstring>
+#include <cmath>
+using namespace std;
+
+#define M 100
+
+bool onb(double a, double b)
+{
+	long sum;// diff;
+	long power_a = 1;
+	long power_b = 1;
+	
+	while (a>16) { 
+		power_a *= (long(pow(2,16))%(M+1));
+		a -= 16;
+	}
+	power_a *= long(pow(2,a))%(M+1);
+
+	while (b>16) { 
+		power_b *= (long(pow(2,16))%(M+1));
+		b -= 16;
+	}
+	power_b *= long(pow(2,b))%(M+1);
+
+	sum = power_a + power_b;
+//	diff = power_a - power_b;
+//	while(diff < 0)
+//		diff += 131;
+	sum = sum%(M+1);
+//	diff = diff%131;
+	if (sum == 1 || sum == 0)
+		return 1;
+	else return 0;
+}
+
+int main()
+{
+	bool l[M][M];
+	int idx[M][2];
+	int cnt=0;
+	int x;
+	double i = 0, j = 0;
+	for(i=0; i<M; i++)
+		memset(l[int(i)],0,M);
+	for(i=0; i<M; i++){
+		idx[int(i)][0] = -1;
+		idx[int(i)][1] = -1;
+	}
+	for(i=0; i < M; i++){
+		cnt = 0;
+		for(j = 0; j<M; j++){
+			if(onb(i,j) == 1){
+				l[int(i)][int(j)] = 1;
+				idx[int(i)][cnt] = int(j);
+				cnt++;
+			}
+		}
+	}
+	for(i=0;i<M;i++){
+		for(j=0;j<M;j++)
+			cout<<l[int(i)][int(j)]<<" ";
+		cout<<endl;
+	}
+	
+	for(x=0;x<M;x++)
+		cout<<idx[x][0]<<" "<<idx[x][1]<<" ";
+	cout<<endl;
+	return 0;
+}
+

### 2014-04-25 sequential/perlfile/11bit.pl

@@ -1,0 +1,98 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+open SING, ">", "11bitnew.sing";
+
+my @idx = qw /1 0 8 6 8 4 5 3 9 3 7 2 9 5 10 1 2 4 6 7 10/;
+
+=thisismultiplicationtableof11bit
+0 1 0 0 0 0 0 0 0 0 0 
+1 0 0 0 0 0 0 0 1 0 0 
+0 0 0 0 0 0 1 0 1 0 0 
+0 0 0 0 1 1 0 0 0 0 0 
+0 0 0 1 0 0 0 0 0 1 0 
+0 0 0 1 0 0 0 1 0 0 0 
+0 0 1 0 0 0 0 0 0 1 0 
+0 0 0 0 0 1 0 0 0 0 1 
+0 1 1 0 0 0 0 0 0 0 0 
+0 0 0 0 1 0 1 0 0 0 0 
+0 0 0 0 0 0 0 1 0 0 1 
+1 0 8 6 8 4 5 3 9 3 7 2 9 5 10 1 2 4 6 7 10
+=cut
+ 
+#print SING "/*\n";
+
+#my $i,$j,$c,$rr,$R;
+my $i = 1;
+my $c = 1;
+while($i < 20){
+ printf SING "array_A[%d]+array_A[%d]+c%d,\n", ($idx[$i]+$c)%11+1, ($idx[$i+1]+$c)%11+1, $c;
+ $i += 2;
+ $c ++;
+}
+print SING "/***************************************************/\n";
+
+print SING "array_A[2]*array_B[1]+r10+R0,\n";
+$i = 1; 
+$c = 1;
+while($i < 11){
+ printf SING "c%d*array_B[%d]+r%d+R%d,\n", $c, ($i+$c)%11+1, $c-1, $c;
+ $i ++;
+ $c ++;
+}
+print SING "/***************************************************/\n";
+
+$i = 1;
+$c = 293;
+print SING "A";
+while($i < 12){
+  printf SING "+a%d*X^%d", $i-1, $c;
+  $i ++;
+  $c = $c*2%2047;
+}
+print SING ",\n";
+
+$i = 1;
+$c = 293;
+print SING "B";
+while($i < 12){
+  printf SING "+b%d*X^%d", $i-1, $c;
+  $i ++;
+  $c = $c*2%2047;
+}
+print SING ",\n";
+
+$i = 1;
+$c = 293;
[…]

### 2014-04-25 sequential/perlfile/33bit_or_more.pl

@@ -1,0 +1,215 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+open SING, ">", "100bitnew.sing"; # if you want keep all copies, change this filename
+
+my $size = 100; # remember to change the size!!
+
+#              1 0 8 6 8 4 5 3 9 3 7 2 9 5 10 1 2 4 6 7 10
+#my @idx = qw /1 0 8 6 8 4 5 3 9 3 7 2 9 5 10 1 2 4 6 7 10/;
+my @idx = qw /50 50 51 19 52 53 59 43 54 34 55 56 97 17 57 58 73 20 59 16 60 57 61 62 87 63 75 28 64 65 92 10 66 7 67 23 68 2 69 9 70 66 71 48 72 18 73 52 74 75 88 46 76 77 85 14 78 42 79 54 80 32 81 31 82 81 83 5 84 47 85 72 86 58 87 53 88 80 89 83 90 91 94 29 92 4 93 84 94 77 95 26 96 35 97 22 98 68 99 0 1 1 69 2 24 3 38 4 30 5 82 6 90 7 11 8 37 3 9 10 93 11 91 12 65 13 71 14 86 15 62 16 21 17 98 18 49 19 51 20 74 21 63 22 36 8 23 24 70 13 25 26 78 27 45 28 76 29 95 30 39 31 33 32 55 33 40 34 44 27 35 36 64 12 37 25 38 39 96 40 56 41 61 15 42 43 60 41 44 45 79 46 89 6 47 48 67 49 99/;
+
+=thisismultiplicationtableof11bit
+0 1 0 0 0 0 0 0 0 0 0 
+1 0 0 0 0 0 0 0 1 0 0 
+0 0 0 0 0 0 1 0 1 0 0 
+0 0 0 0 1 1 0 0 0 0 0 
+0 0 0 1 0 0 0 0 0 1 0 
+0 0 0 1 0 0 0 1 0 0 0 
+0 0 1 0 0 0 0 0 0 1 0 
+0 0 0 0 0 1 0 0 0 0 1 
+0 1 1 0 0 0 0 0 0 0 0 
+0 0 0 0 1 0 1 0 0 0 0 
+0 0 0 0 0 0 0 1 0 0 1 
+1 0 8 6 8 4 5 3 9 3 7 2 9 5 10 1 2 4 6 7 10
+=cut
+ 
+#print SING "/*\n";
+
+#my $i,$j,$c,$rr,$R;
+my $i = 1;
[…]

### 2014-04-25 sequential/perlfile/quick_perl.pl

@@ -1,0 +1,61 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+#open SING, ">", "temp.sing";
+
+=outputStdBasisPoly
+my $i = 1;
+print "f = a0 + a1*X";
+for($i=2;$i<33;$i++)
+{
+  print " + a$i"."*X^$i";
+}
+print "\n";
+=cut
+
+=vanishIdeal
+my $i = 0;
+for($i=0;$i<33;$i++)
+{
+  print "a$i"."^2+a$i, ";
+}
+print "\n";
+=cut
+
+=fromSATcntResult
+my @cnt = qw /1 -2 3 -4 5 -6 7 8 -9 10 -11 -12 -13 -14 15 16 17 -18 -19 20 -21 22 23 -24 -25 -26 27 28 -29 30 31 -32 33/;
+my $i = 1;
+for($i=0;$i<33;$i++)
+{
+  if($cnt[$i] > 0){
+  	print "X^$i+";
+  }
+}
+print "\n";
+=cut
+
+my $i = 0;
+for($i=0;$i<33;$i++)
+{
+  print "R$i,";
+}
+for($i=0;$i<33;$i++)
+{
+  print "r$i,";
+}
+for($i=1;$i<33;$i++)
+{
+  print "c$i,";
+}
+for($i=0;$i<33;$i++)
+{
+  print "b$i,";
+}
+for($i=0;$i<33;$i++)
+{
+  print "a$i,";
+}
+print "R,r,A,B;\n";

### 2014-04-25 sequential/perlfile/transf.pl

@@ -1,0 +1,386 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+open BEDROCK, ">", "var_list.dat";
+open CPLEX, ">", "output.dat";
+
+my $readin;  #read-in lines (currently working on)
+my @var;  #list of i/o and wires
+my @items;  #splitted elements in word within one line
+my $cnt_tmp = 0;  #count of temporary variable
+my $cnt_lines = 0;  #No. of read in lines
+my $i;  #okay we always use it to count
+my $def_var;  #just record i/o or wire name for current processing line
+my $flag_unit = 0;  #judge if now there's any unit calc
+my $cnt_unit = 0;  #count for total potential unit calc 
+my $unit = 99;  # position of current unit operator
+my %var_list;  # hash table used to store exist variables with their bits(width)
+my $subject = 0;
+
+while (<>) {
+	chomp;
+	$cnt_lines ++;
+	print CPLEX "\n\#Constrains for line $cnt_lines :\n";
+	$readin = $_;
+	@items = split /\s+/;
+	$i = 1;
+	my $flag_io = 0;
+	my $flag_eq = 0;
+	my $cnt_eq = 0;
+	next unless $items[1];
+	while ( $items[$i] ) {
+		$_ = $items[$i];
+		if (/input|output/) {$flag_io = 1;}
+		if (/=/ && !$flag_eq)
+		{
+			$flag_eq = 1;
+			$cnt_eq = $i;
+		}
+		$i ++;
+	}
+	$_ = "";  
+	my $leng = $i-1;
+	if ($flag_io) {$_ = $items[$leng]; s/;/\n/; chomp; print BEDROCK "var $_;\n"; } # -1 more is [] info, if no [] then max 1
+	if ($flag_eq) {$_ = $items[$cnt_eq-1]; print BEDROCK "var $_;\n"; }
+	chomp;
+	if ($_) {push @var, $_; $def_var = $_;}
+	$_ = $readin;
+	if (/assign/) {pop @var;}  # remove existed var if assign sentence recognized
+	#before shift, need upper limit constrain based on bits definition
+	elsif ($flag_io || $flag_eq) {
+#		print CPLEX "$items[2]\n";
+		if ($items[2] =~ /\[(\d+)\:/) {
[…]

### 2014-04-25 sequential/perlfile/xorsat.pl

@@ -1,0 +1,63 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+open BEDROCK, ">", "xor_o.dat";
+open LAVA, ">", "new_o.dat";
+
+my %record;
+my $readin;
+my @items;
+my $n = 33; ################## This is ONB size
+
+while(<>){
+	chomp;
+	$readin = $_;
+	@items = split /\+/;
+	print BEDROCK "x-";
+	my $i = 0;
+	while($items[$i])
+	{
+		if($items[$i] =~ /\*/)
+		{
+			if(exists $record{$items[$i]})
+			{
+				print BEDROCK $n+1+$record{$items[$i]};
+				print BEDROCK " ";
+			}
+			else
+			{
+				my $k = &create_new($items[$i]);
+				print BEDROCK $n+1+$k;
+				print BEDROCK " ";
+			}
+		}
+		else
+		{
+			my $idx = substr($items[$i],1);
+			print BEDROCK $idx+1;
+			print BEDROCK " ";
+		}
+		$i++;
+	}
+	print BEDROCK "0\n";
+}
+my $size = keys %record;
+printf LAVA "p cnf %d %d", $n+$size, $n+3*$size;
+
+sub create_new {
+  my $arg = $_[0];
+  $_ = $arg;
+  chomp;
+  my @s0 = split /\*/;
+  my $s1 = substr($s0[0],1);
+  my $s2 = substr($s0[1],1);
+  my $s3 = keys %record;
+  printf LAVA "-%d -%d %d 0\n", $s1+1, $s2+1, $n+$s3+1;
+  printf LAVA "%d -%d 0\n", $s1+1, $n+$s3+1;
+  printf LAVA "%d -%d 0\n", $s2+1, $n+$s3+1;
+  $record{$arg} = $s3;
+  return $s3;
+}

### 2014-04-25 sequential/reduce_tool/alphaPoly.cpp

@@ -1,0 +1,756 @@
+#include "alphaPoly.h"
+#include <sstream>
+#include <map>
+#include <string>
+#include <iostream>
+#include <stdio.h>
+#include <string.h>
+#include <stdlib.h>
+#include <vector>
+#include <ctype.h>
+
+using namespace std;
+
+static unsigned int ringSize = 0;
+static map<string,int> theRing; 
+
+void Ring::setRing(string ringList)
+{
+	size_t index;
+	size_t lastIndex = 0;
+	ringSize = 0;
+	theRing.clear();
+
+	// Remove all spaces
+	index = ringList.find_first_of(" ");
+	while(index != string::npos)
+	{
+		ringList.erase(index,1);
+		index=ringList.find_first_of(" ",index);
+	}
+
+	index = ringList.find_first_of(",");
+	while(index != string::npos)
+	{
+		ringSize++;
+		theRing[ringList.substr(lastIndex, index-lastIndex)] = ringSize;
+		lastIndex=index+1;
+		index=ringList.find_first_of(",",lastIndex);
+	}
+	ringSize++;
+	theRing[ringList.substr(lastIndex)] = ringSize;
+}
+
+int Ring::getTermPosition(string termName)
+{
+	return theRing[termName];
+}
+
+string Ring::getTermName(int position)
+{
+	for (map<string,int>::iterator it=theRing.begin(); it!=theRing.end(); ++it)
+	{
+		if( it->second == position )
+			return it->first;
+	}
+	return "";
+}
+
+void Ring::printRing(void)
+{
+	cout<<"(";
+	for(int i = 1; i <= ringSize; i++)
+	{
+		if( i!=1 )
+			cout<<", ";
+		cout<<Ring::getTermName(i);
+	}
+	cout<<")\n";
+}
+
+
+
+// Word Level if first char is upper case
+bool Ring::isWordLevel(string termName)
+{
+	return (isupper(termName[0]));
+}
+
+
+bool Ring::isWordLevel(int position)
+{
+	return isWordLevel(getTermName(position));
+}
+
+
+
+// Start AlphaPoly functions
+/* These are the static variables of AlphaPoly.
+ * They are set when setting the minpoly
+ * K = Size of field extension as in F(2^k) (i.e. word size )
[…]

### 2014-04-25 sequential/reduce_tool/alphaPoly.h

@@ -1,0 +1,146 @@
+#pragma once
+
+#include <string>
+using namespace std;
+
+/* NOTE!! setMinpoly must be called before any other AlphaPoly function.
+ */
+class AlphaPoly
+{
+	private:
+		/*
+		 * An AlphaPoly is a Galois Field element, such as "X^4+X+1", where X is alpha.
+         * Each element is stored as a bit list of exponents "exp" in increasing order.
+		 * So the first bit of the first byte of exp represents 1, the next bit represents X, etc.
+         * For example, "X^4+X+1" would fit in 1 unsigned char as follows: exp[0] = 1'b00010011
+         * Larger fields require more bytes, 
+		 *  For instance, "X^10+X^6+X+1": exp[1] = 1'b00000100, exp[0] = 1'b01000011
+         * exp does not store any exponents K or greater (where K is defined by the minpoly)
+         * So for minpoly "X^4+X+1", the element "X^4+X^2" would be stored as "X^2+X+1"
+         */
+		unsigned char* exp;
+
+	public:
+
+        // MUST BE CALLED FIRST BEFORE ANY OTHER FUNCTION
+		// poly is a minimum polynomial string.
+		// Term must be X.
+		// Highest term must be the first one (k)
+		// ex: X^3 + X^2 + 1
+		static void setMinpoly(string poly);
+
+        // Returns the size of the field, k
+        // ex: if minpoly was set to X^3 + X^2 + 1, returns 3
+		static int getK(void);
+
+
+		// Create alpha poly object from poly string (where X is alpha)
+		// i.e. "X^4+X+1"
+		AlphaPoly(string poly);
+
+		// Create poly object from exponent
+        // i.e AlphaPoly(4) -> "X^4"
+		AlphaPoly(unsigned long expon);
+
+		// Create poly object by multiplying or adding two different polys
+        // doMult = true - > Multiply one and two
+        // doMult = false -> Add one and two
+        // one and two are not modified 
+		AlphaPoly(AlphaPoly* one, AlphaPoly* two, bool doMult);
+
+		// Create an exact duplicate of the given AlphaPoly
+        // copy is not modified
+		AlphaPoly(AlphaPoly* copy);
+
+		// Create AlphaPoly element 0
+		AlphaPoly(void);
+
+		// Delete AlphaPoly
+		~AlphaPoly(void);
+
+		// Multiply this AlphaPoly by other
+        // other is not modified
+		void mult(AlphaPoly* other);
+
[…]

### 2014-04-25 sequential/reduce_tool/main.cpp

@@ -1,0 +1,39 @@
+#include "alphaPoly.h"
+#include<vector>
+#include<list>
+#include<string>
+#include<sstream>
+#include<map>
+#include<set>
+#include<fstream>
+#include<algorithm>
+#include<iostream>
+using namespace std;
+
+extern void cavmain(char *filename);
+extern double diffms;
+int main (int argc, char* argv[]) 
+{
+	if(argc!=2)
+	{
+		cout<<"Usage: ./cav input.singular"<<endl;
+		return 1;
+	}
+
+	clock_t t;
+	t= clock();
+
+	cavmain(argv[1]);
+
+	t = clock() - t;
+	
+	/*AlphaPoly::setMinpoly("X^8+X^4+X^3+X+1");
+	AlphaPoly one = AlphaPoly("X^6+X^4+X+1");
+
+	one.getInverse();*/
+	
+	cout<<"Total verification time is: "<<((t*1000)/CLOCKS_PER_SEC)<<" ms."<<endl;
+
+	return 0;
+}
+

### 2014-05-18 sequential/latex/ICCAD2014/exp_result.tex
FMCAD latex update

\section{Experimental Results}
\label{sec:result}
Using our approach described above, we have performed experiments to verify SMPO circuits over $\Fkk$ with \textsc{Singular} symbolic algebra computation system [v. 3-1-6]\cite{Singular}. 
Bugs are also introduced into SMPO designs for runtime comparison. To show the advantages of our approach, experiments using traditional methods such as SAT, 
BDD are also conducted. Our experiments run on a desktop with 3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and 64-bit Linux.

\subsection{Evaluation of SAT/BDD based methods}
For every SMPO design, there is an equation in following form (this is a 5-bit example):
\begin{align}
R_i &= b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
\end{align}
This equation can be used as specification of SMPO circuit. For SAT/BDD based methods,
we first unroll our SMPO design, then create a \emph{miter} with the unrolled design and
circuit specification to check if it is unsatisfiable.

The SAT solver we use is \emph{Lingeling}\cite{Lingeling}. Single-output miter circuit is written
in DIMACS CNF\cite{dimacs} format and fed to Lingeling. For BDD based approach, we use the BDD module integrated
with VIS tool. The miter is described by a BLIF file and read by VIS\cite{VIS}. After BDD is built, if there is
no path to leaf "1", then the miter is unsatisfiable.

The runtime results given in seconds include experiments for 11, 18, 23, 33 and 50 bits unrolled SMPO
and their specifications. They show that both SAT and BDD based approach 
cannot verify SMPO beyond 33 bits.

\begin{table}[h]
\centering
\begin{tabular}{|c||c|c|c|c|c|} 
\hline
& \multicolumn{5}{|c|}{Word size of the operands $k$-bits}  \\
\hline
Solver & 11 & 18 & 23 & 33 & 50\\
\hline
\hline
Lingeling & 0.5  & 12.9  & 157.4 & 5931.2 & \emph{TO}\\
\hline
\hline
BDD & 0.02 & 5.7 & 237.9 & ?? & \emph{TO} \\
\hline
\end{tabular}
\caption{Runtime for verification of bug-free SMPO circuits over $\Fkk$ for SAT and BDD based methods. \emph{TO} = timeout of 4 hrs}\label{table:satbdd}  
\end{table}

\subsection{Evaluation of Our Approach}
Our approach requires only one time reduction for each clock cycle using RATO. If we use built-in function in \textsc{Singular} to directly
compute Gr\"obner basis, the high-degree word definition polynomial will exceed \textsc{Singular}'s capacity, even for 11 bits SMPO.

We also introduce bugs to our SMPO designs; these bugs are simply swapping an arbitrary pair of gate output signals, they will make the final result be
a far more complicated polynomial instead of $R+A\cdot B$, so the runtime will be slightly longer than bug-free one.

### 2014-06-06 today/conjecture.tex
modify bugs in 2bit traversal and add conjecture tex

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{fullpage}
\usepackage{fancyhdr}
\usepackage{cases}
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=1cm,bottom=2cm}
\pagestyle{fancyplain}
\headheight 15pt
%\lhead{Xiaojun Sun\ (u0741290)}
%\chead{\textbf{\Large Homework \asgnname}}
%\rhead{\today}
%\headsep 10pt

\begin{document}

\title{Conjecture on multivariate generator ideals' quotient}
\author{Xiaojun Sun}
\maketitle

\section{Problem Description}
From Prof. Enescu's last write-up, our conjecture on univariate generator ideals' quotient has been
proved. Let $K = \mathbb F_q$, where $q=p^e$, $p$ is a prime number. We have:\par
\indent \\
\textbf{Lemma}\ \ Let $f,g\in K[x]$, then $(f:g) = (\frac{f}{gcd(f,g)})$.\par
\indent \\
\textbf{Proposition}\ \ Let $J$ be an ideal of $K[x]$ and $J_0 = (x^q - x)$. Then
$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J).$$
\indent \par
Note above assertions dealt with ideals $J_0, J$ with only 1 generator and this generator is limited
to univariate polynomial about $x$. 
Now we would like to extend this conjecture on ideals with multivariate polynomial generators. Of course
those generators must be written in specific form. Here we give an example, which covers
typical requirement for our verification theory.\par
\indent\\
\textbf{Conjecture}\ \  Assume we have 3 ring variables $x,y,z$, and an ideal $J$ composed by 3 generators:
$J = \langle f, v_x, v_y\rangle$. Here $f$ has the form $f= z+\mathcal F(x,y)$, where $\mathcal F$
is a 2-argument function, for example $f = z+(x^2y+xy^2)$. And $v_x, v_y$ are "vanishing polynomials" about
$x, y$: $v_x = x^q - x, v_y = y^q - y$. Now we have another generator serving as "universal set"
$J_0 = \langle v_z, v_x, v_y\rangle$, $v_z = z^q - z$. (If necessary, $J$ can also include $v_z$ as 4th generator)
We conjecture that
$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J).$$
And we believe this conjecture is scalable such that it is also valid for situation 
$J = \langle f, v_x,v_y,v_u,v_w\rangle$, $f = z + \mathcal F (x,y,u,w)$.
\end{document}

### 2015-04-30 today/GB2newick.pl
push from labmachine

@@ -1,0 +1,67 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+# Note: read_in file format:
+# first line: start with "s", then a single number means # of original polys, then # of total GBs include "1" 
+# followed by n lines means n new polys added to GB
+# those lines: start with index of newly added poly (from s+1 to s+n)
+# then a pair of numbers means spoly pair
+# if more following, they mean polys used for reduction, exactly divided in the same order
+# last line start with "e" means the spoly reduced to "1"
+# start with "c" for comments
+# Edit: 04/20/2015 by Xiaojun Sun
+# Objective: translate the output from singular file (computing GB for UNSAT) showing all original polys involved for
+# spoly and reduction to newick tree format displaying the hierarchy
+
+open TREE, ">", "unsat11.nh";  # change the name
+my $cnt = 0;
+
+my $i;
+my $tmp;
+my %lines;
+my $origin;
+my $n;
+
+while(<>){
+	chomp;
+	my $read_in = $_;
+	my @items = split /\s+/;
+	if(!defined($items[0])) {next;}  # skip empty lines
+	if($items[0] =~ /c/) {next;}  # skip comments
+	if($items[0] =~ /s/){  # starting line
+	  $origin = $items[1];
+	  $n = $items[2];
+	  next;
+	}
+	$cnt++;
+	$tmp = "(".$items[1].":1,".$items[2].":1)";
+	if(defined($items[3]))  # this line has reduction polys
+	{
+	  for($i = 3; $i <= $#items; $i++)
+	  {
+		$tmp = "(".$tmp.":0.5,".$items[$i].":0.5)";
+	  }
+	}
+	$lines{$items[0]} = $tmp;
+	if($items[0] =~ /e/)
+	{
+	  my @keys = sort keys %lines;
+	  print "1..$origin ";
+	  print "@keys\n";
+	  last;
+	}
+}
+
+$tmp = $lines{"e"}.":0.5;";
+for($i = $n-1; $i > $origin; $i--)
+{
+	my $op1 = $i.":";
+	my $op2 = $lines{$i}.":";
+	$tmp =~ s/$op1/$op2/g;
+}
+
+print TREE $tmp;
+close TREE;

### 2015-04-30 today/cnf2poly.pl

@@ -1,0 +1,66 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+# Note: this is a simple version for 3-SAT CNF parser
+# Edit: 03/31/2015 by Xiaojun Sun
+# Objective: do scratch investigation on GB based UNSAT core extraction
+
+open POLY, ">", "unsat65.sing";  # change the name
+my $cnt = 0;
+
+my $i;
+my $tmp;
+my $clauses;
+my $vars;
+
+my @alphabet = qw /a b c d e f g h/;
+my @args;
+
+print POLY "ideal f = ";
+
+while(<>){
+	chomp;
+	my $read_in = $_;
+	my @items = split /\s+/;
+	$cnt++;
+	if(!defined($items[0])) {next;}  # skip empty lines
+	if($items[0] =~ /c/) {next;}  # skip comments
+	if($items[0] =~ /p/){  # write vanishing polys based on inputs
+	  $vars = $items[2];
+	  $clauses = $items[3];
+	  next;
+	}
+	if($items[3] != 0) {  # for 3-SAT problems
+	  print "Error in Line$cnt!\n";
+	  next;
+	}
+	undef @args;
+	for($i = 0; $i < 3; $i++)
+	{
+	  if($items[$i] < 0)
+	  {
+		$tmp = 0 - $items[$i] -1;
+#		print "What is tmp?? It is $tmp!!!!!!!";
+		push @args, "1+".$alphabet[$tmp];
+	  }
+	  else
+	  {
+		$tmp = $items[$i] -1;
+		push @args, $alphabet[$tmp];
+	  }
+	}
+	printf POLY "\n1+or%d(", 1+$#args; # $#args is NOT the number but the last index, which is 3-1=2 here
+	for($i = 0; $i <= $#args; $i++)
+	{
+	  print POLY $args[$i];
+	  if($i < $#args) 
+	  {print POLY ',';}
+	}
+	print POLY "),";
+}
+print POLY ";\n";
+if($cnt != $clauses + 1) 
+{print "Error: # of clauses not match! May exist empty lines\n";}

### 2015-04-30 today/draw.py

@@ -1,0 +1,5 @@
+from ete2 import Tree
+
+t = Tree("unsat11.nh", format = 0)
+
+t.render("mytree2.png", w=183, units="mm")

### 2015-04-30 today/duplicate_remove.cpp

@@ -1,0 +1,23 @@
+#include<iostream>
+#include<vector>
+#include<algorithm>
+using namespace std;
+
+int main()
+{
+	int a;
+	vector<int> array;
+	vector<int>::const_iterator i;
+	while(cin>>a)
+	{
+		if(find(array.begin(), array.end(), a) != array.end())
+			continue;
+		else
+			array.push_back(a);
+	}
+	sort(array.begin(), array.end());
+	for(i = array.begin(); i != array.end(); i++)
+		cout<<*i<<" ";
+	cout<<endl;
+	return 0;
+}

### 2015-04-30 today/slim_cnf.pl

@@ -1,0 +1,29 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+# slim_cnf: abstract certain clauses from a CNF file
+# Edit: 03/31/2015 by Xiaojun Sun
+# Objective: do scratch investigation on GB based UNSAT core extraction
+
+open CNF, ">", "extr65.cnf";  # change the name
+my $cnt = 0;
+my @idx = qw/1 2 3 4 5 6 7 9 10 11 12 13 15 16 18 19 23 24 26 27 29 30 31 36 40 42 44 48 52 53 58 63/;
+
+while(<>){
+	chomp;
+	my $read_in = $_;
+	my @items = split /\s+/;
+	if(!defined($items[0])) {next;}  # skip empty lines
+	if($items[0] =~ /c/) {next;}  # skip comments
+	if($items[0] =~ /p/){  # write vanishing polys based on inputs
+	  print CNF "p cnf $items[2] ";
+	  print CNF $#idx+1;
+	  next;
+	}
+	$cnt++;
+	if($cnt ~~ @idx)
+	{print CNF "\n$read_in";}
+}

### 2015-04-30 today/draw.py
conclude UNSAT core efforts

@@ -1,5 +1,0 @@
-from ete2 import Tree
-
-t = Tree("unsat11.nh", format = 0)
-
-t.render("mytree2.png", w=183, units="mm")

### 2015-04-30 unsat/visualization/draw.py

@@ -1,0 +1,9 @@
+# this is based on ETE2 package, visualizing a tree written in newick format (".nh" file)
+# sudo apt-get install python-qt4 python-mysqldb setuptools (may exist typo..care)
+# sudo easy_install -U ete2
+
+from ete2 import Tree
+
+t = Tree("tier2.nh", format = 0)
+
+t.render("mytree2.png", w=183, units="mm")

### 2015-05-01 techwriting/ASPDAC/abs.tex
all write-up from virtue machine

{\it Abstract -- Sequential Galois field ($\Fkk$) arithmetic circuits
take $k$-bit inputs and produce a $k$-bit result, after $k$-clock
cycles of operation. Formal verification of sequential
arithmetic circuits with large datapath size is beyond the
capabilities of contemporary verification techniques. To address this
problem, this paper describes a verification method based on algebraic
geometry that: i) implicitly unrolls the sequential arithmetic circuit
by multiple ($k$) clock-cycles and ii) represents the function
computed by the state-registers of the circuit, canonically, as a
multi-variate word-level polynomial in $\Fkk$. Our approach employs
the \Grobner basis theory with a specific elimination ideal. Moreover,
an efficient implementation is described to identify the $k$-cycle
computation performed  by the circuit at word-level. We demonstrate
the feasibility of our approach by verifying up to 100-bit sequential
Galois field multipliers, whereas conventional techniques fail beyond
23-bit circuits.  }

### 2015-05-01 techwriting/ASPDAC/append.tex
all write-up from virtue machine

\appendix

\subsection{Abstract Algebra Concepts}
Let $q = 2^k$ and $\mathbb F_{q}[x_1,\dots, x_d]$ be the polynomial
ring with indeterminates $x_1,\dots,x_d$. A polynomial $f = c_1 X_1 +
c_2 X_2 + \dots + c_t X_t$ is a finite sum of terms, where $c_1,
\dots, c_t$ are coefficients and $X_1, \dots, X_t$ are monomials. A
monomial ordering $X_1 > X_2 \dots > X_t$ is imposed on the
polynomials to process them systematically. Then, $LT(f) = c_1 X_1,
LM(f) = X_1$ denote the leading term and the leading monomial of $f$,
respectively. Given polynomials $f, g$, if $cX$ is a term in $f$ that
is divisible by $LT(g)$, 
then $f \xrightarrow{g} r$ denotes a one-step reduction (division) of
$f$ by $g$, resulting in remainder $r = f - {{cX} \over {LT(g)}} \cdot
g$. Similarly, $f$ reduces to $r$ modulo the set of polynomials $F =
\{f_1, \dots, f_s\}$, denoted $f \stackrel{F} {\textstyle
  \longrightarrow}_+ r$, such that no term in $r$ is divisible by the
$LT(f_i)$ of any polynomial in $ f_i \in F$.

Let $f_1,\dots, f_s$ be polynomials in $\mathbb F_q[x_1,\dots,x_d]$,
then we set $J = \langle f_1,\dots, f_s\rangle = \left\{\sum_{i=1}^s
h_if_i:h_1,\dots,h_s \in \mathbb F_q[x_1,\dots,x_d]\right\}$.  $J =
\langle f_1,\dots, f_s\rangle$ forms an ideal in $\mathbb
F_{q}[x_1,\dots, x_d]$, and $f_1,\dots, f_s$ are its generators.  
Let $\mathbf{a} = (a_1, \dots, a_d) \in \Fq^d$ be a point. For any
ideal $J = \langle f_1, \dots, f_s \rangle \subseteq 
\Fq[x_1,\dots, x_d]$, the {\it affine variety} of $J$ over
$\Fq$ is: 
$V(J) = \{\mathbf{a} \in \Fq^d: \forall f \in
J, f(\mathbf{a}) = 0\}.$ The variety $V(J)$ corresponds to
the set of all solutions to $f_1 = \dots = f_s = 0$. Any finite set of
points can be construed as the variety of an ideal --- a concept we
exploit to model the  set of (finite) states of the machine.

An ideal may have many generating sets. The set $G = \{g_1, \dots,
g_t\}$ is called a \Grobner basis of $J$ if and only if the
leading term of all polynomials in $J$ is divisible be the leading
term of some polynomial $g_i$ in $G$: i.e. $\forall f \in J, \exists
g_i \in G \ s.t. \ LT(g_i) ~|~ LT(f)$. The famous Buchberger's
algorithm, given in textbook \cite{ideals:book}, is used to compute a
\Grobner basis (GB). Operating on  input $F = \{f_1, \dots, f_s\}$,
and subject to the imposed term order $>$, it derives $G = GB(J) = \{
g_1, \dots, g_t \}$. Buchberger's algorithm repeatedly computes
$S$-polynomials. For pairs $(f_i, f_j) \in F$, $Spoly(f_i, f_j) =
\frac{L}{lt(f_i)}\cdot f_i - \frac{L}{lt(f_j)}\cdot f_j$, where $L =
LCM(LM(f_i), LM(f_j))$. Reducing $Spoly(f_i, f_j) \xrightarrow{F}_+ r$
cancels the leading terms of $f_i, f_j$ and gives a polynomial $r$
with a new leading term. This remainder $r$ is added to the current
basis and $Spoly(f_i, f_j)$ computations are repeated for all pairs of
polynomials until all $S$-polynomials reduce to 0.

\subsection{Normal Basis Theory}
Let $\beta$ be a element in the Galois field $F_{2^n}$ constructed by primitive element $\alpha$ and minimal polynomial 
$f(\alpha)$. Then a basis in the form $\{\beta, \beta^2, \beta^4, \beta^8, ... ,\beta^{2^{n-1}}\}$ is a
\emph{Normal Basis}; here $\beta$ is called \emph{Normal Element}.\\

For arithmetic in Galois fields, multiplication (with modulus) can be greatly simplified if Normal Basis representation 
is adopted to represent operands, especially in element square and multiplication:\\

\textit{Example B.1:}\ \ Element square: In $F_{2^n}$, we have relation $(a+b)^2 = a^2 + b^2$. Then 
$(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{n-1}\beta^{2^{n-1}})^2 = 
b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{n-1}^2\beta = 
b_{n-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{n-2}\beta^{2^{n-1}}$, which means a simple right-cyclic rotation.
One the other hand, standard basis representation do not have this benefit:\\
In $F_{2^3}$ constructed by $\alpha^3 + \alpha + 1$, standard basis is $\{ 1, \alpha, \alpha^2\}$.
Let $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ is a normal basis. Write down an element with both representations:
$E = a_0 + a_1\alpha + a_2\alpha^2 = b_0\beta + b_1\beta^2 + b_2\beta^4$, and its square in standard basis:
$E^2 = a_0 + a_1\alpha^2 + a_2\alpha^4 = a_0 + a_2\alpha + (a_1 + a_2)\alpha^2$. When it is written in normal basis:
$E^2 = b_2\beta + b_0\beta^2 + b_1\beta^4$.

\textit{Example B.2:}\ \ Normal basis multiplication: Assume there are 2 binary vectors representing 2 operands in normal
basis: $A = (a_0, a_1, \dots, a_{n-1}), B = (b_0, b_1, \dots, b_{n-1})$; similarly the product can also be written
as: $C = A*B = (c_0, c_1, \dots, c_{n-1}).$\\
Then the highest digit of product can be represented by a function: $c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, 
\dots, b_{n-1})$. Square both side: $C^2 = A^2*B^2$, i.e. the second highest digit $c_{n-2} = f(a_{n-1}, a_0, a_1, 
\dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2}).$ By this method it is easy to get all digits of product $C$.\\

\textit{Example B.3:}\ \ $\lambda$-Matrix: We can employ a binary $n\times n$ matrix $M$ to describe the "function"
 mentioned above: $c_{n-1} = f(A, B) = A \cdot M \cdot B^T$, $B^T$ denotes vector transposition. 
More specifically, we denote the matrix by \emph{$k$-th $\lambda$-Matrix}: $c_k = A \cdot M^{(k)} \cdot B^T$.
Then $c_{k-1} = A \cdot M^{(k-1)} \cdot B^T = rotate(A) \cdot M^{(k)} \cdot rotate(B)^T$, which means
by right and down shifting $M^{(k-1)}$ we can get $M^{(k)}$.\\
In $F_{2^3}$ constructed by $\alpha^3 + \alpha + 1$, let $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
is a normal basis. $0$-th $\lambda$-Matrix
\begin{equation}
M^{(0)} = \left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right).
\end{equation}
i.e.,
\begin{equation}
c_0 = (a_0\  a_1\  a_2)\left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right)\left(
\begin{array} {lcr}
b_0\\
b_1\\
b_2
\end{array} \right).
\end{equation}

\indent \\
$\lambda$-Matrix is defined with cross-product terms from multiplication, which is 
\begin{equation}
Product C = (\sum_{i=0}^{n-1}a_i\beta^{2^i})(\sum_{j=0}^{n-1}b_j\beta^{2^j}) = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}a_ib_j\beta^{2^i}\beta^{2^j}
\end{equation}
The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
normal basis, i.e.
\begin{equation}
\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1}\lambda_{ij}^{(k)}\beta^{2^k}, \ \ \lambda_{ij}^{(k)} \in F_2.
\end{equation}
Substitution yields, result is an expression for k-th digit of product as showed in \textit{Example B.3}:
\begin{equation}
c_k = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}\lambda_{ij}^{(k)}a_ib_j
\end{equation}
$\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ in $k$-th $\lambda$-Matrix.\par

\subsection{Optimal Normal Basis (ONB)}

The number of non-zero entries in $\lambda$-Matrix is named as  \emph{Complexity$(C_N)$}. \\
To define optimal normal basis, it is necessary to find out the minimum number
of non-zero terms in $\lambda$-matrix.\\
\textit{Theorem C.1}\ \ If N is a normal basis for $F_{p^n}$ with $\lambda$-matrix $M^{(k)}$, then non-zero entries in 
matrix $C_N\geq 2n-1$.\\
\begin{proof}
Let $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, denote $\beta^{p^i}$ by $\beta_i$. 
Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. \\
Denote $trace\ \beta$ with $b$, consider $n\times n$ matrix $M^{(0)}$. Then\\
$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$.\\
Therefore, the sum of all rows in $M^{(0)}$ is an n-tuple with $b$ as the first element and zeros elsewhere.
So the first column must contain at least 1 nonzero element, and other columns must get a sum of zero 
while keeping every row linearly independent so there should be at least 2 non-zero elements each column.
\end{proof}
If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as
\emph{Optimal Normal Basis}(ONB).\\
%For $F_{2^n}$, easy to figure out the first row of multiplication table must contain one "1" because 
%$\beta \beta^{2^0} = \beta^{2^1}$. for the rest rows $\beta \beta^{2^i} \neq \beta^{2^{i+1}}$,
%so there must be at least 2 "1"s to fulfil the equation. So for multiplication table, $C_N\geq 2n-1$ also holds.\\

\textit{Example C.1:}\ \ In $F_{2^4}$ constructed with $\alpha^4 + \alpha + 1$, 2 sets of normal basis can be found:
$\beta = \alpha^3, N = \{ \alpha^3, \alpha^6, \alpha^{12}, \alpha^9\}$ and 
$\beta = \alpha^7, N = \{\alpha^7, \alpha^{14}, \alpha^{13}, \alpha^{11}\}$. Multiplication table for $\beta = \alpha^3$:
\begin{equation}
T_1 = \left(
\begin{array}{lccr}
0 & 1 & 0 & 0\\
0 & 0 & 0 & 1\\
1 & 1 & 1 & 1\\
0 & 0 & 1 & 0
\end{array} \right)
\end{equation}
Complexity $C_N$ = 7, this is optimal. For $\beta = \alpha^7$:
\begin{equation}
T_2 = \left(
\begin{array}{lccr}
0 & 1 & 0 & 0\\
1 & 1 & 0 & 1\\
1 & 0 & 1 & 0\\
1 & 0 & 1 & 1
\end{array} \right)
\end{equation}
Complexity $C_N$ = 9, this is NOT optimal.\\

There is an efficient method to construct optimal normal basis by transforming minimal polynomials based on
special property of optimal normal basis\cite{ECC}.

\subsection{Construction of Type-I Optimal Normal Basis}
Rules for constructing Type-I ONB in $F_{2^n}$ are:
\begin{itemize}
\item n+1 must be prime.
\item 2 must be primitive in $\mathbb{Z}_{n+1}$.
\end{itemize}
Second rule means 2 raise to any power 0 $\sim$ n-1 (mod n+1) must cover every integer 1 $\sim$ n.\\
$\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ from k-th $\lambda$-Matrix. Then the crossproduct
term can be written as
\begin{equation}
\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1} \lambda_{ij}^{(k)} \beta^{2^k}
\end{equation}
Suppose we only care about k=0. So simplified to following equations:
\begin{numcases}{}
\beta^{2^i}\beta^{2^j} = \beta\\
\beta^{2^i}\beta^{2^j} = 1 \ \ (if\ \  2^i \equiv 2^j\ \  mod\ \  n+1)
\end{numcases}
solution $(i,j)$ implies entries equal to "1" in matrix. If $\beta$ is already an optimal normal element, 
$\beta^{2^i}$ counts through all powers of $\beta$ and generates our basis. So
\begin{numcases}{}
2^i + 2^j = 1 \ \ mod\ \  n+1\\
2^i + 2^j = 0 \ \ mod \ \ n+1
\end{numcases}
have the same solution for $M^{(0)}$. Note this method can be applied even both primitive polynomial and normal element 
are unkown (but existance and adoption of optimal normal basis must be guaranteed).\\
By repeating above steps, equations for multiplication table can also be obtained. Using this multiplication table, 
a transformation relation 
between $\lambda$-Matrix and Multiplication table can be concluded.
Example C.1 shows a type-I ONB in $F_{2^4}$.

	\subsection{Construction of Type-II Optimal Normal Basis}
Rules for constructing Type-II ONB over $F_{2^n}$ are:
\begin{itemize}
\item 2n+1 must be prime. And either
\item 2 must be primitive in $\mathbb{Z}_{2n+1}$, OR
\item $2n+1 \equiv 3 \ \ mod\ \  4$ AND 2 generates the quadratic residues in $\mathbb{Z}_{2n+1}$
\end{itemize}
The last rule means: The last 2 bits are set in the binary representation of prime 2n+1; even if $2^k \ \ mod\ \  2n+1$ does not
generate every element in 0 $\sim$ 2n, we can at least take the square root $mod\ \  2n+1$ of $2^k$. \\
To generate a Type-II ONB, first pick an element $\gamma$ of order 2n+1 in $F_{2^{2n}}$, then use this to find optimal normal 
element from $F_{2^n}$: $\beta = \gamma + \gamma^{-1}$. Crossproduct terms
\begin{equation}
\begin{split}
\beta^{2^i}\beta^{2^j} &= (\gamma^{2^i} + \gamma^{-2^i})(\gamma^{2^j} + \gamma^{-2^j})\\ &= 
(\gamma^{2^i+2^j} + \gamma^{-(2^i+2^j)}) + (\gamma^{2^i-2^j} + \gamma^{-(2^i-2^j)})\\ &= 
\begin{cases}
\beta^{2^k} + \beta^{2^{k\prime}}\ \  if\ 2^i \neq 2^j \ \ mod\ \  2n+1\\
\beta^{2^k}\ \  if\ 2^i \equiv 2^j \ \ mod\ \  2n+1
\end{cases}
\end{split}
\end{equation}
$k$ and $k\prime$ are the 2 possible solutions to multiplication of any 2 basis elements. That is why this normal basis
 is optimal: it has the minimum number of possible terms. In the case of $2^i \neq 2^j \ \ mod\ \  2n+1$, at least one of these
equations:
\begin{equation}
\begin{cases}
2^i + 2^j = 2^k \ \ mod\ \  2n+1\\
2^i + 2^j = -2^k \ \ mod \ \ 2n+1
\end{cases}
\end{equation}
will have a solution, and at least one of these equations
\begin{equation}
\begin{cases}
2^i - 2^j = 2^{k\prime} \ \ mod \ \ 2n+1\\
2^i - 2^j = -2^{k\prime} \ \ mod \ \ 2n+1
\end{cases}
\end{equation}
also has a solution.\\
In the case of $2^i \equiv \pm 2^j \ \ mod\ \  2n+1$, at least one of the following 4 equations has a solution:
\begin{equation}
\begin{cases}
2^i + 2^j = 2^k \ \ mod \ \ 2n+1\\
2^i + 2^j = -2^k \ \ mod \ \ 2n+1\\
2^i - 2^j = 2^k \ \ mod\ \  2n+1\\
2^i - 2^j = -2^k \ \ mod \ \ 2n+1
\end{cases}
\end{equation}
In the first set of equations, there are two possible solutions, and in the second set of equations,
there is only one possible solution. It is easy to see that the equations are all similar, so instead of 
working with 2 different sets we can combine them and work with just one group of 4 equations. To build 
our $\lambda$-Matrix, we set $k=0$ and find solutions to:
\begin{equation}
\begin{cases}
2^i + 2^j = 1 \ \mod\ \  2n+1\\
2^i + 2^j = -1 \ \ mod \ \ 2n+1\\
2^i - 2^j = 1 \ \ mod\ \  2n+1\\
2^i - 2^j = -1 \ \ mod \ \ 2n+1
\end{cases}
\end{equation}

\textit{Example E.1:} \ \ Following this criteria, easy to find $M^{(0)}$ for $F_{2^5}$:
\begin{equation}
M^{(0)} = \left(
\begin{array}{lcccr}
0 & 1 & 0 & 0 & 0\\
1 & 0 & 0 & 1 & 0\\
0 & 0 & 0 & 1 & 1\\
0 & 1 & 1 & 0 & 0\\
0 & 0 & 1 & 0 & 1
\end{array} \right)
\end{equation}

### 2015-05-01 techwriting/ASPDAC/appendix_fmcad.tex
all write-up from virtue machine

--- 2014-04-25 sequential/latex/ICCAD2014/appendix_fmcad.tex
@@
 
-\subsection{Design a Normal Basis Multiplier using $\lambda$-Matrix}
+\subsection{Design Angew's SMPO using $\lambda$-Matrix}
 Imitating the structure of SMPO, just specify the gates' connections for the first clock cycle, then following cycles are
@@
 The special polynomial for type-I and type-II optimal normal basis can be computed from algorithms showed in page 108, IEEE standard 1363-2000 document.
+
+\subsection{Theory about RH-multiplier}
+\label{sec:RHmulti}
+Assume there are 3 word variables from $\mathbb F_{2^k}$: $A = (a_0,a_1,\dots,a_{k-1}), B = (b_0,b_1,\dots,b_{k-1})$
+and $C = (c_0,c_1,\dots,c_{k-1})$, they have relation $C = A\cdot B$. Define function
+$$F_i(A,B) = a_{i-g}b_{i-g}\beta +\sum_{j=1}^{v}z_{i,j}\delta_j, 0\leq i\leq k-1$$
+where $v = \lfloor \frac{k}{2}\rfloor, \delta_j = \beta^{1+2^j}, 1\leq j\leq v$, $\beta$ is normal element.
+$g$ is a binary variable which determines value of $z_{i,j}$ as follows:
+
+(i) $k$ is odd, for $1\leq j < \lceil \frac{k}{2} \rceil$
+\begin{displaymath}
+z_{i,j} = \left\{
+\begin{array}{ll}
+(a_i+a_{i+j})(b_i+b_{i+j}) & g=0\\
+a_ib_{i+j} + a_{i+j}b_i & g=1
+\end{array}\right.
+\end{displaymath}
+
+(2) $k$ is even, for $1\leq j < \lceil \frac{k}{2} \rceil$
+\begin{displaymath}
+z_{i,j} = \left\{
+\begin{array}{ll}
+b_i(a_i+a_{i+v}) & g=0\\
+a_ib_{i+v} & g=1
+\end{array}\right.
+\end{displaymath}
+
+Here $i+j$ and $i+v$ stands for $i+j \pmod k$ and $i+v\pmod k$. With the definition, we can
+introduce a theorem which serves as basic concept of RH-multiplier:
+\begin{Theorem}
+$$C = (((F_{k-1}^2 + F_{k-2})^2 + F_{k-3})^2 + \cdots + F_1)^2 + F_0$$
+where $F_i\in \mathbb F_{2^k}$ is a short form of $F_i(A,B)$ defined above.
+\end{Theorem}
+Proof omitted.
+
+In this paper we choose $g=1$ to build AESMPO. The definition of function $F$ shows exactly
+the way to construct a RH-multiplier, note multiplication in $\mathbb F_2$ is equivalent to
+AND gate in Boolean logic circuits, and addition is equivalent to XOR gate, respectively.

### 2015-05-01 techwriting/ASPDAC/exp_result.tex
all write-up from virtue machine

\section{Experimental Results}
\label{sec:result}
We have implemented our approach within the \textsc{Singular} symbolic
algebra computation system [v. 3-1-6] \cite{DGPS}. Using our
implementation, we have performed experiments to verify two SMPO
architectures --- Agnew-SMPO \cite{agnew1991implementation} and the
RH-SMPO \cite{RHmulti} --- over $\Fkk$, for various datapath/field
sizes. Bugs are also introduced into the SMPO designs by modifying a
few gates in the combinational logic block. Experiments using SAT-,
BDD-, and AIG-based solvers are also conducted and results are
compared against our approach.  Our experiments run on a desktop with
3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and
64-bit Linux.

{\it Evaluation of SAT/ABC/BDD based methods:} To verify circuit $S$
against the polynomial $\F$, we unroll the SMPO over $k$ time-frames,
and construct a miter against a combinational implementation of $\F$.
A (pre-verified) $\Fkk$ Mastrovito multiplier \cite{mastro:1989} is
used as the {\it spec} model. This miter is checked for SAT using the
\emph{Lingeling} \cite{biere2013lingeling} solver. We also experiment
with the Combinational Equivalence Checking (CEC) engine of the ABC
tool \cite{abc}, which uses AIG-based reductions to identify internal
AIG equivalences within the miter to efficiently solve
verification. The BDD-based VIS tool \cite{brayton1996vis} is also used
for equivalence check. The run-times for verification of (unrolled)
RH-SMPO against Mastrovito {\it   spec} are given in Table
\ref{tbl:equiv} -- which shows that the techniques fail beyond 23 bit
fields.

\vspace{-0.1in}
\begin{table}[htb]
\centering
\caption{\small Run-time for verification of bug-free RH-SMPO circuits
  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
\label{tbl:equiv}
\begin{tabular}{|c||c|c|c|c|} 
\hline
& \multicolumn{4}{|c|}{Word size of the operands $k$-bits}  \\
\hline
Solver & 11 & 18 & 23 & 33 \\
\hline
\hline
Lingeling & 593  & \emph{TO}  & \emph{TO}  & \emph{TO}\\
\hline
\hline
ABC & 6.24 & \emph{TO} & \emph{TO} & \emph{TO}\\
\hline
\hline
BDD & 0.1 & 11.7 & 1002.4 & \emph{TO}  \\
\hline
\end{tabular}
\label{table:satbdd}  
\end{table}

CEC between unrolled RH-SMPO and Agnew-SMPO also suffers the same fate
(results omitted). In fact, both SMPO designs are based on slightly
different mathematical concepts and their computations in all
clock-cycles, except for the $k^{th}$ one, are also different. These
designs have no internal logical/structural equivalencies, and
verification with SAT/BDDs/ABC is infeasible. Their dissimilarity is
depicted in Table \ref{tbl:fraig}, where  $N_1$ depicts the number of
AIG nodes in the miter prior to {\it fraig\_sweep}, and the nodes
after {\it fraiging} are recorded as $N_2$; so $\frac{N_1-N_2}{N_1}$
reflects the proportion of equivalent nodes in original miter, which
emphasizes the (lack of) \emph{similarity} between two designs.

\begin{table}[htb]
\centering
\caption{\small Similarity between RH-SMPO and Agnew's SMPO}
\label{tbl:fraig}
\begin{tabular}{|c||c|c|c|c|} 
\hline
Size $k$ & 11 & 18 & 23 & 33 \\
\hline
$N_1$ & 734  & 2011  & 3285  & 6723\\
\hline
$N_2$ & 529 & 1450 & 2347 & 4852\\
\hline
Similarity & $27.9\%$ & $27.9\%$ & $28.6\%$ & $27.8\%$  \\
\hline
\end{tabular}\label{table:similarity}  
\end{table}

{\it Evaluation of Our Approach:} Our algorithm inputs the circuit
given in BLIF format, derives RATO, and constructs the polynomial
ideal from the logic gates and the register/data-word description.  We
perform one $Spoly$ reduction, followed by the bit-level to word-level 
substitution, in each clock cycle. After $k$ iterations, the final
result polynomial $R$ is compared against the spec polynomial. The
run-times for verifying bug-free and buggy RH-SMPO and Agnew-SMPO are
shown in Table \ref{tbl:exp1} and Table \ref{tbl:exp2}, respectively. 
We can verify, as well as catch bugs in, up to 100-bit
multipliers. Beyond 100-bit fields, our approach is infeasible --
mostly due to the fact that the intermediate 
abstraction polynomial $R$ is very dense and contains high-degree
terms, which can be infeasible to compute. However, it should be noted
that if we do not use the proposed bit-level to word-level
substitution, and compute reduced \Grobner bases with RATO, then our
approach does not scale beyond 33-bit datapaths.

\begin{table}[htb]
\centering
\caption{\small Run-time (seconds) for verification of bug-free and
  buggy RH-SMPO using our approach} 
\label{tbl:exp1}  
\begin{tabular}{|c||c|c|c|c|c|c|} 
\hline
Operand size $k$ & 33 & 51 & 65 & 81 & 89 & 99\\
\hline
\#variables & 4785 & 11424 & 18265 & 28512 & 34354 & 42372\\
\hline
\#polynomials & 3630 & 8721 & 13910 & 21789 & 26255 & 32373\\
\hline
\#terms & 13629 & 32793 & 52845 & 82539 & 99591 & 122958\\
\hline
\hline
Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 & 67021\\
\hline
Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120 & 66929\\
\hline
\end{tabular}
%\begin{minipage}{8cm}
%\def\arraystretch{1.5}\tabcolsep 2pt
%% \begin{tabular}{|c||c|c|c|c|c|} 
%% \hline
%% Operand size $k$ & 33 & 51 & 65 & 81 & 89 \\
%% \hline
%% \#variables & 4785 & 11424 & 18265 & 28512 & 34354\\
%% \hline
%% \#polynomials & 3630 & 8721 & 13910 & 21789 & 26255\\
%% \hline
%% \#terms & 13629 & 32793 & 52845 & 82539 & 99591 \\
%% \hline
%% %\hline
%% Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 \\
%% \hline
%% Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120\\
%% \hline
%% \hline
%% \end{tabular}
%% %\end{minipage}
\end{table}

\begin{table}[hbt]
\centering
\caption{\small Run-time (seconds) for verification of
  bug-free and buggy Agnew's SMPO our approach}
\label{tbl:exp2}
\begin{tabular}{|c||c|c|c|c|c|} 
\hline
Operand size $k$ & 36 & 66 & 82 & 89 & 100 \\
\hline
\#variables & 6588 & 21978 & 33866 & 39872 & 50300\\
\hline
\#polynomials & 2700 & 8910 & 13694 & 16109 & 20300\\
\hline
\#terms & 12996 & 43626 & 67322 & 79299 & 100100 \\
\hline
\hline
Runtime(bug-free) & 113 & 3673 & 15117 & 28986 & 50692 \\
\hline
Runtime(buggy) & 118 & 4320 & 15226 & 31571 & 58861\\
\hline
\end{tabular}
\end{table}

\section{Conclusions and Further Work}\label{sec:concl}
This proposal has described a method to verify sequential Galois field
multipliers over $\Fkk$ using computer algebra and algebraic geometry
based approach. As sequential Galois field circuits perform the
computations over $k$ clock-cycles, verification requires an efficient
approach to unroll the computation, and represent it as a canonical
word-level multi-variate polynomial. Using algebraic geometry, we show
that the unrolling of the computation at word-level can be performed by
\Grobner bases and elimination term orders. Subsequently, we show
that the complex \Grobner basis computation can be eliminated by means
of a bit-level to word-level substitution, which is implemented using
the binomial expansion over Galois fields and Gaussian
elimination. Our approach is able to verify up to 100-bit sequential
circuits, whereas contemporary techniques fail beyond 23-bit
datapaths.

Our approach still has following limitations: first, it can only be applied
on XOR-rich circuits, while most industrial designs are AND-OR gates dominant;
second, it only uses naive bit-word abstraction based on functions of arithmetic
circuits, which will be inefficient when the function does not have a straightforward
expression (such as an implicit function); last but not least, Gr\"obner basis computation
in our improved approach still requires a very long time. To overcome these limitations,
further explorations are needed for my research.

One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
Analysis on experiment results shows major time consumption is on ``multi-division" part.
A matrix-based technique named as "F-4 style reduction" \cite{F4reduce} can speed up the procedure
dividing a low-degree polynomial with term-sparse polynomial ideal.

We can also further expand our theory to sequential FSM implicit traversal. The sequential multiplier
verified in our experiment is actually a mealy FSM without primary inputs, i.e., we verify it through 
a single-chain implicit traversal. Our theory can also be applied when there are primary inputs -- a 
generic FSM.

Let us put following 2-bit FSM as an example. According to state transition graph (STG), we can get 
its gate-level implementation.
$\{s_0, s_1\}$ are state/pseudo inputs, $\{t_0,t_1\}$ are state/pseudo outputs, and we have a primary input (PI)  
$x$.
\begin{figure}[hbt]
\centering
\subfigure[State Transition Graph(STG)]{
\label{Fig.sub.1}
\includegraphics[width=0.35\textwidth]{stg_fig.eps}}
\subfigure[Gate-level implementation]{
\label{Fig.sub.2}
\includegraphics[width=0.55\textwidth]{fsm_fig.eps}}
\caption{Example Sequential Circuit}
\label{Fig.lable}
\end{figure}
Following algorithm is a modified breath-first search (BFS) traversal algorithm based on ideals with multivariate 
generators.
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition polynomial $f_t = T + \mathcal F (S,x)$, 
	initial state ideal $from^0 = \langle S+\mathcal G(x), x^{q_1} - x\rangle$}

  $reached = from^0(T\setminus S)$\;
  \Repeat{$GB(new^i) == 1$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$GB$(\langle f_t, from^{i-1}\rangle) \setminus \mathcal H(S)$\;
	$\overline{reached} = \langle T^{q_2}-T, x^{q_1} - x \rangle : reached$\;
	$new^i \gets $GB$(to^i + \overline{reached})$\;
  	$reached \gets $GB$( reached \cdot new^i)$\;
	$from^i \gets new^i(S\setminus T)$\;
  }
\Return{$reached$}
\caption {Algebraic Geometry based Traversal Algorithm (multivariate-generator ideals)}\label{alg:multi}
\end{algorithm}

Inputs of this algorithm are: a transition polynomial, which is the result of word-level abstraction; initial 
states description ideal, contains 2 generators defining PI could be any inputs and state inputs $S$ is a specific
value. (Ex: $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
$\langle S+x\cdot\alpha, x^2+x\rangle$ means initial states$=\{00,10\}$).

Transition polynomial calculation exploits our improved RATO approach. After building
an elimination ideal, use RATO such that \emph{reverse\ topo\ order\ circuit\ variables }$> T > S > x$, the reduction
remainder has the form $T+\mathcal F(s_0,s_1,x)$. From word definition $S+s_0+s_1\alpha$ we get
$$s_0 = \alpha S^2+ (1+\alpha)S, s_1 = S^2+S$$
Do substitution, the transition polynomial of example circuit is 
$$f_T = T+S^3\cdot x+\alpha S^3+(1+\alpha)S^2\cdot x+S^2+S\cdot x+(1+\alpha)x+1$$

Assume we start from state $\{11\}$. First iteration, it will reach state $\{01\}$. Line 4 is to compose an
ideal with 2 generators from $from^0$ and transition polynomial $f_T$, compute its Gr\"obner basis. Note this ideal
has the form
\begin{equation}
I_{tran} = \left\{
             \begin{array}{c}
             T+\mathcal F(S,x) \\
             S + \mathcal G(x) \\
             v_x
             \end{array}  
        \right.
\end {equation}
$v_x$ is a polynomial containing only $x$, initially it should be vanishing polynomial $x^{q_1}-x$, with the program executing
it may be factorized.

Consider Buchberger's algorithm, all generators' leading terms are relatively prime, so it is a GB itself. 
Then we need to reduce this GB, we will find out $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. So, at last we will get a polynomial
$T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
$S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$

Line 5 is the ideal quotient of universal set and reached states. In the first iteration, 
$reached$ is the initial state $\langle T+1+\alpha, x^2+x \rangle$. Result of ideal quotient
is $\langle T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle$ represents $\{00,01,10\}$.

Line 6 is ideals' sum (intersection of their varieties), it is done by combining all generators
from 2 ideals and compute GB. For first iteration result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.

Line 7 is ideals' product (union of their varieties), it is done by multiplying all pairs of
generators from both ideal. For first iteration result is $GB(\langle (T+1)(T+1+\alpha),
(T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$ representing $\{01,11\}$.

The traversal will run 3 iterations and terminate at 4th iteration. We list all intermediate 
results below:
\begin{itemize}
\item Iteration 1: $from^0 = \langle S+1+\alpha, x^2+x\rangle, to^1 = \langle T+1, x^2+x\rangle,
 reached = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$
\item Iteration 2: $from^1 = \langle S+1, x^2+x\rangle, to^2= \langle T+\alpha, x^2+x\rangle,
reached = \langle T^3+1, x^2+x\rangle$
\item Iteration 3: $from^2 = \langle S+\alpha, x^2+x\rangle, to^3 = \langle T+\alpha x, x^2+x
\rangle, reached = \langle T^3\cdot x+x, T^4+T, x^2+x\rangle$
\item Iteration 4: $from^3 = \langle S, x\rangle, to^4 = \langle T+1, x\rangle, new = \langle1\rangle$
\end{itemize}
The final reachable states are represented by ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
which means $\{00,01,10,11\}$.

This example shows a potential application of our proposed technique, which is promising if succeed on 
FSMs with larger word size and state numbers.

### 2015-05-01 techwriting/ASPDAC/exp_xiaojun.tex
all write-up from virtue machine

\section{Experimental Results}
\label{sec:result}
We have implemented our approach within the \textsc{Singular} symbolic
algebra computation system [v. 3-1-6] \cite{DGPS}. Using our
implementation, we have performed experiments to verify two SMPO
architectures \cite{agnew1991implementation} \cite{RHmulti} over
$\Fkk$, for various datapath/field sizes. Bugs are also introduced
into the SMPO designs by modifying a few gates in the combinational
logic block. Experiments using SAT-, BDD-, and AIG-based solvers are
also conducted and results are compared against our approach. 
Our experiments run on a desktop with 3.5GHz Intel
$\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and 64-bit Linux.

{\it Evaluation of SAT/ABC/BDD based methods:} 
For Angew's SMPO
design, there is an equation in following form (this is a 5-bit
example): 
\begin{align}
R_i &= b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
\end{align}
This equation can be used as specification of any unrolled SMPO circuit. For SAT/ABC/BDD based methods,
we first unroll our RH-SMPO design, then create a \emph{miter} with the unrolled design and
circuit specification, which we use unrolled Angew's SMPO in this paper, to check if it is unsatisfiable.

The SAT solver we use is \emph{Lingeling} \cite{biere2013lingeling}. Single-output miter circuit is written
in DIMACS CNF \cite{DIMACSformat} format and fed to Lingeling. In ABC, we use the Combinational Equivalence Checking (CEC) engine 
 to read and compare 2 networks from unrolled RH-SMPO and specification (both written in BLIF format). 
 For BDD based approach, we use the BDD module integrated within VIS tool \cite{brayton1996vis}. The miter is described by a BLIF 
 file and read by VIS. After BDD is built, if there is no path to leaf "1", then the miter is unsatisfiable.

The runtime results given in seconds include experiments for 11, 18, 23 and 33 bits unrolled SMPO
and their specifications. They show that all SAT, ABC and BDD based approach 
cannot verify SMPO beyond 33 bits.

\begin{table}[tb]
\centering
\caption{Runtime for verification of bug-free SMPO circuits over $\Fkk$ for SAT, ABC and BDD based methods. \emph{TO} = timeout of 14 hrs}
\begin{tabular}{|c||c|c|c|c|} 
\hline
& \multicolumn{4}{|c|}{Word size of the operands $k$-bits}  \\
\hline
Solver & 11 & 18 & 23 & 33 \\
\hline
\hline
Lingeling & 593  & \emph{TO}  & \emph{TO}  & \emph{TO}\\
\hline
\hline
ABC & 6.24 & \emph{TO} & \emph{TO} & \emph{TO}\\
\hline
\hline
BDD & 0.1 & 11.7 & 1002.4 & \emph{TO}  \\
\hline
\end{tabular}
\label{table:satbdd}  
\end{table}

The CEC engine embedded within ABC tool \cite{brayton2010abc} combines multiple techniques such as AIG rewriting via structural hashing, simulation and
SAT and has high efficiency when verifying structurally similar netlists, esp. comparing original design with its synthesized and optimized implementation.
However, the RH-SMPO we verify has distinct structure from our specification (Angew's SMPO). We can approximately give the similarity between these 2 designs
using information from networks before and after structural hashing. First, we read in the miter design created from unrolled RH-SMPO and specification,
record the original number of nodes in this network as $N_1$; then we use command \emph{fraig\_sweep} in ABC to reduce this network, equivalent nodes in AIG
are merged, record the number of nodes after reduction as $N_2$; $\frac{N_1-N_2}{N_1}$ reflects the portion of equivalent nodes in original miter, which is
the \emph{similarity} between two designs.

\begin{table}[tb]
\centering
\caption{Similarity between RH-SMPO and Angew's SMPO}
\begin{tabular}{|c||c|c|c|c|} 
\hline
Size $k$ & 11 & 18 & 23 & 33 \\
\hline
$N_1$ & 734  & 2011  & 3285  & 6723\\
\hline
$N_2$ & 529 & 1450 & 2347 & 4852\\
\hline
Similarity & $27.9\%$ & $27.9\%$ & $28.6\%$ & $27.8\%$  \\
\hline
\end{tabular}\label{table:similarity}  
\end{table}

\subsection{Evaluation of Our Approach}
Our approach requires only one time reduction for each clock cycle using RATO. If we use built-in function in \textsc{Singular} to directly
compute Gr\"obner basis, the high-degree word definition polynomial will exceed \textsc{Singular}'s capacity, even for 11 bits SMPO.

We also introduce bugs to our SMPO designs; these bugs are simply swapping an arbitrary pair of gate output signals. Since our approach does not
rely on arithmetic complexity of the design, the runtime for verifying buggy circuits has almost no difference from bug-free ones.

\begin{table}[tb]
\centering
\caption{Runtime (given in seconds) for verification of bug-free and buggy RH-SMPO circuits over $\Fkk$ using our approach}
\begin{minipage}{8cm}
\def\arraystretch{1.5}\tabcolsep 2pt
\begin{tabular}{|c||c|c|c|c|c|} 
\hline
Operand size $k$ & 33 & 51 & 65 & 81 & 89 \\
\hline
\#variables & 4785 & 11424 & 18265 & 28512 & 34354\\
\hline
\#polynomials & 3630 & 8721 & 13910 & 21789 & 26255\\
\hline
\#terms & 13629 & 32793 & 52845 & 82539 & 99591 \\
\hline
\hline
Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 \\
\hline
Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120\\
\hline
\end{tabular}
\end{minipage}
\label{table:SMPO}  
\end{table}

%\begin{table}[h]
%\centering
%\begin{tabular}{|c||c|c|c|c|c|} 
%\hline
%Operand size $k$ & 36 & 66 & 82 & 89 & 100 \\
%\hline
%\#variables & 183 & 333 & 413 & 448 & 503\\
%\hline
%\#polynomials & 2700 & 8910 & 13694 & 16109 & 20300\\
%\hline
%\#terms & 12996 & 43626 & 67322 & 79299 & 100100 \\
%\hline
%\hline
%Runtime(bug-free) & 113 & 3673 & 15117 & 28986 & 50692 \\
%\hline
%Runtime(buggy) & 118 & 4320 & 15226 & 31571 & 58861\\
%\hline
%\end{tabular}
%\caption{Runtime (given in seconds) for verification of bug-free and buggy Angew's SMPO circuits over $\Fkk$ using our approach}\label{table:SMPO}  
%\end{table}

### 2015-05-01 techwriting/ASPDAC/ganeshclass.tex
all write-up from virtue machine

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  IEEEsample.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%    More information: see the header of IEEEtran.sty
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%

\documentclass[conference,10pt]{IEEEtran}
%\documentclass[conference]{IEEEtran}

%%%\IEEEoverridecommandlockouts

\usepackage[ruled]{./algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%% \RequirePackage{times}
%% \RequirePackage{algorithmic}
%% \PassOptionsToPackage{boxed}{algorithm}
%% \RequirePackage{algorithm}
%% \RequirePackage{multicol}
%\renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
%\renewcommand{\algorithmicensure}{\textbf{Outputs:}}
%\DeclareMathAlphabet{\mathtsl}{OT1}{ptm}{m}{sl}

%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em1
%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newtheorem{example}{Example}
%\newtheorem{corollary}{Corollary}

\RequirePackage{amssymb, mathptm}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{comment}
\usepackage{leqno}
\usepackage{cases}

%%indent in algorithm

%\setcounter{page}{1}

% New command for the table notes.
\def\tabnote#1{{\small{#1}}}

% New command for the line spacing.
\newcommand{\ls}[1]
    {\dimen0=\fontdimen6\the\font
     \lineskip=#1\dimen0
     \advance\lineskip.5\fontdimen5\the\font
     \advance\lineskip-\dimen0
     \lineskiplimit=.9\lineskip
     \baselineskip=\lineskip
     \advance\baselineskip\dimen0
     \normallineskip\lineskip
     \normallineskiplimit\lineskiplimit
     \normalbaselineskip\baselineskip
     \ignorespaces
    }
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqarr}{\begin{eqnarray}}
\newcommand{\eeqarr}{\end{eqnarray}}
%\newcommand{\ov}{\overline}
\newcommand{\ov}{\bar}
\newcommand{\xor}{\bigoplus}
\newcommand{\Fm}{{\mathbb{F}}}

%the following is for space before and after align or other equation environment.

%%
\newtheorem{Algorithm}{Algorithm}[section]
\newtheorem{Definition}{Definition}[section]
\newtheorem{Example}{Example}[section]
\newtheorem{Proposition}{Proposition}[section]
\newtheorem{Lemma}{Lemma}[section]
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}{Corollary}[section]
\newtheorem{Conjecture}{Conjecture}[section]
\newtheorem{Problem}{Problem}[section]
\newtheorem{Notation}{Notation}[section]
\newtheorem{Setup}{Problem Setup}[section]
%%%

%%set spacing between table columns
\setlength{\tabcolsep}{3pt}

\begin{document}

%\thispagestyle{empty}
%\pagestyle{empty}

\title{Class Project Report: Boundary-free Technology Mapping using Gr\"obner Basis}

\author{
\IEEEauthorblockN{Xiaojun Sun}
\IEEEauthorblockA{Electrical and Computer Engineering\\
University of Utah\\
Salt Lake City, Utah 84112\\
Email: xiaojuns@ece.utah.edu}
}

%%\author{\IEEEauthorblockN{Jinpeng Lv and Priyank Kalla}\thanks{This work is sponsored in part by a grant from NSF \#CCF-546859.}
%\IEEEauthorblockA{Department of  Electrical and Computer Eng.\\
% University of Utah, Salt Lake City, UT-84112 \vspace{-0.3in}
 %\{lv, kalla\}@eng.utah.edu
% }
%\and
%\IEEEauthorblockN{Florian Enescu} \thanks{\normalsize  978-3-9810801-8-6/DATE12/$\copyright 2012$ EDAA}
%\IEEEauthorblockA{Department of Mathematics and Statistics\\
% Georgia State University,  Atlanta, GA 30302-4038 \vspace{-0.3in}
% fenescu@mathstat.gsu.edu
%} 
%

\maketitle

%\markboth{MS Proposal by Tim Pruss}{}
\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
\newcommand{\B}{{\mathbb{B}}}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\F}{{\mathcal{F}}}
\newcommand{\G}{{\mathcal{G}}}
\newcommand{\R}{\mathbb{R}}
%%%

\newcommand{\debug}[1]{\textcolor{gray}{[ #1 ]}}

%\thispagestyle{empty}

%%%%%%%%%%%%%%%%%%%% Include your files here %%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
As FPGA is widely used in hardware implementation and the size of design is usually large, there 
is an increasing need in industrial domain for an approach that can map macro blocks. Currently
there is no algorithms can map large-scale functions without boundary information, especially
arithmetic blocks in datapath. Our approach provides an alternative inspiration to solve
this problem using Gr\"obner basis based ideal membership test, whose efficiency does not
rely on the complexity of arithmetic block's inner structure.
\end{abstract}

\section{Introduction}
\subsection{Problem Description}
One sentence on this problem is: find a method to map the macro blocks without boundary information.
Mapping is an essential technique used in synthesis and verification. In synthesis, we can map the 
macro functions with smaller and faster implementations to optimize the timing and area; in simulation,
we can map a complicated function to a simple execution to accelerate simulation speed.

Given a gate-level design $D$ and several word-level macro blocks $\{B_i\}$, we need to map macro
blocks $B_i$ into design $D$, and write out the mapped design $D'$ which is equivalent to original
design $D$. The objective is to generate a mapped design $D$ with as many of $B_i$ such that the area
and timing is optimal.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.3]{macro.png}
	\end{center}
	\caption{The outline and flow of mapping for macro blocks}
	\label{fig:macro}
\end{figure}

\section{Preliminaries}
\label{sec:intro}

\subsection{Modeling of constrains by ideal}
A state can be described by a set of constrains. Generally speaking, a constrain is a function mapping 
from variables to boolean values ($f:X\to bool,\ X=\{x_0,\dots,x_n\}$); meanwhile this "function" can be
written in form of polynomial about variables. If the constrain is satisfied, i.e. $f:\sigma(X)\to \perp$,
accordingly the evaluation of this polynomial is 0:
$$f(x_0,\dots,x_n) = a_0x_0+\cdots + a_nx_n = 0$$
If a state is valid under some interpretation, then all of its constrains need to be satisfied:
\begin{equation}
\left\{
\begin{array}{ll}
f_0(x_0\dots x_n) = 0\\
\vdots\\
f_k(x_0\dots x_n) = 0
\end{array}\right.
\nonumber
\end{equation}
It is reasonable to consider the solution vector $\langle x_0,\dots,x_n\rangle$ for this system of 
equations. However in most cases, solving the system is impossible and unnecessary; instead people
tend to generate a simpler constrain/polynomial as an over-approximation (also known as \emph{abstraction})
from heuristic methods and then check if it satisfies all constrains of this state, which means
$$\forall j, f_j(\langle x_i\rangle) = 0\implies\F(\langle x_i\rangle) = 0  $$
One heuristic is to take $\F$ as composition of $\langle f_j\rangle$ as follows: if
$$\F = c_0f_0+c_1f_1+\cdots+c_kf_k$$
then above implication can hold. Here coefficients $c_j$ could be any polynomials
or real numbers. Now review Definition 2 in original paper:
\begin{Definition}
{\bf Ideals:}\ \ An \emph{ideal} $J$ generated by polynomials $f_0,\dots,f_k\in \mathbb{F}_q[x_0,\dots,x_n]$
is defined as $$J = \langle f_0,\dots,f_k\rangle = \{\sum_{i=0}^{k}c_i\cdot f_i\ |\ c_i\in\mathbb{F}_q[x_0,
\dots,x_n]\}$$
$f_0,\dots,f_k$ are called \emph{generators} of ideal $J$.
\end{Definition}
Thus we can formally say
\begin{Theorem}
If $f_0\sim f_k$ are all satisfied and $\F$ belongs to the \emph{ideal} generated by $\{f_0,\dots,f_k\}$, then
$\F$ is also satisfied.
\end{Theorem}

\subsection{Ideal membership and Gr\"obner basis}
From last section, to check if $\F$ is a valid abstraction, it is necessary to check if 
$\F\in J=\langle x_0,\dots,x_n\rangle$, this requires a technique called \emph{ideal membership test}.
To execute this technique, $\langle f_0,\dots,f_k\rangle$ need to be transformed to \emph{Gr\"obner Basis}
$\langle g_0,\dots,g_s\rangle$.

Gr\"obner basis can generate the same ideal as from original generators. Meanwhile it has an important property
that the leading term from an arbitrary polynomial from this ideal can be divided by at least one Gr\"obner
basis polynomial's leading term. More specifically,
\begin{Definition}
{\bf Gr\"obner Basis:}\ \ For a monomial ordering $>$, a set of non-zero polynomials $G=\{g_0,\dots,g_s\}$
contained in an ideal $J$, is called a \emph{Gr\"obner basis} for $J$ if and only if:
$$\forall f\in J, f\neq 0, \exists g_i\in G: lm(g_i)\ |\ lm(f)$$
where $lm$ denotes the leading monomial of a polynomial.
\end{Definition}

This property is vital when operating following "multi-division":
\begin{Definition}
{\bf Multi-division:}\ \ A polynomial $f$ divided by an ideal $J$(actually its generators, a finite set of
polynomials) follows rules from multi-division $f\xrightarrow{J}$:\\
for each term $t$ in $f$, search in all generators of $J$ such that the leading monomial of $g_i$ can divide
this term's monomial, then update $f$:
$$f = f - \frac{t}{lt(g_i)}g_i$$
$lt$ means leading term. Every time $\frac{t}{lt(g_i)}$ is recorded as one term of quotient $c_i$ w.r.t. $g_i$.
When it is impossible to update $f$, the value of $f$ is recorded as \emph{remainder} $r$, denoted as
$f\xrightarrow{J}_{+}r$.
\end{Definition}

If $J = \langle f_0,\dots,f_k\rangle$, $\F\xrightarrow{J}_+r$, apparently we have
$$\F = c_0f_0+c_1f_1+\cdots+c_kf_k + r$$
if the remainder $r=0$, then $\F$ is the desired constrain ($\F\in J$). Now look back to the property of
Gr\"obner basis, when $\F\in J$, then every step updated $\F$ also belongs to $J$, which means for every
term in original $\F$ we can find corresponding $g_i$ whose leading term can divide it, thus the remainder
will be 0. Reversely if remainder is 0, we can also deduce $\F\in J$.

\begin{Theorem}
{\bf Ideal membership:}\ \ For ideal $J=\langle g_0,\dots,g_s\rangle$ and $\{ g_0,\dots,g_S\}$
is a Gr\"obner basis, polynomial $\F\in J$ if and only if $\F\xrightarrow{J}_+0$.
\end{Theorem}
\subsection{Buchberger's Algorithm}
\label{sec:buch}
Buchberger's algorithm
shown in Algorithm \ref{alg:gb},  computes a Gr\"obner
basis over a field. Given polynomials $F = \{f_1, \dots, f_s\}$, 
the algorithm computes the Gr\"obner basis $G = \{g_1, \dots, 
g_t\}$.
% such that ideal $I = \langle f_1, \dots, f_s\rangle = \langle g_1,
% \dots, g_t\rangle$. 
In the algorithm,

\begin{equation}
       Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g \nonumber
      \label{eqn:spoly}
\end{equation}
where $L = \text{LCM}(lm(f), lm(g))$, where $lm(f)$ is the leading
monomial of $f$, and $lt(f)$ is the leading term of $f$.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{$F = \{f_1, \dots, f_s\}$}
 \KwOut{$G = \{g_1,\dots ,g_t\}$\\} %, a Gr\"{o}bner basis
  $G:= F$\;
  \Repeat{$G = G'$}
  {
  	$G' := G$\;
  	\For{ each pair $\{f, g\}, f \neq g$ in $G'$} 
	{
		$Spoly(f, g) \stackrel{G'}{\textstyle\longrightarrow}_+r$ \;
		\If{$r \neq 0$}
		{
			$G:= G \cup \{r\}$ \;
		}
	}
   }
\caption {Buchberger's Algorithm}\label{alg:gb}
\end{algorithm}

\subsection{Elimination using Gr\"obner Basis}
\label{sec:elim}
We are given a circuit $C$ with $k$-bit inputs and outputs that
performs a polynomial computation $Y = \F(A)$ over $\Fq = \Fkk$. Let
$P(x)$ be the {\it given} irreducible or primitive polynomial used for
field construction, and let $\alpha$ be its root, i.e. $P(\alpha) = 0
$. Note that we do not know the polynomial representation
$\F(A)$ and our objective is to identify (the coefficients of)
$\F(A)$. Let $\{a_0, \dots, a_{k-1}\}$ denote the primary inputs and
let $\{y_0, \dots, y_{k-1}\}$ be the primary outputs of $C$. Then, the
word-level and bit-level correspondences are: 
\begin{equation}
\label{eqn:words}
 A = a_0 + a_1 \alpha + \dots + a_{k-1} \alpha^{k-1}; ~~ Y = y_0 +
y_1 \alpha + \dots + y_{k-1} \alpha^{k-1};
\end{equation}

We analyze the circuit and model all the gate-level Boolean operators
as polynomials in ${\mathbb{F}}_2 \subset \Fkk$. To this set of
Boolean polynomials, we append the polynomials of Eqn
(\ref{eqn:words}) that relate the word-level and bit-level
variables. We model this set of polynomials as $F = \{f_1, \dots,
f_s\}$ over the ring $R = \Fq[x_1, \dots, x_d, Y, A]$. Here $x_1,
\dots, x_d$ denote, collectively, all the bit-level variables of the
circuit --- i.e. primary inputs, primary outputs and the intermediate
circuit variables --- and $Y, A$ the word-level variables. Denote the
generated ideal as $J = \langle F \rangle \subset R$. As $Y = \F(A)$
is a polynomial representation of the circuit, represent this (unknown)
``specification'' as a polynomial $f: Y - \F(A)$, or as $f: Y + \F(A)$
and $-1 = +1$ over $\Fkk$.

As the circuit $C$ implements the function $f$, clearly $f$ {\it
  agrees with the solutions} to $f_1 = \dots = f_s = 0$. In computer
algebra terminology, this means that $f$ {\it vanishes on the variety} 
$V_{\Fq}(J)$. If $f$ vanishes on $V_{\Fq}(J)$, then $f$ is a member of
the ideal $I(V_{\Fq}(J))$. Strong Nullstellensatz over Galois fields tells us that $I(V_{\Fq}(J)) =
J + J_0$, where $J_0 = \langle x_1^q - x_1, \dots, x_d^q - x_d, Y^q -
Y, A^q - A \rangle$ is the ideal of vanishing polynomials in
$R$. Consolidating these results, we deduce that the specification
polynomial $f \in (J+J_0)$.

If the specification polynomial is known, then the verification
problem can be solved using membership testing of $f$ in the ideal $(J
+ J_0)$. We will now show
that by computing a Gr\"obner basis of $(J + J_0)$, using a specific
elimination term order, we can also identify the polynomial $f$ which
represents the function implemented by the circuit.

Moreover, Gr\"obner bases may be used to generate an elimination ideal
by using an ``elimination order.''  One such ordering
is a pure lexicographic ordering, which features into a theorem:
%The ideal $I_k$ is spanned by elements which eliminate variables
%$x_1,\dots,x_k$.  This leads to an {\bf elimination theorem} for \Grobner
%bases:
\begin{Theorem} \label{thm:elim}
({\bf Elimination Theorem}) Let $J
  \subset \Fq[x_1,\dots,x_d]$ be an     ideal and let $G$ be a
  Gr\"obner basis of $J$ with respect to a lex ordering where $x_1
  > x_2 > \dots > x_d$.  Then for every $0     \leq l \leq
  d$, the set 
    \begin{equation}
        G_l= G \cap \Fq[x_{l+1},\dots,x_d]
    \end{equation}
    is a Gr\"obner basis of the $l$th elimination ideal $J_l$.
\end{Theorem}

Gr\"obner basis computations w.r.t. pure lexicographic
term orders can be used to eliminate variables from an ideal. The
above example motivates our approach: Since we want to derive a
polynomial representation from a circuit in variables $Y, A$, we can
compute a Gr\"obner basis of $J + J_0$ w.r.t. an elimination order
that eliminates all the ($d$) bit-level variables of the
circuit. Then, the Gr\"obner basis $G_d = G \cap \Fq[x_1, \dots, x_d,
  Y, A]$ of the $d^{th}$ elimination ideal of $(J + J_0)$ will contain
polynomials in only $Y, A$. We now prove that the 
required polynomial representation will be found in $G_d$.

\section{Techniques from Loop Invariant paper}
This section briefly illustrates S. Sankaranarayanan's POPL'04 conference paper.
First, a transition system is modeled by algebraic assertions; then \emph{ideal membership test}
is applied on the set of assertions to help abstract the loop invariant. "Template" is the concept
I borrowed from this paper to apply on my own approach.
\subsection{Template and state constrains}
\label{sec:basic}
To effectively and efficiently utilize the ideal membership test technique, a heuristic to
generate $\F$ is needed. One heuristic can provide better coverage for loop invariant abstraction as well as
relatively small size is \emph{generic quadratic form}.

For example, a pair of state variables $\{x,y\}$'s generic quadratic form is
$$\F = a_0x^2+a_1x+a_2xy+a_3y+a_4y^2+a_5$$
It covers all possible terms with degree at most 2. $a_0\sim a_5$ are usually real number parameters,
a certain assignment about them can turn $\F$ into desired invariant. This parameterized constrain
polynomial covers all combinations of state variables can also be called a \emph{template}.

Finding a proper assignment is the major part of original paper and has potential to be even further explored
beyond that paper, and/or to be applied on different research field.
\subsection{Modeling initial state}
\label{sec:CKR}
For initial state, the constrains are explicit. A template is adopted and refined by Gr\"obner basis 
generated by original constrains, by equaling the remainder to 0 we can get constrains on
parameters from the template.

An example is Fig.\ref{fig:loopinv}. The template is generic quadratic form of $\{s,i,j,j_0\}$, which is
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.3]{../../../Pictures/Selection_006.png}
	\end{center}
	\caption{An example program of 2 Natural numbers' multiplication (Fig.1 from original paper)}
	\label{fig:loopinv}
\end{figure}
\begin{align}
\F =& a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
&a_5i^2+a_6i+a_7ij+a_8ij_0+a_9j^2\nonumber\\
&a_{10}j+a_{11}jj_0+a_{12}j_0^2+a_{13}j_0+a_{14}\nonumber
\end{align}

Constrains of initial state: $s=0\land j=j_0$ can be interpreted to polynomials:
$\{s, j-j_0\}$. Check if it is Gr\"obner basis using \emph{Buchberger's algorithm}. Since
their leading terms are relatively prime, we can write Gr\"obner basis $G=\{s, j-j_0\}$. Its ideal
$J=\langle s,j-j_0\rangle$.

Compute multi-division $\F \xrightarrow{G}_+ r$, the remainder is
$r = a_5i^2+a_6i+(a_7+a_8)ij_0+(a_9+a_{11}+a_{12})j_0^2+(a_{10}+a_{13})j_0+a_{14}$.
Let it equal to 0, each coefficient will generate a constrain, solution to the system is candidate
assignment to generate invariant.
\begin{equation}
\left\{
\begin{array}{l}
a_5=a_6=a_{14}=0\\
a_7+a_8=0\\
a_9+a_{11}+a_{12}=0\\
a_{10}+a_{13}=0
\end{array}\right.
\nonumber
\end{equation}

\subsection{Modeling state transitions}
2 states get involved in a state transition. The technique requires to model the 2 states individually,
do multi-division separately to get remainder $r_1$ and $r_2$. Suppose a constrain polynomial for
this transition is $r_t$, we requires that when invariant for state 1 holds and transition $1\to 2$ stands,
invariant for state 2 should also holds, which is
$$(r_1 = 0)\land (r_t = 0) \implies (r_2 = 0)$$
One reasonable conjecture is
$$r_t = r_1 - \lambda r_2$$
Theoretically $\lambda$ could be any polynomial. Consider the system to be easy to solve, we keep $\lambda$
only taking value from real numbers.

Still take Fig.\ref{fig:loopinv} as example (also refer to Example 10 in original paper). State 1 is initial state we 
just characterized $\F = f(s,i,j,j_0)$, and the post state (state 2) have exactly the same form constrain:
$\F' = f(s',i',j',j_0')$. Considering transition relation, substitute $s'$ with $s+i$, replace $j'$
with $j-1$, and $i'$ for $i$, $j_0'$ for $j_0$, the template for state 2 is expression $f'$ in Example 10 from
original paper.

Do multi-division with Gr\"obner basis generated by transition relation, record its remainder $r_2$.
From equality $r_2 = \lambda r_1$, we get constrains for parameters. Solve the system using covering-problem-like
techniques (Part. \emph{Elimination by Splitting} in original paper). One branching result is:
$a_0,a_2\sim a_6, a_9\sim a_{14}$ are all zero; $a_1=a_7=-a_8$. Then reduced remainder 
$$r_2 = a_1s +(a_1-a_7)i+a_7ij+a_8ij_0 = a_1(s+ij-ij_0) = 0$$
Desired invariant is 
$$s = i(j_0- j)$$
We knew the program calculates $i\times j_0$. Initially $s=0$, $j_0-j=0$, invariant holds;
during each cycle, $s'=s+i, (j_0'-j') = (j_0-j)+1$, the invariant also holds. In conclusion, this is a
loop invariant for program in Fig.\ref{fig:loopinv}.

\section{Our approach on technology mapping}
Our approach borrows inspiration of "templates" from loop invariant paper. The difference is,
we are not using template polynomial to describe the system, instead we add some templates in the ideal
of macro blocks we want to map to simulate all possibilities.
\subsection{System abstraction}
The polynomial we use to test ideal membership should include all information of a circuit partition,
this requires us to abstract information from the system and write it into only 1 polynomial.
Usually this polynomial has the form like:
$$Z + f(i_1,i_2,\dots,i_n)$$
Here $Z$ is the output. When there is only one output, $Z$ will be a bit-level variable. However in most
cases we have multiple outputs, this asks us to write $Z$ as a word-level variable. $f$ is a Boolean
function about all inputs, and $i_1,i_2,\dots,i_n$ are all bit-level inputs.

\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.2]{tobemapped.png}
	\end{center}
	\caption{An example gate-level netlist to be mapped}
	\label{fig:tobemapped}
\end{figure}

Fig.\ref{fig:tobemapped} shows an example circuit partition with $a1,b1,a0,b0$ as inputs and
$Z = \{z0,z1\}$ as outputs. If we use elements from Galois field $F_{2^2}$ to represent word $Z$,
we have $Z = z_0 + \alpha\cdot z_1$.

The abstraction also uses the property of Gr\"obner basis. If we arrange a term ordering as
$$Other\ circuit\ variables > output\ word\ Z > all\ bit\ level\ inputs$$
and include all gates information, word level variable definition and vanishing polynomials,
the reduced Gr\"obner basis will have a polynomial (generator) in the form of $Z + f(a1,b1,a0,b0)$,
this is an application of "elimination using GB", where we eliminate all variables except output
word variable $Z$ and inputs $a1,a0,b1,b0$.

In this example:

\begin{itemize}
\item Term ordering: $z0,z1,t,g1,p1,ci,t1,t2,Z,a0,b0,a1,b1$
\item Gate description (from table \ref{table:booltogalois_op}): $z0+t+g1+t\cdot g1, t+t1+t2+t1\cdot t2, g1+a1\cdot b1,
			t1+a1\cdot ci, t2+b1\cdot ci, ci+a0\cdot b0, z1+p1+ci, p1+a1+b1$
\item Word definition: $Z+z0+z1\cdot \alpha$
\item Vanishing polynomials $(J_0)$: $z0^2+z0, z1^2+z1, t^2+t, g1^2+g1, t1^2+t1, t2^2+t2, p1^2+p1, ci^2+ci,
			a0^2+a0, b0^2+b0, a1^2+a1, b1^2+b1, Z^4+Z$(note Z is 2-bit word)
\end{itemize}

The result is one polynomial in Gr\"obner basis with leading term $Z$:
$$Z+a0\cdot b0\cdot a1+a0\cdot b0\cdot b1+\alpha\cdot a0\cdot b0+a1\cdot b1+\alpha\cdot a1+\alpha\cdot b1$$

\begin{table}
\centering
\begin{tabular}{|c|c|} \hline
Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
$\overline{a}$ & $1 + a$\\ \hline
$a\ and\ b$ & $ab$\\ \hline
$a\ or\ b$ & $a + b + ab$\\ \hline
$a \oplus b$ & $a + b$\\
\hline\end{tabular}
\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
\label{table:booltogalois_op}
\end{table}

\subsection{Templates on Boundary Information}
\begin{figure}[hbt]
	\begin{center}
	\includegraphics[scale=0.2]{template.png}
	\end{center}
	\caption{Standard 2-bit adder with input/output mapping}
	\label{fig:template}
\end{figure}

Fig.\ref{fig:template} shows a 2-bit standard cell. It has 3 outputs mapping to 2 output pins,
and 4 inputs mapped to 3 input pins (the rest pin is mapped to fixed 0/1 signal). Dashed connection
lines show one possible mapping, to find out this kind of feasible mapping, we need to simulate
all possible mappings, where the concept "template" can be used.\vspace{5mm}

{\bf Output template}: $z0+ct0_{z0}\cdot t0+ct1_{z0}\cdot t1+c0_{CO}\cdot CO+cn0$

When $cn0 = 0$, any one of $ct0_{z0},ct1_{z0},c0_{CO}$ could be 1 means mapping to corresponding output
pin. If $cn0 = 1$, mapping to negation of corresponding output pin.

{\bf Input template}: $a0+pa0+c0_{A0}\cdot s0+c1_{A0}\cdot s1+c2_{A0}\cdot s2$

When $pa0 = 0$ and any one of $c0_{A0},c1_{A0},c2_{A0}$ equal to 1 means mapping to corresponding input
pin; if all $c0_{A0},c1_{A0},c2_{A0}$ equal to 0 then means mapping to fixed signal "0" (when $pa0 = 1$ then
mapping to fixed signal "1").

\subsection{Fast GB computing}
We compose an ideal with polynomials from all gates information, input/output templates and word definition,
then compute a GB of this ideal. However, computing GB using algorithms like Buchberger's algorithm is
less efficient. For example, the "slimgb" function in Singular tool cannot even compute a GB for ideal
from Fig.\ref{fig:template}. So finding a fast GB computing method is necessary.

Analyzing Buchberger's algorithm, we find the reason that GB compute takes long time is the size of ideal
will explode when there exists many pairs of polynomial can compute a Spoly. If we arrange the term ordering
carefully to minimize the number of Spoly, then GB computation will be easy.

We find such term ordering can satisfying our requirement:
\begin{align}
&Reverse\ topological\ order\ of\ circuit\ variables > \nonumber\\
& output\ word\ variable > output\ templates\ variables>\nonumber\\
& fixed\ 0/1\ templates\ variables > input\ templates\ var\nonumber\\
& > input pins \nonumber
\end{align}

Under this term ordering, all polynomials' leading term are relatively prime (so Spoly equal to 0)
except only 1 pair of polynomials. So the work load is just compute a Spoly for this pair,
then reduce the Spoly with whole ideal. After adding the remainder to the ideal, it becomes
a Gr\"obner basis.

\subsection{Our Tool Flow}
First, choose a set of independent wires as input pins;

Second, push forward input signal for a certain number of gates (depth), choose a cut set
which fully dependent on these inputs as output pins;

Third, abstract a description polynomial using GB abstraction method;

Fourth, reduce (do multi-division) the description polynomial with the GB we computed with macro block 
and templates, get a remainder polynomial;

Last but not least, abstract all coefficients from the remainder polynomial, set up a system of equations,'
find a solution to this system, then the solution is a feasible mapping; no solution means cannot make a
feasible mapping.

\section{Problems to address}
The final system of equations is not linear, so a 0/1 programming technique cannot be used. However,
an exhaustive enumerating is still practicable:
If we map a marco block with $j$ inputs and $n$ outputs into circuit partition with $i$ inputs
and $m$ outputs, number of total possible assignments is
$$2^{j-i}A_j^iA_n^m$$
when considering mapping to negative signals, it increases to 
$$2^{m+j}A_j^iA_n^m$$

For example, when mapping 3 inputs 2 outputs partition to a 2-bit adder:

Total possible assignments: 9216

Full search space: $2^{24} = 16.8M$

So it only takes $0.05\%$ time to scan all possibilities using our approach.

When the macro block is larger as 4-bit adder, number of total possible assignment is $2^{27}$,
still a number we can accept.
%%%%%%%%%%%%%%%%%%%% The bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{thebibliography}{1}
%\bibitem{ref1}
%\end{thebibliography}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-05-01 techwriting/ASPDAC/intro_aspdac.tex
all write-up from virtue machine

\section{Introduction}
\label{sec:intro}

Galois field (GF) arithmetic finds application in many areas such as 
cryptography, error control coding, VLSI testing, etc. In most
hardware applications, fields of the type $\Fkk$ are widely
chosen. Such {\it binary} GFs are $k$-dimensional extensions
of the base field ${\mathbb{F}}_2$; this allows for the design of
efficient (AND-XOR) arithmetic architectures and algorithms for
hardware design. For example, in elliptic curve cryptography, the
operations of encryption, decryption and authentication rely on
polynomial computations over $\Fkk$. In many applications, however,
the datapath size (bit-vector operand size) $k$ can be very
large. Therefore, for efficient implementation, many custom VLSI
architectures for polynomial GF aritmetic have been devised
\cite{wu:2002} \cite{acar:1998} \cite{agnew1991implementation}
\cite{RHmulti}. Due to their deployment in communications and
security-related applications, there is a critical need to {\it formally
verify} the correctness of hardware implementations of GF arithmetic.

Formal verification of such specialized and custom GF arithmetic
architectures remains a challenge. While there have been recent
advancements in formal verification of {\it combinational} GF
arithmetic circuits \cite{ibm:blueveri} \cite{lv:tcad2013}
\cite{pruss:dac14}, formal verification of sequential GF circuits has
not been addressed.

The high complexity

 This introduce many
design and verification challenges

### 2015-05-01 techwriting/ASPDAC/intro_date.tex
all write-up from virtue machine

\section{Introduction}
\label{sec:intro}

Galois field (GF) arithmetic is part of modern abstract algebra focused on 
computations on modulo-based fields
and finds application in areas such as 
cryptography, error control coding, VLSI testing, etc. In most
hardware applications, fields of the type $\Fkk$ are widely
chosen. Such { binary} GFs are $k$-dimensional extensions
of the base field ${\mathbb{F}}_2$; this allows for the design of
efficient (AND-XOR) arithmetic architectures and algorithms for
hardware design. Due to their deployment in communications and
security-related applications, there is a critical need to {
  formally verify} the correctness of hardware implementations of GF
arithmetic. In most applications, however, the datapath size
(bit-vector operand size) $k$ is too large to use naive enumeration test. Most conventional formal
verification methods are unable to cope with the large size and
complexity of GF circuits.

GF arithmetic circuits in $\Fkk$ take $k$-bit vectors as inputs and
produce $k$-bit outputs. For example, a GF modulo multiplier computes
$R = A \times B \pmod{ P(x)}$, where: i) $A = a_{s(0)} + a_{s(1)}\alpha +
\dots + a_{s(k-1)} \alpha^{k-1} = \sum _{i=0}^{k-1} a_{s(i)}\alpha^i$,
$B = \sum_{i=0}^{k-1}b_{s(i)} \alpha^{i}$ denote the $k$-bit inputs,
$R = \sum _{i=0} ^{k-1} r_{s(i)} \alpha^i$ is the output, and $a_{s(i)},
b_{s(i)}, r_{s(i)} \in \Ftwo$; ii) $P(x)$ is the given primitive
polynomial used to construct $\Fkk$; and iii) $P(\alpha) = 0$,
i.e, $\alpha$ is the primitive element of the field. In the above, the
elements are represented in { standard basis notation} (denoted by
subscript ``s'' on the bits). As the datapath size $k$ increases,
combinational GF designs become prohibitively large;
%\cite{wu:2002} \cite{acar:1998}; 
sequential GF circuits are therefore desirable.

Sequential GF circuits operate as follows: $k$-bit input operands are
loaded into $k$-bit state registers (flip-flops), and the circuit is
executed for $k$ clock-cycles; after which the $k$-bit result is
available in the output registers. Data representation and circuit
design for sequential GF circuits is mostly based on { normal basis}
\cite{gao:phd_normal_basis}. Data is represented as $A = a_{n(0)}
\beta + a_{n(1)} \beta^2 + a_{n(2)} \beta^{2^2} + \dots + a_{n(k-1)}
\beta^{2^{k-1}}$, where $\beta \in \Fkk$ is the { normal
  element}, and $\{\beta, \beta^2, \dots, \beta^{2^i}, \dots,
\beta^{2^{k-1}}\}$ forms the normal basis. Standard and normal basis
representations can be derived from each other, i.e, $\beta$ can be
derived from $\alpha$ and vice-versa. It has been shown that
architectures for multiplication and  squaring can be efficiently
designed using normal bases  \cite{agnew1991implementation} 
\cite{RHmulti} \cite{gao:phd_normal_basis} as sequential circuits. 
\footnote{In fact, some (not all) finite fields have an
  { optimal normal basis}, where the combinational logic part of
  the sequential circuit contains a minimum number of 2-input AND/XOR
  gates.}.

\begin{Example}
\label{ex:nb_sq}
For $a, b \in \Fkk, (a+b)^2 = a^2 + b^2$. Applying this rule for
element squaring: 
\begin{align}
B = & (b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}}) \nonumber\\
B^2 = &b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta^{2^k} \nonumber\\
= &b_{k-1}\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}} \nonumber
\end{align}
as $\beta^{2^k} = \beta$ due to Fermat's little theorem, and $b_i^2 = b_i$. 
\end{Example}

The above example shows that squaring of elements represented in
normal bases can be implemented simply by a cyclic right-shift
operation. However, multiplication of two elements in normal basis is
still a complicated operation --- and its design and verification is still
challenging. Recent literature has addressed formal equivalence proofs
of { combinational} GF arithmetic circuits \cite{ibm:blueveri}
\cite{lv:tcad2013} \cite{pruss:dac14}. Our research aims to address verification
of sequential GF circuits, which has not been addressed before.

{\it Problem Statement:} We are given: 
i) the Galois field $\Fkk = \Ftwo[X] \pmod{ P(X)}, P(\alpha) = 0$,
along with the { normal basis representation}, i.e. $\beta$ is also
known; 
ii) a word-level specification polynomial $R = \F(A, B) \pmod{ P(X)}$;
where $A = \sum _{i=0}^{k-1} a_i \beta^{2^{i}}$, $B = \sum
_{i=0}^{k-1} b_i \beta^{2^{i}}$, $R = \sum _{i=0}^{k-1} r_i
\beta^{2^{i}}$, $ a_i, b_i, r_i \in \Ftwo$, and iii) a sequential
circuit ($S$) implementation of the 
polynomial computation. Our objective is to prove or disprove that $S$
is a $k$-cycle implementation of $R$.

{\it Approach \& Contributions:} The sequential GF arithmetic circuit
can be viewed as a {restricted} Mealy finite state machine (FSM),
as shown in Fig. \ref{fig:sequential}. The FSM contains no
primary inputs or outputs. The operands are loaded into state
registers $A, B$ (as initial states), and after $k$-clock cycles of
operation, the result $R = \F(A, B)$ is stored in the $R$ register.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=2.8in]{./gf_seq_model.eps}
\end{center}
\caption{\small A typical normal basis GF sequential circuit model. $A =
  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
  $A', B', R'$ denote next-state inputs.}
\label{fig:sequential}
\end{figure}

% \vspace{-0.2in}
A straightforward approach to verify such a sequential circuit may
consist of unrolling the circuit for $k$ time-frames, and performing a
(combinational) equivalence check between the unrolled machine and the
specification polynomial. Such a technique is grossly inefficient for
large circuits. Therefore, we propose a method that implicitly
(symbolically) represents the unrolled computation, { canonically,
  as a word-level multi-variate polynomial.} We show that the
$k$-cycle polynomial representation can be derived { iteratively}
by performing a sequence of \Grobner basis ($GB$) computations of the
ideal generated by the polynomials corresponding to the circuit. The
approach  requires the use of a specific elimination term order for
the $GB$ computation, based on the circuit's topology. Once the
canonical polynomial is derived, it can be checked against the
specification polynomial for verification.

Computing \Grobner bases with elimination orders is infeasible for
large circuits. To overcome this complexity, we draw inspiration from
\cite{pruss:dac14}, and exploit the binomial expansion in GFs to
engineer a new, efficient implementation to derive the word-level
polynomial. We demonstrate the feasibility of our approach by
verifying (and also detecting bugs in) up to 100-bit sequential GF
multipliers (containing 300 flip-flops), whereas conventional
techniques fail beyond 23-bit circuits. Finally, our approach can be
construed as a word-level, implicit traversal of the underlying FSM of
the sequential GF circuit, wherein the set of states is encoded as the
variety of an elimination ideal related to the FSM transition
function.

### 2015-05-01 techwriting/ASPDAC/latest_alg.tex
all write-up from virtue machine

\section{Verification of Sequential Arithmetic Circuits}
Fig.\ref{fig:seq_block} shows the basic structure of a sequential arithmetic circuit without primary inputs. 
There are 3 register files representing input operands $A, B$ and result output $R$. After $k$ clock cycles,
data stored in register file $R$ will be the result of an arithmetic operation.

The combinational logic block takes present state variable $A, B$ and $R$ as inputs, and next state variable 
$A', B'$ and $R'$ as outputs. After $k$ clock cycles, the output $R_{final}$ equals to desired function of
initial input operands $A_{init}, B_{init}$. Our approach aims to find a polynomial indicating
this function without unrolling.
An algorithm is used to describe how our approach works to verify this kind 
of sequential arithmetic circuits.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial 
 $\F(R), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from^0(R,A,B) = \F(R), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$(J_{ckt}+ J_0+ from^{i-1}(R,A,B))$ w/ abstraction term order\;
	$to^i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from^i \gets to^i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from^k(R_{final})$}
\caption {Sequential arithmetic circuit inductive verification}\label{alg:modified}
\end{algorithm}

In this algorithm, $from^i$ and $to^i$ are polynomial ideals whose varieties are evaluations of word-level
variables $R, A, B$ and $R',A',B'$ in $i$-th iteration. However in most cases, an arithmetic circuit should give the result $R_{final}$ in certain
function of initial loaded operands $A_{init}$ and $B_{init}$ (they are fixed values during one calculation task);
e.g. for a multiplier, $R_{final} = A_{init}\cdot B_{init}$. So in our approach we record all intermediate relations
$R = \mathcal{F}(A_{init}, B_{init})$ for each clock cycle to evaluate output $R$.
Run algorithm \ref{alg:modified}, the return value should be desired function $R_{final} = \mathcal{F}(A_{init},B_{init})$.

\begin{Example}
\label{ex:RHSMPO}
Fig.\ref{fig:RHmulti} shows the detailed structure of a 5-bit RH-multiplier (AESMPO). The transition function for
operands $A,B$ is doing cyclic shift, while transition function for $R$ has to be computed through Gr\"obner basis
abstraction approach. Following ideal $J_{ckt}$ from line 5 in algorithm \ref{alg:modified} is the ideal for 
all gates in combinational logic block and definition of word-level variables.
\begin{align}
J_{ckt} = & d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4,\nonumber\\
& c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2, \nonumber\\
& R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4,\nonumber\\
& A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},\nonumber\\
		 & B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18},\nonumber\\
		 & R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18},\nonumber\\
		 & R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18};\nonumber
\end{align}
In our implementation here, since we only focus on the output variable $R$, evaluations of intermediate input 
operands $A, B$ are unnecessary. Polynomials about $A$ and $B$ can be removed from $J_{ckt}$, and $R$ is directly
evaluated by initial operands $A_{init}$ and $B_{init}$, which are associated with present state bit-level inputs
$a_0,a_1,\dots,a_4$ and $b_0,b_1,\dots,b_4$ by polynomials in $from^i$.

According to line 5 of algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
Gr\"obner basis with abstraction term order (copy details here). There is a polynomial in form of $R'+\F(A_{init},B_{init})$,
which should be included by $to^{i+1}$. $to^{i+1}$ also exclude next state variable $A'$ and $B'$, instead we 
redefine $A_{init}$ and $B_{init}$ using next state bit-level variables $\{a_i', b_j'\}$. Next state Bit-level variables
$a_i' = a_{i-1\pmod k}, b_j' = b_{j-1\pmod k}$ according to definition of cyclic shift.

Line 7 in algorithm \ref{alg:modified} is implemented by replacing $R'$ with $R$, $\{a_i', b_j'\}$ with $\{a_i,b_j\}$.

All intermediate results for each clock cycle are listed below:
\begin{itemize}
\item Clock 1: $from^0 = \{R, A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\}$, 
	$to^1 = \{R'+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^2) A_{init}^8 B_{init}+(\alpha^4+\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^8+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha^2+\alpha) A_{init}^2 B_{init}^2+(\alpha^4+\alpha) A_{init}^2 B_{init}+(\alpha^4+\alpha^3+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init}^2+(\alpha^3+\alpha+1) A_{init} B_{init},
	A_{init}+a_4'\alpha^5+a_0'\alpha^{10}+a_1'\alpha^{20}+a_2'\alpha^9+a_3'\alpha^{18},
B_{init}+b_4'\alpha^5+b_0'\alpha^{10}+b_1'\alpha^{20}+b_2'\alpha^9+b_3'\alpha^{18}\}$
\item Clock 2: $from^1 = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^2) A_{init}^8 B_{init}+(\alpha^4+\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^8+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha^2+\alpha) A_{init}^2 B_{init}^2+(\alpha^4+\alpha) A_{init}^2 B_{init}+(\alpha^4+\alpha^3+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init}^2+(\alpha^3+\alpha+1) A_{init} B_{init},
	A_{init}+a_4\alpha^5+a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^9+a_3\alpha^{18},
B_{init}+b_4\alpha^5+b_0\alpha^{10}+b_1\alpha^{20}+b_2\alpha^9+b_3\alpha^{18}\}$,
$to^2 = \{R'+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^8+(\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^2+(\alpha^3+1) A_{init}^8 B_{init}+(\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4) A_{init}^4 B_{init}^4+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^2) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}+(\alpha^3+1) A_{init} B_{init}^8+(\alpha) A_{init} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init},
A_{init}+a_3'\alpha^5+a_4'\alpha^{10}+a_0'\alpha^{20}+a_1'\alpha^9+a_2'\alpha^{18},
B_{init}+b_3'\alpha^5+b_4'\alpha^{10}+b_0'\alpha^{20}+b_1'\alpha^9+b_2'\alpha^{18}\}$
\item Clock 3: $from^2 = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^8+(\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^2+(\alpha^3+1) A_{init}^8 B_{init}+(\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4) A_{init}^4 B_{init}^4+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^2) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}+(\alpha^3+1) A_{init} B_{init}^8+(\alpha) A_{init} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init},
A_{init}+a_3\alpha^5+a_4\alpha^{10}+a_0\alpha^{20}+a_1\alpha^9+a_2\alpha^{18},
B_{init}+b_3\alpha^5+b_4\alpha^{10}+b_0\alpha^{20}+b_1\alpha^9+b_2\alpha^{18}\}$, 
$to^3 = \{R'+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}+(\alpha) A_{init}^8 B_{init}^{16}+(\alpha+1) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init},
A_{init}+a_2'\alpha^5+a_3'\alpha^{10}+a_4'\alpha^{20}+a_0'\alpha^9+a_1'\alpha^{18},
B_{init}+b_2'\alpha^5+b_3'\alpha^{10}+b_4'\alpha^{20}+b_0'\alpha^9+b_1'\alpha^{18}\}$
\item Clock 4: $from^3 = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}+(\alpha) A_{init}^8 B_{init}^{16}+(\alpha+1) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init},
A_{init}+a_2\alpha^5+a_3\alpha^{10}+a_4\alpha^{20}+a_0\alpha^9+a_1\alpha^{18},
B_{init}+b_2\alpha^5+b_3\alpha^{10}+b_4\alpha^{20}+b_0\alpha^9+b_1\alpha^{18}\}$,
$to^4 = \{R'+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^8 B_{init}^{16}+(\alpha^3+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^8 B_{init}^4+(\alpha^2+\alpha) A_{init}^8 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}+(\alpha^4+\alpha) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^2+\alpha) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^2) A_{init}^2 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}+(\alpha^3+\alpha+1) A_{init} B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init} B_{init}^8+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^3+\alpha^2+1) A_{init} B_{init}^2+(\alpha) A_{init} B_{init},
A_{init}+a_1'\alpha^5+a_2'\alpha^{10}+a_3'\alpha^{20}+a_4'\alpha^9+a_0'\alpha^{18},
B_{init}+b_1'\alpha^5+b_2'\alpha^{10}+b_3'\alpha^{20}+b_4'\alpha^9+b_0'\alpha^{18}\}$
\item Clock 5: $from^4 = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^8 B_{init}^{16}+(\alpha^3+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^8 B_{init}^4+(\alpha^2+\alpha) A_{init}^8 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}+(\alpha^4+\alpha) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^2+\alpha) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^2) A_{init}^2 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}+(\alpha^3+\alpha+1) A_{init} B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init} B_{init}^8+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^3+\alpha^2+1) A_{init} B_{init}^2+(\alpha) A_{init} B_{init},
A_{init}+a_1\alpha^5+a_2\alpha^{10}+a_3\alpha^{20}+a_4\alpha^9+a_0\alpha^{18},
B_{init}+b_1\alpha^5+b_2\alpha^{10}+b_3\alpha^{20}+b_4\alpha^9+b_0\alpha^{18}\}$,
$to^5 = \{R'+A_{init}B_{init},
A_{init}+a_0'\alpha^5+a_1'\alpha^{10}+a_2'\alpha^{20}+a_3'\alpha^9+a_4'\alpha^{18},
B_{init}+b_0'\alpha^5+b_1'\alpha^{10}+b_2'\alpha^{20}+b_3'\alpha^9+b_4'\alpha^{18}\}$
\end{itemize}
The final result is
$from^5(R_{final}) = R_{final}+A_{init}\cdot B_{init}$
\end{Example}

### 2015-05-01 techwriting/ASPDAC/mult.tex
all write-up from virtue machine

\section{Sequential GF Multiplier Design}
\label{sec:design}

Let us briefly describe the fundamentals behind the design of normal
basis sequential GF multipliers, so as to put in perspective the type
of designs that have been verified in this paper. Let $R =
\sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
\beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$, then 
\[
R = A\cdot B = (\sum_{i=0}^{k-1} a_i \beta^{2^{i}}) (\sum_{j=0}^{k-1}
b_j \beta^{2^{j}})  =
\sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}\nonumber 
\]

The expressions $\beta^{2^i}\beta^{2^j}$ are called cross-product
terms and they can also be represented in normal basis: 
\begin{displaymath}
\beta^{2^i}\beta^{2^j} =
\sum_{n=0}^{k-1}\lambda_{ij}^{(n)}\beta^{2^n}, \ \ \lambda_{ij}^{(n)}
\in \Ftwo. 
\end{displaymath}

From the above two equations, one can see that the expression for the
$n^{th}$ digit of product $R = (r_0, \dots, r_n, \dots r_{k-1})$ is:
\[
r_n = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(n)}a_ib_j = A
\cdot M_n \cdot B^T, ~~0 \leq n \leq k-1
\]

where $M_n = (\lambda_{ij}^{(n)})$ is a binary $k \times k$ matrix over
$\Ftwo$, and it is called the $\lambda$-matrix. 
%The collection of all
%$\{M_n\}$  $\lambda$-matrices is called the multiplication
%table. 
Moreover, let $r_n = A \cdot M^{(n)} \cdot B^T$.
Then $r_{n-1} = A \cdot M^{(n-1)} \cdot B^T = rotate(A) \cdot M^{(n)}
\cdot rotate(B)^T$. This implies that $M^{(n)}$ is generated by right
and down cyclic shifting $M^{(n-1)}$. Therefore, the hardware design
of sequential GF multipliers is based on mappings of $A\cdot M_n
\cdot B^T$ into AND-XOR gates and cyclic shift operations.

This paper verifies the implementation of two distinct 
architectures of {\it sequential multipliers with parallel output
  (SMPO)}, namely: i) the Agnew-SMPO  \cite{agnew1991implementation}
by G. B. Agnew, which is a straight-forward implementation of the
$\lambda$-matrix; and ii) the more recent, more complicated, yet very
efficient RH-SMPO \cite{RHmulti}, by Reyhani-Masoleh and Hasan,
depicted in Fig. \ref{fig:RHmulti}.

\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=4in]{./RH.eps}
\vspace{-0.2in}
\caption{A 5-bit Reyhani-Hasan Sequential Multiplier with Parallel Outputs (RH SMPO)}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}

### 2015-05-01 techwriting/ASPDAC/new_RATO.tex
all write-up from virtue machine

\section{Improving our Approach}
\label{sec:improve}

Computing \Grobner bases with elimination orders is infeasible for
large circuits; the complexity of Buchberger's algorithm to compute
$GB(J + J_0)$ in $\Fq$ is $q^{O(d)}$ \cite{lv:tcad2013}. 
To overcome this complexity, \cite{pruss:dac14} proposed a
\underline{r}efinement of ATO (called RATO), and simplified the
\Grobner basis computation. They exploited Buchberger's product
criteria \cite{productc:1979}, which states that: {\it If the leading  
monomials of $f_i, f_j$ are relatively prime, then $Spoly(f_i, f_j)$
reduces to zero modulo the generating set, i.e. $Spoly(f_i, f_j)
\xrightarrow{F} _+ 0$.} This concept was exploited in RATO as follows:

Perform a reverse topological sorting of the nodes in the
combinational logic, and define a {\it lex} term order by the
following relation $>_{r}$: {\it bit-level circuit variables ordered
  reverse topologically} $>$  {\it word-level output variables} $>$
{\it word-level input   variables}. Representing the polynomial ideal
$J$ in RATO has the effect that there exists {\it one and only one
  pair of polynomials} in $J$ that do not have relatively prime
leading terms (see  Section 5 in \cite{pruss:dac14} for details). All
other polynomial pairs will have leading terms that are 
relatively prime, so these polynomial pairs are not considered in
Buchberger's algorithm.  The authors of \cite{pruss:dac14} exploited
this concept and showed how the \Grobner basis of $(J + J_0)$ can be
computed by a {\it subset} of polynomials, which  improves the
scalability of their approach. Their approach, however, cannot
circumvent the \Grobner basis computation altogether. Consequently, 
their approach fails to derive a canonical polynomial abstraction when
the representation is dense, and contains monomials of high-degrees
(e.g. in case of buggy designs).

It turns out that RATO can be applied to sequential circuits in much
the same way: {\it \{bit-level circuit variables ordered
  reverse topologically} $x_1 > \dots x_d\}$ $>$  {\it \{word-level NS
  variables} $R'> A'> B'\}$ $>$ {\it \{word-level PS variables}
$R>A>B\}$. Importantly, {\it we show that using RATO, the polynomial
abstraction can be derived without resorting to a \Grobner basis
computation.} Perform the following operations:

\begin{enumerate}
\item Represent the polynomials of the sequential circuit $S$ using RATO.
\item Due to RATO, only one pair of polynomials ($f_i, f_j$) will have
  leading terms that will not be relatively prime.
\item Reduce $Spoly(f_i, f_j) \xrightarrow{F}_+ h$.
\end{enumerate}
As described in \cite{pruss:dac14}, remainder $h$ will contain:
  i) the word-level variables, and ii) bit-level inputs to the
  combinational logic, i.e. bit-level present-state variables. All
  other internal circuit variables will not appear in $h$, as they
  will be canceled by division due to RATO.

\begin{Example}
\label{ex:newRATO}
Let us re-visit Example \ref{ex:RHSMPO} and the RH-SMPO circuit of
Fig. \ref{fig:RHmulti}. For this circuit, the term order under
RATO is lex with: $\{r_0',r_1',r_2',r_3',r_4'\} > \{r_0, r_1, r_2,r_3,r_4\}
> \{e_0,e_3,e_4\}, \{d_0,d_1,d_2\}, \{c_1,c_2,c_3,c_4\} 
\{ a_0, a_1, a_2, a_3, a_4, \\
b_0, b_1, b_2,  b_3, b_4 \} > R'>R>\{A,B\}$.

Among all the generators of the ideal $J$ from Ex.\ref{ex:RHSMPO},
using RATO \textbf{we find only one pair of polynomials} whose leading  
monomials are not relatively prime:
$(f_i, f_j), f_i = r_0'+r_4+e_0, f_j
=r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18}
+ R'$. Then: %Buchberger's algorithm will compute $Spoly(f_w,
\vspace{-0.1in}
{\small
\begin{align}
&Spoly(f_i,f_j) \xrightarrow{J+J_0}_{+}\nonumber\\
&(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
&+(\alpha^3+\alpha^2) a_1 b_1+(\alpha^4+\alpha^3+\alpha^2+\alpha) a_1 b_2+(\alpha^2+\alpha) a_1 b_3\nonumber\\
&+(\alpha^2+1) a_1 b_4+(\alpha^4+1) a_1 B+(\alpha^4+\alpha) a_2 b_1+(\alpha^4+\alpha^3+\alpha) a_2 b_2\nonumber\\
&+(\alpha^3+1) a_2 b_3+(\alpha^3+\alpha^2+1) a_2 b_4+(\alpha^3+\alpha^2) a_2 B+(\alpha^2+\alpha) a_3 b_1\nonumber\\
&+(\alpha^3+1) a_3 b_2+(\alpha+1) a_3 b_3+(\alpha^4+\alpha^2+\alpha) a_3 b_4\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) a_3 B+(\alpha^3+1) a_4 b_1+a_4 b_2+(\alpha^4+\alpha^2+\alpha) a_4 b_3\nonumber\\
&+(\alpha^4+\alpha^3+1) a_4 b_4+(\alpha^2+\alpha) a_4 B+(\alpha^4+1) b_1 A+(\alpha^3+\alpha^2) b_2 A\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) b_3 A+(\alpha^2+\alpha) b_4 A+(\alpha^4+\alpha^2+\alpha+1) R'+A B\nonumber
\end{align}
}
\end{Example}
\vspace{-0.1in}
As shown above, the remainder $h$ contains both bit-level variables and
word-level {\it state variables} $a_i, b_i, r_i, R, A, B$. We desire
a polynomial in only word 
level variables (e.g. $R'+\mathcal{F}(A,B)$), without computing a
\Grobner basis. This can be achieved if we  represent the
bit-level state variables $a_i, b_i, r_i$ in terms of their word-level
variables $A, B, R$, respectively. This can be achieved due to 
the following property of finite fields:

\begin{Lemma}
\label{lemma:square}
For $\alpha_1, \dots, \alpha_t \in \Fkk$, $(\alpha_1 + \alpha_2 +
\dots + \alpha_t)^{2^i} = \alpha_1^{2^i} + \alpha_2^{2^i} + \dots +
\alpha_t^{2^i}$, for $i = 1, 2, \ldots$. 
\end{Lemma}

Consider the polynomials that define the relationship between the
word-level and bit-level variables. Let 
$ A = a_0 \beta + a_1\beta^2 + \dots + a_{k-1}\beta^{2^{k-1}}$.
Due to Lemma \ref{lemma:square}, we have $A^2 = a_0 \beta^{2}+ a_1
\beta^{4} + \dots + a_{k-1} \beta^{2\cdot 2^{k-1}}$ (as
$a_i^2 = a_i$). By repeated squaring:
\begin{eqnarray}
A & = & a_0 \beta + a_1\beta^2 + \dots + a_{k-1}\beta^{2^{k-1}}\nonumber\\
A^2 & = & a_0 \beta^{2}+ a_1 \beta^{4} + \dots + a_{k-1}
\beta^{2\cdot 2^{k-1}} \nonumber\\
\vdots & \vdots& \vdots \nonumber\\
A^{2^{k-1}} & = & a_0 \beta^{2^{k-1}}  +  a_1 \beta^{2^{k-1}\cdot 2} +
\dots + a_{k-1}\beta^{2^{2(k-1)}} \nonumber
\end{eqnarray}
Consider the above as $k$ {\it linear equations} with $a_0, \dots, a_{k-1}$
as $k$-unknowns, with $\beta$ and its powers as coefficients, and 
$A, A^2, ... \dots, A^{2^{k-1}}$ as $k$ constants. Then, we can solve
for the unknowns $a_0, \dots, a_{k-1}$ and obtain expressions for them
in terms of $A$ and $\beta$. The problem can be setup in
matrix form:

\begin{displaymath}
\begin{pmatrix}
\beta & \beta^2 & \cdots & \beta^{2^{k-1}} \\
\beta^2 & \beta^4 & \cdots & \beta^{2^{k-1}\cdot 2} \\
\vdots & \vdots & \ddots & \vdots \\
\beta^{2^{k-1}} & \beta^{2^{k-1}\cdot 2} & \cdots & \beta^{2^{2(k-1)}}
\end{pmatrix}
\begin{pmatrix}
a_0\\
a_1\\
\vdots\\
a_{k-1}
\end{pmatrix}
=
\begin{pmatrix}
A\\
A^2\\
\vdots\\
A^{2^{k-1}}
\end{pmatrix}
\end{displaymath}

Gaussian elimination on this system of equations can be 
applied, and each bit-level variable $a_i$ can be represented as a
function of the word-level variables: $a_i = g_i(A)$. These bit-level
variables can be easily substituted in the remainder $h$ obtained by
reduction due to RATO. What we will obtain is a word-level polynomial
of the form $h: R' + \F(A, B)$ which canonically represents the
$k$-cycle function of the circuit. It is also easy to show that this
polynomial is a unique, canonical representation because it is reduced
modulo $J + J_0$. 
% = \langle x_1^2 - x_1, \dots, x_d^2 - x_d, R^q - R, A^q -
%A, B^q - B\rangle$.

%% By replacing all bit-level variables by corresponding word-level
%% variable polynomials, we transform the remainder of $Spoly$ reduction
%% to the form of $R'+R+\mathcal F'(A,B)$. Note $R$ is present state
%% notion of output, which equals to initial value $R=0$ in first clock
%% cycle, or value of $R'$ from last clock cycle. By substituting $R$
%% with its corresponding value ($0$ or a polynomial only about $A$ and
%% $B$), we get the desired polynomial function $R'+\mathcal F(A,B)$.

### 2015-05-01 techwriting/ASPDAC/new_RATO_xiaojun.tex
all write-up from virtue machine

\section{Improvement on RATO Based Abstraction}
A lexicographic order defined by following relation $>_{r}$: "circuit variables ordered reverse topologically" $>$ 
"word-level output" $>$ "word-level inputs" is called the \emph{Refined Abstraction Term Order (RATO)}. Main procedure
of abstraction based on RATO proposed by T. Pruss(cite Tim here)\cite{TimDAC} is: first, in ideal generated by
gates information polynomials and word-level variable definition polynomials, find the unique pair of polynomial
generators with leading monomials not relatively prime to each other; then, compute their specification polynomial
using definition 
$$Spoly(f_w,f_g) = \frac{LCM}{lt(f_w)}\cdot f_w - \frac{LCM}{lt(f_g)}\cdot f_g$$
where $LCM$ is least common multiple of $lm(f_w)$ and $lm(f_g)$, and $lt$ denotes the leading term;
last, reduce $Spoly$ with ideal $J_{ckt}+J_0$, it is possible that the remainder will be canonical polynomial
function of the circuit. This technique
cannot obviate the need of Gr\"obner basis computation under some situations, as discussed
in section \ref{sec:prework}. To better illustrate those limitations, we apply RATO on 5-bit RH-SMPO in Fig.\ref{fig:RHmulti}.

\begin{Example}
\label{ex:newRATO}
Variable order under RATO is:
\begin{align}
&\{r_0',r_1',r_2',r_3',r_4'\}>\{r_0,r_1,r_2,r_3,r_4\}\nonumber\\
&>\{e_0,e_3,e_4\},\{d_0,d_1,d_2\},\{c_1,c_2,c_3,c_4\}\nonumber\\
&>\{a_0,a_1,a_2,a_3,a_4,b_0,b_1,b_2,b_3,b_4\}>R'>R>\{A,B\}\nonumber
\end{align}
Search among all generators of $J_{ckt}$ from Ex.\ref{ex:RHSMPO} using RATO, we find a pair of polynomials whose leading 
monomials are not relatively prime:
$(f_w,f_g), f_w = r_0'+r_4+e_0, f_g =r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18} + R'$.
We calculate $Spoly$ can reduce it by $J_{ckt}+J_0$:
\begin{align}
&Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
&(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
&+(\alpha^3+\alpha^2) a_1 b_1+(\alpha^4+\alpha^3+\alpha^2+\alpha) a_1 b_2+(\alpha^2+\alpha) a_1 b_3\nonumber\\
&+(\alpha^2+1) a_1 b_4+(\alpha^4+1) a_1 B+(\alpha^4+\alpha) a_2 b_1+(\alpha^4+\alpha^3+\alpha) a_2 b_2\nonumber\\
&+(\alpha^3+1) a_2 b_3+(\alpha^3+\alpha^2+1) a_2 b_4+(\alpha^3+\alpha^2) a_2 B+(\alpha^2+\alpha) a_3 b_1\nonumber\\
&+(\alpha^3+1) a_3 b_2+(\alpha+1) a_3 b_3+(\alpha^4+\alpha^2+\alpha) a_3 b_4\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) a_3 B+(\alpha^3+1) a_4 b_1+a_4 b_2+(\alpha^4+\alpha^2+\alpha) a_4 b_3\nonumber\\
&+(\alpha^4+\alpha^3+1) a_4 b_4+(\alpha^2+\alpha) a_4 B+(\alpha^4+1) b_1 A+(\alpha^3+\alpha^2) b_2 A\nonumber\\
&+(\alpha^4+\alpha^3+\alpha) b_3 A+(\alpha^2+\alpha) b_4 A+(\alpha^4+\alpha^2+\alpha+1) R'+R+A B\nonumber
\end{align}
\end{Example}
Above example indicates that RATO based abstraction on 5-bit RH-SMPO will result a remainder contains both
bit-level variables and word-level variables, and the number of remaining variables is still large such that
Gr\"obner basis computation will be inefficient.

\subsection{Improved RATO based Abstraction}
The remainder from \emph{Spoly} reduction contains some bit-level variables, and our objective is to compute
a polynomial contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
this problem is to replace bit-level variable monomials by equivalent polynomials that only contain word-level
variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination-fashion 
approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.

\begin{Example}
{\bf Objective}:\ Compute polynomial $a_i + \mathcal{G}_i(A)$ from $f_0 = a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A
 = g_0 + A$.

First, compute $f_0^2= a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2 = g_0^2 + A^2$;
then $f_0^4, f_0^8, f_0^{16}$. By repeating squaring we get a system of equations:
\begin{displaymath}
  \begin{cases}
  f_0  &=\ \  0\\
  f_0^2 & =\ \  0\\
  f_0^4  &=\ \  0\\
  f_0^8  &=\ \  0\\
  f_0^{16} & =\ \ 0
  \end{cases}
\Longleftrightarrow\ \
\begin{cases}
  g_0 & =\ \  A\\
  g_0^2 & =\ \  A^2\\
  g_0^4 & =\ \  A^4\\
  g_0^8 & =\ \  A^8\\
  g_0^{16} & =\ \  A^{16}
  \end{cases}
\end{displaymath}
Following is the coefficients matrix form of this system of equations:
\begin{displaymath}
\begin{pmatrix}
\alpha^5 & \alpha^{10} & \alpha^{20} & \alpha^9 & \alpha^{18} \\
\alpha^{10} & \alpha^{20} & \alpha^9 & \alpha^{18} & \alpha^5 \\
\alpha^{20} & \alpha^9 & \alpha^{18} & \alpha^5 & \alpha^{10} \\
\alpha^9 & \alpha^{18} & \alpha^5 & \alpha^{10} & \alpha^{20} \\
\alpha^{18} & \alpha^5 & \alpha^{10} & \alpha^{20} & \alpha^9 
\end{pmatrix}
\begin{pmatrix}
a_0\\
a_1\\
a_2\\
a_3\\
a_4
\end{pmatrix}
=
\begin{pmatrix}
A\\
A^2\\
A^4\\
A^8\\
A^{16}
\end{pmatrix}
\end{displaymath}

Use Gaussian elimination on coefficients matrix, for example  
\begin{align}
Row\ 2 =& Row\ 1\times \alpha^5 + Row\ 2: \nonumber\\
&a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
=&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
\end{align}
Recursively eliminate $a_1$ from third row, $a_2$ from fourth row, etc. The final solution to this system of
equations is
%\begin{displaymath}
%\{g_i\ | \ g_i: a_i + \mathcal{G}_i(A)\}
%\end{displaymath}
%\end{Example}
\begin{displaymath}
\left\{
  \begin{array}{lcl}
  a_0 & = & (\alpha+1)A^{16}+(\alpha^4+\alpha^3+\alpha)A^8+(\alpha^3+\alpha^2)A^4\\&&+(\alpha^4+1)A^2+(\alpha^2+1)A\\
  a_1 & = & (\alpha^2+1)A^{16}+(\alpha+1)A^8+(\alpha^4+\alpha^3+\alpha)A^4\\&&+(\alpha^3+\alpha^2)A^2+(\alpha^4+1)A\\
  a_2 & = & (\alpha^4+1)A^{16}+(\alpha^2+1)A^8+(\alpha+1)A^4\\&&+(\alpha^4+\alpha^3+\alpha)A^2+(\alpha^3+\alpha^2)A\\
  a_3 & = & (\alpha^3+\alpha^2)A^{16}+(\alpha^4+1)A^8+(\alpha^2+1)A^4\\&&+(\alpha+1)A^2+(\alpha^4+\alpha^3+\alpha)A\\
  a_4 & = & (\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4\\&&+(\alpha^2+1)A^2+(\alpha+1)A
  \end{array} \right.
\end{displaymath}
\end{Example}
Similarly we can compute equivalent polynomials $\mathcal{H}_i(R)$ for $r_i$, $\mathcal{T}_j(B)$ for $b_j$.
\textbf{(do we need a proof that remainder will only have $r_i,b_j$ here?)}

By replacing all bit-level variables by corresponding word-level variable polynomials, we transform the remainder
of $Spoly$ reduction to the form of $R'+R+\mathcal F'(A,B)$. Note $R$ is present state notion of output, which
equals to initial value $R=0$ in first clock cycle, or value of $R'$ from last clock cycle. By substituting
$R$ with its corresponding value ($0$ or a polynomial only about $A$ and $B$), we get the desired polynomial function
$R'+\mathcal F(A,B)$.

### 2015-05-01 techwriting/ASPDAC/new_theory_example.tex
all write-up from virtue machine

\section{Verification of Sequential GF Circuits}
\label{sec:theory}

We follow  the sequential GF circuit model of
Fig. \ref{fig:sequential}, with word-level variables $A, B, R$
denoting {\it present states (PS)} and $A', B', R'$ denoting {\it next
  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
\beta^{2^i}$ for NS variables, and so on. The normal element $\beta$ is
given. Variables $R (R')$ correspond to those that 
store the result, and $A, B (A', B')$ store input operands. {\it E.g.,}
for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
0$) are the initial values (operands) loaded into the registers,  and
$R = \F(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
result after $k$-cycles. Our approach aims to find this polynomial
representation for $R$.

Each gate in the combinational logic is represented by a Boolean
polynomial. To 
this set of Boolean polynomials, we append the polynomials that define
the word-level to bit-level relations for PS and NS variables ($A =
\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
as ideal $J = \langle 
f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
  B']$. The ideal of vanishing polynomials $J_0$ is also included, and
then the implicit FSM unrolling problem is setup for abstraction.

The configurations of the flip-flops are the states of the
machine. {\it Since the set of states is a finite set of points, we
can consider it as the variety of an ideal related to the circuit
(from Section \ref{sec:prelim})}. Moreover, since we are interested in
the {\it function encoded} by the state variables (over $k$-time
frames), we can {\it project this variety} on the word-level state
variables, starting from the initial state $A_{init}, B_{init}$.
Projection of varieties (geometry) corresponds to elimination ideals
(algebra), and can be analyzed via \Grobner bases. Therefore, we
employ a \Grobner basis computation with ATO: we use a {\it lex term
  order} with {\it bit-level variables} 
$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
allows to eliminate all the bit-level variables 
%(corresponding to the combinational logic and the state variables),
%so as to 
and derives a representation only in terms of words. 
Consequently, $k$-successive \Grobner basis computations implicitly
unroll the machine, and provide word-level algebraic $k$-cycle
abstraction for $R'$ as $R' = \F(A_{init}, B_{init})$.

Algorithm
\ref{alg:modified} describes our approach.  In the algorithm, $from^i$
and $to^i$ are polynomial ideals whose varieties are the valuations of
word-level variables $R, A, B$ and $R',A',B'$ in the $i$-th iteration;
and the notation ``$\setminus$'' signifies that the $NS$ in iteration
$(i)$ becomes the $PS$ in iteration $(i+1)$. 
%The forward image
%$to^{i}$ is computed using \Grobner bases with ATO.

\vspace{-0.1in}
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
   state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from^0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
	$G \gets$GB$( \langle J + J_0+ from^{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to^i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from^i \gets to^i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from^k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}

\vspace{-0.1in}
\begin{Example}
\label{ex:RHSMPO}
We demonstrate our approach to verify the 5-bit RH-SMPO circuit of
Fig.\ref{fig:RHmulti}. The normal element $\beta$ in
$\mathbb{F}_{2^5}$ is known to be $\beta = \alpha^5$, where $\alpha$
is the primitive element. The circuit can be described by the ideal:
\begin{align}
J = & d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4,\nonumber\\
& c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2, \nonumber\\
& R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4,\nonumber\\
& A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},\nonumber\\
& B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18},\nonumber\\
& R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18},\nonumber\\
& R+r_0\alpha^5+r_1\alpha^{10}+r_2\alpha^{20}+r_3\alpha^9+r_4\alpha^{18};\nonumber
\end{align}

In the first iteration, $from^0 = \{R, 
~A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
~B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\}$
denotes the initial state.

After the GB computation is performed with ATO,  we find
$to^1 = \{R' (\alpha^4+\alpha^3+1) A_{init}^{16}
B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^{16}
B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1)
A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^8
B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8
B_{init}^4+(\alpha^3+\alpha+1) A_{init}^8
B_{init}^2+(\alpha^4+\alpha^2) A_{init}^8 B_{init}+(\alpha^4+\alpha^2)
A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4
B_{init}^8+(\alpha^2) A_{init}^4
B_{init}^4+(\alpha^3+\alpha^2+\alpha+1) A_{init}^4
B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4
B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha+1)
A_{init}^2 B_{init}^8+(\alpha^3+\alpha^2+\alpha+1) A_{init}^2
B_{init}^4+(\alpha^3+\alpha^2+\alpha) A_{init}^2
B_{init}^2+(\alpha^4+\alpha) A_{init}^2 B_{init}+(\alpha^4+\alpha^3+1)
A_{init} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}
B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}
B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init}^2+(\alpha^3+\alpha+1)
A_{init} B_{init}\}$ --- a polynomial in variables $R', A_{init}, B_{init}$.

Continuing this way, in the 5$^{th}$ iteration: $from^4 =
\{R+(\alpha^3+\alpha+1) A_{init}^{16} 
B_{init}^{16}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16}
B_{init}^8+(\alpha^4+\alpha) A_{init}^{16} B_{init}^4+(\alpha^3+1)
A_{init}^{16} B_{init}^2+(\alpha^3+\alpha+1) A_{init}^{16}
B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^8
B_{init}^{16}+(\alpha^3+1) A_{init}^8
B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^8
B_{init}^4+(\alpha^2+\alpha) A_{init}^8
B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}+(\alpha^4+\alpha)
A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^2+\alpha) A_{init}^4
B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^4
B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^3+1)
A_{init}^2 B_{init}^{16}+(\alpha^2+\alpha) A_{init}^2
B_{init}^8+(\alpha^4+\alpha^2) A_{init}^2
B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^2
B_{init}+(\alpha^3+\alpha+1) A_{init}
B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init}
B_{init}^8+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^3+\alpha^2+1)
A_{init} B_{init}^2+(\alpha) A_{init} B_{init}, 
~~A_{init}+a_1\alpha^5+a_2\alpha^{10}+a_3\alpha^{20}+a_4\alpha^9+a_0\alpha^{18},
~~B_{init}+b_1\alpha^5+b_2\alpha^{10}+b_3\alpha^{20}+b_4\alpha^9+b_0\alpha^{18}\}$
denotes the current state values.

Finally, by computing GB with ATO, we obtain: $to^5 = \{ \mathbf{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^5+a_1'\alpha^{10}+a_2'\alpha^{20}+a_3'\alpha^9+a_4'\alpha^{18},
~B_{init}+b_0'\alpha^5+b_1'\alpha^{10}+b_2'\alpha^{20}+b_3'\alpha^9+b_4'\alpha^{18}\}$
as the image. The final result is $from^5(R_{final}) = R_{final}+A_{init}\cdot
B_{init}$, which verifies the multiplier. 
\end{Example}

### 2015-05-01 techwriting/ASPDAC/oldtheory.tex
all write-up from virtue machine

--- 2014-04-25 sequential/latex/ICCAD2014/theory.tex
@@
 
-\begin{figure}[hbt]
+\begin{figure}[tb]
 \begin{center}
 \includegraphics[width=3.5in]{./sequential_fig.eps}
+
 \end{center}
@@
 $\mathbb{B}^k \to \mathbb{F}_{2^k}$, which means it is possible to find a set of elements in $\mathbb{F}_{2^k}$ representing
-all Boolean values every bit of a bit vector can take. In Galois field, a as \ref{table:booltogalois} shows.
-\begin{table}[h]
+all Boolean values every bit of a bit vector can take. In Galois field, an example is in table \ref{table:booltogalois}.
+\begin{table}[tb]
 \centering
+\caption{Bit-vector, Exponential and Polynomial representation of
+elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
+\pmod{x^4+x^3+1}$}
 \begin{tabular}{|c|c||c|c|} 
@@
 \end{tabular}
-\caption{Bit-vector, Exponential and Polynomial representation of
-elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
-\pmod{x^4+x^3+1}$}\label{table:booltogalois}  
+\label{table:booltogalois}  
 \end{table}
@@
 \begin{Example}
-\label{ex:nb_multi}
+\label{ex:onb_multi}
 Normal basis multiplication: There are 2 binary vectors representing operands of multiplication in normal
 basis: $$A = (a_0, a_1, \dots, a_{n-1}),\ B = (b_0, b_1, \dots, b_{n-1})$$ 
-the product is also written in normal basis representation: $$C = A*B = (c_0, c_1, \dots, c_{n-1})$$
+the product is also written in normal basis representation: $$C = A\cdot B = (c_0, c_1, \dots, c_{n-1})$$
 Describe calculation procedure for the most significant digit of product with a function: 
 $$c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})$$
-Square both side: $C^2 = A^2*B^2$, i.e. the second significant digit 
+Square both side: $C^2 = A^2\cdot B^2$, i.e. the second significant digit 
 $$c_{n-2} = f(a_{n-1}, a_0, a_1, \dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2})$$ 
@@
 \begin{center}
-\includegraphics[scale=0.3]{./stg_fig.ps}
+\includegraphics[scale=0.3]{./stg_fig.eps}
+
 \end{center}
@@
 \includegraphics[width=3.5in]{./fsm_fig.eps}
+
 \end{center}
@@
 One important technique to check sequential equivalence is \textbf{state space traversal}. In this example,
-Breath-First-Search (BFS) is employed for traversal, which can be described with following algorithm:
+Breath-First-Search (BFS) is employed for traversal, which can be described with following algorithm\cite{KallaPartialScan}:
 
@@
 reversely using Shannon's expansion.
-\begin{table}
+\begin{table}[tb]
 \centering
+\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
 \begin{tabular}{|c|c|} \hline
@@
 \hline\end{tabular}
-\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
 \label{table:booltogalois_op}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{./mySMPO.eps}
+\includegraphics[width=\textwidth]{mySMPO.eps}
+
 \end{center}
-\caption{5-bit SMPO}
+\caption{5-bit SMPO.
+Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, i.e.
+if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
+Index $w = 2i\pmod 5$}
 \label{fig:SMPO}
@@
   	$i \gets i + 1$\;
-	$to^i \gets$GB w/ elimination term order$<I_{ckt}, from^{i-1}>$\;
-	$new^i \gets $generator of $<to^i> + (<T^4-T>:<reached>)$\;
-  	$reached \gets $generator of $<reached> \cdot <new^i>$\;
+	$to^i \gets$GB w/ elimination term order$\langle I_{ckt}, from^{i-1}\rangle$\;
+	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
+  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
 	$from^i \gets new^i(S\setminus T)$\;
@@
 \subsection{Sequential Galois Arithmetic Multiplier}
-Our experiment is to verify the function of a Galois field multiplier after certain clock cycles.
-The following design is Sequential Multiplier with Parallel Output (SMPO),
+Our experiment is to verify the function of a sequential Galois field multiplier, which will
+give the result of multiplication after certain clock cycles.
+Design shown in Fig.\ref{fig:SMPO} is Sequential Multiplier with Parallel Output (SMPO) proposed by 
+G.B. Angew et al.\cite{agnew1991implementation},
 a Normal Basis multiplier based on Massey-Omura algorithm. Inputs and outputs
-are all 5-bit word taking value from ${\mathbb F}_{2^5}$. After loading operands
- A and B, setting all output latches to 0 and running for 5 iterations, the 
+are all 5-bit words taking values from ${\mathbb F}_{2^5}$. After loading operands
+ $A$ and $B$, setting all output latches to 0 and running for 5 iterations, the 
 output should be $R = A\cdot B$ (mod $x^5 + x^2 + 1$).
 
-Similarly, build elimination ideal for all gates/operations and induce
-initial states of latches. However, instead of eliminating all variables
-to one, this example adopts abstraction term order
-and keeps the polynomial contains function between output and inputs.
-Here the lex ordering is $\{all\ bit$-$level\ variables\} > R > \{A, B\}$, and objective
+\begin{figure}[hbt]
+\begin{center}
+\includegraphics[width=3.5in]{./RH.eps}
+
+\end{center}
+\caption{5-bit RH-multiplier (AESMPO)}
+\label{fig:RHmulti}
+\end{figure}
+
+There is another type of SMPO known as RH-multiplier proposed by A. Reyhani-Masoleh and M.A. Hasan
+\cite{RHmulti}. Fig.\ref{fig:RHmulti} is the structure of 5-bit RH-multiplier (AND efficient SMPO, or AESMPO).
+Theory and design methodology about RH-multiplier is discussed in appendix \ref{sec:RHmulti}.
+
+To apply our approach to verify the function of above SMPO, first is to build elimination ideal for 
+all gates/operations and induce initial states of latches. Then, instead of eliminating all variables
+to one, we adopt abstraction term order and deduce abstraction polynomial describing the function 
+between output and inputs.
+Here the lexicographic order is $\{all\ bit$-$level\ variables\} > R > \{A, B\}$, and objective
 polynomial is $R + \mathcal{F}(A, B)$.
@@
 \begin{align}
-R_i &= b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
+c_i &= b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
 &+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
@@
 \end{align}
+based on results from Example \ref{ex:onbmulti}.
+
 It is possible to calculate every conjunctive term in the product simultaneously within one clock cycle,
@@
 \begin{Example}
-Sequential Multiplier Protocol:
+Angew's SMPO Protocol: Initial value of all outputs are 0s. After the first clock cycle, the $i$-th
+term of equation about $c_i$ is computed. In next clock cycle, output $R_i$ becomes state input $r_i$, which
+is used to compute $R_{i+1}: R_{i+1} = r_i + (a_u,a_v)b_w$. In this way, $R_i$ has one more term each clock cycle,
+at last after 5 clock cycles it has 5 terms, and equals to the equation about $R_i$.
 \begin{itemize}

### 2015-05-01 techwriting/ASPDAC/prelim.tex
all write-up from virtue machine

--- 2015-05-01 techwriting/ASPDAC/ganeshclass.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  IEEEsample.tex
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%    More information: see the header of IEEEtran.sty
-%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%
-
-\documentclass[conference,10pt]{IEEEtran}
-%\documentclass[conference]{IEEEtran}
-
-%%%\IEEEoverridecommandlockouts
-
-\usepackage[ruled]{./algorithm2e}
-%%for algorithm2e package, label has to be following caption in the same line!!!
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
-
-%% \RequirePackage{times}
-%% \RequirePackage{algorithmic}
-%% \PassOptionsToPackage{boxed}{algorithm}
-%% \RequirePackage{algorithm}
-%% \RequirePackage{multicol}
-%\renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
-%\renewcommand{\algorithmicensure}{\textbf{Outputs:}}
-%\DeclareMathAlphabet{\mathtsl}{OT1}{ptm}{m}{sl}
-
-%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em1
-%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-%\newtheorem{theorem}{Theorem}
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{example}{Example}
-%\newtheorem{corollary}{Corollary}
-
-\RequirePackage{amssymb, mathptm}
-\usepackage{amsbsy}
-\usepackage{graphicx}
-\usepackage{helvet}
-\usepackage{enumerate}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{subfig}
-\usepackage{comment}
-\usepackage{leqno}
-\usepackage{cases}
-
-%%indent in algorithm
-
-%\setcounter{page}{1}
-
-% New command for the table notes.
-\def\tabnote#1{{\small{#1}}}
-
-% New command for the line spacing.
-\newcommand{\ls}[1]
-    {\dimen0=\fontdimen6\the\font
-     \lineskip=#1\dimen0
-     \advance\lineskip.5\fontdimen5\the\font
-     \advance\lineskip-\dimen0
-     \lineskiplimit=.9\lineskip
-     \baselineskip=\lineskip
-     \advance\baselineskip\dimen0
-     \normallineskip\lineskip
-     \normallineskiplimit\lineskiplimit
-     \normalbaselineskip\baselineskip
-     \ignorespaces
-    }
-%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-%\renewcommand{\algorithmicensure}{\textbf{Output:}}
-
-\newcommand{\beq}{\begin{equation}}
-\newcommand{\eeq}{\end{equation}}
-\newcommand{\beqarr}{\begin{eqnarray}}
-\newcommand{\eeqarr}{\end{eqnarray}}
-%\newcommand{\ov}{\overline}
-\newcommand{\ov}{\bar}
-\newcommand{\xor}{\bigoplus}
-\newcommand{\Fm}{{\mathbb{F}}}
-
-%the following is for space before and after align or other equation environment.
-
-%%
-\newtheorem{Algorithm}{Algorithm}[section]
-\newtheorem{Definition}{Definition}[section]
-\newtheorem{Example}{Example}[section]
-\newtheorem{Proposition}{Proposition}[section]
-\newtheorem{Lemma}{Lemma}[section]
-\newtheorem{Theorem}{Theorem}[section]
-\newtheorem{Corollary}{Corollary}[section]
-\newtheorem{Conjecture}{Conjecture}[section]
-\newtheorem{Problem}{Problem}[section]
-\newtheorem{Notation}{Notation}[section]
-\newtheorem{Setup}{Problem Setup}[section]
-%%%
-
-%%set spacing between table columns
-\setlength{\tabcolsep}{3pt}
-
-\begin{document}
-
-%\thispagestyle{empty}
-%\pagestyle{empty}
-
-\title{Class Project Report: Boundary-free Technology Mapping using Gr\"obner Basis}
-
-\author{
-\IEEEauthorblockN{Xiaojun Sun}
-\IEEEauthorblockA{Electrical and Computer Engineering\\
-University of Utah\\
-Salt Lake City, Utah 84112\\
-Email: xiaojuns@ece.utah.edu}
-}
-
-%%\author{\IEEEauthorblockN{Jinpeng Lv and Priyank Kalla}\thanks{This work is sponsored in part by a grant from NSF \#CCF-546859.}
-%\IEEEauthorblockA{Department of  Electrical and Computer Eng.\\
-% University of Utah, Salt Lake City, UT-84112 \vspace{-0.3in}
- %\{lv, kalla\}@eng.utah.edu
-% }
-%\and
-%\IEEEauthorblockN{Florian Enescu} \thanks{\normalsize  978-3-9810801-8-6/DATE12/$\copyright 2012$ EDAA}
-%\IEEEauthorblockA{Department of Mathematics and Statistics\\
-% Georgia State University,  Atlanta, GA 30302-4038 \vspace{-0.3in}
-% fenescu@mathstat.gsu.edu
-%} 
-%
-
-\maketitle
-
-%\markboth{MS Proposal by Tim Pruss}{}
-\newcommand{\Fq}{{\mathbb{F}}_{q}}
-\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
-\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
-\newcommand{\Grobner}{Gr\"{o}bner\xspace}
-\newcommand{\B}{{\mathbb{B}}}
-\newcommand{\Z}{{\mathbb{Z}}}
-\newcommand{\F}{{\mathcal{F}}}
-\newcommand{\G}{{\mathcal{G}}}
-\newcommand{\R}{\mathbb{R}}
-%%%
-
-\newcommand{\debug}[1]{\textcolor{gray}{[ #1 ]}}
-
-%\thispagestyle{empty}
-
-%%%%%%%%%%%%%%%%%%%% Include your files here %%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-As FPGA is widely used in hardware implementation and the size of design is usually large, there 
-is an increasing need in industrial domain for an approach that can map macro blocks. Currently
-there is no algorithms can map large-scale functions without boundary information, especially
-arithmetic blocks in datapath. Our approach provides an alternative inspiration to solve
-this problem using Gr\"obner basis based ideal membership test, whose efficiency does not
-rely on the complexity of arithmetic block's inner structure.
-\end{abstract}
-
-\section{Introduction}
-\subsection{Problem Description}
-One sentence on this problem is: find a method to map the macro blocks without boundary information.
-Mapping is an essential technique used in synthesis and verification. In synthesis, we can map the 
-macro functions with smaller and faster implementations to optimize the timing and area; in simulation,
-we can map a complicated function to a simple execution to accelerate simulation speed.
-
-Given a gate-level design $D$ and several word-level macro blocks $\{B_i\}$, we need to map macro
-blocks $B_i$ into design $D$, and write out the mapped design $D'$ which is equivalent to original
-design $D$. The objective is to generate a mapped design $D$ with as many of $B_i$ such that the area
-and timing is optimal.
-
-\begin{figure}[hbt]
-	\begin{center}
-	\includegraphics[scale=0.3]{macro.png}
-	\end{center}
-	\caption{The outline and flow of mapping for macro blocks}
-	\label{fig:macro}
-\end{figure}
-
 \section{Preliminaries}
-\label{sec:intro}
-
-\subsection{Modeling of constrains by ideal}
-A state can be described by a set of constrains. Generally speaking, a constrain is a function mapping 
-from variables to boolean values ($f:X\to bool,\ X=\{x_0,\dots,x_n\}$); meanwhile this "function" can be
-written in form of polynomial about variables. If the constrain is satisfied, i.e. $f:\sigma(X)\to \perp$,
+\label{sec:prelim}
+%\subsection{Computer Algebra Preliminaries}
+\subsection{Modeling of Constrains by Ideal}
+A state can be described by a set of constrains. Generally, a constrain is a function mapping 
+from field variables to boolean values ($f:X\to bool,\ X=\{x_0,\dots,x_n\}$); meanwhile this "function" can be
+written in form of polynomial about field variables. If the constrain is satisfied, i.e. $f:\sigma(X)\to \perp$,
 accordingly the evaluation of this polynomial is 0:
@@
 then above implication can hold. Here coefficients $c_j$ could be any polynomials
-or real numbers. Now review Definition 2 in original paper:
+or real numbers. We introduce a new concept to represent this expression:
 \begin{Definition}
@@
 \end{Definition}
-Thus we can formally say
+Thus we conclude following theorem:
 \begin{Theorem}
@@
 
-\subsection{Ideal membership and Gr\"obner basis}
-From last section, to check if $\F$ is a valid abstraction, it is necessary to check if 
-$\F\in J=\langle x_0,\dots,x_n\rangle$, this requires a technique called \emph{ideal membership test}.
+\subsection{Ideal Membership and Gr\"obner Basis}
+As discussed above, in order to check whether $\F$ is a valid abstraction, it is necessary to check if 
+$\F\in J=\langle x_0,\dots,x_n\rangle$, which requires a technique called \emph{ideal membership test}.
 To execute this technique, $\langle f_0,\dots,f_k\rangle$ need to be transformed to \emph{Gr\"obner Basis}
@@
 
-If $J = \langle f_0,\dots,f_k\rangle$, $\F\xrightarrow{J}_+r$, apparently we have
+If $J = \langle f_0,\dots,f_k\rangle$, $\F\xrightarrow{J}_+r$, we can deduce that
 $$\F = c_0f_0+c_1f_1+\cdots+c_kf_k + r$$
-if the remainder $r=0$, then $\F$ is the desired constrain ($\F\in J$). Now look back to the property of
-Gr\"obner basis, when $\F\in J$, then every step updated $\F$ also belongs to $J$, which means for every
+if the remainder $r=0$, then $\F$ is the desired constrain ($\F\in J$). According to the property of
+Gr\"obner basis, if $\F\in J$ then during all steps updated $\F$ also belongs to $J$. Based on this result,
+for every
 term in original $\F$ we can find corresponding $g_i$ whose leading term can divide it, thus the remainder
@@
 Buchberger's algorithm
-shown in Algorithm \ref{alg:gb},  computes a Gr\"obner
-basis over a field. Given polynomials $F = \{f_1, \dots, f_s\}$, 
+shown in Algorithm 1 computes a Gr\"obner
+basis in a field. Given polynomials $F = \{f_1, \dots, f_s\}$, 
 the algorithm computes the Gr\"obner basis $G = \{g_1, \dots, 
@@
 \begin{algorithm}[hbt]
+\label{alg:gb}
 \SetAlgoNoLine
@@
    }
-\caption {Buchberger's Algorithm}\label{alg:gb}
+\caption{Buchberger's Algorithm}
 \end{algorithm}
 
-\subsection{Elimination using Gr\"obner Basis}
-\label{sec:elim}
-We are given a circuit $C$ with $k$-bit inputs and outputs that
-performs a polynomial computation $Y = \F(A)$ over $\Fq = \Fkk$. Let
-$P(x)$ be the {\it given} irreducible or primitive polynomial used for
-field construction, and let $\alpha$ be its root, i.e. $P(\alpha) = 0
-$. Note that we do not know the polynomial representation
-$\F(A)$ and our objective is to identify (the coefficients of)
-$\F(A)$. Let $\{a_0, \dots, a_{k-1}\}$ denote the primary inputs and
-let $\{y_0, \dots, y_{k-1}\}$ be the primary outputs of $C$. Then, the
-word-level and bit-level correspondences are: 
-\begin{equation}
-\label{eqn:words}
- A = a_0 + a_1 \alpha + \dots + a_{k-1} \alpha^{k-1}; ~~ Y = y_0 +
-y_1 \alpha + \dots + y_{k-1} \alpha^{k-1};
-\end{equation}
+{\it Fermat's little theorem:} For any $ \alpha \in \mathbb
+F_{q}, \alpha^q = \alpha$. Therefore, the polynomial $x^q - x$
+vanishes ($=0$) over $\Fq$, and is called a vanishing polynomial. We
+denote by $J_0 = \langle x_1^q - x_1, \dots, x_d^q - x_d \rangle$ the
+ideal of all vanishing polynomials in $\Fq[x_1, \dots, x_d]$. When $q
+= 2^k, x^q - x = x^q + x$ as $-1 = +1$ over $\Fkk$.
 
-We analyze the circuit and model all the gate-level Boolean operators
-as polynomials in ${\mathbb{F}}_2 \subset \Fkk$. To this set of
-Boolean polynomials, we append the polynomials of Eqn
-(\ref{eqn:words}) that relate the word-level and bit-level
-variables. We model this set of polynomials as $F = \{f_1, \dots,
-f_s\}$ over the ring $R = \Fq[x_1, \dots, x_d, Y, A]$. Here $x_1,
-\dots, x_d$ denote, collectively, all the bit-level variables of the
-circuit --- i.e. primary inputs, primary outputs and the intermediate
-circuit variables --- and $Y, A$ the word-level variables. Denote the
-generated ideal as $J = \langle F \rangle \subset R$. As $Y = \F(A)$
-is a polynomial representation of the circuit, represent this (unknown)
-``specification'' as a polynomial $f: Y - \F(A)$, or as $f: Y + \F(A)$
-and $-1 = +1$ over $\Fkk$.
-
-As the circuit $C$ implements the function $f$, clearly $f$ {\it
-  agrees with the solutions} to $f_1 = \dots = f_s = 0$. In computer
-algebra terminology, this means that $f$ {\it vanishes on the variety} 
-$V_{\Fq}(J)$. If $f$ vanishes on $V_{\Fq}(J)$, then $f$ is a member of
-the ideal $I(V_{\Fq}(J))$. Strong Nullstellensatz over Galois fields tells us that $I(V_{\Fq}(J)) =
-J + J_0$, where $J_0 = \langle x_1^q - x_1, \dots, x_d^q - x_d, Y^q -
-Y, A^q - A \rangle$ is the ideal of vanishing polynomials in
-$R$. Consolidating these results, we deduce that the specification
-polynomial $f \in (J+J_0)$.
-
-If the specification polynomial is known, then the verification
-problem can be solved using membership testing of $f$ in the ideal $(J
-+ J_0)$. We will now show
-that by computing a Gr\"obner basis of $(J + J_0)$, using a specific
-elimination term order, we can also identify the polynomial $f$ which
-represents the function implemented by the circuit.
-
-Moreover, Gr\"obner bases may be used to generate an elimination ideal
-by using an ``elimination order.''  One such ordering
-is a pure lexicographic ordering, which features into a theorem:
-%The ideal $I_k$ is spanned by elements which eliminate variables
-%$x_1,\dots,x_k$.  This leads to an {\bf elimination theorem} for \Grobner
-%bases:
-\begin{Theorem} \label{thm:elim}
-({\bf Elimination Theorem}) Let $J
-  \subset \Fq[x_1,\dots,x_d]$ be an     ideal and let $G$ be a
-  Gr\"obner basis of $J$ with respect to a lex ordering where $x_1
-  > x_2 > \dots > x_d$.  Then for every $0     \leq l \leq
-  d$, the set 
-    \begin{equation}
-        G_l= G \cap \Fq[x_{l+1},\dots,x_d]
-    \end{equation}
-    is a Gr\"obner basis of the $l$th elimination ideal $J_l$.
+Gr\"obner bases can be used to {\it eliminate} variables from an
+ideal. Given ideal $J = \langle f_1,\dots,f_s\rangle \subset \mathbb
+F_{q}[x_1,\dots,x_d]$, the $l^{th}$ elimination ideal $J_l$ is the
+ideal of $\Fq[x_{l+1}, \dots, x_d]$ defined by $J_l = J \cap
+\Fq[x_{l+1}, \dots, x_d]$. Variable elimination can be achieved 
+by computing a Gr\"obner basis of $J$ w.r.t. elimination orders: 
+\begin{Theorem}
+(Elimination Theorem\cite{ideals:book}) Let $J\subset \mathbb
+  F_{2^k}[x_1,\dots,x_d]$ be an ideal and let $G$ be a Gr\"obner basis
+  of $J$ with respect to a lexicographic ordering where
+  $x_1>x_2>\cdots>x_d$. Then for every $0\leq l\leq d$, the set $G_l =
+  G\cap\mathbb F_{2^k}[x_{l+1},\dots,x_d]$ is a Gr\"obner basis of
+  the $l$-th elimination ideal $J_l$.
 \end{Theorem}
 
-Gr\"obner basis computations w.r.t. pure lexicographic
-term orders can be used to eliminate variables from an ideal. The
-above example motivates our approach: Since we want to derive a
-polynomial representation from a circuit in variables $Y, A$, we can
-compute a Gr\"obner basis of $J + J_0$ w.r.t. an elimination order
-that eliminates all the ($d$) bit-level variables of the
-circuit. Then, the Gr\"obner basis $G_d = G \cap \Fq[x_1, \dots, x_d,
-  Y, A]$ of the $d^{th}$ elimination ideal of $(J + J_0)$ will contain
-polynomials in only $Y, A$. We now prove that the 
-required polynomial representation will be found in $G_d$.
+\subsection{Polynomial Abstraction with Elimination \& Gr\"obner Bases}
 
-\section{Techniques from Loop Invariant paper}
-This section briefly illustrates S. Sankaranarayanan's POPL'04 conference paper.
-First, a transition system is modeled by algebraic assertions; then \emph{ideal membership test}
-is applied on the set of assertions to help abstract the loop invariant. "Template" is the concept
-I borrowed from this paper to apply on my own approach.
-\subsection{Template and state constrains}
-\label{sec:basic}
-To effectively and efficiently utilize the ideal membership test technique, a heuristic to
-generate $\F$ is needed. One heuristic can provide better coverage for loop invariant abstraction as well as
-relatively small size is \emph{generic quadratic form}.
+The authors of \cite{pruss:dac14} showed that for any combinational
+logic block, a canonical word-level polynomial representation can be
+derived through \Grobner bases computed with elimination
+orders. Our approach is based on their result, which we reproduce
+here:
+\begin{Lemma}
+(From \cite{pruss:dac14}) Given a combinational circuit $C$ with $k$-bit
+  input $A = (a_0, \dots, a_{k-1})$ and $k$-bit output $R = (r_0, \dots,
+  r_{k-1})$. Denote by $x_1, \dots, x_d$ all the bit-level
+  variables of   $C$. Let $J = \langle f_1, \dots, f_s \rangle \subset
+  \Fkk[x_1, \dots, x_d, R, A]$ denote all the polynomials corresponding to the
+  logic gates of the circuit. Let $J_0 = \langle x_1^2 - x_1, \dots,
+  x_d^2 - x_d, R^q - R, A^q - A \rangle$ be the vanishing ideal, so
+  that $J + J_0 = \langle f_1, \dots, f_s, ~~ x_1^2 - x_1, \dots,
+  x_d^2 - x_d, R^q - R, A^q - A \rangle$. Compute \Grobner basis $G =
+  GB(J + J_0)$ w.r.t. lex term order with $x_1 > x_2 > \dots > x_d > R
+  > A$. Then $G_d = G \cap \Fkk[R, A]$ eliminates the internal
+  variables $x_1, \dots, x_d$ of the circuit. $G_d$ also contains the
+  word-level polynomial $R = \F(A)$ which canonically represents the
+  function of the circuit.  
+\end{Lemma}
 
-For example, a pair of state variables $\{x,y\}$'s generic quadratic form is
-$$\F = a_0x^2+a_1x+a_2xy+a_3y+a_4y^2+a_5$$
-It covers all possible terms with degree at most 2. $a_0\sim a_5$ are usually real number parameters,
-a certain assignment about them can turn $\F$ into desired invariant. This parameterized constrain
-polynomial covers all combinations of state variables can also be called a \emph{template}.
-
-Finding a proper assignment is the major part of original paper and has potential to be even further explored
-beyond that paper, and/or to be applied on different research field.
-\subsection{Modeling initial state}
-\label{sec:CKR}
-For initial state, the constrains are explicit. A template is adopted and refined by Gr\"obner basis 
-generated by original constrains, by equaling the remainder to 0 we can get constrains on
-parameters from the template.
-
-An example is Fig.\ref{fig:loopinv}. The template is generic quadratic form of $\{s,i,j,j_0\}$, which is
-\begin{figure}[hbt]
-	\begin{center}
-	\includegraphics[scale=0.3]{../../../Pictures/Selection_006.png}
-	\end{center}
-	\caption{An example program of 2 Natural numbers' multiplication (Fig.1 from original paper)}
-	\label{fig:loopinv}
-\end{figure}
-\begin{align}
-\F =& a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
-&a_5i^2+a_6i+a_7ij+a_8ij_0+a_9j^2\nonumber\\
-&a_{10}j+a_{11}jj_0+a_{12}j_0^2+a_{13}j_0+a_{14}\nonumber
-\end{align}
-
-Constrains of initial state: $s=0\land j=j_0$ can be interpreted to polynomials:
-$\{s, j-j_0\}$. Check if it is Gr\"obner basis using \emph{Buchberger's algorithm}. Since
-their leading terms are relatively prime, we can write Gr\"obner basis $G=\{s, j-j_0\}$. Its ideal
-$J=\langle s,j-j_0\rangle$.
-
-Compute multi-division $\F \xrightarrow{G}_+ r$, the remainder is
-$r = a_5i^2+a_6i+(a_7+a_8)ij_0+(a_9+a_{11}+a_{12})j_0^2+(a_{10}+a_{13})j_0+a_{14}$.
-Let it equal to 0, each coefficient will generate a constrain, solution to the system is candidate
-assignment to generate invariant.
-\begin{equation}
-\left\{
-\begin{array}{l}
-a_5=a_6=a_{14}=0\\
-a_7+a_8=0\\
-a_9+a_{11}+a_{12}=0\\
-a_{10}+a_{13}=0
-\end{array}\right.
-\nonumber
-\end{equation}
-
-\subsection{Modeling state transitions}
-2 states get involved in a state transition. The technique requires to model the 2 states individually,
-do multi-division separately to get remainder $r_1$ and $r_2$. Suppose a constrain polynomial for
-this transition is $r_t$, we requires that when invariant for state 1 holds and transition $1\to 2$ stands,
-invariant for state 2 should also holds, which is
-$$(r_1 = 0)\land (r_t = 0) \implies (r_2 = 0)$$
-One reasonable conjecture is
-$$r_t = r_1 - \lambda r_2$$
-Theoretically $\lambda$ could be any polynomial. Consider the system to be easy to solve, we keep $\lambda$
-only taking value from real numbers.
-
-Still take Fig.\ref{fig:loopinv} as example (also refer to Example 10 in original paper). State 1 is initial state we 
-just characterized $\F = f(s,i,j,j_0)$, and the post state (state 2) have exactly the same form constrain:
-$\F' = f(s',i',j',j_0')$. Considering transition relation, substitute $s'$ with $s+i$, replace $j'$
-with $j-1$, and $i'$ for $i$, $j_0'$ for $j_0$, the template for state 2 is expression $f'$ in Example 10 from
-original paper.
-
-Do multi-division with Gr\"obner basis generated by transition relation, record its remainder $r_2$.
-From equality $r_2 = \lambda r_1$, we get constrains for parameters. Solve the system using covering-problem-like
-techniques (Part. \emph{Elimination by Splitting} in original paper). One branching result is:
-$a_0,a_2\sim a_6, a_9\sim a_{14}$ are all zero; $a_1=a_7=-a_8$. Then reduced remainder 
-$$r_2 = a_1s +(a_1-a_7)i+a_7ij+a_8ij_0 = a_1(s+ij-ij_0) = 0$$
-Desired invariant is 
-$$s = i(j_0- j)$$
-We knew the program calculates $i\times j_0$. Initially $s=0$, $j_0-j=0$, invariant holds;
-during each cycle, $s'=s+i, (j_0'-j') = (j_0-j)+1$, the invariant also holds. In conclusion, this is a
-loop invariant for program in Fig.\ref{fig:loopinv}.
-
-\section{Our approach on technology mapping}
-Our approach borrows inspiration of "templates" from loop invariant paper. The difference is,
-we are not using template polynomial to describe the system, instead we add some templates in the ideal
-of macro blocks we want to map to simulate all possibilities.
-\subsection{System abstraction}
-The polynomial we use to test ideal membership should include all information of a circuit partition,
-this requires us to abstract information from the system and write it into only 1 polynomial.
-Usually this polynomial has the form like:
-$$Z + f(i_1,i_2,\dots,i_n)$$
-Here $Z$ is the output. When there is only one output, $Z$ will be a bit-level variable. However in most
-cases we have multiple outputs, this asks us to write $Z$ as a word-level variable. $f$ is a Boolean
-function about all inputs, and $i_1,i_2,\dots,i_n$ are all bit-level inputs.
-
-\begin{figure}[hbt]
-	\begin{center}
-	\includegraphics[scale=0.2]{tobemapped.png}
-	\end{center}
-	\caption{An example gate-level netlist to be mapped}
-	\label{fig:tobemapped}
-\end{figure}
-
-Fig.\ref{fig:tobemapped} shows an example circuit partition with $a1,b1,a0,b0$ as inputs and
-$Z = \{z0,z1\}$ as outputs. If we use elements from Galois field $F_{2^2}$ to represent word $Z$,
-we have $Z = z_0 + \alpha\cdot z_1$.
-
-The abstraction also uses the property of Gr\"obner basis. If we arrange a term ordering as
-$$Other\ circuit\ variables > output\ word\ Z > all\ bit\ level\ inputs$$
-and include all gates information, word level variable definition and vanishing polynomials,
-the reduced Gr\"obner basis will have a polynomial (generator) in the form of $Z + f(a1,b1,a0,b0)$,
-this is an application of "elimination using GB", where we eliminate all variables except output
-word variable $Z$ and inputs $a1,a0,b1,b0$.
-
-In this example:
-
-\begin{itemize}
-\item Term ordering: $z0,z1,t,g1,p1,ci,t1,t2,Z,a0,b0,a1,b1$
-\item Gate description (from table \ref{table:booltogalois_op}): $z0+t+g1+t\cdot g1, t+t1+t2+t1\cdot t2, g1+a1\cdot b1,
-			t1+a1\cdot ci, t2+b1\cdot ci, ci+a0\cdot b0, z1+p1+ci, p1+a1+b1$
-\item Word definition: $Z+z0+z1\cdot \alpha$
-\item Vanishing polynomials $(J_0)$: $z0^2+z0, z1^2+z1, t^2+t, g1^2+g1, t1^2+t1, t2^2+t2, p1^2+p1, ci^2+ci,
-			a0^2+a0, b0^2+b0, a1^2+a1, b1^2+b1, Z^4+Z$(note Z is 2-bit word)
-\end{itemize}
-
-The result is one polynomial in Gr\"obner basis with leading term $Z$:
-$$Z+a0\cdot b0\cdot a1+a0\cdot b0\cdot b1+\alpha\cdot a0\cdot b0+a1\cdot b1+\alpha\cdot a1+\alpha\cdot b1$$
-
-\begin{table}
-\centering
-\begin{tabular}{|c|c|} \hline
-Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
-$\overline{a}$ & $1 + a$\\ \hline
-$a\ and\ b$ & $ab$\\ \hline
-$a\ or\ b$ & $a + b + ab$\\ \hline
-$a \oplus b$ & $a + b$\\
-\hline\end{tabular}
-\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
-\label{table:booltogalois_op}
-\end{table}
-
-\subsection{Templates on Boundary Information}
-\begin{figure}[hbt]
-	\begin{center}
-	\includegraphics[scale=0.2]{template.png}
-	\end{center}
-	\caption{Standard 2-bit adder with input/output mapping}
-	\label{fig:template}
-\end{figure}
-
-Fig.\ref{fig:template} shows a 2-bit standard cell. It has 3 outputs mapping to 2 output pins,
-and 4 inputs mapped to 3 input pins (the rest pin is mapped to fixed 0/1 signal). Dashed connection
-lines show one possible mapping, to find out this kind of feasible mapping, we need to simulate
-all possible mappings, where the concept "template" can be used.\vspace{5mm}
-
-{\bf Output template}: $z0+ct0_{z0}\cdot t0+ct1_{z0}\cdot t1+c0_{CO}\cdot CO+cn0$
-
-When $cn0 = 0$, any one of $ct0_{z0},ct1_{z0},c0_{CO}$ could be 1 means mapping to corresponding output
-pin. If $cn0 = 1$, mapping to negation of corresponding output pin.
-
-{\bf Input template}: $a0+pa0+c0_{A0}\cdot s0+c1_{A0}\cdot s1+c2_{A0}\cdot s2$
-
-When $pa0 = 0$ and any one of $c0_{A0},c1_{A0},c2_{A0}$ equal to 1 means mapping to corresponding input
-pin; if all $c0_{A0},c1_{A0},c2_{A0}$ equal to 0 then means mapping to fixed signal "0" (when $pa0 = 1$ then
-mapping to fixed signal "1").
-
-\subsection{Fast GB computing}
-We compose an ideal with polynomials from all gates information, input/output templates and word definition,
-then compute a GB of this ideal. However, computing GB using algorithms like Buchberger's algorithm is
-less efficient. For example, the "slimgb" function in Singular tool cannot even compute a GB for ideal
-from Fig.\ref{fig:template}. So finding a fast GB computing method is necessary.
-
-Analyzing Buchberger's algorithm, we find the reason that GB compute takes long time is the size of ideal
-will explode when there exists many pairs of polynomial can compute a Spoly. If we arrange the term ordering
-carefully to minimize the number of Spoly, then GB computation will be easy.
-
-We find such term ordering can satisfying our requirement:
-\begin{align}
-&Reverse\ topological\ order\ of\ circuit\ variables > \nonumber\\
-& output\ word\ variable > output\ templates\ variables>\nonumber\\
-& fixed\ 0/1\ templates\ variables > input\ templates\ var\nonumber\\
-& > input pins \nonumber
-\end{align}
-
-Under this term ordering, all polynomials' leading term are relatively prime (so Spoly equal to 0)
-except only 1 pair of polynomials. So the work load is just compute a Spoly for this pair,
-then reduce the Spoly with whole ideal. After adding the remainder to the ideal, it becomes
-a Gr\"obner basis.
-
-\subsection{Our Tool Flow}
-First, choose a set of independent wires as input pins;
-
-Second, push forward input signal for a certain number of gates (depth), choose a cut set
-which fully dependent on these inputs as output pins;
-
-Third, abstract a description polynomial using GB abstraction method;
-
-Fourth, reduce (do multi-division) the description polynomial with the GB we computed with macro block 
-and templates, get a remainder polynomial;
-
-Last but not least, abstract all coefficients from the remainder polynomial, set up a system of equations,'
-find a solution to this system, then the solution is a feasible mapping; no solution means cannot make a
-feasible mapping.
-
-\section{Problems to address}
-The final system of equations is not linear, so a 0/1 programming technique cannot be used. However,
-an exhaustive enumerating is still practicable:
-If we map a marco block with $j$ inputs and $n$ outputs into circuit partition with $i$ inputs
-and $m$ outputs, number of total possible assignments is
-$$2^{j-i}A_j^iA_n^m$$
-when considering mapping to negative signals, it increases to 
-$$2^{m+j}A_j^iA_n^m$$
-
-For example, when mapping 3 inputs 2 outputs partition to a 2-bit adder:
-
-Total possible assignments: 9216
-
-Full search space: $2^{24} = 16.8M$
-
-So it only takes $0.05\%$ time to scan all possibilities using our approach.
-
-When the macro block is larger as 4-bit adder, number of total possible assignment is $2^{27}$,
-still a number we can accept.
-%%%%%%%%%%%%%%%%%%%% The bibliography %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{thebibliography}{1}
-%\bibitem{ref1}
-%\end{thebibliography}
-\end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%  End of IEEEsample.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%
+The authors referred to the elimination (lex) order with
+$\{\textit{internal variables} \ x_1 > \dots > x_d\} > \{\textit{word-level
+  output} \ R\} > \{\textit{word-level input} \ A\}$ as the {\it abstraction
+  term order (ATO)}. We will now show how to repeatedly apply \Grobner
+basis computations with ATO to verify sequential GF circuits.

### 2015-05-01 techwriting/ASPDAC/prev.tex
all write-up from virtue machine

\section{Review of Previous Work}
\label{sec:prev}

Verification of a combinational GF arithmetic circuit $C$ against a
polynomial specification $\F$ has been previously addressed \cite{ibm:blueveri}
\cite{lv:tcad2013} \cite{pruss:dac14}. Verification problems in
\cite{ibm:blueveri} \cite{lv:tcad2013} are formulated using
Nullstellensatz and decided using the \Grobner basis algorithm.
%% Moreover, in \cite{lv:tcad2013} the authors show that by analyzing the
%% given circuit $C$, a term order ($>_1$) can be derived that renders
%% the set of polynomials itself a \Grobner basis. Consequently, a
%% \Grobner basis computation is not required, and the verification test
%% can be performed simply by dividing the specification polynomial $\F$
%% by polynomials of $C$ (when represented using $>_1$).

The paper 
\cite{pruss:dac14} performs verification by deriving a canonical
word-level polynomial representation $\F$ from the circuit $C$. Their
approach views any arbitrary Boolean function (circuit) $f: \B^k
\rightarrow \B^k$ as a polynomial function $f: \Fkk \rightarrow \Fkk$,
and derives a canonical polynomial representation $\F$ over
$\Fkk$. They show that this can be achieved by computing a reduced 
\Grobner basis w.r.t. an {\it abstraction term order} derived from the
circuit. Subsequently, they propose a \underline{r}efinement of this
\underline{a}bstraction \underline{t}erm \underline{o}rder (called
RATO), that enables to compute the \Grobner basis of a smaller subset
of polynomials. The authors show that their approach can prove
correctness of up to 571-bit combinational GF multipliers.

Since we are also interested in deriving a polynomial representation
of the computation performed by the sequential circuit $S$, we draw
inspirations from \cite{pruss:dac14} -- particularly the use of RATO
-- for sequential verification. However, the approach of
\cite{pruss:dac14} suffers from a few limitations: i) When the GF
polynomial implemented by the circuit is dense (say,  due to the
presence of a bug in the design), their approach is computationally
infeasible; ii) The use of RATO still requires a \Grobner basis
computation (even though on a subset of polynomials) to derive the
polynomial $\F$, which can lead to a memory explosion. Experiments in
\cite{pruss:dac14} are only successful for hierarchically designed and
bug-free GF circuits. {\it While we do employ RATO as the term order
  for sequential verification, we further present an efficient
  symbolic computation approach that does not suffer from these
  limitations}.

The problem addressed in this paper is not suitable to be
solved by conventional bit-level sequential equivalence
\cite{coudert:iccad90} or (bounded) model checking frameworks
based on interpolation \cite{mcmillan:interpolation_cav03} or property
directed reachability \cite{IC3}. This is mostly due to the word-level
and GF polynomial nature of the specification (property) $\F$, which 
is also only valid in one state of the machine. The use of algebraic
geometry has been proposed for model checking 
\cite{Avrunin:CAV} \cite{Vardi:gb_bool_ring} \cite{polybori:2009};
however, these approaches are a straight-forward application of {\it
  bit-level} Boolean \Grobner basis engines in lieu of binary decision diagrams (BDDs) or 
  satisfiability (SAT)
solvers.  
%computations to the model checking algorithm.

### 2015-05-01 techwriting/ASPDAC/theory.tex
all write-up from virtue machine

--- 2015-05-01 techwriting/ASPDAC/oldtheory.tex
@@
 \subsection{Polynomial Representation}
-A typical sequential circuit is composed as figure \ref{fig:sequential}. The combinational logic has primary input $x$, state
-inputs $\{s_0, s_1, \dots, s_{k-1}\}$, primary output $z$ and state outputs $t_0, t_1, \dots, t_{k-1}$. State outputs
+A typical sequential circuit model is showed as blocks in figure \ref{fig:sequential}. The combinational logic block 
+has primary inputs $X = \{x_0, x_1, \dots, x_{m-1}\}$, state
+inputs $S = \{s_0, s_1, \dots, s_{k-1}\}$, primary output $Z$ and state outputs $T = \{t_0, t_1, \dots, t_{k-1}\}$. State outputs
 are latched into registers, which will be fed back to state inputs in next clock cycle.
 
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[width=3.5in]{./sequential_fig.eps}
+%% \begin{figure}[tb]
+%% \begin{center}
+%% \includegraphics[width=3.5in]{./sequential_fig.eps}
+%% \caption{A typical sequential circuit block view}
+%% \end{center}
+%% \label{fig:sequential}
+%% \end{figure}
 
-\end{center}
-\caption{A typical sequential circuit block view}
-\label{fig:sequential}
-\end{figure}
-
-A transition function $\Delta$ is a Boolean function about state outputs and all inputs, i.e.
+A transition function $\Delta$ is a Boolean function giving state outputs from state inputs and primary inputs, i.e.
 \begin{displaymath}
@@
 $\mathbb{B}^k \to \mathbb{F}_{2^k}$, which means it is possible to find a set of elements in $\mathbb{F}_{2^k}$ representing
-all Boolean values every bit of a bit vector can take. In Galois field, an example is in table \ref{table:booltogalois}.
-\begin{table}[tb]
+all Boolean values that a $k$-bit bit-vector can take. In Galois
+field, a as \ref{table:booltogalois} shows.
+
+\begin{table}[htb]
 \centering
@@
 \end{table}
-A word-level representation on state inputs/outputs can be attained through above correspondence. If the set of 
-state inputs $\{s_0, s_1, \dots, s_{k-1}\}$ is denoted by $S$ taking values from $\mathbb{F}_{2^k}$, and
-set of outputs $\{t_0, t_1, \dots, t_{k-1}\}$ is denoted by $T$, a word-level transition function can be written as
+A word-level representation of state inputs/outputs can be obtained through above correspondence. If the set of 
+state inputs $\{s_0, s_1, \dots, s_{k-1}\}$ is denoted by $S$ taking values from $\mathbb{F}_{2^k}$, $X$ stands
+for a word taking values from $\mathbb F_{2^m}$ and
+the set of outputs $\{t_0, t_1, \dots, t_{k-1}\}$ is denoted by $T$, a word-level transition function can be written as
 \begin{displaymath}
-T - \Delta(x, S) = 0
+T - \Delta(X, S) = 0
 \end{displaymath}
-If $T - \Delta(x, S)$ is taken as a polynomial $f$, then $f \in \mathbb{F}_{2^k}[x]\ mod\ P(\alpha)$.
+Let polynomial $f = T - \Delta(x, S)$, then $f \in \mathbb{F}_{2^k}[x]\ mod\ P(\alpha)$.
 \subsection{Gr\"obner Basis Theory}
@@
 
-Calculate Reduced Gr\"obner Basis (RGB) from ideal $I$ with \textbf{elimination term order}, there exists a polynomial contains
+Calculate Reduced Gr\"obner Basis (redGB) from ideal $I$ with \textbf{elimination term order}, there exists a polynomial contains
 only $T$, if the values of states input $S$ and primary input $x$ are given in the form of polynomials and included by $I$,
@@
 
-Otherwise, if RGB calculation is manipulated under \textbf{abstraction term order}, there exists a polynomial in
+Otherwise, if redGB calculation is manipulated under \textbf{abstraction term order}, there exists a polynomial in
 the form of $T - \mathcal{F}(x, S)$ representing transition function. In next clock cycle, assign new state input $S'$ with $T$, the next
@@
 \includegraphics[scale=0.3]{./stg_fig.eps}
-
+\caption{State transition graph of sample FSM}
 \end{center}
-\caption{State transition graph of sample FSM}
 \label{fig:stg}
@@
 \includegraphics[width=3.5in]{./fsm_fig.eps}
-
+\caption{Gate-level circuit of combinational logic block of sample FSM}
 \end{center}
-\caption{Gate-level circuit of combinational logic block of sample FSM}
 \label{fig:fsm}
@@
 
-The first example is a mealy finite state machine (FSM). Gate-level circuits of combinational logic
-is figure \ref{fig:fsm}, $x$ is primary input,
+The first example is a mealy Finite State Machine (FSM). Gate-level circuits of combinational logic
+is figure \ref{fig:fsm}, $x$ is primary input (one bit),
 $\{s_0, s_1\}$ are state inputs, $Z$ is primary output and $\{t_0, t_1\}$ are state outputs.
@@
 $\mathbb{F}_{2^2}$, bit-level variables $x, s_0, s_1, Z, t_0, t_1$ can take values from $\{0, 1\}$, while
-word-level variables $S, T$ may take values from $\{0, 1, \alpha, 1 + \alpha\}$. State transition graph (STG)
+word-level variables $S, T$ may take values from $\{0, 1, \alpha, 1 + \alpha\}$. State Transition Graph (STG)
 showed in figure \ref{fig:stg} uses 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2, S_3\}$, which
-could be converted to elements in $\mathbb{F}_{2^2}$ similarly as table \ref{table:booltogalois} shows.
+could be mapped to elements in $\mathbb{F}_{2^2}$ similarly as table \ref{table:booltogalois} shows.
 
 One important technique to check sequential equivalence is \textbf{state space traversal}. In this example,
-Breath-First-Search (BFS) is employed for traversal, which can be described with following algorithm\cite{KallaPartialScan}:
+Breath-First-Search (BFS) is employed for traversal procedure, which can be described by following algorithm\cite{KallaPartialScan}:
 
@@
 \includegraphics[width=\textwidth]{mySMPO.eps}
-
-\end{center}
 \caption{5-bit SMPO.
@@
 Index $w = 2i\pmod 5$}
+\end{center}
 \label{fig:SMPO}
@@
 \begin{figure}[hbt]
-\begin{center}
-\includegraphics[width=3.5in]{./RH.eps}
-
-\end{center}
+\centering
+\begin{minipage}{12cm}
+\includegraphics[width=4.5in]{./RH.eps}
 \caption{5-bit RH-multiplier (AESMPO)}
+\end{minipage}
 \label{fig:RHmulti}
@@
 
-Apply this approach to calculate image in BFS traversal, but modify algorithm
-\ref{alg:BFS} to make it adapt 5 steps reachable states enumeration.
-The result is $R + AB$, which validates the function of this circuit.
+If we apply this approach to calculate "image function" in BFS traversal, but modify algorithm
+\ref{alg:BFS} to make it adapt 5 steps reachable states enumeration,
+the result will be $R + AB$ which validates the function of this circuit.
 
@@
 \begin{Example}
-{\bf Objective}:\ Abstract polynomial $a_i + \mathcal{G}_i(A)$ from $f_0: a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A$.
+{\bf Objective}:\ Abstract polynomial $a_i + \mathcal{G}_i(A)$ from $f_0 = a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A
+ = g_0 + A$.
 
-First, compute $f_0^2: a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2$. Apparently variable $a_0$ can be
-eliminated by operation 
+First, compute $f_0^2= a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2 = g_0^2 + A^2$;
+then $f_0^4, f_0^8, f_0^{16}$. By repeating squaring we get a system of equations:
+\begin{displaymath}
+  \begin{cases}
+  f_0  &=\ \  0\\
+  f_0^2 & =\ \  0\\
+  f_0^4  &=\ \  0\\
+  f_0^8  &=\ \  0\\
+  f_0^{16} & =\ \ 0
+  \end{cases}
+\Longleftrightarrow\ \
+\begin{cases}
+  g_0 & =\ \  A\\
+  g_0^2 & =\ \  A^2\\
+  g_0^4 & =\ \  A^4\\
+  g_0^8 & =\ \  A^8\\
+  g_0^{16} & =\ \  A^{16}
+  \end{cases}
+\end{displaymath}
+Following is the coefficients matrix form of this system of equations:
+\begin{displaymath}
+\begin{pmatrix}
+\alpha^5 & \alpha^{10} & \alpha^{20} & \alpha^9 & \alpha^{18} \\
+\alpha^{10} & \alpha^{20} & \alpha^9 & \alpha^{18} & \alpha^5 \\
+\alpha^{20} & \alpha^9 & \alpha^{18} & \alpha^5 & \alpha^{10} \\
+\alpha^9 & \alpha^{18} & \alpha^5 & \alpha^{10} & \alpha^{20} \\
+\alpha^{18} & \alpha^5 & \alpha^{10} & \alpha^{20} & \alpha^9 
+\end{pmatrix}
+\begin{pmatrix}
+a_0\\
+a_1\\
+a_2\\
+a_3\\
+a_4
+\end{pmatrix}
+=
+\begin{pmatrix}
+A\\
+A^2\\
+A^4\\
+A^8\\
+A^{16}
+\end{pmatrix}
+\end{displaymath}
+
+Use Gaussian elimination on coefficients matrix, for example  
 \begin{align}
-f_1 =& f_0\times \alpha^5 + f_0^2: \nonumber\\
+Row\ 2 =& Row\ 1\times \alpha^5 + Row\ 2: \nonumber\\
 &a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
-+&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
+=&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
 \end{align}
-Recursively eliminate $a_1$ from $f_1$, $a_2$ from $f_2$, etc. The final polynomial $f_4$ has the form 
-$a_4 + \mathcal{G}_4(A) = a_4+(\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4+(\alpha^2+1)A^2+(\alpha+1)A$. Recursively substitute
-$\mathcal{G}_4(A)$ back to $f_3$, etc. The result is a set of polynomials
+Recursively eliminate $a_1$ from third row, $a_2$ from fourth row, etc. The final solution to this system of
+equations is
+%\begin{displaymath}
+%\{g_i\ | \ g_i: a_i + \mathcal{G}_i(A)\}
+%\end{displaymath}
+%\end{Example}
 \begin{displaymath}
-\{g_i\ | \ g_i: a_i + \mathcal{G}_i(A)\}
+\left\{
+  \begin{array}{lcl}
+  a_0 & = & (\alpha+1)A^{16}+(\alpha^4+\alpha^3+\alpha)A^8+(\alpha^3+\alpha^2)A^4\\&&+(\alpha^4+1)A^2+(\alpha^2+1)A\\
+  a_1 & = & (\alpha^2+1)A^{16}+(\alpha+1)A^8+(\alpha^4+\alpha^3+\alpha)A^4\\&&+(\alpha^3+\alpha^2)A^2+(\alpha^4+1)A\\
+  a_2 & = & (\alpha^4+1)A^{16}+(\alpha^2+1)A^8+(\alpha+1)A^4\\&&+(\alpha^4+\alpha^3+\alpha)A^2+(\alpha^3+\alpha^2)A\\
+  a_3 & = & (\alpha^3+\alpha^2)A^{16}+(\alpha^4+1)A^8+(\alpha^2+1)A^4\\&&+(\alpha+1)A^2+(\alpha^4+\alpha^3+\alpha)A\\
+  a_4 & = & (\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4\\&&+(\alpha^2+1)A^2+(\alpha+1)A
+  \end{array} \right.
 \end{displaymath}

### 2015-05-01 techwriting/IBM-Watson/f4.tex
all write-up from virtue machine

\section{Improving Polynomial Division using $F4$-style Reduction}
\label{sec:f4}

The most intensive computational step in our approach is that of
polynomial division $f \stackrel{F,F_0^{PI}}{\longrightarrow}_+
r$. When the circuit $C$ is very large, the polynomial set
$\{F, F_0^{PI}\}$ also becomes extremely large. This division
procedure then becomes the bottleneck in verifying the equivalence.  
In principle, this reduction can be performed using contemporary
computer-algebra systems --- {\it e.g.},  the {\sc Singular}
\cite{DGPS} tool, which is widely used within the verification
community \cite{wienand:cav08} \cite{wedler:date11}
\cite{lv:date2012}. In our work, we have also performed experiments 
with {\sc Singular}. However, as in any ``general-purpose''
computer algebra tool, the data-structures are not specifically
optimized for circuit verification problems. Moreover, {\sc
  Singular} also limits the number of variables ($d$) that it can
accommodate in the system to $d < 32767$; this limits its
application to large circuits. Therefore, to further improve our
approach, we exploit the relatively recent concept of $F4$-style
polynomial reduction \cite{f4} --- which implements polynomial division
using row-reductions on a matrix --- to develop a custom
verification tool to perform this Gr\"obner basis reduction
efficiently.

$Faug\grave{e}re$'s $F4$ approach \cite{f4} presents a new algorithm to
compute a Gr\"obner basis. It uses the same mathematical principles as
Buchberger's algorithm. However, instead of computing and reducing one 
$S$-polynomial at a time, it computes many $S$-polynomials in one
step and reduces them simultaneously using sparse linear algebra on a
matrix (triangulation). In our formulation, since we already have a
Gr\"obner basis, no S-polynomials are computed. We only need to
perform the subsequent Gr\"obner basis reduction, for which the $F4$
technique can be very efficient.  We now show how our reduction problem
$f\stackrel{F,F_0^{PI}}{\longrightarrow}_+r$ can be represented and
solved on a matrix. 
%, based on an $F4$-style approach. 
First, let us consider the following example that demonstrates the
main concepts behind the reduction approach of $F4$.

\begin{Example}
{\it
Consider the {\it lex} term order with $x>y>z$ on the ring
${\mathbb{Q}}[x, y, z]$.  Given $F = \{f_1 = 2x^2 + y, f_2 = 3xy^2
-xy, f_3 = 4y^3 -1\}$, consider one step of Buchberger's algorithm:
$S(f_1, f_2) \xrightarrow{f_1, f_2, f_3}_+r$. We have, $Spoly(f_1,
f_2) = \frac{1}{3}x^2y + \frac{1}{2}y^3 = f_4$. The reduction
$Spoly(f_1, f_2)\stackrel{f_1, f_2, f_3}{\longrightarrow}_+
(-\frac{1}{6}y^2 + \frac{1}{8})$ is done as follows: 
Since $lt(f_1) ~|~ lt(f_4), ~f_4 \xrightarrow{f_1} h$ is computed as:
\[
h = f_4 - {{lt(f_4)} \over {lt(f_1)}} f_1 = f_4 - \frac{1}{6}y f_1 =
\frac{1}{2}y^3 - \frac{1}{6}y^2; 
\]
Now $lt(f_2)$ does not divide any term in $h$, but $lt(f_3) ~|~ lt(h)$,
so $f \xrightarrow{f_3} r$:
\[
r = h - {{lt(h)} \over {lt(f_3)}} f_3 =  \frac{1}{2}y^3 - \frac{1}{6}y^2 -
\frac{1}{8}f_3 = -\frac{1}{6}y^2 + \frac{1}{8} 
\]

This reduction procedure can also be simulated on a matrix using
Gaussian elimination. 
%Matrix triangulation can also be used for this purpose. 
The reduction above requires the computation of $\frac{1}{6}y f_1$ and 
$\frac{1}{8}f_3$. Ignoring the coefficients $\frac{1}{6},
\frac{1}{8}$, 
we can generate all the monomials  required in the reduction process: 
i.e.  monomials of $f_4, yf_1, f_3$, and setup the problem of
cancellation of terms as Gaussian elimination on a matrix. Monomials
of $f_4, yf_1, f_3$ are, respectively, $\{x^2y, y^3\}, \{x^2y, y^2\}, \{y^3, 1\}$.
Let the rows of a matrix $M$ correspond to polynomials $\left[ f_4,
  yf_1, f_3   \right]$, and columns correspond to all the monomials
(in {\it lex} order) $\left[x^2y, y^3, y^2,  1 \right]$. Then the
matrix $M$ shows the representation of these polynomials where
the entry $M(i, j)$ is the coefficient of monomial of column $j$
present in the polynomial of row $i$.

\[
%% M = \begin{pmatrix*}[l]
%%  & x^2y & y^3 & y^2 & y\\
%% f_4 & \frac{1}{3} & \frac{1}{2} & 0 & 0 \\
%% yf_1 & 2 &0 & 1 & 0 \\
%% yf_3 & 0 & 4 & 0 & -1
%% \end{pmatrix*}
M = \bordermatrix{
~ & x^2y & y^3 & y^2 & 1\cr
f_4 & \frac{1}{3} & \frac{1}{2} & 0 & 0 \cr
yf_1 & 2 &0 & 1 & 0 \cr
f_3 & 0 & 4 & 0 & -1 \cr
}
\]

Now, reducing $M$ to a row echelon form using Gaussian elimination gives:
\[
M = \bordermatrix{
~ & x^2y & y^3 & y^2 & 1\cr
f_4 & \frac{1}{3} & \frac{1}{2} & 0 & 0 \cr 
h = f_4 - \frac{1}{6}yf_1 & 0 &\frac{1}{3}  & -\frac{1}{6} & 0 \cr 
r = h - \frac{1}{8}f_3 & 0 & 0 & -\frac{1}{6} & \frac{1}{8} \cr 
}
\]

The last row $(0, 0, -\frac{1}{6}, \frac{1}{8})$ accounts for
polynomial $-\frac{1}{6}y^2 + \frac{1}{8}y$ which is equal to the
reduction result $r$ obtained before. 
}
\end{Example}

This approach generates all the monomial terms that are required
in the division process, 
%i.e. $\frac{lm(f_i)}{lm(f_j)} f_j$, 
and the coefficients required for cancellation of terms are accounted
for by elementary row reductions in the subsequent Gaussian elimination. 
Based on the above concepts, a matrix can be constructed for our
problem: $f\stackrel{F,F_0^{PI}}{\longrightarrow}_+r$.

\begin{Definition}
Let $L = \left [f_1, \dots, f_m \right]$ be a list of $m$ polynomials. Let
$M_L$ be an ordered list of monomials of elements of $L$ and let $n$
be the number of elements in $M_L$. Define $M$ as the $m \times n$
matrix which associates the polynomials of $L$ to rows and monomials
of $M_L$ to columns. Entry in row $i$, column $j$ is the coefficient of the
$j^{th}$ element of $M_L$ in $f_i$. 
\end{Definition}

\begin{algorithm}[hbt]
\SetAlgoNoLine

 \KwIn{$f, F=\{f_1,\dots,f_s$\}, term order $>$ } %with $f_{1}>f_{2}>\dots>f_{s}$. }
 \KwOut{A matrix $M$ representing $f \xrightarrow{f_1,\dots,f_s}_+r$}
  %%%%%%%%%%%%%%%%%%%%
        \CommentSty{/*$L=$ set of polynomials, rows of $M$*/\;}
        L:=\{f\} \;{}
%        \CommentSty{/*The index of polynomials in $L$*/\;}
        i:=1\;
%       \CommentSty{/*Initial remainder $r$ is set as $f$*/\;}
%       r:=f\;
        \CommentSty{/*$M_{L} = $ the set of monomials, columns of $M$ */\;}
        $M_{L}$:=\{ monomials of f\} \;{}
%       \CommentSty{/*Let $Done$ be the set of monomials that have been handled*/\;}
%       $Done:=\{\}$\;
%       \CommentSty{/*The first $i-1$ monomials of $M_{L}$ have been handled*/\;}
        mon:= the $i^{th}$ monomial of $M_{L}$\;
        \While { mon $\notin PrimaryInputs$ }
                {
                        Identify $f_{k} \in F$ satisfying: $lm(f_{k})$ can divide $mon$ \;
                        \CommentSty{/*add polynomial $f_k$ to L as a
                          new row in $M$ */\;}
                        $L:=L \cup \frac{mon}{lm(f_{k})}\cdot f_{k}$ \;
                        \CommentSty{/*Add monomials to $M_{L}$ as new
                          columns in $M$  */\;}
                        $M_{L}$:=$M_{L} \cup \{ \text{monomials of }
                        \frac{mon}{lm(f_{k})}\cdot f_{k}\}$ \;{} 
                        %\CommentSty{/*Cancel adjacent monomials if they are the same*/\;}
                        %CancelAdjMons($M_{L}$)\;
                        $i:=i+1$\;{}
                        mon:= the $i^{th}$ monomial of $M_{L}$\;
                }
         Gaussian Elimination on $M$;\\
         \Return $r =$ last row of $M$;
\caption{Generating the Matrix for Polynomial Reduction}\label{alg:matrix}
\end{algorithm}

\begin{figure*}[hbt]
\centering
\subfloat[Matrix $M$ generated by Algorithm \ref{alg:matrix}]
[Matrix $M$ generated by Algorithm \ref{alg:matrix}]
{ \label{subfig:matrix}
\begin{math}
\bordermatrix{
~ & Z  & AB & Ba_0 & Ba_1 & z_0 & z_1 & r_0 & a_0b_0 & a_0b_1 & a_1b_0& a_1b_1\cr
f & 1 &1 &0 &0 &0 &0 &0 &0 &0 &0 &0\cr
f_3 & 1 &0 &0 & 0 &1 &\alpha &0 &0 &0 &0 &0 \cr
Bf_1 & 0 &1 &1 &\alpha &0 &0 &0 &0 &0 &0 &0\cr
a_0f_2 &0  &0 &1 &0 &0 & 0& 0&1 &\alpha &0 &0\cr
a_1f_2 &0  &0 &0 &1 &0 &0 &0 &0 &0 &1 &\alpha\cr
f_5  & 0 &0 &0 &0 &1 &0 &0 &1 &0 &0 &1\cr
f_6  & 0 &0 &0 &0 &0 &1 &1 &0 &0 &0 &1\cr
f_4 & 0 &0 &0 &0 &0 &0 &1 &0 &1 &1 &0\cr
}
\end{math}
}

\ \\
\subfloat[$M$ reduced to row echelon form via Gaussian Elimination]
[$M$ reduced to row echelon form via Gaussian Elimination]
{
\begin{math}
M = \bordermatrix{
~ & Z  & AB & Ba_0 & Ba_1 & z_0 & z_1 & r_0 & a_0b_0 & a_0b_1 & a_1b_0& a_1b_1\cr
row_1 = f & 1 &1 &0 &0 &0 &0 &0 &0 &0 &0 &0\cr
row_2 = f_3 - row_1 & 0 &1 &0 & 0 &1 &\alpha &0 &0 &0 &0 &0 \cr
row_3 = Bf_1 - row_2& 0 &0 &1 &\alpha &1 &\alpha &0 &0 &0 &0 &0\cr
row_4 = a_0f_2 - row_3&0  &0 &0 &\alpha &1 & \alpha& 0&1 &\alpha &0 &0\cr
row_5 = \alpha a_1f_2 - row_4&0  &0 &0 &0 &1 &\alpha &0 &1 &\alpha &\alpha &\alpha^2\cr
row_6 = f_5 - row_5 & 0 &0 &0 &0 &0 &\alpha &0 &0 &\alpha &\alpha &\alpha^2+1\cr
row_7 = \alpha f_6 -row_6 & 0 &0 &0 &0 &0 &0 &\alpha &0 &\alpha &\alpha &\alpha^2+\alpha+1\cr
r = \alpha f_4 - row_7& 0 &0 &0 &0 &0 &0 &0 &0 &0 &0 &\alpha^2+\alpha+1\cr
}
\end{math}
\label{subfig:triangular}
}
\label{fig:matrix}
\caption{$F_4$-style Polynomial reduction on a matrix for Example \ref{ex:f4}.}
\end{figure*}

Algorithm \ref{alg:matrix} describes our procedure to generate the
matrix $M$ of polynomials corresponding to our verification instance. 
The main idea  is to setup the rows  and columns of the matrix in a
way that polynomial division can be subsequently performed by
applying Gaussian elimination on $M$. 
%subtracting row $i$ from row $i-1$. 
In the algorithm, the set of polynomials $F = \{f_1, \dots, f_s\}$ 
correspond to the circuit constraints. The term ordering derived from
the topological analysis of the circuit is imposed to represent the
polynomials. The specification polynomial $f$ is to be reduced
w.r.t. $F= \{f_1, \dots, f_s\}$. Initially, $L = \{f\}$ is inserted as
the first row of the matrix and $M_L$ constitutes the (ordered) list
of monomials of $f$. Then, in every iteration $i$, a polynomial $f_k \in
F$ is identified such that $lm(f_k)$ divides the $i^{th}$ monomial
($mon$) of $M_L$; this is to enable cancellation of the corresponding
monomial term. The computation $L:=L \cup \frac{mon}{lm(f_{k})}\cdot
f_{k}$ in the while-loop, generates the polynomials required for
reduction\footnote{Recall that the division $\frac{f_i}{f_k} = f_i -
  \frac{lt(f_i)}{lt(f_{k})}\cdot f_{k} = f_i - \frac{lc(f_i)}{lc(f_k)} \cdot
  \frac{lm(f_i)}{lm(f_k)}\cdot f_k$. In the algorithm, the computation
  $\frac{mon}{lm(f_{k})}\cdot f_{k}$ corresponds to
  $\frac{lm(f_i)}{lm(f_k)}\cdot f_k$ used in the division.}.  
The list $M_L$ is updated to include monomials of
$\frac{mon}{lm(f_{k})}\cdot f_{k}$. Finally, the iteration in the
loop terminates when monomial $mon$ consists solely of primary input
variables. This is because primary inputs are never a leading term of
any polynomial; therefore, no polynomial $f_k \in F$ exists which can
divide $mon$. Moreover, due to our term order, once $mon$ consists of
only primary inputs, all remaining monomials will also contain 
only primary input variables. Clearly, no more polynomials $f_k$ need
to be generated in $L$, and the loop terminates.

Using the set $L$ as rows and $M_L$ as columns, a matrix $M$ is
constructed and Gaussian elimination is applied to reduce it to
row-echelon form. The last row in the reduced matrix 
corresponds to the reduction result $r$. 
%Equation \ref{eqn:mat-red}. 
Let us describe the approach using an example.

\begin{Example}
\label{ex:f4}
{\it 
Consider the reduction related to verification of the ${\mathbb{F}}_{2^2}$
multiplier circuit of Fig. \ref{fig:2bitmul}. Given specification $f:
Z + A B$, and the circuit polynomials $f_1:A + a_0 + a_1 \alpha,
~f_2: B + b_0 + b_1 \alpha,
~f_3: Z + z_0 + z_1 \alpha,
~f_4: r_0 + a_0b_1 + a_1b_0,
~f_5: z_0 + a_0b_0 + a_1b_1,
~f_6: z_1 + r_0 +a_1b_1.$
Here  $P(x) = x^2 + x + 1$, and  $P(\alpha) = 0$. We have to compute
$f \xrightarrow{f_1, \dots, f_6}_+r$.  Note that, for simplicity, 
variables $c_0, c_1, c_2, c_3$ from Fig. \ref{fig:2bitmul} have been
substituted by functions on primary inputs. The imposed ordering
corresponds to the one obtained due to Proposition
\ref{prop:top-order}: i.e. {\it lex} with $Z > A > B > z_0 > z_1 > r_0
> a_0 > a_1 > b_0 > b_1$. The algorithm constructs the matrix as
follows:

\begin{enumerate}

\item Initialization: $L = \{f\} = \{Z + AB\}. ~~M_L = \{Z, AB\}, i =
  1, mon = Z$ ($i^{th}$ monomial of $M_L$).
\item Iteration 1: Identify a polynomial $f_k \in F$ s.t. $lm(f_k) ~|~
  mon$. Clearly, $f_k = f_3 = Z + z_0 + z_1 \alpha$. Then, $L = L \cup
  \frac{mon}{lt(f_k)} \cdot f_k = L \cup f_3$. Therefore, $L = \{f,
  f_3\}$ and $M_L = \{Z, AB, z_0, z_1\}$, $ i = 2$ and $mon = AB$.
\item Iteration 2: $f_k = f_1 = A + a_0 + a_1\alpha$ because $lm(f_1) ~|~
  mon$. Therefore, $L = L \cup \frac{AB}{A} \cdot f_1 = L \cup Bf_1 =
  \{f, f_3, Bf_1\}$ and $M_L = \{Z, AB, Ba_0, Ba_1, z_0 , z_1\}, i =
  3, mon = Ba_0$. 
\item Iteration 3: $~f_k = f_2 = B + b_0 + b_1\alpha$ as $lt(f_2)
  ~|~ mon$. Therefore, $L = L \cup \frac{mon}{lt(f_k)} \cdot f_k = L
  \cup \frac{Ba_0}{B} \cdot f_2 = L \cup a_0f_2$. So, $L = \{f, f_3,
  Bf_1, a_0f_2\}$ and the monomial set $M_L = \{Z, AB, Ba_0, Ba_1,
  z_0, z_1, a_0b_0, a_0b_1\}, ~i = 4, mon = Ba_1$.
\item Continuing in this fashion $\dots$

%% \begin{center}
%% $\vdots$
%% \end{center}

\item Iteration 7: $L = \{f, f_3, Bf_1, a_0f_2, a_1f_2, f_5, f_6,
  f_4\}, ~M_L = \{Z, AB, Ba_0, Ba_1, z_0, z_1, r_0, a_0b_0, a_0b_1,
  a_1b_0, a_1b_1\}, i = 8, mon = a_0b_0$.

\item Iteration 8: Since $mon = a_0b_0$ contains only the
  primary inputs, no polynomial in $F$ has a leading term that can
  cancel $mon$, so the loop terminates. The matrix $M$ can be
  constructed using $L$ as rows and $M_L$ as columns.

\end{enumerate}

Fig. \ref{subfig:matrix} shows the matrix $M$, and its subsequent
Gaussian elimination is shown in Fig. \ref{subfig:triangular}. The
last row of the reduced matrix corresponds to the reduction
$f\stackrel{f_1, \dots,   f_s}{\longrightarrow}_+r$, where $r =
(\alpha^2 + \alpha + 1)\cdot (a_1b_1)$. Note that since $\alpha$ is the
root of the irreducible polynomial $P(x) = x^2 + x + 1$, we have that
$\alpha^2 + \alpha + 1 = 0$ in ${\mathbb{F}}_{2^2}$, and hence $r =
0$. In conclusion, $f\stackrel{f_1, \dots,
  f_s}{\longrightarrow}_+0$, and the circuit correctly implements
$f$. 
}
\end{Example}

{\bf Other implementation issues:} Recall that in our problem, we also
have to account for bit-level vanishing polynomials of primary input 
variables: $F_0^{PI} = \{x_i^2 - x_i: x_i \in X_{PI}\}$. In our
algorithm, if we encounter a bit-level variable $x_i$ with degree $n
\geq 2$, we replace $x_i^n = x_i$. Consequently, no specific entries 
for vanishing polynomials in the matrix need to be created. Moreover,
it is also possible to encounter coefficients $\alpha^n$, where $n
\geq k$ in $\Fkk$. We precompute $\alpha^k, \dots, \alpha^{2k-2}$
modulo the primitive polynomial $P(x)$, store them in a table, and use
these reduced values as coefficients in the matrix for reduction.

The above approach is implemented as a standalone, custom verification
tool, which inputs a Boolean gate-level circuit netlist $C$ and a
specification polynomial $f$. The tool performs a reverse topological
traversal of the circuit and derives the term order to represent
polynomials. Then, Algorithm \ref{alg:matrix} is invoked to construct
the matrix $M$; Gaussian elimination is finally performed on $M$ to
obtain the verification result via reduction:
$f\stackrel{F,F_0^{PI}}{\longrightarrow}_+r$. Our tool is written in
C++.

### 2015-05-01 techwriting/UNSAT-CORE/complex.tex
all write-up from virtue machine

\section{Some Questions on Gr\"obner Basis Complexity}

Another problem that we need to find an answer to is one of 
computational complexity of our approach of \cite{lv:tcad2013} (and
its derivatives \cite{pruss:dac14}) that we have been using to verify
Galois field circuits. Here are some interesting results that are
known:

First of all, computing a Gr\"obner Basis $G$ of an ideal $I$ is known
to be doubly-exponential in $n$, the number of variables, and
polynomial in $d$, the degree of the ideal. A tight bound was derived
by T. W. Dube in \cite{Dube:gb-complexity}. I read his technical
report to try to understand the results at a high-level. Dube performs
a disjoint decomposition/partitioning on a homogeneous ideal, uses the
concepts of the Hilbert function and Macaulay constants (I don't
really know what they are) and derives the bound using combinatorial
arguments. He gives a long sequence of deductions which I glossed
over, and I just went to his final result. The result, given in the last
paragraph of the technical report, is summarized below:

\begin{Result}
Let $F = \{f_1, \dots, f_r\}$ be a homogeneous basis for an ideal $I
\subset K[x_1, \dots, x_n]$. Assume without loss of generality that
$f_1$ has the highest degree $deg(f_1) = d$. Then the degree of
polynomials in the Gr\"obner basis $G$ of $I$ is bounded by 
$2(\frac{d^2}{2}+d)^{2^{n-2}}$. This provides an upper bound on the
{\it degree of power products} which are required to generate the
leading term ideal $LT(I)$, and these power products have to appear in
the Gr\"obner basis\footnote{You may recall that $LT(I)$ is the set of
leading terms of all elements of $I$, i.e. $LT(I) = \{cx^{\alpha}:
\exists f \in I ~\text{with} ~ LT(f) = cx^{\alpha}\}$. Then $\langle
LT(I) \rangle = \langle LT(f_1), \dots, LT(f_r) \rangle$ if and only
if the set $\{f_1, \dots, f_r\}$ is a Gr\"obner basis --- this is one
of the many definitions of a Gr\"obner basis.}. 
\end{Result}

Clearly, the expression $2(\frac{d^2}{2}+d)^{2^{n-2}}$ is polynomial
in $d$ and double-exponential in $n$, and represents the {\it
complexity of the Gr\"obner basis problem}. This should not be
confused for the complexity of Buchberger's algorithm. Dube's
argument is combinatorial --- it just states that in the worst case
$G$ may require polynomials of degree $\leq
2(\frac{d^2}{2}+d)^{2^{n-2}}$ in the Gr\"obner basis. Buchberger's
algorithm, however, solves the problem using S-polynomials and their
reduction $Spoly(f_i, f_j) \xrightarrow{I}_+ r$, and keeps on pairing
the remainder $r$ with the current basis. To the best of my knowledge,
{\it the complexity of Buchberger's algorithm is still a mystery.}
This is because no one has been able to ``predict'' how many new
remainders $r$ may be generated by Buchberger's
algorithm. Buchberger's algorithm will definitely terminate --- thanks
to Dickson's Lemma --- but it is hard to predict when it will
terminate. I am unfamiliar with any new developments in the past few
years, but any serious development would have made headlines.... so I
am assuming this is still a correct assessment.

The above degree bound of $2(\frac{d^2}{2}+d)^{2^{n-2}}$ for the
Gr\"obner basis is in a {\it general setting} of a ring with
coefficients from a field $K[x_1, \dots, x_n]$. It turns out that life
might be simpler when working over zero-dimensional ideals (ideals
with finite varieties). Papers such as \cite{Lakshman:complexity_gb},
and more importantly \cite{Lakshman:complexity_gb_91}, state that the
number of arithmetic operations required to compute a reduced
Gr\"obner basis of a zero-dimensional ideal is {\it polynomial in
  $d^n$} --- this is a single exponential bound. Lakshman
\cite{Lakshman:complexity_gb_91} works over rationals (ring $R =
\Q[x_1, \dots, x_n]$) and says that {\it ``the   number of arithmetic
  operations required to compute reduced GB of $I   \subset R$ is
  polynomially bounded by the dimension of the residue   class ring
  $R/I$ seen as a rational vector space''}.

We are interested in Galois fields $\Fkk$. In \cite{gao:gf-gb-ms},
Sicun Gao shows that the complexity of computing a $GB(J + \langle
x_i^q - x_i\rangle)$ over $\Fq[x_1, \dots, x_n]$ is $q^{O(n)}$.
Analogous to Lakshman's result, this is also single exponential in
$n$. [See  section 4.1 in his MS thesis   \cite{gao:gf-gb-ms}]. This
is not surprising because $J + \langle x_i^q - x_i \rangle$ is
zero-dimensional.  In our case,  $q = 2^k$, so a complexity of
$2^{k\cdot n}$ is still prohibitive for practical verification.

To overcome this complexity, Jinpeng came up with a specialized term
ordering derived from the given circuit. In this term ordering, the
variables of the circuit are ordered in a {\it reverse topological
  order} and a lex term order is imposed. As a result, the polynomials
of our circuit themselves constitute a Gr\"obner basis. We formulate
the verification problem as an ideal membership test: is $f 
\xrightarrow{J+J_0}_+0$? Since we already have a Gr\"obner basis of
$J+J_0$ directly from the circuit, our approach obviates the need to
construct a Gr\"obner basis and performs verification only by a
sequence of polynomial divisions. Please refer to Section VI,
Proposition 6.1, and Theorem 6.1, in our paper \cite{lv:tcad2013} for
more details.

Note that every polynomial in our Gr\"obner basis is of the form 
$f_i = x_i + P_i$ where: 
\bi
\item $x_i$ is the output of a logic gate in the
design; 
\item $P_i = \text{tail}(f_i)$ corresponds to the function
implemented by the gate; and 
\item each variable $x_j$ that appears in $P_i$ is less than
  $x_i$. 
\ei

So, if the Gr\"obner basis is of the form $f_i: x_i + P_i$, then what
is the complexity of the division: $f \xrightarrow{x_i + P_i}_+r$?  
It turns out that in our experiments in \cite{lv:tcad2013}, the
polynomial $f$ (given specification polynomial to be verified against
a circuit) is a ``sparse'' polynomial of low degrees, e.g. $f: Z -
A*B$ for a multiplier circuit has degree = 2. Dividing low
degree polynomials $f \pmod{ x_i + P_i}$ might not be too
complicated. Jinpeng once (wrongfully) claimed that this complexity is
always bounded polynomially in the size of the Gr\"obner basis. This
is not true: If $f = x^n$, then dividing $x_i^n$ by $x_i + P_i$
requires raising $P_i$ to the $n^{th}$ power --- which is not
polynomial in the size of the Gr\"obner basis. So we need to analyze
this complexity further:

\begin{Problem}
\label{prob:complex}
Let $F = \{f_1, \dots, f_s\}$ be a set of polynomials in the ring $R =
\Fkk[x_1, \dots, x_n]$ where each polynomial $f_i$ is of the form 
$f_i = x_i + P_i$ where $P_i = \text{tail}(f_i)$. Let ideal 
$J = \langle F \rangle$. Let 
$F_0 = \{x_i^{2^k} - x_i: ~i = 1, \dots,n\}$ and let 
$J_0 = \langle F_0 \rangle$ denote the ideal of vanishing
polynomials. Let $f$ be a polynomial of total degree $d$. What is  the
computational complexity of the multivarite division $f \xrightarrow{F
  \cup F_0}_+ r$?
\end{Problem}

{\bf Here is one way to approach this problem:} 
Since multi-variate division $f\xrightarrow{f_1, \dots, f_s}_+ r$ is
modeled as a series of cancellation of terms $r = f -
\frac{LT(f)}{LT(f_i)} f_i$, it might suffice to derive an upper
bound on the number of terms that would be canceled in the division
process. In the worst case, all terms in $f$ can be
canceled. However, while a term is canceled, other terms may also
be added in the intermediate remainders. So we can further refine
Problem \ref{prob:complex} under these restrictions:

\begin{itemize}
\item First of all, we may assume that in each monomial, {\it each
  variable is raised to a degree  $\leq 2^k -1$}; this is due to the
  set of vanishing polynomials $x_i^{2^k}-x_i$ of $\Fkk$. 
\item Based on our approach \cite{lv:tcad2013}, we may assume a LEX
  term   order. On the other hand, if you prefer to use a degree-based
  (DEGLEX) order, feel free to do so, if it simplifies the analysis.
\item The polynomial $f$ is over $n$ variables $x_1, \dots, x_n$, and
  so the maximum total degree is $d = n(2^k - 1)$. So how many terms,
  at the most, can $f$ have? %Each of these will be canceled. 
\item In the set of polynomials $F = \{f_1, \dots, f_s\}$, we may
  assume that the size (number of terms) of each polynomial $f_i$ is
  bounded by  a known integer $t$. In other words, each $f_i$ has no
  more than $t$ terms. [So, at each ``one-step'' division
    $f\xrightarrow{f_i} r$, a term from $f$ is canceled, and no
    more than $t$ terms are added to $r$.]
\end{itemize}

\begin{Problem}
\label{prob:complex-refine}
Refining Problem \ref{prob:complex} further: If we were to use the
above restrictions, {\it what is the upper bound   on the number of
  terms that are canceled in the division $f   \xrightarrow{F \cup
    F_0}_+ r$?} This gives an estimate of the complexity of
solving verification using Jinpeng's and Tim's approaches
\cite{lv:tcad2013} \cite{pruss:dac14}. 
\end{Problem}

Is it very hard to find the bound in Problem
\ref{prob:complex-refine}? We need to solve this problem.

### 2015-05-01 techwriting/UNSAT-CORE/core.tex
all write-up from virtue machine

\section{On Unsatisfiable Cores of Ideals with Empty Variety}

Consider the polynomial ring $\Fkk[x_1, \dots, x_n]$, and an ideal $J$
generated by a set of polynomials $F = \{f_1, \dots, f_s\},
\idealj$. Suppose that it is known that $V(J) = \emptyset$, or
otherwise it is determined so by applying the Gr\"obner basis
algorithm. Throughout this document, we will assume that the set $F$
is such that $V(f_1, \dots, f_s) = \emptyset$, unless stated
otherwise. An interesting (and important) problem to solve is as
follows:

\begin{Problem}
Identify a subset of polynomials $F_c \subseteq F, J_c = \langle F_c
\rangle$, such that $V(J_c) = \emptyset$ too. For lack of a better
term, let us call $F_c$ the infeasible core or the unsatisfiable
core (UNSAT core) of $F$. The terminology is borrowed from the domain of
Boolean Satisfiability solvers. If you can think of a better
terminology, please let me know. 
%Here $F_c = \{f_i : f_i \in F, i\leq s\}$.
\end{Problem}

UNSAT cores have important uses in post-verification debugging, and in
many abstraction-and-refinement techniques used in the analysis of
circuits and systems. Any contribution here will be important.

\begin{Example}\label{ex1}
Consider the following set of polynomials:
\begin{align*}
f_1: & ab\\
f_2: & ab+a\\
f_3: & ab+b\\
f_4: & ab+a+b+1\\
f_5: & xy+y+x+1\\
f_6: & yz+y+z+1\\
f_7: & c + d\\
f_8: & c + d + 1
\end{align*}

Let $F = \{f_1, \dots, f_8\}$ and $J = \langle F \rangle \subseteq
\F_2[a, b, c, d, x, y]$ where $\F_2 = \{0, 1\}$. Then
$V_{\overline{\F_2}}(J) = \emptyset$ as $GB(J) = \{1\}$.

Now consider the subset $F_c = \{f_1, \dots, f_4\}, J_c = \langle
F_c\rangle$; it turns out that $V_{\overline{\F_2}}(J)= \emptyset$ as
well, so $F_c$ is an UNSAT core of $F$. Likewise, $\{f_7, f_8\}$ also
form a core. 
\end{Example}

Any set of polynomials $F$ with empty variety is an UNSAT core in
itself. There may be more than one UNSAT cores in $F$, some
polynomials may or may not be common to multiple cores (i.e. these
cores may be non-disjoint), and these cores may have
different cardinalities (different $|F_c|$). The UNSAT core problems
are therefore further classified as: 
\bi
\item Identify a {\it minimum} core, i.e. a core of minimum
  cardinality. In Ex. \ref{ex1}, the set $\{f_7, f_8\}$ is a minimum
  core. [Hardest problem to solve?]
\item Identify a {\it minimal} core. An UNSAT core $F_{c}$ is minimal
  with respect to the property that while the variety of $F_{c}$ is
  empty, the variety of {\it any} subset of $F_c$ is {\it non-empty},
  i.e. $V( F_{c} - \{f_i\} ) \neq \emptyset$ for any $f_i \in
  F_c$. The set $\{f_1, \dots, f_4\}$ constitutes a minimal
  core. [Very important problem, but I guess still a Hard Problem to
    solve!] 
\item Identify any UNSAT core disregarding minimality; i.e. find any
  subset $F_c$ of $F$. For very large problems, sometimes it is
  beneficial to solve this problem, so long as we can experimentally
  demonstrate that the obtained cores are small and their size is close to
  minimal cores --- but maybe algorithmically they can be found very
  quickly/efficiently, whereas finding minimal/minimum cores may be
  computationally infeasible. 
\ei

\begin{Conjecture}
Buchberger's algorithm, and only 1 iteration of it, could be used to
identify a {\it minimal} core. 
\end{Conjecture}

\begin{Conjecture}
Buchberger's algorithm may also be used to identify a {\it minimum}
core, but it will probably require the enumeration of all cores. 
\end{Conjecture}

So,
let us ignore the {\bf minimum} core problem for now, and concentrate
on the {\bf minimal} one.

The UNSAT core problem should be related to Gr\"obner bases. Suppose
that we are given $F_c = F - \{f_j\}$. Then, if $GB(F) = \{1\}$ and if
$GB(F_c) = \{1\}$ too, then doesn't it imply that $f_j \in \langle F_c
\rangle$? My reasoning is as follows: Denote $J = \langle F \rangle,
J_c = \langle F_c \rangle$. Since $V(J) = V(J_c) = \emptyset$, from
the Weak Nullstellensatz we have $J = J_c = \Fkk[x_1, \dots,
  x_n]$. This implies that $f_j$ can be composed of polynomials of
$F_c$, or $f_j \in F_c$. So $f_j$ is not part of the UNSAT core and
that can be identified using Gr\"obner bases. [Florian: Do you have a
  better/cleaner argument for this?]

\section{Investigating Algorithms to find UNSAT cores $F_c$ from $F$}

A n\"aive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is given in Algorithm 1. In the algorithm, we
pick a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
(i.e. $F_c - \{f_i\}$ is also infeasible/UNSAT). If so, discard $f_i$
from the core; otherwise keep $f_i$ in the 
core. Pick a different $f_i$ and continue until all polynomials $f_i$
are visited. This algorithm {\it will produce a minimal core}, as we
have tested each polynomial $f_i$ for inclusion in the core. This
requires $O(|F|)$ calls to the GB engine, which is really impractical.

\begin{algorithm}[hbt]
\caption{Find a minimal UNSAT core}\label{alg:MUS}
\label{alg1}
\begin{algorithmic}

 \REQUIRE{$F=\{f_1,\dots,f_s$\}, $J = \langle F \rangle$ } 
 \ENSURE{The minimal core $F_c \subset F$}
  %%%%%%%%%%%%%%%%%%%%
 \STATE{/* First of all, set $F$ as the core itself*/\;}
 \STATE{$F_c \gets F$;}

 \FORALL{$f_i \in F_c$}
   \IF{ $\text{Gr\"obnerBasis}(F_c - \{f_i\}) = \{1\}$ } 
     \STATE{/* Discard polynomial $f_i$ as it does not belong to the core*/\;}
     \STATE{$F_c \leftarrow F_c - \{f_i\}$\;}
   \ELSE 
     \STATE{ Keep $f_i$ in the minimal core\;}
   \ENDIF
 \ENDFOR
 \RETURN {$F_c$ as the core}
\end{algorithmic}
\end{algorithm}

The question is if we can identify a core by analyzing the
$Spoly(f_i,f_j)\xrightarrow{F}_+1$ computation itself? Note here
that since $F$ is an UNSAT core, there definitely exists one iteration
of Buchberger's algorithm where $Spoly(f_i, f_j) \xrightarrow{F}_+ 1$
is achieved. We think that a core can definitely be found, and we will
show you how. Then we need to investigate further whether our approach
can be used to identify minimal cores without invoking the GB engine
multiple times.

Recall that each step of division $f \xrightarrow{f_i}_+r$ operates as
$r = f - \frac{lt(f)}{lt(f_i)} f_i$. If $f$ is to be divided by (say)
two polynomials $f_i, f_j$, then: i) first we see if $lt(f_i) ~|~
lt(f)$. If so, then $lt(f)$ is canceled by $lt(f_i)$. ii) Otherwise,
we check to see if $lt(f_j)$ cancels $lt(f)$. This implies that while
performing the division $f \xrightarrow{f_1, \dots, f_s}_+r$, we can
identify (or keep track of) which polynomials in $f_1, \dots, f_s$ are
used to cancel terms of $f$.

I will demonstrate the core identification with an Example.

\begin{Example}
Consider the set $F = \{f_1, \dots, f_8\}$ given in Ex
\ref{ex1}. Assume that our term order is LEX with $a > b > c > d > x >
y$.  Apply Buchberger's algorithm to $F$.

\ben
\item Suppose that in the first
iteration $Spoly(f_1, f_4) \xrightarrow{F}_+r_1$ is computed. 
$Spoly(f_1, f_4) = a+b+1 = r_1$. Mark $f_1, f_2$ and keep them in the
set $F_m = \{f_1, f_4\}$. Since no polynomial
in $F$ divides $r_1$, we have $Spoly(f_1, f_4) \xrightarrow{F}_+ r_1 =
a+b+1$. No other polynomials (divisors) are marked.

\item Update $F = F \cup r_1$

\item In the next iteration of Buchberger's algorithm, suppose we pair
  $r_1$ with $f_4$, and obtain  $Spoly(r_1, f_4)  = a + b^2 +
  1$. Polynomial $a+b^2+1$ is divisible by $r_1$, so $Spoly(r_1,
  f_4)\xrightarrow{F}_+r_2$, where $r_2 = b^2 + b$. Since $r_1$ is not
  a polynomial from the original set (it is derived from $f_1, f_4$),
  it is not marked. Update $F = F \cup r_2$, and the set of marked
  polynomials is still $F_m = \{f_1, f_4\}$.

\item Let us draw a graph representing this sequence of
  computations. This is shown in Fig. \ref{fig:trace} (a). The
  vertices (nodes) in the graph represent polynomials (marked or
  unmarked). Every vertex (polynomial) has either 0 or 2 incoming
  edges representing the $Spoly$ pair that derived it. It may also
  represent the remainder obtained from the dividend and the divisor
  (e.g. $a+b^2+1 \xrightarrow{a+b+1}_+ b^2+b$). [This graph is just to
    illustrate a book-keeping procedure].

\item We continue: Suppose that Buchberger's algorithm picks
  $S(f_1, f_2)  = a$. Now $r_1$ divides $a$, so $S(f_1, f_2)
  \xrightarrow{r_1}_+ r_3 = b+1$. $F = \{f_1, \dots, f_8, r_1, r_2,
  r_3\}$. Mark $f_1, f_2$ so that $F_m = \{f_1, f_4, f_2\}$. 
\item Next, $Spoly(f_1, f_3) = b$, and $b\xrightarrow{r_3}_+1$. Bingo!
  Mark $f_1, f_3$, $F_m = \{f1, \dots, f_4\}$.  Note that  $r_3$ is not marked as
  it is a derived polynomial.

\item Update the graph, as  shown in Fig. \ref{fig:trace} (b).

\item Now, starting from the \textbf{1} node, traverse the graph
  backwards, to reach/identify all the vertices that are sources,
  i.e. those that have no incoming edges. Clearly, vertices 1, 2, 3, 4
  are the source vertices in this case. Of these vertices, identify
  only those that are marked in $F_m$. In this special case, all 4 of
  them are marked; so the core is obtained as $F_c = F_m$; in general, 
  we would only chose the marked vertices in the core.

\item Just for fun, and to prove a point, compute $S(f_5, f_6)
  \xrightarrow{F}_+0$, so nothing is added to the Spoly graph. On the
  other hand, $Spoly(f_7, f_8)$ would have given us the   minimum
  core. 
\een

\end{Example}

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.6]{trace.eps}
\caption{Keeping track of polynomials involved in $GB(J) = \{1\}$}
\label{fig:trace}
\end{figure}

Coming to the point: i) Keep track of all $Spoly(f_i,
f_j)\xrightarrow{f_1, \dots, f_s}_+ r$; ii) mark $f_i, f_j$ if $r\neq
0$; iii) Mark those $f_l$'s that have leading terms that cancel
monomial terms in $Spoly(f_i, f_j)$; iv) Build this Spoly graph (I
don't yet have a catchy name for this graph); v) Once $1\in J$ is
detected, traverse the graph to identify all the marked polynomials that were
used in obtaining this unit element. These polynomials constitute the
core.

\begin{Problem}
Does the approach guarantee a minimal core? If yes, how do we show it?
If not, we need to generate an example to understand it better. The
example that I have derived is too trivial. Probably we need to find
more complicated examples to test the idea. 
\end{Problem}

Questions and comments are welcome.

%% Clauses:
%% \begin{align*}
%% \bar{a}\lor\bar{b}\\
%% a\lor\bar{b}\\
%% \bar{a}\lor b\\
%% a\lor b\\
%% x\lor y\\
%% y\lor z
%% \end{align*}

### 2015-05-01 techwriting/UNSAT-CORE/counterexample.tex
all write-up from virtue machine

\section{Counterexample}

Recall the conjecture we made:
\begin{Conjecture}
\label{conj:original}
If a polynomial $f$, that is not part of a minimal core, is not on the longest path in our graph, 
then it would never be found on a shorter path.
\end{Conjecture}
However, it is not always the case. Look at the following example:

\begin{Example}
\label{exi}
Consider the following set of polynomials:
\begin{align*}
f_1: & abc+ab+ac+bc+a+b+c+1\\
f_2: & b\\
f_3: & ac\\
f_4: & ac+a\\
f_5: & bc+c\\
f_6: & abd+ad+bd+d\\
f_7: & cd\\
f_8: & abd+ab+ad+bd+a+b+d+1\\
f_9: & abd+ab+bd+b
\end{align*}

Here we use degree lexicographical term order (Dp) $a>b>c>d$. Let $F = \{f_1, \dots, f_9\}$ and $J = \langle F \rangle \subseteq
\F_2[a, b, c, d]$ where $\F_2 = \{0, 1\}$. Then
$V_{\overline{\F_2}}(J) = \emptyset$ as $GB(J) = \{1\}$.

During GB computation we use no optimized s-poly picking strategy, just following the order of polynomials given.
In this example we choose pairs in following order ($1,2,\dots$ means $f_1,f_2$,etc.):
$$(1,2) \to (1,3)(2,3) \to (1,4)(2,4)(3,4) \to \dots$$
The s-poly reduction leading to "1" can be depicted by a tree:
\begin{figure}[hbt]
\centering
\includegraphics[scale=0.6]{tier1.png}
\caption{GB-UNSAT tree in original polynomial order}
\label{fig:tier1}
\end{figure}

In this graph, longer edges (1 unit length) denote s-poly pairs and shorter edges (0.5 unit length) denote
polynomials used for division/reduction. Illustration of this graph: s-poly of $(1,2)$ could be divided by $f_2,f_3$
and s-poly generated by $(3,4)$ and $(2,5)$. UNSAT core attained using this GB computation method is $\{1,2,3,4,5\}$.
However, the minimal core is $\{1,2,4,5\}$ so $f_3$ is "redundant" here, but it is not 
on the longest path.
\end{Example}

This example violates conjecture \ref{conj:original}. Still, we notice the "redundant" $f_3$ could be found in other
minimal cores. In fact, there are at least 4 minimal cores for this set of polynomials:
\begin{align*}
& \{1,2,3,4,7,8\}
& \{2,4,5,6,8\}
& \{2,3,4,6,8\}
& \{1,2,4,5\}
\end{align*}

Maybe we can refine our conjecture like this:
\begin{Conjecture}
\label{conj:refine}
Any polynomial which is included by GB computation method but not in any minimal core will be on the longest path.
\end{Conjecture}

To verify this conjecture we will need to perform more non-trivial but reasonably-small experiments.

\section{Smaller UNSAT core extraction}
If an analytical method for minimal UNSAT core extraction is intractable, we may want to get UNSAT core as small as possible.
To achieve this goal with limited number of iterations, we may need following conjecture:
\begin{Conjecture}
By carefully arranging the order of picking s-poly pairs, we can always get a new core whose size is no larger than
last iteration.
\end{Conjecture}

Or we can loose the constrain:
\begin{Conjecture}
By carefully arranging the order of picking s-poly pairs, the probability that we get a smaller core
is much higher than the probability that we get a larger core.
\end{Conjecture}

Here I will propose a simple rearrange strategy, which works at least for Ex.\ref{exi}:

\emph{After getting an UNSAT core, we can rearrange the order of polynomials by following criteria:\\
i. Polynomials in the core come first;\\
ii. Polynomial with a shorter distance to refutation "1" comes first;\\
iii. When there are 2 polynomials with the same distance, the one appears more frequently in the tree comes first.}

\begin{Example}
Let us make Ex.\ref{exi} as the first iteration, we get a core $\{1,2,3,4,5\}$. Check out Fig.\ref{fig:tier1},
we can rearrange polynomials as
$$\{2,5,4,3,1,6,7,8,9\}$$
Run our approach using this order, we get a new core $\{1,2,4,5\}$, as Fig.\ref{fig:tier2} shows:
\begin{figure}[hbt]
\centering
\includegraphics[scale=0.6]{tier2.png}
\caption{GB-UNSAT tree after rearranging polynomial order}
\label{fig:tier2}
\end{figure}
Note "1 2 3 5" here are new indices corresponding to original $f_1,f_2,f_4,f_5$. Applying the criteria
we further rearrange polynomials as
$$\{2,5,4,1,3,6,7,8,9\}$$
Run our approach again, the result is still core $\{1,2,4,5\}$, which is minimal. Notice we arrive a fixpoint.
\end{Example}

Again, my conjecture need more experiments. I will keep updating and discussions are welcome.

### 2015-05-01 techwriting/present/presentation.tex
all write-up from virtue machine

%\documentclass[pdf,final,colorBG,slideColor]{prosper}
\documentclass[xcolor=dvipsnames]{beamer}
%%\usetheme{default}
\usecolortheme[named=Maroon]{structure}
%\usetheme{Boadilla}
\usetheme{Madrid}
\useoutertheme{default}%[footline=empty]{infolines}

% \usepackage{helvet}
% \usepackage{enumerate}
% \usepackage{amsmath}
% \usepackage{amsfonts}
% \usepackage{graphicx}
% \usepackage{ulem}
% \usepackage{multirow}
\usepackage{comment}
\usepackage{xspace}

\usepackage[absolute,overlay]{textpos}
\usepackage[ruled]{./algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\RequirePackage{algorithmic}
%\RequirePackage{algorithm}
% \renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
% \renewcommand{\algorithmicensure}{\textbf{Outputs:}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{lemma}
%\newtheorem{corollary}{Corollary}
%\newtheorem{proposition}{Proposition}
%\newtheorem{Q}{Question}
%\newtheorem{Exa}{Example}
%\newtheorem{Definition}{Definition}

\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Zkk}{{\mathbb{Z}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
%\newcommand{\Grobner}{Gr\"{o}bner}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\F}{{\mathcal{F}}}
\newcommand{\B}{{\mathbb{B}}}

\title[Ph.D Proposal]{Sequential Circuit Verification at Word Level using Algebraic Geometry}

\author[Xiaojun Sun]{Xiaojun Sun}

%\email{rostamian@umbc.edu}
\institute[Univ. of Utah]{
%\includegraphics[height=17mm]{/Users/Kalla/teaching/Comp-Algebra-Course/lectures/old_ulogo.eps}\\
Ph.D Candidate\\
Electrical and Computer Engineering, University of Utah\\
xiaojuns@ece.utah.edu\\
\ \\
\ \\
{\bf Ph.D's Dissertation Proposal}\\
}

\date{}
%\slideCaption{}

%% Images
%\pgfdeclareimage[width=.4in]{fg:logo}{/Users/Kalla/teaching/Comp-Algebra-Course/lectures/old_ulogo.eps}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%----------- titlepage ----------------------------------------------%
\begin{frame}[plain]
  \titlepage

\end{frame}

%\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Agenda}}

\begin{itemize}
\item Focus
	\begin{itemize}
	\item Implicitly analyze the reachability of a sequential circuit at word level
	\item Use algebraic geometry to assist in sequential circuits abstraction refinement
	\end{itemize}
\item Motivation
	\begin{itemize}
	\item Bit-level to word-level abstraction, BFS traversal of FSMs
	\end{itemize}
\item Target problems 
	\begin{itemize}
	\item Given a sequential FSM, with $k$-bit state variables,
     perform an implicit state enumeration at word level, i.e. use word-level variables from $\Fkk$ to represent reachable states
        \item Extract UNSAT cores efficiently from a given set of CNF clauses
	\end{itemize}
\item Approach: {\bf Algebraic geometry techniques}
	\begin{itemize}
	\item  \Grobner basis methods + Elimination ordering + BFS traversal
        \item  Challenge: Discover efficient algorithmic techniques to implement image computations, 
        set operations, UNSAT proofs, etc. at word level
        \item  Proposed Contribution: Polynomial abstraction as well as algebraic geometry techniques 
        is applied to reachability analysis; A new algorithm based on \Grobner basis computation is explored to extract UNSAT cores
	\end{itemize}
%\item Results \& Conclusions
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large {Motivation}}
\vspace{-0.2in}
%\ptsize{10}

\begin{itemize}
\item Importance of reachability analysis in sequential circuits verification
	\begin{itemize}
	\item Circuits $\to$ state machines; errors $\to$ bad states
	\item Bad states are reachable $implies$ errors affect circuit behavior
	\end{itemize}	
\end{itemize}

\begin{itemize}
\item Advantages exploiting word-level verification
	\begin{itemize}
	\item Many circuit datapaths/system models are described at word level
	\item Reduce state space, avoid ``bit-blasting"
	\end{itemize}	
\end{itemize}
% 
% \bi
% \item Bugs in hardware can leak secret keys [{\it Biham et al.}, ``Bug
%   Attacks'', Crypto 2008] 
% \ei

% \begin{itemize}
% \item Type of circuits in ECC
% 	\begin{itemize}
% 	\item Multiplication dominates Galois field computations
% 	\item In cryptography: $99\%$ time in encryption and decryption
% 	\item 
% 	\end{itemize}
% \end{itemize}

%  \begin{itemize}
%  \item Data-path size in ECC crypto-systems can be very large
%  	\begin{itemize}
%         \item In $\mathbb{F}_{2^k}$, $k = 163, 233, \dots$ (NIST standard)
%         \item ECC-point addition for encryption, decryption, authentication
%  	\item Custom arithmetic architectures -- hard to verify
%         \item Synthesized circuits are ``easier'' to verify
%  	\end{itemize}
%  \end{itemize}

\begin{itemize}
\item Why use algebraic geometry?
	\begin{itemize}
        \item Provide a symbolic representation for both bit and word level variables in one unified framework
	\item Recent work shows it is practical to apply algebraic geometry to circuit verification
	\end{itemize}
\end{itemize}

\end{frame}

\end{document}

### 2015-05-01 techwriting/proposal/abs.tex
all write-up from virtue machine

\begin{abstract}
This PhD dissertation proposal addresses the problem of sequential circuit verification at the word-level and is based on concepts derived from algebraic geometry. Analyzing the sequential circuits at word level is an efficient way
of {\it abstraction}, which may  lower the complexity of verification by efficient representation of the state-space. We propose to model the verification properties and the gate-level sequential circuit implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, techniques from algebraic geometry can be used to reason about the state-space by analyzing varieties of these ideals.

We propose to apply these techniques to traverse a finite state machine (FSM) for reachability analysis at the word-level, and also to 
implicitly unroll sequential arithmetic circuits to verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern abstraction-refinement techniques for verification, we propose to investigate a word-level analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The proposal will not only derive new algorithms and techniques, but will also consider efficient CAD implementations to target sequential equivalence and model checking problems. 
\end{abstract}

### 2015-05-01 techwriting/proposal/abs_geo.tex
all write-up from virtue machine

\section{Combine algebraic geometry with abstraction techniques}
\subsection{History of abstraction on sequential circuits}
Traditional method to analyze a sequential circuit is based on explicit state variable encoding
and full states enumerating from initial state and explicit transition function.
With the size of modern sequential circuits increasing, it is infeasible to apply old methods
any longer. Abstraction is a technique to minimize the state space by combining states with certain 
similarities. Sometimes it can effectively lower down the number of states by orders of magnitude,
without affecting the properties we need to verify.

At first abstraction is done manually by designers. In 2000 E. Clarke \cite{clarke2000counterexample}
proposed an automated abstraction: first,
initial abstraction is set up by dividing variables to different clusters based on transition conditions;
then, an ordinary model checking is operated on this abstraction. If a counterexample is given, 
it is further checked to be concrete or not. For a false abstract state or a spurious path 
, the initial abstraction is refined by further splitting the false abstract state. 
A heuristic algorithm keeps refining the abstraction until a real counterexample is found or 
property verified. This technique is based on binary decision diagrams (BDDs), which means
potential risk of memory explosion; meanwhile it still requires an initial abstraction to start with.
So in 2004 L. Zhang and M. Hsiao \cite{zhang2005design} proposed another abstraction method based on CNF-SAT.
This paper uses simplest latch abstraction -- by removing "irrelevant" latches and change them to primary inputs. 
The key idea is to pick those ``irrelevant" latches (variables in $k$-bounded model checking, $k$-BMC) 
according to SAT/UNSAT analysis from $k$-BMC. It improves the $k$-BMC procedure\cite{biere1999symbolic}: 
LTL properties to check for a $k$-bounded machine can be translated to a big CNF and fed to SAT solver. 
If SAT, a concrete counterexample is found. 
If UNSAT, the original method will just increase bound k. In Zhang's paper, the clauses for UNSAT 
(UNSAT cores) are checked for variables do not appear, they are referring to "irrelevant" latches 
at least for first $k$ iterations. 
Thus an abstracted model is obtained and checked with original property. If the abstracted model checking fails, 
then bound $k$ will increase. This method still has disadvantages: it is counterexample-independent,
which means it cannot utilize information from invariants.

Abstraction usually means over-approximation, i.e. properties are true on abstracted machine
are also valid for original states. Craig's interpolation is a sort of abstraction which could be both
over-approximation and under-approximation. In 2003 K. McMillan used Craig's interpolation to improve 
$k$-BMC \cite{mcmillan2003interpolation}. Initially a $k$-length path to failure state $F$ is picked and transformed to 
CNF formula, then split at first transition into prefix and suffix. Then an interpolation $P$ is computed
between prefix and suffix. $P$ over-approximates 1-step reachable states, and under-approximates states 
that cannot reach $F$ in $k$ steps. If no counterexample exposed, path will be split again at second
transition; in this way a precise abstraction can be found for all reachable states.

$k$-BMC with interpolation is a purely incremental model checking approach, and interpolation procedure relies
on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
\cite{bradley2011sat} \cite{bradley2011incremental}. (Not related to abstraction really..)
\subsection{Word-level abstraction: Naive abstraction for arithmetic circuits}
Our approach is an abstraction based on word operand definition of datapaths in arithmetic circuits,
and it can be expanded to generic FSM by bundling a set of bit-level variables to one or several word-level variables.
The abstraction polynomial is obtained through computing GB of an elimination ideal using a special term order.

The authors of \cite{pruss:dac14} showed that for any combinational
logic block, a canonical word-level polynomial representation can be
derived through \Grobner bases computed with elimination
orders. Our approach is based on their result, which we reproduce
here:
\begin{Lemma}
(From \cite{pruss:dac14}) Given a combinational circuit $C$ with $k$-bit
  input $A = (a_0, \dots, a_{k-1})$ and $k$-bit output $R = (r_0, \dots,
  r_{k-1})$. Denote by $x_1, \dots, x_d$ all the bit-level
  variables of   $C$. Let $J = \langle f_1, \dots, f_s \rangle \subset
  \Fkk[x_1, \dots, x_d, R, A]$ denote all the polynomials corresponding to the
  logic gates of the circuit. Let $J_0 = \langle x_1^2 - x_1, \dots,
  x_d^2 - x_d, R^q - R, A^q - A \rangle$ be the vanishing ideal, so
  that $J + J_0 = \langle f_1, \dots, f_s, ~~ x_1^2 - x_1, \dots,
  x_d^2 - x_d, R^q - R, A^q - A \rangle$. Compute \Grobner basis $G =
  GB(J + J_0)$ w.r.t. lex term order with $x_1 > x_2 > \dots > x_d > R
  > A$. Then $G_d = G \cap \Fkk[R, A]$ eliminates the internal
  variables $x_1, \dots, x_d$ of the circuit. $G_d$ also contains the
  word-level polynomial $R = \F(A)$ which canonically represents the
  function of the circuit.  
\end{Lemma}

The authors referred to the elimination (lex) order with
$\{\textit{internal variables} \ x_1 > \dots > x_d\} > \{\textit{word-level
  output} \ R\} > \{\textit{word-level input} \ A\}$ as the {\it abstraction
  term order (ATO)}.

One-step transition can also be implemented on word level using ATO. We represent present state (PS) 
with a univariate polynomial about word-level state variable $f(S)$, its variety intersecting with variety of
circuit ideal $J+J_0$ results to polynomial about word-level next state (NS) $g(T)$. Recall theorem \ref{thm:unionintersect},
$$NS = {\bf V}(\langle g\rangle) = {\bf V}(\langle f\rangle + J+J_0) = GB\ with\ ATO(\langle f\rangle + J+J_0)$$
Note that ideal corresponding to next state is an ideal with single univariate polynomial generator.

Reachability analysis is a useful tool when checking sequential equivalence as well as other invariants.
With word-level state variables, we can do implicit state enumeration to provide a picture of 
reachable states. Following algorithm is breadth-first search style state enumeration \cite{KallaPartialScan}:
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}\label{alg:BFS}
\end{algorithm}
\begin{figure}[hbt]
\centering{
\includegraphics[scale=0.3]{./stg_fig.eps}
\caption{State transition graph of sample FSM}
\label{fig:stg}}
\end{figure}

\begin{figure}[hbt]
\centering{
\includegraphics[width=3.5in]{./fsm_fig.eps}
\caption{Gate-level circuit of combinational logic block of sample FSM}
\label{fig:fsm}}
\end{figure}
Here we put a 2-bit Mealy machine as an example showing how to use univariate polynomial ideal
and elimination ideal with ATO to implement this algorithm.
\begin{Example}
\label{ex:2bit}

We are given a mealy Finite State Machine (FSM). Gate-level circuits of combinational logic
is figure \ref{fig:fsm}, $x$ is primary input (one bit),
$\{s_0, s_1\}$ are state inputs, $Z$ is primary output and $\{t_0, t_1\}$ are state outputs.
Word-level state input is $S = s_0 + s_1\cdot\alpha$, output is $T = t_0 + t_1\cdot\alpha$. In Galois field
$\mathbb{F}_{2^2}$, bit-level variables $x, s_0, s_1, Z, t_0, t_1$ can take values from $\{0, 1\}$, while
word-level variables $S, T$ may take values from $\{0, 1, \alpha, 1 + \alpha\}$. State Transition Graph (STG)
showed in figure \ref{fig:stg} uses 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2, S_3\}$, which
could be mapped to elements in $\mathbb{F}_{2^2}$.

For example in Line 1 of BFS algorithm, assume initial state is $S_0$ in figure \ref{fig:stg}, then corresponding
polynomial $f = \mathcal{F}(S^0) = S^0 - 0$. Consider an ideal $I$ with only one generator $f$, its variety
$V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 0\}$, which equals to $\{0\}$, the only
valid value $S^0$ can take.

If an ideal contains multiple polynomial specifications, it is necessary to compute Gr\"obner basis with elimination term
order to get one polynomial only on desired variable. In first iteration of Line 4, $to^i$ has multiple specifications,
some of them are transition functions on bit-level:
\begin{displaymath}
\begin{array}{ll}
f_1:& t_0 - (\overline{x}\ \land\ \overline{s_0}\ \land\ \overline{s_1})\ \lor\ (s_0\ \land\ s_1) \\
f_2:& t_1 - (\overline{s_0}\ \land\ x)\ \lor\ (s_0\ \land\ \overline{s_1})\
\end{array}
\end{displaymath}
And some are bits-to-word definitions using polynomial representation of elements in $\mathbb{F}_{2^2}$:
\begin{displaymath}
\begin{array}{ll}
f_3:& S - s_0 - s_1\alpha \\
f_4:& T - t_0 - t_1\alpha
\end{array}
\end{displaymath}
And an polynomial about initial state as mentioned above:
\begin{displaymath}
f_5:\  S
\end{displaymath}
And the rests are vanishing polynomials for every variable, bit-level and word level:
$f_6: x^2 - x; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: s_0^2 - s_0;
f_{11}: s_1^2 - s_1; f_{12}: T^4 - T$

Note that there is no specification on initial primary input $x$, in Gr\"obner basis approach this means $x$ is smoothed by
reversely using Shannon's expansion.

Employing elimination term order: $intermediate\ bit$-$level\ signals\ >\ bit$-$level\ primary\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
$J = \langle f_1, f_2, \dots, f_{12}\rangle $, the result will include one polynomial $f_T$ contains only variable $T$. In this example,
$f_T = T^2+(\alpha+1)T+\alpha$. This polynomial equals to $T^2-(\alpha+1)T+\alpha$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
limited within $\mathbb{F}_{2}$, where $x \equiv -x\ (mod\ 2)$ for any element $x$ in the field. Solution to $f_T = 0$ is $T = 1\ or \ T = \alpha$, which shows that next
state the machine just reached is $\{S_1(01), S_2(10)\}$.

Continue to run this traversal algorithm, it will terminates with final reachable states $f_T = T^3+(\alpha+1)T^2+\alpha T$, its solution is $T=0\ or\ T=1\ or\ T=\alpha$,
which shows within total 4 states, state $S_3$ is unreachable from initial state $S_0$.

When executing Line 5 and Line 6, use \textbf{union}, \textbf{intersection} and \textbf{complement} of varieties over $\mathbb{F}_{2^k}$. Since
state set variables $from^i, to^i, reached$ all have single polynomial generator, consider an ideal $I$ with single generator $I = \langle f\rangle $.
For example, assume $I = \langle f\rangle  = \langle T^2 + (1+\alpha)\cdot T+\alpha\rangle $, its variety $V(I) = \{a\ |\ a \in \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$.

For varieties intersection $\{1\}\bigcap\{1, \alpha\}$, calculate ideal sum $\langle T+1, T^2 + (1+\alpha)\cdot T+\alpha\rangle  = \langle T+1\rangle $,
its variety is $\{1\}$; for varieties union $\{1,\alpha\}\bigcup\{1+\alpha\}$, calculate
ideal product $\langle (T+1+\alpha)\cdot(T^2 + (1+\alpha)\cdot T+\alpha)\rangle  = \langle T^3 + 1\rangle $, its variety
is $\{1, \alpha, 1+\alpha\}$; for complement set of variety $\{1, \alpha\}$, the universal set is
the variety of ideal of vanishing polynomial $V(\langle T^4-T\rangle ) = \{0,1,\alpha,1+\alpha\}$,
so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
the result is $\{0,1+\alpha\}$.

From discussion above, the BFS traversal is completely implemented in Galois field.
If the initial input is $S_3$: $S + 1 + \alpha$, the final return value will be set of
reachable states: $T^4 + T$, i.e. the universal set, $\{S_0, S_1, S_2, S_3\}$. Modified algorithm could be
written in following interpretation:

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Input-output circuit characteristic polynomial ideal $I_{ckt}$, initial state polynomial $\F(S)$}

  $from^0 = reached = \F(S)$\;
  \Repeat{$new^i == 1$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$GB w/ elimination term order$\langle I_{ckt}, from^{i-1}\rangle$\;
	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
	$from^i \gets new^i(S\setminus T)$\;
  }
\Return{$reached$}
\caption {Algebraic Geometry based Traversal Algorithm}\label{alg:univa}
\end{algorithm}

Here $from^i, to^i, new^i$ are all univariate polynomials about $S$ or $T$.
\end{Example}

\subsection{Obviate GB computation}
% Introduce: RATO
% Get word-level output -- bit-level inputs function
% Use bit variable substitution to finish word-level abstraction
Elimination ideal with ATO technique requires a GB computation, which is usually infeasible on large designs.
To obviate GB computation, we introduce a refinement on ATO as well as bit-word substitution method.

To overcome high computational complexity of GB calculation, \cite{pruss:dac14} proposed a
\underline{r}efinement of ATO (called RATO), and simplified the
\Grobner basis computation. They exploited Buchberger's product
criteria \cite{productc:1979}, which states that: {\it If the leading  
monomials of $f_i, f_j$ are relatively prime, then $Spoly(f_i, f_j)$
reduces to zero modulo the generating set, i.e. $Spoly(f_i, f_j)
\xrightarrow{F} _+ 0$.} This concept was exploited in RATO as follows:

Perform a reverse topological sorting of the nodes in the
combinational logic, and define a {\it lex} term order by the
following relation $>_{r}$: {\it bit-level circuit variables ordered
  reverse topologically} $>$  {\it word-level output variables} $>$
{\it word-level input   variables}. Representing the polynomial ideal
$J$ in RATO has the effect that there exists {\it one and only one
  pair of polynomials} in $J$ that do not have relatively prime
leading terms (see  Section 5 in \cite{pruss:dac14} for details). All
other polynomial pairs will have leading terms that are 
relatively prime, so these polynomial pairs are not considered in
Buchberger's algorithm.  The authors of \cite{pruss:dac14} exploited
this concept and showed how the \Grobner basis of $(J + J_0)$ can be
computed by a {\it subset} of polynomials, which  improves the
scalability of their approach. Their approach, however, cannot
circumvent the \Grobner basis computation altogether. Consequently, 
their approach fails to derive a canonical polynomial abstraction when
the representation is dense, and contains monomials of high-degrees
(e.g. in case of buggy designs).

It turns out that RATO can be applied to sequential circuits in much
the same way: {\it \{bit-level circuit variables ordered
  reverse topologically} $x_1 > \dots x_d\}$ $>$  {\it \{word-level NS
  variables} $R'> A'> B'\}$ $>$ {\it \{word-level PS variables}
$R>A>B\}$. Importantly, {\it we show that using RATO, the polynomial
abstraction can be derived without resorting to a \Grobner basis
computation.} Perform the following operations:

\begin{enumerate}
\item Represent the polynomials of the sequential circuit $S$ using RATO.
\item Due to RATO, only one pair of polynomials ($f_i, f_j$) will have
  leading terms that will not be relatively prime.
\item Reduce $Spoly(f_i, f_j) \xrightarrow{F}_+ h$.
\end{enumerate}
As described in \cite{pruss:dac14}, remainder $h$ will contain:
  i) the word-level variables, and ii) bit-level inputs to the
  combinational logic, i.e. bit-level present-state variables. All
  other internal circuit variables will not appear in $h$, as they
  will be canceled by division due to RATO.

However with only RATO, the final remainder $h$ contains both bit-level variables and
word-level {\it state variables}. We desire
a polynomial in only word 
level variables without computing a
\Grobner basis. This can be achieved if we  represent the
bit-level state variables in terms of their word-level
variables. We exploit 
the following property of finite fields:

\begin{Lemma}
\label{lemma:square}
For $\alpha_1, \dots, \alpha_t \in \Fkk$, $(\alpha_1 + \alpha_2 +
\dots + \alpha_t)^{2^i} = \alpha_1^{2^i} + \alpha_2^{2^i} + \dots +
\alpha_t^{2^i}$, for $i = 1, 2, \ldots$. 
\end{Lemma}

Consider the polynomials that define the relationship between the
word-level and bit-level variables. Let 
$ A = a_0 \beta + a_1\beta^2 + \dots + a_{k-1}\beta^{2^{k-1}}$.
Due to Lemma \ref{lemma:square}, we have $A^2 = a_0 \beta^{2}+ a_1
\beta^{4} + \dots + a_{k-1} \beta^{2\cdot 2^{k-1}}$ (as
$a_i^2 = a_i$). By repeated squaring:
\begin{eqnarray}
A & = & a_0 \beta + a_1\beta^2 + \dots + a_{k-1}\beta^{2^{k-1}}\nonumber\\
A^2 & = & a_0 \beta^{2}+ a_1 \beta^{4} + \dots + a_{k-1}
\beta^{2\cdot 2^{k-1}} \nonumber\\
\vdots & \vdots& \vdots \nonumber\\
A^{2^{k-1}} & = & a_0 \beta^{2^{k-1}}  +  a_1 \beta^{2^{k-1}\cdot 2} +
\dots + a_{k-1}\beta^{2^{2(k-1)}} \nonumber
\end{eqnarray}
% \alert{Some concern that we do not show that this system of equations is always
% uniquely solvable.}
Consider the above as $k$ {\it linear equations} with $a_0, \dots, a_{k-1}$
as $k$-unknowns, with $\beta$ and its powers as coefficients, and 
$A, A^2, ... \dots, A^{2^{k-1}}$ as $k$ constants. Then, we can solve
for the unknowns $a_0, \dots, a_{k-1}$ and obtain expressions for them
in terms of $A$ and $\beta$. The problem can be setup in
matrix form:

\begin{displaymath}
\begin{pmatrix}
\beta & \beta^2 & \cdots & \beta^{2^{k-1}} \\
\beta^2 & \beta^4 & \cdots & \beta^{2^{k-1}\cdot 2} \\
\vdots & \vdots & \ddots & \vdots \\
\beta^{2^{k-1}} & \beta^{2^{k-1}\cdot 2} & \cdots & \beta^{2^{2(k-1)}}
\end{pmatrix}
\begin{pmatrix}
a_0\\
a_1\\
\vdots\\
a_{k-1}
\end{pmatrix}
=
\begin{pmatrix}
A\\
A^2\\
\vdots\\
A^{2^{k-1}}
\end{pmatrix}
\end{displaymath}

Gaussian elimination on this system of equations can be 
applied, and each bit-level variable $a_i$ can be represented as a
function of the word-level variables: $a_i = g_i(A)$. These bit-level
variables can be easily substituted in the remainder $h$ obtained by
reduction due to RATO. What we will obtain is a word-level polynomial
 which canonically represents the function of the circuit.

Using RATO to get word-level transition function may result in multivariate polynomial generators
instead of single univariate generator. To compute NS, we intersects variety of transition ideal 
with PS (output word's function about input words, plus
evaluations of input words), then result NS may have multiple multivariate generators. 
Complement also requires multivariate ideal quotient.

\begin{Example}
Consider 2-bit FSM traversal (Ex:\ref{ex:2bit}) using RATO and multivariate polynomial generators.
Following algorithm is a modified breath-first search (BFS) traversal algorithm based on ideals with multivariate 
generators.
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition polynomial $f_t = T + \mathcal F (S,x)$, 
	initial state ideal $from^0 = \langle S+\mathcal G(x), x^{q_1} - x\rangle$}

  $reached = from^0(T\setminus S)$\;
  \Repeat{$GB(new^i) == 1$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$GB$(\langle f_t, from^{i-1}\rangle) \setminus \mathcal H(S)$\;
	$\overline{reached} = \langle T^{q_2}-T, x^{q_1} - x \rangle : reached$\;
	$new^i \gets $GB$(to^i + \overline{reached})$\;
  	$reached \gets $GB$( reached \cdot new^i)$\;
	$from^i \gets new^i(S\setminus T)$\;
  }
\Return{$reached$}
\caption {Algebraic Geometry based Traversal Algorithm (multivariate-generator ideals)}\label{alg:multi}
\end{algorithm}

Inputs of this algorithm are: a transition polynomial, which is the result of word-level abstraction; initial 
states description ideal, contains 2 generators defining PI could be any inputs and state inputs $S$ is a specific
value. (Ex: $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
$\langle S+x\cdot\alpha, x^2+x\rangle$ means initial states$=\{00,10\}$).

Transition polynomial calculation exploits our improved RATO approach. After building
an elimination ideal, use RATO such that \emph{reverse\ topo\ order\ circuit\ variables }$> T > S > x$, the reduction
remainder has the form $T+\mathcal F(s_0,s_1,x)$. From word definition $S+s_0+s_1\alpha$ we get
$$s_0 = \alpha S^2+ (1+\alpha)S, s_1 = S^2+S$$
Do substitution, the transition polynomial of example circuit is 
$$f_T = T+S^3\cdot x+\alpha S^3+(1+\alpha)S^2\cdot x+S^2+S\cdot x+(1+\alpha)x+1$$

Assume we start from state $\{11\}$. First iteration, it will reach state $\{01\}$. Line 4 is to compose an
ideal with 2 generators from $from^0$ and transition polynomial $f_T$, compute its Gr\"obner basis. Note this ideal
has the form
\begin{equation}
I_{tran} = \left\{
             \begin{array}{c}
             T+\mathcal F(S,x) \\
             S + \mathcal G(x) \\
             v_x
             \end{array}  
        \right.
\end {equation}
$v_x$ is a polynomial containing only $x$, initially it should be vanishing polynomial $x^{q_1}-x$, with the program executing
it may be factorized.

Consider Buchberger's algorithm, all generators' leading terms are relatively prime, so it is a GB itself. 
Then we need to reduce this GB, we will find out $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. So, at last we will get a polynomial
$T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
$S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$

Line 5 is the ideal quotient of universal set and reached states. In the first iteration, 
$reached$ is the initial state $\langle T+1+\alpha, x^2+x \rangle$. Result of ideal quotient
is $\langle T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle$ represents $\{00,01,10\}$.

Line 6 is ideals' sum (intersection of their varieties), it is done by combining all generators
from 2 ideals and compute GB. For first iteration result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.

Line 7 is ideals' product (union of their varieties), it is done by multiplying all pairs of
generators from both ideal. For first iteration result is $GB(\langle (T+1)(T+1+\alpha),
(T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$ representing $\{01,11\}$.

The traversal will run for 3 iterations and terminate at 4th iteration. We list all intermediate 
results below:
\begin{itemize}
\item Iteration 1: $from^0 = \langle S+1+\alpha, x^2+x\rangle, to^1 = \langle T+1, x^2+x\rangle,
 reached = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$
\item Iteration 2: $from^1 = \langle S+1, x^2+x\rangle, to^2= \langle T+\alpha, x^2+x\rangle,
reached = \langle T^3+1, x^2+x\rangle$
\item Iteration 3: $from^2 = \langle S+\alpha, x^2+x\rangle, to^3 = \langle T+\alpha x, x^2+x
\rangle, reached = \langle T^3\cdot x+x, T^4+T, x^2+x\rangle$
\item Iteration 4: $from^3 = \langle S, x\rangle, to^4 = \langle T+1, x\rangle, new = \langle1\rangle$
\end{itemize}
The final reachable states are represented by ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
which means $\{00,01,10,11\}$.
\end{Example}

\subsection{A simplified traversal on multivariate polynomial generators: SMPO}
% 1. How to design SMPO
% 2. Improved algorithm: why this can verify arithmetic function
% 3. Example of 3-bit SMPO
Sequential GF multiplier has no primary input, so it can be regarded as Moore FSM. 
The states are encoded by evaluations of state variables, which can be implicitly verified 
by checking output-input function ($R = A\cdot B \pmod P(\alpha)$). Additionally
the inner structure of GF multiplier is XOR-rich such that we can test our approach on very large
(>100 bits datapath) circuits.

Let us briefly describe the fundamentals behind the design of normal
basis sequential GF multipliers, so as to put in perspective the type
of designs that have been verified in this paper. Let $R =
\sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
\beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$, then 
\[
R = A\cdot B = (\sum_{i=0}^{k-1} a_i \beta^{2^{i}}) (\sum_{j=0}^{k-1}
b_j \beta^{2^{j}})  =
\sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}\nonumber 
\]

The expressions $\beta^{2^i}\beta^{2^j}$ are called cross-product
terms and they can also be represented in normal basis: 
\begin{displaymath}
\beta^{2^i}\beta^{2^j} =
\sum_{n=0}^{k-1}\lambda_{ij}^{(n)}\beta^{2^n}, \ \ \lambda_{ij}^{(n)}
\in \Ftwo. 
\end{displaymath}

From the above two equations, one can see that the expression for the
$n^{th}$ digit of product $R = (r_0, \dots, r_n, \dots r_{k-1})$ is:
\[
r_n = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(n)}a_ib_j = A
\cdot M_n \cdot B^T, ~~0 \leq n \leq k-1
\]

where $M_n = (\lambda_{ij}^{(n)})$ is a binary $k \times k$ matrix over
$\Ftwo$, and it is called the $\lambda$-matrix. 
%The collection of all
%$\{M_n\}$  $\lambda$-matrices is called the multiplication
%table. 
Moreover, let $r_n = A \cdot M_n \cdot B^T$.
Then $r_{n-1} = A \cdot M_{n-1} \cdot B^T = rotate(A) \cdot M_n
\cdot rotate(B)^T$. This implies that $M_n$ is generated by right
and down cyclic shifting $M_{n-1}$. Therefore, the hardware design
of sequential GF multipliers is based on mappings of $A\cdot M_n
\cdot B^T$ into AND-XOR gates and cyclic shift operations.

This paper verifies the implementation of two distinct 
architectures of {\it sequential multipliers with parallel output
  (SMPO)}, namely: i) the Agnew-SMPO  \cite{agnew1991implementation}
by G. B. Agnew, which is a straight-forward implementation of the
$\lambda$-matrix; and ii) the more recent, more complicated, yet very
efficient RH-SMPO \cite{RHmulti}, by Reyhani-Masoleh and Hasan,
depicted in Fig. \ref{fig:RHmulti}.

\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=2.25in]{./RH3.eps}
% \vspace{-0.2in}
\caption{A 3-bit RH-SMPO}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}

We follow  the typical sequential GF circuit model with word-level variables $A, B, R$
denoting {\it present states (PS)} and $A', B', R'$ denoting {\it next
  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
store the result, and $A, B\ (A', B')$ store input operands. {\it E.g.,}
for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
0$) are the initial values (operands) loaded into the registers,  and
$R = \F(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
result after $k$-cycles. Our approach aims to find this polynomial
representation for $R$.

Each gate in the combinational logic is represented by a Boolean
polynomial. To 
this set of Boolean polynomials, we append the polynomials that define
the word-level to bit-level relations for PS and NS variables ($A =
\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
as ideal $J = \langle 
f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
  of the circuit. The ideal of vanishing polynomials $J_0$ is also included, and
then the implicit FSM unrolling problem is setup for abstraction.

The configurations of the flip-flops are the states of the
machine. {\it Since the set of states is a finite set of points, we
can consider it as the variety of an ideal related to the circuit
}. Moreover, since we are interested in
the {\it function encoded} by the state variables (over $k$-time
frames), we can {\it project this variety} on the word-level state
variables, starting from the initial state $A_{init}, B_{init}$.
Projection of varieties (geometry) corresponds to elimination ideals
(algebra), and can be analyzed via \Grobner bases. Therefore, we
employ a \Grobner basis computation with ATO: we use a {\it lex term
  order} with {\it bit-level variables} 
$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
allows to eliminate all the bit-level variables 
%(corresponding to the combinational logic and the state variables),
%so as to 
and derives a representation only in terms of words. 
Consequently, $k$-successive \Grobner basis computations implicitly
unroll the machine, and provide word-level algebraic $k$-cycle
abstraction for $R'$ as $R' = \F(A_{init}, B_{init})$.

Algorithm
\ref{alg:modified} describes our approach.  In the algorithm, $from_i$
and $to_i$ are polynomial ideals whose varieties are the valuations of
word-level variables $R, A, B$ and $R',A',B'$ in the $i$-th iteration;
and the notation ``$\setminus$'' signifies that the $NS$ in iteration
$(i)$ becomes the $PS$ in iteration $(i+1)$. Line 5 computes the Gr\"obner 
basis with the abstraction term order.  Line 6 computes the elimination 
ideal, eliminating the bit-level variables and representing the set of 
reachable states up to iteration $i$ in terms of the elimination ideal. 
These computations are analogous to those of image computations performed in FSM reachability. 
%The forward image
%$to^{i}$ is computed using \Grobner bases with ATO.

\vspace{-0.1in}
\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
   state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
\DecMargin{1em}
\vspace{-0.1in}
\begin{Example}
\label{ex:RHSMPO}

We demonstrate our approach to verify the 3-bit RH-SMPO circuit of
Fig.\ref{fig:RHmulti}. The normal element $\beta$ in
$\mathbb{F}_{2^3}$ is known to be $\beta = \alpha^3$, where $\alpha$
is the primitive element. The circuit can be described with an ideal by translating
AND and XOR gates accordingly. For the first iteration:
\begin{align*}
J = &d_0+b_2\cdot a_2,
c_1+a_0+a_2,
c_2+b_0+b_2,
d_1+c_1\cdot c_2,\\
&e_0+d_0+d_1,
e_2+d_1,
r_0'+r_2+e_0,
r_1'+r_0,
r_2'+r_1+e_2,\\
&A+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
B+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12},\\
&R+r_0\alpha^3+r_1\alpha^6+r_2\alpha^{12},
R'+r_0'\alpha^3+r_1'\alpha^6+r_2'\alpha^{12};
\end{align*}
The last 4 polynomials are translated from the definition of word-level variables.
This represents ideal ``$J$" from line 5 in Algorithm
\ref{alg:modified}. ``$J_0$" is the ideal of vanishing polynomials in all bit-level
variables (e.g. $a_0^2-a_0$) and word-level variables (e.g. $A^8-A$). ``$from_{i-1}$"
% contains evaluation of PS word-level output in form of
% $R+\mathcal F(A,B)$. Since the definition of $A$ and $B$ changes after every iteration
% ($A',B'$ are different from $A, B$) and the multiplier's function only concerns about 
% initial value of $A,B$, we include the definition of $A_{init}, B_{init}$ into $from_i$:
represents the set of current states for iteration $i$.
In the first iteration, $from_0 = \{R, A_{init}+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
B_{init}+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12}\}$.

After the GB computation is performed with ATO, as line 6 in Algorithm \ref{alg:modified},
we find a polynomial in variables $R', A_{init}, B_{init}$ in $to_1 : 
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
$.

Line 7 in Algorithm \ref{alg:modified} simply replaces NS output $R'$ with PS output
$R$ in this example; so in second iteration $from_1 = \{
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
, A_{init}+a_2\alpha^3+a_0\alpha^6+a_1\alpha^{12},
B_{init}+b_2\alpha^3+b_0\alpha^6+b_1\alpha^{12}\}$.

% By continue executing the loop,
Finally, after 3 iterations we obtain: $to_3 = \{ \mathbf{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
as the image. The final result is $from_3(R_{final}) = R_{final}+A_{init}\cdot
B_{init}$, which verifies the multiplier. 
\end{Example}

\subsection{Limitations: complexity of reduction (multivariate division)}
Although we gain success on verifying sequential GF multipliers, there are still problems when
doing implicit state enumerating on generic FSM. For example, when we tried to apply our approach
on ISCAS'89 benchmark s953 with 29 DFFs and 311 gates, RATO spent too much time on reducing
Spoly with ideal generators. By analyzing the process of polynomial reduction, we conjecture that
\begin{Conjecture}
Long chains of AND/OR gates will make polynomial reduction infeasible.
\end{Conjecture}

\begin{figure}[hbt]
\centering{
\includegraphics[scale=0.3]{./fig_chain.eps}
\caption{An example of AND-OR-gate chain}
\label{fig:chain}}
\end{figure}

\begin{Example}
Consider a chain (Fig.\ref{fig:chain}) of 3 gates: 2 AND gates and 1 OR gate. Write them in polynomials:
\begin{align*}
&h_1:g+ef+e+f\\
&h_2:e+cd\\
&h_3:c+ab
\end{align*}
Impose RATO: $g>e>c>\{a,b,d,f\}$. Reduce a simple polynomial $g$ with this set of polynomials:
$$g\xrightarrow{h_1}_{+} ef+e+f$$
$$g\xrightarrow{h_2}_{+} cdf+cd+f$$
$$g\xrightarrow{h_3}_{+} abdf+abd+f$$
We can conclude that:
\begin{itemize}
\item When divided by polynomial from XOR gate, the length of remainder will increase by 1 term;\\
\item When divided by polynomial from AND gate, the degree of remainder will increase by 1;\\
\item When divided by polynomial from OR gate, the length and degree of remainder will increase by 1.
\end{itemize}
Since all gate polynomials have leading term with degree 1, so increase of remainder degree means
iterations of division will at least increase by 1 (while there is still possibility to cancel multiple terms
at one time). Consider the querying time, a chain with AND/OR gates will blow up computational complexity of
reduction.

Sequential GF multiplier has only limited levels of gate chains (4 for RH-multiplier). Although the size of 
datapath is large, actual complexity for reduction is relatively low.
\end{Example}

To lower down the complexity of reduction, we can introduce a matrix-based technique named as "F-4 style reduction" \cite{F4reduce}. 
It can speed up the procedure dividing a low-degree polynomial with term-sparse polynomial ideal.

Another option is to cut back the cost of each univariate division. Horner's expansion diagram (HED)\cite{alizadeh2010modular}
is a graphical method to turn operations such as uniting like terms to lower complexity DAG manipulations.

### 2015-05-01 techwriting/proposal/abs_refine.tex
all write-up from virtue machine

\section{Using Algebraic Geometry to Assist in Abstraction refinement}
\subsection{Previous work}
Traditional methods to analyze sequential circuits require state variable encoding
as well as a complete/exact state enumeration with the initial state and transition function.
With the size of modern sequential circuits increasing, it is infeasible to apply conventional methods
any longer. Abstraction is a technique to minimize the state space representation by combining states with certain 
similarities. Sometimes it can effectively lower the number of states that require analysis by orders of magnitude,
without affecting the properties we need to verify.

At first, abstraction was done manually by designers. In 2000, E. Clarke \cite{clarke2000counterexample}
proposed an automated abstraction: first,
initial abstraction is set up by dividing variables to different clusters based on transition conditions;
then, an ordinary model checking is applied on this abstraction. If a counterexample is detected, 
it is further checked to be concrete or not. For a false abstract state or a spurious path 
, the initial abstraction is refined by further splitting the false abstract state. 
A heuristic algorithm keeps refining the abstraction until a real counterexample is found or the 
property is verified. This technique is based on binary decision diagrams (BDDs), which implies a
potential risk of memory explosion; meanwhile it still requires an initial abstraction to start with.

In 2004 L. Zhang and M. Hsiao \cite{zhang2005design} proposed another abstraction method based on CNF-SAT.
This paper uses simplest latch abstraction -- by removing "irrelevant" latches from the analysis. 
The key idea is to pick those ``irrelevant" latches (variables in $k$-bounded model checking, $k$-BMC) 
according to SAT/UNSAT analysis from the $k$-BMC. It improves the $k$-BMC procedure\cite{biere1999symbolic}: 
linear temporal logic (LTL) properties to check for can be translated to a (big) CNF formula and fed to a SAT solver. 
If SAT, a concrete counterexample is found. 
If UNSAT, the original method increases the bound $k$. In Zhang's paper, the clauses for UNSAT 
(UNSAT cores\footnote{UNSAT core is a subset of the original CNF clauses which are also unsatisfiable.})
are checked for variables that do not appear.
Thus an abstracted model is obtained and checked with the original property. If the abstracted model checking fails, 
then bound $k$ is increased. This method still has disadvantages: it is counterexample-independent,
which means it cannot utilize information from invariants.

In 2005, H. Jain et al.\cite{HimanshuDAC2005} improved the abstraction refinement technique of \cite{clarke2000counterexample},
where they use CNF-SAT to perform the refinement instead of using BDDs. The new approach is applied to verify RTL Verilog
and was known to be successful.

Abstraction usually means over-approximation, i.e. properties that are true on the abstracted machine
are also valid for the original states. Craig's interpolation is a sort of abstraction which may be either
over-approximation or under-approximation. In 2003, K. McMillan used Craig's interpolation to improve 
$k$-BMC \cite{mcmillan2003interpolation}. Initially a $k$-length path to failure state $F$ is picked and transformed to 
a CNF formula, then split at first transition into a prefix and a suffix. An interpolation $P$ is computed
between the prefix and suffix. $P$ over-approximates 1-step reachable states, and under-approximates states 
that cannot reach $F$ in $k$ steps. If no counterexample is exposed, the path is split again at the second
transition; in this way a precise abstraction can be found for all reachable states.

$k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
\cite{bradley2011sat} \cite{bradley2011incremental}. IC3 works incrementally to find out inductive subclauses
of negations of reached states, meanwhile it is monolithic when computing over-approximations to sets of reachable
states within $1,2,\dots,k$ steps. It is proved to be more efficient than interpolation based model checking
although using similar mechanisms.

\subsection{Exploiting UNSAT cores for abstraction refinement}
By exploring previous work, we learn that most state-of-art abstraction refinement techniques require 
information mining from UNSAT proofs of intermediate abstractions. 
Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain that how an UNSAT proof is
utilized in such techniques.

\begin{algorithm}[hbt]
\SetAlgoNoLine
	\KwIn{$M$ is the original machine, $p$ is the property to check, $k$ is the number of steps in $k$-BMC}
  $k = $ InitValue\;
  \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT}}
  {
	\Return{``Found error trace"}
  }
  {
	Extract UNSAT proof $\mathcal P$ of $k$-BMC\;
	$M' = $ \textit{ABSTRACT}$(M,\mathcal P)$\;
  }
  \eIf{MODEL-CHECK$(M',p)$ returns \textbf{PASS}}
  {
	\Return{``Passing property"}
  }
  {
	Increase bound $k$\;
	goto Line 2\;
  }
\caption {Abstraction refinement using $k$-BMC}\label{alg:absrefine}
\end{algorithm}
\begin{figure}[hbt]
\centering{
\includegraphics[width=5.5in]{./refine.eps}
\caption{Abstraction by reducing latches}
\label{fig:refine}}
\end{figure}
Assume that we are given a sequential circuit with $n$ latches as shown in Fig.\ref{fig:refine}(a). 
This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
$l_1,\dots,l_n$. Algorithm \ref{alg:absrefine} describes an approach to check if machine $M$ violates property $p$.
This algorithm relies on $k$-BMC technique, which works on the basis of CNF-SAT solving.
The $k$-BMC represents the initial states $I$, the transition relation $T$ and property $p$ as CNF formulas.

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is $SAT$ or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
$SAT$, then a counterexample is found that violates property $p$. If the result is $UNSAT$, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fixpoint detected.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is $UNSAT$.
If there are some latch variables not included in this UNSAT core (denoted by $L_{abs}$), then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Fig.\ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

The second ``if-else" branch means: if we do an ordinary model checking on the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason of this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an over-approximation
of $M$, such that $(M'\implies p) \implies (M\implies p)$. If we find a violation on abstracted machine, then this
abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which will encounter the ``bit-blasting" problem
when the size of datapath (number of latches in Fig.\ref{fig:refine}) is very large. \textbf{Here we propose a totally new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation based on our experience:}

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While computing GB over finite fields is exponential in the number of variables, GB  computation 
is observed to be more efficient for UNSAT problems.} 
The reason is discussed as follows:
\begin{Theorem}
{\bf Weak Nullstellensatz:} Given ideal $J\subset \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow reducedGB(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high
time complexity and is usually not practicable. One reason is that the size of GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is 1, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in non-reduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

\subsection{A demonstration of our conjecture}
One  important research topic about UNSAT problems is to identify UNSAT cores efficiently.
An UNSAT core in a CNF formula denotes a subset of clauses which is still unsatisfiable. Here
we re-define this concept in algebraic geometry:
\begin{Definition}
Assume a set of polynomials $F$ and its subset $F_s\subset F$. If ${\bf V}(\langle F\rangle) = {\bf V}
(\langle F_s\rangle) = \emptyset$, we call $F_s$ an {\bf UNSAT core} of $F$. Additionally, if
$F_s$ contains no other UNSAT core, we call it a {\bf minimal} UNSAT core.
\end{Definition}

We  conjecture that based on observation of Buchberger's algorithm's execution, an UNSAT core can be identified.
\begin{Conjecture}
Buchberger's algorithm picks pairs of polynomials from a given set, computes their S-poly, then reduces this S-poly
with the given set of polynomials. If the remainder is non-zero,
it is added to the set of polynomials.
By tracking S-poly computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.
\end{Conjecture}
\begin{Example}
A SAT problem is described with 8 CNF clauses:

\begin{minipage}[h]{0.3\textwidth}
\begin{align*}
&c_1: \bar{a}\lor\bar{b}\\
&c_2: a\lor\bar{b}\\
&c_3: \bar{a}\lor b\\
&c_4: a\lor b
\end{align*}
\end{minipage}
\begin{minipage}[h]{0.7\textwidth}
\begin{align*}
&c_5: x\lor y\\
&c_6: y\lor z\\
&c_7: b\lor \neg y\\
&c_8: a\lor x\lor \neg z
\end{align*}
\end{minipage}

% \begin{align*}
% &\bar{a}\lor\bar{b}\\
% &a\lor\bar{b}\\
% &\bar{a}\lor b\\
% &a\lor b\\
% &x\lor y\\
% &y\lor z\\
% &b\lor \neg y\\
% &a\lor x\lor \neg z
% \end{align*}
Using Boolean to polynomial mappings given in Table \ref{tab:booltof4}, we can transform them to a set of
polynomials $F$ over ring $\mathbb F_2[a,b,x,y,z]$:

\begin{minipage}[h]{0.4\textwidth}
\begin{align*}
&f_1:ab\\
&f_2:ab+a\\
&f_3:ab+b\\
&f_4:ab+a+b+1
\end{align*}
\end{minipage}
\begin{minipage}[h]{0.6\textwidth}
\begin{align*}
&f_5:xy+y+x+1\\
&f_6:yz+y+z+1\\
&f_7:by+y\\
&f_8:axz+az+xz+z
\end{align*}
\end{minipage}
% \begin{align*}
% &f_1:ab\\
% &f_2:ab+a\\
% &f_3:ab+b\\
% &f_4:ab+a+b+1\\
% &f_5:xy+y+x+1\\
% &f_6:yz+y+z+1\\
% &f_7:by+y\\
% &f_8:axz+az+xz+z
% \end{align*}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Fig.\ref{fig:UNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Fig.\ref{fig:UNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of unsatisfiability of original set of clauses.
\end{enumerate}
\end{Example}
\begin{figure}[hbt]
\centering{
\includegraphics[width=4.5in]{./UNSAT.eps}
\caption{DAG representing Spoly computations and multivariate divisions}
\label{fig:UNSAT}}
\end{figure}
% We conclude our approach as follows: i) Keep track of all $Spoly(f_i,
% f_j)\xrightarrow{f_1, \dots, f_s}_+ r$; ii) mark $f_i, f_j$ if $r\neq
% 0$; iii) Mark those $f_l$'s that have leading terms that cancel
% monomial terms in $Spoly(f_i, f_j)$; iv) Build a DAG recording all reductions including Spoly computations
% and multivariate divisions; v) Once $1\in J$ is
% detected, traverse the graph to identify all the marked polynomials that were
% used in obtaining this unit element. These polynomials constitute the
% core.

We conclude our approach as a conjecture algorithm (Algorithm \ref{alg:UNSAT}).

\begin{algorithm}[hbt]
\SetAlgoNoLine
	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
	\KwOut{An UNSAT core $\{f_{m_1},f_{m_2},\dots,f_{m_t}\}$}
\Repeat{$r_l == 1$}
{
	Pick a pair $f_i,f_j\in F$ that has never been computed S-poly\;
	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
	{
		$F = F\cup r_l$\;
		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the S-poly, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the S-poly\;
	}
}
Backward traverse the DAG for remainder ``1", replace $r_l$ with corresponding DAG $G_l$\;
\Return{All roots}
\caption {Extract UNSAT core using a variation of Buchberger's algorithm}\label{alg:UNSAT}
\end{algorithm}

\subsection{Research objective}
i) To prove that the algorithm is concrete and complete: it can always abstract an UNSAT core;

ii) To design an algorithm to find a minimal UNSAT core;

iii) Run the algorithm at word level, to find UNSAT proofs for word-level variables in $\Fkk$, which 
can help word-level abstraction refinement procedure.

% If we start with other term orderings, we may get a larger UNSAT core. By dynamically adjusting
% term ordering and re-calculate GB, we can get new UNSAT core with smaller size, until
% we find minimal UNSAT core or expire runtime limit.

### 2015-05-01 techwriting/proposal/contribute.tex
all write-up from virtue machine

\section{Proposed Objectives}
\begin{itemize}
\item[-] Explore an implementation of algebraic geometry based reachability analysis for sequential circuits verification;
\item[-] Currently our approach is implemented within the {\sc Singular} tool, which is not efficient enough to execute our
proposed algorithms, because its data-structures are not efficient for circuit verification problems. We plan to design a standalone CAD tool implemented in C++, which specifically aims to solve word-level sequential
verification problems. We will borrow techniques from \cite{timDAC}, \cite{jinpeng} and \cite{f4reduce}, and tailor them to sequential circuit analysis, to lower the computational
complexity;
\item[-] Fine-tune the tool for sequential Galois field arithmetic circuits verification, because arithmetic circuit
verification greatly benefits from the abstraction of word-level operands in the datapath;
\item[-] Explore a new abstraction-refinement paradigm, based on information from UNSAT cores, which is also 
attained using a Gr\"obner basis approach.
\end{itemize}

\section{Proposed Timetable}
\begin{itemize}
\item[-] Current status: Experiments have been performed to run implicit state enumeration on sequential circuits benchmarks
such as ISCAS'89 circuits. Current problem is the algorithm spends too much time on multivariate division procedure;
\item[-] Spring 2015: Implement the tool which can efficiently do multivariate division and test the performance of 
our tool based on T. Pruss et al's \cite{timDAC} approach;
\item[-] Summer 2015: Integrate the refined multivariate division routine into our verification tool, test its
performance on circuits with various sizes;
\item[-] Fall 2015: Evaluate data and write the dissertation.
\end{itemize}

### 2015-05-01 techwriting/proposal/gb_sat.tex
all write-up from virtue machine

\section{Algebraic Geometry Concepts and Boolean Satisfiability}
\subsection{History}
Boolean satisfiability (SAT) problem is one of most basic problems in computer science.
It is the essential problem for a large number of circuit testing and verification problems.
Fig.\ref{fig:SAT} shows a simple example of SAT problem on circuits verification:
we need to verify whether subcircuit $A$ and subcircuit $B$ has the same function, so
we build a miter circuit for their outputs $X$ and $Y$, and the equivalence checking 
problem is turned into a SAT problem as follows:
\begin{align*}
&Is\ subcircuit\ A\ functionally\ equivalent\ to\ subcircuit\ B?\\
&\Longleftrightarrow
Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ primary\ inputs\ a,b,c\ such\ that\ Z=1?
\end{align*}

\begin{figure}[hbt]
\centering{
\includegraphics[scale=0.3]{./fig_SAT.eps}
\caption{An example of Boolean satisfiability problem on circuits}
\label{fig:SAT}}
\end{figure}

In most cases, SAT problems are modeled as CNF formulas, and solved by procedures based on 
Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of DPLL algorithm
is recursive branching and backtrack searching. To improve the efficiency of DPLL-based
SAT solver, efforts are made on minimizing number of branches, accelerating unit propagation
and modifying backtrack algorithm. Recently state-of-art SAT solvers developed conflict-driven
clause learning (CDCL) technique to prune the search space, which is effective to reduce
result search time.

On the other hand, people are seeking an alternative solution for SAT problem using a totally
different method from ``old-fashioned" DPLL algorithm. One promising option is polynomial calculus (PC),
mapping Boolean variables and connectors in CNF formulas to variables and operators in 
Galois fields. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
valid assignments or proof of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
from computer algebra.

The inspiration that use PC to solve SAT problems first comes from \cite{ceiSTOC96}.
Using PC and its variations, researchers develop many SAT solvers \cite{STABLE} \cite{BLUEVERI} \cite{PolyBoRi} 
Besides, people borrow concepts from PC, combine them with traditional DPLL and clause learning techniques
to build hybrid SAT solvers. \cite{condratTACAS07} \cite{Zengler2010}
\subsection{Define SAT problem in computer algebra}
\label{sec:mapping}
In this part, we formally define SAT problems in language of computer algebra by mapping CNF formulas to polynomials.
First, Boolean variables are mapped to field/ring elements, so it is necessary to introduce the concept of field
and ring.
\begin{Definition}
A {\bf ring} is a set with binary operators ``$+$" and ``$\cdot$" satisfying:\\
(1.a) ``$+$" is associative and commutative;\\
(1.b) There exist element $0$ as additive identity, and every element has additive inverse;\\
(2.a) ``$\cdot$" is associative;\\
(2.b) There exist element $1$ as multiplicative identity;\\
(3) Multiplication distributes over addition.
\par
A {\bf field} is a ring satisfying:\\
(1) every nonzero element has multiplicative inverse;\\
(2) ``$\cdot$" is commutative.
\par
A field containing finite number of elements is called {\bf finite field}, or {\bf Galois field}.
\end{Definition}
\begin{Example}
Consider the set of all integers $\mathbb Z$ and usual definition of addition ``$+$" and multiplication ``$\cdot$".
Then $\mathbb Z$ is a ring. But it is not a field because some integers do not have multiplicative inverse: 
$\frac{1}{2} \notin \mathbb Z$. The set of all rational numbers $\mathbb Q$ is a field, but not a Galois
field since it has infinite number of elements. An example of Galois field is the set of integers modulo a prime
($\mathbb Z/n\mathbb Z,\ n$ is prime), e.g. Galois filed $\mathbb F_5$ can be interpreted as set of 5 integers
$\{0,1,2,3,4\}$, and addition/multiplication defined as $a+b\pmod 5,\ a\cdot b\pmod 5$.
\end{Example}
In some cases a field $\mathbb F$ can be extended to a ring. If we take indeterminates $x_1,x_2,\dots,x_n$ 
(usually not from field $\mathbb F$), an arbitrary combination of their finite product 
$x_1^{d_1}\cdot x_2^{d_2}\cdots x_n^{d_n}$ is a {\bf monomial}. A {\bf polynomial} $f = c_1 X_1 +
c_2 X_2 + \dots + c_t X_t$ is a finite sum of terms, where $c_1,
\dots, c_t$ are coefficients and $X_1, \dots, X_t$ are monomials.
These polynomials can form a ring, if coefficients come from elements of $\mathbb F$, we call this ring as
a {\bf multivariate polynomial ring} $\mathbb F[x_1,\dots,x_n]$.
\begin{Example}
\label{ex:polyring}
Ring $\mathbb Z[a,b,c]$ contains elements such as $4a^3b+7ab^2c+2c^2$, and ring $\mathbb F_2[a,b,c]$
contains elements such as $a+bc+1$, which can be mapped to a Boolean CNF formula!
\end{Example}
$k$-variable CNF clauses for SAT are defined over Boolean ring $\mathbb B^k$. We can define a unique 
mapping $\mathbb B^k \to \mathbb F_2[x_1,\dots,x_k]$ to transform a CNF formula to polynomial,
including:
\begin{align*}
T &\to 1\\
F &\to 0\\
\land &\to \cdot\\
\oplus &\to +\\
\text{Boolean variables} &\to \text{ field } \mathbb F_2 \text{ elements }
\end{align*}
Namely we can represent Boolean connectors $\neg,\ \bar{\oplus},\ \lor$ as:
\begin{align*}
&\neg a \to 1+a\\
&a \bar{\oplus} b \to 1+a+b\\
&a \lor b \to a+b+a\cdot b
\end{align*}
Using these mappings it is easy to write CNF clauses to polynomials accordingly.
The most straightforward way to solve SAT problem in form of polynomials is to create
equations that each polynomial equals to 1, then solve the system of equations.
Solutions to the system is defined as affine varieties:
\begin{Definition}
Given a set of polynomials $f_1,\dots,f_k$ over ring $\mathbb F_q[x_1,\dots,x_n]$, their 
{\bf affine variety} 
$$V(f_1,\dots,f_k) = \{(a_1,\dots,a_n)\in \mathbb F_{q^n} | f_1(a_1,\dots,a_n) = \cdots = f_k(a_1,\dots,a_n) = 0\}$$
\end{Definition}
Different sets of polynomials may have same variety. There is a method to find out the set of
all polynomials with the same variety, which is 
\begin{Definition}
{\bf Ideal of Polynomials:} Let $f_1,f_2,\dots,f_s\in \mathbb F[x_1,\dots,x_n]$.
Define an ideal
$$J = \langle f_1,f_2,\dots,f_s\rangle = \{f_1\cdot h_1 + f_2\cdot h_2 +\cdots + f_s\cdot h_s : h_1,\dots,h_s\in \mathbb F[x_1,\dots,x_n]\}$$
We call $J = \langle f_1,f_2,\dots,f_s\rangle$ an ideal generated by $f_1,\dots,f_s$ and these polynomials 
the {\bf generators} of ideal $J$.
\end{Definition}
On the other hand, if given another polynomial $f$, we need to judge whether it belongs to $J$.
\begin{Definition}
{\bf Ideal membership:} Let $f_1,f_2,\dots,f_s\in \mathbb F[x_1,\dots,x_n]$, and $J = \langle f_1,f_2,\dots,f_s\rangle$
be an ideal over ring $\mathbb F[x_1,\dots,x_n]$. If 
$$f = f_1h_1 + f_2h_2 + \cdots + f_sh_s$$
then $f\in J$.
\end{Definition}
We can easily deduce varieties of $J$ will also vanish on polynomial $f$, i.e. if 
$f_1({\bf a}) = f_2({\bf a}) = \cdots = f_s({\bf a}) = 0$, then $f({\bf a}) = 0$.
\subsection{The structure of affine variety and algebraic Geometry}
The structure of affine variety is straightforward when analyzing from geometry view. Consider
ring $\mathbb F[x_1,\dots,x_n]$, it can be interpreted as a $n$-dimension space, and
coordinates are elements from field $\mathbb F$. Affine varieties on this ring are
sets of dots in the space.
\begin{Example}
(Fig.\ref{fig:variety}) Consider ring $\mathbb R[x,y]$, it is a plane with Cartesian coordinates $(x,y)$. Variety of a single-generator 
ideal $J = \langle x^2+y^2-1\rangle$ is all points on center $(0,0)$ radius $1$ circle.
$$V_{\mathbb R}(\langle x^2+y^2-1\rangle) = \{(x,y)|x^2+y^2=1,\ x,y\in \mathbb R\}$$
Meanwhile variety depends on the ring where ideal locates, e.g. for ring $\mathbb R[x]$
$$V_{\mathbb R}(\langle x^2+1\rangle) = \emptyset$$
while for ring allowing complex numbers $\mathbb C[x]$
$$V_{\mathbb C}(\langle x^2+1\rangle) = \{\pm i\}$$
\end{Example}

\begin{figure}[hbt]
\centering{
\includegraphics[scale=0.5]{./fig_variety.eps}
\caption{Examples of affine varieties in affine space}
\label{fig:variety}}
\end{figure}
Applying set manipulations such as {\bf union}, {\bf intersection} and {\bf complement} to varieties 
and ideals is an important topic to explore. First, we introduce following concepts from algebraic geometry:
\begin{Definition}
\label{def:sum}
({\bf Sum of Ideals}) If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then the 
{\bf sum} of $I$ and $J$, denoted by $I + J$, is the set
  \begin{equation}
  I + J = \{f + g\ |\ f \in I \ and\  g \in J\}.
  \end{equation}
Furthermore, if $I = \langle f_1, \dots, f_r\rangle$ and 
$J = \langle g_1, \dots, g_s\rangle$, then 
$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$.
\end{Definition}
\begin{Definition}
\label{def:prod}
({\bf Product of Ideals}) If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then the
{\bf product} of $I$ and $J$, denoted by $I \cdot J$, is defined to be the ideal generated 
by all polynomials $f \cdot g$ where $f \in I$ and $g \in J$. Furthermore, let
$I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$, then
  \begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle .
  \end{equation}
\end{Definition}
\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then $I:J$
is the set
  \begin{equation}
  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}
These concepts can lead the way to solution by adopting following theorems:
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
\bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) = {\bf V}(I) \bigcup {\bf V}(J)$.
\end{Theorem}
However, complement of variety cannot be easily dealt by simple arithmetic on polynomial generators.
For non-trivial cases we can only prove that 
\begin{Theorem}
Let $I, J$ be ideals in $\mathbb F[x_1,\dots,x_n]$, then 
$${\bf V}(I:J) \supset {\bf V}({\bf I}({\bf V}(I) - {\bf V}(J)))$$
\end{Theorem}

Fortunately for most hardware verification cases, the form of polynomial generators are restricted.
A proposition has been proved by \cite{jinpeng} that after adding {\bf vanishing polynomials ideal}
the new composed ideal is {\bf radical}, which implies following corollary can be employed:
\begin{Corollary}
\label{thm:quotient}
Let $I, J$ be radical ideals in $\mathbb F[x_1,\dots,x_n]$, then 
$${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$.
\end{Corollary}
The variety of vanishing polynomial ideal contains all possible evaluations of variables, so it
is the {\bf universal set}. Subsequently, the {\bf absolute complement} can be written as
$$\overline{{\bf V}(J)} = {\bf V}(J_0:J)$$

\subsection{Gr\"obner basis and UNSAT}
Different generators may generate the same ideal, which means we can equivalently transform one set of generators
to another for computational or theoretical convenience.
Some generators are "optimal" representation of the ideal, such as {\bf Gr\"obner basis}.

Gr\"obner basis contains generators whose leading monomials can divide the leading monomials of all polynomials 
in the ideal:
\begin{Definition}
A set of polynomials $G = \{g_1,\dots,g_t\}$ is the Gr\"obner basis (GB) of ideal $J$ if and only if
$$\forall f\in J, \exists g_i,\ s.t.\ LM(g_i)\mid LM(f)$$
\end{Definition}
The advantage of representing an ideal with GB is that it can serve as decision procedure for ideal membership
test when dividing polynomial by GB, i.e.
$$G = GB(J) \Longleftrightarrow \forall f\in J, f\xrightarrow{g_1,g_2,\dots,g_t}_{+} 0$$
Gr\"obner basis can be reduced by eliminating redundant elements. A reduced GB is a canonical representation of 
the ideal under some monomial orderings.

When an ideal has empty variety, its reduce GB has a very special form:
\begin{Theorem}
{\bf Weak Nullstellensatz:} Given ideal $J\subset \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure
of field $\mathbb F$ is null if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow reducedGB(J) = \{1\}$$
\end{Theorem}

It is well learned that using Buchberger's algorithm and its variations to compute a GB has a very high
time complexity and is usually not practicable. One reason is that the size of GB may explode even the term
ordering is carefully chosen. However if the reduced GB is 1, which means every term in original polynomials
will be canceled, the degree of newly added GB will be limited. Thus the size of GB is less possible to grow 
too fast. Instead of applying polynomial calculus to SAT solving, it may be more efficient to try UNSAT
problems.

One of most important research topics about UNSAT problem is to find out UNSAT core efficiently.
\begin{Definition}
Assume a set of polynomials $F$ and its subset $F_s\subset F$. If ${\bf V}(\langle F\rangle) = {\bf V}
(\langle F_s\rangle) = \emptyset$, we call $F_s$ an {\bf UNSAT core} of $F$. Additionally, if
$F_s$ has no UNSAT core of itself, we call it a {\bf minimal} UNSAT core.
\end{Definition}

We propose a conjecture based on observation of Buchberger's algorithm's execution.
\begin{Conjecture}
By tracking s-polynomial computations and multivariate divisions that lead to remainder (newly added GB)
1, we can obtain an UNSAT core; we can figure out a minimal UNSAT core at high probability with limited
iterations of execution of Buchberger's algorithm.
\end{Conjecture}
\begin{Example}
A SAT problem is described with 8 CNF clauses:
\begin{align*}
&\bar{a}\lor\bar{b}\\
&a\lor\bar{b}\\
&\bar{a}\lor b\\
&a\lor b\\
&x\lor y\\
&y\lor z\\
&b\lor \neg y\\
&a\lor x\lor \neg z
\end{align*}
Using mappings mentioned in section \ref{sec:mapping}, we can transform them to
polynomials over ring $\mathbb F_2[a,b,x,y,z]$:
\begin{align*}
&f_1:ab\\
&f_2:ab+a\\
&f_3:ab+b\\
&f_4:ab+a+b+1\\
&f_5:xy+y+x+1\\
&f_6:yz+y+z+1\\
&f_7:by+y\\
&f_8:axz+az+xz+z
\end{align*}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
First we compute $Spoly(f_1,f_2)\xrightarrow{\ }_{+} r_1$, remainder $r_1$ equals to $a$;
Next we compute $Spoly(f_1,f_3)\xrightarrow{\ }_{+} r_2$, remainder $r_2$ equals to $b$;
Then we compute $Spoly(f_1,f_4) = a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ and $r_2$, and they can be backtracked to $f_1,f_2,f_3$. The remainder is ``1", so we
conclude that $f_1,f_2,f_3,f_4$ is an UNSAT core for this problem.
\end{Example}
If we start with other term orderings, we may get a larger UNSAT core. By dynamically adjust
term ordering and re-calculate GB, we can get new UNSAT core with smaller size, until
we find minimal UNSAT core or expire runtime limit. (Need more experiments)

### 2015-05-01 techwriting/proposal/intro.tex
all write-up from virtue machine

\section{Introduction}
Verification of sequential circuits is challenging, and therefore a topic of continuous research. With
the increasing size of new 
integrated circuits, sequential circuit designers face much more complicated problems of design errors in
specification models and implementations. These errors are usually modeled as ``bad" states,
and the circuits/functional components are modeled as state machines. Once state reachability is analyzed,
the existence of errors can be identified by judging whether ``bad" states are reachable from certain initial states.
Reachability analysis can be performed by various techniques such as extensive simulation, temporal logic model checking
and property directed reachability (PDR). In this proposal we discuss a new promising approach that applies implicitly on the
word level and borrows concepts from commutative algebra and algebraic geometry.

There are several advantages in exploiting word-level verification: on the one hand, a number of circuit designs have 
their datapaths and/or system-level models described  at the word-level, which include many register transfer level 
(RTL) descriptions,
arithmetic components and cryptographic systems. Using word-level information to do verification is convenient 
and beneficial for interpreting the circuit's function. On the other hand, using word-level instead of bit-level
information is one way of ``abstraction". Abstraction is a key technique to reduce the state space of a finite state machine 
when analyzing a sequential circuit. It is usually done by combining sets of states with similar properties.
During state reachability analysis, if we use bit-level variables to  represent the states, 
the representations may become too large to handle (e.g. BDD explosion). However, when a ``bundle" of bit-level variables are represented as only one
word-level variable, the set of reachable states can be represented by a  word-level constraint 
expression. Word-level 
verification techniques are effective in lowering the complexity of state reachability analysis.

Algebraic geometry can provide a symbolic representation of both bit-level (Boolean) variables and 
word-level variables \textit{in one unified framework}. The basic idea is that values of state variables can be represented as solutions to a 
finite set of polynomials. Gr\"obner basis (GB) techniques can be subsequently
applied to transform such polynomial systems into a unique canonical form, thus facilitating verification.
To represent Boolean circuits at word level, polynomial functions over Galois fields provide a {\it bit-precise
word-level model}.
%bit, word ,\Fkk
Particularly, arbitrary word-level datapaths can be modeled as polynomial
functions over $\Fkk$ -- i.e. the Galois field of $2^k$ elements, where $k$ represents the bit-vector word length. In this fashion,
%\footnote{currently no theory can guarantee unique one-to-one mapping for polynomials 
%from $\mathbb B^k$ to $\mathbb Z_{2^k}$ yet}.
%Need introduce \Fkk first
the states and transition relations can be represented as polynomial functions over $\Fkk$ and the power of algebraic geometry can be applied to
implement symbolic reasoning about state reachability. This kind of implicit state
analysis and abstraction can overcome scalability issues encountered by explicit bit-blasting.
Recent work of T. Pruss \cite{timDAC}, J. Lv \cite{jinpeng}, A. Lvov \cite{BLUEVERI} and M. Brickenstein \cite{PolyBoRi}
shows that it is possible to make application of algebraic geometry practical for circuits.

\textbf{This proposal investigates a wide gamut of symbolic computation and algebraic geometry techniques for reachability analysis, abstraction and
verification of sequential circuits applied at the word level.}

%Above example shows how algebraic geometry can be applied to induce one-step reachability. 
% : every function on $k$-bit Boolean vector $f:\mathbb B^k\to \mathbb B^k$ 
% can be uniquely mapped to a polynomial function about $k$-bit word from Galois field $\mathcal F:\Fkk \to \Fkk$,
% this word represents a word-level state variable. Thus, the state space with explicit Boolean bits encoding
% can be uniquely mapped to solution (variety) of a set of polynomials (ideal). Then other concepts from algebraic
% geometry are introduced to solve the reachability problem.

The essence of our approach relies on  bit-level to word-level abstraction.
Any Boolean function over $k$-bit vectors $f:\mathbb B^k\to \mathbb B^k$ 
can be uniquely mapped to a polynomial function over $k$-bit words in the Galois field $f:\Fkk \to \Fkk$ \cite{timDAC}. 
Therefore, the transition relation of sequential circuits can be modeled at word level. The set of
states (Boolean encoding) can be represented as (word-level) elements over Galois fields. Verification
requires operations such as image computations, set intersections, unions and complements, unsatisfiability 
(UNSAT) proofs and analysis, etc. The challenge is to discover efficient algorithmic techniques to perform all such 
operations at the word level. While this challenging task is the topic of this research, we
demonstrate the feasibility of this approach in the example below.

\begin{figure}[hbt]
\centering{
\includegraphics[width=5.5in]{./fig_fsm.eps}
\caption{2-bit FSM example circuit}
\label{fig:fsm}}
\end{figure}

\begin{Example}
\label{ex:motiv}
{\bf Motivating example:} Consider Fig.\ref{fig:fsm}(a), a Mealy finite state machine (FSM) with
4 states encoded by a 2-bit Boolean vector, and Fig.\ref{fig:fsm}(b) as its gate-level circuit implementation.
Variable $x$ is the primary input (PI), $\{s_0,s_1\}$ are present state (PS) variables, $Z$ is primary
output (PO) and $\{t_0, t_1\}$ are next state (NS) variables. The 4 states are encoded as
$\{00,01,10,11\}$, which can be interpreted as elements in $\mathbb F_4$ according to Table \ref{tab:booltof4}.
We can also translate all gates (Boolean operators) to polynomial functions over $\mathbb F_4$.
\vspace{-0.1in}
\begin{table}[htb]
\centering
\caption{Mapping bit-vector state encoding to elements in $\mathbb F_4$, Boolean operators to functions over $\mathbb F_4$}
\begin{tabular}{|c|c||c|c|} 
\hline
Bit-vector & Element in $\mathbb F_4$ & Boolean operator & Function over $\mathbb F_4$ \\
\hline
\hline
00 & 0 & $a\land b$ & $a\cdot b$ \\
\hline
01 & 1 & $a\oplus b$ & $a+b$ \\
\hline
10 & $\alpha$ & $\bar{a}$ & $1+a$ \\
\hline
11 & $1+\alpha$ & $a\lor b$ & $a+b+a\cdot b$\\
\hline
\end{tabular}
\label{tab:booltof4}  
\end{table}

We first obtain polynomials $f_1$ and $f_2$, denoting transition relations for NS $\{t_0,t_1\}$:
\begin{displaymath}
\begin{array}{ll}
f_1:& t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1) \\
f_2:& t_1 - (xs_0+x+s_0s_1+s_0)\
\end{array}
\end{displaymath}
Since we model reachability at word level, let $S = \{s_0,s_1\}$ represent a word-level PS variable in $\mathbb F_4$,
its relation to bit-level variables is $S = s_0 + s_1\cdot \alpha$, where $\alpha$ is the primitive element
of $\mathbb F_4$. Similarly we can define a word-level NS variable $T$. They can be both interpreted in polynomial form:
\begin{displaymath}
\begin{array}{ll}
f_3:& S - s_0 - s_1\alpha \\
f_4:& T - t_0 - t_1\alpha
\end{array}
\end{displaymath}
The initial state can also be represented in polynomial form. Assume the initial state is $\{00\}$ ($0$ in $\mathbb F_4$),
it corresponds to the solution of the polynomial equation $(S-0) = 0$. So its polynomial form is
\begin{displaymath}
f_5:\  S
\end{displaymath}
We can then use the Gr\"obner basis (GB) computation to compute a canonical form of this system of polynomials
$f_1,\dots, f_5$ representing possible value of NS variable $T$. Note that
if we eliminate the PI variable $x$ from the ideal $\langle f_1,\dots,f_5\rangle$ while computing GB, the result
implies an existential quantification on $x$.

After computing GB with a lexicographic term order $\{x,s_0,s_1,t_0,t_1\} > S > T$, the result contains a polynomial
only in the word-level variable $T$: ${\bf g_T:T^2+(\alpha+1)T+\alpha}$. 
Notice that $g_T$ is a polynomial in word-level variable $T$, it has 2 roots which encode 2 states of the machine  $T = 1(01)\ and 
\ T = \alpha(10)$.
\end{Example}
% \begin{Example}
% \label{ex:motiv}
% {\bf Motivating example:} Fig.\ref{fig:fsm}(a) is a simple 2-bit sequential circuit, it can be modeled
% as a Mealy finite state machine (FSM). $x$ is primary input (one bit),
% $\{s_0, s_1\}$ are state inputs (present state variables), $Z$ is primary output and $\{t_0, t_1\}$ are 
% state outputs (next state variables). Word-level state input (2-bit vector) is $S$, output is $T$.
% Fig.\ref{fig:fsm}(b) shows its state transition graph (STG), states are encoded explicitly with
% present state variables $\{s_0,s_1\}$ taking values from $\{00,01,10,11\}$. Using one-to-one mapping
% $\mathbb B^k \to \Fkk$, we can write word $S$ as element in $\mathbb F_4$: $S = s_0 + s_1\cdot\alpha$, 
% and $T = t_0 + t_1\cdot\alpha$. Here $P(\alpha) = \alpha^2+\alpha+1 = 0$ is the primitive polynomial
% of $\mathbb F_4$.
% 
% We can translate all gates (Boolean operators) to polynomial functions and compose an ideal for this circuit.
% First, transition functions at bit level:
% \begin{displaymath}
% \begin{array}{ll}
% f_1:& t_0- (xs_0^2\cdot s_1^2+xs_0^2\cdot s_1+xs_0\cdot s_1^2+xs_0+xs_1+x+s_0^2s_1^2+s_0^2s_1+s_0s_1^2+s_0s_1+s_0+s_1+1) \\
% f_2:& t_1 - (xs_0^2\cdot s_1+xs_0^2+xs_0\cdot s_1+x+s_0s_1+s0)\
% \end{array}
% \end{displaymath}
% Followed by bits-to-word definitions using polynomial representation of elements in $\mathbb{F}_4$:
% \begin{displaymath}
% \begin{array}{ll}
% f_3:& S - s_0 - s_1\alpha \\
% f_4:& T - t_0 - t_1\alpha
% \end{array}
% \end{displaymath}
% And an polynomial about initial state ($\{00\}$, i.e. $0$ in $\mathbb F_4$):
% \begin{displaymath}
% f_5:\  S
% \end{displaymath}
% And the rests are vanishing polynomials for each variable:
% $f_6: x^2 - x; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: s_0^2 - s_0;
% f_{11}: s_1^2 - s_1; f_{12}: T^4 - T$
% 
% Note that there is no specification on initial primary input $x$, in Gr\"obner basis (GB) approach this means $x$ is smoothed by
% reversely using Shannon's expansion.
% 
% We use elimination term order: $intermediate\ bit$-$level\ signals\ >\ bit$-$level\ primary\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
% $J = \langle f_1, f_2, \dots, f_{12}\rangle $ to compute GB, the result will include one polynomial $f_T$ contains only variable $T$. In this example,
% GB we get is: 
% $g_1:S;
% g_2:T+t_0+t_1\alpha;
% g_3:S+s_0+s_1\alpha;
% g_4:s_1;
% g_5:t_1+T^2+T;
% g_6:x+T^2+T;
% {\bf g_7:T^2+(\alpha+1)T+\alpha}$.
% Note the last polynomial equals to $T^2-(\alpha+1)T+\alpha$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
% limited within $\mathbb{F}_{2}$, where $x \equiv -x\ (mod\ 2)$ for any element $x$ in the field. 
% Solution to $g_7 = 0$ is $T = 1\ or \ T = \alpha$, which shows that next
% state the machine just reached is $\{S_1(01), S_2(10)\}$.
% \end{Example}

The above example shows how algebraic geometry can be applied to induce one-step reachability at the word level. 
It is possible to combine our approach in this example with other techniques inspired by concepts and theorems
from algebraic geometry to help reachability analysis for FSMs, as well as further formal verification problems
on sequential circuits.

{\bf Contributions of the dissertation:} Our approach to solve the sequential circuit verification problem and contribution
of the thesis can be described as follows:
\begin{itemize}
\item[-] This proposal will investigate techniques to perform sequential circuit verification at the word level
using polynomial abstractions of Boolean circuits over Galois fields. The applications considered are exact
reachability analysis and property directed reachability \cite{bradley2011sat} \cite{bradley2011incremental}. Both sequential datapath circuits
as well as random logic controllers will be considered. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\item[-] Using polynomial abstractions, we analyze the circuit, and model the gate-level Boolean operators as elements of a
%multivariate polynomial ring with coefficients in $\Fkk$.
\item[-] Algebraic geometry based techniques will be researched for reachability analysis. 
The concepts of Nullstellensatz, Gr\"obner bases and elimination ideals will be explored for 
the purpose of image computations. The canonical representations of polynomial ideals provided
by Gr\"obner bases will be exploited for verification.
%deduce that the one-step reachability problem can be formulated as one of computing a Gr\"obner basis of the set of
%polynomials derived from the given circuit netlist, using a specific elimination term order.
\item[-] Computing Gr\"obner bases for large circuits is infeasible due to its high computational complexity.
Therefore, based on inspirations from \cite{timDAC} and \cite{jinpeng}, a significant part of this work will focus on improving the scalability of our approach.
%\item[-] Apply our approach for one-step reachability to a breath-first search (BFS) state space traversal
%algorithm, the reachability of system can be analyzed when reaching a fixpoint.
%\item[-] Computing Gr\"obner bases using elimination term orders is infeasible for large circuits. 
%To overcome this limitation we introduced improved approach using new term order and variable substitution, 
%as well as F-4 algorithm to accelerate multivariate division.
\item[-] New paradigms for abstraction refinement using algebraic geometry will be explored.
Such techniques require analysis of UNSAT proofs of the verification instances.
UNSAT cores in polynomial calculus have not been researched before. In this proposal,
we will investigate a new technique to identify minimal UNSAT cores using the Gr\"obner basis algorithm 
and apply it to verification.
%and we propose a new
%GB based technique to compute UNSAT core to assist abstraction refinement procedure.
\end{itemize}
%In this proposal, I will show that conceptually our proposed approach is feasible and we will research
%efficient symbolic computing techniques for sequential verification based on algebraic geometry.
All in all, this proposal aims to discover an altogether new paradigm for formal verification/model checking
of sequential circuits at the word level using a combination of design automation, algebraic geometry and 
symbolic computation techniques. Bit-level Gr\"obner basis computation has been applied to model 
checking \cite{avrunin1996symbolic} \cite{tran2007groebner} and satisfiability problems \cite{PolyBoRi},
but Gr\"obner basis at word level has never been explored before for circuit verification.

This proposal is organized as follows: 
Section II reviews preliminary finite state machine and algebraic geometry concepts related to ideals, varieties, Gr\"obner bases, elimination ideals, etc.
Section III describes our proposed approach to perform implicit state enumeration at word level.
An algorithm to extract UNSAT cores based on algebra geometry techniques is conjectured in Section IV. 
Section V outlines the proposed objectives. Finally, Section VI outlines the proposed research timeline.

### 2015-05-01 techwriting/proposal/limit.tex
all write-up from virtue machine

--- 2015-05-01 techwriting/proposal/abs_geo.tex
@@
-\section{Combine algebraic geometry with abstraction techniques}
-\subsection{History of abstraction on sequential circuits}
-Traditional method to analyze a sequential circuit is based on explicit state variable encoding
-and full states enumerating from initial state and explicit transition function.
-With the size of modern sequential circuits increasing, it is infeasible to apply old methods
-any longer. Abstraction is a technique to minimize the state space by combining states with certain 
-similarities. Sometimes it can effectively lower down the number of states by orders of magnitude,
-without affecting the properties we need to verify.
-
-At first abstraction is done manually by designers. In 2000 E. Clarke \cite{clarke2000counterexample}
-proposed an automated abstraction: first,
-initial abstraction is set up by dividing variables to different clusters based on transition conditions;
-then, an ordinary model checking is operated on this abstraction. If a counterexample is given, 
-it is further checked to be concrete or not. For a false abstract state or a spurious path 
-, the initial abstraction is refined by further splitting the false abstract state. 
-A heuristic algorithm keeps refining the abstraction until a real counterexample is found or 
-property verified. This technique is based on binary decision diagrams (BDDs), which means
-potential risk of memory explosion; meanwhile it still requires an initial abstraction to start with.
-So in 2004 L. Zhang and M. Hsiao \cite{zhang2005design} proposed another abstraction method based on CNF-SAT.
-This paper uses simplest latch abstraction -- by removing "irrelevant" latches and change them to primary inputs. 
-The key idea is to pick those ``irrelevant" latches (variables in $k$-bounded model checking, $k$-BMC) 
-according to SAT/UNSAT analysis from $k$-BMC. It improves the $k$-BMC procedure\cite{biere1999symbolic}: 
-LTL properties to check for a $k$-bounded machine can be translated to a big CNF and fed to SAT solver. 
-If SAT, a concrete counterexample is found. 
-If UNSAT, the original method will just increase bound k. In Zhang's paper, the clauses for UNSAT 
-(UNSAT cores) are checked for variables do not appear, they are referring to "irrelevant" latches 
-at least for first $k$ iterations. 
-Thus an abstracted model is obtained and checked with original property. If the abstracted model checking fails, 
-then bound $k$ will increase. This method still has disadvantages: it is counterexample-independent,
-which means it cannot utilize information from invariants.
-
-Abstraction usually means over-approximation, i.e. properties are true on abstracted machine
-are also valid for original states. Craig's interpolation is a sort of abstraction which could be both
-over-approximation and under-approximation. In 2003 K. McMillan used Craig's interpolation to improve 
-$k$-BMC \cite{mcmillan2003interpolation}. Initially a $k$-length path to failure state $F$ is picked and transformed to 
-CNF formula, then split at first transition into prefix and suffix. Then an interpolation $P$ is computed
-between prefix and suffix. $P$ over-approximates 1-step reachable states, and under-approximates states 
-that cannot reach $F$ in $k$ steps. If no counterexample exposed, path will be split again at second
-transition; in this way a precise abstraction can be found for all reachable states.
-
-$k$-BMC with interpolation is a purely incremental model checking approach, and interpolation procedure relies
-on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
-\cite{bradley2011sat} \cite{bradley2011incremental}. (Not related to abstraction really..)
-\subsection{Word-level abstraction: Naive abstraction for arithmetic circuits}
-Our approach is an abstraction based on word operand definition of datapaths in arithmetic circuits,
-and it can be expanded to generic FSM by bundling a set of bit-level variables to one or several word-level variables.
-The abstraction polynomial is obtained through computing GB of an elimination ideal using a special term order.
-
-The authors of \cite{pruss:dac14} showed that for any combinational
-logic block, a canonical word-level polynomial representation can be
-derived through \Grobner bases computed with elimination
-orders. Our approach is based on their result, which we reproduce
-here:
-\begin{Lemma}
-(From \cite{pruss:dac14}) Given a combinational circuit $C$ with $k$-bit
-  input $A = (a_0, \dots, a_{k-1})$ and $k$-bit output $R = (r_0, \dots,
-  r_{k-1})$. Denote by $x_1, \dots, x_d$ all the bit-level
-  variables of   $C$. Let $J = \langle f_1, \dots, f_s \rangle \subset
-  \Fkk[x_1, \dots, x_d, R, A]$ denote all the polynomials corresponding to the
-  logic gates of the circuit. Let $J_0 = \langle x_1^2 - x_1, \dots,
-  x_d^2 - x_d, R^q - R, A^q - A \rangle$ be the vanishing ideal, so
-  that $J + J_0 = \langle f_1, \dots, f_s, ~~ x_1^2 - x_1, \dots,
-  x_d^2 - x_d, R^q - R, A^q - A \rangle$. Compute \Grobner basis $G =
-  GB(J + J_0)$ w.r.t. lex term order with $x_1 > x_2 > \dots > x_d > R
-  > A$. Then $G_d = G \cap \Fkk[R, A]$ eliminates the internal
-  variables $x_1, \dots, x_d$ of the circuit. $G_d$ also contains the
-  word-level polynomial $R = \F(A)$ which canonically represents the
-  function of the circuit.  
-\end{Lemma}
-
-The authors referred to the elimination (lex) order with
-$\{\textit{internal variables} \ x_1 > \dots > x_d\} > \{\textit{word-level
-  output} \ R\} > \{\textit{word-level input} \ A\}$ as the {\it abstraction
-  term order (ATO)}.
-
-One-step transition can also be implemented on word level using ATO. We represent present state (PS) 
-with a univariate polynomial about word-level state variable $f(S)$, its variety intersecting with variety of
-circuit ideal $J+J_0$ results to polynomial about word-level next state (NS) $g(T)$. Recall theorem \ref{thm:unionintersect},
-$$NS = {\bf V}(\langle g\rangle) = {\bf V}(\langle f\rangle + J+J_0) = GB\ with\ ATO(\langle f\rangle + J+J_0)$$
-Note that ideal corresponding to next state is an ideal with single univariate polynomial generator.
-
-Reachability analysis is a useful tool when checking sequential equivalence as well as other invariants.
-With word-level state variables, we can do implicit state enumeration to provide a picture of 
-reachable states. Following algorithm is breadth-first search style state enumeration \cite{KallaPartialScan}:
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
- \KwIn{Transition functions $\Delta$, initial state $S^0$}
-
-  $from^0 = reached = S^0$\;
-  \Repeat{$new^i == 0$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$Img$(\Delta, from^{i-1})$\;
-	$new^i \gets to^i \cap \overline{reached}$\;
-  	$reached \gets reached \cup new^i$\;
-	$from^i \gets new^i$\;
-  }
-\Return{$reached$}
-\caption {Breadth-first Traversal Algorithm}\label{alg:BFS}
-\end{algorithm}
-\begin{figure}[hbt]
-\centering{
-\includegraphics[scale=0.3]{./stg_fig.eps}
-\caption{State transition graph of sample FSM}
-\label{fig:stg}}
-\end{figure}
-
-\begin{figure}[hbt]
-\centering{
-\includegraphics[width=3.5in]{./fsm_fig.eps}
-\caption{Gate-level circuit of combinational logic block of sample FSM}
-\label{fig:fsm}}
-\end{figure}
-Here we put a 2-bit Mealy machine as an example showing how to use univariate polynomial ideal
-and elimination ideal with ATO to implement this algorithm.
-\begin{Example}
-\label{ex:2bit}
-
-We are given a mealy Finite State Machine (FSM). Gate-level circuits of combinational logic
-is figure \ref{fig:fsm}, $x$ is primary input (one bit),
-$\{s_0, s_1\}$ are state inputs, $Z$ is primary output and $\{t_0, t_1\}$ are state outputs.
-Word-level state input is $S = s_0 + s_1\cdot\alpha$, output is $T = t_0 + t_1\cdot\alpha$. In Galois field
-$\mathbb{F}_{2^2}$, bit-level variables $x, s_0, s_1, Z, t_0, t_1$ can take values from $\{0, 1\}$, while
-word-level variables $S, T$ may take values from $\{0, 1, \alpha, 1 + \alpha\}$. State Transition Graph (STG)
-showed in figure \ref{fig:stg} uses 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2, S_3\}$, which
-could be mapped to elements in $\mathbb{F}_{2^2}$.
-
-For example in Line 1 of BFS algorithm, assume initial state is $S_0$ in figure \ref{fig:stg}, then corresponding
-polynomial $f = \mathcal{F}(S^0) = S^0 - 0$. Consider an ideal $I$ with only one generator $f$, its variety
-$V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 0\}$, which equals to $\{0\}$, the only
-valid value $S^0$ can take.
-
-If an ideal contains multiple polynomial specifications, it is necessary to compute Gr\"obner basis with elimination term
-order to get one polynomial only on desired variable. In first iteration of Line 4, $to^i$ has multiple specifications,
-some of them are transition functions on bit-level:
-\begin{displaymath}
-\begin{array}{ll}
-f_1:& t_0 - (\overline{x}\ \land\ \overline{s_0}\ \land\ \overline{s_1})\ \lor\ (s_0\ \land\ s_1) \\
-f_2:& t_1 - (\overline{s_0}\ \land\ x)\ \lor\ (s_0\ \land\ \overline{s_1})\
-\end{array}
-\end{displaymath}
-And some are bits-to-word definitions using polynomial representation of elements in $\mathbb{F}_{2^2}$:
-\begin{displaymath}
-\begin{array}{ll}
-f_3:& S - s_0 - s_1\alpha \\
-f_4:& T - t_0 - t_1\alpha
-\end{array}
-\end{displaymath}
-And an polynomial about initial state as mentioned above:
-\begin{displaymath}
-f_5:\  S
-\end{displaymath}
-And the rests are vanishing polynomials for every variable, bit-level and word level:
-$f_6: x^2 - x; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: s_0^2 - s_0;
-f_{11}: s_1^2 - s_1; f_{12}: T^4 - T$
-
-Note that there is no specification on initial primary input $x$, in Gr\"obner basis approach this means $x$ is smoothed by
-reversely using Shannon's expansion.
-
-Employing elimination term order: $intermediate\ bit$-$level\ signals\ >\ bit$-$level\ primary\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
-$J = \langle f_1, f_2, \dots, f_{12}\rangle $, the result will include one polynomial $f_T$ contains only variable $T$. In this example,
-$f_T = T^2+(\alpha+1)T+\alpha$. This polynomial equals to $T^2-(\alpha+1)T+\alpha$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
-limited within $\mathbb{F}_{2}$, where $x \equiv -x\ (mod\ 2)$ for any element $x$ in the field. Solution to $f_T = 0$ is $T = 1\ or \ T = \alpha$, which shows that next
-state the machine just reached is $\{S_1(01), S_2(10)\}$.
-
-Continue to run this traversal algorithm, it will terminates with final reachable states $f_T = T^3+(\alpha+1)T^2+\alpha T$, its solution is $T=0\ or\ T=1\ or\ T=\alpha$,
-which shows within total 4 states, state $S_3$ is unreachable from initial state $S_0$.
-
-When executing Line 5 and Line 6, use \textbf{union}, \textbf{intersection} and \textbf{complement} of varieties over $\mathbb{F}_{2^k}$. Since
-state set variables $from^i, to^i, reached$ all have single polynomial generator, consider an ideal $I$ with single generator $I = \langle f\rangle $.
-For example, assume $I = \langle f\rangle  = \langle T^2 + (1+\alpha)\cdot T+\alpha\rangle $, its variety $V(I) = \{a\ |\ a \in \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$.
-
-For varieties intersection $\{1\}\bigcap\{1, \alpha\}$, calculate ideal sum $\langle T+1, T^2 + (1+\alpha)\cdot T+\alpha\rangle  = \langle T+1\rangle $,
-its variety is $\{1\}$; for varieties union $\{1,\alpha\}\bigcup\{1+\alpha\}$, calculate
-ideal product $\langle (T+1+\alpha)\cdot(T^2 + (1+\alpha)\cdot T+\alpha)\rangle  = \langle T^3 + 1\rangle $, its variety
-is $\{1, \alpha, 1+\alpha\}$; for complement set of variety $\{1, \alpha\}$, the universal set is
-the variety of ideal of vanishing polynomial $V(\langle T^4-T\rangle ) = \{0,1,\alpha,1+\alpha\}$,
-so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
-which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
-the result is $\{0,1+\alpha\}$.
-
-From discussion above, the BFS traversal is completely implemented in Galois field.
-If the initial input is $S_3$: $S + 1 + \alpha$, the final return value will be set of
-reachable states: $T^4 + T$, i.e. the universal set, $\{S_0, S_1, S_2, S_3\}$. Modified algorithm could be
-written in following interpretation:
-
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
- \KwIn{Input-output circuit characteristic polynomial ideal $I_{ckt}$, initial state polynomial $\F(S)$}
-
-  $from^0 = reached = \F(S)$\;
-  \Repeat{$new^i == 1$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$GB w/ elimination term order$\langle I_{ckt}, from^{i-1}\rangle$\;
-	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
-  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
-	$from^i \gets new^i(S\setminus T)$\;
-  }
-\Return{$reached$}
-\caption {Algebraic Geometry based Traversal Algorithm}\label{alg:univa}
-\end{algorithm}
-
-Here $from^i, to^i, new^i$ are all univariate polynomials about $S$ or $T$.
-\end{Example}
-
+\section{Overcoming the Limitations}
 \subsection{Obviate GB computation}
-% Introduce: RATO
-% Get word-level output -- bit-level inputs function
-% Use bit variable substitution to finish word-level abstraction
 Elimination ideal with ATO technique requires a GB computation, which is usually infeasible on large designs.
@@
 
-It turns out that RATO can be applied to sequential circuits in much
-the same way: {\it \{bit-level circuit variables ordered
-  reverse topologically} $x_1 > \dots x_d\}$ $>$  {\it \{word-level NS
-  variables} $R'> A'> B'\}$ $>$ {\it \{word-level PS variables}
-$R>A>B\}$. Importantly, {\it we show that using RATO, the polynomial
-abstraction can be derived without resorting to a \Grobner basis
-computation.} Perform the following operations:
-
-\begin{enumerate}
-\item Represent the polynomials of the sequential circuit $S$ using RATO.
-\item Due to RATO, only one pair of polynomials ($f_i, f_j$) will have
-  leading terms that will not be relatively prime.
-\item Reduce $Spoly(f_i, f_j) \xrightarrow{F}_+ h$.
-\end{enumerate}
-As described in \cite{pruss:dac14}, remainder $h$ will contain:
-  i) the word-level variables, and ii) bit-level inputs to the
-  combinational logic, i.e. bit-level present-state variables. All
-  other internal circuit variables will not appear in $h$, as they
-  will be canceled by division due to RATO.
-
-However with only RATO, the final remainder $h$ contains both bit-level variables and
+Additionally, further improvements can be made to RATO:
+With only RATO, the final remainder $h$ contains both bit-level variables and
 word-level {\it state variables}. We desire
@@
 bit-level state variables in terms of their word-level
-variables. We exploit 
-the following property of finite fields:
+variables. The basic idea of this improvement is: exploit the property of
+polynomial squaring in $\Fkk$ to list $k$ polynomial equations,
+use a Gaussian-elimination style technique to solve the system of equations.
+Then we can substitute all bit-level variables in the remainder polynomial from RATO.
+Details are described in our accepted DATE 2015 conference paper 
+``{\it Formal Verification of Sequential Galois Field Arithmetic Circuits using Algebraic Geometry}".
 
-\begin{Lemma}
-\label{lemma:square}
-For $\alpha_1, \dots, \alpha_t \in \Fkk$, $(\alpha_1 + \alpha_2 +
-\dots + \alpha_t)^{2^i} = \alpha_1^{2^i} + \alpha_2^{2^i} + \dots +
-\alpha_t^{2^i}$, for $i = 1, 2, \ldots$. 
-\end{Lemma}
-
-Consider the polynomials that define the relationship between the
-word-level and bit-level variables. Let 
-$ A = a_0 \beta + a_1\beta^2 + \dots + a_{k-1}\beta^{2^{k-1}}$.
-Due to Lemma \ref{lemma:square}, we have $A^2 = a_0 \beta^{2}+ a_1
-\beta^{4} + \dots + a_{k-1} \beta^{2\cdot 2^{k-1}}$ (as
-$a_i^2 = a_i$). By repeated squaring:
-\begin{eqnarray}
-A & = & a_0 \beta + a_1\beta^2 + \dots + a_{k-1}\beta^{2^{k-1}}\nonumber\\
-A^2 & = & a_0 \beta^{2}+ a_1 \beta^{4} + \dots + a_{k-1}
-\beta^{2\cdot 2^{k-1}} \nonumber\\
-\vdots & \vdots& \vdots \nonumber\\
-A^{2^{k-1}} & = & a_0 \beta^{2^{k-1}}  +  a_1 \beta^{2^{k-1}\cdot 2} +
-\dots + a_{k-1}\beta^{2^{2(k-1)}} \nonumber
-\end{eqnarray}
-% \alert{Some concern that we do not show that this system of equations is always
-% uniquely solvable.}
-Consider the above as $k$ {\it linear equations} with $a_0, \dots, a_{k-1}$
-as $k$-unknowns, with $\beta$ and its powers as coefficients, and 
-$A, A^2, ... \dots, A^{2^{k-1}}$ as $k$ constants. Then, we can solve
-for the unknowns $a_0, \dots, a_{k-1}$ and obtain expressions for them
-in terms of $A$ and $\beta$. The problem can be setup in
-matrix form:
-
-\begin{displaymath}
-\begin{pmatrix}
-\beta & \beta^2 & \cdots & \beta^{2^{k-1}} \\
-\beta^2 & \beta^4 & \cdots & \beta^{2^{k-1}\cdot 2} \\
-\vdots & \vdots & \ddots & \vdots \\
-\beta^{2^{k-1}} & \beta^{2^{k-1}\cdot 2} & \cdots & \beta^{2^{2(k-1)}}
-\end{pmatrix}
-\begin{pmatrix}
-a_0\\
-a_1\\
-\vdots\\
-a_{k-1}
-\end{pmatrix}
-=
-\begin{pmatrix}
-A\\
-A^2\\
-\vdots\\
-A^{2^{k-1}}
-\end{pmatrix}
-\end{displaymath}
-
-Gaussian elimination on this system of equations can be 
-applied, and each bit-level variable $a_i$ can be represented as a
-function of the word-level variables: $a_i = g_i(A)$. These bit-level
-variables can be easily substituted in the remainder $h$ obtained by
-reduction due to RATO. What we will obtain is a word-level polynomial
- which canonically represents the function of the circuit.
-
-Using RATO to get word-level transition function may result in multivariate polynomial generators
-instead of single univariate generator. To compute NS, we intersects variety of transition ideal 
-with PS (output word's function about input words, plus
-evaluations of input words), then result NS may have multiple multivariate generators. 
-Complement also requires multivariate ideal quotient.
-
-\begin{Example}
-Consider 2-bit FSM traversal (Ex:\ref{ex:2bit}) using RATO and multivariate polynomial generators.
-Following algorithm is a modified breath-first search (BFS) traversal algorithm based on ideals with multivariate 
-generators.
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
- \KwIn{Transition polynomial $f_t = T + \mathcal F (S,x)$, 
-	initial state ideal $from^0 = \langle S+\mathcal G(x), x^{q_1} - x\rangle$}
-
-  $reached = from^0(T\setminus S)$\;
-  \Repeat{$GB(new^i) == 1$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$GB$(\langle f_t, from^{i-1}\rangle) \setminus \mathcal H(S)$\;
-	$\overline{reached} = \langle T^{q_2}-T, x^{q_1} - x \rangle : reached$\;
-	$new^i \gets $GB$(to^i + \overline{reached})$\;
-  	$reached \gets $GB$( reached \cdot new^i)$\;
-	$from^i \gets new^i(S\setminus T)$\;
-  }
-\Return{$reached$}
-\caption {Algebraic Geometry based Traversal Algorithm (multivariate-generator ideals)}\label{alg:multi}
-\end{algorithm}
-
-Inputs of this algorithm are: a transition polynomial, which is the result of word-level abstraction; initial 
-states description ideal, contains 2 generators defining PI could be any inputs and state inputs $S$ is a specific
-value. (Ex: $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
-$\langle S+x\cdot\alpha, x^2+x\rangle$ means initial states$=\{00,10\}$).
-
-Transition polynomial calculation exploits our improved RATO approach. After building
-an elimination ideal, use RATO such that \emph{reverse\ topo\ order\ circuit\ variables }$> T > S > x$, the reduction
-remainder has the form $T+\mathcal F(s_0,s_1,x)$. From word definition $S+s_0+s_1\alpha$ we get
-$$s_0 = \alpha S^2+ (1+\alpha)S, s_1 = S^2+S$$
-Do substitution, the transition polynomial of example circuit is 
-$$f_T = T+S^3\cdot x+\alpha S^3+(1+\alpha)S^2\cdot x+S^2+S\cdot x+(1+\alpha)x+1$$
-
-Assume we start from state $\{11\}$. First iteration, it will reach state $\{01\}$. Line 4 is to compose an
-ideal with 2 generators from $from^0$ and transition polynomial $f_T$, compute its Gr\"obner basis. Note this ideal
-has the form
-\begin{equation}
-I_{tran} = \left\{
-             \begin{array}{c}
-             T+\mathcal F(S,x) \\
-             S + \mathcal G(x) \\
-             v_x
-             \end{array}  
-        \right.
-\end {equation}
-$v_x$ is a polynomial containing only $x$, initially it should be vanishing polynomial $x^{q_1}-x$, with the program executing
-it may be factorized.
-
-Consider Buchberger's algorithm, all generators' leading terms are relatively prime, so it is a GB itself. 
-Then we need to reduce this GB, we will find out $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
-and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. So, at last we will get a polynomial
-$T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
-$S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$
-
-Line 5 is the ideal quotient of universal set and reached states. In the first iteration, 
-$reached$ is the initial state $\langle T+1+\alpha, x^2+x \rangle$. Result of ideal quotient
-is $\langle T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle$ represents $\{00,01,10\}$.
-
-Line 6 is ideals' sum (intersection of their varieties), it is done by combining all generators
-from 2 ideals and compute GB. For first iteration result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.
-
-Line 7 is ideals' product (union of their varieties), it is done by multiplying all pairs of
-generators from both ideal. For first iteration result is $GB(\langle (T+1)(T+1+\alpha),
-(T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$ representing $\{01,11\}$.
-
-The traversal will run for 3 iterations and terminate at 4th iteration. We list all intermediate 
-results below:
-\begin{itemize}
-\item Iteration 1: $from^0 = \langle S+1+\alpha, x^2+x\rangle, to^1 = \langle T+1, x^2+x\rangle,
- reached = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$
-\item Iteration 2: $from^1 = \langle S+1, x^2+x\rangle, to^2= \langle T+\alpha, x^2+x\rangle,
-reached = \langle T^3+1, x^2+x\rangle$
-\item Iteration 3: $from^2 = \langle S+\alpha, x^2+x\rangle, to^3 = \langle T+\alpha x, x^2+x
-\rangle, reached = \langle T^3\cdot x+x, T^4+T, x^2+x\rangle$
-\item Iteration 4: $from^3 = \langle S, x\rangle, to^4 = \langle T+1, x\rangle, new = \langle1\rangle$
-\end{itemize}
-The final reachable states are represented by ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
-which means $\{00,01,10,11\}$.
-\end{Example}
-
-\subsection{A simplified traversal on multivariate polynomial generators: SMPO}
-% 1. How to design SMPO
-% 2. Improved algorithm: why this can verify arithmetic function
-% 3. Example of 3-bit SMPO
-Sequential GF multiplier has no primary input, so it can be regarded as Moore FSM. 
-The states are encoded by evaluations of state variables, which can be implicitly verified 
-by checking output-input function ($R = A\cdot B \pmod P(\alpha)$). Additionally
-the inner structure of GF multiplier is XOR-rich such that we can test our approach on very large
-(>100 bits datapath) circuits.
-
-Let us briefly describe the fundamentals behind the design of normal
-basis sequential GF multipliers, so as to put in perspective the type
-of designs that have been verified in this paper. Let $R =
-\sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
-\beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$, then 
-\[
-R = A\cdot B = (\sum_{i=0}^{k-1} a_i \beta^{2^{i}}) (\sum_{j=0}^{k-1}
-b_j \beta^{2^{j}})  =
-\sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}\nonumber 
-\]
-
-The expressions $\beta^{2^i}\beta^{2^j}$ are called cross-product
-terms and they can also be represented in normal basis: 
-\begin{displaymath}
-\beta^{2^i}\beta^{2^j} =
-\sum_{n=0}^{k-1}\lambda_{ij}^{(n)}\beta^{2^n}, \ \ \lambda_{ij}^{(n)}
-\in \Ftwo. 
-\end{displaymath}
-
-From the above two equations, one can see that the expression for the
-$n^{th}$ digit of product $R = (r_0, \dots, r_n, \dots r_{k-1})$ is:
-\[
-r_n = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(n)}a_ib_j = A
-\cdot M_n \cdot B^T, ~~0 \leq n \leq k-1
-\]
-
-where $M_n = (\lambda_{ij}^{(n)})$ is a binary $k \times k$ matrix over
-$\Ftwo$, and it is called the $\lambda$-matrix. 
-%The collection of all
-%$\{M_n\}$  $\lambda$-matrices is called the multiplication
-%table. 
-Moreover, let $r_n = A \cdot M_n \cdot B^T$.
-Then $r_{n-1} = A \cdot M_{n-1} \cdot B^T = rotate(A) \cdot M_n
-\cdot rotate(B)^T$. This implies that $M_n$ is generated by right
-and down cyclic shifting $M_{n-1}$. Therefore, the hardware design
-of sequential GF multipliers is based on mappings of $A\cdot M_n
-\cdot B^T$ into AND-XOR gates and cyclic shift operations.
-
-This paper verifies the implementation of two distinct 
-architectures of {\it sequential multipliers with parallel output
-  (SMPO)}, namely: i) the Agnew-SMPO  \cite{agnew1991implementation}
-by G. B. Agnew, which is a straight-forward implementation of the
-$\lambda$-matrix; and ii) the more recent, more complicated, yet very
-efficient RH-SMPO \cite{RHmulti}, by Reyhani-Masoleh and Hasan,
-depicted in Fig. \ref{fig:RHmulti}.
-
-\begin{figure}[hbt]
-\centering{
-%\begin{minipage}{12cm}
-\includegraphics[width=2.25in]{./RH3.eps}
-% \vspace{-0.2in}
-\caption{A 3-bit RH-SMPO}
-%\end{minipage}
-\label{fig:RHmulti}}
-\end{figure}
-
-We follow  the typical sequential GF circuit model with word-level variables $A, B, R$
-denoting {\it present states (PS)} and $A', B', R'$ denoting {\it next
-  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
-for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
-\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
-store the result, and $A, B\ (A', B')$ store input operands. {\it E.g.,}
-for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
-0$) are the initial values (operands) loaded into the registers,  and
-$R = \F(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
-result after $k$-cycles. Our approach aims to find this polynomial
-representation for $R$.
-
-Each gate in the combinational logic is represented by a Boolean
-polynomial. To 
-this set of Boolean polynomials, we append the polynomials that define
-the word-level to bit-level relations for PS and NS variables ($A =
-\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
-as ideal $J = \langle 
-f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
-  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
-  of the circuit. The ideal of vanishing polynomials $J_0$ is also included, and
-then the implicit FSM unrolling problem is setup for abstraction.
-
-The configurations of the flip-flops are the states of the
-machine. {\it Since the set of states is a finite set of points, we
-can consider it as the variety of an ideal related to the circuit
-}. Moreover, since we are interested in
-the {\it function encoded} by the state variables (over $k$-time
-frames), we can {\it project this variety} on the word-level state
-variables, starting from the initial state $A_{init}, B_{init}$.
-Projection of varieties (geometry) corresponds to elimination ideals
-(algebra), and can be analyzed via \Grobner bases. Therefore, we
-employ a \Grobner basis computation with ATO: we use a {\it lex term
-  order} with {\it bit-level variables} 
-$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
-allows to eliminate all the bit-level variables 
-%(corresponding to the combinational logic and the state variables),
-%so as to 
-and derives a representation only in terms of words. 
-Consequently, $k$-successive \Grobner basis computations implicitly
-unroll the machine, and provide word-level algebraic $k$-cycle
-abstraction for $R'$ as $R' = \F(A_{init}, B_{init})$.
-
-Algorithm
-\ref{alg:modified} describes our approach.  In the algorithm, $from_i$
-and $to_i$ are polynomial ideals whose varieties are the valuations of
-word-level variables $R, A, B$ and $R',A',B'$ in the $i$-th iteration;
-and the notation ``$\setminus$'' signifies that the $NS$ in iteration
-$(i)$ becomes the $PS$ in iteration $(i+1)$. Line 5 computes the Gr\"obner 
-basis with the abstraction term order.  Line 6 computes the elimination 
-ideal, eliminating the bit-level variables and representing the set of 
-reachable states up to iteration $i$ in terms of the elimination ideal. 
-These computations are analogous to those of image computations performed in FSM reachability. 
-%The forward image
-%$to^{i}$ is computed using \Grobner bases with ATO.
-
-\vspace{-0.1in}
-\IncMargin{1em}
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
-\LinesNumbered
- \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
-   state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}
-
-  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
-  $i = 0$\;
-  \Repeat{$i == k$}
-  {
-  	$i \gets i + 1$\;
-%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
-%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
-	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
-    )$ with ATO\;
-	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
-	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
-  }
-\Return{$from_k(R_{final})$}
-\caption {Abstraction via implicit unrolling for Sequential GF circuit
-  verification}
-\label{alg:modified}
-\end{algorithm}
-\DecMargin{1em}
-\vspace{-0.1in}
-\begin{Example}
-\label{ex:RHSMPO}
-
-We demonstrate our approach to verify the 3-bit RH-SMPO circuit of
-Fig.\ref{fig:RHmulti}. The normal element $\beta$ in
-$\mathbb{F}_{2^3}$ is known to be $\beta = \alpha^3$, where $\alpha$
-is the primitive element. The circuit can be described with an ideal by translating
-AND and XOR gates accordingly. For the first iteration:
-\begin{align*}
-J = &d_0+b_2\cdot a_2,
-c_1+a_0+a_2,
-c_2+b_0+b_2,
-d_1+c_1\cdot c_2,\\
-&e_0+d_0+d_1,
-e_2+d_1,
-r_0'+r_2+e_0,
-r_1'+r_0,
-r_2'+r_1+e_2,\\
-&A+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
-B+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12},\\
-&R+r_0\alpha^3+r_1\alpha^6+r_2\alpha^{12},
-R'+r_0'\alpha^3+r_1'\alpha^6+r_2'\alpha^{12};
-\end{align*}
-The last 4 polynomials are translated from the definition of word-level variables.
-This represents ideal ``$J$" from line 5 in Algorithm
-\ref{alg:modified}. ``$J_0$" is the ideal of vanishing polynomials in all bit-level
-variables (e.g. $a_0^2-a_0$) and word-level variables (e.g. $A^8-A$). ``$from_{i-1}$"
-% contains evaluation of PS word-level output in form of
-% $R+\mathcal F(A,B)$. Since the definition of $A$ and $B$ changes after every iteration
-% ($A',B'$ are different from $A, B$) and the multiplier's function only concerns about 
-% initial value of $A,B$, we include the definition of $A_{init}, B_{init}$ into $from_i$:
-represents the set of current states for iteration $i$.
-In the first iteration, $from_0 = \{R, A_{init}+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
-B_{init}+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12}\}$.
-
-After the GB computation is performed with ATO, as line 6 in Algorithm \ref{alg:modified},
-we find a polynomial in variables $R', A_{init}, B_{init}$ in $to_1 : 
-R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
-$.
-
-Line 7 in Algorithm \ref{alg:modified} simply replaces NS output $R'$ with PS output
-$R$ in this example; so in second iteration $from_1 = \{
-R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
-, A_{init}+a_2\alpha^3+a_0\alpha^6+a_1\alpha^{12},
-B_{init}+b_2\alpha^3+b_0\alpha^6+b_1\alpha^{12}\}$.
-
-% By continue executing the loop,
-Finally, after 3 iterations we obtain: $to_3 = \{ \mathbf{R'+A_{init}B_{init},}
-~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
-~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
-as the image. The final result is $from_3(R_{final}) = R_{final}+A_{init}\cdot
-B_{init}$, which verifies the multiplier. 
-\end{Example}
-
-\subsection{Limitations: complexity of reduction (multivariate division)}
+\subsection{Complexity of multivariate division}
+Using RATO and bit-word substitution, we can turn GB computation problem to a single multivariate division.
 Although we gain success on verifying sequential GF multipliers, there are still problems when
@@
 \begin{itemize}
-\item When divided by polynomial from XOR gate, the length of remainder will increase by 1 term;\\
-\item When divided by polynomial from AND gate, the degree of remainder will increase by 1;\\
-\item When divided by polynomial from OR gate, the length and degree of remainder will increase by 1.
+\item[-] When divided by polynomial from XOR gate, the length of remainder will increase by 1 term;
+\item[-] When divided by polynomial from AND gate, the degree of remainder will increase by 1;
+\item[-] When divided by polynomial from OR gate, the length and degree of remainder will increase by 1.
 \end{itemize}
-Since all gate polynomials have leading term with degree 1, so increase of remainder degree means
-iterations of division will at least increase by 1 (while there is still possibility to cancel multiple terms
-at one time). Consider the querying time, a chain with AND/OR gates will blow up computational complexity of
+Since all gate polynomials have leading term with degree 1, a higher degree of remainder polynomial means
+ more iterations of division (while a longer remainder does not always imply more divisions because
+ there is still possibility to cancel multiple terms at one time). 
+ Consider the querying time, a chain with AND/OR gates will blow up computational complexity of
 reduction.
 
-Sequential GF multiplier has only limited levels of gate chains (4 for RH-multiplier). Although the size of 
+Sequential GF multiplier has only limited levels of gate chains (4 for RH-multiplier example). Although the size of 
 datapath is large, actual complexity for reduction is relatively low.

### 2015-05-01 techwriting/proposal/prelim.tex
all write-up from virtue machine

%\input{myformulae.tex}

\section{Preliminaries}

In this section, we will briefly review the preliminary concepts
related to finite state machine models, Galois fields, algebraic geometry and Gr\"obner bases, as
our approaches are based on a combination of these concepts.

\subsection{FSM model for sequential circuits}
A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
the formal definition is as follows:
\begin{Definition}
A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
\begin{itemize}
\item $\Sigma$ is the input label, $O$ is the output label;
\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
\end{itemize}
\end{Definition}
The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
its primary outputs only depend on the present states, i.e. the output function is defined as
$$\Lambda:\ S \to O$$
Typical sequential circuits can be depicted as Fig.\ref{fig:seqmodel}(a). Primary inputs
$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
at bit level are defined as $\Delta_i: t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$.
\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{./seqmodel.eps}
% \vspace{-0.2in}
\caption{FSM models of sequential circuits}
%\end{minipage}
\label{fig:seqmodel}}
\end{figure}
In some cases, arithmetic computations are implemented as Moore machines where input operands
are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
We can simplify them to the model in Fig.\ref{fig:seqmodel}(b).

\subsection{Commutative algebra and algebraic geometry preliminaries}
A {\bf field} $\mathbb{F}$ is a set of elements, including 0 and 1 (unity),
allowing for associative and commutative addition and multiplication;
and every non-zero element has a multiplicative inverse.  A {\bf
  finite field} or {\bf Galois filed} is a field with a finite number
($q$) of elements, and is denoted by $\Fq$, where $q=p^k$ is a power of
a prime integer $p$. In our work, $q = 2^k$ for
a given $k$, where $k$ represents the datapath (bit-vector)
word-lengths, or the number of memory elements (state registers) in
finite state machines.

Let the field $\mathbb{F}_2 = \{0, 1\} ~(\equiv \B)$, and let
$\mathbb{F}_2[X]$ denote the set (ring) of all univariate polynomials
in variable $X$ with coefficients from $\mathbb{F}_2$. Then, the
Galois field $\Fkk$ is constructed as $\Fkk = \mathbb{F}_2[X] \pmod{
  P(X)}$, where $P(X)$ is an irreducible polynomial over
$\mathbb{F}_2$. Let $\alpha$ be a root of the irreducible polynomial
$P(X)$, i.e. $P(\alpha) = 0$. Any element $A \in \Fkk$ can be
represented as $A = \sum_{i=0}^{k-1} a_i \alpha^i$, where $a_i \in
\mathbb{F}_2$. The field $\Fkk$ is therefore, a $k$-dimensional {\it
  extension} of the base field $\mathbb{F}_2$: so,  $\mathbf{\Fkk \supset
\mathbb{F}_2}$. Consequently, all operations of addition and
multiplication in $\Fkk$ are performed modulo the irreducible
polynomial $P(\alpha)$ and coefficients are reduced modulo 2.

Boolean variables in field $\mathbb B$ can be easily mapped to
elements in $\mathbb F_2$. Since $\mathbb F_2 \subset \Fkk$, these 
%Namely $k$-bit Boolean bit-vector defined
%over Boolean ring $\mathbb B^k$ can also be mapped uniquely  to $\Fkk
%= \mathbb F_2[x_1,\dots,x_k]$. 
Boolean operators are interpreted as functions over $\Fkk$ (where $+$
and $\cdot$ are addition and multiplication performed modulo 2): 
\begin{align*}
&a\land b \to a\cdot b\\
&a\oplus b \to a+b\\
&\neg a \to 1+a\\
&a \bar{\oplus} b \to 1+a+b\\
&a \lor b \to a+b+a\cdot b
\end{align*}

Using these mappings we can write Boolean functions in form of
polynomials over $\mathbb F_2\subset \Fkk$. 
These concepts provide a mechanism to represent and
manipulate both bit-level ($\mathbb{F}_2$) and $k$-bit word-level
constraints {\bf in one unified mathematical domain $\Fkk$ --- a
  concept we   exploit for abstraction. }
  Consider Ex.\ref{ex:motiv},
polynomials for transition functions (bit-level outputs) $f_1,f_2$
are over $\mathbb F_2$, i.e. $f_1,f_2\subseteq \mathbb F_2 \subset \Fkk$,
and polynomials containing word-level variables $f_3,f_4,f_5 \subseteq \Fkk$.
All polynomials belong to unified domain $f_1,f_2,\dots,f_5 \subseteq \Fkk$.

It is well-known that every Boolean mapping between $k$ dimensional
Boolean spaces $f: \B^k \rightarrow \B^k$ can be construed as a
function over Galois fields $f: \Fkk \rightarrow \Fkk$. Moreover,
every function $f: \Fkk \rightarrow \Fkk$ is a polynomial function: 
i.e. $f$ can be represented by way of a unique, minimal, canonical
polynomial $\F(X)$, and the work of \cite{timDAC} shows how to
efficiently derive such polynomial representations from circuits ---
another concept that makes our approach feasible.

We represent Boolean circuits by way of polynomials over
$\Fkk$. If we take indeterminates $x_1,x_2,\dots,x_n$, an arbitrary
combination of their finite product  
$x_1^{d_1}\cdot x_2^{d_2}\cdots x_n^{d_n}, d_i\geq 0$ 
is a {\bf monomial}. A {\bf polynomial} $f = c_1 X_1 + c_2 X_2 + \dots
+ c_t X_t$ is a finite sum of terms, where $c_1, \dots, c_t$ are
coefficients and $X_1, \dots, X_t$ are monomials. The set of {\it all}
such polynomials with coefficients from $\Fkk$ forms a {\bf
  multivariate polynomial ring} denoted $\Fkk[x_1,\dots,x_n]$. 
A monomial ordering $X_1 > X_2 \dots > X_t$ is imposed on the
polynomials to process them systematically. Then, $LT(f) = c_1 X_1,
LM(f) = X_1$ denote the leading term and the leading monomial of $f$,
respectively.

Multivariate polynomial division will play a key role in our
algorithmic techniques. Division is implemented as {\it cancellation
  of terms.} Given polynomials $f, g$, if $cX$ is a term in
$f$ that is divisible by $LT(g)$, then $f \xrightarrow{g} r$ denotes a
one-step reduction (division) of $f$ by $g$, resulting in remainder $r
= f - {{cX} \over {LT(g)}} \cdot g$. This has the effect of cancelling
the term $cX$ from $f$. 
\begin{Example}
\label{ex:multidiv}
If $f = e + cd$ and $g = c + ab$,
then the term $cd$ in $f$ can be canceled by $LT(g) = c$: $r = f - {cd
  \over c} g = e + abd$. 
\end{Example}
  Similarly, $f$ reduces to $r$ modulo the set
of polynomials $F = \{f_1, \dots, f_s\}$, denoted $f \stackrel{F}
{\textstyle   \longrightarrow}_+ r$, such that no term in $r$ is
divisible (cancellable) by the $LT(f_i)$ of any polynomial in $ f_i
\in F$.

In verification, we have to analyze the {\it solutions to a
set of polynomials.} The set of all solutions to a system of
polynomial equations $f_1 = \dots = f_s = 0$ is defined as the affine variety:
\begin{Definition}
Given a set of polynomials $f_1,\dots,f_s$ over ring $\mathbb F_q[x_1,\dots,x_n]$, their 
{\bf affine variety} 
$$V(f_1,\dots,f_s) = \{(a_1,\dots,a_n)\in  (\mathbb F_q)^n |
f_1(a_1,\dots,a_n) = \cdots = f_s(a_1,\dots,a_n) = 0\}$$ 
\end{Definition}

Generally we can find many sets of polynomials with the same variety, which are linear combinations
of given set of polynomials. This set is defined as follows:
\begin{Definition}
{\bf Ideal of Polynomials:} Let $f_1,f_2,\dots,f_s\in \mathbb F[x_1,\dots,x_n]$.
Define an ideal
$$J = \langle f_1,f_2,\dots,f_s\rangle = \{f_1\cdot h_1 + f_2\cdot h_2 +\cdots + f_s\cdot h_s : h_1,\dots,h_s\in \mathbb F[x_1,\dots,x_n]\}$$
We call $J = \langle f_1,f_2,\dots,f_s\rangle$ an ideal generated by $f_1,\dots,f_s$ and these polynomials 
the {\bf generators} of ideal $J$.
\end{Definition}
% On the other hand, if given another polynomial $f$, we need to judge whether it belongs to $J$.
A practical problem is: given an ideal $J = \langle f_1,f_2,\dots,f_s\rangle$ and a polynomial $f$,
we need to check if the variety of $J$ can make the evaluation of $f$ equal to 0, i.e. $f$ vanishes on $V(J)$.
This problem is usually described as ideal membership checking problem.
\begin{Definition}
{\bf Ideal membership:} Let $f_1,f_2,\dots,f_s\in \mathbb F[x_1,\dots,x_n]$, and $J = \langle f_1,f_2,\dots,f_s\rangle$
be an ideal over ring $\mathbb F[x_1,\dots,x_n]$. If 
$$f = f_1h_1 + f_2h_2 + \cdots + f_sh_s$$
then $f\in J$.
\end{Definition}

An ideal may have many generating sets. For example, we may have different set of polynomial generators denoting
the same ideal, where they have the same variety: $\langle f_1,\dots,f_s\rangle = \langle h_1,\dots,h_r\rangle
= \langle g_1,\dots,g_t\rangle$ such that $V(f_1,\dots,f_s) = V(h_1,\dots,h_r) = V(g_1,\dots,g_t)$.
Therefore a canonical representation of an ideal is needed, which leads to the concept of Gr\"obner bases.

\begin{Definition}
The set $G = \{g_1, \dots,
g_t\}$ is called a \textbf{\Grobner basis} of $J$ if and only if the
leading term of all polynomials in $J$ is divisible be the leading
term of some polynomial $g_i$ in $G$: i.e. $\forall f \in J, \exists
g_i \in G \ s.t. \ LT(g_i) ~|~ LT(f)$. 
\end{Definition}
% The famous Buchberger's
% algorithm, given in textbook \cite{ideals:book}, is used to compute a
% \Grobner basis (GB). Operating on  input $F = \{f_1, \dots, f_s\}$,
% and subject to the imposed term order $>$, it derives $G = GB(J) = \{
% g_1, \dots, g_t \}$. Buchberger's algorithm repeatedly computes
% $S$-polynomials. For pairs $(f_i, f_j) \in F$, $Spoly(f_i, f_j) =
% \frac{L}{lt(f_i)}\cdot f_i - \frac{L}{lt(f_j)}\cdot f_j$, where $L =
% LCM(LM(f_i), LM(f_j))$. Reducing $Spoly(f_i, f_j) \xrightarrow{F}_+ r$
% cancels the leading terms of $f_i, f_j$ and gives a polynomial $r$
% with a new leading term. This remainder $r$ is added to the current
% basis and $Spoly(f_i, f_j)$ computations are repeated for all pairs of
% polynomials until all $S$-polynomials reduce to 0.

An advantage of representing an ideal with GB is that it can serve as a decision procedure for ideal membership
test when dividing a polynomial $f$ by a GB, i.e.
$$G = GB(J) \Longleftrightarrow \forall f\in J, f\xrightarrow{g_1,g_2,\dots,g_t}_{+} 0$$
Gr\"obner basis can be reduced by eliminating redundant elements. \textbf{A reduced GB is a canonical representation of 
the ideal under a given monomial ordering}. Given an ideal $J = \langle f_1,\dots,f_s\rangle, ~G = 
\{g_1,\dots,g_t\}$ is the GB of $J$, it can be computed by Buchberger's algorithm (refer to textbook \cite{ideal:book}).

Another advantage of using GB representation is that GB computation can work as a {\it quantification procedure}.
In the following part we will introduce the concept of \textit{vanishing polynomials}, \textit{elimination ideal}, etc.
as the bases of this theory.

{\it Fermat's little theorem over $\Fq$:} For any $ \alpha \in \mathbb
F_{q}, \alpha^q = \alpha$. Therefore, the polynomial $x^q - x$
vanishes ($=0$) over $\Fq$, and is called a vanishing polynomial. We
denote by $J_0 = \langle x_1^q - x_1, \dots, x_d^q - x_d \rangle$ the
ideal of all vanishing polynomials in $\Fq[x_1, \dots, x_d]$. When $q
= 2^k, x^q - x = x^q + x$ as $-1 = +1$ over $\Fkk$.

Gr\"obner bases can be used to {\it eliminate} (i.e. quantify) variables from an
ideal. Given ideal $J = \langle f_1,\dots,f_s\rangle \subset \mathbb
F_{q}[x_1,\dots,x_d]$, the $l^{th}$ elimination ideal $J_l$ is the
ideal of $\Fq[x_{l+1}, \dots, x_d]$ defined by $J_l = J \cap
\Fq[x_{l+1}, \dots, x_d]$. Variable elimination can be achieved 
by computing a Gr\"obner basis of $J$ w.r.t. elimination orders: 
\begin{Theorem}
\label{thm:elim}
(Elimination Theorem\cite{ideals:book}) Let $J\subset \mathbb
  F_{2^k}[x_1,\dots,x_d]$ be an ideal and let $G$ be a Gr\"obner basis
  of $J$ with respect to a lexicographic (LEX) ordering where
  $x_1>x_2>\cdots>x_d$. Then for every $0\leq l\leq d$, the set $G_l =
  G\cap\mathbb F_{2^k}[x_{l+1},\dots,x_d]$ is a Gr\"obner basis of
  the $l$-th elimination ideal $J_l$.
\end{Theorem}
We describe an application of elimination ideals using following example borrowed from \cite{ideals:book}:
\begin{Example}
Consider polynomials $f_1: x^2-y-z-1;\ f_2:x-y^2-z-1;\ f_3:x-y-z^2-1$ and ideal $J = \langle f_1,f_2,f_3\rangle
\subset \mathbb C[x,y,z]$. Gr\"obner basis $G = GB(J)$ w.r.t. LEX term order equals to 
$g_1:x-y-z^2-1;\ g_2:y^2-y-z^2-z;\ g_3: 2yz^2-z^4-z^2;\ g_4:z^6-4z^4-4z^3-z^2$. From observation,
we find that the polynomial $g_4$ only contains variable $z$ ($x,y$ eliminated), and polynomials $g_2,g_3,g_4$ only contain variables
$y,z$ ($x$ eliminated). According to Theorem \ref{thm:elim}, $G_1 = G\cap\mathbb C[y,z] = \{g_2,g_3,g_4\}$
is the Gr\"obner basis of the $1^{st}$ elimination ideal of $J$ and $G_2 = G\cap\mathbb C[z] = \{g_4\}$ is the 
$2^{nd}$ elimination ideal of $J$, respectively.
\end{Example}

\subsection{Application of elimination theorem on circuit verification}
\label{sec:elim}
Assume that we are given a circuit (combinational component) with input $A = (a_0,\dots,a_{k-1})$ and output 
$R = (r_0,\dots,r_{k-1})$ (both can be represented
by word level variables in $\Fkk$). We can describe this circuit with an elimination ideal $J+J_0$, where
$J$ is the ideal generated by the polynomials corresponding to circuit gates and $J_0$ is the ideal of vanishing polynomials.
The authors of \cite{timDAC} showed that for any combinational
logic block, a canonical word-level polynomial representation can be
derived through \Grobner bases computed with elimination orders:
\begin{Lemma}
(From \cite{timDAC}) Given a combinational circuit $C$ with $k$-bit
  input $A = (a_0, \dots, a_{k-1})$ and $k$-bit output $R = (r_0, \dots,
  r_{k-1})$. Denote by $x_1, \dots, x_d$ all the bit-level
  variables of   $C$. Let $J = \langle f_1, \dots, f_s \rangle \subset
  \Fkk[x_1, \dots, x_d, R, A]$ denote all the polynomials corresponding to the
  logic gates of the circuit. Let $J_0 = \langle x_1^2 - x_1, \dots,
  x_d^2 - x_d, R^q - R, A^q - A \rangle$ be the vanishing ideal, so
  that $J + J_0 = \langle f_1, \dots, f_s, ~~ x_1^2 - x_1, \dots,
  x_d^2 - x_d, R^q - R, A^q - A \rangle$. Compute \Grobner basis $G =
  GB(J + J_0)$ w.r.t. lex term order with $x_1 > x_2 > \dots > x_d > R
  > A$. Then $G_d = G \cap \Fkk[R, A]$ eliminates the internal
  variables $x_1, \dots, x_d$ of the circuit. $G_d$ also contains the
  word-level polynomial $R = \F(A)$ which canonically represents the
  function of the circuit with only word level variables $R$ and $A$.
\end{Lemma}
This lemma shows an application of GB computations over an elimination ideal.
Since it abstracts the function of a combinational circuit, we call the term order
$primary~inputs~and~intermediate~variables~>~word~level~output~>~word~level~inputs$
as \textit{abstraction term order} (ATO).
If we further eliminate word-level input, the result will be a polynomial containing only 
the word-level output variable. In a sequential circuit
such as Ex.\ref{ex:motiv}, the output of combinational logic serves as the next state variable. Polynomial $g_T$ in
the example is the desired projection; i.e. using GB computation on elimination ideal and eliminating to NS
variables provides us the canonical representation of reachable states in next time frame.

% Machine traversal is key for many verification techniques, e.g. to check
% the equivalence of 2 FSMs, we can observe whether the output responses are the same at
% every step of traversal. An explicit traversal is usually infeasible, here we use 
% implicit state enumeration (BFS traversal) based on Boolean formulas to implement a machine traversal.
% The algorithm is as follows:
% \begin{algorithm}[hbt]
% \SetAlgoNoLine
%  \KwIn{Transition functions $\Delta$, initial state $S^0$}
% 
%   $from^0 = reached = S^0$\;
%   \Repeat{$new^i == 0$}
%   {
%   	$i \gets i + 1$\;
% 	$to^i \gets$Img$(\Delta, from^{i-1})$\;
% 	$new^i \gets to^i \cap \overline{reached}$\;
%   	$reached \gets reached \cup new^i$\;
% 	$from^i \gets new^i$\;
%   }
% \Return{$reached$}
% \caption {Breadth-first Traversal Algorithm for Reachability Analysis of FSMs}\label{alg:BFS}
% \end{algorithm}
% The main computation in this algorithm is the \textit{image function}. Img$(\Delta,from^{i-1})$
% denotes the forward image of the set $from^{i-1}$ under the transition function $\Delta$.
% Let  $\Delta_i$ denote the transition relation 
% for $i^{th}$ bit of output $T$ (denoted by $t_i$), and it is described by a Boolean function. We can obtain the transition relation 
% for bit-vector $T$: $Tran(s_0,s_1,x,t_0,t_1) = \bigwedge_{i=1}^{2}(t_i\ \bar{\oplus}\ \Delta_i)$. Assume present states
% are represent by Boolean formulas $PS(s_0,s_1)$, then the image function is written as
% $\text{Img}(Tran,\ PS) = \exists_{s_0,s_1}\exists_{x}[Tran(s_0,s_1,x,t_0,t_1)\land PS(s_0,s_1)]$, where
% $\exists_x f$ denotes the existential quantification of $f$ w.r.t. $x$.
% \begin{Example}
% We use implicit state enumeration based on Boolean formulas to traverse FSM in Fig.\ref{fig:fsm}.
% Initial state $\{00\}$ can be represented by Boolean formula $C(s) = \overline{s_0}\cdot \overline{s_1}$.
% Transition function for $NS$ variables are 
% $$t_0\overline{\oplus}\Delta_0 = t_0\overline{\oplus}(\overline{x}\overline{s_0}\overline{s_1}+s_0s_1)$$
% $$t_1\overline{\oplus}\Delta_1 = t_1\overline{\oplus}(x\overline{s_0}+s_0\overline{s_1})$$
% \end{Example}

### 2015-05-01 techwriting/proposal/reacha.tex
all write-up from virtue machine

\section{Reachability Analysis of Finite State Machines (FSMs) using Algebraic Geometry}
% \subsection{An example illustrating our proposed approach}
In our approach we use symbolic state reachability with algebraic
geometry concepts. It is an abstraction based on word operand
definition of datapaths in circuits, and it can be applied
to arbitrary FSMs by bundling a set of bit-level variables together as
one or several word-level variables.  The abstraction polynomial,
encoding the reachable state space of the FSM, is obtained through
computing a GB of an elimination ideal using elimination term order
based on Theorem \ref{thm:elim}.

The motivation behind our approach can be described as follows:
\begin{itemize}
\item Any finite set of points is the variety (solutions) of a
  polynomial ideal. Since the state-space of a FSM is finite, it can
  be construed as the variety of an ideal corresponding to the
  function (circuit implementation, gates) of the   sequential circuit.
\item Using canonical representations of polynomial ideals (Gr\"obner
  bases), the transition functions and the sets of states can be used
  for FSM traversal and property checking. 
\item In our work, Gr\"obner bases are considered in lieu of BDDs or
  SAT-solvers. Furthermore, instead of working in the Boolean domain
  $\B$, we model the circuit and the properties over the Galois field
  $\Fkk$. This allows to perform reachability analysis at the level of
  $k$-bit words, as $\mathbb{F}_2 \subset \Fkk$.

\end{itemize}

Reachability analysis is a useful tool when checking sequential
equivalence as well as other invariants. 
%With word-level state
%variables, we can do implicit state enumeration to provide a picture
%of  reachable states. 
Conceptually, the state-space of a FSM is traversed in a breadth-first
manner, as shown in Algorithm \ref{alg:BFS}. % \cite{KallaPartialScan}: 
In general, the algorithm operates on the FSM 
$\mathcal{M} = (\sum, O, S, S^0, \Delta, \Lambda)$ underlying a
sequential circuit. In such cases, the transition function $\Delta$
and the initial states are represented and manipulated using Boolean
representations such as BDDs or SAT solvers. The variables $from,
reached, to, new$ represent characteristic functions of sets of
states. Starting from the initial state $from^i = S^0$, the algorithm
computes the states reachable in 1-step from $from^i$ in each iteration.
In line 4 of algorithm \ref{alg:BFS}, the {\bf image computation} is
used to compute the 1-step next reachable states. In
\cite{KallaPartialScan}, the author implicitly represents the states
and transition relations by Boolean functions and Boolean vectors,
respectively.

Let us describe the application of the algorithm on the  FSM circuit
of Fig.\ref{fig:fsm}. {\it We will first describe its operation at the
Boolean level, and then describe how this algorithm can be implemented
using algebraic geometry at word level.}

In Line 1 of the BFS algorithm, assume that the initial state
is $S_3$ in Fig.\ref{fig:fsm}(b), which is encoded as 
$S_3 = \{11\}$. Using Boolean variables $s_0, s_1$ for the present
states, $from^0 = s_0\cdot s_1$ is represented as a Boolean formula.

The {\it transition function} $\Delta$ is given by Boolean equations
  of the flip-flops of the circuit: $t_i = \Delta_i(s, x)$, where
  $t_i$ is the next state variable, $s$ represents the present   state
  variables and $x$ represents the input variables. 
The {\it Transition relation of the FSM} is then represented as: 
\begin{equation} 
T(s, x, t) =   \prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)
\end{equation}
where $n$ is the number of flip flops, and $\overline{\oplus}$ is XNOR
operation. Let $from$ denote the set of initial states, then the
image of the initial states, under the transition function $\Delta$ is
finally computed as:
\begin{align}
\label{img}
to = \text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
  \cdot from ] = \exists _s ~\exists _x ~\prod_{i=1}^{n} (t_i
\overline{\oplus } \Delta_i)\cdot from
\end{align}

Here, the notation $\exists x (f)$ represents the {\it existential
  quantification of $f$ w.r.t. variable $x$}.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm for Reachability Analysis of FSMs}\label{alg:BFS}
\end{algorithm}

\begin{Example}
For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
functions of the machine as:

\begin{align*}
\Delta_1: & ~t_0 \overline{\oplus} ((\overline{x + s_0 + s_1}) + s_0 s_1)\\
\Delta_2: & ~t_0 \overline{\oplus} (\overline{s_0}x + \overline{s_1}s_0)\\
from:     & ~from^0 = s_0\cdot s_1
\end{align*}

When the formula of Eqn. (\ref{img}) is applied to compute 1-step
reachability, $to = \exists _{s_0, s_1, x} (\Delta_1 \cdot \Delta_2
\cdot from^0)$, we obtain $to = \overline{t_0}\cdot t_1$, which denotes
the state $S_1 = \{01\}$ reached in 1-step from $S_3$.

In the next iteration, the algorithm uses state $S_1 = \{01\}$ as the
current (initial) state, and computes $S_2 = \{10\} = t_0\cdot
\overline{t_1}$ as the next reachable state, and so on. 
\end{Example}

%% Let  $\Delta_i$ denote the transition function for $i^{th}$ bit of the
%% output $T$
%% (denoted by $t_i$), and it is described by a Boolean function. We can
%% obtain the transition relation  for bit-vector $T$:
%% $Tran(s_0,s_1,x,t_0,t_1) =
%% \bigwedge_{i=1}^{2}(t_i\ \bar{\oplus}\ \Delta_i)$. Assume present
%% states are represent by Boolean formulas $PS(s_0,s_1)$, then the image
%% function is written as $\text{Img}(Tran,\ PS) =
%% \exists_{s_0,s_1}\exists_{x}[Tran(s_0,s_1,x,t_0,t_1)\land
%%   PS(s_0,s_1)]$, where $\exists_x f$ denotes the existential
%% quantification of $f$ w.r.t. $x$.

Our objective is to model the transition functions $\Delta$ as a
polynomial ideal $J$, and to perform the image computations (Algorithm
\ref{alg:BFS}, line 4) using Gr\"obner bases over Galois fields. {\bf
This requires to perform quantifier elimination; which can be
accomplished using the Gr\"obner basis computation over Galois fields
by using elimination ideals} \cite{gao:qe-gf-gb}. Finally, the set
union,  intersection and complement operations are also to be
implemented in algebraic geometry.

{\bf FSM Traversal at word-level over $\Fkk$:} 
The state transition graph (STG) shown in Fig.\ref{fig:fsm}(a) uses a
2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
S_3\}$. We map these states to elements in $\mathbb{F}_{2^2}$, where
$S_0 = 0, S_1 = 1, S_2 = \alpha, S_3 = \alpha+1$ where $\alpha^2 +
\alpha + 1 = 0$.

{\it Initial state:} In Line 1 of the algorithm, the initial state is
specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
S - 1 - \alpha$. Notice that if we consider the ideal generated by the
initial state polynomial, $I = \langle f\rangle$, then its variety
$V(I) = 1+\alpha$, corresponds to the state encoding $S_3 = \{11\} =
1+\alpha$, where a polynomial in word-level variable $S$ encodes the
initial state. 
% with only one generator $f$, its variety
% $V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 1+\alpha\}$, which equals to $\{1+\alpha\}$, the only
% valid value $S_3$ can take.

%Using theorems from Section \ref{sec:elim}, we implement the image
%function by a GB computation on elimination ideal. Ex.\ref{ex:motiv}
%is an example for our implementation of image function, i.e. one-step
%reachability.

{\bf Set operations:} When executing Line 5 and Line 6, we use
\textbf{union}, \textbf{intersection} and \textbf{complement} of
varieties over $\mathbb{F}_{2^k}$:

\begin{Definition}
\label{def:sum}
({\bf Sum/Product of Ideals}) If $I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are 
ideals in $\mathbb F[x_1, \dots, x_n]$, then the {\bf sum} of $I$ and $J$ is defined as
$$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$$ And the {\bf product} of $I$ and $J$ is defined
as
\begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle \nonumber
  \end{equation}
\end{Definition}

With concepts of ideal sums and products, we can obtain the
intersection and union of affine varieties as:
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf
  V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) =
{\bf V}(I) \bigcup {\bf V}(J)$. 
\end{Theorem}

In Line 5 of Alg.\ref{alg:BFS}, we need to compute the complement of a
set of states. Assume that $J$ denotes a polynomial ideal whose
variety $V(J)$ denotes a set of states. We require the computation of
another polynomial ideal $J'$, such that $V(J') =
\overline{V(J)}$. We have recently proved (the proofs are omitted)
that this computation can be performed using the concept of {\bf ideal
  quotient:}

\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
  F[x_1, \dots, x_n]$, then $I:J$ is the set
  \begin{equation}
  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}\nonumber
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}

% However, complement of variety cannot be easily dealt by simple arithmetic on polynomial generators.
% For non-trivial cases we can only prove that 
% \begin{Theorem}
% Let $I, J$ be ideals in $\mathbb F[x_1,\dots,x_n]$, then 
% $${\bf V}(I:J) \supset {\bf V}({\bf I}({\bf V}(I) - {\bf V}(J)))$$
% \end{Theorem}
% Fortunately for most hardware verification cases, the form of polynomial generators are restricted.
% A proposition has been proved by \cite{jinpeng} that after adding {\bf vanishing polynomials ideal}
% the new composed ideal implies following corollary:

For elimination ideals in circuit verification problems, we can obtain
the complement of a variety through the following result:

\begin{Theorem}
\label{thm:quotient}
Let $I, J$ be ideals with vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
$${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$
\end{Theorem}

Let $J_0 = \langle x_1^{2^k} - x_1, \dots, x_n^{2^k} - x_n \rangle$
denote the ideal of all vanishing polynomials in $\Fkk$. Then, we have
$V(J_0) = (\Fkk)^{n}$; i.e. the variety of vanishing ideal contains
all possible valuations of variables, so it constitutes the {\bf
  universal set}. Subsequently, the {\bf absolute complement} can be
computed as:

\begin{Corollary}
$$V(J') = \overline{{\bf V}(J)} = {\bf V}(J_0:J)$$

\end{Corollary}

In other words, the ideal $J'$, computed as $J' = J_0:J$ is such that
$V(J') =\overline{V(J)}$.

Since we can add vanishing polynomials for state sets
$from^i, to^i, reached$, it is feasible to use above corollary. We
will demonstrate the application of this important concept through an
example. 
% 
% {\bf Now, STEP BY STEP, SHOW THE FULL Blown traversal COMPUTATION
%   HERE.... Make sure to also show how elimination order with GB
%   computation is also used...} \vspace{0.5in}

Assume $I = \langle f\rangle  = \langle T^2 +
(1+\alpha)\cdot T+\alpha\rangle $,  we can add vanishing polynomial
$T^4 - T$ to this ideal such that its variety  $V(I) = \{a\ |\ a \in
\mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$. For complement set
of variety $\{1, \alpha\}$, the universal set is the variety of ideal
of vanishing polynomial $V(\langle T^4-T\rangle ) =
\{0,1,\alpha,1+\alpha\}$, 
so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
the result is $\{0,1+\alpha\}$.

According to the discussion above, the BFS traversal can be completely implemented using algebraic geometry.
as in Algorithm \ref{alg:univa}.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\F(S)$}

  $from^0 = reached = \F(S)$\;
  \Repeat{$new^i == 1$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$GB w/ elimination term order$\langle J_{ckt},J_0, from^{i-1}\rangle$\;
	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
	$from^i \gets new^i(S\setminus T)$\;
  }
\Return{$reached$}
\caption {Algebraic Geometry based Traversal Algorithm}\label{alg:univa}
\end{algorithm}

Here $from^i, to^i, new^i$ are all univariate polynomials in variables $S$ or $T$, due to GB computation
with elimination term order.

\begin{Example}
We apply Algorithm \ref{alg:univa} to Ex.\ref{ex:motiv} to execute the FSM traversal. Let the initial state
$from^0 = S$ (i.e. state $\{00\}$), in the first iteration we compose an elimination ideal $J$ with

\begin{minipage}[h]{0.4\textwidth}
\begin{align*}
&f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
&f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
&f_3: S - s_0 - s_1\\
&f_4: T - t_0 - t_1
\end{align*}
$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$
\end{minipage}
\begin{minipage}[h]{0.6\textwidth}
\begin{align*}
&f_5: x^2-x\\
&f_6: s_0^2-s_0, ~~f_7: s_1^2-s_1\\
&f_8: t_0^2-t_0, ~~~f_9: t_1^2-t_1\\
&f_{10}: S^4-S, ~f_{11}:T^4-T
\end{align*}
$\ \ \ \ \ \ \ \ \ \ \  \ J_0 = \langle f_5,f_6,\dots,f_{11}\rangle$
\end{minipage}

and $from^0$. Compute the reduced GB for $J = J_{ckt}+J_0+\langle from^0\rangle$,
with elimination term order
$$\{x,s_0,s_1,t_0,t_1\}~(all~PI~and~bit~level~variables)~>~S~(PS~word)~ >~ T~(NS~word)$$
the result contains
a polynomial generator with only $T$, we assign it to next state $$to^1 = T^2+(\alpha+1)T+\alpha$$ denoting
$\{01,10\}$. Since formerly reached state $reach = T$, its complement is $$\langle T^4-T\rangle:\langle T\rangle
= T^3+1$$ (i.e. states $\{01,10,11\}$). Then the newly reached state in this iteration is
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle$$ We add these states
to formerly reached states $$reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$
(i.e. states $\{00,01,10\}$). We update the present states for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha$$

In the second iteration, we compute the reduced GB with the same term order for ideal $J = J_{ckt}+J_0+\langle from^1\rangle$.
It includes a polynomial generator $$to^2 = T^2+\alpha T$$ denotes states
$\{00,10\}$. The complement of $reached$ is $$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
= T + 1+\alpha$$ (i.e. states $\{11\}$). We compute the newly reached state 
$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$$ 
It means the newly reached state is empty, thus a fixpoint has been detected. The algorithm terminates and returns
$$reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$ as final reachable states.

% 
% $J = \langle t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1), t_1 - (xs_0+x+s_0s_1+s_0), S - s_0 - s_1, T - t_0 - t_1, S \rangle$
% and a vanishing polynomial ideal $J_0 = \langle x^2-x, s_0^2-s_0, s_1^2-s_1, t_0^2-t_0, t_1^2-t_1, S^4-S, T^4-T\rangle$
% The elimination term order is $\{x,s_0,s_1,t_0,t_1\} > S > T$. Compute a reduced GB for $J+J_0$, the result has 
% a polynomial generator with only $T$, we assign it to next state $to^1 = T^2+(\alpha+1)T+\alpha$ denoting
% $\{01,10\}$. Since formerly reached state $reach = T$, its complement is $\langle T^4-T\rangle:\langle T\rangle
% = T^3+1$ (i.e. states $\{01,10,11\}$). Then the newly reached state in this iteration is
% $\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle$. We add these states
% to formerly reached states $reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$
% (i.e. states $\{00,01,10\}$). We update the present states for next iteration $from^1 = S^2+(\alpha+1)S+\alpha$.
% In the second iteration, we just replace 1 generator in $J+J_0$: $S$ with $S^2+(\alpha+1)S+\alpha$, and elimination
% term order keeps the same. The reduced GB includes a polynomial generator $to^2 = T^2+\alpha T$ denotes states
% $\{00,10\}$. The complement of $reached$ is $\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
% = T + 1+\alpha$ (i.e. states $\{11\}$). We compute the newly reached state 
% $\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$, then the algorithm terminates and returns
% $reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$ as final reachable states.
\end{Example}

{\bf Significance for using GB:} Reduced GB is a unique, minimal and
{\bf canonical} representation of the circuit's function. Starting
from a certain initial state and using a reduced GB to represent the
transition function, reachable states can be computed and represented
canonically. Then it becomes possible to identify  when a fixpoint
is reached (termination of the algorithm) by performing an equality
check of polynomial ideals.

\subsection{Research objective: To make the approach scalable}

At the time of writing this proposal, the theoretical concepts behind
the word-level FSM traversal have been studied and validated. A
completed algorithm has been devised and preliminary experiments were
conducted for reachability analysis using the {\sc Singular} computer
algebra tool \cite{DGPS}. While the approach is shown to work
correctly and completely, it is practically infeasible. The complexity
lies in the Gr\"obner basis computation.

Over Galois fields of the type $\Fkk$, computing the reduced Gr\"obner
basis of the ideal $(J + J_0)$ requires both memory and space
${2^{k}}^{O(n)}$; where $J$ is an arbitrary polynomial ideal, $J_0$ is
the vanishing ideal, and $n$ is the number of variables in the system
(circuit). In this research, I wish to draw inspirations from the work
of \cite{jinpeng} and \cite{timDAC} to overcome the
complexity of GB computations.

In \cite{jinpeng}, the authors analyze the given circuit topology,
and derive a term order to represent the polynomials. They show that
this term order renders the set of polynomials itself a Gr\"obner
basis of the ideal. In \cite{timDAC}, the authors use a similar
concept to derive a {\it refinement of the abstraction term order}
(RATO) to simplify the Gr\"obner basis computation. However, the above
concepts are applied to combinational circuit verification. I will
investigate how similar techniques can be explored for the
computations that are needed in Algorithm \ref{alg:univa} for
reachability analysis to make the approach scalable. I believe that it
is possible to make the approach scalable, as I have recently
demonstrated in a paper \cite{myDATE} %{\bf refer to the DATE paper here?}
which has been accepted for publication at the {\it Design Automation
  and Test in Europe Conference}, 2015.

\subsection{Research objective : Application to Unrolling Sequential Circuits for Sequential Arithmetic Verification}
In above discussions, we propose an implicit state enumeration algorithm implemented by GB computations over
an elimination ideal, which is usually applied on FSMs working as control units. For many arithmetic
datapath components, state enumeration is not the best way to verify the function of circuit. On the one hand, the signals preloaded
to the state variables in FSM are usually not specified such that we can hardly encode the initial states; on the other
hand, instead of the set of globally reachable states, arithmetic verification focuses on exact reached states
after $k$ clock cycles which reflects desired arithmetic function. Considering these reasons, we modify
our algebraic geometry based approach and apply it to implicit FSM unrolling.

% In above discussions, we propose an implicit state enumeration algorithm implemented by GB computation over
% an elimination ideal, which is usually applied on Mealy FSMs working as control unit. For many arithmetic
% datapath components, they should be modeled as Moore FSMs according to their structures. On the one hand, the signals preloaded
% to a Moore FSM are usually not specified such that we can hardly encode the initial states; on the other
% hand, instead of the set of globally reachable states, arithmetic verification focuses on exact reached states
% after $k$ clock cycles which reflects desired arithmetic function. Considering these reasons, we modify
% our algebraic geometry based approach and apply it to Moore FSM unrolling.

% % Figure Moore.eps
% \begin{figure}[hbt]
% \centering{
% %\begin{minipage}{12cm}
% \includegraphics[width=7.5in]{./Moore.eps}
% % \vspace{-0.2in}
% \caption{A simple Moore FSM and its state transition graph}
% %\end{minipage}
% \label{fig:Moore}}
% \end{figure}

Fig.\ref{fig:seqmodel}(b) shows a typical Moore FSM without primary inputs, $R$ is the word-level state variable. Let word-level indeterminate 
$R_{init}$ denote the initial preloaded value of $R$, and $R_{i}$ denote the evaluation of $R$ after $i$ clock cycles
(transitions). Transition relation $Tr:R_{i-1}\to R_i$ can be modeled similar to
the image function in Algorithm \ref{alg:BFS}. Assume that after $k$ clock cycles, the machine should give a state output
$R_{k} = \mathcal F(R_{init})$. To verify this result, we can implicitly unroll this FSM and execute transition
function at word level for $k$ times, which is $R_{k} = Tr(R_{k-1}) = Tr(Tr(\cdots Tr(R_{init})\cdots)) = Tr^k(R_{init})$.

Sequential Galois field (GF) multiplier is a typical sequential arithmetic component widely used in cryptographic systems.
Since there is no primary input in the design, it can be regarded as Moore FSM (Fig.\ref{fig:RHmulti}(a))
where $R = \sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
 \beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$. Notice that $\beta$ is the {\it normal element}
 which can be represented by a power of primitive element $\alpha:~\beta = \alpha^t$.
The states are encoded by evaluations of state variables, which can be implicitly verified 
by checking output-input function ($R = A\cdot B \pmod{P(\alpha)}$).

% Additionally
% the inner structure of GF multiplier is XOR-rich such that we can test our approach on very large
% (>100 bits datapath) circuits.
% 
% Let us briefly describe the fundamentals behind the design of normal
% basis sequential GF multipliers, so as to put in perspective the type
% of designs that have been verified in this paper. Let $R =
% \sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
% \beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$, then 
% \[
% R = A\cdot B = (\sum_{i=0}^{k-1} a_i \beta^{2^{i}}) (\sum_{j=0}^{k-1}
% b_j \beta^{2^{j}})  =
% \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}\nonumber 
% \]
% 
% The expressions $\beta^{2^i}\beta^{2^j}$ are called cross-product
% terms and they can also be represented in normal basis: 
% \begin{displaymath}
% \beta^{2^i}\beta^{2^j} =
% \sum_{n=0}^{k-1}\lambda_{ij}^{(n)}\beta^{2^n}, \ \ \lambda_{ij}^{(n)}
% \in \Ftwo. 
% \end{displaymath}
% 
% From the above two equations, one can see that the expression for the
% $n^{th}$ digit of product $R = (r_0, \dots, r_n, \dots r_{k-1})$ is:
% \[
% r_n = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(n)}a_ib_j = A
% \cdot M_n \cdot B^T, ~~0 \leq n \leq k-1
% \]
% 
% where $M_n = (\lambda_{ij}^{(n)})$ is a binary $k \times k$ matrix over
% $\Ftwo$, and it is called the $\lambda$-matrix. 
%The collection of all
%$\{M_n\}$  $\lambda$-matrices is called the multiplication
%table. 
% Moreover, let $r_n = A \cdot M_n \cdot B^T$.
% Then $r_{n-1} = A \cdot M_{n-1} \cdot B^T = rotate(A) \cdot M_n
% \cdot rotate(B)^T$. This implies that $M_n$ is generated by right
% and down cyclic shifting $M_{n-1}$. Therefore, the hardware design
% of sequential GF multipliers is based on mappings of $A\cdot M_n
% \cdot B^T$ into AND-XOR gates and cyclic shift operations.

% This paper verifies the implementation of two distinct 
% architectures of {\it sequential multipliers with parallel output
%   (SMPO)}, namely: i) the Agnew-SMPO  \cite{agnew1991implementation}
% by G. B. Agnew, which is a straight-forward implementation of the
% $\lambda$-matrix; and ii) the more recent, more complicated, yet very
% efficient RH-SMPO \cite{RHmulti}, by Reyhani-Masoleh and Hasan,
% depicted in Fig. \ref{fig:RHmulti}.

In this proposal we use the verification of a sequential GF multiplier -- 
a 3-bit RH-SMPO invented by Reyhani-Masoleh and Hasan\cite{RHmulti} -- to illustrate our approach.
The normal element is given as $\beta = \alpha^3$, and primitive polynomial $P(\alpha) = \alpha^3+\alpha+1$.
The gate-level circuit is depicted in Fig.\ref{fig:RHmulti}(b).

\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=7.5in]{./new_multi.eps}
% \vspace{-0.2in}
\caption{A 3-bit RH-SMPO and its Moore FSM model}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}

We follow  the typical sequential GF circuit model with word-level variables $A, B, R$
denoting {\it present states (PS)} and $A', B', R'$ denoting {\it next
  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
store the result, and $A, B\ (A', B')$ store input operands. {\it E.g.,}
for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
0$) are the initial values (operands) loaded into the registers,  and
$R = \F(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
result after $k$-cycles. Our approach aims to find this polynomial
representation for $R$.

Each gate in the combinational logic is represented by a Boolean
polynomial. To 
this set of Boolean polynomials, we append the polynomials that define
the word-level to bit-level relations for PS and NS variables ($A =
\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
as ideal $J = \langle 
f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
  of the circuit. The ideal of vanishing polynomials $J_0$ is also included, and
then the implicit FSM unrolling problem is setup for abstraction.

The configurations of the flip-flops are the states of the
machine. {\it Since the set of states is a finite set of points, we
can consider it as the variety of an ideal related to the circuit
}. Moreover, since we are interested in
the {\it function encoded} by the state variables (over $k$-time
frames), we can {\it project this variety} on the word-level state
variables, starting from the initial state $A_{init}, B_{init}$.
Projection of varieties (geometry) corresponds to elimination ideals
(algebra), and can be analyzed via \Grobner bases. Therefore, we
employ a \Grobner basis computation with abstraction term order (ATO)\cite{timDAC}: we use a {\it lex term
  order} with {\it bit-level variables} 
$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
allows to eliminate all the bit-level variables 
%(corresponding to the combinational logic and the state variables),
%so as to 
and derives a representation only in terms of words. 
Consequently, $k$-successive \Grobner basis computations implicitly
unroll the machine, and provide word-level algebraic $k$-cycle
abstraction for $R'$ as $R' = \F(A_{init}, B_{init})$.

Algorithm
\ref{alg:modified} describes our approach.  In the algorithm, $from_i$
and $to_i$ are polynomial ideals whose varieties are the valuations of
word-level variables $R, A, B$ and $R',A',B'$ in the $i$-th iteration;
and the notation ``$\setminus$'' signifies that the $NS$ in iteration
$(i)$ becomes the $PS$ in iteration $(i+1)$. Line 5 computes the Gr\"obner 
basis with the abstraction term order.  Line 6 computes the elimination 
ideal, eliminating the bit-level variables and representing the set of 
reachable states up to iteration $i$ in terms of the elimination ideal. 
These computations are analogous to those of image computations performed in FSM reachability. 
%The forward image
%$to^{i}$ is computed using \Grobner bases with ATO.

\vspace{-0.1in}
\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
   state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
\DecMargin{1em}
\vspace{-0.1in}
\begin{Example}
\label{ex:RHSMPO}

We demonstrate our approach to verify the 3-bit RH-SMPO circuit of
Fig.\ref{fig:RHmulti}. The normal element $\beta$ in
$\mathbb{F}_{2^3}$ is known to be $\beta = \alpha^3$, where $\alpha$
is the primitive element. The circuit can be described with an ideal by translating
AND and XOR gates accordingly. For the first iteration:
\begin{align*}
J = &d_0+b_2\cdot a_2,
c_1+a_0+a_2,
c_2+b_0+b_2,
d_1+c_1\cdot c_2,\\
&e_0+d_0+d_1,
e_2+d_1,
r_0'+r_2+e_0,
r_1'+r_0,
r_2'+r_1+e_2,\\
&A+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
B+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12},\\
&R+r_0\alpha^3+r_1\alpha^6+r_2\alpha^{12},
R'+r_0'\alpha^3+r_1'\alpha^6+r_2'\alpha^{12};
\end{align*}
The last 4 polynomials are translated from the definition of word-level variables.
This represents ideal ``$J$" from line 5 in Algorithm
\ref{alg:modified}. ``$J_0$" is the ideal of vanishing polynomials in all bit-level
variables (e.g. $a_0^2-a_0$) and word-level variables (e.g. $A^8-A$). ``$from_{i-1}$"
represents the set of current states for iteration $i$.
In the first iteration, $from_0 = \{R, A_{init}+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
B_{init}+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12}\}$.

After the GB computation is performed with ATO, as line 6 in Algorithm \ref{alg:modified},
we find a polynomial in variables $R', A_{init}, B_{init}$ in $to_1 : 
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
$.

Line 7 in Algorithm \ref{alg:modified} simply replaces NS output $R'$ with PS output
$R$ in this example; so in second iteration $from_1 = \{
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
, A_{init}+a_2\alpha^3+a_0\alpha^6+a_1\alpha^{12},
B_{init}+b_2\alpha^3+b_0\alpha^6+b_1\alpha^{12}\}$.

% By continue executing the loop,
Finally, after 3 iterations we obtain: $to_3 = \{ \mathbf{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
as the image. The final result is $from_3(R_{final}) = R_{final}+A_{init}\cdot
B_{init}$, which verifies the multiplier. 
\end{Example}
Using our approach we successfully verified the function of an 100-bit RH-multiplier,
which implies its capability to handle the verification of large designs. The results
are published in \cite{myDATE}.

### 2015-05-01 techwriting/proposal/slides.tex
all write-up from virtue machine

%\documentclass[pdf,final,colorBG,slideColor]{prosper}
\documentclass[xcolor=dvipsnames]{beamer}
%%\usetheme{default}
\usecolortheme[named=Maroon]{structure}
%\usetheme{Boadilla}
\usetheme{Madrid}
\useoutertheme{default}%[footline=empty]{infolines}

% \usepackage{helvet}
% \usepackage{enumerate}
% \usepackage{amsmath}
% \usepackage{amsfonts}
% \usepackage{graphicx}
% \usepackage{ulem}
% \usepackage{multirow}
\usepackage{comment}
\usepackage{xspace}
%\usepackage{columns}

\usepackage[absolute,overlay]{textpos}
\usepackage[ruled,linesnumbered]{./algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\RequirePackage{algorithmic}
%\RequirePackage{algorithm}
% \renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
% \renewcommand{\algorithmicensure}{\textbf{Outputs:}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{lemma}
%\newtheorem{corollary}{Corollary}
%\newtheorem{proposition}{Proposition}
%\newtheorem{Q}{Question}
%\newtheorem{Exa}{Example}
%\newtheorem{Definition}{Definition}

\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Zkk}{{\mathbb{Z}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
%\newcommand{\Grobner}{Gr\"{o}bner}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\F}{{\mathcal{F}}}
\newcommand{\B}{{\mathbb{B}}}

\title[Ph.D Proposal]{Sequential Circuit Verification at Word Level using Algebraic Geometry}

\author[Xiaojun Sun]{Xiaojun Sun}

%\email{rostamian@umbc.edu}
\institute[Univ. of Utah]{
%\includegraphics[height=17mm]{/Users/Kalla/teaching/Comp-Algebra-Course/lectures/old_ulogo.eps}\\
Ph.D Candidate\\
Electrical and Computer Engineering, University of Utah\\
xiaojuns@ece.utah.edu\\
\ \\
\ \\
{\bf Ph.D's Dissertation Proposal}\\
}

\date{}
%\slideCaption{}

%% Images
%\pgfdeclareimage[width=.4in]{fg:logo}{/Users/Kalla/teaching/Comp-Algebra-Course/lectures/old_ulogo.eps}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%----------- titlepage ----------------------------------------------%
\begin{frame}[plain]
  \titlepage

\end{frame}

%\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item Sequential circuit verification requires reachability analysis
\item From bit-level to word-level
\item One step back: Sequential arithmetic circuit verification
\item New inspiration: UNSAT core extraction using algebraic geometry
\item The plan
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Agenda}}

\begin{itemize}
\item Focus
	\begin{itemize}
	\item Implicitly analyze the reachability of a sequential circuit at word level
	\item Algebraic geometry to assist in sequential circuits verification and abstraction refinement
	\end{itemize}
\item Motivation
	\begin{itemize}
	\item Data flow = word-level info
	\item Conventional techniques are bit-level
	\item Need bit-level to word-level abstraction
	\item Verification $\hookleftarrow$ Reachability $\hookleftarrow$ Abstraction (at word level?)
	\end{itemize}
\item Target problems 
	\begin{itemize}
	\item Given: sequential circuit (FSM), with $k$-bit state variables, property for verification
     \bi
		\item Perform sequential equivalence/property checking at word level
        \item Identify UNSAT cores in word-level problems
     \ei
	\end{itemize}
	\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large {Agenda(2)}}
\bi
\item Approach: {\bf Algebraic geometry techniques}
	\begin{itemize}
	\item Model system over $\Fkk~~\implies$ Represent at word-level; compatible with bit-level:$\mathbb F_2\subset \Fkk$ 
	\item  \Grobner basis methods + Elimination ordering + BFS traversal
	\ei
\item  Challenge: Discover efficient algorithm to implement image computations, 
        set operations, UNSAT proofs, etc. at word level
\item  Contributions: 
    \bi
    \item Polynomial abstraction + algebraic geometry techniques 
         applied to reachability analysis
    \item A new algorithm based on \Grobner basis computation  to extract UNSAT cores
    \ei
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large {Motivation}}
\vspace{-0.2in}
%\ptsize{10}

\begin{itemize}
\item Importance of reachability analysis in sequential circuits verification
	\begin{itemize}
	\item Circuits $\to$ state machines; errors $\to$ bad states
	\item Bad states are reachable $implies$ errors affect circuit behavior
	\end{itemize}
\end{itemize}

\begin{itemize}
\item Advantages exploiting word-level verification
	\begin{itemize}
	\item Many circuit datapaths/system models are described at word level
	\item Reduce state space, avoid ``bit-blasting"
	\end{itemize}	
\end{itemize}

\begin{itemize}
\item Why use algebraic geometry?
	\begin{itemize}
    \item A symbolic representation for bit-level \& word-level: $\Fkk$
    \item Algebraic geometry: allows reasoning on solutions by analyzing polynomials, solution $\gets$ states
	\item Recent work [Tim, {\it Abstraction using GB}, DAC'14] [Lv, {\it Equivalence}, TCAD'13] [M. Brickenstein, {\it GB for Boolean Poly}, J.Symb.Comp.'09]
	shows it is practical to apply algebraic geometry to circuit verification
	\end{itemize}
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item \alert{Sequential circuit verification requires reachability analysis}
\item From bit-level to word-level
\item One step back: Sequential arithmetic circuit verification
\item New inspiration: UNSAT core extraction using algebraic geometry
\item The plan
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Finite state machine (FSM)}}
\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{./seqmodel.eps}
 \vspace{-0.2in}
\caption{FSM models of sequential circuits}
\label{fig:seqmodel}}
\end{figure}
\bi
\item (a) A typical model for sequential circuits
\item (b) sequential datapath without primary inputs
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large {State transition graph (STG) and breadth-first search (BFS) state space traversal}}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{stg_fig.eps}
\caption{State Transition Graph}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\bi
\item Initial state: $\{00\}$
\item Iteration 1: 
	\bi
	\item Start from $\{00\}$
	\item One-step transition: $\{01,10\}$
	\item Newly reached: $\{01,10\}$
	\ei
\item Iteration 2: 
	\bi
	\item Start from $\{01,10\}$
	\item One-step transition: $\{00,10\}$
	\item Newly reached: $\emptyset$
	\ei
\item All reachable states detected. Final reached states: $\{00,01,10\}$
\ei
\end{column}
\end{columns}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{BFS traversal with gate-level circuits}}
\begin{figure}[hbt]
\centerline{
\includegraphics[scale=0.35]{fsm_fig.eps}
}
\caption{A Gate-level Circuit corresponding to last STG}
\end{figure}
An algorithmic approach is needed to implement BFS traversal
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Breadth-First Traversal Algorithm based on Boolean formulas}}
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Breadth-First Traversal Algorithm based on Boolean formulas(2)}}
\bi
\item Image function: $\text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
  \land from ] = \exists _s ~\exists _x ~\bigwedge_
  {i=1}^{n} (t_i
\overline{\oplus } \Delta_i)\land from$
\item Initial state: $from^0 = \overline{s_0}\land\overline{s_1}~(\{00\})$
\item Transition function: 
\begin{align*}
\Delta_1: & ~t_0 \overline{\oplus} ((\overline{x \lor s_0 \lor s_1}) ~\lor~ s_0\land s_1)\\
\Delta_2: & ~t_1 \overline{\oplus} (\overline{s_0}\land x ~\lor~ \overline{s_1}\land s_0)
\end{align*}
\item Iteration 1: 
	\bi
	\item One-step transition $to^1 = \exists _{s_0, s_1, x} (\Delta_1 \land \Delta_2
	\land from^0) = \overline{t_0}\land t_1~\lor~t_0\land \overline{t_1}~(\{01,10\})$
	\item Newly reached $new^1 = (\overline{t_0}\land t_1~\lor~t_0\land\overline{t_1})~\land ~
	\overline{(\overline{t_0}\land\overline{t_1})} = \overline{t_0}\land t_1~\lor~t_0\land\overline{t_1}$
	\ei
\item Iteration 2: same fashion
\item Return value: $reached = \{\overline{t_0}\lor \overline{t_1}\}~(\{00,01,10\})$
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item Sequential circuit verification requires reachability analysis
\item From bit-level to word-level
	\bi
	\item \alert{All about math}
	\item Utilize the math tool: what does Tim get?
	\item My approach
	\ei
\item One step back: Sequential arithmetic circuit verification
\item New inspiration: UNSAT core extraction using algebraic geometry
\item The plan
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large {Galois Field Overview}}
%\vspace{-0.2in}
%\ptsize{10}
\textbf {Galois field}(GF) $\mathbb{F}_q$ is a finite field with $q$
elements, $q = p^k$
\begin{itemize}
\item Commutative Ring with unity, associate, distributive laws
\item Closure property: $+,-,\times$, inverse ($\div$)
\end{itemize}

\begin{itemize}
\item $\mathbb{F}_p \equiv (\mathbb{Z} ~\pmod{ p })$, where $p = $ prime, is a field
\begin{itemize}
\item $\mathbb{Z}_7=\{0,1,2,3,4,5,6\}$
\end{itemize}

\end{itemize}

Our interest: $\mathbb{F}_{q} = \mathbb{F}_{2^k}$, i.e. $q = 2^k$
\begin{itemize}
\item  $\mathbb{F}_{2^k}$: $k$-dimensional extension of  $\mathbb{F}_{2}$
	\begin{itemize}
	\item $k$-bit bit-vector, AND/XOR arithmetic
	\end{itemize}
\end{itemize}

To construct $\mathbb{F}_{2^k}$
\begin{itemize}
\item $\mathbb{F}_{2^k} \equiv \mathbb{F}_{2}[x] \pmod {P(x)}$
\item $P(x) \in \mathbb{F}_{2}[x]$, irreducible polynomial of degree $k$
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{comment}

\begin{frame}{\large {Field Elements: {\it e.g.} $\mathbb{F}_8$}}
%\ptsize{10}

Consider: $\mathbb{F}_{2^3}= \mathbb{F}_{2}[x] \pmod {x^3 + x +
  1}$\\ $A \in \mathbb{F}_{2}[x]$ \\ 
$A \pmod {x^3 + x + 1} = a_2 x^2 + a_1 x + a_0$. Let $P(\alpha) = 0$:

\begin{itemize}
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 0, 0\rangle = 0$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 0, 1\rangle = 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 1, 0\rangle = \alpha$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 1, 1\rangle = \alpha
  + 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 0, 0\rangle = \alpha^2$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 0, 1\rangle =
  \alpha^2 + 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 1, 0\rangle =
  \alpha^2 + \alpha$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 1, 1\rangle =
  \alpha^2 + \alpha + 1$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Polynomial functions in Galois field}}
\bi
\item Polynomial functions over $\Fkk \Leftrightarrow$ Boolean functions on $k$-bit vectors
$$f:\B^k \to \B^k \Leftrightarrow \mathcal F:\Fkk \to \Fkk$$
\item Example: Lagrange's interpolation
\begin{figure}[H]
\centerline{
\includegraphics[scale=0.45]{lagrange.png}
}
%\caption{A Gate-level Circuit corresponding to last STG}
\end{figure}
\ei
\end{frame}
\begin{frame}{\large{Polynomial functions in Galois field(2)}}
\bi
\item 8 pairs of $(A,Z)$, use Lagrange's interpolation to abstract a polynomial function
$$Z = \F(A) = \sum_{n=1}^{8}\left[ \frac{\prod_{i\neq n}(A-A_i)}{\prod_{i\neq n}(A_n-A_i)}\cdot Z_n \right]$$
\item Result = $(\alpha^2+1)A^4+(\alpha^2+1)A^2~ implies$ \alert{polynomial representation of the function} 
\item $P(\alpha) = \alpha^3+\alpha+1 = 0$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Algebraic Geometry Terminology}}
\vspace{-0.2in}
%\ptsize{10}
Let $\mathbb{F}_q = GF(2^k)$:
\begin{itemize}
\item $\mathbb{F}_q[x_1, \ldots, x_n]$: ring of all polynomials with
  coefficients in $\mathbb{F}_q$ 
\item Given a set of polynomials:
\begin{itemize}
\item $f, f_1, f_2, \dots , f_s \in \mathbb{F}_q[x_1, \dots, x_n]$
\item Find solutions to $f_1 = f_2 = \dots = f_s = 0$
\end{itemize}
\item \alert{Variety:} Set of ALL solutions to a given system of polynomial equations: $V(f_1, \dots, f_s)$
	\begin{itemize}
	\item In $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2-1)=\{all\  points\  on\ circle: x^2+y^2-1=0\}$
	\item In $\mathbb{R}[x]$, $V(x^2+1)=\emptyset$
	\item In $\mathbb{C}[x]$, $V(x^2+1)=\{(\pm i)\}$
	\end{itemize}
\item Variety depends on the \alert{ideal} generated by the polynomials.
\item Reason about the Variety by analyzing the Ideals
\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large {Ideals and \Grobner basis}}
\vspace{-0.1in}

\begin{Definition}
{\bf Ideals of Polynomials:} Let $f_1, f_2, \ldots, f_s \in
\mathbb{F}_q[x_1, \dots, x_n]$. Let 
\begin{eqnarray}
J = \langle f_1, f_2 \ldots, f_s\rangle = \{f_1 h_1 + f_2 h_2 + \dots + f_s h_s\} \nonumber 
\end{eqnarray}

$J = \langle f_1, f_2 \ldots, f_s\rangle$ is an ideal generated by
$f_1, \ldots, f_s$ and the polynomials are called the generators. 
\end{Definition}

\vspace{-0.1in}
%\ptsize{10}

\begin{itemize}
\item Different generators can generate the same ideal
\item $\langle f_1,\cdots,f_s \rangle=\cdots=\langle g_1,\cdots,g_t
  \rangle$
\item Some generators are a ``better'' representation of the ideal
\item A (reduced){\bf Gr\"obner basis} is a ``canonical'' representation of an ideal
\end{itemize}

Given $F = \{f_1, f_2,\cdots, f_s\}$, Compute a Gr\"obner Basis (using Buchberger's algorithm) $G =
\{g_1,g_2,\cdots,g_t\}$, such that $I = \langle F \rangle = \langle G \rangle$

\begin{center}
$V(F)=V(G)$
\end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item Sequential circuit verification requires reachability analysis
\item From bit-level to word-level
	\bi
	\item All about math
	\item \alert{Utilize the math tool: what does Tim get?}
	\item My approach
	\ei
\item One step back: Sequential arithmetic circuit verification
\item New inspiration: UNSAT core extraction using algebraic geometry
\item The plan
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large {Elimination Theorem [Tim, {\it Abstraction using GB}, DAC'14]}}
\begin{itemize}
\item Let $J \subset \Fq[x_1,\dots,x_d]$ be an ideal
\end{itemize}
\begin{itemize}
\item Let $G$ be a \Grobner basis of $J$ with respect to a lex ordering where $x_1
> x_2 > \dots > x_d$.
\end{itemize}
\begin{itemize}
\item Then for every $0 \leq l \leq d$:
	\begin{itemize}	
	\item The set $G_l= G \cap \Fq[x_{l+1},\dots,x_d]$ is a Gr\"obner basis of the $l$th elimination ideal $J_l$.
	\end{itemize}
\end{itemize}

The $l$th elimination ideal does not contain variables
$x_1,\dots,x_l$, nor do the generators of it.

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large {Elimination Term Ordering Example}}
%\ptsize{10}
\begin{itemize}
\item Let ideal $I = \langle f_1, f_2, f_3 \rangle$ where
	\begin{itemize}
	\item $f_1 = x^2 + y + z - 1$
	\item $f_2 = x + y^2 + z - 1$
	\item $f_3 = x + y + z^2 - 1$
	\end{itemize}
\item The \Grobner basis of $I$ with lex order $(x > y > z)$ is
	\begin{itemize}
	\item $g_1 = x + y + z^2 - 1$
	\item $g_2 = y^2 - y - z^2 + z$
	\item $g_3 = 2yz^2 + z^4 - z^2$
	\item $g_4 = z^6 - 4z^4 + 4z^3 - z^2$
	\end{itemize}
\item Notice that $g_2$ and $g_3$ only contain variables $y$ and $z$
	\begin{itemize}
	\item Eliminates variable $x~\Leftrightarrow  ~\exists_x$ in Boolean formula!
	\end{itemize}
\item Similarly, $g_4$ only contains the variable $z$ and eliminates $x$ and $y$
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large {Abstraction Term Ordering[Tim, {\it Abstraction using GB}, DAC'14]}}

Derived from applying elimination theorem to our problem set
\begin{itemize}
\item Given a circuit $C$ implementing $Z = \F(A)$ over $\Fq$
\item Using the variable order $x_1 > x_2 > \dots > x_d > Z > A$
	\begin{itemize}
	\item $x_1, \dots ,x_d$ are the circuit variables
	\end{itemize}
\item Impose a lex term order $>$ on the polynomial ring $R = \Fq[x_1,
  \dots, x_d, Z, A]$.
\item This elimination term order $>$ is defined as
the {\bf Abstraction Term Order}.

\item Compute a Gr\"obner basis $G$ of ideal $(J
+ J_0)$ using $>$
	\begin{itemize}
	\item $G$ will contain a polynomial of the form $Z + \F(A)$ ($Z = \F(A)$)
	\item $Z = \F(A)$ is a \alert{unique, canonical, polynomial}
  representation of $C$ over $\Fq$ 
	\end{itemize}
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Abstraction Term Order Example}}

\centerline{
\includegraphics[scale=0.4]{2bitmult.eps}
}
%\vspace{-0.3in}
{\bf $(z_0 > z_1 > r_0 > s_0 > s_3 > s_1 > s_2 > a_0 > a_1 >  b_0 > b_1 > Z > A > B)$}
\begin{align*}
f_1: s_0+a_0 \cdot b_0; ~~f_2: s_1+a_0 \cdot b_1;  ~~f_3: s_2+a_1 \cdot b_0; ~~f_4: s_3+a_1 \cdot b_1  \nonumber \\
f_5: r_0+s_1 + s_2; ~~f_6: z_0+s_0 + s_3; ~~f_7: z_1+r_0+s_3; ~~f_8: a_0 + a_1 \alpha + A \nonumber \\ 
f_9: b_0 + b_1 \alpha + B; ~~f_{10}: z_0 + z_1 \alpha + Z \nonumber
\end{align*}
\vspace{-0.3in}
{\bf $J$ = $\langle f_1, \dots, f_{10} \rangle$}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Abstraction Term Order Example}}

\centerline{
\includegraphics[scale=0.4]{2bitmult.eps}
}
%\vspace{-0.3in}
{\bf $(z_0 > z_1 > r_0 > s_0 > s_3 > s_1 > s_2 > a_0 > a_1 >  b_0 > b_1 > Z > A > B)$}
\begin{align*}
f_{11}: a_0^2+a_0; ~~f_{12}: a_1^2+a_1 \alpha; ~~f_{13}: b_0^2+b_0; ~~f_{14}: b_1^2+b_1; ~~f_{15}: s_0^2+s_0;  \nonumber \\
f_{16}: s_1^2+s_1; ~~f_{17}: s_2^2+s_2; f_{18}: s_3^2+s_3; ~~f_{19}: r_0^2+r_0; ~~f_{20}: z_0^2+z_0 \nonumber \\
f_{21}: z_1^2+z_1; ~~f_{22}: A^4+A; ~~f_{23}: B^4+B; ~~f_{24}:Z^4+Z \nonumber
\end{align*}
\vspace{-0.3in}
{\bf $J_0$ = $\langle f_{11}, \dots, f_{24} \rangle$}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Abstraction Term Order Example}}

\centerline{
\includegraphics[scale=0.4]{2bitmult.eps}
}
%\vspace{-0.3in}
{\bf $(z_0 > z_1 > r_0 > s_0 > s_3 > s_1 > s_2 > a_0 > a_1 >  b_0 > b_1 > Z > A > B)$}

Compute the \Grobner basis, $G$, of \{$J + J_0$\} with respect to abstraction term ordering $>$.
$G$ = \{$g_1, \dots, g_{14}$\}
\begin{align*}
g_1: B^4+B; ~~g_2: b_0+b_1 \alpha + B; ~~g_3: a_0+a_1 \alpha + A; ~~g_4: A^4+A; \nonumber \\
g_5: s_0+s_1 \alpha + s_2 (\alpha + 1) + Z; ~~g_6: r_0+s_1+s_2; ~~g_7: z_1+r_0+s_3 \nonumber \\
g_7: z_0+z_1 \alpha +Z; ~~\alert{\bf g_9: Z+A*B} ; ~~g_{10}: b_1+B^2+B; ~~g_{11}: a_1+A^2+A \nonumber \\
g_{12}: s_3+a_1 b_1; ~~g_{13}: s_2+a_1 b_1 \alpha +a_1 B; ~~g_{14}: s_1+a_1 b_1 \alpha +b_1 A  \nonumber
\end{align*}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item Sequential circuit verification requires reachability analysis
\item From bit-level to word-level
	\bi
	\item All about math
	\item Utilize the math tool: what does Tim get?
	\item \alert{My approach}
	\ei
\item One step back: Sequential arithmetic circuit verification
\item New inspiration: UNSAT core extraction using algebraic geometry
\item The plan
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Inspirations from Tim's work}}
 \vspace{-0.1in}
\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{./comb2.eps}
%\caption{FSM models of sequential circuits}
\label{fig:comb}}
\end{figure}
 \vspace{-1in}
\bi
\item From Tim's work I learned:
	\bi
	\item Compute GB with term order {\it intermediate $>$ Z $>$ A}
	\item Obtain $Z = \mathcal F(A)$
	\ei
\item Inspiration: what if we impose term order {\it intermediate $>$ A $>$ Z} ?
\item Proposed approach: add the evaluation of $A$ into elimination ideal, then eliminate all but $Z$, we will get the 
\alert{evaluation of $Z$}!
	\bi
	\item Present state $\gets A$, next state $\gets Z$
	\item It is a way to do \alert{one-step reachability}!
	\ei
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Recall: Breadth-First Traversal Algorithm}}
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}
\vspace{0.2in}
\textbf{We want to implement this algorithm using algebraic geometry approach}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Implement Image Function in Algebraic Geometry}}
\begin{itemize}
\item State variables (word-level) $S, T$ and sets of states such as
$from^i, to^i$ can always be represented as varieties of ideals.
\item Boolean operators can always be converted to operations in $\mathbb F_2$
\begin{table}
\centering
\begin{tabular}{|c|c|} \hline
Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
$\overline{a}$ & $1 + a$\\ \hline
$a\ and\ b$ & $ab$\\ \hline
$a\ or\ b$ & $a + b + ab$\\ \hline
$a \oplus b$ & $a + b$\\
\hline\end{tabular}
\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
\label{table:booltogalois_op}
\end{table}
\item An \emph{elimination ideal} can be built from circuit gates, PS/NS
word definition and vanishing polynomials
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large {Example Formulation}}
\vspace{-0.1in}
\begin{figure}[hbt]
\centerline{
\includegraphics[scale=0.25]{fsm_fig.eps}
}
%\caption{A multiplier over GF($2^2$)}
%\label{fig:fsm}
\end{figure}
\vspace{-0.4in}
Model circuit as polynomials in $\mathbb{F}_2 \subset \mathbb{F}_{2^k}$:
\begin{align*}
&t_0 = (\overline{x}\ and\ \overline{s_0}\ and\ \overline{s_1})\ or\ (s_0\ and\ s_1)\\
&\implies f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
&t_1 = (\overline{s_0}\ and\ x)\ or\ (s_0\ and\ \overline{s_1})~~~~\\
&\implies f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
&f_3: S + s_0 + s_1\alpha, ~~f_4:T + t_0 + t_1\alpha
\end{align*}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Example Formulation(2)}}
\bi
\item Elimination ideal to model Image function for example circuit:
\begin{itemize}
\item Transition functions (bit-level): $f_1,f_2$\\

\item Word variable definitions: $f_3,f_4$\\

\item Vanishing polynomials: $f_6: x^2 - x; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: s_0^2 - s_0;
f_{11}: s_1^2 - s_1; f_{12}: T^4 - T$
\end{itemize}
\item Add the present state (e.g. initial states in first iteration $f_5: S$), compute Gr\"obner basis
for ideal $J = \langle f_1,\dots, f_{12}\rangle$ under elimination term order $$intermediate\ bit\text{-}level\ signals >\ bit\text{-}level\ PIs/POs >\ S >\ T$$
\item  Result will include a univariate polynomial about \emph{next states} $T$, e.g. $T^2+(\alpha+1)T+\alpha$ 
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Recall: Breadth-First Traversal Algorithm}}
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}
\vspace{0.2in}
\textbf{We want to implement this algorithm using algebraic geometry approach}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Intersection and union in algebraic geometry}}
\begin{Definition}
\label{def:sum}
({\bf Sum/Product of Ideals}) If $I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are 
ideals in $\mathbb F[x_1, \dots, x_n]$, then the {\bf sum} of $I$ and $J$ is defined as
$$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$$ And the {\bf product} of $I$ and $J$ is defined
as
\begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle \nonumber
  \end{equation}
\end{Definition}
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf
  V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) =
{\bf V}(I) \bigcup {\bf V}(J)$. 
\end{Theorem}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Complement set in algebraic geometry}}
\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
  F[x_1, \dots, x_n]$, then $I:J$ is the set
  \begin{equation}
  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}\nonumber
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}
\begin{Theorem}
\label{thm:quotient}
Let $I, J$ be ideals with vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
$${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$
\end{Theorem}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = pptalg]{\large{Our proposed algorithm of BFS traversal based on algebraic geometry}}
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\F(S)$}
  $from^0 = reached = \F(S)$\;
  \Repeat{$new^i == 1$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$GB w/ elimination term order$\langle J_{ckt},J_0, from^{i-1}\rangle$\;
	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
	$from^i \gets new^i(S\setminus T)$\;
  }
\Return{$reached$}
\caption{Algebraic Geometry based Traversal Algorithm}
\end{algorithm}

\hyperlink{pptpage2}{\beamergotobutton{Go to example page 2}}

\hyperlink{pptpage3}{\beamergotobutton{Go to example page 3}}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Initial state $from^0 = S(\{00\})$
\item {\bf Iteration 1:}Compose an elimination ideal $J$ 
\vspace{-0.2in}\\
\begin{align*}
&f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
&f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
&f_3: S - s_0 - s_1\\
&f_4: T - t_0 - t_1
\end{align*}
$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$
\begin{align*}
&f_5: x^2-x\\
&f_6: s_0^2-s_0, f_7: s_1^2-s_1\\
&f_8: t_0^2-t_0, f_9: t_1^2-t_1\\
&f_{10}: S^4-S, f_{11}:T^4-T
\end{align*}
$\ \ \ \ \ \ \ \ \ \ \  \ J_0 = \langle f_5,f_6,\dots,f_{11}\rangle$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = pptpage2]
\bi
\item Elimination term order: 
$$\{x,s_0,s_1,t_0,t_1\}~(all~bits)~>~S~(PS~word)~ >~ T~(NS~word)$$
\item Compute the reduced GB for $J = J_{ckt}+J_0+\langle from^0\rangle$
\item Next state 
$$to^1 = \langle T^2+(\alpha+1)T+\alpha\rangle~(\{01,10\})$$
\item Complement of formerly reached state:
$$\langle T^4-T\rangle:\langle T\rangle = \langle T^3+1\rangle ~(\{01,10,11\})$$
\item Newly reached state:
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle$$ 
\item Update current reached states 
$$reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle~(\{00,01,10\})$$
\item Update the present states for next iteration
$$from^1 = \langle S^2+(\alpha+1)S+\alpha\rangle$$
\ei
\hyperlink{pptalg}{\beamergotobutton{The algorithm}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = pptpage3]{\large{Full Blown traversal of example circuit using algebraic geometry(3)}}
\bi
\item Iteration 2:
	\bi
	\item Next state: $to^2 = \langle T^2+\alpha T\rangle~(\{00,10\})$
	\item The complement of $reached$:
	$$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
= \langle T + 1+\alpha\rangle ~(\{11\})$$ 
	\item Newly reached state: 
$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle \alert{1}\rangle$$ 
	\ei
\item Algorithm terminates
\item Return value (final reachable states):
$$reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$
\ei
\hyperlink{pptalg}{\beamergotobutton{The algorithm}}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item Sequential circuit verification requires reachability analysis
\item From bit-level to word-level
\item \alert{One step back: Sequential arithmetic circuit verification}
\item New inspiration: UNSAT core extraction using algebraic geometry
\item The plan
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Application on implicit unrolling}}
\vspace{-0.1in}
\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=4.5in]{./Moore.eps}
% \vspace{-0.2in}
%\caption{A simple Moore FSM and its state transition graph}
%\end{minipage}
\label{fig:Moore}}
\end{figure}
\vspace{-0.2in}
\bi
\item State enumeration cannot address sequential arithmetic circuits verification
	\bi
	\item Need to consider various initial states
	\item Exact reached states after $k$ clock cycles directly implies desired arithmetic function
	\ei
\item State transitions on simplified model
$$R_{k} = Tr(R_{k-1}) = Tr(Tr(\cdots Tr(R_{init})\cdots)) = Tr^k(R_{init})$$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Verification of a sequential Galois field multiplier (Normal basis)}}
SPEC: $R = A_{init}\cdot B_{init} \pmod {P(\alpha)}$ after $k$ clock cycles
\hspace{-0.3in}\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{./new_multi.eps}
% \vspace{-0.2in}
\caption{A 3-bit RH-SMPO and its Moore FSM model}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Our approach vs Conventional approach}}
\vspace{-0.1in}
\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{./convention.eps}
}
\end{figure}
\vspace{-0.4in}
\bi
\item Conventional: explicitly unroll $k$ time-frames
	\bi
	\item Bit-blasting!
	\ei
\item New: \alert{Implicitly} unroll
	\bi
	\item Keep only $k$ polynomials ($R = \F(A,B)$) when unrolling
	\item $R_1 = \F(A_{init},B_{init}),~R_2 = \F(A_{1},B_{1}) = \F^2(A_{init},B_{init}),~ 
	\cdots, ~ R_k = \F^k(A_{init},B_{init}) = A_{init}\cdot B_{init}$
	\ei
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment results}}
\begin{table}[htb]
\centering
\caption{\small Run-time (seconds) for verification of bug-free and
  buggy RH-SMPO using our approach} 
\label{tbl:exp1}  
\begin{tabular}{|c||c|c|c|c|c|c|} 
\hline
Operand size $k$ & 33 & 51 & 65 & 81 & 89 & 99\\
\hline
\#variables & 4785 & 11424 & 18265 & 28512 & 34354 & 42372\\
\hline
\#polynomials & 3630 & 8721 & 13910 & 21789 & 26255 & 32373\\
\hline
\#terms & 13629 & 32793 & 52845 & 82539 & 99591 & 122958\\
\hline
\hline
Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 & 67021\\
\hline
Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120 & 66929\\
\hline
\end{tabular}
\end{table}
* \textit{Results from X. Sun, et al. ``Formal Verification of Sequential Galois Field Arithmetic 
Circuits using Algebraic Geometry", to be presented in Grenoble, DATE'15}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{What I have done for this experiment}}
\bi
\item Sequential arithmetic circuits over normal basis
\item Standard basis $\rightleftharpoons$ normal basis
\item Optimal normal basis
\item Design circuit(multiplier) over optimal normal basis
\item Verify the function of the circuit
\ei
\hyperlink{pptnodetail}{\beamergotobutton{No more details!}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Normal basis representation}}
\bi
\item Normal basis representation: $A(a_0,\dots,a_{k-1}) = \sum_{i=0}^{k-1}a_{n(i)}\beta^{2^i}$
\item Normal element: $\beta = \alpha^t$
\item Squaring of elements represented in
normal bases can be implemented simply by a cyclic right-shift
operation.
\ei
\begin{Example}
\label{ex:nb_sq}
For $a, b \in \Fkk, (a+b)^2 = a^2 + b^2$. Applying this rule for
element squaring: 
\begin{align}
B = & (b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}}) \nonumber\\
B^2 = &b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta^{2^k} \nonumber\\
= &b_{k-1}\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}} \nonumber
\end{align}
as $\beta^{2^k} = \beta$ by applying Fermat's little theorem to $\Fkk$, and $b_i^2 = b_i$. 
\end{Example}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Galois field multiplication with normal basis}}
\bi
\item Let $R =
\sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
\beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$, then 
\[
R = A\cdot B = (\sum_{i=0}^{k-1} a_i \beta^{2^{i}}) (\sum_{j=0}^{k-1}
b_j \beta^{2^{j}})  =
\sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}\nonumber 
\]

\item Expressions $\beta^{2^i}\beta^{2^j}$ are called cross-product
terms. Their normal basis representations are: 
\begin{displaymath}
\beta^{2^i}\beta^{2^j} =
\sum_{n=0}^{k-1}\lambda_{ij}^{(n)}\beta^{2^n}, \ \ \lambda_{ij}^{(n)}
\in \mathbb F_2. 
\end{displaymath}
\ei
\end{frame}
\begin{frame}{\large{Galois field multiplication with normal basis(2)}}
\bi
\item The expression for the
$n^{th}$ digit of product $R = (r_0, \dots, r_n, \dots r_{k-1})$ is:
\[
r_n = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(n)}a_ib_j = A
\cdot M_n \cdot B^T, ~~0 \leq n \leq k-1
\]
\item $M_n = (\lambda_{ij}^{(n)})$ is a binary $k \times k$ matrix over
$\mathbb F_2$, and it is called the $\lambda$-matrix. 
\item $\lambda$-matrix is \alert{unique} when $k$ and $\beta$ are given!
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Modified algorithm to verify the function of sequential GF multipliers}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle)$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit verification}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment on 3-bit RH-SMPO}}
\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=1.5in]{./RH3.eps}
% \vspace{-0.2in}
%\caption{Gate-level circuit of 3-bit RH-SMPO}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}
\vspace{-0.2in}
\bi
\item The elimination ideal (first iteration):
\begin{align*}
J = &d_0+b_2\cdot a_2,
c_1+a_0+a_2,
c_2+b_0+b_2,
d_1+c_1\cdot c_2,\\
&e_0+d_0+d_1,
e_2+d_1,
r_0'+r_2+e_0,
r_1'+r_0,
r_2'+r_1+e_2,\\
&A+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
B+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12},\\
&R+r_0\alpha^3+r_1\alpha^6+r_2\alpha^{12},
R'+r_0'\alpha^3+r_1'\alpha^6+r_2'\alpha^{12};
\end{align*}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment on 3-bit RH-SMPO(2)}}
\bi
\item ``$J_0$" is the ideal of vanishing polynomials in all bit-level
variables (e.g. $a_0^2-a_0$) and word-level variables (e.g. $A^8-A$).
\item $from_0 = \{R, A_{init}+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
B_{init}+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12}\}$
\item $to_1 : 
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
$
\item $from_1 = \{
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
, A_{init}+a_2\alpha^3+a_0\alpha^6+a_1\alpha^{12},
B_{init}+b_2\alpha^3+b_0\alpha^6+b_1\alpha^{12}\}$
\item $\cdots$
\item After 3 iterations: $to_3 = \{ \alert{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item Sequential circuit verification requires reachability analysis
\item From bit-level to word-level
\item One step back: Sequential arithmetic circuit verification
\item \alert{New inspiration: UNSAT core extraction using algebraic geometry}
\item The plan
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{An example of abstraction refinement algorithm}}
\begin{algorithm}[H]
\SetAlgoNoLine
	\KwIn{$M$ is the original machine, $p$ is the property to check, $k$ is the number of steps in $k$-BMC}
  $k = $ InitValue\;
  \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT}}
  {
	\Return{``Found error trace"}
  }
  {
	Extract UNSAT proof $\mathcal P$ of $k$-BMC\;
	$M' = $ \textit{ABSTRACT}$(M,\mathcal P)$\;
  }
  \eIf{MODEL-CHECK$(M',p)$ returns \textbf{PASS}}
  {
	\Return{``Passing property"}
  }
  {
	Increase bound $k$\;
	goto Line 2\;
  }
\caption{$k$-BMC with Abstraction Refinement (L. Zhang Thesis)}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{An example of abstraction refinement algorithm(2)}}
\bi
\item $k$-BMC: unroll the machine for $k$ times, represent reachable states with CNF formula to check property $p$
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
\item UNSAT core: a subset of clauses that is still UNSAT
\item State variables not in UNSAT core: no matter what their values are, $p$ will NOT be violated
\item In abstracted model, ignore these ``irrelevant" variables
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{An example of abstraction refinement algorithm(3)}}
\begin{figure}[hbt]
\centering{
\includegraphics[width=4.5in]{./refine.eps}
\caption{Abstraction by reducing latches}
\label{fig:refine}}
\end{figure}
\bi
\item Remove ``irrelevant" latches, reduce state space
\item Provide an over-approximation
\item This algorithm requires \alert{UNSAT core extraction}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Buchberger's Algorithm Computes a Gr\"obner Basis}}
\vspace{-0.3in}
%{\small
%\begin{algorithm}
%\caption {Buchberger's Algorithm}
{\bf Buchberger's Algorithm}\\
%\label{alg:gb}
%\begin{algorithmic}
% \REQUIRE : $F = \{f_1, \dots, f_s\}$
 INPUT : $F = \{f_1, \dots, f_s\}$\\
% \ENSURE : $G = \{g_1,\dots ,g_t\}$\\ %, a Gr\"{o}bner basis \\
 OUTPUT : $G = \{g_1,\dots ,g_t\}$\\ %, a Gr\"{o}bner basis \\
  $G:= F$; \\
  REPEAT\\
  \hspace{0.1in} $G' := G$\\
  \hspace{0.1in} For each pair $\{f, g\}, f \neq g$ in $G'$ DO\\
\hspace{0.2in}  $S(f, g) \stackrel{G'}{\textstyle\longrightarrow}_+
r$ \\
\hspace{0.2in}  IF $r \neq 0$ THEN $G:= G \cup \{r\}$ \\
%\hspace{0.2in}  $G(x):=G(x) / x$
UNTIL $G = G'$
%\end{algorithmic}
%\end{algorithm}
\bi
\item S-poly: $S(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g$\par
$L = \text{LCM}(lm(f), lm(g))$, ~~$lm(f)$: leading monomial of $f$
\item Multivariate division: $f\stackrel{g}{\textstyle\longrightarrow}r:~lm(g)|lm(f)\to r = f-\frac{lt(f)}{lt(g)}g$
\ei
%}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Observation when executing Buchberger's algorithm}}
\bi
\item Translate CNF clauses to polynomials in $\mathbb F_2$
\ei
\begin{columns}[onlytextwidth]
\begin{column}{0.3\textwidth}
\begin{align*}
&c_1: \bar{a}\lor\bar{b}\\
&c_2: a\lor\bar{b}\\
&c_3: \bar{a}\lor b\\
&c_4: a\lor b\\
&c_5: x\lor y\\
&c_6: y\lor z\\
&c_7: b\lor \neg y\\
&c_8: a\lor x\lor \neg z
\end{align*}
\end{column}
\begin{column}{0.3\textwidth}

$$\implies$$
\end{column}
\begin{column}{0.4\textwidth}
\begin{align*}
&f_1:ab\\
&f_2:ab+a\\
&f_3:ab+b\\
&f_4:ab+a+b+1\\
&f_5:xy+y+x+1\\
&f_6:yz+y+z+1\\
&f_7:by+y\\
&f_8:axz+az+xz+z
\end{align*}
\end{column}
\end{columns}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Observation when executing Buchberger's algorithm(2)}}
\bi
\item Compute a GB using Buchberger's algorithm
\item Term ordering: $a>b>x>y>z$
\item
\begin{enumerate}
\item Compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1 = a$
\item Update $F=F\cup r_1$
\item Compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2=b$
\item Update $F=F\cup r_2$
\item Use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$
\item Compute $Spoly(f_1,f_4) = s_3= a+b+1$, $a+b+1$ can be reduced 
$r_1$ , the intermediate remainder $r_3 = b+1$. Then divided by $r_2$, and the remainder is ``1", 
\alert{terminate the Buchberger's algorithm}
\item Draw a DAG depicting the process through which we obtain remainder ``1". 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. 
\end{enumerate}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Observation when executing Buchberger's algorithm(3)}}
\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{./UNSAT.eps}
\caption{DAG representing Spoly computations and multivariate divisions}
\label{fig:UNSAT}}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Observation when executing Buchberger's algorithm(3)}}
\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{./UNSAT2.eps}
\caption{DAG representing Spoly computations and multivariate divisions}
}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Proposed algorithm to extract UNSAT core}}
\begin{algorithm}[H]
\SetAlgoNoLine
	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
	\KwOut{An UNSAT core $\{f_{m_1},f_{m_2},\dots,f_{m_t}\}$}
\Repeat{$r_l == 1$}
{
	Pick a pair $f_i,f_j\in F$ that has never been computed S-poly\;
	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
	{
		$F = F\cup r_l$\;
		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the S-poly, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the S-poly\;
	}
}
Backward traverse the DAG for remainder ``1", replace $r_l$ with corresponding DAG $G_l$\;
\Return{All roots}
\caption {Extract UNSAT core using a variation of Buchberger's algorithm}
\end{algorithm}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Introduction
\item Sequential circuit verification requires reachability analysis
\item From bit-level to word-level
\item One step back: Sequential arithmetic circuit verification
\item New inspiration: UNSAT core extraction using algebraic geometry
\item \alert{The plan}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Objectives}}
\bi
\item Explore an implementation of algebraic geometry based reachability analysis
\item CAD tool design
	\bi
	\item {\sc Singular}'s data structure not optimized $\to$ design a standalone CAD tool implemented in C++
	\item Specifically aims to solve word-level sequential verification problems
	\item Lower complexity: borrow techniques from [T. Pruss, {\it Abstraction}, DAC'14], [J. Lv, {\it Equivalence}, TCAD'13]
		and [C. Eder, {\it F-4 reduction}, ISSAC'11]
	\ei
\item Fine-tune the tool for sequential GF arithmetic circuits verification
\item Explore a new abstraction-refinement paradigm based on information from UNSAT cores
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Time table}}
\begin{itemize}
\item Current status: Experiments have been performed to run implicit state enumeration on sequential circuits benchmarks
such as ISCAS'89 circuits. Current problem is the algorithm spends too much time on multivariate division procedure;
\item Spring 2015: Implement the tool which can efficiently do multivariate division and test the performance of 
our tool based on [Tim, {\it Abstraction}, DAC'14] approach;
\item Summer 2015: Integrate the refined multivariate division routine into our verification tool, test its
performance on circuits with various sizes;
\item Fall 2015: Evaluate data and write the dissertation.
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{frame}{\large{Obviate GB computation}}
% 
% \end{frame}

\end{document}

### 2015-05-03 techwriting/CAI/prelim.tex
copy original prelim & theory

--- 2015-05-01 techwriting/proposal/prelim.tex
@@
 \section{Preliminaries}
-
-In this section, we will briefly review the preliminary concepts
-related to finite state machine models, Galois fields, algebraic geometry and Gr\"obner bases, as
-our approaches are based on a combination of these concepts.
 
@@
 the formal definition is as follows:
-\begin{Definition}
+\begin{definition}
 A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
@@
 \end{itemize}
-\end{Definition}
+\end{definition}
 The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
@@
 the term $cX$ from $f$. 
-\begin{Example}
+\begin{example}
 \label{ex:multidiv}
@@
   \over c} g = e + abd$. 
-\end{Example}
+\end{example}
   Similarly, $f$ reduces to $r$ modulo the set
@@
 polynomial equations $f_1 = \dots = f_s = 0$ is defined as the affine variety:
-\begin{Definition}
+\begin{definition}
 Given a set of polynomials $f_1,\dots,f_s$ over ring $\mathbb F_q[x_1,\dots,x_n]$, their 
@@
 f_1(a_1,\dots,a_n) = \cdots = f_s(a_1,\dots,a_n) = 0\}$$ 
-\end{Definition}
+\end{definition}
 
@@
 of given set of polynomials. This set is defined as follows:
-\begin{Definition}
+\begin{definition}
 {\bf Ideal of Polynomials:} Let $f_1,f_2,\dots,f_s\in \mathbb F[x_1,\dots,x_n]$.
@@
 the {\bf generators} of ideal $J$.
-\end{Definition}
+\end{definition}
 % On the other hand, if given another polynomial $f$, we need to judge whether it belongs to $J$.
@@
 This problem is usually described as ideal membership checking problem.
-\begin{Definition}
+\begin{definition}
 {\bf Ideal membership:} Let $f_1,f_2,\dots,f_s\in \mathbb F[x_1,\dots,x_n]$, and $J = \langle f_1,f_2,\dots,f_s\rangle$
@@
 then $f\in J$.
-\end{Definition}
+\end{definition}
 
@@
 
-\begin{Definition}
+\begin{definition}
 The set $G = \{g_1, \dots,
@@
 g_i \in G \ s.t. \ LT(g_i) ~|~ LT(f)$. 
-\end{Definition}
+\end{definition}
 % The famous Buchberger's
@@
 by computing a Gr\"obner basis of $J$ w.r.t. elimination orders: 
-\begin{Theorem}
+\begin{theorem}
 \label{thm:elim}
-(Elimination Theorem\cite{ideals:book}) Let $J\subset \mathbb
+(Elimination theorem\cite{ideals:book}) Let $J\subset \mathbb
   F_{2^k}[x_1,\dots,x_d]$ be an ideal and let $G$ be a Gr\"obner basis
@@
   the $l$-th elimination ideal $J_l$.
-\end{Theorem}
+\end{theorem}
 We describe an application of elimination ideals using following example borrowed from \cite{ideals:book}:
-\begin{Example}
+\begin{example}
 Consider polynomials $f_1: x^2-y-z-1;\ f_2:x-y^2-z-1;\ f_3:x-y-z^2-1$ and ideal $J = \langle f_1,f_2,f_3\rangle
@@
 we find that the polynomial $g_4$ only contains variable $z$ ($x,y$ eliminated), and polynomials $g_2,g_3,g_4$ only contain variables
-$y,z$ ($x$ eliminated). According to Theorem \ref{thm:elim}, $G_1 = G\cap\mathbb C[y,z] = \{g_2,g_3,g_4\}$
+$y,z$ ($x$ eliminated). According to theorem \ref{thm:elim}, $G_1 = G\cap\mathbb C[y,z] = \{g_2,g_3,g_4\}$
 is the Gr\"obner basis of the $1^{st}$ elimination ideal of $J$ and $G_2 = G\cap\mathbb C[z] = \{g_4\}$ is the 
 $2^{nd}$ elimination ideal of $J$, respectively.
-\end{Example}
+\end{example}
 
@@
 derived through \Grobner bases computed with elimination orders:
-\begin{Lemma}
+\begin{lemma}
 (From \cite{timDAC}) Given a combinational circuit $C$ with $k$-bit
@@
   function of the circuit with only word level variables $R$ and $A$.
-\end{Lemma}
+\end{lemma}
 This lemma shows an application of GB computations over an elimination ideal.
@@
 % $\exists_x f$ denotes the existential quantification of $f$ w.r.t. $x$.
-% \begin{Example}
+% \begin{example}
 % We use implicit state enumeration based on Boolean formulas to traverse FSM in Fig.\ref{fig:fsm}.
@@
 % $$t_1\overline{\oplus}\Delta_1 = t_1\overline{\oplus}(x\overline{s_0}+s_0\overline{s_1})$$
-% \end{Example}
+% \end{example}

### 2015-05-03 techwriting/CAI/reacha.tex
copy original prelim & theory

--- 2015-05-01 techwriting/proposal/reacha.tex
@@
 computing a GB of an elimination ideal using elimination term order
-based on Theorem \ref{thm:elim}.
+based on theorem \ref{thm:elim}.
 
@@
 
-\begin{Example}
+\begin{example}
 For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
@@
 \overline{t_1}$ as the next reachable state, and so on. 
-\end{Example}
+\end{example}
 
@@
 
-\begin{Definition}
+\begin{definition}
 \label{def:sum}
@@
   \end{equation}
-\end{Definition}
+\end{definition}
 
@@
 intersection and union of affine varieties as:
-\begin{Theorem}
+\begin{theorem}
 \label{thm:unionintersect}
@@
 {\bf V}(I) \bigcup {\bf V}(J)$. 
-\end{Theorem}
+\end{theorem}
 
@@
 
-\begin{Definition}
+\begin{definition}
 ({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
@@
 and is called the {\bf ideal quotient} of $I$ by $J$.
-\end{Definition}
+\end{definition}
 
@@
 % For non-trivial cases we can only prove that 
-% \begin{Theorem}
+% \begin{theorem}
 % Let $I, J$ be ideals in $\mathbb F[x_1,\dots,x_n]$, then 
 % $${\bf V}(I:J) \supset {\bf V}({\bf I}({\bf V}(I) - {\bf V}(J)))$$
-% \end{Theorem}
+% \end{theorem}
 % Fortunately for most hardware verification cases, the form of polynomial generators are restricted.
@@
 
-\begin{Theorem}
+\begin{theorem}
 \label{thm:quotient}
@@
 $${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$
-\end{Theorem}
+\end{theorem}
 
@@
 
-\begin{Corollary}
+\begin{corollary}
 $$V(J') = \overline{{\bf V}(J)} = {\bf V}(J_0:J)$$
 
-\end{Corollary}
+\end{corollary}
 
@@
 
-\begin{Example}
+\begin{example}
 We apply Algorithm \ref{alg:univa} to Ex.\ref{ex:motiv} to execute the FSM traversal. Let the initial state
@@
 % $reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$ as final reachable states.
-\end{Example}
+\end{example}
 
@@
 %\begin{minipage}{12cm}
-\includegraphics[width=7.5in]{./new_multi.eps}
+\includegraphics[width=4.5in]{./new_multi.eps}
 % \vspace{-0.2in}
@@
 \vspace{-0.1in}
-\begin{Example}
+\begin{example}
 \label{ex:RHSMPO}
@@
 B_{init}$, which verifies the multiplier. 
-\end{Example}
+\end{example}
 Using our approach we successfully verified the function of an 100-bit RH-multiplier,

### 2015-05-13 techwriting/CAI_FMCAD/prelim.tex
update FMCAD2015 first draft

--- 2015-05-01 techwriting/proposal/prelim.tex
@@
 %\begin{minipage}{12cm}
-\includegraphics[width=5.5in]{./seqmodel.eps}
+\includegraphics[width=3.5in]{./seqmodel.eps}
 % \vspace{-0.2in}
@@
 \label{thm:elim}
-(Elimination Theorem\cite{ideals:book}) Let $J\subset \mathbb
+(Elimination theorem\cite{ideals:book}) Let $J\subset \mathbb
   F_{2^k}[x_1,\dots,x_d]$ be an ideal and let $G$ be a Gr\"obner basis
@@
 we find that the polynomial $g_4$ only contains variable $z$ ($x,y$ eliminated), and polynomials $g_2,g_3,g_4$ only contain variables
-$y,z$ ($x$ eliminated). According to Theorem \ref{thm:elim}, $G_1 = G\cap\mathbb C[y,z] = \{g_2,g_3,g_4\}$
+$y,z$ ($x$ eliminated). According to theorem \ref{thm:elim}, $G_1 = G\cap\mathbb C[y,z] = \{g_2,g_3,g_4\}$
 is the Gr\"obner basis of the $1^{st}$ elimination ideal of $J$ and $G_2 = G\cap\mathbb C[z] = \{g_4\}$ is the 

### 2015-05-13 techwriting/CAI_FMCAD/reacha.tex
update FMCAD2015 first draft

--- 2015-05-01 techwriting/proposal/reacha.tex
@@
 \label{thm:quotient}
-Let $I, J$ be ideals with vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
-$${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$
+Let $J$ be ideals with single univariate polynomial about variable $x$ over $\Fkk[x]$, vanishing ideal $J_0 = \langle x^{2^k}-x\rangle$. then 
+$${\bf V}(J_0:J) = {\bf V}(J_0) - {\bf V}(J)$$
 \end{Theorem}
+
+To prove this theorem, we need following lemma:
+\begin{Lemma}
+\label{lem:gcd}
+Let $f,g\in \Fkk[x]$, then $(f:g) = \left(\frac{f}{gcd(f,g)}\right)$.
+\end{Lemma}
+
+\begin{proof}
+Let $d = gcd(f, g)$. So, $f = df_1 , g = dg_1$ with $gcd(f_1 , g_1 ) = 1$.
+
+Note that $f_1 = \frac{f}{gcd(f,g)}$.
+
+Take $h \in (f : g)$. So, $hg \in (f )$, that is $hg = f \cdot r$ with $r \in \Fkk[x]$.
+
+Therefore, $hdg_1 = df_1 r$, and so $hg_1 = f_1 r$. But since $gcd(g_1 , f_1 ) = 1$ we have that $f_1$
+divides $h$. Hence $h \in (f_1 )$.
+Conversely, let $h \in (f_1 )$. Then $h = s \cdot f_1$ , with $s \in \Fkk[x]$. So, $hg = hdg_1 = sf_1 dg_1 = sg_1 f \in
+(f )$. Therefore, $h \in (f : g)$.
+\end{proof}
+Using this lemma we can easily prove the proposition:
+
+\begin{proof}
+Since $\Fkk[x]$ is a principal ideal domain, then$ J = (g)$ for some polynomial$ g \in \Fkk[x]$.
+
+Let $d = gcd(g, x^{2^k} - x)$. So, $g = dg_1 , x^{2^k} - x = df_1$ , with $gcd(f_1 , g_1 ) = 1$.
+
+Then $J_0 : J = (f_1 )$ by Lemma \ref{lem:gcd}.
+
+Let $x \in{\bf V}_{\Fkk} (J_0 ) - {\bf V}_{\Fkk} (J)$. Then $x \in \Fkk$, but $g(x) \neq 0$.
+
+Since $x^{2^k} = x$, we see that either $d(x) = 0 or f_1 (x) = 0$. Note that since $g(x) \neq 0$, we can
+conclude that $d(x) = 0$. In conclusion, $f_1 (x) = 0$ and $x \in {\bf V}_{\Fkk} (f_1 )$.
+
+Now let $x \in V{\bf V}_{\Fkk} (f_1 )$. Therefore $f_1 (x) = 0$. So, $x^{2^k} - x = 0$ which gives $x \in {\bf V}_{\Fkk} (J_0 ) = \Fkk$.
+
+Assume now that $x ∈ {\bf V}_{\Fkk} (g)$. Then $0 = g(x) = d(x)g_1 (x)$.
+
+So, $d(x) = 0$ or $g_1 (x) = 0$.
+
+If $g_1 (x) = 0$, since $f_1 (x) = 0$ we get that $f_1 , g_1$ share a root and this contradicts the fact
+that $gcd(f_1 , g_1 ) = 1$.
+
+If $d(x) = 0$ then since $f_1 (x) = 0$ and $x^{2^k} - x = df_1$ we get that $x^{2^k} - x$ has a double root.
+
+But this is impossible since the derivative of $x^{2^k} - x$ is $-1$.
+
+So, $x \notin {\bf V}_{\Fkk} (g)$ and this concludes the proof.
+\end{proof}
 
@@
 	$to^i \gets$GB w/ elimination term order$\langle J_{ckt},J_0, from^{i-1}\rangle$\;
-	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
+	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
   	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
@@
 
-\subsection{Research objective: To make the approach scalable}
-
-At the time of writing this proposal, the theoretical concepts behind
-the word-level FSM traversal have been studied and validated. A
-completed algorithm has been devised and preliminary experiments were
-conducted for reachability analysis using the {\sc Singular} computer
-algebra tool \cite{DGPS}. While the approach is shown to work
-correctly and completely, it is practically infeasible. The complexity
-lies in the Gr\"obner basis computation.
-
-Over Galois fields of the type $\Fkk$, computing the reduced Gr\"obner
-basis of the ideal $(J + J_0)$ requires both memory and space
-${2^{k}}^{O(n)}$; where $J$ is an arbitrary polynomial ideal, $J_0$ is
-the vanishing ideal, and $n$ is the number of variables in the system
-(circuit). In this research, I wish to draw inspirations from the work
-of \cite{jinpeng} and \cite{timDAC} to overcome the
-complexity of GB computations.
-
-In \cite{jinpeng}, the authors analyze the given circuit topology,
-and derive a term order to represent the polynomials. They show that
-this term order renders the set of polynomials itself a Gr\"obner
-basis of the ideal. In \cite{timDAC}, the authors use a similar
-concept to derive a {\it refinement of the abstraction term order}
-(RATO) to simplify the Gr\"obner basis computation. However, the above
-concepts are applied to combinational circuit verification. I will
-investigate how similar techniques can be explored for the
-computations that are needed in Algorithm \ref{alg:univa} for
-reachability analysis to make the approach scalable. I believe that it
-is possible to make the approach scalable, as I have recently
-demonstrated in a paper \cite{myDATE} %{\bf refer to the DATE paper here?}
-which has been accepted for publication at the {\it Design Automation
-  and Test in Europe Conference}, 2015.
-
-\subsection{Research objective : Application to Unrolling Sequential Circuits for Sequential Arithmetic Verification}
-In above discussions, we propose an implicit state enumeration algorithm implemented by GB computations over
-an elimination ideal, which is usually applied on FSMs working as control units. For many arithmetic
-datapath components, state enumeration is not the best way to verify the function of circuit. On the one hand, the signals preloaded
-to the state variables in FSM are usually not specified such that we can hardly encode the initial states; on the other
-hand, instead of the set of globally reachable states, arithmetic verification focuses on exact reached states
-after $k$ clock cycles which reflects desired arithmetic function. Considering these reasons, we modify
-our algebraic geometry based approach and apply it to implicit FSM unrolling.
-
-% In above discussions, we propose an implicit state enumeration algorithm implemented by GB computation over
-% an elimination ideal, which is usually applied on Mealy FSMs working as control unit. For many arithmetic
-% datapath components, they should be modeled as Moore FSMs according to their structures. On the one hand, the signals preloaded
-% to a Moore FSM are usually not specified such that we can hardly encode the initial states; on the other
-% hand, instead of the set of globally reachable states, arithmetic verification focuses on exact reached states
-% after $k$ clock cycles which reflects desired arithmetic function. Considering these reasons, we modify
-% our algebraic geometry based approach and apply it to Moore FSM unrolling.
-
-% % Figure Moore.eps
-% \begin{figure}[hbt]
-% \centering{
-% %\begin{minipage}{12cm}
-% \includegraphics[width=7.5in]{./Moore.eps}
-% % \vspace{-0.2in}
-% \caption{A simple Moore FSM and its state transition graph}
-% %\end{minipage}
-% \label{fig:Moore}}
-% \end{figure}
-
-Fig.\ref{fig:seqmodel}(b) shows a typical Moore FSM without primary inputs, $R$ is the word-level state variable. Let word-level indeterminate 
-$R_{init}$ denote the initial preloaded value of $R$, and $R_{i}$ denote the evaluation of $R$ after $i$ clock cycles
-(transitions). Transition relation $Tr:R_{i-1}\to R_i$ can be modeled similar to
-the image function in Algorithm \ref{alg:BFS}. Assume that after $k$ clock cycles, the machine should give a state output
-$R_{k} = \mathcal F(R_{init})$. To verify this result, we can implicitly unroll this FSM and execute transition
-function at word level for $k$ times, which is $R_{k} = Tr(R_{k-1}) = Tr(Tr(\cdots Tr(R_{init})\cdots)) = Tr^k(R_{init})$.
-
-Sequential Galois field (GF) multiplier is a typical sequential arithmetic component widely used in cryptographic systems.
-Since there is no primary input in the design, it can be regarded as Moore FSM (Fig.\ref{fig:RHmulti}(a))
-where $R = \sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
- \beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$. Notice that $\beta$ is the {\it normal element}
- which can be represented by a power of primitive element $\alpha:~\beta = \alpha^t$.
-The states are encoded by evaluations of state variables, which can be implicitly verified 
-by checking output-input function ($R = A\cdot B \pmod{P(\alpha)}$).
-
-% Additionally
-% the inner structure of GF multiplier is XOR-rich such that we can test our approach on very large
-% (>100 bits datapath) circuits.
+% \subsection{Research objective: To make the approach scalable}
 % 
-% Let us briefly describe the fundamentals behind the design of normal
-% basis sequential GF multipliers, so as to put in perspective the type
-% of designs that have been verified in this paper. Let $R =
-% \sum_{i=0}^{k-1} r_i \beta^{2^{i}}, ~A = \sum_{i=0}^{k-1} a_i
-% \beta^{2^{i}}, ~B = \sum_{i=0}^{k-1} b_i \beta^{2^{i}}$, then 
-% \[
-% R = A\cdot B = (\sum_{i=0}^{k-1} a_i \beta^{2^{i}}) (\sum_{j=0}^{k-1}
-% b_j \beta^{2^{j}})  =
-% \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}\nonumber 
-% \]
+% At the time of writing this proposal, the theoretical concepts behind
+% the word-level FSM traversal have been studied and validated. A
+% completed algorithm has been devised and preliminary experiments were
+% conducted for reachability analysis using the {\sc Singular} computer
+% algebra tool \cite{DGPS}. While the approach is shown to work
+% correctly and completely, it is practically infeasible. The complexity
+% lies in the Gr\"obner basis computation. 
 % 
-% The expressions $\beta^{2^i}\beta^{2^j}$ are called cross-product
-% terms and they can also be represented in normal basis: 
-% \begin{displaymath}
-% \beta^{2^i}\beta^{2^j} =
-% \sum_{n=0}^{k-1}\lambda_{ij}^{(n)}\beta^{2^n}, \ \ \lambda_{ij}^{(n)}
-% \in \Ftwo. 
-% \end{displaymath}
+% Over Galois fields of the type $\Fkk$, computing the reduced Gr\"obner
+% basis of the ideal $(J + J_0)$ requires both memory and space
+% ${2^{k}}^{O(n)}$; where $J$ is an arbitrary polynomial ideal, $J_0$ is
+% the vanishing ideal, and $n$ is the number of variables in the system
+% (circuit). In this research, I wish to draw inspirations from the work
+% of \cite{jinpeng} and \cite{timDAC} to overcome the
+% complexity of GB computations. 
 % 
-% From the above two equations, one can see that the expression for the
-% $n^{th}$ digit of product $R = (r_0, \dots, r_n, \dots r_{k-1})$ is:
-% \[
-% r_n = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(n)}a_ib_j = A
-% \cdot M_n \cdot B^T, ~~0 \leq n \leq k-1
-% \]
 % 
-% where $M_n = (\lambda_{ij}^{(n)})$ is a binary $k \times k$ matrix over
-% $\Ftwo$, and it is called the $\lambda$-matrix. 
-%The collection of all
-%$\{M_n\}$  $\lambda$-matrices is called the multiplication
-%table. 
-% Moreover, let $r_n = A \cdot M_n \cdot B^T$.
-% Then $r_{n-1} = A \cdot M_{n-1} \cdot B^T = rotate(A) \cdot M_n
-% \cdot rotate(B)^T$. This implies that $M_n$ is generated by right
-% and down cyclic shifting $M_{n-1}$. Therefore, the hardware design
-% of sequential GF multipliers is based on mappings of $A\cdot M_n
-% \cdot B^T$ into AND-XOR gates and cyclic shift operations.
-
-% This paper verifies the implementation of two distinct 
-% architectures of {\it sequential multipliers with parallel output
-%   (SMPO)}, namely: i) the Agnew-SMPO  \cite{agnew1991implementation}
-% by G. B. Agnew, which is a straight-forward implementation of the
-% $\lambda$-matrix; and ii) the more recent, more complicated, yet very
-% efficient RH-SMPO \cite{RHmulti}, by Reyhani-Masoleh and Hasan,
-% depicted in Fig. \ref{fig:RHmulti}.
-
-In this proposal we use the verification of a sequential GF multiplier -- 
-a 3-bit RH-SMPO invented by Reyhani-Masoleh and Hasan\cite{RHmulti} -- to illustrate our approach.
-The normal element is given as $\beta = \alpha^3$, and primitive polynomial $P(\alpha) = \alpha^3+\alpha+1$.
-The gate-level circuit is depicted in Fig.\ref{fig:RHmulti}(b).
-
-\begin{figure}[hbt]
-\centering{
-%\begin{minipage}{12cm}
-\includegraphics[width=7.5in]{./new_multi.eps}
-% \vspace{-0.2in}
-\caption{A 3-bit RH-SMPO and its Moore FSM model}
-%\end{minipage}
-\label{fig:RHmulti}}
-\end{figure}
-
-We follow  the typical sequential GF circuit model with word-level variables $A, B, R$
-denoting {\it present states (PS)} and $A', B', R'$ denoting {\it next
-  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
-for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
-\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
-store the result, and $A, B\ (A', B')$ store input operands. {\it E.g.,}
-for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
-0$) are the initial values (operands) loaded into the registers,  and
-$R = \F(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
-result after $k$-cycles. Our approach aims to find this polynomial
-representation for $R$.
-
-Each gate in the combinational logic is represented by a Boolean
-polynomial. To 
-this set of Boolean polynomials, we append the polynomials that define
-the word-level to bit-level relations for PS and NS variables ($A =
-\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
-as ideal $J = \langle 
-f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
-  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
-  of the circuit. The ideal of vanishing polynomials $J_0$ is also included, and
-then the implicit FSM unrolling problem is setup for abstraction.
-
-The configurations of the flip-flops are the states of the
-machine. {\it Since the set of states is a finite set of points, we
-can consider it as the variety of an ideal related to the circuit
-}. Moreover, since we are interested in
-the {\it function encoded} by the state variables (over $k$-time
-frames), we can {\it project this variety} on the word-level state
-variables, starting from the initial state $A_{init}, B_{init}$.
-Projection of varieties (geometry) corresponds to elimination ideals
-(algebra), and can be analyzed via \Grobner bases. Therefore, we
-employ a \Grobner basis computation with abstraction term order (ATO)\cite{timDAC}: we use a {\it lex term
-  order} with {\it bit-level variables} 
-$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
-allows to eliminate all the bit-level variables 
-%(corresponding to the combinational logic and the state variables),
-%so as to 
-and derives a representation only in terms of words. 
-Consequently, $k$-successive \Grobner basis computations implicitly
-unroll the machine, and provide word-level algebraic $k$-cycle
-abstraction for $R'$ as $R' = \F(A_{init}, B_{init})$.
-
-Algorithm
-\ref{alg:modified} describes our approach.  In the algorithm, $from_i$
-and $to_i$ are polynomial ideals whose varieties are the valuations of
-word-level variables $R, A, B$ and $R',A',B'$ in the $i$-th iteration;
-and the notation ``$\setminus$'' signifies that the $NS$ in iteration
-$(i)$ becomes the $PS$ in iteration $(i+1)$. Line 5 computes the Gr\"obner 
-basis with the abstraction term order.  Line 6 computes the elimination 
-ideal, eliminating the bit-level variables and representing the set of 
-reachable states up to iteration $i$ in terms of the elimination ideal. 
-These computations are analogous to those of image computations performed in FSM reachability. 
-%The forward image
-%$to^{i}$ is computed using \Grobner bases with ATO.
-
-\vspace{-0.1in}
-\IncMargin{1em}
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
-\LinesNumbered
- \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
-   state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}
-
-  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
-  $i = 0$\;
-  \Repeat{$i == k$}
-  {
-  	$i \gets i + 1$\;
-%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
-%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
-	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
-    )$ with ATO\;
-	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
-	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
-  }
-\Return{$from_k(R_{final})$}
-\caption {Abstraction via implicit unrolling for Sequential GF circuit
-  verification}
-\label{alg:modified}
-\end{algorithm}
-\DecMargin{1em}
-\vspace{-0.1in}
-\begin{Example}
-\label{ex:RHSMPO}
-
-We demonstrate our approach to verify the 3-bit RH-SMPO circuit of
-Fig.\ref{fig:RHmulti}. The normal element $\beta$ in
-$\mathbb{F}_{2^3}$ is known to be $\beta = \alpha^3$, where $\alpha$
-is the primitive element. The circuit can be described with an ideal by translating
-AND and XOR gates accordingly. For the first iteration:
-\begin{align*}
-J = &d_0+b_2\cdot a_2,
-c_1+a_0+a_2,
-c_2+b_0+b_2,
-d_1+c_1\cdot c_2,\\
-&e_0+d_0+d_1,
-e_2+d_1,
-r_0'+r_2+e_0,
-r_1'+r_0,
-r_2'+r_1+e_2,\\
-&A+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
-B+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12},\\
-&R+r_0\alpha^3+r_1\alpha^6+r_2\alpha^{12},
-R'+r_0'\alpha^3+r_1'\alpha^6+r_2'\alpha^{12};
-\end{align*}
-The last 4 polynomials are translated from the definition of word-level variables.
-This represents ideal ``$J$" from line 5 in Algorithm
-\ref{alg:modified}. ``$J_0$" is the ideal of vanishing polynomials in all bit-level
-variables (e.g. $a_0^2-a_0$) and word-level variables (e.g. $A^8-A$). ``$from_{i-1}$"
-represents the set of current states for iteration $i$.
-In the first iteration, $from_0 = \{R, A_{init}+a_0\alpha^3+a_1\alpha^6+a_2\alpha^{12},
-B_{init}+b_0\alpha^3+b_1\alpha^6+b_2\alpha^{12}\}$.
-
-After the GB computation is performed with ATO, as line 6 in Algorithm \ref{alg:modified},
-we find a polynomial in variables $R', A_{init}, B_{init}$ in $to_1 : 
-R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
-$.
-
-Line 7 in Algorithm \ref{alg:modified} simply replaces NS output $R'$ with PS output
-$R$ in this example; so in second iteration $from_1 = \{
-R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
-, A_{init}+a_2\alpha^3+a_0\alpha^6+a_1\alpha^{12},
-B_{init}+b_2\alpha^3+b_0\alpha^6+b_1\alpha^{12}\}$.
-
-% By continue executing the loop,
-Finally, after 3 iterations we obtain: $to_3 = \{ \mathbf{R'+A_{init}B_{init},}
-~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
-~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
-as the image. The final result is $from_3(R_{final}) = R_{final}+A_{init}\cdot
-B_{init}$, which verifies the multiplier. 
-\end{Example}
-Using our approach we successfully verified the function of an 100-bit RH-multiplier,
-which implies its capability to handle the verification of large designs. The results
-are published in \cite{myDATE}.
+% In \cite{jinpeng}, the authors analyze the given circuit topology,
+% and derive a term order to represent the polynomials. They show that
+% this term order renders the set of polynomials itself a Gr\"obner
+% basis of the ideal. In \cite{timDAC}, the authors use a similar
+% concept to derive a {\it refinement of the abstraction term order}
+% (RATO) to simplify the Gr\"obner basis computation. However, the above
+% concepts are applied to combinational circuit verification. I will
+% investigate how similar techniques can be explored for the
+% computations that are needed in Algorithm \ref{alg:univa} for
+% reachability analysis to make the approach scalable. I believe that it
+% is possible to make the approach scalable, as I have recently
+% demonstrated in a paper \cite{myDATE} %{\bf refer to the DATE paper here?}
+% which has been accepted for publication at the {\it Design Automation
+%   and Test in Europe Conference}, 2015.

### 2015-05-13 techwriting/CAI_FMCAD/ref_alg.tex
update FMCAD2015 first draft

\section{Refined Algorithm}
Within the naive algorithm described in last section, Gr\"obner basis is computed for each
iteration to attain the word-level next state polynomial representation. In practice, for 
a sequential circuit with complicated structure and large datapath, Gr\"obner basis computation
is intractable. To overcome the high computational complexity brought by GB computing, 
we managed to find a method obviating directly computing GB; more specifically, we turn 
GB computation into one-step multivariate polynomial division. In this way, our proposed 
approach can be used on larger sequential circuit verification.

\subsection{Refined abstraction term order (RATO)}
A lexicographic order constrained by following relation $>_{r}$: "circuit variables ordered reverse topologically" $>$ 
"designated word-level output" $>$ "word-level inputs" is called the \emph{Refined Abstraction Term Order (RATO)}.

In Buchberger's algorithm, the specification polynomial (Spoly) for each pair is calculated. In RATO, most polynomials
have relatively prime leading terms/monomials (which means $Spoly \xrightarrow{J+J_0}_{+} 0$) except one pair:
word-level polynomial corresponding to outputs and its leading bit-level variable's gate description polynomial.
Its remainder $r$ lets following lemma hold:

\begin{Lemma}
\label{lem:1}
$r$ will only contain primary inputs (bit-level) and word-level output; furthermore, there will be one and
only one term containing word-level output whose monomial is word-level output itself rather than higher order form.
\end{Lemma}

\begin{proof}
First proposition is easy to prove by contradiction. Second part, the candidate pair of polynomials only have one term of
single word-level output variable (say it is $R$) and this term is the last term under RATO, which means there is only one term with
$R$ in Spoly. Meanwhile in other polynomials from $J+J_0$ there is no such term containing $R$, so this term will be
kept to remainder $r$, in first degree.
\end{proof}

\begin{Example}
After adding intermediate bit-level signal $a,b,c,d$, the elimination ideal for example circuit (Ex.\ref{ex:SMPO}) 
could be rewritten under RATO:
\begin{align}
&(t_0,t_1)>(a,b,c,d)\nonumber\\&>T>(x,s_0,s_1)\nonumber
\end{align}
Thus the candidate pair is
$(f_w,f_g), f_w = t_0+a\cdot b+a+b, f_g =t_0+t_1\alpha + T$.
Result after reduction is:
\begin{align}
&Spoly(f_w,f_g) \xrightarrow{J+J_0}_{+}\nonumber\\
&T + s_0 s_1 x+s_0 s_1 \alpha+(1+\alpha)s_0 x+(1+\alpha) s_0+s_1 x+s_1+(1+\alpha) x+1\nonumber
\end{align}
The remainder satisfies Lemma \ref{lem:1}.
\end{Example}

\subsection{Bit-level Variable Substitution (BLVS)}
\label{sec:blvs}
The remainder from \emph{Spoly} contains bit-level PS variables, and our objective is to get a polynomial contains only word-level PS variables
. One possible method is to rewrite bit-level variables in term of function on word-level
variables, i.e. $s_i = \mathcal{G}(S)$, then do substitution. A Gaussian-elimination-fashion approach could be applied to
compute corresponding $\mathcal{G}(S)$ efficiently.

\begin{Example}
{\bf Objective}:\ Abstract polynomial $s_i + \mathcal{G}_i(S)$ from $f_0: s_0+s_1\alpha+S$.

First, compute $f_0^2: s_0+s_1\alpha^2+S^2$. Apparently variable $s_0$ can be
eliminated by operation 
\begin{align}
f_1 =& f_0 + f_0^2: \nonumber\\
&(\alpha^2+\alpha)s_1+S^2+S\nonumber
\end{align}
Now we can solve univariate polynomial equation $f_1 = 0$ and get solution
$$s_1 = S^2 + S$$
Use this solution we can easily solve equation $f_0 = 0$. The result is
$$s_0 = \alpha S^2+(1+\alpha)S$$
\end{Example}
In this approach it is easy to get word-level variable representation for each bit-level PS variables. By substitution, a new polynomial in word-level PS/NS variables
could be attained.

After processing with RATO and BLVS techniques, we get a polynomial in form of $f_T = T+\mathcal{F}(S,x)$
denoting the transition function. We add a polynomial about $S$ to define the present states $f_S$,
as well as a polynomial guarantee free-end signals for primary inputs $f_x = x^2+x$, using elimination term order $S>x>T$,
we can generate a GB out of elimination ideal $\langle f_T,f_S,f_x\rangle$, which contains a univariate
polynomial denoting next states. The new scalable algorithm can be depicted as following:

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\F(S)$}

  $from^0 = reached = \F(S)$\;
  $f_T = $Reduce w/ RATO$(J_{ckt})$\;
  \Repeat{$new^i == 1$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$GB w/ elimination TO$\langle f_T, from^{i-1}, f_x\rangle$\;
	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
	$from^i \gets new^i(S\setminus T)$\;
  }
\Return{$reached$}
\caption {Refined Algebraic Geometry based Traversal Algorithm}\label{alg:refined}
\end{algorithm}

\subsection{PI Partition}
Using above techniques we can get a remainder polynomial with only word-level PS/NS variables. However in most 
cases, the number of bit-level PIs will be too large for the last-step Gr\"obner basis computation. Therefore it is necessary
to convert bit-level PIs to word-level PI variables.

Assume we have $k$-bit datapath and $n$-bit PIs. In Galois field, we need to carefully partition $n$ PIs
to guarantee states of all partitions can be covered by single univariate polynomials.

\begin{Proposition}
Divide $n$-bit PIs to partitions $n_1,n_2,\dots, n_s$ and let $n_1,n_2,\dots,n_s$ bit word-level variables
represent their evaluations in $\mathbb F_{2^k}$ accordingly. Then there always be $n_1,n_2,\dots,n_s~|~k$.
\end{Proposition}

Again, assume a partition $n_i~|~k$ and corresponding word-level variable is $P$. Then we can use polynomial
$P^{2^n_i}-P$ to represent all signals at free-end PIs, according to following theorem about {\bf composite
field}:
\begin{Theorem}
\label{thm:composite}
Let $k = m\dot n_i$, such that $\mathbb F_{2^k} = \mathbb F_{(2^{n_i})^m}$. Let $\alpha$ be primitive root of 
$\mathbb F_{2^k}$, $\beta$ be primitive root of the ground field $\mathbb F_{2^{n_i}}$. Then
$$\beta = \alpha^\omega,~~\text{where}~~\omega = \frac{2^k-1}{2^{n_i}-1}$$
\end{Theorem}
\begin{Example}
In a sequential circuit, PS/NS inputs/outputs are 4-bit signals, which means we will use $\mathbb F_{2^4}$
as working field. PIs are partitioned to 2-bit vectors, which means the ground field is $\mathbb F_{2^2}$.
In ground field we can represent all possible evaluations of this PI partition $\{p_0,p_1\}$ with
$$P^4+P,~~\text{where}~~P=p_0+p_1\cdot\beta$$
Using Thm.\ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
$$P = p_0 + p_1\cdot\alpha^5$$
\end{Example}
Using this method we can efficiently partition large size PIs to small number of word-level PI variables.
One limitation of this approach is PIs cannot be partitioned when $k$ is prime.

### 2015-05-14 techwriting/CAI_FMCAD/ref_alg.tex
fixing something in FMCAD

@@
   $from^0 = reached = \F(S)$\;
-  $f_T = $Reduce w/ RATO$(J_{ckt})$\;
-  \Repeat{$new^i == 1$}
+  $f_T = $Reduce(spoly$(f_w,f_s), J_{ckt})$\;
+  \tcc{Compute S-poly for the critical pair, then reduce it with circuit ideal under RATO}
+  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
   {
   	$i \gets i + 1$\;
-	$to^i \gets$GB w/ elimination TO$\langle f_T, from^{i-1}, f_x\rangle$\;
-	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
-  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
+  	$G \gets$GB($\langle f_T ,f_x, from^{i-1}\rangle$)\;
+  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest; $f_x$ is polynomial covers all possible inputs from PIs}
+	$\langle to^i \rangle \gets G\cup \Fkk[T]$\;
+	\tcc{There will be a univariate polynomial in $G$ denoting next state in word-level variable $T$}
+	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
+	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
+  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
+  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
 	$from^i \gets new^i(S\setminus T)$\;
+	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
   }
-\Return{$reached$}
+  \tcc{Loop until a fixpoint reached: newly reached state is empty}
+\Return{$\langle reached\rangle$}
 \caption {Refined Algebraic Geometry based Traversal Algorithm}\label{alg:refined}

### 2016-02-25 sequential/RH_related/SMPO_miter.pl
update SMPO_miter generator for type-1 ONB

@@ -4,17 +4,20 @@ use 5.014;
 use strict;
 use warnings;
 
-open MTR, ">", "miter_65bit.blif"; # change filename
+open MTR, ">", "SATmiter4.blif"; # change filename
 
-my $size = 65; # change this size
+my $size = 4; # change this size
 
 my $d = 1;
 my $c = 1;
 my $b = 1;
 my $i = 0;
-my @idx = qw /1 0 7 7 46 14 31 43 53 29 63 45 64 1 2 31 46 18 56 23 27 40 62 16 44 30 55 3 47 35 43 12 41 52 55 9 32 21 27 26 50 19 33 26 37 10 57 40 60 36 49 20 22 10 19 50 62 5 54 13 45 3 8 18 53 21 51 37 42 15 48 25 41 22 34 48 57 59 61 11 24 16 36 34 52 4 15 12 63 6 30 2 8 14 56 35 38 25 61 20 28 33 54 17 42 4 32 29 51 13 17 9 47 23 38 59 60 39 58 24 58 39 49 11 28 5 44 6 64/;
+my @idx = qw /2 2 3 0 1 1 3/;
 
-my @array = qw /0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64/;
+my @array = qw /0 1 2 3/;
+
+
+#my @array = qw /0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64/;
 
 #my @idx = qw /1 0 3 3 4 1 2 2 4/;
 my $j = 1;
@@ -23,15 +26,11 @@ print MTR ".model SMPOmiter\n.inputs ";
 for($i=0;$i<$size;$i++){
 print MTR "a$i b$i ";
 }
-print MTR "\n.outputs";
-for($i=0;$i<$size;$i++)
-{
-  print MTR " mo$i";
-}
-print MTR "\n";
+print MTR "\n.outputs Z\n";
+
 my $k;
 for($k = 0; $k < $size; $k++){
[…]

### 2016-02-25 sequential/RH_related/bigmiter.pl

@@ -1,137 +1,0 @@
-#!/usr/bin/perl -w
-
-use 5.014;
-use strict;
-use warnings;
-
-open MTR, ">", "SATmiter33.blif"; # change filename
-
-my $size = 33; # change this size
-
-my $d = 1;
-my $c = 1;
-my $b = 1;
-my $i = 0;
-my @idx = qw / 1 0 6 6 15 12 23 21 31 14 32 1 2 15 23 19 26 28 30 13 17 22 25 3 16 10 31 5 22 2 7 12 26 10 29 20 25 8 24 18 30 4 24 11 14 3 7 19 21 11 18 8 16 28 29 9 27 17 27 9 20 4 13 5 32 /;
-my @array = qw /0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32/;
-
-
-#my @array = qw /0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64/;
-
-#my @idx = qw /1 0 3 3 4 1 2 2 4/;
-my $j = 1;
-
-print MTR ".model SMPOmiter\n.inputs ";
-for($i=0;$i<$size;$i++){
-print MTR "a$i b$i ";
-}
-print MTR "\n.outputs Z\n";
-
-my $k;
-for($k = 0; $k < $size; $k++){
-printf MTR ".names b%d a%d d0\_r%d\n11 1\n", $k, ($k+1)%$size, $k;
-$j = 1; $b = 1; $c = 1; $d = 1;
-while ($j < $size) {
-printf MTR (".names a%d a%d c%d\_r%d\n01 1\n10 1\n", ($idx[$j*2-1]+$k)%$size, ($idx[$j*2]+$k)%$size, $c, $k); $c++;
-printf MTR (".names b%d c%d\_r%d d%d\_r%d\n11 1\n", ($b+$k)%$size, $c-1, $k, $d, $k); $b++; $d++;
-$j ++;
-}
-$i = 0;
-printf MTR ".names d0\_r%d d1\_r%d tmp0\_r%d\n01 1\n10 1\n", $k, $k, $k;
-for($i = 2; $i < $size - 1; $i++){
[…]

### 2016-02-27 techwriting/UNSAT-CORE/experiment.tex
SAT2016 write-up -- Xiaojun part

\section{Experiment results}
We have implemented our approach using the \textsc{Singular} symbolic
algebra computation system [v. 3-1-6] \cite{DGPS}. With our
tool implementation, we have performed experiments to extract a minimal UNSAT core from a given set of
polynomials. Our experiments run on a desktop with
3.5GHz Intel $\text{Core}^\text{TM}$ i7-4770K Quad-core CPU, 16 GB RAM and
64-bit Ubuntu Linux OS.

Nowadays most SAT benchmarks are huge, which choke our GB engine. In our experiment we use our customized 
benchmark library: "aim-100" is revised from ramdon 3-SAT benchmark "aim-50/100", "subset" is generated for
some random subset sum problem, "cocktail" is revised from a mixed factorization/random 3-SAT benchmark,
and "phole4/5" are generated by added some redundancies to pigeon hole benchmarks. Meanwhile we create "$3\times3$/$5\times5$ SMPO" benchmarks
by translating miter circuits of unrolled sequential normal basis multiplier \cite{SMPO}
and its specification circuit, and "2$\times$2 MasVMon" benchmark is the miter circuit of Mastrovito multiplier and Montgomery 
multiplier. These benchmarks are in moderate size for our GB engine but not too trivial.

\vspace{-0.1in}
\begin{table}[htb]
\centering
\caption{\small Results of running benchmarks using our tool}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
\multirow{4}{2cm}{Benchmark} 
& \multirow{4}{0.9cm}{\# Polys} 
& \multirow{4}{0.8cm}{\# MUS} 
& \multirow{4}{1.8cm}{Size of core by our tool}
 & \multirow{4}{1.6cm}{\# calls of GB engine}
 & \multirow{4}{2cm}{\# redundant polynomials by quotient cancellation}
 & \multirow{4}{1.3cm}{Runtime (sec)} \\
  & & & & & & \\
    & & & & & & \\
      & & & & & & \\
\hline
\hline
$5\times 5$ SMPO & 240  & 137  & 137  & 8  & 57 & 3160\\
\hline
aim-100 & 79 & 22 & 22 & 1  & 0 & 43\\
\hline
$3 \times 3$ SMPO & 17 & 2 & 2 & 1 & 0 & 0.07  \\
$2 \times 2$ MasVMont & 27 & 23 & 23 & 2 & 0 & 2.3 \\
subset & 100 & 6 & 6 & 1 & 0 & 2.4 \\
cocktail & 135 & 4 & 4 & 2 & 1 & 5 \\
phole4 & 104 & 10 & 16 & 1 & 0 & 4.4 \\
phole5 & 169 & 19 & 25 & 3 & 2 & 14 \\
\hline
\end{tabular}
\label{tab:result}  
\end{table}

In table \ref{tab:result}, we list the details of our experiment results. It shows in most cases, our tool
can get to the minimal core when hitting a fixpoint. Meanwhile, with a limited number of calls to our GB
engine, our tool can remove a major part of redundancies which indicates our tool
can extract a relatively small core efficiently.

From the experiment results we can also make several observations. One observation is our tool generates
minimal core when reaching fixpoint, which reflects the effectiveness of our heuristic. Another fact is,
for some test instances, quite a few redundant polynomials are removed before we start calling
the GB engine. This shows the strong potential for eliminating generator polynomials by our
"cancellation due to quotients" heuristic.

### 2016-04-27 techwriting/journal/application.tex
start journal writing

--- 2015-05-01 techwriting/proposal/reacha.tex
@@
-\section{Reachability Analysis of Finite State Machines (FSMs) using Algebraic Geometry}
-% \subsection{An example illustrating our proposed approach}
-In our approach we use symbolic state reachability with algebraic
-geometry concepts. It is an abstraction based on word operand
-definition of datapaths in circuits, and it can be applied
-to arbitrary FSMs by bundling a set of bit-level variables together as
-one or several word-level variables.  The abstraction polynomial,
-encoding the reachable state space of the FSM, is obtained through
-computing a GB of an elimination ideal using elimination term order
-based on Theorem \ref{thm:elim}.
-
-The motivation behind our approach can be described as follows:
-\begin{itemize}
-\item Any finite set of points is the variety (solutions) of a
-  polynomial ideal. Since the state-space of a FSM is finite, it can
-  be construed as the variety of an ideal corresponding to the
-  function (circuit implementation, gates) of the   sequential circuit.
-\item Using canonical representations of polynomial ideals (Gr\"obner
-  bases), the transition functions and the sets of states can be used
-  for FSM traversal and property checking. 
-\item In our work, Gr\"obner bases are considered in lieu of BDDs or
-  SAT-solvers. Furthermore, instead of working in the Boolean domain
-  $\B$, we model the circuit and the properties over the Galois field
-  $\Fkk$. This allows to perform reachability analysis at the level of
-  $k$-bit words, as $\mathbb{F}_2 \subset \Fkk$.
-
-\end{itemize}
-
-Reachability analysis is a useful tool when checking sequential
-equivalence as well as other invariants. 
-%With word-level state
-%variables, we can do implicit state enumeration to provide a picture
-%of  reachable states. 
-Conceptually, the state-space of a FSM is traversed in a breadth-first
-manner, as shown in Algorithm \ref{alg:BFS}. % \cite{KallaPartialScan}: 
-In general, the algorithm operates on the FSM 
-$\mathcal{M} = (\sum, O, S, S^0, \Delta, \Lambda)$ underlying a
-sequential circuit. In such cases, the transition function $\Delta$
-and the initial states are represented and manipulated using Boolean
-representations such as BDDs or SAT solvers. The variables $from,
-reached, to, new$ represent characteristic functions of sets of
-states. Starting from the initial state $from^i = S^0$, the algorithm
-computes the states reachable in 1-step from $from^i$ in each iteration.
-In line 4 of algorithm \ref{alg:BFS}, the {\bf image computation} is
-used to compute the 1-step next reachable states. In
-\cite{KallaPartialScan}, the author implicitly represents the states
-and transition relations by Boolean functions and Boolean vectors,
-respectively.
-
-Let us describe the application of the algorithm on the  FSM circuit
-of Fig.\ref{fig:fsm}. {\it We will first describe its operation at the
-Boolean level, and then describe how this algorithm can be implemented
-using algebraic geometry at word level.}
-
-In Line 1 of the BFS algorithm, assume that the initial state
-is $S_3$ in Fig.\ref{fig:fsm}(b), which is encoded as 
-$S_3 = \{11\}$. Using Boolean variables $s_0, s_1$ for the present
-states, $from^0 = s_0\cdot s_1$ is represented as a Boolean formula.
-
-The {\it transition function} $\Delta$ is given by Boolean equations
-  of the flip-flops of the circuit: $t_i = \Delta_i(s, x)$, where
-  $t_i$ is the next state variable, $s$ represents the present   state
-  variables and $x$ represents the input variables. 
-The {\it Transition relation of the FSM} is then represented as: 
-\begin{equation} 
-T(s, x, t) =   \prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)
-\end{equation}
-where $n$ is the number of flip flops, and $\overline{\oplus}$ is XNOR
-operation. Let $from$ denote the set of initial states, then the
-image of the initial states, under the transition function $\Delta$ is
-finally computed as:
-\begin{align}
-\label{img}
-to = \text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
-  \cdot from ] = \exists _s ~\exists _x ~\prod_{i=1}^{n} (t_i
-\overline{\oplus } \Delta_i)\cdot from
-\end{align}
-
-Here, the notation $\exists x (f)$ represents the {\it existential
-  quantification of $f$ w.r.t. variable $x$}.
-
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
- \KwIn{Transition functions $\Delta$, initial state $S^0$}
-
-  $from^0 = reached = S^0$\;
-  \Repeat{$new^i == 0$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$Img$(\Delta, from^{i-1})$\;
-	$new^i \gets to^i \cap \overline{reached}$\;
-  	$reached \gets reached \cup new^i$\;
-	$from^i \gets new^i$\;
-  }
-\Return{$reached$}
-\caption {Breadth-first Traversal Algorithm for Reachability Analysis of FSMs}\label{alg:BFS}
-\end{algorithm}
-
-\begin{Example}
-For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
-functions of the machine as:
-
-\begin{align*}
-\Delta_1: & ~t_0 \overline{\oplus} ((\overline{x + s_0 + s_1}) + s_0 s_1)\\
-\Delta_2: & ~t_0 \overline{\oplus} (\overline{s_0}x + \overline{s_1}s_0)\\
-from:     & ~from^0 = s_0\cdot s_1
-\end{align*}
-
-When the formula of Eqn. (\ref{img}) is applied to compute 1-step
-reachability, $to = \exists _{s_0, s_1, x} (\Delta_1 \cdot \Delta_2
-\cdot from^0)$, we obtain $to = \overline{t_0}\cdot t_1$, which denotes
-the state $S_1 = \{01\}$ reached in 1-step from $S_3$.
-
-In the next iteration, the algorithm uses state $S_1 = \{01\}$ as the
-current (initial) state, and computes $S_2 = \{10\} = t_0\cdot
-\overline{t_1}$ as the next reachable state, and so on. 
-\end{Example}
-
-%% Let  $\Delta_i$ denote the transition function for $i^{th}$ bit of the
-%% output $T$
-%% (denoted by $t_i$), and it is described by a Boolean function. We can
-%% obtain the transition relation  for bit-vector $T$:
-%% $Tran(s_0,s_1,x,t_0,t_1) =
-%% \bigwedge_{i=1}^{2}(t_i\ \bar{\oplus}\ \Delta_i)$. Assume present
-%% states are represent by Boolean formulas $PS(s_0,s_1)$, then the image
-%% function is written as $\text{Img}(Tran,\ PS) =
-%% \exists_{s_0,s_1}\exists_{x}[Tran(s_0,s_1,x,t_0,t_1)\land
-%%   PS(s_0,s_1)]$, where $\exists_x f$ denotes the existential
-%% quantification of $f$ w.r.t. $x$.
-
-Our objective is to model the transition functions $\Delta$ as a
-polynomial ideal $J$, and to perform the image computations (Algorithm
-\ref{alg:BFS}, line 4) using Gr\"obner bases over Galois fields. {\bf
-This requires to perform quantifier elimination; which can be
-accomplished using the Gr\"obner basis computation over Galois fields
-by using elimination ideals} \cite{gao:qe-gf-gb}. Finally, the set
-union,  intersection and complement operations are also to be
-implemented in algebraic geometry.
-
-{\bf FSM Traversal at word-level over $\Fkk$:} 
-The state transition graph (STG) shown in Fig.\ref{fig:fsm}(a) uses a
-2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
-S_3\}$. We map these states to elements in $\mathbb{F}_{2^2}$, where
-$S_0 = 0, S_1 = 1, S_2 = \alpha, S_3 = \alpha+1$ where $\alpha^2 +
-\alpha + 1 = 0$.
-
-{\it Initial state:} In Line 1 of the algorithm, the initial state is
-specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
-S - 1 - \alpha$. Notice that if we consider the ideal generated by the
-initial state polynomial, $I = \langle f\rangle$, then its variety
-$V(I) = 1+\alpha$, corresponds to the state encoding $S_3 = \{11\} =
-1+\alpha$, where a polynomial in word-level variable $S$ encodes the
-initial state. 
-% with only one generator $f$, its variety
-% $V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 1+\alpha\}$, which equals to $\{1+\alpha\}$, the only
-% valid value $S_3$ can take.
-
-%Using theorems from Section \ref{sec:elim}, we implement the image
-%function by a GB computation on elimination ideal. Ex.\ref{ex:motiv}
-%is an example for our implementation of image function, i.e. one-step
-%reachability.
-
-{\bf Set operations:} When executing Line 5 and Line 6, we use
-\textbf{union}, \textbf{intersection} and \textbf{complement} of
-varieties over $\mathbb{F}_{2^k}$:
-
-\begin{Definition}
-\label{def:sum}
-({\bf Sum/Product of Ideals}) If $I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are 
-ideals in $\mathbb F[x_1, \dots, x_n]$, then the {\bf sum} of $I$ and $J$ is defined as
-$$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$$ And the {\bf product} of $I$ and $J$ is defined
-as
-\begin{equation}
-  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle \nonumber
-  \end{equation}
-\end{Definition}
-
-With concepts of ideal sums and products, we can obtain the
-intersection and union of affine varieties as:
-\begin{Theorem}
-\label{thm:unionintersect}
-If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf
-  V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) =
-{\bf V}(I) \bigcup {\bf V}(J)$. 
-\end{Theorem}
-
-In Line 5 of Alg.\ref{alg:BFS}, we need to compute the complement of a
-set of states. Assume that $J$ denotes a polynomial ideal whose
-variety $V(J)$ denotes a set of states. We require the computation of
-another polynomial ideal $J'$, such that $V(J') =
-\overline{V(J)}$. We have recently proved (the proofs are omitted)
-that this computation can be performed using the concept of {\bf ideal
-  quotient:}
-
-\begin{Definition}
-({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
-  F[x_1, \dots, x_n]$, then $I:J$ is the set
-  \begin{equation}
-  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}\nonumber
-  \end{equation}
-and is called the {\bf ideal quotient} of $I$ by $J$.
-\end{Definition}
-
-% However, complement of variety cannot be easily dealt by simple arithmetic on polynomial generators.
-% For non-trivial cases we can only prove that 
-% \begin{Theorem}
-% Let $I, J$ be ideals in $\mathbb F[x_1,\dots,x_n]$, then 
-% $${\bf V}(I:J) \supset {\bf V}({\bf I}({\bf V}(I) - {\bf V}(J)))$$
-% \end{Theorem}
-% Fortunately for most hardware verification cases, the form of polynomial generators are restricted.
-% A proposition has been proved by \cite{jinpeng} that after adding {\bf vanishing polynomials ideal}
-% the new composed ideal implies following corollary:
-
-For elimination ideals in circuit verification problems, we can obtain
-the complement of a variety through the following result:
-
-\begin{Theorem}
-\label{thm:quotient}
-Let $I, J$ be ideals with vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
-$${\bf V}(I:J) = {\bf V}(I) - {\bf V}(J)$$
-\end{Theorem}
-
-Let $J_0 = \langle x_1^{2^k} - x_1, \dots, x_n^{2^k} - x_n \rangle$
-denote the ideal of all vanishing polynomials in $\Fkk$. Then, we have
-$V(J_0) = (\Fkk)^{n}$; i.e. the variety of vanishing ideal contains
-all possible valuations of variables, so it constitutes the {\bf
-  universal set}. Subsequently, the {\bf absolute complement} can be
-computed as:
-
-\begin{Corollary}
-$$V(J') = \overline{{\bf V}(J)} = {\bf V}(J_0:J)$$
-
-\end{Corollary}
-
-In other words, the ideal $J'$, computed as $J' = J_0:J$ is such that
-$V(J') =\overline{V(J)}$.
-
-Since we can add vanishing polynomials for state sets
-$from^i, to^i, reached$, it is feasible to use above corollary. We
-will demonstrate the application of this important concept through an
-example. 
-% 
-% {\bf Now, STEP BY STEP, SHOW THE FULL Blown traversal COMPUTATION
-%   HERE.... Make sure to also show how elimination order with GB
-%   computation is also used...} \vspace{0.5in}
-
-Assume $I = \langle f\rangle  = \langle T^2 +
-(1+\alpha)\cdot T+\alpha\rangle $,  we can add vanishing polynomial
-$T^4 - T$ to this ideal such that its variety  $V(I) = \{a\ |\ a \in
-\mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$. For complement set
-of variety $\{1, \alpha\}$, the universal set is the variety of ideal
-of vanishing polynomial $V(\langle T^4-T\rangle ) =
-\{0,1,\alpha,1+\alpha\}$, 
-so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
-which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
-the result is $\{0,1+\alpha\}$.
-
-According to the discussion above, the BFS traversal can be completely implemented using algebraic geometry.
-as in Algorithm \ref{alg:univa}.
-
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
- \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\F(S)$}
-
-  $from^0 = reached = \F(S)$\;
-  \Repeat{$new^i == 1$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$GB w/ elimination term order$\langle J_{ckt},J_0, from^{i-1}\rangle$\;
-	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^4-T\rangle:\langle reached\rangle)$\;
-  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
-	$from^i \gets new^i(S\setminus T)$\;
-  }
-\Return{$reached$}
-\caption {Algebraic Geometry based Traversal Algorithm}\label{alg:univa}
-\end{algorithm}
-
-Here $from^i, to^i, new^i$ are all univariate polynomials in variables $S$ or $T$, due to GB computation
-with elimination term order.
-
-\begin{Example}
-We apply Algorithm \ref{alg:univa} to Ex.\ref{ex:motiv} to execute the FSM traversal. Let the initial state
-$from^0 = S$ (i.e. state $\{00\}$), in the first iteration we compose an elimination ideal $J$ with
-
-\begin{minipage}[h]{0.4\textwidth}
-\begin{align*}
-&f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
-&f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
-&f_3: S - s_0 - s_1\\
-&f_4: T - t_0 - t_1
-\end{align*}
-$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$
-\end{minipage}
-\begin{minipage}[h]{0.6\textwidth}
-\begin{align*}
-&f_5: x^2-x\\
-&f_6: s_0^2-s_0, ~~f_7: s_1^2-s_1\\
-&f_8: t_0^2-t_0, ~~~f_9: t_1^2-t_1\\
-&f_{10}: S^4-S, ~f_{11}:T^4-T
-\end{align*}
-$\ \ \ \ \ \ \ \ \ \ \  \ J_0 = \langle f_5,f_6,\dots,f_{11}\rangle$
-\end{minipage}
-
-and $from^0$. Compute the reduced GB for $J = J_{ckt}+J_0+\langle from^0\rangle$,
-with elimination term order
-$$\{x,s_0,s_1,t_0,t_1\}~(all~PI~and~bit~level~variables)~>~S~(PS~word)~ >~ T~(NS~word)$$
-the result contains
-a polynomial generator with only $T$, we assign it to next state $$to^1 = T^2+(\alpha+1)T+\alpha$$ denoting
-$\{01,10\}$. Since formerly reached state $reach = T$, its complement is $$\langle T^4-T\rangle:\langle T\rangle
-= T^3+1$$ (i.e. states $\{01,10,11\}$). Then the newly reached state in this iteration is
-$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle$$ We add these states
-to formerly reached states $$reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$
-(i.e. states $\{00,01,10\}$). We update the present states for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha$$
-
-In the second iteration, we compute the reduced GB with the same term order for ideal $J = J_{ckt}+J_0+\langle from^1\rangle$.
-It includes a polynomial generator $$to^2 = T^2+\alpha T$$ denotes states
-$\{00,10\}$. The complement of $reached$ is $$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
-= T + 1+\alpha$$ (i.e. states $\{11\}$). We compute the newly reached state 
-$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$$ 
-It means the newly reached state is empty, thus a fixpoint has been detected. The algorithm terminates and returns
-$$reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$ as final reachable states.
-
-% 
-% $J = \langle t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1), t_1 - (xs_0+x+s_0s_1+s_0), S - s_0 - s_1, T - t_0 - t_1, S \rangle$
-% and a vanishing polynomial ideal $J_0 = \langle x^2-x, s_0^2-s_0, s_1^2-s_1, t_0^2-t_0, t_1^2-t_1, S^4-S, T^4-T\rangle$
-% The elimination term order is $\{x,s_0,s_1,t_0,t_1\} > S > T$. Compute a reduced GB for $J+J_0$, the result has 
-% a polynomial generator with only $T$, we assign it to next state $to^1 = T^2+(\alpha+1)T+\alpha$ denoting
-% $\{01,10\}$. Since formerly reached state $reach = T$, its complement is $\langle T^4-T\rangle:\langle T\rangle
-% = T^3+1$ (i.e. states $\{01,10,11\}$). Then the newly reached state in this iteration is
-% $\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle$. We add these states
-% to formerly reached states $reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$
-% (i.e. states $\{00,01,10\}$). We update the present states for next iteration $from^1 = S^2+(\alpha+1)S+\alpha$.
-% In the second iteration, we just replace 1 generator in $J+J_0$: $S$ with $S^2+(\alpha+1)S+\alpha$, and elimination
-% term order keeps the same. The reduced GB includes a polynomial generator $to^2 = T^2+\alpha T$ denotes states
-% $\{00,10\}$. The complement of $reached$ is $\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
-% = T + 1+\alpha$ (i.e. states $\{11\}$). We compute the newly reached state 
-% $\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$, then the algorithm terminates and returns
-% $reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$ as final reachable states.
-\end{Example}
-
-{\bf Significance for using GB:} Reduced GB is a unique, minimal and
-{\bf canonical} representation of the circuit's function. Starting
-from a certain initial state and using a reduced GB to represent the
-transition function, reachable states can be computed and represented
-canonically. Then it becomes possible to identify  when a fixpoint
-is reached (termination of the algorithm) by performing an equality
-check of polynomial ideals.
-
-\subsection{Research objective: To make the approach scalable}
-
-At the time of writing this proposal, the theoretical concepts behind
-the word-level FSM traversal have been studied and validated. A
-completed algorithm has been devised and preliminary experiments were
-conducted for reachability analysis using the {\sc Singular} computer
-algebra tool \cite{DGPS}. While the approach is shown to work
-correctly and completely, it is practically infeasible. The complexity
-lies in the Gr\"obner basis computation.
-
-Over Galois fields of the type $\Fkk$, computing the reduced Gr\"obner
-basis of the ideal $(J + J_0)$ requires both memory and space
-${2^{k}}^{O(n)}$; where $J$ is an arbitrary polynomial ideal, $J_0$ is
-the vanishing ideal, and $n$ is the number of variables in the system
-(circuit). In this research, I wish to draw inspirations from the work
-of \cite{jinpeng} and \cite{timDAC} to overcome the
-complexity of GB computations.
-
-In \cite{jinpeng}, the authors analyze the given circuit topology,
-and derive a term order to represent the polynomials. They show that
-this term order renders the set of polynomials itself a Gr\"obner
-basis of the ideal. In \cite{timDAC}, the authors use a similar
-concept to derive a {\it refinement of the abstraction term order}
-(RATO) to simplify the Gr\"obner basis computation. However, the above
-concepts are applied to combinational circuit verification. I will
-investigate how similar techniques can be explored for the
-computations that are needed in Algorithm \ref{alg:univa} for
-reachability analysis to make the approach scalable. I believe that it
-is possible to make the approach scalable, as I have recently
-demonstrated in a paper \cite{myDATE} %{\bf refer to the DATE paper here?}
-which has been accepted for publication at the {\it Design Automation
-  and Test in Europe Conference}, 2015.
-
-\subsection{Research objective : Application to Unrolling Sequential Circuits for Sequential Arithmetic Verification}
+\section{Application: Unrolling Sequential Circuits for Sequential Arithmetic Verification}
 In above discussions, we propose an implicit state enumeration algorithm implemented by GB computations over

### 2016-04-27 techwriting/journal/prelim.tex
start journal writing

--- 2015-05-13 techwriting/CAI_FMCAD/prelim.tex
@@
 \subsection{FSM model for sequential circuits}
+$\Mi$ Hey Hey
 A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 

### 2016-04-27 techwriting/journal/reacha.tex
start journal writing

--- 2015-05-13 techwriting/CAI_FMCAD/reacha.tex
@@
 
+The following example shows a simple ideal quotient operation:
+\begin{Example}
+In $\F[x,y,z]$, ideal $I = \langle xz,yz\rangle$, ideal $J = \langle z\rangle$. Then
+\begin{align*}
+I:J &= \{f\in\F[x,y,z]~|~f\cdot z \in \langle xz,yz\rangle\} \\
+&= \{f\in\F[x,y,z]~|~f\cdot z = Axz+Byz\}\\
+&= \{f\in\F[x,y,z]~|~f= Ax+By\}\\
+&= \langle x,y\rangle
+\end{align*}
+\end{Example}
+
 % However, complement of variety cannot be easily dealt by simple arithmetic on polynomial generators.
@@
 \label{thm:quotient}
-Let $J$ be ideals with single univariate polynomial about variable $x$ over $\Fkk[x]$, vanishing ideal $J_0 = \langle x^{2^k}-x\rangle$. then 
+Let $J$ be an ideal generated by a single univariate polynomial in variable $x$ over $\Fkk[x]$, vanishing ideal $J_0 = \langle x^{2^k}-x\rangle$. then 
 $${\bf V}(J_0:J) = {\bf V}(J_0) - {\bf V}(J)$$
@@
 \label{lem:gcd}
-Let $f,g\in \Fkk[x]$, then $(f:g) = \left(\frac{f}{gcd(f,g)}\right)$.
+Let $f,g\in \Fkk[x]$, then $\langle f:g\rangle = \left\langle\frac{f}{gcd(f,g)}\right\rangle$.
 \end{Lemma}
@@
 
-Take $h \in (f : g)$. So, $hg \in (f )$, that is $hg = f \cdot r$ with $r \in \Fkk[x]$.
+Take $h \in \langle f : g\rangle$. According to the definition, $hg \in \langle f \rangle$, which means $hg = f \cdot r$ with $r \in \Fkk[x]$.
 
-Therefore, $hdg_1 = df_1 r$, and so $hg_1 = f_1 r$. But since $gcd(g_1 , f_1 ) = 1$ we have that $f_1$
-divides $h$. Hence $h \in (f_1 )$.
-Conversely, let $h \in (f_1 )$. Then $h = s \cdot f_1$ , with $s \in \Fkk[x]$. So, $hg = hdg_1 = sf_1 dg_1 = sg_1 f \in
-(f )$. Therefore, $h \in (f : g)$.
+Therefore, $hdg_1 = df_1 r$ and $hg_1 = f_1 r$. But considering $gcd(g_1 , f_1 ) = 1$ we have the fact that $f_1$
+divides $h$. Hence $h \in \langle f_1\rangle$.
+
+Conversely, let $h \in \langle f_1 \rangle$. Then $h = s \cdot f_1$ ,where $s \in \Fkk[x]$. So, $hg = hdg_1 = sf_1 dg_1 = sg_1 f \in
+\langle f \rangle$. Therefore, $h \in \langle f : g\rangle$.
 \end{proof}
@@
 \begin{proof}
-Since $\Fkk[x]$ is a principal ideal domain, then$ J = (g)$ for some polynomial$ g \in \Fkk[x]$.
+Since $\Fkk[x]$ is a principal ideal domain, $ J = \langle g\rangle$ for some polynomial $ g \in \Fkk[x]$.
 
@@
 
-Then $J_0 : J = (f_1 )$ by Lemma \ref{lem:gcd}.
+Then $J_0 : J = \langle f_1 \rangle$ by Lemma \ref{lem:gcd}.
 
-Let $x \in{\bf V}_{\Fkk} (J_0 ) - {\bf V}_{\Fkk} (J)$. Then $x \in \Fkk$, but $g(x) \neq 0$.
+Let $x \in{\bf V}_{\Fkk} (J_0 ) - {\bf V}_{\Fkk} (J)$. From the definition of set complement, we get $x \in \Fkk$ while $g(x) \neq 0$.
 
-Since $x^{2^k} = x$, we see that either $d(x) = 0 or f_1 (x) = 0$. Note that since $g(x) \neq 0$, we can
-conclude that $d(x) = 0$. In conclusion, $f_1 (x) = 0$ and $x \in {\bf V}_{\Fkk} (f_1 )$.
+Since $x^{2^k} = x$, we see that either $d(x) = 0$ or $f_1 (x) = 0$. Considering $g(x) \neq 0$, we can
+assert that $d(x) = 0$. In conclusion, $f_1 (x) = 0$ and $x \in {\bf V}_{\Fkk} (f_1 )$.
 
-Now let $x \in V{\bf V}_{\Fkk} (f_1 )$. Therefore $f_1 (x) = 0$. So, $x^{2^k} - x = 0$ which gives $x \in {\bf V}_{\Fkk} (J_0 ) = \Fkk$.
+Now let $x \in {\bf V}_{\Fkk} (f_1 )$, we get $f_1 (x) = 0$. So, $x^{2^k} - x = 0$ which gives $x \in {\bf V}_{\Fkk} (J_0 ) = \Fkk$ which 
+contains all elements in the field.
 
-Assume now that $x ∈ {\bf V}_{\Fkk} (g)$. Then $0 = g(x) = d(x)g_1 (x)$.
-
-So, $d(x) = 0$ or $g_1 (x) = 0$.
+Now we make an assumption that $x \in {\bf V}_{\Fkk} (g)$. Then $0 = g(x) = d(x)g_1 (x)$ which means either $d(x) = 0$ or $g_1 (x) = 0$.
 
@@
   $from^0 = reached = \F(S)$\;
-  \Repeat{$new^i == 1$}
+  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
   {
   	$i \gets i + 1$\;
-	$to^i \gets$GB w/ elimination term order$\langle J_{ckt},J_0, from^{i-1}\rangle$\;
-	$new^i \gets $generator of $\langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
-  	$reached \gets $generator of $\langle reached\rangle \cdot \langle new^i\rangle$\;
+  	$G \gets$GB($\langle J_{ckt},J_0, from^{i-1}\rangle$)\;
+  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest}
+	$\langle to^i \rangle \gets G\cup \Fkk[T]$\;
+	\tcc{There will be a univariate polynomial in $G$ denoting next state in word-level variable $T$}
+	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
+	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
+  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
+  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
 	$from^i \gets new^i(S\setminus T)$\;
+	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
   }
-\Return{$reached$}
+  \tcc{Loop until a fixpoint reached: newly reached state is empty}
+\Return{$\langle reached\rangle$}
 \caption {Algebraic Geometry based Traversal Algorithm}\label{alg:univa}

### 2016-04-27 techwriting/journal/ref_alg.tex
start journal writing

--- 2015-05-14 techwriting/CAI_FMCAD/ref_alg.tex
@@
 . One possible method is to rewrite bit-level variables in term of function on word-level
-variables, i.e. $s_i = \mathcal{G}(S)$, then do substitution. A Gaussian-elimination-fashion approach could be applied to
-compute corresponding $\mathcal{G}(S)$ efficiently.
+variables, i.e.
+\begin{equation}
+\label{eqn:BLVS}
+s_i = \mathcal{G}(S)
+\end{equation}
+then do substitution. One way to obtain Eqn.\ref{eqn:BLVS} is to apply a Gaussian-elimination-fashion approach, which could 
+compute corresponding $\mathcal{G}(S)$ efficiently with time complexity $O(k^3)$.
 
@@
 \end{Example}
-In this approach it is easy to get word-level variable representation for each bit-level PS variables. By substitution, a new polynomial in word-level PS/NS variables
-could be attained.
+
+Another formalized method is to solve following system of equations
+\begin{align}
+\label{eqn:betamat}
+\begin{bmatrix}
+S \\
+S^2 \\
+S^{2^2} \\
+\vdots \\
+S^{2^{k-1}}
+\end{bmatrix}
+&=
+\begin{bmatrix}
+\beta & \beta^{2} & \beta^{2^2} & \cdots & \beta^{2^{k-1}}\\
+\beta^{2} & \beta^{2^2} & \beta^{2^3} & \cdots & \beta \\
+\beta^{2^2} & \beta^{2^3} & \beta^{2^4} & \cdots & \beta^2\\
+\vdots & \vdots & \vdots & \ddots & \vdots \\
+\beta^{2^{k-1}} & \beta & \beta^2 & \cdots & \beta^{2^{k-2}}
+\end{bmatrix}
+\begin{bmatrix}
+s_0\\
+s_1\\
+s_2\\
+\vdots\\
+s_{k-1}
+\end{bmatrix}
+\end{align}
+Let ${\bf s}$ be a vector of $k$ unknowns $s_0,\dots,s_{k-1}$,
+then Eqn.\ref{eqn:betamat} can be solved by using Cramer's rule:
+\begin{equation}
+\label{eqn:Cramer}
+s_i = \frac{\Mi}{\M}, \ \ 0\leq i\leq k-1, \M \neq 0
+\end{equation}
+where ${\bf M_i}$ denotes a coefficient matrix replacing $i$-th column in ${\bf M}$ with
+vector ${\bf S} = [S~~S^2~~\cdots~~S^{2^{k-1}}]^T$.
+
+Notice that ${\bf M}$ is constructed by squaring a row and assigning it to next row,
+therefore its determinant belongs to a special sort of determinants:
+\begin{Definition}
+Let $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ be a set of $k$ elements of $\Fpk$. Then the determinant
+\begin{align}
+\det{M(\alpha_0,\dots,\alpha_{k-1})} &= 
+\begin{vmatrix}
+\alpha_0 & \alpha_1 & \cdots & \alpha_{k-1} \\
+\alpha_0^p & \alpha_1^p & \cdots & \alpha_{k-1}^p \\
+\vdots & \vdots & \ddots & \vdots \\
+\alpha_0^{p^{k-1}} & \alpha_1^{p^{k-1}} & \cdots & \alpha_{k-1}^{p^{k-1}}
+\end{vmatrix}
+\end{align}
+is called the {\bf Moore determinant} of set $\{\alpha_0,\dots,\alpha_{k-1}\}$.
+\end{Definition}
+Moore determinant can be written as an explicit expression
+\begin{equation}
+\label{eqn:Moore}
+\det{M(\alpha_0,\dots,\alpha_{k-1})} = \alpha_0 \prod_{i=1}^{k-1} \prod_{c_0,\dots,c_{i-1}\in \Fp} (\alpha_{i} - \sum_{j=0}^{i-1}c_j\alpha_j)
+\end{equation}
+We use an example to help understanding the notations in Eqn.\ref{eqn:Moore}:
+\begin{Example}
+Let $\{\alpha_0,\alpha_1,\alpha_2\}$ be a set of elements of $\mathbb F_{2^3}$. Then
+\begin{align}
+\det{M(\alpha_0,\alpha_1,\alpha_2)} &= 
+\begin{vmatrix}
+\alpha_0 & \alpha_1 & \alpha_2 \\
+\alpha_0^2 & \alpha_1^2 & \alpha_2^2 \\
+\alpha_0^4 & \alpha_1^4 & \alpha_2^4
+\end{vmatrix} \\ \nonumber
+=& \alpha_0 \prod_{i=1}^{2} \prod_{c_0,\dots,c_{i-1}\in \Ftwo} (\alpha_{i} - \displaystyle\sum_{j=0}^{i-1}c_j\alpha_j)
+\end{align}
+First, let $i=1$,  we obtain $c_0\in \Ftwo$. When $c_0=0$, the product term equals to $\alpha_1$; when $c_0=1$ it equals to $(\alpha_1-\alpha_0)$.
+Then let $i=2$, we obtain ${c_0,c_1}\in \Ftwo$, they can take value from $\{0,0\},\{0,1\},\{1,0\}$ and $\{1,1\}$.
+We add 4 more product terms $\alpha_2,(\alpha_2-\alpha_1),(\alpha_2-\alpha_0),(\alpha_2-\alpha_0-\alpha_1)$ respectively.
+
+Thus, the result is
+\begin{equation}
+\det{M(\alpha_0,\alpha_1,\alpha_2)} = \alpha_0\alpha_1(\alpha_1-\alpha_0)\alpha_2(\alpha_2-\alpha_1)(\alpha_2-\alpha_0)(\alpha_2-\alpha_0-\alpha_1)
+\end{equation}
+\end{Example}
+We discover through investigation that $\M$ has a special property when the set of elements forms a basis. The proof is given below:
+\begin{Lemma}
+\label{lemma:Moore}
+Let $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ be a normal basis of $\Fpk$ over $\Fp$. Then
+\begin{equation}
+\det{M(\alpha_0,\alpha_1,\dots,\alpha_{k-1})} = 1
+\end{equation}
+\end{Lemma}
+\begin{proof}
+According to the definition Eqn.\ref{eqn:Moore}, the Moore determinant consists of all possible linear combinations of 
+$\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ with coefficients over $\Fp$. If $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$
+is a (normal) basis, then all product terms are distinct and represents all elements in the field $\Fpk$. Since the product 
+of all elements of a field equals to 1, the Moore determinant $\M=1$.
+\end{proof}
+Applying Lemma \ref{lemma:Moore} to Eqn.\ref{eqn:Cramer} gives
+\begin{equation}
+s_i = \Mi,~~0\leq i\leq k-1
+\end{equation}
+where $\Mi$ can be easily computed using Laplace expansion method, with complexity $O(k!)$.
+
+In this approach it is easy to get word-level variable representation for each bit-level PS variables. 
+By substitution, a new polynomial in word-level PS/NS variables could be attained.
 

### 2016-05-12 unsat/paper/BuchbergerAlg2.tex
Finalize UNSAT core staff(hopefully)

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

\theoremstyle{plain}
\newtheorem{defn}{Definition}[section]
%\usepackage{datetime}
\newtheorem{thm}[defn]{Theorem} % definition numbers are dependent on
% theorem
% %
\newtheorem{prop}[defn]{Proposition} % definition numbers are dependent on
\newtheorem{dem}[defn]{Proof} % definition numbers are dependent on
              % numbers
\newtheorem{exemp}{Example}[defn]
\newtheorem{lema}{\bf Lemma}[defn]

\def\opn#1#2{\def#1{\operatorname{#2}}} % to make operators

\opn\Ker{Ker} \opn\Coker{Coker}  \opn\Hom{Hom} \opn\Im{Im}
\opn\End{End} \opn\Aut{Aut} \opn\defect{def} \opn\ord{ord}
\opn\id{id} \opn\dim{dim} \opn\det{det} \opn\tr{tr} \opn\grad{grad} \opn\lcm{lcm}
\opn\min{min} \opn\max{max} \opn\det{det}
%\def\Frob{{\mathcal F}}
\opn\Span{Span}   \opn\rang{rang}  \opn\id{id} \opn\dim{dim} \opn\ad{ad} \opn\tr{tr} \opn\ker{ker}
\opn\GL{GL} \opn\SL{SL} \opn\mod{mod} \opn\diag{diag}
\opn\min{min} \opn\sgn{sgn} \opn\ini{in_<}  \opn\Mon{Mon} \opn\LC{LC_<}
\opn\LT{LT_<}
\opn\s{supp}

\title{On Cores of Ideals with Empty Variety}

\begin{document}

\maketitle

Let $F=\{f_1,f_2,\ldots,f_m\}$ be a set of $m$ polynomials in $\mathbb{F}_2[x_1,\ldots,x_n]$ and $I=(F)$ the ideal generated by the set $F$. Assume that our term order is LEX with $x_1>x_2>...>x_n$. Suppose that it is known that $V(I)=\emptyset$.

We can now apply the Buchberger's algorithm \cite{HH} to the system of generators $F$ of $I$.

Next, we will keep track of the S-polynomials that give non-zero remainder when divided by the system of generators of $I$ at that moment.
\begin{center}
$g_{ij}:= S(f_i,f_j)-\displaystyle\sum_{k=1}^l c_kf_k$,
\end{center}
where $0\neq c_k\in\mathbb{F}_2[x_1,\ldots,x_n]$ and $\{f_1,\ldots,f_l\}$ is the current system of generators of $I$.

For each non-zero $g_{ij}$, we will record the following data:
$$((g_{ij})(f_i,f_j)(h_{ij},h_{ji})| (f_1,\ldots,f_l),(c_1,\ldots,c_l))$$

 Note that here $g_{ij}$ denotes the remainder of the $S$-polynomial $S(f_i,f_j)$ modulo the current system of generators $f_1,\ldots,f_l$ of the ideal $I$ and we denote by 
$$h_{ij}:=\displaystyle\frac{\lcm(\ini(f_i),\ini(f_j))}{\LT(f_i)}, h_{ji}:=\displaystyle\frac{\lcm(\ini(f_i),\ini(f_j))}{\LT(f_j)}$$
the coefficients of $f_i$, respectively $f_j$ in the $S$-polynomial $S(f_i,f_j)$.

I recall that we have $$S(f_i,f_j)=\displaystyle\frac{\lcm(\ini(f_i),\ini(f_j))}{\LT(f_i)}f_i-\frac{\lcm(\ini(f_i),\ini(f_j))}{\LT(f_j)}f_j$$.

The last part contains the $f_i$'s that appear in the division process and the second parentheses has the corresponding coefficients.

 When we obtain $1$ as a remainder of a $S$-polynomial to the current system of generators, we stop.

As an output of the Buckberger's Algorithm, we can obtain not only the Gr\"obner bases, let's call it $\{g_1,\ldots,g_t\}$, but also a matrix $M$ such that
\begin{center}
\begin{align}
   \begin{bmatrix}
           g_{1} \\
           g_{2} \\
           \vdots \\
           g_{t}
         \end{bmatrix}
    &= M \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{m}
         \end{bmatrix}
  \end{align}

\end{center}
\begin{thm}
With the notations above, we have that a core for the system of generators $F$ of the ideal $I$ is given by the union of those $f_i$'s from $F$ that appear in the data recorded above and correspond to the nonzero entries in the matrix $M$. 
\end{thm}

\begin{proof}
In our case, $g_1=1$ and $t=1$, since the variety is empty, and hence the ideal is the unit ideal. Say $M= [a_1, \ldots, a_m]$. Then the output of the algorithm gives

$$1 = a_1f_1+\cdots + a_m f_m.$$ Clearly, if $a_i=0$ for some $i$ then $f_i$ does not appear in this equation and should not be included in the core.

\end{proof}

%
%

 Our next goal will be to describe an algorithm that allows us to compute a core of the ideal and the matrix $M$.

INITIAL DATA: $S_0=\{f_1,\ldots,f_m\}$ system of generators of the ideal $I$, monomial order $<$ on $\mathbb{F}_2[x_1,\ldots,x_n]$.

STEP 1: We start computing the S-polynomials of the system of generators $\{f_1,\ldots,f_m\}$ as in the Buchberger's algorithm \cite{HH}. After computing the first one, we divide it by the system of generators $S_0$. In this way, we will obtain the following expression
$$g_{i_1i_2}:= h_{i_1i_2}f_{i_1}-h_{i_2i_1}f_{i_2}-\displaystyle\sum_{k=1}^m c_kf_k$$

If the remainder $g_{i_1i_2}$ is non-zero, we will denote it by $f_{m+1}$ and we will add it to our initial system of generators. We will also record the data as follows
$$((f_{m+1}:=g_{i_1i_2})(f_{i_1},f_{i_2})(h_{i_1i_2},h_{i_2i_1})| (f_1,\ldots,f_m),(c_1,\ldots,c_m))$$ 
Then we will consider $S_1:=\{f_1,\ldots,f_m,f_{m+1}\}$ to be the new system of generators of $I$.

STEP s:  If $S_s:=\{f_1,\ldots,f_{s}\}$ is the current system of generators of $I$ and we have the following relation for $f_s$
$$f_s=g_{ij}=h_{ij}f_i-h_{ji}f_j-\displaystyle\sum_{k=1}^{s-1} a_kf_k$$

and the recorded data 
$$((f_{s}:=g_{ij})(f_{i},f_{j})(h_{ij},h_{ji})| (f_1,\ldots,f_{s-1}),(a_1,\ldots,a_{s-1}))$$

We will compute the next S-polynomial which gives a non-zero remainder when divided by our current system of generators $S_s$
$$f_{s+1}=g_{pq}=h_{pq}f_p-h_{qp}f_q-\displaystyle\sum_{k=1}^s b_kf_k$$

By substituting $f_s$ in the expression of $f_{s+1}$ we get 
$$f_{s+1}= h_{pq}f_p-h_{qp}f_q-\displaystyle\sum_{k=1}^{s-1} b_kf_k-b_s(h_{ij}f_i-h_{ji}f_j-\displaystyle\sum_{k=1}^{s-1} a_kf_k)$$
$$= h_{pq}f_p-h_{qp}f_q-\displaystyle\sum_{k=1}^{s-1} b_kf_k-b_sh_{ij}f_i+b_sh_{ji}f_j+\displaystyle\sum_{k=1}^{s-1} b_sa_kf_k$$
$$= h_{pq}f_p-h_{qp}f_q-b_sh_{ij}f_i+b_sh_{ji}f_j+\displaystyle\sum_{k=1}^{s-1} (b_sa_k-b_k)f_k$$

 Next, we will record the data for $f_{s+1}$
$$((f_{s+1}:=g_{pq})(f_{p},f_{q})(h_{pq},h_{qp})| (f_1,\ldots,f_{s-1},f_i,f_j),(b_sa_1-b_1,\ldots,b_sa_{s-1}-b_{s-1},b_sh_{ij},b_sh_{ji}))$$

The algorithm will stop when we will obtain the last non-zero remainder $f_l=1$. After using the previous relations, we will have the following
$$1=f_l=\displaystyle\sum_{k=1}^{m} d_kf_k$$
OUTPUT:
the core of the system of generators $S_0$ is $\{f_k: k\in\overline{1,m}, d_k\neq 0\}$

the matrix with the coefficients  $$M=(d_1,\ldots,d_k)$$.
	\begin{exemp}
Consider the following set of polynomials:
\begin{center}
$f_1: abc + ab + ac + bc + a + b + c + 1$
\end{center}
\begin{center}
$f_2: b$
\end{center}
\begin{center}
$f_3: ac$
\end{center}
\begin{center}
$f_4: ac + a$
\end{center}
\begin{center}
$f_5: bc + c$
\end{center}
\begin{center}
$f_6: abd + ad + bd + d$
\end{center}
\newpage
\begin{center}
$f_7: cd$
\end{center}
\begin{center}
$f_8: abd + ab + ad + bd + a + b + d + 1$
\end{center}
\begin{center}
$f_9: abd + ab +bd + b$
\end{center}

Let $S_0=\{f_1,\ldots,f_9\}$ and $I=(S_0)$ ideal in $\mathbb{F}_2[a,b,c,d]$. Then $V(I)=\emptyset$ as $GB(I)=\{1\}$.

INITIAL DATA: $S_0=\{f_1,\ldots,f_9\}$ system of generators of the ideal $I$, monomial order $<=<_{LEX}$ on $\mathbb{F}_2[a,b,c,d]$

STEP 1:

 We start computing the S-polynomials and their corresponding remaiders modulo the current system of generators.
 We keep track of the first non-zero remaider
$$f_{10}=g_{12}=f_1+acf_2+af_2+f_3+cf_2+f_2$$
recording the data as follows

\begin{center}
$((f_{10}=g_{12}), (f_1,f_2)(1,ac)|(f_2,f_3,f_2,f_2)(a,1,c,1))$
\end{center}
STEP 2:
 The second non-zero remainder is given by
$$f_{11}=g_{18}=df_1+cf_8+f_1+adf_2+af_2+df_{10}+df_2+f_7+f_{10}+f_{2}$$
and the corresponding recorded data 
$$((f_{11}=g_{18}), (f_1,f_8)(d,c)|(f_1,f_2,f_2,f_{10},f_2,f_7,f_{10},f_{2})(1,ad,a,d,d,1,1,1))$$

  Now by substituting $f_{10}$ in the expression of $f_{11}$, we obtain

$$f_{11}=(acd+ac+cd+c)f_2+(d+1)f_3+f_7+cf_8$$
so we have the following data
$$((f_{11}=g_{18}), (f_1,f_8)(0,c)|(f_2,f_3,f_7)(acd+ac+cd+c,d+1,1)$$

STEP 3:
The last non-zero remaider comes from the $S$-polynomial of $f_3$ and $f_4$
$$1=f_{12}=g_{34}=1=f_3+f_4+f_{10}+f_{11}$$
so we record the data 
$$((f_{12}=g_{34}=1), (f_3,f_4)(1,1)|(f_{10},f_{11})(1,1))$$
and after making the substitutions we obtain
$$1= f_1 + (acd + cd + a + 1)f_2 + (d+1)f_3 + f_4 + f_7 + c f_8$$

OUTPUT:

the core of the $S_0$ is given by  $\{f_1,f_2,f_3,f_4,f_7,f_8\}$

the matrix $M$ is given by $$M=(1,acd + cd + a + 1,d+1,1,1,c)$$

\end{exemp}				
\begin{thebibliography}{99}

\bibitem{HH}  J\"urgen Herzog, Viviana Ene, \emph{Gr\"obner Bases in Commutative Algebra}, American Mathematical Society, 2012.
\end{thebibliography}

\end{document}

### 2016-05-12 unsat/paper/SAT2016_xiaojun/core_refine.tex
Finalize UNSAT core staff(hopefully)

\section{UNSAT core refinement}
As all other MUS extractors, our algorithm cannot generate a minimal (irreducible) core in one shot.
To get a smaller core, we need to restart our tool with the current core we have. In this section several heuristics
are applied to our algorithm to increase the probability to generate a smaller core in next iteration.

\subsection{Leaf polynomial cancellation due to quotients}
First of all, before we start a new iteration of our GB engine, we can analyze the information we get from last 
iteration and remove redundancy.
In the last section we extract an UNSAT core by including all leaf polynomials in our tree to refutation "1".
The tree represents a linear combination with polynomial coefficients
of leaf polynomials. For example, in Fig.\ref{fig:tree} refutation "1" can be written as equation
$$1 = c\cdot f_2 + f_5 + {\bf \it f_3} + f_4 + f_1 + ac\cdot f_2 + a\cdot f_2 + {\bf \it f_3} + c\cdot f_2 + f_2$$
By combining like terms, the RHS polynomial can be simplified as
$$1 = f_1 + (ac+a+1)f_2 + f_4 + f_5$$
Note that polynomial $f_3$ is eliminated as $f_3+f_3 = 0 \pmod 2$. 
Thus $f_3$ is redundant in the core we previously extracted.

\subsection{Refining heuristics}
After eliminating all redundant polynomials, we can call our GB engine with the new core.
A effective heuristic should enhance the probability that the refutation "1" is composed by fewer polynomials.
In our GB computing algorithm, we use a strategy to pick critical pairs such that polynomials with 
larger indices get involved as late as possible:
$$(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$$
Meanwhile, for the reduction process $Spoly(f_i,f_j)\xrightarrow{F}_+ f_{new}$, we pick divisor
polynomials from $F$ following the order of polynomial indices. Therefore, by renaming the polynomial indices
and moving specific polynomials to small indices, we can increase the likelihood they are selected into the UNSAT core.

A straightforward inspiration is, by moving polynomials that are most likely included in the 
refutation tree to small indices, we can reach a fixpoint with fewer calls to GB engine.
In this paper, we use 2 criteria to increase the probability that a polynomial is included in the 
refutation tree.  The one is the "distance" to refutation "1",
the other is the "frequency" a polynomial appears in refutation tree.

"Distance" in a refutation tree corresponds to the number of arcs on the shortest path from refutation "1"
to leaf polynomial
Usually polynomials with shorter distance to "1" are used as divisors in later steps of polynomial reductions,
which indicates they will more generally have lower degree leading terms. 
If we use a degree-lexicographic order, then term cancellations reduce the degree of the polynomial. So polynomials with lower degree leading terms are more 
likely to be used for reduction, such that the possibility they appear in refutation tree is larger. 
Similarly, the motivation for "frequency" of appearance is as follows:
polynomials appearing frequently in refutation tree also indicates they have some properties make
them "favourable" in UNSAT core selection. For example, their leading terms may contain a variable that suppose to be included in a necessary clause of a minimal core.

Our tool applied both heuristics: we sort the polynomials in the core by "distance" criterion,
then use frequency as tiebreaker -- if there are multiple polynomials with same "distance", we put the one appears most frequently in last refutation tree first. 
We use an example to illustrate our heuristic.
\vspace{5mm}\\
{\it Example:}\ \ Consider a set of 6 polynomials which is UNSAT:
\begin{align*}
f_1 &: x_1x_3+x_3\\
f_2 &: x_2 + 1\\
f_3 &: x_2x_3+x_2\\
f_4 &: x_2x_3\\
f_5 &: x_2x_3 + x_2 + x_3 + 1\\
f_6 &: x_1x_2x_3 +x_1x_3
\end{align*}
After the first iteration, we get a tree of refutation in Fig.\ref{fig:refine}(a). 
\begin{figure}[hbt]
\centering
\includegraphics[scale=0.25]{core_refine.eps}
\caption{Refutation trees of core refinement example}
\label{fig:refine}
\end{figure}
"Distance" from $f_2$ to "1" is 2, so we put it in first place. The "distance" and "frequency" for other 
polynomials are identical, so we keep their ordering untouched. We re-index the polynomial set
$$f_1'=f_2, f_2' = f_1, f_3' = f_3, f_4' = f_4$$
and apply our GB-core algorithm on $\{f_1',f_2',f_3',f_4'\}$. The result is as
Fig.\ref{fig:refine}(b) shows. We reach a fixpoint, and the core $\{f_1', f_3', f_4'\}$ i.e. $\{f_2,f_3,f_4\}$ is proved to be minimal.

### 2016-05-12 unsat/paper/abstract.tex
Finalize UNSAT core staff(hopefully)

\begin{abstract}
Finding small unsatisfiable subformulas (unsat cores) of infeasible
constraint problems is an active area of research. The problem has
been widely investigated in the propositional SAT domain. Analogous
investigations in the polynomial algebra domain are, however,
somewhat lacking. This paper investigates an algorithmic approach to
identify a small unsatisfiable core of an ideal in polynomial rings
with coefficients from any field, where the ideal is found to have an
empty variety. We show that such a core can be identified by means the
Buchberger's algorithm and its variations. We identify certain
conditions that are helpful in ascertaining whether or not a
polynomial from the given generating set is a part of the unsat
core. Our algorithm cannot guarantee a minimal unsat core; hence the
paper discusses opportunities for refinement of the identified
core. The approach is demonstrated on a few examples, and experiments
are conducted using an implementation of our algorithm within the
SINGULAR computer-algebra tool. 
\end{abstract}

### 2016-05-12 unsat/paper/cnfsat_prev.tex
Finalize UNSAT core staff(hopefully)

\section{UNSAT core extraction through CNF-SAT solvers}
There are 2 major kinds of minimal UNSAT core extraction algorithms,
one is based on deletion,  the other one is based on insertion. Both
algorithms require multiple calls of SAT solver to verify if a clause
is necessary,  so the major efforts to enhance efficiency are to
reduce the number of calls. Currently a large portion of SAT solvers
work on DPLL algorithm and its variations,  which use resolution --
SAT solver can produce an UNSAT core by reading resolution refutation
-- and are usually incremental,  thus it is possible to re-use
previously learnt information from SAT solving.

One of the state-of-art MUS extractors based on resolution is
Haifa-MUC \cite{Haifa2014}, it gives a proof of UNSAT to incremental
SAT solver rather than the original formula. Moreover,  several
optimizations are implemented within this tool to boost its
performance: maintaining a partial resolution proof instead of whole
proof can reduce processing time; applying Eager model rotation can
mark multiple clauses as necessary without calling  SAT solver. AMUSE
\cite{AMUSE} is a typical MUS extractor using insertion-based
algorithm. It adds selector variables to original formula,  and
constructs an unsatisfiable subformula by making assumptions.

Recently A. Belov et al. \cite{Joao2012} introduced hybrid MUS
extraction algorithm combining methods from  both deletion-based and
insertion-based extractors. It not only grows UNSAT core from small
subsets but also call SAT solvers for its clause removing
strategy. Recursive model rotation (RMR) technique is also used to
learn transition clauses to save the cost of calling SAT solvers.

### 2016-05-12 unsat/paper/core.tex
Finalize UNSAT core staff(hopefully)

\section{Finding the Unsat Core}
With the aforementioned concepts and notations, we can formally state
the unsat core problem over polynomial rings as follows:

\begin{Problem}
Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
the ring $R = \F[x_1,\dots,x_n]$ that generate ideal $J = \langle
f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
\emptyset$, or it is determined to be so by applying the Gr\"obner
basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
too. Borrowing the terminology from the Boolean SAT domain, we
call $F_c$ the infeasible core or the unsat core of $F$. 
\end{Problem}

Any set of polynomials $F$ with empty variety is an unsat core in
itself. There may be more than one unsat cores in $F$, some
polynomials may be common to multiple cores (i.e. the
cores may be non-disjoint), and these cores may have
different cardinalities $|F_c|$. The unsat core problems
are therefore further classified as: 
\bi
\item Identify a {\it minimum} core, i.e. a core of minimum
  cardinality. 
\item Identify a {\it minimal} core. An unsat core $F_{c}$ is minimal
  with respect to the property that while the variety of $F_{c}$ is
  empty, the variety of {\it any} subset of $F_c$ is {\it non-empty},
  i.e. $V( F_{c} - \{f_i\} ) \neq \emptyset$ for any $f_i \in
  F_c$. 
\item Identify any unsat core disregarding minimality; i.e. find any
  subset $F_c$ of $F$.
\ei

Let us first consider the problem of finding any $F_c \subset F$,
disregarding minimality. It is not hard to motivate that an unsat
core should be identifiable using the Gr\"obner basis
algorithm: Assume that $F_c = F - \{f_j\}$. If $GB(F) = GB(F_c) =
\{1\}$, then it implies that $f_j \in \langle F - \{f_j\}\rangle$. In
other words, 
$f_j$ can be composed of the other polynomials of $F_c$, so $f_j$ is
not a part of the unsat core, and it can be safely discarded from
$F_c$. This can be identified by means of the GB algorithm for this
ideal membership test (due to Def. \ref{def:gb}(ii)).

A n\"aive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
selectr a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

\subsection{Investigating algorithmic techniques to find $F_c$ from $F$}

We investigate if it is possible to identify a core by analyzing the
$Spoly(f_i,f_j)\xrightarrow{F}_+ r$ reductions in Buchberger's
algorithm. Since $F$ is an unsat core, 
definitely {\it there exists  
a sequence of Spoly reductions in Buchberger's algorithm where
$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
polynomial reduction algorithms can be suitably modified to record
which polynomials from $F$ are used in the division leading to
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that by recording
the {\it ``data''} generated by Buchberger's algorithm --- i.e. the
critical pair polynomials ($f_i,f_j$) used in the Spoly computations,
and the polynomials from $F$ used to cancel terms in the reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+1$ --- we should be able to identify a
core.

We will use an example to demonstrate our approach of identifying $F_c
\subseteq F$ using this ``data'':

\begin{example}
\label{ex:1}
Consider the following set of polynomials:
\begin{center}
$f_1: abc + ab + ac + bc + a + b + c + 1$
\end{center}
\begin{center}
$f_2: b$
\end{center}
\begin{center}
$f_3: ac$
\end{center}
\begin{center}
$f_4: ac + a$
\end{center}
\begin{center}
$f_5: bc + c$
\end{center}
\begin{center}
$f_6: abd + ad + bd + d$
\end{center}
%\newpage
\begin{center}
$f_7: cd$
\end{center}
\begin{center}
$f_8: abd + ab + ad + bd + a + b + d + 1$
\end{center}
\begin{center}
$f_9: abd + ab +bd + b$
\end{center}

Assume $>_{DEGLEX}$ monomial ordering with $a>b>c>d$. 
Let $F = \{f_1,\dots,f_9\}$ and 
$J = \langle F \rangle \subset \mathbb{F}_2[a,b,c,d]$ where
$\mathbb{F}_2 = \{0, 1\}$ is the finite field of 2 elements. Then 
$V(J) = \emptyset$ as $GB(J) = 1$.  The set $F$ consists of 4 {\it minimal}
cores: $F_{c1} = \{ f_1,f_2,f_3,f_4,f_7,f_8\}, F_{c2} = \{
f_2,f_4,f_5,f_6,f_8\}, F_{c3} = \{ f_2,f_3,f_4,f_6,f_8\},$ and $F_{c4}
= \{ f_1,f_2,f_4,f_5\}$. 
\end{example}

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs are selected and reduced as
$Spoly(f_i,f_j)\xrightarrow{F}_+r$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and $(f_2,f_5)$. It turns
out that the Spoly-reductions corresponding to these 3 pairs lead to
the unit ideal. Recording the ``data'' corresponding to this sequence
of reductions is depicted in Fig. \ref{fig:refute}. We call this
graph a refutation tree.

\begin{figure}[hbt]
\centering
\includegraphics[scale=0.25]{./SAT2016_xiaojun/refutation_tree.eps}
\caption{Generating refutation trees to record unsat cores.}
\label{fig:refute}
\end{figure}

In the figure, the nodes of the graph correspond to the polynomials
utilized in Buchberger's algorithm. The leaf nodes always correspond
to polynomials $f_i \in F$. An edge $e_{ij}$ from node $i$ to node $j$
signifies that the polynomial at node $j$ resulted from polynomial
$i$. For example, consider the computation
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
following sequence of 1-step divisions:
$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} \xrightarrow{f_3}
\xrightarrow{c\cdot f_2}  \xrightarrow{f_2} f_{10}$. This is depicted
as the bottom subtree in the figure, terminating at polynomial
$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
division by $f_2$ resulted in the quotient $a$. The refutation tree of
Fig. \ref{fig:refute} shows further that
$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.

To identify an $F_c \subset F$, we start from the refutation node
$''1''$, and traverse the graph in reverse, all the way upto the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an unsat core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Fig. \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an unsat core of $F$.

\subsubsection{Ordering critical pairs in Buchberger's algorithm:} It
is clear that the inclusion of polynomials in the obtained core
depends on the order in which the critical pairs are
reduced. This is a well-known intractable problem, and various
heurstics have been proposed to favourably influence the termination
of the Gr\"obner basis algorithm -- these include the sugar
strategy \cite{GBSelectionStrategies}, the preconditioning used by
{\it Faug\~{e}re} in the $F4$ algorithm \cite{f4}, among others. Any
such selection strategy can be also be applied in our setting to
identify a smaller core.

\section{Reducing the Size of the Infeasible Core $F_c$}

The core $F_c$ obtained from the aforementioned procedure is not
guaranteed to be minimal. For example, let us revisit the the core
$F_c=\{f_1,\dots, f_5\}$ generated in the previous section. Note that
while $F_c$ is a smaller infeasible core of $F$, it is not minimal. In
fact, Example 1 shows that $F_{c4} =\{f_1,f_2,f_4,f_5\}$ is the
minimal core where $F_{c4} \subset F_{c}$. Clearly, the polynomial
$f_3 \in F_c$ is a redundant element of the core. It could be
discarded. We will now describe how the size of this core could be
reduced further.  For this purpose, we will have to perform a more
systematic book-keeping of the data generated during the execution of
Buchberger's algorithm and the refutation tree.

\subsection{Identifying redundant polynomials from the refutation tree}

We will keep track of the S-polynomials that give a non-zero remainder
when divided by the system of polynomials $F$ at that moment:
\begin{equation}
\label{eqn1}
g_{ij}= S(f_i,f_j)+\displaystyle\sum_{k=1}^m c_kf_k,
\end{equation}
where $0\neq c_k\in\mathbb{F}[x_1,\ldots,x_n]$ and
$\{f_1,\ldots,f_l\}$ is the ``current'' system of polynomials
$F$. For each non-zero $g_{ij}$, we will record the following data:

\begin{equation}
\label{data1}
((g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{equation}

Note that in (\ref{eqn1}) and (\ref{data1}), $g_{ij}$ denotes the
remainder of the $S$-polynomial $S(f_i,f_j)$ modulo the current system
of polynomials $f_1,\ldots,f_m$, and we denote by

$$h_{ij}:=\displaystyle\frac{\lcm(lm(f_i),lm(f_j))}{lt(f_i)},
h_{ji}:=\displaystyle\frac{\lcm(lm(f_i),lm(f_j))}{lt(f_j)}$$ 
the coefficients of $f_i$, respectively $f_j$, in the $S$-polynomial
$S(f_i,f_j)$. Furthermore, in (\ref{data1}), $(c_{k1}, \dots c_{kl})$ are the
polynomial coefficients of $(f_{k1}, \dots,f_{kl})$ that appear in the
division process.

\begin{example}
Revisiting Example \ref{ex:1},  the data corresponding to
$Spoly(f_1,f_2)\xrightarrow{F}_+  g_{12} = f_{10}$ reduction is
obtained as the following sequence of computations:
$$f_{10}=g_{12}=f_1-acf_2-af_2-f_3-cf_2-f_2.$$ As the coefficient
field is $\mathbb{F}_2$ in this example, $-1 = +1$, so:
$$f_{10}=g_{12}=f_1+acf_2+af_2+f_3+cf_2+f_2$$ is obtained.
The data is recorded according to (\ref{data1}):

\begin{center}
$((f_{10}=g_{12}), (f_1,1)(f_2,ac)|(a,f_2),(1,f_3),(c,f_2),(1,f_2))$
\end{center}

\end{example}

Of course, our approach and the book-keeping stops when we obtain
1 as the remainder of the S-polynomial modulo the current system of
generators. As an output of the Buchberger's algorithm, we can obtain
not only the Gr\"obner bases,  $G = \{g_1,\ldots,g_t\}$, but also a
matrix $M$ such that: 
\begin{center}
\begin{align}
   \begin{bmatrix}
           g_{1} \\
           g_{2} \\
           \vdots \\
           g_{t}
         \end{bmatrix}
    &= M \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
  \end{align}

\end{center}

Each element $g_i$ of $G$ is a polynomial combination of $f_1, \dots,
f_s\}$. Moreover, this matrix $M$ is constructed precisely using the
data that is recorded in the form of (\ref{data1}). We now a condition
when the matrix $M$ may identify some redundant elements.

\begin{theorem}
\label{thm}
With the notations above, we have that a core for the system of
generators $F = \{f_1,\dots,f_s\}$ of the ideal $J$ is given by the
union of those $f_i$'s from $F$ that appear in the data recorded above
and correspond to the nonzero entries in the matrix $M$.  
\end{theorem}

\begin{proof}
In our case, since the variety is empty, and hence the ideal is the
unit ideal, we have that $G = \{g_1=1\}$ and $t=1$. Therefore $M=
[a_1, \ldots, a_s]$ is a vector. Then the output of the algorithm
gives:

$$1 = a_1f_1+\cdots + a_s f_s.$$ Clearly, if $a_i=0$ for some $i$ then
$f_i$ does not appear in this equation and should not be included in
the infeasible core of $F$.

\end{proof}

\begin{example}

Corresponding to Example \ref{ex:1} and the refutation tree shown in
Fig. \ref{fig:refute}, we discover that the polynomial $f_3$ is used
only twice in the division process. In both occasions, the quotient of
the division is 1. Therefore, from Fig. \ref{fig:refute}, it follows that:
\begin{equation}
1 = (f_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot f_3}+ \dots (f_1
+ f_2)
\end{equation}

Since $1 + 1 = 0$ over $\F_2$, we have that the entry in $M$
corresponding to $f_3$ is 0, and so $f_3$ can be discarded from the
core. 
\end{example}

\subsection{The GB-Core Algorithm Outline}

The following steps describe an algorithm (GB-Core) that allows us to compute a
refutation tree of the polynomial set and corresponding matrix $M$.

{\bf Inputs:} Given a system of polynomials $F=\{f_1,\ldots,f_s\}$
, a monomial order $>$ on $\mathbb{F}[x_1,\ldots,x_n]$.

{\bf S-polynomial reduction:} We start computing the S-polynomials of the system of
generators $\{f_1,\ldots,f_s\}$, then divide each of them by the system of
generators $G=\{f_1,\dots,f_s,\dots,f_m\}$ which is the intermediate result of Buchberger's algorithm. 
In this way, we obtain the following expression:
\begin{equation}
\label{eqn:red}
g_{ij}= h_{ij}f_{i}+h_{ji}f_{j}+\displaystyle\sum_{k=1}^m c_kf_k
\end{equation}
If the remainder $g_{ij}$ is non-zero, we will denote it by
$f_{m+1}$ and we will add it to intermediate system of generators $G$. We
will also record the data as (\ref{data1}): 
\begin{displaymath}
((f_{m+1}=g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
It forms a tree with root $f_{m+1}$.

{\bf Coefficients recording:} In (\ref{eqn:red}) we get a coefficient vector including coefficients
$c_k$ where $k>s$. These coefficients are associated with new generators in the Gr\"obner basis,
which cannot benefit the UNSAT core extraction. Therefore we need to rewrite the vector to
one with only $c_k, k=1\dots s$, which can be achieved by substituting $f_k,k>s$ by $f_1,\dots,f_s$:
initially we have $f_{s+1}= c_1f_1 + c_2f_2 + \cdots + (h_{ij}+c_i)f_{i}+\cdots+(h_{ji}+c_j)f_{j}+\cdots+c_sf_s$,
inductively if we can present $f_{s+1}$ to $f_{s+k-1}$ by $f_1,\dots,f_s$, we can also rewrite
$f_{s+k} = d_1f_1+\cdots+d_sf_s$. Then we add a new row $ (d_1,\dots,d_s ) $ to coefficient matrix $M$.

{\bf Termination and refutation tree construction:} By repeating S-polynomial reduction and coefficients 
recording until we get remainder $f_m = 1$, the algorithm
will generate a set of new generators $G$ and coefficient matrix $M$.
Meanwhile we can construct the refutation tree with the recorded data.
We start with 
\begin{displaymath}
((f_{m}=1)(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
It represents a tree start from leaves $f_{i}, f_j$ then internal nodes $f_{k1},\dots,f_{kl}$
and ``1" as the root. Next we substitute nodes $f_{m-1}, f_{m-2}, \dots, f_{s+1}$ in order
with corresponding subtrees. The algorithm returns the complete refutation tree as well as
matrix $M$.

The complete algorithm is concluded as pseudo code below:

\begin{algorithm}[H]
 \caption{GB-core algorithm (based on Buchberger's algorithm)}
 \label{algo:gbcore}
 \begin{algorithmic}[1]

 \REQUIRE $F = \{f_1, \dots, f_s\} \in \F[x_1, \dots, x_n], f_i\neq 0$
 \ENSURE Refutation tree $T$ and coefficients matrix $M$
 \STATE{ {Initialize: list $G \gets F$; Dataset $D\gets \emptyset$; $M\gets s\times s$ unit matrix} }
 \FOR {{ each pair $(f_i,f_j)\in G$  }}
 	\STATE  $f_{sp},(f_{i},h_{ij})(f_{j},h_{ji}) \gets$ spoly-pair($f_i,f_j$) \COMMENT{$f_{sp}$ is the S-polynomial}
 	\STATE{{ $g_{ij}|(c_{k1},f_{k1}),\dots,(c_{kl},f_{kl}) \gets (sp\xrightarrow{G}_+  g_{ij})$}}
 	\IF{{ $g_{ij} \neq 0$ }}
 		\STATE {{$G \gets G \cup g_{ij}$}}
 		\STATE {{Update $D$ and $M$}}
 	\ENDIF
 	\IF {{$g_{ij} = 1$}}
 		\STATE {{Construct $T$ from $D$}}
 		\STATE {{ Return($T,M$) }}
 	\ENDIF
 \ENDFOR
 \end{algorithmic}
 \end{algorithm}

\section{Iterative Refinement of the Unsat Core}
Similar to all other core extractors, our algorithm cannot generate a
minimal core in one execution. To obtain a smaller core, we need to
re-execute our tool with the core obtained in the current
iteration. We describe two heuristics that are applied to our
algorithm to increase the probability to generate a smaller core in
the next iteration.

After eliminating all redundant polynomials, we can call our GB engine
with the new core. A effective heuristic should enhance the
probability that the refutation "1" is composed by fewer polynomials. 
In our GB computing algorithm, we use a strategy to pick critical
pairs such that polynomials with  larger indexes get involved as late
as possible: 
$$(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$$

Meanwhile, for the reduction process $Spoly(f_i,f_j)\xrightarrow{F}_+
f_{new}$, we pick divisor polynomials from $F$ following the order of
polynomial indexes. Therefore, by renaming the polynomial indexes, we
can increase the likelihood they are selected into the unsat core. 
We use 2 criteria to increase the probability that a polynomial is
included in the refutation tree.  One corresponds to the \emph{refutation distance},
whereas the other corresponds to the frequency that
a polynomial appears in refutation tree.

\begin{definition}
\emph{Refutation distance} in a refutation tree corresponds to the number of arcs on
the shortest path from refutation "1" to a leaf
polynomial.
\end{definition}

Polynomials with shorter refutation distance are used as
divisors in later stages of polynomial reductions, which indicates
they will generally have lower degree leading terms.  
This is because we are assuming a degree-lexicographic order, and
successive term cancellations reduce the degree of the remainders. So
polynomials with lower degree leading terms are more  likely to be
used for reduction, such that the probability that they appear in
the refutation tree is larger.

Similarly, the motivation for \emph{frequency of appearance} of $f_i$ in
the refutation tree is as follows: polynomials appearing frequently in
the refutation tree also indicates that they have some properties
(leading terms) that make them "favourable" in the unsat core
selection. For example, their leading terms may contain variables
that are require them to be included in the minimal core.

We apply both heuristics: we sort the polynomials in the core
by the refutation distance criterion, and use the frequency criterion as
the tiebreaker. The following example illustrates our heuristic.

%\vspace{5mm}\\
\begin{example} 
Consider a set of 6 polynomials over $\F_2$ of an
infeasible instance.
\begin{align*}
f_1 &: x_1x_3+x_3\\
f_2 &: x_2 + 1\\
f_3 &: x_2x_3+x_2\\
f_4 &: x_2x_3\\
f_5 &: x_2x_3 + x_2 + x_3 + 1\\
f_6 &: x_1x_2x_3 +x_1x_3
\end{align*}

After the first iteration, we get a tree of refutation as shown in Fig.\ref{fig:refine}(a). 
\begin{figure}[hbt]
\centering
\includegraphics[scale=0.25]{SAT2016_xiaojun/core_refine.eps}
\caption{Refutation trees of core refinement example}
\label{fig:refine}
\end{figure}

The shortest "Distance" corresponds to that of $f_2$ to "1", which is
2 levels. So, we reorder $f_2$ to be the polynomial with the shortest
index. The "distance" and "frequency" for other polynomials are
identical, so we keep their ordering untouched. We re-index the
polynomial set 
$$f_1'=f_2, f_2' = f_1, f_3' = f_3, f_4' = f_4$$
and apply our GB-core algorithm on $\{f_1',f_2',f_3',f_4'\}$. The
result is shown in Fig.\ref{fig:refine}(b). We reach a fixpoint, and
the core $\{f_1', f_3', f_4'\} = \{f_2,f_3,f_4\}$ is proved to be
minimal. 
\end{example}

### 2016-05-12 unsat/paper/experiment.tex
Finalize UNSAT core staff(hopefully)

\section{Experiment results}
We have implemented our core extraction approach (the GB-Core
algorithm) using the \textsc{Singular} symbolic
algebra computation system [v. 3-1-6] \cite{DGPS}. With our
tool implementation, we have performed experiments to extract a
minimal unsat core from a given set of  
polynomials. Our experiments run on a desktop with
3.5GHz Intel $\text{Core}^\text{TM}$ i7-4770K Quad-core CPU, 16 GB RAM and
64-bit Ubuntu Linux OS.

%Nowadays most SAT benchmarks are huge, which choke our GB engine. 
Gr\"obner basis is not an efficient engine for solving contemporary
CNF-SAT benchmarks that are of large size. In order to validate our
approach, we use a customized  benchmark library: "aim-100" is revised
from ramdon 3-SAT benchmark "aim-50/100", "subset" series are generated for
random subset sum problems, "cocktail" is revised from a mixed
factorization/random 3-SAT benchmark, and "phole4/5" are generated by
adding redundandant clauses to pigeon hole benchmarks. Moreover, we
created SMPO/RH benchmarks that correspond to
equivalence checking instances of sequential Galois field normal basis
modulo multiplier implementations (also known as Agnew's multiplier \cite{SMPO}
and Reyhani-Hassan multiplier \cite{RHmulti} respectively) against the golden model spec. 
Similarly, the "MasVMon" benchmarks are the miter circuits of
 Mastrovito multipliers compared against  Montgomery
multipliers. The CNF formulae are translated as polynomial constraints
over $\mathbb{F}_2$ (as shown in \cite{condrat-tacas07}), and the
GB-Core algorithm is applied. 
%These benchmarks are in moderate size for our GB engine but not too trivial.

\vspace{-0.1in}
\begin{table}[htb]
\centering
\caption{\centering Results of running benchmarks using our tool\\
\small{Asterisk($^*$) denotes a benchmark which is not converted from CNF benchmark}}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
\multirow{4}{2.2cm}{\centering Benchmark} 
& \multirow{4}{0.9cm}{\centering \# Polys} 
& \multirow{4}{0.8cm}{\centering \# MUS} 
& \multirow{4}{1.9cm}{\centering Size of core w/ iterative refinement heuristic}
 & \multirow{4}{1.6cm}{\centering \# GB-core iterations}
 & \multirow{4}{2cm}{\centering \# Size of core after using syzygy heuristic}
 & \multirow{4}{1.3cm}{\centering Runtime (sec)} \\
  & & & & & & \\
    & & & & & & \\
      & & & & & & \\
\hline
\hline
$5\times 5$ SMPO & 240  & 137  & 137  & 8  &  & 3160\\
$4\times 4$ SMPO$^*$ & 84  & 21  & 21  & 1  &  & 125\\
$3\times 3$ SMPO$^*$ & 45  & 15  & 15  & 1  &  & 6.6\\
$3 \times 3$ SMPO & 17 & 2 & 2 & 1 &  & 0.07  \\
$4 \times 4$ MasVMont$^*$ & 148 & 83 & 83 & 1 &  & 23 \\
$3 \times 3$ MasVMont$^*$ & 84 & 53 & 53 & 1 &  & 4.3 \\
$2 \times 2$ MasVMont & 27 & 23 & 23 & 2 &  & 2.3 \\
$5\times 5$ RH$^*$ & 142  & 34  & 35  & 4  &  & 1001\\
$4\times 4$ RH$^*$ & 104  & 35  & 36  & 3  &  & 98\\
$3\times 3$ RH$^*$ & 50  & 20  & 20  & 1  &  & 2.9\\
aim-100 & 79 & 22 & 22 & 1  &  & 43\\
cocktail & 135 & 4 & 4 & 2 &  & 5 \\
subset-1 & 100 & 6 & 6 & 1 &  & 2.4 \\
subset-2 & 141 & 19 & 23 & 2 & 21 & 12 \\
subset-3 & 118 & 16 & 12 & 2 & 11 & 8.7 \\
phole4 & 104 & 10 & 16 & 1 & 10 & 4.4 \\
phole5 & 169 & 19 & 25 & 3 & 19 & 14 \\
\hline
\end{tabular}
\label{tab:result}  
\end{table}

In Table \ref{tab:result}, we list the details of our experimental
results. In the table, \#Polys denotes the given number of polynomials
from which a core is to be extracted. \#MUS is the minimal core
extracted by PicoMUS. \#GB-core iterations corresponds to the
number of calling GB-core engine to arrive at an unsatisfiable core. The
second last column is showing the improvement of minimal core
size applying syzygy heuristic on cases which are not able to be
refined further with the iterative heuristic.
The data shows that in most of these
cases, our tool can produce a minimal core when our iterations
converge. From the experimental results we can also make the
observation that the GB-core algorithm, particularly Theorem
\ref{thm}, offers quite a lot of scope for identifying redundant
polynomials that can be eliminated from the core --- without resorting
to a brute-force membership check of every polynomial $f_i \in F -
\{f_i\}$.

\section{Conclusions}
This paper addressed the problem of identifying an infeasible core of
a set of multivariate polynomials, with coefficients from a field,
that have no common zeros. The problem is posed in the context of
computational algebraic geometry and solved using the Gr\"obner basis
algorithm. We show that by recording the data produced by the
Buchberger's algorithm -- the $Spoly(f_i,f_j)$ pairs, as well as the
polynomials of $F$ used in the division process
$Spoly(f_i,f_j)\xrightarrow{F}_+ 1$ -- we can identify certain
conditions under which a polynomial can be discarded from a core. An
algorithm was implemented within the Singular computer algebra tool
and some experiments were conducted to validate the approach. While
the use of GB engines for SAT solving has a rich history, the problem
of unsat core identification hasn't been fully addressed by the SAT
community. We hope that this paper will kindle some interest in this
topic which is worthy of attention from the SAT community.

### 2016-05-12 unsat/paper/intro.tex
Finalize UNSAT core staff(hopefully)

\section{Introduction}

The Boolean Satisfiability Problem (SAT) is formulated as finding
solutions satisfying a set of Boolean equations, or to show that no
such solutions exist (unsat). Such problems are often represented in
Conjunctive Normal Form (CNF), whereby sets of literal-disjunctions
(clauses) must be simultaneously satisfied through some variable
assignment. When no solutions exist, it is possible to identify a
subset $C_c$ of the original set of clauses $C$ such that $C_c$ is
unsat too. This set $C_c$ is called the unsat core of the
problem. Unsat cores have various uses in system design and
verification, where they can identify the causes of unsat in the
problem and restore satisfiability. Various unsat core extractors are
available \cite{Haifa2014} \cite{Joao2012} that can identify smaller
cores -- and {\it minimal} unsat cores -- from very large unsat
problems.

The problem has analogous manifestations in the polynomial algebra
domain. Suppose that we are given a set of polynomials $F =
\{f_1,\dots,f_s\}$ with coefficients from any field. Suppose, further,
that the system of equations $f_1 = f_2 = \dots = f_s = 0$ has no
common zero -- i.e. the system of polynomial constraints is infeasible
(or unsat). Can we identify a subset $F_c \subseteq F$ such that the
system of polynomials of $F_c$ also has no common
zeros? In other words, we wish to devise algorithmic techniques to
identify infeasible or unsat cores $F_c$ of $F$.

Clearly, computational algebraic geometry -- particularly the theory
and technology of Gr\"obner bases -- provides a mechanism to answer
the polynomial unsat core question. The given set of polynomials $F$
generates an ideal, and the celebrated Hilbert's Nullstellensatz
\cite{ideals:book} provides all the tools to characterize the zero-set
(varieties) of polynomial ideals. Most Nullstellensatz-related
polynomial decision questions can be answered by the Gr\"obner basis
algorithm \cite{gb_book}. This motivates our investigation into
extraction of infeasible cores of the set of polynomials $F$ using
Gr\"obner basis techniques.

\subsubsection{Previous Work:}

Moreover, as Gr\"obner basis techniques
find more applications in SAT solving \cite{CEI:stoc-96} \cite{condrat-tacas07}
\cite{zengler:GB-SAT}, unsat cores of polynomial systems will also find

in hardware and software verification
\cite{lv:tcad2013}, as well as

### 2016-05-12 unsat/paper/irina.tex
Finalize UNSAT core staff(hopefully)

--- 2016-05-12 unsat/paper/BuchbergerAlg2.tex
@@
-%%This is a very basic article template.
-%%There is just one section and two subsections.
-\documentclass{article}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{amssymb}
+\section{Introduction}
 
-\theoremstyle{plain}
-\newtheorem{defn}{Definition}[section]
-%\usepackage{datetime}
-\newtheorem{thm}[defn]{Theorem} % definition numbers are dependent on
-% theorem
-% %
-\newtheorem{prop}[defn]{Proposition} % definition numbers are dependent on
-\newtheorem{dem}[defn]{Proof} % definition numbers are dependent on
-              % numbers
-\newtheorem{exemp}{Example}[defn]
-\newtheorem{lema}{\bf Lemma}[defn]
-
-\def\opn#1#2{\def#1{\operatorname{#2}}} % to make operators
-
-\opn\Ker{Ker} \opn\Coker{Coker}  \opn\Hom{Hom} \opn\Im{Im}
-\opn\End{End} \opn\Aut{Aut} \opn\defect{def} \opn\ord{ord}
-\opn\id{id} \opn\dim{dim} \opn\det{det} \opn\tr{tr} \opn\grad{grad} \opn\lcm{lcm}
-\opn\min{min} \opn\max{max} \opn\det{det}
-%\def\Frob{{\mathcal F}}
-\opn\Span{Span}   \opn\rang{rang}  \opn\id{id} \opn\dim{dim} \opn\ad{ad} \opn\tr{tr} \opn\ker{ker}
-\opn\GL{GL} \opn\SL{SL} \opn\mod{mod} \opn\diag{diag}
-\opn\min{min} \opn\sgn{sgn} \opn\ini{in_<}  \opn\Mon{Mon} \opn\LC{LC_<}
-\opn\LT{LT_<}
-\opn\s{supp}
-
-\title{On Cores of Ideals with Empty Variety}
-
-\begin{document}
-
-\maketitle
+{\bf This is the dummy submission uploaded on Feb 21. The final paper will be uploaded on Feb 24, 2016.}
 
@@
 \end{exemp}
-\begin{thebibliography}{99}
-
-\bibitem{HH}  J\"urgen Herzog, Viviana Ene, \emph{Gr\"obner Bases in Commutative Algebra}, American Mathematical Society, 2012.
-\end{thebibliography}
-
-\end{document}

### 2016-05-12 unsat/paper/prelim.tex
Finalize UNSAT core staff(hopefully)

\section{Preliminaries}

We denote by $C$ the set of clauses representing the CNF-SAT
problem. The problem is assumed to be unsat, and $C_c$ denotes
the set of clauses that constitute the unsat core.

Let $\F$ be any field, $\Fc$ its algebraic closure, and $R = \R$ 
the polynomial ring in variables $x_1, \dots, x_d$ with coefficients
from $\F$. A monomial in
variables $x_1, \dots, x_d$  is a power product of the form  $X =
x_1^{e_{1}}\cdot x_2^{e_{2}}\cdots x_d^{e_{d}}$, where 
$e_i \in \Z_{\geq 0}, i\in \{1, \dots, d\}$. A {\it polynomial} 
$f \in R$ is written as a finite sum of terms 
$f = c_1 X_1 + c_2 X_2 + \dots + c_t X_t$.  Here $c_1, \dots, c_t$ are
coefficients and $X_1, \dots, X_t$ are monomials. To systematically
manipulate the polynomials, a monomial order $>$ (or a term order) is
imposed on the ring --- i.e. a total order on the monomials
s.t. multiplication with another monomial preserves the ordering.
The monomials of $f = c_1 X_1 + c_2 X_2 + \dots + c_t X_t$  are
ordered w.r.t. to $>$, such that  $X_1 > X_2 > \dots >  X_t$.  Subject
to such a term order, $lt(f) = c_1 X_1, ~lm(f) = X_1, ~lc(f) = c_1$,
are the {\it leading term}, {\it leading monomial} and {\it   leading
  coefficient} of $f$, respectively. 
%We also denote tail($f$) = $f - lt(f) = c_2X_2 + \dots + c_t X_t$. 
While our approach works for any permissible term order, in the paper,
we consider terms ordered {\it degree-lexicographically}, where the
monomials are ordered according to their descending total degree, and
ties are broken lexicographically.

{\bf Polynomial Reduction:} Let $f, g$ be polynomials. If a non-zero
term $cX$ of $f$ is divisible by the leading term of $g$, then we say
that $f$ {\it is reducible to} $r$ modulo $g$, denoted $f
\stackrel{g}{\textstyle\longrightarrow} r$, where $r = f - {cX   \over
  lt(g)} \cdot g$. Similarly, $f$ can be {\it reduced (divided) 
w.r.t. a set of polynomials}  $F = \{f_1, \dots, f_s\}$ to obtain a
remainder $r$. This reduction is denoted $f \stackrel{F} {\textstyle
  \longrightarrow}_+ r$, and the remainder $r$ has the property that
no term in $r$ is divisible by the leading term of any polynomial
$f_i$ in $F$. The division algorithm ({\it e.g.} Alg. ??? \ref{alg:div})
records the {\it data} related to both the remainders and the quotients in
the division process. We will utilize this data to identify the core.

\subsubsection{Ideals, Varieties and Nullstellensatz:} 
Let $F = \{f_1, \dots, f_s\}$ denote the given set of polynomials. An
{\it ideal} $J \subseteq R$ generated by polynomials $f_1, \dots, f_s \in R$
is: \vspace{-0.1in} 
$$J = \langle f_1, \dots, f_s \rangle = \{\sum_{i=1}^{s} h_i\cdot f_i:
~h_i \in \F[x_1,\dots,   x_d]\}.$$  %\vspace{-0.1in}
The polynomials $f_1, \dots, f_s$ form the {\it basis} or the {\it
  generators} of $J$.

We have to consider the set of solutions to the system of polynomials
equations $f_1 = \dots = f_s = 0$. The set of all solutions to a
given system of polynomial equations $f_1 = \dots = f_s = 0$ is called
the {\it variety}, which is uniquely defined by the ideal 
$J = \langle f_1,\dots,f_s\rangle$ generated by
the polynomials. The variety is denoted by $V(J)=V(f_1,\dots,f_s)$ and
defined as: 
$$V(J) = V(f_1, \dots, f_s) = \{\mathbf{a} \in \F^d: \forall
f \in J, f(\mathbf{a}) = 0\},$$ 
where $\mathbf{a} = (a_1, \dots, a_d) \in \F ^d$ denotes a point in
the affine space.

\begin{theorem} \label{thm:weak-ns}
{\it The Weak Nullstellensatz} \cite{ideals:book}: 
%Let $\F$ be a field and $\Fc$ be its algebraic closure. 
Let $J = \langle f_1,\dots, f_s \rangle \subseteq \F[x_1, \dots, x_d]$ be
an ideal and $V_{\Fc}(J)$ be its variety over $\Fc$. Then
$V_{\Fc}(J) = \emptyset \iff J = \F[x_1,\dots,x_d] \iff 1 \in J$. 
\end{theorem}

The Weak Nullstellensatz provides a mechanism to ascertain whether or not
a given system of polynomials has a feasible solution. This is deduced
by testing whether the unit element is a member of the ideal $J$. This
{\it ideal membership test} requires the computation of a Gr\"obner
basis.

\subsubsection{Gr\"obner bases:} An ideal $J$ may have many bases
(generators). A {\it Gr\"obner basis} (abbreviated as GB) is a generating set for $J$
with special properties that allow to solve many polynomial decision
problems, including ideal membership testing. From among the many
{\it equivalent} definitions of a Gr\"obner basis, we will consider the
following two:

\begin{definition}
\label{def:gb}
$\bf{\left[Gr\ddot{o}bner\ Basis\right]}$ \cite{gb_book}:\\
(i) For a monomial ordering $>$, a set  of non-zero polynomials $G =
\{g_1,g_2,\cdots,g_t\}$ contained in an ideal $J$, is called a
Gr\"{o}bner basis for $J$ iff $\forall f \in J$, $f\neq 0$ there
exists $i \in \{1,\cdots, t\}$ such that $lm(g_i)$ divides $lm(f)$;
i.e., $G = GB(J) \Leftrightarrow\  \forall f \in J : f \neq 0, \exists
g_i \in G : lm(g_i)\mid lm(f)$. Or equivalently, \\
(ii) $G = \{g_1,g_2,\cdots,g_t\}$ is a Gr\"{o}bner basis for $J$ iff
division by $G$ of any polynomial in $J$ gives the remainder 0;
i.e. $G = GB(J) \iff \forall f \in J, f\xrightarrow{g_1,\dots,g_t}_+ 0$.
\end{definition}

Buchberger's algorithm 
\cite{buchberger_thesis},
shown in Algorithm \ref{alg:gb},  computes a Gr\"obner
basis over a field. Given polynomials $F = \{f_1, \dots, f_s\}$, 
the algorithm computes the Gr\"obner basis $G = \{g_1, \dots, 
g_t\}$. 
% such that ideal $I = \langle f_1, \dots, f_s\rangle = \langle g_1,
% \dots, g_t\rangle$. 
The algorithm takes pairs of polynomials $(f, g)$, and computes their
{\it S-polynomial} ($Spoly(f, g)$):

\begin{equation}
       Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g \nonumber
      \label{eqn:spoly}
\end{equation}
where $L = \text{LCM}(lm(f), lm(g))$. 
%, where $lm(f)$ is the leading monomial of $f$, and $lt(f)$ is the
%leading term of $f$. 
$Spoly(f,g)$ cancels the leading terms of $f$ and $g$. Therefore, the 
computation $Spoly(f,g) \xrightarrow{G'}_+r$ results in a remainder $r$,
which if non-zero, provides an element with new leading term in the
generating set.  The Gr\"obner basis algorithm terminates when for all pairs
$(f, g)$, $Spoly(f,g)\xrightarrow{G'}_+0$.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{$F = \{f_1, \dots, f_s\}$}
 \KwOut{$G = \{g_1,\dots ,g_t\}$\\} %, a Gr\"{o}bner basis
  $G:= F$\;
  \Repeat{$G = G'$}
  {
        $G' := G$\;
        \For{ each pair $\{f, g\}, f \neq g$ in $G'$} 
        {
                $Spoly(f, g) \stackrel{G'}{\textstyle\longrightarrow}_+r$ \;
                \If{$r \neq 0$}
                {
                        $G:= G \cup \{r\}$ \;
                }
        }
   }
\caption {Buchberger's Algorithm}\label{alg:gb}
\end{algorithm}

A Gr\"obner basis $G$ may contain redundant elements. To remove
these redundant elements, $G$ is first made {\it minimal} and
subsequently {\it reduced.} To obtain a minimal GB, all polynomials
$g_j$ are removed from $G$ if there exists a $g_i$ such that 
$lm(g_i) ~|~ lm(g_j)$. Then the remaining elements ($g_i$'s) are made
monic by dividing each $g_i$ by $lc(g_i)$. This minimal basis is
further reduced by ensuring that no term in $g_j$ is divisible by the
leading term $lt(g_i)$ for all $i \neq j$. Subject to %the given term order 
$>$, the reduced Gr\"obner   basis $G_r = \{g_1, \dots, g_t\}$ is a
unique, canonical representation of ideal $J$.

In the context of Nullstellensatz, $V_{\Fc}(J) = \emptyset \iff G_r =
\{1\}$. This implies that for ideals with empty variety, there exists
a sequence of $Spoly(f, g)$ reductions in the reduced GB computation
that leads to the unit element. We now show that by analyzing this
data, the unsat core $F_c$ of the given generating set $F =
\{f_1,\dots,f_s\}$ can be identified.

### 2016-05-12 unsat/paper/section4_xiaojun.tex
Finalize UNSAT core staff(hopefully)

--- 2016-05-12 unsat/paper/core.tex
@@
-\section{Finding the Unsat Core}
-With the aforementioned concepts and notations, we can formally state
-the unsat core problem over polynomial rings as follows:
-
-\begin{Problem}
-Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
-the ring $R = \F[x_1,\dots,x_n]$ that generate ideal $J = \langle
-f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
-\emptyset$, or it is determined to be so by applying the Gr\"obner
-basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
-J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
-too. Borrowing the terminology from the Boolean SAT domain, we
-call $F_c$ the infeasible core or the unsat core of $F$. 
-\end{Problem}
-
-Any set of polynomials $F$ with empty variety is an unsat core in
-itself. There may be more than one unsat cores in $F$, some
-polynomials may be common to multiple cores (i.e. the
-cores may be non-disjoint), and these cores may have
-different cardinalities $|F_c|$. The unsat core problems
-are therefore further classified as: 
-\bi
-\item Identify a {\it minimum} core, i.e. a core of minimum
-  cardinality. 
-\item Identify a {\it minimal} core. An unsat core $F_{c}$ is minimal
-  with respect to the property that while the variety of $F_{c}$ is
-  empty, the variety of {\it any} subset of $F_c$ is {\it non-empty},
-  i.e. $V( F_{c} - \{f_i\} ) \neq \emptyset$ for any $f_i \in
-  F_c$. 
-\item Identify any unsat core disregarding minimality; i.e. find any
-  subset $F_c$ of $F$.
-\ei
-
-Let us first consider the problem of finding any $F_c \subset F$,
-disregarding minimality. It is not hard to motivate that an unsat
-core should be identifiable using the Gr\"obner basis
-algorithm: Assume that $F_c = F - \{f_j\}$. If $GB(F) = GB(F_c) =
-\{1\}$, then it implies that $f_j \in \langle F - \{f_j\}\rangle$. In
-other words, 
-$f_j$ can be composed of the other polynomials of $F_c$, so $f_j$ is
-not a part of the unsat core, and it can be safely discarded from
-$F_c$. This can be identified by means of the GB algorithm for this
-ideal membership test (due to Def. \ref{def:gb}(ii)).
-
-A n\"aive way (and inefficient way) to identify {\it a minimal core}
-using the GB computation is as follows:
-selectr a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
-(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
-from the core; otherwise retain $f_i$ in $F_c$. Select a different
-$f_i$ and continue until all polynomials $f_i$ are visited for
-inclusion in $F_c$. This approach will produce a minimal core, as we
-would have tested each polynomial $f_i$ for inclusion in the
-core. This requires $O(|F|)$ calls to the GB engine, which is really
-impractical.
-
-\subsection{Investigating algorithmic techniques to find $F_c$ from $F$}
-
-We investigate if it is possible to identify a core by analyzing the
-$Spoly(f_i,f_j)\xrightarrow{F}_+ r$ reductions in Buchberger's
-algorithm. Since $F$ is an unsat core, 
-definitely {\it there exists  
-a sequence of Spoly reductions in Buchberger's algorithm where
-$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
-polynomial reduction algorithms can be suitably modified to record
-which polynomials from $F$ are used in the division leading to
-$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that by recording
-the {\it ``data''} generated by Buchberger's algorithm --- i.e. the
-critical pair polynomials ($f_i,f_j$) used in the Spoly computations,
-and the polynomials from $F$ used to cancel terms in the reduction
-$Spoly(f_i,f_j)\xrightarrow{F}_+1$ --- we should be able to identify a
-core.
-
-We will use an example to demonstrate our approach of identifying $F_c
-\subseteq F$ using this ``data'':
-
-\begin{example}
-\label{ex:1}
-Consider the following set of polynomials:
-\begin{center}
-$f_1: abc + ab + ac + bc + a + b + c + 1$
-\end{center}
-\begin{center}
-$f_2: b$
-\end{center}
-\begin{center}
-$f_3: ac$
-\end{center}
-\begin{center}
-$f_4: ac + a$
-\end{center}
-\begin{center}
-$f_5: bc + c$
-\end{center}
-\begin{center}
-$f_6: abd + ad + bd + d$
-\end{center}
-%\newpage
-\begin{center}
-$f_7: cd$
-\end{center}
-\begin{center}
-$f_8: abd + ab + ad + bd + a + b + d + 1$
-\end{center}
-\begin{center}
-$f_9: abd + ab +bd + b$
-\end{center}
-
-Assume $>_{DEGLEX}$ monomial ordering with $a>b>c>d$. 
-Let $F = \{f_1,\dots,f_9\}$ and 
-$J = \langle F \rangle \subset \mathbb{F}_2[a,b,c,d]$ where
-$\mathbb{F}_2 = \{0, 1\}$ is the finite field of 2 elements. Then 
-$V(J) = \emptyset$ as $GB(J) = 1$.  The set $F$ consists of 4 {\it minimal}
-cores: $F_{c1} = \{ f_1,f_2,f_3,f_4,f_7,f_8\}, F_{c2} = \{
-f_2,f_4,f_5,f_6,f_8\}, F_{c3} = \{ f_2,f_3,f_4,f_6,f_8\},$ and $F_{c4}
-= \{ f_1,f_2,f_4,f_5\}$. 
-\end{example}
-
-Buchberger's algorithm terminates to a unique reduced GB, irrespective
-of the order in which the critical pairs are selected and reduced as
-$Spoly(f_i,f_j)\xrightarrow{F}_+r$. Let us suppose that in the GB
-computation corresponding to Example \ref{ex:1}, the first 3 critical
-pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and $(f_2,f_5)$. It turns
-out that the Spoly-reductions corresponding to these 3 pairs lead to
-the unit ideal. Recording the ``data'' corresponding to this sequence
-of reductions is depicted in Fig. \ref{fig:refute}. We call this
-graph a refutation tree.
-
-\begin{figure}[hbt]
-\centering
-\includegraphics[scale=0.25]{./SAT2016_xiaojun/refutation_tree.eps}
-\caption{Generating refutation trees to record unsat cores.}
-\label{fig:refute}
-\end{figure}
-
-In the figure, the nodes of the graph correspond to the polynomials
-utilized in Buchberger's algorithm. The leaf nodes always correspond
-to polynomials $f_i \in F$. An edge $e_{ij}$ from node $i$ to node $j$
-signifies that the polynomial at node $j$ resulted from polynomial
-$i$. For example, consider the computation
-$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
-1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
-corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
-$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
-following sequence of 1-step divisions:
-$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} \xrightarrow{f_3}
-\xrightarrow{c\cdot f_2}  \xrightarrow{f_2} f_{10}$. This is depicted
-as the bottom subtree in the figure, terminating at polynomial
-$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
-division by $f_2$ resulted in the quotient $a$. The refutation tree of
-Fig. \ref{fig:refute} shows further that
-$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
-$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.
-
-To identify an $F_c \subset F$, we start from the refutation node
-$''1''$, and traverse the graph in reverse, all the way upto the
-leaves. Then, all the leaves in the transitive fanin of ``1''
-constitute an unsat core. The polynomials (nodes) that do not lie in
-the transitive fanin of ``1'' can be safely discarded from $F_c$. From
-Fig. \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
-as an unsat core of $F$.
-
-\subsubsection{Ordering critical pairs in Buchberger's algorithm:} It
-is clear that the inclusion of polynomials in the obtained core
-depends on the order in which the critical pairs are
-reduced. This is a well-known intractable problem, and various
-heurstics have been proposed to favourably influence the termination
-of the Gr\"obner basis algorithm -- these include the sugar
-strategy \cite{GBSelectionStrategies}, the preconditioning used by
-{\it Faug\~{e}re} in the $F4$ algorithm \cite{f4}, among others. Any
-such selection strategy can be also be applied in our setting to
-identify a smaller core.
-
 \section{Reducing the Size of the Infeasible Core $F_c$}
@@
  \end{algorithm}
-
-\section{Iterative Refinement of the Unsat Core}
-Similar to all other core extractors, our algorithm cannot generate a
-minimal core in one execution. To obtain a smaller core, we need to
-re-execute our tool with the core obtained in the current
-iteration. We describe two heuristics that are applied to our
-algorithm to increase the probability to generate a smaller core in
-the next iteration.
-
-After eliminating all redundant polynomials, we can call our GB engine
-with the new core. A effective heuristic should enhance the
-probability that the refutation "1" is composed by fewer polynomials. 
-In our GB computing algorithm, we use a strategy to pick critical
-pairs such that polynomials with  larger indexes get involved as late
-as possible: 
-$$(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$$
-
-Meanwhile, for the reduction process $Spoly(f_i,f_j)\xrightarrow{F}_+
-f_{new}$, we pick divisor polynomials from $F$ following the order of
-polynomial indexes. Therefore, by renaming the polynomial indexes, we
-can increase the likelihood they are selected into the unsat core. 
-We use 2 criteria to increase the probability that a polynomial is
-included in the refutation tree.  One corresponds to the \emph{refutation distance},
-whereas the other corresponds to the frequency that
-a polynomial appears in refutation tree.
-
-\begin{definition}
-\emph{Refutation distance} in a refutation tree corresponds to the number of arcs on
-the shortest path from refutation "1" to a leaf
-polynomial.
-\end{definition}
-
-Polynomials with shorter refutation distance are used as
-divisors in later stages of polynomial reductions, which indicates
-they will generally have lower degree leading terms.  
-This is because we are assuming a degree-lexicographic order, and
-successive term cancellations reduce the degree of the remainders. So
-polynomials with lower degree leading terms are more  likely to be
-used for reduction, such that the probability that they appear in
-the refutation tree is larger.
-
-Similarly, the motivation for \emph{frequency of appearance} of $f_i$ in
-the refutation tree is as follows: polynomials appearing frequently in
-the refutation tree also indicates that they have some properties
-(leading terms) that make them "favourable" in the unsat core
-selection. For example, their leading terms may contain variables
-that are require them to be included in the minimal core.
-
-We apply both heuristics: we sort the polynomials in the core
-by the refutation distance criterion, and use the frequency criterion as
-the tiebreaker. The following example illustrates our heuristic.
-
-%\vspace{5mm}\\
-\begin{example} 
-Consider a set of 6 polynomials over $\F_2$ of an
-infeasible instance.
-\begin{align*}
-f_1 &: x_1x_3+x_3\\
-f_2 &: x_2 + 1\\
-f_3 &: x_2x_3+x_2\\
-f_4 &: x_2x_3\\
-f_5 &: x_2x_3 + x_2 + x_3 + 1\\
-f_6 &: x_1x_2x_3 +x_1x_3
-\end{align*}
-
-After the first iteration, we get a tree of refutation as shown in Fig.\ref{fig:refine}(a). 
-\begin{figure}[hbt]
-\centering
-\includegraphics[scale=0.25]{SAT2016_xiaojun/core_refine.eps}
-\caption{Refutation trees of core refinement example}
-\label{fig:refine}
-\end{figure}
-
-The shortest "Distance" corresponds to that of $f_2$ to "1", which is
-2 levels. So, we reorder $f_2$ to be the polynomial with the shortest
-index. The "distance" and "frequency" for other polynomials are
-identical, so we keep their ordering untouched. We re-index the
-polynomial set 
-$$f_1'=f_2, f_2' = f_1, f_3' = f_3, f_4' = f_4$$
-and apply our GB-core algorithm on $\{f_1',f_2',f_3',f_4'\}$. The
-result is shown in Fig.\ref{fig:refine}(b). We reach a fixpoint, and
-the core $\{f_1', f_3', f_4'\} = \{f_2,f_3,f_4\}$ is proved to be
-minimal. 
-\end{example}

### 2016-05-12 unsat/paper/section5_xiaojun.tex
Finalize UNSAT core staff(hopefully)

--- 2016-05-12 unsat/paper/core.tex
@@
-\section{Finding the Unsat Core}
-With the aforementioned concepts and notations, we can formally state
-the unsat core problem over polynomial rings as follows:
-
-\begin{Problem}
-Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
-the ring $R = \F[x_1,\dots,x_n]$ that generate ideal $J = \langle
-f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
-\emptyset$, or it is determined to be so by applying the Gr\"obner
-basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
-J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
-too. Borrowing the terminology from the Boolean SAT domain, we
-call $F_c$ the infeasible core or the unsat core of $F$. 
-\end{Problem}
-
-Any set of polynomials $F$ with empty variety is an unsat core in
-itself. There may be more than one unsat cores in $F$, some
-polynomials may be common to multiple cores (i.e. the
-cores may be non-disjoint), and these cores may have
-different cardinalities $|F_c|$. The unsat core problems
-are therefore further classified as: 
-\bi
-\item Identify a {\it minimum} core, i.e. a core of minimum
-  cardinality. 
-\item Identify a {\it minimal} core. An unsat core $F_{c}$ is minimal
-  with respect to the property that while the variety of $F_{c}$ is
-  empty, the variety of {\it any} subset of $F_c$ is {\it non-empty},
-  i.e. $V( F_{c} - \{f_i\} ) \neq \emptyset$ for any $f_i \in
-  F_c$. 
-\item Identify any unsat core disregarding minimality; i.e. find any
-  subset $F_c$ of $F$.
-\ei
-
-Let us first consider the problem of finding any $F_c \subset F$,
-disregarding minimality. It is not hard to motivate that an unsat
-core should be identifiable using the Gr\"obner basis
-algorithm: Assume that $F_c = F - \{f_j\}$. If $GB(F) = GB(F_c) =
-\{1\}$, then it implies that $f_j \in \langle F - \{f_j\}\rangle$. In
-other words, 
-$f_j$ can be composed of the other polynomials of $F_c$, so $f_j$ is
-not a part of the unsat core, and it can be safely discarded from
-$F_c$. This can be identified by means of the GB algorithm for this
-ideal membership test (due to Def. \ref{def:gb}(ii)).
-
-A n\"aive way (and inefficient way) to identify {\it a minimal core}
-using the GB computation is as follows:
-selectr a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
-(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
-from the core; otherwise retain $f_i$ in $F_c$. Select a different
-$f_i$ and continue until all polynomials $f_i$ are visited for
-inclusion in $F_c$. This approach will produce a minimal core, as we
-would have tested each polynomial $f_i$ for inclusion in the
-core. This requires $O(|F|)$ calls to the GB engine, which is really
-impractical.
-
-\subsection{Investigating algorithmic techniques to find $F_c$ from $F$}
-
-We investigate if it is possible to identify a core by analyzing the
-$Spoly(f_i,f_j)\xrightarrow{F}_+ r$ reductions in Buchberger's
-algorithm. Since $F$ is an unsat core, 
-definitely {\it there exists  
-a sequence of Spoly reductions in Buchberger's algorithm where
-$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
-polynomial reduction algorithms can be suitably modified to record
-which polynomials from $F$ are used in the division leading to
-$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that by recording
-the {\it ``data''} generated by Buchberger's algorithm --- i.e. the
-critical pair polynomials ($f_i,f_j$) used in the Spoly computations,
-and the polynomials from $F$ used to cancel terms in the reduction
-$Spoly(f_i,f_j)\xrightarrow{F}_+1$ --- we should be able to identify a
-core.
-
-We will use an example to demonstrate our approach of identifying $F_c
-\subseteq F$ using this ``data'':
-
-\begin{example}
-\label{ex:1}
-Consider the following set of polynomials:
-\begin{center}
-$f_1: abc + ab + ac + bc + a + b + c + 1$
-\end{center}
-\begin{center}
-$f_2: b$
-\end{center}
-\begin{center}
-$f_3: ac$
-\end{center}
-\begin{center}
-$f_4: ac + a$
-\end{center}
-\begin{center}
-$f_5: bc + c$
-\end{center}
-\begin{center}
-$f_6: abd + ad + bd + d$
-\end{center}
-%\newpage
-\begin{center}
-$f_7: cd$
-\end{center}
-\begin{center}
-$f_8: abd + ab + ad + bd + a + b + d + 1$
-\end{center}
-\begin{center}
-$f_9: abd + ab +bd + b$
-\end{center}
-
-Assume $>_{DEGLEX}$ monomial ordering with $a>b>c>d$. 
-Let $F = \{f_1,\dots,f_9\}$ and 
-$J = \langle F \rangle \subset \mathbb{F}_2[a,b,c,d]$ where
-$\mathbb{F}_2 = \{0, 1\}$ is the finite field of 2 elements. Then 
-$V(J) = \emptyset$ as $GB(J) = 1$.  The set $F$ consists of 4 {\it minimal}
-cores: $F_{c1} = \{ f_1,f_2,f_3,f_4,f_7,f_8\}, F_{c2} = \{
-f_2,f_4,f_5,f_6,f_8\}, F_{c3} = \{ f_2,f_3,f_4,f_6,f_8\},$ and $F_{c4}
-= \{ f_1,f_2,f_4,f_5\}$. 
-\end{example}
-
-Buchberger's algorithm terminates to a unique reduced GB, irrespective
-of the order in which the critical pairs are selected and reduced as
-$Spoly(f_i,f_j)\xrightarrow{F}_+r$. Let us suppose that in the GB
-computation corresponding to Example \ref{ex:1}, the first 3 critical
-pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and $(f_2,f_5)$. It turns
-out that the Spoly-reductions corresponding to these 3 pairs lead to
-the unit ideal. Recording the ``data'' corresponding to this sequence
-of reductions is depicted in Fig. \ref{fig:refute}. We call this
-graph a refutation tree.
-
-\begin{figure}[hbt]
-\centering
-\includegraphics[scale=0.25]{./SAT2016_xiaojun/refutation_tree.eps}
-\caption{Generating refutation trees to record unsat cores.}
-\label{fig:refute}
-\end{figure}
-
-In the figure, the nodes of the graph correspond to the polynomials
-utilized in Buchberger's algorithm. The leaf nodes always correspond
-to polynomials $f_i \in F$. An edge $e_{ij}$ from node $i$ to node $j$
-signifies that the polynomial at node $j$ resulted from polynomial
-$i$. For example, consider the computation
-$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
-1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
-corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
-$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
-following sequence of 1-step divisions:
-$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} \xrightarrow{f_3}
-\xrightarrow{c\cdot f_2}  \xrightarrow{f_2} f_{10}$. This is depicted
-as the bottom subtree in the figure, terminating at polynomial
-$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
-division by $f_2$ resulted in the quotient $a$. The refutation tree of
-Fig. \ref{fig:refute} shows further that
-$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
-$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.
-
-To identify an $F_c \subset F$, we start from the refutation node
-$''1''$, and traverse the graph in reverse, all the way upto the
-leaves. Then, all the leaves in the transitive fanin of ``1''
-constitute an unsat core. The polynomials (nodes) that do not lie in
-the transitive fanin of ``1'' can be safely discarded from $F_c$. From
-Fig. \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
-as an unsat core of $F$.
-
-\subsubsection{Ordering critical pairs in Buchberger's algorithm:} It
-is clear that the inclusion of polynomials in the obtained core
-depends on the order in which the critical pairs are
-reduced. This is a well-known intractable problem, and various
-heurstics have been proposed to favourably influence the termination
-of the Gr\"obner basis algorithm -- these include the sugar
-strategy \cite{GBSelectionStrategies}, the preconditioning used by
-{\it Faug\~{e}re} in the $F4$ algorithm \cite{f4}, among others. Any
-such selection strategy can be also be applied in our setting to
-identify a smaller core.
-
-\section{Reducing the Size of the Infeasible Core $F_c$}
-
-The core $F_c$ obtained from the aforementioned procedure is not
-guaranteed to be minimal. For example, let us revisit the the core
-$F_c=\{f_1,\dots, f_5\}$ generated in the previous section. Note that
-while $F_c$ is a smaller infeasible core of $F$, it is not minimal. In
-fact, Example 1 shows that $F_{c4} =\{f_1,f_2,f_4,f_5\}$ is the
-minimal core where $F_{c4} \subset F_{c}$. Clearly, the polynomial
-$f_3 \in F_c$ is a redundant element of the core. It could be
-discarded. We will now describe how the size of this core could be
-reduced further.  For this purpose, we will have to perform a more
-systematic book-keeping of the data generated during the execution of
-Buchberger's algorithm and the refutation tree.
-
-\subsection{Identifying redundant polynomials from the refutation tree}
-
-We will keep track of the S-polynomials that give a non-zero remainder
-when divided by the system of polynomials $F$ at that moment:
-\begin{equation}
-\label{eqn1}
-g_{ij}= S(f_i,f_j)+\displaystyle\sum_{k=1}^m c_kf_k,
-\end{equation}
-where $0\neq c_k\in\mathbb{F}[x_1,\ldots,x_n]$ and
-$\{f_1,\ldots,f_l\}$ is the ``current'' system of polynomials
-$F$. For each non-zero $g_{ij}$, we will record the following data:
-
-\begin{equation}
-\label{data1}
-((g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
-\end{equation}
-
-Note that in (\ref{eqn1}) and (\ref{data1}), $g_{ij}$ denotes the
-remainder of the $S$-polynomial $S(f_i,f_j)$ modulo the current system
-of polynomials $f_1,\ldots,f_m$, and we denote by
-
-$$h_{ij}:=\displaystyle\frac{\lcm(lm(f_i),lm(f_j))}{lt(f_i)},
-h_{ji}:=\displaystyle\frac{\lcm(lm(f_i),lm(f_j))}{lt(f_j)}$$ 
-the coefficients of $f_i$, respectively $f_j$, in the $S$-polynomial
-$S(f_i,f_j)$. Furthermore, in (\ref{data1}), $(c_{k1}, \dots c_{kl})$ are the
-polynomial coefficients of $(f_{k1}, \dots,f_{kl})$ that appear in the
-division process.
-
-\begin{example}
-Revisiting Example \ref{ex:1},  the data corresponding to
-$Spoly(f_1,f_2)\xrightarrow{F}_+  g_{12} = f_{10}$ reduction is
-obtained as the following sequence of computations:
-$$f_{10}=g_{12}=f_1-acf_2-af_2-f_3-cf_2-f_2.$$ As the coefficient
-field is $\mathbb{F}_2$ in this example, $-1 = +1$, so:
-$$f_{10}=g_{12}=f_1+acf_2+af_2+f_3+cf_2+f_2$$ is obtained.
-The data is recorded according to (\ref{data1}):
-
-\begin{center}
-$((f_{10}=g_{12}), (f_1,1)(f_2,ac)|(a,f_2),(1,f_3),(c,f_2),(1,f_2))$
-\end{center}
-
-\end{example}
-
-Of course, our approach and the book-keeping stops when we obtain
-1 as the remainder of the S-polynomial modulo the current system of
-generators. As an output of the Buchberger's algorithm, we can obtain
-not only the Gr\"obner bases,  $G = \{g_1,\ldots,g_t\}$, but also a
-matrix $M$ such that: 
-\begin{center}
-\begin{align}
-   \begin{bmatrix}
-           g_{1} \\
-           g_{2} \\
-           \vdots \\
-           g_{t}
-         \end{bmatrix}
-    &= M \begin{bmatrix}
-           f_{1} \\
-           f_{2} \\
-           \vdots \\
-           f_{s}
-         \end{bmatrix}
-  \end{align}
-
-\end{center}
-
-Each element $g_i$ of $G$ is a polynomial combination of $f_1, \dots,
-f_s\}$. Moreover, this matrix $M$ is constructed precisely using the
-data that is recorded in the form of (\ref{data1}). We now a condition
-when the matrix $M$ may identify some redundant elements.
-
-\begin{theorem}
-\label{thm}
-With the notations above, we have that a core for the system of
-generators $F = \{f_1,\dots,f_s\}$ of the ideal $J$ is given by the
-union of those $f_i$'s from $F$ that appear in the data recorded above
-and correspond to the nonzero entries in the matrix $M$.  
-\end{theorem}
-
-\begin{proof}
-In our case, since the variety is empty, and hence the ideal is the
-unit ideal, we have that $G = \{g_1=1\}$ and $t=1$. Therefore $M=
-[a_1, \ldots, a_s]$ is a vector. Then the output of the algorithm
-gives:
-
-$$1 = a_1f_1+\cdots + a_s f_s.$$ Clearly, if $a_i=0$ for some $i$ then
-$f_i$ does not appear in this equation and should not be included in
-the infeasible core of $F$.
-
-\end{proof}
-
-\begin{example}
-
-Corresponding to Example \ref{ex:1} and the refutation tree shown in
-Fig. \ref{fig:refute}, we discover that the polynomial $f_3$ is used
-only twice in the division process. In both occasions, the quotient of
-the division is 1. Therefore, from Fig. \ref{fig:refute}, it follows that:
-\begin{equation}
-1 = (f_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot f_3}+ \dots (f_1
-+ f_2)
-\end{equation}
-
-Since $1 + 1 = 0$ over $\F_2$, we have that the entry in $M$
-corresponding to $f_3$ is 0, and so $f_3$ can be discarded from the
-core. 
-\end{example}
-
-\subsection{The GB-Core Algorithm Outline}
-
-The following steps describe an algorithm (GB-Core) that allows us to compute a
-refutation tree of the polynomial set and corresponding matrix $M$.
-
-{\bf Inputs:} Given a system of polynomials $F=\{f_1,\ldots,f_s\}$
-, a monomial order $>$ on $\mathbb{F}[x_1,\ldots,x_n]$.
-
-{\bf S-polynomial reduction:} We start computing the S-polynomials of the system of
-generators $\{f_1,\ldots,f_s\}$, then divide each of them by the system of
-generators $G=\{f_1,\dots,f_s,\dots,f_m\}$ which is the intermediate result of Buchberger's algorithm. 
-In this way, we obtain the following expression:
-\begin{equation}
-\label{eqn:red}
-g_{ij}= h_{ij}f_{i}+h_{ji}f_{j}+\displaystyle\sum_{k=1}^m c_kf_k
-\end{equation}
-If the remainder $g_{ij}$ is non-zero, we will denote it by
-$f_{m+1}$ and we will add it to intermediate system of generators $G$. We
-will also record the data as (\ref{data1}): 
-\begin{displaymath}
-((f_{m+1}=g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
-\end{displaymath}
-It forms a tree with root $f_{m+1}$.
-
-{\bf Coefficients recording:} In (\ref{eqn:red}) we get a coefficient vector including coefficients
-$c_k$ where $k>s$. These coefficients are associated with new generators in the Gr\"obner basis,
-which cannot benefit the UNSAT core extraction. Therefore we need to rewrite the vector to
-one with only $c_k, k=1\dots s$, which can be achieved by substituting $f_k,k>s$ by $f_1,\dots,f_s$:
-initially we have $f_{s+1}= c_1f_1 + c_2f_2 + \cdots + (h_{ij}+c_i)f_{i}+\cdots+(h_{ji}+c_j)f_{j}+\cdots+c_sf_s$,
-inductively if we can present $f_{s+1}$ to $f_{s+k-1}$ by $f_1,\dots,f_s$, we can also rewrite
-$f_{s+k} = d_1f_1+\cdots+d_sf_s$. Then we add a new row $ (d_1,\dots,d_s ) $ to coefficient matrix $M$.
-
-{\bf Termination and refutation tree construction:} By repeating S-polynomial reduction and coefficients 
-recording until we get remainder $f_m = 1$, the algorithm
-will generate a set of new generators $G$ and coefficient matrix $M$.
-Meanwhile we can construct the refutation tree with the recorded data.
-We start with 
-\begin{displaymath}
-((f_{m}=1)(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
-\end{displaymath}
-It represents a tree start from leaves $f_{i}, f_j$ then internal nodes $f_{k1},\dots,f_{kl}$
-and ``1" as the root. Next we substitute nodes $f_{m-1}, f_{m-2}, \dots, f_{s+1}$ in order
-with corresponding subtrees. The algorithm returns the complete refutation tree as well as
-matrix $M$.
-
-The complete algorithm is concluded as pseudo code below:
-
-\begin{algorithm}[H]
- \caption{GB-core algorithm (based on Buchberger's algorithm)}
- \label{algo:gbcore}
- \begin{algorithmic}[1]
-
- \REQUIRE $F = \{f_1, \dots, f_s\} \in \F[x_1, \dots, x_n], f_i\neq 0$
- \ENSURE Refutation tree $T$ and coefficients matrix $M$
- \STATE{ {Initialize: list $G \gets F$; Dataset $D\gets \emptyset$; $M\gets s\times s$ unit matrix} }
- \FOR {{ each pair $(f_i,f_j)\in G$  }}
- 	\STATE  $f_{sp},(f_{i},h_{ij})(f_{j},h_{ji}) \gets$ spoly-pair($f_i,f_j$) \COMMENT{$f_{sp}$ is the S-polynomial}
- 	\STATE{{ $g_{ij}|(c_{k1},f_{k1}),\dots,(c_{kl},f_{kl}) \gets (sp\xrightarrow{G}_+  g_{ij})$}}
- 	\IF{{ $g_{ij} \neq 0$ }}
- 		\STATE {{$G \gets G \cup g_{ij}$}}
- 		\STATE {{Update $D$ and $M$}}
- 	\ENDIF
- 	\IF {{$g_{ij} = 1$}}
- 		\STATE {{Construct $T$ from $D$}}
- 		\STATE {{ Return($T,M$) }}
- 	\ENDIF
- \ENDFOR
- \end{algorithmic}
- \end{algorithm}
-
 \section{Iterative Refinement of the Unsat Core}

### 2016-05-12 unsat/paper/syzygy.tex
Finalize UNSAT core staff(hopefully)

\section{Reduce the size of UNSAT core further}
In some
cases our iterative core refining algorithm cannot
give us the minimal core when it hits the fixpoint.
It indicates the limitation of using re-labelling 
strategy, therefore a new method to identify the 
redundancy in UNSAT core is needed.

The UNSAT core we obtained through our algorithm
is by nature a refutation polynomial that equals to 1:
$$1 = \sum_{i=1}^s c_i\cdot f_i$$
where $c_i \neq 0$ and $f_1,\dots,f_s$ form a core.
If by any means we find a relation between $f_k$ and set $\{f_j\}$
 ($f_k,f_j\in \{f_1,\dots,f_s\}$) and other polynomials
 in the core such that:
 $$f_k = \sum_{j\neq k} c_j'f_j$$
 Then we can substitute $f_k$ with other polynomials
 in the refutation polynomial. Thus, $f_k$ can be
 recognized as redundant polynomial in the core.

 Since coefficients $\{c_i\}$ are also polynomials in $\mathbb F[x_1,\dots,x_n]$, we can find 
 enormous number of combinations of $\{f_i\}$. Consider finding an algorithmic solution to 
 this problem without introducing extra search,
 we propose to utilize the \emph{syzygies} generated during
 executing the GB-core algorithm.

 In Buchberger's algorithm, if the polynomial division
 gives a nonzero remainder, then the information is 
 recorded. However, if the polynomial division results
 in a zero remainder nothing will be recorded. In order
 to gather more information from one iteration, we
 need to also record all quotients from s-polynomial
 reductions with 0 remainder. Those quotients (as well 
 as coefficients when pairing s-polynomials) form
 a syzygy:
\[
 \begin{cases}
 c_1^1f_1+c_2^1f_2+\cdots+c_s^1f_s = 0\\
 c_1^2f_1+c_2^2f_2+\cdots+c_s^2f_s  = 0\\
 \ \ \ \ \ \  \vdots \\
 c_1^mf_1+c_2^mf_2+\cdots+c_s^mf_s = 0  
 \end{cases}
\]
The matrix form of the syzygy is:
\begin{center}
\begin{align}
\label{mat:syzygy}
   \begin{bmatrix}
           c_1^1 & c_2^1 & \cdots & c_s^1 \\
           c_1^2 & c_2^2 & \cdots & c_s^2 \\
           \vdots & \vdots & \ddots & \vdots \\
           c_1^m & c_2^m & \cdots & c_s^m
         \end{bmatrix}
    \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
         &= 0
  \end{align}

\end{center}
 Here $\{f_1,f_2,\dots,f_s\}$ is the given core.
 Take one column of the syzygy matrix (e.g. $c_1^1, c_1^2, \dots, c_1^m$)
 and compute its minimal Gr\"obner basis. If the result is 1, then there exists some polynomial vector
 $r_1,r_2,\dots,r_m$ such that
 $$1 = r_1c_1^1 + r_2c_1^2 + \cdots + r_mc_1^m = \sum_{i=1}^m r_ic_1^i$$
 If we multiply the $i$-th row in above matrix with $r_i$ then add them up, 
 we get following equation
 \begin{center}
\begin{align}
   \begin{bmatrix}
           1 & \sum_{i=1}^m r_ic_2^i & \cdots & \sum_{i=1}^m r_ic_s^i
         \end{bmatrix}
    \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
         &= 0
  \end{align}

\end{center}
Which implies $f_1$ is a combination of $f_2,\dots,f_s$:
 $$f_1 = \sum_{i=1}^m r_ic_2^if_2 + \sum_{i=1}^m r_ic_3^if_3
 +\cdots + \sum_{i=1}^m r_ic_s^if_s$$
 Then we can remove $f_1$ from the core. By repeating
 this procedure, some redundant polynomials can be identified
 and size of UNSAT core will get closer to minimal.
 \begin{example}
 Assume we are given 9 polynomials from Example\ref{ex:1},
 execute GB-core algorithm and record syzygy. We
 write the coefficients as entries in following matrix:
 \[
 \begin{blockarray}{cccccccccccc}
  && f_1 & f_2 & f_3 & f_4 & f_5 & f_6 & f_7 & f_8 & f_9 & f_{10} \\
  \begin{block}{cc(cccccccccc)}
  Spoly(f_1,f_3)\ & & 1 & a+c+1 & b+1 &0&0&0&0&0&0&1\\
  Spoly(f_2,f_3)\ & & 0 & ac & b &0&0&0&0&0&0&0\\
  Spoly(f_1,f_4)\ & & 1 & c+1 & 1 &b&0&0&0&0&0&1\\
  Spoly(f_2,f_4)\ & & 0 & ac+a & 0 &b&0&0&0&0&0&0\\
  Spoly(f_1,f_5)\ & & 1 & a+c+1 & 0 &0&a&0&0&0&0&1\\
  \end{block}
  \end{blockarray}
 \]
 Usually we will get extra columns comparing to syzygy matrix in (\ref{mat:syzygy}).
 In this example we add an extra column for coefficient of $f_{10}$
 since some s-polynomial pairs require new generator $f_{10}$
 as divisor during reduction. In order to remove this
 extra column, we need to turn nonzero entries in this column
 to 0 through matrix manipulations. Recall that
 we record $f_{10}$ as nonzero remainder when reducing s-polynomial
 pair $Spoly(f_1,f_2)$, we extract corresponding submatrix from the coefficient matrix $M$
 $$(1 ac+a+c+1 1 0 0 0 0 0 0 )$$
 It represents $f_{10}$ is a combination of $f_1$ to $f_9$:
 $$f_{10} = f_1 + (ac+a+c+1)f_2 + f_3$$
 It can be written in matrix with column $f_{10}$ as follows:
  \[
 \begin{blockarray}{cccccccccccc}
  && f_1 & f_2 & f_3 & f_4 & f_5 & f_6 & f_7 & f_8 & f_9 & f_{10} \\
  \begin{block}{cc(cccccccccc)}
  Spoly(f_1,f_2)& & 1 & ac+a+c+1 & 1 &0&0&0&0&0&0&1\\
  \end{block}
  \end{blockarray}
 \]
 By adding this row to corresponding rows in the matrix,
 we can get syzygy matrix only for polynomials
 in the core:
  \[
 \begin{blockarray}{ccccccccccc}
  && f_1 & f_2 & f_3 & f_4 & f_5 & f_6 & f_7 & f_8 & f_9  \\
  \begin{block}{cc(ccccccccc)}
  Spoly(f_1,f_3)\ & & 0 & ac & b &0&0&0&0&0&0\\
  Spoly(f_2,f_3)\  && 0 & ac & b &0&0&0&0&0&0\\
  Spoly(f_1,f_4)\  && 0 & ac+a & 0 &b&0&0&0&0&0\\
  Spoly(f_2,f_4)\  && 0 & ac+a & 0 &b&0&0&0&0&0\\
  Spoly(f_1,f_5)\  && 0 & ac & 1 &0&a&0&0&0&0\\
  \end{block}
  \end{blockarray}
 \]
 We find out there is a ``1" entry in column $f_3$, so 
 its minimal GB equals to 1. Therefore $f_3$ can be
 recognized as redundant polynomial and removed from the UNSAT core.

 \end{example}

### 2016-05-12 unsat/FINAL/SMPO_poly.pl

@@ -1,0 +1,148 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+open MTR, ">", "SMPOmiter3.sing"; # change filename
+
+my $size = 3; # change this size
+
+my $d = 1;
+my $c = 1;
+my $b = 1;
+my $i = 0;
+#my @idx = qw /2 2 3 0 1 1 3/;
+
+my @array = qw /0 1 2/;
+
+
+#my @array = qw /0 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 21 22 23 24 25 26 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49 50 51 52 53 54 55 56 57 58 59 60 61 62 63 64/;
+
+#my @idx = qw /1 0 3 3 4 1 2 2 4/;
+my @idx = qw/1 0 2 1 2/;
+my $j = 1;
+my @names;
+print MTR "ideal I = \n";
+for($i=0;$i<$size;$i++){
+push @names, "a".$i;
+push @names, "b".$i;
+push @names, "Rs".$i;
+push @names, "Ru".$i;
+}
+
+my $k;
+for($k = 0; $k < $size; $k++){
+printf MTR "b%d*a%d+d0\_r%d,\n", $k, ($idx[0]+$k)%$size, $k;
+push @names, "d0\_r".$k;
+$j = 1; $b = 1; $c = 1; $d = 1;
+while ($j < $size) {
+printf MTR ("a%d+a%d+c%d\_r%d,\n", ($idx[$j*2-1]+$k)%$size, ($idx[$j*2]+$k)%$size, $c, $k); 
+push @names, "c".$c."\_r".$k;$c++;
+printf MTR ("b%d*c%d\_r%d+d%d\_r%d,\n", ($b+$k)%$size, $c-1, $k, $d, $k); 
+push @names, "d".$d."\_r".$k; $b++; $d++;
+$j ++;
+}
+$i = 0;
+printf MTR "d0\_r%d+d1\_r%d+tmp0\_r%d,\n", $k, $k, $k;
+push @names, "tmp0\_r".$k;
+for($i = 2; $i < $size - 1; $i++){
+printf MTR ("tmp%d\_r%d+d%d\_r%d+tmp%d\_r%d,\n", $i-2,$k, $i,$k, $i-1,$k);
+push @names, "tmp".($i-1)."\_r".$k;
+}
+
[…]

### 2016-05-12 unsat/FINAL/cnf2poly.pl

@@ -1,0 +1,94 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+# Minor Update: 02/22/2016 by Xiaojun
+# Note: Now support dummy 9-SAT (You should define or6+() proc in singular script separately)
+# Note2: Add an item output, printing init intvec "rd" for consistency
+# Note3: trim spaces at head
+
+# Update: 02/18/2016 by Xiaojun Sun
+# Note: Expand capability to 5-SAT
+# Note2: cancel alphabet, use v1 v2 v3... instead
+# Note3: add a line as list of vars for Singular use
+
+# Note: this is a simple version for 3-SAT CNF parser
+# Edit: 03/31/2015 by Xiaojun Sun
+# Objective: do scratch investigation on GB based UNSAT core extraction
+
+open CNF, "<", $ARGV[0];   # first argument as input
+open POLY, ">", $ARGV[1];  # second argument as output
+my $cnt = 0;
+
+my $i;
+my $tmp;
+my $clauses;
+my $vars;
+
+#my @alphabet = qw /a b c d e f g h/;
+my @args;
+
+while(<CNF>){
+	chomp;
+	my $read_in = $_;
+	$read_in =~ s/^\s+//; # remove spaces at head
+	my @items = split /\s+/, $read_in;
+	$cnt++;
+	if(!defined($items[0])) {next;}  # skip empty lines
+	if($items[0] =~ /c/) {next;}  # skip comments
+	if($items[0] =~ /p/){  # write vanishing polys based on inputs
+	  $vars = $items[2];
+	  $clauses = $items[3];
+	  print POLY "ring Q = 2, (";
+	  foreach $i (1..($vars-1)) {
+		print POLY "v$i,";
+	  }
+	  print POLY "v$vars), Dp;\n\nideal I = ";
+	  next;
+	}
+	if(defined($items[9]) && $items[9] != 0) {  # for max-9-SAT problems
+	  print "Error in Line$cnt!\n";
+	  next;
+	}
+	undef @args;
+	for($i = 0; $i < 9; $i++)
+	{
+	  if($items[$i] < 0)
+	  {
+		$tmp = 0 - $items[$i];
+#		print "What is tmp?? It is $tmp!!!!!!!";
+		push @args, "1+v".$tmp;
+	  }
+	  elsif($items[$i] > 0)
+	  {
[…]

### 2016-05-12 unsat/FINAL/miterRH_poly.pl

@@ -1,0 +1,265 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+# Update: 02/23/2016 by Xiaojun Sun
+# Note: Revised according to major update of autogen_RHmulti.pl
+# NOW support Type-1 ONB!!!!
+
+my $size = 3; # change this size
+open POLY, ">", "RH$size.tmpsing";
+my @names;
+
+my $v = int $size/2;
+my $d = 1;
+my $c = 1;
+my $b = 1;
+my $i = 0;
+#my @idx = qw /2 2 3 0 1 1 3/; # this is M^0, by default first row only have 1 entry
+#my @idx = qw/1 0 3 3 4 1 2 2 4/;
+my @idx = qw/1 0 2 1 2/;
+my @array;
+foreach(1..$size)
+{
+  push @array, $_-1;
+}
+
+my $j = 1;
+
+my $tmp;
+my $k;
+my $strid = 'head 0'; # e0 is always needed because of special beta monomial
+my @elist; #new
+my $cnt; #new
+
+my (@Mat, @tmprow);
+foreach(1..$size)
+{ push @tmprow, 0;}
+
+for($j = 0; $j < $size; $j++){ # init multitable
+	push @Mat, [@tmprow];
+}
+# $Mat[0][0] = 1; # obsolete code now, because we take care of d0 separately
+# Following part we calculate multi table. If we use type-2 ONB, it is same with M^0
+# However for generalization with type-1 ONB, we calculate using general rule for all rows
+$i = 0; $j = $idx[0];
+$Mat[($j-$i)%$size][(0-$i)%$size] = 1;
+for($i = 1; $i < $size; $i++)
+{
+	$j = $idx[2*$i-1]; $Mat[($j-$i)%$size][(0-$i)%$size] = 1;
+	$j = $idx[2*$i]; $Mat[($j-$i)%$size][(0-$i)%$size] = 1;
+}
+# Check those one who need "e" layer
+push @elist, 0; # e0 is always needed...
+for($i = 1; $i <= $v; $i++)
+{
+	for($j = 0; $j < $size; $j++)
+	{
+		if($Mat[$i][$j] == 1) {
+			if($strid !~ /\b$j\b/)
+			{
[…]

### 2016-05-12 unsat/FINAL/parser_nh_dist.pl

@@ -1,0 +1,175 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+use List::MoreUtils qw(uniq);
+
+# Modification from version A: Xiaojun Sun, 04/05/2016
+# Remove coef output file, restore nh output file
+
+# Update: Xiaojun Sun, 02/19/2016
+# Change 1: Input format change, now first line add #var as 3rd arg, then each line index and coeff are alternating
+# e.g. line "10 2 1 5 ab 4 c" means f_10 = 1*f_2 + ab*f_5 + c*f_4
+# Change 2: Besides newick tree output, added another output file listing coefficients for each poly
+# involved with s-poly/reduction (not simplified), it is meant to be fed to Singular to check if any of
+# them reduced to 0. Excluding those, remaining nonzero coeffs represents the core. Note for fitting
+# the size of total poly, coeff[last] must be defined (if not used, then define as 0)
+# Change 3: add another output file for dist&freq
+
+# Note: read_in file format:
+# first line: start with "s", then a single number means # of original polys, then # of total GBs include "1" 
+# followed by n lines means n new polys added to GB
+# those lines: start with index of newly added poly (from s+1 to s+n)
+# then a pair of numbers means spoly pair
+# if more following, they mean polys used for reduction, exactly divided in the same order
+# last line start with "e" means the spoly reduced to "1"
+# start with "c" for comments
+# Edit: 04/20/2015 by Xiaojun Sun
+# Objective: translate the output from singular file (computing GB for UNSAT) showing all original polys involved for
+# spoly and reduction to newick tree format displaying the hierarchy
+
+open TREE, ">", $ARGV[0].".nh";  # auto-assign name
+#open COEF, ">", "tmp.coef";
+open DIST, ">", "tmp.dist";
+my $cnt = 0;
+
+my $i;
+my $tmp;
+my %lines;
+my $origin;
+my $n;
+my $nvars;
[…]

### 2016-05-12 unsat/FINAL/slim_cnf.pl

@@ -1,0 +1,33 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+# Update: 02/20/2016 by Xiaojun Sun
+# Add second arg for output filename
+
+# slim_cnf: abstract certain clauses from a CNF file
+# Edit: 03/31/2015 by Xiaojun Sun
+# Objective: do scratch investigation on GB based UNSAT core extraction
+
+open RIN, "<", $ARGV[0];
+open CNF, ">", $ARGV[1];
+my $cnt = 0;
+my @idx = qw/116 3 1 5 35 12 23 13 41 84 58/;
+
+while(<RIN>){
+	chomp;
+	my $read_in = $_;
+	my @items = split /\s+/;
+	if(!defined($items[0])) {next;}  # skip empty lines
+	if($items[0] =~ /c/) {next;}  # skip comments
+	if($items[0] =~ /p/){  # write vanishing polys based on inputs
+	  print CNF "p cnf $items[2] ";
+	  print CNF $#idx+1;
+	  next;
+	}
+	$cnt++;
+	if($cnt ~~ @idx)
+	{print CNF "\n$read_in";}
+}

### 2016-05-22 unsat/response_CP2016.txt
response to reviewers for CP2016

Review 1:
-- Originality issue?
Our heuristics are original, and proved to be effective by our experiments.

Review 2:
-- Any kind of theoretical guarantees?
No, because the procedure of computing a Groebner basis does not guarantee that we can deduce all polynomial dependences in the original set.
--  In your experiments the GB-core algorithm runs only once and the syzygy heuristic does not provide further improvements?
See above.
-- It would be interesting to see how using the syzygy heuristic compares to re-running GB-core?
Syzygy heuristic requires GB computation, which is inefficient. For example, single run of GB computation for 200 polynomials could be hours. So we use syzygy heuristic as complementation to re-running GB-core.

Review 3:
-- The paper addresses the problem of finding small unsatisfiable cores in the context of Boolean formulas presented in XNF? 
Not exactly. Both Groebner basis and unsatisfiable cores are concepts not limited to XNF,  so our approach can be applied to any decision problem formalized by polynomials over any field.
-- Benchmarks must have been translated in the opposite direction too, for use by picoMUS?
The size of MUS for benchmarks not translated from CNF is not obtained by using picoMUS. Instead, we first get a smaller core using our approach, then reduce it to minimal using exhaustive deletion-based techniques.
-- The result for GB-core without the post-processing?
Okay.
-- The result for subset-3 must be wrong, as it suggests the proposed method finds a core that is smaller than the optimal result?
No, the smaller core obtained by our approach is not a subset of the core from picoMUS.
-- Consider publishing all benchmarks used, including those that have been modified by adding redundant clauses. I assume StarExec is the right venue for this?
Okay. But not StarExec.
-- Page 14, mid: "translation to CNF" -> "translation from CNF"?
Correct.
-- (Important) Page 14 (mid) suggests that any scalability issues stem from translation to/from CNF, but in [12] those issues must have been overcome?
(Need further discussion) Reference [12] does not overcome the scalability issue for GB computation. In that paper, the author divided the whole set of clauses into lots of small subsets, and GB computation is operated on small sets. In our case, partitioning is not helpful for identifying smaller cores because UNSAT problem is related to a global property rather than local ones.

Review 4:
-- Is the refutation tree really helpful in the construction? Is it used for storing the computed polynomials in a more efficient way?
The refutation tree is a structure helpful in illustrating the resolution-style refutation when running the Buchberger's algorithm. We do not use a tree-like data structure to actually storing the data. We have introduced the data structure in section 4.2,  especially in Algorithm 2.
-- What is the criterion used to stop the refinement procedure (besides considering the size of the unsat core computed by another tool)?
We stop the refinement when the size of core do not decrease after more GB-core iterations are taken. It is a fixpoint detection mechanism that not rely on the MUS we learn in advance.
-- How do the execution times compare with those of picoMUS?
As the GB computation is hard and our tool is not well specialized,  the runtime is much larger than conventional tools -- their runtimes are usually less than 1 ms.
-- Is picoMUS the only solver to compare with? Why?
No,  we also compare the results from MUSer2 and HaifaMUC. They are identical for our benchmarks. We choose these 3 solvers because they have best performance on SAT 2011 competition,  which is the latest SAT competition with a MUS track.
(URL for HaifaMUC: http://ie.technion.ac.il/~ofers/haifasolvers/)

### 2016-05-24 unsat/response_CP2016.txt
update response to CP2016 reviewers

@@
 -- Any kind of theoretical guarantees?
-No, because the procedure of computing a Groebner basis does not guarantee that we can deduce all polynomial dependences in the original set.
+No.
+This is a hard theoretical problem in algebra, because a redundant polynomial can always be written as combination of necessary polynomials, while the procedure of computing a Groebner basis does not guarantee that we can deduce all polynomial dependences in the original set.
+The GB computation procedure can be controlled by both monomial order and strategy to pick critical s-poly pairs, which is infeasible for exhaustive method.
 --  In your experiments the GB-core algorithm runs only once and the syzygy heuristic does not provide further improvements?
-See above.
+Although syzygy heuristic gather extra information from the GB computation, it is still not sufficient to derive all polynomial dependencies.
 -- It would be interesting to see how using the syzygy heuristic compares to re-running GB-core?
-Syzygy heuristic requires GB computation, which is inefficient.
-For example, single run of GB computation for 200 polynomials could be hours.
+Syzygy heuristic collects extra information from GB-core algorithm, so the number of rows of syzygy matrix will be much greater than the size of original polynomial set.
+Since it requires a full GB computation on each column of syzygy matrix, it will take too much time.
+For example, if we directly apply syzygy heuristic on aim-100 benchmark (79 polynomials), there are 426 rows in syzygy matrix, it is too large such that GB computation for a single column takes hours.
 So we use syzygy heuristic as complementation to re-running GB-core.
@@
 -- The result for GB-core without the post-processing?
-Okay.
+We have the data and will be happy to share with the PC.
 -- The result for subset-3 must be wrong, as it suggests the proposed method finds a core that is smaller than the optimal result?
-No, the smaller core obtained by our approach is not a subset of the core from picoMUS.
+No, the smaller core obtained by our approach is not a subset of the core from picoMUS. Actually we obtained a different minimal core using another MUS extractor (HaifaMUC).
 -- Consider publishing all benchmarks used, including those that have been modified by adding redundant clauses. I assume StarExec is the right venue for this?
-Okay. But not StarExec.
+We would love to share our benchmarks and tool with the PC and the public.
 -- Page 14, mid: "translation to CNF" -> "translation from CNF"?
-Correct.
+We accept this suggestion.
 -- (Important) Page 14 (mid) suggests that any scalability issues stem from translation to/from CNF, but in [12] those issues must have been overcome?
-(Need further discussion) Reference [12] does not overcome the scalability issue for GB computation.
+Reference [12] does not overcome the scalability issue for GB computation.
 In that paper, the author divided the whole set of clauses into lots of small subsets, and GB computation is operated on small sets.
@@
 -- Is the refutation tree really helpful in the construction? Is it used for storing the computed polynomials in a more efficient way?
-The refutation tree is a structure helpful in illustrating the resolution-style refutation when running the Buchberger's algorithm.
+The refutation tree is a structure helpful in illustrating the resolution-style refutation when running the Buchberger's algorithm, as well as introducing concepts such as refutation distance and frequency of occurence.
 We do not use a tree-like data structure to actually storing the data.
-We have introduced the data structure in section 4.2,  especially in Algorithm 2.
+In Algorithm 2, data set D is necessary, while the construction of refutation tree T is optional and it has similar cost with extracting distance and frequency from D.
 -- What is the criterion used to stop the refinement procedure (besides considering the size of the unsat core computed by another tool)?
@@
 -- How do the execution times compare with those of picoMUS?
-As the GB computation is hard and our tool is not well specialized,  the runtime is much larger than conventional tools -- their runtimes are usually less than 1 ms.
+As the GB computation is hard and our tool is implemented through Singular which is a universal symbolic computing system that not well specialized,  the runtime is much larger than conventional tools -- their runtimes are usually less than 0.1 sec.
 -- Is picoMUS the only solver to compare with? Why?
-No,  we also compare the results from MUSer2 and HaifaMUC.
-They are identical for our benchmarks.
-We choose these 3 solvers because they have best performance on SAT 2011 competition,  which is the latest SAT competition with a MUS track.
-(URL for HaifaMUC: http://ie.technion.ac.il/~ofers/haifasolvers/)
+No,  we also compare the results from MUSer2 and HaifaMUC and they show identical results for most of our benchmarks.
+For several benchmarks (subset1~3, phole4) they give different solutions, the results from picoMUS are usually the same or a subset of the output from our tool (for subset3 our result is better than all other ones).
+We choose these 3 solvers because they have best performance on SAT 2011 competition (ranking 1,2,4,5 on number of solved instances and 1,2,3,5 on average CPU time for solving single instance),  which is the latest SAT competition with a MUS track.
+(word count: 865)

### 2016-09-21 dissertation/abstract.tex
init dissertation writing

%\begin{abstract}
\begin{center}{\bf ABSTRACT}\end{center}

Abstraction plays an important role in digital design,
analysis, and verification, as it allows for the refinement of
functions through different levels of conceptualization. This dissertation
introduces a new method to compute a symbolic, canonical, word-level abstraction of the function implemented
by a combinational logic circuit. This abstraction provides a
representation of the function as a {\it polynomial} $Z = F (A)$
over the Galois field $\Fkk$, expressed over the $k$-bit input to the circuit, $A$.
This representation is easily utilized for formal verification (equivalence checking) of
combinational circuits.

The approach to abstraction is based upon concepts from
commutative algebra and algebraic geometry, notably the
\Grobner basis theory. It is shown that the polynomial $F(A)$ can
be derived {\it by computing a \Grobner basis of the polynomials}
corresponding to the circuit, using a specific elimination term
order based on the circuit’s topology. However, computing \Grobner bases
using elimination term orders is infeasible for large circuits.
To overcome these limitations, this work introduces {\it an efficient symbolic
computation} to derive the word-level polynomial. The presented algorithms
exploit i) the structure of the circuit, ii) the properties of \Grobner bases,
iii) characteristics of Galois fields $\Fkk$, and iv) modern algorithms from symbolic computation.

While the concept is
applicable to any arbitrary combinational logic circuit, it 
is particularly powerful in verification and equivalence checking of 
hierarchical, custom-designed and structurally dissimilar Galois field arithmetic circuits. 
%These designs are prevalent in cryptography, error correction codes, and signal processing.
In most applications, the field size and the data-path size $k$
in the circuits is very large, up to $1024$ bits.
%The U.S. National
%Institute for Standards and Technology (\emph{NIST}), recommends Galois
%fields corresponding to operand-sizes $k = 163, 233, 283, 409$ and
%$571$ bits wide.

The proposed abstraction procedure can exploit the hierarchy of the
given Galois field arithmetic circuits.
% by computing the abstraction
%of each block in parallel before combining the representations together. 
A custom abstraction %software
tool is designed to efficiently implement the abstraction procedure.
Preliminary experiments show that, using the proposed approach, our
tool can abstract and verify Galois field arithmetic circuits
{\it up to $1024$ bits in size}. %, which is the largest NIST standard. 
Contemporary
techniques fail to verify these types of circuits beyond $163$ bits
and cannot abstract a canonical representation beyond $32$ bits.

%To efficiently implement the
%procedure, we design a custom abstraction software tool. 
%Modern abstraction techniques of
%these circuits fail beyond $32$-bits.

%Custom arithmetic circuits designed over Galois fields $\Fkk$ are
%prevalent in cryptography, error correction codes, signal processing,
%etc. In most applications, the field size and the data-path size $k$
%in the circuits can be very large. For example, the U.S. National
%Institute for Standards and Technology (NIST), recommends Galois
%fields corresponding to operand-sizes $k = 163, 233, 283, 409$ and
%$571$ bits wide. The complexity of such large architectures 
%necessitates custom and hierarchical design. Custom design can raise
%the potential for bugs in the implementation. As arithmetic bugs can
%compromise the security of crypto-systems, formal verification of
%Galois field circuits is an imperative.

%This thesis presents an approach for formal verification and
%equivalence checking of large Galois field arithmetic circuits based
%on word-level canonical polynomial abstraction from gate-level
%circuits. The word-level abstraction is based on concepts from
%commutative algebra and algebraic geometry, and relies on the
%following mathematical insights.

%A combinational circuit with $k$-inputs and $k$-outputs implements
%Boolean functions $f: \B^k \rightarrow \B^k$, where $\B = \{0, 1\}$.
%Such functions can also be construed as a mapping $f: \Fkk
%\rightarrow \Fkk$,  where $\Fkk$ denotes the Galois field of $2^k$
%elements. Every function over $\Fkk$ is a polynomial function ---
%i.e. there exists a unique, minimal, canonical polynomial
%$\Func$ that describes $f$.  This thesis derives the polynomial
%representation of the circuit as $Z = \Func(A)$ over $\Fkk$, where $A$
%and $Z$ denote, respectively, the input and output bit-vectors of the
%circuit. This can be achieved by computing a Gr\"obner basis of a set
%of polynomials derived from the circuit, using a specific elimination
%(abstraction) term order.

%Computing Gr\"obner bases using elimination orders is, however,
%practically infeasible for large circuits. To overcome this
%limitation, we devise an approach to guide the Gr\"obner basis
%computation to search for the abstraction polynomial. The approach is
%implemented in a custom tool using modern $F_4$-style polynomial
%reduction procedures. Using our approach, we are able to verify the
%correctness of up to $571$-bit Galois field multiplier circuits,
%whereas contemporary techniques prove to be infeasible beyond
%$163$-bit circuits. 
%\end{abstract}

### 2016-09-21 dissertation/abstraction.tex
init dissertation writing

\chapter{Word-Level Abstraction of Combinational Circuits} \label{ch:abstract}

In this chapter, we introduce our approach to abstract word-level
canonical representations of combinational circuits using methods based on 
computer-algebra and algebraic-geometry.  
%For simplicity, the focus is on arithmetic circuits over Galois fields 
%of the type $\Fkk$, but the methods presented are directly applicable
%using methods based on 
%computer-algebra and algebraic-geometry. 
Given a bit-level implementation of a combinational, acyclic circuit $C$ that
implements some unknown function $f:\Fkk^n\rightarrow\Fkk$, where
$Z$ is the $k$-bit output and $A_1,\dots,A_n$ are the $k$-bit inputs,
find the canonical word-level representation $Z=\Func(A_1,\dots,A_n)$ 
implemented by $C$; that is, find a canonical representation of the
polynomial $\Func$ in terms of $A_1,\dots,A_n$. For example, a combinational
circuit with one word-level $k$-bit input $A$ and $k$-bit output $Z$, 
which computes $Z=\Func(A)$ over $\Fkk$,
is shown in Fig. \ref{fig:abstractA_Z2}.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/interpolate}
}
\caption{Derive the abstraction $Z=\Func(A)$}
\label{fig:abstractA_Z2}
\end{figure}
}

By modelling the the arithmetic circuit as a polynomial system in
$\Fkk[x_1,x_2,\cdots,x_d]$, the abstraction problem 
can be solved using a \Grobner basis computation using an
abstraction term order. 
However, as computing a \Grobner basis is computationally
expensive, this approach is directly  applicable only to Galois field arithmetic
circuits no larger than $40$-bits.

\section{Problem Statement}

\begin{itemize}
\item  Given a gate-level combinational, acyclic circuit $C$ with $n$ word-level
$k$-bit inputs,
$A_1,\dots, A_n \in \Fkk$
and one $k$-bit output $Z$.
\item Pick a primitive, irreducible polynomial $P(x)$ over $\F_2[x]$ of degree $k$ 
to construct $\Fkk$ (these polynomials are known).
Let $P(\alpha) = 0$, where $\alpha \in \Fkk$ is a root of the irreducible
polynomial $P$.
\item The bit-level primary inputs of the circuit are denoted
$\{a_{0}^{i},a_{1}^{i},\dots,a_{k-1}^{i}\}$, for $1\leq i \leq n$;
the bit-level primary outputs are $\{z_0, \dots, z_{k-1}\} = Z$. 
Note that all $a_{j}^{i}, z_j \in \F_2$ for $0\leq j < k$.
\item Find the word-level polynomial function $\Func$ over $\Fkk$ 
computed by $C$ in the form of $Z=\Func(A_1, \dots, A_n)$.
\end{itemize}

In order for a combinational circuit $C$ to compute a function $f$ over the 
Galois field $\Fkk$, $C$ must have any number of 
$k$-bit word-level inputs and one $k$-bit word-level output.
Since every function over $\Fkk$ is a polynomial function, 
$C$ has a word-level polynomial representation over $\Fkk$.
The goal is to derive this word-level polynomial representation $\Func$ 
computed by a given combinational circuit over a 
given $\Fkk$.

%The proposed approach is generic enough to abstract the implementation of any
%combinational Galois field arithmetic circuit.
For the purpose of explaining the proposed abstraction approach, 
this chapter explores its application to Galois field multiplier circuits, 
which are described in Chapter \ref{ch:prelim}, 
as they form the core of most cryptographic computations in ECC and are 
notoriously hard to verify. In this case, $P(x)$ is chosen to be the same
primitive polynomial over which the circuit was designed.

\begin{Example}
Consider our problem statement as it applies to a multiplier 
circuit over $\Fkk$. The specification of the circuit is unknown.
\begin{itemize}

\item Given the Galois field $\Fkk$ and the corresponding
  irreducible polynomial $P(x)$. Let $P(\alpha) = 0$.

\item  Given a gate-level combinational circuit. The bit-level primary
  inputs of the circuit are $\{a_0, \dots, a_{k-1}, ~b_0, 
  \dots, b_{k-1}\}$, and the bit-level primary outputs are 
$\{z_0, \dots, z_{k-1}\}$; thus all $a_i, b_i, z_i \in \F_2$.

\item $A$ and $B$ denote the $k$-bit word-level inputs and $Z$ is the
$k$-bit word-level output. Therefore, 
$A = a_0 + a_1\alpha + \dots + a_{k-1}\alpha^{k-1}$,
$B = b_0 + b_1\alpha + \dots + b_{k-1}\alpha^{k-1}$, and 
$Z = z_0 + z_1\alpha + \dots + z_{k-1}\alpha^{k-1}$ with $A,B,Z \in \Fkk$.

\item Find the word-level polynomial function $\Func$ that this circuit 
implements over $\Fkk$.
The polynomial must be in the form of $Z=\Func(A,B)$. Since, in this case, 
the circuit is a multiplier, this resulting polynomial will be $Z=A\cdot B$.

\end{itemize}
\end{Example}

\section{Circuit Polynomial Modeling}

Given a gate-level implementation of a circuit,
we map each gate-level
Boolean operator in the circuit ($NOT$, $AND$, $OR$, $XOR$) to a polynomial 
over $\F_2$ using the following
one-to-one mapping over $\mathbb{B} \rightarrow \mathbb{F}_2$ : 
\begin{equation}
\label{b2poly}
\begin{split}
NOT :&~\neg a \rightarrow a+1 \pmod 2  \\   
AND :&~a \wedge b \rightarrow a\cdot b \pmod 2  \\ 
OR :&~a \vee b \rightarrow a+b+a\cdot b \pmod 2  \\
XOR :&~a \oplus b \rightarrow a+b \pmod 2 
\end{split}
\end{equation}
where $a,b \in \mathbb{F}_{2}=\{0,1\}$.
Note that the equation $c=\mathcal{F}(a,b)$ is written in polynomial
form as $c-\mathcal{F}(a,b)=c+\mathcal{F}(a,b)=0$, ~as $-1 \equiv +1
\pmod 2$.

\begin{Example}
Consider the equation with Boolean operators: 
\begin{equation}
z=a \oplus (b \vee c). \nonumber
\end{equation}

This equation modeled over $\F_2$ is:
\begin{equation}
z+a + b + c + b\cdot c=0 \nonumber
\end{equation}

Notice that the left-hand side expression is a polynomial in
$\F_{2}\left[a,b,c,z\right] \subset
\Fkk\left[a,b,c,z\right]$
\end{Example}

Secondly, we model the $k$-bit word-level inputs
and the $k$-bit word-level output of the given circuit as polynomial
expressions in $\mathbb{F}_{2^{k}}$ as shown in the problem statement.
If the $k$-bit word-level output of the circuit is denoted $Z$ which is 
composed of bit-level outputs 
$z_0, \dots , z_{k-1}$, the corresponding equation
is:
\begin{equation}
 Z = z_0 + z_1 \alpha + \dots + z_{k-1} \alpha^{k-1} \nonumber
\end{equation}
Once again, since $-Z = +Z \pmod 2$, this equation is modelled as:
\begin{equation}
Z + z_0 + z_1 \alpha + \dots + z_{k-1} \alpha^{k-1} = 0 \nonumber
\end{equation}

Likewise, for all word-level inputs $A_1,\dots,A_n$ we have
\begin{eqnarray}
A_1 + a_0^1 \alpha + \dots + a_{k-1}^1 = 0 \nonumber \\
\vdots \nonumber \\
A_n + a_0^n \alpha + \dots + a_{k-1}^n = 0 \nonumber 
\end{eqnarray}

Overall, a combinational circuit composed of $s$ Boolean gates
with $n$ $k$-bit inputs, $A_1, \dots , A_n$, and one $k$-bit output $Z$, 
is modeled as a polynomial system over $\mathbb{F}_{2^{k}}$ as follows:

\begin{eqnarray}
 \left .  \begin{aligned}
f_1(x_1,x_2,\cdots, x_d)=0  \\
f_2(x_1,x_2,\cdots, x_d)=0  \\
\vdots  \\
f_s(x_1,x_2,\cdots, x_d)=0 \\
\end{aligned} 
\ \right\}
 &\qquad&  \text{\it Bit-level circuit constraints} \nonumber \\
 \left . \begin{aligned}
f_{Z}: Z+z_{0}+z_{1}\cdot \alpha,\cdots,{z_{k-1}}\cdot \alpha^{k-1}=0    \\
f_{A_1}:A_1+a_{0}^1+a_{1}^1\cdot \alpha+\cdots+a_{k-1}^1\cdot \alpha^{k-1}=0   \\
\vdots  \\
f_{A_n}:A_n+a_{0}^n+a_{1}^n\cdot \alpha+\cdots+a_{k-1}^n\cdot \alpha^{k-1}=0   \\
\end{aligned} 
\right\}
 &\qquad&  \text{\it Word-level designation} \nonumber \\
\end{eqnarray}

\begin{Example}
\label{exp:mul2bit}
Consider a 2-bit multiplier over ${\mathbb{F}}_{2^2}$ with
$P(x)=x^{2}+x+1$, given in Fig. \ref{fig:2bitmul}. Variables $a_0,
a_1, b_0, b_1$ are primary inputs, $z_0, z_1$ are primary outputs, and
$c_0, c_1, c_2, c_3, r_0$ are intermediate variables. 
%The gate $\otimes$
%corresponds to AND-gate, i.e. bit-level multiplication modulo 2. The
%gate $\oplus$ corresponds to XOR-gate, i.e. addition modulo 2.

\begin{figure}[H]
\centerline{
\includegraphics[scale=0.4]{./figures/2bitmasmult.eps}
}
\caption{ A 2-bit multiplier over ${\mathbb{F}}_(2^2)$.}
\label{fig:2bitmul}
\end{figure}

The circuit can be described using the following Boolean equations:
\begin{align*}
c_0=a_0 \wedge b_0,   \nonumber \\
c_1=a_0 \wedge b_1,   \nonumber \\
c_2=a_1 \wedge b_0, \nonumber \\
c_3=a_1 \wedge b_1,   \nonumber \\
r_0=c_1 \oplus c_2 ,      \nonumber \\
z_0=c_0 \oplus c_3,   \nonumber \\
z_1=r_0 \oplus c_3,    \nonumber
\end{align*}

With the mapping rules given in Equation \ref{b2poly}, 
the above equations are transformed into the following polynomials:
\begin{align*}
c_0+a_0 \cdot b_0,   \nonumber \\
c_1+a_0 \cdot b_1,   \nonumber \\
c_2+a_1 \cdot b_0,   \nonumber \\
c_3+a_1 \cdot b_1,   \nonumber \\
r_0+c_1 + c_2 ,      \nonumber \\
z_0+c_0 + c_3,   \nonumber \\
z_1+r_0 + c_3,   \nonumber
\end{align*}

Therefore, our overall polynomial system is:
\begin{eqnarray}
 \left .  \begin{aligned}
f_1: c_0+a_0 \cdot b_0  \\
f_2: c_1+a_0 \cdot b_1  \\
f_3: c_2+a_1 \cdot b_0  \\
f_4: c_3+a_1 \cdot b_1  \\
f_5: r_0+c_1 + c_2		\\
f_6: z_0+c_0 + c_3		\\
f_7: z_1+r_0 + c_3		   
 \end{aligned} 
\ \right\}
 &\qquad&  {\text {\it Bit-level circuit constraints}} \nonumber \\
 \left . \begin{aligned}
f_{A}: A+a_0+a_1\cdot \alpha  \\ 
f_{B}: B+b_0+b_1\cdot  \alpha  \\
f_{Z}: Z+z_0+z_1\cdot \alpha
 \end{aligned} 
\right\}
 &\qquad&  {\text {\it Word-Level designation}} \nonumber \\
\end{eqnarray}

\end{Example}

\section{Abstraction Formulation}

Let $S$ be the system of 
polynomials, 
$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
derived from the hardware
implementation of the Galois field arithmetic circuit over $\Fkk$.
This circuit performs some unknown function $f$ over 
$\Fkk$ in the form of $Z=\Func(A_1,\dots,A_n)$, where $Z$ is the $k$-bit 
output and $A_1,\dots,A_n$ are the $k$-bit inputs.
The polynomial representation of $\Func$ over $\Fkk$ is thus:

\begin{equation}
f_{\Func}: Z+\Func(A_1,\dots,A_n) \nonumber
\end{equation}

Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
it agrees with the solution to the system of polynomials $\{S\}=0$, i.e.:
\begin{equation}
f_1=\dots=f_s=f_{A_1}=\dots=f_{A_n}=f_{Z}=0 \nonumber
\end{equation}
Thus, if we let $J=\langle f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\rangle$ 
be the ideal generated by $S$, 
$f_{\Func}$ {\bf vanishes} on the variety $V_{\Fkk}(J)$. 
Therefore, due to Proposition \ref{pro:iofv}, 
$f_{\Func}$ must be contained in the ideal of polynomials that vanish on
this variety, $f_{\Func} \in I(V_{\Fkk}(J))$.

By applying Strong Nullstellensatz over $\Fkk$ (Theorem \ref{thm:sns}), 
$I(V_{\Fkk}(J))=J+J_0$ where
$J_0$ is the ideal generated by all vanishing polynomials in $\Fkk$. 
Recall that a vanishing polynomial in $\Fkk[x]$ is $x^q-x=x^q+x$. 
In our case, 
$\{x_1,\dots,x_d\} \in \F_2$ and $\{A_1,\dots,A_n,Z\} \in \Fkk$.
Thus, for $\Fkk[x_1,\dots,x_d,A_1,\dots,A_n,Z]$: 
\begin{equation}
J_0 = \langle x_1^2+x_1,\dots,x_d^2+x_d, A_1^{2^k}+A_1,\dots,A_n^{2^k}+A_n,Z^{2^k}+Z\rangle \nonumber
\end{equation}

The generators of the ideal sum $J+J_0$ are simply the combination of the 
generators of $J$ and the generators $J_0$.

\begin{Example}
\label{exp:finalMul2Bit}
Let us re-consider Example \ref{exp:mul2bit}. First, polynomials are
extracted from the circuit implementation as
shown in Example \ref{exp:mul2bit}. These polynomials represent the
ideal $J$. Along with the ideal of vanishing polynomials $J_0$, 
the following polynomials represent the generators 
of $J+J_0$ for the multiplier circuit.

\begin{eqnarray}
 \left .  
	\begin{aligned}
		f_1:c_0+a_0 \cdot b_0  \\
		f_2:c_1+a_0 \cdot b_1  \\
		f_3:c_2+a_1 \cdot b_0  \\
		f_4:c_3+a_1 \cdot b_1  \\
		f_5:r_0+c_1 + c_2		\\
		f_6:z_0+c_0 + c_3		\\
		f_7:z_1+r_0 + c_3		
	\end{aligned} 
 \ \right\}
 &\qquad&  {\it  \text{Bit-level circuit constraints} ~(\subset J)} \nonumber \\
 \left . 
	\begin{aligned}
		f_{A}:A+a_0+a_1\cdot \alpha   \\ 
		f_{B}:B+b_0+b_1\cdot \alpha  \\ 
		f_{Z}:Z+z_0+z_1\cdot \alpha   
	\end{aligned} 
 \right\}
 &\qquad&  {\it  \text{Word-level designation} ~(\subset J)} \nonumber \\
  \left . 
	\begin{aligned}
		a_0^2-a_0, ~a_1^2-a_1,~b_0^2-b_0, ~b_1^2-b_1   \\ 
		c_0^2-c_0, ~c_1^2-c_1,~c_2^2-c_2, ~c_3^2-c_3  \\ 
		r_0^2-r_0, ~z_0^2-z_0,~z_1^2-z_1    \\ 
		A^4-A, ~B^4-B ,~Z^4-Z		  
	\end{aligned} 
 \right\}
 &\qquad&  {\it \text{vanishing polynomials} (J_0)} \nonumber
\end{eqnarray}

\end{Example}

The variety $V_{\Fq}(J)$ is
the set of all consistent assignments to the nets (signals) in the
circuit $C$. If we {\it project this variety on the word-level input and
output variables of the circuit $C$, we essentially generate the
function $\Func$ implemented by the circuit.} Projection of varieties from
$d$-dimensional space $\Fq^d$ onto a lower dimensional subspace
$\Fq^{d-l}$ is equivalent to {\it eliminating $l$ variables} from the
corresponding ideal. This can be done by computing a \Grobner basis
of the ideal with elimination ordering, as described in the 
Elimination Theorem (Theorem \ref{thm:elimth}).
Thus, we can find the polynomial 
$f_\Func:Z+\Func(A_1,\dots,A_n)$ by computing the \Grobner
basis of $J+J_0$ using the proper elimination ordering.

The proposed elimination order for 
abstraction is defined as the {\bf abstraction term order}.

\begin{Definition}
\label{def:ato}
Given a circuit $C$,
let $x_1, \dots, x_d$ denote all the bit-level variables, 
let $A_1,\dots,A_n$ denote the $k$-bit word-level inputs, 
and let $Z$ denote the $k$-bit word-level output. 
Using the partial variable order 
$\{x_1, \dots, x_d\} > Z > \{A_1, \dots, A_n\}$,
where any refinement of the order will do,
impose a lex term order $>$ on the polynomial ring 
$R = \Fq[x_1, \dots, x_d, Z, A_1, \dots, A_n]$. 
This elimination term order $>$ is defined as
the {\bf Abstraction Term Order}. The relative ordering among $x_1,
\dots, x_d$ is not important and can be chosen arbitrarily. Likewise,
the relative ordering among $A_1, \dots, A_n$ is also unimportant.
\end{Definition}

\begin{Theorem}
{\bf Abstraction Theorem:} Using the setup and notations above,
compute a \Grobner basis $G$ of ideal $(J+J_0)$ using the abstraction term 
order $>$. Then: \\
(i) For every word-level input $A_i$, $G$ must contain the vanishing 
polynomial $A_i^q - A_i$ as the only polynomial with $A_i$ as its only 
variable;\\
(ii) $G$ must contain a polynomial of the form 
$Z + \mathcal{G}(A_1,\dots,A_n)$; and\\ 
(iii) $Z + \mathcal{G}(A_1,\dots,A_n)$ is such that 
$\Func(A_1,\dots,A_n) = {\mathcal{G}}(A_1,\dots,A_n),
\forall A_1,\dots,A_n \in \Fq$. 
In other words, ${\mathcal{G}}(A_1,\dots,A_n)$ and $\Func(A_1,\dots,A_n)$ are
equal as polynomial functions over $\Fq$.
\label{thm:abs}
\end{Theorem}

\begin{Proof}\
(i) For $A_i$, $A_i^q-A_i$ is a given generator of $J_0$. 
  $A_1,\dots,A_n$ are also 
  the last variables in the abstraction term order. Moreover, $A_i$ is an
  input to the circuit, so $A_i$ is an independent variable. As a result, 
  $G_{d+1} = G \cap\Fkk[A_1,\dots,A_n] = \{A_1^q - A_1,\dots,A_n^q-A_n\}$.

(ii) Since $f:Z + \Func(A_1,\dots,A_n)$ is a polynomial representation of
  the function of the circuit, $Z + \Func(A_1,\dots,A_n) \in J + J_0$, as
  described above. Therefore, according to the definition of a
  \Grobner basis, the leading term of $Z +
  \Func(A_1,\dots,A_n)$ (which is $Z$) should be divisible by the leading
  term of some polynomial $g_i \in G$. The only way $lt(g_i)$ can
  divide $Z$ is when $lt(g_i) = Z$ itself. Moreover, due to our
  abstraction (lex) term order, $Z > A > \dots > A_n$, so this polynomial 
  must be of the form $Z + {\mathcal{G}}(A_1,\dots,A_n)$.

(iii) As $Z = \Func(A_1,\dots,A_n)$ represents the function of the circuit, 
  $Z + \Func(A_1,\dots,A_n) \in J + J_0$. 
  Moreover, $V(J + J_0) \subset V(Z + \Func(A_1,\dots,A_n))$. 
  By projecting this variety $V(J + J_0)$ onto the co-ordinates
  corresponding to $(A_1,\dots,A_n, Z)$, we obtain the {\it graph of the
  function} $(A_1,\dots,A_n) \mapsto \Func(A_1,\dots,A_n)$ 
  from $\Fkk \rightarrow \Fkk$. Since $Z + {\mathcal{G}}(A)$ is an element 
  of the \Grobner basis of $J + J_0$, 
  $V(J + J_0) \subset V(Z + {\mathcal{G}}(A_1,\dots,A_n))$. Therefore,
  $Z = {\mathcal{G}}(A_1,\dots,A_n)$ gives the same function as 
  $Z = \Func(A_1,\dots,A_n)$, for all $A_i \in \Fkk$.
\end{Proof}

As a consequence of the Abstraction Theorem, computing a \Grobner
basis $G$ of $J + J_0$ using the abstraction term order finds
a polynomial of the form $Z + \G(A_1,\dots,A_n)$ in the \Grobner basis, 
such that $Z = \G(A_1,\dots,A_n)$ is a polynomial representation of the 
circuit. However, if the \Grobner basis is not reduced, it is possible to 
obtain multiple polynomials in $G$ of the form 
$Z + \G _1(A_1,\dots,A_n), Z + \G _2(A_1,\dots,A_n), \dots,$;
all of which correspond to the same function.

\begin{Corollary}
By computing a {\bf reduced} \Grobner basis $G_r$ of $J + J_0$, 
$G_r$ will contain one and only one polynomial in of the form 
$Z + \G(A_1,\dots,A_n)$, such that $Z = \G(A_1,\dots,A_n)$ is the 
{\bf unique, minimal, canonical} representation of the function 
$\Func$ implemented by the circuit.  
\end{Corollary}

\begin{Proof}
Any function $f: \Fkk^d \to \Fkk$ has a unique canonical 
representation as polynomial $P_f \in \Fkk[x_1, ..., x_d]$ such that all 
its nonzero monomials are of the form $x_1^{i_1}\cdots x_d^{i_d}$ where $0 
\leq i_j \leq q-1$, for all $j=1, \ldots d$.

%Let $q_i$ be a power of $2$ for $i=1, \ldots, d$. 
Let $J_0$ be the 
ideal of all polynomials that vanish over $\Fkk[x_1, \ldots, x_d]$.
The generators of $J_0$ are polynomials in the form $x_i^{2^q_i}-x_i$,
where $q_i$ is the datapath size of $x_i$.
Then these generators also form a reduced Gr\"obner basis for $J_0$.
This implies that the elements $A_h^{2^k}-A_h, 1 \leq h \leq n$ will 
have to be part of the reduced Gr\"obner basis of $J+J_0$.

Corollary 1.8.6 in~\cite{gb_book} shows that the obtained element $\Func
(A_1,..., A_n)$ that is reduced modulo
$A_h^{2^k}-A_h, 1 \leq h \leq n$. Thus, the polynomial representation of 
$\Func$ in the reduced \Grobner basis is the unique canonical 
representation.
\end{Proof}

\begin{Example}
Consider the $2$-bit multiplier Example 
\ref{exp:finalMul2Bit}, for which we have already generated $J+J_0$. 
We apply abstraction term order $>$, i.e a lex order with
"bit-level variables" $>$ "Output Z" $>$ "Inputs A, B".

When we compute the reduced \Grobner basis, $G_r$, of \{$J + J_0$\} with 
respect to this ordering, $G_r$ = \{$g_1, \dots, g_{14}\}:$
\begin{align*}
g_1: B^4+B; 
~~g_2: b_0+b_1 \alpha + B; 
~~g_3: a_0+a_1 \alpha + A;  \nonumber \\
~~g_4: c_0+c_1 \alpha + c_2 \alpha + c_3(\alpha+1)+Z;
g_5: r_0+c_1+c_2; 
~~g_6: z_0+c_0+c_3; \nonumber \\
~~g_7: z_1+r_0+c_3; 
~~{\bf g_8: Z+A\cdot B};
~~g_9: b_1+B^2+B; 
~~g_{10}: a_1+A^2+A; \nonumber \\
~~g_{11}: c_3+a_1\cdot b_1
g_{12}: c_2+a_1\cdot b_1 \alpha + a_1 \cdot B; 
~~g_{13}: c_1+a_1\cdot b_1 \alpha +b_1 A; 
~~g_{14}: A^4+A  \nonumber
\end{align*}

$g_8=Z+A\cdot B$ is the {\bf canonical, word-level polynomial } 
representing the function performed by the multiplier $Z=A\cdot B$.
\end{Example}

Consolidating our results, the proposed abstraction approach is described
as follows:

\begin{enumerate}
\item Given a bit-level implementation of a Galois field arithmetic circuit 
$C$ over a given $\Fkk$, 
with $k$-bit output $Z$ and $k$-bit inputs $A_1,\dots,A_n$. 
\item $C$ performs some unknown function $Z=\Func(A_1,\dots,A_n)$.
\item Model the the circuit as a system of polynomials 
$\{f_1,\dots,f_s\}\subset \Fkk[x_1,\dots,x_d, Z, A_1, \dots, A_n]$ as 
described above and let $J$ be the ideal generated by these polynomials. 
\item Let $J_0$ be the ideal generated by all vanishing polynomials of 
$\Fkk$. 
\item By computing a reduced \Grobner basis $G_r$ of ideal $J+J_0$ using 
{\bf abstraction term order}, 
the word-level polynomial $Z+\Func(A_1,\dots,A_n)$ will be found in $G_r$.
\end{enumerate}

\section{Experimental Results: Validation of the Approach}

Our experiments take, as inputs, Mastrovito \cite{mastro:1989} multiplier 
circuits of various word sizes $k$. Each multiplier performs the polynomial 
function $Z = \Func(A,B) = A \cdot B$ over a Galois field, 
where $Z$ is the $k$-bit 
output and $A$ and $B$ are the $k$-bit inputs. We extract the 
Boolean gate-level operators $J$ and vanishing polynomials $J_0$.
Then we compute the reduced \Grobner basis $G_r$ of $J + J_0$
with respect to abstraction term order $>$. The resulting $G_r$ contains a 
polynomial $Z + A\cdot B$.

The experiments were designed as scripts in the computer algebra tool, 
{\sc Singular} \cite{DGPS}, which provides functionality for polynomial
computations over rings and fields. This tool provides a number of efficient 
polynomial algorithms. The ring $\Fkk[x_1,\dots,x_f,\allowbreak Z,A,B]$ is defined over
the abstraction ordering, using the same primitive polynomial 
(``minpoly'' in Singular) $P(X)$ that was used to design the Galois field multiplier. 
Ideals  $J$ and $J_0$ were provided using their generating polynomials and the
the reduced \Grobner basis computation of $J+J_0$ was performed
using the ``slimgb'' command.

The experiments were conducted on a 64-bit Ubuntu machine running a 2.4GHz 
processor with 8GB of memory. We applied our approach to
abstract the canonical, polynomial representation of Mastrovito 
multipliers of various sizes.
Our machine was unable to perform the 
computations of the Gr\"obner basis of multipliers beyond $40$-bit word 
inputs.

\begin{table}[H]
\label{tab:slimgb}
	\begin{center}
	    \caption{Run-time of Gr\"obner Basis Computation of Mastrovito Multipliers in Singular using Abstraction Term Order $>$.}\label{tab:sT}
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (minutes)   \\
		\hline
	        $16$	&  $1,871$  & $2.4$ \\
		$24$	&  $3,135$  & $12$  \\
	        $32$	&  $5,549$  & $22.6$ \\
	        $40$	&  $8,587$  & $266$ \\
		$48$	& $12,327$  & NA (Out of Memory) \\
	        \hline
	    \end{tabular}
	\end{center} 
\end{table}

\section{Conclusions}
The above approach is guaranteed to find a canonical, word-level 
representation of the function $\Func$ performed by a circuit $C$ over 
$\Fkk$. However, the \Grobner basis computation is prohibitively complex for
circuits of practical sizes. The next chapter proposes a 
method to overcome the complexity of the \Grobner
basis computation in order to make this abstraction approach scalable.

### 2016-09-21 dissertation/algprelim.tex
init dissertation writing

\chapter{Computer Algebra Fundamentals} \label{ch:ideals}
This chapter reviews fundamental concepts of commutative and 
computer algebra which are used in this work. 
Specifically, this chapter covers monomial ordering, polynomial ideals and 
varieties, and the computation of \Grobner bases.
It also overviews elimination theory as well as Hilbert's Nullstellensatz 
theorems and how
they apply to Galois fields. The results of these theorems are used in polynomial
abstraction and formal verification of Galois field circuits and 
are discussed 
in subsequent chapters. The material of this chapter is
mostly referred from the textbooks \cite{ideals:book} \cite{gb_book} and 
previous work by {\it Lv} \cite{lv:phd}.

\section{Monomials, Polynomials, and Term Orderings}

\begin{Definition} \label{def:mono}
A {\bf monomial} in variables $x_1,x_2,\cdots,x_d$ is a product of the form:
\begin{equation}
x_1^{{\alpha}_1} \cdot x_2^{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d},
\end{equation}
where $\alpha_i \ge 0, i\in\{1,\cdots,d\}$. 
The total degree of the monomial is $\alpha_{1}+\cdots+\alpha_{d}$.
\end{Definition}

Thus, $x^2\cdot y$ is a monomial in variables $x,y$ with total degree $3$.
For simplicity, we will henceforth denote a 
monomial $x_1^{{\alpha}_1} \cdot x_2^
{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d}$ as $x^{\alpha}$, 
where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$. In the

\begin{Definition}
A {\bf multivariate polynomial} $f$ in variables $x_1, x_2, \ldots, x_d$ 
with coefficients in any given field $\K$ is a finite linear 
combination of monomials with coefficients in $\K$: 
\begin{equation}
	f=\sum_{\alpha}a_{\alpha}\cdot x^{\alpha}, ~~a_{\alpha}\in \K \nonumber
\end{equation}

The set of all polynomials in $x_1, x_2, \ldots, x_d$ with coefficients in 
field $\K$ is denoted by $\K[x_1, x_2, \ldots, x_d]$. 
Thus, $f \in \K[x_1, x_2, \ldots, x_d]$

\begin{enumerate}
\item We refer to the constant $a_{\alpha} \in \K$ as the 
{\bf coefficient} of the monomial $a_{\alpha} x^{\alpha}$.
\item If $a_{\alpha} \neq 0$, we call $a_{\alpha} x^{\alpha}$ a term of $f$.
\end{enumerate}
\end{Definition}

As an example, $2x^2+y$ is a polynomial with two terms $2x^2$ and $y$, with 
$2$ and $1$ as coefficients respectively. 
In contrast, $x+y^{-1}$ is not a polynomial because the exponent of $y$ is 
less than $0$.

Since a polynomial is a sum of its terms,
these terms have to be arranged unambiguously so that they can be 
manipulated in a consistent manner.
Therefore, we need to establish a concept of 
{\bf term ordering} (also called monomial ordering).
A term ordering, represented by $>$, defines how
terms in a polynomial are ordered.
%%%%%%%%%%%%%%%%%%%   monomial    Ordering   %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{Definition}
Let $\mathbb{T}^{d}=\{x^{\alpha}: \alpha\in \mathbb{Z}_{\ge 0}^{d}\}$ be the set of all monomials in $x_{1},\dots,x_{d}$.
A {\bf monomial order} $>$ on $\mathbb{T}^{d}$ is a total well-ordering satisfying:
\begin{itemize}
        \item For all $x^{\alpha}, x^{\beta} \in \mathbb{T}^{d}$, $x^{\alpha}$ and $x^{\beta}$ are comparable
	\item For any $x^{\alpha} \in \mathbb{T}^{d}$, $x^{\alpha}>1$
	\item For all $x^{\alpha}, x^{\beta}, x^{\gamma} \in \mathbb{T}^{d}$, $x^{\alpha}>x^{\beta} \Rightarrow x^{\alpha} \cdot x^{\gamma}> x^{\beta} \cdot x^{\gamma}$
\end{itemize}
\end{Definition}

Term-orderings are totally ordered, i.e. anti-symmetric with constant terms
last in the ordering.
A total-order ensures that there is no ambiguity with respect to where a 
term is found in the term-ordering.  Total orderings for monomials come in 
different forms, notably 
lexicographic orderings (lex), and its variants: degree-lexicographic ordering (deglex) 
and reverse degree-lexicographic ordering (degrevlex).

A {\bf lexicographic ordering} (lex) is a total-ordering $>$ such that 
variables in the terms are lexicographically ordered, i.e. simply based on 
when the variables appear in the ordering.
Higher variable-degrees take 
precedence over lower degrees for equivalent variables (e.g. $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
\begin{Definition}
{\bf Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
& \text{Starting  from the  left, the first co-ordinates of $\alpha_i, \beta_i$} \\
& \text{that are different satisfy $\alpha_i > \beta_i$}

\end{cases}
\end{equation}
\end{Definition}

A {\bf degree-lexicographic ordering} (deglex) is a total-ordering $>$ such 
that the total degree of a term takes precedence over the lexicographic 
ordering.  
A {\bf degree-reverse-lexicographic ordering} (degrevlex) is the same as a
deglex ordering, however terms are lexed in reverse.

\begin{Definition}
{\bf Degree Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
\sum_{i=1}^{d}\alpha_i > \sum_{i=1}^{d} \beta_i & \text{ or }\\
\sum_{i=1}^{d}\alpha_i = \sum_{i=1}^{d} \beta_i  \text{ and }
x^{\alpha} > x^{\beta} & \text{w.r.t. lex order}
\end{cases}
\end{equation}
\end{Definition}

\begin{Definition}
{\bf Degree Reverse Lexicographic order:} Let $x_1 > x_2 > \dots > x_d$
lexicographically. Also let $\alpha = (\alpha_1, \dots, \alpha_d);
~\beta = (\beta_1, \dots, \beta_d) \in \mathbb{Z}^d_{\geq 0}$. Then we
have: 
\begin{equation}
x^{\alpha} > x^{\beta} \iff 
\begin{cases}
\sum_{i=1}^{d}\alpha_i > \sum_{i=1}^{d} \beta_i  \text{ or }\\
\sum_{i=1}^{d}\alpha_i = \sum_{i=1}^{d} \beta_i  \text{ and the first co-ordinates}\\
\text{$\alpha_i, \beta_i$ from the right, which are different, satisfy $\alpha_i < \beta_i$}
\end{cases}
\end{equation}

\end{Definition}

Applying these term orderings, we have the following relations, 
where $a > b > c$.

\begin{eqnarray}
    \eqntext{lex:}  
    a^2b > a^2 > abc > ab > ac^2 > ac > b^2c > b^2 > bc^3 > 1 
    \label{ex:ordering:lex}\\
    \eqntext{deglex:} 
    bc^3 > a^2b > abc > ac^2 > b^2c > a^2 > ab >
    ac > b^2 > 1
    \label{ex:ordering:deglex}\\
    \eqntext{degrevlex:}  
    bc^3 > a^2b > abc > b^2c > ac^2 > a^2 > ab >
    b^2 > ac > 1
    \label{ex:ordering:degrevlex}
\end{eqnarray}

The difference between the {\it lex} and two {\it deg-} orderings is
obvious, while the difference between the two degree-based orderings can be
seen by considering from which direction the term is lexed, e.g. $a\cdot c\cdot c > b\cdot b\cdot c$ (deglex, left-to-right)
versus $b\cdot b\cdot c > a \cdot c\cdot c$ (degrevlex, right-to-left).

\begin{Example}
Let $f = 2x^2yz + 3xy^3 - 2x^3$. The effects of different term orderings on 
$f$ are:
\begin{itemize}
\item lex $x> y> z$: $f = -2x^3 + 2x^2yz + 3xy^3$
\item deglex $x>y>z$:  $f = 2x^2yz + 3xy^3 -2x^3$
\item degrevlex $x>y>z$: $f = 3xy^3 + 2x^2yz - 2x^3$
\end{itemize}
\end{Example}

%Based on the {\it monomial ordering}, we have the following concepts:

\begin{Definition}
The {\bf leading term} is the first term in a term-ordered polynomial.
Likewise, the {\bf leading coefficient} is the coefficient of
the leading term. 
Finally, a {\bf leading monomial} is the leading term 
lacking the coefficient.  We use the following notation:
\begin{eqnarray}
     lt(f)&& \text{--- Leading Term} \\
     lc(f)&& \text{--- Leading Coefficient} \\
     lm(f)&& \text{--- Leading Monomial} \\
     tail(f)&& f - lt(f)
\end{eqnarray}
\end{Definition}

\begin{Example}
\begin{eqnarray}
     f      &=& 3a^2b + 2ab + 4bc \\
     lt(f)  &=& 3a^2b \\
     lc(f)  &=& 3 \\
     lm(f)  &=& a^2b \\
     tail(f) &=& 2ab+4bc
\end{eqnarray}
\end{Example}

{\bf Polynomial division} is an operation over polynomials that is dependent on
the imposed monomial ordering. Dividing a polynomial $f$ by another polynomial
$g$ cancels the leading term of $f$ to derive a new polynomial.

\begin{Definition}
Let $\K$ be a field and let $f, g \in \K[x_1, x_2, \ldots, x_d]$ be polynomials
over the field. {\bf Polynomial division} of $f$ by $g$ is computes following:
\begin{equation}
f-\frac{lt(f)}{lt(g)}\cdot g
\end{equation}
This polynomial division is denoted
\begin{equation}
f\xrightarrow{g} r
\end{equation}
where $r$ is the resulting polynomial of the division.
If $\frac{lt(f)}{lt(g)}$ is non-zero, then $f$ is considered divisible by $g$, 
i.e. $g \mid f$.
\end{Definition}
Notice that if $g \nmid f$, that is if $f$ is not divisible by $g$,
then the division operation gives $r = f$.
\begin{Example}
Over $\R[x,y,z]$, set the lex term order $x > y > z$.
Let $f = -2x^3 + 2x^2yz + 3xy^3$ and $g = x^2+yz$.
\begin{equation}
\frac{lt(f)}{lt(g)} = \frac{-2x^3}{x^2} = -2x
\end{equation}
Since $\frac{lt(f)}{lt(g)}$ is non-zero $g|f$. The division, $f\xrightarrow{g} r$, 
is computed as:
\begin{eqnarray}
r &=& f-\frac{lt(f)}{lt(g)}\cdot g = -2x^3 + 2x^2yz + 3xy^3 - (-2x \cdot (x^2+yz)) \nonumber \\
&=& -2x^3 + 2x^2yz + 3xy^3 - (-2x^3-2xyz) = 2x^2yz + 3xy^3 + 2xyz
\end{eqnarray}
Notice that the division cancels the leading term of $f$.
\end{Example}

\section{Varieties and Ideals}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In computer-algebra based formal verification,
it is often necessary to analyze the presence 
or absence of solutions to a given system of constraints.
In our applications, these constraints are polynomials and their solutions 
are modeled as {\bf varieties}.

\begin{Definition}
Let $\K$ be a field, and let $f_1, \ldots, f_s 
\in \K[x_1, x_2, \ldots, x_d]$. 
We call $V(f_1, \dots, f_s)$ the {\bf affine variety} 
defined by $f_1, \dots, f_s$ as:
\begin{equation}
V(f_1, \ldots, f_s)= \{(a_1, \ldots, a_{d})\in \K^d:f_i(a_1, \ldots, a_d)=0, \forall{i},1\le i \le s\}
\end{equation}
\end{Definition}

$V(f_1, \dots, f_s)\in \K^d$ is {\bf the set of all solutions} in $\K^d$ 
of the system of equations: 
$f_1(x_1,\ldots,x_d)=\dots=f_s(x_1,\dots,x_d)=0$.

\begin{Example}
Given $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2)$ is the set of all elements
that satisfy $x^2+y^2=0$ over $\mathbb{R}^2$. So $V(x^2+y^2)=\{(0,0)\}$. 
Similarly, in $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2-1)=\{all\  points\  on\ the\ circle: x^2+y^2-1=0\}$.
Note that varieties depend on which field we are operating on. 
For the same polynomial $x^2+1$, we have:
\begin{itemize}
\item In $\mathbb{R}[x]$, $V(x^2+1)=\emptyset$.
\item In $\mathbb{C}[x]$, $V(x^2+1)=\{(\pm i)\}$.
\end{itemize}
\end{Example}

The above example shows the variety can be infinite, finite (non-empty set) 
or empty. It is interesting to note that since we will be operating over
finite fields $\F_{q}$, and any finite set of points is a variety. Likewise,
any variety over $\F_{q}$ is finite (or empty).
Consider the points $\{(a_1,\dots, a_d): a_1, \dots, a_d \in \F_q\}$
in $\F_q^d$. Any single point is a variety of some polynomial system:
e.g. $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
varieties are also varieties.

\begin{Example}
Let $U = V(f_1, \dots, f_s)$ and $W =
V(g_1, \dots, g_t)$ in $\F_{q}$. Then:  
\begin{itemize}
\item $U \cap W = V(f_1, \dots, f_s, g_1, \dots, g_t)$
\item $U \cup W = V(f_i g_j: 1 \leq i \leq s, 1 \leq j \leq t)$
\end{itemize}
\end{Example}

One important distinction we need to make about varieties is that a
variety depends not just on the given system of polynomial 
equations, but rather on the {\bf ideal} generated by the polynomials.

%%%%%%%%%%%%%%%%%%%%%  ideal %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{Definition} 
A subset $I \subset \K[x_1, x_2, \ldots, x_d]$ is an {
\bf ideal} if it satisfies:
\begin{itemize}
\item $0 \in I$
\item $I$ is closed under addition: $x, y \in I \Rightarrow x+y \in I$
\item If $x \in \K[x_1, x_2, \ldots, x_d]$ and $y \in I$, then $x\cdot y \in I$ and $y\cdot x \in  I$.
\end{itemize}
\end{Definition}

An ideal is generated by its {\it basis} or {\it generators}.

%%%%%%%%%%%%%%%%%%%%%  ideal basis%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{Definition}
Let $f_1, f_2, \ldots, f_s$ be 
polynomials of the ring $\K[x_1, x_2, \ldots, x_d]$. 
Let $I$ be an ideal generated by $f_1, f_2, \ldots, f_s$. Then:
\begin{equation}
I = \langle f_1,\dots,f_s \rangle=\{h_1 f_1 + h_2 f_2 + \ldots + h_s f_s : h_1,\dots,h_s\in\K[x_1, \dots, x_d]\} \nonumber
\end{equation}
then, $f_1, \ldots, f_s$ are called the {\bf basis (or generators)} of the 
ideal $I$ and correspondingly $I$ is denoted as $I = \langle f_1, f_2, 
\ldots, f_s \rangle$. 
\end{Definition}

\begin{Example}
The set of even integers, which is a subset of the ring of integers
$Z$, forms an ideal of $Z$. This can be seen from the following;
\begin{itemize}
\item $0$ belongs to the set of even integers.
\item The sum of two even integers $x$ and $y$ is always an even
  integer.
\item The product of any integer $x$ with an even integer $y$ is
  always an even integer.
\end{itemize}
\end{Example}

\begin{Example}
Given $\mathbb{R}\left[x,y\right]$, $I = \langle x, y \rangle$ is an 
ideal containing all polynomials generated by $x$ and $y$, 
such as $x^2+y$ and $x+x\cdot y$.
$J = \langle x^2, y^2 \rangle$ is an ideal containing all polynomials 
generated by $x^2$ and $y^2$, such as $x^2+y^3$ and $x^{10}+x^2\cdot y^2$. 
Notice that $J\subset I$ because every polynomial generated by $J$ can
be generated by $I$. 
But $I\neq J$ because $x+y$ can only be generated by $I$.
\end{Example}

The same ideal may have many different bases.
For instance, it is possible to have different sets of polynomials
$\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
ideal, i.e., 
$\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
variety depends on the ideal, these sets of polynomials have the same 
solutions.

\begin{Proposition}
If $f_1,\dots,f_{s}$ and $g_{1},\dots,g_{t}$ are bases of the same ideal 
in $\mathbb{F}[x_{1},\dots,x_{d}]$,
so that $\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$, 
then $V(f_{1},\dots,f_{s})=V(g_{1},\dots,g_{t})$.
\end{Proposition}

\begin{Example}
	Consider the two bases $F_{1}=\{(2x^{2}+3y^{2}-11,x^{2}-y^{2}-3\}$ and $F_{2}=\{x^{2}-4,y^{2}-1\}$.
	These two bases generate the same ideal, i.e., $\langle F_{1}\rangle= \langle F_{2} \rangle$.{}
	Therefore, they represent the same variety, i.e., 
	\begin{equation}
		V(F_{1})= V( F_{2})=\{\pm 2, \pm 1\}.
	\end{equation}
\end{Example}

Ideals and their varieties are a key part of computer-algebra based formal
verification. A given hardware design can be transformed into a set
of polynomials over a field, $f_1, \ldots, f_s \in F$ 
(we showed how this is done for Galois field arithmetic 
circuits in the previous chapter).
This set of polynomials gives the system of equations:
\begin{eqnarray}
f_1 = 0 \nonumber \\
\vdots \nonumber \\
f_s = 0 \nonumber
\end{eqnarray}
Using algebra, it is possible to derive new equations from the original 
system.
The ideal $\langle f_1,\ldots, f_s \rangle$ provides a way of analyzing 
such {\it consequences} of a system of polynomials.

\begin{Example}
Given two equations in $\mathbb{R}[x,y,z]$:
\begin{eqnarray}
x=z+1 \nonumber \\
y=x^2+1 \nonumber 
\end{eqnarray}
we can eliminate $x$ to obtain a new equation:
\begin{equation}
y=(z+1)^2+1=z^2+2z+2 \nonumber 
\end{equation}
Let $f_1, f_2, h \in \mathbb{R}[x,y,z]$ be polynomials based on these 
equations:
\begin{eqnarray}
f_1 = x-z-1 &= 0 \nonumber \\
f_2 = y-x^2-1 &= 0 \nonumber \\
h   = y-z^2-2z-2 &= 0 \nonumber
\end{eqnarray}
If $I$ is the ideal generated by $f_1$ and $f_2$, i.e. 
$I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
\begin{eqnarray}
g_1 = x+z+1 \nonumber \\
g_2 = 1     \nonumber \\
h = g_1\cdot f_1+g_2\cdot f_2  = y-z^2-2z-2 \nonumber
\end{eqnarray}
where $g_1, g_2 \in \mathbb{R}[x,y,z]$.
Thus, we call $h$ a {\bf member of the ideal} $I$.
\end{Example}

%\subsection{Ideals of Varieties}

Let $\mathbb{K}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{K}^d$ be a point, and $f \in
\mathbb{K}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ if $f(\mathbf{a}) = 0$, i.e.,
$\mathbf{a}$ is in the variety of $f$.

\begin{Definition}
For any variety $V$ of $\mathbb{K}^d$, the ideal of polynomials that vanish on $V$,
called the {\it vanishing ideal of $V$}, is defined as $I(V) = \{f\in
\mathbb{F}[x_1,\dots, x_d]: \forall \mathbf{a} \in V, f(\mathbf{a}) =
0\}$. 
\end{Definition}

\begin{Proposition}\label{pro:iofv}
	If a polynomial $f$ vanishes on a variety $V$, then $f \in I(V)$. 
\end{Proposition}

\begin{Example}
	Let ideal $J=\langle x^{2},y^{2}\rangle$. Then $V(J)=\{(0,0)\}$.
	All polynomials in $J$ will obviously agree with the solution and vanish on this variety.
	However, the polynomials $x,y$ are not in $J$ but they also vanish on this variety. 
	Therefore, $I(V(J))$ is the set of all polynomials that vanish on $V(J)$, and the polynomials
	$x,y$ are members of $I(V(J))$.
\end{Example}

\begin{Definition}\label{def:radical}
Let $J \subset \mathbb{K}[x_1,\dots, x_d]$ be an ideal. The {\it radical of $J$} is defined as $\sqrt{J} = \{f \in
\mathbb{K}[x_1,\dots, x_d]: \exists m \in \mathbb{N}, f^m \in J\}$. 
\end{Definition}

\begin{Example}
Let $J=\langle x^2,y^2\rangle \subset \mathbb{K}\left[x,y\right]$.
Note neither $x$ nor $y$ belongs to $J$, but they belong to $\sqrt J$.
Similarly, $x\cdot y \notin J$, but since $(x \cdot y)^{2}=x^{2}\cdot y^{2}\in J$, therefore,
$x\cdot y \in \sqrt J$. 
\end{Example}

When $J = \sqrt J$, then $J$ is said to be a 
{\it radical ideal}. Moreover, $I(V)$ is a radical ideal.
By analyzing the ideal $J$, generated by a system of polynomials derived 
from a hardware design, its variety $V(J)$, and the ideal of 
polynomials that vanish over this variety, $I(V(J))$, we can reason about the 
existence of certain properties of the design. To check for the existence of 
a property, we formulate the 
property as a polynomial and then perform an {\bf ideal membership test} to 
determine if this polynomial is contained within the ideal $I(V(J))$. 
A {\bf \Grobner basis} provides a decision procedure for performing this 
test, which is described in the following section. 
A future section focuses on {\bf Hilbert's Nullstellensatz}, 
which describes the properties of the ideal of a variety, $I(V(J))$.

%%%%%%%%%%%%%%%%%%%%	Grobner bases	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\Grobner Bases}

As mentioned earlier, different polynomial sets may generate the same 
ideal. Some of these generating sets may be a better representation of 
the ideal, and thus provide more information and insight into the properties 
of ideal. One such ideal representation is a {\bf \Grobner basis}, which has
a number of important properties that can solve numerous polynomial 
decision questions:

\begin{itemize}
\item Presence or absence of solutions (varieties)
\item Dimension of the varieties
\item Ideal membership of a polynomial
\end{itemize}

In essence, a \Grobner basis is a canonical representation of an ideal.
There are many equivalent definitions of \Grobner bases, so we start with 
the definition that best describes their properties:

\begin{Definition}
A set of non-zero polynomials $G=\{g_1,\dots,g_t\}$ which generate the 
ideal $I=\langle g_1,\dots,g_t\rangle$, is called a 
{\bf Gr\"obner basis} for $I$ if and only if 
for all $f \in I$ where $f \neq 0$, there exists a $g_i \in G$ such that $lm(g_i)$ divides $lm(f)$.
\begin{eqnarray}
G = \text{Gr\"obner{Basis}} (I) \iff 
\forall f \in I: f \neq 0, \exists g_i \in G: lm(g_i)\ |\ lm(f)
        \label{eqn:groebnermin}
    \end{eqnarray}

\end{Definition}

The foundation for computing the \Grobner basis of an ideal was laid out 
by Buchberger\cite{buchberger_thesis}.
Given a set of polynomials $F=\{f_{1},\dots,f_{s}\}$ that generate ideal $I=
\langle f_{1},\dots,f_{s} \rangle$, 
Buchberger gives an algorithm to compute a Gr\"obner basis $G=\langle g_{1},
\dots,g_{t}\rangle$. This algorithm relies on the notions of $S$-polynomials 
and polynomial reduction.

\begin{Definition}
For $f, g \in \K[x_1,\dots,x_d]$, an {\bf S-polynomial} $Spoly(f,g)$ is 
defined as:
\begin{equation}
    Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g
    \label{eqn:spoly}
\end{equation}
\begin{equation}
\text{where }L = lcm\left(lt(f), lt(g)\right) \nonumber
\end{equation}
Note $lcm$ denotes least common multiple.
\end{Definition}

\begin{Definition}
    The {\bf reduction} of a polynomial $f$, by another polynomial $g$, to
    a reduced polynomial $r$ is denoted:
    \begin{equation*}
        f\stackrel{g}{\textstyle\longrightarrow}r
    \end{equation*}
    Reduction is carried out using multivariate, polynomial long division. 
    % The long division is performed according to a term-ordering on polynomials, and the division algorithm 
    %terminates when the leading term of the divisor does not divide any 
    %other term in the dividend.

    For sets of polynomials, the notation 
    \begin{equation*}
    f\stackrel{F}{\textstyle\longrightarrow}_+r    
    \end{equation*}
    represents the reduced polynomial $r$ resulting from $f$ as reduced by a 
    set of non-zero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
    $r = 0$  or no term in $r$ is divisible  by a $lm(f_i), \forall f_i \in F$.
\end{Definition}

The reduction process $f\stackrel{F}{\textstyle\longrightarrow}_+r$, of 
dividing a polynomial $f$ by a set of polynomials of $F$, can be modeled as
repeated long-division of $f$ by each of the polynomials in $F$ until no
further reductions can be made. The result of this process is then $r$.
This reduction process is shown in Algorithm \ref{alg:polydiv}.

\begin{algorithm}[H]
\SetAlgoNoLine

 \KwIn{$f,f_{1},\dots,f_{s}$}
 \KwOut{$r,a_{1},\dots,a_{s}$, such that $f=a_{1}\cdot f_{1}+\dots+a_{s}\cdot f_{s}+r$.}

 $a_{1}=a_{2}=\dots=a_{s}=0$; $r=0$\;
 $p:=f$\;

 \While { $p \neq 0$ }
 {
	i=1\;
	divisionmark = false\;
	\While { $i\le s  $ \&\& divisionmark = false }
	{
		\eIf {$f_{i}$ can divide $p$}
		{
			$a_{i}=a_{i}+lt(p)/lt(f_{i})$\;
			$p=p-lt(p)/lt(f_{i}) \cdot f_{i}$\;
			divisionmark = true\;
		}
		{
			i=i+1\;
		}
	}

	\If {divisionmark = false}
	{
		$r=r+lt(p)$\;
		$p=p-lt(p)$\;
	}

 }
\caption{Polynomial Reduction}\label{alg:polydiv}
\end{algorithm}

The reduction algorithm keeps canceling the leading terms of polynomials 
until no more leading terms can be further canceled.
So the key step is $p=p-lt(p)/lt(f_{i}) \cdot f_{i}$, as the following 
example shows.
\begin{Example}
Given $f = y^{2}-x$ and $f_{1} = y - x$ in $\mathbb{Q}[x,y]$ with $deglex$: 
$y>x$, perform $f\stackrel{f_1}{\textstyle\longrightarrow}_+r$:

\begin{enumerate}
\item $f=y^{2}-x$, $f/f_{1}=f-lt(f)/lt(f_{1}) \cdot f_{1}=y^{2}-x-(y^{2} /y) \cdot (y-x)=y\cdot x-x$
\item $f=y\cdot x-x$, $f/f_{1}=f-lt(f)/lt(f_{1}) \cdot f_{1}=(y\cdot x-x)/f_{1}=x^{2}-x$
\item $f=x^{2}-x$, no more operations possible, so $r=x^{2}-x$
\end{enumerate}
\end{Example}

With the notions of $S$-polynomials and polynomial reduction in place,
we can now present Buchberger's Algorithm 
for computing Gr\"obner bases \cite{buchberger_thesis}. Note that a fixed 
monomial (term) ordering is required for a \Grobner basis 
computation to ensure that polynomials are manipulated in a consistent 
manner.

\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{$F = \{f_1, \dots, f_s\}$, such that $I=\langle f_1, \dots, f_s\rangle$}, and term order $>$
 \KwOut{$G = \{g_1,\dots ,g_t\}$, a Gr\"{o}bner basis of $I$ }
  $G:= F$\;
  \Repeat{$G = G'$}
  {
  	$G' := G$\;
  	\For{ each pair $\{f_{i}, f_{j}\}, i \neq j$ in $G'$} 
	{
		$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$ \;
		\If{$r \neq 0$}
		{
			$G:= G \cup \{r\}$ \;
		}
	}
   }
\caption {Buchberger's Algorithm}\label{alg:gb}
\end{algorithm}

Buchberger's algorithm takes pairs of polynomials ($f_{i}, f_{j}$) in 
the basis $G$ and combines them into ``$S$-polynomials'' 
($Spoly(f_{i}, f_{j})$) to cancel leading terms. The $S$-polynomial is then 
reduced (divided) by all elements of $G$ to a remainder $r$, denoted as  
$Spoly(f_{i}, f_{j}) \stackrel{G}{\textstyle\longrightarrow}_+r$. This
process is repeated for all unique pairs of polynomials, including
those created by newly added elements, until no new polynomials are
generated; ultimately constructing the \Grobner basis.
\begin{Example}\label{exp:gbsimple}
Consider the ideal $I \subset \mathbb{Q}[x, y]$, $I = \langle f_1, f_2 
\rangle$, where $f_1 = yx - y, ~f_2 = y^2 - x$. 
Assume a degree-lexicographic term ordering with $y > x$ is imposed.

First, we need to compute $Spoly(f_{1},f_{2})=x\cdot f_{2}-y\cdot f_{1}=y^{2}-x^{2}$.
Then we conduct a polynomial reduction 
$y^{2}-x^{2}\stackrel{f_{2}}{\textstyle\longrightarrow}x^{2}-x \stackrel{f_{1}}{\textstyle\longrightarrow}x^{2}-x$.
Let $f_{3}=x^{2}-x$. Then $G$ is updated as $\{f_{1},f_{2},f_{3}\}$. Next we compute $Spoly(f_{1},f_{3})=0$. So there
is no new polynomial generated. Similarly, we compute $Spoly(f_{2},f_{3})=x\cdot y^{2}-x^{3}$, followed by 
$x\cdot y^{2}-x^{3}\stackrel{f_{1}}{\textstyle\longrightarrow}y^{2}-x^{3} \stackrel{f_{2}}{\textstyle\longrightarrow}x-x^{3}
\stackrel{f_{2}}{\textstyle\longrightarrow}0$. Again, no polynomial is generated. Finally, $G=\{f_{1,}f_{2},f_{3}\}$.

\end{Example}

When computing a \Grobner basis, it's important to note that if $lt(f_i)$ 
and $lt(f_j)$ have no common variables, the S-poly reduction step in 
Buchberger's algorithm,
$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$,
will produce $r=0$.

\begin{Proof}
If $lt(f)$ and $lt(g)$ have no common variables,  
$L=lcm(lt(f),lt(g))=lt(f)\cdot lt(g)$. Then: 
\begin{equation}
    Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g=
\frac{lt(f)\cdot lt(g)}{lt(f)}\cdot f - \frac{lt(f)\cdot lt(g)}{lt(g)}\cdot g
= lt(g)\cdot f - lt(f)\cdot g \nonumber
\end{equation}
Thus, every monomial in $Spoly(f, g)$ is divisible by either $lt(f)$ 
or $lt(g)$, so computing 
$Spoly(f, g) \stackrel{f,g}{\textstyle\longrightarrow}_+r$ will give $r=0$.
\end{Proof}

As mentioned previously, a \Grobner basis gives a decision procedure to test 
for polynomial membership in an ideal. This is explained in the following 
Theorem.

\begin{Theorem}\label{the:membership}
	{\bf Ideal Membership Test}
 Let $G = \{g_1,\cdots,g_t \}$ be a Gr\"obner basis for an ideal $I \subset \mathbb{K}[x_1,\cdots,x_d ]$
	and let $f \in \mathbb{K}[x_{1},\dots, x_{d}]$. Then $f \in I$ if and only if the remainder on division of $f$ by
	$G$ is zero.
\end{Theorem}
In other words, 
\begin{equation}
f \in I \iff f \stackrel{G}{\textstyle\longrightarrow}_+0
\end{equation}

\begin{Example}
Consider Example \ref{exp:gbsimple}. Let $f = y^2x - x$ be another
polynomial. Note that $f = yf_1 + f_2$, so $f \in I$. If we divide $f$
by $f_1$ first and then by $f_2$, we will obtain a zero
remainder. However, since the set $\{f_1, f_2\}$ is not a Gr\"{o}bner
basis, we find that the reduction $f
\stackrel{f_2}{\textstyle\longrightarrow} x^2 - x
\stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
i.e. dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
$I$, $G = \{x^2 - x, yx - y, y^2 - x\}$, dividing $f$ by polynomials
in $G$ in any order will always lead to the zero remainder. Therefore,
one can decide ideal membership unequivocally using the Gr\"{o}bner
basis. 
\end{Example}

A \Grobner basis is not a canonical representation of an ideal, but a
{\bf reduced \Grobner basis} is. To compute a reduced \Grobner basis, we
first must compute a minimal \Grobner basis.

\begin{Definition}\label{def:minigb}
A {\bf minimal Gr\"obner basis} for a polynomial ideal $I$ is a \Grobner basis $G$ for $I$ such that
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$,  $lt(g_{i}) \notin \langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
A {\bf minimal} \Grobner basis is a \Grobner basis such that all polynomials
have a coefficient of $1$ and no leading term of any element in $G$ divides 
another in $G$.
Given a \Grobner basis $G$, a minimal \Grobner basis can be
computed as follows:
\begin{enumerate}
\item Minimize every $g_i \in G$, i.e $g_i=g_i/lc(g_i)$
\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, i.e. remove every polynomial in $G$ whose leading term is divisible by the leading term of some other polynomial in $G$.
\end{enumerate}

A minimal Gr\"obner basis can then be further reduced.
\begin{Definition}
	A {\bf reduced Gr\"obner basis} for a polynomial ideal $I$ is a Gr\"obner basis $G=\{g_{1},\dots,g_{t}\}$ such that:
	\begin{itemize}
		\item $lc(g_{i})=1,\forall g_{i}\in G$
		\item $\forall g_{i} \in G$, no monomial of $g_{i}$ lies in $\langle lt(G-\{g_{i}\})\rangle$
	\end{itemize}
\end{Definition}
$G$ is a reduced Gr\"obner basis when no monomial of any element in $G$ 
divides the leading term of another element. 
This reduction is achieved as follows:

\begin{Definition}
Let $H = \{h_1, \ldots, h_t\}$ be a minimal Gr\"obner basis.  Apply
the following reduction process: 
\begin{itemize}
\item $h_1 \stackrel{G_1}{\textstyle\longrightarrow}_+ g_1$, where
  $g_1$ is reduced w.r.t. $G_1 = \{h_2, \ldots, h_t\}$

\item $h_2 \stackrel{G_2}{\textstyle\longrightarrow}_+ g_2$, where
  $g_2$ is reduced w.r.t. $G_2 = \{g_1, h_3, \ldots, h_t\}$
\item $h_3 \stackrel{G_3}{\textstyle\longrightarrow}_+ g_3$, where
  $g_3$ is reduced w.r.t. $G_3 = \{g_1, g_2, h_4, \ldots, h_t\}$

\hspace{0.25in} $\vdots$
\vspace{0.1in}
\item $h_t \stackrel{G_t}{\textstyle\longrightarrow}_+ g_t$, where
  $g_t$ is reduced w.r.t. $G_t = \{g_1, g_2, g_3, \ldots, g_{t-1}\}$
\end{itemize}
Then $G = \{g_1, \ldots, g_t\}$ is a {\bf reduced Gr\"obner basis.}
\end{Definition}

Subject to the given term order $>$, such a reduced Gr\"obner
  basis $G = \{g_1, \dots, g_t\}$ is a {\bf unique canonical
    representation of the ideal}, as 
given by Proposition \ref{pro:unique} below.

\begin{Proposition}\label{pro:unique} \cite{gb_book} 
Let $I \neq \{0\}$ be a polynomial ideal. Then, for a given monomial ordering, $I$ has a unique reduced Gr\"obner basis.
\end{Proposition}

\Grobner basis computation depends on the $Spoly$ computation, which in turn 
depends on the leading terms of polynomials. Thus, different monomial 
orderings can result in different \Grobner basis computations for the 
same ideal. Computation using a degrevlex ordering tends to be least 
difficult, while lex ordering tends to be computationally complex. However, 
lex ordering used in the computation of \Grobner basis is an {\bf elimination
ordering}; that is, the polynomials contained in the resulting \Grobner basis
have continuously eliminated variables in the ordering. This is the topic of 
elimination theory, which is described in the following section.

\section{Elimination Theory}

Elimination theory uses {\bf elimination ordering} 
to systematically eliminate variables from a system of polynomial
equations.

\begin{Definition}
Let $I$ be an ideal in  $\K[x_1,\dots,x_k]$. The $i$-th 
{\bf elimination ideal} $I_i$ is the ideal of $\K[x_{i+1},\dots,x_k]$ defined
by
\begin{equation}
I_k = I \cap \K[x_{i+1},\dots,x_k]
\end{equation}
\end{Definition}

The elimination ideal $I_i$ has eliminated all the variables 
$x_1,\dots,x_i$, i.e. it only contains polynomials with variables in
$x_{i+1},\dots,x_k$. 
We can generate elimination ideals by computing
\Grobner bases using elimination orderings.

\begin{Theorem}
$\left[\bf{Elimination\  Theorem}\right]$
Let $I$ be an ideal in $\K[x_1,\dots,x_k]$ and let $G$ be the \Grobner 
basis of $I$ with respect to the lex order (elimination order) 
$x_1>x_2>\dots>x_k$. Then, for every $0\leq i\leq k$,
\begin{equation}
G_k=G\cap \K[x_{i+1},\dots,x_k]
\end{equation}
is a \Grobner basis of the $i$-th elimination ideal $I_i$.
\label{thm:elimth}
\end{Theorem}

This can be better visualized using the following example.

\begin{Example}
Given the following equations in $\R[x,y,z]$
\begin{eqnarray}
x^2+y+z&=1 \nonumber \\
x+y^2+z&=1 \nonumber \\
x+y+z^2&=1 \nonumber
\end{eqnarray}
let $I$ be the ideal generated by these equations:
\begin{equation}
I=\langle x^2+y+z-1, x+y^2+z-1, x+y+z^2-1\rangle \nonumber
\end{equation}
The \Grobner basis for $I$ with respect to lex order $x>y>z$ is 
found to be $G=\{g_1,g_2,g_3,g_4\}$ where
\begin{eqnarray}
g_1&=&x+y+z^2-1 \nonumber \\
g_2&=&y^2-y-z^2+z \nonumber \\
g_3&=&2yz^2+z^4-z^2 \nonumber \\
g_4&=&z^6-4z^4+4z^3-z^2 \nonumber
\end{eqnarray}

Notice that while $g_1$ has variables in $\R[x,y,z]$, $g_2$ and $g_3$ only 
have variables in $\R[y,z]$ and $g_4$ only has variables in $\R[z]$. Thus, 
$G_1=G\cap \R[y,z]=\{g_2,g_3,g_4\}$ and $G_2=G\cap \R[z]=\{g_4\}$

Also notice that since $g_4$ only contains variable $z$, and since $g_4=0$, 
a solution for $z$ can be obtained. This solution can then be applied to 
$g_2$ and $g_3$ to obtain solutions for $y$, and so on.
\end{Example}

Elimination theory provides the basis for our abstraction approach.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Hilbert's Nullstellensatz}

In this section, we further describe some correspondence between ideals and 
varieties in the context of algebraic geometry. The celebrated results of 
Hilbert's Nullstellensatz establish these correspondences.

%%%%%%%%%%%%%%%%%algebraically closed field%%%%%%%%%%%%%
\begin{Definition}\label{def:acf}
A field $\overline {\K}$ is an {\bf algebraically closed} field if every  
polynomial in one variable with degree at least $1$, with coefficients 
in $\overline {\K}$, has a root in $\overline {\K}$. 
\end{Definition}
In other words, any non-constant polynomial equation over 
$\overline {\K}\left[x\right]$ always has at least one root 
in $\overline {\K}$. Every field $\K$ is contained in an algebraically 
closed one $\overline {\K}$. 
For example, the field of real numbers $\mathbb{R}$ is not an algebraically closed 
field, because $x^2+1=0$ has no root in $\mathbb{R}$. 
However, $x^2+1=0$ has roots in the field of 
complex numbers $\mathbb{C}$, which is an algebraically closed field. 
In fact, $\mathbb{C}$ is the algebra closure of $\mathbb{R}$. 
Every algebraically closed field is an infinite field.

%%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
%\begin{Theorem}
%$\left[\bf{Weak\  Nullstellensatz}\right]$ Let 
%$I \subset \overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$ 
%be an ideal satisfying $V(I)=\emptyset$. 
%Then $I=\overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
%\begin{equation}
%V(I)=\emptyset\ \iff\ I=\overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
%\end{equation}
%\end{Theorem}
%
%\begin{Corollary}
%	Let $I=\langle f_{1},\dots,f_{s} \rangle \subset \overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$. 
%	Let $G$ be the reduced Gr\"obner basis of $I$. Then $V(I)=0 \iff G=\{1\}$.
%\end{Corollary}

%{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
%system of multivariate polynomial equations (ideal $I$) has common solutions 
%in ${\overline {\mathbb{K}}}^d$. For this purpose, we only need to check if 
%the ideal is generated by the unit element, i.e., $1\in I$. 
%This approach can be used to evaluate the feasibility of constraints in 
%verification problems.
%%%%%%%%%%%%%%%%%%%%%%%%%strong Nullstellensatz%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
An interesting result is one of {\bf Strong 
Nullstellensatz}.
The strong
Nullstellensatz establishes the correspondence between radical ideals
and varieties.

\begin{Theorem}\label{thm:sns}
({\it The Strong Nullstellensatz} \cite{gb_book}) 
Let $\overline{\mathbb{K}}$ be an algebraically closed field, and let $J$
be an ideal in $\overline{\mathbb{K}}[x_1,\dots, x_d]$. 
Then we have $I(V_{\overline{\mathbb{K}}}(J)) =\sqrt{J}$. 
\end{Theorem}

%\subsection{Nullstellensatz over Galois fields}

Strong Nullstellensatz holds a special form over Galois fields $\Fq$.
Recall the notion of vanishing polynomials over Galois fields from the 
previous chapter: for every element $A \in \Fq$, $A-A^q = 0$; then the 
polynomial $x^q-x$ in $\Fq[x]$ vanishes over $\Fq$. Thus, if 
$J_0=\langle x^q-x \rangle$ is the ideal generated by the vanishing 
polynomial, $V(J_0)=\Fq$. Similarly, over $\Fq[x_1,\dots,x_d]$, $J_0$ is 
$\langle x_1^q-x_1,\dots,x_d^q-x_d \rangle$ and $V(J_0)=(F_q)^d$.

\begin{Definition}
Given two ideals, $I_1=\langle f_1, \dots,f_s \rangle$ and 
$I_2=\langle g_1,\dots g_t\rangle$, then the {\bf sum of ideals} 
$I_1+I_2=\langle f_1,\dots,f_s,g_1,\dots g_t\rangle$
\end{Definition}

\begin{Theorem}
({\it Strong Nullstellensatz over $\Fq$})
For any Galois field $\Fq$, let $J \subset \Fq[x_1,\dots,x_d]$ be any ideal 
and let $J_0 = \langle x_1^q-x_1, x_d^q-x_d \rangle$ be the ideal of all
vanishing polynomials. Let $V_{\Fq}(J)$ denote the variety of $J$ over $\Fq$.
Then, $I(V_{\Fq}(J))=J+J_0$.
\end{Theorem}

The proof is given in \cite{gao:gf-gb-ms}. Here, we provide a proof outline.

\begin{Proof}
\begin{enumerate}
\item $\sqrt{J+J_0} = J+J_0$. That is, $J+J_0$ is a radical ideal.
\item $V_{\Fq}(J)=V_{\overline{\Fq}}(J+J_0)$.
\item Due to (2), $I(V_{\Fq}(J)) = I(V_{\overline{\Fq}}(J+J_0))$. 
By Strong Nullstellensatz, this is equivalent to $\sqrt{J+J_0}$.
Finally, due to (1), this is equivalent to $J+J_0$.
\end{enumerate}
\end{Proof}

%% 
%% Using this result, Weak Nullstellensatz can be 
%% modified to be applicable over finite fields $\Fq$.
%% %%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
%% \begin{Theorem}\label{wnull:ff}
%% $[\bf{Weak~Nullstellensatz~in~\Fq}]$\cite{null:1890}\\
%% Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
%% Let $J=\langle f_1,f_2,\cdots,f_s\rangle \subset \Fq[x_1,
%% x_2, \cdots, x_d]$ be an ideal. Let $J_0 = \langle 
%% x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
%% of vanishing polynomials in $\Fq$. Then
%% $V_{\Fq}(J) = V_{\overline {\Fq}}(J +
%% J_0)=\emptyset$,  if and only if the reduced
%% Gr\"obnerBasis$(J+J_{0})=\{1\}$. 
%% \end{Theorem}

%% The proof is given in \cite{null:1890}. Here, we provide a proof outline.

%% \begin{Proof}
%% The variety of $J$ over $\Fq[x_1,x_2,\cdots,x_d]$ 
%% is equivalent to the variety over the algebraic closure of $\Fq$ 
%% intersected by the entire field $\Fq$. That is, $V_{\Fq}(J)=V_{\overline 
%% {\Fq}}(J) \cap \Fq$.

%% Let $J_0 = \langle 
%% x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
%% generated by all vanishing polynomials in $\Fq[x_1,x_2,\cdots,x_d]$.
%% Then $V_{\overline{\Fq}}(J_0)=\Fq$.

%% Thus, $V_{\Fq}(J) = V_{\overline{\Fq}}(J)\cap V_{\overline{\Fq}}(J_0)
%% = V_{\overline{\Fq}}(J+J_0)$.
%% \end{Proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Concluding Remarks}

Our approach to word-level abstraction of Galois field arithmetic 
circuits applies concepts of polynomial ideals, varieties, \Grobner basis, 
and elimination theory to abstract a word-level representation of the 
circuit. This approach is described in the next chapter. However, a \Grobner
basis computation is prohibitively expensive; thus we propose improvements 
to our original approach in a subsequent chapter.

%Once we have a word-level representation, we can apply these representations
%to perform equivalence checking of Galois field arithmetic circuits.
%The verification problem is formulated using Weak
%Nullstellensatz, as it applies over Galois fields, and subsequently solved 
%using a \Grobner basis approach. This approach is described in a future 
%chapter.

### 2016-09-21 dissertation/appendix.tex
init dissertation writing

\numberofappendices=1
\appendix
\chapter{}
\fixchapterheading
\section{Representations of Base Field Elements over Extension Fields}\label{append:Fpk}
Consider a field $\F_q$ and a $k$-bit extension of this field, $\F_{q^k}$.
This extension is created using a primitive, irreducible polynomial $P(x)$ of
degree $k$ over $\F_q[x]$.
Any element $A\in \F_{q^k}$ can be represented as
\begin{equation}
A=a_0+a_1\alpha+\dots+a_{k-1}\alpha^{k-1} \label{eqn:aToAFq}
\end{equation}
where $\{a_0,\dots,a_{k-1}\}\in\F_q$ and $\alpha$ is the primitive element of $\F_{q^k}$, i.e. 
$P(\alpha)=0$. The goal is to derive the polynomial functions $a_i=\Func(A)$ for all
$0\leq i < k$. Note that $\F_q$ could itself be an extension field of a base field $\F_p$ 
where $q=p^l$ for some $l\geq 1$.

Each $a_i$ has the following property:
\begin{equation}
a_i=a_i^{q} \label{eqn:appendixaiq}
\end{equation}

Element $A$ has a similar property:
\begin{equation}
A=A^{q^k} \label{eqn:appendixAqk}
\end{equation}

Examine what happens when Equation \ref{eqn:aToAFq} is raised by the power $q$.
The following lemma is applied.

\begin{Lemma}\label{lemma:raiseToPq}\cite{galois_field:mceliece}
{\it Let $\alpha_1, \dots, \alpha_t$ be any elements in $\F_{p^k}$. Then
\begin{equation}
(\alpha_1+\alpha_2+\cdots+\alpha_t)^{p^i}=\alpha_1^{p^i}+\alpha_2^{p^i}+\dots+\alpha_t^{p^i}
\end{equation}
for all integers $i \geq 0$. }
\end{Lemma}

\begin{Corollary}\label{corollary:raiseqk}
Since $q=p^l$ for some $l \geq 1$, Lemma \ref{lemma:raiseToPq} is applicable to
$\F_{q^k}$.
Let $\alpha_1, \dots, \alpha_t$ be any elements in $\F_{q^k}$. Then
\begin{equation}
(\alpha_1+\alpha_2+\cdots+\alpha_t)^{q^i}=\alpha_1^{q^i}+\alpha_2^{q^i}+\dots+\alpha_t^{q^i}
\end{equation}
for all integers $i \geq 0$.
\end{Corollary}
Applying Corollary \ref{corollary:raiseqk} to Equation \label{eqn:aToFq} gives:
\begin{equation}
A^{q^i}=a_0^{q^i}+a_1^{q^i}\alpha^{q^i}+\dots+a_{k-1}^{q^i}\alpha^{(k-1)q^i} \label{eqn:appendixAqi}
\end{equation}
Then, applying 
Equation \ref{eqn:appendixaiq} to this result gives:
\begin{equation}
A^{q^i}=a_0+a_1\alpha^{q^i}+\dots+a_{k-1}\alpha^{(k-1)q^i} \label{eqn:appendixAqiRewrite}
\end{equation}

In this way, $k$ unique equations are derived, $\{A,A^q,\dots,A^{q^{k-1}}\}$,
along with $k$ unknowns, $\{a_0,\dots,a_{k-1}\}$.
These equations can be represented in matrix form, $\mathbf{A = M a}$, where 
$\mathbf{A}=\{A,A^q,\dots,A^{q^{k-1}}\}^T$, 
$\mathbf{M}$ is
a $k$ by $k$ matrix of coefficients $\in \F_{q^k}$, and $\mathbf{a}=\{a_0,\dots,a_{k-1}\}^T$:
\begin{equation}
\begin{bmatrix}
1      &   \alpha           & \alpha^2           & \dots & \alpha^{k-1}\\
1      &   \alpha^q         & \alpha^{2q}        & \dots & \alpha^{(k-1)q}\\
\vdots & \vdots             & \vdots             & ~     & \vdots \\
1      &   \alpha^{q^{k-1}} & \alpha^{2q^{k-1}} & \dots & \alpha^{(k-1)q^{k-1}}
\end{bmatrix}
\begin{bmatrix}
a_0 \\ a_1 \\ \vdots \\ a_{k-1}
\end{bmatrix}
=
\begin{bmatrix}
A \\ A^q \\ \vdots \\ A^{q^{k-1}}
\end{bmatrix}
\end{equation}

Treat $\mathbf{a}$ as a vector of unknowns, $\mathbf{M}$ and $\mathbf{A}$ as constants. 
This system of equations can be solved using Gaussian 
elimination. However, this system also has a special structure which can
be exploited. $\mathbf{M}$ is a $k$ by $k$ Vandermonde matrix of the form 
$V(\alpha,\alpha^q,\dots,\alpha^{q^{k-1}})$.
Due to the property of Vandermonde matrices:
\begin{equation}
|\mathbf{M}| = \prod\limits_{0\leq i < j < k}(\alpha^{q^j} - \alpha^{q^i})
\end{equation}
Since the elements $\{\alpha,\alpha^q,\dots,\alpha^{q^{k-1}}\}$ are distinct:
\begin{equation}
|\mathbf{M}| \neq 0
\end{equation}
Thus, Cramer's rule can be applied to derive 
an equation for every $a_i$:
\begin{eqnarray}
a_i = \frac{|\mathbf{M_i}|}{|\mathbf{M}|} \label{eqn:cramerformFq}%\\
\end{eqnarray}
where $\mathbf{M_i}$ is $\mathbf{M}$ with the column $\{\alpha^i,\alpha^{iq},\dots,\alpha^{iq^{k-1}}\}^T$ 
replaced by $\mathbf{A}$.
\begin{equation}
\mathbf{M_i} = \begin{bmatrix}
1      &   \alpha            & \dots  & \alpha^{i-1}          & A           & \alpha^{i+1}              & \dots & \alpha^{k-1} \\
1      &   \alpha^q         & \dots  & \alpha^{(i-1)q}       & A^q         & \alpha^{(i+1)q}       & \dots & \alpha^{(k-1)q} \\
\vdots & \vdots             & \cdot  & \vdots                & \vdots      & \vdots                & \cdot & \vdots \\
1      &   \alpha^{q^{k-1}} & \dots  & \alpha^{(i-1)q^{k-1}} & A^{q^(k-1)} & \alpha^{(i+1)q^(k-1)} & \dots & \alpha^{(k-1)q^(k-1)}
\end{bmatrix}
\end{equation}

Computing $|\mathbf{M_i}|$ by interpolating along the $A'$ column gives
\begin{equation}
|\mathbf{M_i}|=\sum\limits_{j=0}^{k-1}(-1)^{(i+j)}A^{q^j}
|V_{i+1}(\alpha,\dots,\alpha^{q^(j-1)},\alpha^{q^(j+1)},\dots,\alpha^{q^(k-1)})|
\end{equation}
where $V_i(x_1,\dots,x_n)$ is the Vandermonde matrix $V(x_1,\dots,x_n)$ with the
$i$-th column skipped and an extra column added to the end.
\begin{equation}
V_i(x_1,\dots,x_n) 
=
\begin{bmatrix}
1 & x_1 & x_1^2  & \dots  & x_1^{i-1} & x_1^{i+1} & \dots & x_1^n \\
1 & x_2 & x_2^2  & \dots  & x_2^{i-1} & x_2^{i+1} & \dots & x_2^n \\
\vdots & \vdots  & \vdots & \cdot  & \vdots    & \vdots    & \cdot & \vdots    \\
1 & x_n & x_n^2  & \dots  & x_n^{i-1} & x_n^{i+1} & \dots & x_n^n
\end{bmatrix} 
\end{equation}
The computation of determinant of $V_i$ is known to the linear algebra community to
be
\begin{equation}
|V_i(x_1,\dots,x_n)| = |V(x_1\dots,x_n)|\cdot S_{n-i}(x_1,\dots,x_n)
\end{equation}
where $S_i(x_1,\dots,x_n)$ is the $i$-th fundamental symmetric polynomial in $\{x_1,\dots,x_n\}$.
Thus, $|\mathbf{M_i}|$ is computed as.
\begin{eqnarray}
|\mathbf{M_i}| &=& \sum\limits_{j=0}^{k-1}[(-1)^{j}A^{q^j} \nonumber \\
& & \cdot |V(\alpha,\dots,\alpha^{q^(j-1)},\alpha^{q^(j+1)},\dots,\alpha^{q^(k-1)})| \nonumber \\
& & \cdot S_{n-1-i}(\alpha,\dots,\alpha^{q^(j-1)},\alpha^{q^(j+1)},\dots,\alpha^{q^(k-1)})] \label{eqn:MireducedFq}
\end{eqnarray}

Lastly, this work has discovered the following proposition for the determinant $|\mathbf{M}|$.
\begin{Proposition}
The determinant $|\mathbf{M}|$ has the property
\begin{equation}
|\mathbf{M}|^q=(-1)^{k-1}|\mathbf{M}|
\end{equation}
\end{Proposition}
\begin{Proof}
\it Since $\mathbf{M}$ is a Vandermonde matrix of the form $V(\alpha,\alpha^q,\dots,\alpha^q{k-1})$, 
then 
\begin{equation}
|\mathbf{M}| = \prod\limits_{0\leq i < j < k}(\alpha^{q^j}-\alpha^{q^i}) \label{proofeqn:detMFq}
\end{equation}
Computing $|\mathbf{M}|^q$ gives
\begin{equation} \label{proof:MsquareFq}
|\mathbf{M}|^q = [\prod\limits_{0\leq i < j < k}(\alpha^{q^j}-\alpha^{q^i})]^q
\end{equation}
Applying Corollary \ref{corollary:raiseqk} to Equation \ref{proof:MsquareFq} gives
\begin{equation}
|\mathbf{M}|^q = \prod\limits_{0\leq i < j < k}(\alpha^{q^{j+1}}-\alpha^{q^{i+1}})
\end{equation}
When $j=k-1$, the product term is in the form $(\alpha^{q^k}-\alpha^{q^{i+1}})$. 
Since $\alpha^{q^k}=\alpha$ over $\F_{q^k}$, this term equivalent to  $-(\alpha^{q^{i+1}}-\alpha)$.
This gives the property:
\begin{equation}
|\mathbf{M}|^q=(-1)^{k-1}|\mathbf{M}|
\end{equation}
\end{Proof}
%Appendix \ref{append:Fpk}

### 2016-09-21 dissertation/applic.tex
init dissertation writing

\chapter{Applications of Word-Level Abstraction of Galois Field 
Circuits}
\label{ch:prop}

In this chapter, we propose two additional applications of word-level 
abstractions
of Galois field circuits. First, we propose an approach to partition
any Galois field arithmetic circuit over $\Fkk$ into sub-circuit  
blocks, each of which can then be analyzed over the same field $\Fkk$.
As each block is a smaller partition of the overall circuit, the 
abstraction complexity of each partition is smaller. Thus, 
word-level abstraction
of the sub-circuits is faster than abstraction of the original circuit.
Abstraction of each sub-circuit can be performed independent of each other,
after which the word-level representations can be combined to find the 
overall function of the entire circuit.
Second, we propose a method for applying our abstraction approach to 
combinational circuits whose function maps one Galois field to 
a different one,
i.e. $f:F_{2^k} \rightarrow F_{2^j}$ where $k \neq j$.

\section{Partitioning Algorithm}
Galois field arithmetic circuits of practical sizes are very large.
As the size of these circuits increase, the computational complexity of a
word-level abstraction increases exponentially.

However, if a large circuit is partitioned, that is 
seperated into smaller seperate pieces, the complexity of each partition is 
smaller and and thus each partition is more manageable.
For word-level abstraction of a circuit, an abstraction of each partition
could be performed and then the word-level representations could be combined
to find the overall word-level representation on the entire circuit.
In addition, each piece
could be abstracted in parallel, thus decreasing the time to
find the overall abstraction.
Recall our abstractions results of the Montgomery multiplier circuits, shown
in Table \ref{tab:absmontflatresults} for a 'flat' abstraction
and Table \ref{tab:absmontresults} for block abstractions. The Montgomery
circuit structure is shown in Figure \ref{fig:montmont}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.50]{figures/mmcircuit}
	\end{center}
	\caption{Montgomery multiplier over $\mathbb{F}_{2^k}$}
	\label{fig:montblock}
\end{figure}

Each block of the multiplier could be considered a partition. By 
comparing our results for the flat multiplier to the results of each 
block, the power of partitioning becomes apparent. Since the block 
abstractions can be performed in parallel, the time for an 
overall word-level abstraction is based on the time of the slowest block
abstraction.

\subsection{Partition Problem Statement}

The goal of our proposed partitioning algorithm 
is to quickly find partitions for any circuit. 
Given a circuit over $\Fkk$, each partition must have $k$-bit inputs and 
one $k$-bit output for simplicity in calculation.
These restrictions allow us to calculate the polynomial representation of 
each partition over the same field $\Fkk$. Thus, our problem statement is
as follows:

\begin{itemize}
\item Given a Galois field $\Fkk$, i.e. given the corresponding
irreducible polynomial $P(x)$ in $\F[x]$ of degree $k$,
let $P(\alpha) = 0$ where $\alpha \in \Fkk$ is the root of the irreducible
polynomial $P$.
\item  Given a gate-level combinational circuit $C$ with $n$ word-level
$k-bit$ inputs,
$A_1,\dots, A_n \in \Fkk$
and one $k-bit$ output $Z \in \Fkk$.
\item The bit-level primary inputs of the circuit are denoted
$\{a_{0}^{i},a_{1}^{i},\dots,a_{k-1}^{i}\}$, for $i=1,\dots, n$;
the primary bit-level outputs are $\{z_0, \dots, z_{k-1}\} = Z$. 
Note that all $a_{i}^{j}, z_i \in \F_2$.
\item Split the circuit into any number of partitions, $\{P_1,\dots,P_d\}$.
\item Each partition $P$ can have any number of $k-bit$ word-level inputs and
one $k$-bit word-level output.
\item The circuit $C$ must be fully represented by the partitions 
$\{P_1,\dots,P_d\}$.
\end{itemize}

After we have the partitions, $\{P_1,\dots,P_d\}$, if we can abstract a 
word-level polynomial of each partition, we can find the overall word-level 
polynomial function of the circuit $C$ using substitution.

\subsection{Partitioning Approach}

{\it Topological location} from Definition \ref{def:topord} which 
is the maximum number of Boolean logic gates that any path from any input 
must pass to reach a certain node.

\begin{Definition}
The {\bf topological depth} of a combinational circuit is maximum 
topological location of the bit-level outputs of the circuit.
\end{Definition}

First, we find the topological depth $D$ of the circuit $C$. 
Then, we split the circuit approximately in half by finding all nodes $M$ of
depth $\frac{D}{2}$. Ultimately, $M$ must contain exactly
one node in every path from any input to any output of the circuit. 
Thus, for every path which does not contain a node in $M$, 
select a node with the next smallest depth and add it to $M$.
These nodes split the circuit approximately in half. One 
half of the circuit treats $M$ as the bit-level outputs and the other treats 
them as the inputs.

\begin{Example}
Consider the 4-bit Mastrovito multiplier circuit shown
previously in Figure \ref{fig:mas4}. We find the maximum topological depth
of this circuit to be $5$. Thus, we find all nodes with topological
depth $2$ and add them to our nodelist $M$. For every path that does not
contain a node with depth $2$, we add the node with depth $1$ from the path
to $M$.

The topological depth of each node of this circuit is shown in 
Figure \ref{fig:nodes}. Every node in $M$ is highlighted. Note that there
is exactly one node in $M$ for every path through the circuit.

\begin{figure}
	\begin{center}
	\includegraphics[scale=0.7]{figures/mul4bitTopo2.eps}
	\end{center}
	\caption{Topological Depth and Node Selection}
	\label{fig:nodes}
\end{figure}
\label{exp:mul4topex}
\end{Example}

Now, we the half of the circuit where the nodes $M$ act as the
bit-level outputs.
Select exactly $k$ nodes in $M$, where $k$ is the size of the field $\Fkk$; 
we call these nodes $M_1$. $M_1$ can be considered a $k$-bit word-level 
output. By including every path from $M_1$ to the primary inputs, we 
create a partition $P_1$. Thus, the $k$-bit word-level output of $P_1$ 
are the nodes $M_1$. The $k-bit$ word-level inputs to $P_1$ are the same 
inputs as to the circuit $C$. Thus, $P_1$ is a partition which can be 
analyzed over $\Fkk$.

With $P_1$ designated, select the next $k$ nodes $M_2\subset M$, where $M_2$
contains none of the same nodes as $M_1$. 
Create another partition, $P_2$, where $M_2$ is the $k-bit$ output of $P_2$.
Continue to do this 
until all nodes $M$ have been used in at least one partition.
For the last partition, it is likely that there will be fewer
than $k$ of nodes remaining in $M$ which have not yet been used as the 
output of some partition. 
At this point, select all of the remaining nodes in $M$ which have not been
used
and add in extra nodes in $M$ that have been previously used in a 
partition in order to bring the total to number $k$. Use this collection
of nodes to create the last partition.
Thus, we allow some nodes in $M$ may be outputs of more than one partition, 
although we try to minimize this.
We now have designated $i$ partitions, $P_1 \cdots P_i$, each with a $k$ 
bit output.

\begin{Example}
Refer to the same 4-bit Mastrovito multiplier circuit from Example 
\ref{exp:mul4topex}. We have already found every node $M$. Since $k=4$,
we need to seperate these nodes into groups of $4$ and create partitions
that span from the nodes $M$ to the bit-level inputs 
$\{a_0,\dots,a_3,b_0,\dots,b_3\}$. As there are $10$ nodes in $M$, this will
create $3$ partitions with some nodes in $M$ being
used as the output of more than one partition.

By incrementally selecting $4$ nodes at a time starting at the top, and
labelling the $4$-bit word-level outputs as $H,G,F$, we create partitions
as shown in Figures \ref{fig:part1}, \ref{fig:part2}, and \ref{fig:part3}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.8]{figures/part1.eps}
	\end{center}
	\caption{First Partition, output $H$}
	\label{fig:part1}
\end{figure}

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.8]{figures/part2.eps}
	\end{center}
	\caption{Second Partition, output $G$}
	\label{fig:part2}
\end{figure}

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.8]{figures/part3.eps}
	\end{center}
	\caption{Third Partition, output $F$}
	\label{fig:part3}
\end{figure}

Note that $f_0=g_2$ and $f_1=g_3$.
\label{exp:mul4inputparts}
\end{Example}

With one half of the circuit partitioned, we look at the other half 
where the nodes in $M$ act as the bit-level inputs. 
The partitions $P_1, \cdots, P_i$ now act as $k$ bit inputs into the second 
half of the circuit. Since there is only one $k$ bit output in the circuit, 
$Z$, we can take the whole second half of the circuit and create one 
partition with $i$ $k$-bit inputs from the outputs of $P_1...P_i$, and one 
$k$ bit output, $Z$. Thus we have partitioned the entire circuit. Since each
partition contains one $k$-bit word-level output and only $k$-bit word-level
inputs, each partition can be analyzed over the same Galois field $\Fkk$ as
the original circuit.

\begin{Example}
Continue the partitioning approach to the 4-bit multiplier
in Example \ref{exp:mul4inputparts}. By using the outputs of the $3$ 
partitions we already have, $\{h_0,\dots,h_1,g_0,\dots,g_3,f_0,\dots,f_3\}$, 
as inputs, we can create the last partition as shown in 
Figure \ref{fig:part4}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.7]{figures/part4.eps}
	\end{center}
	\caption{Fourth Partition}
	\label{fig:part4}
\end{figure}

Thus, the $4$-bit Mastrovito multiplier is fully represented by the four
partitions, as shown in Figure \ref{fig:partitions}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.7]{figures/partitions.eps}
	\end{center}
	\caption{Partitioned Circuit Structure}
	\label{fig:partitions}
\end{figure}
\label{exp:mul4partlast}
\end{Example}

\subsection{Overall Approach}
Algorithm \ref{alg:part} is our overall algorithm for creating partitions of 
Galois field arithmetic circuit over $\Fkk$.

\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{$C = $ circuit over $\Fkk$ with $k$ bit input $A$ and $k$ bit output $Z$}
 \KwOut{$P =$ list of circuit partitions}
  $M = $\xspace\{\}\;
  $P = $\xspace\{\}\;
  $D = $ max topological depth of $C$\;
  \For{ every path $T$ in $C$ }
  {
    \For{ every node $N$ in $T$ in reverse topological order }
    {
      \If{ $N \in M$ }
      {
        break\;
      }
      \If{ $N$.depth $<= \frac{D}{2}$ }
      {
        $M = M \cup N$\;
        break\;
      }
    }
  }
  \For{ every $k$ nodes in $M$ as $m_i$ }
  {
  	$P = P$ \xspace$\cup$ partition from $A$ to $m_i$
  }
  $P = P$ \xspace$\cup$ partition from $M$ to $Z$
\caption {General Partitioning Algorithm}\label{alg:part}
\end{algorithm}

Note that while our algorithm proposes to cut the circuit approximately in 
half by finding all nodes of depth $\frac{D}{2}$, this is merely for 
simplicity's sake. This cut can be shifted by changing the depth at which
to search for these nodes. By changing this depth we change the 
relative sizes of the input and output partitions.

\subsection{Applying Abstraction to Circuit Partitions}

Applying abstraction to partitions is similar to applying abstraction to
Montgomery multiplier blocks. We abstract a word-level polynomial 
representation of each partition in parallel. The polynomial abstraction
of the output partition is
\begin{equation}
f_Z: Z + \Func(G_1,\dots,G_i) \nonumber
\end{equation}
where $Z$ is the word-level output of the entire circuit and 
$\{G_1,\dots,G_i\}$ are the word-level outputs of the input partitions
$\{P_1,\dots,P_i\}$. The word-level abstractions of the input partitions are
\begin{eqnarray}
f_1: G_1 + \Func(A_1,\dots,A_n) \nonumber \\
\vdots \\
f_i: G_i + \Func(A_1,\dots,A_n) \nonumber
\end{eqnarray}
where $\{A_1,\dots,A_n\}$.
We are looking for a polynomial in the form of $Z+\Func(A_1,\dots,A_n)$.
Thus, using lex abstraction order $G_1>\dots>G_i>Z>A_1>\dots>A_n$, if we 
perform the reduction $f_Z\stackrel{f_1,\dots,f_i}{\longrightarrow}_+ r$,
$r$ will contain the word-level polynomial $Z+\Func(A_1,\dots,A_n)$.

There are a few issues, however. Every function over $\Fkk$ is a polynomial
function, but for arbitrary logic these polynomial functions can be quite
large. For instance, in $4$-bit multiplier in Example \ref{exp:mul4partlast},
the polynomial word-level abstraction of the partition with output $F$, 
in the form of $F+\Func(A,B)$, is

\begin{eqnarray}
F+(\alpha^3+1)A^8 B^8+(\alpha^2+\alpha)A^8 B^4+A^8 B^2+(\alpha^3+\alpha+1) A^8 B+(\alpha^3+\alpha+1) A^4 B^8 \nonumber \\
+(\alpha+1) A^4 B^4+(\alpha^2+\alpha) A^4 B^2+A^4 B+(\alpha^3+\alpha+1) A^2 B^8+(\alpha^3) A^2 B^2+(\alpha^3+\alpha^2) A B^8 \nonumber \\
+(\alpha^3+\alpha) A B^4+(\alpha^3+\alpha^2) A B^2+(\alpha^3+\alpha+1) A B \nonumber 
\end{eqnarray}

Thus, as the circuit grows in the size, the polynomial abstraction of each
partition will become quite large. This could make reduction time-consuming.
Also, since every partition is random logic over $\Fkk$, our 
abstraction approach that obviates the complexity of \Grobner basis 
computation will always produce a representation that includes bit-
level input variables. Thus, the simple reduction approach shown above will
not work. Thus, in order to seriously consider abstraction with 
partitioning for circuits of applicable sizes, we need to an abstraction 
solution that obviates \Grobner basis complexity and always computes a word-
level representation only.

\section{Analyzing Circuits Over Two Different Galois Fields}

Thus far, our abstraction technique only worked for circuits that perform
a function $\Func$ over $\Fkk$. That is, a circuit which mapped inputs
over $\Fkk$ to outputs over $\Fkk$, $\Func:\Fkk\rightarrow \Fkk$. This
limits the approach only to circuits with inputs and outputs of equal size 
$k$.

In this section, we present a method for abstracting a representation of a
circuit which maps inputs over a field $\F_{2^k}$ to outputs over another
field $F_{2^j}$, where $k \neq j$.

\subsection{Problem Statement}
A circuit which performs a functional mapping from $\F_{2^k}$ to $\F_{2^n}$
must have $k$-bit word-level inputs and a $j$-bit word-level output. 
Thus, our problem statement is as follows:

\begin{itemize}
\item Given a Galois field $\F_{2^k}$, i.e. given the corresponding
irreducible polynomial $P_k(x)$ in $\F[x]$ of degree $k$,
let $P_k(\beta) = 0$ where $\beta \in \F_{2^k}$ is the root of the 
irreducible polynomial $P_k$.
\item Given another Galois field $\F_{2^j}$, where $j\neq k$, i.e. 
given the corresponding
irreducible polynomial $P_j(x)$ in $\F[x]$ of degree $j$,
let $P_j(\gamma) = 0$ where $\gamma \in \F_{2^j}$ is the root of the 
irreducible polynomial $P_j$.
\item Given another Galois field $\F_{2^h}$, where $h=lcm(j,k)$, i.e. 
given the corresponding
irreducible polynomial $P_h(x)$ in $\F[x]$ of degree $h$,
let $P_h(\alpha) = 0$ where $\alpha \in \F_{2^h}$ is the root of the 
irreducible polynomial $P_h$.
\item Given a gate-level combinational circuit $C$ with
word-level $k$-bit input $A \in \Fkk$, and
and $j$-bit output $Z \in \F_{2^j}$.
\item The bit-level primary input of the circuit is denoted
$\{a_0,\dots,a_{k-1}\} \in \F_2$,
the primary bit-level outputs are $\{z_0, \dots, z_{j-1}\} \in \F_2$.
\item The word-level primary input is denoted 
$A = a_0 + a_1\beta + \dots + a_{k-1}\beta^{k-1}$ and the word-level 
primary output is denoted $Z = z_0 + z_1\gamma + \dots + z_{j-1}\gamma^{j-1}$
\item Find the word-level polynomial function $\Func$ over $\F_{2^h}$
performed by $C$ in the form of $Z=\Func(A_1, \dots, A_n)$.
\end{itemize}

Since the word-level inputs $\{A_1, \dots, A_n\} \in \Fkk$ and the
word-level output $Z \in \F_{2^j}$, the abstraction polynomial 
$Z+\Func(A_1, \dots, A_n)$ must be in $\F_{2^h}$ where 
$\Fkk \subset \F_{2^h}$ and $\F_{2^j} \subset \F_{2^h}$. In other words,
the word-level polynomial is contained in the smallest field that contains 
both the input field $\Fkk$ and output field $\F_{2^j}$. Thus, $h=lcm(k,j)$.

\subsection{Problem Formulation}

Since the abstraction polynomial representation is over the field 
$\F_{2^h}$, we need to model the entire circuit as a system of polynomials 
over this field (as shown in our abstraction chapter).
The only polynomials that cannot be 
immediately modelled over this field are the word-level designation 
polynomials, $f_A$ and $f_Z$.
\begin{eqnarray}
f_A&:&a_0 + a_1\beta + \dots + a_{k-1}\beta^{k-1} \nonumber \\
f_Z&:&z_0 + z_1\gamma + \dots + z_{j-1}\gamma^{j-1} \nonumber
\end{eqnarray}
$f_A$ and $f_Z$ have coefficients $\beta$ and $\gamma$
which are over $\Fkk$ and $\F_{2^j}$ respectively.
Thus, we need to represent $\beta$ and $\gamma$ in terms of 
$\alpha \in \F_{2^h}$.
For this, we
use the following result from \cite{cf:2003} related to 
{\bf composite fields}, which states the following:

\begin{Theorem}
Let $k = m\cdot n$, such that $\Fkk = \F_{(2^m)^n}$. Let $\alpha$ be
the primitive root of $\F_{(2^m)^n}$, and let $\beta$ be the
primitive root of the ground field $\F_{2^m}$. Then,
$\beta=\alpha^{\omega}$, where $\omega=(2 ^{m \cdot n}-1)/(2^m-1)$. 
In other words: 
\begin{equation}
\beta=\alpha^{(2^{m \cdot n}-1)/(2^m-1)} \nonumber
\label{eqn:relation}
\end{equation}
\label{thm:2to3thm}
\end{Theorem}
Since $h=lcm(k,j)$, there exists a $m$ such that $h = k\cdot m$.
Likewise, there exists a $n$ such that $h = j\cdot n$.
Thus, using the above theorem, we can find $\beta = \alpha^\omega$ and 
$\gamma = \alpha^\psi$, where
\begin{eqnarray}
\omega = (2^{k \cdot m}-1)/(2^k-1) \nonumber \\
\psi = (2^{j \cdot n}-1)/(2^j-1) \nonumber 
\end{eqnarray}
Thus, we can write $A$ and $Z$ in terms of $\alpha$ as follows:
\begin{eqnarray}
A = a_0+a_1\cdot\alpha^\omega+\dots+a_{k-1}\cdot\alpha^{(k-1)\omega} \nonumber \\
Z = z_0+z_1\cdot\alpha^\psi+\dots+z_{j-1}\cdot\alpha^{(j-1)\psi} \nonumber
\end{eqnarray}

\begin{Example}
Consider the following circuit shown in Figure \ref{fig:2to3fig}.
\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.80]{figures/2to3.eps}
	\end{center}
	\caption{Circuit with function $\F_{2^3} \rightarrow \F_{2^2}$}
	\label{fig:2to3fig}
\end{figure}

This circuit has a $3$-bit input $A$ and a $2$-bit output $Z$. Thus, it can
be considered to be a functional mapping over two fields 
$\F_{2^3} \rightarrow \F_{2^2}$. Let $\beta$ be the root of the irreducible 
polynomial $P_k$ of degree $3$ which is generates $\F_{2^3}$. 
Let $\gamma$ be the root of
the irreducible polynomial $P_j$ of degree $2$
which generates $\F_{2^2}$. Then the 
word-level designation polynomials of the input and output, $f_A$ and $f_Z$, 
are:
\begin{eqnarray}
f_A&:&A+a_0+a_1\cdot\beta+a_2\cdot\beta^2 \nonumber \\
f_Z&:&Z+z_0+z_1\cdot\gamma \nonumber
\end{eqnarray}

As $lcm(2,3)=6$, the smallest field that contains both 
$\F_{2^3}$ and $\F_{2^2}$ is $\F_{2^6}$. Thus,
the word-level polynomial $Z+\Func(A)$ is a polynomial over $\F_{2^6}$.
Let $\alpha$ be the root of the 
irreducible polynomial $P_h$ of degree $6$ which generates $\F_{2^6}$. 
Using Theorem \ref{thm:2to3thm}, we find $\beta$ and $\gamma$ in terms of 
$\alpha$, as follows: 
\begin{eqnarray}
\beta=\alpha^\omega, \omega={2^{2\cdot3}-1 \over 2^3-1}=9 \nonumber \\
\gamma=\alpha^\psi, \psi={2^{3\cdot2}-1 \over 2^2-1}=21 \nonumber
\end{eqnarray}
thus, $\beta=\alpha^9$ and $\gamma=\alpha^{21}$. 
The polynomials $f_A$ and $f_Z$ are now be represented as:
\begin{eqnarray}
f_A&:&A+a_0+a_1\cdot\alpha^9+a_2\cdot\alpha^{18} \nonumber \\
f_Z&:&Z+z_0+z_1\cdot\alpha^{21} \nonumber
\end{eqnarray}
We can now analyze the entire circuit over $\F_{2^6}$.
\label{exp:2to3ex1}
\end{Example}

\subsection{Abstraction of Circuits over Different Galois Fields}

With all polynomials modelled over $\F_{2^h}$, we can apply our abstraction
approach using abstraction ordering to find the word-level polynomial
function $Z+\F(A)$ over $\F_{2^h}$, as shown in a previous chapter. Note that
vanishing polynomials for the word-level variables, $A$ and $Z$ are formed 
from their corresponding fields, i.e. $A^{2^k}-A$ and $Z^{2^j}-Z$.

\begin{Example}
Consider the same circuit from Example \ref{exp:2to3ex1}, which performs a
function $\F_{2^3} \rightarrow \F_{2^2}$.
Let $\alpha$ be the root of the irreducible polynomial $P_h=X^6+X+1$, which
is used to construct $\F_{2^6}$.
We model the entire circuit as a system of polynomials $J$ over $\F_{2^6}$.
We find the relevant vanishing polynomials $J_0$. Abstraction ordering of 
this circuit is
\begin{equation}
z_0>z_1>s_0>s_1>s_2>a_0>a_1>a_2>Z>A \nonumber
\end{equation}
The system of polynomials $J+J_0$ is

\begin{eqnarray}\label{eqn:miterbit}
 \left .  \begin{aligned}
f_1&:&s_0+a_0\cdot a_1\nonumber\\
f_2&:&s_1+a_1\cdot a_2\nonumber\\
f_3&:&s_2+a_0\cdot a_2\nonumber\\
f_4&:&z_0+s_0+s_1\nonumber\\
f_5&:&z_1+s_1+s_2\nonumber\\
\end{aligned}
\ \right\}
&\qquad&  \text{\it Bit-level implementation} (J) \nonumber \\
 \left .  \begin{aligned}
f_A&:&a_0+a_1\cdot\alpha^9+a_2\cdot\alpha^{18}+A\nonumber\\
f_Z&:&z_0+z_1\cdot\alpha^{21}+Z\nonumber\\ 
\end{aligned}
\ \right\}
 &\qquad&  \text{\it Word-level designation} (J) \nonumber \\
\left . \begin{aligned}
a_0^2-a_0\nonumber\\
a_1^2-a_1\nonumber\\
a_2^2-a_2\nonumber\\
z_0^2-z_0\nonumber\\
z_1^2-z_1\nonumber\\
s_0^2-s_0\nonumber\\
s_1^2-s_1\nonumber\\
s_2^2-s_2\nonumber\\
A^8-A\nonumber\\
Z^4-Z\nonumber
\end{aligned}
\ \right\}
&\qquad&  \text{\it Vanishing Polynomials} (J_0) \nonumber 
\end{eqnarray}

By computing a reduced \Grobner basis, the word-level abstraction polynomial
in the form of $Z+\Func(A)$ is found in the basis:
\begin{equation}
Z+(\alpha^2+\alpha)A^6+(\alpha^4+\alpha^3+\alpha)A^5+(\alpha^2+\alpha)A^4+(\alpha^4+\alpha^3+\alpha^2)A^3+(\alpha^4+\alpha^3+\alpha^2)A^2+(\alpha^4+\alpha^3+\alpha)A \nonumber
\end{equation}
\end{Example}

This method allows us to abstract a word-level polynomial representation
of circuits which have inputs and outputs of different datapath sizes, where
previously abstraction was only possible if the datapath of the inputs and
outputs was the same.
However, if this circuit with function $\Fkk \rightarrow \F_{2^j}$ does not 
perform a
simple function over the give $\F_{2^h}$, like in the example above, then 
the word-level polynomial representation over $Z+\Func(A)$ over $\F_{2^h}$ 
could be quite large. In this case, our approach to obviate the complexity
of the \Grobner basis computation will derive a representation with bit-level
input variables.
Also, depending on the size of the inputs, $k$, and the 
size of the outputs, $j$, the field $\F_{2^h}$ could be very large if $j$ and
$k$ are relatively prime to each other, i.e. $lcm(j,k)=j*k$.

\section{Conclusions}

In this chapter we proposed two more applications of word-level abstraction 
over Galois Fields: a partitioning algorithm over $\Fkk$ and an approach to
analyze circuits over $\Fkk \rightarrow \F_{2^j}$.
Notably, these proposals are applicable not just to 
abstraction, but to all computer-algebra based formal verification of Galois 
field circuits. Currently, there are a few limitations of these proposals 
when applied to our abstraction approach. When applied to partitions, our 
abstraction approach with obviation of \Grobner basis computation produces a 
result with bit-level input variables. Thus, we cannot apply substitution to
abstract the full representation of the partitioned circuit. However, 
we could potentially apply these mixed-level abstractions to Lv's
equivalence approach, as described in the previous chapter.

There are a number of unexplored applications of the proposals put forth in 
this chapter. If our abstraction approach could guarantee a word-level 
abstraction of any circuit logic, then the partitioning approach could 
very-well allows us to abstract even larger Galois field circuits. Our 
proposal of abstracting representations of circuits over 
$\Fkk \rightarrow \F_{2^j}$ could also then be applied to allow for 
partitions of varying inputs and outputs. These applications seem 
promising, and thus motivate further research.

### 2016-09-21 dissertation/comp.tex
init dissertation writing

\chapter{Generalizing the Approach to Arbitrary Combinational Circuits}\label{ch:generalize}
The abstraction approach presented in the previous chapter is
directly applicable only when the word-size of the operands 
of the given circuit are the same, $k$ bits in size. 
In this case, the circuit computes a function over $\F^n_{2^k} \rightarrow \Fkk$, 
where $n$ is the number of word-level inputs,
and the abstraction is thus analyzed over $\Fkk$. 
When the input and output sizes vary, the analysis must be performed 
over an overarching field.
This chapter shows how to suitably modify the approach for
abstracting word-level representations of circuits with varying input and output sizes.

As an application of this generalization, design and verification methodologies 
for composite field multipliers over $\Fkk$ 
are also explored.
These designs decompose the field $\Fkk$ to $\F_{(2^m)^n}$, where 
$k=m\cdot n$. 
Internally, the circuit is then composed of an $n$-interconnection of 
$m$-bit multipliers and adders over $\F_{2^m}$.
This chapter describes how this hierarchy can be exploited by the abstraction
approach.

\section{Circuits with Varying Input and Output Sizes}
When the word size of the inputs and output of the circuits vary, the 
functionality of the circuit must be analyzed over an encompassing composite field. 
Given a circuit with a word-level input $A$ and word-level output $Z$,
let $t$ be the bit size of the $A$ and $u$ be the bit size of $Z$.
Input $A$ can be represented as an element over the field $\F_{2^{t}}$; likewise,
$Z$ is an element of $\F_{2^u}$.
Thus, this circuit computes some function $f: \F_{2^t} \rightarrow \F_{2^u}$.
If $t \neq u$, there is no guarantee that $A$ can be represented over $\F_{2^u}$
and or conversely $Z$ over $\F_{2^t}$. Thus, the analysis must be performed over some 
$\Fkk$ such that $\F_{2^t} \subset \Fkk$ and $\F_{2^u} \subset \Fkk$. 
That is, the function is mapped to a larger field $\Fkk \rightarrow \Fkk$.
The smallest such field $\Fkk$ is constructed where $k = LCM(t,u)$.

Given the primitive polynomial $P(x)$ of degree $k$, let $\alpha$ be the primitive 
element of $\Fkk$, i.e. $P(\alpha)=0$.
Let $\beta$ be some primitive
element of $\F_{2^t}$. Then the polynomial $f_A$ is denoted 
\begin{equation}
f_A: a_0+a_1\beta+\dots+a_{t-1}\beta^{t-1}+A 
\end{equation}
Since the analysis is over $\Fkk$, $\beta$ must be mapped
directly to $\alpha$. Here, a generalized result from \cite{cf:2003} gives the relation.
%applying Equation \label{eqn:relation} gives the relation.

\begin{equation}
\beta=\alpha^{(2^k-1)/(2^t-1)} \label{eqn:betaToAlpha}
\end{equation}

All powers of $\beta$ in $f_A$ are replaced by their representation in $\alpha$.
Similarly, let $\gamma$ be the primitive element of $\F_{2^u}$, such that
\begin{equation}
f_Z: z_0+z_1\gamma+\dots+z_{u-1}\gamma^{u-1}+Z
\end{equation}
A mapping from $\gamma$ to $\alpha$ is found in the same way and applied to $f_Z$.
All polynomials in the ideal $J$ now have coefficients in $\Fkk$. The polynomials in 
$J_0$ derived from $A$ and from $Z$ are modified to reflect their corresponding fields:
\begin{eqnarray}
A^{2^t}+A=0 & & Z^{2^u}+Z=0
\end{eqnarray}
Now the reduction procedure $f_Z\xrightarrow{F-{f_z},F_0}_+ r$ is 
computed normally over $\Fkk$.

%Some changes still need to be made to
%functionally map each $\{a_0,\dots,a_{t-1}\}$ to $A$ over $\Fkk$. This is 
%accomplished by first setting up the bit-level to word-level mapping over $\F_{2^t}$ using
%the same approach described in Chapter \ref{ch:improv}. That is, construct the matrix
%\mathbf{M}, which is composed of variable $A$ and coefficients in $\beta$.
%\begin{equation}
%\mathbf{M} = \begin{bmatrix}
%1      &   \beta          & \beta^2           & \dots & \beta^{t-1}\\
%1      &   \beta^2        & \beta^{4}        & \dots & \beta^{(t-1)*2}\\
%1      &   \beta^4        & \beta^{8}        & \dots & \beta^{(t-1)*4}\\
%\vdots & \vdots           & \vdots        & \cdot     & \vdots \\
%1      &   \beta^{2^{t-1}}   & \beta^{2*2^{t-1}} & \dots & \beta^{(t-1)*2^{t-1}}
%\end{bmatrix}
%\end{equation}
%Then,
%map each $\beta$ coefficient in $\mathbf{M}$ to $\alpha$ using 
%Equation \ref{eqn:betaToAlpha}. Since $\beta$ is a primitive element of $\F_{2^t}$,
%each $\{\beta,\beta^2,\dots\beta^{2^{t-1}}\}$ is a distinct element. It follows that,
%once mapped to $\alpha$, each element stays distinct since $\alpha$ is a primitive 
%element of a larger field. Thus, $\mathbf{|M|}\neq 0$, Cramer's rule can be applied, and
%the rest of the approach proceeds as before
%to derive polynomials the $f_{a_0},\dots,f_{a_{t-1}}\}$ over $\Fkk$.
%\begin{eqnarray}
%&f_{a_0} : a_0 + & \nonumber \\
%&\vdots& \nonumber \\
%&f_{a_{k-1}} : a_{k-1} + \Func_{a_{k-1}}(A)& \nonumber
%\end{eqnarray}

Some changes still need to be made to
functionally map each $\{a_0,\dots,a_{t-1}\}$ to $A$ over $\Fkk$. 
This is accomplished by first computing the bit-level to word-level mapping over 
$\F_{2^t}$ using the same approach described in Chapter \ref{ch:improv}, which will provide
the polynomials $f_{a_0},\dots,f_{a_t}$ with coefficients in $\beta$. Then, map each $\beta$ 
coefficient in $F_a = \{f_{a_0},\dots,f_{a_t}\}$ to $\alpha$ using Eqn.(\ref{eqn:betaToAlpha}). 
Polynomials of $F_a$ are now in $\Fkk$ and
the final substitution 
$r\xrightarrow{F_a+J_0}_+r_w$ is computed as normal.

This allows the word-level abstraction of any combinational circuit with one
word-level input of any size and one word-level output. In the case of multiple
word-level inputs, let $k$ be the $LCM$ of all the bit-sizes of the inputs and
outputs. Then the abstraction proceeds as normal.

\begin{Example}
\begin{figure}[!hbt]
\centerline{
\includegraphics[scale=0.4]{./figures/3to2.eps}
}
\caption{ Circuit with varying word sizes.}
\label{fig:3to2}
\end{figure}
{\it 
Consider the circuit shown in Fig. \ref{fig:3to2}. The input, $A$, is 3 bits 
wide while the output, $Z$ is 2 bits. Thus, $A\in {\mathbb{F}}_{2^3}$ and 
$Z \in {\mathbb{F}}_{2^2}$. Let $\beta$ be the primitive element of 
${\mathbb{F}}_{2^3}$ and $\gamma$ be the primitive element of 
${\mathbb{F}}_{2^2}$, i.e. $A=a_0+a_1\beta+a_2\beta^2$ and $Z=z_0+z_1\gamma$.
Thus, the circuit computes the function ${\mathbb{F}}_{2^3}\rightarrow{\mathbb{F}}_{2^2}$. Since $LCM(2,3)=6$, then $\F_{2^2} \subset \F_{2^6}$ and $\F_{2^3} \subset
\F_{2^6}$. So this function mapped to $\F_{2^6} \rightarrow \F_{2^6}$.
Choose $P(X)=X^6+X+1$ as the irreducible
polynomial to construct $\F_{2^6}$, where $P(\alpha)=0$. Then $\beta$ and
$\gamma$ in can be represented in terms of $\alpha$:
\begin{eqnarray}
\beta=\alpha^{(2^6-1)/(2^3-1)}=\alpha^9 \nonumber \\
\gamma=\alpha^{(2^6-1)/(2^2-1)}=\alpha^{21} 
\end{eqnarray}
So the word-level polynomials are:
\begin{eqnarray}
&&f_A: a_0+a_1\alpha^9+a_2\alpha^{18}+A \nonumber \\
&&f_Z: z_0+z_1\alpha^{21}+Z 
\end{eqnarray}
The rest of the polynomials in $F$ are derived from the circuit:
\begin{eqnarray}
f_1: z_0+s_0+s_1 & f_2: z_1+s_1+s_2 & f_3: s_0+a_0\cdot a_1 \nonumber \\
f_4: s_1+a_1\cdot a_2 & f_5: s_2+a_0\cdot a_2 &
\end{eqnarray}
where the RATO ordering is:
\begin{equation}
z_0 > z_1 > s_0 > s_1 > s_2 > a_0 > a_1 > a_2 > Z > A
\end{equation}
$F_0$ is defined as before, except with a change to the polynomials derived from word-level 
variables $A$ and $Z$:
\begin{eqnarray}
f_6: z_0^2+z_0 & f_7: z_1^2+z_1 & f_8: s_0^2+s_0 \nonumber \\
f_9: s_1^2+s_1 & f_{10}: s_2^2+s_2 & f_{11}: a_0^2+a_0 \nonumber \\
f_{12}: a_1^2+a_2 & f_{13}: a_2^2+a_2 & f_{14}: Z^4+Z \nonumber \\
f_{15}: A^8+A & & 
\end{eqnarray}

Computing $f_Z\xrightarrow{F-{f_z},F_0}_+ r$ gives:
\begin{eqnarray}
r&=& (\alpha^2+\alpha)\cdot a_1\cdot a_2+a_1\cdot A+(\alpha^4+\alpha^3)\cdot a_1 \nonumber \\
&&+(\alpha^5+\alpha^4+\alpha^3+\alpha+1)\cdot a_2\cdot A+(\alpha^5+\alpha^4+\alpha^2+\alpha)\cdot a_2+Z
\end{eqnarray}

This result must be further reduced by $f_{a_1}: a_1 + \Func_{a_1}(A)$ and $f_{a_2}: a_2 + \Func-{a_2}(A)$. 
First, find $\mathbf{M}$ in terms of $\beta$ as derived from $A=a_0+a_1\beta+a_2\beta^2$.
Then map it to $\alpha$.

\begin{eqnarray}
\mathbf{M}=
\begin{bmatrix}
1      &   \beta   & \beta^2  \\
1      &   \beta^2 & \beta^4  \\
1      &   \beta^4 & \beta^8
\end{bmatrix} & 
\mathbf{M_1}=
\begin{bmatrix}
1      &   A   & \beta^2  \\
1      &   A^2 & \beta^4  \\
1      &   A^4 & \beta^8
\end{bmatrix} &
\mathbf{M_2}=
\begin{bmatrix}
1      &   \beta   & A  \\
1      &   \beta^2 & A^2  \\
1      &   \beta^4 & A^4
\end{bmatrix}
\end{eqnarray}

Computing the determinants in $f_{a_1}: a_1 + |\mathbf{M_1}|$ and $f_{a_2}: a_2 + |\mathbf{M_2}|$ gives:
\begin{eqnarray}
f_{a_1}: a_1+A^4\cdot (\beta^4+\beta^2)+A^2\cdot (\beta^8+\beta^2)+A\cdot (\beta^8+\beta^4) \nonumber \\
f_{a_2}: a_2+A^4\cdot (\beta^2+\beta)+A^2\cdot (\beta^4+\beta)+A\cdot (\beta^4+\beta^2)
\end{eqnarray}
Replacing all $\beta$ in $f_{a_1}$ and $f_{a_2}$ with $\alpha^9$ gives their proper form over $\F_{2^6}$.
\begin{eqnarray}
f_{a_1}: a_1+A^4\cdot(\alpha^4+\alpha^3+1)+A^2\cdot(\alpha^4+\alpha^2+\alpha+1)+A\cdot(\alpha^3+\alpha^2+\alpha) \nonumber \\
f_{a_2}: a_2+A^4\cdot(\alpha^4+\alpha^2+\alpha+1)+A^2\cdot(\alpha^3+\alpha^2+\alpha)+A\cdot(\alpha^4+\alpha^3+1)
\end{eqnarray}

Finally, computing the reduction $r\xrightarrow{f_{a_1},f_{a_2},F_0}_+ r_w$ gives the word-level polynomial
abstraction of the circuit.
%Applying the abstraction approach, the word-level polynomial $Z+\F(A)$ found is

\begin{eqnarray}
r_w:&&Z+A^6(\alpha^2+\alpha)+A^5(\alpha^4+\alpha^3+\alpha)+A^4(\alpha^2+\alpha) \nonumber \\
&&+A^3(\alpha^4+\alpha^3+\alpha^2)+A^2(\alpha^4+\alpha^3+\alpha^2)+A(\alpha^4+\alpha^3+\alpha) \nonumber
\end{eqnarray}
}
\end{Example}

This result allows the abstraction to be computed over fields of different sizes. 
An important application of this result is that of modularly-designed composite 
field arithmetic circuits.
These circuits compute operations over very large fields (i.e. $k=1024$) by combining 
operations over smaller sub-fields (i.e. $k=32$).
The abstraction approach can be efficiently applied to these types of circuits by
exploiting the hierarchy found in these designs.

\section{Composite Field Arithmetic Circuits}

A Galois field multiplier over $\Fkk$ can be composed over the composite field $\F_{(2^m)^n}$ 
where $k=m\cdot n$ \cite{phdpaar:1994}. Similarly to how $\Fkk$ is a $k$-dimensional 
extension of the subfield $\F_2$, $\F_{(2^m)^n}$ is also an $n$-dimensional extension of $\F_{2^m}$.
A {\bf composite field multiplier} lifts the ground field from $\F_2$ to $\F_{2^m}$ and
computes the multiplication over $\Fkk$ as a collection of operations
over $\F_{2^m}$. Thus, a composite field multiplier over $\Fkk$ is composed internally as a
collection of {\it multipliers} and {\it adders} over $\F_{2^m}$.

This hierarchy found in composite field multipliers can be exploited by the proposed 
abstraction approach. Abstraction of these types of multipliers is composed of
two steps:
\begin{enumerate}
\item Compute the canonical word-level polynomial representation of each $\F_{2^m}$ multiplier and adder.
\begin{itemize}
  \item These abstractions are independent of one another. Thus, they are computed in parallel.
  \item In the case of an adder, the abstraction is trivial.
\end{itemize}
\item Compute the overall abstraction of the $\Fkk$ multiplier.
\end{enumerate}
The first step utilizes the proposed abstraction approach over $\F_{2^m}$ with no changes.
Once these word-level abstractions are known, they replace the gate-level
implementations for the final word-level abstraction of the multiplier over $\Fkk$.
%Design of composite field multipliers is explored in Chapter \label{ch:prelim}.

\subsection{Design of Composite Field Multipliers}

The following adapts principles of composite fields from \cite{phdpaar:1994} and 
explains how they are applied to construct Galois field multipliers.
Consider the element $A \in \Fkk$ and its representation over 
$\F_{(2^m)^n}$. Let $\alpha$ be the primitive element of $\Fkk$ and let $\gamma$
be the primitive element of $\F_{(2^m)^n}$. Then any element $A \in \mathbb{F}_{2^k}$ 
is represented as:
\begin{equation}
A=a_0+a_1\alpha+\dots+a_{k-1}\alpha^{k-1},\text{where } a_i  \in \mathbb{F}_{2} \label{eqn:compAf2k}
\end{equation}
This same element $A \in \mathbb{F}_{(2^m)^n}$ is represented as:
\begin{equation}
A=A_0+A_1\gamma+\dots+A_{n-1}\gamma^{n-1},\text{where } A_i \in \mathbb{F}_{2^m} \label{eqn:compAf2m}
\end{equation}
Let $\beta$ be the primitive element of $\F_{2^m}$. Then each $A_i$ is represented as
\begin{equation}
A_i=a_{i0}+a_{i1}\beta+\dots+a_{i\{m-1\}}\beta^{m-1},\text{where } a_{ij}\in\F_2
\end{equation}

Since there always exists a unique field with $p^k$ elements, the field $\Fkk$ is isomorphic to 
the field $\F_{(2^m)^n}$. Due to this, $\gamma = \alpha$. Furthermore, $\beta$ can be derived 
from $\alpha$ using Eqn.(\ref{eqn:betaToAlpha}) since $\F_{2^m} \subset \Fkk$, i.e. 
$\beta=\alpha^w$ for some $w$. Thus, all
that is required to construct a composite field multiplier over $\F_{(2^m)^n}$ is 
the primitive polynomial $P(x)$ which generates $\Fkk$, with $P(\alpha)=0$, 
where $\beta$ is known.

In order to lift the ground field $\F_2$ to $\F_{2^m}$, the variables $\{a_{00},\dots,a_{\{n-1\}\{m-1\}}\}$
must be derived in terms of $\{a_0,\dots,a_{k-1}\}$. Equating the representation of $A\in\Fkk$
with the representation of $A\in\F_{(2^m)^n}$ from Eqns.(\ref{eqn:compAf2k}) and (\ref{eqn:compAf2m}) 
gives the following:
\begin{eqnarray}
& &a_0+a_1\alpha+\dots+a_{k-1}\alpha^{k-1} \nonumber \\
&=&A_0+A_1\alpha+\dots+A_{n-1}\alpha^{n-1} \nonumber \\
&=&\sum_{i=0}^{i=n-1}(\sum_{j=0}^{j=m-1}a_{ij} \cdot \alpha^{wj}) \cdot \alpha^i
\end{eqnarray}
%
Here, every $A_0,\dots,A_{n-1}$ is replaced by its representation over $\F_{2^m}$. Analyzing
the coefficients
gives $\{a_0,\dots,a_{k-1}\}$ in terms of $\{a_{00},\dots,a_{\{n-1\}\{m-1\}}\}$. This mapping can be
depicted as a matrix multiplication.

\begin{equation}
\begin{bmatrix} a_0\\ \vdots \\ a_{k-1}\end{bmatrix}
=\mathbf{T}
\begin{bmatrix} a_{00}\\ \vdots \\ a_{\{n-1\}\{m-1\}}\end{bmatrix}
\end{equation}
where $\mathbf{T}$ is a $k$ by $k$ matrix consisting of elements in $\F_2$. Inverting 
$\mathbf{T}$ gives a mapping from $\{a_{00},\dots,a_{\{n-1\}\{m-1\}}\}$ to $\{a_0,\dots,a_{k-1}\}$.

\begin{Example}\label{ex:comp22}
An example composite field multiplier $\F_{(2^2)^2}$, which computes a multiplication
over $\F_{2^4}$, is shown in Figure \ref{fig:comp4ex}. Notice that, after the transformation,
all additions and multiplications are computed over the base field $\F_{2^2}$.

\begin{figure}[t]
        \centering
        \includegraphics[width=.9\linewidth]{./figures/compMineSmall}
        \caption{$4$-bit composite multiplier designed over $\F_{(2^2)^2}$}\label{fig:comp4ex}
\end{figure}

Let $P(x) = x^4 + x^3 + 1$ and $P(\alpha)=0$. 
Representation of element $A \in \mathbb{F}_{(2^2)^2}$ is:
\begin{eqnarray}
A& = A_0 +A_1 \cdot \alpha
\end{eqnarray}
Representation of $A_0, A_1$ in $\mathbb{F}_{2^m}$ is:
\begin{eqnarray}
A_0=a_{00}+a_{01} \cdot \beta \nonumber \\
A_1=a_{10}+a_{11} \cdot \beta
\end{eqnarray}
where $a_{ij}\in \mathbb{F}_2$. From Eqn.(\ref{eqn:betaToAlpha}), $\beta=\alpha^5$
Now $A_0, A_1$ can be substituted into $A$ as follows:
\begin{eqnarray}
A&=&a_{00}+a_{01}\cdot \alpha^5+(a_{10}+a_{11}\cdot \alpha^5)\cdot \alpha
\end{eqnarray}
Since $P(x)=x^4+x^3+1$ with $P(\alpha)=0$,
\begin{equation}\label{a}
A \pmod {P(\alpha)}=a_{00}+a_{01}+a_{11}+(a_{01}+a_{10}+a_{11})\cdot \alpha+a_{1
1} \cdot \alpha^2+(a_{01}+a_{11})\cdot \alpha^3   
\end{equation}
The same element $A \in \mathbb{F}_{2^4}$ is represented as:
\begin{equation}\label{aa}
A=a_0+a_1\cdot \alpha+a_2\cdot \alpha^2+a_3\cdot \alpha^3 
\end{equation}

\item Since Eqns. (\ref{a}) and (\ref{aa}) represent the same element, we
  can match the coefficients of the the polynomials to obtain:
\begin{eqnarray}
a_0&=&a_{00}+a_{01}+a_{11} \nonumber \\
a_1&=&a_{01}+a_{10}+a_{11} \nonumber \\
a_2&=&a_{11} \nonumber \\
a_3&=&a_{01}+a_{11} \nonumber
\end{eqnarray}

This mapping can also be reversed and represented as a matrix $T^{-1}$:
\begin{equation}
\begin{bmatrix} a_{00}\\ a_{01} \\a_{10} \\ a_{11}\end{bmatrix}
=
\begin{bmatrix} 1 & 0 & 0 & 1\\ 0 & 0 & 1 & 1\\ 0 & 1 & 0 & 1\\ 0 & 0
  & 1 & 0 \end{bmatrix} 
\begin{bmatrix} a_0\\ a_1 \\a_2 \\ a_3\end{bmatrix}
\end{equation}
Thus, $A$ is represented in $\mathbb{F}_{(2^2)^2}$ as:
\begin{eqnarray}
A&=&A_0+A_1 \cdot \alpha \nonumber \\
A_0&=&a_{00}+a_{01} \cdot \alpha^5 \nonumber \\
A_1&=&a_{10}+a_{11} \cdot \alpha^5 \nonumber \\
a_{00}&=&a_0+a_3 \nonumber \\
a_{01}&=&a_2+a_3 \nonumber \\
a_{10}&=&a_1+a_3 \nonumber \\
a_{11}&=&a_2 \nonumber 
\end{eqnarray}
$B$ is similarly represented in $\mathbb{F}_{(2^2)^2}$.
\end{Example}

\subsection{Abstraction of Composite Field Multipliers}

The first step in abstracting the word-level polynomial representation of the composite field
multiplier is to abstract every internal $\F_{2^m}$ computational block. In the case of a $\F_{2^m}$
adder block ($\{Z_0,Z_1,C_4\}$ from Fig.(\ref{fig:comp4ex})), this abstraction is trivial, as the adder is just a bit-wise XOR computation. In the case
of a multiplier block ($\{C_0,C_1,C_2,C_3,C_5,C_6\}$ from Fig.(\ref{fig:comp4ex})), 
the abstraction approach presented in the previous chapter is directly 
applicable. Each abstraction can be computed independently, so these computations can be 
performed in parallel. This set of abstracted polynomials, $F$, generates the ideal $J$.

Once the word-level abstraction of each $\F_{2^m}$ sub-block is known, the final abstraction of the 
entire design can be computed using only word-level variables. Set a RATO ordering. Then
reduce the word level polynomial by $F+F_0$:
\begin{equation}
f_Z: Z_0+Z_1\alpha+\dots+Z_{n-1}\alpha^{n-1}+Z
\end{equation}
\begin{equation}
f_Z\xrightarrow{F+F_0}_+ r
\end{equation}
Here, $F$ is the set of word-level
polynomial abstractions from each $\F_{2^m}$ block and $F_0$ is the set of vanishing polynomials.
Every variable $X$, apart from the word-level inputs $A$ and $B$ and word-level output $Z$, is an element of
$\F_{2^m}$, so its corresponding vanishing polynomial is $X^{2^m}+X$. Computing the reduction
gives remainder $r$ containing the elements 
\begin{equation}
A_0,A_1,\dots,A_{n-1},B_0,B_1,\dots,B_{n-1},Z
\end{equation}
Similarly, the mapping from $\{A_0,\dots,A_{n-1}\}$ to $A$ now needs to be derived (along with
the mapping from $\{B_0,\dots,B_{n-1}\}$ to $B$). That is, the next step is to find 
\begin{eqnarray}
& &F_A=\{F_{A_0},F_{A_1},\dots,F_{A_{n-1}}\} \nonumber \\
\text{where }& &F_{A_i}=A_i+\Func_{A_i}(A)
\end{eqnarray}
$F_A$ is derived from the polynomial
\begin{equation}
A=A_0+A_1\alpha+\dots+A_{n-1}\alpha^{n-1}
\end{equation}
Compute $A^{2^m}$ as follows:
\begin{eqnarray}
A^{2^m}: & & (A_0+A_1\alpha+\dots+A_{n-1}\alpha^{n-1})^{2^m} \nonumber \\
&&=A_0^{2^m}+A_1^{2^m}\alpha^{2^m}+\dots+A_{n-1}^{2^m}\alpha^{2^m(n-1)} \nonumber \\
&&=A_0+A_1\alpha^{2^m}+\dots+A_{n-1}\alpha^{2^m(n-1)}
\end{eqnarray}
Here, the property is $A_i^{2^m}=A_i$ is exploited. Continually raise this result by $2^m$ 
to obtain $A^{2^{jm}}$ for all $0\leq j < n$. This gives a system of $n$ equations and $n$
unknowns $\{A_0,\dots,A_{n-1}\}$. As before, this system of equations can be represented
in matrix form, $\mathbf{A = M a}$, where 
$\mathbf{A}=\{A,A^{2^m},\dots,A^{2^{(n-1)m}}\}^T$, 
$\mathbf{M}$ is
a $n$ by $n$ matrix of coefficients $\in \Fkk$, and $\mathbf{a}=\{A_0,\dots,A_{n-1}\}^T$:

\begin{eqnarray}
\begin{bmatrix}
A \\ A^{2^m} \\ \vdots \\ A^{2^{(n-1)m}}
\end{bmatrix}  &=&
\begin{bmatrix}
1      &   \alpha          & \alpha^2         & \dots & \alpha^{n-1}\\
1      &   \alpha^{2^m}     & \alpha^{2\cdot2^m}    & \dots & \alpha^{(n-1)\cdot 2^m}\\
1      &   \alpha^{2^{2m}}   & \alpha^{2\cdot2^{2m}}    & \dots & \alpha^{(n-1)\cdot 2^{2m}}\\
\vdots & \vdots            & \vdots           & \cdot & \vdots \\
1      &   \alpha^{2^{(n-1)m}} & \alpha^{2\cdot2^{(n-1)m}} & \dots & \alpha^{(n-1)\cdot 2^{(n-1)m}}
\end{bmatrix}
\begin{bmatrix}
A_0 \\ A_1 \\ \vdots \\ A_{n-1}
\end{bmatrix} \label{eqn:matrixForm}
\end{eqnarray}

The matrix $\mathbf{M}$ is the Vandermonde matrix $V(\alpha,\alpha^{2^m},\dots,\alpha^{2^{(n-1)m}})$.
Since $\alpha$ is a primitive element, 
all elements $\alpha,\dots,\alpha^{2^{k-1}}$ are unique, where $k=m\cdot n$. Thus, 
all elements $\alpha,\alpha^{2^m},\dots,\alpha^{2^{(n-1)m}}$ are unique, so
$|\mathbf{M}| \neq 0$. Then, Cramer's rule can be applied to derive each $F_{A_i}$ as 
\begin{equation}
F_{A_i}=A_i+\frac{|\mathbf{M_i}|}{|\mathbf{M}|}
\end{equation}
where $\mathbf{M_i}$ is $\mathbf{M}$ with the $i$-th column replaced
by $\mathbf{A}$. Here, $|\mathbf{M}|$ is not guaranteed to be equal to $1$,  
%Rather, $|\mathbf{M}|$ is a non-zero element of $\F_{2^m}$ (proof in Appendix \ref{append:Fpk}),
so it must be computed. 
$F_B=\{F_{B_0},\dots,F_{B_{n-1}}\}$ is similarly derived.
Computing $r\xrightarrow{F_A,F_B,J_0}_+ r_w$ gives $r_w: Z+\Func(A,B)$ which is the 
canonical word-level polynomial abstraction of the composite field design.

\begin{Example}
Consider the composite field multiplier over $\F_{{2^2}^2}$ from Example \ref{ex:comp22}.
Abstracting every multiplier and adder over the base field $\F_{2^2}$ gives the following
polynomials:
\begin{eqnarray}
f_1: Z_0+C_6+C_2      & f_2: Z_1+C_5+C_4      & f_3: C_6+\alpha^5C_3 \nonumber \\
f_4: C_5+\alpha^5C_3  & f_5: C_4+C_1+C_0      & f_6: C_3+A_1\cdot B_1 \nonumber \\
f_7: C_2+A_0\cdot B_0 & f_8: C_1+A_1\cdot B_0 & f_9: C_0+A_0\cdot B_1 
\end{eqnarray}
Set the following RATO ordering:
\begin{eqnarray}
Z_0 > Z_1 > C_6 > C_5 > C_4 > C_3 > C_2 > C_1 > C_0 \nonumber \\
> A_0 > A_1 > B_0 > B_1 > Z > A > B
\end{eqnarray}
The vanishing polynomials are:
\begin{eqnarray}
f_{10}: Z_0^4+Z_0 & f_{11}: Z_1^4+Z_1 & f_{12}: C_6^4+C_6 \nonumber \\
f_{13}: C_5^4+C_5 & f_{14}: C_4^4+C_4 & f_{15}: C_3^4+C_3 \nonumber \\
f_{16}: C_2^4+C_2 & f_{17}: C_1^4+C_1 & f_{18}: C_0^4+C_0 \nonumber \\
f_{19}: A_0^4+A_0 & f_{20}: A_1^4+A_1 & f_{21}: B_0^4+B_0 \nonumber \\
f_{22}: B_1^4+B_1 & f_{23}: Z^{16}+Z  & f_{24}: A^{16}+A \nonumber \\
f_{25}: B^{16}+B
\end{eqnarray}
Then $F=f_1,\dots,f_9$ and 
$F_0=f_{10},\dots,f_{25}$.
Here, $f_Z: Z_0+Z_1\alpha+Z$. Computing $f_Z\xrightarrow{F,F_0}_+ r$ gives:
\begin{equation}
r= A_0\cdot B_0+\alpha A_0\cdot B_1+\alpha A_1\cdot B_0+\alpha^2A_1\cdot B_1+Z
\end{equation}

Next, $F_A=\{F_{A_0},F_{A_1}\}$ needs to be derived. Here, $A=A_0+A_1\alpha$ and $A^{4}=A_0+A_1\alpha^4$,
which gives
\begin{eqnarray}
\mathbf{M}=
\begin{bmatrix}
1 & \alpha \\
1 & \alpha^4
\end{bmatrix}; &
\mathbf{M_0}=
\begin{bmatrix}
A & \alpha \\
A^4 & \alpha^4
\end{bmatrix}; &
\mathbf{M_1}=
\begin{bmatrix}
1 & A \\
1 & A^4
\end{bmatrix}
\end{eqnarray}
After minimizing by the primitive polynomial $P(x)=x^4+x^3+1$, 
the determinants of these matrices are:
\begin{eqnarray}
|\mathbf{M}|=\alpha^3+\alpha+1; &
|\mathbf{M_0}|=\alpha A^4+(\alpha^3+1)A; &
|\mathbf{M_1}|=A^4+A
\end{eqnarray}
%Notice that $|\mathbf{M}|=\alpha^5=\beta\in\F_{2^2}$.%% NOTE: haven't figured out if this is always the case 
Now $F_{A_0}$ and $F_{A_1}$ are derived:
\begin{eqnarray}
F_{A_0} = &A_0+\frac{|\mathbf{M_0}|}{|\mathbf{M}|}& = A_0+(\alpha^3+\alpha^2+1)A^4+(\alpha^3+\alpha^2)A \\
F_{A_1} = &A_1+\frac{|\mathbf{M_1}|}{|\mathbf{M}|}& = A_1+(\alpha^3+\alpha)A^4+(\alpha^3+\alpha)A
\end{eqnarray}
Since $F_B=\{F_{B_0},F_{B_1}\}$ is derived from $B=B_0+B_1\alpha$, which is isomorphic to
$A=A_0+A_1\alpha$, then $F_B$ can be derived from $F_A$ by substituting corresponding
variables.
\begin{eqnarray}
F_{B_0}& = &B_0+(\alpha^3+\alpha^2+1)B^4+(\alpha^3+\alpha^2)B \\
F_{B_1}& = &B_1+(\alpha^3+\alpha)B^4+(\alpha^3+\alpha)B
\end{eqnarray}
Finally, computing $r\xrightarrow{F_A,F_B,F_0}_+ r_w$ gives the word-level abstraction of
the circuit.
\begin{equation}
r_w: Z+A\cdot B
\end{equation}

\end{Example}

\section{Conclusion}

This chapter described how to generalize the abstraction approach to any arbitrary
combinational circuit. This method works best when the derived operand-width $k$, from the $LCM$ of the 
word-lengths of all inputs and output, is not very large, for instance, when the output size is a 
multiple of the input sizes. When all word-sizes are relatively prime,
$k$ is a product of all sizes, which can make the analysis overly bulky.

An abstraction technique for exploiting the hierarchy of composite field multiplier circuits over
$\F_{(2^m)^n} $
was also examined. The approach abstracts all internal multipliers and adders over $\F_{2^m}$ in 
parallel and uses these abstractions to compute the final abstraction completely over
word-level variables. Experimental results for abstractions of composite field multipliers using a
custom-built tool is presented in the next chapter.

### 2016-09-21 dissertation/conclude.tex
init dissertation writing

\chapter{Conclusions and Future Work}
\label{ch:concl}

A combinational circuit with $k$-inputs and $k$-outputs implements
Boolean functions  $f: \B ^k \rightarrow \B ^k$, where $\B = \{0,
1\}$. The function can also be construed as a mapping  $f: \Fkk
\rightarrow \Fkk$, where $\Fkk$ denotes the Galois field of
$2^k$ elements. A circuit with differing input and output sizes 
computes $f: \B ^m \rightarrow \B ^n$, which can be represented
as a function over Galois fields $f: \F_{2^m} \rightarrow \F_{2^n}$.
This circuit can also be analyzed as the function 
$f: \Fkk \rightarrow \Fkk$, where $\Fkk \supset \F_{2^m}$ 
and $\Fkk \supset \F_{2^m}$.

Every function $f$ over $\Fkk$ is a polynomial
function --- i.e., there exists a unique,  minimal, canonical
polynomial $\Func$ that describes $f$. This dissertation presented novel
techniques based on computer-algebra and algebraic-geometry to
derive the canonical (word-level) polynomial representation from the
circuit as  $Z = \Func(A)$ over $\Fkk$, where $A$ and $Z$ denote, 
respectively, the input and output bit-vectors of the circuit.

A theory for word-level polynomial abstraction of bit-level circuits 
over Galois fields is first developed.
This theory is derived using techniques from computer-algebra, notably the theory of
\Grobner basis. However, due to the computational complexity of computing a \Grobner
basis, the solution is not scalable to large designs.
In order to overcome these limitations, new symbolic computational
algorithms are developed and refined. The algorithms employ techniques
from the binomial expansion over $\Fkk$ and $\F_4$-style reduction and can exploit
hierarchy in a given circuit. Finally, an efficient implementation of the algorithmic
approach is presented.

Experiments show that the proposed approach works exceptionally well
for abstracting word-level Galois field arithmetic circuits. It has been
shown that the approach can abstract and verify these types of circuits with up to 
$1024$-bit datapaths. Other contemporary techniques
cannot to verify these types for circuits beyond $163$-bits and fail to
abstract them beyond $32$-bits.

However, in cases of random logic, the abstraction approach
can generate high-degree polynomials:
\begin{equation}
X^{q-1}+X^{q-2}\dots
\end{equation}
In these cases, the polynomials derived during the computation are dense, 
and the computational complexity of manipulating such polynomials
makes abstraction infeasible.

\section{Future Work}

Due to the modular nature of the proposed solution, there are many
potential future research directions that can be explored.

\subsection{Hardware Acceleration}
The first reduction step, $f_Z\xrightarrow{F-f_{z_i},F_0}_+ r$
is the most computationally complex part of the proposed abstraction approach. This
reduction could be implemented using a hardware accelerator.
Significant speed-ups have 
been observed in GPU implementations of circuit simulation algorithms 
\cite{PengLi:GPU}. 
Furthermore, this work has shown cases where multiple, independent abstractions 
need to be computed at the same time, such as when abstracting a word-level 
representation of a composite field multiplier.

These abstractions can be computed in parallel with one another, and this
parallelism could then be exploited using a GPU. Furthermore, our 
approach to compute the abstractions uses an $F4$-style reduction procedure, 
which performs many complex computations over a large matrix. 
Operations over matrices can be suitably implemented using a GPU.
Lastly, the substitution by $a_i=\Func(A)$ is trivially parallelized. 
Thus, further study is proposed to implement word-level abstraction 
on a general purpose GPU.

\subsection{Integration with EDA Tools} 
The proposed canonical word-level abstraction approach is a full, 
self-contained solution.
It can thus be integrated into other EDA tools. There are direct
applications of word-level abstractions to design synthesis. For instance, 
the approach can compute a functional decomposition of a logic or it could 
be used in high-level RTL synthesis. Since the derived abstraction 
is canonical, it can also be used in verification engines such as SMT solvers.
The abstraction approach most efficiently handles AND/XOR logic, so it
could be used to complement approaches in the mentioned tools that are
efficient over AND/OR logic.

\subsection{Polynomial Reductions using Data-Structures}
The abstraction approach poorly handles chains of OR gates
due to their representation as polynomials of Galois fields.
Other polynomial-based tools \cite{polybori:2009} have shown that it 
can be beneficial to represent polynomials internally as decision diagrams.
Thus, it is worthwhile to explore whether it is possible to
implement the algorithmic approach in a different data-structure that
is better-suited for handling this type of logic.
One candidate data-structure is the And-Invert-Graph, as this structure efficiently
handles OR gates. The widely-used tool ABC\cite{abc} provides a very efficient, 
flexible, and open-source implementation of the AIG data structure.

Recall that a one-step reduction  of the polynomial $f$ by polynomial $g$, 
$f\xrightarrow{g} r$, is computed as:
\begin{equation}
r = f-\frac{LT(f)}{LT(g)}\cdot g
\end{equation}
Over Boolean circuits, $\B \equiv \F_2$. Since the leading term of
any monomial in $\B$ is $1$, and $-1\equiv +1$, then
\begin{equation}
f-(\frac{LT(f)}{LT(g)}\cdot g)\equiv f+(\frac{LM(f)}{LM(g)}\cdot g)
\end{equation}
which computes an XOR operation
\begin{equation}
f\oplus(\frac{LM(f)}{LM(g)}\cdot g)
\end{equation}
while the $\cdot$ operator acts as an AND operation $\wedge$.
So one step of the reduction procedure can be computed as the following 
AND/XOR operation:
\begin{equation}
r = f \oplus \frac{LM(f)}{LM(g)}\wedge g
\end{equation}

Thus, we propose an investigation into implementing the algorithms presented
in this dissertation over AIGs.

\subsection{Application to Sequential Circuit Verification}
Sequential Galois field arithmetic circuits over $\Fkk$
take k-bit inputs and produce a k-bit result after k-clock cycles 
of operation. Formal verification of sequential arithmetic
circuits with large datapath sizes is beyond the capabilities of
contemporary verification techniques. To address this problem,
we described a verification method in \cite{sun:date15} 
which uses the presented abstraction approach to implicitly 
unroll the sequential arithmetic circuit over multiple (k) clock-cycles.
The resulting function computed by the state-registers of the circuit 
is represented canonically as a multi-variate word-level polynomial over
$\Fkk$. While directly applicable to sequential Galois field arithmetic circuits, 
this work needs to be further generalized in order to make applicable to
any sequential state machine.

\subsection{Application to Formal Software Verification}
Computer algebra techniques based on \Grobner basis theory have been used 
in formal software verification \cite{manna:program}.
In this work, a \Grobner basis computation is used to derive {\it loop invariants}.
However, the derived invariants are not bit-precise, so not every invariant that is
computed can be applied to the verification. As our approach maintains
the input-output relationship in the abstraction, it could be applied to
find bit-precise invariants.

\subsection{Application to Integer Arithmetic Circuits}
The abstraction approach derives a
word-level representation of circuits over Galois fields, $\Fkk$. 
In order to expand its usability, we conjecture whether it is possible to 
apply concepts from this approach to {\bf abstract word-level representations 
of circuits over integer rings}, $\Zkk$. 
As any function over a Galois 
field $\Fkk$ is a polynomial function, there exists a polynomial which 
describes the word-level function of a given circuit over $\Fkk$. 
%If the word-level 
%input and output of the circuit is $A$ and $Z$, respectively, this 
%polynomial exists in the form $Z=\Func(A)$.
However, not every function over an integer ring $\Zkk$ is a polynomial 
function. 
Thus, a single polynomial which describes the function of a circuit over 
$\Zkk$ is not guaranteed to exist. 
%Instead, there could exist a set of
%polynomials $\{f_1,\dots,f_n\}$ which describes this function.
%These polynomials are in following form:
%\begin{eqnarray}
%f_1:Z=\Func_1(A) \nonumber \\
%\vdots \nonumber \\
%f_n:Z=\Func_n(A) \nonumber
%\end{eqnarray}
Even though there may not exist a single polynomial which describes the
entire function over $\Zkk$, elimination theory and \Grobner basis still
applies over this ring. Thus, it may be possible to modify the theory and 
implementation of our word-level 
abstraction approach in order to abstract a set of word-level 
polynomials over $\Zkk$.

### 2016-09-21 dissertation/gfprelim.tex
init dissertation writing

\chapter{Galois Fields Preliminaries and Application in Hardware Design} \label{ch:prelim}
This chapter provides a mathematical background for understanding 
Galois fields and explains how to design Galois field arithmetic circuits.
We first introduce the mathematical concepts of groups, rings, fields, and 
polynomials. 
We then apply these concepts to create Galois field arithmetic functions and 
explain how to map them to a Boolean circuit implementation.
The material is referred from \cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997} for Galois field concepts and 
\cite{mastro:1989} \cite{PT:1985} \cite{acar:1998} \cite{wu:2002} \cite{Knezevic:2008} for hardware design over Galois fields and previous work
by {\it Lv} \cite{lv:phd}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Rings, Fields and Polynomials}

\begin{Definition}
An {\bf abelian group} is a set $\mathbb{S}$ with a binary operation $'+'$
which satisfies the following properties: 
\begin{itemize}
\item {\it Closure Law:} For every $a, b \in \mathbb{S}, a + b \in \mathbb{S}$  
\item {\it Associative Law:} For every $a, b, c \in \mathbb{S}, (a + b) + c = a + (b + c)$
\item {\it Commutativity:} For every $a, b \in \mathbb{S}, a + b = b + a$. 
\item {\it Additive Identity:} There is an identity element $0 \in \mathbb{S}$
such that for all $a \in \mathbb{S};$ $a + 0 = a$.
\item {\it Additive Inverse:} If $a \in \mathbb{S}$, then there is an
element $a^{-1} \in \mathbb{S}$ such that $ a + a^{-1} = 0$.
\end{itemize}
\end{Definition}

The set of integers $\mathbb{Z}$ forms an abelian group under the addition operation.

\begin{Definition}
Given a set $\mathbb{R}$ with two binary operations, $'+'$ and $'\cdot'$, 
and element $0 \in \mathbb{R}$, the system $\mathbb{R}$ is called a {\bf commutative ring with unity} if the following properties hold:
\begin{itemize}
\item $\mathbb{R}$ forms an abelian group under the '+' operation with additive identity element $0$.
\item {\it Multiplicative Distributive Law}: For all $a, b, c \in$ $\mathbb{R}$, $a\cdot (b + c) = a\cdot b + a\cdot c$.
\item {\it Multiplicative Associative Law}: For every $a, b, c\in \mathbb{R}$, $a\cdot (b\cdot c) = (a\cdot b)\cdot c$. 
\item {\it Multiplicative Commutative Law}: For every $a,b \in \mathbb{R}$, $a\cdot b = b\cdot a$
\item {\it Identity Element}: There exists an element $1 \in$ $\mathbb{R}$ 
such that for all $a \in \mathbb{R}$, $a\cdot 1 = a =1\cdot a$
\end{itemize}
\end{Definition}

For the purpose of this dissertation, any time we refer to a {\bf ring}, we are 
specifically referring to a {\bf commutative ring with unity}. Two common 
examples of such rings are the set of integers, $\mathbb{Z}$, and the set of 
rational numbers, $\mathbb{Q}$. Note that while both of these examples are
rings with an infinite number of elements, the number of elements in a ring 
can also be finite.

\begin{Definition}
The {\bf modular number system} with base $n$ is a set of positive
integers $Z_n = \{0, 1, \ldots, n-1\}$, with the two operations $+$
and $\cdot$ satisfying the properties below:
\begin{eqnarray}
(a + b)\pmod{ n } &\equiv& ((a ~\pmod {n}) + (b ~\pmod {n})) ~\pmod {n} \nonumber\\
(a\cdot b) ~\pmod {n} &\equiv& ((a ~\pmod {n}) \cdot (b ~\pmod {n})) ~\pmod {n} \label{eq:modmult}\nonumber\\
(-a) ~\pmod {n} &\equiv& (n-a) ~\pmod {n}\nonumber 
\end{eqnarray}
\end{Definition}

\begin{Example}
The set $Z_8 = \{0, 1, \ldots, 7\}$ denotes the modular number system
with base $8$. Examples of some operations performed $\pmod {8}$
are:
\begin{eqnarray} \nonumber
        3 + 6   &~=& 9  ~\pmod{8} ~= 1 \nonumber \\
        3 \cdot 6   &~=& 18 ~\pmod{8} ~= 2 \nonumber \\
        (-3)    &~=& 8-3  ~\pmod{8} ~= 5 \nonumber
\end{eqnarray} \nonumber
\end{Example}

The modular number system $\mathbb{Z}_n = \{0, 1, \ldots, n-1\}$, where $n$ 
is a positive integer, forms a ring.
Since this type of ring contains a finite number of elements $n$,
it is termed a {\it finite integer ring}, where addition and multiplication 
are computed {\it modulo n} $\pmod {n}$. 
In hardware applications, arithmetic over $k$-bit vectors manifests itself 
as algebra over the finite integer ring $\mathbb{Z}_{2^k}$, where the $k$-
bit vector represents integer values from $\{0, ...., 2^k-1\}$.

\begin{Example}
\label{exp:4bitadder}
Consider the following arithmetic circuit:

\begin{figure}[!h]
\centerline{
\includegraphics[width=0.3\textwidth]{figures/4bitMultTop}
}
%\caption{Typical hardware design flow.}
%\label{fig:cadflow}
\end{figure}

This circuit takes two 4-bit inputs, $A$ and $B$, and computes a 4-bit sum $C$. 
Since $A$, $B$, and $C$ are all bit-vectors of size $4$, 
the addition computation this circuit performs is modulo $2^4$.
Hence, this circuit exemplifies arithmetic computations over the ring $Z_{2^4}$.

Some examples of possible inputs and outputs of the circuit:
\begin{table}[!h]
	\centering
	\begin{tabular}{|lll|lll|}
	\hline
	\multicolumn{3}{|c|}{Addition over $\mathbb{Z}_{2^4}$} & \multicolumn{3}{c|}{Boolean Circuit Implementation} \\
	\hline
	$5 + 8$ & $~=$ & $13 ~\pmod{16} ~= 13$ & $A=0101$, $B=1000$ & $\rightarrow$ &  $C=1101$ \\
	$10 + 9$ & $~=$ & $ 19 ~\pmod{16} ~= 3$ & $A=1010$, $B=1001$ & $\rightarrow$ &  $C=0011$ \\
	$12 + 4$ & $~=$ & $ 16 ~\pmod{16} ~= 0$ & $A=1100$, $B=0100$ & $\rightarrow$ &  $C=0000$ \\
	\hline
	\end{tabular}
\end{table}
\end{Example}

\begin{Definition}\label{def:poly}
Let $\mathbb{R}$ be a ring. A {\bf polynomial} over $\mathbb{R}$ in the 
indeterminate $x$ is an expression of the form:
\begin{equation} \label{eq:poly1}
a_0 + a_1 x + a_2 x^2 + \cdots + a_k x^k = \sum_{i=0}^{k} a_i x^i, \forall a_i \in \mathbb{R}. 
\end{equation}

\end{Definition}

The constants $a_i$ are the coefficients and $k$ is the degree of the polynomial. 
For example, $8x^3 + 6x + 1$ is a polynomial in $x$ over $\mathbb{Z}$, with
coefficients $8$, $6$, and $1$ and degree $3$.

\begin{Definition}
The set of all polynomials in the indeterminate
$x$ with coefficients in the ring $\mathbb{R}$ forms a {\bf
ring of polynomials} $\mathbb{R}[x]$. 
Similarly, $\mathbb{R}[x_1,x_{2},\cdots, x_{n}]$ 
represents the ring of multivariate polynomials with coefficients in $\mathbb{R}$.
\end{Definition}

For example, $\mathbb{Z}_{2^4}[x]$ stands for the set of all polynomials in
$x$ with coefficients in $\mathbb{Z}_{2^4}$. $8x^3 + 6x + 1$ is an instance 
of a polynomial contained in $\mathbb{Z}_{2^4}[x]$.

\begin{Definition}
A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
non-zero element in $\mathbb{F}$ has a multiplicative inverse; i.e. $\forall
a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
\hat{a} = 1$.
\end{Definition}

A field is defined as a ring with one extra condition: the presence of a 
multiplicative inverse for all non-zero elements.
Therefore, a field must be a ring while a ring is not necessarily a field.
For example, the set $\mathbb{Z}_{2^k} = \{0,1,\cdots, 2^k-1\}$ forms a finite ring.
However, $\mathbb{Z}_{2^k}$ is not a field because not every element in
$\mathbb{Z}_{2^k}$ has a multiplicative inverse. 
In the ring $\mathbb{Z}_{2^3}$, for 
instance, the element $5$ has an inverse ($5\cdot5\pmod{8}=1$) but the element $4$
does not.

The main concept of field theory is {\bf Field Extensions}. The idea behind a
field extension is to take a base field and construct a larger field which 
contains the base field as well as satisfies additional properties. For example,
the set of real numbers $\mathbb{R}$ forms a field; one common extension of 
$\mathbb{R}$ is the set of complex numbers $\mathbb{C}=\mathbb{R}(i)$. Every
element of $\mathbb{C}$ can be represented as $a+b\cdot i$ where $a,b \in \mathbb{R}$,
hence $\mathbb{C}$ is a two-dimensional extension of $\mathbb{R}$.

Like rings, fields can also contain either an infinite or a finite number of 
elements. 
In this dissertation we focus on finite fields, also known as Galois fields, and 
the construction of their field extensions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Galois Fields}\label{sec:ff}
Galois fields, also known as finite fields, find widespread applications in 
many areas of electrical engineering and computer science such as error-
correcting codes, elliptic curve cryptography, digital signal processing, 
testing of VLSI circuits, among others.
In this dissertation, we specifically focus on their application to 
Elliptic Curve Cryptography as Galois field arithmetic circuits.
This section describes the relevant Galois field concepts
\cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997}
and hardware arithmetic designs over such fields \cite{mastro:1989} \cite{PT:1985} 
\cite{acar:1998} \cite{wu:2002} \cite{Knezevic:2008}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{Definition} 
A {\bf Galois field}, denote $\Fq$, is a field with a finite
number of elements, $q$. The number of elements $q$ of the Galois field is
a power of a prime integer, i.e. $q = p^k$, where $p$ is a prime
integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
$\F_{p^{k}}$.
\end{Definition}

Fields in the form $\F_{p^{k}}$ are called Galois extension fields.
We are specifically interested in extension fields of type 
$\Fkk$, where $k > 1$. These are extensions of the binary
field $\F_2$.
\begin{Example}
Addition and multiplication operations over $\F_2$:
\begin{table}[!h]
	\centering
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
	\hhline{~---~}
	\multirow{3}{*}{} & $+$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $1$ & \\
	& $1$ & $1$ & $0$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Addition over $\F_2$}\\
	\end{tabular}
	\quad
	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
	\hhline{~---~}
	\multirow{3}{*}{} & $\cdot$ & $0$ & $1$ & \multirow{3}{*}{} \\
	\hhline{~---~}
	& $0$ & $0$ & $0$ & \\
	& $1$ & $0$ & $1$ & \\
	\hhline{~---~}
	\multicolumn{5}{c}{}\\
	\multicolumn{5}{c}{Multiplication over $\F_2$}\\
	\end{tabular}
\end{table}

Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
because it is performed modulo $2$.
Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.
\end{Example}

Algebraic extensions of the binary field $\F_{2}$  
are generally termed as {\it binary extension fields} $\Fkk$.
Where elements in $\F_2$ can only represent $1$ bit, elements in $\Fkk$ 
represent a $k$-bit vector.
This allows them to be widely used in digital hardware applications.
In order to construct a Galois field of the form $\Fkk$, 
an {\bf irreducible polynomial} is required:
\begin{Definition}
A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$ is {\bf irreducible} 
if $P(x)$ is non-constant with degree $k$ and cannot be 
factored into a product of polynomials of lower degree in $\mathbb{F}_2[x]$.
\end{Definition}

Therefore, the polynomial $P(x)$ with degree $k$ is irreducible over 
$\mathbb{F}_{2}$ if and only if it has no roots in $\mathbb{F}_{2}$,
i.e if $\forall a \in \mathbb{F}_{2}$, $P(a)\neq 0$.
For example, $x^2+x+1$ is an irreducible polynomial over $\mathbb{F}_{2}$
because it has no solutions in $\mathbb{F}_{2}$, i.e. $(0)^2+(0)+1=1\neq0$ 
and $(1)^2+(1)+1=1\neq0$ over $\F_2$.
Irreducible polynomials exist for any degree $\geq 2$ in $\mathbb{F}_2[x]$.

Given an irreducible polynomial $P(x)$ of degree $k$ in the polynomial ring 
$\mathbb{F}_2[x]$, we can construct a binary extension field 
$\mathbb{F}_{2^k} \equiv \mathbb{F}_2[x] \pmod{P(x)}$.
Let $\alpha$ be a root of $P(x)$, i.e., $P(\alpha)=0$.
Since $P(x)$ is irreducible over
$\mathbb{F}_2[x]$, $\alpha \notin \mathbb{F}_2$. 
Instead, $\alpha$ is an element in $\mathbb{F}_{2^k}$. 
Any element $A \in \mathbb{F}_{2^k}$ is then represented as: 
\begin{equation}\label{rep:poly}
A= \sum_{i=0}^{k-1} (a_i \cdot \alpha^i) = a_0 + a_1\cdot\alpha + \cdots + a_{k-1}\cdot \alpha^{k-1}\nonumber
\end{equation}
where $a_i \in \mathbb{F}_2$ are the coefficients and $P(\alpha)=0$.

To better understand this field extension, compare its similarities to another
common-place
field extension $\C$, the set of complex numbers. $\C$ is an extension of the field 
of real numbers $\R$ with an additional element $i=\sqrt{-1}$, which is an imaginary
root in $\R$.
Thus $i \notin \R$, rather $i \in \C$.
Every element $A \in \mathbb{C}$ can be represented as:
\begin{equation}\label{rep:polyC}
A=\sum_{j=0}^{1} (a_j \cdot i^j)=a_0+a_1\cdot i
\end{equation}
where $a_j \in \R$ are coefficients. Similarly, $\Fkk$ is an extension of $\F_2$ with 
an additional element $\alpha$, which is the ``imaginary root'' of an irreducible 
polynomial $P$ in $\F_2[x]$.

Every element $A \in \Fkk$ has a degree less than $k$ because 
$A$ is always computed modulo $P(x)$, which has degree $k$. 
Thus, $A\pmod {P(x)}$ can be of degree at most $k-1$ and at least $0$.
For this reason, the field $\mathbb{F}_{2^k}$ can be viewed as a $k$
dimensional vector space over $\mathbb{F}_{2}$. 
The equivalent bit vector representation for element $A$ is:
\begin{equation}
A=(a_{k-1} a_{k-2} \cdots a_{0})
\end{equation}

\begin{Example}
A 4-bit Boolean vector, $(a_{3} a_{2} a_{1} a_{0})$
can be presented over $\F_{2^4}$ as: 
\begin{equation}
a_3 \cdot \alpha^3+a_2 \cdot \alpha^2+a_1 \cdot \alpha+a_0
\end{equation}
For instance, the Boolean vector $1011$ is represented as the element 
$\alpha^3+\alpha+1$.
\end{Example}

\begin{Example}\label{exp:1}
Let us construct $\mathbb{F}_{2^4}$ as $\mathbb{F}_2[x] \pmod{ P(x)}$, where
$P(x)=x^4+x^3+1 \in \mathbb{F}_2[x]$ is an irreducible polynomial of degree $k=4$. 
Let $\alpha$ be the root of $P(x)$, i.e. $P(\alpha)=0$.

Any element $A \in \mathbb{F}_2[x] \pmod{ x^4 + x^3 + 1}$
has a representation of the type: $A = a_3 x^3 + a_2 x^2 +
a_1 x + a_0$ (degree $< 4$) where the coefficients $a_3, \dots, a_0$ are in $\F_2 =
\{0, 1\}$. Since there are only $16$ such polynomials, we obtain
$16$ elements in the field $\mathbb{F}_{2^4}$. Each element in
$\mathbb{F}_{2^4}$ can then be viewed as a $4$-bit vector over $\mathbb{F}_{2}$. 
Each element also has an exponential $\alpha$
representation. All three representations are shown in Table
\ref{tab:gfelement}.

\begin{table}[h]
\begin{center}
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$}\label{tab:gfelement} 
\begin{tabular}{|c|c|c||c|c|c|} 
\hline
$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
\hline
$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
\hline
$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
\hline
$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
\hline
$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
\hline
$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
\end{center}
\end{table}

We can compute the polynomial representation from the exponential representation.
Since every element is computed $\pmod{P(\alpha)} = \pmod{\alpha^4+\alpha^3+1}$, 
we compute the element $\alpha^{4}$ as 
\begin{equation}
\alpha^{4} \pmod{ \alpha^4+\alpha^3+1} = -\alpha^3 - 1 = \alpha^3+1
\end{equation}
Recall that all coefficients of $\F_{2^4}$ 
are in $\F_{2}$ where $-1 = +1$ modulo 2.
The next element $\alpha^{5}$ can be computed as 
\begin{equation}
\alpha^{5} = \alpha^{4}\cdot \alpha = (\alpha^3+1)\cdot \alpha = \alpha^4+\alpha = \alpha^3+\alpha+1 
\end{equation}
Then $\alpha^6$ can be computed as $\alpha^{5}*\alpha$ and so on.
\end{Example}

An irreducible polynomial can also be a primitive polynomial.

\begin{Definition}
A {\bf primitive polynomial} $P(x)$ is a polynomial with coefficients in $\mathbb{F}_2$ 
which has a root $\alpha$ $\in$ $\mathbb{F}_{2^k}$
such that \{$0$, $1(=\alpha^{{2^k}-1})$, $\alpha$, $\alpha^2$, $\cdots$, $\alpha^{2^k-2}$\} is the set of 
all elements in $\mathbb{F}_{2^k}$, 
where $\alpha$ is a {\bf primitive element} of $\mathbb{F}_{2^k}$. 
\end{Definition}

A primitive polynomial is guaranteed to generate all distinct elements 
of a finite field $\mathbb{F}_{2^k}$ while an irreducible polynomial
has no such guarantee.
Often, there exists more than one irreducible polynomial of degree $k$.
In such cases, any degree $k$ irreducible polynomial can be 
used for field construction. For example, both $x^3+x+1$ and $x^3+x^2+1$ 
are irreducible in $\mathbb{F}_2$ and either one can be used
to construct $\mathbb{F}_{2^3}$. This is due to the following:

\begin{Theorem}\label{the:unique}
There exist a {\bf unique} field $\mathbb{F}_{p^k}$, for any prime $p$ and any positive integer $k$.
\end{Theorem}

Theorem \ref{the:unique} implies that Galois fields with the same number of elements are 
{\bf isomorphic} to each other up to the labeling of the elements.

Theorem \ref{the:fer} provides an important property for investigating solutions to
polynomial equations in $\Fq$.

\begin{Theorem}\label{the:fer}
 $\left[Generalized\  Fermat's\  Little\  Theorem \right]$ Given a
 Galois field $\mathbb{F}_{q}$, each element $A \in \mathbb{F}_{q}$ satisfies: 
\begin{eqnarray}\label{fe}
 A^{q} & \equiv & A  \nonumber \\
 A^{q} - A & \equiv& 0  
\end{eqnarray}
\end{Theorem}

We can extend Theorem \ref{the:fer} to polynomials in $\mathbb{F}_{q}[x]$ as 
follows: 
\begin{Definition}
Let $x^q-x$ be a polynomial in $\mathbb{F}_{q}[x]$.
Every element $A \in \mathbb{F}_{q}$ is a solution to  $x^q-x=0$. 
Therefore, $x^{q} - x$ always {\it vanishes} in $\mathbb{F}_{q}$. Such 
polynomials are called {\bf vanishing polynomials} of the field $\mathbb{F}_{q}$.
\end{Definition}

\begin{Example}
Given $\mathbb{F}_{2^2} =\{0,1,\alpha,\alpha+1\}$ with $P(x)=x^2+x+1$, where $P(\alpha)=0$. 
 \begin{eqnarray}
 0^{2^2}&=&0 \nonumber \\
 1^{2^2}&=&1 \nonumber \\
 \alpha^{2^2}&=&\alpha \pmod {\alpha^2+\alpha+1}\nonumber \\
 (\alpha+1)^{2^2}&=&\alpha+1 \pmod {\alpha^2+\alpha+1} \nonumber 
 \end{eqnarray}
\end{Example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Containment of Galois Fields}
A Galois field $\F_q$ can be fully contained within a larger field $\F_{q^k}$.
That is, $\F_q \subset \F_{q^k}$.
For example, Fig \ref{fig:contain2_4_16} shows the containment of the fields 
$\F_2 \subset \F_4 \subset \F_{16}$. It's easy to see that since $\F_4=\F_{2^2}$, it
contains $\F_2$. Likewise $\F_{16}=\F_{4^2}=\F_{2^4}$ contains $\F_4$ and $\F_2$.
The elements $\{0,1,\alpha,\dots,\alpha^{14}\}$
designate $\F_{16}$. Of these, $\{0,1,\alpha^5,\alpha^{10}\}$ create $\F_4$.
From these, only $\{0,1\}$ exist in $\F_2$.

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.7]{./figures/field-cont}
\end{center}
\caption{Containment of Fields: $\F_2 \subset \F_4 \subset \F_{16}$}
\label{fig:contain2_4_16}
\end{figure}

Consider the element $\alpha^5$ of $\F_{16}$. 
Deriving all $(\alpha^5)^i$ for $i \geq 0$ 
over $\F_{16}$ gives the following recurrence:
\begin{eqnarray}
(\alpha^5)^0&=&1\nonumber \\
(\alpha^5)^1 &=& \alpha^5\nonumber \\
(\alpha^5)^2 &=& \alpha^{10}\nonumber \\
(\alpha^5)^3 &=& \alpha^{15} = 1
\end{eqnarray}
The only elements that are generated in this recurrence are $\{1,\alpha^5,\alpha^{10}\}$.
Every field contains $\{0,1\}$, so the elements $\{0,1,\alpha^5,\alpha^{10}\}$ form $\F_4$.
Let $P(x)=x^4+x^3+1$ be the primitive polynomial used to generate
$\F_{2^4}=\F_{16}$. A primitive polynomial of degree $2$ used to generate $\F_{2^2}=\F_4$
can be found as follows:
\begin{eqnarray}
& &(x+\alpha^5)\cdot(x+\alpha^{10}) \mod P(x) \nonumber \\
&=& x^2+(\alpha^{10}+\alpha^5)x+\alpha^{15} \mod P(x) \nonumber \\
&=& x^2+x+1
\end{eqnarray}

\begin{Theorem}
$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, i.e. if $n$ divides $m$.
\end{Theorem}

Therefore:
\begin{itemize}
\item $\F_2 \subset \F_{2^2} \subset \F_{2^4} \subset \F_{2^8} \subset \dots$
\item $\F_2 \subset \F_{2^3} \subset \F_{2^9} \subset \F_{2^{27}} \subset \dots$
\item $\F_2 \subset \F_{2^5} \subset \F_{2^{25}} \subset \F_{2^{125}} \subset \dots,$ and so on
\end{itemize}

\begin{Definition}
The {\bf algebraic closure} of the Galois field $\F_{2^k}$, denoted $\overline{\Fkk}$, is the 
union of all fields $\F_{2^n}$ such that $k \mid n$.
\end{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Polynomial Interpolation over Galois Fields}

In the construction of digital circuits, arbitrary mappings between 
two bit-vectors of size $k$ can be constructed. Each such
mapping generates a function $f: \B^k \rightarrow \B^k$.
As every $k$-bit vector can be construed as an element in $\Fkk$ 
(as shown in the previous section), 
every such function also corresponds to a function over a 
Galois field: $f: \Fkk \rightarrow \Fkk$.

\begin{Definition}
A function $f: \mathbb{R} \rightarrow \mathbb{R}$ over a ring $R$ is 
considered a {\bf polynomial function} if there exists a polynomial
$\Func \in \mathbb{R}[x_1,\dots,x_d]$ such that 
$\Func(x_1,\dots,x_d) = f(x_1,\dots,x_d)$.
\end{Definition}

\begin{Theorem}
From \cite{ff:1997}: 
Let $\Fq$ be a Galois field of $q$ elements where $q$ is a power of a
prime integer. Given any function $f: \Fq \to \Fq$, there exists a 
polynomial 
$\Func \in \Fq [x]$ such that $f(a) = \Func(a)$, for all $a \in \Fq$.
Thus, every function $f: \Fq \to \Fq$ is a polynomial function.
\end{Theorem}

Thus, since every function over a Galois field, $f: \Fkk \rightarrow \Fkk$, 
is a polynomial function, 
every mapping between two bit-vectors of size $k$ 
is a polynomial function over $\Fkk$.
Furthermore, every polynomial can be derived using Lagrange interpolation.

\begin{Theorem} ({\bf Lagrange Interpolation}): \\
Given a set of $k$ data points over a function $f$,
\begin{equation}
(x_0,f(x_0)),\dots,(x_{k-1},f(x_{k-1})) \nonumber
\end{equation}
where no two $x_i \in \{x_0,\dots,x_{k-1}\}$ are the same elements,
the polynomial representation of $f$, $\Func(x)$, can be interpolated as 
follows:  
\begin{eqnarray}
\Func(x) = \sum_{i=0}^{k-1} f(x_i)\cdot L_i(x) \nonumber \\
L_i(x) =  \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \nonumber  
\end{eqnarray}
\end{Theorem}

By applying Lagrange interpolation over every element in the Galois
field $\Fkk$, 
we can derive the polynomial representation $\Func$ of any function 
$f:\Fkk \rightarrow \Fkk$. Furthermore, $\Func$ is a polynomial of degree 
at most $2^k-1$ in $x$ and $\Func(a)=f(a)$ for all $a \in \Fkk$.

While every function over a Galois field is a polynomial function,
not every function over the integer ring $\mathbb{Z}$ is a polynomial 
function.

%\begin{equation}
%\Func(x) = \sum_{k=1} ^q  \frac{ \prod_{i \neq k}  (x -x_i)}{\prod_{i \neq k}(x_k -x_i)} \cdot \f(x_k)
%\label{eqn:lagrange}
%\end{equation}

%As every function over a Galois field is also a polynomial 
%function, any mapping between two $k$-bit vectors is also a 
%polynomial function when analyzed over $\Fkk$.

%By analyzing $\Func$ over each point in $\Fq$ and applying 
%{\bf Lagrange's interpolation formula}, shown in 
%Equation \ref{eqn:lagrange}, one can interpolate the 
%polynomial representation, $f_p$, of the function $\Func$.

%Then, $f_p$ is a polynomial of degree at most $q-1$ in $x$ and
%$f_p = \F(a)$ for all $a \in \Fq$, and $\F(x)$.
%Thus, $f_p$ is a polynomial function representation of $\Func$.

\begin{Example} {\it
Let $A = \{a_2, a_1, a_0\}$ and $Z = \{z_2,z_1,z_0\}$ be 3-bit vectors.
Thus, $A$ and $Z \in \mathbb{B}^3$. 
Consider the following function:
\begin{equation}
f:Z[2:0] = A[2:0]>>1 \nonumber
\end{equation}
$f$ is a {\bf bit-vector right shift} operation on $A$. 
This function can be analyzed as a mapping over different forms: 
$\mathbb{B}^3 \rightarrow \mathbb{B}^3$,
$\mathbb{Z}_8 \rightarrow \mathbb{Z}_8$, and
$\F_{2^3} \rightarrow \F_{2^3}$. These mappings from $A$ to $Z$ 
are:

\begin{center}
{\small
\begin{tabular}{c|c|ccc|c|c|} 
$\{a_2a_1a_0\}\in\mathbb{B}^3$  & $A\in \mathbb{Z}_8$ & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$ &$Z\in \mathbb{Z}_8$ & $Z\in \F_{2^3}$ \\
\hline
000  &0&0 &$\rightarrow$&000 &0& 0 \\
001  &1&1 &$\rightarrow$&000 &0& 0 \\
010  &2&$\alpha$ & $\rightarrow$ & 001&1& 1 \\
011  &3&$\alpha + 1$ &$\rightarrow$& 001&1 &1 \\
100  &4&$\alpha^2$ &$\rightarrow$& 010 &2&  $\alpha$ \\
101  &5&$\alpha^2 + 1$ &$\rightarrow$&010 &2& $\alpha$ \\
110  &6&$\alpha^2 + \alpha$&$\rightarrow$& 011 &3&$\alpha + 1$ \\
111  &7&$\alpha^2 + \alpha + 1$ &$\rightarrow$& 011 &3&$\alpha + 1$\\
\hline
\end {tabular}
}
\end{center}

$f: \mathbb{Z}_8 \rightarrow \mathbb{Z}_8$ is not a polynomial function 
(this can be verified using the results of \cite{singmaster}
\cite{chen_95} \cite{chen_96}). However, $f: \F_{2^3} \rightarrow \F_{2^3}$
is a polynomial function.
By applying Lagrange's interpolation formula to $f$ over $\F_{2^3}$ for 
every element in $\F_{2^3}$, 
we obtain the following polynomial function: $Z =
(\alpha^2+1)A^4+(\alpha^2+1)A^2$, where $P(\alpha) = \alpha^3 +
\alpha + 1 = 0$. 
}
\end{Example}

Since every function over $\Fkk$ is a polynomial function, the 
functional mapping of a Galois field arithmetic circuit over
$\Fkk$ must exist in polynomial form. 
Construction of these arithmetic circuits is described next.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Hardware Implementations of Arithmetic Operations Over Galois Fields}

There are two main applications of hardware implementations of Galois field 
arithmetic.
In the first case, Galois field arithmetic computations, such as {\sc add or mul},  
are implemented in hardware, and  
algorithms are then implemented in software 
(e.g. cryptoprocessors \cite{ST23} \cite{kobayashi}). 
In other cases, the entire design can be implemented in hardware, such as a one-shot 
Reed-Solomon encoder-decoder chip \cite{reed-solo-chip} \cite{ecc163}, or point 
multiplication circuitry \cite{ecc:software} used in elliptic curve cryptosystems. 
Therefore, there has been extensive research in efficient hardware design of 
primitive arithmetic computations over Galois fields.
In this section, we describe the design principles of such circuits with focus 
on their architecture and verification complexity.

{\bf Addition} in $\Fkk$ is performed by correspondingly adding the
polynomials together and reducing the coefficients of the result modulo $2$.
\begin{Example}
Given $A=\alpha^3+\alpha^2+1=(1101) $ and $B=\alpha^2+1=(0101)$ in $\mathbb{F}_{2^4}$, 
\begin{equation}
A+B=(\alpha^3+\alpha^2+1)+(\alpha^2+1)=(\alpha^3) + (\alpha^2+\alpha^2) +(1+1)=\alpha^3=(1000). \nonumber
 \end{equation}
\end{Example}

Effectively, the addition operation is only performed on the coefficients, 
which are in $\F_2$. 
As addition over $\F_2$ performs an {\it XOR} operation,
constructing an addition circuit over $\Fkk$ is trivial as it
only consists of $k$ number of {\it XOR} gates. 
A $4$-bit adder over $\F_{2^4}$ is shown in
Fig. \ref{fig:adder4}.
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.6]{figures/adder4bit}
\end{center}
\caption{$4$-bit adder over $\mathbb{F}_{2^4}$.}
\label{fig:adder4}
\end{figure}

{\bf Multiplication} $Z =A\times B \pmod{ P(x) }$ in $\Fkk$ conceptually consists of 
two steps.
In the first step, the multiplication $A\times B$ is performed. In the second 
step, the result is reduced
modulo the irreducible polynomial $P(x)$.
This multiplication procedure is shown in Example \ref{exp:mul}.

%\ref{exp1}, 
\begin{Example}
\label{exp:mul}
Consider the field $\mathbb{F}_{2^4}$ with the irreducible polynomial 
$P(x)=x^4+x^3+1$ and $P(\alpha)=0$. We take as inputs:
$A=a_0+a_1\cdot \alpha+a_2\cdot \alpha^2+a_3\cdot \alpha^3$ and
$B=b_0+b_1\cdot \alpha+b_2\cdot \alpha^2+b_3\cdot \alpha^3$. 
We have to perform the multiplication $Z =A\times B \pmod{ P(x) }$. The coefficients
of $A = \{a_0, \dots, a_3\}, B = \{b_0, \dots, b_3\}$ are in
$\mathbb{F}_2 = \{0, 1\}$. This multiplication can be performed as
shown:

{\begin{tabular}{c c c c c c c c}
  &   &   & $a_3$ & $a_2$ & $a_1$ & $a_0$  \\ 
 $\times$&   &   & $b_3$ & $b_2$ & $b_1$ & $b_0$  \\ 
 \hline
 &   &   & $a_3\cdot b_0$ & $a_2 \cdot b_0$ & $a_1\cdot b_0$ & $a_0\cdot b_0$ \\
 &  & $a_3\cdot b_1$ & $a_2\cdot b_1$ & $a_1 \cdot b_1$ & $a_0\cdot b_1$ &   \\
 & $a_3\cdot b_2$ & $a_2\cdot b_2$ & $a_1\cdot b_2$ & $a_0\cdot b_2$ &  &   \\
 $a_3\cdot b_3$ & $a_2\cdot b_3$ & $a_1\cdot b_3$ & $a_0\cdot b_3$ &  &  &   \\
 \hline
 $s_6$& $s_5$  & $s_4$  & $s_3$ & $s_2$  & $s_1$   & $s_0$ 
\end{tabular}}

The result $Sum = s_0+s_1\cdot \alpha + s_2\cdot \alpha^2 + s_3\cdot
\alpha^3 + s_4\cdot \alpha^4 + s_5\cdot \alpha^5 + s_6\cdot \alpha^6$,
where
\begin{eqnarray}
	s_0  &=&  a_0\cdot b_0 \nonumber \\ 
	s_1  &=&  a_0\cdot b_1 + a_1\cdot b_0 \nonumber \\
	s_2  &=&  a_0\cdot b_2 + a_1\cdot b_1 + a_2\cdot b_0 \nonumber \\
	s_3  &=&  a_0\cdot b_3 + a_1\cdot b_2 + a_2\cdot b_2 +  a_3\cdot b_1\nonumber \\
	s_4  &=&  a_1\cdot  b_3 + a_2\cdot b_1 + a_3\cdot b_1 \nonumber \\
	s_5  &=&  a_2\cdot b_3 + a_3\cdot b_2  \nonumber \\
	s_6  &=&  a_3\cdot b_3   \nonumber
\end{eqnarray}
 Here the multiply ``$\cdot$'' and add ``$+$'' operations are performed
modulo 2, so they can be implemented in a circuit using AND and XOR
gates respectively. Note that unlike integer multipliers, there are no carry-chains
in the design, as the coefficients are always reduced modulo $2$. 
However, the result is yet to be reduced modulo the primitive
polynomial $P(x) = x^4 + x^3 + 1$. This transforms every exponent representation, 
$\alpha^d$, to a polynomial representation where $d\geq k=4$.

{\begin{tabular}{|c c c c | l }
 \multicolumn{5}{c}{} \\
 \multicolumn{1}{c}{$\alpha^3$} & $\alpha^2$ & $\alpha$ & \multicolumn{1}{c}{$1$} &  \\
\hhline{----~}
  $s_3$ 	&$s_2$  	&$s_1$   &$s_0$ 	&   \\
 \hline
 $s_4$ 		&$0$		&$0$ 	 &$s_4$  	&$s_4\cdot \alpha^4 \pmod{P(\alpha)} = s_4 \cdot (\alpha^3 + 1)$\\
 $s_5$ 		&$0$		&$s_5$   &$s_5$     &$s_5\cdot \alpha^5 \pmod{P(\alpha)} = s_5\cdot (\alpha^3+ \alpha + 1)$\\
 $s_6$ 		&$s_6$		&$s_6$   &$s_6$     &$s_6\cdot \alpha^6 \pmod{ P(\alpha)} = s_6\cdot( \alpha^3 + \alpha^2 + \alpha + 1)$\\
 \hline
 $z_3$ 		&$z_2$ 		&$z_1$   &$z_0$ 	&\\
 \multicolumn{5}{c}{} \\
 \end{tabular}\par}

The final result (output) of the circuit is: $Z = z_0 + z_1 \alpha + z_2
\alpha^2 + z_3 \alpha^3$; where  $z_0=s_0+s_4+s_5+s_6; ~~z_1=s_1+s_5+s_6;
~~z_2=s_2+s_6; ~~z_3=s_3+s_4+s_5+s_6$. 
\end{Example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The above multiplier design is called the {\it Mastrovito multiplier} \cite{mastro:1989} 
which is the most straightforward way to design a multiplier over $\mathbb{F}_{2^k}$. 
A logic circuit for a $4$-bit {\it Mastrovito} multiplier over {\it Galois field} $\mathbb{F}_{2^4}$ is illustrated in Fig. \ref{fig:mas4}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.60]{figures/mul4bit.eps}
	\end{center}
	\caption{Mastrovito multiplier over $\mathbb{F}_{2^4}$.}
	\label{fig:mas4}
\end{figure}

Modular multiplication is at the heart of many public-key cryptosystems, 
such as Elliptic Curve Cryptography (ECC) \cite{ecc:1986}. 
Due to the very large field size (and hence the data-path width) used in these cryptosystems, 
the above {\it Mastrovito} multiplier architecture is inefficient, especially when 
exponentiation and repeat multiplications are performed.
Therefore, efficient hardware and software implementations of modular multiplication 
algorithms are used to overcome the complexity of such operations. 
One such algorithm which we will focus on is the Montgomery reduction \cite{PT:1985} \cite{acar:1998}.

\subsection{Montgomery Multipliers}
Montgomery Reduction (MR) computes:

\begin{equation}
G=MR(A,B)=A\cdot B \cdot R^{-1} \pmod {P(x)}
\end{equation}
where $A,B$ are $k$-bit inputs, $R={\alpha}^k$, $R^{-1}$ is multiplicative
inverse of $R$ in $\mathbb{F}_{2^k}$, and $P(x)$ is the irreducible polynomial for
$\mathbb{F}_{2^k}$. Since Montgomery reduction cannot directly compute $A\cdot B$, 
we need to pre-compute $A\cdot R$ and $B\cdot R$,
as shown in Fig. \ref{fig:mm4}.

\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=0.50]{figures/mmcircuit}
	\end{center}
	\caption{Montgomery multiplier over $\mathbb{F}_{2^k}$}
	\label{fig:mm4}
\end{figure}

Each $\it MR$ block in Fig. \ref{fig:mm4} represents a Montgomery reduction step 
which is a hardware implementation of the algorithm shown in 
Algorithm \ref{alg:mont}.

\begin{algorithm}
\SetAlgoNoLine

 \KwIn{$A(x), B(x)\in \mathbb{F}_{2^k}$; irreducible polynomial $P(x)$.}
 \KwOut{$G(x)=A(x)\cdot B(x)\cdot x^{-k} \pmod {P(x)}$.}
%%%%%%%%%%%%%%%%%%%%
  $G(x):=$0 \\
  \For { ($i=0$;   $i \le k-1$; ++i ) }
  {
	$G(x):=G(x)+A_i\cdot B(x)$ \CommentSty{/*$A_{i}$ is the $i^{th}$ bit of $A$*/\;}
	$G(x):=G(x)+G_0\cdot P(x)$ \CommentSty{/*$G_{0}$ is the lowest bit of $G$*/\;}
	$G(x):=G(x) / x$ \CommentSty{/*Right shift $1$ bit*/\;}
  }
\caption{Montgomery Reduction Algorithm \cite{acar:1998}}\label{alg:mont}
\end{algorithm}

The design of Fig. \ref{fig:mm4} is not efficient to computing
$A\cdot B \pmod{ P(x)}$ when compared to the Mastrovito implementation.
However, when these multiplications are
performed repeatedly, such as in iterative squaring, then the
Montgomery approach speeds-up the computation. 
As shown in \cite{wu:2002}, the critical path delay and gate counts of a squarer 
designed using the Montgomery approach are much smaller than the traditional 
approaches.

\subsection{Circuit Designs over Composite Fields}
The Galois field $\mathbb{F}_{2^k}$ is a $k$-dimensional vector space over the
sub-field $\mathbb{F}_2$. If $k = m\cdot n$, the field $\mathbb{F}_{2^k}$
can be decomposed as $\mathbb{F}_{(2^m)^n}$. Such a field representation is
called a {\bf composite field}, and it is constructed as a $n$-dimensional 
extension of the sub-field $\mathbb{F}_{2^m}$. The sub-field $\mathbb{F}_{2^m}$ is
called the ground field. Note that we have $\mathbb{F}_2 \subset \mathbb{F}_{2^m}
\subset \mathbb{F}_{(2^m)^n}$.

A Galois field arithmetic circuit over $\Fkk$ can thus be composed
as circuit over $\F_{(2^m)^n}$ if $k=m\cdot n$.
Since the base field is $\F_{2^m}$, this composite field circuit is
composed of blocks of $m$-bit multipliers and adders, along with 
$m$-bit buses that act as the inputs and outputs of these blocks.
A $\F_{2^4}$ Galois field multiplier designed over the composite field $\F_{(2^2)^2}$ is shown in 
Fig. \ref{fig:comp4exPrelim}.
Design methodologies of these circuits are examined more closely in Chapter \ref{ch:generalize}.

\begin{figure}[t]
        \centering
        \includegraphics[width=.9\linewidth]{./figures/compMineSmall}
        \caption{$4$-bit composite multiplier designed over $\F_{(2^2)^2}$}\label{fig:comp4exPrelim}
\end{figure}

%% According to Theorem \ref{the:unique}, there exists an unique field of size $p^{k}$. 
%% This implies that $\mathbb{F}_{2^k}$ is isomorphic to
%% $\mathbb{F}_{(2^m)^n}$ when $k = m\cdot n$, and due to this isomorphism,
%% it is possible to derive one field representation from the other. 
%% The principle of constructing a composite field is described in \cite{phdpaar:1994}. 
%% We derive concrete steps for constructing circuits over composite fields.

%% To construct $\mathbb{F}_{(2^m)^n}$, we require a primitive polynomial of 
%% degree $n$, with coefficients from the ground field $\mathbb{F}_{2^m}$. 
%% Let $k=m\cdot n$. Given $\mathbb{F}_{2^k}$ and its primitive polynomial $P(x)$
%% of degree k, 
%% the primitive polynomial of the composite field can be easily
%% derived. We use the following notation:

%% \begin{itemize}
%% \item Let $P(x)$ denote the given primitive polynomial of general
%%   field $\mathbb{F}_{2^k}$. Let $\alpha$ be a root of the 
%%   this polynomial, i.e. $P(\alpha)=0$.  
%% \item Let $Q(x)$ denote the primitive polynomial of ground field
%%   $\mathbb{F}_{2^m}$. Let $\beta$ be a root of $\mathbb{F}_{2^m}$,
%%   i.e. $Q(\beta)=0$. Note that $Q(x)$ is a degree $m$ primitive
%%   polynomial over $\mathbb{F}_{2}$ so it is also known. 
%% \item Let $R(x)$ denote the primitive polynomial of composite field
%%   $\mathbb{F}_{(2^m)^n}$. Let $\gamma$ be a root,
%%   i.e. $R(\gamma)=0$. This polynomial $R(x)$ has to be derived. 
%% \end{itemize}

%% %To construct $R(x)$, we hae the following result 
%% \begin{Lemma}
%% From \cite{cf:2003}: Let $\mathbb{F}_{2^k}$ be decomposed as $\mathbb{F}_{(2^m)^n}$
%% where $k = m\cdot n$. Let $\gamma$ be the primitive root of the field $\mathbb{F}_{(2^m)^n}$. 
%% Then 
%% \begin{equation}
%% R(x)=\prod_{i=0}^{i=n-1}(x_i+\gamma^{2^{m \cdot i}})
%% \end{equation}
%% \end{Lemma}

%% Since $\mathbb{F}_{2^k}$ is isomorphic to $\mathbb{F}_{(2^m)^n}$, $\alpha$ and
%% $\gamma$ are the same elements ($\alpha=\gamma$).
%% Consider the representation of an element $A$ in $\mathbb{F}_{2^k}$ and its corresponding
%% representation in the composite field.

%% \begin{itemize}
%% \item Any element $A \in \mathbb{F}_{2^k}$ is represented as:
%% \begin{equation}
%% A=\sum_{i=0}^{i=k-1}a_i \cdot \alpha^i, a_i \in \mathbb{F}_{2}, \text{and}\	 P(\alpha) = 0
%% \end{equation}
%% \item The same element $A \in \mathbb{F}_{(2^m)^n}$ is represented as:
%% \begin{equation}
%% A=\sum_{i=0}^{i=n-1}A_i \cdot \gamma^i, A_i \in \mathbb{F}_{2^m}, \text{and} \	 R(\gamma) = 0
%% \end{equation}
%% \item The element $A_i$ needs to be represented in the ground field $\mathbb{F}_{2^m}$:
%% \begin{equation}
%% A_i=\sum_{j=0}^{j=m-1}a_{ij} \cdot \beta^j, a_{ij} \in \mathbb{F}_{2}, \text{and} \	 Q(\beta) = 0
%% \end{equation}
%% \end{itemize}

%% Thus, we need to find the relationship between the primitive roots
%% $\alpha$ and $\beta$ (or between $\gamma$ and $\beta$, since $\alpha =\gamma$), 
%% so as to be able to map the elements from $\mathbb{F}_{2^k}$ to $\mathbb{F}_{(2^m)^n}$. 
%% We use the following result \cite{cf:2003}:

%% \begin{Theorem}\label{thm:gamma}
%% For $\gamma \in \mathbb{F}_{(2^m)^n}$, and $\beta=\gamma^{\omega}$, where $\omega=(2
%% ^{m \cdot n}-1)/(2^m-1)$, then we have $\beta \in \mathbb{F}_{2^m}$. In other
%% words: 
%% \begin{equation}
%% \beta=\alpha^{(2^{m \cdot n}-1)/(2^m-1)}=\gamma^{(2^{m \cdot n}-1)/(2^m-1)} \label{eqn:relation}
%% \end{equation}

%% \end{Theorem}

%% The above result states the following: Since $\gamma$ is a primitive
%% root, it can be used to generate all the non-zero elements of $\mathbb{F}_{(2^m)^n}$. 
%% Moreover, $\beta$ is a primitive root of the ground field $\mathbb{F}_{2^m}$, 
%% which is a sub-field of $\mathbb{F}_{(2^m)^n}$ ( i.e. $\mathbb{F}_{2^m}
%% \subset \mathbb{F}_{(2^m)^n}$); so $\beta \in \mathbb{F}_{(2^m)^n}$. Therefore
%% an exponent of $\gamma$ can be used to generate $\beta$ as
%% $\beta=\gamma^{\omega}$, where $\omega$ is given in Theorem
%% \ref{thm:gamma}. Now that we have all the relationships between $\alpha,
%% \beta, \gamma$, it is possible to perform the decomposition.

%% \begin{Example}
%% Consider the field $\mathbb{F}_{2^4}$ 
%% with $P(x) = x^4 + x^3 + 1$ and $P(\alpha)=0$.
%% In order to decompose it as $\mathbb{F}_{(2^2)^2}$,
%% perform the following steps:

%% \begin{enumerate}
%% \item 
%% Derivation of $R(x)$:
%% \begin{eqnarray}
%% R(x)&=&\prod_{i=0}^{i=1}(x+\gamma^{2^{2 \cdot i}}) \nonumber \\
%% &=&(x+\gamma)\cdot (x+\gamma^{2^2})               \nonumber \\
%% &=&x^2+(\gamma^4+\gamma) \cdot x+\gamma^5          
%% \end{eqnarray}
%% Notice that $R(\gamma) = \gamma^2 + (\gamma^4+\gamma) \cdot \gamma+\gamma^5 =0$.
%% \item 
%% Representation of element $A \in \mathbb{F}_{(2^2)^2}$:
%% \begin{eqnarray}
%% A& = & \sum_{i=0}^{i=1}A_i \cdot \gamma^i, A_i \in \mathbb{F}_{2^2}\nonumber \\
%%  & = & A_0 +A_1 \cdot \gamma
%% \end{eqnarray}
%% \item 
%% Representation of $A_0, A_1$ in $\mathbb{F}_{2^m}$:
%% \begin{eqnarray}
%% A_0=a_{00}+a_{01} \cdot \beta \nonumber \\
%% A_1=a_{10}+a_{11} \cdot \beta
%% \end{eqnarray}
%% where $a_{ij}\in \mathbb{F}_2$. $Q(x)$ can be any degree $m=2$ primitive
%% polynomial in the ground field  $\mathbb{F}_{2^2}$. For the sake of
%% this example, let $Q(x)=x^2+x+1$. 
%% \item Substitute $A_0, A_1$ into $A$:
%% \begin{eqnarray}
%% A&=&\sum_{i=0}^{i=1}(\sum_{j=0}^{j=1}a_{ij} \cdot \beta^j) \cdot \gamma^i \nonumber \\
%% &=&a_{00}+a_{01}\cdot \beta+(a_{10}+a_{11}\cdot \beta)\cdot \gamma   \end{eqnarray}
%% where each $a_{ij} \in \mathbb{F}_2$. From Equation (\ref{eqn:relation}), 
%% $\beta=\alpha^5=\gamma^5$. Substitute $\beta$ and
%% $\gamma$ with $\alpha$ to obtain:
%% \begin{eqnarray}
%% A&=&\sum_{i=0}^{i=1}(\sum_{j=0}^{j=1}a_{ij} \cdot \beta^j) \cdot \gamma^i \nonumber \\
%% &=&a_{00}+a_{01}\cdot \alpha^5+(a_{10}+a_{11}\cdot \alpha^5)\cdot \alpha \nonumber 
%% \end{eqnarray} 
%% Since $P(x)=x^4+x^3+1$ with $P(\alpha)=0$, then
%% \begin{equation}\label{a}
%% A \pmod {P(\alpha)}=a_{00}+a_{01}+a_{11}+(a_{01}+a_{10}+a_{11})\cdot \alpha+a_{1
%% 1} \cdot \alpha^2+(a_{01}+a_{11})\cdot \alpha^3
%% \end{equation}

%% \item The same element $A \in \mathbb{F}_{2^4}$ is represented as:
%% \begin{equation}\label{aa}
%% A=a_0+a_1\cdot \alpha+a_2\cdot \alpha^2+a_3\cdot \alpha^3 
%% \end{equation}

%% \item Since Eqns. \ref{a} and \ref{aa} represent the same element, we
%%   can match the coefficients of the the polynomials to obtain:
%% \begin{eqnarray}
%% a_0&=&a_{00}+a_{01}+a_{11} \nonumber \\
%% a_1&=&a_{01}+a_{10}+a_{11} \nonumber \\
%% a_2&=&a_{11} \nonumber \\
%% a_3&=&a_{01}+a_{11} \nonumber
%% \end{eqnarray}

%% This mapping can also be reversed and represented as a matrix $T$:
%% \begin{center}
%% $\begin{bmatrix} a_{00}\\ a_{01} \\a_{10} \\ a_{11}\end{bmatrix}
%% =
%% \begin{bmatrix} 1 & 0 & 0 & 1\\ 0 & 0 & 1 & 1\\ 0 & 1 & 0 & 1\\ 0 & 0
%%   & 1 & 0 \end{bmatrix} 
%% \begin{bmatrix} a_0\\ a_1 \\a_2 \\ a_3\end{bmatrix}$
%% \end{center}
%% \end{enumerate}

%% Now we have successfully derived the composite field representation
%% $\mathbb{F}_{(2^2)^2}$ from $\mathbb{F}_{2^4}$. The element $A \in \mathbb{F}_{2^4}$ is
%% represented as $A = a_0 + a_1 \alpha + a_2 \alpha^2 + a_3 \alpha^3$,
%% where $P(\alpha) = 0$. The same element $A$ is represented in $\mathbb{F}_{(2^2)^2}$ as:
%% \begin{eqnarray}
%% A&=&A_0+A_1 \cdot \alpha \nonumber \\
%% A_0&=&a_{00}+a_{01} \cdot \alpha^5 \nonumber \\
%% A_1&=&a_{10}+a_{11} \cdot \alpha^5 \nonumber \\
%% a_{00}&=&a_0+a_3 \nonumber \\
%% a_{01}&=&a_2+a_3 \nonumber \\
%% a_{10}&=&a_1+a_3 \nonumber \\
%% a_{11}&=&a_2 \nonumber 
%% \end{eqnarray}

%% In the above equations, $\alpha = \gamma$ and $R(\gamma) = 0$. 
%% \end{Example}

%% A Galois field multiplier circuit over $\Fkk$ can thus be composed
%% as multiplier over $\F_{(2^m)^n}$ if $k=m\cdot n$.
%% Since the base field is $\F_{2^m}$, this composite field circuit is
%% composed of blocks of $m$-bit multipliers and adders, along with 
%% $m$-bit buses that act as the inputs and outputs of these blocks.

%% \begin{figure}[h!]
%% \centerline{
%% \includegraphics[width=\linewidth]{./figures/compMineSmall.eps}
%% }
%% \caption{Galois field multiplier over the composite field $\mathbb{F}_{(2^2)^2}$}
%% \label{fig:mas22}
%% \end{figure}

%% \begin{Example}
%% Consider the $4$-bit Mastrovito multiplier circuit over $\F_{2^4}$
%% shown in Figure \ref{fig:mas4}, where $A$ and $B$ are the $4$-bit 
%% word-level inputs and $\{a_0,a_1,a_2,a_3,b_0,b_1,b_2,b_3\}$ are
%% primary inputs, thus
%% \begin{eqnarray}
%% A=a_0+a_1\alpha+a_2\alpha^2+a_3\alpha^3 \nonumber \\
%% B=b_0+b_1\alpha+B_2\alpha^2+B_3\alpha^3 \nonumber
%% \end{eqnarray}

%% Since the field $\F_{2^4}$ can be decomposed as $\F_{(2^2)^2}$, 
%% this multiplier can be constructed as a composite field multiplier
%% over $\F_{(2^2)^2}$ with the base field $\F_{2^2}$.
%% Such a multiplier is shown Figure \ref{fig:mas22} and
%% is represented as follows:
%% \begin{eqnarray}
%% A = A_0+A_1\gamma \nonumber \\
%% B = B_0+B_1\gamma \nonumber \\
%% A_0=a_{00}+a_{01} \cdot \beta \nonumber \\
%% A_1=a_{10}+a_{11} \cdot \beta \nonumber \\
%% B_0=b_{00}+b_{01} \cdot \beta \nonumber \\
%% B_1=b_{10}+b_{11} \cdot \beta \nonumber
%% \end{eqnarray}
%% where $A_0,A_1,B_0,B_1$ are the $2$-bit inputs $\in \F_{2^2}$
%% composed of $a_{00},a_{01},a_{10},a_{11},b_{00},b_{01},b_{10},b_{11}$, 
%% which are the bit-level inputs to the composite field multiplier
%% derived from the transformation of $\{a_0,a_1,a_2,a_3,b_0,b_1,b_2,b_3\}$.
%% Correspondingly, each block in Figure \ref{fig:mas22}
%% internally represents a $2$-bit operation: $\times$ represents $2$-bit
%% {\it multiplication} and $+$ represents $2$-bit {\it addition} over
%% the ground field.
%% \end{Example}

%\afterpage{%
%    \clearpage% Flush earlier floats (otherwise order might not be correct)
%    \thispagestyle{empty}% empty page style (?)
%	\global\pdfpageattr\expandafter{\the\pdfpageattr/Rotate 90}
%    \begin{landscape}% Landscape page

%\begin{sidewaysfigure}
%\begin{figure}[h!]
%\centerline{
%\includegraphics[width=.9\linewidth]{./figures/compMineSmall.eps}
%}
%\caption{Mastrovito multiplier over the composite field $\mathbb{F}_{(2^2)^2}$}
%\label{fig:mas22}
%\end{figure}
%\end{sidewaysfigure}
%  
%    \end{landscape}
%    \clearpage% Flush page
%	\global\pdfpageattr\expandafter{\the\pdfpageattr/Rotate 0}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Applications to Elliptic Curve Cryptography}
Elliptic Curve Cryptography (ECC) is one of the most influential applications 
of Galois fields.
ECC is an approach to public-key (or asymmetric-key) cryptography based on the 
algebraic structure of elliptic curves over Galois fields. Due to the complex 
nature of these curves, key sizes in ECC can be smaller than other public-key
cryptography techniques while providing the same level of security \cite{ecc:book}.
The main operations of encryption, decryption and authentication in ECC 
rely on {\it point multiplications}.

Point multiplication involves a series of addition and doubling of points on the 
elliptic curve.
A drawback of traditional point multiplication is that each point addition and 
doubling require a multiplicative inverse operation over Galois fields, 
the computation of which is costly.
Modern methods, however, represent the points in 
projective coordinate systems \cite{ecc:software},
which has eliminated the need for a multiplicative inverse operation by replacing it 
with addition and multiplication operations over Galois fields.
This has increased the efficiency of point multiplication operations, but it has also
increased the need for fast, custom hardware designs of Galois field arithmetic.

In-depth analysis of elliptic curve theory is beyond the scope of this 
dissertation. Instead, we will look at some examples of point addition and point doubling to
give a general idea of the operations involved in ECC and how they apply to Galois 
field arithmetic. Our experiments use custom Galois field arithmetic designs based 
on L$\acute{o}$pez-Dahab (LD) coordinate system \cite{eccld}, so these examples will
use the same coordinate system.

\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=0.50]{figures/ecc}
	\end{center}
	\caption{Point addition over an Elliptic Curve (R=P+Q)}
	\label{fig:ecc}
\end{figure}

\begin{Example}
Consider point addition in a LD projective coordinate system, 
as seen in Fig. \ref{fig:ecc}.

Given an elliptic curve: $Y^2 + XYZ = X^3Z + aX^2Z^2 + bZ^4$ over $\mathbb{F}_{2^k}$, 
where $X,Y,Z$ are $k$-bit vectors
that are elements in $\mathbb{F}_{2^k}$ and similarly, 
$a, b$ are constants from the field. 
Let $P + Q = R$ represent point addition over the elliptic curve.
$P=(X_1, Y_1, Z_1)$ and $Q=(X_2, Y_2, 1)$ are given.
Then $R=(X_3, Y_3, Z_3)$ can be computed as follows:

\begin{align*}
A &= Y_2 \cdot Z_1^2 + Y_1 \\
B &= X_2 \cdot Z_1 + X_1 \\
C &= Z_1 \cdot B \\
D &= B^2 \cdot(C + a Z_1^2) \\
Z_3 &= C^2 \\
E &= A \cdot C  \\
X_3 &= A^2 + D + E  \\
F &= X_3 + X_2 \cdot Z_3 \\
G &= X_3 + Y_2\cdot Z_3 \\
Y_3 &= E\cdot F + Z_3 \cdot G \\
\end{align*}
\end{Example}

\begin{Example}
Consider point doubling in a LD projective coordinate system. 
Given an elliptic curve: $Y^2 + XYZ = X^3Z + aX^2Z^2 + bZ^4$. 
Let 2($X_1$, $Y_1$, $Z_1$) = ($X_3$, $Y_3$, $Z_3$), then

\begin{align*}
X_3 &= X_1^4 + b \cdot Z_1^4  \\
Z_3 &= X_1^2 \cdot Z_1^2 \\
Y_3 &= b Z_1^4 \cdot Z_3 + X_3 \cdot (aZ_3 + Y_1^2 + bZ_1^4 ) \\
\end{align*}
\end{Example}

In the above examples, polynomial multiplication and squaring operations can be 
implemented in hardware using Montgomery reductions over Galois fields $\mathbb{F}_
{2^k}$. In practical applications, the field size $k$ of $\mathbb{F}_{2^k}$ is $163$,
or larger. However, there are no word-level 
abstraction techniques applicable to circuits of such size, so hardware 
implementations of Galois field arithmetic circuits 
cannot benefit from the many advantages of abstraction.
Thus, we propose a computer-algebra approach to word-level polynomial abstractions 
of Galois field arithmetic circuits.
Recent computer-algebra formal verification techniques \cite{lv:phd} 
have been able to verify these circuits up to $163$ bits. 
We propose an application of our abstraction approach to improve these techniques. 
These improvements allow us to perform formal verification of these
circuits up to $571$ bits. These proposals are described in detail in 
subsequent chapters.

### 2016-09-21 dissertation/implement.tex
init dissertation writing

\chapter{Implementation of the Custom Abstraction Software and Experimental Results} \label{ch:implement}

The abstraction procedure can be fully scripted using the computer algebra tool 
{\sc Singular}. However, {\sc Singular} has limitations which make abstraction
of large circuits {\it impossible}. This is due to:
\begin{itemize}
\item A limit on the number of ring variables allowed. As of {\sc Singular} release 4.0.1, 
a ring cannot be declared with more than $32,767$ variables. This limits the number
of gates that can be present in the design.
\item The size of an exponent $(n)$ of a variable $x$, is limited $n<2^{32}$.
\item Large amount of memory usage and slow computation time. 
{\sc Singular} uses a dense-distributive structure for polynomials, which is a poor 
representation of sparse polynomials over rings with many variables.
\end{itemize}
The limit on the size of exponents prohibits an abstraction beyond $32$-bit circuits,
as larger circuits require manipulating word-level variables with exponents 
larger than $2^{32}$. Even
if this limitation is overcome, Singular uses an enormous amount of memory.
For instance, preliminary experiments show that using Singular to compute the
initial reduction of a $163$-bit Mastrovito multiplier uses $41.6$ GB of memory!
Thus, deriving abstractions using Singular is infeasible on 
desktop workstations.

In order to overcome these limitations, a custom C++ tool is developed to 
compute the word-level abstraction of circuits quickly and efficiently. 
This chapter describes the implementation details of this tool.

\section{Data Structures and Algorithms}

Computing a word-level abstraction requires the representation and manipulation 
of polynomials in $\Fkk[x_1,\dots,x_d]$ over a lex ordered ring. 
Intermediate polynomials can become 
very large during the abstraction procedure, so the data structure to represent them 
is designed to 
conserve memory while allowing for fast manipulation during the reduction 
procedures. The backbone of the tool is a custom library composed of three main
sections:
\begin{enumerate}
\item Galois field elements
\item Monomials and rings
\item Polynomials and division procedures
\end{enumerate}
These are built on top of each other. The starting point is the custom 
Galois field element section, which facilitates the construction of 
Galois field elements over $\Fkk$.

\subsection{Galois Field Elements}

The Galois field section of the library is initialized by parsing a given primitive polynomial $P(x)$
of degree $k$ which constructs $\Fkk$. Any element $C \in \Fkk$ can be represented 
in the form 
\begin{equation}
C = c_{k-1}\cdot\alpha^{k-1}+c_{k-2}\cdot\alpha^{k-2}+\cdots+c_2\cdot\alpha^2+c_1\cdot\alpha+c_0
\end{equation}
where 
$\{c_0,\dots,c_{k-1}\}\in\F_2$ and $\alpha$ is the primitive element. This {\it structure}
is stored as an unsigned byte array containing $\{c_0,\dots,c_{k-1}\}$, 
as shown in Figure \ref{fig:gfStruct}. Thus, each element uses $\left\lfloor\frac{k-1}{8}\right\rfloor+1$ 
bytes of memory. 
Any leading bits after $c_{k-1}$ in the last byte are set to $0$.
\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=1]{figures/gfElementStructure}
	\end{center}
	\caption{Object structure of a Galois field element}
	\label{fig:gfStruct}
\end{figure}

{\it Addition} between two elements
\begin{eqnarray}
C&=&c_{k-1}\cdot\alpha^{k-1}+\cdots+c_2\cdot\alpha^2+c_1\cdot\alpha+c_0 \\
D&=&d_{k-1}\cdot\alpha^{k-1}+\cdots+d_2\cdot\alpha^2+d_1\cdot\alpha+d_0
\end{eqnarray}
is simply a combination of like terms
\begin{equation}
C+D=(c_{k-1}+d_{k-1})\cdot\alpha^{k-1}+\cdots+(c_2+d_2)\cdot\alpha^2+(c_1+d_1)\cdot\alpha+(c_0+d_0)
\end{equation}
Since addition over $\F_2$ is computed as a bit-wise XOR, the library's Galois field element 
structure allows addition to be trivially performed as a byte-wise XOR operation. 
Furthermore, this structure makes it easy to check if a given element is equal 
to $0$ or $1$, which is used in deciding when a term is to be removed ($0$) and when 
a division can be ignored ($1$).

During {\it library initialization}, the elements 
$\alpha^k,\alpha^{k+1},\dots,\alpha^{2k-2}$ are pre-computed and cached. First,
$\alpha^k$ is derived directly from the given primitive polynomial, 
\begin{equation}
P(x)=x^k+c_{k-1}\cdot x^{k-1}+\cdots+c_1\cdot x+1
\end{equation}
for $\{c_1,\dots,c_{k-1}\}\in\F_2$. Since $P(\alpha)=0$,
\begin{equation}
\alpha^k=c_{k-1}\cdot\alpha^{k-1}+\cdots+c_1\cdot\alpha+1
\end{equation}
To compute $\alpha^{k+1}$, first compute a $1$-bit left shift of
$\alpha^k$.
\begin{equation}
\alpha^{k+1}=c_{k-1}\cdot\alpha^k+c_{k-2}\cdot\alpha^{k-1}+\cdots+c_1\cdot\alpha^2+\alpha
\end{equation}
This element can contain the term $\alpha^k$, which must be minimized by the primitive
polynomial.
Thus, if $c_{k-1}$ is $1$, the $c_{k-1}\alpha^k$ term is removed and the minimized 
form of $\alpha^k$ is added. This derives the minimized form for $\alpha^{k+1}$. 
Computation continues in this fashion (shift by 1, minimize if needed) until all
$\alpha^k,\alpha^{k+1},\dots,\alpha^{2k-2}$ have been derived.
These elements are later used during the multiplication procedure.

\begin{Example}
\label{ex:gf4init}
Given the primitive polynomial $P(x)=x^4+x^3+1$, initialize the library by computing
$\alpha^4,\alpha^5,$ and $\alpha^6$. Here, $k=4$, and each element of $\F_{2^4}$ can be 
represented as
\begin{equation}
c_3\cdot\alpha^3+c_2\cdot\alpha^2+c_1\cdot\alpha+c_0
\end{equation}
which is stored as one byte in the following form.
\begin{equation}
\begin{tabular}{|c|c|c|c|c|c|c|c|} 
\hline
$0$ & $0$ & $0$ & $0$ & $\mathbf{c_3}$ & $\mathbf{c_2}$ & $\mathbf{c_1}$ & $\mathbf{c_0}$ \\
\hline
\end{tabular}
\end{equation}
Notice that there are $4$ leading bits which are unused; these are always set to $0$.

$P(\alpha)=\alpha^4+\alpha^3+1=0$. Hence, $\alpha^4=\alpha^3+1$, 
which is stored as
\begin{equation}
\begin{tabular}{cc|c|c|c|c|c|c|c|c|} 
\multicolumn{6}{c}{}&\multicolumn{1}{c}{$c_3$}&\multicolumn{1}{c}{$c_2$}&\multicolumn{1}{c}{$c_1$}&\multicolumn{1}{c}{$c_0$} \\
\hhline{~~--------}
$\alpha^4$&$=$&$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{0}$ & $\mathbf{1}$ \\
\hhline{~~--------}
\end{tabular}
\end{equation}
To compute $\alpha^5$, take the $\alpha^4$ element and
shift the result left 1 bit. The $c_3$ term is dropped since the leading $4$ 
bits are always $0$.
\begin{equation}
\begin{tabular}{|c|c|c|c|c|c|c|c|} 
\hline
$0$ & $0$ & $0$ & $0$ & $\mathbf{0}$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{0}$ \\
\hline
\end{tabular}
\end{equation}

Then, since $c_3$ was $1$, add $\alpha^4$.
\begin{equation}
\begin{tabular}{c|c|c|c|c|c|c|c|c|} 
\hhline{~--------}
&$0$ & $0$ & $0$ & $0$ & $\mathbf{0}$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{0}$ \\
\hhline{~========}
$+$&$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{0}$ & $\mathbf{1}$ \\
\hhline{~--------}\multicolumn{9}{c}{}\\
\hhline{---------}\multicolumn{9}{c}{}\\
\hhline{~--------}
&$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{1}$ \\
\hhline{~--------}
\end{tabular}
\end{equation}
This gives $\alpha^5=\alpha^3+\alpha+1$. Similarly, $\alpha^6$ is derived as
\begin{equation}
\begin{tabular}{c|c|c|c|c|c|c|c|c|} 
\hhline{~--------}
&$0$ & $0$ & $0$ & $0$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{0}$ \\
\hhline{~========}
$+$&$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{0}$ & $\mathbf{1}$ \\
\hhline{~--------}\multicolumn{9}{c}{}\\
\hhline{---------}\multicolumn{9}{c}{}\\
\hhline{~--------}
&$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{1}$ \\
\hhline{~--------}
\end{tabular}
\end{equation}
\end{Example}

{\it Multiplication} requires temporarily increasing the size of the byte-array to 
store the intermediate result, which can have values up to $\alpha^{2(k-1)}$.
\begin{equation}
c_{2k-2}\cdot\alpha^{2k-2}+\cdots+c_k\cdot\alpha^k+c_{k-1}\cdot\alpha^{k-1}+\cdots+c_2\cdot\alpha^2+c_1\cdot\alpha+c_0
\end{equation}
This result needs to be divided by the given minimum polynomial. Each $\alpha$ 
term with an exponent of $k$ or larger is replaced by its minimized equivalent,
which was computed during initialization. That is, for $i\geq k$, each term for which $c_{i}=1$ 
is removed and the minimized form of $\alpha^i$ added in.

\begin{Example}
\label{ex:gf4mult}
Consider again the setup for $\F_{2^4}$ from Example \ref{ex:gf4init}. Compute the
product of the following two elements:
\begin{eqnarray}
\alpha^3+\alpha^2+1 \\
\alpha^2+\alpha
\end{eqnarray}
These elements are stored respectively as
\begin{eqnarray}
\begin{tabular}{|c|c|c|c|c|c|c|c|} 
\hline
$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{1}$ \\
\hline
\end{tabular} \\
\begin{tabular}{|c|c|c|c|c|c|c|c|} 
\hline
$0$ & $0$ & $0$ & $0$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{0}$ \\
\hline
\end{tabular}
\end{eqnarray}
The intermediate result is computed using the basic shift-and-add procedure.
\begin{equation}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hhline{~~~~----}
\multicolumn{3}{c}{} & & $1$ & $1$ & $0$ & $1$ \\
\hhline{~~~~----}
\multicolumn{3}{c}{} & x & $0$ & $1$ & $1$ & $0$ \\
\hhline{~~~~----}\multicolumn{8}{c}{}\\
\hhline{--------}\multicolumn{8}{c}{}\\
\hhline{~~~~----}
\multicolumn{3}{c}{} & & $0$ & $0$ & $0$ & $0$ \\
\hhline{~~~-----}
\multicolumn{2}{c}{} & & $1$ & $1$ & $0$ & $1$ & \multicolumn{1}{c}{} \\
\hhline{~~-----~}
\multicolumn{1}{c}{} & & $1$ & $1$ & $0$ & $1$ & \multicolumn{2}{c}{} \\
\hhline{~-----~~}
\multicolumn{1}{c|}{$+$} & $0$ & $0$ & $0$ & $0$ & \multicolumn{3}{c}{} \\
\hhline{~----~~~}\multicolumn{8}{c}{}\\
\hhline{--------}\multicolumn{8}{c}{}\\
\hhline{--------}
$0$& $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{0}$ \\
\hline
\end{tabular}
\end{equation}
The intermediate result of the multiplication is $\alpha^5+\alpha^3+\alpha^2+\alpha$, 
which needs to be further minimized. The value of $\alpha^5$ was determined during
initialization to be $\alpha^3+\alpha+1$. The $\alpha^5$ term from the intermediate
result is removed and the minimized form is added.

\begin{equation}
\begin{tabular}{cc|c|c|c|c|c|c|c|c|} 
\hhline{~~--------}
&&$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{0}$ \\
\hhline{~~========}
&$+$&$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{1}$ \\
\hhline{~~--------}\\
\hhline{----------}\\
\hhline{~~--------}
&&$0$ & $0$ & $0$ & $0$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{1}$ \\
\hhline{~~--------}
\end{tabular}
\end{equation}

So the minimized result of the product is $\alpha^2+1$.

\end{Example}

{\it Division} of two Galois field elements, $C=\frac{B}{A}$, requires finding 
the multiplicative inverse of the divisor: $C=B\cdot A^{-1}$. To find the inverse, 
the library implements the extended Euclidean algorithm over $\Fkk$, depicted in
Algorithm \ref{alg:eucInv}. The algorithm requires a non-minimized 
representation of the element $P(\alpha)$, so the size of object is temporarily 
increased to allow the storage of the $\alpha^k$ bit.
The function $DIV$ returns the quotient and 
remainder of a Euclidean division; that is, $DIV(A,B)$ returns 
$\{Q,R\}$, where $A=B\cdot Q + R$. This procedure is described in Algorithm \ref{alg:eucDiv};
here, $DEG$ returns the highest degree of a given element in $\Fkk$, 
i.e. $DEG(\alpha^4+\alpha^3+1)$ would return $4$.

\begin{algorithm}[hbt]
\SetAlgoNoLine

 \KwIn{$M:=P(\alpha)$ where $P(x)$ was used to generate $\Fkk$, $A\in\Fkk$}
 \KwOut{$A^{-1}$ over $\Fkk$}
  %%%%%%%%%%%%%%%%%%%%
        $\{Q_0,Q_1\}$ := $\{0,0\}$\;
        $\{R_0,R_1\}$ := $\{M,A\}$\;
        $\{U_0,U_1\}$ := $\{0,1\}$\;
        $i$ := $1$\;
        \While { $R_i \neq 1$ }
        {
                \If { $R_i == 0$ }{{\bf ERROR}: No inverse exists}
                $\{Q_{i+1},R_{i+1}\} := DIV(R_{i-1},R_i)$\;
                $U_{i+1}$ := $(Q_{i+1} \cdot U_i) + U_{i-1}$\;
                $i$ := $i+1$\;
        }
        \Return $U_i$\;
\caption{Inverse of an element over $\Fkk$}\label{alg:eucInv}
\end{algorithm}

\begin{algorithm}[hbt]
\SetAlgoNoLine

 \KwIn{$A,B \in \Fkk$}
 \KwOut{$\{Q,R\}$ such that $A = B\cdot Q + R$}
        $\{Q, R\} := \{0,A\}$\;
        %$d := DEG(B)$\;
        \While { $DEG(R) \geq DEG(B)$ }
        {
                $S$ := $\alpha^{DEG(R)-DEG(B)}$\;
                $Q$ := $Q + S$\;
                $R$ := $R + S\cdot B$\;
        }
        \Return $\{Q,R\}$\;
\caption{$DIV$ (Euclidean Division over $\Fkk$)}\label{alg:eucDiv}
\end{algorithm}

\begin{Example}
Given $P(x)=x^8+x^4+x^3+x+1$ which generates $\F_{2^8}$, find $A^{-1}$ where
$A=\alpha^6+\alpha^4+\alpha+1$. Table \ref{tab:divExTab} shows the steps 
Algorithm \ref{alg:eucDiv} goes through to find the inverse.

\begin{table}[h]
\begin{center}
\caption{Steps to derive the inverse of $\alpha^6+\alpha^4+\alpha+1$ }
\label{tab:divExTab}
%\noindent\resizebox{\textwidth}{!}{%
\begin{tabular}{|c|l|l|l|} 
\hline
$\mathbf{i}$ & \multicolumn{1}{c|}{$\mathbf{Q_i}$} & \multicolumn{1}{c|}{$\mathbf{R_i}$} & \multicolumn{1}{c|}{$\mathbf{U_i}$} \\
\hline
$0$ & $0$                 & $\alpha^8+\alpha^4+\alpha^3+\alpha+1$ & $0$                                 \\
\hline
$1$ & $0$                 & $\alpha^6+\alpha^4+\alpha+1$          & $1$                                 \\
\hline
$2$ & $\alpha^2+1$        & $\alpha^2$                            & $\alpha^2+1$                        \\
\hline
$3$ & $\alpha^4+\alpha^2$ & $\alpha+1$                            & $\alpha^6+\alpha^2+1$               \\
\hline
$4$ & $\alpha$            & $1$                                   & $\alpha^7+\alpha^6+\alpha^3+\alpha$ \\
\hline
\end{tabular}
%}
\end{center}
\end{table}

The derived inverse is $A^{-1}=\alpha^7+\alpha^6+\alpha^3+\alpha$. Correctness
can be checked by computing $A \cdot A^{-1}$ and verifying that the result is $1$.
\end{Example}

\subsection{Rings and Monomials over Galois Fields}

A monomial $M$ over the ring $\Fkk[x_1,\dots,x_d]$ is a power-product of variables from 
the ring along with a coefficient $C \in \Fkk$.
\begin{equation}
M = C \cdot {x_1}^{e_1} \cdot {x_2}^{e_2} \cdots {x_d}^{e_d}; \quad e_i \geq 0
\end{equation}
Ring variables can either be bit-level (representing a single wire within a circuit) or 
word-level (representing a word input or output).
If $x_i$ is a bit-level variable then $x_i \in \F_2$; thus it has the property 
$x_i^2=x_i$, so its exponent $e_i \in \{0,1\}$. 
If the variable is word-level, then $e_i < 2^k$ due to the property $x_i^{2^k}=x_i$.

Lex ordering is the only monomial ordering used for abstraction, and hence it is
the only ordering implemented in the tool. Ring variables are added as strings,
one at a time, along with an argument stating whether the variable is bit-level
or word-level. Each variable is given a unique unsigned integer id, which is 
continuously incremented with each added variable. Thus, ids of two variables 
can be compared to quickly distinguish which variable appears earlier in the 
ordering.

Three static objects are created during initialization of the 
ring:
\bi
\item {\bf strToId} - a map of each variable name (string) to its id (unsigned int)
\item {\bf idToStr} - a map of each id (unsigned int) to its variable name (string)
\item {\bf wordSet} - a set of ids (unsigned int) of all variables which are word-level
\ei
Here, ``map'' and ``set'' are classes of the standard C++ library.
It's important to note that a C++ ``set'' is a container of unique, ordered 
elements (this property is exploited later).
The ``strToId'' map object
is used when constructing monomials to quickly find the id of a parsed variable 
name. The ``idToStr'' map object allows a monomial object to be printed to
the user. The ``wordSet'' object is used to quickly check whether a given variable
is word-level, which determines how it is handled during monomial operations.

Once the ring has been initialized, monomials can be generated and manipulated.
Internally, all monomial variables are manipulated using their ids. 
Each monomial object contains the following:
\bi
\item {\bf coef} - a Galois field object (as described in the previous subsection)
\item {\bf idSet} - a set of ids (unsigned int) of all variables in the monomial
\item {\bf idToExp} - a map of variable ids (unsigned int) to their exponents (BigUnsigned)
\ei
During monomial creation, string variable names are parsed and mapped to their 
corresponding ids, which are then added to the set.
The exponent map is only filled for variables that are word-level. As most 
variables are bit-level in a circuit, most monomials will have a completely empty
map. Since exponents can be much larger than what can be stored in a primitive 
data structure, each exponent is stored as a BigUnsigned object of the open 
source library {\it BigInt} \cite{BigInt}. This is a library which provides
basic functionality for signed and unsigned integers of unbounded size.

{\it Monomial comparison} is required for proper monomial ordering, which is
necessary for implementation of polynomial procedures such as reduction.
The comparison procedure compares the id sets of two monomials, one variable at 
a time. If the variables differ, the smaller id appears earlier in the ordering.
If they are the same, exponents are checked if the variable is word-level. 
This procedure is shown in Algorithm \ref{alg:monComp}.

\begin{algorithm}[hbt]
\SetAlgoNoLine

\KwIn{Monomials $M_1$ and $M_2$}
\KwOut{$<0$ if $M_2>M_1$, $>0$ if $M_1>M_2$, $0$ if $M_1==M_2$}
    id$_1$ := $M_1$.idSet.begin()\;
    id$_2$ := $M_2$.idSet.begin()\;
    \While{id$_1 \neq \varnothing$ $\&\&$ id$_2 \neq \varnothing$}
    {
        \If{id$_1 \neq$ id$_2$}
        {
            \Return id$_2$-id$_1$\;
        }
        \If{id$_1 \in $ wordSet}
        {
            \If{$M_1$.idToExp[id$_1$] $\neq$ $M_2$.idToExp[id$_2$]}
            {
                \Return $M_1$.idToExp[id$_1$] - $M_2$.idToExp[id$_2$]\;
            }
        }
        id$_1$ := $M_1$.idSet.next()\;
        id$_2$ := $M_2$.idSet.next()\;
    }
    \If{id$_1$ == $\varnothing$ $\&\&$ id$_2$ == $\varnothing$}{\Return $0$\;}
    \If{id$_1$ == $\varnothing$}{\Return $-1$\;}
    \Return $1$\;
\caption{Monomial Comparison}\label{alg:monComp}

\end{algorithm}
%\begin{enumerate}
%\item Compare the current ids.
%\begin{enumerate}
%\item If they are different, monomial with the smaller id appears earlier.
%\end{enumerate}
%\item If the current id is word-level, compare exponents.
%\begin{enumerate}
%\item If they are different, monomial with the greater exponent appears earlier.
%\end{enumerate}
%\item Continue to next id in each set.
%\begin{enumerate}
%\item If exactly one set has runs out of ids, the other monomial appears earlier.
%\item If both sets have run out of ids, the monomials are equal.
%\item Otherwise, go back to Step 1.
%\end{enumerate}
%\end{enumerate}

{\it Multiplication} of two monomials is the main function of this portion of
the tool. First, the two Galois field objects are multiplied together using the previously described
method. Then, the two sets of ids are merged together. Since sets can only 
contain unique values, duplicates are discarded; this is done automatically using
the standard set::insert operation. 
%Note that if a monomial is squared, this operation can be ignored.
In the common case, both 
monomials only contain bit-level variables and the multiplication would be
complete. If there are word-level variables in the monomials, the mapped 
exponents of each such variable would be added together and then minimized if the
resulting exponent is $\geq2^k$.

\begin{Example}
\label{ex:gf4MonMult}
Consider again the setup for $\F_{2^4}$ from Example \ref{ex:gf4init}.
Construct the ring $\F_{2^4}[a,b,c,Z]$ with the lex ordering 
$a > b > c > Z$, where $\{a,b,c\}$ are bit-level variables and $Z$ is a 
word-level variable. The initialized monomial static library objects are:
\begin{equation}
\begin{tabular}{lcrc|clcrc|cc} 
\multicolumn{3}{c}{\bf strToId} & & & \multicolumn{3}{c}{\bf idToStr} & & & {\bf wordSet} \\
\hline
``$a$'' & $\rightarrow$ & $0$ & & & $0$ & $\rightarrow$ & ``$a$'' & & & \\
``$b$'' & $\rightarrow$ & $1$ & & & $1$ & $\rightarrow$ & ``$b$'' & & & $\{3\}$\\
``$c$'' & $\rightarrow$ & $2$ & & & $2$ & $\rightarrow$ & ``$c$'' & & & \\
``$Z$'' & $\rightarrow$ & $3$ & & & $3$ & $\rightarrow$ & ``$Z$'' & & &
\end{tabular}
\end{equation}
Let $M_1$ and $M_2$ be the following monomials:
\begin{eqnarray}
M_1=(\alpha^3+\alpha^2+1)a b Z^{10} & & M_2=(\alpha^2+\alpha) b c Z^7
\end{eqnarray}
These are stored internally by the tool as:
\begin{equation}
\begin{tabular}{|c|c|c|c|c|c|c|c|cccc} 
\multicolumn{12}{c}{$\mathbf{M_1}$} \\
\multicolumn{8}{c}{\bf coef} & & {\bf idSet} & & {\bf idToExp}\\
\hhline{--------~~~~}
$0$ & $0$ & $0$ & $0$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{1}$ & & $\{0, 1, 3\}$ & & $3 \rightarrow 10$ \\
\hhline{--------~~~~}
\multicolumn{12}{c}{}\\
\end{tabular}
\end{equation}
\begin{equation}
\begin{tabular}{|c|c|c|c|c|c|c|c|cccc} 
\multicolumn{12}{c}{$\mathbf{M_2}$} \\
\multicolumn{8}{c}{\bf coef} & & {\bf idSet} & & {\bf idToExp}\\
\hhline{--------~~~~}
$0$ & $0$ & $0$ & $0$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{1}$ & $\mathbf{0}$ & & $\{1, 2, 3\}$ & & $3 \rightarrow 7$ \\
\hhline{--------~~~~}
\end{tabular}
\end{equation}
It's easy to see that $M_1 > M_2$ in the given ordering, since the 
first element of ``idSet'' in $M_1$ is $0$ while in $M_2$ it is $1$.
Multiplying $M_1$ by $M_2$ is computed by first multiplying the Galois field 
elements (coef) together (as shown in Example \ref{ex:gf4mult}). Then, the two
sets (idSet) are merged (union). 
Notice that, although variable $b$ appears in both monomials, $b^2=b$ due to it
being a bit-level variable. This is handled automatically by the set class 
(duplicates thrown out).
Finally, the two corresponding exponents of variable $Z$ (idToExp) are added together. 
Since this new exponent of $Z$ is $17$, and $17\geq 2^4$, the exponent is 
minimized by the property $Z^{16}=Z$. Thus, the new exponent of $Z$ is $2$.
\begin{equation}
\begin{tabular}{|c|c|c|c|c|c|c|c|cccc} 
\multicolumn{12}{c}{$\mathbf{M_1}\cdot\mathbf{M_2}$} \\
\multicolumn{8}{c}{\bf coef} & & {\bf idSet} & & {\bf idToExp}\\
\hhline{--------~~~~}
$0$ & $0$ & $0$ & $0$ & $\mathbf{0}$ & $\mathbf{1}$ & $\mathbf{0}$ & $\mathbf{1}$ & & $\{0, 1, 2, 3\}$ & & $3 \rightarrow 2$ \\
\hhline{--------~~~~}
\end{tabular}
\end{equation}
So $M_1\cdot M_2 = (\alpha^2+1) a b c Z^2$
\end{Example}

{\it Monomial division} is a procedure mainly used during polynomial reduction. 
Given two monomials 
$M_1$ and $M_2$, compute $\frac{M_1}{M_2}$. That is, find a monomial $M_3$ 
such that $M_1 = M_2 \cdot M_3$.
Monomial division is described in Algorithm \ref{alg:monDiv}. The 
division procedure loops over all variables in $M_2$ and removes them from $M_1$ 
if they are bit-level.
For word-level variables, exponents are subtracted from each other.
If a variable exists in $M_2$ but not in $M_1$, or if the exponent of a 
word-level variable in $M_2$ is larger than in $M_1$, the division is $0$.
Finally at the end, the Galois field elements are divided by each other.

\begin{algorithm}[hbt]
\SetAlgoNoLine

 \KwIn{Monomials $M_1$ and $M_2$}
 \KwOut{$\frac{M_1}{M_2}$ if it exists}
        $M_3$ := $M_1$\;
        \ForEach{id $\in M_2$.idSet}
        {
            \If{id $\notin M_3$ }{\Return NULL\;}
            \eIf{id $\notin$ wordSet}
            {
                $M_3$.idSet.erase(id)\;
            }
            {
                \If{$M_2$.idToExp[id] $> M_3$.idToExp[id]}
                {
                    \Return NULL\;
                }
                $M_3$.idToExp[id] -= $M_2$.idToExp[id]\;
                \If{ $M_3$.idToExp[id] == $0$}{$M_3$.idSet.erase(id)\;}
            }
        }
        $M_3$.coef /= $M_2$.coef\;
        \Return $M_3$\;
\caption{Monomial Division}\label{alg:monDiv}
\end{algorithm}

\subsection{Polynomials and Polynomial Division}
With the monomial structure defined, a polynomial is simply a C++ vector of
monomial objects. These monomial objects are ordered by the given ring ordering,
which is imposed at all times, using monomial comparisons.

{\it Polynomial addition} is computed by simply merging the vector lists of two 
polynomials together, since the two polynomial vectors are already sorted. If two 
monomials are found to be equal, their Galois field coefficients are added 
together and the resulting monomial added to the sum if the new coefficient is 
not $0$.

{\it Multiplication of a polynomial by a monomial}, $P_2 = M_1 \cdot P_1$ is 
detailed in Algorithm \ref{alg:monPolyMult}. Each monomial in $P_1$ is 
iteratively multiplied by $M_1$ to derive a temporary monomial $M_{temp}$, 
which is added to $P_2$. 
Let $M_{last}$ denote the last monomial in $P_2$ at any given time. 
Due to the ordering, typically $M_{last}\geq M_{temp}$ during the procedure.
In cases where it is not, which only happens when exponents have been minimized, 
$M_{temp}$ falls not far earlier than $M_{last}$.
Thus, to add $M_{temp}$ to $P_2$, $M_{temp}$ is compared to monomials in $P_2$ 
in reverse order.
\begin{algorithm}[hbt]
\SetAlgoNoLine

 \KwIn{Monomial $M_1$, Polynomial $P_1$.}
 \KwOut{$M_1 \cdot P_1$}
        $P_2$ := $\varnothing$\;
        \ForEach{Monomial $M_p \in P_1$}
        {
            $M_{temp}$ := $M_p \cdot M_1$\;
            \ForEach{Monomial $M_{p2} \in P_2$ in reverse order}
            {
                \If{$M_{p2} > M_{temp}$}
                {
                    $P_2$.insertAfter($M_{p2}$,$M_{temp}$)\;
                    break\;
                }
                \If{$M_{p2} == M_{temp}$}
                {
                    $M_{p2}$.coef += $M_{temp}$.coef\;
                    \If{$M_{p2}$.coef == $0$}{$P_2$.pop()\;}
                    break\;
                }
            }
            \If{$M_{temp}$ hasn't been inserted}
            {
                $P_2$.insertToFront($M_{temp}$)\;
            }
        }
        \Return $P_2$\;
\caption{Multiplication of a Polynomial by a Monomial}\label{alg:monPolyMult}
\end{algorithm}

{\it Multiplication of polynomials}, $P_3 = P_1 \cdot P_2$, is computed as
numerous monomial-by-polynomial multiplications. Each monomial in $P_1$
is multiplied by the entire polynomial $P_2$ to derive a temporary polynomial
$P_{temp}$. Then, $P_{temp}$ is added to a growing $P_3$. Order is maintained
by these sub-procedures, so no further ordering logic is needed.

\begin{Example}
\label{ex:polyMultAdd}
Assume the environment has been set up over the ring $\F_{2^4}[a,b,c,Z]$ as in
Example \ref{ex:gf4MonMult}. Let $P_1$ and $P_2$ be the following polynomials
\begin{eqnarray}
P_1=(\alpha) ab +bZ^3 && P_2=abc+ab+b
\end{eqnarray}
$P_1+P_2$ is computed by merging the polynomials together. Two monomials exist
with the same order, $(\alpha) a b$ in $P_1$ and $ab$ in $P_2$, so here only the
coefficients are merged.
\begin{equation}
P_1+P_2 = abc + (\alpha+1)ab + bZ^3+b
\end{equation}
$P_1 \cdot P_2$ is computed by taking each monomial of $P_1$ and multiplying it
by $P_2$. The first temporary polynomial generated is:
\begin{equation}
(\alpha)ab \cdot P_2 = (\alpha)ab \cdot (abc+ab+b) = (\alpha)abc+(\alpha)ab+(\alpha)ab=(\alpha)abc
\end{equation}
Notice that, since two equivalent terms were generated, their coefficients were
added together creating $(0)\cdot ab$, so this term was removed. The second
multiplication is
\begin{equation}
bZ^3 \cdot P_2 = bZ^3\cdot(abc+ab+b) = abcZ^3+abZ^3+bZ^3
\end{equation}
Finally, these two polynomials are added together to obtain the final result.
\begin{equation}
P_1\cdot P_2 = abcZ^3+(\alpha)abc+abZ^3+bZ^3
\end{equation}
\end{Example}

{\it Polynomial reduction} is the main procedure computed by the abstraction tool.
To reduce a polynomial $P_1$ by polynomial $P_2$, one reduction step
is computed as
\begin{equation}
P_1\stackrel{P_2}{\textstyle\longrightarrow} P_1 + \frac{LT(P_1)}{LT(P_2)} \cdot P_2
\end{equation}
where $LT$ is the first monomial object of the given polynomial. $\frac{LT(P_1)}{LT(P_2)} P_2$ 
modifies $P_2$ so that it has the same leading term as $P_1$. Thus, when the two 
polynomials are added together, the leading terms are cancelled out. Note that reduction
is only possible when $LT(P_1)$ is divisible by $LT(P_2)$.
However, one reduction step
may not be sufficient to compute a full reduction, i.e. it's possible that the 
resulting polynomial could be reduced further by $P_2$. 
One can reapply the reduction steps until no more reductions are possible.
A more efficient method is to collect all monomials in $P_1$ that are divisible
by $LT(P_2)$, add the results of each division to $P_{temp}$, and then compute
$P_1+P_{temp}\cdot P_2$. 
Due to the ordering, if $LT(P_2) > M_i$ where $M_i$ is the $i^{th}$ monomial in $P_1$, 
then all $M_j \in P_1$ for $j \geq i$ are not divisible by $LT(P_2)$. 
Thus, the divisions are performed in monomial order 
and stopped as soon as this condition holds. The overall procedure is described in 
Algorithm \ref{alg:polyReduction}. Polynomial reduction makes use of 
monomial division, monomial-by-polynomial multiplication,
and polynomial addition, which were all detailed previously.

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Polynomial $P_1$, Polynomial $P_2$.}
 \KwOut{$r$ where $P_1\stackrel{P_2}{\textstyle\longrightarrow}_+ r$}
        $P_{temp}$ := $\varnothing$\;
        \ForEach{Monomial $M_p \in P_1$}
        {
            \If{ $LT(P_2) > M_p$ }
            {
                break\;
            }
            $M_{div}$ := $M_p / LT(P_2)$\;
            \If{ $M_{div}$ $!=$ $\varnothing$ }
            {
                $P_{temp}$.push\_back($M_{div}$)\;
            }
        }
        \Return $P_1 + (P_{temp} \cdot P_2)$\;
\caption{Polynomial Reduction}\label{alg:polyReduction}
\end{algorithm}

\begin{Example}
Consider again the $\F_{2^4}[a,b,c,Z]$ setup from Example \ref{ex:polyMultAdd} 
with
\begin{eqnarray}
P_1=(\alpha) ab +bZ^3 && P_2=abc+ab+b
\end{eqnarray}
Compute $P_2\stackrel{P_1}{\textstyle\longrightarrow}_+ r$. Attempt to divide 
each monomial of $P_2$ by $LT(P_1)=(\alpha)ab$.
The first monomial
division is:
\begin{equation}
\frac{abc}{(\alpha)ab} = (\alpha^3+\alpha^2)c
\end{equation}
Note that over this field, $(\alpha)^{-1}=(\alpha^3+\alpha^2)$.
The next monomial division is:
\begin{equation}
\frac{ab}{(\alpha)ab} = (\alpha^3+\alpha^2)
\end{equation}
The last monomial division is not possible,
\begin{equation}
\frac{b}{(\alpha)ab} = \varnothing
\end{equation}
so $P_{temp}=(\alpha^3+\alpha^2)c+(\alpha^3+\alpha^2)$. The final reduction
is then computed as
\begin{eqnarray}
&&P_2 + (P_{temp} \cdot P_1) \nonumber \\
&&= abc+ab+c+((\alpha^3+\alpha^2)c+(\alpha^3+\alpha^2))\cdot((\alpha) ab +bZ^3) \nonumber \\
&&=abc+ab+c+(abc+ab+(\alpha^3+\alpha^2)bcZ^3+(\alpha^3+\alpha^2)bZ^3) \nonumber \\
&&=(\alpha^3+\alpha^2)bcZ^3+(\alpha^3+\alpha^2)bZ^3+c 
\end{eqnarray}
Notice that the two monomials of $P_2$ that were divisible by $LT(P_1)$ are 
cancelled out.
\end{Example}

\section{Abstraction Tool Flow and Results}

The tool takes the circuit as input
and applies the approach presented in Chapter \ref{ch:improv}
to derive the polynomial representation of the circuit.
The most computationally intensive procedures in the approach are
\begin{enumerate}
\item Initial reduction of the word-level polynomial, 
$f_z \xrightarrow{F-\{f_z\},F_0}_+ r$ %(Step \ref{alg:initialred})
\item Derivation of the polynomials for the second reduction procedure, 
$F_a=\{a_0=\Func_0(A), a_1=\Func_1(A), \dots\}$ %(Step \ref{alg:secondpoly})
\item Computation of the second reduction (substitution), $r\xrightarrow{F_A,F_0}_+ Z+\Func(A)$ %(Step \ref{alg:secondred}).
\end{enumerate}
Of these, the first two are computed in parallel as they are independent of each 
other. Step $1$ orders the polynomials in $J+J_0$ by their
monomial order and reduces $f_Z$ in that order. In our experiments, this step typically
takes longer than step $2$.

%\subsection{Experimental Results}

All experiments are run on a 64-bit Linux desktop with a
3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU and 16 GB of RAM.
Table \ref{tbl:mastroToolresults} depicts the time and memory required to derive
 the polynomial abstraction
from bug-free and buggy Mastrovito multiplier circuits using our custom tool. 
This circuit is provided as a bit-blasted/flattened gate-level netlist. These 
circuits compute $Z = A \cdot B$ over some field $\Fkk$, so the analysis is performed over
this same field, abstracting this word-level representation.
The bug introduced is a swapping of two output nodes of the given circuit,
ensuring that the effect propagates down during the reduction process.
Similarly, Table \ref{tbl:montFlatToolResults} depicts the results for abstracting flattened Montgomery
multipliers.

%The tool takes the circuit as input,
%performs a reverse topological traversal to determine RATO, 
%applies the approach presented in Section \ref{sec:improve}
%and derives the polynomial representation $Z + A\cdot B$ (in the case of a bug-free circuit). 
%For comparison, Table \ref{table:masSing} depicts the time and memory required to compute the 
%abstraction of bug-free circuits using Singular scripts. Our tool is noticeably 
%faster and can compute word-level abstractions of bug-free Mastrovito circuits 
%up to 571-bits while Singular quickly becomes infeasible due to memory constraints.

\setlength{\abovecaptionskip}{3pt}
\begin{table}[hbt]
\begin{center}
{\small
\caption{Abstraction of Mastrovito multipliers. Time given in seconds, memory given in MB. $TO$ = $3$ days ($259,200$ seconds.)}
\label{tbl:mastroToolresults}
%\noindent\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|r|c|c|c|c|c|c|c|} 
\hline
\multicolumn{2}{|c|}{Size (k)}      &  163  &     233 &     283 &     409 & 571 \\
\multicolumn{2}{|c|}{\# of Gates}   & 153K  &    167K &    399K &    508K & 1.6M \\
%% \hline
%% \multirow{2}{*}{Singular} & Time    &    -     &    -    &    -    &    -    &    -      \\
%%                           & Max Mem &    -     &    -    &    -    &    -    &    -      \\
\hline
\multirow{2}{*}{Time (s)} & Bug Free &   1,443  &   1,913 &  11,116 & 17,848  & 192,032 \\
                      & Buggy    &   1,487  &   2,106 &  11,606 & 20,263  & 204,194 \\
\hline
\multicolumn{2}{|c|}{Max Memory (MB)} &     213  &     269 &     561 &    845  &   2,855 \\
\hline
\end{tabular}
%}
}
\end{center}
\vspace{-0.15in}
\end{table}

%For bug-free circuits up to
%409-bit multipliers, with 508K gates, our approach is
%successful.
%These also compute a multiplication function over a finite field, 
%but they are designed hierarchically. Montgomery multipliers are composed of 
%four Montgomery Reduce (MR) blocks, which takes $k$-bit inputs $A$ and $B$ and
%computes $ABR^{-1}$ (where $R = \alpha^k$ is a constant). The overall design is shown in
%Figure \ref{fig:mm4}.
%\begin{figure}[h]
%	\begin{center}
%	\includegraphics[scale=0.40]{figures/mmcircuit}
%	\end{center}
%	\caption{Montgomery multiplier over $\mathbb{F}_{2^k}$}
%	\label{fig:mm4}
%\end{figure}
Montgomery multipliers are typically designed hierarchically, as shown in Fig. \ref{fig:mm4}. 
If the hierarchy is known, it can be
exploited by computing the abstraction of each MR block in parallel, as shown in 
Table \ref{tbl:montBlkToolResults}.
In this table, 'BLK A' and 'B' denote the input MR blocks, 'BLK
Mid' denotes the middle block and 'BLK Out' is the output block. While
each block is an MR block, some have been simplified by
constant-propagation, hence they have
different sizes. First, a polynomial is extracted for each MR block
(gate-level to word-level abstraction), and then the approach is
re-applied at word-level to derive the input-output relation (solved
trivially in $< 1$ second). Our approach can extract the word-level
polynomial for up to 571-bit circuits!

%\setlength{\abovecaptionskip}{3pt}
\begin{table}[hbt]
\begin{center}
{\small
\caption{Abstraction of flat Montgomery multipliers. Time given in seconds, memory given in MB. $TO$ = $3$ days ($259,200$ seconds.}
\label{tbl:montFlatToolResults}
%\noindent\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|r|c|c|c|c|c|c|c|} 
\hline
\multicolumn{2}{|c|}{Size (k)}      &  163  &     233 &     283 &     409 & 571 \\
\multicolumn{2}{|c|}{\# of Gates}   & 184K  &    329K &    488K &    1.0M & 1.97M \\
%% \hline
%% \multirow{2}{*}{Singular} & Time    &    -     &    -    &    -    &    -    &    -      \\
%%                           & Max Mem &    -     &    -    &    -    &    -    &    -      \\
\hline
\multirow{2}{*}{Time} & Bug Free &   6,897  &  63,805 &  TO & TO  &   TO \\
                      & Buggy    &   6,961  &  64,009 &  TO & TO  &   TO      \\
\hline
\multicolumn{2}{|c|}{Max Memory} &     153  &     325 &     505 &    971  &   2,240   \\
\hline
\end{tabular}
%}
}
\end{center}
\vspace{-0.15in}
\end{table}

\begin{table}[hbt]
\begin{center}
{\small
\caption{Abstraction of Montgomery blocks. Time given in seconds, memory is given in MB. TO = 3 days (259,200 seconds)}
\label{tbl:montBlkToolResults}
%\noindent\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|r|c|c|c|c|c|c|} 
\hline
\multicolumn{3}{|c|}{Circuit Size (k)}                          &    163 &      233 &     283 &     409 &     571 \\
\hline
\multicolumn{2}{|c|}{\multirow{4}{*}{\# of Gates}}    & Blk A   & 33K &   55K &  82K & 168K & 330K \\
\multicolumn{2}{|c|}{}                                & Blk B   & 33K &   55K &  82K & 168K & 330K \\
\multicolumn{2}{|c|}{}                                & Blk Mid & 85K &  163K & 241K & 502K & 980K \\
\multicolumn{2}{|c|}{}                                & Blk Out & 32K &   54K &  81K & 168K & 328K \\
\hline
%% \multirow{6}{*}{Singular}    & \multirow{4}{*}{Time}  & Blk A   & 22,860  &    -    &    -    &    -    &    -    \\
%%                              &                        & Blk B   & 19,476  &    -    &    -    &    -    &    -    \\
%%                              &                        & Blk Mid & 18,010  &    -    &    -    &    -    &    -    \\
%%                              &                        & Blk Out & 17,604  &    -    &    -    &    -    &    -    \\
%% \hhline{~-----------}
%%                              & \multicolumn{2}{|r|}{Total Time} & 77,950  &    -    &    -    &    -    &    -    \\
%%                              & \multicolumn{2}{|r|}{Max Mem}    & 58,266  &    -    &    -    &    -    &    -    \\
%% \hline
%                                                             163       233       283      409       571
\multirow{8}{*}{Time} &\multirow{4}{*}{Bug Free}& Blk A   &    25 &      142 &     330 &   1,322 &   5,371 \\
                      &                         & Blk B   &    25 &      141 &     329 &   1,335 &   5,241 \\
                      &                         & Blk Mid &    73 &      408 &     883 &   4,471 &  19,942 \\
                      &                         & Blk Out &    24 &      140 &     321 &   1,338 &   5,532 \\
\hhline{~-------}
                      &\multirow{4}{*}{Buggy}   & Blk A   &    26 &      142 &     331 &   1,323 &   5,372 \\
                      &                         & Blk B   &    26 &      141 &     330 &   1,336 &   5,421 \\
                      &                         & Blk Mid &   111 &      580 &   1,411 &   6,829 &  37,804 \\
                      &                         & Blk Out &    25 &      141 &     322 &   1,339 &   5,539 \\
\hline%{~-----------}
%\multicolumn{3}{|r|}{Total Time}                          &    636 &    1,909 &   8,186 &  34,002 &  87,458 \\
\multicolumn{3}{|r|}{Max Mem Per Blk}                      &    80 &      168 &    254 &     538 &   1,129 \\
\hline

\end{tabular}
%}
}
\end{center}
\vspace{-0.2in}
\end{table}

%A multiplier over $\Fkk$ can also be composed over the composite field
%$\Fm_{(2^m)^n}$, where $k = m\cdot n$. The input $A=a_0+a_1\alpha+\dots+a_{k-1}\alpha^{k-1}$ is 
%transformed into multiple word-level input blocks over $\Fm_{2^m}$, $A_0,A_1,\dots,A_{n-1}$,
%where each $A_i=a_{i0}+a_{i1}\beta+\dots+a_{i(m-1)}\beta^{m-1}$ 
%and $\beta$ is the primitive element of $\Fm_{2^m}$. The multiplier is then 
%composed of m-bit blocks, each of which is either an m-bit adder or an m-bit 
%multiplier over $\Fm_{2^m}$. 
%An example $4$-bit multiplier with designed over $\Fm_{(2^2)^2}$ is 
%seen in Figure \ref{fig:comp4ex}. 
To abstract a word-level representation of the 
composite field multiplier $\F_{(2^m)^n}$ [Fig. \ref{fig:comp4ex}], we
first apply our approach to abstract a word-level representation of each m-bit block. 
In the case of an adder, this abstraction is trivially computed in $1$ second. 
In the case of a multiplier, 
refer to our experimental results for Mastrovito multipliers for comparison (Table \ref{tbl:mastroToolresults})
as these are designed as $m$-bit Mastrovito blocks. Each abstraction can be computed independently. 
Once these word-level abstractions are known, the final abstraction over $\F_{(2^m)^n}$ is performed
using only word-level variables. 
%For this, a functional mapping from each 
%$A_i$ to $A$ is needed in the form $A_i=\F_i(A)$. This is computed in a similarly 
%to the bit-level functional mapping $a_i=\F_i(A)$ described in 
%Section \ref{sec:improve}, except here the property $A_i^{2^m}=A_i$ is exploited 
%instead of the property $a_i^2=a_i$.
The results of this final word-level abstraction of 
buggy and bug-free multipliers over composite fields are shown in 
Table \ref{tbl:compositeToolResults}.

%\centering
%\begin{subfigure}{.4\columnwidth}
%\includegraphics[width=\columnwidth]{example-image-a}%
%\caption{cap a}%
%\label{subfiga}%
%\end{subfigure}\hfill%
%\begin{subfigure}{.4\columnwidth}
%\includegraphics[width=\columnwidth]{example-image-b}%
%\caption{cap b}%
%\label{subfigb}%
%\end{subfigure}\hfill%
%\begin{subfigure}{.4\columnwidth}
%\includegraphics[width=\columnwidth]{example-image-c}%
%\caption{cap c}%
%\label{subfigc}%
%\end{subfigure}%
%\caption{The proper caption}
%\label{figabc}
%\end{figure*}

%\begin{figure*}[t]
%        \centering
%        \includegraphics[width=.75\linewidth]{./figures/compMineSmall}
%        \caption{$4$-bit composite multiplier designed over $\Fm_{(2^2)^2}$}\label{fig:comp4ex}
%\end{figure*}

\begin{table}[!h]
\centering
{
\small
%\begin{tabular}{||c|c|c|c|c|c||c|c|c|c|c|c||c|c|c|c|c|c||} 
%\hline
%\multicolumn{6}{||c||}{64}&\multicolumn{6}{c||}{128}&\multicolumn{6}{c||}{256}\\
%\hline
%\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Bug Free} & \multicolumn{2}{c||}{Buggy} &
%\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Bug Free} & \multicolumn{2}{c||}{Buggy} &
%\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Bug Free} & \multicolumn{2}{c||}{Buggy}   \\
%\hhline{~~----~~----~~----}
%     &       & Time    &   Mem   & Time     &   Mem &
%     &       & Time    &   Mem   & Time     &   Mem &
%     &       & Time    &   Mem   & Time     &   Mem\\
%\hline
%$2$  & $32$  & $4$     &  $4$    & $4$      & $10$  &
%$2$  & $64$  & $188$   & $11$    & $118$    & $32$  &
%$2$  & $128$ & $10,602$& $45$    & $10,617$ & $126$ \\
%\hline
%$4$  & $16$  & $7$     &   $5$   & $7$      & $11$  &
%$4$  & $32$  & $271$   &   $15$  &  $272$   & $36$  &
%$4$  & $64$  & $23,876$&   $61$  & $23,910$ & $142$ \\
%\hline
%$8$  & $8$  &  $9$     &  $5$    & $9$      & $11$  &
%$8$  & $16$ & $412$    &  $17$   &  $414$   & $38$ &
%$8$  & $32$ & $32,247$ &  $69$   & $32,293$ & $150$ \\
%\hline
%$16$ & $4$  & $11$     &  $5$    & $11$     & $11$  &
%$16$ & $8$  & $513$    & $19$    & $514$    & $40$ &
%$16$ & $16$ & $41,594$ & $74$    & $41,651$ & $155$ \\
%\hline
%$32$ & $2$  & $12$     &  $5$    & $12$     & $12$  &
%$32$ & $4$  & $545$    & $19$    & $546$    & $40$ &
%$32$ &  $8$ & $46,223$ & $75$    & $46,285$ & $156$ \\
%\hline
%   - &   -  &   -      &      -  &        - &      - &
%$64$ & $2$  & $563$    & $19$    &  $565$   & $40$&
%$64$ &  $4$ & $49,466$ & $75$    & $49,532$ &  $156$ \\
%\hline
%  -   &   - &   -      &    -    &    -     &     -  &
%  -   &   - &   -      &    -    &    -     &     -  &
%$128$ & $2$ & $55,440$ & $76$    & $55,513$ &  $157$ \\
%\hline
%\end{tabular}
%\begin{tabular}{|c|c|c|c|c|}
%\hline
%\multicolumn{5}{|c|}{64}\\
%\hline
%\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Time} & Max \\
%\hhline{~~--~}
%     &       & Bug Free    &   Buggy   &   Mem \\
%\hline
%$2$  & $32$  & $4$     & $4$      & $10$  \\
%\hline
%$4$  & $16$  & $7$     & $7$      & $11$  \\
%\hline
%$8$  & $8$  &  $9$     & $9$      & $11$  \\
%\hline
%$16$ & $4$  & $11$     & $11$     & $11$  \\
%\hline
%$32$ & $2$  & $12$     & $12$     & $12$  \\
%\hline
%   - &   -  &   -      &        - &      -\\
%\hline
%  -   &   - &   -      &    -     &     - \\
%\hline
%\end{tabular}

\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{128}\\
\hline
\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Time} & Max \\
\hhline{~~--~}
     &      & Bug Free & Buggy & Mem \\
\hline
 $2$ & $64$ &      $1$ &   $1$ & $4$ \\
\hline
 $4$ & $32$ &      $1$ &   $1$ & $2$ \\
\hline
 $8$ & $16$ &      $1$ &   $1$ & $2$ \\
\hline
$16$ &  $8$ &      $1$ &   $1$ & $2$ \\
\hline
$32$ &  $4$ &      $1$ &   $1$ & $2$ \\
\hline
$64$ &  $2$ &      $1$ &   $1$ & $3$ \\
\hline
  -  &   -  &   -      &   -   &  -  \\
\hline
  -  &   -  &   -      &   -   &  -  \\
\hline
  -  &   -  &   -      &   -   &  -  \\
\hline
\end{tabular}
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{256}\\
\hline
\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Time} & Max \\
\hhline{~~--~}
     &        & Bug Free & Buggy & Mem  \\
\hline
  $2$ & $128$ &     $15$ &  $15$ & $23$ \\
\hline
  $4$ &  $64$ &      $2$ &   $2$ &  $4$ \\
\hline
  $8$ &  $32$ &      $1$ &   $1$ &  $3$ \\
\hline
 $16$ &  $16$ &      $1$ &   $1$ &  $2$ \\
\hline
 $32$ &   $8$ &      $1$ &   $1$ &  $2$ \\
\hline
 $64$ &   $4$ &      $1$ &   $1$ &  $2$ \\
\hline
$128$ &   $2$ &      $1$ &   $1$ &  $2$ \\
\hline
  -   &   -   &       -  &    -  &   -  \\
\hline
  -   &   -   &       -  &    -  &   -  \\
\hline
\end{tabular}
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{512}\\
\hline
\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Time} & Max \\
\hhline{~~--~}
     &        & Bug Free & Buggy & Mem  \\
\hline
  $2$ & $256$ &    $406$ & $408$ & $90$ \\
\hline
  $4$ & $128$ &     $53$ &  $53$ & $25$ \\
\hline
  $8$ &  $64$ &      $8$ &   $8$ &  $4$ \\
\hline
 $16$ &  $32$ &      $2$ &   $2$ &  $4$ \\
\hline
 $32$ &  $16$ &      $1$ &   $1$ &  $3$ \\
\hline
 $64$ &   $8$ &      $1$ &   $1$ &  $3$ \\
\hline
$128$ &   $4$ &      $1$ &   $1$ &  $2$ \\
\hline
$256$ &   $2$ &      $1$ &   $1$ &  $2$ \\
\hline
  -   &   -   &      -   &   -   &   -  \\
\hline
\end{tabular}
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{1024}\\
\hline
\multirow{2}{*}{$m$} &  \multirow{2}{*}{$n$} & \multicolumn{2}{c|}{Time} & Max \\
\hhline{~~--~}
      &       & Bug Free &  Buggy   &   Mem \\
\hline
  $2$ & $512$ & $11,883$ & $12,050$ & $414$ \\
\hline
  $4$ & $256$ &  $1,520$ &  $1,536$ & $106$ \\
\hline
  $8$ & $128$ &    $209$ &    $211$ &  $29$ \\
\hline
 $16$ &  $64$ &     $38$ &     $37$ &  $10$ \\
\hline
 $32$ &  $32$ &     $10$ &     $10$ &   $5$ \\
\hline
 $64$ &  $16$ &      $4$ &      $4$ &   $3$ \\
\hline
$128$ &   $8$ &      $2$ &      $2$ &   $3$ \\
\hline
$256$ &   $4$ &      $1$ &      $1$ &   $3$ \\
\hline
$512$ &   $2$ &      $1$ &      $1$ &   $3$ \\
\hline
\end{tabular}
}
\caption{Abstraction of bug-free Mastrovito multipliers over $\mathbb{F}_{(2^m)^n}$. Time is given in seconds. Memory is given in MB. $TO$ = more than $24$ hours = $86,400$ seconds. 
Note that abstractions of the $m$-bit blocks which compose the circuits are already known.}
\label{tbl:compositeToolResults}
\end{table}

\begin{table}[!hbt]
\begin{center}
{\small
\caption{Statistics of Designs over $\mathbb{F}_{(2^m)^n}$}
\label{tbl:stats}
\begin{tabular}{|c|c|c|c|c|c|c|c|} 
\hline
$n$ & 2 & 4 & 8 & 16 & 32 \\
\hline
\# of $\F_{2^m}$ Multipliers  & $6$ & $36$ & $168$ & $720$ & $2976$ \\
\hline
\# of $\F_{2^m}$ Adders       & $3$ & $27$ & $147$ & $675$ & $2883$  \\
\hline
\end{tabular}
}
\end{center}
\end{table}

%% For these
%% experiments, we used the Singular \cite{DGPS} computer algebra
%% tool. The results are shown in Table \ref{tab:ours}. The row
%% ``Compute-GB'' depicts the time required to compute the Gr\"obner
%% basis using the abstraction term order; whereas the row ``Improved-GB''
%% depicts the result of our improved, guided, approach using RATO. We
%% correctly identified $Z + A\cdot B$ as the only word-level polynomial
%% in the Gr\"obner basis $G'$, for up to 163-bit circuits over
%% ${\mathbb{F}}_{2^{163}}$. Similarly, the results for Montgomery
%% multipliers are shown in Table \ref{tab:mont}. Montgomery multipliers
%% are significantly larger than Mastrovito ones. The 163-bit circuit has
%% more than $91,000$ variables, which Singular cannot accommodate, as
%% its capacity is limited to $< 32767$ variables. Therefore, the largest
%% experiment was restricted to 128-bit Montgomery multipliers. 
%% These experiments were conducted  on a desktop with $2.40$GHz Intel
%% Core \texttrademark{$2$} Quad CPU with $8$GB memory running $64$-bit
%% Linux.

The above experiments also demonstrate that we can perform equivalence
checking between Mastrovito (golden model) and Montgomery multiplier
(implementation) circuits, by deriving a canonical polynomial ($Z_1,
Z_2$)from each circuit independently and then checking if $Z_1 =
Z_2$, for up to 571-bit circuits. Our experiments have shown that contemporary 
approaches (BDDs, SAT, SMT, and AIG/ABC) show some success in bug-catching for these
kinds of circuits (particularly ABC). These experiments are shown in Table \ref{tab:satsmtbugcatch}.
Note that ABC uses random simulation for FRAIGing (functionally reducing AIGs), which
can catch a bug early if it is lucky.
However, full equivalence checking using any of these techniques failed even for 
16-bit circuits, as shown in Table \ref{tab:equivCheckFail}.

\begin{table}[!hbt]
\begin{center}
{\small
\caption{Bug-catching between a golden-model Mastrovito and buggy Montgomery circuit. 
Time given in seconds. TO = 3 days (259,200 seconds)}
\label{tab:satsmtbugcatch}
%\noindent\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|c|c|c|c|c|c|} 
\hline
Circuit Size (k) &  64  &    163 &      233 &     283 &     409 &    571  \\
\hline
ABC              &  1     &    32  &     6    &   96    &     217 &    401  \\
\hline
Lingeling        &  1     &     8  &    362   & 12,728  &   3,323 & 23,298  \\
\hline
Picosat          & 15,235 &    TO  &    TO    &    TO   &     TO  &   TO    \\
\hline
Boolector        &  4     &    30  &    41    &   105   &    152  & 19,113  \\
\hline
CVC4             &  2     &    11  &    64    &  8,660  &    280  &    TO   \\
\hline
Z3               &  1     &    12  &    55    & 10,169  &    335  &    TO   \\
\hline
Yices            &  1     &     6  &     7    &   618   &    578  & 11,568  \\
\hline
\end{tabular}
%}
}
\end{center}
\vspace{-0.2in}
\end{table}

\begin{table}[!hbt]
\begin{center}
{\small
\caption{Equivalence checking between a golden-model Mastrovito and a bug-free Montgomery circuit. 
TO = 3 days (259,200 seconds)}
\label{tab:equivCheckFail}
%\noindent\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|c|} 
\hline
Circuit Size (k) &  16  \\
\hline
ABC              &  TO \\
\hline
Lingeling        &  TO \\
\hline
Picosat          &  TO \\
\hline
Boolector        &  TO \\
\hline
CVC4             &  TO \\
\hline
Z3               &  TO \\
\hline
Yices            &  TO \\
\hline
\end{tabular}
%}
}
\end{center}
\vspace{-0.2in}
\end{table}

%between f, 
%as shown in Table III, these tools have shown
% In \cite{lv:date2012}, the authors
%have shown that this same equivalence checking experiment using
%contemporary approaches (BDDs, SAT, SMT, and AIG/ABC) is infeasible
%beyond 16-bit circuits (cf. Table I in \cite{lv:date2012}).

%, the table is not reproduced here due to lack of space).

%% \begin{table}[hbt!]
%% {\myfontsize
%% \begin{center}
%% \caption{\small Run-time for word-level abstraction from Mastrovito
%%   multipliers using our approach.  {\sc Singular} computer algebra
%%   tool is used for this experiments. Time is given in seconds. MO = out
%%   of memory.}  
%% \label{tab:ours}
%% \begin{tabular}{|l||c|c|c|c|c|c|c|} \hline 
%% Size $k$-bits& 16 & 32  & 64 & 128  &163\\
%% \hline
%% \#variables &$323$ &$1155$ &$4355$ &$16899$ &$27224$ \\
%% \hline
%% \#polynomials &$291$ &$1091$  &$4227$ &$16643$ &$26989$ \\
%% \hline
%% \#terms &$1793$ &$7169$  &$28673$ &$114689$ &$185984$ \\
%% \hline
%% Compute-GB & $124$ &$720$  &$MO$ &$MO$ &$MO$ \\
%% \hline
%% Improved-GB &$0.04$ &$1.6$ &$78$ &$4500$ &$18424$ \\
%% \hline
%% %Our approach: Bugs (s) & $0.04$ &$1.43$  &$114.86$ &$788.65$ &$3061$  &$9384$
%% % &$16368$\\
%% \hline
%% \end{tabular}
%% \end{center}
%% \vspace{-0.2in}
%% }
%% \end{table}

%% \begin{table}[htb!]
%% {\myfontsize
%% \begin{center}
%% \caption{ Runtime for word-level abstraction from Montgomery
%%   multipliers. TO = timeout of 10hrs. Time is given
%%   in seconds. $\ast$ denotes {\sc singular}'s capacity exceeded.} 
%% \label{tab:mont}
%% \begin{tabular}{|c||c|c|c|c|c|c|} \hline 
%% Operand size $k$ &  32 & 48 & 64 & 128 &163\\
%% \hline
%% \#variables     &$1194$     &$2280$     &$4395$     &$14122$  &$91246$\\
%% \hline
%% \#polynomials    &$1130$     &$2184$     &$4267$     &$13866$  &$89917$\\
%% \hline
%% \#terms          &$10741$    &$18199$    &$40021$    &$134887$ &$484738$\\
%% \hline \hline
%% %Compute-GB &$1.50$  &$11.03$    &$27.70$  &$10919$  &$\ast$\\
%% %\hline
%% Improved-GB    &$25$     &$1023$     &$3060$     &$11047$    &$\ast$\\
%% \hline
%% \end{tabular}
%% \end{center}
%% }
%% \end{table}

%\afterpage{\clearpage}
\FloatBarrier

\section{Limitations of the Abstraction Approach}

Our tools and approach performs very favorable for $\Fkk$ multiplier circuits and other 
functions designed over fields. These types of circuits are based on AND-XOR gate logic. 
Thus, the polynomials 
derived during the reduction procedures are very sparse. Since the  
complexity of the algorithm heavily depends on the density of the polynomials,
the worst-case is avoided in such designs.
In the case of bugs within these circuits, these polynomials increase in size, but are
still easily manageable. This is why the approach is applicable to
very large Galois field multipliers, both buggy and bug-free.

However, for random logic, especially logic containing chains of
OR gates, the polynomial becomes very dense. As a result, the algorithm begins 
to encounter the computational worst-case.  
This is best shown with a small example.

\begin{Example}
\begin{figure}[h]
        \centering
        \begin{subfloat}[][XOR logic]{%{0.25\textwidth}
                \includegraphics[width=0.25\textwidth]{./figures/xor-example}
                %\caption{XOR logic}
                \label{fig:xorEx}}
        \end{subfloat}%
        ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfloat}[][OR logic]{%{0.25\textwidth}
                \includegraphics[width=0.25\textwidth]{./figures/or-example}
                %\caption{OR logc}
                \label{fig:orEx}}
        \end{subfloat}
        \caption{Logic comparisons}\label{fig:logicExmp}
\end{figure}

%\begin{figure}[!hbt]
%\centerline{
%\includegraphics[scale=0.5]{}
%}
%\caption{ XOR-based logic.}
%\label{fig:xorEx}
%\end{figure}
{\it 
Consider the circuit in Figure \ref{fig:xorEx}, which performs a 4-input
XOR function. Due to RATO, the monomial ordering of the variables is 
$z > f > e > d > c> b >a$. Thus, $z$ will be reduced in terms of the rest of the 
variables. The polynomials derived from the design are:
\begin{eqnarray}
f_1:z+f+d & f_2:f+e+c & f_3:e+b+a \nonumber
\end{eqnarray}
The reduction procedure $z \xrightarrow{f_1,f_2,f_3}_+ r$ will be computed as follows:
\begin{enumerate}
\item $z \xrightarrow{z+f+d} f+d$
\item $(f+d)\xrightarrow{f+e+c}e+d+c$
\item $(e+d+c)\xrightarrow{e+b+a}d+c+b+a$
\end{enumerate}
In each reduction, the output gate variable is removed and one copy of each 
input variable is added, leaving a sparse polynomial. Now consider the same 
circuit with the XOR gates replaced by OR gates, as shown in Figure \ref{fig:orEx}.

%\end{figure}
%\begin{figure}[!hbt]
%\centerline{
%\includegraphics[scale=0.5]{./figures/or-example}
%}
%\caption{ OR-based logic.}
%\label{fig:orEx}
%\end{figure}
The monomial ordering stays the same, but the polynomials derived from each gate
have changed:

\begin{eqnarray}
f_1:z+fd+f+d & f_2:f+ec+e+c & f_3:e+ba+b+a \nonumber
\end{eqnarray}
The reduction procedure, $z \xrightarrow{f_1,f_2,f_3}_+ r$ is now computed as:
\begin{enumerate}
\item $z \xrightarrow{z+fd+f+d}fd+f+d$
\item $(fd+f+d)\xrightarrow{f+ec+e+c}f+edc+ed+dc+d$;\\
$(f+edc+ed+dc+d)\xrightarrow{f+ec+e+c}edc+ed+ec+e+dc+d+c$
\item $(edc+ed+ec+e+dc+d+c)\xrightarrow{e+ba+b+a}_+$\\ $dcba+dcb+dca+dba+dc+db+da+d+cba+cb+ca+c+ba+b+a$
\end{enumerate}
Each pass removes an output variable of the gate, but replaces it with two instances of each input variable.
This increases the density of the resulting polynomial exponentially.
}
\end{Example}

\section{Conclusions}

This chapter examined the data structures and algorithms implemented in a
custom software abstraction tool.
Abstraction of Galois field circuits using the custom tool has 
greatly better performance compared to using {\sc Singular} scripts. With it, we can
abstract circuits up to $1024$-bits, buggy or bug-free. Although a bug will 
generally substantially inflate the size of the resulting polynomial abstraction, 
the tool is not greatly hindered by the presence of bugs in a circuit.
However, for random logic, especially OR-based logic, the size of the polynomials tends 
to grow exponentially during the abstraction. Thus, the approach is infeasible for
circuits with OR-gate chains.

### 2016-09-21 dissertation/improve.tex
init dissertation writing

\chapter{Overcoming Grobner Basis Complexity for Abstraction} \label{ch:improv}

Computing a \Grobner basis is prohibitively expensive for large circuits. 
%The worst-case complexity of computing $GB(J + J_0)$
%in $\Fq[x_1, \dots, x_d]$ is known to be bounded by $q^{O(d)}$
%\cite{gao:gf-gb-ms}, which is prohibitive over large
%fields. 
The approach from the last 
chapter is limited only to small circuits, with data-paths no larger than 
$40$-bits. 
A full \Grobner basis computation results in numerous polynomials, but
the abstraction approach ``searches'' for only one
polynomial ($Z + \G(A)$) in the basis. This motivates an investigation into
whether it is possible to {\it  guide a sequence of $Spoly(f,
  g)\xrightarrow{J+J_0}_+ r$  computations} to arrive at the desired
word-level polynomial. This chapter describes this 
{\it smaller subset} of computations, which are derived from a \Grobner basis 
analysis, to find the word-level polynomial of the function performed by a 
given circuit. The improved approach can abstract canonical word-level 
representations of circuits up to $571$ bits, 
corresponding to the largest NIST-specified ECC standard.

\section{Improving the Abstraction Approach}
Consider the word-level abstraction problem formulation from 
Chapter \ref{ch:abstract}. $J$ is the ideal 
generated by all polynomials derived from the circuit implementation and 
$J_0$ is the ideal of all the vanishing polynomials of 
every variable in the ring. 
The computation of the reduced \Grobner basis of $J+J_0$ over $\Fq$ 
has the following known complexity\cite{gao:gf-gb-ms}:

\begin{Theorem}
Let $J+J_0 = \langle f_1, \dots, f_s, ~x_1^q - x_1, \dots, x_d^q -
x_d\rangle \subset \Fq [x_1, \dots, x_d]$ be an ideal. The time and
space complexity of Buchberger's algorithm to compute a \Grobner
basis of $J+J_0$ is bounded by $q^{O(d)}$.
\end{Theorem}

In our case $q = 2^k$, and when $k$ and $d$ are large, this complexity 
makes abstraction infeasible.

Recall that Buchberger's algorithm \cite{buchberger_thesis} for 
computing \Grobner bases depends on the computation of an $S$-polynomial, 
which is then reduced by all the polynomials in the basis.

\begin{equation}
Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r \nonumber
\end{equation}
where
\begin{eqnarray}
Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g \nonumber \\ 
\nonumber \\
L=lcm(lt(f),lt(g)) \nonumber
\end{eqnarray}

A new polynomial is added to the basis when the remainder of the $Spoly$
reduction, $r$, is non-zero.

Notice that our approach searches
for only one polynomial $f_\Func:Z+\Func(A_1,\dots,A_n)$, and it does
by computing the entire reduced \Grobner basis, $G=\{g_1,\dots,g_m\}$ and
finding $f_\Func\in \{g_1,\dots,g_m\}$.
This motivates us to
investigate whether it's possible to {\it  guide a sequence of 
$Spoly(f,g)\xrightarrow{J+J_0}_+ r$  computations} to arrive at the desired
word-level polynomial, without considering other polynomials in the
generating set.

Numerous improvements have been introduced to improve the efficiency of 
Buchberger's algorithm. One of these is the product criterion, the results
of which we exploit for our approach.

\begin{Lemma}
\label{lemma:prodcriteria}
[Product Criterion \cite{productc:1979}] Let $\mathbb{F}$ be any
field, and $f, g \in \mathbb{F}[x_1,\cdots,x_d]$ be polynomials. If
the equality $lm(f) \cdot lm(g) = LCM(lm(f), lm(g))$ holds, then
$Spoly(f,g)\stackrel{G}{\textstyle\longrightarrow}_+ 0.$ 
\end{Lemma}

The above result states that when the leading monomials of $f, g$ are
relatively prime then 
$Spoly(f, g)$ always reduces to 0 modulo $G$. In this case,
$Spoly(f, g)$ need not be considered in Buchberger's algorithm, and thus
the computation is avoided. 
Recall that in the Abstraction Term Order (Definition \ref{def:ato}),
we have ``circuit variables $x_1, \dots, x_d$'' $>$ ``word-level output''
$>$ ``word-level inputs'', where the
relative ordering among  $x_1, \dots, x_d$ is not important. This ordering
is now further refined to exploit the product criteria.

Given an acyclic combinational circuit, an ordering can be applied to the 
bit-level variables, $\{x_1,\dots,x_d\} \in \F_2$, 
based on their topological position in the circuit. In a {\it reverse topological
ordering}, the output variable of the gate will always come earlier in the
ordering than any of its input variables.
%This is denoted {\bf topological ordering}.

%\begin{Definition}
%Given a Boolean combinational circuit, let the topological location of the 
%bit-level variable $x$, denoted as $TL(x)$, be the 
%maximum number of Boolean logic gates that any path from any input must 
%pass to reach the variable. Then the variable ordering 
%\begin{equation}
%x>y : TL(x)<TL(y) 
%\end{equation}
%is denoted {\bf topological ordering}. Likewise, the ordering 
%\begin{equation}
%x>y : TL(x)>TL(y) 
%\end{equation}
%is a {\bf reverse topological ordering}.
%\label{def:topord}
%\end{Definition}
%\begin{Example}
%Consider the combinational circuit shown in Figure \ref{fig:topo}.
%This circuit has two bit-level inputs
%$a$ and $b$, intermediate variable $c$; and output $z$.

%\begin{figure}[H]
%\centerline{
%\includegraphics[scale=0.6]{./figures/topoExample.png}
%}
%\caption{Boolean combinational circuit.}
%\label{fig:topo}
%\end{figure}

%Then:
%\begin{eqnarray}
%TL(a)=0 \nonumber \\
%TL(b)=0 \nonumber \\
%TL(c)=1 \nonumber \\
%TL(z)=2 \nonumber
%\end{eqnarray}

%Thus, $a > b > c > z$ is a topological ordering. 
%Since $TL(a)=TL(b)$, $b > a > c > z$ is also a topological 
%ordering.

%Likewise, $z > c > a > b$ and $z > c > b > a$ are considered reverse
%topological orderings.
%\end{Example}

\begin{Definition}
\label{def:rato}
{\bf Refined Abstraction Term Order (RATO) $>_r$:} 
%Starting from the primary
%outputs of the circuit $C$, perform a reverse topological traversal
%toward the primary inputs. Order each variable of the circuit
%according to its reverse topological level: i.e. $x_i > x_j$ if $x_i$
%appears earlier in the reverse topological order.
Given a circuit $C$, apply a reverse topological ordering to the bit-level
variables $\{x_1, \dots, x_d\}$.
Then, impose a lex term
order $>_r$ on $\Fq[x_1, \dots, x_d, Z, A_1, \dots, A_n]$ with ``circuit 
variables ordered reverse topologically'' $>$ 
``output word-level variable'' $>$ ``input word-level variables''. 
%This term order $>_r$ is called the {\bf refined abstraction term order (RATO)}.
\end{Definition}

Let $F$ be the set of polynomials which generate $J$, i.e. $J=<F>$. 
When RATO is applied, 
we find that all 
{\it bit-level circuit constraint
polynomials} in $F$ have leading terms that are relatively prime to each 
other. Since we are using a
reverse topological variable ordering with lex term ordering, these 
polynomials are in the 
form of $f_i = x_i + \text{tail}(f_i)$, where $x_i$ is the output of a gate $f_i$, 
and thus the following
proposition from {\it Lv's} work \cite{lv:phd} can be applied.

\begin{Proposition} \label{prop:top-order}
Let $C$ be any arbitrary combinational circuit. Let $\{x_1, \dots,
x_d\}$ denote the set of all variables (signals) in the circuit,
i.e. the primary input, intermediate and primary output
variables. Perform a {\bf reverse topological traversal} of the
circuit and order the variables such that $x_i > x_j$ if $x_i$ appears
earlier in the reverse topological order. Impose a lex term order to
represent the Boolean expression for each gate as a polynomial $f_i$;
then $f_i = x_i + \text{tail}(f_i)$. Then the set of all polynomials
$\{f_1, \dots, f_s\}$ forms a Gr\"obner basis, as $lt(f_i)$ and $
lt(f_j)$ for $i\neq j$ are relatively prime. 
\end{Proposition}

\begin{Example}
Consider the $2$-bit multiplier from Example 
\ref{exp:finalMul2Bit}. With RATO applied, the bit-level circuit constraint
polynomials in $F$, where $J=<F>$, are:
\begin{eqnarray}
		f_1:c_0+a_0 \cdot b_0 \nonumber \\
		f_2:c_1+a_0 \cdot b_1 \nonumber \\
		f_3:c_2+a_1 \cdot b_0 \nonumber \\
		f_4:c_3+a_1 \cdot b_1 \nonumber \\
		f_5:r_0+c_1 + c_2		\nonumber \\
		f_6:z_0+c_0 + c_3	\nonumber	\\
		f_7:z_1+r_0 + c_3	\nonumber	
\end{eqnarray}
The leading terms of $f_1,\dots,f_7$ are relatively prime to each other.
\label{exp:2bitmulrato}
\end{Example}

Let $F_0$ be the set of polynomials which generate $J_0$, i.e. $J_0=<F_0>$.
In $F\cup F_0$, for every polynomial $f_i = x_i + \text{tail}(f_i)$ 
in $F$ there is a vanishing polynomial $v_i = x_i^2+x_i$ in $F_0$; 
$f_i$ and $v_i$ have leading terms which are not relatively prime.
In this case, 
\cite{lv:phd} shows that $Spoly(x_i+\text{tail}(f_i), x_i^{2^k}-x_i)
\stackrel{J,J_0}{\longrightarrow}_+ r$ always produces $r=0$, and thus can
be excluded from the \Grobner basis computation.

\begin{Theorem}
\label{thm:lvcontrib}
Let $q = 2^k$, and let $\Fq[x_1, \ldots, x_d]$ be a ring on
which we have a reverse topological lex order. Let $I$ be a subset of $\{1,
\ldots, d\}$. For all $i \in I$, let $f_i = x_i +P_i$ (where $P_i =
\text{tail}(f_i)$) such that  all indeterminates $x_j$  that appear in
$P_i$ satisfy $x_i > x_j$.  Then the set $G = \{f_i :  i  \in I\} \cup
\{x_1^q-x_1, \ldots, x_d^q-x_d\}$ is a Gr\"obner basis. 
\end{Theorem}

The proof is given in \cite{lv:phd} and is reproduced here:

\begin{Proof}
Given a system of polynomials derived from a circuit over 
$\Fq[x_1, \ldots, x_d]$, where $\{x_1,\dots,x_d\}$ are bit-level variables.
Apply a reverse topological lex ordering to $\{x_1,\dots,x_d\}$. 
Let $x_i$ be the output of a Boolean logic gate for some $1\leq i \leq d$.
Let $f = x_i + P_i$ be the polynomial derived from this logic gate and
$g = x_i^q-x_i$ be vanishing polynomial of $x_i$.
Then $Spoly(f,g)=
x_i^{q-1} f - g = x_i^{q-1}P_i + x_i$. In what follows, it is important
to note that the indeterminates appearing in $P_i$ are all less than
$x_i$ over the given ordering.

First,  $x_i^{q-1}P_i +x_i - x_i^{q-2}P_i(x_i+P_i)=x_i^{q-2}P_i^2 +x_i,$ 
which shows that 
$x_i^{q-1}P_i +x_i \stackrel{x_i+P_i}{\longrightarrow}  x_i^{q-2}P_i^2
+x_i.$

Next, $x_i^{q-2}P_i^2 + x_i - x_i^{q-3}P_i^2(x_i+P_i)=
x_i^{q-3}P_i^3+ x_i.$ Continuing in this fashion, we get $x_iP_i^{q-1}
+x_i-P_i^{q-1}(x_i+P_i) = x_i + P_i^q,$ and finally 
$x_i+P_i^q -(x_i+P_i) = P_i^q-P_i.$ Hence, 
$$x_i^{q-1}P_i +x_i \stackrel{x_i+P_i}{\longrightarrow} x_i^{q-2}P_i^2
+x_i \stackrel{x_i+P_i}{\longrightarrow} x_i^{q-3}
+x_i \stackrel{x_i+P_i}{\longrightarrow} \cdots$$
$$\cdots \stackrel{x_i+P_i}{\longrightarrow}
P_i^q+x_i\stackrel{x_i+P_i}{\longrightarrow} P_i^q-P_i.$$

Over the finite field $\mathbb{F}_{q}$, $P_i^q-P_i$ is a vanishing
polynomial. Therefore, $P_i^q-P_i \in I(V(J_0))= \langle
x_1^q-x_1, \ldots, x_d^q-x_d\rangle$. Due to the product criterion
(Lemma \ref{lemma:prodcriteria}), $G_0=\{x_1^q-x_1, \ldots, x_d^q-x_d\}$ is
Gr\"obner basis. Therefore 
$P_i^q-P_i \stackrel{G_0}\rightarrow_+ 0$. 
\end{Proof}

Due to RATO, there exists a polynomial $f_{z_i}\in F$ which is the 
polynomial derived from a Boolean logic gate, where $z_i$ is the first 
variable in the ordering for some $0\leq i <k$. That is, $f_{z_i}: z_i+\text{tail}(f_{z_i})$.
\begin{Proposition}
Over RATO, the polynomial pair $(f_Z, f_{z_i})$ is the only critical pair
at the start of the \Grobner basis computation of $J+J_0$, where 
$f_{z_i}$ is the polynomial derived from the gate
\end{Proposition}
\begin{Proof}
Due to Theorem \ref{thm:lvcontrib} and the product criterion, 
a critical pair must come from a word level designation polynomial, 
$\{f_Z,f_{A_1},\dots,f_{A_n}\}$, and a polynomial derived from a Boolean logic
gate. The leading terms of the polynomials $\{f_{A_1},\dots,f_{A_n}\}$ are 
bit-level inputs to the circuit and thus are not the outputs of any gate.
Thus, the only critical pair if $f_Z$ and $f_{z_i}$, where $z_i$ is the first
variable in the ordering and is thus the leading monomial of $f_Z$.
\end{Proof}
%Thus, we are left with only a few options for an $Spoly$ at the  
%beginning of a \Grobner basis computation of $J+J_0$ with abstraction order
%$>$. Consider the {\it word-level designation polynomials} 
%$\{f_{Z},f_{A_1},\dots,f_{A_n}\}$ in $J$. Since all word-level variables 
%are found last in abstraction order, the leading monomial of each of these
%polynomials is some bit-level $x_d$. Thus there is a vanishing polynomial
%$x_d^2+x_d$ with the same leading monomial. Take $f_Z$ for example:

%\begin{eqnarray}
%f_{Z}: z_{0}+z_{1}\cdot \alpha,\cdots,{z_{k-1}}\cdot \alpha^{k-1}+Z \nonumber \\
%f_{v}: z_0^2+z
%\end{eqnarray}

%Thus, there are no valid $S$-polynomials which can be generated using
%the set $\{$bit-level constraints $\in J$,  $J_0\}$.
%So a valid $S$-polynomials must be generated using at least 
%one {\it word-level designation polynomial}, $f_Z,f_{A_1},\dots,f_{A_n}$.
%Consider the output word-level designation polynomial $f_Z$:
%\begin{equation}
%f_Z: Z+z_0+z_1\alpha+\dots+z_{k-1}\alpha^{k-1}
%\end{equation}
%Since all word-level variables 
%are found last in RATO, the leading monomial of $f_Z$ 
%has to be a bit-level output variable, $z_i$, for some $0\leq i < k$. 
%Since $z_i$ is the output of a Boolean logic gate, there 
%exists a polynomial $f_{z_i} \in J$ with $z_i$ as the leading monomial, 
%$f_{z_i}:z_i + \text{tail}(f_{z_i})$. $Spoly(f_Z,f_{z_i})$ is 
%a valid $S$-polynomial, and it has
%some interesting properties.

Thus, the first computation of the \Grobner basis is guaranteed to be 
$Spoly(f_Z,f_{z_i})\xrightarrow{F,F_0}_+ r$.
This computation has the following interesting property.

\begin{Proposition} \label{prop:reduce}
Normalize $f_Z$, i.e. $f_Z$=$f_Z/LC(f_Z)$. Then
\begin{eqnarray}
Spoly(f_Z,f_{z_i})\xrightarrow{F,F_0}_+ r & \text{is equivalent to} &
f_Z\xrightarrow{F-\{f_Z\},F_0} + r
\end{eqnarray}
\end{Proposition}

%% \begin{Proposition} \label{prop:reduce}
%% Let $C$ be any arbitrary combinational circuit with $k$-bit inputs 
%% $A_1,\dots,A_n$ and $k$-bit output $Z$. Model the circuit as a system of 
%% polynomials as described earlier.
%% Let $f_Z$ be the word-level 
%% designation polynomial $Z+z_0+z_1\alpha+\dots+z_{k-1}\alpha^{k-1}$. By
%% applying RATO, the leading monomial of $f_Z$ will be $z_i$ for some
%% $0\leq i < k$.
%% There exists a polynomial $f_{z_i}$ in the form of $z_i+\text{tail}$, 
%% which is the polynomial representation of the 
%% Boolean gate with output $z_i$.
%% %Then if $f_Z$ and $f_i$ is first minimized, i.e. $f_Z$=$f_Z/LC(f_Z)$, 
%% Then 
%% \begin{equation}
%% Spoly(f_Z,f_{z_i})\stackrel{J,J_0}{\longrightarrow}_+ r
%% \end{equation}
%% is equivalent to
%% \begin{equation}
%% %f_Z\stackrel{J-\{f_Z\},J_0}{\longrightarrow}_+ r
%% f_Z\xrightarrow{J-\{f_Z\},J_0} + r
%% \end{equation}
%% \end{Proposition}

\begin{Proof}
Assuming both $f_{z_i}$ and $f_Z$ are minimized, i.e. 
$LC(f_{z_i})=LC(f_Z)=1$, 
they are both in the form $z_i+P$ for some polynomial $P$.
Let $f=z_i+P_f$ and $g=z_i+P_g$ represent $f_{z_i}$ and $f_Z$ respectively.

(i) For $Spoly(f,g)$
\begin{equation}
L=LCM(LT(f),LT(g))=LM(f)=LM(g)=z_i
\end{equation}
so
\begin{eqnarray}
Spoly(f,g)&=&\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g \nonumber \\
&=&\frac{z_i}{z_i}\cdot f - \frac{z_i}{z_i}\cdot g \nonumber \\
&=&f + g \nonumber
\end{eqnarray}
Thus 
$Spoly(f_Z,f_{z_i})\stackrel{F,F_0}{\longrightarrow}_+ r$
is equivalent to 
$f+g\stackrel{F,F_0}{\longrightarrow}_+ r$.

(ii) For $f_Z\xrightarrow{F-\{f_Z\},F_0} + r$, since the leading term 
of $f_Z$ is $z_i$, the only polynomial in the set $\{F-\{f_Z\},F_0\}$
which can perform the first division is $f_{z_i}$. 
Again, denote $f_Z$ as $f$ and $f_{z_i}$ 
as $g$. According to the reduction algorithm, the remainder $r$ of 
$f\stackrel{g}{\longrightarrow}_+ r$ is:
\begin{eqnarray}
r &=& f-lt(f)/lt(g) \cdot g \nonumber \\
  &=& f-z_i/z_i \cdot g \nonumber \\
  &=& f-g \nonumber \\
  &=& f+g
\end{eqnarray}
Thus 
$f_Z\xrightarrow{F-\{f_Z\},F_0} + r$
is equivalent to 
$(f+g)\xrightarrow{F-\{f_Z\},F_0} + r$

(iii) Consider $(f+g)\xrightarrow{F,F_0} + r$. $f=z_i+P_f$ and 
$g=z_i+P_g$. So $f+g=2z_i+P_g+P_f=P_g+P_f$ no longer contains $z_i$ as the 
leading monomial. As a consequence of RATO, as 
reduction proceeds the remainder will never again contain 
$z_i$ since $z_i$ is the first variable in the ordering. 
Since the leading monomial of $f_Z$ is $z_i$, $f_Z$ will never be 
used for reduction. Therefore 
$(f+g)\xrightarrow{F,F_0}+ r$
is equivalent to 
$(f+g)\xrightarrow{F-\{f_Z\},F_0}+ r$

Thus, 
$Spoly(f_Z,f_{z_i})\stackrel{F,F_0}{\longrightarrow}_+ r$
is equivalent to
$f_Z\xrightarrow{F-\{f_Z\},F_0}_+ r$
\end{Proof}

%Due to the reverse-topological abstraction order, 
%since there is a $x_i + \text{tail}$ for every $x_i \in \{x_1,\dots,x_d\}$ 
%except the bit level inputs $a_0,\dots,a_k$,
%$f_Z\xrightarrow{J-\{f_Z\},J_0}+ r$ can only contain the bit-level
%inputs $a_0,\dots,a_k$, the word level inputs $A_1,\dots,A_n$, and the 
%word-level output $Z$. 
Thus, $f_Z\xrightarrow{F-\{f_Z\},F_0}_+ r$ is the first computational step 
of the abstraction.
The polynomial remainder $r$
will {\it not} contain any bit-level variable corresponding to the
output of any gate in the design;  i.e. primary output bits and
intermediate variables of the circuit do not appear in $r$. To prove
this, assume that a non-primary-input variable $x_j$ appears in a
monomial term $m_j$ in $r$. Since there always exists a polynomial
$f_j$ such that $f_j = x_j + \text{tail}(f_j)$, $lt(f_j)$
divides monomial $m_j$ and $m_j$ can be canceled. Therefore, all
such terms $m_j$ with non-primary-input bit-level variables can be
eliminated.

Two cases need to be considered: 
\begin{enumerate}
\item Remainder $r$ only contains word-level variables:
  word-level output $Z$ and the word-level 
  inputs $A_1,\dots,A_n$. 
  Since RATO is lex with $Z > \{A_1,\dots,A_n\}$, the remainder $r$
  is the desired canonical polynomial representation,
  $Z + \Func(A_1,\dots,A_n)$.
\item Remainder $r$ contains both the bit-level primary
  input variables, as well as the word-level
  variables.  
\end{enumerate}

%We've found that for bug-free Galois field
%arithmetic circuits,
%when we compute $Spoly(f_{Z},f_{z_i})\stackrel{J,J_0}{\longrightarrow}_+ r$, 
%$r$ only contains variables in $A_1,\dots,A_n,Z$ and is thus {\bf canonical, 
%word-level polynomial } in the form of 
%$Z+\Func(A_1,\dots,A_n)$.

\begin{Example}
\label{exp:absGood}
Again consider the $2$-bit multiplier from Example 
\ref{exp:2bitmulrato}.
RATO for this example is
\begin{equation}
z_1 > z_0 > r_0 > c_0 > c_1 > c_2 > c_3 > a_0 > a_1 > b_0 > b_1 > Z > A > B 
\end{equation}
$F+F_0$ with RATO applied is:
\begin{eqnarray}
 \left .  
	\begin{aligned}
		f_1:c_0+a_0 \cdot b_0  \\
		f_2:c_1+a_0 \cdot b_1  \\
		f_3:c_2+a_1 \cdot b_0  \\
		f_4:c_3+a_1 \cdot b_1  \\
		f_5:r_0+c_1 + c_2		\\
		f_6:z_0+c_0 + c_3		\\
		f_7:z_1+r_0 + c_3		
	\end{aligned} 
 \ \right\}
 &\qquad&  {\it  \text{Bit-level circuit constraints} ~(\subset J)} \nonumber \\
 \left . 
	\begin{aligned}
		f_{A}:a_0+a_1\cdot \alpha + A  \\ 
		f_{B}:b_0+b_1\cdot \alpha + B \\ 
		f_{Z}:z_1\cdot \alpha + z_0 + Z  
	\end{aligned} 
 \right\}
 &\qquad&  {\it  \text{Word-level designation} ~(\subset J)} \nonumber \\
  \left . 
	\begin{aligned}
		a_0^2-a_0, ~a_1^2-a_1,~b_0^2-b_0, ~b_1^2-b_1   \\ 
		c_0^2-c_0, ~c_1^2-c_1,~c_2^2-c_2, ~c_3^2-c_3  \\ 
		r_0^2-r_0, ~z_0^2-z_0,~z_1^2-z_1    \\ 
		A^4-A, ~B^4-B ,~Z^4-Z		  
	\end{aligned} 
 \right\}
 &\qquad&  {\it \text{vanishing polynomials} (J_0)} \nonumber
\end{eqnarray}
Notice that the leading monomial of $f_Z$ is $z_1$, which is also the leading monomial 
of $f_7$.
Minimize $f_Z$, 
$f_{Zmin}=f_Z/\alpha = z_1 +z_0\cdot(\alpha+1)+Z\cdot(\alpha+1)$.
By computing the $S$-polynomial of $f_{Zmin}$ and $f_7$:
\begin{equation}
Spoly(f_{Zmin},f_{7})\stackrel{F,F_0}{\longrightarrow}_+ r \nonumber
\end{equation}
the remainder $r$ is $Z+A\cdot B$.

Likewise, if we reduce $f_{Z}$ by $F-\{f_Z\},F_0$, that is reduce it by all
generators in $J+J_0$ except $f_Z$:
\begin{equation}
f_{Z}\xrightarrow{F-\{f_Z\},F_0} + r \nonumber
\end{equation}
the remainder $r$ is $Z+A\cdot B$. 
\end{Example}

%As this abstraction process can be computed as a reduction procedure, we
%created a highly efficient reduction tool over $Fkk$. Using this tool, we
%are able to abstract word-level polynomial representations for bug-free 
%circuits up
%to $407$-bits in size. When applied to buggy circuits, however, 
%approach yields a polynomial which includes bit-level input variables.

\begin{Example}\label{ex:improve1}
Now consider a $2$-bit multiplier which has a {\bf bug}.
The output lines, $z_0$ and $z_1$, have been swapped, as shown in 
Figure \ref{fig:2bitmulbug} below:

\begin{figure}[H]
\centerline{
\includegraphics[scale=0.4]{./figures/2bitmasmultBUGswap}
}
\caption{ A buggy 2-bit multiplier over ${\mathbb{F}}(2^2)$.}
\label{fig:2bitmulbug}
\end{figure}

The polynomials in $F+F_0$ are the same as in Example \ref{exp:absGood} 
except for the following changes:
\begin{eqnarray}
f_6:z_1+c_0 + c_3	\nonumber	\\
f_7:z_0+r_0 + c_3	\nonumber	\\
f_Z:z_1 + z_0\cdot \alpha + Z  \nonumber	
\end{eqnarray}

$f_Z$ has common leading terms with $f_6$. Computing 
\begin{equation}
Spoly(f_Z,f_{6})\stackrel{F,F_0}{\longrightarrow}_+ r \nonumber
\end{equation}
gives remainder $r: a_1\cdot b_1 \alpha + a_1\cdot B \alpha + b_1 \cdot A \alpha + Z + A\cdot B$, which is the same result as computing
$f_Z\xrightarrow{F-\{f_Z\},F_0}_+ r$
\label{exp:2bitbugmulex}
\end{Example}

%%%%%%%%%%%%%%%%%%%%%%F4-STYLE REDUCTION%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Improving Polynomial Division using $F4$-style Reduction}

The most intensive computational step in our proposed improvement 
is that of polynomial division $f_Z \xrightarrow{F-\{f_Z\},F_0}_+ r$. 
When the circuit $C$ is very large, the polynomial set
$\{F-\{f_Z\},F_0\}$ also becomes extremely large. This division
procedure then becomes the bottleneck in our abstraction approach.  
In principle, this reduction can be performed using contemporary
computer-algebra systems --- {\it e.g.},  the {\sc Singular}
\cite{DGPS} tool, which is widely used within the verification
community \cite{wienand:cav08} \cite{wedler:date11}
\cite{lv:date2012}. In our work, we have also performed experiments 
with {\sc Singular}. However, as in any ``general-purpose''
computer algebra tool, the data-structures are not specifically
optimized for circuit verification problems. Moreover, {\sc
  Singular} also limits the number of variables ($d$) that it can
accommodate in the system to $d < 32767$; this limits its
application to large circuits. Recent symbolic computation techniques \cite{lv:phd}
have shown improvements from employing the concept of $F4$-style 
polynomial reduction \cite{f4}.
Therefore, to further improve our
approach, we exploit this relatively recent concept, 
which implements polynomial division
using row-reductions on a matrix, to develop a custom
verification tool to perform this reduction
efficiently.

$Faug\grave{e}re$'s $F4$ approach \cite{f4} presents a new algorithm to
compute a Gr\"obner basis. It uses the same mathematical principles as
Buchberger's algorithm. However, instead of computing and reducing one 
$S$-polynomial at a time, it computes many $S$-polynomials in one
step and reduces them simultaneously using sparse linear algebra on a
matrix (triangulation). We can use this efficient reduction technique
to perform our reduction, $f_Z \xrightarrow{F-\{f_Z\},F_0}_+ r$, by 
representing and solving it on a matrix.
First, let us consider the following example that demonstrates the
main concepts behind the reduction approach of $F4$.

\begin{Example}
{\it
Consider the {\it lex} term order with $x>y>z$ on the ring
${\mathbb{Q}}[x, y, z]$.  Given $F = \{f_1 = 2x^2 + y, f_2 = 3xy^2
-xy, f_3 = 4y^3 -1\}$, consider one step of Buchberger's algorithm:
$S(f_1, f_2) \xrightarrow{f_1, f_2, f_3}_+r$. We have, $Spoly(f_1,
f_2) = \frac{1}{3}x^2y + \frac{1}{2}y^3 = f_4$. The reduction
$Spoly(f_1, f_2)\stackrel{f_1, f_2, f_3}{\longrightarrow}_+
(-\frac{1}{6}y^2 + \frac{1}{8})$ is done as follows: 
Since $lt(f_1) ~|~ lt(f_4), ~f_4 \xrightarrow{f_1} h$ is computed as:
\[
h = f_4 - {{lt(f_4)} \over {lt(f_1)}} f_1 = f_4 - \frac{1}{6}y f_1 =
\frac{1}{2}y^3 - \frac{1}{6}y^2; 
\]
Now $lt(f_2)$ does not divide any term in $h$, but $lt(f_3) ~|~ lt(h)$,
so $f \xrightarrow{f_3} r$:
\[
r = h - {{lt(h)} \over {lt(f_3)}} f_3 =  \frac{1}{2}y^3 - \frac{1}{6}y^2 -
\frac{1}{8}f_3 = -\frac{1}{6}y^2 + \frac{1}{8} 
\]

This reduction procedure can also be simulated on a matrix using
Gaussian elimination. 
%Matrix triangulation can also be used for this purpose. 
The reduction above requires the computation of $\frac{1}{6}y f_1$ and 
$\frac{1}{8}f_3$. Ignoring the coefficients $\frac{1}{6},
\frac{1}{8}$, 
we can generate all the monomials  required in the reduction process: 
i.e.  monomials of $f_4, yf_1, f_3$, and setup the problem of
cancellation of terms as Gaussian elimination on a matrix. Monomials
of $f_4, yf_1, f_3$ are, respectively, $\{x^2y, y^3\}, \{x^2y, y^2\}, \{y^3, 1\}$.
Let the rows of a matrix $M$ correspond to polynomials $\left[ f_4,
  yf_1, f_3   \right]$, and columns correspond to all the monomials
(in {\it lex} order) $\left[x^2y, y^3, y^2,  1 \right]$. Then the
matrix $M$ shows the representation of these polynomials where
the entry $M(i, j)$ is the coefficient of monomial of column $j$
present in the polynomial of row $i$.

\[
M = \bordermatrix{
~ & x^2y & y^3 & y^2 & 1\cr
f_4 & \frac{1}{3} & \frac{1}{2} & 0 & 0 \cr
yf_1 & 2 &0 & 1 & 0 \cr
f_3 & 0 & 4 & 0 & -1 \cr
}
\]

Now, reducing $M$ to a row echelon form using Gaussian elimination gives:
\[
M = \bordermatrix{
~ & x^2y & y^3 & y^2 & 1\cr
f_4 & \frac{1}{3} & \frac{1}{2} & 0 & 0 \cr 
h = f_4 - \frac{1}{6}yf_1 & 0 &\frac{1}{3}  & -\frac{1}{6} & 0 \cr 
r = h - \frac{1}{8}f_3 & 0 & 0 & -\frac{1}{6} & \frac{1}{8} \cr 
}
\]

The last row $(0, 0, -\frac{1}{6}, \frac{1}{8})$ accounts for
polynomial $-\frac{1}{6}y^2 + \frac{1}{8}y$ which is equal to the
reduction result $r$ obtained before. 
}
\end{Example}

This approach generates all the monomial terms that are required
in the division process, 
%i.e. $\frac{lm(f_i)}{lm(f_j)} f_j$, 
and the coefficients required for cancellation of terms are accounted
for by elementary row reductions in the subsequent Gaussian elimination. 
Based on the above concepts, a matrix can be constructed for our
problem: $f_Z \xrightarrow{F-\{f_Z\},F_0}+r$.

\begin{Definition}
Let $L = \left [f_1, \dots, f_m \right]$ be a list of $m$ polynomials. Let
$M_L$ be an ordered list of monomials of elements of $L$ and let $n$
be the number of elements in $M_L$. Define $M$ as the $m \times n$
matrix which associates the polynomials of $L$ to rows and monomials
of $M_L$ to columns. Entry in row $i$, column $j$ is the coefficient of the
$j^{th}$ element of $M_L$ in $f_i$. 
\end{Definition}

\begin{algorithm}[hbt]
\SetAlgoNoLine

 \KwIn{$f_Z, \{F-\{f_Z\},F_0\}$ as $\{f_1,\dots,f_s\}$, RATO $>_r$ } %with $f_{1}>f_{2}>\dots>f_{s}$. }
 \KwOut{Remainder $r$ of $f_Z \xrightarrow{f_1,\dots,f_s}_+r$}
  %%%%%%%%%%%%%%%%%%%%
        \CommentSty{/*$L=$ set of polynomials, rows of $M$*/\;}
        L:=\{$f_Z$\} \;{}
%        \CommentSty{/*The index of polynomials in $L$*/\;}
%       \CommentSty{/*Initial remainder $r$ is set as $f$*/\;}
%       r:=f\;
        \CommentSty{/*$M_{L} = $ the set of monomials, columns of $M$ */\;}
        $M_{L}$:=\{ monomials of f\} \;{}
%       \CommentSty{/*Let $Done$ be the set of monomials that have been handled*/\;}
%       $Done:=\{\}$\;
%       \CommentSty{/*The first $i-1$ monomials of $M_{L}$ have been handled*/\;}
        \For { ($i=0$; $i \leq $num monomials in $M_L$ ; i++) }
        {
                mon:= the $i^{th}$ monomial of $M_{L}$\;
                Identify $f_{k} \in F$ satisfying: $lm(f_{k})$ can divide $mon$ \;
                \CommentSty{/*add polynomial $f_k$ to L as a
                  new row in $M$ */\;}
                $L:=L \cup \frac{mon}{lm(f_{k})}\cdot f_{k}$ \;
                \CommentSty{/*Add monomials to $M_{L}$ as new
                  columns in $M$  */\;}
                $M_{L}$:=$M_{L} \cup \{ \text{monomials of }
                \frac{mon}{lm(f_{k})}\cdot f_{k}\}$ \;{} 
                %\CommentSty{/*Cancel adjacent monomials if they are the same*/\;}
                %CancelAdjMons($M_{L}$)\;
        }
         Gaussian Elimination on $M$;\\
         \Return $r =$ last row of $M$;
\caption{Generating the Matrix for Polynomial Reduction}\label{alg:matrix}
\end{algorithm}

Algorithm \ref{alg:matrix} describes our procedure to generate the 
matrix $M$ of polynomials corresponding to our reduction procedure. 
The main idea is to setup the rows and columns of the matrix in a 
way that polynomial division can be subsequently performed by 
applying Gaussian elimination on $M$. 
%subtracting row $i$ from row $i-1$. 
In the algorithm, the set of polynomials 
$\{F-\{f_Z\},F_0\} = \{f_1, \dots, f_s\}$ 
correspond to the circuit constraints and RATO 
is imposed on the polynomials. 
The output word-level polynomial $f_Z$ is to be reduced
w.r.t. $\{f_1, \dots, f_s\}$. Initially, $L = \{f_Z\}$ is inserted as
the first row of the matrix and $M_L$ constitutes the (ordered) list
of monomials of $f_Z$. Then, in every iteration $i$, a polynomial $f_k \in
\{f_1, \dots, f_s\}$ is identified such that $lm(f_k)$ divides the $i^{th}$ monomial
($mon$) of $M_L$; this is to enable cancellation of the corresponding
monomial term. The computation $L:=L \cup \frac{mon}{lm(f_{k})}\cdot
f_{k}$ in the while-loop, generates the polynomials required for
reduction.

%\footnote{Recall that the division $\frac{f_i}{f_k} = f_i -
%  \frac{lt(f_i)}{lt(f_{k})}\cdot f_{k} = f_i - \frac{lc(f_i)}{lc(f_k)} \cdot
%  \frac{lm(f_i)}{lm(f_k)}\cdot f_k$. In the algorithm, the computation
%  $\frac{mon}{lm(f_{k})}\cdot f_{k}$ corresponds to
%  $\frac{lm(f_i)}{lm(f_k)}\cdot f_k$ used in the division.}.

The list $M_L$ is updated to include monomials of
$\frac{mon}{lm(f_{k})}\cdot f_{k}$. Finally, the iteration in the
loop terminates when all monomials of $M_L$ have been analyzed.
The loop is guaranteed to terminate once $mon$ contains only
word-level variables, as no polynomials in $\{f_1,\dots,f_s\}$ have a leading
term that contains a word-level variable over RATO.

Using the set $L$ as rows and $M_L$ as columns, a matrix $M$ is
constructed and Gaussian elimination is applied to reduce it to
row-echelon form. The last row in the reduced matrix 
corresponds to the reduction result $r$. 
%Equation \ref{eqn:mat-red}. 
Let us describe the approach using an example.

\begin{Example}
\label{ex:f4}
Consider the reduction related to the abstraction of the 
${\mathbb{F}}_{2^2}$
multiplier circuit from Example \ref{exp:absGood}. The word-level output 
designation polynomial $f_Z$ is $z_1\alpha+z_0+Z$, and the circuit 
polynomials are
\begin{eqnarray}
f_1:a_0 + a_1 \alpha + A \nonumber \\
f_2:b_0 + b_1 \alpha + B \nonumber \\
f_3:r_0 + a_0b_1 + a_1b_0 \nonumber \\
f_4:z_0 + a_0b_0 + a_1b_1 \nonumber \\
f_5:z_1 + r_0 +a_1b_1 \nonumber
\end{eqnarray}
Here  $P(x) = x^2 + x + 1$, and  $P(\alpha) = 0$. We have to compute
$f_Z \xrightarrow{f_1, \dots, f_5}_+r$.  Note that, for simplicity, 
variables $c_0, c_1, c_2, c_3$ from Example \ref{exp:mul2bit} have been
substituted by functions on primary inputs. Impose RATO on the polynomials
as follows:
\begin{equation}
z_1 > z_0 > r_0 > a_0 > a_1 > b_0 > b_1 > Z > A > B
\end{equation}
The algorithm constructs the matrix as follows:

\begin{enumerate}

\item Initialization: $L = \{f_Z\} = \{z_1\alpha+z_0+Z\}. ~~M_L = \{z_1, z_0, Z\}, i =
  1, mon = z_1$ ($i^{th}$ monomial of $M_L$).
\item Iteration 1: Identify a polynomial $f_k \in \{f_1,\dots,f_s\}$ s.t. $lm(f_k) ~|~
  mon$. Clearly, $f_k = f_5 =z_1 + r_0 +a_1b_1$. Then, $L = L \cup
  \frac{mon}{lt(f_k)} \cdot f_k = L \cup f_5$. Therefore, $L = \{f,
  f_5\}$ and $M_L = \{z_1, z_0, r_0, a_1b_1, Z\}$, $ i = 2$ and $mon = z_0$.
\item Iteration 2: $f_k = f_4 = z_0 + a_0b_0 + a_1b_1$ because $lm(f_4) ~|~
  mon$. Therefore, $L = L \cup f_4$
  and $M_L = \{z_1, z_0, r_0, a_0b_0, a_1b_1, Z\}, i =
  3, mon = r_0$. 
\item Iteration 3: $~f_k = f_3 = r_0 + a_0b_1 + a_1b_0$ as $lt(f_3)
  ~|~ mon$. Therefore, $L = L \cup f_3$ and 
  $M_L = \{z_1, z_0, r_0, a_0b_0, a_0b_1, a_1b_0, a_1b_1, Z\}, ~i = 4, mon = a_0b_0$.
\item Iteration 4: $~f_k = f_1 = a_0 + a_1 \alpha + A$ because
$lm(f_1) ~|~ mon$. Then $L = L \cup \frac{a_0b_0}{a_0} \cdot f_1 = L \cup
b_0\cdot f_1=\{f_5,f_4,f_3,b_0f_1\}$ and $M_L = \{z_1, z_0, r_0, a_0b_0, a_0b_1, a_1b_0, a_1b_1, b_0A, Z\}$.
\item Continuing in this fashion $\dots$

\item Iteration 8: $L = \{f_Z, f_5, f_4, f_3, b_0f_1, b_1f_1, a_1f_2,
  Af_2\}$,
  $\\M_L = \{z_1, z_0, r_0, a_0b_0, a_0b_1, a_1b_0, a_1b_1, a_1B, b_0A, b_1A, Z, AB\}$, 
  $\\i = 9$, $mon = AB$.

\item Iteration 8: Since $mon = AB$ contains only the
  word-level inputs, no polynomial in $F$ has a leading term that can
  cancel $mon$, so the loop terminates. The matrix $M$ can be
  constructed using $L$ as rows and $M_L$ as columns.

\end{enumerate}

Figure \ref{subfig:matrix} shows the matrix $M$, and its subsequent
Gaussian elimination is shown in Fig. \ref{subfig:triangular}. The
last row of the reduced matrix corresponds to the reduction
$f_Z \xrightarrow{f_1, \dots, f_s}_+r$, 
where $r = Z+A\cdot B$.
\end{Example}

\begin{figure}[hbt]
\centering
\subfloat[Matrix $M$ generated by Algorithm \ref{alg:matrix}]
[Matrix $M$ generated by Algorithm \ref{alg:matrix}]
{ \label{subfig:matrix}
\begin{math}
M = \bordermatrix{
~      &    z_1 & z_0 & r_0 & a_0b_0 & a_0b_1 & a_1b_0 & a_1b_1 & a_1B & b_0A &   b_1A & Z & AB \cr
f_Z    & \alpha &   1 &   0 &      0 &      0 &      0 &      0 &    0 &    0 &      0 & 1 &  0 \cr
f_5    &      1 &   0 &   1 &      0 &      0 &      0 &      1 &    0 &    0 &      0 & 0 &  0 \cr
f_4    &      0 &   1 &   0 &      1 &      0 &      0 &      1 &    0 &    0 &      0 & 0 &  0 \cr
f_3    &      0 &   0 &   1 &      0 &      1 &      1 &      0 &    0 &    0 &      0 & 0 &  0 \cr
b_0f_1 &      0 &   0 &   0 &      1 &      0 & \alpha &      0 &    0 &    1 &      0 & 0 &  0 \cr
b_1f_1 &      0 &   0 &   0 &      0 &      1 &      0 & \alpha &    0 &    0 &      1 & 0 &  0 \cr
a_1f_2 &      0 &   0 &   0 &      0 &      0 &      1 & \alpha &    1 &    0 &      0 & 0 &  0 \cr
Af_2   &      0 &   0 &   0 &      0 &      0 &      0 &      0 &    0 &    1 & \alpha & 0 &  1 \cr
}
\end{math}
}

\ \\
\subfloat[$M$ reduced to row echelon form via Gaussian Elimination]
[$M$ reduced to row echelon form via Gaussian Elimination]
{
\begin{math}
M = \bordermatrix{
~                  &    z_1 & z_0 &    r_0 & a_0b_0 & a_0b_1 & a_1b_0 &   a_1b_1 & a_1B & b_0A & b_1A   & Z & AB \cr
f_Z                & \alpha &   1 &      0 &      0 &      0 &      0 &        0 &    0 &    0 &      0 & 1 &  0 \cr
\alpha f_5-row1    &      0 &   1 & \alpha &      0 &      0 &      0 &   \alpha &    0 &    0 &      0 & 1 &  0 \cr
f_4-row2           &      0 &   0 & \alpha &      1 &      0 &      0 & \alpha+1 &    0 &    0 &      0 & 1 &  0 \cr
\alpha f_3-row3    &      0 &   0 &      0 &      1 & \alpha & \alpha & \alpha+1 &    0 &    0 &      0 & 1 &  0 \cr
b_0f_1-row4        &      0 &   0 &      0 &      0 & \alpha &      0 & \alpha+1 &    0 &    1 &      0 & 1 &  0 \cr
\alpha b_1f_1-row5 &      0 &   0 &      0 &      0 &      0 &      0 &        0 &    0 &    1 & \alpha & 1 &  0 \cr
Af_2-row6          &      0 &   0 &      0 &      0 &      0 &      0 &        0 &    0 &    0 &      0 & 1 &  1 \cr
}
\end{math}
\label{subfig:triangular}
}
\label{fig:matrix}
\caption{$F_4$-style polynomial reduction on a matrix for Example \ref{ex:f4}.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%NOW: BIT LEVEL INPUTS%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reducing Bit-Level Inputs}

When the remainder $r$ only contains word-level variables, the problem
of word-level abstraction is solved. Thus, the focus now is to 
efficiently obtain a word-level abstraction when $r$ also contains 
bit-level input variables. 
In this case, a functional mapping is needed from each bit-level input 
variable $\{a_0, \dots, a_{k-1}\}\in\F_2$ to the word-level input variable $A\in\Fkk$.
\begin{eqnarray}
&a_0 = \Func_{a_0}(A)& \nonumber \\
&\vdots&  \label{eqn:bitLevelMappings}\\
&a_{k-1} = \Func_{a_{k-1}}(A)& \nonumber
\end{eqnarray}
where each $\Func_{a_i}$ is some function of $A$, which needs to be derived.
Here, we present the derivation when $\{a_0,\dots,a_{k-1}\}\in\F_2$ and $A\in\Fkk$.
However, this result is applicable from any field $\F_q$ to any extension
of the field $\F_{q^k}$, i.e. when $\{a_0,\dots,a_{k-1}\}\in\F_q$ and $A\in\F_{q^k}$. 
This generalized derivation is presented in Appendix \ref{append:Fpk}.

These mappings from $\{a_0, \dots, a_{k-1}\}$ to $A$ in Eqn.(\ref{eqn:bitLevelMappings})
are represented as polynomial functions $f_{a_0}, \dots, f_{a_{k-1}}$ in the
following form:
\begin{eqnarray}
&f_{a_0} : a_0 + \Func_{a_0}(A)& \nonumber \\
&\vdots& \label{eqn:bitLevelPolynomials} \\
&f_{a_{k-1}} : a_{k-1} + \Func_{a_{k-1}}(A)& \nonumber
\end{eqnarray}
Due to RATO, 
$\{a_{0}, \dots, a_{k-1}\} > A$, thus the leading terms of 
$f_{a_0},\dots,f_{a_{k-1}}$ are $a_{0}, \dots, a_{k-1}$ respectively.
Let $F_a = \{f_{a_0},\dots,f_{a_{k-1}}\}$. 
Then computing 
$r \xrightarrow{F_a,F_0}_+ r_w$ ensures that the new 
remainder $r_w$ must only contain word-level variables. In other words, 
$r_w$ must be in the form $Z + \Func(A)$ and is thus the word-level polynomial 
representation of the circuit.

Over $\Fkk$, $A=a_0+a_1\alpha+\cdots+a_{k-1}\alpha^{k-1}$. To compute $A^2$,
a special property of Galois fields, dealing with powers of elements, can be applied.
\begin{Lemma}\label{lemma:raiseToP}(from \cite{galois_field:mceliece})
{\it Let $\alpha_1, \dots, \alpha_t$ be any elements in $\F_{p^k}$. Then
\begin{equation}
(\alpha_1+\alpha_2+\cdots+\alpha_t)^{p^i}=\alpha_1^{p^i}+\alpha_2^{p^i}+\cdots+\alpha_t^{p^i}
\end{equation}
for all integers $i \geq 1$. }
\end{Lemma}
Lemma \ref{lemma:raiseToP} can be applied to compute $A^2$:
\begin{equation}
A^2=a_0^2+a_1^2\alpha^2+\cdots+a_{k-1}^2\alpha^{2(k-1)}
\end{equation}
Since each $a_i \in \F_2$, for $0\leq i<k$, then $a_i^2=a_i$. This is applied 
to find the final form for $A^2$:
\begin{equation}
A^2=a_0+a_1\alpha^2+\cdots+a_{k-1}\alpha^{2(k-1)}
\end{equation}
Similarly, $A^4$ can be derived as $(A^2)^2$:
\begin{equation}
A^4=a_0+a_1\alpha^4+\cdots+a_{k-1}\alpha^{4(k-1)}
\end{equation}
%\begin{Corollary} 
%\end{Corollary}
%As a consequence of Lemma \ref{lemma:raiseToP},

%The technique for deriving this functional mapping applies to any $\Fpk$ constructed 
%from a primitive polynomial $P(x)$ over $\F_p$, not just when $p=2$. 
%Let $P(\alpha)=0$; then any 
%element $A \in \Fpk$ can be represented as 
%$A=a_0+a_1\alpha+a_2\alpha^2+\cdots+a_{k-1}\alpha^{k-1}$ where 
%$\{a_0,\dots,a_{k-1}\} \in \F_p$.

%Since $a_i^p=a_i$ for all $i\in\{0,\dots,k-1\}$, 
%then 
%\begin{equation}
%A^{p^j}=a_0+a_1\alpha^{p^j}+a_2\alpha^{2{p^j}}+\cdots+a_{k-1}\alpha^{(k-1){p^j}} \nonumber
%\end{equation}

Deriving $A^{2^j}$ in this manner for all $0\leq j<k$ gives a system
of $k$ equations.
These equations can be represented in matrix form, $\mathbf{A = M a}$, where 
$\mathbf{A}=\{A,A^2,\dots,A^{2^{k-1}}\}^T$, 
$\mathbf{M}$ is
a $k$ by $k$ matrix of coefficients, and $\mathbf{a}=\{a_0,\dots,a_{k-1}\}^T$:
%\begin{equation}
%\begin{bmatrix}
%1      &   \alpha           & \alpha^2           & \dots & \alpha^{k-1}\\
%1      &   \alpha^p         & \alpha^{2p}        & \dots & \alpha^{(k-1)p}\\
%\vdots & \vdots             & \vdots             & ~     & \vdots \\
%1      &   \alpha^{p^{k-1}} & \alpha^{2p^{k-1}} & \dots & \alpha^{(k-1)p^{k-1}}
%\end{bmatrix}
%\begin{bmatrix}
%a_0 \\ a_1 \\ \vdots \\ a_{k-1}
%\end{bmatrix}
%=
%\begin{bmatrix}
%A \\ A^p \\ \vdots \\ A^{p^{k-1}}
%\end{bmatrix}
%\end{equation}
\begin{eqnarray}
\begin{bmatrix}
A \\ A^2 \\ \vdots \\ A^{2^{k-1}}
\end{bmatrix}  &=&
\begin{bmatrix}
1      &   \alpha           & \alpha^2           & \dots & \alpha^{k-1}\\
1      &   \alpha^2         & \alpha^{4}        & \dots & \alpha^{(k-1)\cdot 2}\\
1      &   \alpha^4         & \alpha^{8}        & \dots & \alpha^{(k-1)\cdot 4}\\
\vdots & \vdots             & \vdots             & \cdot     & \vdots \\
1      &   \alpha^{2^{k-1}} & \alpha^{2\cdot 2^{k-1}} & \dots & \alpha^{(k-1)\cdot 2^{k-1}}
\end{bmatrix}
\begin{bmatrix}
a_0 \\ a_1 \\ \vdots \\ a_{k-1}
\end{bmatrix} \label{eqn:matrixFormF2k}
\end{eqnarray}

Note that $\mathbf{M}$ is a matrix of constants and $\mathbf{A}$ and $\mathbf{a}$ are
vectors of variables.
However, by interpreting $\mathbf{a}$ as a vector of unknowns, $\mathbf{M}$ and $\mathbf{A}$ as constants. 
then $F_a$ can be derived by solving Eqn.(\ref{eqn:matrixFormF2k}) using Gaussian 
elimination. However, this system of equations also has a special structure which can
be exploited to further simplify the abstraction procedure.

%Alternatively, every $a_i\in\{a_0,\dots,a_{k-1}\}$ also has a closed form expression under Cramer's Rule.
\begin{Definition}
{\it {\bf Cramer's Rule}: Consider a system of $n$ linear equations and $n$ unknowns,
$x_1,\dots,x_n$, expressed in matrix form as $\mathbf{Mx=b}$:
%\begin{eqnarray}
%a_{11}x_1+a_{12}x_2+\cdots+a_{1n}x_n&=&b_1 \nonumber \\
%a_{21}x_1+a_{22}x_2+\cdots+a_{1n}x_n&=&b_2 \nonumber \\
% & \vdots &  \nonumber \\
%a_{n1}x_1+a_{n2}x_2+\cdots+a_{nn}x_n&=&b_n \nonumber
%\end{eqnarray}
%This is expressed in matrix form as $Ax=b$:
\begin{equation}
\begin{bmatrix}
m_{11} & m_{12} & \dots  & m_{1n} \\
m_{21} & m_{22} & \dots  & m_{2n} \\
\vdots & \vdots & \ddots & \vdots \\
m_{n1} & m_{2n} & \dots  & m_{nn}
\end{bmatrix}
\begin{bmatrix}
x_1 \\ x_2 \\ \vdots \\ x_n
\end{bmatrix}
 = 
\begin{bmatrix}
b_1 \\ b_2 \\ \vdots \\ b_n
\end{bmatrix}
\end{equation}
If the determinant $|\mathbf{M}|$ is non-zero, then for $1\leq i \leq n$,
\begin{equation}
x_i = \frac{|\mathbf{M_i}|}{|\mathbf{M}|}
\end{equation}
where $\mathbf{M_i}$ is $\mathbf{M}$ with the $i$-th column replaced with 
vector $\mathbf{b}$:
\begin{eqnarray}
\mathbf{M_i} = 
\begin{bmatrix}
m_{11} & m_{12} & \dots  & m_{1i-1} & b_{1}  & m_{1i+1} & \dots & m_{1n} \\
m_{21} & m_{22} & \dots  & m_{2i-1} & b_{2}  & m_{2i+1} & \dots & m_{2n} \\
\vdots & \vdots & \cdot  & \vdots   & \vdots & \vdots   & \cdot & \vdots \\
m_{n1} & m_{n2} & \dots  & m_{ni-1} & b_{n}  & m_{ni+1} & \dots & m_{nn}
\end{bmatrix} & &
\end{eqnarray}
}
\end{Definition}

\begin{Definition}\label{def:vandermonde}
{\it {\bf Vandermonde Matrix}: Let $V(x_1,\dots,x_n)$ denote a square $n$ x $n$ matrix of the form
\begin{equation}
\begin{bmatrix}
1 & x_1 & x_1^2  & \dots  & x_1^{n-1} \\
1 & x_2 & x_2^2  & \dots  & x_2^{n-1} \\
\vdots& \vdots  & \vdots & \cdot & \vdots    \\
1 & x_n & x_n^2  & \dots  & x_n^{n-1}
\end{bmatrix}
\end{equation}
where elements of each row are presented in a geometric progression.
Then $V(x_1,\dots,x_n)$ is a {\bf Vandermonde Matrix}, 
the determinant of which can be computed as:
\begin{equation} \label{eqn:vandet}
|V(x_1,\dots,x_n)| = \prod\limits_{1\leq i < j \leq n}(x_j - x_i)
\end{equation}
This determinant is non-zero if each $x_i \in \{x_1,\dots,x_n\}$ is a distinct
element.
}
\end{Definition}

Notice that $\mathbf{M}$ in Eqn.(\ref{eqn:matrixFormF2k}) is a square Vandermonde matrix 
of the form $V(\alpha,\alpha^2,\dots,\alpha^{2^{k-1}})$. 
\begin{Lemma}\label{lemma:nonzero}{\it The determinant of $\mathbf{M}$ as in Eqn.(\ref{eqn:matrixFormF2k}) is non-zero.}\end{Lemma}
\begin{Proof}{
\it Since $\mathbf{M}$ in Eqn.(\ref{eqn:matrixFormF2k}) is the Vandermonde matrix $V(\alpha,\alpha^2,\alpha^4,\dots,\alpha^{2^{k-1}})$, 
\begin{equation}
|\mathbf{M}|=\prod\limits_{0\leq i < j <k}{(\alpha^{2^j}-\alpha^{2^i})}
\end{equation}
Since $\Fkk$ is constructed from a primitive polynomial,
every $\alpha^i$ is a distinct element for $0\leq i < 2^k$.  
Thus, $|\mathbf{M}|$ is non-zero as it is a product of non-zero elements.}
\end{Proof}

Since $|\mathbf{M}|$ is non-zero, Cramer's rule can be applied to derive 
an equation for every $a_i$, $0\leq i < k$:
\begin{eqnarray}
a_i = \frac{|\mathbf{M_i}|}{|\mathbf{M}|} \label{eqn:cramerform}%\\
\end{eqnarray}
where $\mathbf{M_i}$ is $\mathbf{M}$ with the column $\{\alpha^i,\alpha^{i\cdot 2},\dots,\alpha^{i\cdot 2^{k-1}}\}^T$ 
replaced by $\mathbf{A}$.
\begin{equation}\label{eqn:M_i}
\mathbf{M_i} = \begin{bmatrix}
1      &   \alpha           & \alpha^2          & \dots  & \alpha^{i-1}          & A           & \alpha^{i+1}          & \dots & \alpha^{k-1} \\
1      &   \alpha^2         & \alpha^{4}       & \dots  & \alpha^{(i-1)\cdot 2}       & A^2         & \alpha^{(i+1)\cdot 2}       & \dots & \alpha^{(k-1)\cdot 2} \\
\vdots & \vdots             & \vdots            & \cdot  & \vdots                & \vdots      & \vdots                & \cdot & \vdots \\
1      &   \alpha^{2^{k-1}} & \alpha^{2\cdot 2^{k-1}} & \dots  & \alpha^{(i-1)\cdot 2^{k-1}} & A^{2^{k-1}} & \alpha^{(i+1)\cdot 2^{k-1}} & \dots & \alpha^{(k-1)\cdot 2^{k-1})}
\end{bmatrix}
\end{equation}

%The formulation in Eqn.(\ref{eqn:cramerform}) can be further simplified by reducing $|\mathbf{M}|$. 
%The following lemma is a new mathematical result, presented for the first time here.
\begin{Lemma}\label{lemma:MIs1}
{\it Over $\Fkk$, $|\mathbf{M}|=1$.}
\end{Lemma}
\begin{Proof}
{\it Since $\mathbf{M}$ is a Vandermonde matrix of the form $V(\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1})$, 
then from Eqn.(\ref{eqn:vandet})
\begin{equation}
|\mathbf{M}| = \prod\limits_{0\leq i < j < k}(\alpha^{2^j}-\alpha^{2^i}) \label{proofeqn:detM}
\end{equation}
Over $\Fkk$, $-1=1$, so Equation \ref{proofeqn:detM} is rewritten as
\begin{equation}
|\mathbf{M}| = \prod\limits_{0\leq i < j < k}(\alpha^{2^j}+\alpha^{2^i}) \label{proofeqn:detMrewrite}
\end{equation}
Computing $|\mathbf{M}|^2$ gives
\begin{equation} \label{proof:Msquare}
|\mathbf{M}|^2 = [\prod\limits_{0\leq i < j < k}(\alpha^{2^j}+\alpha^{2^i})]^2
\end{equation}
Applying Lemma \ref{lemma:raiseToP} to Equation \ref{proof:Msquare} gives
\begin{equation}
|\mathbf{M}|^2 = \prod\limits_{0\leq i < j < k}(\alpha^{2^{j+1}}+\alpha^{2^{i+1}})
\end{equation}
When $j=k-1$, the product term is in the form $(\alpha^{2^k}+\alpha^{2^{i+1}})$. 
Since $\alpha^{2^k}=\alpha$ over $\Fkk$, this term equivalent to  $(\alpha^{2^{i+1}}+\alpha)$.
This gives the property:
%\begin{equation}
%|\mathbf{M}|^2=(-1)^{k-1}|\mathbf{M}|\label{eqn:proofMp}
%\end{equation}
%Since $-1=1$ over $\Fkk$, Equation \ref{eqn:proofMp} can be simplified.
\begin{equation}
|\mathbf{M}|^2=|\mathbf{M}|\label{eqn:proofM2}
\end{equation}
$|\mathbf{M}|\in \Fkk$, and only two elements of $\Fkk$ satisfy Eqn.(\ref{eqn:proofM2}): 
$0$ and $1$. From Lemma \ref{lemma:nonzero}, $|\mathbf{M}|\neq 0$. So $|\mathbf{M}|=1$.}
\end{Proof}

The proof can be further explained through the help of an example.
\begin{Example}
{\it Over $\F_{2^3}$:
\begin{eqnarray}
A&=&a_0+a_1\alpha+a_2\alpha^2 \nonumber \\
A^2&=&a_0+a_1\alpha^2+a_2\alpha^4 \nonumber \\
A^4&=&a_0+a_1\alpha^4+a_2\alpha^8
\end{eqnarray}
From these equations, $\mathbf{M}$ is derived:
\begin{equation}
\mathbf{M}=
\begin{bmatrix}
1 & \alpha   & \alpha^2 \\
1 & \alpha^2 & \alpha^4 \\
1 & \alpha^4 & \alpha^8
\end{bmatrix}
\end{equation}
Since $\mathbf{M}$ is a Vandermonde matrix of the form $V(\alpha,\alpha^2,\alpha^4)$, 
its determinant is found by applying Eqn.(\ref{eqn:vandet}).
\begin{equation}
\mathbf{|M|}=(\alpha^4-\alpha^2)\cdot(\alpha^4-\alpha)\cdot(\alpha^2-\alpha) \label{ex:detM1}
\end{equation}
Over any $\Fkk$, $-1=1$, so Equation \ref{ex:detM1} is rewritten as
\begin{equation}
\mathbf{|M|}=(\alpha^4+\alpha^2)\cdot(\alpha^4+\alpha)\cdot(\alpha^2+\alpha) \label{ex:detM2}
\end{equation}
Note that $\mathbf{|M|}$ is non-zero since it is a product of non-zero terms.
Now compute $\mathbf{|M|^2}$ while applying Lemma \ref{lemma:raiseToP}:
\begin{eqnarray}
|\mathbf{M}|^2&=&[(\alpha^4+\alpha^2)\cdot(\alpha^4+\alpha)\cdot(\alpha^2+\alpha)]^2\\ \nonumber
&=&(\alpha^8+\alpha^4)\cdot(\alpha^8+\alpha^2)\cdot(\alpha^4+\alpha^2) \label{ex:M2eqn}
\end{eqnarray}
Over $\F_{2^3}$, $\alpha^8=\alpha$, so Equation \ref{ex:M2eqn} is further simplified.
\begin{equation}
|\mathbf{M}|^2=(\alpha+\alpha^4)\cdot(\alpha+\alpha^2)\cdot(\alpha^4+\alpha^2)
\end{equation}
Notice that $|\mathbf{M}|^2=|\mathbf{M}|$. Since $|\mathbf{M}|\neq 0$, $|\mathbf{M}|$ 
must equal $1$, as no other element of $\F_{2^3}$ can satisfy this condition.
Indeed, evaluating Equation \ref{ex:detM2} and minimizing the result based on 
the primitive polynomial, $P(x)$, that was used to construct $\F_{2^3}$ will 
always give the result $1$ regardless of which $P(x)$ is chosen.
}
%
%To the skeptic, however, Equation \ref{ex:detM2} may not look like it evaluates 
%to $1$, so let's evaluate it.
%\begin{eqnarray}
%\mathbf{|M|}&=&(\alpha^4+\alpha^2)\cdot(\alpha^4+\alpha)\cdot(\alpha^2+\alpha) \nonumber \\
%&=&\alpha^{10}+\alpha^9+\alpha^7+\alpha^6+\alpha^8+\alpha^7+\alpha^5+\alpha^4 \nonumber \\
%&=&\alpha^{10}+\alpha^9+\alpha^8+\alpha^6+\alpha^5+\alpha^4 \label{ex:detMwork}
%\end{eqnarray}
%Apply $\alpha^8=\alpha$ to Equation \ref{ex:detMwork}
%\begin{equation}
%\mathbf{|M|}=\alpha^6+\alpha^5+\alpha^4+\alpha^3+\alpha^2+\alpha
%\end{equation}
%The terms $\alpha^6$, $\alpha^5$, $\alpha^4$, and $\alpha^3$ are minimized based on
%which of the two possible primitive polynomials, $P(x)$, constructed $\F_{2^3}$.
\end{Example}

Applying Lemma \ref{lemma:MIs1} to Eqn.(\ref{eqn:cramerform}) gives the 
equation for $a_i$,
\begin{equation}
a_i = |\mathbf{M_i}|
\end{equation}
The determinant $|\mathbf{M_i}|$ can be computed symbolically as described below.
%where Eqn.(\ref{eqn:Mireduced}) gives a representation for $|\mathbf{M_i}|$.

%Thus, $F_a=\{f_{a_0},\dots,f_{a_{k-1}}\}$ is constructed as
%\begin{eqnarray}
%&f_{a_0} : a_0 + |\mathbf{M_0}|& \nonumber \\
%&\vdots& \nonumber \\
%&f_{a_{k-1}} : a_{k-1} + |\mathbf{M_{k-1}}|&
%\end{eqnarray}
%
%Afterwhich, computing $r \xrightarrow{F_a}_+r_w$ gives the final word-level 
%abstraction of the circuit, $r_w$.

\subsection{Symbolically Computing the Bit-Level Mapping}
%We know that $a_i=|\mathbf{M_i}|$ for $0\leq i<k$.
The refinement of the determinant $|\mathbf{M_i}|$ uses fundamental
symmetric polynomials.

\begin{Definition}
{\it For $x_1, \dots, x_n$ and $0 \leq j \leq n$, let $S_j(x_1, \dots, x_n )$ be the $j$-th 
{\bf Fundamental Symmetric Polynomial} in $\{x_1, \dots, x_n\}$:
\begin{equation}
S_j(x_1,\dots,x_n) = \sum\limits_{i_1 < \dots < i_j}x_{i_1}x_{i_2}\cdots x_{i_j}
\end{equation}
}
\end{Definition}

Informally, $S_j$ is the sum of all unique monomials of exactly $j$ variables, with no variable
having an exponent greater than $1$.
\begin{Example}
{\it The possible fundamental symmetric polynomials over $\{x_1,x_2,x_3\}$ are:
\begin{eqnarray}
S_0(x_1,x_2,x_3) &=& 1 \nonumber \\
S_1(x_1,x_2,x_3) &=& x_1+x_2+x_3 \nonumber \\
S_2(x_1,x_2,x_3) &=& x_1x_2+x_1x_3+x_2x_3 \nonumber \\
S_3(x_1,x_2,x_3) &=& x_1x_2x_3 \nonumber 
\end{eqnarray}
}
\end{Example}

\begin{Proposition}
{\it
Let $V_i(x_1,\dots,x_n)$, $0\leq i\leq n$, be a square Vandermonde-like matrix
derived similarly to $V(x_1,\dots,x_n)$ but with the column $\{x_1^i,\dots,x_n^i\}^T$ skipped and 
a column $\{x_1^n,\dots,x_n^n\}^T$ appended to the end:

\begin{equation}
V_i(x_1,\dots,x_n) 
=
\begin{bmatrix}
1 & x_1 & x_1^2  & \dots  & x_1^{i-1} & x_1^{i+1} & \dots & x_1^n \\
1 & x_2 & x_2^2  & \dots  & x_2^{i-1} & x_2^{i+1} & \dots & x_2^n \\
\vdots & \vdots  & \vdots & \cdot  & \vdots    & \vdots    & \cdot & \vdots    \\
1 & x_n & x_n^2  & \dots  & x_n^{i-1} & x_n^{i+1} & \dots & x_n^n
\end{bmatrix} 
\end{equation}

It is known that
\begin{equation}
|V_i(x_1,\dots,x_n)| = |V(x_1\dots,x_n)|\cdot S_{n-i}(x_1,\dots,x_n)
\end{equation}
}
\end{Proposition}

Computing $|\mathbf{M_i}|$ by interpolating along the $\mathbf{A}^T$ column in Eqn.(\ref{eqn:M_i}) gives
\begin{equation}
|\mathbf{M_i}|=\sum\limits_{j=0}^{k-1}(-1)^{(i+j)}A^{2^j}
|V_{i+1}(\alpha,\dots,\alpha^{2^(j-1)},\alpha^{2^(j+1)},\dots,\alpha^{2^(k-1)})|
\end{equation}
the final form of which is
\begin{eqnarray}
|\mathbf{M_i}| &=& \sum\limits_{j=0}^{k-1}(-1)^{j}A^{2^j} \nonumber \\
& & \cdot |V(\alpha,\dots,\alpha^{2^(j-1)},\alpha^{2^(j+1)},\dots,\alpha^{2^(k-1)})| \nonumber \\
& & \cdot S_{n-1-i}(\alpha,\dots,\alpha^{2^(j-1)},\alpha^{2^(j+1)},\dots,\alpha^{2^(k-1)}) \label{eqn:Mireduced}
\end{eqnarray}

%Applying this formulation of $|\mathbf{M_i}|$ to Equation \ref{eqn:cramerform} gives the
%final computable formula for $a_i$, $0\leq i<k$ over any $\Fpk$. However, over
%$\Fkk$, $|\mathbf{M}|=1$ which simplifies this equation to:

\section{Overall Approach}
The entirety of the word-level 
abstraction approach for a circuit with $k$-bit input $A$ and $k$-bit 
output $Z$ is summarized as follows:
\begin{enumerate}
\item Given a combinational circuit $C$, with word-level $k$-bit inputs $A$ and word-level output $Z$.
\item Select a primitive polynomial $P(x)$ of degree $k$ and construct $\Fkk$.
\item Perform a reverse-topological traversal of $C$ to find RATO, $\{x_1>x_2>\cdots>x_d>Z>A\}$, where $\{x_1,\dots,x_d\}$ are bit-level variables with $x_i$ appearing earlier in traversal than $x_j$ if $i<j$.
\item Derive the bit-level polynomials $\{f_1,\dots,f_s\}$ from $C$. These will be in the form $f_i:x_i+TAIL(f_i)$ where $x_i$ is the output of a Boolean logic gate.
\item Compose the word-level polynomials which correspond bit-level and word-level input and output:
\begin{eqnarray}
&f_A:a_0+a_1\alpha+\dots+a_{k-1}\alpha^{k-1}+A\\
&f_Z:z_0+z_1\alpha+\dots+z_{k-1}\alpha^{k-1}+Z
\end{eqnarray}
\item \label{alg:initialred}Compute the reduction $f_Z\xrightarrow{f_1,\dots,f_s,f_A}+r$.
\item If $r$ does not contain bit-level variables, then $r$ is the word-level abstraction of $C$ over $\Fkk$. Otherwise, continue to step \ref{alg:secondpoly}.
%\item \label{alg:stepCont} Construct $\mathbf{M_0},\dots,\mathbf{M_{k-1}}$ where 
%$\mathbf{M_i}$ is $\mathbf{M}$ with the column $\{\alpha^i,\alpha^{i*2},\dots,\alpha^{i*2^{k-1}}\}^T$ replaced by $\{A,A^2,\dots,A^{2^{k-1}}\}$
%\begin{equation}
%\mathbf{M}=
%\begin{bmatrix}
%1      &   \alpha           & \alpha^2           & \dots & \alpha^{k-1}\\
%1      &   \alpha^2         & \alpha^{4}        & \dots & \alpha^{(k-1)*2}\\
%1      &   \alpha^4         & \alpha^{8}        & \dots & \alpha^{(k-1)*4}\\
%\vdots & \vdots             & \vdots             & \cdot     & \vdots \\
%1      &   \alpha^{2^{k-1}} & \alpha^{2*2^{k-1}} & \dots & \alpha^{(k-1)*2^{k-1}}
%\end{bmatrix}
%\end{equation}
\item \label{alg:secondpoly} Compute $F_a=\{f_{a_0},\dots,f_{a_{k-1}}\}$ as
\begin{eqnarray}
&f_{a_0} : a_0 + |\mathbf{M_0}|& \nonumber \\
&\vdots& \nonumber \\
&f_{a_{k-1}} : a_{k-1} + |\mathbf{M_{k-1}}|&
\end{eqnarray}
where each $|\mathbf{M_i}|$ for $0\leq i<k$ is given by Eqn.(\ref{eqn:Mireduced}).
\item \label{alg:secondred}Compute $r\xrightarrow{F_a,F_0}_+ r_w$. Then $r_w$ is the word-level abstraction of $C$ over $\Fkk$.
\end{enumerate}
This approach can be easily extended to circuits with multiple word-level inputs as 
well as circuits with varying word-sizes amongst the word-level inputs and output.
\begin{Example}
\label{ex:improve2}
{\it 
Consider, again, the buggy example shown in Example \ref{ex:improve1}, 
corresponding to a buggy version of the multiplier circuit of
Fig. \ref{fig:2bitmul}. We already found 
\begin{equation}
r = (\alpha)a_1b_1 + (\alpha+1)a_1B+b_1A +
Z + (\alpha+1)AB
\end{equation}
Since $r$ contains the bit-level variable $a_1$, find 
$f_{a_1}: a_1+|\mathbf{M_1}|$. In this example, $f_A:a_0 + a_1 \alpha + A$, so
\begin{equation}
\mathbf{M}=
\begin{bmatrix}
1      &   \alpha           \\
1      &   \alpha^2         
\end{bmatrix}
\end{equation}
and
\begin{equation}
\mathbf{M_1}=
\begin{bmatrix}
1      &   A           \\
1      &   A^2
\end{bmatrix}
\end{equation}
Computing $|\mathbf{M_1}|$ finds
\begin{equation}
f_{a_1}: a_1 + A^2 + A
\end{equation}
As $r$ also contains the bit-level input $b_1$, the polynomial $f_{b_1}$ is also required. 
Since $f_B:b_0+b_1\alpha+B$ is 
isomorphic to $f_A$, $f_{b_1}$ can be derived by performing the corresponding 
substitutions in $f_{a_1}$.
\begin{equation}
f_{b_1}: b_1 + B^2 + B
\end{equation}
Now, computing $r \xrightarrow{f_{a_1},f_{b_1}} _+ r_w$ finds
\begin{equation}
r_w=Z+(\alpha) A^2 B^2+ A^2 B+(\alpha+1) A 
 B^2+(\alpha+1) A B
\end{equation}
which is indeed the polynomial representation of the buggy circuit.
}
\end{Example}

\section{Complexity Analysis}

The worst case complexity of abstracting a combinational circuit over $\Fkk[x_1,\dots,x_n]$
using the proposed approach is now analyzed. 
For simplicity, a generic circuit with a $k$-bit input $A$ and a $k$-bit output $Z$ is
examined. No assumptions are made about the type of internal gates of the circuit; 
any gate can have an arbitrary number of inputs and an arbitrary representation over $\Fkk$.

The initial step of the algorithm is to compute:
\begin{eqnarray}
& & f_z\xrightarrow{F-\{f_z\},F_0}_+ r \\
\text{where } & & f_z: z_0+z_1\alpha+\dots+z_{k-1}\alpha^{k-1}+Z
\end{eqnarray}
where $\{z_0,\dots,z_{k-1}\}$ are the bit-level outputs, $F$ is the set of polynomials
derived from the circuit, and $F_0$ is the set of vanishing polynomials.

\begin{Lemma}\label{lem:degreeOne}
No intermediate polynomial $r_i$ during the reduction process $f_z\xrightarrow{F-\{f_z\},F_0}_+ r$
will ever contain a variable with a degree larger than $1$.
\end{Lemma}
\begin{Proof}
Any intermediate result can contain $3$ types of variables:
\begin{itemize}
\item Bit-level variables: Any intermediate division which results in a polynomial containing a variable 
      $x\in\mathbb{F}_2$ with degree higher than $1$ is immediately divided by $\{x^2+x\}\in F_0$.
\item $Z$: 
As reduction proceeds,  
the term $Z$ is never modified since no polynomial other than $f_Z$ contains the variable $Z$ and $f_Z$ is 
never used during the reduction process.
\item $A$: This term is contained in $f_A:a_0+a_1\alpha+\dots+a_{k-1}\alpha^{k-1}+A$. 
Notice that $LT(f_A)=a_0\in\mathbb{F}_2$; 
since an intermediate polynomial will never contain the variable $a_0$ with a degree higher than $1$, 
reducing by $f_A$ will not create a variable $A$ with degree higher than $1$.
\end{itemize}
\end{Proof}

\begin{Lemma}\label{lem:numTerms}
Let $C\cdot M$ be a term where $C$ is a coefficient in $\Fkk$ and $M$ is a monomial
with variables in $\{x_1,\dots,x_n\}$. Since the maximum degree of any variable of 
any intermediate polynomial $r_i$ is $1$,
the maximum number of terms in any $r_i$ is $2^n$.
Similarly, the maximum number of terms of any polynomial $\in F$ is $2^n$ since these
polynomials are also guaranteed not to have any variables with degree greater than $1$.
\end{Lemma}

The order in which polynomials in $F$ and $F_0$ are used to divide $r$ is based on RATO.
Each division process divides the leading term of the intermediate polynomial.

\begin{Lemma}\label{lem:maxDivsPerLT}
Each term in an intermediate result $r_i$ is reduced at most once.
\end{Lemma}
\begin{Proof}
Assume the division is being computed by some $f\in F$. 
This division is computed as $\frac{LT(r_i)}{LT(f)}\cdot f+r_i$.
Since $LT(\frac{LT(r_i)}{LT(f)}\cdot f)=LT(r_i)$, the leading terms 
are cancelled. 
The leading term of the resulting
polynomial is strictly smaller than $LT(r_i)$ in the ordering.
As every subsequent division produces a leading term strictly smaller
than the last, $LT(r_i)$ never appears again.
Thus, each term is divided at most once.
%: once by a
%polynomial in $J$ and once by a polynomial in $J_0$.%with $LT(r) \geq LT(f)$ and $LT(r) > \frac{LT(r)}{LT(f)}$.
\end{Proof}

\begin{Lemma}
Since the maximum number of terms in any resulting intermediate polynomial $r_i$ is 
$2^n$ and each term is divided at most once, the maximum number of divisions computed
during the reduction is $2^n$.
\end{Lemma}

Assume that a monomial multiplication and monomial addition can be computed
in constant time. Each division is computed as $\frac{LT(r_i)}{LT(f)}\cdot f+r_i$.
Here, $f$ can have at most $2^n$ terms. $\frac{LT(r_i)}{LT(f)}$ is computed in
constant time. Then, $(\frac{LT(r_i)}{LT(f)})\cdot f$ requires at most $2^n$ 
monomial multiplications. As the result from a monomial multiplication may contain variables with
degrees higher than $1$, each monomial has its variables minimized, for a maximum of $2^n$ minimizations.
Finally, this result is added to $r_i$ using a maximum of $2^n$ monomial additions.
Thus each division uses at a maximum $2^n+2^n+2^n+1$ monomial operations.

\begin{Lemma}
The complexity of the reduction $f_z\xrightarrow{F-\{f_z\},J_0}_+ r$ is $O(2^{2n})$.
\end{Lemma}
\begin{Proof}
This is computed as maximum number of divisions multiplied by maximum number of
monomial additions/multiplications per division.
\begin{equation}
2^{n}\cdot (2^n+2^n+2^n+1)=2^{2n}+2^{2n}+2^{2n}+2^{n}<4\cdot 2^{2n}=O(2^{2n}).
\end{equation}
\end{Proof}

The next step is to derive the bit-level to word-level mapping $F_A$. Without 
any optimizations, this can be derived by computing Gaussian elimination of a 
$k$ by $k$ matrix. The worst case arithmetic complexity here is $O(k^3)$. 
Furthermore, this is done
in parallel with the reduction. As $k$ is much smaller than $n$, the $O(2^{2n})$
reduction easily absorbs the complexity of deriving $F_A$, i.e. $O(2^{2n})>>O(k^3)$.

The last step is computing the final reduction $r\xrightarrow{F_A,F_0}_+ r_w$.
\begin{Lemma}Any intermediate result $r_j$ during the final reduction $r\xrightarrow{F_A,F_0}_+ r_w$ 
will contain at most $2^{2k}+1$ terms.
\end{Lemma}
\begin{Proof}
Any intermediate polynomial will only contain the variables 
$\{a_0,\dots,a_{k-1},Z,A\}$. From Lemma \ref{lem:degreeOne},
all variables $\{a_0,\dots,a_{k-1}\}$ can be at most degree $1$.
Hence, there can be at most $2^k$ terms 
containing only variables $\{a_0,\dots,a_{k-1}\}$.
The variable $A$ can have a degree of at
most $2^k-1$, as any higher degree is immediately divided by $\{A^{2^k}+A\}\in J_0$.
Thus there can be at most $2^k$ terms containing only the variable $A$, which
means that there are $2^k\cdot 2^k=2^{2k}$ terms containing variables in 
$\{a_0,\dots,a_{k-1},A\}$.
The variable $Z$ is only found in $1$ term: $Z$. This term is never divided
and never modified since the only polynomial within the set of divisors
which contain the variable $Z$ is $\{Z^{2^k}+Z\}\in J_0$. So the maximum 
number of monomials in an intermediate result is $2^{2k}+1$.
\end{Proof}

\begin{Lemma}
Due to Lemma \ref{lem:maxDivsPerLT}, each term will be divided at most
once. $Z$ is never divided, so at most $2^{2k}$ divisions are computed.
\end{Lemma}

In each division, $\frac{LT(r_i)}{LT(f)}\cdot f+r_j$, $f$ has at most 
$2^k+1$ terms, since it is of the form $a_i+\Func(A)$. 
Thus, there is $1$ monomial division, at most $2^k+1$ monomial multiplications.
As before, the degree of each resulting monomial needs to be minimized, 
which performs a maximum of $2^k+1$ minimizations. 
Finally, at most and $2^k+1$ monomial additions are computed when adding this 
result to $r_j$. Thus, the total number of monomial operations per division is:
\begin{equation}
1+(2^k+1)+(2^k+1)+(2^k+1)=3\cdot 2^k+4
\end{equation}

\begin{Lemma}
The complexity of the final substitution, which is computed as the reduction
$r\xrightarrow{F_a,F_0}_+ r_w$, is $O(2^{3k})$.
\end{Lemma}
\begin{Proof}
This is computed as the maximum number of divisions multiplied by the maximum 
number of monomial operations per division.
\begin{equation}
(2^{2k})\cdot(3\cdot 2^{k}+4)=3\cdot 2^{3k}+\cdot 2^{2k+2} %\nonumber
%\end{equation}
%\begin{equation}
<3\cdot 2^{3k+2}+2^{3k+2}=2^{3k+5}=O(2^{3k})
\end{equation}
\end{Proof}

\begin{Theorem}
The worst-case complexity for the abstraction algorithm is
\begin{equation}
O(2^{2n})+O(2^{3k})
\end{equation}
\end{Theorem}

The first reduction is the main bottleneck in the abstraction computation 
since $n$ is much larger than $k$.

\section{Conclusions}
This chapter proposed an approach which solves the problem of word-level abstraction
from bit-level circuits using symbolic computation. 
Using an improved ordering (RATO) a reduction is computed 
to derive a polynomial, $r$, which contains only bit-level inputs $\{a_0,\dots,a_{k-1}\}$ 
and the word-level datapaths $A$ and $Z$. This reduction is efficiently computed 
using an $F4$-style reduction engine. Next, an equation of the form $a_i=\Func_{a_i}(A)$, which
maps each bit-level input to its corresponding word-level representation, is computed
using a binomial expansion over $\Fkk$. Substituting variable each $a_i$ in $r$ by $\Func_{a_i}(A)$
gives $r_w$, which is the {\it canonical, word-level, polynomial representation of the circuit}.
Finally, a complexity analysis of the overall approach is provided.

The abstraction approach is directly applicable only to circuits with equivalent
data-path sizes amongst the inputs and output, i.e. every word input and output 
is of size $k$. The next chapter generalizes the abstraction approach to be applicable
to any arbitrary combinational circuit.

### 2016-09-21 dissertation/intro.tex
init dissertation writing

\chapter{Introduction} \label{ch:intro}
There is an ever-increasing need for secure communication within information 
technology. Security of sensitive information relies more and more heavily 
on encryption methodologies implemented in hardware by cryptographic 
circuits. One of the most prominent of these methodologies is Elliptical 
Curve Cryptography (\emph{ECC}), which provides more strength per encryption bit 
than other encryption methodologies.
The main building blocks of ECC hardware implementations 
are fast, \emph{custom-built Galois field arithmetic circuits}.
These circuits are notoriously hard to verify, yet their correctness is 
vitally important in critical applications. In \cite{crypto:bug_attacks}, 
for example, it is shown that a bug in the hardware could lead to the full 
leakage of the secret cryptographic key, which could compromise the entire
system. Thus, formal verification is imperative in Electronic Design 
Automation (\emph{EDA}) when dealing with cryptographic circuits.

To facilitate this verification, it is highly desirable to obtain a {\it word-level 
representation of the datapath of the ECC arithmetic block} from its 
bit-level implementation. Ideally, this abstraction should be {\it canonical},
as this allows the it to be directly applicable to equivalence checking. 
Such a canonical, word-level abstraction of the 
Galois field arithmetic block would not only make it easier to verify and 
reason 
about the cryptographic system as a whole, but also enable the use of 
higher level abstraction and synthesis tools. 
As arithmetic circuits are custom-designed, often modularly, using Galois field arithmetic blocks, 
the abstraction should also exploit the hierarchical nature of the circuitry.
Due to the modular circuit structure, abstraction of each arithmetic block
becomes the key in verification of the full circuit. 
Practical 
applications of ECC dictate a datapath of a minimum of $163$-bits, up to 
$571$-bits, as 
designated by the National Institute for Standards and Technology (NIST).
However abstraction of 
Galois field arithmetic circuits has been infeasible for data-paths beyond 
$16$ bits.

This dissertation proposes an algebraic geometry based approach to abstract canonical, 
word-level representations of bit-level Galois field arithmetic circuits.
The approach is able to abstract representations
for circuits up to 571 bits in size, which is the largest NIST
standard for datapath size in ECC. 
Verification of circuits for which this abstraction has been computed is shown to 
be trivial; thus, the focus is on deriving the abstraction quickly and efficiently.
%Furthermore, we show how we can use these abstract representations to 
%perform improved formal verification of these circuits.

\section{Hardware Design and Verification Overview}
The typical design flow of a hardware system, 
as shown in Fig. \ref{fig:cadflow}, starts with a hardware system 
specification, which describes the necessary functions and parameters that 
the system must perform and adhere to. The specification is typically
modelled using a transaction-level model (\emph{TLM}), which describes 
communication details between large circuit modules. 
The TLM is then translated 
into a register-transfer-level (\emph{RTL}) description, which is composed of abstracted, 
interconnected circuit blocks that compose the entire system. RTL is 
typically implemented 
in hardware description languages (\emph{HDL}) such as Verilog and VHDL, 
which are the most popular choices in the industry.
Next, the RTL is optimized and converted into a \emph{netlist}, i.e. a large 
collection of small physical blocks (MOSFET, Boolean logic gates, etc.) and 
the inter-connections (wires) between them. 
Lastly, the netlist is further optimized and then mapped onto a physical space 
on a chip, which is then sent off for fabrication. This entire design flow
is automated by Computer-Aided Design (\emph{CAD}) tools.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/designFlow}
}
\caption{Typical hardware design flow.}
\label{fig:cadflow}
\end{figure}
}

%\subsection{The Hardware Verification Imperative}
When moving from one abstraction level of the hardware design process to the 
next, an important issue arises: how can one ensure that the functionality of 
the optimized design matches original spec? 
Bugs in hardware design which are 
not caught early can have costly effects later, such as the need for a 
redesign. Bugs in arithmetic circuits can be especially catastrophic.
One infamous example is the 1994 floating point division (FDIV)
bug that affected the Intel Pentium chip \cite{nicely:FDIV}, 
and subsequently cost the company \$475 million because it was  
discovered after the chip's release. In another more fatal case, during the
Gulf war, an American Patriot Missile battery failed to intercept an incoming
enemy missile due to an arithmetic error \cite{arnold:patriot}.
Since hardware bugs can have significant consequences, 
there has been extensive work in field of hardware verification to find and
eliminate bugs prior to fabrication.

The two main methodologies used in hardware verification are simulation and 
formal verification. \emph{Simulation} checks correctness by applying exhaustive 
assignments to the circuit inputs and verifying correctness of the output. 
This ensures that the circuit performs as designed under all possible 
inputs. Such exhaustive testing is quite effective for smaller circuits. 
However, as the size of the circuit increases, it becomes 
computationally infeasible to simulate all possible test vectors. This is the 
case with Galois field arithmetic circuits, which are commonly very large in 
real-world applications. Often for such large circuits, simulations of a smaller and more 
manageable subset of test vectors are employed to catch bugs. While these tests
can increase confidence in the correctness of the design,  
{\it they do not guarantee correctness} since every data-flow of the design hasn't
been analyzed.
%Such is the case with large Galois field arithmetic circuits, so we instead 
%focus on formal verification.

\section{Formal Verification}
Instead of simulating input vectors, \emph{formal verification} utilizes 
mathematical theory to reason about the correctness of hardware designs.
%which overcomes some limitations of simulation. 
Formal verification has two main forms: property checking and equivalence 
checking.

{\it Property checking} (or property verification) verifies
that a design satisfies certain given properties. Property checking is done mainly 
in the form of theorem proving, model checking, or approaches which 
combine the two.
\begin{enumerate}
\item \emph{Theorem proving} \cite{theoremproving:91} requires the existence of
mathematical descriptors of the specification and implementation of the 
circuit. Theorem provers apply mathematical rules to these descriptors to
derive new properties of the specification. In this way, the tool can reduce
a proof goal to simpler sub-goals, which can be automatically verified.
However, generating the initial proof-goal requires extensive guidance from
the user, so there is an overall lack of automation in theorem 
proving.
\item \emph{Model checking} \cite{modelcheck:99} is an approach
to verifying finite-state systems where specification 
properties are modeled as a system of 
logic formulas. The design is then traversed to check if the 
properties hold. If the design is found to violate a
particular property, a counter-example is generated which exercises the
incorrect behavior in the design. Such counter-examples allow the designer
to trace the behavior and find where the error in the design lies.
Modern model checking techniques use the result to automatically refine
the system and perform further checking.
These tools are typically automated, and thus have found widespread 
use in CAD tool suites.
\end{enumerate}

{\it Equivalence Checking} verifies that two different representations of
a circuit design have equivalent functionality. An example of equivalence
checking as it applies to the hardware design flow is shown in
Fig. \ref{fig:equivflow}.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/designVerification}
}
\caption{Equivalence checking as applied to the hardware design flow.}
\label{fig:equivflow}
\end{figure}
}

There are two major
equivalence checking techniques: graph-based
and satisfiability-based.
\begin{enumerate}
\item \emph{Graph-based} techniques construct a canonical graph 
representation, such as a Binary Decision Diagram (\emph{BDD}) or one of
its many variants, of each circuit. A linear comparison is then conducted to 
determine whether the two graphs are isomorphic. Since the graph 
representation is canonical, the graphs of the two circuits will be 
equivalent if and only if the circuits perform the same function.
\item \emph{Satisfiability} techniques construct a miter of the two circuits,
typically in a graph such as an And-Inverter graph (\emph{AIG}). A
\emph{miter} is a combination of the two circuits with one bit-level output, which 
is only in a "1" state when the outputs of the circuits differ given 
the same given 
input, as shown in Fig. \ref{fig:miter}. 
A satisfiability (\emph{SAT}) tool \cite{csat} 
is then employed to simplify the graph and find a solution to the miter, 
i.e. find an input for which the 
miter output is "1". If a solution is found, this solution acts as a 
counter-example of when the circuit outputs differ; otherwise the circuits
are functionally equivalent.
\end{enumerate}

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.8\textwidth]{./figures/betterMiter}
}
\caption{A miter of two circuits.}
\label{fig:miter}
\end{figure}
}

%\subsection{Computer-Algebra Based Formal Verification}
Certain formal verification methods use \emph{computer-algebra} and \emph{algebraic
geometry} techniques based on mathematical theories.
Unlike SAT-based verification, modern algebraic geometry 
techniques do not explicitly solve the constraints to find a solution; 
rather, they reason about the presence or absence of solutions, or explore
the geometry of the solutions.
These methods \cite{Avrunin:CAV} \cite{condrat-tacas07} \cite{gbverify:2007} 
transform the circuit design into a polynomial system. Typically, this system
of polynomials is then used to compute a Gr\"obner basis \cite{gb_book}. 
Computation of Gr\"obner bases allows for 
the easy deduction of important properties of a polynomial system, 
such as the presence or absence of 
solutions. These properties are then leveraged to perform 
verification. Unfortunately, such a computation 
has been shown to be doubly 
exponential in the worst case, and thus these methods have not been 
practical for real-world applications. However, recent
breakthroughs in computer-algebra hardware verification have shown
that it is possible to overcome the complexity of this computation while
still utilizing the beneficial properties of a Gr\"obner bases
\cite{lv:phd}.

\section{Importance of Word-level Abstraction}
Most formal verification techniques can benefit from word-level abstractions 
of the circuits they verify.
Abstraction is defined as state-space reduction, i.e{\text . }abstraction
reduces state-space by mapping the set of states of a system to a smaller 
set of states. Because the new representation contains fewer states, it
is easier to comprehend and thus easier to use. 
Word-level abstraction focuses specifically on abstracting a word-level
representation of a circuit out of a bit-level representation. For example,
a bit-level representation of an integer multiplier is represented by a
collection of Boolean inputs and outputs, whereas a word-level
abstraction hides the underlying logic and represents the circuit as two 
integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
multiplier increases, the logical implementation of the multiplier grows (typically
exponentially) while the word-level abstraction stays the same.

Word-level abstractions have a wide variety applications in formal 
verification. Theorem proving techniques can leverage abstraction as an 
automatic decision procedure or as a canonical reduction 
engine. For example, since RTL is composed of circuit blocks that represent 
the underlying circuit, RTL verification methods can exploit 
abstractions of these blocks.
This is seen in the following RTL verification methods:
\begin{itemize}  
\item Model checking \cite{kroening:model}, 
where an approximation abstraction of RTL blocks is generated and then 
refined.
\item Graph-based equivalence 
checking \cite{WLS} \cite{arditi:bmd}, where abstraction methods are used
to generate a canonical word-level graph representation of the circuit.
\item Satisfiability-based equivalence checking \cite{lpsat}, where 
abstractions are used identify symmetrics and similarities in order to 
minimize the amount of logic that is sent to the 
SAT tool. 
\end{itemize}

Other equivalence checking techniques that employ abstractions 
include satisfiability modulo theory ({\it SMT}) techniques \cite{boolector} \cite{bryant:tacas07}, 
which are similar to SAT except they operate on higher-level data
structures (integers, reals, bit vectors, etc.), as well as 
constraint solving techniques \cite {ms:research} \cite{tew:iccad08}.
In general, RTL equivalence checking approaches would ideally maintain a 
high-level of abstraction while still retaining sufficient lower-level 
functional details  (such as bit-vector size, precision, etc) 
\cite{gupta_survey}.

Word-level hardware abstractions also have applications in RTL and datapath 
synthesis \cite{demicheli:iccad_98} \cite{demicheli:dac_99}
\cite{demicheli:tcad_03}. 
Abstractions of circuits allow for design reuse, which allows for tool-automated 
synthesis of larger circuit blocks.
Since hardware design specifications tend to be word-level, synthesis tools 
can use these larger circuit blocks to generate and optimize the
datapaths and create the RTL of the system. Thus, in order for a circuit to 
be used by these automated synthesis tools, its word-level abstraction must
be known.

Finally, abstractions can also be applied to detect malicious 
modifications to a circuit, potentially inserted as a hardware trojan horse.
Hardware trojans, a relatively new security concern in the hardware 
industry, use certain techniques to add incorrect behavior to a 
design. 
This behavior is only activated under certain rare circumstances that only 
the mal-intent designer has knowledge of.
The behavior is purposely hidden and is very difficult to encounter during 
simulation of the design. A manufactured chip with a subsystem 
that contains a hardware trojan could compromise the entire system in which 
it is used.
In some hardware trojan cases, formal verification techniques may be applied 
to catch a bug in a design and provide a counter-example which exercises it. 
However, it can be difficult to tell whether the bug in the design was 
introduced intentionally of not. On the other hand, word-level abstractions 
of bit-level circuits {\it effectively reverse-engineer the true function 
implemented by the circuit}, which could be used to determine the designer's 
true intention.

\section{Dissertation Objective, Motivation, and Contributions}
This dissertation focuses on abstracting a canonical, 
word-level representation of hardware (bit-level) implementations of 
combinational circuits. The proposed technique is a full abstraction solution which can be 
applied to any arbitrary acyclic combinational circuit. 
It is particularly efficient when applied to Galois field arithmetic circuits.
Using this technique, if the abstraction of the circuit's implementation and its
specification are found, they can be easily compared to determine equivalence.
Implementation of a custom software tool, developed to compute the abstractions, is
also described.

\subsection{Motivating Application}
The motivation for this work comes from applications of Galois field 
arithmetic circuits in elliptical curve cryptography ({\it ECC}) hardware systems.
The main operations of encryption, decryption, and 
authentication in ECC rely on operations performed on elliptic curves, which 
are implemented in hardware as polynomial functions over Galois fields. 
To be applicable in real-world situations, ECC data-paths
should be a minimum of $163$-bits wide, which is the minimum NIST standard, 
up to a recommended size of $571$-bit operand widths. Many non-ECC cryptosystems
have datapaths on the order of $1000$-bits.

A Galois field arithmetic circuit with a datapath size of $k$ 
is built as a Boolean function: $\mathbb{B}^k \rightarrow \mathbb{B}^k$. 
This function is mapped to an operation 
$f:\mathbb{F}_{2^k} \rightarrow \mathbb{F}_{2^k}$ 
over the Galois field $\mathbb{F}_{2^k}$. 
These circuits are custom-built, modular
systems which cannot be synthesized due to their complex nature. Thus, 
formal verification is needed to ensure they operate correctly.

Recent computer-algebra based formal verification techniques have been
able to perform verification of Galois field arithmetic circuits with
a datapath size up to $163$-bits \cite{lv:phd}. Word-level abstractions 
of Galois
field arithmetic circuits could be used to further improve these formal 
verification techniques to allow for verification of larger circuits, as
well as provide the other benefits of word-level abstraction.
However, there is currently no technique for computing word-level 
abstractions of Galois field circuits of any practical size.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/interpolate}
}
\caption{Circuit with $k$-bit input $A$ and $k$-bit output $Z$. 
Abstraction to be derived as $Z=\Func(A)$.}
\label{fig:abstractA_Z}
\end{figure}
}

While the motivation comes from the need to verify Galois field arithmetic
circuits, the presented approach can be generalized to be applicable to any
combinational acyclic circuit.
Any such circuit with a $k$-bit input $A$ and a $k$-bit output $Z$, such
as the one shown in Fig. \ref{fig:abstractA_Z}, computes 
$f: \mathbb{B}^k \rightarrow \mathbb{B}^k$ and can thus be analyzed as the
function $f: \Fkk \rightarrow \Fkk$. Over $\Fkk$ this function can be 
represented as the polynomial $Z=\Func(A)$. This is trivially generalized
when there are multiple $k$-bit inputs $A_1,A_2,\dots,A_i$, i.e. $Z=\Func(A_1,\dots,A_i)$.
Now assume the word-size of the input differs from the output, that is the circuit
computes $f: \mathbb{B}^m \rightarrow \mathbb{B}^n$ for $m \neq n$. This can 
be represented as a function over Galois fields as $f: \F_{2^m} \rightarrow \F_{2^n}$.
This function can be analyzed over the field $\Fkk$ such that $\Fkk \supset \F_{2^m}$ 
and $\Fkk \supset \F_{2^m}$, where $k=LCM(m,n)$.

\subsection{Dissertation Contributions}
To solve the problem of word-level abstraction, this dissertation proposes 
a full solution consisting of three main contributions.

\begin{enumerate}
\item A theory for finding the word-level abstraction from a bit-level circuit over Galois fields is created.
The given bit-level circuit implementation is modelled as a system of
polynomials over the field.
This theory is derived using techniques from computer-algebra, notably the theory of
\Grobner basis \cite{pruss:iwls13}.
\item Using this theory, new algorithms based on symbolic computation are developed to 
derive the word-level abstraction. The algorithms are designed to be applicable to 
industry-size arithmetic circuits over Galois fields \cite{pruss:dac14}\cite{sun:date15}.
A complexity analysis of the algorithmic approach is also presented. Furthermore, the
approach is also generalized to make it applicable to arbitrary combinational circuits.
Finally, we show how the approach can be used to exploit the hierarchical structure of
large Galois field multipliers designed over composite fields.
\item A custom software tool implementation of the algorithmic approach is described, including
an analysis of efficient data structures designed for this purpose \cite{pruss:tcad15}.
\end{enumerate}

Experiments show that the proposed solution can abstract canonical, word-level, 
polynomial representations of Galois field arithmetic circuits up to $1024$-bits in
size, while other contemporary approaches are infeasible beyond a $32$-bit designs.

%an 
%approach based on computer-algebra, algebraic geometry - notably the theory of
%{\it Gr\"obner bases} \cite{gb_book} \cite{buchberger_thesis}.
%and {\it elimination ideals} \cite{ideals:book}, as our approach for 
%abstracting canonical word-level representations from bit-level Galois 
%field arithmetic circuits. 
%Computer-algebra techniques easily integrate with 
%Galois field theory and allow for polynomial abstraction of the circuit over 
%the field itself. Thus, if the circuit computes a function over some field $
%\mathbb{F}_{2^k}$, the resulting abstraction is a word-level polynomial 
%function over the same field $\mathbb{F}_{2^k}$.

%The given bit-level circuit implementation is first modeled as a system of
%polynomials over the field. Using the theory of elimination ideals, we 
%derive an elimination term ordering and prove that by using this ordering we 
%can obtain a canonical word-level representation of a circuit by 
%computing a Gr\"obner basis of the polynomials. However, complexity 
%of the computation of a Gr\"obner basis proves to be prohibitive for circuits
%of practical size. Therefore, we employ a select sub-set of computations 
%from the Gr\"obner basis theory to overcome this complexity and obtain the
%abstraction. We prove that this simpler computation can be 
%performed via a polynomial reduction process, and engineer a custom 
%verification tool for this purpose. Using this approach, we are able to 
%successfully 
%abstract word-level representations of Galois field arithmetic circuits up 
%to $571$-bits, which is the {\it largest recommended NIST standard} for ECC.

\section{Dissertation Organization}
The rest of this dissertation is organized as follows. Chapter
\ref{ch:prev} reviews previous applicable work and highlights their
drawbacks with respect to the canonical, word-level abstraction problem. 
Chapter \ref{ch:prelim} describes the properties of Galois fields, 
$\mathbb{F}_{2^k}$, and explains the process of constructing them.
It also describes how to design arithmetic circuits over such fields, their 
complexities, and the role of these circuits in Elliptic Curve Cryptography.
Chapter \ref{ch:ideals} provides a theoretical background of 
computer-algebra and Gr\"obner bases and explains their application
to Galois fields. 
Chapter \ref{ch:abstract} describes an approach to abstract 
word-level polynomial representations of combinational circuits using a 
Gr\"obner basis computation.
Chapter \ref{ch:improv} improves on this word-level abstraction approach to 
make it applicable to much larger circuits.
Chapter \ref{ch:generalize} generalizes the abstraction approach to make
it applicable to circuits with varying operand word-lengths. It also describes how the 
approach can take advantage of the hierarchy of arithmetic circuits designed over
composite fields.
Chapter \ref{ch:implement} describes the implementation details of a 
custom abstraction tool and gives experimental results of
abstracting large Galois field multiplier circuits.
Chapter \ref{ch:concl} concludes the dissertation and outlines potential future 
research for continuation of this work.

### 2016-09-21 dissertation/normal.tex
init dissertation writing

\chapter{Functional Verification of Sequential Normal Basis Multiplier}
\label{ch:normal}
In order to utilize our traversal algorithm, it is necessary to find out
a sort of suitable circuit benchmarks which is easy to compute its
Gr\"bner basis (GB). From the work of Lv et al. \cite{lv_dissertation},
we learn that arithmetic circuits in Galois field (GF) is
convertible to an ideal of circuit polynomials, and the 
ideal generators form a GB itself with reverse topological
term order. Furthermore, according to the work of Pruss et al.
\cite{tim_dissertation}, with a limited computation complexity,
we can abstract the word-level signature of an arithmetic 
component working in GF. Thus, we consider the possibility 
of applying our traversal algorithm on sequential Galois
field circuits. In each frame, we can use the techniques 
from \cite{tim_dissertation} to abstract the word-level
signature of the combinational logic, which corresponds
to the transition function in our traversal algorithm.
As a result, we manage to find a type of sequential GF multiplier
which we can apply our traversal algorithm to actually 
verify its functional correctness.

\section{Motivation}
\label{sec:normal_motiv}

### 2016-09-25 dissertation/normal.tex
dissertation -- NB part 1

@@
 convertible to an ideal of circuit polynomials, and the 
-ideal generators form a GB itself with reverse topological
+ideal generators form a GB themselves when applying reverse topological
 term order. Furthermore, according to the work of Pruss et al.
@@
 \label{sec:normal_motiv}
+From the preliminaries (Chapter \ref{ch:prelim}) about FSMs, we learn that the
+Moore machine does not rely on inputs for state transitions. 
+As depicted in Figure \ref{fig:Moore}(a), a typical Moore machine implementation
+consists of combinational logic component and register files, where
+$r_0,\dots,r_k$ are present state (PS) variables 
+standing for state inputs (SI), and $r_0',\dots,r_k'$ are next state (NS) variables standing for
+state outputs (SO). Figure \ref{fig:Moore}(b) shows the state transition graph (STG) of 
+a Moore machine with $k+1$ distinct states. We notice that it forms a simple chain,
+with $k$ consecutive transitions the machine reaches final state $R_k$.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=3.5in]{newfig/Moore.eps}
+\caption{A typical Moore machine and its state transition graph}
+\label{fig:Moore}}
+\end{figure}
+
+In practice,
+some arithmetic components are designed in sequential circuits similar to the structure in 
+Figure \ref{fig:Moore}(a). Initially the operands are loaded into the registers, 
+then the closed circuit executes without taking any additional information from outside,
+and store the results in registers after $k$ clock cycles. Its behavior can be described using
+STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, $R_init$ is the initial
+state (usually reset to all zeros), $R_1$ to $R_{k-1}$ are intermediate results stored as SO of current state and SI
+for next state, and $R_k$ (or $R_{final}$) is the final result given by arithmetic circuits (and equals to the answer
+to arithmetic function when circuit is working functional correctly).
+This kind of design results in 
+reusing a smaller combinational logic component such that the area cost is greatly optimized.
+However, it also brings difficulties in verifying the the circuit functions.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=3.5in]{newfig/convention.eps}
+\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking}
+\label{fig:convention}}
+\end{figure}
+
+Conventional methods to such a sequential circuit may consist of unrolling the circuit for 
+$k$ time-frames, and performing an equivalence checking between the unrolled machine and
+the specification function. However, the number of gates will grow fast when doing unrolling
+on bit-level. Meanwhile the structural similarity based equivalence checking techniques 
+will fail when the sequential circuit is highly customized and optimized from the naive specification 
+function. As a result, conventional techniques is grossly inefficient for large circuits.
+Therefore, a new method based on our proposed word-level FSM traversal technique is worthy to be explored.
+
+\section{Galois Field multiplier and Normal Basis}
+\subsection{Multiplier in Galois Field}
+
+\subsection{Introduction to Normal Basis}

### 2016-09-28 dissertation/normal.tex
dissertation:part normal update 2

@@
 
-\section{Galois Field multiplier and Normal Basis}
-\subsection{Multiplier in Galois Field}
+\section{Normal Basis Multiplier over Galois Field}
+Given a Galois field (GF) $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
+Its elements can be written in polynomials of $\alpha$, when there is an irreducible polynomial $p(\alpha)$
+defined.
 
-\subsection{Introduction to Normal Basis}
+If we use a basis $(1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1})$, we can easily transform polynomial representations
+to binary bit-vector representations by recording the coefficients. For example,
+
+\begin{table}[H]
+\centering
+\caption{Bit-vector, Exponential and Polynomial representation of
+elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
+\pmod{x^4+x^3+1}$}
+\begin{tabular}{|c|c||c|c|} 
+\hline
+$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
+\hline
+$0000$        & $0$           & $1000$  &$\alpha^3$\\
+\hline
+$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
+\hline
+$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
+\hline
+$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
+\hline
+$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
+\hline
+$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
+\hline
+$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
+\hline
+$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
+\hline
+\end{tabular}
+\label{table:booltogalois}  
+\end{table}
+
+Basis $(1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1})$ is called {\bf standard basis} (StB), which results in
+a straightforward representation for elements, and operations of elements such as addition and subtraction.
+The addition/subtraction of GF elements in StB follows the rules of polynomial addition/subtraction
+where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} in
+Boolean algebra, element $A$ add/subtract by element $B$ in StB is defined as
+\begin{align}
+A+B = A-B &= (a_0,a_1,\dots,a_{k-1}) \xor (b_0,b_1,\dots,b_{k-1}) \nonumber\\
+&=(a_0\oplus b_0, a_1\oplus b_1,\dots,a_{k-1}\oplus b_{k-1}) 
+\end{align}

### 2016-09-30 dissertation/normal.tex
dissertation: part normal update 3

@@
 
-If we use a basis $(1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1})$, we can easily transform polynomial representations
+If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
 to binary bit-vector representations by recording the coefficients. For example,
@@
 
-Basis $(1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1})$ is called {\bf standard basis} (StB), which results in
+Basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$ is called {\bf standard basis} (StdB), which results in
 a straightforward representation for elements, and operations of elements such as addition and subtraction.
-The addition/subtraction of GF elements in StB follows the rules of polynomial addition/subtraction
+The addition/subtraction of GF elements in StdB follows the rules of polynomial addition/subtraction
 where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} in
-Boolean algebra, element $A$ add/subtract by element $B$ in StB is defined as
-\begin{align}
-A+B = A-B &= (a_0,a_1,\dots,a_{k-1}) \xor (b_0,b_1,\dots,b_{k-1}) \nonumber\\
-&=(a_0\oplus b_0, a_1\oplus b_1,\dots,a_{k-1}\oplus b_{k-1}) 
+Boolean algebra, element $A$ add/subtract by element $B$ in StdB is defined as
+\begin{align}\label{eqn:StdB}
+A+B = A-B &= (a_0,a_1,\dots,a_{k-1})_{StdB} \xor (b_0,b_1,\dots,b_{k-1})_{StdB} \nonumber\\
+&=(a_0\oplus b_0, a_1\oplus b_1,\dots,a_{k-1}\oplus b_{k-1})_{StdB} 
 \end{align}
+
+Besides addition/subtraction, multiplication is also very common in arithmetic circuit design.
+The multiplication of GF elements in $\Fkk$ in StdB follows the rule of polynomial multiplication.
+However, it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
+in bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
+the area and delay of circuit will be costly.
+
+In order to lower down the complexity of arithmetic circuit design, Massey and Omura \cite{MasseyOmura} % ref 7 in RH paper
+use a new basis to represent GF elements, which is called {\bf normal basis} (NB).
+A normal basis over $\Fkk$ is written in the form of
+\begin{equation*}
+N.B. ~~~ \N = \{\beta,\beta^2,\beta^4,\beta^8,\dots,\beta^{2^{k-1}}\}
+\end{equation*}
+Respectively, a field element in NB representation is actually
+\begin{align*}
+A &= (a_0,a_1,\dots,a_{k-1})_{NB} \\
+  &= a_0\beta+a_1\beta^2+\cdots+a_{k-1}\beta^{2^{k-1}} \\
+  &= \sum_{i=0}^{k-1} a_i\beta^{2^i}
+\end{align*}
+
+According to the definition, a normal basis is a vector where the next entry is the square of the former one.
+We note that the vector is cyclic, i.e. $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
+{\bf Normal element} $\beta$ is an element from the field which is used to construct the normal basis,
+and can be represent as a power of primitive element $\alpha$: 
+\begin{equation*}
+\beta = \alpha^t, ~~~ 1\leq t<2^k
+\end{equation*}
+
+The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
+However, what makes NB powerful is its property when doing multiplications and exponentiations.
+The following lemmas and examples illustrate this fabulous property very well.
+\begin{Lemma}[Square of NB]
+In $\Fkk$, equation 
+\begin{equation*}
+(a+b)^2 = a^2 + b^2
+\end{equation*}
+has been proved. According to the \textbf{binomial theorem}, it can be extended to
+\begin{align*}
+&(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{n-1}\beta^{2^{k-1}})^2 \\
+&= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
+&= b_{k-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}}
+\end{align*}
+\end{Lemma}
+This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
+Obviously, StdB representation does not have this benefit.
+
+\begin{Example}[Square of NB]
+In GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $x^3 + x + 1$, the standard basis is denoted as 
+$\{ 1, \alpha, \alpha^2\}$ where $\alpha^3+\alpha+1=0$.
+Let $\beta = \alpha^3$, then $\N = \{ \beta, \beta^2, \beta^4\}$ forms a normal basis. 
+Write down element $E$ using both representations:
+\begin{align*}
+E &= (a_0,a_1,a_2)_{StdB} = (b_0,b_1,b_2)_{NB} \\
+  &= a_0 + a_1\alpha + a_2\alpha^2 = b_0\beta + b_1\beta^2 + b_2\beta^4
+\end{align*}
+Compute the square of $E$ in StdB first:
+\begin{align*}
+E^2 &= a_0 + a_1\alpha^2 + a_2\alpha^4 \\
+    &= a_0 + a_2\alpha + (a_1 + a_2)\alpha^2 \\
+    &= (a_0,a_2,a_1+a_2)_{StdB}
+\end{align*}
+When it is computed in NB, we can make it very simple:
+\begin{align*}
+E^2 &= \overset{\xrightarrow{Cyclic~~shift}}{(b_0,b_1,b_2)}_{NB} \\
+	&= (b_2,b_0,b_1)_{NB}
+\end{align*}
+\end{Example}
+
+\begin{Example}
+\emph{NB multiplication:} Assume there are 2 binary vectors representing 2 operands in normal
+basis: $A = (a_0, a_1, \dots, a_{n-1}), B = (b_0, b_1, \dots, b_{n-1})$; similarly the product can also be written
+as: $C = A*B = (c_0, c_1, \dots, c_{n-1}).$
+
+Then the highest digit of product can be represented by a function: $c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, 
+\dots, b_{n-1})$. Square both side: $C^2 = A^2*B^2$, i.e. the second highest digit $c_{n-2} = f(a_{n-1}, a_0, a_1, 
+\dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2}).$ By this method it is easy to get all digits of product $C$.
+\end{Example}
+
+\begin{Example}
+\emph{$\lambda$-Matrix:} We can employ a binary $n\times n$ matrix $M$ to describe the "function"
+ mentioned above: $c_{n-1} = f(A, B) = A \cdot M \cdot B^T$, $B^T$ denotes vector transposition. 
+More specifically, we denote the matrix by \emph{$k$-th $\lambda$-Matrix}: $c_k = A \cdot M^{(k)} \cdot B^T$.
+Then $c_{k-1} = A \cdot M^{(k-1)} \cdot B^T = rotate(A) \cdot M^{(k)} \cdot rotate(B)^T$, which means
+by right and down shifting $M^{(k-1)}$ we can get $M^{(k)}$.\\
+In $F_{2^3}$ constructed by $\alpha^3 + \alpha + 1$, let $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
+is a normal basis. $0$-th $\lambda$-Matrix
+\begin{equation}
+M^{(0)} = \left(
+\begin{array} {lcr}
+0 & 1 & 0\\
+1 & 0 & 1\\
+0 & 1 & 1
+\end{array} \right).
+\end{equation}
+i.e.,
+\begin{equation}
+c_0 = (a_0\  a_1\  a_2)\left(
+\begin{array} {lcr}
+0 & 1 & 0\\
+1 & 0 & 1\\
+0 & 1 & 1
+\end{array} \right)\left(
+\begin{array} {lcr}
+b_0\\
+b_1\\
+b_2
+\end{array} \right).
+\end{equation}
+\end{Example}
+
+$\lambda$-Matrix is defined with cross-product terms from multiplication, which is 
+\begin{equation}
+Product C = (\sum_{i=0}^{n-1}a_i\beta^{2^i})(\sum_{j=0}^{n-1}b_j\beta^{2^j}) = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}a_ib_j\beta^{2^i}\beta^{2^j}
+\end{equation}
+The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
+normal basis, i.e.
+\begin{equation}
+\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1}\lambda_{ij}^{(k)}\beta^{2^k}, \ \ \lambda_{ij}^{(k)} \in F_2.
+\end{equation}
+Substitution yields, result is an expression for k-th digit of product as showed in \textit{Example B.3}:
+\begin{equation}
+c_k = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}\lambda_{ij}^{(k)}a_ib_j
+\end{equation}
+$\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ in $k$-th $\lambda$-Matrix.
+
+% After fixing this, add a whole-piece StdB vs NB example, list their cost as the conclusion

### 2016-10-07 dissertation/normal.tex
dissertation chapter Normal: update 4

@@
 \centering{
-\includegraphics[width=3.5in]{newfig/Moore.eps}
+\includegraphics[width=5in]{newfig/Moore.eps}
 \caption{A typical Moore machine and its state transition graph}
@@
 \centering{
-\includegraphics[width=3.5in]{newfig/convention.eps}
+\includegraphics[width=5in]{newfig/convention.eps}
 \caption{Conventional verification techniques based on bit-level unrolling and equivalence checking}
@@
 \begin{Lemma}[Square of NB]
+\label{lem:squareNB}
 In $\Fkk$, equation 
@@
 \begin{align*}
-&(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{n-1}\beta^{2^{k-1}})^2 \\
+&(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}})^2 \\
 &= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
@@
 
-\begin{Example}
-\emph{NB multiplication:} Assume there are 2 binary vectors representing 2 operands in normal
-basis: $A = (a_0, a_1, \dots, a_{n-1}), B = (b_0, b_1, \dots, b_{n-1})$; similarly the product can also be written
-as: $C = A*B = (c_0, c_1, \dots, c_{n-1}).$
+This example shows that convenience to use NB when computing $2^k$ power of an element.
+Multiplication is a bit complicated than squaring; but when it is decomposed as bit-wise
+operations, the property in lemma \ref{lem:squareNB} can be well utilized.
 
-Then the highest digit of product can be represented by a function: $c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, 
-\dots, b_{n-1})$. Square both side: $C^2 = A^2*B^2$, i.e. the second highest digit $c_{n-2} = f(a_{n-1}, a_0, a_1, 
-\dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2}).$ By this method it is easy to get all digits of product $C$.
+\begin{Example}[Bit-wise NB multiplication]
+Assume there are 2 binary vectors representing 2 operands in NB over $\Fkk$: 
+$A = (a_0, a_1, \dots, a_{k-1}), B = (b_0, b_1, \dots, b_{k-1})$. Note that in this example, 
+by default we use normal basis representation so subscript ``NB" is skipped. Their product can also be written
+as: $$C = A\times B = (c_0, c_1, \dots, c_{k-1})$$
+
+Assume the most significant bit (MSB) of the product can be represented by a function $f_{mult}$: 
+\begin{equation}
+\label{eqn:multiMSB}
+c_{n-1} = f_{mult}(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})
+\end{equation}
+Before discussing the details of the function $f_{mult}$, we can take 
+a square on both side of equation \ref{eqn:multiMSB}, i.e. $C^2 = A^2\times B^2$.
+Obviously, using the property in lemma \ref{lem:squareNB}, the original second most significant bit 
+becomes the new MSB because of right-cyclic shifting. 
+Concretely, 
+$$(c_{k-1},c_0,c_1,\dots,c_{k-2}) = (a_{k-1},a_0,a_1,\dots,a_{k-2})\times(b_{k-1},b_0,b_1,\dots,b_{k-2})$$
+Note $A^2, B^2$ and $C^2$ still belong to $\Fkk$, thus as a universal function implementing MSB multiplication
+over $\Fkk$, $f_{mult}$ still keeps the same. As a result, the new MSB can be written as 
+$$c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
+\dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})$$ 
+Similarly, if we take a square again on the new equation, we can get $c_{k-3}$.
+Successively we can derive all bits of product $C$ using the same function $f_{mult}$, and the only
+adjustment we need to make is to right-cyclic shift 2 operands by 1 bit each time.
 \end{Example}
+
+From above example, it is proved that a universal structure that implements $f_{mult}$ can be reused
+for $k$ times in NB multiplication over $\Fkk$. Comparing to StdB, which requires distinct design 
+for every bit of multiplication, NB is less costly if we can prove $f_{mult}$ will not result in 
+a structure with $O(k^2)$ complexity. So our next mission is to explore the details of $f_{mult}$
+to prove it will be a relatively simple design with complexity lower than $O(k^2)$.
+
+If we want to make the complexity of $f_{mult}$ lower than $O(k^2)$, then the best choice is to try out 
+linear functions. As we know, matrix multiplication can simulate all possible combinations of 
+linear functions (which is also the reason it is used as basic model to simulate the behavior of a neuron
+in neural network machine learning algorithms). Imagine $A$ is a $k$-bit row vector and $B$ is a $k$-bit 
+column vector, then the single bit product can be written as the product of matrix multiplication
+$$c_{i} = A\times C\times B$$
+where $C$ is a $k\times k$ square matrix.
 

### 2016-10-12 dissertation/normal.tex
dissertation: part normal update 5:NB intro completed

@@
 a sort of suitable circuit benchmarks which is easy to compute its
-Gr\"bner basis (GB). From the work of Lv et al. \cite{lv_dissertation},
+Gr\"obner basis (GB). From the work of Lv et al. \cite{lv_dissertation},
 we learn that arithmetic circuits in Galois field (GF) is
@@
 \section{Normal Basis Multiplier over Galois Field}
+\subsection{Normal Basis}
 Given a Galois field (GF) $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
@@
 
+\subsection{Multiplication using Normal Basis}
 Besides addition/subtraction, multiplication is also very common in arithmetic circuit design.
@@
 over $\Fkk$, $f_{mult}$ still keeps the same. As a result, the new MSB can be written as 
-$$c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
-\dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})$$ 
+\begin{equation}
+\label{eqn:shiftMSB}
+c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
+\dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})
+\end{equation}
 Similarly, if we take a square again on the new equation, we can get $c_{k-3}$.
@@
 column vector, then the single bit product can be written as the product of matrix multiplication
-$$c_{i} = A\times C\times B$$
+$$c_{l} = A\times C\times B$$
 where $C$ is a $k\times k$ square matrix.
 
-\begin{Example}
-\emph{$\lambda$-Matrix:} We can employ a binary $n\times n$ matrix $M$ to describe the "function"
- mentioned above: $c_{n-1} = f(A, B) = A \cdot M \cdot B^T$, $B^T$ denotes vector transposition. 
-More specifically, we denote the matrix by \emph{$k$-th $\lambda$-Matrix}: $c_k = A \cdot M^{(k)} \cdot B^T$.
-Then $c_{k-1} = A \cdot M^{(k-1)} \cdot B^T = rotate(A) \cdot M^{(k)} \cdot rotate(B)^T$, which means
-by right and down shifting $M^{(k-1)}$ we can get $M^{(k)}$.\\
-In $F_{2^3}$ constructed by $\alpha^3 + \alpha + 1$, let $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
-is a normal basis. $0$-th $\lambda$-Matrix
+\begin{Definition}[$\lambda$-Matrix]
+A binary $k\times k$ matrix $M$ is used to describe the bit-wise normal basis multiplication function $f_{mult}$ where
 \begin{equation}
+\label{eqn:def_lambda}
+c_{l} = f_{mult}(A, B) = A \times M \times B^T
+\end{equation}
+$B^T$ denotes vector transposition. Matrix $M$ is called $\lambda$-Matrix of
+$k$-bit NB multiplication over $\Fkk$.
+\end{Definition}
+
+When taking different bits $l$ of the product in equation \ref{eqn:def_lambda}, 
+we obtain a series of conjugate matrices of $M$. 
+Which means instead of shifting operands $A$ and $B$, we can also shift the 
+matrix.
+
+More specifically, we denote the matrix by \emph{$l$-th $\lambda$-Matrix} as 
+$$c_l = A \times M^{(l)} \cdot B^T$$
+Meanwhile, the operator shifting rule in equation \ref{eqn:shiftMSB} still holds. Then we have relation 
+$$c_{l-1} = A \cdot M^{(l-1)} \cdot B^T = shift(A) \cdot M^{(l)} \cdot shift(B)^T$$ which means
+by right and down cyclically shifting $M^{(l-1)}$, we can get $M^{(l)}$.
+
+\begin{Example}[NB multiplication using $\lambda$-Matrix]
+Over GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $\alpha^3 + \alpha + 1$, let normal element $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
+forms a normal basis. Corresponding $0$-th $\lambda$-Matrix is
+\begin{equation*}
 M^{(0)} = \left(
@@
 \end{array} \right).
-\end{equation}
+\end{equation*}
 i.e.,
-\begin{equation}
+\begin{equation*}
 c_0 = (a_0\  a_1\  a_2)\left(
@@
 b_2
-\end{array} \right).
-\end{equation}
+\end{array} \right)
+\end{equation*}
+From $0$-th $\lambda$-Matrix we can directly write down all remaining $\lambda$-Matrices:
+\begin{equation*}
+M^{(1)} = \left(
+\begin{array} {lcr}
+1 & 0 & 1\\
+0 & 0 & 1\\
+1 & 1 & 0
+\end{array} \right)~~~~~~
+M^{(2)} = \left(
+\begin{array} {lcr}
+0 & 1 & 1\\
+1 & 1 & 0\\
+1 & 0 & 0
+\end{array} \right)
+\end{equation*}
 \end{Example}
 
-$\lambda$-Matrix is defined with cross-product terms from multiplication, which is 
+If we generalize the definition and explore the nature of $\lambda$-Matrix, it is defined as cross-product terms from multiplication, which is 
 \begin{equation}
-Product C = (\sum_{i=0}^{n-1}a_i\beta^{2^i})(\sum_{j=0}^{n-1}b_j\beta^{2^j}) = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}a_ib_j\beta^{2^i}\beta^{2^j}
+Product~vector~C = (\sum_{i=0}^{k-1}a_i\beta^{2^i})(\sum_{j=0}^{k-1}b_j\beta^{2^j}) = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}
 \end{equation}
 The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
-normal basis, i.e.
+NB, i.e.
 \begin{equation}
-\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1}\lambda_{ij}^{(k)}\beta^{2^k}, \ \ \lambda_{ij}^{(k)} \in F_2.
+\beta^{2^i}\beta^{2^j} = \sum_{l=0}^{k-1}\lambda_{ij}^{(l)}\beta^{2^l}, \ \ \lambda_{ij}^{(l)} \in \mathbb F_2.
 \end{equation}
-Substitution yields, result is an expression for k-th digit of product as showed in \textit{Example B.3}:
+Substitution yields, result is an expression for l-th digit of product as showed in equation \ref{eqn:multiMSB}:
 \begin{equation}
-c_k = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}\lambda_{ij}^{(k)}a_ib_j
+c_l = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(l)}a_ib_j
 \end{equation}
-$\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ in $k$-th $\lambda$-Matrix.
+$\lambda_{ij}^{(l)}$ is the entry with coordinate $(i,j)$ in $l$-th $\lambda$-Matrix.
+
+The $\lambda$-Matrix can be implemented with XOR and AND gates in circuit design.
+The very naive implementation requires $O(C_N)$ gates, where $C_N$ is the number of
+nonzero entries in $\lambda$-Matrix.
+There usually exists multiple NBs in $\Fkk, k>3$. If we employ a random NB, there is no mathematical
+guarantee that $C_N \sim o(k)$ (symbol $o$ denotes ``strictly lower than bound"). However,
+Mullin et al. \cite{mullinONB} % This citation is valid
+proves that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
+its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
+called optimal normal basis (ONB), details are introduced in appendix.
+
+In practice, large size NB multipliers are usually designed in $\Fkk$ when ONB exists
+to minimized the number of gates. So in the following part of this chapter and our experiments,
+we only focus on ONB multipliers instead of general NB multipliers.
 
 % After fixing this, add a whole-piece StdB vs NB example, list their cost as the conclusion
+\subsection{Comparison between Standard Basis and Normal Basis}
+At the end of this section, a detailed example is used to make a comparison between StdB multiplication
+and NB multiplication.
+\begin{Example}[Rijndael's finite field]
+Rijndael uses a characteristic 2 finite field with 256 elements, which can also be called the GF $\mathbb F_{2^8}$.
+Let us define the primitive element $\alpha$ using irreducible polynomial 
+$\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$. Coincidently, $\alpha$ is also a normal element,
+i.e. $\beta = \alpha$ can construct a NB $\{\alpha,\alpha^2,\alpha^4,\alpha^8,\alpha^{16},\alpha^{32},\alpha^{64},
+\alpha^{128}\}$.
+
+We pick a pair of elements from the Rijndael's field: $A=(0100~1011)_{StdB} = (4B)_{StdB},~B=(1100~1010)_{StdB} = (CA)_{StdB}$.
+First let us compute their product in StdB, the rule follows ordinary polynomial multiplication.
+
+\begin{align*}
+A\cdot B &= (\alpha^6+\alpha^3+\alpha+1)(\alpha^7+\alpha^6+\alpha^3+\alpha)\\
+&= (\alpha^{13}+\alpha^{10}+\alpha^8+\alpha^7)+(\alpha^{12}+\alpha^9+\alpha^7+\alpha^6)+
+	(\alpha^9+\alpha^6+\alpha^4+\alpha^3)\\
+	&~~~+(\alpha^7+\alpha^4+\alpha^2+\alpha) \\
+&= \alpha^{13}+\alpha^{12}+\alpha^{10}+\alpha^8+\alpha^7+\alpha^3+\alpha^2+\alpha
+\end{align*}
+Note that this polynomial is not the final form of the product because it needs to be
+reduced modulo irreducible polynomial $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$.
+This can be done using base-2 long division. Note the dividend and divisor are written in pseudo Boolean
+vectors, not real Boolean vectors in any kind of bases.
+
+% The folowing part is customized latex code mimicing binary long division
+% Note \divrule coordinates may not reflect the intuitive column # in tabular. Adjust by urself!!!
+
+\vspace{1cm}
+
+\newdimen\digitwidth
+\settowidth\digitwidth{0}
+\def~{\hspace{\digitwidth}}
+
+\def\divrule#1#2{%
+\noalign{\moveright#1\digitwidth%
+\vbox{\hrule width#2\digitwidth}}}
+111010111\,\begin{tabular}[b]{@{}r@{}}
+101001 \\ \hline
+\big)\begin{tabular}[t]{@{}l@{}}
+11010110001110 \\
+111010111 \\ \divrule{0}{14}
+~~111101101 \\
+~~111010111 \\ \divrule{2}{12}
+~~~~~111010110 \\
+~~~~~111010111 \\ \divrule{5}{9}
+~~~~~~~~~~~~~1
+\end{tabular}
+\end{tabular}
+\vspace{0.5cm}
+
+The final remainder is $1$, i.e. the product equals to 1 in StdB.
+
+On the other hand, operands $A$ and $B$ can be written in NB as
+$$A = (0010~1001)_{NB},~~B = (0100~0010)_{NB}$$
+The $\lambda$-Matrix for $\mathbb F_{2}[x] \pmod{x^8+x^7+x^6+x^4+x^2+x+1}$
+is
+(Computation of $\lambda$-Matrix refers to appendix)
+\begin{equation*}
+M^{(0)} = \left(\begin{array}{lccccccr}
+0 &0 &0 &0 &1 &0 &1 &1 \\
+0 &0 &1 &1 &1 &1 &0 &0 \\
+0 &1 &0 &0 &0 &0 &1 &0 \\
+0 &1 &0 &0 &1 &1 &0 &1 \\
+1 &1 &0 &1 &0 &1 &0 &0 \\
+0 &1 &0 &1 &1 &0 &0 &1 \\
+1 &0 &1 &0 &0 &0 &0 &0 \\
+1 &0 &0 &1 &0 &1 &0 &1
+\end{array}\right)
+\end{equation*}
+Taking matrix multiplication $c_0 = A\times M^{(0)}\times B^T$,
+the result is $c_0 = 1$. Then by cyclic shifting $A$ and $B$
+(or shifting $M^{(0)}$, either is applicable), we can successively
+obtain other bits of product. The final answer is
+$$C = (0000~0001)_{NB}$$
+It is equivalent to the result in StdB.
+\end{Example}
+
+From the intuition of humans, StdB multiplication is straightforward and easier to understand
+while NB is difficult to comprehend. However, if we implement both multiplications to 
+hardware multipliers, it will be clear which side a circuit designer prefers.
+
+% May consider providing details?
+
+Mastrovito multiplier and Montgomery multiplier are 2 common designs
+of GF multipliers using StdB. As a naive implementation of GF multiplication,
+Mastrovito multiplier uses most number of gates:
+$k^2$ AND gates plus $k^2-\Delta$ XOR gates \cite{Mastrovito}. Montgomery multiplier 
+applies lazy reduction techniques and results in a better latency performance, while the number of gates are about
+the same with Mastrovito multiplier:
+$k^2$ AND gates plus $k^2-k/2$ XOR gates \cite{Montgomery}. 
+Concretely, typical design of Mastrovito multiplier consists of 218 logic gates, while 
+Montgomery multiplier needs 198 gates. However, the NB multiplier reuses the $\lambda$-Matrix 
+logic, so this component will only need to be implemented for once. 
+Consider the definition of matrix multiplication, it needs $C_N$ AND gates to apply 
+bit-wise multiplication and $C_N-1$ XOR gates to sum them up. The number of nonzero entries
+in the $\lambda$-Matrix can be counted: $C_N = 27$.
+As a result, the most naive NB multiplier design (or Massey-Omura multiplier \cite{MasseyOmura})
+contains 53 gates in total, which is a great saving in area cost comparing to StdB multipliers.

### 2016-10-12 dissertation/scripts/multitable_to_lambdaMat.pl

@@ -1,0 +1,62 @@
+#!/usr/bin/perl -w
+
+use 5.014;
+use strict;
+use warnings;
+
+# Create by Xiaojun Sun @ OCT 12, 2016
+# Read-in multiplication table manually, transform to lambda-Matrix
+
+my $size = 8; # remember to change the size!!
+
+my @row1 = qw /0 1 0 0 0 0 0 0/; # Read-in M^0 for generating multi table
+my @row2 = qw /0 1 0 0 1 1 0 1/;
+my @row3 = qw /0 0 1 1 0 1 0 1/;
+my @row4 = qw /0 0 0 0 0 0 0 1/;
+my @row5 = qw /1 1 1 1 1 1 1 1/;
+my @row6 = qw /0 0 0 0 1 0 0 0/;
+my @row7 = qw /1 1 0 1 0 1 0 0/;
+my @row8 = qw /1 0 0 1 1 0 1 0/;
+my @mtable;
+push @mtable, [@row1];
+push @mtable, [@row2];
+push @mtable, [@row3];
+push @mtable, [@row4];
+push @mtable, [@row5];
+push @mtable, [@row6];
+push @mtable, [@row7];
+push @mtable, [@row8];
+
+
+my $tmp;
+my $i;
+my $j;
+my $c;
+my $k;
+
+my (@Mat, @tmprow);
+foreach(1..$size)
+{ 
+  push @tmprow, 0;
+}
+
+for($j = 0; $j < $size; $j++){ # init lambda-mat
+	push @Mat, [@tmprow];
+}
+
+#print $Mat[0][0];
+
+for($i = 0; $i < $size; $i++)
+{	
+    for($j = 0; $j < $size; $j++)
+    {
+	if($mtable[($j-$i)%$size][(0-$i)%$size] == 1) 
+		{$Mat[$i][$j]=1;}
+    }
+}
+for($i=0;$i<$size;$i++){
+for($j = 0; $j < $size; $j++){
+	print "$Mat[$i][$j] ";
+}
+print "\n";
+}

### 2016-10-14 dissertation/normal.tex
dissertation part normal: update 6

@@
 \section{Normal Basis Multiplier over Galois Field}
+From algebraic view, a field is a space, and field elements are dots in the space. Those elements can be 
+represented with unique coordinates, which requires the pre-definition of a basis vector. In this
+section, we discuss a special basis called normal basis, as well as the advantages adopting it in 
+GF operations esp. multiplication.
 \subsection{Normal Basis}
@@
 Consider the definition of matrix multiplication, it needs $C_N$ AND gates to apply 
-bit-wise multiplication and $C_N-1$ XOR gates to sum them up. The number of nonzero entries
+bit-wise multiplication and $C_N-1$ XOR gates to sum the intermediate products up. The number of nonzero entries
 in the $\lambda$-Matrix can be counted: $C_N = 27$.
@@
 contains 53 gates in total, which is a great saving in area cost comparing to StdB multipliers.
+
+\section{Design a Normal Basis Multiplier on Gate Level}
+The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
+we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
+very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
+test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
+routine of several kinds of modern NB multipliers.
+\subsection{Sequential Multiplier with Parallel Outputs}
+The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
+the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.
+
+% fig:MasseyOmura
+
+Figure \ref{fig:MasseyOmura} shows the basic architecture of a Massey-Omura multiplier. The operands
+$A$ and $B$ are 2 arrays of flip-flops which allow 1-bit right-cyclic shift every clock cycle.
+The logic gates in the boxes implements the matrix multiplication with $\lambda$-Matrix $M^{(0)}$, 
+while each AND gate corresponds to term $a_ib_j$ and each XOR gate corresponds to addition $a_ib_j+a_{i'}b_{j'}$. 
+The XOR layer has only 1 output, giving out 1 bit of product $C$ every clock cycle.
+
+The behavior of Massey-Omura multiplier can be concluded as: pre-load operands $A,B$ and reset $C$ to 0, after 
+executing for $k$ clock cycles, the data stored in flip-flop array $C$ is the product $A\times B$.
+We note that there is only one output giving 1 bit of the product each clock cycle, which matches the 
+definition of serial output to communication channel. Therefore this type of design is named as 
+sequential multiplier with serial output (SMSO).
+The SMSO architecture need $C_N$ AND gates and $C_N-1$ XOR gates, which equals to $2k-1$ AND gates and $2k-2$ XOR gates
+if it is designed using ONB. In fact, the number of gates can be reduced if the multiplication is 
+implemented using a conjugate of SMSO.
+
+The gate-level logic boxes are implementing following function:
+\begin{equation}
+\label{eqn:SMPOterms}
+c_l = row_1(A\times M^{(l)}) \times B + row_2(A\times M^{(l)}) \times B + \cdots + row_{k}(A\times M^{(l)}) \times B
+\end{equation}
+It can be decomposed into $k$ terms. If we only compute one term for each $c_l,~0\leq l\leq k-1$ in one 
+clock cycle, make $k$ outputs and add them up using shift register after $k$ clock cycles, 
+it will generate the same result with SMSO. This kind of architecture is named as 
+sequential multiplier with parallel outputs (SMPO). The basic SMPO, as a conjugate of Massey-Omura multiplier,
+is invented by Agnew et al. \cite{agnew1991implementation}.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/mySMPO.eps}
+\caption{5-bit Agnew's SMPO.
+Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, i.e.
+if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
+Index $w = 2i\pmod 5$}
+\label{fig:SMPO}}
+\end{figure}
+
+\begin{Example}[5-bit Agnew's SMPO]
+Given GF $\mathbb F_{2^5}$ and primitive element $\alpha$ defined by irreducible polynomial 
+$\alpha^5+\alpha^2+1=0$, normal element $\beta = \alpha^5$ constructs an ONB $\{\beta,\beta^2,\beta^4,\beta^8,\beta^{16}\}$.
+The $0$-th $\lambda$-Matrix for this ONB is
+\begin{equation*}
+M^{(0)} = \left(\begin{array}{lcccr}
+0 &1 &0 &0 &0 \\
+1 &0 &0 &1 &0 \\
+0 &0 &0 &1 &1 \\
+0 &1 &1 &0 &0 \\
+0 &0 &1 &0 &1
+\end{array}\right)
+\end{equation*}
+Then a typical design of 5-bit Agnew's SMPO is depicted in figure \ref{fig:SMPO}.
+
+The operands part of this circuit is the same with Massey-Omura multiplier. The differences are on 
+the matrix multiplication part, while it is implemented as separate logic blocks for 5 outputs,
+and the 5 blocks are connected in a shift register fashion. By analyzing the detailed function of 
+logic blocks, we can reveal the mechanism of Agnew's SMPO.
+
+Suppose we implement $M^{(0)}$ as the logic block in SMSO. In the first clock cycle, the output is
+\begin{equation}
+\label{eqn:5bitSMPO_c0}
+c_0 = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4
+\end{equation}
+Note it is written in the form of equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
+remaining bits of the product, which can be written in following general form polynomial:
+\begin{align}
+c_i =& b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
+&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
+(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
+\end{align}
+Note all index calculations are reduced modulo 5.
+
+Now let us observe the behavior of 5-bit Agnew's SMPO. Initially all DFFs are reset to 0. 
+In the first clock cycle,
+signal sent to the flip-flop in block $R_0$ denotes function:
+$$R_0^{(1)} = a_1b_0$$
+It equals to the first term of equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
+this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
+from operands (shifted by 1 bit), generating signal $a_u,a_v$ and $b_w$. Concretely,
+signal sent to flip-flop in block $R_1$ is:
+$$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
+which forms first 2 terms of equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
+on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
+$$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
+which equals to $c_0$ in equation \ref{eqn:5bitSMPO_c0}.
+After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
+the result of $c_0$ is computed after 5 clock cycles and gives on $r_0$.
+
+If we track $R_1\to R_2\to R_3 \to \cdots \to R_0$, we can obtain $c_1$ respectively.
+Thus we conclude that Agnew's SMPO functions the same with Massey-Omura multiplier.
+\end{Example}
+
+The design of Agnew's SMPO guarantees that there is only one AND gate in each $R_i$ block.
+For ONB, adopting Agnew's SMPO will reduce the number of AND gates from $2k-1$ to $k$.
+
+\subsection{Multiplier not based on $\lambda$-Matrix}

### 2016-10-16 dissertation/normal.tex
dissertation: part normal update 6: finish NB multiplier design

@@
 After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
-the result of $c_0$ is computed after 5 clock cycles and gives on $r_0$.
+the result of $c_0$ is computed after 5 clock cycles and given on $r_0$.
 
@@
 \subsection{Multiplier not based on $\lambda$-Matrix}
+Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
+which means that they will be identical if unrolled to full combinational circuits. 
+After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
+some of them jump out of the circle and are independent from $\lambda$-Matrix.
+One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
+\cite{RHmulti}, which is therefore called RH-SMPO.
+
+% fig:5bitRH
+
+Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
+in figure \ref{fig:SMPO}. A brief proof is as follows:
+
+\begin{Proof}
+First, we define an auxiliary function for $i$-th bit 
+\begin{equation}
+\label{eqn:aux}
+F_i(A,B) = a_ib_i\beta + \sum_{j=1}^v d_{i,j}\beta^{1+2^j}
+\end{equation}
+where $0\leq i\leq k-1, v = \lfloor k/2\rfloor, 1\leq j \leq v$.
+The $d$-layer index $d_{i,j}$ is defined as
+\begin{equation}
+\label{eqn:auxDC}
+d_{i,j} = c_{a,i} c_{b,i} = (a_i+a_{i+j})(b_i+b_{i+j}),~~1\leq j\leq v
+\end{equation}
+$i+j$ here is the result reduced modulo $k$. Note that there is a special boundary case when
+$k$ is an even number ($v = \frac{k}{2}$):
+$$d_{i,v} = (a_i+a_{i+v})b_i$$
+With the auxiliary function, we can utilize following theorem (proof refers to \cite{RHmulti}):
+\begin{Theorem}
+Consider three elements $A,B$ and $R$ such that $R=A\times B$ over $\Fkk$. Then,
+$$R=(((F_{k-1}^2+F_{k-2})^2+F_{k-3})^2+\cdots+F_1)^2+F_0$$
+\end{Theorem}
+This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
+$R_i$ flip-flops. Concretely, the multiplier behavior is an implementation of 
+following algorithm:
+
+\begin{algorithm}[H]
+\SetAlgoNoLine
+
+ \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
+ \KwOut{$R=A\times B$}
+%%%%%%%%%%%%%%%%%%%%
+  Initialize $A,B$ and aux var $X$ to 0\;
+  \For { ($i=0$;   $i < k$; ++i ) }
+  {
+	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Eq.\ref{eqn:aux}*/\;}
+	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
+  }
+  $R\gets X$
+\caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
+\end{algorithm}
+
+In this algorithm, we use a fixed auxiliary function $F_{k-1}$ inside the loop.
+This is because of equation
+$$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
+So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
+$F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
+\end{Proof}
+
+To better understand the mechanism of RH-SMPO, we will use this 5-bit 
+RH-SMPO as an example and introduce the details on how to design it.
+\begin{Example}[Designing a 5-bit RH-SMPO]
+From equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
+and XOR gates in $c$-layer according to equation \ref{eqn:auxDC}. Concretely, as algorithm \ref{alg:RHmulti}
+describes, we implement auxiliary function $F_{k-1}$ in the logic:
+$$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
+\begin{equation}
+\label{eqn:5bitRHaux}
+F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
+\end{equation}
+Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
+$$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
+$$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
+The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
+As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
+$F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
+turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
+$i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.
+
+In order to obtain information 
+indicating interconnections between $d$-layer and $e$-layer, we need to interpret $\beta^{1+2^j}$ 
+to NB representation.
+There is a concept called {\bf multiplication table} (M-table) which can assist this interpretation. It is defined as 
+a $k\times k$ matrix $T$ over $\mathbb F_2$:
+\begin{equation}
+\label{eqn:multitable}
+\begin{bmatrix}
+\beta^{1+2^0} \\ \beta^{1+2^1} \\ \beta^{1+2^2} \\ \vdots \\ \beta^{1+2^{k-1}}
+\end{bmatrix}
+= \beta
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+\end{bmatrix}
+=
+\begin{bmatrix}
+T_{0,0}      &   T_{0,1}        & \dots & T_{0,k-1}\\
+T_{1,0}    &   T_{1,1}           & \dots & T_{1,k-1}\\
+T_{2,0}    &   T_{2,1}           & \dots & T_{2,k-1}\\
+\vdots & \vdots              & \ddots     & \vdots \\
+T_{k-1,0}    &   T_{k-1,1}           & \dots & T_{k-1,k-1}
+\end{bmatrix}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+\end{bmatrix}
+= {\bf T}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+\end{bmatrix}
+\end{equation}
+
+It is a known fact that M-table $T$ can be converted from $\lambda$-Matrix $M$:
+$$M_{i,j}^{(0)} = T_{j-i,-i}$$
+with indices reduced modulo $k$. Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
+
+% fig:multitable
+
+Note that we only use row 1 and row 2 from the M-table since range $1\leq j \leq 2$.
+All nonzero entries in these 2 rows corresponds to the interconnections between $d$-layer and 
+$e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
+between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in equation \ref{eqn:multitable}:
+\begin{equation*}
+\beta\cdot\beta^2 = 
+\begin{bmatrix}
+1 & 0 & 0 & 1 & 0
+\end{bmatrix}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
+\end{bmatrix}
+= \beta + \beta^{2^3}
+\end{equation*}
+
+Similarly, from row 2 of M-table we derive that $d_2$ has fanouts $e_3,e_4$:
+\begin{equation*}
+\beta\cdot\beta^{2^2} = 
+\begin{bmatrix}
+0 & 0 & 0 & 1 & 1
+\end{bmatrix}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
+\end{bmatrix}
+= \beta^{2^3} + \beta^{2^4}
+\end{equation*}
+
+Let us look back at equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
+The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
+After gathering all interconnection information, we can translate it to gate-level circuit implementation:
+$$e_0 = d_0+d_1,~e_3=d_1+d_2,~e_4=d_2$$
+
+Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
+$R_{i-1}$ in last clock cycle, we can connect using relation 
+$$R_i = r_{i-1} + e_i$$
+In this example, according to the M-table in figure \ref{fig:multitable}, columns $e_1,e_2$
+have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.
+
+This finishes the full design procedure for a 5-bit RH-SMPO.
+\end{Example}
+
+The area cost of RH-SMPO is even smaller than Agnew's SMPO. XOR gates corresponds to all nonzero entries 
+in M-table, which is with the same number of nonzero entries in $\lambda$-Matrix ($C_N$). The number of AND gates 
+equals to $v$ plus 1 (for gate $d_0$). When using ONB ($C_N = 2k-1$), the total number of gates 
+is $2k+\lfloor \frac{k}{2}\rfloor$.
+
+\subsection{Customized Galois Field for Optimal Normal Basis Multiplier Design}

### 2016-10-17 dissertation/appendix.tex
dissertation part normal:rough finish

@@
 \fixchapterheading
-\section{Representations of Base Field Elements over Extension Fields}\label{append:Fpk}
-Consider a field $\F_q$ and a $k$-bit extension of this field, $\F_{q^k}$.
-This extension is created using a primitive, irreducible polynomial $P(x)$ of
-degree $k$ over $\F_q[x]$.
-Any element $A\in \F_{q^k}$ can be represented as
-\begin{equation}
-A=a_0+a_1\alpha+\dots+a_{k-1}\alpha^{k-1} \label{eqn:aToAFq}
-\end{equation}
-where $\{a_0,\dots,a_{k-1}\}\in\F_q$ and $\alpha$ is the primitive element of $\F_{q^k}$, i.e. 
-$P(\alpha)=0$. The goal is to derive the polynomial functions $a_i=\Func(A)$ for all
-$0\leq i < k$. Note that $\F_q$ could itself be an extension field of a base field $\F_p$ 
-where $q=p^l$ for some $l\geq 1$.
+\section{Normal Basis Theory}
+\label{append:NB}
+\subsection{Characterization of Normal Basis}
+\subsection{Construction of General Normal Basis}
+\subsection{Bases Conversion}
+% General \lambda matrix computation
 
-Each $a_i$ has the following property:
-\begin{equation}
-a_i=a_i^{q} \label{eqn:appendixaiq}
-\end{equation}
-
-Element $A$ has a similar property:
-\begin{equation}
-A=A^{q^k} \label{eqn:appendixAqk}
-\end{equation}
-
-Examine what happens when Equation \ref{eqn:aToAFq} is raised by the power $q$.
-The following lemma is applied.
-
-\begin{Lemma}\label{lemma:raiseToPq}\cite{galois_field:mceliece}
-{\it Let $\alpha_1, \dots, \alpha_t$ be any elements in $\F_{p^k}$. Then
-\begin{equation}
-(\alpha_1+\alpha_2+\cdots+\alpha_t)^{p^i}=\alpha_1^{p^i}+\alpha_2^{p^i}+\dots+\alpha_t^{p^i}
-\end{equation}
-for all integers $i \geq 0$. }
-\end{Lemma}
-
-\begin{Corollary}\label{corollary:raiseqk}
-Since $q=p^l$ for some $l \geq 1$, Lemma \ref{lemma:raiseToPq} is applicable to
-$\F_{q^k}$.
-Let $\alpha_1, \dots, \alpha_t$ be any elements in $\F_{q^k}$. Then
-\begin{equation}
-(\alpha_1+\alpha_2+\cdots+\alpha_t)^{q^i}=\alpha_1^{q^i}+\alpha_2^{q^i}+\dots+\alpha_t^{q^i}
-\end{equation}
-for all integers $i \geq 0$.
-\end{Corollary}
-Applying Corollary \ref{corollary:raiseqk} to Equation \label{eqn:aToFq} gives:
-\begin{equation}
-A^{q^i}=a_0^{q^i}+a_1^{q^i}\alpha^{q^i}+\dots+a_{k-1}^{q^i}\alpha^{(k-1)q^i} \label{eqn:appendixAqi}
-\end{equation}
-Then, applying 
-Equation \ref{eqn:appendixaiq} to this result gives:
-\begin{equation}
-A^{q^i}=a_0+a_1\alpha^{q^i}+\dots+a_{k-1}\alpha^{(k-1)q^i} \label{eqn:appendixAqiRewrite}
-\end{equation}
-
-In this way, $k$ unique equations are derived, $\{A,A^q,\dots,A^{q^{k-1}}\}$,
-along with $k$ unknowns, $\{a_0,\dots,a_{k-1}\}$.
-These equations can be represented in matrix form, $\mathbf{A = M a}$, where 
-$\mathbf{A}=\{A,A^q,\dots,A^{q^{k-1}}\}^T$, 
-$\mathbf{M}$ is
-a $k$ by $k$ matrix of coefficients $\in \F_{q^k}$, and $\mathbf{a}=\{a_0,\dots,a_{k-1}\}^T$:
-\begin{equation}
-\begin{bmatrix}
-1      &   \alpha           & \alpha^2           & \dots & \alpha^{k-1}\\
-1      &   \alpha^q         & \alpha^{2q}        & \dots & \alpha^{(k-1)q}\\
-\vdots & \vdots             & \vdots             & ~     & \vdots \\
-1      &   \alpha^{q^{k-1}} & \alpha^{2q^{k-1}} & \dots & \alpha^{(k-1)q^{k-1}}
-\end{bmatrix}
-\begin{bmatrix}
-a_0 \\ a_1 \\ \vdots \\ a_{k-1}
-\end{bmatrix}
-=
-\begin{bmatrix}
-A \\ A^q \\ \vdots \\ A^{q^{k-1}}
-\end{bmatrix}
-\end{equation}
-
-Treat $\mathbf{a}$ as a vector of unknowns, $\mathbf{M}$ and $\mathbf{A}$ as constants. 
-This system of equations can be solved using Gaussian 
-elimination. However, this system also has a special structure which can
-be exploited. $\mathbf{M}$ is a $k$ by $k$ Vandermonde matrix of the form 
-$V(\alpha,\alpha^q,\dots,\alpha^{q^{k-1}})$.
-Due to the property of Vandermonde matrices:
-\begin{equation}
-|\mathbf{M}| = \prod\limits_{0\leq i < j < k}(\alpha^{q^j} - \alpha^{q^i})
-\end{equation}
-Since the elements $\{\alpha,\alpha^q,\dots,\alpha^{q^{k-1}}\}$ are distinct:
-\begin{equation}
-|\mathbf{M}| \neq 0
-\end{equation}
-Thus, Cramer's rule can be applied to derive 
-an equation for every $a_i$:
-\begin{eqnarray}
-a_i = \frac{|\mathbf{M_i}|}{|\mathbf{M}|} \label{eqn:cramerformFq}%\\
-\end{eqnarray}
-where $\mathbf{M_i}$ is $\mathbf{M}$ with the column $\{\alpha^i,\alpha^{iq},\dots,\alpha^{iq^{k-1}}\}^T$ 
-replaced by $\mathbf{A}$.
-\begin{equation}
-\mathbf{M_i} = \begin{bmatrix}
-1      &   \alpha            & \dots  & \alpha^{i-1}          & A           & \alpha^{i+1}              & \dots & \alpha^{k-1} \\
-1      &   \alpha^q         & \dots  & \alpha^{(i-1)q}       & A^q         & \alpha^{(i+1)q}       & \dots & \alpha^{(k-1)q} \\
-\vdots & \vdots             & \cdot  & \vdots                & \vdots      & \vdots                & \cdot & \vdots \\
-1      &   \alpha^{q^{k-1}} & \dots  & \alpha^{(i-1)q^{k-1}} & A^{q^(k-1)} & \alpha^{(i+1)q^(k-1)} & \dots & \alpha^{(k-1)q^(k-1)}
-\end{bmatrix}
-\end{equation}
-
-Computing $|\mathbf{M_i}|$ by interpolating along the $A'$ column gives
-\begin{equation}
-|\mathbf{M_i}|=\sum\limits_{j=0}^{k-1}(-1)^{(i+j)}A^{q^j}
-|V_{i+1}(\alpha,\dots,\alpha^{q^(j-1)},\alpha^{q^(j+1)},\dots,\alpha^{q^(k-1)})|
-\end{equation}
-where $V_i(x_1,\dots,x_n)$ is the Vandermonde matrix $V(x_1,\dots,x_n)$ with the
-$i$-th column skipped and an extra column added to the end.
-\begin{equation}
-V_i(x_1,\dots,x_n) 
-=
-\begin{bmatrix}
-1 & x_1 & x_1^2  & \dots  & x_1^{i-1} & x_1^{i+1} & \dots & x_1^n \\
-1 & x_2 & x_2^2  & \dots  & x_2^{i-1} & x_2^{i+1} & \dots & x_2^n \\
-\vdots & \vdots  & \vdots & \cdot  & \vdots    & \vdots    & \cdot & \vdots    \\
-1 & x_n & x_n^2  & \dots  & x_n^{i-1} & x_n^{i+1} & \dots & x_n^n
-\end{bmatrix} 
-\end{equation}
-The computation of determinant of $V_i$ is known to the linear algebra community to
-be
-\begin{equation}
-|V_i(x_1,\dots,x_n)| = |V(x_1\dots,x_n)|\cdot S_{n-i}(x_1,\dots,x_n)
-\end{equation}
-where $S_i(x_1,\dots,x_n)$ is the $i$-th fundamental symmetric polynomial in $\{x_1,\dots,x_n\}$.
-Thus, $|\mathbf{M_i}|$ is computed as.
-\begin{eqnarray}
-|\mathbf{M_i}| &=& \sum\limits_{j=0}^{k-1}[(-1)^{j}A^{q^j} \nonumber \\
-& & \cdot |V(\alpha,\dots,\alpha^{q^(j-1)},\alpha^{q^(j+1)},\dots,\alpha^{q^(k-1)})| \nonumber \\
-& & \cdot S_{n-1-i}(\alpha,\dots,\alpha^{q^(j-1)},\alpha^{q^(j+1)},\dots,\alpha^{q^(k-1)})] \label{eqn:MireducedFq}
-\end{eqnarray}
-
-Lastly, this work has discovered the following proposition for the determinant $|\mathbf{M}|$.
-\begin{Proposition}
-The determinant $|\mathbf{M}|$ has the property
-\begin{equation}
-|\mathbf{M}|^q=(-1)^{k-1}|\mathbf{M}|
-\end{equation}
-\end{Proposition}
-\begin{Proof}
-\it Since $\mathbf{M}$ is a Vandermonde matrix of the form $V(\alpha,\alpha^q,\dots,\alpha^q{k-1})$, 
-then 
-\begin{equation}
-|\mathbf{M}| = \prod\limits_{0\leq i < j < k}(\alpha^{q^j}-\alpha^{q^i}) \label{proofeqn:detMFq}
-\end{equation}
-Computing $|\mathbf{M}|^q$ gives
-\begin{equation} \label{proof:MsquareFq}
-|\mathbf{M}|^q = [\prod\limits_{0\leq i < j < k}(\alpha^{q^j}-\alpha^{q^i})]^q
-\end{equation}
-Applying Corollary \ref{corollary:raiseqk} to Equation \ref{proof:MsquareFq} gives
-\begin{equation}
-|\mathbf{M}|^q = \prod\limits_{0\leq i < j < k}(\alpha^{q^{j+1}}-\alpha^{q^{i+1}})
-\end{equation}
-When $j=k-1$, the product term is in the form $(\alpha^{q^k}-\alpha^{q^{i+1}})$. 
-Since $\alpha^{q^k}=\alpha$ over $\F_{q^k}$, this term equivalent to  $-(\alpha^{q^{i+1}}-\alpha)$.
-This gives the property:
-\begin{equation}
-|\mathbf{M}|^q=(-1)^{k-1}|\mathbf{M}|
-\end{equation}
-\end{Proof}
-%Appendix \ref{append:Fpk}
+\section{Optimal Normal Basis}
+\label{append:ONB}
+\subsection{Construction of Optimal Normal Basis}
+\subsection{Optimal Normal Basis Multiplier Design}
+% Customized Galois Field IEEE1363-2000
+% Multi table <=> \lambda matrix

### 2016-10-17 dissertation/normal.tex
dissertation part normal:rough finish

@@
 \centering{
-\includegraphics[width=5in]{newfig/Moore.eps}
+\includegraphics[width=\textwidth]{newfig/Moore.eps}
 \caption{A typical Moore machine and its state transition graph}
@@
 \centering{
-\includegraphics[width=5in]{newfig/convention.eps}
+\includegraphics[width=\textwidth]{newfig/convention.eps}
 \caption{Conventional verification techniques based on bit-level unrolling and equivalence checking}
@@
 The addition/subtraction of GF elements in StdB follows the rules of polynomial addition/subtraction
-where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} in
+where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} (XOR) in
 Boolean algebra, element $A$ add/subtract by element $B$ in StdB is defined as
@@
 its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
-called optimal normal basis (ONB), details are introduced in appendix.
+called optimal normal basis (ONB), details are introduced in appendix \ref{append:ONB}.
 
@@
 is
-(Computation of $\lambda$-Matrix refers to appendix)
+(Computation of $\lambda$-Matrix refers to appendix \ref{append:NB})
 \begin{equation*}
@@
 $$M_{i,j}^{(0)} = T_{j-i,-i}$$
-with indices reduced modulo $k$. Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
+with indices reduced modulo $k$ (proof refers to appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
 
@@
 
-\subsection{Customized Galois Field for Optimal Normal Basis Multiplier Design}
+\section{Full-Blown Verification Procedure for Normal Basis Multiplier Functional Correctness Checking}
+Once a gate-level design of a NB multiplier is generated, it is ready to be verified using 
+similar approach appear in section \ref{sec:recha_approach}. The following part borrows contents
+from my own conference paper \cite{myDATE}.
+\subsection{Implicit Unrolling based on Abstraction with ATO}
+In preliminaries we talk about the abstraction basics. If we use elimination term order 
+$intermediate~variables~>~R~>~A,B$, the function of the combinational logic component 
+can be abstracted as 
+$$R = \Func(A,B)$$
+If we are verifying the functional correctness of a combinational NB multiplier (e.g. Mastrovito multiplier
+or Montgomery multiplier), the function given by abstraction will be $R=AB$.
+While in the sequential case, the function of combinational logic only fulfills a part of the multiplication,
+such as $F_{k-1}$ in equation \ref{eqn:aux}. Nevertheless, the abstraction still provides a word-level 
+representation which works definitely better on unrolling than bit-level expressions.
+In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
+explicit unrolling and avoid bit-blasting problem.
+
+For 2-input sequential NB multipliers, abstraction is utilized to implement following algorithm:
+\vspace{-0.1in}
+\IncMargin{1em}
+\begin{algorithm}[hbt]
+\SetAlgoNoLine
+\LinesNumbered
+ \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
+   state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}
+
+  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
+  $i = 0$\;
+  \Repeat{$i == k$}
+  {
+  	$i \gets i + 1$\;
+%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
+%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
+	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
+    )$ with ATO\;
+	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
+	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
+  }
+\Return{$from_k(R_{final})$}
+\caption {Abstraction via implicit unrolling for Sequential GF circuit
+  verification}
+\label{alg:modified}
+\end{algorithm}
+\DecMargin{1em}
+\vspace{-0.1in}
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=5in]{newfig/gf_seq_model.eps}
+\caption{A typical normal basis GF sequential circuit model. $A =
+  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
+  $A', B', R'$ denote next-state inputs.}
+\label{fig:sequential}}
+\end{figure}
+
+We follow  the sequential GF circuit model of
+figure \ref{fig:sequential}, with word-level variables $A, B, R$
+denoting {\it present states (PS)} and $A', B', R'$ denoting {\it next
+  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
+for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
+\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
+store the result, and $A, B\ (A', B')$ store input operands. E.g.,
+for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
+0$) are the initial values (operands) loaded into the registers,  and
+$R = \Func(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
+result after $k$-cycles. Our approach aims to find this polynomial
+representation for $R$.
+
+Each gate in the combinational logic is represented by a Boolean
+polynomial. To 
+this set of Boolean polynomials, we append the polynomials that define
+the word-level to bit-level relations for PS and NS variables ($A =
+\sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
+as ideal $J = \langle 
+f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
+  B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
+  of the circuit. The ideal of vanishing polynomials $J_0$ is also included, and
+then the implicit FSM unrolling problem is setup for abstraction.
+
+The configurations of the flip-flops are the states of the
+machine. Since the set of states is a finite set of points, we
+can consider it as the variety of an ideal related to the circuit.
+Moreover, since we are interested in
+the function encoded by the state variables (over $k$-time
+frames), we can project this variety on the word-level state
+variables, starting from the initial state $A_{init}, B_{init}$.
+Projection of varieties (geometry) corresponds to elimination ideals
+(algebra), and can be analyzed via \Grobner bases. Therefore, we
+employ a \Grobner basis computation with ATO: we use a {\it lex term
+  order} with {\it bit-level variables} 
+$>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
+allows to eliminate all the bit-level variables 
+%(corresponding to the combinational logic and the state variables),
+%so as to 
+and derives a representation only in terms of words. 
+Consequently, $k$-successive \Grobner basis computations implicitly
+unroll the machine, and provide word-level algebraic $k$-cycle
+abstraction for $R'$ as $R' = \F(A_{init}, B_{init})$.
+
+Algorithm
+\ref{alg:modified} describes our approach.  In the algorithm, $from_i$
+and $to_i$ are polynomial ideals whose varieties are the valuations of
+word-level variables $R, A, B$ and $R',A',B'$ in the $i$-th iteration;
+and the notation ``$\setminus$'' signifies that the $NS$ in iteration
+$(i)$ becomes the $PS$ in iteration $(i+1)$. Line 5 computes the Gr\"obner 
+basis with the abstraction term order.  Line 6 computes the elimination 
+ideal, eliminating the bit-level variables and representing the set of 
+reachable states up to iteration $i$ in terms of the elimination ideal. 
+These computations are analogous to those of image computations performed in FSM reachability. 
+\begin{Example}[Functional verification of 5-bit RH-SMPO]
+\label{ex:RHSMPO}
+Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO. The transition function for
+operands $A,B$ is doing cyclic shift, while transition function for $R$ has to be computed through Gr\"obner basis
+abstraction approach. Following ideal $J_{ckt}$ from line 5 in algorithm \ref{alg:modified} is the ideal for 
+all gates in combinational logic block and definition of word-level variables.
+\begin{align}
+J_{ckt} = & d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4,\nonumber\\
+& c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2, \nonumber\\
+& R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4,\nonumber\\
+& A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},\nonumber\\
+		 & B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18},\nonumber\\
+		 & R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18},\nonumber\\
+		 & R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18};\nonumber
+\end{align}
+In our implementation here, since we only focus on the output variable $R$, evaluations of intermediate input 
+operands $A, B$ are unnecessary. Polynomials about $A$ and $B$ can be removed from $J_{ckt}$, and $R$ is directly
+evaluated by initial operands $A_{init}$ and $B_{init}$, which are associated with present state bit-level inputs
+$a_0,a_1,\dots,a_4$ and $b_0,b_1,\dots,b_4$ by polynomials in $from^i$.
+
+According to line 5 of algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
+Gr\"obner basis with abstraction term order (copy details here). There is a polynomial in form of $R'+\Func(A_{init},B_{init})$,
+which should be included by $to^{i+1}$. $to^{i+1}$ also exclude next state variable $A'$ and $B'$, instead we 
+redefine $A_{init}$ and $B_{init}$ using next state bit-level variables $\{a_i', b_j'\}$. Next state Bit-level variables
+$a_i' = a_{i-1\pmod k}, b_j' = b_{j-1\pmod k}$ according to definition of cyclic shift.
+
+Line 7 in algorithm \ref{alg:modified} is implemented by replacing $R'$ with $R$, $\{a_i', b_j'\}$ with $\{a_i,b_j\}$.
+
+All intermediate results for each clock cycle are listed below:
+\begin{itemize}
+\item Clock 1: ${\bf from^0} = \{R, A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
+B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\}$, \\
+	${\bf to^1} = \{R'+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^2) A_{init}^8 B_{init}+(\alpha^4+\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^8+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha^2+\alpha) A_{init}^2 B_{init}^2+(\alpha^4+\alpha) A_{init}^2 B_{init}+(\alpha^4+\alpha^3+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init}^2+(\alpha^3+\alpha+1) A_{init} B_{init},
+	A_{init}+a_4'\alpha^5+a_0'\alpha^{10}+a_1'\alpha^{20}+a_2'\alpha^9+a_3'\alpha^{18},
+B_{init}+b_4'\alpha^5+b_0'\alpha^{10}+b_1'\alpha^{20}+b_2'\alpha^9+b_3'\alpha^{18}\}$
+\item Clock 2: ${\bf from^1} = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^2) A_{init}^8 B_{init}+(\alpha^4+\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^8+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha^2+\alpha) A_{init}^2 B_{init}^2+(\alpha^4+\alpha) A_{init}^2 B_{init}+(\alpha^4+\alpha^3+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init}^2+(\alpha^3+\alpha+1) A_{init} B_{init},
+	A_{init}+a_4\alpha^5+a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^9+a_3\alpha^{18},
+B_{init}+b_4\alpha^5+b_0\alpha^{10}+b_1\alpha^{20}+b_2\alpha^9+b_3\alpha^{18}\}$, \\
+${\bf to^2} = \{R'+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^8+(\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^2+(\alpha^3+1) A_{init}^8 B_{init}+(\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4) A_{init}^4 B_{init}^4+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^2) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}+(\alpha^3+1) A_{init} B_{init}^8+(\alpha) A_{init} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init},
+A_{init}+a_3'\alpha^5+a_4'\alpha^{10}+a_0'\alpha^{20}+a_1'\alpha^9+a_2'\alpha^{18},
+B_{init}+b_3'\alpha^5+b_4'\alpha^{10}+b_0'\alpha^{20}+b_1'\alpha^9+b_2'\alpha^{18}\}$
+\item Clock 3: ${\bf from^2} = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^8+(\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^2+(\alpha^3+1) A_{init}^8 B_{init}+(\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4) A_{init}^4 B_{init}^4+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^2) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}+(\alpha^3+1) A_{init} B_{init}^8+(\alpha) A_{init} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init},
+A_{init}+a_3\alpha^5+a_4\alpha^{10}+a_0\alpha^{20}+a_1\alpha^9+a_2\alpha^{18},
+B_{init}+b_3\alpha^5+b_4\alpha^{10}+b_0\alpha^{20}+b_1\alpha^9+b_2\alpha^{18}\}$, \\
+${\bf to^3} = \{R'+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}+(\alpha) A_{init}^8 B_{init}^{16}+(\alpha+1) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init},
+A_{init}+a_2'\alpha^5+a_3'\alpha^{10}+a_4'\alpha^{20}+a_0'\alpha^9+a_1'\alpha^{18},
+B_{init}+b_2'\alpha^5+b_3'\alpha^{10}+b_4'\alpha^{20}+b_0'\alpha^9+b_1'\alpha^{18}\}$
+\item Clock 4: ${\bf from^3} = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}+(\alpha) A_{init}^8 B_{init}^{16}+(\alpha+1) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init},
+A_{init}+a_2\alpha^5+a_3\alpha^{10}+a_4\alpha^{20}+a_0\alpha^9+a_1\alpha^{18},
+B_{init}+b_2\alpha^5+b_3\alpha^{10}+b_4\alpha^{20}+b_0\alpha^9+b_1\alpha^{18}\}$, \\
+${\bf to^4} = \{R'+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^8 B_{init}^{16}+(\alpha^3+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^8 B_{init}^4+(\alpha^2+\alpha) A_{init}^8 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}+(\alpha^4+\alpha) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^2+\alpha) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^2) A_{init}^2 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}+(\alpha^3+\alpha+1) A_{init} B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init} B_{init}^8+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^3+\alpha^2+1) A_{init} B_{init}^2+(\alpha) A_{init} B_{init},
+A_{init}+a_1'\alpha^5+a_2'\alpha^{10}+a_3'\alpha^{20}+a_4'\alpha^9+a_0'\alpha^{18},
+B_{init}+b_1'\alpha^5+b_2'\alpha^{10}+b_3'\alpha^{20}+b_4'\alpha^9+b_0'\alpha^{18}\}$
+\item Clock 5: ${\bf from^4} = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^8 B_{init}^{16}+(\alpha^3+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^8 B_{init}^4+(\alpha^2+\alpha) A_{init}^8 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}+(\alpha^4+\alpha) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^2+\alpha) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^2) A_{init}^2 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}+(\alpha^3+\alpha+1) A_{init} B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init} B_{init}^8+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^3+\alpha^2+1) A_{init} B_{init}^2+(\alpha) A_{init} B_{init},
+A_{init}+a_1\alpha^5+a_2\alpha^{10}+a_3\alpha^{20}+a_4\alpha^9+a_0\alpha^{18},
+B_{init}+b_1\alpha^5+b_2\alpha^{10}+b_3\alpha^{20}+b_4\alpha^9+b_0\alpha^{18}\}$, \\
+${\bf to^5} = \{{\bf R'+A_{init}B_{init}},
+A_{init}+a_0'\alpha^5+a_1'\alpha^{10}+a_2'\alpha^{20}+a_3'\alpha^9+a_4'\alpha^{18},
+B_{init}+b_0'\alpha^5+b_1'\alpha^{10}+b_2'\alpha^{20}+b_3'\alpha^9+b_4'\alpha^{18}\}$
+\end{itemize}
+The final result is
+$from^5(R_{final}) = R_{final}+A_{init}\cdot B_{init}$
+\end{Example}
+\subsection{Overcome Computational Complexity using RATO}
+Similar to our improvements in section \ref{sec:reachaRATO}, RATO \cite{TimDAC} is also available to 
+accelerate the GB computation here. More specifically, we obviate GB computation by turning it 
+into a single-step multivariate polynomial division: first, in ideal generated by
+gates information polynomials and word-level variable definition polynomials, find the unique pair of polynomial
+generators with leading monomials not relatively prime to each other; then, compute their specification polynomial
+using definition 
+$$Spoly(f_w,f_g) = \frac{LCM}{lt(f_w)}\cdot f_w - \frac{LCM}{lt(f_g)}\cdot f_g$$
+where $LCM$ is least common multiple of $lm(f_w)$ and $lm(f_g)$, and $lt$ denotes the leading term;
+last, reduce $Spoly$ with ideal $J_{ckt}+J_0$, it is possible that the remainder will be canonical polynomial
+function of the circuit. We will illustrate the whole improved procedure by applying RATO on 
+5-bit RH-SMPO in figure \ref{fig:5bitRH}.
+
+\begin{Example}
+\label{ex:newRATO}
+Variable order under RATO is:
+\begin{align}
+&\{r_0',r_1',r_2',r_3',r_4'\}>\{r_0,r_1,r_2,r_3,r_4\}\nonumber\\
+&>\{e_0,e_3,e_4\},\{d_0,d_1,d_2\},\{c_1,c_2,c_3,c_4\}\nonumber\\
+&>\{a_0,a_1,a_2,a_3,a_4,b_0,b_1,b_2,b_3,b_4\}>R'>R>\{A,B\}\nonumber
+\end{align}
+Search among all generators of $J_{ckt}$ from Ex.\ref{ex:RHSMPO} using RATO, we find a pair of polynomials whose leading 
+monomials are not relatively prime:
+$(f_w,f_g), f_w = r_0'+r_4+e_0, f_g =r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18} + R'$.
+We calculate $Spoly$ can reduce it by $J_{ckt}+J_0$:
+\begin{align}
+&Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
+&(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
+&+(\alpha^3+\alpha^2) a_1 b_1+(\alpha^4+\alpha^3+\alpha^2+\alpha) a_1 b_2+(\alpha^2+\alpha) a_1 b_3\nonumber\\
+&+(\alpha^2+1) a_1 b_4+(\alpha^4+1) a_1 B+(\alpha^4+\alpha) a_2 b_1+(\alpha^4+\alpha^3+\alpha) a_2 b_2\nonumber\\
+&+(\alpha^3+1) a_2 b_3+(\alpha^3+\alpha^2+1) a_2 b_4+(\alpha^3+\alpha^2) a_2 B+(\alpha^2+\alpha) a_3 b_1\nonumber\\
+&+(\alpha^3+1) a_3 b_2+(\alpha+1) a_3 b_3+(\alpha^4+\alpha^2+\alpha) a_3 b_4\nonumber\\
+&+(\alpha^4+\alpha^3+\alpha) a_3 B+(\alpha^3+1) a_4 b_1+a_4 b_2+(\alpha^4+\alpha^2+\alpha) a_4 b_3\nonumber\\
+&+(\alpha^4+\alpha^3+1) a_4 b_4+(\alpha^2+\alpha) a_4 B+(\alpha^4+1) b_1 A+(\alpha^3+\alpha^2) b_2 A\nonumber\\
+&+(\alpha^4+\alpha^3+\alpha) b_3 A+(\alpha^2+\alpha) b_4 A+(\alpha^4+\alpha^2+\alpha+1) R'+R+A B\nonumber
+\end{align}
+\end{Example}
+Above example indicates that RATO based abstraction on 5-bit RH-SMPO will result a remainder contains both
+bit-level variables and word-level variables, and the number of remaining variables is still large such that
+Gr\"obner basis computation will be inefficient.
+
+Since the remainder from \emph{Spoly} reduction contains some bit-level variables, our objective is to compute
+a polynomial contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
+this problem is to replace bit-level variable monomials by equivalent polynomials that only contain word-level
+variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination-fashion 
+approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.
+
+\begin{Example}
+\label{ex:BLVS}
+{\bf Objective}:\ Compute polynomial $a_i + \mathcal{G}_i(A)$ from $f_0 = a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A
+ = g_0 + A$.
+
+First, compute $f_0^2= a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2 = g_0^2 + A^2$;
+then $f_0^4, f_0^8, f_0^{16}$. By repeating squaring we get a system of equations:
+\begin{displaymath}
+  \begin{cases}
+  f_0  &=\ \  0\\
+  f_0^2 & =\ \  0\\
+  f_0^4  &=\ \  0\\
+  f_0^8  &=\ \  0\\
+  f_0^{16} & =\ \ 0
+  \end{cases}
+\Longleftrightarrow\ \
+\begin{cases}
+  g_0 & =\ \  A\\
+  g_0^2 & =\ \  A^2\\
+  g_0^4 & =\ \  A^4\\
+  g_0^8 & =\ \  A^8\\
+  g_0^{16} & =\ \  A^{16}
+  \end{cases}
+\end{displaymath}
+Following is the coefficients matrix form of this system of equations:
+\begin{displaymath}
+\begin{pmatrix}
+\alpha^5 & \alpha^{10} & \alpha^{20} & \alpha^9 & \alpha^{18} \\
+\alpha^{10} & \alpha^{20} & \alpha^9 & \alpha^{18} & \alpha^5 \\
+\alpha^{20} & \alpha^9 & \alpha^{18} & \alpha^5 & \alpha^{10} \\
+\alpha^9 & \alpha^{18} & \alpha^5 & \alpha^{10} & \alpha^{20} \\
+\alpha^{18} & \alpha^5 & \alpha^{10} & \alpha^{20} & \alpha^9 
+\end{pmatrix}
+\begin{pmatrix}
+a_0\\
+a_1\\
+a_2\\
+a_3\\
+a_4
+\end{pmatrix}
+=
+\begin{pmatrix}
+A\\
+A^2\\
+A^4\\
+A^8\\
+A^{16}
+\end{pmatrix}
+\end{displaymath}
+
+Use Gaussian elimination on coefficients matrix, for example  
+\begin{align}
+Row\ 2 =& Row\ 1\times \alpha^5 + Row\ 2: \nonumber\\
+&a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
+=&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
+\end{align}
+Recursively eliminate $a_1$ from third row, $a_2$ from fourth row, etc. The final solution to this system of
+equations is
+%\begin{displaymath}
+%\{g_i\ | \ g_i: a_i + \mathcal{G}_i(A)\}
+%\end{displaymath}
+%\end{Example}
+\begin{displaymath}
+\left\{
+  \begin{array}{lcl}
+  a_0 & = & (\alpha+1)A^{16}+(\alpha^4+\alpha^3+\alpha)A^8+(\alpha^3+\alpha^2)A^4\\&&+(\alpha^4+1)A^2+(\alpha^2+1)A\\
+  a_1 & = & (\alpha^2+1)A^{16}+(\alpha+1)A^8+(\alpha^4+\alpha^3+\alpha)A^4\\&&+(\alpha^3+\alpha^2)A^2+(\alpha^4+1)A\\
+  a_2 & = & (\alpha^4+1)A^{16}+(\alpha^2+1)A^8+(\alpha+1)A^4\\&&+(\alpha^4+\alpha^3+\alpha)A^2+(\alpha^3+\alpha^2)A\\
+  a_3 & = & (\alpha^3+\alpha^2)A^{16}+(\alpha^4+1)A^8+(\alpha^2+1)A^4\\&&+(\alpha+1)A^2+(\alpha^4+\alpha^3+\alpha)A\\
+  a_4 & = & (\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4\\&&+(\alpha^2+1)A^2+(\alpha+1)A
+  \end{array} \right.
+\end{displaymath}
+\end{Example}
+Similarly we can compute equivalent polynomials $\mathcal{H}_i(R)$ for $r_i$, $\mathcal{T}_j(B)$ for $b_j$.
+Using those polynomial equations, it is sufficient to translate all bit-level inputs in the remainder polynomial
+because of following lemma:
+\begin{Lemma}
+\label{lem:bitlevelremainder}
+Remainder of S-poly reduction will only contain primary inputs (bit-level) and word-level output; furthermore, there will be one and
+only one term containing word-level output whose monomial is word-level output itself rather than higher order form.
+\end{Lemma}
+
+\begin{Proof}
+First proposition is easy to prove by contradiction: assume there exists an intermediate bit-level variable $v$ in the remainder,
+then this remainder must be divided further by a polynomial with leading term $v$. 
+Since the remainder cannot be divided by any other polynomials in $J_{ckt}$, the assumption does not hold.
+
+Second part, the candidate pair of polynomials only have one term of
+single word-level output variable (say it is $R$) and this term is the last term under RATO, which means there is only one term with
+$R$ in Spoly. Meanwhile in other polynomials from $J_{ckt}+J_0$ there is no such term containing $R$, so this term will be
+kept to remainder $r$, with exponent equals to 1.
+\end{Proof}
+
+By replacing all bit-level variables by corresponding word-level variable polynomials, we transform the remainder
+of $Spoly$ reduction to the form of $R'+R+\mathcal F'(A,B)$. Note $R$ is present state notion of output, which
+equals to initial value $R=0$ in first clock cycle, or value of $R'$ from last clock cycle. By substituting
+$R$ with its corresponding value ($0$ or a polynomial only about $A$ and $B$), we get the desired polynomial function
+$R'+\mathcal F(A,B)$.
+
+\subsection{Solving Linear System for Bit-to-Word Substitution}
+In example \ref{ex:BLVS} we use a Gaussian-elimination-fashion method to solve the system of polynomial
+equations. There is another formalized method to solve following system of equations
+\begin{align}
+\label{eqn:betamat}
+\begin{bmatrix}
+S \\
+S^2 \\
+S^{2^2} \\
+\vdots \\
+S^{2^{k-1}}
+\end{bmatrix}
+&=
+\begin{bmatrix}
+\beta & \beta^{2} & \beta^{2^2} & \cdots & \beta^{2^{k-1}}\\
+\beta^{2} & \beta^{2^2} & \beta^{2^3} & \cdots & \beta \\
+\beta^{2^2} & \beta^{2^3} & \beta^{2^4} & \cdots & \beta^2\\
+\vdots & \vdots & \vdots & \ddots & \vdots \\
+\beta^{2^{k-1}} & \beta & \beta^2 & \cdots & \beta^{2^{k-2}}
+\end{bmatrix}
+\begin{bmatrix}
+s_0\\
+s_1\\
+s_2\\
+\vdots\\
+s_{k-1}
+\end{bmatrix}
+\end{align}
+Let ${\bf s}$ be a vector of $k$ unknowns $s_0,\dots,s_{k-1}$,
+then equation \ref{eqn:betamat} can be solved by using Cramer's rule:
+\begin{equation}
+\label{eqn:Cramer}
+s_i = \frac{\Mi}{\M}, \ \ 0\leq i\leq k-1, \M \neq 0
+\end{equation}
+where ${\bf M_i}$ denotes a coefficient matrix replacing $i$-th column in ${\bf M}$ with
+vector ${\bf S} = [S~~S^2~~\cdots~~S^{2^{k-1}}]^T$.
+
+Notice that ${\bf M}$ is constructed by squaring a row and assigning it to next row,
+therefore its determinant belongs to a special sort of determinants:
+\begin{Definition}
+Let $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ be a set of $k$ elements of $\Fpk$. Then the determinant
+\begin{align}
+\det{M(\alpha_0,\dots,\alpha_{k-1})} &= 
+\begin{vmatrix}
+\alpha_0 & \alpha_1 & \cdots & \alpha_{k-1} \\
+\alpha_0^p & \alpha_1^p & \cdots & \alpha_{k-1}^p \\
+\vdots & \vdots & \ddots & \vdots \\
+\alpha_0^{p^{k-1}} & \alpha_1^{p^{k-1}} & \cdots & \alpha_{k-1}^{p^{k-1}}
+\end{vmatrix}
+\end{align}
+is called the {\bf Moore determinant} of set $\{\alpha_0,\dots,\alpha_{k-1}\}$.
+\end{Definition}
+Moore determinant can be written as an explicit expression
+\begin{equation}
+\label{eqn:Moore}
+\det{M(\alpha_0,\dots,\alpha_{k-1})} = \alpha_0 \prod_{i=1}^{k-1} \prod_{c_0,\dots,c_{i-1}\in \Fp} (\alpha_{i} - \sum_{j=0}^{i-1}c_j\alpha_j)
+\end{equation}
+We use an example to help understanding the notations in Eqn.\ref{eqn:Moore}:
+\begin{Example}
+Let $\{\alpha_0,\alpha_1,\alpha_2\}$ be a set of elements of $\mathbb F_{2^3}$. Then
+\begin{align}
+\det{M(\alpha_0,\alpha_1,\alpha_2)} &= 
+\begin{vmatrix}
+\alpha_0 & \alpha_1 & \alpha_2 \\
+\alpha_0^2 & \alpha_1^2 & \alpha_2^2 \\
+\alpha_0^4 & \alpha_1^4 & \alpha_2^4
+\end{vmatrix} \\ \nonumber
+=& \alpha_0 \prod_{i=1}^{2} \prod_{c_0,\dots,c_{i-1}\in \mathbb F_2} (\alpha_{i} - \displaystyle\sum_{j=0}^{i-1}c_j\alpha_j)
+\end{align}
+First, let $i=1$,  we obtain $c_0\in \mathbb F_2$. When $c_0=0$, the product term equals to $\alpha_1$; when $c_0=1$ it equals to $(\alpha_1-\alpha_0)$.
+Then let $i=2$, we obtain ${c_0,c_1}\in \mathbb F_2$, they can take value from $\{0,0\},\{0,1\},\{1,0\}$ and $\{1,1\}$.
+We add 4 more product terms $\alpha_2,(\alpha_2-\alpha_1),(\alpha_2-\alpha_0),(\alpha_2-\alpha_0-\alpha_1)$ respectively.
+
+Thus, the result is
+\begin{equation}
+\det{M(\alpha_0,\alpha_1,\alpha_2)} = \alpha_0\alpha_1(\alpha_1-\alpha_0)\alpha_2(\alpha_2-\alpha_1)(\alpha_2-\alpha_0)(\alpha_2-\alpha_0-\alpha_1)
+\end{equation}
+\end{Example}
+We discover through investigation that $\M$ has a special property when the set of elements forms a basis. The proof is given below:
+\begin{Lemma}
+\label{lemma:Moore}
+Let $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ be a normal basis of $\Fpk$ over $\Fp$. Then
+\begin{equation}
+\det{M(\alpha_0,\alpha_1,\dots,\alpha_{k-1})} = 1
+\end{equation}
+\end{Lemma}
+\begin{proof}
+According to the definition Eqn.\ref{eqn:Moore}, the Moore determinant consists of all possible linear combinations of 
+$\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ with coefficients over $\Fp$. If $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$
+is a (normal) basis, then all product terms are distinct and represents all elements in the field $\Fpk$. Since the product 
+of all elements of a field equals to 1, the Moore determinant $\M=1$.
+\end{proof}
+Applying Lemma \ref{lemma:Moore} to Eqn.\ref{eqn:Cramer} gives
+\begin{equation}
+s_i = \Mi,~~0\leq i\leq k-1
+\end{equation}
+where $\Mi$ can be easily computed using Laplace expansion method, with complexity $O(k!)$.
+
+However, we observe a fact that solution, if written as the inversion of matrix $M$, is a circulant matrix itself.
+We have following proposition:
+$$\dots$$
+Therefore, for a certain type of ONB, we can directly write down the inversion of corresponding matrix $M$
+as 
+$$\dots$$
+% TODO to be continued...
+
+\section{Software Implementation of Implicit Unrolling Approach}
+% TODO this can be divided into Singular part and C++ part, with detailed example on the trick
+Our experiment on different size of SMPO designs is performed with both \textsc{Singular} \cite{DGPS}
+symbolic algebra computation system and our customized toolset deployed using C++.
+The SMPO designs are given as gate-level netlists with registers, then translated to polynomials to compose
+elimination ideal for Gr\"obner basis calculation. The experiment is conducted on desktop with
+3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and running 64-bit Linux OS.
+\subsection{Architecture in Singular}
+The Singular tool can read in scripts written in its own format similar to ANSI-C. For SMPO experiment, the main 
+loop of our script file performs the same function as algorithm \ref{alg:modified} describes, while Gr\"obner basis
+computation in main loop can be divided into 4 different function parts:
+
+(i) Pre-process:
+
+This step is executed only once before main loop starts. The function of pre-process is to compute following system
+of equations for bit-level inputs $a_0 \sim a_{k-1}$:
+\begin{displaymath}
+\left\{
+  \begin{array}{ll}
+  a_0 & = f_0(A)\\
+  a_1 & = f_1(A)\\
+  \vdots & \ \\
+  a_{k-1} & = f_{k-1}(A)
+  \end{array} \right.
+\end{displaymath}
+The methodology has been discussed in section \ref{sec:blvs}. For 5-bit SMPO example, we start from word-level
+expression polynomial
+\begin{displaymath}
+A + a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}
+\end{displaymath}
+and the result is
+\begin{displaymath}
+\left\{
+  \begin{array}{lcl}
+  a_0 & = & (\alpha+1)A^{16}+(\alpha^4+\alpha^3+\alpha)A^8+(\alpha^3+\alpha^2)A^4\\&&+(\alpha^4+1)A^2+(\alpha^2+1)A\\
+  a_1 & = & (\alpha^2+1)A^{16}+(\alpha+1)A^8+(\alpha^4+\alpha^3+\alpha)A^4\\&&+(\alpha^3+\alpha^2)A^2+(\alpha^4+1)A\\
+  a_2 & = & (\alpha^4+1)A^{16}+(\alpha^2+1)A^8+(\alpha+1)A^4\\&&+(\alpha^4+\alpha^3+\alpha)A^2+(\alpha^3+\alpha^2)A\\
+  a_3 & = & (\alpha^3+\alpha^2)A^{16}+(\alpha^4+1)A^8+(\alpha^2+1)A^4\\&&+(\alpha+1)A^2+(\alpha^4+\alpha^3+\alpha)A\\
+  a_4 & = & (\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4\\&&+(\alpha^2+1)A^2+(\alpha+1)A
+  \end{array} \right.
+\end{displaymath}
+By replacing bit-level variable $a_i$ with $b_i, r_i$ or $R_i$, and word-level variable $A$ with $B, r, R$ respectively,
+we can directly get bit-word relation functions for another operand input, pseudo input and pseudo output.
+
+One limitation to Singular tool is the exponential cannot exceed $2^{63}$, so when doing experiments for SMPO larger than
+62 bits, we use a little trick (the feasibility of this trick can also be verified in following steps). Since the BLVS method
+only requires squaring of equations each time, the exponential of word $A$ can only be in the form $2^{i-1}$, i.e. 
+$A^{2^0},A^{2^1},\dots,A^{2^{k-1}}$. To minimize the exponential presenting in Singular tool, we rewrite $2^{i-1}$ to $i$, i.e.
+$(A^{2^0},A^{2^1},\dots,A^{2^{k-1}}) \to (A, A^2, \dots, A^k)$. In this way result is rewritten to be
+$$a_0 = (\alpha+1)A^5+(\alpha^4+\alpha^3+\alpha)A^4+(\alpha^3+\alpha^2)A^3+(\alpha^4+1)A^2+(\alpha^2+1)A$$
+Thus the exponential will not exceed the Singular data size limit.
+
+This step requires limited substitution operations in Singular, so although we use the naive Gaussian elimination
+method (whose time complexity is $O(k^3)$), the time cost is trivial comparing to following steps. For 33 bits experiment,
+pre-process execution time is 2.7 sec; while for 100 bits experiment time cost is 36 sec.
+
+(ii) Spoly reduction:
+
+First, Spoly is calculated based on RATO, then reduced with the ideal composed by circuit description polynomials ($J$).
+For already finished experiments, naive reduction (multi-division) is adopted, and this step takes largest portion of total 
+time consumption.
+
+For SMPO experiments, reduced Spoly has following generic form (all coefficients are omitted):
+$$redSpoly = \sum r_i + \sum a_ib_i + \sum a_iB + \sum b_iA + R + r$$
+Note there is no cross-term for bit-level or word-level variables from same side such as $a_ia_j, a_iA$, etc. Consider the necessary
+condition of our trick, this property of reduced Spoly guarantees the word level variable can only exist in the form $A^{2^{i-1}}$,
+after substituting bit-level variables with corresponding word-level variable.
+
+(iii) Substitute bit-level variables in reduced Spoly:
+
+Use the result from pre-process, get rid of $r_i, a_i$ and $b_i$ through substitution. This step yields following polynomial (consider the trick we used):
+$$R + \sum r^i + \sum A^iB^j$$
+all coefficients omitted.
+
+(iv) Substitute present state word-level variable $r$ with inputs $A$ and $B$:
+
+According to section \ref{sec:SMPOexperiment}, there is still a polynomial $r_{in}$ in the ideal we want to compute Gr\"obner basis.
+This polynomial has form $r+f'(A,B)$, which is last clock cycle's output ($R+f'(A,B)$) with only leading term replaced in step "$from^i\gets to^i$" in 
+algorithm \ref{alg:modified}. Basically this step has nothing different from last one, however, it must be taken good care of when using our
+trick. There is power of $r$, $r^m$ is originally $r^{2^{m-1}}$; so if $r+f'(A,B)$ contains terms $A^iB^j$, the correct result after doing power
+is $$(A^{2^{i-1}}B^{2^{j-1}})^{2^{m-1}} = A^{2^{((i+m-2)\bmod k)+1}}B^{2^{((j+m-2)\bmod k)+1}}$$
+So the correct exponential for $A$ and $B$ in $(A^iB^j)^m$ should be $((i+m-2)\bmod k)+1$ and $((j+m-2)\bmod k)+1$, respectively.
+
+Within one main loop, after finishing steps (ii) to (iv), the output should be intermediate multiplication result $R+f(A,B)$. After $k$ loops,
+the output is $R+A\cdot B$ when SMPO is bug-free.
+
+\subsection{Architecture in Customized C++ Toolset}
+% what kind of package, what kind of data structure, what kind of algorithm we use (refer to preliminaries part)
+
+\section{Experimental Results}
+\label{sec:result}
+% TODO revise repeated parts and references, add C++ results
+
+We have implemented our approach within the \textsc{Singular} symbolic
+algebra computation system [v. 3-1-6] \cite{DGPS}. Using our
+implementation, we have performed experiments to verify two SMPO
+architectures --- Agnew-SMPO \cite{agnew1991implementation} and the
+RH-SMPO \cite{RHmulti} --- over $\Fkk$, for various datapath/field
+sizes. Bugs are also introduced into the SMPO designs by modifying a
+few gates in the combinational logic block. Experiments using SAT-,
+BDD-, and AIG-based solvers are also conducted and results are
+compared against our approach.  Our experiments run on a desktop with
+3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and
+64-bit Linux.
+
+{\it Evaluation of SAT/ABC/BDD based methods:} To verify circuit $S$
+against the polynomial $\F$, we unroll the SMPO over $k$ time-frames,
+and construct a miter against a combinational implementation of $\F$.
+A (pre-verified) $\Fkk$ Mastrovito multiplier \cite{mastro:1989} is
+used as the {\it spec} model. This miter is checked for SAT using the
+\emph{Lingeling} \cite{biere2013lingeling} solver. We also experiment
+with the Combinational Equivalence Checking (CEC) engine of the ABC
+tool \cite{abc}, which uses AIG-based reductions to identify internal
+AIG equivalences within the miter to efficiently solve
+verification. The BDD-based VIS tool \cite{brayton1996vis} is also used
+for equivalence check. The run-times for verification of (unrolled)
+RH-SMPO against Mastrovito {\it   spec} are given in Table
+\ref{tbl:equiv} -- which shows that the techniques fail beyond 23 bit
+fields.
+
+\vspace{-0.1in}
+\begin{table}[htb]
+\centering
+\caption{\small Run-time for verification of bug-free RH-SMPO circuits
+  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
+\label{tbl:equiv}
+\begin{tabular}{|c||c|c|c|c|} 
+\hline
+& \multicolumn{4}{|c|}{Word size of the operands $k$-bits}  \\
+\hline
+Solver & 11 & 18 & 23 & 33 \\
+\hline
+\hline
+Lingeling & 593  & \emph{TO}  & \emph{TO}  & \emph{TO}\\
+\hline
+\hline
+ABC & 6.24 & \emph{TO} & \emph{TO} & \emph{TO}\\
+\hline
+\hline
+BDD & 0.1 & 11.7 & 1002.4 & \emph{TO}  \\
+\hline
+\end{tabular}
+\label{table:satbdd}  
+\end{table}
+
+CEC between unrolled RH-SMPO and Agnew-SMPO also suffers the same fate
+(results omitted). In fact, both SMPO designs are based on slightly
+different mathematical concepts and their computations in all
+clock-cycles, except for the $k^{th}$ one, are also different. These
+designs have no internal logical/structural equivalencies, and
+verification with SAT/BDDs/ABC is infeasible. Their dissimilarity is
+depicted in Table \ref{tbl:fraig}, where  $N_1$ depicts the number of
+AIG nodes in the miter prior to {\it fraig\_sweep}, and the nodes
+after {\it fraiging} are recorded as $N_2$; so $\frac{N_1-N_2}{N_1}$
+reflects the proportion of equivalent nodes in original miter, which
+emphasizes the (lack of) \emph{similarity} between two designs.
+
+\begin{table}[htb]
+\centering
+\caption{\small Similarity between RH-SMPO and Agnew's SMPO}
+\label{tbl:fraig}
+\begin{tabular}{|c||c|c|c|c|} 
+\hline
+Size $k$ & 11 & 18 & 23 & 33 \\
+\hline
+$N_1$ & 734  & 2011  & 3285  & 6723\\
+\hline
+$N_2$ & 529 & 1450 & 2347 & 4852\\
+\hline
+Similarity & $27.9\%$ & $27.9\%$ & $28.6\%$ & $27.8\%$  \\
+\hline
+\end{tabular}\label{table:similarity}  
+\end{table}
+
+{\it Evaluation of Our Approach:} Our algorithm inputs the circuit
+given in BLIF format, derives RATO, and constructs the polynomial
+ideal from the logic gates and the register/data-word description.  We
+perform one $Spoly$ reduction, followed by the bit-level to word-level 
+substitution, in each clock cycle. After $k$ iterations, the final
+result polynomial $R$ is compared against the spec polynomial. The
+run-times for verifying bug-free and buggy RH-SMPO and Agnew-SMPO are
+shown in Table \ref{tbl:exp1} and Table \ref{tbl:exp2}, respectively. 
+We can verify, as well as catch bugs in, up to 100-bit
+multipliers. Beyond 100-bit fields, our approach is infeasible --
+mostly due to the fact that the intermediate 
+abstraction polynomial $R$ is very dense and contains high-degree
+terms, which can be infeasible to compute. However, it should be noted
+that if we do not use the proposed bit-level to word-level
+substitution, and compute reduced \Grobner bases with RATO, then our
+approach does not scale beyond 33-bit datapaths.
+
+\begin{table}[htb]
+\centering
+\caption{\small Run-time (seconds) for verification of bug-free and
+  buggy RH-SMPO using our approach} 
+\label{tbl:exp1}  
+\begin{tabular}{|c||c|c|c|c|c|c|} 
+\hline
+Operand size $k$ & 33 & 51 & 65 & 81 & 89 & 99\\
+\hline
+\#variables & 4785 & 11424 & 18265 & 28512 & 34354 & 42372\\
+\hline
+\#polynomials & 3630 & 8721 & 13910 & 21789 & 26255 & 32373\\
+\hline
+\#terms & 13629 & 32793 & 52845 & 82539 & 99591 & 122958\\
+\hline
+\hline
+Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 & 67021\\
+\hline
+Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120 & 66929\\
+\hline
+\end{tabular}
+%\begin{minipage}{8cm}
+%\def\arraystretch{1.5}\tabcolsep 2pt
+%% \begin{tabular}{|c||c|c|c|c|c|} 
+%% \hline
+%% Operand size $k$ & 33 & 51 & 65 & 81 & 89 \\
+%% \hline
+%% \#variables & 4785 & 11424 & 18265 & 28512 & 34354\\
+%% \hline
+%% \#polynomials & 3630 & 8721 & 13910 & 21789 & 26255\\
+%% \hline
+%% \#terms & 13629 & 32793 & 52845 & 82539 & 99591 \\
+%% \hline
+%% %\hline
+%% Runtime(bug-free) & 112.6 & 1129 & 5243 & 20724 & 36096 \\
+%% \hline
+%% Runtime(buggy) & 112.7 & 1129 & 5256 & 20684 & 36120\\
+%% \hline
+%% \hline
+%% \end{tabular}
+%% %\end{minipage}
+\end{table}
+
+\begin{table}[hbt]
+\centering
+\caption{\small Run-time (seconds) for verification of
+  bug-free and buggy Agnew's SMPO our approach}
+\label{tbl:exp2}
+\begin{tabular}{|c||c|c|c|c|c|} 
+\hline
+Operand size $k$ & 36 & 66 & 82 & 89 & 100 \\
+\hline
+\#variables & 6588 & 21978 & 33866 & 39872 & 50300\\
+\hline
+\#polynomials & 2700 & 8910 & 13694 & 16109 & 20300\\
+\hline
+\#terms & 12996 & 43626 & 67322 & 79299 & 100100 \\
+\hline
+\hline
+Runtime(bug-free) & 113 & 3673 & 15117 & 28986 & 50692 \\
+\hline
+Runtime(buggy) & 118 & 4320 & 15226 & 31571 & 58861\\
+\hline
+\end{tabular}
+\end{table}
+
+\section{Conclusions and Further Work}\label{sec:concl}
+%TODO some of this should be moved to future work chapter
+
+This proposal has described a method to verify sequential Galois field
+multipliers over $\Fkk$ using computer algebra and algebraic geometry
+based approach. As sequential Galois field circuits perform the
+computations over $k$ clock-cycles, verification requires an efficient
+approach to unroll the computation, and represent it as a canonical
+word-level multi-variate polynomial. Using algebraic geometry, we show
+that the unrolling of the computation at word-level can be performed by
+\Grobner bases and elimination term orders. Subsequently, we show
+that the complex \Grobner basis computation can be eliminated by means
+of a bit-level to word-level substitution, which is implemented using
+the binomial expansion over Galois fields and Gaussian
+elimination. Our approach is able to verify up to 100-bit sequential
+circuits, whereas contemporary techniques fail beyond 23-bit
+datapaths.
+
+Our approach still has following limitations: first, it can only be applied
+on XOR-rich circuits, while most industrial designs are AND-OR gates dominant;
+second, it only uses naive bit-word abstraction based on functions of arithmetic
+circuits, which will be inefficient when the function does not have a straightforward
+expression (such as an implicit function); last but not least, Gr\"obner basis computation
+in our improved approach still requires a very long time. To overcome these limitations,
+further explorations are needed for my research.
+
+One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
+Analysis on experiment results shows major time consumption is on ``multi-division" part.
+A matrix-based technique named as "F-4 style reduction" \cite{F4reduce} can speed up the procedure
+dividing a low-degree polynomial with term-sparse polynomial ideal.

### 2016-10-19 dissertation/recha.tex
dissertation:part reacha:update 1

\chapter{Word-level Traversal of Finite State Machines using Algebraic Geometry}
\label{ch:reacha}
Reachability analysis is a basic component of sequential circuit verification, esp. for 
formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
tools, in order to improve various performance indicators such as latency, clock skew or power
density, some major refactoring and attachments are made on original designs. 
Those modifications may introduce malfunctions or glitches to the whole logic. 
In localized simulation or formal verification (esp. equivalence checking), the modifications may be 
denied since the malfunctions or glitches are considered as ``faults" in this circuit.
However, if the circuit behavior is carefully investigated, it may come to a verdict that 
those ``faults" will never be activated/excitated during a restricted execution starting 
from legal initial states and with legal inputs. Thus we will call those ``faults" as 
{\bf spurious faults}, since they will not affect the circuit's normal behavior.

Almost all practical sequential logic components can be modeled as finite state machines (FSMs). 
If we apply constrains upon the machine to make it start from designated initial states and 
take in specific legal inputs, a set of reachable states can be derived. 
As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
spurious faults by checking if they sit in the reachable states. From the spurious fault validation 
perspective, reachability analysis is a must when developing full set of sequential circuit verification
techniques.

There are quite a few methods to perform reachability checking on FSMs. One among them is 
state space traversal. Conventionally the algorithm is based on bit-level techniques such as
binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
which bring critical advantages. In this chapter the approach will be described and discussed in depth, 
with examples and experiments showing its feasibility when applied on general circuit benchmarks.

\section{Motivation}
The inspiration of this research mainly comes from a journal paper \cite{KallaPartialScan}. 
In that paper, the author proposed an traversal algorithm using concept ``implicit state enumeration".
Concretely, the algorithm is written as follows:

\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Sequential circuit, number of registers to scan}
 \KwOut{Scan registers listed in decreasing order of their non-controllability}

  $from^0 = reached = S^0$\;
  $i = 0$\;
  \While{TRUE}
  {
  	$i++$\;
	$to^i = $IMAGE$(\Delta,from^{i-1})$\;
	$new^i = to^i \cdot \overline{reached}$\;
	\For{each state variable $r_j$}
	{
		record\_if\_transitions\_present\_or\_missing$(r_j,new^i)$\;
		compute\_degree\_of\_unsettability$(r_j,new^i)$\;
	}
	\If{$new^i == 0$}
	{
		break\;
	}
	$reached = reached + new^i$\;
	$from^i = new^i$\;
  }\
  \tcc{\ \ \ \ \ \ \ \ \ \ \ \ \ \ FSM traversal completed}\
  \For{each state variable $r_j$}
  {
	\If{missing transition for $r_j$}
	{
		scan state variable $r_j$\;
	}
  }
  $illegal\_states = $bdd\_complement$(reached)$\;
  \For{each state variable $r_j$}
  {
	compute\_degree\_of\_unateness$(r_j,illegal\_states)$\;
	non-controllability$(r_j)=$degree\_of\_unsettability$(r_j)+$degree\_of\_unateness$(r_j)$\;
  }
  order state variables in terms of their non-controllabilities;\tcc{\ Sorting}\
  output the required scan registers\;
%\Return{$from_k(R_{final})$}
\caption {SIMPSON: Scan Register Selection using Implicit State Enumeration\cite{KallaPartialScan}}
\label{alg:SIMPSON}
\end{algorithm}
\DecMargin{1em}

### 2016-10-20 dissertation/recha.tex
dissertation:part reacha update 1.5

@@
 \section{Motivation}
+\subsection{Limitations of Full Scan Algorithms}
 The inspiration of this research mainly comes from a journal paper \cite{KallaPartialScan}. 
@@
 \DecMargin{1em}
+
+Note that the former part of this algorithm describes a breath-first-search (BFS) traversal in
+state space. The traversal algorithm is a simple variation of BFS algorithm in graph theory where 
+states are nodes and transitions are arcs. Each state is uniquely encoded by a combination
+of a set of register data, which is usually represented by a Boolean vector.
+
+Since a typical sequential circuit usually contains an functionally independent combinational logic
+component, the traversal algorithm analyzes the combinational logic and derive the transition 
+function for one-step reachability within current time-frame, and extend the result to complete 
+execution paths through unrolling. If each state encoding (i.e. exact values in the selected registers) is
+explicitly exposed and counted during the unrolling 
+procedure, this unrolling is called {\bf explicit unrolling}. In the SIMPSON algorithm, the states cannot be 
+directly read in the execution; instead, they are implicitly represented using a conjuction of several 
+Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.
+
+However, the SIMPSON algorithm proposed by the author is usually not practical. The conjuctions of 
+Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
+the size of BDD is easy to explode when the formulas become too long and too complicated. In the paper,
+the author make a compromise between accuracy and cost, and turn to a partial scan technique rather than 
+a full scan. In my research work, the aim is to explore a word-level technique which can make a full scan 
+available.
+
+\subsection{Word-level Data Flow on Modern Datapaths}
+
+\subsection{Prerequisites of Word-level Techniques}
+
+\section{FSM Reachability using Algebraic Geometry}
+\label{sec:reach}
+% \subsection{An example illustrating our proposed approach}
+We use symbolic state reachability with algebraic
+geometry concepts. It is an abstraction based on word operand
+definition of datapaths in circuits, and it can be applied
+to arbitrary FSMs by bundling a set of bit-level variables together as
+one or several word-level variables.  The abstraction polynomial,
+encoding the reachable state space of the FSM, is obtained through
+computing a GB over $\Fkk$ of the polynomials of the circuit using an
+elimination term order based on Theorem \ref{thm:elim}.
+
+Conceptually, the state-space of a FSM is traversed in a breadth-first
+manner, as shown in Algorithm \ref{alg:BFS}. % \cite{KallaPartialScan}: 
+The algorithm operates on the FSM $\mathcal{M} = (\sum, O, S, S^0,
+\Delta, \Lambda)$ underlying a sequential circuit. In such cases, the
+transition function $\Delta$ and the initial states are represented
+and manipulated using Boolean representations such as BDDs or SAT
+solvers. The variables $from, reached, to, new$ represent
+characteristic functions of sets of states. Starting from the initial
+state $from^i = S^0$, the algorithm computes the states reachable in
+1-step from $from^i$ in each iteration. In line 4 of algorithm
+\ref{alg:BFS}, the {\it image computation} is 
+used to compute the reachable states in every execution step.
+
+The {\it transition function} $\Delta$ is given by Boolean equations
+  of the flip-flops of the circuit: $t_i = \Delta_i(s, x)$, where
+  $t_i$ is a next state variable, $s$ represents the present   state
+  variables and $x$ represents the input variables. The {\it
+    transition relation of the FSM} is then represented as:  
+\begin{equation} 
+T(s, x, t) =   \prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)
+\end{equation}
+where $n$ is the number of flip flops, and $\overline{\oplus}$ is XNOR
+operation. Let $from$ denote the set of initial states, then the
+image of the initial states, under the transition function $\Delta$ is
+finally computed as:
+\begin{align}
+\label{img}
+to = \text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
+  \cdot from ] 
+%= \exists _s ~\exists _x ~\prod_{i=1}^{n} (t_i \overline{\oplus } \Delta_i)\cdot from
+\end{align}
+
+Here, $\exists x (f)$ represents the {\it existential quantification
+  of $f$ w.r.t. variable $x$}.
+
+\IncMargin{1em}
+\begin{algorithm}[hbt]
+\SetAlgoNoLine
+\Indm
+ \KwIn{Transition functions $\Delta$, initial state $S^0$}
+\Indp
+
+  $from^0 = reached = S^0$\;
+  \Repeat{$new^i == 0$}
+  {
+  	$i \gets i + 1$\;
+	$to^i \gets$Img$(\Delta, from^{i-1})$\;
+	$new^i \gets to^i \cap \overline{reached}$\;
+  	$reached \gets reached \cup new^i$\;
+	$from^i \gets new^i$\;
+  }
+\Return{$reached$}
+\caption {BFS Traversal for FSM Reachability}\label{alg:BFS}
+\end{algorithm}
+\DecMargin{1em}
+
+Let us describe the application of the algorithm on the  FSM circuit
+of Fig. \ref{fig:fsm}. {\it We will first describe its operation at the
+Boolean level, and then describe how this algorithm can be implemented
+using algebraic geometry at word level.}
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/new_stg.eps}
+\caption{The example FSM and the gate-level implementation. }
+\label{fig:fsm}}
+\end{figure}
+
+In Line 1 of the BFS algorithm, assume that the initial state
+is $S_3$ in Fig.\ref{fig:fsm}(b), which is encoded as 
+$S_3 = \{11\}$. Using Boolean variables $s_0, s_1$ for the present
+states, $from^0 = s_0\cdot s_1$ is represented as a Boolean formula.
+
+\begin{Example}
+For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
+functions of the machine as:
+
+\begin{align*}
+\Delta_1: & ~t_0 \overline{\oplus} ((\overline{x \vee s_0 \vee s_1}) \vee s_0 s_1)\\
+\Delta_2: & ~t_0 \overline{\oplus} (\overline{s_0}x \vee \overline{s_1}s_0)\\
+from:     & ~from^0 = s_0\cdot s_1
+\end{align*}
+
+When the formula of Eqn. (\ref{img}) is applied to compute 1-step
+reachability, $to = \exists _{s_0, s_1, x} (\Delta_1 \cdot \Delta_2
+\cdot from^0)$, we obtain $to = \overline{t_0}\cdot t_1$, which denotes
+the state $S_1 = \{01\}$ reached in 1-step from $S_3$.
+In the next iteration, the algorithm uses state $S_1 = \{01\}$ as the
+current (initial) state, and computes $S_2 = \{10\} = t_0\cdot
+\overline{t_1}$ as the next reachable state, and so on. 
+\end{Example}
+
+%% Let  $\Delta_i$ denote the transition function for $i^{th}$ bit of the
+%% output $T$
+%% (denoted by $t_i$), and it is described by a Boolean function. We can
+%% obtain the transition relation  for bit-vector $T$:
+%% $Tran(s_0,s_1,x,t_0,t_1) =
+%% \bigwedge_{i=1}^{2}(t_i\ \bar{\oplus}\ \Delta_i)$. Assume present
+%% states are represent by Boolean formulas $PS(s_0,s_1)$, then the image
+%% function is written as $\text{Img}(Tran,\ PS) =
+%% \exists_{s_0,s_1}\exists_{x}[Tran(s_0,s_1,x,t_0,t_1)\land
+%%   PS(s_0,s_1)]$, where $\exists_x f$ denotes the existential
+%% quantification of $f$ w.r.t. $x$.
+
+Our objective is to model the transition functions $\Delta$ as a
+polynomial ideal $J$, and to perform the image computations 
+%(Algorithm \ref{alg:BFS}, line 4) 
+using Gr\"obner bases over Galois fields. {\it
+This requires to perform quantifier elimination; which can be
+accomplished using the GB computation over $\Fkk$ using elimination
+ideals} \cite{gao:qe-gf-gb}. Finally, the set union, intersection and
+complement operations are also to be implemented in algebraic geometry.
+
+{\bf FSM Traversal at word-level over $\Fkk$:} 
+The state transition graph (STG) shown in Fig.\ref{fig:fsm}(a) uses a
+2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
+S_3\}$. We map these states to elements in $\mathbb{F}_{2^2}$, where
+$S_0 = 0, S_1 = 1, S_2 = \alpha, S_3 = \alpha+1$. Here, we take 
+$P(X) = X^2+X+1$ as the irreducible polynomial to construct
+$\mathbb{F}_4$, and $P(\alpha) = 0$ so that $\alpha^2 + \alpha + 1 =
+0$.
+
+{\it Initial state:} In Line 1 of Alg.\ref{alg:BFS}, the initial state is
+specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
+S - 1 - \alpha$. Notice that if we consider the ideal generated by the
+initial state polynomial, $I = \langle f\rangle$, then its variety
+$V(I) = 1+\alpha$ corresponds to the state encoding $S_3 = \{11\} =
+1+\alpha$, where a polynomial in word-level variable $S$ encodes the
+initial state. 
+% with only one generator $f$, its variety
+% $V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 1+\alpha\}$, which equals to $\{1+\alpha\}$, the only
+% valid value $S_3$ can take.
+
+%Using theorems from Section \ref{sec:elim}, we implement the image
+%function by a GB computation on elimination ideal. Ex.\ref{ex:motiv}
+%is an example for our implementation of image function, i.e. one-step
+%reachability.
+
+{\bf Set operations:} In Lines 5 and 6 of
+Alg. \ref{alg:BFS}, we need  \textbf{union}, 
+\textbf{intersection} and \textbf{complement} of varieties over
+$\mathbb{F}_{2^k}$, for which we again use algebraic geometry concepts.
+
+\begin{Definition}
+\label{def:sum}
+({\bf Sum/Product of Ideals} \cite{ideals:book}) If $I = \langle f_1,
+\dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are
+ideals in $R$, then the {\bf sum} of $I$ and $J$ is defined as $I + J
+= \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$. Similarly, the
+{\bf product} of $I$ and $J$ is $I \cdot J = \langle
+f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle$. 
+\end{Definition}
+
+%With concepts of ideal sums and products, we can obtain the
+%intersection and union of affine varieties as:
+\begin{Theorem}
+\label{thm:unionintersect}
+If $I$ and $J$ are ideals in $R$, then 
+${\bf V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot
+J) = {\bf V}(I) \bigcup {\bf V}(J)$. 
+\end{Theorem}
+
+In Line 5 of Alg. \ref{alg:BFS}, we need to compute the complement of a
+set of states. Assume that $J$ denotes a polynomial ideal whose
+variety $V(J)$ denotes a set of states. We require the computation of
+another polynomial ideal $J'$, such that $V(J')=\overline{V(J)}$. We
+show that this computation can be performed using the concept of {\bf
+  ideal quotient}:
+
+\begin{Definition}
+\label{def:quo}
+({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in a ring $R$, then $I:J$ is the set
+%  \begin{equation}
+  $\{f \in R \ |\ f\cdot g \in I, \forall g \in J\}$ %\nonumber
+%  \end{equation}
+and is called the {\bf ideal quotient} of $I$ by $J$.
+\end{Definition}
+
+%The following example shows a simple ideal quotient operation:
+\begin{Example}
+In $\Fq[x,y,z]$, ideal $I = \langle xz,yz\rangle$, ideal $J = \langle z\rangle$. Then
+\begin{align*}
+I:J &= \{f\in\Fq[x,y,z]~|~f\cdot z \in \langle xz,yz\rangle\} \\
+&= \{f\in\Fq[x,y,z]~|~f\cdot z = Axz+Byz\}\\
+&= \{f\in\Fq[x,y,z]~|~f= Ax+By\}\\
+&= \langle x,y\rangle
+\end{align*}
+\end{Example}
+
+% However, complement of variety cannot be easily dealt by simple arithmetic on polynomial generators.
+% For non-trivial cases we can only prove that 
+% \begin{Theorem}
+% Let $I, J$ be ideals in $\mathbb F[x_1,\dots,x_n]$, then 
+% $${\bf V}(I:J) \supset {\bf V}({\bf I}({\bf V}(I) - {\bf V}(J)))$$
+% \end{Theorem}
+% Fortunately for most hardware verification cases, the form of polynomial generators are restricted.
+% A proposition has been proved by \cite{jinpeng} that after adding {\bf vanishing polynomials ideal}
+% the new composed ideal implies following corollary:
+
+We can now obtain the complement of a variety through the following
+results which are stated and proved below:
+
+\begin{Lemma}
+\label{lem:gcd}
+Let $f,g\in \Fkk[x]$, then $\langle f:g\rangle = \left\langle\frac{f}{gcd(f,g)}\right\rangle$.
+\end{Lemma}
+
+\begin{proof}
+Let $d = gcd(f, g)$. So, $f = df_1 , g = dg_1$ with $gcd(f_1 , g_1 ) =
+1$. Note that $f_1 = \frac{f}{gcd(f,g)}$.
+
+Take $h \in \langle f : g\rangle$. According to the
+Def. \ref{def:quo}, $hg \in \langle f \rangle$, which means $hg = f
+\cdot r$ with $r \in \Fkk[x]$. Therefore, $hdg_1 = df_1 r$ and $hg_1 =
+f_1 r$. But considering $gcd(g_1 , f_1 ) = 1$ we have the fact that
+$f_1$ divides $h$. Hence $h \in \langle f_1\rangle$.
+
+Conversely, let $h \in \langle f_1 \rangle$. Then $h = s \cdot f_1$,
+where $s \in \Fkk[x]$. So, $hg = hdg_1 = sf_1 dg_1 = sg_1 f \in 
+\langle f \rangle$. Therefore, $h \in \langle f : g\rangle$.
+\end{proof}
+
+\begin{Theorem}
+\label{thm:quotient}
+Let $J$ be an ideal generated by a single univariate polynomial in
+variable $x$ over $\Fkk[x]$, and let the vanishing ideal $J_0 = \langle
+x^{2^k}-x\rangle$. Then  
+$${ V}(J_0:J) = { V}(J_0) - { V}(J),$$ where all the varieties are
+considered over the field $\Fkk$. 
+\end{Theorem}
+
+\begin{proof}
+Since $\Fkk[x]$ is a principal ideal domain, $ J = \langle g\rangle$
+for some polynomial $ g \in \Fkk[x]$. Let $d = gcd(g, x^{2^k} -
+x)$. So, $g = dg_1 , x^{2^k} - x = df_1$, with $gcd(f_1 , g_1 ) =
+1$. Then $J_0 : J = \langle f_1 \rangle$ by Lemma \ref{lem:gcd}.
+
+Let $x \in V(J_0 ) - V(J)$. From the definition of set
+complement, we get $x \in \Fkk$ while $g(x) \neq 0$.
+
+Since $x^{2^k} = x$, we see that either $d(x) = 0$ or $f_1 (x) =
+0$. Considering $g(x) \neq 0$, we can assert that $d(x) \neq 0$. In
+conclusion, $f_1 (x) = 0$ and $x \in V(f_1 )$.
+
+Now let $x \in V(f_1 )$, we get $f_1 (x) = 0$. So, $x^{2^k} - x = 0$
+gives $x \in V(J_0) = \Fkk$ which  contains all elements in the
+field.
+
+Now we make an assumption that $x \in V(g)$. Then $g(x) = 0 =
+d(x)g_1(x)$ which means either $d(x) = 0$ or $g_1 (x) = 0$.
+
+If $g_1 (x) = 0$, then since $f_1 (x) = 0$ we get that $f_1 , g_1$
+share a root. This contradicts the fact that $gcd(f_1 , g_1 ) = 1$.
+
+On the other hand, if $d(x) = 0$, then since $f_1 (x) = 0$ and
+$x^{2^k} - x = df_1$, we get that $x^{2^k} - x$ has a double root. 
+But this is impossible since the derivative of $x^{2^k} - x$ is $-1$.
+
+So, $x \notin V(g)$ and this concludes the proof.
+\end{proof}
+
+% Let $J_0 = \langle x_1^{2^k} - x_1, \dots, x_n^{2^k} - x_n \rangle$
+% denote the ideal of all vanishing polynomials in $\Fkk$. Then, we have
+% $V(J_0) = (\Fkk)^{n}$; i.e. the variety of vanishing ideal contains
+% all possible valuations of variables, so it constitutes the {\bf
+%   universal set}. Subsequently, based on Theorem \ref{thm:quotient},
+% the {\bf absolute complement} $V(J')$ of a variety $V(J)$ can be
+% computed as:
+
+Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$
+i.e. the variety of vanishing ideal contains
+ all possible valuations of variables, so it constitutes the {\bf
+   universal set}. Subsequently, based on Theorem \ref{thm:quotient},
+ the {\bf absolute complement} $V(J')$ of a variety $V(J)$ can be
+ computed as:
+
+\begin{Corollary} \label{cor:complement}
+Let $J\subseteq \Fkk[x]$ be an ideal, and $J_0=\langle x^{2^k}-x\rangle$. Let $J'$ be an ideal 
+computed as $J' = J_0:J$. Then $$V(J') = \overline{V(J)} = {V}(J_0:J)$$
+\end{Corollary}
+
+With Corollary \ref{cor:complement}, we are ready to demonstrate the
+concept of word-level FSM traversal over $\Fkk$ using algebraic
+geometry. The algorithm is given in Alg. \ref{alg:univa}. Note that in
+the algorithm, $from^i, to^i, new^i$ are {\it univariate polynomials in variables $S$
+or $T$} only, due to the fact that they are the result of a GB
+computation with an elimination term order, where the bit-level 
+variables are abstracted and quantified away.
+
+%% Assume $I = \langle f\rangle  = \langle T^2 +
+%% (1+\alpha)\cdot T+\alpha\rangle $,  we can add vanishing polynomial
+%% $T^4 - T$ to this ideal such that its variety  $V(I) = \{a\ |\ a \in
+%% \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$. For complement set
+%% of variety $\{1, \alpha\}$, the universal set is the variety of ideal
+%% of vanishing polynomial $V(\langle T^4-T\rangle ) =
+%% \{0,1,\alpha,1+\alpha\}$, 
+%% so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
+%% which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
+%% the result is $\{0,1+\alpha\}$.
+
+\IncMargin{1em}
+\begin{algorithm}[hbt]
+\SetAlgoNoLine
+\Indm
+ \KwIn{The circuit's characteristic polynomial ideal $J_{ckt}$,
+   initial state polynomial $\F(S)$, and LEX term order: bit-level
+   variables $x, s, t$ $>$ PS word $S$ $>$ NS word $T$}
+\Indp
+
+  $from^0 = reached = \F(S)$\;
+  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
+  {
+  	$i \gets i + 1$\;
+  	$G \gets$GB($\langle J_{ckt},J_v, from^{i-1}\rangle$)\;
+  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest}
+	$to^i \gets G\cap \Fkk[T]$\;
+	\tcc{There will be a univariate polynomial in $G$ denoting the
+          set of next states in word-level variable $T$}
+	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
+	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
+  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
+  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
+	$from^i \gets new^i(S\setminus T)$\;
+	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
+  }
+  \tcc{Loop until a fixpoint reached: newly reached state is empty}
+\Return{$\langle reached\rangle$}
+\caption {Algebraic Geometry based FSM Traversal}\label{alg:univa}
+\end{algorithm}
+\DecMargin{1em}
+
+\begin{Example}
+\label{ex:SMPO}
+We apply Algorithm \ref{alg:univa} to the example shown in
+Fig. \ref{fig:fsm} to execute the FSM traversal. Let the initial state
+$from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F _4$. Polynomially, it is
+written as $from^0 = S - 0$. In the first iteration, we compose
+an ideal $J$ with
+
+%\begin{minipage}[h]{0.4\textwidth}
+\begin{align*}
+&f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
+&f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
+&f_3: S - s_0 - s_1\alpha; ~~~f_4: T - t_0 - t_1\alpha
+\end{align*}
+$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$, and the vanishing
+polynomials: 
+%\end{minipage}
+%\begin{minipage}[h]{0.6\textwidth}
+
+\begin{align*}
+&f_5: x^2-x; ~~f_6: s_0^2-s_0, ~~f_7: s_1^2-s_1\\
+&f_8: t_0^2-t_0, ~~f_9: t_1^2-t_1; ~~f_{10}: S^4-S, ~~f_{11}:T^4-T
+\end{align*}
+with $J_v = \langle f_5,f_6,\dots,f_{11}\rangle$.
+%\end{minipage}
+
+%and $from^0$.
+
+Compute $G = GB(J)$ for $J = J_{ckt}+J_0+\langle from^0\rangle$,
+with an elimination term order 
+$$ \underbrace{\{x,s_0,s_1,t_0,t_1\}}_{\text{all bit-level variables}} 
+> \underbrace{S}_{\text{(PS~word)}} > \underbrace{T}_{\text{(NS~word)}}.$$
+
+The resulting GB $G$ contains a polynomial generator with only $T$ as
+the variable. In Line 5, assign it to the next state $$to^1 =
+T^2+(\alpha+1)T+\alpha.$$ Note that the roots or variety of
+$T^2+(\alpha+1)T+\alpha$ is $\{1, \alpha\}$, denoting the states
+$\{01,10\}$.
+
+Since the formerly reached state ``$reached = T$'', its complement is
+computed using Corollary \ref{cor:complement} 
+$$\langle T^4-T\rangle:\langle T\rangle
+= \langle T^3+1\rangle.$$
+
+$V(\langle T^3 + 1\rangle) = \{1, \alpha, \alpha+1\}$ denoting the states
+$\{01,10,11\}$. Then the newly reached state set in this iteration is 
+$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle
+T^2+(\alpha+1)T+\alpha \rangle$$ We add these states 
+to formerly reached states 
+\begin{align*}
+reach &= \langle T\rangle \cdot \langle T^2+(\alpha+1)T+\alpha \rangle \\
+&= \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle \\
+&= \langle T^3+(\alpha+1)T^2+\alpha T\rangle
+\end{align*}
+ i.e. states $\{00,01,10\}$. We update the present states
+for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha.$$
+
+In the second iteration, we compute the reduced GB with the same term
+order for ideal $J = J_{ckt}+J_v+\langle from^1\rangle$. 
+It includes a polynomial generator $$to^2 = T^2+\alpha T$$ denotes states
+$\{00,10\}$. The complement of $reached$ is $$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
+= \langle T + 1+\alpha\rangle$$ (i.e. states $\{11\}$). We compute the newly reached state 
+$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$$
+
+Since the GB contains the unit ideal, it means the newly reached state
+set is empty, thus a fixpoint has been reached. The algorithm
+terminates and returns $$reached = \langle T^3+(\alpha+1)T^2+\alpha
+T\rangle$$ which, as a Gr\"obner basis of the elimination ideal,
+canonically encodes the final reachable state set. 
+\end{Example}
+
+{\it Significance of using GB:} A reduced GB is a unique, minimal and
+{\it canonical} representation of the circuit's function. Starting
+from a certain initial state and using a reduced GB to represent the
+transition function, reachable states can be computed and represented
+canonically. Then it becomes possible to identify  when a fixpoint
+is reached (termination of the algorithm) by performing an equality
+check of polynomial ideals. Moreover, the GB computation is also used
+as a quantification procedure. As the GB is computed w.r.t. an
+elimination term order with ``bit-level variables'' $>$
+``present-state word'' $S$ $>$ ``next-state word'' $T$, the set of
+reachable states are encoded, canonically, using a univariate
+polynomial in $T$, quantifying away the rest of the variables.

### 2016-10-24 dissertation/abstract.tex
dissertation: milestone 1: first draft

@@
-%\begin{abstract}
 \begin{center}{\bf ABSTRACT}\end{center}
 
-Abstraction plays an important role in digital design,
-analysis, and verification, as it allows for the refinement of
-functions through different levels of conceptualization. This dissertation
-introduces a new method to compute a symbolic, canonical, word-level abstraction of the function implemented
-by a combinational logic circuit. This abstraction provides a
-representation of the function as a {\it polynomial} $Z = F (A)$
-over the Galois field $\Fkk$, expressed over the $k$-bit input to the circuit, $A$.
-This representation is easily utilized for formal verification (equivalence checking) of
-combinational circuits.
+Verification of sequential circuits remains to be a topic of all time. With increasing size of new
+integrated circuits, sequential circuit designers may face much more complicated problems on 
+logic errors, unexpected delays and failures on critical path. 
+This dissertation addresses the problem of sequential circuit verification at the word-level
+and is based on concepts derived from algebraic geometry.
 
-The approach to abstraction is based upon concepts from
-commutative algebra and algebraic geometry, notably the
-\Grobner basis theory. It is shown that the polynomial $F(A)$ can
-be derived {\it by computing a \Grobner basis of the polynomials}
-corresponding to the circuit, using a specific elimination term
-order based on the circuit’s topology. However, computing \Grobner bases
-using elimination term orders is infeasible for large circuits.
-To overcome these limitations, this work introduces {\it an efficient symbolic
-computation} to derive the word-level polynomial. The presented algorithms
-exploit i) the structure of the circuit, ii) the properties of \Grobner bases,
-iii) characteristics of Galois fields $\Fkk$, and iv) modern algorithms from symbolic computation.
+Analyzing the sequential circuits at word level is an efficient way
+of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
+state-space. We manage to model the verification properties and the gate-level sequential circuit 
+implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
+their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
+techniques from algebraic geometry can be used to reason about the state-space on high-level.
+In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
+polynomial ideals. Without actually solving the system, the states in state space can be investigated 
+by manipulating these ideals.
 
-While the concept is
-applicable to any arbitrary combinational logic circuit, it 
-is particularly powerful in verification and equivalence checking of 
-hierarchical, custom-designed and structurally dissimilar Galois field arithmetic circuits. 
-%These designs are prevalent in cryptography, error correction codes, and signal processing.
-In most applications, the field size and the data-path size $k$
-in the circuits is very large, up to $1024$ bits.
-%The U.S. National
-%Institute for Standards and Technology (\emph{NIST}), recommends Galois
-%fields corresponding to operand-sizes $k = 163, 233, 283, 409$ and
-%$571$ bits wide.
-
-The proposed abstraction procedure can exploit the hierarchy of the
-given Galois field arithmetic circuits.
-% by computing the abstraction
-%of each block in parallel before combining the representations together. 
-A custom abstraction %software
-tool is designed to efficiently implement the abstraction procedure.
-Preliminary experiments show that, using the proposed approach, our
-tool can abstract and verify Galois field arithmetic circuits
-{\it up to $1024$ bits in size}. %, which is the largest NIST standard. 
-Contemporary
-techniques fail to verify these types of circuits beyond $163$ bits
-and cannot abstract a canonical representation beyond $32$ bits.
-
-%To efficiently implement the
-%procedure, we design a custom abstraction software tool. 
-%Modern abstraction techniques of
-%these circuits fail beyond $32$-bits.
-
-%Custom arithmetic circuits designed over Galois fields $\Fkk$ are
-%prevalent in cryptography, error correction codes, signal processing,
-%etc. In most applications, the field size and the data-path size $k$
-%in the circuits can be very large. For example, the U.S. National
-%Institute for Standards and Technology (NIST), recommends Galois
-%fields corresponding to operand-sizes $k = 163, 233, 283, 409$ and
-%$571$ bits wide. The complexity of such large architectures 
-%necessitates custom and hierarchical design. Custom design can raise
-%the potential for bugs in the implementation. As arithmetic bugs can
-%compromise the security of crypto-systems, formal verification of
-%Galois field circuits is an imperative.
-
-%This thesis presents an approach for formal verification and
-%equivalence checking of large Galois field arithmetic circuits based
-%on word-level canonical polynomial abstraction from gate-level
-%circuits. The word-level abstraction is based on concepts from
-%commutative algebra and algebraic geometry, and relies on the
-%following mathematical insights.
-
-%A combinational circuit with $k$-inputs and $k$-outputs implements
-%Boolean functions $f: \B^k \rightarrow \B^k$, where $\B = \{0, 1\}$.
-%Such functions can also be construed as a mapping $f: \Fkk
-%\rightarrow \Fkk$,  where $\Fkk$ denotes the Galois field of $2^k$
-%elements. Every function over $\Fkk$ is a polynomial function ---
-%i.e. there exists a unique, minimal, canonical polynomial
-%$\Func$ that describes $f$.  This thesis derives the polynomial
-%representation of the circuit as $Z = \Func(A)$ over $\Fkk$, where $A$
-%and $Z$ denote, respectively, the input and output bit-vectors of the
-%circuit. This can be achieved by computing a Gr\"obner basis of a set
-%of polynomials derived from the circuit, using a specific elimination
-%(abstraction) term order.
-
-%Computing Gr\"obner bases using elimination orders is, however,
-%practically infeasible for large circuits. To overcome this
-%limitation, we devise an approach to guide the Gr\"obner basis
-%computation to search for the abstraction polynomial. The approach is
-%implemented in a custom tool using modern $F_4$-style polynomial
-%reduction procedures. Using our approach, we are able to verify the
-%correctness of up to $571$-bit Galois field multiplier circuits,
-%whereas contemporary techniques prove to be infeasible beyond
-%$163$-bit circuits. 
-%\end{abstract}
+We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
+analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
+verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
+abstraction-refinement techniques for verification, we propose to investigate a word-level 
+analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
+not only derive new algorithms and techniques, but will also consider efficient CAD 
+implementations to target sequential equivalence and model checking problems. 
+% \end{abstract}

### 2016-10-24 dissertation/conclude.tex
dissertation: milestone 1: first draft

@@
 \chapter{Conclusions and Future Work}
-\label{ch:concl}
-
-A combinational circuit with $k$-inputs and $k$-outputs implements
-Boolean functions  $f: \B ^k \rightarrow \B ^k$, where $\B = \{0,
-1\}$. The function can also be construed as a mapping  $f: \Fkk
-\rightarrow \Fkk$, where $\Fkk$ denotes the Galois field of
-$2^k$ elements. A circuit with differing input and output sizes 
-computes $f: \B ^m \rightarrow \B ^n$, which can be represented
-as a function over Galois fields $f: \F_{2^m} \rightarrow \F_{2^n}$.
-This circuit can also be analyzed as the function 
-$f: \Fkk \rightarrow \Fkk$, where $\Fkk \supset \F_{2^m}$ 
-and $\Fkk \supset \F_{2^m}$.
-
-Every function $f$ over $\Fkk$ is a polynomial
-function --- i.e., there exists a unique,  minimal, canonical
-polynomial $\Func$ that describes $f$. This dissertation presented novel
-techniques based on computer-algebra and algebraic-geometry to
-derive the canonical (word-level) polynomial representation from the
-circuit as  $Z = \Func(A)$ over $\Fkk$, where $A$ and $Z$ denote, 
-respectively, the input and output bit-vectors of the circuit.
-
-A theory for word-level polynomial abstraction of bit-level circuits 
-over Galois fields is first developed.
-This theory is derived using techniques from computer-algebra, notably the theory of
-\Grobner basis. However, due to the computational complexity of computing a \Grobner
-basis, the solution is not scalable to large designs.
-In order to overcome these limitations, new symbolic computational
-algorithms are developed and refined. The algorithms employ techniques
-from the binomial expansion over $\Fkk$ and $\F_4$-style reduction and can exploit
-hierarchy in a given circuit. Finally, an efficient implementation of the algorithmic
-approach is presented.
-
-Experiments show that the proposed approach works exceptionally well
-for abstracting word-level Galois field arithmetic circuits. It has been
-shown that the approach can abstract and verify these types of circuits with up to 
-$1024$-bit datapaths. Other contemporary techniques
-cannot to verify these types for circuits beyond $163$-bits and fail to
-abstract them beyond $32$-bits.
-
-However, in cases of random logic, the abstraction approach
-can generate high-degree polynomials:
-\begin{equation}
-X^{q-1}+X^{q-2}\dots
-\end{equation}
-In these cases, the polynomials derived during the computation are dense, 
-and the computational complexity of manipulating such polynomials
-makes abstraction infeasible.
+\label{ch:conclude}
 
 \section{Future Work}
-
-Due to the modular nature of the proposed solution, there are many
-potential future research directions that can be explored.
-
-\subsection{Hardware Acceleration}
-The first reduction step, $f_Z\xrightarrow{F-f_{z_i},F_0}_+ r$
-is the most computationally complex part of the proposed abstraction approach. This
-reduction could be implemented using a hardware accelerator.
-Significant speed-ups have 
-been observed in GPU implementations of circuit simulation algorithms 
-\cite{PengLi:GPU}. 
-Furthermore, this work has shown cases where multiple, independent abstractions 
-need to be computed at the same time, such as when abstracting a word-level 
-representation of a composite field multiplier.
-
-These abstractions can be computed in parallel with one another, and this
-parallelism could then be exploited using a GPU. Furthermore, our 
-approach to compute the abstractions uses an $F4$-style reduction procedure, 
-which performs many complex computations over a large matrix. 
-Operations over matrices can be suitably implemented using a GPU.
-Lastly, the substitution by $a_i=\Func(A)$ is trivially parallelized. 
-Thus, further study is proposed to implement word-level abstraction 
-on a general purpose GPU.
-
-\subsection{Integration with EDA Tools} 
-The proposed canonical word-level abstraction approach is a full, 
-self-contained solution.
-It can thus be integrated into other EDA tools. There are direct
-applications of word-level abstractions to design synthesis. For instance, 
-the approach can compute a functional decomposition of a logic or it could 
-be used in high-level RTL synthesis. Since the derived abstraction 
-is canonical, it can also be used in verification engines such as SMT solvers.
-The abstraction approach most efficiently handles AND/XOR logic, so it
-could be used to complement approaches in the mentioned tools that are
-efficient over AND/OR logic.
-
-\subsection{Polynomial Reductions using Data-Structures}
-The abstraction approach poorly handles chains of OR gates
-due to their representation as polynomials of Galois fields.
-Other polynomial-based tools \cite{polybori:2009} have shown that it 
-can be beneficial to represent polynomials internally as decision diagrams.
-Thus, it is worthwhile to explore whether it is possible to
-implement the algorithmic approach in a different data-structure that
-is better-suited for handling this type of logic.
-One candidate data-structure is the And-Invert-Graph, as this structure efficiently
-handles OR gates. The widely-used tool ABC\cite{abc} provides a very efficient, 
-flexible, and open-source implementation of the AIG data structure.
-
-Recall that a one-step reduction  of the polynomial $f$ by polynomial $g$, 
-$f\xrightarrow{g} r$, is computed as:
-\begin{equation}
-r = f-\frac{LT(f)}{LT(g)}\cdot g
-\end{equation}
-Over Boolean circuits, $\B \equiv \F_2$. Since the leading term of
-any monomial in $\B$ is $1$, and $-1\equiv +1$, then
-\begin{equation}
-f-(\frac{LT(f)}{LT(g)}\cdot g)\equiv f+(\frac{LM(f)}{LM(g)}\cdot g)
-\end{equation}
-which computes an XOR operation
-\begin{equation}
-f\oplus(\frac{LM(f)}{LM(g)}\cdot g)
-\end{equation}
-while the $\cdot$ operator acts as an AND operation $\wedge$.
-So one step of the reduction procedure can be computed as the following 
-AND/XOR operation:
-\begin{equation}
-r = f \oplus \frac{LM(f)}{LM(g)}\wedge g
-\end{equation}
-
-Thus, we propose an investigation into implementing the algorithms presented
-in this dissertation over AIGs.
-
-\subsection{Application to Sequential Circuit Verification}
-Sequential Galois field arithmetic circuits over $\Fkk$
-take k-bit inputs and produce a k-bit result after k-clock cycles 
-of operation. Formal verification of sequential arithmetic
-circuits with large datapath sizes is beyond the capabilities of
-contemporary verification techniques. To address this problem,
-we described a verification method in \cite{sun:date15} 
-which uses the presented abstraction approach to implicitly 
-unroll the sequential arithmetic circuit over multiple (k) clock-cycles.
-The resulting function computed by the state-registers of the circuit 
-is represented canonically as a multi-variate word-level polynomial over
-$\Fkk$. While directly applicable to sequential Galois field arithmetic circuits, 
-this work needs to be further generalized in order to make applicable to
-any sequential state machine.
-
-\subsection{Application to Formal Software Verification}
-Computer algebra techniques based on \Grobner basis theory have been used 
-in formal software verification \cite{manna:program}.
-In this work, a \Grobner basis computation is used to derive {\it loop invariants}.
-However, the derived invariants are not bit-precise, so not every invariant that is
-computed can be applied to the verification. As our approach maintains
-the input-output relationship in the abstraction, it could be applied to
-find bit-precise invariants.
-
-\subsection{Application to Integer Arithmetic Circuits}
-The abstraction approach derives a
-word-level representation of circuits over Galois fields, $\Fkk$. 
-In order to expand its usability, we conjecture whether it is possible to 
-apply concepts from this approach to {\bf abstract word-level representations 
-of circuits over integer rings}, $\Zkk$. 
-As any function over a Galois 
-field $\Fkk$ is a polynomial function, there exists a polynomial which 
-describes the word-level function of a given circuit over $\Fkk$. 
-%If the word-level 
-%input and output of the circuit is $A$ and $Z$, respectively, this 
-%polynomial exists in the form $Z=\Func(A)$.
-However, not every function over an integer ring $\Zkk$ is a polynomial 
-function. 
-Thus, a single polynomial which describes the function of a circuit over 
-$\Zkk$ is not guaranteed to exist. 
-%Instead, there could exist a set of
-%polynomials $\{f_1,\dots,f_n\}$ which describes this function.
-%These polynomials are in following form:
-%\begin{eqnarray}
-%f_1:Z=\Func_1(A) \nonumber \\
-%\vdots \nonumber \\
-%f_n:Z=\Func_n(A) \nonumber
-%\end{eqnarray}
-Even though there may not exist a single polynomial which describes the
-entire function over $\Zkk$, elimination theory and \Grobner basis still
-applies over this ring. Thus, it may be possible to modify the theory and 
-implementation of our word-level 
-abstraction approach in order to abstract a set of word-level 
-polynomials over $\Zkk$.
+\subsection{Multivariate polynomial ideal based FSM Traversal}
+\subsection{New Diagram Structure accelerating Polynomial Reduction}
+\subsection{Interpolation extraction using GB Algorithm}
+\subsection{Verification of Integer Arithmetic Circuits}

### 2016-10-24 dissertation/intro.tex
dissertation: milestone 1: first draft

@@
-\chapter{Introduction} \label{ch:intro}
-There is an ever-increasing need for secure communication within information 
-technology. Security of sensitive information relies more and more heavily 
-on encryption methodologies implemented in hardware by cryptographic 
-circuits. One of the most prominent of these methodologies is Elliptical 
-Curve Cryptography (\emph{ECC}), which provides more strength per encryption bit 
-than other encryption methodologies.
-The main building blocks of ECC hardware implementations 
-are fast, \emph{custom-built Galois field arithmetic circuits}.
-These circuits are notoriously hard to verify, yet their correctness is 
-vitally important in critical applications. In \cite{crypto:bug_attacks}, 
-for example, it is shown that a bug in the hardware could lead to the full 
-leakage of the secret cryptographic key, which could compromise the entire
-system. Thus, formal verification is imperative in Electronic Design 
-Automation (\emph{EDA}) when dealing with cryptographic circuits.
+\chapter{Introduction}
+\label{ch:intro}
+\section{Design, Test and Verification of Modern Sequential Circuit Components}
 
-To facilitate this verification, it is highly desirable to obtain a {\it word-level 
-representation of the datapath of the ECC arithmetic block} from its 
-bit-level implementation. Ideally, this abstraction should be {\it canonical},
-as this allows the it to be directly applicable to equivalence checking. 
-Such a canonical, word-level abstraction of the 
-Galois field arithmetic block would not only make it easier to verify and 
-reason 
-about the cryptographic system as a whole, but also enable the use of 
-higher level abstraction and synthesis tools. 
-As arithmetic circuits are custom-designed, often modularly, using Galois field arithmetic blocks, 
-the abstraction should also exploit the hierarchical nature of the circuitry.
-Due to the modular circuit structure, abstraction of each arithmetic block
-becomes the key in verification of the full circuit. 
-Practical 
-applications of ECC dictate a datapath of a minimum of $163$-bits, up to 
-$571$-bits, as 
-designated by the National Institute for Standards and Technology (NIST).
-However abstraction of 
-Galois field arithmetic circuits has been infeasible for data-paths beyond 
-$16$ bits.
+\section{Formal Verification Techniques}
+\subsection{Conventional Formal Verification}
+\subsection{Word-level Verification Techniques}
 
-This dissertation proposes an algebraic geometry based approach to abstract canonical, 
-word-level representations of bit-level Galois field arithmetic circuits.
-The approach is able to abstract representations
-for circuits up to 571 bits in size, which is the largest NIST
-standard for datapath size in ECC. 
-Verification of circuits for which this abstraction has been computed is shown to 
-be trivial; thus, the focus is on deriving the abstraction quickly and efficiently.
-%Furthermore, we show how we can use these abstract representations to 
-%perform improved formal verification of these circuits.
-
-\section{Hardware Design and Verification Overview}
-The typical design flow of a hardware system, 
-as shown in Fig. \ref{fig:cadflow}, starts with a hardware system 
-specification, which describes the necessary functions and parameters that 
-the system must perform and adhere to. The specification is typically
-modelled using a transaction-level model (\emph{TLM}), which describes 
-communication details between large circuit modules. 
-The TLM is then translated 
-into a register-transfer-level (\emph{RTL}) description, which is composed of abstracted, 
-interconnected circuit blocks that compose the entire system. RTL is 
-typically implemented 
-in hardware description languages (\emph{HDL}) such as Verilog and VHDL, 
-which are the most popular choices in the industry.
-Next, the RTL is optimized and converted into a \emph{netlist}, i.e. a large 
-collection of small physical blocks (MOSFET, Boolean logic gates, etc.) and 
-the inter-connections (wires) between them. 
-Lastly, the netlist is further optimized and then mapped onto a physical space 
-on a chip, which is then sent off for fabrication. This entire design flow
-is automated by Computer-Aided Design (\emph{CAD}) tools.
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.5\textwidth]{./figures/designFlow}
-}
-\caption{Typical hardware design flow.}
-\label{fig:cadflow}
-\end{figure}
-}
-
-%\subsection{The Hardware Verification Imperative}
-When moving from one abstraction level of the hardware design process to the 
-next, an important issue arises: how can one ensure that the functionality of 
-the optimized design matches original spec? 
-Bugs in hardware design which are 
-not caught early can have costly effects later, such as the need for a 
-redesign. Bugs in arithmetic circuits can be especially catastrophic.
-One infamous example is the 1994 floating point division (FDIV)
-bug that affected the Intel Pentium chip \cite{nicely:FDIV}, 
-and subsequently cost the company \$475 million because it was  
-discovered after the chip's release. In another more fatal case, during the
-Gulf war, an American Patriot Missile battery failed to intercept an incoming
-enemy missile due to an arithmetic error \cite{arnold:patriot}.
-Since hardware bugs can have significant consequences, 
-there has been extensive work in field of hardware verification to find and
-eliminate bugs prior to fabrication.
-
-The two main methodologies used in hardware verification are simulation and 
-formal verification. \emph{Simulation} checks correctness by applying exhaustive 
-assignments to the circuit inputs and verifying correctness of the output. 
-This ensures that the circuit performs as designed under all possible 
-inputs. Such exhaustive testing is quite effective for smaller circuits. 
-However, as the size of the circuit increases, it becomes 
-computationally infeasible to simulate all possible test vectors. This is the 
-case with Galois field arithmetic circuits, which are commonly very large in 
-real-world applications. Often for such large circuits, simulations of a smaller and more 
-manageable subset of test vectors are employed to catch bugs. While these tests
-can increase confidence in the correctness of the design,  
-{\it they do not guarantee correctness} since every data-flow of the design hasn't
-been analyzed.
-%Such is the case with large Galois field arithmetic circuits, so we instead 
-%focus on formal verification.
-
-\section{Formal Verification}
-Instead of simulating input vectors, \emph{formal verification} utilizes 
-mathematical theory to reason about the correctness of hardware designs.
-%which overcomes some limitations of simulation. 
-Formal verification has two main forms: property checking and equivalence 
-checking.
-
-{\it Property checking} (or property verification) verifies
-that a design satisfies certain given properties. Property checking is done mainly 
-in the form of theorem proving, model checking, or approaches which 
-combine the two.
-\begin{enumerate}
-\item \emph{Theorem proving} \cite{theoremproving:91} requires the existence of
-mathematical descriptors of the specification and implementation of the 
-circuit. Theorem provers apply mathematical rules to these descriptors to
-derive new properties of the specification. In this way, the tool can reduce
-a proof goal to simpler sub-goals, which can be automatically verified.
-However, generating the initial proof-goal requires extensive guidance from
-the user, so there is an overall lack of automation in theorem 
-proving.
-\item \emph{Model checking} \cite{modelcheck:99} is an approach
-to verifying finite-state systems where specification 
-properties are modeled as a system of 
-logic formulas. The design is then traversed to check if the 
-properties hold. If the design is found to violate a
-particular property, a counter-example is generated which exercises the
-incorrect behavior in the design. Such counter-examples allow the designer
-to trace the behavior and find where the error in the design lies.
-Modern model checking techniques use the result to automatically refine
-the system and perform further checking.
-These tools are typically automated, and thus have found widespread 
-use in CAD tool suites.
-\end{enumerate}
-
-{\it Equivalence Checking} verifies that two different representations of
-a circuit design have equivalent functionality. An example of equivalence
-checking as it applies to the hardware design flow is shown in
-Fig. \ref{fig:equivflow}.
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.5\textwidth]{./figures/designVerification}
-}
-\caption{Equivalence checking as applied to the hardware design flow.}
-\label{fig:equivflow}
-\end{figure}
-}
-
-There are two major
-equivalence checking techniques: graph-based
-and satisfiability-based.
-\begin{enumerate}
-\item \emph{Graph-based} techniques construct a canonical graph 
-representation, such as a Binary Decision Diagram (\emph{BDD}) or one of
-its many variants, of each circuit. A linear comparison is then conducted to 
-determine whether the two graphs are isomorphic. Since the graph 
-representation is canonical, the graphs of the two circuits will be 
-equivalent if and only if the circuits perform the same function.
-\item \emph{Satisfiability} techniques construct a miter of the two circuits,
-typically in a graph such as an And-Inverter graph (\emph{AIG}). A
-\emph{miter} is a combination of the two circuits with one bit-level output, which 
-is only in a "1" state when the outputs of the circuits differ given 
-the same given 
-input, as shown in Fig. \ref{fig:miter}. 
-A satisfiability (\emph{SAT}) tool \cite{csat} 
-is then employed to simplify the graph and find a solution to the miter, 
-i.e. find an input for which the 
-miter output is "1". If a solution is found, this solution acts as a 
-counter-example of when the circuit outputs differ; otherwise the circuits
-are functionally equivalent.
-\end{enumerate}
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.8\textwidth]{./figures/betterMiter}
-}
-\caption{A miter of two circuits.}
-\label{fig:miter}
-\end{figure}
-}
-
-%\subsection{Computer-Algebra Based Formal Verification}
-Certain formal verification methods use \emph{computer-algebra} and \emph{algebraic
-geometry} techniques based on mathematical theories.
-Unlike SAT-based verification, modern algebraic geometry 
-techniques do not explicitly solve the constraints to find a solution; 
-rather, they reason about the presence or absence of solutions, or explore
-the geometry of the solutions.
-These methods \cite{Avrunin:CAV} \cite{condrat-tacas07} \cite{gbverify:2007} 
-transform the circuit design into a polynomial system. Typically, this system
-of polynomials is then used to compute a Gr\"obner basis \cite{gb_book}. 
-Computation of Gr\"obner bases allows for 
-the easy deduction of important properties of a polynomial system, 
-such as the presence or absence of 
-solutions. These properties are then leveraged to perform 
-verification. Unfortunately, such a computation 
-has been shown to be doubly 
-exponential in the worst case, and thus these methods have not been 
-practical for real-world applications. However, recent
-breakthroughs in computer-algebra hardware verification have shown
-that it is possible to overcome the complexity of this computation while
-still utilizing the beneficial properties of a Gr\"obner bases
-\cite{lv:phd}.
-
-\section{Importance of Word-level Abstraction}
-Most formal verification techniques can benefit from word-level abstractions 
-of the circuits they verify.
-Abstraction is defined as state-space reduction, i.e{\text . }abstraction
-reduces state-space by mapping the set of states of a system to a smaller 
-set of states. Because the new representation contains fewer states, it
-is easier to comprehend and thus easier to use. 
-Word-level abstraction focuses specifically on abstracting a word-level
-representation of a circuit out of a bit-level representation. For example,
-a bit-level representation of an integer multiplier is represented by a
-collection of Boolean inputs and outputs, whereas a word-level
-abstraction hides the underlying logic and represents the circuit as two 
-integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
-multiplier increases, the logical implementation of the multiplier grows (typically
-exponentially) while the word-level abstraction stays the same.
-
-Word-level abstractions have a wide variety applications in formal 
-verification. Theorem proving techniques can leverage abstraction as an 
-automatic decision procedure or as a canonical reduction 
-engine. For example, since RTL is composed of circuit blocks that represent 
-the underlying circuit, RTL verification methods can exploit 
-abstractions of these blocks.
-This is seen in the following RTL verification methods:
-\begin{itemize}  
-\item Model checking \cite{kroening:model}, 
-where an approximation abstraction of RTL blocks is generated and then 
-refined.
-\item Graph-based equivalence 
-checking \cite{WLS} \cite{arditi:bmd}, where abstraction methods are used
-to generate a canonical word-level graph representation of the circuit.
-\item Satisfiability-based equivalence checking \cite{lpsat}, where 
-abstractions are used identify symmetrics and similarities in order to 
-minimize the amount of logic that is sent to the 
-SAT tool. 
-\end{itemize}
-
-Other equivalence checking techniques that employ abstractions 
-include satisfiability modulo theory ({\it SMT}) techniques \cite{boolector} \cite{bryant:tacas07}, 
-which are similar to SAT except they operate on higher-level data
-structures (integers, reals, bit vectors, etc.), as well as 
-constraint solving techniques \cite {ms:research} \cite{tew:iccad08}.
-In general, RTL equivalence checking approaches would ideally maintain a 
-high-level of abstraction while still retaining sufficient lower-level 
-functional details  (such as bit-vector size, precision, etc) 
-\cite{gupta_survey}.
-
-Word-level hardware abstractions also have applications in RTL and datapath 
-synthesis \cite{demicheli:iccad_98} \cite{demicheli:dac_99}
-\cite{demicheli:tcad_03}. 
-Abstractions of circuits allow for design reuse, which allows for tool-automated 
-synthesis of larger circuit blocks.
-Since hardware design specifications tend to be word-level, synthesis tools 
-can use these larger circuit blocks to generate and optimize the
-datapaths and create the RTL of the system. Thus, in order for a circuit to 
-be used by these automated synthesis tools, its word-level abstraction must
-be known.
-
-Finally, abstractions can also be applied to detect malicious 
-modifications to a circuit, potentially inserted as a hardware trojan horse.
-Hardware trojans, a relatively new security concern in the hardware 
-industry, use certain techniques to add incorrect behavior to a 
-design. 
-This behavior is only activated under certain rare circumstances that only 
-the mal-intent designer has knowledge of.
-The behavior is purposely hidden and is very difficult to encounter during 
-simulation of the design. A manufactured chip with a subsystem 
-that contains a hardware trojan could compromise the entire system in which 
-it is used.
-In some hardware trojan cases, formal verification techniques may be applied 
-to catch a bug in a design and provide a counter-example which exercises it. 
-However, it can be difficult to tell whether the bug in the design was 
-introduced intentionally of not. On the other hand, word-level abstractions 
-of bit-level circuits {\it effectively reverse-engineer the true function 
-implemented by the circuit}, which could be used to determine the designer's 
-true intention.
-
-\section{Dissertation Objective, Motivation, and Contributions}
-This dissertation focuses on abstracting a canonical, 
-word-level representation of hardware (bit-level) implementations of 
-combinational circuits. The proposed technique is a full abstraction solution which can be 
-applied to any arbitrary acyclic combinational circuit. 
-It is particularly efficient when applied to Galois field arithmetic circuits.
-Using this technique, if the abstraction of the circuit's implementation and its
-specification are found, they can be easily compared to determine equivalence.
-Implementation of a custom software tool, developed to compute the abstractions, is
-also described.
-
-\subsection{Motivating Application}
-The motivation for this work comes from applications of Galois field 
-arithmetic circuits in elliptical curve cryptography ({\it ECC}) hardware systems.
-The main operations of encryption, decryption, and 
-authentication in ECC rely on operations performed on elliptic curves, which 
-are implemented in hardware as polynomial functions over Galois fields. 
-To be applicable in real-world situations, ECC data-paths
-should be a minimum of $163$-bits wide, which is the minimum NIST standard, 
-up to a recommended size of $571$-bit operand widths. Many non-ECC cryptosystems
-have datapaths on the order of $1000$-bits.
-
-A Galois field arithmetic circuit with a datapath size of $k$ 
-is built as a Boolean function: $\mathbb{B}^k \rightarrow \mathbb{B}^k$. 
-This function is mapped to an operation 
-$f:\mathbb{F}_{2^k} \rightarrow \mathbb{F}_{2^k}$ 
-over the Galois field $\mathbb{F}_{2^k}$. 
-These circuits are custom-built, modular
-systems which cannot be synthesized due to their complex nature. Thus, 
-formal verification is needed to ensure they operate correctly.
-
-Recent computer-algebra based formal verification techniques have been
-able to perform verification of Galois field arithmetic circuits with
-a datapath size up to $163$-bits \cite{lv:phd}. Word-level abstractions 
-of Galois
-field arithmetic circuits could be used to further improve these formal 
-verification techniques to allow for verification of larger circuits, as
-well as provide the other benefits of word-level abstraction.
-However, there is currently no technique for computing word-level 
-abstractions of Galois field circuits of any practical size.
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.5\textwidth]{./figures/interpolate}
-}
-\caption{Circuit with $k$-bit input $A$ and $k$-bit output $Z$. 
-Abstraction to be derived as $Z=\Func(A)$.}
-\label{fig:abstractA_Z}
-\end{figure}
-}
-
-While the motivation comes from the need to verify Galois field arithmetic
-circuits, the presented approach can be generalized to be applicable to any
-combinational acyclic circuit.
-Any such circuit with a $k$-bit input $A$ and a $k$-bit output $Z$, such
-as the one shown in Fig. \ref{fig:abstractA_Z}, computes 
-$f: \mathbb{B}^k \rightarrow \mathbb{B}^k$ and can thus be analyzed as the
-function $f: \Fkk \rightarrow \Fkk$. Over $\Fkk$ this function can be 
-represented as the polynomial $Z=\Func(A)$. This is trivially generalized
-when there are multiple $k$-bit inputs $A_1,A_2,\dots,A_i$, i.e. $Z=\Func(A_1,\dots,A_i)$.
-Now assume the word-size of the input differs from the output, that is the circuit
-computes $f: \mathbb{B}^m \rightarrow \mathbb{B}^n$ for $m \neq n$. This can 
-be represented as a function over Galois fields as $f: \F_{2^m} \rightarrow \F_{2^n}$.
-This function can be analyzed over the field $\Fkk$ such that $\Fkk \supset \F_{2^m}$ 
-and $\Fkk \supset \F_{2^m}$, where $k=LCM(m,n)$.
-
-\subsection{Dissertation Contributions}
-To solve the problem of word-level abstraction, this dissertation proposes 
-a full solution consisting of three main contributions.
-
-\begin{enumerate}
-\item A theory for finding the word-level abstraction from a bit-level circuit over Galois fields is created.
-The given bit-level circuit implementation is modelled as a system of
-polynomials over the field.
-This theory is derived using techniques from computer-algebra, notably the theory of
-\Grobner basis \cite{pruss:iwls13}.
-\item Using this theory, new algorithms based on symbolic computation are developed to 
-derive the word-level abstraction. The algorithms are designed to be applicable to 
-industry-size arithmetic circuits over Galois fields \cite{pruss:dac14}\cite{sun:date15}.
-A complexity analysis of the algorithmic approach is also presented. Furthermore, the
-approach is also generalized to make it applicable to arbitrary combinational circuits.
-Finally, we show how the approach can be used to exploit the hierarchical structure of
-large Galois field multipliers designed over composite fields.
-\item A custom software tool implementation of the algorithmic approach is described, including
-an analysis of efficient data structures designed for this purpose \cite{pruss:tcad15}.
-\end{enumerate}
-
-Experiments show that the proposed solution can abstract canonical, word-level, 
-polynomial representations of Galois field arithmetic circuits up to $1024$-bits in
-size, while other contemporary approaches are infeasible beyond a $32$-bit designs.
-
-%an 
-%approach based on computer-algebra, algebraic geometry - notably the theory of
-%{\it Gr\"obner bases} \cite{gb_book} \cite{buchberger_thesis}.
-%and {\it elimination ideals} \cite{ideals:book}, as our approach for 
-%abstracting canonical word-level representations from bit-level Galois 
-%field arithmetic circuits. 
-%Computer-algebra techniques easily integrate with 
-%Galois field theory and allow for polynomial abstraction of the circuit over 
-%the field itself. Thus, if the circuit computes a function over some field $
-%\mathbb{F}_{2^k}$, the resulting abstraction is a word-level polynomial 
-%function over the same field $\mathbb{F}_{2^k}$.
-
-%The given bit-level circuit implementation is first modeled as a system of
-%polynomials over the field. Using the theory of elimination ideals, we 
-%derive an elimination term ordering and prove that by using this ordering we 
-%can obtain a canonical word-level representation of a circuit by 
-%computing a Gr\"obner basis of the polynomials. However, complexity 
-%of the computation of a Gr\"obner basis proves to be prohibitive for circuits
-%of practical size. Therefore, we employ a select sub-set of computations 
-%from the Gr\"obner basis theory to overcome this complexity and obtain the
-%abstraction. We prove that this simpler computation can be 
-%performed via a polynomial reduction process, and engineer a custom 
-%verification tool for this purpose. Using this approach, we are able to 
-%successfully 
-%abstract word-level representations of Galois field arithmetic circuits up 
-%to $571$-bits, which is the {\it largest recommended NIST standard} for ECC.
+\section{Objective, Motivation and Contribution of this Dissertation}
 
 \section{Dissertation Organization}
-The rest of this dissertation is organized as follows. Chapter
-\ref{ch:prev} reviews previous applicable work and highlights their
-drawbacks with respect to the canonical, word-level abstraction problem. 
-Chapter \ref{ch:prelim} describes the properties of Galois fields, 
-$\mathbb{F}_{2^k}$, and explains the process of constructing them.
-It also describes how to design arithmetic circuits over such fields, their 
-complexities, and the role of these circuits in Elliptic Curve Cryptography.
-Chapter \ref{ch:ideals} provides a theoretical background of 
-computer-algebra and Gr\"obner bases and explains their application
-to Galois fields. 
-Chapter \ref{ch:abstract} describes an approach to abstract 
-word-level polynomial representations of combinational circuits using a 
-Gr\"obner basis computation.
-Chapter \ref{ch:improv} improves on this word-level abstraction approach to 
-make it applicable to much larger circuits.
-Chapter \ref{ch:generalize} generalizes the abstraction approach to make
-it applicable to circuits with varying operand word-lengths. It also describes how the 
-approach can take advantage of the hierarchy of arithmetic circuits designed over
-composite fields.
-Chapter \ref{ch:implement} describes the implementation details of a 
-custom abstraction tool and gives experimental results of
-abstracting large Galois field multiplier circuits.
-Chapter \ref{ch:concl} concludes the dissertation and outlines potential future 
-research for continuation of this work.

### 2016-10-24 dissertation/prelim.tex
dissertation: milestone 1: first draft

--- 2016-09-21 dissertation/algprelim.tex
@@
-\chapter{Computer Algebra Fundamentals} \label{ch:ideals}
-This chapter reviews fundamental concepts of commutative and 
-computer algebra which are used in this work. 
+\chapter{Preliminaries}
+\label{ch:prelim}
+This chapter provides a mathematical background for understanding 
+Galois fields and explains how to design Galois field arithmetic circuits.
 Specifically, this chapter covers monomial ordering, polynomial ideals and 
 varieties, and the computation of \Grobner bases.
-It also overviews elimination theory as well as Hilbert's Nullstellensatz 
-theorems and how
-they apply to Galois fields. The results of these theorems are used in polynomial
-abstraction and formal verification of Galois field circuits and 
-are discussed 
-in subsequent chapters. The material of this chapter is
-mostly referred from the textbooks \cite{ideals:book} \cite{gb_book} and 
-previous work by {\it Lv} \cite{lv:phd}.
-
-\section{Monomials, Polynomials, and Term Orderings}
-
+Moreover, we introduce the mathematical concepts of groups, rings, fields, and 
+polynomials. 
+We then apply these concepts to create Galois field arithmetic functions and 
+explain how elimination and abstraction on Galois field works.
+The material is referred from \cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997} for Galois field concepts,
+\cite{ideals:book} \cite{gb_book} for commutative algebra basics and 
+previous work by {\it Lv} \cite{lv:phd} and {\it Pruss} \cite{tim:phd}.
+
+\section{Commutative Algebra}
+\label{sec:algebra}
+\subsection{Group, ring and field}
+\begin{Definition}
+An {\bf abelian group} is a set $\mathbb{S}$ with a binary operation $'+'$
+which satisfies the following properties: 
+\begin{itemize}
+\item {\it Closure Law:} For every $a, b \in \mathbb{S}, a + b \in \mathbb{S}$  
+\item {\it Associative Law:} For every $a, b, c \in \mathbb{S}, (a + b) + c = a + (b + c)$
+\item {\it Commutativity:} For every $a, b \in \mathbb{S}, a + b = b + a$. 
+\item {\it Additive Identity:} There is an identity element $0 \in \mathbb{S}$
+such that for all $a \in \mathbb{S};$ $a + 0 = a$.
+\item {\it Additive Inverse:} If $a \in \mathbb{S}$, then there is an
+element $a^{-1} \in \mathbb{S}$ such that $ a + a^{-1} = 0$.
+\end{itemize}
+\end{Definition}
+
+The set of integers $\mathbb{Z}$ forms an abelian group under the addition operation.
+
+\begin{Definition}
+Given a set $\mathbb{R}$ with two binary operations, $'+'$ and $'\cdot'$, 
+and element $0 \in \mathbb{R}$, the system $\mathbb{R}$ is called a {\bf commutative ring with unity} if the following properties hold:
+\begin{itemize}
+\item $\mathbb{R}$ forms an abelian group under the '+' operation with additive identity element $0$.
+\item {\it Multiplicative Distributive Law}: For all $a, b, c \in$ $\mathbb{R}$, $a\cdot (b + c) = a\cdot b + a\cdot c$.
+\item {\it Multiplicative Associative Law}: For every $a, b, c\in \mathbb{R}$, $a\cdot (b\cdot c) = (a\cdot b)\cdot c$. 
+\item {\it Multiplicative Commutative Law}: For every $a,b \in \mathbb{R}$, $a\cdot b = b\cdot a$
+\item {\it Identity Element}: There exists an element $1 \in$ $\mathbb{R}$ 
+such that for all $a \in \mathbb{R}$, $a\cdot 1 = a =1\cdot a$
+\end{itemize}
+\end{Definition}
+
+For the purpose of this dissertation, any time we refer to a {\bf ring}, we are 
+specifically referring to a {\bf commutative ring with unity}. Two common 
+examples of such rings are the set of integers, $\mathbb{Z}$, and the set of 
+rational numbers, $\mathbb{Q}$. Note that while both of these examples are
+rings with an infinite number of elements, the number of elements in a ring 
+can also be finite.
+
+\begin{Definition}
+A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
+non-zero element in $\mathbb{F}$ has a multiplicative inverse; i.e. $\forall
+a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
+\hat{a} = 1$.
+\end{Definition}
+
+A field is defined as a ring with one extra condition: the presence of a 
+multiplicative inverse for all non-zero elements.
+Therefore, a field must be a ring while a ring is not necessarily a field.
+For example, the set $\mathbb{Z}_{2^k} = \{0,1,\cdots, 2^k-1\}$ forms a finite ring.
+However, $\mathbb{Z}_{2^k}$ is not a field because not every element in
+$\mathbb{Z}_{2^k}$ has a multiplicative inverse. 
+In the ring $\mathbb{Z}_{2^3}$, for 
+instance, the element $5$ has an inverse ($5\cdot5\pmod{8}=1$) but the element $4$
+does not.
+
+The main concept of field theory is {\bf Field Extensions}. The idea behind a
+field extension is to take a base field and construct a larger field which 
+contains the base field as well as satisfies additional properties. For example,
+the set of real numbers $\mathbb{R}$ forms a field; one common extension of 
+$\mathbb{R}$ is the set of complex numbers $\mathbb{C}=\mathbb{R}(i)$. Every
+element of $\mathbb{C}$ can be represented as $a+b\cdot i$ where $a,b \in \mathbb{R}$,
+hence $\mathbb{C}$ is a two-dimensional extension of $\mathbb{R}$.
+
+Like rings, fields can also contain either an infinite or a finite number of 
+elements. 
+In this dissertation we focus on finite fields, also known as Galois fields, and 
+the construction of their field extensions.
+
+\subsection{Galois Field}
+Galois fields, also known as finite fields, find widespread applications in 
+many areas of electrical engineering and computer science such as error-
+correcting codes, elliptic curve cryptography, digital signal processing, 
+testing of VLSI circuits, among others.
+In this dissertation, we specifically focus on their application to 
+Elliptic Curve Cryptography as Galois field arithmetic circuits.
+This section describes the relevant Galois field concepts
+\cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997}
+and hardware arithmetic designs over such fields \cite{mastro:1989} \cite{PT:1985} 
+\cite{acar:1998} \cite{wu:2002} \cite{Knezevic:2008}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{Definition} 
+A {\bf Galois field}, denote $\Fq$, is a field with a finite
+number of elements, $q$. The number of elements $q$ of the Galois field is
+a power of a prime integer, i.e. $q = p^k$, where $p$ is a prime
+integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
+$\F_{p^{k}}$.
+\end{Definition}
+
+Fields in the form $\F_{p^{k}}$ are called Galois extension fields.
+We are specifically interested in extension fields of type 
+$\Fkk$, where $k > 1$. These are extensions of the binary
+field $\F_2$.
+\begin{Example}
+Addition and multiplication operations over $\F_2$:
+\begin{table}[!h]
+	\centering
+	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
+	\hhline{~---~}
+	\multirow{3}{*}{} & $+$ & $0$ & $1$ & \multirow{3}{*}{} \\
+	\hhline{~---~}
+	& $0$ & $0$ & $1$ & \\
+	& $1$ & $1$ & $0$ & \\
+	\hhline{~---~}
+	\multicolumn{5}{c}{}\\
+	\multicolumn{5}{c}{Addition over $\F_2$}\\
+	\end{tabular}
+	\quad
+	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
+	\hhline{~---~}
+	\multirow{3}{*}{} & $\cdot$ & $0$ & $1$ & \multirow{3}{*}{} \\
+	\hhline{~---~}
+	& $0$ & $0$ & $0$ & \\
+	& $1$ & $0$ & $1$ & \\
+	\hhline{~---~}
+	\multicolumn{5}{c}{}\\
+	\multicolumn{5}{c}{Multiplication over $\F_2$}\\
+	\end{tabular}
+\end{table}
+
+Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
+because it is performed modulo $2$.
+Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.
+\end{Example}
+
+Algebraic extensions of the binary field $\F_{2}$  
+are generally termed as {\it binary extension fields} $\Fkk$.
+Where elements in $\F_2$ can only represent $1$ bit, elements in $\Fkk$ 
+represent a $k$-bit vector.
+This allows them to be widely used in digital hardware applications.
+In order to construct a Galois field of the form $\Fkk$, 
+an {\bf irreducible polynomial} is required:
+\begin{Definition}
+A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$ is {\bf irreducible} 
+if $P(x)$ is non-constant with degree $k$ and cannot be 
+factored into a product of polynomials of lower degree in $\mathbb{F}_2[x]$.
+\end{Definition}
+
+Therefore, the polynomial $P(x)$ with degree $k$ is irreducible over 
+$\mathbb{F}_{2}$ if and only if it has no roots in $\mathbb{F}_{2}$,
+i.e if $\forall a \in \mathbb{F}_{2}$, $P(a)\neq 0$.
+For example, $x^2+x+1$ is an irreducible polynomial over $\mathbb{F}_{2}$
+because it has no solutions in $\mathbb{F}_{2}$, i.e. $(0)^2+(0)+1=1\neq0$ 
+and $(1)^2+(1)+1=1\neq0$ over $\F_2$.
+Irreducible polynomials exist for any degree $\geq 2$ in $\mathbb{F}_2[x]$.
+
+Given an irreducible polynomial $P(x)$ of degree $k$ in the polynomial ring 
+$\mathbb{F}_2[x]$, we can construct a binary extension field 
+$\mathbb{F}_{2^k} \equiv \mathbb{F}_2[x] \pmod{P(x)}$.
+Let $\alpha$ be a root of $P(x)$, i.e., $P(\alpha)=0$.
+Since $P(x)$ is irreducible over
+$\mathbb{F}_2[x]$, $\alpha \notin \mathbb{F}_2$. 
+Instead, $\alpha$ is an element in $\mathbb{F}_{2^k}$. 
+Any element $A \in \mathbb{F}_{2^k}$ is then represented as: 
+\begin{equation}\label{rep:poly}
+A= \sum_{i=0}^{k-1} (a_i \cdot \alpha^i) = a_0 + a_1\cdot\alpha + \cdots + a_{k-1}\cdot \alpha^{k-1}\nonumber
+\end{equation}
+where $a_i \in \mathbb{F}_2$ are the coefficients and $P(\alpha)=0$.
+
+To better understand this field extension, compare its similarities to another
+common-place
+field extension $\C$, the set of complex numbers. $\C$ is an extension of the field 
+of real numbers $\R$ with an additional element $i=\sqrt{-1}$, which is an imaginary
+root in $\R$.
+Thus $i \notin \R$, rather $i \in \C$.
+Every element $A \in \mathbb{C}$ can be represented as:
+\begin{equation}\label{rep:polyC}
+A=\sum_{j=0}^{1} (a_j \cdot i^j)=a_0+a_1\cdot i
+\end{equation}
+where $a_j \in \R$ are coefficients. Similarly, $\Fkk$ is an extension of $\F_2$ with 
+an additional element $\alpha$, which is the ``imaginary root'' of an irreducible 
+polynomial $P$ in $\F_2[x]$.
+
+Every element $A \in \Fkk$ has a degree less than $k$ because 
+$A$ is always computed modulo $P(x)$, which has degree $k$. 
+Thus, $A\pmod {P(x)}$ can be of degree at most $k-1$ and at least $0$.
+For this reason, the field $\mathbb{F}_{2^k}$ can be viewed as a $k$
+dimensional vector space over $\mathbb{F}_{2}$. 
+The equivalent bit vector representation for element $A$ is:
+\begin{equation}
+A=(a_{k-1} a_{k-2} \cdots a_{0})
+\end{equation}
+
+\begin{Example}
+A 4-bit Boolean vector, $(a_{3} a_{2} a_{1} a_{0})$
+can be presented over $\F_{2^4}$ as: 
+\begin{equation}
+a_3 \cdot \alpha^3+a_2 \cdot \alpha^2+a_1 \cdot \alpha+a_0
+\end{equation}
+For instance, the Boolean vector $1011$ is represented as the element 
+$\alpha^3+\alpha+1$.
+\end{Example}
+
+\begin{Example}\label{exp:1}
+Let us construct $\mathbb{F}_{2^4}$ as $\mathbb{F}_2[x] \pmod{ P(x)}$, where
+$P(x)=x^4+x^3+1 \in \mathbb{F}_2[x]$ is an irreducible polynomial of degree $k=4$. 
+Let $\alpha$ be the root of $P(x)$, i.e. $P(\alpha)=0$.
+
+Any element $A \in \mathbb{F}_2[x] \pmod{ x^4 + x^3 + 1}$
+has a representation of the type: $A = a_3 x^3 + a_2 x^2 +
+a_1 x + a_0$ (degree $< 4$) where the coefficients $a_3, \dots, a_0$ are in $\F_2 =
+\{0, 1\}$. Since there are only $16$ such polynomials, we obtain
+$16$ elements in the field $\mathbb{F}_{2^4}$. Each element in
+$\mathbb{F}_{2^4}$ can then be viewed as a $4$-bit vector over $\mathbb{F}_{2}$. 
+Each element also has an exponential $\alpha$
+representation. All three representations are shown in Table
+\ref{tab:gfelement}.
+
+\begin{table}[h]
+\begin{center}
+\caption{Bit-vector, Exponential and Polynomial representation of
+elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$}\label{tab:gfelement} 
+\begin{tabular}{|c|c|c||c|c|c|} 
+\hline
+$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
+\hline
+$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
+\hline
+$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
+\hline
+$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
+\hline
+$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
+\hline
+$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
+\hline
+$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
+\hline
+$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
+\hline
+$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+We can compute the polynomial representation from the exponential representation.
+Since every element is computed $\pmod{P(\alpha)} = \pmod{\alpha^4+\alpha^3+1}$, 
+we compute the element $\alpha^{4}$ as 
+\begin{equation}
+\alpha^{4} \pmod{ \alpha^4+\alpha^3+1} = -\alpha^3 - 1 = \alpha^3+1
+\end{equation}
+Recall that all coefficients of $\F_{2^4}$ 
+are in $\F_{2}$ where $-1 = +1$ modulo 2.
+The next element $\alpha^{5}$ can be computed as 
+\begin{equation}
+\alpha^{5} = \alpha^{4}\cdot \alpha = (\alpha^3+1)\cdot \alpha = \alpha^4+\alpha = \alpha^3+\alpha+1 
+\end{equation}
+Then $\alpha^6$ can be computed as $\alpha^{5}*\alpha$ and so on.
+\end{Example}
+
+An irreducible polynomial can also be a primitive polynomial.
+
+\begin{Definition}
+A {\bf primitive polynomial} $P(x)$ is a polynomial with coefficients in $\mathbb{F}_2$ 
+which has a root $\alpha$ $\in$ $\mathbb{F}_{2^k}$
+such that \{$0$, $1(=\alpha^{{2^k}-1})$, $\alpha$, $\alpha^2$, $\cdots$, $\alpha^{2^k-2}$\} is the set of 
+all elements in $\mathbb{F}_{2^k}$, 
+where $\alpha$ is a {\bf primitive element} of $\mathbb{F}_{2^k}$. 
+\end{Definition}
+
+A primitive polynomial is guaranteed to generate all distinct elements 
+of a finite field $\mathbb{F}_{2^k}$ while an irreducible polynomial
+has no such guarantee.
+Often, there exists more than one irreducible polynomial of degree $k$.
+In such cases, any degree $k$ irreducible polynomial can be 
+used for field construction. For example, both $x^3+x+1$ and $x^3+x^2+1$ 
+are irreducible in $\mathbb{F}_2$ and either one can be used
+to construct $\mathbb{F}_{2^3}$. This is due to the following:
+
+\begin{Theorem}\label{the:unique}
+There exist a {\bf unique} field $\mathbb{F}_{p^k}$, for any prime $p$ and any positive integer $k$.
+\end{Theorem}
+
+Theorem \ref{the:unique} implies that Galois fields with the same number of elements are 
+{\bf isomorphic} to each other up to the labeling of the elements.
+
+Theorem \ref{the:fer} provides an important property for investigating solutions to
+polynomial equations in $\Fq$.
+
+\begin{Theorem}\label{the:fer}
+ $\left[Generalized\  Fermat's\  Little\  Theorem \right]$ Given a
+ Galois field $\mathbb{F}_{q}$, each element $A \in \mathbb{F}_{q}$ satisfies: 
+\begin{eqnarray}\label{fe}
+ A^{q} & \equiv & A  \nonumber \\
+ A^{q} - A & \equiv& 0  
+\end{eqnarray}
+\end{Theorem}
+
+We can extend Theorem \ref{the:fer} to polynomials in $\mathbb{F}_{q}[x]$ as 
+follows: 
+\begin{Definition}
+Let $x^q-x$ be a polynomial in $\mathbb{F}_{q}[x]$.
+Every element $A \in \mathbb{F}_{q}$ is a solution to  $x^q-x=0$. 
+Therefore, $x^{q} - x$ always {\it vanishes} in $\mathbb{F}_{q}$. Such 
+polynomials are called {\bf vanishing polynomials} of the field $\mathbb{F}_{q}$.
+\end{Definition}
+
+\begin{Example}
+Given $\mathbb{F}_{2^2} =\{0,1,\alpha,\alpha+1\}$ with $P(x)=x^2+x+1$, where $P(\alpha)=0$. 
+ \begin{eqnarray}
+ 0^{2^2}&=&0 \nonumber \\
+ 1^{2^2}&=&1 \nonumber \\
+ \alpha^{2^2}&=&\alpha \pmod {\alpha^2+\alpha+1}\nonumber \\
+ (\alpha+1)^{2^2}&=&\alpha+1 \pmod {\alpha^2+\alpha+1} \nonumber 
+ \end{eqnarray}
+\end{Example}
+
+A Galois field $\F_q$ can be fully contained within a larger field $\F_{q^k}$.
+That is, $\F_q \subset \F_{q^k}$.
+For example, Fig \ref{fig:contain2_4_16} shows the containment of the fields 
+$\F_2 \subset \F_4 \subset \F_{16}$. It's easy to see that since $\F_4=\F_{2^2}$, it
+contains $\F_2$. Likewise $\F_{16}=\F_{4^2}=\F_{2^4}$ contains $\F_4$ and $\F_2$.
+The elements $\{0,1,\alpha,\dots,\alpha^{14}\}$
+designate $\F_{16}$. Of these, $\{0,1,\alpha^5,\alpha^{10}\}$ create $\F_4$.
+From these, only $\{0,1\}$ exist in $\F_2$.
+
+\begin{Theorem}
+$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, i.e. if $n$ divides $m$.
+\end{Theorem}
+
+Therefore:
+\begin{itemize}
+\item $\F_2 \subset \F_{2^2} \subset \F_{2^4} \subset \F_{2^8} \subset \dots$
+\item $\F_2 \subset \F_{2^3} \subset \F_{2^9} \subset \F_{2^{27}} \subset \dots$
+\item $\F_2 \subset \F_{2^5} \subset \F_{2^{25}} \subset \F_{2^{125}} \subset \dots,$ and so on
+\end{itemize}
+
+\begin{Definition}
+The {\bf algebraic closure} of the Galois field $\F_{2^k}$, denoted $\overline{\Fkk}$, is the 
+union of all fields $\F_{2^n}$ such that $k \mid n$.
+\end{Definition}
+
+% TODO \Fkk operations example
+
+\subsection{Monomials, Polynomials and Polynomial Arithmetics}
 \begin{Definition} \label{def:mono}
@@
 where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
-integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$. In the
+integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.
+
+\begin{Definition}\label{def:poly}
+Let $\mathbb{R}$ be a ring. A {\bf polynomial} over $\mathbb{R}$ in the 
+indeterminate $x$ is an expression of the form:
+\begin{equation} \label{eq:poly1}
+a_0 + a_1 x + a_2 x^2 + \cdots + a_k x^k = \sum_{i=0}^{k} a_i x^i, \forall a_i \in \mathbb{R}. 
+\end{equation}
+
+\end{Definition}
+
+The constants $a_i$ are the coefficients and $k$ is the degree of the polynomial. 
+For example, $8x^3 + 6x + 1$ is a polynomial in $x$ over $\mathbb{Z}$, with
+coefficients $8$, $6$, and $1$ and degree $3$.
+
+\begin{Definition}
+The set of all polynomials in the indeterminate
+$x$ with coefficients in the ring $\mathbb{R}$ forms a {\bf
+ring of polynomials} $\mathbb{R}[x]$. 
+Similarly, $\mathbb{R}[x_1,x_{2},\cdots, x_{n}]$ 
+represents the ring of multivariate polynomials with coefficients in $\mathbb{R}$.
+\end{Definition}
+
+For example, $\mathbb{Z}_{2^4}[x]$ stands for the set of all polynomials in
+$x$ with coefficients in $\mathbb{Z}_{2^4}$. $8x^3 + 6x + 1$ is an instance 
+of a polynomial contained in $\mathbb{Z}_{2^4}[x]$.
 
@@
 terms in a polynomial are ordered.
-%%%%%%%%%%%%%%%%%%%   monomial    Ordering   %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{Definition}
-Let $\mathbb{T}^{d}=\{x^{\alpha}: \alpha\in \mathbb{Z}_{\ge 0}^{d}\}$ be the set of all monomials in $x_{1},\dots,x_{d}$.
-A {\bf monomial order} $>$ on $\mathbb{T}^{d}$ is a total well-ordering satisfying:
-\begin{itemize}
-        \item For all $x^{\alpha}, x^{\beta} \in \mathbb{T}^{d}$, $x^{\alpha}$ and $x^{\beta}$ are comparable
-	\item For any $x^{\alpha} \in \mathbb{T}^{d}$, $x^{\alpha}>1$
-	\item For all $x^{\alpha}, x^{\beta}, x^{\gamma} \in \mathbb{T}^{d}$, $x^{\alpha}>x^{\beta} \Rightarrow x^{\alpha} \cdot x^{\gamma}> x^{\beta} \cdot x^{\gamma}$
-\end{itemize}
-\end{Definition}
-
-Term-orderings are totally ordered, i.e. anti-symmetric with constant terms
-last in the ordering.
-A total-order ensures that there is no ambiguity with respect to where a 
-term is found in the term-ordering.  Total orderings for monomials come in 
-different forms, notably 
-lexicographic orderings (lex), and its variants: degree-lexicographic ordering (deglex) 
-and reverse degree-lexicographic ordering (degrevlex).
+
+Common term orderings are lexicographic ordering (LEX) and its variants: 
+degree-lexicographic ordering (DEGLEX) and reverse degree-lexicographic ordering (DEGREVLEX).
 
@@
 
-Applying these term orderings, we have the following relations, 
-where $a > b > c$.
-
-\begin{eqnarray}
-    \eqntext{lex:}  
-    a^2b > a^2 > abc > ab > ac^2 > ac > b^2c > b^2 > bc^3 > 1 
-    \label{ex:ordering:lex}\\
-    \eqntext{deglex:} 
-    bc^3 > a^2b > abc > ac^2 > b^2c > a^2 > ab >
-    ac > b^2 > 1
-    \label{ex:ordering:deglex}\\
-    \eqntext{degrevlex:}  
-    bc^3 > a^2b > abc > b^2c > ac^2 > a^2 > ab >
-    b^2 > ac > 1
-    \label{ex:ordering:degrevlex}
-\end{eqnarray}
-
-The difference between the {\it lex} and two {\it deg-} orderings is
-obvious, while the difference between the two degree-based orderings can be
-seen by considering from which direction the term is lexed, e.g. $a\cdot c\cdot c > b\cdot b\cdot c$ (deglex, left-to-right)
-versus $b\cdot b\cdot c > a \cdot c\cdot c$ (degrevlex, right-to-left).
-
-\begin{Example}
-Let $f = 2x^2yz + 3xy^3 - 2x^3$. The effects of different term orderings on 
-$f$ are:
-\begin{itemize}
-\item lex $x> y> z$: $f = -2x^3 + 2x^2yz + 3xy^3$
-\item deglex $x>y>z$:  $f = 2x^2yz + 3xy^3 -2x^3$
-\item degrevlex $x>y>z$: $f = 3xy^3 + 2x^2yz - 2x^3$
-\end{itemize}
-\end{Example}
-
-%Based on the {\it monomial ordering}, we have the following concepts:
+Based on the {\it monomial ordering}, we have the following concepts:
 
@@
 
-\section{Varieties and Ideals}
+Similarly, we can also define when a polynomial is divided (reduced) by a set of polynomials.
+\begin{Definition}
+    The {\bf reduction} of a polynomial $f$, by another polynomial $g$, to
+    a reduced polynomial $r$ is denoted:
+    \begin{equation*}
+        f\stackrel{g}{\textstyle\longrightarrow}r
+    \end{equation*}
+    Reduction is carried out using multivariate, polynomial long division. 
+    % The long division is performed according to a term-ordering on polynomials, and the division algorithm 
+    %terminates when the leading term of the divisor does not divide any 
+    %other term in the dividend.
+
+    For sets of polynomials, the notation 
+    \begin{equation*}
+    f\stackrel{F}{\textstyle\longrightarrow}_+r    
+    \end{equation*}
+    represents the reduced polynomial $r$ resulting from $f$ as reduced by a 
+    set of non-zero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
+    $r = 0$  or no term in $r$ is divisible  by a $lm(f_i), \forall f_i \in F$.
+\end{Definition}
+
+The reduction process $f\stackrel{F}{\textstyle\longrightarrow}_+r$, of 
+dividing a polynomial $f$ by a set of polynomials of $F$, can be modeled as
+repeated long-division of $f$ by each of the polynomials in $F$ until no
+further reductions can be made. The result of this process is then $r$.
+This reduction process is shown in Algorithm \ref{alg:polydiv}.
+
+\begin{algorithm}[H]
+\SetAlgoNoLine
+
+ \KwIn{$f,f_{1},\dots,f_{s}$}
+ \KwOut{$r,a_{1},\dots,a_{s}$, such that $f=a_{1}\cdot f_{1}+\dots+a_{s}\cdot f_{s}+r$.}
+
+ $a_{1}=a_{2}=\dots=a_{s}=0$; $r=0$\;
+ $p:=f$\;
+
+ \While { $p \neq 0$ }
+ {
+	i=1\;
+	divisionmark = false\;
+	\While { $i\le s  $ \&\& divisionmark = false }
+	{
+		\eIf {$f_{i}$ can divide $p$}
+		{
+			$a_{i}=a_{i}+lt(p)/lt(f_{i})$\;
+			$p=p-lt(p)/lt(f_{i}) \cdot f_{i}$\;
+			divisionmark = true\;
+		}
+		{
+			i=i+1\;
+		}
+	}
+
+	\If {divisionmark = false}
+	{
+		$r=r+lt(p)$\;
+		$p=p-lt(p)$\;
+	}
+
+ }
+\caption{Polynomial Reduction}\label{alg:polydiv}
+\end{algorithm}
+
+The reduction algorithm keeps canceling the leading terms of polynomials 
+until no more leading terms can be further canceled.
+So the key step is $p=p-lt(p)/lt(f_{i}) \cdot f_{i}$, as the following 
+example shows.
+\begin{Example}
+Given $f = y^{2}-x$ and $f_{1} = y - x$ in $\mathbb{Q}[x,y]$ with $deglex$: 
+$y>x$, perform $f\stackrel{f_1}{\textstyle\longrightarrow}_+r$:
+
+\begin{enumerate}
+\item $f=y^{2}-x$, $f/f_{1}=f-lt(f)/lt(f_{1}) \cdot f_{1}=y^{2}-x-(y^{2} /y) \cdot (y-x)=y\cdot x-x$
+\item $f=y\cdot x-x$, $f/f_{1}=f-lt(f)/lt(f_{1}) \cdot f_{1}=(y\cdot x-x)/f_{1}=x^{2}-x$
+\item $f=x^{2}-x$, no more operations possible, so $r=x^{2}-x$
+\end{enumerate}
+\end{Example}
+
+\section{Algebraic Geometry}
+\label{sec:geo}
+\subsection{Varieties and Ideals}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A {\bf \Grobner basis} provides a decision procedure for performing this 
-test, which is described in the following section. 
-A future section focuses on {\bf Hilbert's Nullstellensatz}, 
-which describes the properties of the ideal of a variety, $I(V(J))$.
-
-%%%%%%%%%%%%%%%%%%%%	Grobner bases	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\Grobner Bases}
+test, which is described in the following part.
+
+\subsection{\Grobner Bases}
 
@@
 \end{Definition}
-
-\begin{Definition}
-    The {\bf reduction} of a polynomial $f$, by another polynomial $g$, to
-    a reduced polynomial $r$ is denoted:
-    \begin{equation*}
-        f\stackrel{g}{\textstyle\longrightarrow}r
-    \end{equation*}
-    Reduction is carried out using multivariate, polynomial long division. 
-    % The long division is performed according to a term-ordering on polynomials, and the division algorithm 
-    %terminates when the leading term of the divisor does not divide any 
-    %other term in the dividend.
-
-    For sets of polynomials, the notation 
-    \begin{equation*}
-    f\stackrel{F}{\textstyle\longrightarrow}_+r    
-    \end{equation*}
-    represents the reduced polynomial $r$ resulting from $f$ as reduced by a 
-    set of non-zero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
-    $r = 0$  or no term in $r$ is divisible  by a $lm(f_i), \forall f_i \in F$.
-\end{Definition}
-
-The reduction process $f\stackrel{F}{\textstyle\longrightarrow}_+r$, of 
-dividing a polynomial $f$ by a set of polynomials of $F$, can be modeled as
-repeated long-division of $f$ by each of the polynomials in $F$ until no
-further reductions can be made. The result of this process is then $r$.
-This reduction process is shown in Algorithm \ref{alg:polydiv}.
-
-\begin{algorithm}[H]
-\SetAlgoNoLine
-
- \KwIn{$f,f_{1},\dots,f_{s}$}
- \KwOut{$r,a_{1},\dots,a_{s}$, such that $f=a_{1}\cdot f_{1}+\dots+a_{s}\cdot f_{s}+r$.}
-
- $a_{1}=a_{2}=\dots=a_{s}=0$; $r=0$\;
- $p:=f$\;
-
- \While { $p \neq 0$ }
- {
-	i=1\;
-	divisionmark = false\;
-	\While { $i\le s  $ \&\& divisionmark = false }
-	{
-		\eIf {$f_{i}$ can divide $p$}
-		{
-			$a_{i}=a_{i}+lt(p)/lt(f_{i})$\;
-			$p=p-lt(p)/lt(f_{i}) \cdot f_{i}$\;
-			divisionmark = true\;
-		}
-		{
-			i=i+1\;
-		}
-	}
-
-	\If {divisionmark = false}
-	{
-		$r=r+lt(p)$\;
-		$p=p-lt(p)$\;
-	}
-
- }
-\caption{Polynomial Reduction}\label{alg:polydiv}
-\end{algorithm}
-
-The reduction algorithm keeps canceling the leading terms of polynomials 
-until no more leading terms can be further canceled.
-So the key step is $p=p-lt(p)/lt(f_{i}) \cdot f_{i}$, as the following 
-example shows.
-\begin{Example}
-Given $f = y^{2}-x$ and $f_{1} = y - x$ in $\mathbb{Q}[x,y]$ with $deglex$: 
-$y>x$, perform $f\stackrel{f_1}{\textstyle\longrightarrow}_+r$:
-
-\begin{enumerate}
-\item $f=y^{2}-x$, $f/f_{1}=f-lt(f)/lt(f_{1}) \cdot f_{1}=y^{2}-x-(y^{2} /y) \cdot (y-x)=y\cdot x-x$
-\item $f=y\cdot x-x$, $f/f_{1}=f-lt(f)/lt(f_{1}) \cdot f_{1}=(y\cdot x-x)/f_{1}=x^{2}-x$
-\item $f=x^{2}-x$, no more operations possible, so $r=x^{2}-x$
-\end{enumerate}
-\end{Example}
 
@@
 
-\section{Elimination Theory}
+\section{Elimination and Abstraction Theory}
+\label{sec:abstraction}
+\subsection{Elimination Theory}
 
@@
 
-Elimination theory provides the basis for our abstraction approach.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Hilbert's Nullstellensatz}
-
-In this section, we further describe some correspondence between ideals and 
-varieties in the context of algebraic geometry. The celebrated results of 
-Hilbert's Nullstellensatz establish these correspondences.
-
-%%%%%%%%%%%%%%%%%algebraically closed field%%%%%%%%%%%%%
-\begin{Definition}\label{def:acf}
-A field $\overline {\K}$ is an {\bf algebraically closed} field if every  
-polynomial in one variable with degree at least $1$, with coefficients 
-in $\overline {\K}$, has a root in $\overline {\K}$. 
-\end{Definition}
-In other words, any non-constant polynomial equation over 
-$\overline {\K}\left[x\right]$ always has at least one root 
-in $\overline {\K}$. Every field $\K$ is contained in an algebraically 
-closed one $\overline {\K}$. 
-For example, the field of real numbers $\mathbb{R}$ is not an algebraically closed 
-field, because $x^2+1=0$ has no root in $\mathbb{R}$. 
-However, $x^2+1=0$ has roots in the field of 
-complex numbers $\mathbb{C}$, which is an algebraically closed field. 
-In fact, $\mathbb{C}$ is the algebra closure of $\mathbb{R}$. 
-Every algebraically closed field is an infinite field.
-
-%%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
-%\begin{Theorem}
-%$\left[\bf{Weak\  Nullstellensatz}\right]$ Let 
-%$I \subset \overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$ 
-%be an ideal satisfying $V(I)=\emptyset$. 
-%Then $I=\overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
-%\begin{equation}
-%V(I)=\emptyset\ \iff\ I=\overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
-%\end{equation}
-%\end{Theorem}
-%
-%\begin{Corollary}
-%	Let $I=\langle f_{1},\dots,f_{s} \rangle \subset \overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$. 
-%	Let $G$ be the reduced Gr\"obner basis of $I$. Then $V(I)=0 \iff G=\{1\}$.
-%\end{Corollary}
-
-%{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
-%system of multivariate polynomial equations (ideal $I$) has common solutions 
-%in ${\overline {\mathbb{K}}}^d$. For this purpose, we only need to check if 
-%the ideal is generated by the unit element, i.e., $1\in I$. 
-%This approach can be used to evaluate the feasibility of constraints in 
-%verification problems.
-%%%%%%%%%%%%%%%%%%%%%%%%%strong Nullstellensatz%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-An interesting result is one of {\bf Strong 
-Nullstellensatz}.
-The strong
-Nullstellensatz establishes the correspondence between radical ideals
-and varieties.
-
-\begin{Theorem}\label{thm:sns}
-({\it The Strong Nullstellensatz} \cite{gb_book}) 
-Let $\overline{\mathbb{K}}$ be an algebraically closed field, and let $J$
-be an ideal in $\overline{\mathbb{K}}[x_1,\dots, x_d]$. 
-Then we have $I(V_{\overline{\mathbb{K}}}(J)) =\sqrt{J}$. 
+Elimination theory provides the basis for following abstraction approach.
+\subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
+Let $S$ be the system of 
+polynomials, 
+$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
+derived from the hardware
+implementation of the Galois field arithmetic circuit over $\Fkk$.
+This circuit performs some unknown function $f$ over 
+$\Fkk$ in the form of $Z=\Func(A_1,\dots,A_n)$, where $Z$ is the $k$-bit 
+output and $A_1,\dots,A_n$ are the $k$-bit inputs.
+The polynomial representation of $\Func$ over $\Fkk$ is thus:
+
+\begin{equation}
+f_{\Func}: Z+\Func(A_1,\dots,A_n) \nonumber
+\end{equation}
+
+Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
+it agrees with the solution to the system of polynomials $\{S\}=0$, i.e.:
+\begin{equation}
+f_1=\dots=f_s=f_{A_1}=\dots=f_{A_n}=f_{Z}=0 \nonumber
+\end{equation}
+Thus, if we let $J=\langle f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\rangle$ 
+be the ideal generated by $S$, 
+$f_{\Func}$ {\bf vanishes} on the variety $V_{\Fkk}(J)$. 
+Therefore, due to Proposition \ref{pro:iofv}, 
+$f_{\Func}$ must be contained in the ideal of polynomials that vanish on
+this variety, $f_{\Func} \in I(V_{\Fkk}(J))$.
+
+By applying Strong Nullstellensatz over $\Fkk$ (Theorem \ref{thm:sns}), 
+$I(V_{\Fkk}(J))=J+J_0$ where
+$J_0$ is the ideal generated by all vanishing polynomials in $\Fkk$. 
+Recall that a vanishing polynomial in $\Fkk[x]$ is $x^q-x=x^q+x$. 
+In our case, 
+$\{x_1,\dots,x_d\} \in \F_2$ and $\{A_1,\dots,A_n,Z\} \in \Fkk$.
+Thus, for $\Fkk[x_1,\dots,x_d,A_1,\dots,A_n,Z]$: 
+\begin{equation}
+J_0 = \langle x_1^2+x_1,\dots,x_d^2+x_d, A_1^{2^k}+A_1,\dots,A_n^{2^k}+A_n,Z^{2^k}+Z\rangle \nonumber
+\end{equation}
+
+The generators of the ideal sum $J+J_0$ are simply the combination of the 
+generators of $J$ and the generators $J_0$.
+
+The variety $V_{\Fq}(J)$ is
+the set of all consistent assignments to the nets (signals) in the
+circuit $C$. If we {\it project this variety on the word-level input and
+output variables of the circuit $C$, we essentially generate the
+function $\Func$ implemented by the circuit.} Projection of varieties from
+$d$-dimensional space $\Fq^d$ onto a lower dimensional subspace
+$\Fq^{d-l}$ is equivalent to {\it eliminating $l$ variables} from the
+corresponding ideal. This can be done by computing a \Grobner basis
+of the ideal with elimination ordering, as described in the 
+Elimination Theorem (Theorem \ref{thm:elimth}).
+Thus, we can find the polynomial 
+$f_\Func:Z+\Func(A_1,\dots,A_n)$ by computing the \Grobner
+basis of $J+J_0$ using the proper elimination ordering.
+
+The proposed elimination order for 
+abstraction is defined as the {\bf abstraction term order} (ATO). 
+\begin{Definition}
+\label{def:ato}
+Given a circuit $C$,
+let $x_1, \dots, x_d$ denote all the bit-level variables, 
+let $A_1,\dots,A_n$ denote the $k$-bit word-level inputs, 
+and let $Z$ denote the $k$-bit word-level output. 
+Using the partial variable order 
+$\{x_1, \dots, x_d\} > Z > \{A_1, \dots, A_n\}$,
+where any refinement of the order will do,
+impose a lex term order $>$ on the polynomial ring 
+$R = \Fq[x_1, \dots, x_d, Z, A_1, \dots, A_n]$. 
+This elimination term order $>$ is defined as
+the {\bf Abstraction Term Order}. The relative ordering among $x_1,
+\dots, x_d$ is not important and can be chosen arbitrarily. Likewise,
+the relative ordering among $A_1, \dots, A_n$ is also unimportant.
+\end{Definition}
+
+\begin{Theorem}
+{\bf Abstraction Theorem:} Using the setup and notations above,
+compute a \Grobner basis $G$ of ideal $(J+J_0)$ using the abstraction term 
+order $>$. Then: \\
+(i) For every word-level input $A_i$, $G$ must contain the vanishing 
+polynomial $A_i^q - A_i$ as the only polynomial with $A_i$ as its only 
+variable;\\
+(ii) $G$ must contain a polynomial of the form 
+$Z + \mathcal{G}(A_1,\dots,A_n)$; and\\ 
+(iii) $Z + \mathcal{G}(A_1,\dots,A_n)$ is such that 
+$\Func(A_1,\dots,A_n) = {\mathcal{G}}(A_1,\dots,A_n),
+\forall A_1,\dots,A_n \in \Fq$. 
+In other words, ${\mathcal{G}}(A_1,\dots,A_n)$ and $\Func(A_1,\dots,A_n)$ are
+equal as polynomial functions over $\Fq$.
+\label{thm:abs}
 \end{Theorem}
 
-%\subsection{Nullstellensatz over Galois fields}
-
-Strong Nullstellensatz holds a special form over Galois fields $\Fq$.
-Recall the notion of vanishing polynomials over Galois fields from the 
-previous chapter: for every element $A \in \Fq$, $A-A^q = 0$; then the 
-polynomial $x^q-x$ in $\Fq[x]$ vanishes over $\Fq$. Thus, if 
-$J_0=\langle x^q-x \rangle$ is the ideal generated by the vanishing 
-polynomial, $V(J_0)=\Fq$. Similarly, over $\Fq[x_1,\dots,x_d]$, $J_0$ is 
-$\langle x_1^q-x_1,\dots,x_d^q-x_d \rangle$ and $V(J_0)=(F_q)^d$.
-
-\begin{Definition}
-Given two ideals, $I_1=\langle f_1, \dots,f_s \rangle$ and 
-$I_2=\langle g_1,\dots g_t\rangle$, then the {\bf sum of ideals} 
-$I_1+I_2=\langle f_1,\dots,f_s,g_1,\dots g_t\rangle$
-\end{Definition}
-
-\begin{Theorem}
-({\it Strong Nullstellensatz over $\Fq$})
-For any Galois field $\Fq$, let $J \subset \Fq[x_1,\dots,x_d]$ be any ideal 
-and let $J_0 = \langle x_1^q-x_1, x_d^q-x_d \rangle$ be the ideal of all
-vanishing polynomials. Let $V_{\Fq}(J)$ denote the variety of $J$ over $\Fq$.
-Then, $I(V_{\Fq}(J))=J+J_0$.
-\end{Theorem}
-
-The proof is given in \cite{gao:gf-gb-ms}. Here, we provide a proof outline.
-
-\begin{Proof}
-\begin{enumerate}
-\item $\sqrt{J+J_0} = J+J_0$. That is, $J+J_0$ is a radical ideal.
-\item $V_{\Fq}(J)=V_{\overline{\Fq}}(J+J_0)$.
-\item Due to (2), $I(V_{\Fq}(J)) = I(V_{\overline{\Fq}}(J+J_0))$. 
-By Strong Nullstellensatz, this is equivalent to $\sqrt{J+J_0}$.
-Finally, due to (1), this is equivalent to $J+J_0$.
-\end{enumerate}
-\end{Proof}
-
-%% 
-%% Using this result, Weak Nullstellensatz can be 
-%% modified to be applicable over finite fields $\Fq$.
-%% %%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
-%% \begin{Theorem}\label{wnull:ff}
-%% $[\bf{Weak~Nullstellensatz~in~\Fq}]$\cite{null:1890}\\
-%% Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
-%% Let $J=\langle f_1,f_2,\cdots,f_s\rangle \subset \Fq[x_1,
-%% x_2, \cdots, x_d]$ be an ideal. Let $J_0 = \langle 
-%% x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
-%% of vanishing polynomials in $\Fq$. Then
-%% $V_{\Fq}(J) = V_{\overline {\Fq}}(J +
-%% J_0)=\emptyset$,  if and only if the reduced
-%% Gr\"obnerBasis$(J+J_{0})=\{1\}$. 
-%% \end{Theorem}
-
-%% The proof is given in \cite{null:1890}. Here, we provide a proof outline.
-
-%% \begin{Proof}
-%% The variety of $J$ over $\Fq[x_1,x_2,\cdots,x_d]$ 
-%% is equivalent to the variety over the algebraic closure of $\Fq$ 
-%% intersected by the entire field $\Fq$. That is, $V_{\Fq}(J)=V_{\overline 
-%% {\Fq}}(J) \cap \Fq$.
-
-%% Let $J_0 = \langle 
-%% x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
-%% generated by all vanishing polynomials in $\Fq[x_1,x_2,\cdots,x_d]$.
-%% Then $V_{\overline{\Fq}}(J_0)=\Fq$.
-
-%% Thus, $V_{\Fq}(J) = V_{\overline{\Fq}}(J)\cap V_{\overline{\Fq}}(J_0)
-%% = V_{\overline{\Fq}}(J+J_0)$.
-%% \end{Proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Concluding Remarks}
-
+\begin{Corollary}
+By computing a {\bf reduced} \Grobner basis $G_r$ of $J + J_0$, 
+$G_r$ will contain one and only one polynomial in of the form 
+$Z + \G(A_1,\dots,A_n)$, such that $Z = \G(A_1,\dots,A_n)$ is the 
+{\bf unique, minimal, canonical} representation of the function 
+$\Func$ implemented by the circuit.  
+\end{Corollary}
+
+\begin{Example}
+\label{exp:finalMul2Bit}
+Consider a 2-bit multiplier over ${\mathbb{F}}_{2^2}$ with
+$P(x)=x^{2}+x+1$, given in Fig. \ref{fig:2bitmul}. Variables $a_0,
+a_1, b_0, b_1$ are primary inputs, $z_0, z_1$ are primary outputs, and
+$c_0, c_1, c_2, c_3, r_0$ are intermediate variables. 
+%The gate $\otimes$
+%corresponds to AND-gate, i.e. bit-level multiplication modulo 2. The
+%gate $\oplus$ corresponds to XOR-gate, i.e. addition modulo 2.
+
+\begin{figure}[H]
+\centerline{
+\includegraphics[scale=0.5]{./figures/2bitmasmult.eps}
+}
+\caption{ A 2-bit multiplier over ${\mathbb{F}}_{2^2}$.}
+\label{fig:2bitmul}
+\end{figure}
+
+Polynomials extracted from the circuit implementation represent the
+ideal $J$. Along with the ideal of vanishing polynomials $J_0$, 
+the following polynomials represent the generators 
+of $J+J_0$ for the multiplier circuit.
+
+\begin{eqnarray}
+ \left .  
+	\begin{aligned}
+		f_1:c_0+a_0 \cdot b_0  \\
+		f_2:c_1+a_0 \cdot b_1  \\
+		f_3:c_2+a_1 \cdot b_0  \\
+		f_4:c_3+a_1 \cdot b_1  \\
+		f_5:r_0+c_1 + c_2		\\
+		f_6:z_0+c_0 + c_3		\\
+		f_7:z_1+r_0 + c_3		
+	\end{aligned} 
+ \ \right\}
+ &\qquad&  {\it  \text{Bit-level circuit constraints} ~(\subset J)} \nonumber \\
+ \left . 
+	\begin{aligned}
+		f_{A}:A+a_0+a_1\cdot \alpha   \\ 
+		f_{B}:B+b_0+b_1\cdot \alpha  \\ 
+		f_{Z}:Z+z_0+z_1\cdot \alpha   
+	\end{aligned} 
+ \right\}
+ &\qquad&  {\it  \text{Word-level designation} ~(\subset J)} \nonumber \\
+  \left . 
+	\begin{aligned}
+		a_0^2-a_0, ~a_1^2-a_1,~b_0^2-b_0, ~b_1^2-b_1   \\ 
+		c_0^2-c_0, ~c_1^2-c_1,~c_2^2-c_2, ~c_3^2-c_3  \\ 
+		r_0^2-r_0, ~z_0^2-z_0,~z_1^2-z_1    \\ 
+		A^4-A, ~B^4-B ,~Z^4-Z		  
+	\end{aligned} 
+ \right\}
+ &\qquad&  {\it \text{vanishing polynomials} (J_0)} \nonumber
+\end{eqnarray}
+We apply abstraction term order $>$, i.e a lex order with
+"bit-level variables" $>$ "Output Z" $>$ "Inputs A, B".
+
+When we compute the reduced \Grobner basis, $G_r$, of \{$J + J_0$\} with 
+respect to this ordering, $G_r$ = \{$g_1, \dots, g_{14}\}:$
+\begin{align*}
+g_1: B^4+B; 
+~~g_2: b_0+b_1 \alpha + B; 
+~~g_3: a_0+a_1 \alpha + A;  \nonumber \\
+~~g_4: c_0+c_1 \alpha + c_2 \alpha + c_3(\alpha+1)+Z;
+g_5: r_0+c_1+c_2; 
+~~g_6: z_0+c_0+c_3; \nonumber \\
+~~g_7: z_1+r_0+c_3; 
+~~{\bf g_8: Z+A\cdot B};
+~~g_9: b_1+B^2+B; 
+~~g_{10}: a_1+A^2+A; \nonumber \\
+~~g_{11}: c_3+a_1\cdot b_1
+g_{12}: c_2+a_1\cdot b_1 \alpha + a_1 \cdot B; 
+~~g_{13}: c_1+a_1\cdot b_1 \alpha +b_1 A; 
+~~g_{14}: A^4+A  \nonumber
+\end{align*}
+
+$g_8=Z+A\cdot B$ is the {\bf canonical, word-level polynomial } 
+representing the function performed by the multiplier $Z=A\cdot B$.
+\end{Example}
+
+\section{Conclusion Remarks}
 Our approach to word-level abstraction of Galois field arithmetic 
 circuits applies concepts of polynomial ideals, varieties, \Grobner basis, 
-and elimination theory to abstract a word-level representation of the 
-circuit. This approach is described in the next chapter. However, a \Grobner
+and abstraction theory to implement verifications on sequential circuits. 
+These approaches are described in the following chapters. However, a \Grobner
 basis computation is prohibitively expensive; thus we propose improvements 
 to our original approach in a subsequent chapter.
-
-%Once we have a word-level representation, we can apply these representations
-%to perform equivalence checking of Galois field arithmetic circuits.
-%The verification problem is formulated using Weak
-%Nullstellensatz, as it applies over Galois fields, and subsequently solved 
-%using a \Grobner basis approach. This approach is described in a future 
-%chapter.

### 2016-10-24 dissertation/prev.tex
dissertation: milestone 1: first draft

\chapter{Previous Work}
\label{ch:prev}
\section{SAT,DD,AIG-based Sequential Circuit Verification}

\section{Bounded Model Checking}

\section{Model Checking with Abstraction Refinement}

\section{Word-level Techniques in Sequential Circuit Verification}
Term rewriting?

\section{Verification of Galois Field Circuits}

\section{Conventional UNSAT Core Extraction}

\section{Conclusion Remarks}

### 2016-10-24 dissertation/prev_tim.tex
dissertation: milestone 1: first draft

\chapter{Previous Work} \label{ch:prev}

This chapter covers previous work in the area of canonical
representations of functions, word-level abstractions and their
application to design verification. Since the application of our
approach is targeted towards formal equivalence verification, modern
combinational equivalence checking techniques are also
reviewed. Finally, formal verification techniques using computer
algebra, algebraic geometry, and polynomial interpolation are also
considered.

\section{Canonical Decision Diagrams}
Canonical representations of Boolean functions have been the subject
of extensive investigation for logic synthesis and design
verification. The Reduced Ordered Binary Decision Diagram (ROBBD)
\cite{BRYA86} was the first significant contribution in this area. 
Efficient implementation of ROBDDs as a software package \cite{brace}
allowed for efficient formal verification of combinational and
sequential circuits.  ROBDDs represent a Boolean function as an
implicit set of points on a canonical directed acyclic graph
(DAG). Manipulation of Boolean functions can then be carried out as
composition operations on their respective DAGs.  The decomposition
principle behind BDDs is one of Shannon's expansion, i.e.
\begin{equation}
f(x, y, \dots) = x f_x + x' f_{x'}
\end{equation}

where $f_x = f(x = 1)$ and $f_{x'} = f(x = 0)$ denote the positive and
negative co-factors of $f$ w.r.t. $x$, respectively. Motivated by the
success of BDDs,  variants of the Shannon's decomposition principle
(Davio, Reed-Muller, etc.) were explored to develop other functional
decision diagrams. For example, the AND-OR-NOT logic based Shannon's
expansion is transformed into an AND-XOR logic based decomposition,
termed as the Davio's decomposition:

\begin{eqnarray}
f(x, y, \dots) &=& x f_x + x' f_{x'}\\
& = & x f_x \oplus x' f_{x'}\\
& = & x f_x \oplus (1 \oplus x) f_{x'}\\
& = & f_{x'} \oplus x (f_x \oplus f_{x'})
\end{eqnarray}

Decision diagrams based on such decompositions include FDDs
\cite{okfdd}, ADDs \cite{add}, MTBDDs \cite{mtbdd}, and their hybrid 
edge-valued counterparts, HDDs \cite{hdd} and EVBDDs \cite{evbdd}. 
While these are referred to as {\it Word-Level Decision Diagrams}
\cite{WLS}, the decomposition is still point-wise, binary, 
w.r.t. each Boolean variable. These representations do not
serve the purpose of word-level abstraction from bit-level
representations.

Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivatives K*BMDs
\cite{kbmd} and *PHDDs \cite{phdd}, depart from the Boolean
decomposition and perform the decomposition of a {\it linear} function
based on its two moments. BMDs provide a compact representation for
integer arithmetic circuits such as multipliers and squarers. However,
these are inapplicable to word-level abstraction of modulo-arithmetic
circuits over Galois fields. 
%in our application, we encounter modulo-arithmetic circuits and that
%too over Galois fields. For such applications, these representations
%are inapplicable. Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp}
%generalize the linear moment decomposition of BMDs into a Taylor
%series expansion --- this allows the representation of RTL
%descriptions as polynomials over bit-vectors. However,

Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp} are a
word-level canonical representation of a {\it polynomial expression},
based on the Taylor's series expansion of a polynomial. However, they do
not represent a {\it polynomial function} canonically. For example,
$f_1 = 0$ and $f_2 = 2x^2 - 2x \pmod{ 4}$ are two different polynomial
representations of the zero function over $\Z_4$; but they are
symbolically different polynomials and they have non-isomorphic TED
DAGs.  While \cite{namrata:phd}  and \cite{alizadeh:tcad2010} provide 
canonical representations of polynomial functions, they do so over
finite integer rings $\Z_{2^k}$ and not over Galois fields $\Fkk$.

MODDs \cite{modd} \cite{modd_tcomp} are a DAG representation of the
characteristic function of a circuit over Galois fields $\Fkk$. MODDs
come very close to satisfying our requirements as a canonical
word-level representation that can be employed over Galois fields, as
it essentially interpolates a polynomial from the characteristic 
function. However, MODDs do not scale very well for large circuits ---
this is because every node in the DAG can have up to $k$ children and
the normalization operations are very complicated for MODDs. They also
suffer from the size explosion problem during intermediate
computations. They are known to be infeasible in representing
functions over $32$-bit operand words.

\section{Word-Level Techniques in RTL Synthesis and Verification}

Other attempts to derive high-level representations of functions,
along with associated decision procedures, can be found in
the rich domain of formal model checking \cite{BHEL96} \cite{SMV},
theorem proving \cite{arditi:bmd}, bit-vector SMT-solvers
\cite{boolector} \cite{cvc3} \cite{z3} \cite{bitvector98}, automated
decision procedures for Presburger arithmetic \cite{presburger}
\cite{bultan:mixed_verification}, algebraic manipulation techniques 
\cite{devadas:algebraic_manipulation_iccd91}, or the ones based on
term re-writing \cite{AST}, etc. 
%Other forms of high-level logics,
%such as EUF \cite{Burch:CAV94}, PEUFM \cite{Velev:JSC03}
%\cite{Velev:CL01}, CLU \cite{Lahiri:CAV02} have been derived for
%verifying micro-processor pipeline control. 
Polynomial, integer and other non-linear representations have also
been researched: Difference Decision Diagrams (DDDs) \cite{ddd-csl99} \cite{ddd-mt-98}, interval
diagrams \cite{interval_dd}, interval analysis using polynomials
\cite{polynomial_sanchez99}, etc. Most of these have found 
application in constraint satisfaction for simulation-based
validation:  \cite{Ritter99} \cite{hsat} \cite{lpsat}
\cite{brinkmann:asp-dac} \cite{Huang:tcad01} \cite{bitvector98}. Among
these, \cite{brinkmann:asp-dac} \cite{Huang:tcad01} \cite{bitvector98}
have been used to {\it solve} integer modular arithmetic on linear
expressions - a different application from {\it representing}
finite field modulo-arithmetic on polynomials in a canonical form.

\section{Combinational Equivalence Checking}

The verification problem addressed in this dissertation is a
manifestation of the combinational equivalence checking (CEC) problem,
where the specification (polynomial) and  the implementation (circuit)
are custom-designed, structurally very dissimilar circuits. To make
use of contemporary gate-level CEC tools, we can take the 
specification circuit (``golden model'') and check its equivalence
against the implementation circuit. Canonical decision diagrams (BDDs
\cite{BRYA86} and their word-level variants \cite{WLS}),
And-Invert-Graph (AIG) based reductions \cite{AIG:2002}
\cite{alanmi:cec:iccad2006}, circuit-SAT solvers \cite{csat}, etc.,
are among the many techniques that can be employed for this CEC.  
When one circuit is synthesized from the 
other, this problem can be efficiently solved using AIG-based
reductions (e.g. the ABC tool \cite{abc}) and circuit-SAT solvers
(e.g., CSAT \cite{csat}). 
Synthesized circuits generally contain many
sub-circuit equivalences which AIG and CSAT based tools can identify
and exploit for verification. However, when the circuits are
functionally equivalent but structurally very dissimilar (e.g.,
Mastrovito \cite{mastro:1989} versus Montgomery implementations
\cite{acar:1998} of Galois field circuits), none of the 
  contemporary techniques, including ABC and CSAT, offer a practical
  solution. Automatic formal verification of large {\it
    custom-designed modulo-arithmetic circuits} largely remains
  unsolved today.

This verification problem is very hard for SAT solvers and also for
quantifier-free bit-vector (QF-BV) theory based SMT-solvers, due to
the large circuit size, and the presence of AND-XOR-SHIFT structures. 
Similarly, the Cryptol tool-set \cite{Cryptol:fmcad09} also employs
AIG-based reductions (SAT-sweeping) and SAT/SMT-solvers for
verification of crypto-protocols. For applications where
AIGs/SAT/SMT-techniques fail, the {\it Cryptol} tool-set also does not
deliver. As shown in \cite{lv:phd}, {\it none of BDDs, SAT, SMT
  solvers, nor the  ABC tool can prove design equivalence beyond
  $16$-bit circuits.}

%%%%%%%%%%MACIEJ HERE%%%%%%%%%%%%%%%
In \cite{ciesielski:flow}, the authors present a method
for verification of integer multipliers using a data-flow
approach. This work abstracts a polynomial function
of the given multiplier and then solves the network flow problem 
using algebraic techniques. However, the abstraction is 
solely bit-level, and is thus not applicable to deriving 
a word-level representation of a given design.

\section{Verification of Galois Field Circuits}

Symbolic computer algebra techniques have been employed for formal
verification of circuits over $\Z_{2^k}$ and also over
Galois fields $\Fkk$. 
The work of \cite{gao:gf-gb-ms} shows how to use Gr\"obner basis
techniques to count the zeros of an ideal $J$ over ${\mathbb{F}}_q$
(i.e. count $V_{{\mathbb{F}}_q}(J)$). The authors then follow-up with
an approach for {\it quantifier elimination} over Galois fields
${\mathbb{F}}_q$ \cite{gao:qe-gf-gb}. 
However, computing a \Grobner basis is computationally
expensive.
While these works present
the proper theory and algorithms, 
efficiency/improvements to the Gr\"obner
basis computation is not addressed. This is also the case with other
general verification techniques using Gr\"obner bases
\cite{Avrunin:CAV} \cite{gbverify:2007} \cite{manna:program}, etc. 
%The paper \cite{wienand:cav08} addresses verification of %-TIM
%finite-precision integer datapath circuits using the concepts of %-TIM
%Gr\"obner bases over the ring ${\mathbb{Z}}_{2^k}$. %-TIM
%They model the
%circuit constraints by way of arithmetic-bit-level (ABL) polynomials
%($\{G\}$) and formulate the verification test as an equivalent
%{\it variety subset problem}. To solve this, first they derive a term
%order to represent the circuit polynomials $\{G\}$ with a Gr\"obner
%basis. Then they compute a normal form $f$ of the specification $g$
%w.r.t. $\{G\}$. They show that, under certain constraints that are
%satisfied by their term order, the circuit is correct if and only if
%$f$ is a vanishing polynomial over ${\mathbb{Z}}_{2^k}$
%\cite{shekhar:tcad07}. 
%In \cite{wedler:date11}, the authors further                 %-TIM
%show an integration of their approach within an SMT-solver for %-TIM
%arithmetic instances.                        %-TIM
%show that  the vanishing polynomial test can be
%omitted by formulating the problem directly over $Q :=
%{\mathbb{Z}}_{2^k} [X]/\langle x^2-x : x \in X \rangle$.

In \cite{ibm:ecc-ita11} \cite{blueveri:ita} \cite{ibm:blueveri},
%from IBM needs special mention: T
the authors present the {\sc BLUEVERI} tool from IBM for verification
of Galois field circuits for error correcting codes against an
algorithmic spec. The implementation consists of a set of
(pre-designed and verified) circuit blocks that are interconnected to
form the error correcting system. The spec is given as a set of design
constraints on a ``check file''. Their objective is to prove the 
equivalence of the implementation against this check file.
They model the verification instance as a data-flow graph, represent each
sub-circuit block with its known (word-level) polynomial over $\Fq$,
and formulate the verification problem using the {\it Weak
  Nullstellensatz} --- i.e. to check if the {\it variety} of the
algebraic system ``{\it spec} $\neq$ {\it implementation}'' is empty
for which they employ a Nullstellensatz formulation.
%for which they use a Gr\"obner basis engine. 
Their main
contributions are: i) a ``term re-writing'' to specify the algorithmic
description using polynomials (ideal); and ii) integrating an
AIG-style \cite{AIG:2002} Boolean solver with their word-level
decision procedure, with lazy signal computations and Boolean
reasoning. For final verification, the polynomial system is given
to a computer algebra tool ({\sc Singular} \cite{DGPS}) to {\it
  compute} a reduced Gr\"obner basis. However, improvements to the
core Gr\"obner basis computational engine are not the subject of their
work.

In \cite{lv:phd} 
\cite{lv:tcad2013}
\cite{lv:date2012}
\cite{lv:hldvt2011} \cite{lv:vlsi2012}, {\it Lv et al.} present
computer algebra techniques for formal verification of Galois field
arithmetic circuits. Given a specification polynomial $f$, and a
circuit $C$, they formulate the verification problems as an ideal
membership test using the Strong Nullstellensatz and Gr\"obner
bases. In \cite{lv:date2012}, the authors show that for any
combinational circuit, {\it there exists a term order $>_1$ that
  renders the set of polynomials of the circuit itself a Gr\"obner
  basis --- and this term order can be easily derived 
by performing a topological traversal of the circuit}. By exploiting
this term order, verification can be significantly scaled
to $163$-bit (NIST-specified) cryptography circuits. 
In contrast to the work of \cite{lv:phd}, we are not given a
specification polynomial. Instead, given the circuit $C$, we want
to derive (extract) the word-level specification $f$. In our work, we
borrow and further build upon the results of \cite{gao:gf-gb-ms}
\cite{gao:qe-gf-gb} %\cite{modd_tcomp} 
\cite{lv:date2012} \cite{lv:phd}.

\section{Verification of Integer Arithmetic Circuits using Gr\"obner Bases}
Symbolic computer algebra techniques have been used for verification
of integer arithmetic circuits \cite{wienand:cav08} and also for
decision procedures over Galois fields \cite{gao:gf-gb-ms}. The paper
\cite{wienand:cav08} addresses verification of finite precision
integer datapath circuits using the concepts of  Gr\"obner bases over
the ring ${\mathbb{Z}}_{2^k}$. This work models the circuit constraints by
way of arithmetic-bit-level (ABL) polynomials ($\{G\}$), and formulates
the verification test as an equivalent variety subset problem. This problem
is solved by deriving a term order that already makes $\{G\}$
a Gr\"obner basis, then computing a normal form $f$ of the
specification $g$ w.r.t. $\{G\}$. Circuit correctness is established
by testing whether or not $f$ is a vanishing
polynomial over ${\mathbb{Z}}_{2^k}$ \cite{shekhar:tcad07}.
In \cite{wedler:date11}, the
authors further show that the vanishing polynomial test can be omitted
by formulating the problem directly over $Q := {\mathbb{Z}}_{2^k}
[X]/\langle x^2-x : x \in X \rangle$.  
However, in these works, the problem requires that the word-level 
abstraction of the circuit be known, whereas our approach derives 
this abstraction polynomial.

\section{Polynomial Interpolation in Symbolic Computation}

The problem of polynomial interpolation is a fundamental problem in
symbolic and algebraic computing which finds application in modular
algorithms, such as the GCD computation and polynomial
factorization. The problem is stated as follows: Given $n$ distinct
data points $x_1, \dots, x_n$, and their evaluations at these points
$y_1, \dots, y_n$,  {\it interpolate} a polynomial $\Func(X)$ of degree
$n-1$ (or less) such that $\Func(x_i) = y_i$ for $1 \leq i \leq n$. 
Let $t$ be the number of non-zero terms in $\Func$ and let $T$ be the
total number of possible terms. When ${{t} \over {T}} << 1$, the
polynomial $\Func$ is {\it sparse}, otherwise it is {\it dense}. Much of
the work in polynomial interpolation addresses sparse 
interpolation using the ``black-box'' model (also called the
algebraic circuit model) as shown in Fig. \ref{fig:blackbox}.

\begin{figure}[hbt]
\centerline{
%\includegraphics[scale=0.3]{../figures/2bitmultiplier.eps}
\includegraphics[scale=0.45]{figures/blackbox.eps}
}
\caption{The black-box or the algebraic circuit representation.}
\label{fig:blackbox}
\end{figure}

Let $\Func$ be a multivariate polynomial in $n$ variables $\{x_1, \dots,
x_n\}$, with $t$ non-zero terms ($0 < t < T$), represented with a
black-box $B$. On input $(x_1, \dots, x_n)$,
the black-box evaluates $y_i = \Func(x_1, \dots, x_n)$. Given also a
degree bound $d$ on $\Func$, the goal is to interpolate the polynomial
$\Func$ with a minimum number of {\it probes} to the black-box. The early
work of Zippel \cite{zippel:interpolate} and
Ben-Or/Tiwari \cite{ben-or-tiwari:interpolate} require $O(ndt)$ and
$O(T \log n)$ probes, respectively, to the black-box. These bounds
have since been improved significantly; the recent algorithm of
\cite{monagan:interpolate} interpolates with $O(nt)$ probes.

Our problem of polynomial abstractions of Galois field circuits 
falls into the category of dense interpolation, as we
require a polynomial that describes the function at each of the $q$
points of the field $\Fq$. Newton's interpolation technique, with the
black-box model, bounds the number of probes by $(d+1)^n$ --- which 
exhibits very high complexity. In the logic synthesis area, the work of
\cite{zilic:interpolate} investigates dense interpolation. Due to this
high-complexity, their approach is feasible only for applications over
small fields, {\it e.g.} computing  Reed-Muller forms for multi-valued
logic over $\F_2$.

For our problem,
we can also employ the black-box model by replacing the black-box
(algebraic circuit) by the given circuit $C$; then every {\it probe}
of the black-box would correspond to a {\it simulation of the
  circuit}. However, as we desire a polynomial representation of the entire
function over the Galois field, exhaustive simulation would be required, 
which is infeasible.

\section{Concluding Remarks}

For the problem of word-level, canonical, polynomial abstractions
of Galois field arithmetic circuits over $\Fkk$, previous related work 
is either inapplicable or only applicable to circuits no larger than
$32$-bits in size.
Therefore, we propose a {\it symbolic approach} to
polynomial interpolation from a circuit using the Gr\"obner basis
computation. However, the complexity of a \Grobner basis computation is 
prohibitively expensive; thus, we propose further improvements to this 
approach by deriving a smaller subset of computations based on a \Grobner
basis analysis. These improvements allow for abstractions of 
flattened Galois field circuits up to $571$-bits, which is the largest NIST 
standard for ECC, or up to $1024$-bits when a hierarchy is given. 
Furthermore, we propose applications of this 
approach to allow for formal verification of flattened Galois field circuits up to 
$1024$-bits, where current techniques are only applicable for circuits up 
to $163$-bits.

### 2016-10-24 dissertation/unsat.tex
dissertation: milestone 1: first draft

\chapter{Finding Unsatisfiable Cores Extraction for a Set of Polynomials using the Gr\"obner Basis Algorithm}
\label{ch:UNSAT}
Besides elimination ideal based abstraction, Gr\"obner basis can also be applied to a vast field
about other techniques in formal verification. Satisfiability (SAT) problem is the basis of modern 
computation theories, as well as the origin of most formal verification problem. 
In this chapter, we will discuss a sort of problems branching out from SAT. They are about a situation 
when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
a set of constrains (e.g. clauses, formulas or polynomials) which is unsatisfiable, sometimes 
it is worthy to explore a smaller subset (core) that keeps UNSAT. From the execution of GB algorithm, 
an auxiliary structure can be obtained to help UNSAT core extraction. This chapter will introduce the 
details about the motivation, mechanism and implementation.

\section{Motivation}
\subsection{Exploiting UNSAT cores for abstraction refinement}
By exploring previous work, we learn that most state-of-art abstraction refinement techniques require 
information mining from UNSAT proofs of intermediate abstractions. 
Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain that how an UNSAT proof is
utilized in such techniques.

\begin{algorithm}[hbt]
\SetAlgoNoLine
	\KwIn{$M$ is the original machine, $p$ is the property to check, $k$ is the number of steps in $k$-BMC}
  $k = $ InitValue\;
  \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT}}
  {
	\Return{``Found error trace"}
  }
  {
	Extract UNSAT proof $\mathcal P$ of $k$-BMC\;
	$M' = $ \textit{ABSTRACT}$(M,\mathcal P)$\;
  }
  \eIf{MODEL-CHECK$(M',p)$ returns \textbf{PASS}}
  {
	\Return{``Passing property"}
  }
  {
	Increase bound $k$\;
	goto Line 2\;
  }
\caption {Abstraction refinement using $k$-BMC}\label{alg:absrefine}
\end{algorithm}
\begin{figure}[hbt]
\centering{
\includegraphics[width=5.5in]{newfig/refine.eps}
\caption{Abstraction by reducing latches}
\label{fig:refine}}
\end{figure}
Assume that we are given a sequential circuit with $n$ latches as shown in Fig.\ref{fig:refine}(a). 
This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
$l_1,\dots,l_n$. Algorithm \ref{alg:absrefine} describes an approach to check if machine $M$ violates property $p$.
This algorithm relies on $k$-BMC technique, which works on the basis of CNF-SAT solving.
The $k$-BMC represents the initial states $I$, the transition relation $T$ and property $p$ as CNF formulas.

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is $SAT$ or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
$SAT$, then a counterexample is found that violates property $p$. If the result is $UNSAT$, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fixpoint detected.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is $UNSAT$.
If there are some latch variables not included in this UNSAT core (denoted by $L_{abs}$), then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Fig.\ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

The second ``if-else" branch means: if we do an ordinary model checking on the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason of this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an over-approximation
of $M$, such that $(M'\implies p) \implies (M\implies p)$. If we find a violation on abstracted machine, then this
abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which will encounter the ``bit-blasting" problem
when the size of datapath (number of latches in Fig.\ref{fig:refine}) is very large. \textbf{Here we propose a totally new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation based on our experience:}

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While computing GB over finite fields is exponential in the number of variables, GB  computation 
is observed to be more efficient for UNSAT problems.} 
The reason is discussed as follows:
\begin{Theorem}
{\bf Weak Nullstellensatz:} Given ideal $J\subset \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow reducedGB(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high
time complexity and is usually not practicable. One reason is that the size of GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is 1, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in non-reduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

\subsection{A Demonstration of Motivating Example}
One  important research topic about UNSAT problems is to identify UNSAT cores efficiently.
An UNSAT core in a CNF formula denotes a subset of clauses which is still unsatisfiable. Here
we re-define this concept in algebraic geometry:
\begin{Definition}
Assume a set of polynomials $F$ and its subset $F_s\subset F$. If ${\bf V}(\langle F\rangle) = {\bf V}
(\langle F_s\rangle) = \emptyset$, we call $F_s$ an {\bf UNSAT core} of $F$. Additionally, if
$F_s$ contains no other UNSAT core, we call it a {\bf minimal} UNSAT core.
\end{Definition}

We  conjecture that based on observation of Buchberger's algorithm's execution, an UNSAT core can be identified.
\begin{Proposition}
Buchberger's algorithm picks pairs of polynomials from a given set, computes their S-poly, then reduces this S-poly
with the given set of polynomials. If the remainder is non-zero,
it is added to the set of polynomials.
By tracking S-poly computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.
\end{Proposition}
\begin{Example}
A SAT problem is described with 8 CNF clauses:

\begin{minipage}[h]{0.3\textwidth}
\begin{align*}
&c_1: \bar{a}\lor\bar{b}\\
&c_2: a\lor\bar{b}\\
&c_3: \bar{a}\lor b\\
&c_4: a\lor b
\end{align*}
\end{minipage}
\begin{minipage}[h]{0.7\textwidth}
\begin{align*}
&c_5: x\lor y\\
&c_6: y\lor z\\
&c_7: b\lor \neg y\\
&c_8: a\lor x\lor \neg z
\end{align*}
\end{minipage}

% \begin{align*}
% &\bar{a}\lor\bar{b}\\
% &a\lor\bar{b}\\
% &\bar{a}\lor b\\
% &a\lor b\\
% &x\lor y\\
% &y\lor z\\
% &b\lor \neg y\\
% &a\lor x\lor \neg z
% \end{align*}
Using Boolean to polynomial mappings given in Table \ref{tab:booltof4}, we can transform them to a set of
polynomials $F$ over ring $\mathbb F_2[a,b,x,y,z]$:

\begin{minipage}[h]{0.4\textwidth}
\begin{align*}
&f_1:ab\\
&f_2:ab+a\\
&f_3:ab+b\\
&f_4:ab+a+b+1
\end{align*}
\end{minipage}
\begin{minipage}[h]{0.6\textwidth}
\begin{align*}
&f_5:xy+y+x+1\\
&f_6:yz+y+z+1\\
&f_7:by+y\\
&f_8:axz+az+xz+z
\end{align*}
\end{minipage}
% \begin{align*}
% &f_1:ab\\
% &f_2:ab+a\\
% &f_3:ab+b\\
% &f_4:ab+a+b+1\\
% &f_5:xy+y+x+1\\
% &f_6:yz+y+z+1\\
% &f_7:by+y\\
% &f_8:axz+az+xz+z
% \end{align*}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Fig.\ref{fig:UNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Fig.\ref{fig:UNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of unsatisfiability of original set of clauses.
\end{enumerate}
\end{Example}
\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}
\caption{DAG representing Spoly computations and multivariate divisions}
\label{fig:UNSAT}}
\end{figure}
% We conclude our approach as follows: i) Keep track of all $Spoly(f_i,
% f_j)\xrightarrow{f_1, \dots, f_s}_+ r$; ii) mark $f_i, f_j$ if $r\neq
% 0$; iii) Mark those $f_l$'s that have leading terms that cancel
% monomial terms in $Spoly(f_i, f_j)$; iv) Build a DAG recording all reductions including Spoly computations
% and multivariate divisions; v) Once $1\in J$ is
% detected, traverse the graph to identify all the marked polynomials that were
% used in obtaining this unit element. These polynomials constitute the
% core.

We conclude our approach as a conjecture algorithm (Algorithm \ref{alg:UNSAT}).

\begin{algorithm}[hbt]
\SetAlgoNoLine
	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
	\KwOut{An UNSAT core $\{f_{m_1},f_{m_2},\dots,f_{m_t}\}$}
\Repeat{$r_l == 1$}
{
	Pick a pair $f_i,f_j\in F$ that has never been computed S-poly\;
	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
	{
		$F = F\cup r_l$\;
		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the S-poly, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the S-poly\;
	}
}
Backward traverse the DAG for remainder ``1", replace $r_l$ with corresponding DAG $G_l$\;
\Return{All roots}
\caption {Extract UNSAT core using a variation of Buchberger's algorithm}\label{alg:UNSAT}
\end{algorithm}

\section{Formalize the Buchberger's Algorithm based UNSAT Core Identification}
\label{sec:core}

\textbf{Problem: }
Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
the ring $R = \F[x_1,\dots,x_d]$ that generate ideal $J = \langle
f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
\emptyset$, or it is determined to be so by applying the Gr\"obner
basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
too. Borrowing the terminology from the Boolean SAT domain, we
call $F_c$ the infeasible core or the unsat core of $F$.

%% Any set of polynomials $F$ with empty variety is an unsat core in
%% itself. There may be more than one unsat cores in $F$, some
%% polynomials may be common to multiple cores (i.e. the
%% cores may be non-disjoint), and these cores may have
%% different cardinalities $|F_c|$. The unsat core problems
%% are therefore further classified as: 
%% \bi
%% \item Identify a {\it minimum} core, i.e. a core of minimum
%%   cardinality. 
%% \item Identify a {\it minimal} core. An unsat core $F_{c}$ is minimal
%%   with respect to the property that while the variety of $F_{c}$ is
%%   empty, the variety of {\it any} subset of $F_c$ is {\it non-empty},
%%   i.e. $V( F_{c} - \{f_i\} ) \neq \emptyset$ for any $f_i \in
%%   F_c$. 
%% \item Identify any unsat core disregarding minimality; i.e. find any
%%   subset $F_c$ of $F$.
%% \ei

%% Let us first consider the problem of finding any $F_c \subset F$,
%% disregarding minimality. 
It is not hard to motivate that an unsat core should be identifiable
using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
the ideal generated by $(F - \{f_j\})$, i.e. $f_j \in \langle F -
\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
polynomials of $F_c$, so $f_j$ is  not a part of the unsat core, and
it can be safely discarded from $F_c$. This can be identified by means
of the GB algorithm for this ideal membership test.% (Def. \ref{def:gb}(ii)).

A n\"aive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

\subsection{The Refutation Tree of the GB Algorithm: Find $F_c$ from $F$}

We investigate if it is possible to identify a core by analyzing the
$Spoly(f_i,f_j)\xrightarrow{F}_+ g_{ij}$ reductions in Buchberger's
algorithm. Since $F$ is itself an unsat core, 
definitely {\it there exists  
a sequence of Spoly reductions in Buchberger's algorithm where
$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
polynomial reduction algorithms can be suitably modified to record
which polynomials from $F$ are used in the division leading to
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that 
we should be able to identify a core by recording
the {\it data} generated by Buchberger's algorithm --- namely, the
critical pairs($f_i,f_j$) used in the Spoly computations,
and the polynomials from $F$ used to cancel terms in the reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. The following example motivates
our approach to identify  $F_c \subseteq F$ using this data:

\begin{Example}
\label{ex:1}
Consider the following set of polynomials $F = \{f_1,\dots, f_9\}$:

\begin{minipage}{2in}
\begin{align*}
f_1&: abc + ab + ac + bc \\
   & + a + b + c + 1\\
f_2&: b\\
f_3&: ac\\
f_4&: ac + a
\end{align*}
\end{minipage}
%\hspace{0.1in}
\begin{minipage}{3in}
\begin{align*}
f_5&: bc + c\\
f_6&: abd + ad + bd + d\\
f_7&: cd\\
f_8&: abd + ab + ad + bd + a + b + d + 1\\
f_9&: abd + ab +bd + b
\end{align*}
\end{minipage}

Assume $>_{DEGLEX}$ monomial ordering with $a>b>c>d$. 
Let $F = \{f_1,\dots,f_9\}$ and 
$J = \langle F \rangle \subset \mathbb{F}_2[a,b,c,d]$ where
$\mathbb{F}_2 = \{0, 1\}$ is the finite field of 2 elements. Then 
$V(J) = \emptyset$ as $GB(J) = 1$.  The set $F$ consists of 4 {\it minimal}
cores: $F_{c1} = \{ f_1,f_2,f_3,f_4,f_7,f_8\}, F_{c2} = \{
f_2,f_4,f_5,f_6,f_8\}, F_{c3} = \{ f_2,f_3,f_4,f_6,f_8\},$ and $F_{c4}
= \{ f_1,f_2,f_4,f_5\}$. 
\end{Example}

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs $(f_i,f_j)$ are selected and reduced by operation
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
$(f_2,f_5)$. It turns out that the Spoly-reductions corresponding to
these 3 pairs lead to the unit ideal. Recording the data
corresponding to this sequence of reductions is depicted by means of a
graph in Fig. \ref{fig:refute}. We call this graph a {\it refutation tree}.

\begin{figure}[hbt]
\centering
\includegraphics[width=\textwidth]{newfig/refutation_tree.eps}
\caption{Generating refutation trees to record unsat cores}
\label{fig:refute}
\end{figure}

In the figure, the nodes of the graph correspond to the polynomials
utilized in Buchberger's algorithm. The leaf nodes always correspond
to polynomials from the given generating set. An edge $e_{ij}$ from
node $i$ to node $j$ signifies that the polynomial at node $j$ results
from polynomial at node $i$. For example, consider the computation
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
following sequence of 1-step divisions:
$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} ~\xrightarrow{f_3}~
\xrightarrow{c\cdot f_2}  ~\xrightarrow{f_2} f_{10}$. This is depicted
as the bottom subtree in the figure, terminating at polynomial
$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
division by $f_2$ resulted in the quotient $a$. The refutation tree of
Fig. \ref{fig:refute} shows further that
$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.

To identify an $F_c \subset F$, we start from the refutation node
``1'', and traverse the graph in reverse, all the way up to the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an unsat core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Fig. \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an unsat core of $F$.

\section{Reducing the Size of the Infeasible Core $F_c$}
\label{sec:alg}
The core $F_c$ obtained from the aforementioned procedure may
contain redundant elements which could be discarded. For
example, consider the core $F_c=\{f_1,\dots, f_5\}$ generated in the
previous section. While $F_c$ is a smaller infeasible core of $F$, it
is not minimal. In fact, Example 1 shows that $F_{c4} =\{f_1,f_2,f_4,f_5\}$ is the
minimal core, where $F_{c4} \subset F_{c}$. Clearly, the polynomial
$f_3 \in F_c$ is a redundant element of the core and can be 
discarded. We will now describe techniques to further reduce the size
of the unsat core by identifying such redundant elements. 
%how the size of this core could be reduced further.  
For this purpose, we will have to perform a more 
systematic book-keeping of the data generated during the execution of
Buchberger's algorithm and the refutation tree.

\subsection{Identifying redundant polynomials from the refutation tree}

We record the S-polynomial reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+{g_{ij}}$ that give a non-zero
remainder when divided by the system of polynomials $F$ at that
moment. The remainder $g_{ij}$ is a polynomial combination of
$Spoly(f_i,f_j)$ and the current basis $F$; thus, it can be
represented as:
\begin{equation}
\label{eqn1}
g_{ij}= S(f_i,f_j)+\displaystyle\sum_{k=1}^m c_kf_k,
\end{equation}
where $0\neq c_k\in\mathbb{F}[x_1,\ldots,x_d]$ and
$\{f_1,\ldots,f_m\}$ is the ``current'' system of polynomials
$F$. For each non-zero $g_{ij}$, we will record the following data:

\begin{equation}
\label{data1}
((g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{equation}

In Eqns. (\ref{eqn1}) and (\ref{data1}), $g_{ij}$ denotes the
remainder of the $S$-polynomial $Spoly(f_i,f_j)$ modulo the current system
of polynomials $f_1,\ldots,f_m$, and we denote by

$$h_{ij}:=\displaystyle\frac{LCM(lm(f_i),lm(f_j))}{lt(f_i)},
h_{ji}:= - \displaystyle\frac{LCM(lm(f_i),lm(f_j))}{lt(f_j)}$$ 
the coefficients of $f_i$, respectively $f_j$, in the $S$-polynomial
$Spoly(f_i,f_j)$. Furthermore, in Eqn. (\ref{data1}), $(c_{k1}, \dots
c_{kl})$ are the respective quotients of division by
polynomials $(f_{k1},\dots,f_{kl})$, generated during the $Spoly$ reduction.  
%polynomial coefficients of  that appear in the
%division process.

\begin{Example}
Revisiting Ex. \ref{ex:1}, and Fig. \ref{fig:refute}, the data
corresponding to $Spoly(f_1,f_2)$\\$\xrightarrow{F}_+  g_{12} = f_{10}$
reduction is obtained as the following sequence of computations:
$$f_{10}=g_{12}=f_1-acf_2-af_2-f_3-cf_2-f_2.$$ As the coefficient
field is $\mathbb{F}_2$ in this example, $-1 = +1$, so:
$$f_{10}=g_{12}=f_1+acf_2+af_2+f_3+cf_2+f_2$$ is obtained.
The data is recorded according to Eqn. (\ref{data1}):

\begin{center}
$((f_{10}=g_{12}), (f_1,1)(f_2,ac)|(a,f_2),(1,f_3),(c,f_2),(1,f_2))$
\end{center}

\end{Example}

Our approach and the book-keeping terminates when we obtain ``1'' as the
remainder of some S-polynomial modulo the current system of 
generators. As an output of the Buchberger's algorithm, we can obtain
not only the Gr\"obner basis $G = \{g_1,\ldots,g_t\}$, but also a
matrix $M$ of polynomials such that:

\vspace{-0.1in}
\begin{center}
\begin{align}
   \begin{bmatrix}
           g_{1} \\
           g_{2} \\
           \vdots \\
           g_{t}
         \end{bmatrix}
    &= M \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
  \end{align}

\end{center}

Each element $g_i$ of $G$ is a polynomial combination of $\{f_1, \dots,
f_s\}$. Moreover, this matrix $M$ is constructed precisely using the
data that is recorded in the form of Eqn. (\ref{data1}). We now give a condition
when the matrix $M$ may identify some redundant elements.

\begin{Theorem}
\label{thm}
With the notations above, we have that a core for the system of
generators $F = \{f_1,\dots,f_s\}$ of the ideal $J$ is given by the
union of those $f_i$'s from $F$ that appear in the data recorded above
and correspond to the nonzero entries in the matrix $M$.  
\end{Theorem}

\begin{Proof}
In our case, since the variety is empty, and hence the ideal is
unit, we have that $G = \{g_1=1\}$ and $t=1$. Therefore $M=
[a_1, \ldots, a_s]$ is a vector. Then the output of the algorithm
gives: $1 = a_1f_1+\cdots + a_s f_s.$ Clearly, if $a_i=0$ for some $i$ then
$f_i$ does not appear in this equation and should not be included in
the infeasible core of $F$.

\end{Proof}

\begin{Example}

Corresponding to Example \ref{ex:1} and the refutation tree shown in
Fig. \ref{fig:refute}, we discover that the polynomial $f_3$ is used
only twice in the division process. In both occasions, the quotient of
the division is 1. From Fig. \ref{fig:refute}, it follows that:
\begin{equation}
1 = (f_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot
  f_3}+ \dots + (f_1 + f_2)
\end{equation}

Since $1 + 1 = 0$ over $\F_2$, we have that the entry in $M$
corresponding to $f_3$ is 0, and so $f_3$ can be discarded from the
core. 
\end{Example}

\subsection{The GB-Core Algorithm Outline}

The following steps describe an algorithm (GB-Core) that allows us to compute a
refutation tree of the polynomial set and corresponding matrix $M$.

{\bf Inputs:} Given a system of polynomials $F=\{f_1,\ldots,f_s\}$, a
monomial order $>$ on $\mathbb{F}[x_1,\ldots,x_d]$.

{\bf S-polynomial reduction:} We start computing the S-polynomials of the system of
generators $\{f_1,\ldots,f_s\}$, then divide each of them by the
current basis $G=\{f_1,\dots,f_s,\dots,f_m\}$, which is the
intermediate result of Buchberger's algorithm.  
In this way, we obtain expressions of the following type:
\begin{equation}
\label{eqn:red}
g_{ij}= \underbrace{h_{ij}f_{i}+h_{ji}f_{j}}_{Spoly(f_i,f_j)}+\displaystyle\sum_{k=1}^m c_kf_k
\end{equation}
If the remainder $g_{ij}$ is non-zero, we denote it by
$f_{m+1}$ and add it to the current set of generators $G$. We
also record the data as in Eqn. (\ref{data1}): 
\begin{displaymath}
((f_{m+1}=g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
This data forms a part of the refutation tree rooted at node $f_{m+1}$.

{\bf Recording the coefficients:} In Eqn. (\ref{eqn:red}) we obtain a
vector of polynomial coefficients $c_k$ where $k>s$. These
coefficients are associated with new elements (remainders) in the
Gr\"obner basis, that are not a part of the unsat core. 
%which cannot benefit the UNSAT core extraction. 
Since each polynomial $f_k$, ($k>s$) is generated by
$\{f_1,\dots,f_s\}$, we can re-express $f_k$ in terms of $\{f_1,\dots,
f_s\}$. Thus, each $f_k, k>s$ can be written as $f_k = d_1f_1 + \dots
+ d_sf_s$. This process adds a new row $(d_1,\dots,d_s)$ to the
coefficient matrix $M$.

%% Therefore we need to rewrite the vector to
%% one with only $c_k, k=1\dots s$, which can be achieved by substituting $f_k,k>s$ by $f_1,\dots,f_s$:
%% initially we have $f_{s+1}= c_1f_1 + c_2f_2 + \cdots + (h_{ij}+c_i)f_{i}+\cdots+(h_{ji}+c_j)f_{j}+\cdots+c_sf_s$,
%% inductively if we can present $f_{s+1}$ to $f_{s+k-1}$ by $f_1,\dots,f_s$, we can also rewrite
%% $f_{s+k} = d_1f_1+\cdots+d_sf_s$. Then we add a new row $ (d_1,\dots,d_s ) $ to coefficient matrix $M$.

{\bf Termination and refutation tree construction:} We perform
S-polynomial reductions and record these coefficients  generated
during the division until the remainder $f_m = 1$ is encountered. The
corresponding data is stored in a data-structure $D$ corresponding to
Eqn. (\ref{data1}). The matrix $M$ is also constructed. From this
recorded data the refutation tree can be easily derived.

%the algorithm
%will generate a set of new generators $G$ and coefficient matrix $M$.
%Meanwhile we can construct the refutation tree with the recorded data.
We start with the refutation node ``$f_m=1$'':
\begin{displaymath}
((f_{m}=1)(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
and recursively substitute the expressions for the polynomials $f_k$
($k>s$) until we obtain the tree with all the leaf nodes corresponding
to the original set of polynomials $\{f_1,\dots,f_s\}$. Algorithm
\ref{algo:gbcore} describes this data recording through which the
refutation tree $T$ and the matrix $M$ is derived.

\begin{algorithm}[H]
 \caption{GB-core algorithm (based on Buchberger's algorithm)}
 \label{algo:gbcore}
 \begin{algorithmic}[1]

 \REQUIRE $F = \{f_1, \dots, f_s\} \in \F[x_1, \dots, x_d], f_i\neq 0$
 \ENSURE Refutation tree $T$ and coefficients matrix $M$
 \STATE{ {Initialize: list $G \gets F$; Dataset $D\gets \emptyset$; $M\gets s\times s$ unit matrix} }
 \FOR {{ each pair $(f_i,f_j)\in G$  }}
 	\STATE  $f_{sp},(f_{i},h_{ij})(f_{j},h_{ji}) \gets$ Spoly($f_i,f_j$) \COMMENT{$f_{sp}$ is the S-polynomial}
 	\STATE{{ $g_{ij}|(c_{k1},f_{k1}),\dots,(c_{kl},f_{kl}) \gets (f_{sp}\xrightarrow{G}_+  g_{ij})$}}
 	\IF{{ $g_{ij} \neq 0$ }}
 		\STATE {{$G \gets G \cup g_{ij}$}}
 		\STATE {{$D \gets D \cup
                    ((g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})|
                    (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))$}}
                \STATE{{Update matrix $M$}}
 	\ENDIF
 	\IF {{$g_{ij} = 1$}}
 		\STATE {{Construct $T$ from $D$}}
 		\STATE {{ Return($T,M$) }}
 	\ENDIF
 \ENDFOR
 \end{algorithmic}
 \end{algorithm}

Notice that the core can actually be derived directly from the matrix 
$M$. However, we also construct the refutation tree $T$ as it
facilitates an iterative refinement of the core, which is described
in the next section.

\section{Iterative Refinement of the Unsat Core}
\label{sec:iter}

As with most other unsat core extractors, our algorithm also cannot
generate a minimal core in one execution. To obtain a smaller core, we
 re-execute our algorithm with the core obtained in the current
iteration. We describe two heuristics that are applied to our
algorithm to increase the likelihood of generating a smaller core in
the next iteration.

%After eliminating all redundant polynomials, we can call our GB engine
%with the new core. 
An effective heuristic should increase the chances that the refutation
``1'' is composed of fewer polynomials.  In our GB-core algorithm, we
use a strategy to pick critical pairs such that polynomials with
larger indexes get paired {\it later} in the order:

$$(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$$

Moreover, for the reduction process
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$, we pick divisor polynomials from
$F$ following the increasing order of polynomial indexes. Therefore,
by relabeling the polynomial indexes, we can affect their
chances of being selected in the unsat core. We use two criteria to
to affect the polynomial selection in the unsat core. One corresponds
to the \emph{refutation distance}, whereas the other corresponds to
the {\it frequency} with which a polynomial appears in the refutation
tree.

\begin{Definition}[Refutation Distance]
Refutation distance of a polynomial $f_i$ in a refutation tree
corresponds to the number of edges on the shortest path from refutation
node ``1'' to any leaf node that represents polynomial $f_i$. 
\end{Definition}

On a given refutation tree, polynomials with shorter refutation
distances are used as divisors in later stages of polynomial
reductions; which implies that they may generally have lower-degree
leading terms. This is because we impose a degree-lexicographic term
order, and successive divisions (term cancellations) reduce the degree
of the remainders. However, what is more desirable is to use these
polynomials with lower-degree leading terms earlier in the reduction,
as they can cancel more terms. This may prohibit other (higher-degree)
polynomials from being present in the unsat core.

%% So polynomials with lower degree
%% leading terms are more  likely to be used as divisors in later steps
%% of reduction.  
%% In other words, polynomials with shorter refutation
%% distance may have lower-degree leading terms. 
%% If lower degree
%% polynomials are used earlier in the division process, they can cancel
%% more terms, and they will appear in the tree more often --- which may
%% imply that they are a part of the core. 
% %such that the probability that they appear in
%the refutation tree is larger.

Similarly, the motivation for using the \emph{frequency of occurrence}
of $f_i$ in the refutation tree is as follows: polynomials that appear
frequently in the refutation tree may imply that they have certain
properties (leading terms) that give them a higher likelihood of being
present in the unsat core.

%make them "favourable" in the unsat core
%selection. For example, their leading terms may contain variables
%that are require them to be included in the minimal core.

We apply both heuristics: after the first iteration of the GB-core
algorithm, we analyze the refutation tree $T$ and sort the polynomials
in the core by the refutation distance criterion, and use the
frequency criterion as the tie-breaker. The following example
illustrates our heuristic.

%\vspace{5mm}\\
\begin{Example} 
Consider a set of 6 polynomials over $\F_2$ of an
infeasible instance.
\begin{align*}
f_1: x_1x_3+x_3; & ~~~f_2: x_2 + 1\\
f_3: x_2x_3+x_2; & ~~~f_4: x_2x_3\\
f_5: x_2x_3 + x_2 + x_3 + 1; & ~~~f_6 : x_1x_2x_3 +x_1x_3
\end{align*}

After the first iteration of the GB-core algorithm, the core is
identified as $\{f_1, f_2,f_3,f_4\}$, and we obtain a
refutation tree as shown in Fig.\ref{fig:refine}(a).  
\begin{figure}[hbt]
\centering
\includegraphics[width=\textwidth]{newfig/core_refine.eps}
\caption{Refutation trees of core refinement example}
\label{fig:refine}
\end{figure}

The refutation distance corresponding to polynomial $f_2$ is equal to
2 levels. Note that while three leaf nodes in Fig. \ref{fig:refine} (a)
correspond to $f_2$, the shortest distance from ``1'' to any $f_2$
node is 2 levels. The refutation distance and frequency measures of
other polynomials are identical -- equal to 3 and 1, respectively --
so their relative ordering is unchanged. We reorder $f_2$ to be the
polynomial with the  smallest index. We re-index the polynomial set 
$f_1'=f_2, f_2' = f_1, f_3' = f_3, f_4' = f_4$
and apply our GB-core algorithm on the core
$\{f_1',f_2',f_3',f_4'\}$. The result is shown in
Fig.\ref{fig:refine}(b) with the core identified as $\{f_1', f_3',
f_4'\} = \{f_2,f_3,f_4\}$. Further iterations do not refine the core
-- i.e. a fix point is reached. 
\end{Example}

\section{Refining the Unsat core using Syzygies}
\label{sec:syz}

%% In some
%% cases our iterative core refining algorithm cannot
%% give us the minimal core when it hits the fixpoint.
%% It indicates the limitation of using re-labelling 
%% strategy, therefore a new method to identify the 
%% redundancy in UNSAT core is needed.

The unsat core obtained through our GB-core algorithm is by nature a
refutation polynomial that equals to 1:  
$$1 = \sum_{i=1}^s c_i\cdot f_i$$
where $0 \neq c_i \in \F[x_1,\dots,x_d]$ and the polynomials
$F = \{f_1,\dots,f_s\}$ form a core. Suppose that a polynomial 
$f_k \in F$ can be represented using a combination of the rest of the
polynomials of the core, e.g.:
$$f_k = \sum_{j\neq k} c_j'f_j.$$

Then we can substitute $f_k$ in terms of the other polynomials in the
refutation. Thus, $f_k$ can be dropped from the core as it is 
redundant. One of the limitations of the GB-core algorithm and the
re-labeling/refinement strategy is that they cannot easily identify
such polynomials $f_k$ in the generating set $F$ that can be composed
of the other polynomials in the basis, i.e. 
$f_k \in \langle F-\{f_k\} \rangle$. We present an approach targeted
to identify such combinations to further refine the core.

 %% Since coefficients $\{c_i\}$ are also polynomials in $\mathbb F[x_1,\dots,x_n]$, we can find 
 %% enormous number of combinations of $\{f_i\}$. Consider finding an algorithmic solution to 
 %% this problem without introducing extra search,
 %% we propose to utilize the \emph{syzygies} generated during
 %% executing the GB-core algorithm.

During the execution of Buchberger's algorithm, many critical pairs
$(f_i,f_j)$ do not add any new polynomials in the basis when
$Spoly(f_i,f_j)\xrightarrow{F}_+0$ gives zero remainder. Naturally,
for the purpose of the GB computation, this data is
discarded. However, our objective is to gather more information from
each GB iteration so as to refine the core. Therefore, we further
record the quotient-divisor data from S-polynomial reductions that
result in the remainder 0. Every $Spoly(f_i,f_j)\xrightarrow{F}_+0$
implies that some polynomial combination of $\{f_1,\dots,f_s\}$
vanishes: i.e. $c_1f_1 + c_2f_2 + \dots+c_sf_s=0$, for some
$c_1,\dots,c_s$. These elements ($c_1,\dots,c_s$) form a syzygy on $f_1,\dots,f_s$.

\begin{Definition}[Syzygy \cite{ideals:book}]
Let $F = \{f_1,\dots,f_s\}$. A syzygy on $f_1,\dots,f_s$ is an
$s$-tuple of polynomials $(c_1,\dots,c_s)\in (\F[x_1,\dots,x_d])^s$
such that $\sum_{i=1}^s c_i\cdot f_i = 0$.
\end{Definition}

For each $Spoly(f_i,f_j)\xrightarrow{F}_+0$ reduction, we record the
information on corresponding syzygies as in Eqn. (\ref{eqn:syz}), also
represented in matrix form in Eqn. (\ref{mat:syzygy}):

\begin{equation} \label{eqn:syz}
%\[
 \begin{cases}
 c_1^1f_1+c_2^1f_2+\cdots+c_s^1f_s = 0\\
 c_1^2f_1+c_2^2f_2+\cdots+c_s^2f_s  = 0\\
 \ \ \ \ \ \  \vdots \\
 c_1^mf_1+c_2^mf_2+\cdots+c_s^mf_s = 0  
 \end{cases}
%\]
\end{equation}

%which can be represented in matrix form as:
\begin{center}
\begin{align}
\label{mat:syzygy}
   \begin{bmatrix}
           c_1^1 & c_2^1 & \cdots & c_s^1 \\
           c_1^2 & c_2^2 & \cdots & c_s^2 \\
           \vdots & \vdots & \ddots & \vdots \\
           c_1^m & c_2^m & \cdots & c_s^m
         \end{bmatrix}
    \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
         &= 0
  \end{align}

\end{center}

\ \\
 Here $\{f_1,f_2,\dots,f_s\}$ is the given core.
 Take one column of the syzygy matrix (e.g. the set of polynomials in $j$-th column
 $c_j^1, c_j^2, \dots, c_j^m$)  and compute its reduced Gr\"obner
 basis $G_r$. If $G_r = \{1\}$, then it means that there exists some
 polynomial vector  $[r_1,r_2,\dots,r_m]$ such that $1 = r_1c_j^1 +
 r_2c_j^2 + \cdots + r_mc_j^m = \sum_{i=1}^m r_ic_j^i.$ 
 If we multiply each row $i$ in the matrix of Eqn. (\ref{mat:syzygy})
 with $r_i$, and sum up all the rows, we will obtain the
 following equation: 
\vspace{-0.2in}
 \begin{center}
\begin{align}
   \begin{bmatrix}
           \sum_{i=1}^m r_ic_1^i & ~~\cdots & ~~ 1 ~~ & ~~ \cdots ~~ & ~~\sum_{i=1}^m r_ic_s^i
         \end{bmatrix}
    \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
         &= 0
  \end{align}

\end{center}

This implies that 
 $$\sum_{i=1}^m r_ic_1^if_1 + \cdots + f_j
 +\cdots + \sum_{i=1}^m r_ic_s^if_s = 0,$$
or that $f_j$ is a polynomial combination of
$f_1,\dots,f_s$ (excluding $f_j$). Subsequently, we can deduce that $f_j$ can be
discarded from the core. By repeating this procedure, some redundant
polynomials can be identified and size of unsat core can be reduced
further.

 \begin{Example}
 Revisiting Example\ref{ex:1}, execute the GB-core algorithm and
 record the syzygies on $f_1,\dots,f_s$ corresponding to the
 S-polynomials that give 0 remainder. The coefficients can be
 represented as entries in matrix shown below. For example, the first
 row in the matrix corresponds to the syzygies generated by
 $Spoly(f_1,f_3)\xrightarrow{F}_+0$.

\begin{equation}\label{eqn:sm}
% \[
 \begin{blockarray}{cccccccccccc}
  && f_1 & f_2 & f_3 & f_4 & f_5 & f_6 & f_7 & f_8 & f_9 & f_{10} \\
  \begin{block}{cc(cccccccccc)}
  Spoly(f_1,f_3)\ & & 1 & a+c+1 & b+1 &0&0&0&0&0&0&1\\
  Spoly(f_2,f_3)\ & & 0 & ac & b &0&0&0&0&0&0&0\\
  Spoly(f_1,f_4)\ & & 1 & c+1 & 1 &b&0&0&0&0&0&1\\
  Spoly(f_2,f_4)\ & & 0 & ac+a & 0 &b&0&0&0&0&0&0\\
  Spoly(f_1,f_5)\ & & 1 & a+c+1 & 0 &0&a&0&0&0&0&1\\
  \end{block}
  \end{blockarray}
% \]
\end{equation}

Usually, we need to generate extra columns compared to the syzygy
matrix of Eqn. (\ref{mat:syzygy}). In this example, we need to add an
extra column for the coefficient of $f_{10}$. This is because $f_{10}$
is not among the original generating set; however, some S-polynomial
pairs require this new remainder $f_{10}$ as a divisor during
reduction. In order to remove this extra column, we need to turn the
non-zero entries in this column to 0 through standard matrix
manipulations.

Recall that we record $f_{10}$ in $M$ as a nonzero remainder when
reducing S-polynomial pair $Spoly(f_1,f_2)\xrightarrow{F}_+f_{10}$. We
extract this information from the coefficient matrix $M$:
 $$(1 ~~ac+a+c+1 ~~1 ~~0 ~~0 ~~0 ~~0 ~~0 ~~0 )$$

 It represents $f_{10}$ is a combination of $f_1$ to $f_9$:
 $$f_{10} = f_1 + (ac+a+c+1)f_2 + f_3$$
 It can be written in the same syzygy matrix form (with column
 $f_{10}$ present) as follows:

\begin{equation}\label{sr}%  \[
 \begin{blockarray}{cccccccccccc}
  && f_1 & f_2 & f_3 & f_4 & f_5 & f_6 & f_7 & f_8 & f_9 & f_{10} \\
  \begin{block}{cc(cccccccccc)}
  Spoly(f_1,f_2)& & 1 & ac+a+c+1 & 1 &0&0&0&0&0&0&1\\
  \end{block}
  \end{blockarray}
 \end{equation}

 By adding this row vector (Eqn. (\ref{sr})) to the rows in
 Eqn. (\ref{eqn:sm}) corresponding to the non-zero entries in the
 column for $f_{10}$, we obtain the syzygy matrix only for the
 polynomials in the core:
  \[
 \begin{blockarray}{ccccccccccc}
  && f_1 & f_2 & f_3 & f_4 & f_5 & f_6 & f_7 & f_8 & f_9  \\
  \begin{block}{cc(ccccccccc)}
  Spoly(f_1,f_3)\ & & 0 & ac & b &0&0&0&0&0&0\\
  Spoly(f_2,f_3)\  && 0 & ac & b &0&0&0&0&0&0\\
  Spoly(f_1,f_4)\  && 0 & ac+a & 0 &b&0&0&0&0&0\\
  Spoly(f_2,f_4)\  && 0 & ac+a & 0 &b&0&0&0&0&0\\
  Spoly(f_1,f_5)\  && 0 & ac & 1 &0&a&0&0&0&0\\
  \end{block}
  \end{blockarray}
 \]

 We find out there is a ``1" entry in the $f_3$ column. The last row
 implies that $f_3$ is a combination of $f_2, f_5$ ($f_3 = ac f_2 + a
 f_5$), so $f_3$ can be discarded from the core.

 \end{Example}

The syzygy heuristic gathers extra information from the GB
computation, it is still not sufficient to derive all polynomial
dependencies. In Buchberger's algorithm,  many S-polynomials reduce to
zero, so the number of rows of the syzygy matrix can be much larger than
the size of original generating set. Full GB computation on each
column of the syzygy matrix can become prohibitive to apply
iteratively. For this reason, we only apply the syzygy heuristic
on the smaller reduced core given by our iterative refinement algorithm.

\textbf{Our Overall Approach for Unsat Core Extraction:} i) Given
the set $F = \{f_1,\dots,f_s\}$, we apply the GB-core algorithm,
record the data $D, M$ (Section 4) and the syzygies $S$ on
$f_1,\dots,f_s$. ii) From $M$, we obtain a core $F_c \subseteq
F$. iii) Iteratively refine $F_c$ (Section 5) until $|F_c|$ cannot be
reduced further. iv) Apply the syzygy-heuristic (Section 6) to
identify if some $f_k \in F_c$ is a combination of other polynomials
in $F_c$; all such $f_k$ are discarded from $F_c$. This gives us the
final unsat core $F_c$.

\section{Experiment results}
\label{sec:exp}

We have implemented our core extraction approach (the GB-Core
and the refinement algorithms) using the \textsc{Singular} symbolic
algebra computation system [v. 3-1-6] \cite{DGPS}. With our
tool implementation, we have performed experiments to extract a
minimal unsat core from a given set of  
polynomials. Our experiments run on a desktop with
3.5GHz Intel $\text{Core}^\text{TM}$ i7-4770K Quad-core CPU, 16 GB RAM and
64-bit Ubuntu Linux OS. The experiments are shown in Table \ref{tab:result}.

%Nowadays most SAT benchmarks are huge, which choke our GB engine. 
Gr\"obner basis is not an efficient engine for solving contemporary
industry-size CNF-SAT benchmarks, as the translation from CNF introduces
too many variables and clauses for GB engines to handle. In order to
validate our approach, we use a somewhat customized  benchmark
library: i) "aim-100" is a modified version  of the random 3-SAT
benchmark "aim-50/100", modified by adding some redundant clauses; ii)
The "subset" series are generated for random subset sum problems; iii)
"cocktail" is similarly revised from a combination of factorization
and a random 3-SAT benchmark; iv) and "phole4/5" are generated by
adding redundant clauses to pigeon hole benchmarks; v) Moreover,
SMPO and RH benchmarks correspond to hardware equivalence checking
instances of sequential Galois field normal basis modulo multiplier
circuits \cite{SMPO,RHmulti}, compared against a golden model
{\it spec}. Similarly, the "MasVMon" benchmarks are the equivalence
checking  circuits corresponding to Mastrovito multipliers compared
against  Montgomery multipliers \cite{lv:tcad2013}. Some of these are
available as CNF formulae, whereas others were available directly as
polynomials over finite fields. The CNF formulae are translated as
polynomial constraints over $\mathbb{F}_2$ (as shown in
\cite{condrat-tacas07}), and the GB-Core algorithm and the refinement
approach is applied.   
%These benchmarks are in moderate size for our GB engine but not too trivial.

%\vspace{-0.1in}
\begin{table}[htb]
\centering
\caption{Results of running benchmarks using our tool. 
\small{Asterisk($^*$) denotes that the benchmark was not translated from CNF.} 
Our tool is composed by 3 parts: part I runs a single GB-core algorithm,
part II applies the iterative refinement heuristic to run the GB-core algorithm
iteratively, part III applies the syzygy heuristic.}
{\small 
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|}
\hline
\multirow{3}{2.2cm}{\centering Benchmark} 
& \multirow{3}{0.9cm}{\centering \# Polys} 
& \multirow{3}{0.8cm}{\centering \# MUS} 
& \multicolumn{3}{c|}{\multirow{2}{1.8cm}{\centering Size of core}}
 & \multirow{3}{1.6cm}{\centering \# GB-core iterations}
 & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
 & \multirow{3}{1.8cm}{\centering Runtime of PicoMUS (sec)} \\
  & & &\multicolumn{3}{c|}{}& &\multicolumn{3}{c|}{}& \\
  \cline{4-6} \cline{8-10}
    & & & I & II & III & & I & II & III & \\
\hline
\hline
$5\times 5$ SMPO & 240  & 137  & 169 & 137 & 137  & 8  & 1222 & 1938 & 1698 & $<0.1$\\
$4\times 4$ SMPO$^*$ & 84  & 21  & 21 & 21 & 21 & 1  & 125 & 0.3 & 29  & - \\
$3\times 3$ SMPO$^*$ & 45  & 15  & 15 & 15 & 15  & 1  & 6.6 & 0.2 & 5.7 & - \\
$3 \times 3$ SMPO & 17 & 2 & 2 &2 & 2 & 1 & 0.07 & 0.01 & 0.01  & $<0.1$  \\
$4 \times 4$ MasVMont$^*$ & 148 & 83 & 83 & 83 & 83 & 1 & 23 & 139 & 12 & - \\
$3 \times 3$ MasVMont$^*$ & 84 & 53 & 53 & 53 & 53  & 1 & 4.3 & 4.6 & 0.9  & - \\
$2 \times 2$ MasVMont & 27 & 23 & 24 & 23 & 23 & 2 & 1.3 & 1.0 & 80  & $<0.1$ \\
$5\times 5$ RH$^*$ & 142  & 34  & 48 & 35 & 35  & 4  & 997 & 1.0 & 80 & - \\
$4\times 4$ RH$^*$ & 104  & 35  & 43 & 36 & 36  & 3  & 96 & 5.7 & 0.6 & -\\
$3\times 3$ RH$^*$ & 50  & 20  & 20 & 20 & 20  & 1  &2.9 & 3.5 & 10  & -\\
aim-100 & 79 & 22 & 22 & 22 & 22 & 1  & 43 & 0.7 & 0.2 & $<0.1$\\
cocktail & 135 & 4 & 6 & 4 & 4 & 2 & 51 & 0.01 & 0.01  & $<0.1$ \\
subset-1 & 100 & 6 & 6 & 6 & 6 & 1 & 2.4 & 0.01 & 0.01 & $<0.1$ \\
subset-2 & 141 & 19 & 37 & 23 & 21 & 2 & 12 & 1.6 & 1.1 & $<0.1$  \\
subset-3 & 118 & 16 & 13 & 12 & 11 & 2 & 8.6 & 0.2 & 0.07 & $<0.1$ \\
phole4 & 104 & 10 & 16 & 16 & 10 & 1 & 4.3 & 0.2 & 0.5 &  $<0.1$\\
phole5 & 169 & 19 & 30 & 25 & 19 & 3 & 12 & 3.2 & 2.7 & $<0.1$ \\
\hline
\end{tabular}
}
\label{tab:result}  
\end{table}

In Table \ref{tab:result}, 
%we list the details of our experimental results. In the table, 
\#Polys denotes the given number of polynomials
from which a core is to be extracted. \#MUS is the {\it minimal} core either
extracted by PicoMUS (for CNF benchmarks) or exhaustive deletion method (for non-CNF bencmarks).
 \#GB-core iterations corresponds to the
number of calls to the GB-core engine to arrive at the reduced unsat
core. The second last column shows the improvement in the minimal core
size by applying the syzygy heuristic on those cases which cannot be
iteratively refined further. 
We choose PicoMUS as a comparison to our tool because it is a state-of-art MUS extractor,
and the results it returned for our set of benchmarks are proved to be minimal. 
The data shows that in most of these
cases, our tool can produce a minimal core. For the subset-3
benchmark, we obtain another core with even smaller size than the one from PicoMUS. The results
demonstrate the power of the Gr\"obner basis technique to identify the
causes of unsatisfiability. 
%From the
%experimental results we can also make the observation that the GB-core
%algorithm, particularly Theorem \ref{thm}, offers quite a lot of scope for identifying redundant
%polynomials that can be eliminated from the core --- without resorting
%to a brute-force membership check of every polynomial $f_i \in F -
%\{f_i\}$.

\section{Conclusions}\label{sec:conc}
This paper addresses the problem of identifying an infeasible core of
a set of multivariate polynomials, with coefficients from a field,
that have no common zeros. The problem is posed in the context of
computational algebraic geometry and solved using the Gr\"obner basis
algorithm. We show that by recording the data produced by the
Buchberger's algorithm -- the $Spoly(f_i,f_j)$ pairs, as well as the
polynomials of $F$ used in the division process
$Spoly(f_i,f_j)\xrightarrow{F}_+ 1$ -- we can identify certain
conditions under which a polynomial can be discarded from a core. An
algorithm was implemented within the Singular computer algebra tool
and some experiments were conducted to validate the approach. While
the use of GB engines for SAT solving has a rich history, the problem
of unsat core identification using GB-engines has not been addressed
by the SAT community. We hope that this paper will kindle some
interest in this topic which is worthy of attention from the SAT
community -- particularly when there seems to be a renewal of interest
in the use of Gr\"obner bases for formal verification
\cite{lv:tcad2013,gao:qe-gf-gb,kalla:fmcad_tut2015,rolf:date16}.

### 2016-11-09 dissertation/prelim_GF.tex
dissertation v2_chap 3 4_update1

\chapter{Galois Field and Sequential Arithmetic Circuits}

### 2016-11-10 dissertation/prelim_GF.tex
dissertation v2: part prelim update 2

@@
 \chapter{Galois Field and Sequential Arithmetic Circuits}
+This chapter provides a mathematical background for understanding 
+finite fields (Galois fields) and explains how to design Galois field arithmetic circuits.
+We first introduce the mathematical concepts of groups, rings, fields, and 
+polynomials. 
+We then apply these concepts to create Galois field arithmetic functions and 
+explain how to map them to a Boolean circuit implementation.
+Additonally, we introduce a special type of sequential arithmetic hardware based on normal basis, as well
+as the normal basis theory behind the designing such hardware.
+
+The material is referred from \cite{galois_field:mceliece, ftheory:2006, ff:1997} for Galois field concepts, 
+\cite{mastro:1989, PT:1985, acar:1998, wu:2002, Knezevic:2008} for hardware design over Galois fields 
+and previous work by {\it Lv} \cite{lv:phd}.
+Normal basis theory in this section refers to \cite{normal_book, gao:phd_normal_basis} and sequential
+normal basis arithmetic hardware designs come from \cite{mullinONB,MasseyOmura,agnew1991implementation, RHmulti}.
+
+\section{Commutative Algebra}
+\label{sec:algebra}
+\subsection{Group, ring and field}
+\begin{Definition}
+An {\bf abelian group} is a set $\mathbb{S}$ with a binary operation $'+'$
+which satisfies the following properties: 
+\begin{itemize}
+\item {\it Closure Law:} For every $a, b \in \mathbb{S}, a + b \in \mathbb{S}$  
+\item {\it Associative Law:} For every $a, b, c \in \mathbb{S}, (a + b) + c = a + (b + c)$
+\item {\it Commutativity:} For every $a, b \in \mathbb{S}, a + b = b + a$. 
+\item {\it Additive Identity:} There is an identity element $0 \in \mathbb{S}$
+such that for all $a \in \mathbb{S};$ $a + 0 = a$.
+\item {\it Additive Inverse:} If $a \in \mathbb{S}$, then there is an
+element $a^{-1} \in \mathbb{S}$ such that $ a + a^{-1} = 0$.
+\end{itemize}
+\end{Definition}
+
+The set of integers $\mathbb{Z}$ forms an abelian group under the addition operation.
+
+\begin{Definition}
+Given a set $\mathbb{R}$ with two binary operations, $'+'$ and $'\cdot'$, 
+and element $0 \in \mathbb{R}$, the system $\mathbb{R}$ is called a {\bf commutative ring with unity} if the following properties hold:
+\begin{itemize}
+\item $\mathbb{R}$ forms an abelian group under the '+' operation with additive identity element $0$.
+\item {\it Multiplicative Distributive Law}: For all $a, b, c \in$ $\mathbb{R}$, $a\cdot (b + c) = a\cdot b + a\cdot c$.
+\item {\it Multiplicative Associative Law}: For every $a, b, c\in \mathbb{R}$, $a\cdot (b\cdot c) = (a\cdot b)\cdot c$. 
+\item {\it Multiplicative Commutative Law}: For every $a,b \in \mathbb{R}$, $a\cdot b = b\cdot a$
+\item {\it Identity Element}: There exists an element $1 \in$ $\mathbb{R}$ 
+such that for all $a \in \mathbb{R}$, $a\cdot 1 = a =1\cdot a$
+\end{itemize}
+\end{Definition}
+
+{\bf Ring} is a broad algebraic concept. In this dissertation, this word is used to refer a special 
+sort of ring -- {\bf commutative ring with unity}. Two common 
+examples of such rings are the set of integers, $\mathbb{Z}$, and the set of 
+rational numbers, $\mathbb{Q}$. Note that while both of these examples are
+rings with an infinite number of elements, the number of elements in a ring 
+can also be finite.
+
+\begin{Definition}
+A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
+non-zero element in $\mathbb{F}$ has a multiplicative inverse; i.e. $\forall
+a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
+\hat{a} = 1$.
+\end{Definition}
+
+A field is defined as a ring with one extra condition: the presence of a 
+multiplicative inverse for all non-zero elements.
+Therefore, a field must be a ring while a ring is not necessarily a field.
+For example, the set $\mathbb{Z}_{2^k} = \{0,1,\cdots, 2^k-1\}$ forms a finite ring.
+However, $\mathbb{Z}_{2^k}$ is not a field because not every element in
+$\mathbb{Z}_{2^k}$ has a multiplicative inverse. 
+In the ring $\mathbb{Z}_{2^3}$, for 
+instance, the element $5$ has an inverse ($5\cdot5\pmod{8}=1$) but the element $4$
+does not.
+
+The main concept of field theory is {\bf Field Extensions}. The idea behind a
+field extension is to take a base field and construct a larger field which 
+contains the base field as well as satisfies additional properties. For example,
+the set of real numbers $\mathbb{R}$ forms a field; one common extension of 
+$\mathbb{R}$ is the set of complex numbers $\mathbb{C}=\mathbb{R}(i)$. Every
+element of $\mathbb{C}$ can be represented as $a+b\cdot i$ where $a,b \in \mathbb{R}$,
+hence $\mathbb{C}$ is a two-dimensional extension of $\mathbb{R}$.
+
+Like rings, fields can also contain either an infinite or a finite number of 
+elements. 
+In this dissertation we focus on finite fields, also known as Galois fields, and 
+the construction of their field extensions.
+
+\subsection{Galois Field}
+Galois fields, also known as finite fields, find widespread applications in 
+many areas of electrical engineering and computer science such as error-
+correcting codes, elliptic curve cryptography, digital signal processing, 
+testing of VLSI circuits, among others.
+In this dissertation, we specifically focus on their application to 
+Elliptic Curve Cryptography as Galois field arithmetic circuits.
+This section describes the relevant Galois field concepts
+\cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997}
+and hardware arithmetic designs over such fields \cite{mastro:1989} \cite{PT:1985} 
+\cite{acar:1998} \cite{wu:2002} \cite{Knezevic:2008}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{Definition} 
+A {\bf Galois field}, denote $\Fq$, is a field with a finite
+number of elements, $q$. The number of elements $q$ of the Galois field is
+a power of a prime integer, i.e. $q = p^k$, where $p$ is a prime
+integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
+$\F_{p^{k}}$.
+\end{Definition}
+
+Fields in the form $\F_{p^{k}}$ are called Galois extension fields.
+We are specifically interested in extension fields of type 
+$\Fkk$, where $k > 1$. These are extensions of the binary
+field $\F_2$.
+\begin{Example}
+Addition and multiplication operations over $\F_2$:
+\begin{table}[!h]
+	\centering
+	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
+	\hhline{~---~}
+	\multirow{3}{*}{} & $+$ & $0$ & $1$ & \multirow{3}{*}{} \\
+	\hhline{~---~}
+	& $0$ & $0$ & $1$ & \\
+	& $1$ & $1$ & $0$ & \\
+	\hhline{~---~}
+	\multicolumn{5}{c}{}\\
+	\multicolumn{5}{c}{Addition over $\F_2$}\\
+	\end{tabular}
+	\quad
+	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
+	\hhline{~---~}
+	\multirow{3}{*}{} & $\cdot$ & $0$ & $1$ & \multirow{3}{*}{} \\
+	\hhline{~---~}
+	& $0$ & $0$ & $0$ & \\
+	& $1$ & $0$ & $1$ & \\
+	\hhline{~---~}
+	\multicolumn{5}{c}{}\\
+	\multicolumn{5}{c}{Multiplication over $\F_2$}\\
+	\end{tabular}
+\end{table}
+
+Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
+because it is performed modulo $2$.
+Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.
+\end{Example}
+
+Algebraic extensions of the binary field $\F_{2}$  
+are generally termed as {\it binary extension fields} $\Fkk$.
+Where elements in $\F_2$ can only represent $1$ bit, elements in $\Fkk$ 
+represent a $k$-bit vector.
+This allows them to be widely used in digital hardware applications.
+In order to construct a Galois field of the form $\Fkk$, 
+an {\bf irreducible polynomial} is required:
+\begin{Definition}
+A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$ is {\bf irreducible} 
+if $P(x)$ is non-constant with degree $k$ and cannot be 
+factored into a product of polynomials of lower degree in $\mathbb{F}_2[x]$.
+\end{Definition}
+
+Therefore, the polynomial $P(x)$ with degree $k$ is irreducible over 
+$\mathbb{F}_{2}$ if and only if it has no roots in $\mathbb{F}_{2}$,
+i.e if $\forall a \in \mathbb{F}_{2}$, $P(a)\neq 0$.
+For example, $x^2+x+1$ is an irreducible polynomial over $\mathbb{F}_{2}$
+because it has no solutions in $\mathbb{F}_{2}$, i.e. $(0)^2+(0)+1=1\neq0$ 
+and $(1)^2+(1)+1=1\neq0$ over $\F_2$.
+Irreducible polynomials exist for any degree $\geq 2$ in $\mathbb{F}_2[x]$.
+
+Given an irreducible polynomial $P(x)$ of degree $k$ in the polynomial ring 
+$\mathbb{F}_2[x]$, we can construct a binary extension field 
+$\mathbb{F}_{2^k} \equiv \mathbb{F}_2[x] \pmod{P(x)}$.
+Let $\alpha$ be a root of $P(x)$, i.e., $P(\alpha)=0$.
+Since $P(x)$ is irreducible over
+$\mathbb{F}_2[x]$, $\alpha \notin \mathbb{F}_2$. 
+Instead, $\alpha$ is an element in $\mathbb{F}_{2^k}$. 
+Any element $A \in \mathbb{F}_{2^k}$ is then represented as: 
+\begin{equation}\label{rep:poly}
+A= \sum_{i=0}^{k-1} (a_i \cdot \alpha^i) = a_0 + a_1\cdot\alpha + \cdots + a_{k-1}\cdot \alpha^{k-1}\nonumber
+\end{equation}
+where $a_i \in \mathbb{F}_2$ are the coefficients and $P(\alpha)=0$.
+
+To better understand this field extension, compare its similarities to another
+common-place
+field extension $\C$, the set of complex numbers. $\C$ is an extension of the field 
+of real numbers $\R$ with an additional element $i=\sqrt{-1}$, which is an imaginary
+root in $\R$.
+Thus $i \notin \R$, rather $i \in \C$.
+Every element $A \in \mathbb{C}$ can be represented as:
+\begin{equation}\label{rep:polyC}
+A=\sum_{j=0}^{1} (a_j \cdot i^j)=a_0+a_1\cdot i
+\end{equation}
+where $a_j \in \R$ are coefficients. Similarly, $\Fkk$ is an extension of $\F_2$ with 
+an additional element $\alpha$, which is the ``imaginary root'' of an irreducible 
+polynomial $P$ in $\F_2[x]$.
+
+Every element $A \in \Fkk$ has a degree less than $k$ because 
+$A$ is always computed modulo $P(x)$, which has degree $k$. 
+Thus, $A\pmod {P(x)}$ can be of degree at most $k-1$ and at least $0$.
+For this reason, the field $\mathbb{F}_{2^k}$ can be viewed as a $k$
+dimensional vector space over $\mathbb{F}_{2}$. 
+The equivalent bit vector representation for element $A$ is:
+\begin{equation}
+A=(a_{k-1} a_{k-2} \cdots a_{0})
+\end{equation}
+
+\begin{Example}
+A 4-bit Boolean vector, $(a_{3} a_{2} a_{1} a_{0})$
+can be presented over $\F_{2^4}$ as: 
+\begin{equation}
+a_3 \cdot \alpha^3+a_2 \cdot \alpha^2+a_1 \cdot \alpha+a_0
+\end{equation}
+For instance, the Boolean vector $1011$ is represented as the element 
+$\alpha^3+\alpha+1$.
+\end{Example}
+
+\begin{Example}\label{exp:1}
+Let us construct $\mathbb{F}_{2^4}$ as $\mathbb{F}_2[x] \pmod{ P(x)}$, where
+$P(x)=x^4+x^3+1 \in \mathbb{F}_2[x]$ is an irreducible polynomial of degree $k=4$. 
+Let $\alpha$ be the root of $P(x)$, i.e. $P(\alpha)=0$.
+
+Any element $A \in \mathbb{F}_2[x] \pmod{ x^4 + x^3 + 1}$
+has a representation of the type: $A = a_3 x^3 + a_2 x^2 +
+a_1 x + a_0$ (degree $< 4$) where the coefficients $a_3, \dots, a_0$ are in $\F_2 =
+\{0, 1\}$. Since there are only $16$ such polynomials, we obtain
+$16$ elements in the field $\mathbb{F}_{2^4}$. Each element in
+$\mathbb{F}_{2^4}$ can then be viewed as a $4$-bit vector over $\mathbb{F}_{2}$. 
+Each element also has an exponential $\alpha$
+representation. All three representations are shown in Table
+\ref{tab:gfelement}.
+
+\begin{table}[h]
+\begin{center}
+\caption{Bit-vector, Exponential and Polynomial representation of
+elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$}\label{tab:gfelement} 
+\begin{tabular}{|c|c|c||c|c|c|} 
+\hline
+$a_3a_2a_1a_0$ & Exponential & Polynomial     &$a_3a_2a_1a_0$ & Exponential & Polynomial  \\
+\hline
+$0000$        & $0$         & $0$            & $1000$ & $\alpha^3$ &  $\alpha^3$\\
+\hline
+$0001$        & $1$         & $1$            & $1001$ & $\alpha^4$ & $\alpha^3 + 1$\\
+\hline
+$0010$        & $\alpha$    & $\alpha$       & $1010$ & $\alpha^{10}$&$\alpha^3 + \alpha$  \\
+\hline
+$0011$        & $\alpha^{12}$& $\alpha + 1$   & $1011$ & $\alpha^5$ & $\alpha^3+\alpha+1$\\
+\hline
+$0100$        & $\alpha^2$  & $\alpha^2$     &  $1100$ & $\alpha^{14}$ & $\alpha^3 + \alpha^2$\\
+\hline
+$0101$        & $\alpha^9$   &$\alpha^2 + 1$ & $1101$  &$\alpha^{11}$  & $\alpha^3+\alpha^2+1$\\
+\hline
+$0110$        & $\alpha^{13}$& $\alpha^2 + \alpha$ & $1110$ & $\alpha^8$& $\alpha^3+\alpha^2+\alpha$\\
+\hline
+$0111$        &$\alpha^7 $ & $\alpha^2+\alpha+1$ & $1111$ &$\alpha^6$ & $\alpha^3+\alpha^2+\alpha+1$\\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+We can compute the polynomial representation from the exponential representation.
+Since every element is computed $\pmod{P(\alpha)} = \pmod{\alpha^4+\alpha^3+1}$, 
+we compute the element $\alpha^{4}$ as 
+\begin{equation}
+\alpha^{4} \pmod{ \alpha^4+\alpha^3+1} = -\alpha^3 - 1 = \alpha^3+1
+\end{equation}
+Recall that all coefficients of $\F_{2^4}$ 
+are in $\F_{2}$ where $-1 = +1$ modulo 2.
+The next element $\alpha^{5}$ can be computed as 
+\begin{equation}
+\alpha^{5} = \alpha^{4}\cdot \alpha = (\alpha^3+1)\cdot \alpha = \alpha^4+\alpha = \alpha^3+\alpha+1 
+\end{equation}
+Then $\alpha^6$ can be computed as $\alpha^{5}*\alpha$ and so on.
+\end{Example}
+
+An irreducible polynomial can also be a primitive polynomial.
+
+\begin{Definition}
+A {\bf primitive polynomial} $P(x)$ is a polynomial with coefficients in $\mathbb{F}_2$ 
+which has a root $\alpha$ $\in$ $\mathbb{F}_{2^k}$
+such that \{$0$, $1(=\alpha^{{2^k}-1})$, $\alpha$, $\alpha^2$, $\cdots$, $\alpha^{2^k-2}$\} is the set of 
+all elements in $\mathbb{F}_{2^k}$, 
+where $\alpha$ is a {\bf primitive element} of $\mathbb{F}_{2^k}$. 
+\end{Definition}
+
+A primitive polynomial is guaranteed to generate all distinct elements 
+of a finite field $\mathbb{F}_{2^k}$ while an irreducible polynomial
+has no such guarantee.
+Often, there exists more than one irreducible polynomial of degree $k$.
+In such cases, any degree $k$ irreducible polynomial can be 
+used for field construction. For example, both $x^3+x+1$ and $x^3+x^2+1$ 
+are irreducible in $\mathbb{F}_2$ and either one can be used
+to construct $\mathbb{F}_{2^3}$. This is due to the following:
+
+\begin{Theorem}\label{the:unique}
+There exist a {\bf unique} field $\mathbb{F}_{p^k}$, for any prime $p$ and any positive integer $k$.
+\end{Theorem}
+
+Theorem \ref{the:unique} implies that Galois fields with the same number of elements are 
+{\bf isomorphic} to each other up to the labeling of the elements.
+
+Theorem \ref{the:fer} provides an important property for investigating solutions to
+polynomial equations in $\Fq$.
+
+\begin{Theorem}\label{the:fer}
+ $\left[Generalized\  Fermat's\  Little\  Theorem \right]$ Given a
+ Galois field $\mathbb{F}_{q}$, each element $A \in \mathbb{F}_{q}$ satisfies: 
+\begin{eqnarray}\label{fe}
+ A^{q} & \equiv & A  \nonumber \\
+ A^{q} - A & \equiv& 0  
+\end{eqnarray}
+\end{Theorem}
+
+We can extend Theorem \ref{the:fer} to polynomials in $\mathbb{F}_{q}[x]$ as 
+follows: 
+\begin{Definition}
+Let $x^q-x$ be a polynomial in $\mathbb{F}_{q}[x]$.
+Every element $A \in \mathbb{F}_{q}$ is a solution to  $x^q-x=0$. 
+Therefore, $x^{q} - x$ always {\it vanishes} in $\mathbb{F}_{q}$. Such 
+polynomials are called {\bf vanishing polynomials} of the field $\mathbb{F}_{q}$.
+\end{Definition}
+
+\begin{Example}
+Given $\mathbb{F}_{2^2} =\{0,1,\alpha,\alpha+1\}$ with $P(x)=x^2+x+1$, where $P(\alpha)=0$. 
+ \begin{eqnarray}
+ 0^{2^2}&=&0 \nonumber \\
+ 1^{2^2}&=&1 \nonumber \\
+ \alpha^{2^2}&=&\alpha \pmod {\alpha^2+\alpha+1}\nonumber \\
+ (\alpha+1)^{2^2}&=&\alpha+1 \pmod {\alpha^2+\alpha+1} \nonumber 
+ \end{eqnarray}
+\end{Example}
+
+A Galois field $\F_q$ can be fully contained within a larger field $\F_{q^k}$.
+That is, $\F_q \subset \F_{q^k}$.
+For example, Fig \ref{fig:contain2_4_16} shows the containment of the fields 
+$\F_2 \subset \F_4 \subset \F_{16}$. It's easy to see that since $\F_4=\F_{2^2}$, it
+contains $\F_2$. Likewise $\F_{16}=\F_{4^2}=\F_{2^4}$ contains $\F_4$ and $\F_2$.
+The elements $\{0,1,\alpha,\dots,\alpha^{14}\}$
+designate $\F_{16}$. Of these, $\{0,1,\alpha^5,\alpha^{10}\}$ create $\F_4$.
+From these, only $\{0,1\}$ exist in $\F_2$.
+
+\begin{Theorem}
+$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, i.e. if $n$ divides $m$.
+\end{Theorem}
+
+Therefore:
+\begin{itemize}
+\item $\F_2 \subset \F_{2^2} \subset \F_{2^4} \subset \F_{2^8} \subset \dots$
+\item $\F_2 \subset \F_{2^3} \subset \F_{2^9} \subset \F_{2^{27}} \subset \dots$
+\item $\F_2 \subset \F_{2^5} \subset \F_{2^{25}} \subset \F_{2^{125}} \subset \dots,$ and so on
+\end{itemize}
+
+\begin{Definition}
+The {\bf algebraic closure} of the Galois field $\F_{2^k}$, denoted $\overline{\Fkk}$, is the 
+union of all fields $\F_{2^n}$ such that $k \mid n$.
+\end{Definition}

### 2016-11-15 dissertation/unsat.tex
dissertation version 2_part prelim update 3

@@
 
-\begin{algorithm}[hbt]
+\IncMargin{1em}
+\begin{algorithm}[H]
 \SetAlgoNoLine
-	\KwIn{$M$ is the original machine, $p$ is the property to check, $k$ is the number of steps in $k$-BMC}
+\LinesNumbered
+	\KwIn{$M$ -- original machine, $p$ -- property to check, $k$ -- \# of steps in $k$-BMC}
+	\KwOut{If $p$ is violated, return error trace; otherwise $p$ is valid on $M$}
   $k = $ InitValue\;
+
   \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT}}
@@
 \end{algorithm}
+\DecMargin{1em}
+
 \begin{figure}[hbt]

### 2016-11-21 dissertation/normal.tex
dissertation:part reacha:update 1
dissertation: milestone 1: first draft
dissertation v2_finish prelims

@@
 
-\section{Normal Basis Multiplier over Galois Field}
-From algebraic view, a field is a space, and field elements are dots in the space. Those elements can be 
-represented with unique coordinates, which requires the pre-definition of a basis vector. In this
-section, we discuss a special basis called normal basis, as well as the advantages adopting it in 
-GF operations esp. multiplication.
-\subsection{Normal Basis}
-Given a Galois field (GF) $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
-Its elements can be written in polynomials of $\alpha$, when there is an irreducible polynomial $p(\alpha)$
-defined.
-
-If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
-to binary bit-vector representations by recording the coefficients. For example,
-
-\begin{table}[H]
-\centering
-\caption{Bit-vector, Exponential and Polynomial representation of
-elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
-\pmod{x^4+x^3+1}$}
-\begin{tabular}{|c|c||c|c|} 
-\hline
-$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
-\hline
-$0000$        & $0$           & $1000$  &$\alpha^3$\\
-\hline
-$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
-\hline
-$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
-\hline
-$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
-\hline
-$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
-\hline
-$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
-\hline
-$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
-\hline
-$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
-\hline
-\end{tabular}
-\label{table:booltogalois}  
-\end{table}
-
-Basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$ is called {\bf standard basis} (StdB), which results in
-a straightforward representation for elements, and operations of elements such as addition and subtraction.
-The addition/subtraction of GF elements in StdB follows the rules of polynomial addition/subtraction
-where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} (XOR) in
-Boolean algebra, element $A$ add/subtract by element $B$ in StdB is defined as
-\begin{align}\label{eqn:StdB}
-A+B = A-B &= (a_0,a_1,\dots,a_{k-1})_{StdB} \xor (b_0,b_1,\dots,b_{k-1})_{StdB} \nonumber\\
-&=(a_0\oplus b_0, a_1\oplus b_1,\dots,a_{k-1}\oplus b_{k-1})_{StdB} 
-\end{align}
-
-\subsection{Multiplication using Normal Basis}
-Besides addition/subtraction, multiplication is also very common in arithmetic circuit design.
-The multiplication of GF elements in $\Fkk$ in StdB follows the rule of polynomial multiplication.
-However, it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
-in bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
-the area and delay of circuit will be costly.
-
-In order to lower down the complexity of arithmetic circuit design, Massey and Omura \cite{MasseyOmura} % ref 7 in RH paper
-use a new basis to represent GF elements, which is called {\bf normal basis} (NB).
-A normal basis over $\Fkk$ is written in the form of
-\begin{equation*}
-N.B. ~~~ \N = \{\beta,\beta^2,\beta^4,\beta^8,\dots,\beta^{2^{k-1}}\}
-\end{equation*}
-Respectively, a field element in NB representation is actually
-\begin{align*}
-A &= (a_0,a_1,\dots,a_{k-1})_{NB} \\
-  &= a_0\beta+a_1\beta^2+\cdots+a_{k-1}\beta^{2^{k-1}} \\
-  &= \sum_{i=0}^{k-1} a_i\beta^{2^i}
-\end{align*}
-
-According to the definition, a normal basis is a vector where the next entry is the square of the former one.
-We note that the vector is cyclic, i.e. $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
-{\bf Normal element} $\beta$ is an element from the field which is used to construct the normal basis,
-and can be represent as a power of primitive element $\alpha$: 
-\begin{equation*}
-\beta = \alpha^t, ~~~ 1\leq t<2^k
-\end{equation*}
-
-The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
-However, what makes NB powerful is its property when doing multiplications and exponentiations.
-The following lemmas and examples illustrate this fabulous property very well.
-\begin{Lemma}[Square of NB]
-\label{lem:squareNB}
-In $\Fkk$, equation 
-\begin{equation*}
-(a+b)^2 = a^2 + b^2
-\end{equation*}
-has been proved. According to the \textbf{binomial theorem}, it can be extended to
-\begin{align*}
-&(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}})^2 \\
-&= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
-&= b_{k-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}}
-\end{align*}
-\end{Lemma}
-This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
-Obviously, StdB representation does not have this benefit.
-
-\begin{Example}[Square of NB]
-In GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $x^3 + x + 1$, the standard basis is denoted as 
-$\{ 1, \alpha, \alpha^2\}$ where $\alpha^3+\alpha+1=0$.
-Let $\beta = \alpha^3$, then $\N = \{ \beta, \beta^2, \beta^4\}$ forms a normal basis. 
-Write down element $E$ using both representations:
-\begin{align*}
-E &= (a_0,a_1,a_2)_{StdB} = (b_0,b_1,b_2)_{NB} \\
-  &= a_0 + a_1\alpha + a_2\alpha^2 = b_0\beta + b_1\beta^2 + b_2\beta^4
-\end{align*}
-Compute the square of $E$ in StdB first:
-\begin{align*}
-E^2 &= a_0 + a_1\alpha^2 + a_2\alpha^4 \\
-    &= a_0 + a_2\alpha + (a_1 + a_2)\alpha^2 \\
-    &= (a_0,a_2,a_1+a_2)_{StdB}
-\end{align*}
-When it is computed in NB, we can make it very simple:
-\begin{align*}
-E^2 &= \overset{\xrightarrow{Cyclic~~shift}}{(b_0,b_1,b_2)}_{NB} \\
-	&= (b_2,b_0,b_1)_{NB}
-\end{align*}
-\end{Example}
-
-This example shows that convenience to use NB when computing $2^k$ power of an element.
-Multiplication is a bit complicated than squaring; but when it is decomposed as bit-wise
-operations, the property in lemma \ref{lem:squareNB} can be well utilized.
-
-\begin{Example}[Bit-wise NB multiplication]
-Assume there are 2 binary vectors representing 2 operands in NB over $\Fkk$: 
-$A = (a_0, a_1, \dots, a_{k-1}), B = (b_0, b_1, \dots, b_{k-1})$. Note that in this example, 
-by default we use normal basis representation so subscript ``NB" is skipped. Their product can also be written
-as: $$C = A\times B = (c_0, c_1, \dots, c_{k-1})$$
-
-Assume the most significant bit (MSB) of the product can be represented by a function $f_{mult}$: 
-\begin{equation}
-\label{eqn:multiMSB}
-c_{n-1} = f_{mult}(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})
-\end{equation}
-Before discussing the details of the function $f_{mult}$, we can take 
-a square on both side of equation \ref{eqn:multiMSB}, i.e. $C^2 = A^2\times B^2$.
-Obviously, using the property in lemma \ref{lem:squareNB}, the original second most significant bit 
-becomes the new MSB because of right-cyclic shifting. 
-Concretely, 
-$$(c_{k-1},c_0,c_1,\dots,c_{k-2}) = (a_{k-1},a_0,a_1,\dots,a_{k-2})\times(b_{k-1},b_0,b_1,\dots,b_{k-2})$$
-Note $A^2, B^2$ and $C^2$ still belong to $\Fkk$, thus as a universal function implementing MSB multiplication
-over $\Fkk$, $f_{mult}$ still keeps the same. As a result, the new MSB can be written as 
-\begin{equation}
-\label{eqn:shiftMSB}
-c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
-\dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})
-\end{equation}
-Similarly, if we take a square again on the new equation, we can get $c_{k-3}$.
-Successively we can derive all bits of product $C$ using the same function $f_{mult}$, and the only
-adjustment we need to make is to right-cyclic shift 2 operands by 1 bit each time.
-\end{Example}
-
-From above example, it is proved that a universal structure that implements $f_{mult}$ can be reused
-for $k$ times in NB multiplication over $\Fkk$. Comparing to StdB, which requires distinct design 
-for every bit of multiplication, NB is less costly if we can prove $f_{mult}$ will not result in 
-a structure with $O(k^2)$ complexity. So our next mission is to explore the details of $f_{mult}$
-to prove it will be a relatively simple design with complexity lower than $O(k^2)$.
-
-If we want to make the complexity of $f_{mult}$ lower than $O(k^2)$, then the best choice is to try out 
-linear functions. As we know, matrix multiplication can simulate all possible combinations of 
-linear functions (which is also the reason it is used as basic model to simulate the behavior of a neuron
-in neural network machine learning algorithms). Imagine $A$ is a $k$-bit row vector and $B$ is a $k$-bit 
-column vector, then the single bit product can be written as the product of matrix multiplication
-$$c_{l} = A\times C\times B$$
-where $C$ is a $k\times k$ square matrix.
-
-\begin{Definition}[$\lambda$-Matrix]
-A binary $k\times k$ matrix $M$ is used to describe the bit-wise normal basis multiplication function $f_{mult}$ where
-\begin{equation}
-\label{eqn:def_lambda}
-c_{l} = f_{mult}(A, B) = A \times M \times B^T
-\end{equation}
-$B^T$ denotes vector transposition. Matrix $M$ is called $\lambda$-Matrix of
-$k$-bit NB multiplication over $\Fkk$.
-\end{Definition}
-
-When taking different bits $l$ of the product in equation \ref{eqn:def_lambda}, 
-we obtain a series of conjugate matrices of $M$. 
-Which means instead of shifting operands $A$ and $B$, we can also shift the 
-matrix.
-
-More specifically, we denote the matrix by \emph{$l$-th $\lambda$-Matrix} as 
-$$c_l = A \times M^{(l)} \cdot B^T$$
-Meanwhile, the operator shifting rule in equation \ref{eqn:shiftMSB} still holds. Then we have relation 
-$$c_{l-1} = A \cdot M^{(l-1)} \cdot B^T = shift(A) \cdot M^{(l)} \cdot shift(B)^T$$ which means
-by right and down cyclically shifting $M^{(l-1)}$, we can get $M^{(l)}$.
-
-\begin{Example}[NB multiplication using $\lambda$-Matrix]
-Over GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $\alpha^3 + \alpha + 1$, let normal element $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
-forms a normal basis. Corresponding $0$-th $\lambda$-Matrix is
-\begin{equation*}
-M^{(0)} = \left(
-\begin{array} {lcr}
-0 & 1 & 0\\
-1 & 0 & 1\\
-0 & 1 & 1
-\end{array} \right).
-\end{equation*}
-i.e.,
-\begin{equation*}
-c_0 = (a_0\  a_1\  a_2)\left(
-\begin{array} {lcr}
-0 & 1 & 0\\
-1 & 0 & 1\\
-0 & 1 & 1
-\end{array} \right)\left(
-\begin{array} {lcr}
-b_0\\
-b_1\\
-b_2
-\end{array} \right)
-\end{equation*}
-From $0$-th $\lambda$-Matrix we can directly write down all remaining $\lambda$-Matrices:
-\begin{equation*}
-M^{(1)} = \left(
-\begin{array} {lcr}
-1 & 0 & 1\\
-0 & 0 & 1\\
-1 & 1 & 0
-\end{array} \right)~~~~~~
-M^{(2)} = \left(
-\begin{array} {lcr}
-0 & 1 & 1\\
-1 & 1 & 0\\
-1 & 0 & 0
-\end{array} \right)
-\end{equation*}
-\end{Example}
-
-If we generalize the definition and explore the nature of $\lambda$-Matrix, it is defined as cross-product terms from multiplication, which is 
-\begin{equation}
-Product~vector~C = (\sum_{i=0}^{k-1}a_i\beta^{2^i})(\sum_{j=0}^{k-1}b_j\beta^{2^j}) = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}
-\end{equation}
-The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
-NB, i.e.
-\begin{equation}
-\beta^{2^i}\beta^{2^j} = \sum_{l=0}^{k-1}\lambda_{ij}^{(l)}\beta^{2^l}, \ \ \lambda_{ij}^{(l)} \in \mathbb F_2.
-\end{equation}
-Substitution yields, result is an expression for l-th digit of product as showed in equation \ref{eqn:multiMSB}:
-\begin{equation}
-c_l = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(l)}a_ib_j
-\end{equation}
-$\lambda_{ij}^{(l)}$ is the entry with coordinate $(i,j)$ in $l$-th $\lambda$-Matrix.
-
-The $\lambda$-Matrix can be implemented with XOR and AND gates in circuit design.
-The very naive implementation requires $O(C_N)$ gates, where $C_N$ is the number of
-nonzero entries in $\lambda$-Matrix.
-There usually exists multiple NBs in $\Fkk, k>3$. If we employ a random NB, there is no mathematical
-guarantee that $C_N \sim o(k)$ (symbol $o$ denotes ``strictly lower than bound"). However,
-Mullin et al. \cite{mullinONB} % This citation is valid
-proves that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
-its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
-called optimal normal basis (ONB), details are introduced in appendix \ref{append:ONB}.
-
-In practice, large size NB multipliers are usually designed in $\Fkk$ when ONB exists
-to minimized the number of gates. So in the following part of this chapter and our experiments,
-we only focus on ONB multipliers instead of general NB multipliers.
-
-% After fixing this, add a whole-piece StdB vs NB example, list their cost as the conclusion
-\subsection{Comparison between Standard Basis and Normal Basis}
-At the end of this section, a detailed example is used to make a comparison between StdB multiplication
-and NB multiplication.
-\begin{Example}[Rijndael's finite field]
-Rijndael uses a characteristic 2 finite field with 256 elements, which can also be called the GF $\mathbb F_{2^8}$.
-Let us define the primitive element $\alpha$ using irreducible polynomial 
-$\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$. Coincidently, $\alpha$ is also a normal element,
-i.e. $\beta = \alpha$ can construct a NB $\{\alpha,\alpha^2,\alpha^4,\alpha^8,\alpha^{16},\alpha^{32},\alpha^{64},
-\alpha^{128}\}$.
-
-We pick a pair of elements from the Rijndael's field: $A=(0100~1011)_{StdB} = (4B)_{StdB},~B=(1100~1010)_{StdB} = (CA)_{StdB}$.
-First let us compute their product in StdB, the rule follows ordinary polynomial multiplication.
-
-\begin{align*}
-A\cdot B &= (\alpha^6+\alpha^3+\alpha+1)(\alpha^7+\alpha^6+\alpha^3+\alpha)\\
-&= (\alpha^{13}+\alpha^{10}+\alpha^8+\alpha^7)+(\alpha^{12}+\alpha^9+\alpha^7+\alpha^6)+
-	(\alpha^9+\alpha^6+\alpha^4+\alpha^3)\\
-	&~~~+(\alpha^7+\alpha^4+\alpha^2+\alpha) \\
-&= \alpha^{13}+\alpha^{12}+\alpha^{10}+\alpha^8+\alpha^7+\alpha^3+\alpha^2+\alpha
-\end{align*}
-Note that this polynomial is not the final form of the product because it needs to be
-reduced modulo irreducible polynomial $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$.
-This can be done using base-2 long division. Note the dividend and divisor are written in pseudo Boolean
-vectors, not real Boolean vectors in any kind of bases.
-
-% The folowing part is customized latex code mimicing binary long division
-% Note \divrule coordinates may not reflect the intuitive column # in tabular. Adjust by urself!!!
-
-\vspace{1cm}
-
-\newdimen\digitwidth
-\settowidth\digitwidth{0}
-\def~{\hspace{\digitwidth}}
-
-\def\divrule#1#2{%
-\noalign{\moveright#1\digitwidth%
-\vbox{\hrule width#2\digitwidth}}}
-111010111\,\begin{tabular}[b]{@{}r@{}}
-101001 \\ \hline
-\big)\begin{tabular}[t]{@{}l@{}}
-11010110001110 \\
-111010111 \\ \divrule{0}{14}
-~~111101101 \\
-~~111010111 \\ \divrule{2}{12}
-~~~~~111010110 \\
-~~~~~111010111 \\ \divrule{5}{9}
-~~~~~~~~~~~~~1
-\end{tabular}
-\end{tabular}
-\vspace{0.5cm}
-
-The final remainder is $1$, i.e. the product equals to 1 in StdB.
-
-On the other hand, operands $A$ and $B$ can be written in NB as
-$$A = (0010~1001)_{NB},~~B = (0100~0010)_{NB}$$
-The $\lambda$-Matrix for $\mathbb F_{2}[x] \pmod{x^8+x^7+x^6+x^4+x^2+x+1}$
-is
-(Computation of $\lambda$-Matrix refers to appendix \ref{append:NB})
-\begin{equation*}
-M^{(0)} = \left(\begin{array}{lccccccr}
-0 &0 &0 &0 &1 &0 &1 &1 \\
-0 &0 &1 &1 &1 &1 &0 &0 \\
-0 &1 &0 &0 &0 &0 &1 &0 \\
-0 &1 &0 &0 &1 &1 &0 &1 \\
-1 &1 &0 &1 &0 &1 &0 &0 \\
-0 &1 &0 &1 &1 &0 &0 &1 \\
-1 &0 &1 &0 &0 &0 &0 &0 \\
-1 &0 &0 &1 &0 &1 &0 &1
-\end{array}\right)
-\end{equation*}
-Taking matrix multiplication $c_0 = A\times M^{(0)}\times B^T$,
-the result is $c_0 = 1$. Then by cyclic shifting $A$ and $B$
-(or shifting $M^{(0)}$, either is applicable), we can successively
-obtain other bits of product. The final answer is
-$$C = (0000~0001)_{NB}$$
-It is equivalent to the result in StdB.
-\end{Example}
-
-From the intuition of humans, StdB multiplication is straightforward and easier to understand
-while NB is difficult to comprehend. However, if we implement both multiplications to 
-hardware multipliers, it will be clear which side a circuit designer prefers.
-
-% May consider providing details?
-
-Mastrovito multiplier and Montgomery multiplier are 2 common designs
-of GF multipliers using StdB. As a naive implementation of GF multiplication,
-Mastrovito multiplier uses most number of gates:
-$k^2$ AND gates plus $k^2-\Delta$ XOR gates \cite{Mastrovito}. Montgomery multiplier 
-applies lazy reduction techniques and results in a better latency performance, while the number of gates are about
-the same with Mastrovito multiplier:
-$k^2$ AND gates plus $k^2-k/2$ XOR gates \cite{Montgomery}. 
-Concretely, typical design of Mastrovito multiplier consists of 218 logic gates, while 
-Montgomery multiplier needs 198 gates. However, the NB multiplier reuses the $\lambda$-Matrix 
-logic, so this component will only need to be implemented for once. 
-Consider the definition of matrix multiplication, it needs $C_N$ AND gates to apply 
-bit-wise multiplication and $C_N-1$ XOR gates to sum the intermediate products up. The number of nonzero entries
-in the $\lambda$-Matrix can be counted: $C_N = 27$.
-As a result, the most naive NB multiplier design (or Massey-Omura multiplier \cite{MasseyOmura})
-contains 53 gates in total, which is a great saving in area cost comparing to StdB multipliers.
-
-\section{Design a Normal Basis Multiplier on Gate Level}
-The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
-we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
-very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
-test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
-routine of several kinds of modern NB multipliers.
-\subsection{Sequential Multiplier with Parallel Outputs}
-The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
-the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.
-
-% fig:MasseyOmura
-
-Figure \ref{fig:MasseyOmura} shows the basic architecture of a Massey-Omura multiplier. The operands
-$A$ and $B$ are 2 arrays of flip-flops which allow 1-bit right-cyclic shift every clock cycle.
-The logic gates in the boxes implements the matrix multiplication with $\lambda$-Matrix $M^{(0)}$, 
-while each AND gate corresponds to term $a_ib_j$ and each XOR gate corresponds to addition $a_ib_j+a_{i'}b_{j'}$. 
-The XOR layer has only 1 output, giving out 1 bit of product $C$ every clock cycle.
-
-The behavior of Massey-Omura multiplier can be concluded as: pre-load operands $A,B$ and reset $C$ to 0, after 
-executing for $k$ clock cycles, the data stored in flip-flop array $C$ is the product $A\times B$.
-We note that there is only one output giving 1 bit of the product each clock cycle, which matches the 
-definition of serial output to communication channel. Therefore this type of design is named as 
-sequential multiplier with serial output (SMSO).
-The SMSO architecture need $C_N$ AND gates and $C_N-1$ XOR gates, which equals to $2k-1$ AND gates and $2k-2$ XOR gates
-if it is designed using ONB. In fact, the number of gates can be reduced if the multiplication is 
-implemented using a conjugate of SMSO.
-
-The gate-level logic boxes are implementing following function:
-\begin{equation}
-\label{eqn:SMPOterms}
-c_l = row_1(A\times M^{(l)}) \times B + row_2(A\times M^{(l)}) \times B + \cdots + row_{k}(A\times M^{(l)}) \times B
-\end{equation}
-It can be decomposed into $k$ terms. If we only compute one term for each $c_l,~0\leq l\leq k-1$ in one 
-clock cycle, make $k$ outputs and add them up using shift register after $k$ clock cycles, 
-it will generate the same result with SMSO. This kind of architecture is named as 
-sequential multiplier with parallel outputs (SMPO). The basic SMPO, as a conjugate of Massey-Omura multiplier,
-is invented by Agnew et al. \cite{agnew1991implementation}.
-
-\begin{figure}[H]
-\centering{
-\includegraphics[width=\textwidth]{newfig/mySMPO.eps}
-\caption{5-bit Agnew's SMPO.
+% \section{Normal Basis Multiplier over Galois Field}
+% From algebraic view, a field is a space, and field elements are dots in the space. Those elements can be 
+% represented with unique coordinates, which requires the pre-definition of a basis vector. In this
+% section, we discuss a special basis called normal basis, as well as the advantages adopting it in 
+% GF operations esp. multiplication.
+% \subsection{Normal Basis}
+% Given a Galois field (GF) $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
+% Its elements can be written in polynomials of $\alpha$, when there is an irreducible polynomial $p(\alpha)$
+% defined.
+% 
+% If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
+% to binary bit-vector representations by recording the coefficients. For example,
+% 
+% \begin{table}[H]
+% \centering
+% \caption{Bit-vector, Exponential and Polynomial representation of
+% elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
+% \pmod{x^4+x^3+1}$}
+% \begin{tabular}{|c|c||c|c|} 
+% \hline
+% $a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
+% \hline
+% $0000$        & $0$           & $1000$  &$\alpha^3$\\
+% \hline
+% $0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
+% \hline
+% $0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
+% \hline
+% $0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
+% \hline
+% $0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
+% \hline
+% $0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
+% \hline
+% $0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
+% \hline
+% $0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
+% \hline
+% \end{tabular}
+% \label{table:booltogalois}  
+% \end{table}
+% 
+% Basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$ is called {\bf standard basis} (StdB), which results in
+% a straightforward representation for elements, and operations of elements such as addition and subtraction.
+% The addition/subtraction of GF elements in StdB follows the rules of polynomial addition/subtraction
+% where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} (XOR) in
+% Boolean algebra, element $A$ add/subtract by element $B$ in StdB is defined as
+% \begin{align}\label{eqn:StdB}
+% A+B = A-B &= (a_0,a_1,\dots,a_{k-1})_{StdB} \xor (b_0,b_1,\dots,b_{k-1})_{StdB} \nonumber\\
+% &=(a_0\oplus b_0, a_1\oplus b_1,\dots,a_{k-1}\oplus b_{k-1})_{StdB} 
+% \end{align}
+% 
+% \subsection{Multiplication using Normal Basis}
+% Besides addition/subtraction, multiplication is also very common in arithmetic circuit design.
+% The multiplication of GF elements in $\Fkk$ in StdB follows the rule of polynomial multiplication.
+% However, it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
+% in bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
+% the area and delay of circuit will be costly.
+% 
+% In order to lower down the complexity of arithmetic circuit design, Massey and Omura \cite{MasseyOmura} % ref 7 in RH paper
+% use a new basis to represent GF elements, which is called {\bf normal basis} (NB).
+% A normal basis over $\Fkk$ is written in the form of
+% \begin{equation*}
+% N.B. ~~~ \N = \{\beta,\beta^2,\beta^4,\beta^8,\dots,\beta^{2^{k-1}}\}
+% \end{equation*}
+% Respectively, a field element in NB representation is actually
+% \begin{align*}
+% A &= (a_0,a_1,\dots,a_{k-1})_{NB} \\
+%   &= a_0\beta+a_1\beta^2+\cdots+a_{k-1}\beta^{2^{k-1}} \\
+%   &= \sum_{i=0}^{k-1} a_i\beta^{2^i}
+% \end{align*}
+% 
+% According to the definition, a normal basis is a vector where the next entry is the square of the former one.
+% We note that the vector is cyclic, i.e. $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
+% {\bf Normal element} $\beta$ is an element from the field which is used to construct the normal basis,
+% and can be represent as a power of primitive element $\alpha$: 
+% \begin{equation*}
+% \beta = \alpha^t, ~~~ 1\leq t<2^k
+% \end{equation*}
+% 
+% The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
+% However, what makes NB powerful is its property when doing multiplications and exponentiations.
+% The following lemmas and examples illustrate this fabulous property very well.
+% \begin{Lemma}[Square of NB]
+% \label{lem:squareNB}
+% In $\Fkk$, equation 
+% \begin{equation*}
+% (a+b)^2 = a^2 + b^2
+% \end{equation*}
+% has been proved. According to the \textbf{binomial theorem}, it can be extended to
+% \begin{align*}
+% &(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}})^2 \\
+% &= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
+% &= b_{k-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}}
+% \end{align*}
+% \end{Lemma}
+% This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
+% Obviously, StdB representation does not have this benefit.
+% 
+% \begin{Example}[Square of NB]
+% In GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $x^3 + x + 1$, the standard basis is denoted as 
+% $\{ 1, \alpha, \alpha^2\}$ where $\alpha^3+\alpha+1=0$.
+% Let $\beta = \alpha^3$, then $\N = \{ \beta, \beta^2, \beta^4\}$ forms a normal basis. 
+% Write down element $E$ using both representations:
+% \begin{align*}
+% E &= (a_0,a_1,a_2)_{StdB} = (b_0,b_1,b_2)_{NB} \\
+%   &= a_0 + a_1\alpha + a_2\alpha^2 = b_0\beta + b_1\beta^2 + b_2\beta^4
+% \end{align*}
+% Compute the square of $E$ in StdB first:
+% \begin{align*}
+% E^2 &= a_0 + a_1\alpha^2 + a_2\alpha^4 \\
+%     &= a_0 + a_2\alpha + (a_1 + a_2)\alpha^2 \\
+%     &= (a_0,a_2,a_1+a_2)_{StdB}
+% \end{align*}
+% When it is computed in NB, we can make it very simple:
+% \begin{align*}
+% E^2 &= \overset{\xrightarrow{Cyclic~~shift}}{(b_0,b_1,b_2)}_{NB} \\
+% 	&= (b_2,b_0,b_1)_{NB}
+% \end{align*}
+% \end{Example}
+% 
+% This example shows that convenience to use NB when computing $2^k$ power of an element.
+% Multiplication is a bit complicated than squaring; but when it is decomposed as bit-wise
+% operations, the property in lemma \ref{lem:squareNB} can be well utilized.
+% 
+% \begin{Example}[Bit-wise NB multiplication]
+% Assume there are 2 binary vectors representing 2 operands in NB over $\Fkk$: 
+% $A = (a_0, a_1, \dots, a_{k-1}), B = (b_0, b_1, \dots, b_{k-1})$. Note that in this example, 
+% by default we use normal basis representation so subscript ``NB" is skipped. Their product can also be written
+% as: $$C = A\times B = (c_0, c_1, \dots, c_{k-1})$$
+% 
+% Assume the most significant bit (MSB) of the product can be represented by a function $f_{mult}$: 
+% \begin{equation}
+% \label{eqn:multiMSB}
+% c_{n-1} = f_{mult}(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})
+% \end{equation}
+% Before discussing the details of the function $f_{mult}$, we can take 
+% a square on both side of equation \ref{eqn:multiMSB}, i.e. $C^2 = A^2\times B^2$.
+% Obviously, using the property in lemma \ref{lem:squareNB}, the original second most significant bit 
+% becomes the new MSB because of right-cyclic shifting. 
+% Concretely, 
+% $$(c_{k-1},c_0,c_1,\dots,c_{k-2}) = (a_{k-1},a_0,a_1,\dots,a_{k-2})\times(b_{k-1},b_0,b_1,\dots,b_{k-2})$$
+% Note $A^2, B^2$ and $C^2$ still belong to $\Fkk$, thus as a universal function implementing MSB multiplication
+% over $\Fkk$, $f_{mult}$ still keeps the same. As a result, the new MSB can be written as 
+% \begin{equation}
+% \label{eqn:shiftMSB}
+% c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
+% \dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})
+% \end{equation}
+% Similarly, if we take a square again on the new equation, we can get $c_{k-3}$.
+% Successively we can derive all bits of product $C$ using the same function $f_{mult}$, and the only
+% adjustment we need to make is to right-cyclic shift 2 operands by 1 bit each time.
+% \end{Example}
+% 
+% From above example, it is proved that a universal structure that implements $f_{mult}$ can be reused
+% for $k$ times in NB multiplication over $\Fkk$. Comparing to StdB, which requires distinct design 
+% for every bit of multiplication, NB is less costly if we can prove $f_{mult}$ will not result in 
+% a structure with $O(k^2)$ complexity. So our next mission is to explore the details of $f_{mult}$
+% to prove it will be a relatively simple design with complexity lower than $O(k^2)$.
+% 
+% If we want to make the complexity of $f_{mult}$ lower than $O(k^2)$, then the best choice is to try out 
+% linear functions. As we know, matrix multiplication can simulate all possible combinations of 
+% linear functions (which is also the reason it is used as basic model to simulate the behavior of a neuron
+% in neural network machine learning algorithms). Imagine $A$ is a $k$-bit row vector and $B$ is a $k$-bit 
+% column vector, then the single bit product can be written as the product of matrix multiplication
+% $$c_{l} = A\times C\times B$$
+% where $C$ is a $k\times k$ square matrix.
+% 
+% \begin{Definition}[$\lambda$-Matrix]
+% A binary $k\times k$ matrix $M$ is used to describe the bit-wise normal basis multiplication function $f_{mult}$ where
+% \begin{equation}
+% \label{eqn:def_lambda}
+% c_{l} = f_{mult}(A, B) = A \times M \times B^T
+% \end{equation}
+% $B^T$ denotes vector transposition. Matrix $M$ is called $\lambda$-Matrix of
+% $k$-bit NB multiplication over $\Fkk$.
+% \end{Definition} 
+% 
+% When taking different bits $l$ of the product in equation \ref{eqn:def_lambda}, 
+% we obtain a series of conjugate matrices of $M$. 
+% Which means instead of shifting operands $A$ and $B$, we can also shift the 
+% matrix.
+% 
+% More specifically, we denote the matrix by \emph{$l$-th $\lambda$-Matrix} as 
+% $$c_l = A \times M^{(l)} \cdot B^T$$
+% Meanwhile, the operator shifting rule in equation \ref{eqn:shiftMSB} still holds. Then we have relation 
+% $$c_{l-1} = A \cdot M^{(l-1)} \cdot B^T = shift(A) \cdot M^{(l)} \cdot shift(B)^T$$ which means
+% by right and down cyclically shifting $M^{(l-1)}$, we can get $M^{(l)}$.
+% 
+% \begin{Example}[NB multiplication using $\lambda$-Matrix]
+% Over GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $\alpha^3 + \alpha + 1$, let normal element $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
+% forms a normal basis. Corresponding $0$-th $\lambda$-Matrix is
+% \begin{equation*}
+% M^{(0)} = \left(
+% \begin{array} {lcr}
+% 0 & 1 & 0\\
+% 1 & 0 & 1\\
+% 0 & 1 & 1
+% \end{array} \right).
+% \end{equation*}
+% i.e.,
+% \begin{equation*}
+% c_0 = (a_0\  a_1\  a_2)\left(
+% \begin{array} {lcr}
+% 0 & 1 & 0\\
+% 1 & 0 & 1\\
+% 0 & 1 & 1
+% \end{array} \right)\left(
+% \begin{array} {lcr}
+% b_0\\
+% b_1\\
+% b_2
+% \end{array} \right)
+% \end{equation*}
+% From $0$-th $\lambda$-Matrix we can directly write down all remaining $\lambda$-Matrices:
+% \begin{equation*}
+% M^{(1)} = \left(
+% \begin{array} {lcr}
+% 1 & 0 & 1\\
+% 0 & 0 & 1\\
+% 1 & 1 & 0
+% \end{array} \right)~~~~~~
+% M^{(2)} = \left(
+% \begin{array} {lcr}
+% 0 & 1 & 1\\
+% 1 & 1 & 0\\
+% 1 & 0 & 0
+% \end{array} \right)
+% \end{equation*}
+% \end{Example}
+% 
+% If we generalize the definition and explore the nature of $\lambda$-Matrix, it is defined as cross-product terms from multiplication, which is 
+% \begin{equation}
+% Product~vector~C = (\sum_{i=0}^{k-1}a_i\beta^{2^i})(\sum_{j=0}^{k-1}b_j\beta^{2^j}) = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}
+% \end{equation}
+% The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
+% NB, i.e.
+% \begin{equation}
+% \beta^{2^i}\beta^{2^j} = \sum_{l=0}^{k-1}\lambda_{ij}^{(l)}\beta^{2^l}, \ \ \lambda_{ij}^{(l)} \in \mathbb F_2.
+% \end{equation}
+% Substitution yields, result is an expression for l-th digit of product as showed in equation \ref{eqn:multiMSB}:
+% \begin{equation}
+% c_l = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(l)}a_ib_j
+% \end{equation}
+% $\lambda_{ij}^{(l)}$ is the entry with coordinate $(i,j)$ in $l$-th $\lambda$-Matrix.
+% 
+% The $\lambda$-Matrix can be implemented with XOR and AND gates in circuit design.
+% The very naive implementation requires $O(C_N)$ gates, where $C_N$ is the number of
+% nonzero entries in $\lambda$-Matrix.
+% There usually exists multiple NBs in $\Fkk, k>3$. If we employ a random NB, there is no mathematical
+% guarantee that $C_N \sim o(k)$ (symbol $o$ denotes ``strictly lower than bound"). However,
+% Mullin et al. \cite{mullinONB} % This citation is valid
+% proves that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
+% its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
+% called optimal normal basis (ONB), details are introduced in appendix \ref{append:ONB}.
+% 
+% In practice, large size NB multipliers are usually designed in $\Fkk$ when ONB exists
+% to minimized the number of gates. So in the following part of this chapter and our experiments,
+% we only focus on ONB multipliers instead of general NB multipliers.
+% 
+% 
+% % After fixing this, add a whole-piece StdB vs NB example, list their cost as the conclusion
+% \subsection{Comparison between Standard Basis and Normal Basis}
+% At the end of this section, a detailed example is used to make a comparison between StdB multiplication
+% and NB multiplication.
+% \begin{Example}[Rijndael's finite field]
+% Rijndael uses a characteristic 2 finite field with 256 elements, which can also be called the GF $\mathbb F_{2^8}$.
+% Let us define the primitive element $\alpha$ using irreducible polynomial 
+% $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$. Coincidently, $\alpha$ is also a normal element,
+% i.e. $\beta = \alpha$ can construct a NB $\{\alpha,\alpha^2,\alpha^4,\alpha^8,\alpha^{16},\alpha^{32},\alpha^{64},
+% \alpha^{128}\}$.
+% 
+% We pick a pair of elements from the Rijndael's field: $A=(0100~1011)_{StdB} = (4B)_{StdB},~B=(1100~1010)_{StdB} = (CA)_{StdB}$.
+% First let us compute their product in StdB, the rule follows ordinary polynomial multiplication.
+% 
+% \begin{align*}
+% A\cdot B &= (\alpha^6+\alpha^3+\alpha+1)(\alpha^7+\alpha^6+\alpha^3+\alpha)\\
+% &= (\alpha^{13}+\alpha^{10}+\alpha^8+\alpha^7)+(\alpha^{12}+\alpha^9+\alpha^7+\alpha^6)+
+% 	(\alpha^9+\alpha^6+\alpha^4+\alpha^3)\\
+% 	&~~~+(\alpha^7+\alpha^4+\alpha^2+\alpha) \\
+% &= \alpha^{13}+\alpha^{12}+\alpha^{10}+\alpha^8+\alpha^7+\alpha^3+\alpha^2+\alpha
+% \end{align*}
+% Note that this polynomial is not the final form of the product because it needs to be
+% reduced modulo irreducible polynomial $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$.
+% This can be done using base-2 long division. Note the dividend and divisor are written in pseudo Boolean
+% vectors, not real Boolean vectors in any kind of bases.
+% 
+% % The folowing part is customized latex code mimicing binary long division
+% % Note \divrule coordinates may not reflect the intuitive column # in tabular. Adjust by urself!!!
+% 
+% \vspace{1cm}
+% 
+% \newdimen\digitwidth
+% \settowidth\digitwidth{0}
+% \def~{\hspace{\digitwidth}}
+% 
+% 
+% \def\divrule#1#2{%
+% \noalign{\moveright#1\digitwidth%
+% \vbox{\hrule width#2\digitwidth}}}
+% 111010111\,\begin{tabular}[b]{@{}r@{}}
+% 101001 \\ \hline
+% \big)\begin{tabular}[t]{@{}l@{}}
+% 11010110001110 \\
+% 111010111 \\ \divrule{0}{14}
+% ~~111101101 \\
+% ~~111010111 \\ \divrule{2}{12}
+% ~~~~~111010110 \\
+% ~~~~~111010111 \\ \divrule{5}{9}
+% ~~~~~~~~~~~~~1
+% \end{tabular}
+% \end{tabular}
+% \vspace{0.5cm}
+% 
+% The final remainder is $1$, i.e. the product equals to 1 in StdB.
+% 
+% On the other hand, operands $A$ and $B$ can be written in NB as
+% $$A = (0010~1001)_{NB},~~B = (0100~0010)_{NB}$$
+% The $\lambda$-Matrix for $\mathbb F_{2}[x] \pmod{x^8+x^7+x^6+x^4+x^2+x+1}$
+% is
+% (Computation of $\lambda$-Matrix refers to appendix \ref{append:NB})
+% \begin{equation*}
+% M^{(0)} = \left(\begin{array}{lccccccr}
+% 0 &0 &0 &0 &1 &0 &1 &1 \\
+% 0 &0 &1 &1 &1 &1 &0 &0 \\
+% 0 &1 &0 &0 &0 &0 &1 &0 \\
+% 0 &1 &0 &0 &1 &1 &0 &1 \\
+% 1 &1 &0 &1 &0 &1 &0 &0 \\
+% 0 &1 &0 &1 &1 &0 &0 &1 \\
+% 1 &0 &1 &0 &0 &0 &0 &0 \\
+% 1 &0 &0 &1 &0 &1 &0 &1
+% \end{array}\right)
+% \end{equation*}
+% Taking matrix multiplication $c_0 = A\times M^{(0)}\times B^T$,
+% the result is $c_0 = 1$. Then by cyclic shifting $A$ and $B$
+% (or shifting $M^{(0)}$, either is applicable), we can successively
+% obtain other bits of product. The final answer is
+% $$C = (0000~0001)_{NB}$$
+% It is equivalent to the result in StdB.
+% \end{Example}
+% 
+% From the intuition of humans, StdB multiplication is straightforward and easier to understand
+% while NB is difficult to comprehend. However, if we implement both multiplications to 
+% hardware multipliers, it will be clear which side a circuit designer prefers.
+% 
+% % May consider providing details?
+% 
+% Mastrovito multiplier and Montgomery multiplier are 2 common designs
+% of GF multipliers using StdB. As a naive implementation of GF multiplication,
+% Mastrovito multiplier uses most number of gates:
+% $k^2$ AND gates plus $k^2-\Delta$ XOR gates \cite{Mastrovito}. Montgomery multiplier 
+% applies lazy reduction techniques and results in a better latency performance, while the number of gates are about
+% the same with Mastrovito multiplier:
+% $k^2$ AND gates plus $k^2-k/2$ XOR gates \cite{Montgomery}. 
+% Concretely, typical design of Mastrovito multiplier consists of 218 logic gates, while 
+% Montgomery multiplier needs 198 gates. However, the NB multiplier reuses the $\lambda$-Matrix 
+% logic, so this component will only need to be implemented for once. 
+% Consider the definition of matrix multiplication, it needs $C_N$ AND gates to apply 
+% bit-wise multiplication and $C_N-1$ XOR gates to sum the intermediate products up. The number of nonzero entries
+% in the $\lambda$-Matrix can be counted: $C_N = 27$.
+% As a result, the most naive NB multiplier design (or Massey-Omura multiplier \cite{MasseyOmura})
+% contains 53 gates in total, which is a great saving in area cost comparing to StdB multipliers.
+% 
+% \section{Design a Normal Basis Multiplier on Gate Level}
+% The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
+% we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
+% very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
+% test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
+% routine of several kinds of modern NB multipliers.
+% \subsection{Sequential Multiplier with Parallel Outputs}
+% The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
+% the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.
+% 
+% \begin{figure}[H]
+% \centering{
+% \includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
+% \caption{A typical SMSO structure of Massey-Omura multiplier}
+% \label{fig:MasseyOmura}}
+% \end{figure}
+% 
+% Figure \ref{fig:MasseyOmura} shows the basic architecture of a Massey-Omura multiplier. The operands
+% $A$ and $B$ are 2 arrays of flip-flops which allow 1-bit right-cyclic shift every clock cycle.
+% The logic gates in the boxes implements the matrix multiplication with $\lambda$-Matrix $M^{(0)}$, 
+% while each AND gate corresponds to term $a_ib_j$ and each XOR gate corresponds to addition $a_ib_j+a_{i'}b_{j'}$. 
+% The XOR layer has only 1 output, giving out 1 bit of product $C$ every clock cycle.
+% 
+% The behavior of Massey-Omura multiplier can be concluded as: pre-load operands $A,B$ and reset $C$ to 0, after 
+% executing for $k$ clock cycles, the data stored in flip-flop array $C$ is the product $A\times B$.
+% We note that there is only one output giving 1 bit of the product each clock cycle, which matches the 
+% definition of serial output to communication channel. Therefore this type of design is named as 
+% sequential multiplier with serial output (SMSO).
+% The SMSO architecture need $C_N$ AND gates and $C_N-1$ XOR gates, which equals to $2k-1$ AND gates and $2k-2$ XOR gates
+% if it is designed using ONB. In fact, the number of gates can be reduced if the multiplication is 
+% implemented using a conjugate of SMSO.
+% 
+% The gate-level logic boxes are implementing following function:
+% \begin{equation}
+% \label{eqn:SMPOterms}
+% c_l = row_1(A\times M^{(l)}) \times B + row_2(A\times M^{(l)}) \times B + \cdots + row_{k}(A\times M^{(l)}) \times B
+% \end{equation}
+% It can be decomposed into $k$ terms. If we only compute one term for each $c_l,~0\leq l\leq k-1$ in one 
+% clock cycle, make $k$ outputs and add them up using shift register after $k$ clock cycles, 
+% it will generate the same result with SMSO. This kind of architecture is named as 
+% sequential multiplier with parallel outputs (SMPO). The basic SMPO, as a conjugate of Massey-Omura multiplier,
+% is invented by Agnew et al. \cite{agnew1991implementation}.
+% 
+% \begin{figure}[H]
+% \centering{
+% \includegraphics[width=\textwidth]{newfig/mySMPO.eps}
+% \caption{5-bit Agnew's SMPO.
 Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, i.e.
@@
 Index $w = 2i\pmod 5$}
-\label{fig:SMPO}}
-\end{figure}
-
-\begin{Example}[5-bit Agnew's SMPO]
-Given GF $\mathbb F_{2^5}$ and primitive element $\alpha$ defined by irreducible polynomial 
-$\alpha^5+\alpha^2+1=0$, normal element $\beta = \alpha^5$ constructs an ONB $\{\beta,\beta^2,\beta^4,\beta^8,\beta^{16}\}$.
-The $0$-th $\lambda$-Matrix for this ONB is
-\begin{equation*}
-M^{(0)} = \left(\begin{array}{lcccr}
-0 &1 &0 &0 &0 \\
-1 &0 &0 &1 &0 \\
-0 &0 &0 &1 &1 \\
-0 &1 &1 &0 &0 \\
-0 &0 &1 &0 &1
-\end{array}\right)
-\end{equation*}
-Then a typical design of 5-bit Agnew's SMPO is depicted in figure \ref{fig:SMPO}.
-
-The operands part of this circuit is the same with Massey-Omura multiplier. The differences are on 
-the matrix multiplication part, while it is implemented as separate logic blocks for 5 outputs,
-and the 5 blocks are connected in a shift register fashion. By analyzing the detailed function of 
-logic blocks, we can reveal the mechanism of Agnew's SMPO.
-
-Suppose we implement $M^{(0)}$ as the logic block in SMSO. In the first clock cycle, the output is
-\begin{equation}
-\label{eqn:5bitSMPO_c0}
-c_0 = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4
-\end{equation}
-Note it is written in the form of equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
-remaining bits of the product, which can be written in following general form polynomial:
-\begin{align}
-c_i =& b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
-&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
-(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
-\end{align}
-Note all index calculations are reduced modulo 5.
-
-Now let us observe the behavior of 5-bit Agnew's SMPO. Initially all DFFs are reset to 0. 
-In the first clock cycle,
-signal sent to the flip-flop in block $R_0$ denotes function:
-$$R_0^{(1)} = a_1b_0$$
-It equals to the first term of equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
-this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
-from operands (shifted by 1 bit), generating signal $a_u,a_v$ and $b_w$. Concretely,
-signal sent to flip-flop in block $R_1$ is:
-$$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
-which forms first 2 terms of equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
-on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
-$$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
-which equals to $c_0$ in equation \ref{eqn:5bitSMPO_c0}.
-After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
-the result of $c_0$ is computed after 5 clock cycles and given on $r_0$.
-
-If we track $R_1\to R_2\to R_3 \to \cdots \to R_0$, we can obtain $c_1$ respectively.
-Thus we conclude that Agnew's SMPO functions the same with Massey-Omura multiplier.
-\end{Example}
-
-The design of Agnew's SMPO guarantees that there is only one AND gate in each $R_i$ block.
-For ONB, adopting Agnew's SMPO will reduce the number of AND gates from $2k-1$ to $k$.
-
-\subsection{Multiplier not based on $\lambda$-Matrix}
-Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
-which means that they will be identical if unrolled to full combinational circuits. 
-After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
-some of them jump out of the circle and are independent from $\lambda$-Matrix.
-One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
-\cite{RHmulti}, which is therefore called RH-SMPO.
-
-% fig:5bitRH
-
-Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
-in figure \ref{fig:SMPO}. A brief proof is as follows:
-
-\begin{Proof}
-First, we define an auxiliary function for $i$-th bit 
-\begin{equation}
-\label{eqn:aux}
-F_i(A,B) = a_ib_i\beta + \sum_{j=1}^v d_{i,j}\beta^{1+2^j}
-\end{equation}
-where $0\leq i\leq k-1, v = \lfloor k/2\rfloor, 1\leq j \leq v$.
-The $d$-layer index $d_{i,j}$ is defined as
-\begin{equation}
-\label{eqn:auxDC}
-d_{i,j} = c_{a,i} c_{b,i} = (a_i+a_{i+j})(b_i+b_{i+j}),~~1\leq j\leq v
-\end{equation}
-$i+j$ here is the result reduced modulo $k$. Note that there is a special boundary case when
-$k$ is an even number ($v = \frac{k}{2}$):
-$$d_{i,v} = (a_i+a_{i+v})b_i$$
-With the auxiliary function, we can utilize following theorem (proof refers to \cite{RHmulti}):
-\begin{Theorem}
-Consider three elements $A,B$ and $R$ such that $R=A\times B$ over $\Fkk$. Then,
-$$R=(((F_{k-1}^2+F_{k-2})^2+F_{k-3})^2+\cdots+F_1)^2+F_0$$
-\end{Theorem}
-This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
-$R_i$ flip-flops. Concretely, the multiplier behavior is an implementation of 
-following algorithm:
-
-\begin{algorithm}[H]
-\SetAlgoNoLine
-
- \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
- \KwOut{$R=A\times B$}
-%%%%%%%%%%%%%%%%%%%%
-  Initialize $A,B$ and aux var $X$ to 0\;
-  \For { ($i=0$;   $i < k$; ++i ) }
-  {
-	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Eq.\ref{eqn:aux}*/\;}
-	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
-  }
-  $R\gets X$
-\caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
-\end{algorithm}
-
-In this algorithm, we use a fixed auxiliary function $F_{k-1}$ inside the loop.
-This is because of equation
-$$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
-So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
-$F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
-\end{Proof}
-
-To better understand the mechanism of RH-SMPO, we will use this 5-bit 
-RH-SMPO as an example and introduce the details on how to design it.
-\begin{Example}[Designing a 5-bit RH-SMPO]
-From equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
-and XOR gates in $c$-layer according to equation \ref{eqn:auxDC}. Concretely, as algorithm \ref{alg:RHmulti}
-describes, we implement auxiliary function $F_{k-1}$ in the logic:
-$$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
-\begin{equation}
-\label{eqn:5bitRHaux}
-F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
-\end{equation}
-Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
-$$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
-$$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
-The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
-As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
-$F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
-turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
-$i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.
-
-In order to obtain information 
-indicating interconnections between $d$-layer and $e$-layer, we need to interpret $\beta^{1+2^j}$ 
-to NB representation.
-There is a concept called {\bf multiplication table} (M-table) which can assist this interpretation. It is defined as 
-a $k\times k$ matrix $T$ over $\mathbb F_2$:
-\begin{equation}
-\label{eqn:multitable}
-\begin{bmatrix}
-\beta^{1+2^0} \\ \beta^{1+2^1} \\ \beta^{1+2^2} \\ \vdots \\ \beta^{1+2^{k-1}}
-\end{bmatrix}
-= \beta
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
-\end{bmatrix}
-=
-\begin{bmatrix}
-T_{0,0}      &   T_{0,1}        & \dots & T_{0,k-1}\\
-T_{1,0}    &   T_{1,1}           & \dots & T_{1,k-1}\\
-T_{2,0}    &   T_{2,1}           & \dots & T_{2,k-1}\\
-\vdots & \vdots              & \ddots     & \vdots \\
-T_{k-1,0}    &   T_{k-1,1}           & \dots & T_{k-1,k-1}
-\end{bmatrix}
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
-\end{bmatrix}
-= {\bf T}
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
-\end{bmatrix}
-\end{equation}
-
-It is a known fact that M-table $T$ can be converted from $\lambda$-Matrix $M$:
-$$M_{i,j}^{(0)} = T_{j-i,-i}$$
-with indices reduced modulo $k$ (proof refers to appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
-
-% fig:multitable
-
-Note that we only use row 1 and row 2 from the M-table since range $1\leq j \leq 2$.
-All nonzero entries in these 2 rows corresponds to the interconnections between $d$-layer and 
-$e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
-between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in equation \ref{eqn:multitable}:
-\begin{equation*}
-\beta\cdot\beta^2 = 
-\begin{bmatrix}
-1 & 0 & 0 & 1 & 0
-\end{bmatrix}
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
-\end{bmatrix}
-= \beta + \beta^{2^3}
-\end{equation*}
-
-Similarly, from row 2 of M-table we derive that $d_2$ has fanouts $e_3,e_4$:
-\begin{equation*}
-\beta\cdot\beta^{2^2} = 
-\begin{bmatrix}
-0 & 0 & 0 & 1 & 1
-\end{bmatrix}
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
-\end{bmatrix}
-= \beta^{2^3} + \beta^{2^4}
-\end{equation*}
-
-Let us look back at equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
-The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
-After gathering all interconnection information, we can translate it to gate-level circuit implementation:
-$$e_0 = d_0+d_1,~e_3=d_1+d_2,~e_4=d_2$$
-
-Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
-$R_{i-1}$ in last clock cycle, we can connect using relation 
-$$R_i = r_{i-1} + e_i$$
-In this example, according to the M-table in figure \ref{fig:multitable}, columns $e_1,e_2$
-have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.
-
-This finishes the full design procedure for a 5-bit RH-SMPO.
-\end{Example}
-
-The area cost of RH-SMPO is even smaller than Agnew's SMPO. XOR gates corresponds to all nonzero entries 
-in M-table, which is with the same number of nonzero entries in $\lambda$-Matrix ($C_N$). The number of AND gates 
-equals to $v$ plus 1 (for gate $d_0$). When using ONB ($C_N = 2k-1$), the total number of gates 
-is $2k+\lfloor \frac{k}{2}\rfloor$.
+% \label{fig:SMPO}}
+% \end{figure}
+% 
+% \begin{Example}[5-bit Agnew's SMPO]
+% Given GF $\mathbb F_{2^5}$ and primitive element $\alpha$ defined by irreducible polynomial 
+% $\alpha^5+\alpha^2+1=0$, normal element $\beta = \alpha^5$ constructs an ONB $\{\beta,\beta^2,\beta^4,\beta^8,\beta^{16}\}$.
+% The $0$-th $\lambda$-Matrix for this ONB is
+% \begin{equation*}
+% M^{(0)} = \left(\begin{array}{lcccr}
+% 0 &1 &0 &0 &0 \\
+% 1 &0 &0 &1 &0 \\
+% 0 &0 &0 &1 &1 \\
+% 0 &1 &1 &0 &0 \\
+% 0 &0 &1 &0 &1
+% \end{array}\right)
+% \end{equation*}
+% Then a typical design of 5-bit Agnew's SMPO is depicted in figure \ref{fig:SMPO}.
+% 
+% The operands part of this circuit is the same with Massey-Omura multiplier. The differences are on 
+% the matrix multiplication part, while it is implemented as separate logic blocks for 5 outputs,
+% and the 5 blocks are connected in a shift register fashion. By analyzing the detailed function of 
+% logic blocks, we can reveal the mechanism of Agnew's SMPO.
+% 
+% Suppose we implement $M^{(0)}$ as the logic block in SMSO. In the first clock cycle, the output is
+% \begin{equation}
+% \label{eqn:5bitSMPO_c0}
+% c_0 = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4
+% \end{equation}
+% Note it is written in the form of equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
+% remaining bits of the product, which can be written in following general form polynomial:
+% \begin{align}
+% c_i =& b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
+% &+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
+% (a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
+% \end{align}
+% Note all index calculations are reduced modulo 5.
+% 
+% Now let us observe the behavior of 5-bit Agnew's SMPO. Initially all DFFs are reset to 0. 
+% In the first clock cycle,
+% signal sent to the flip-flop in block $R_0$ denotes function:
+% $$R_0^{(1)} = a_1b_0$$
+% It equals to the first term of equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
+% this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
+% from operands (shifted by 1 bit), generating signal $a_u,a_v$ and $b_w$. Concretely,
+% signal sent to flip-flop in block $R_1$ is:
+% $$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
+% which forms first 2 terms of equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
+% on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
+% $$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
+% which equals to $c_0$ in equation \ref{eqn:5bitSMPO_c0}.
+% After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
+% the result of $c_0$ is computed after 5 clock cycles and given on $r_0$.
+% 
+% If we track $R_1\to R_2\to R_3 \to \cdots \to R_0$, we can obtain $c_1$ respectively.
+% Thus we conclude that Agnew's SMPO functions the same with Massey-Omura multiplier.
+% \end{Example}
+% 
+% The design of Agnew's SMPO guarantees that there is only one AND gate in each $R_i$ block.
+% For ONB, adopting Agnew's SMPO will reduce the number of AND gates from $2k-1$ to $k$.
+% 
+% \subsection{Multiplier not based on $\lambda$-Matrix}
+% Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
+% which means that they will be identical if unrolled to full combinational circuits. 
+% After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
+% some of them jump out of the circle and are independent from $\lambda$-Matrix.
+% One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
+% \cite{RHmulti}, which is therefore called RH-SMPO.
+% 
+% \begin{figure}[H]
+% \centering{
+% \includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
+% \caption{5-bit RH-SMPO}
+% \label{fig:5bitRH}}
+% \end{figure}
+% 
+% Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
+% in figure \ref{fig:SMPO}. A brief proof is as follows: 
+% 
+% \begin{Proof}
+% First, we define an auxiliary function for $i$-th bit 
+% \begin{equation}
+% \label{eqn:aux}
+% F_i(A,B) = a_ib_i\beta + \sum_{j=1}^v d_{i,j}\beta^{1+2^j}
+% \end{equation}
+% where $0\leq i\leq k-1, v = \lfloor k/2\rfloor, 1\leq j \leq v$.
+% The $d$-layer index $d_{i,j}$ is defined as
+% \begin{equation}
+% \label{eqn:auxDC}
+% d_{i,j} = c_{a,i} c_{b,i} = (a_i+a_{i+j})(b_i+b_{i+j}),~~1\leq j\leq v
+% \end{equation}
+% $i+j$ here is the result reduced modulo $k$. Note that there is a special boundary case when
+% $k$ is an even number ($v = \frac{k}{2}$):
+% $$d_{i,v} = (a_i+a_{i+v})b_i$$
+% With the auxiliary function, we can utilize following theorem (proof refers to \cite{RHmulti}):
+% \begin{Theorem}
+% Consider three elements $A,B$ and $R$ such that $R=A\times B$ over $\Fkk$. Then,
+% $$R=(((F_{k-1}^2+F_{k-2})^2+F_{k-3})^2+\cdots+F_1)^2+F_0$$
+% \end{Theorem}
+% This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
+% $R_i$ flip-flops. Concretely, the multiplier behavior is an implementation of 
+% following algorithm:
+% 
+% \begin{algorithm}[H]
+% \SetAlgoNoLine
+% 
+%  \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
+%  \KwOut{$R=A\times B$}
+% %%%%%%%%%%%%%%%%%%%%
+%   Initialize $A,B$ and aux var $X$ to 0\;
+%   \For { ($i=0$;   $i < k$; ++i ) }
+%   {
+% 	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Eq.\ref{eqn:aux}*/\;}
+% 	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
+%   }
+%   $R\gets X$
+% \caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
+% \end{algorithm}
+% 
+% In this algorithm, we use a fixed auxiliary function $F_{k-1}$ inside the loop.
+% This is because of equation
+% $$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
+% So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
+% $F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
+% \end{Proof}
+% 
+% To better understand the mechanism of RH-SMPO, we will use this 5-bit 
+% RH-SMPO as an example and introduce the details on how to design it.
+% \begin{Example}[Designing a 5-bit RH-SMPO]
+% From equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
+% and XOR gates in $c$-layer according to equation \ref{eqn:auxDC}. Concretely, as algorithm \ref{alg:RHmulti}
+% describes, we implement auxiliary function $F_{k-1}$ in the logic:
+% $$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
+% \begin{equation}
+% \label{eqn:5bitRHaux}
+% F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
+% \end{equation}
+% Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
+% $$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
+% $$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
+% The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
+% As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
+% $F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
+% turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
+% $i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.
+% 
+% In order to obtain information 
+% indicating interconnections between $d$-layer and $e$-layer, we need to interpret $\beta^{1+2^j}$ 
+% to NB representation.
+% There is a concept called {\bf multiplication table} (M-table) which can assist this interpretation. It is defined as 
+% a $k\times k$ matrix $T$ over $\mathbb F_2$:
+% \begin{equation}
+% \label{eqn:multitable}
+% \begin{bmatrix}
+% \beta^{1+2^0} \\ \beta^{1+2^1} \\ \beta^{1+2^2} \\ \vdots \\ \beta^{1+2^{k-1}}
+% \end{bmatrix}
+% = \beta
+% \begin{bmatrix}
+% \beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+% \end{bmatrix}
+% =
+% \begin{bmatrix}
+% T_{0,0}      &   T_{0,1}        & \dots & T_{0,k-1}\\
+% T_{1,0}    &   T_{1,1}           & \dots & T_{1,k-1}\\
+% T_{2,0}    &   T_{2,1}           & \dots & T_{2,k-1}\\
+% \vdots & \vdots              & \ddots     & \vdots \\
+% T_{k-1,0}    &   T_{k-1,1}           & \dots & T_{k-1,k-1}
+% \end{bmatrix}
+% \begin{bmatrix}
+% \beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+% \end{bmatrix}
+% = {\bf T}
+% \begin{bmatrix}
+% \beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+% \end{bmatrix}
+% \end{equation}
+% 
+% It is a known fact that M-table $T$ can be converted from $\lambda$-Matrix $M$:
+% $$M_{i,j}^{(0)} = T_{j-i,-i}$$
+% with indices reduced modulo $k$ (proof refers to appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
+% 
+% \begin{figure}[H]
+% \centering{
+% \includegraphics[width=3in]{newfig/multitable.pdf}
+% \caption{A $5\times 5$ multiplication table}
+% \label{fig:multitable}}
+% \end{figure}
+% 
+% Note that we only use row 1 and row 2 from the M-table since range $1\leq j \leq 2$.
+% All nonzero entries in these 2 rows corresponds to the interconnections between $d$-layer and 
+% $e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
+% between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in equation \ref{eqn:multitable}:
+% \begin{equation*}
+% \beta\cdot\beta^2 = 
+% \begin{bmatrix}
+% 1 & 0 & 0 & 1 & 0
+% \end{bmatrix}
+% \begin{bmatrix}
+% \beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
+% \end{bmatrix}
+% = \beta + \beta^{2^3}
+% \end{equation*}
+% 
+% Similarly, from row 2 of M-table we derive that $d_2$ has fanouts $e_3,e_4$:
+% \begin{equation*}
+% \beta\cdot\beta^{2^2} = 
+% \begin{bmatrix}
+% 0 & 0 & 0 & 1 & 1
+% \end{bmatrix}
+% \begin{bmatrix}
+% \beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
+% \end{bmatrix}
+% = \beta^{2^3} + \beta^{2^4}
+% \end{equation*}
+% 
+% Let us look back at equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
+% The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
+% After gathering all interconnection information, we can translate it to gate-level circuit implementation:
+% $$e_0 = d_0+d_1,~e_3=d_1+d_2,~e_4=d_2$$
+% 
+% Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
+% $R_{i-1}$ in last clock cycle, we can connect using relation 
+% $$R_i = r_{i-1} + e_i$$
+% In this example, according to the M-table in figure \ref{fig:multitable}, columns $e_1,e_2$
+% have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.
+% 
+% This finishes the full design procedure for a 5-bit RH-SMPO.
+% \end{Example}
+% 
+% The area cost of RH-SMPO is even smaller than Agnew's SMPO. XOR gates corresponds to all nonzero entries 
+% in M-table, which is with the same number of nonzero entries in $\lambda$-Matrix ($C_N$). The number of AND gates 
+% equals to $v$ plus 1 (for gate $d_0$). When using ONB ($C_N = 2k-1$), the total number of gates 
+% is $2k+\lfloor \frac{k}{2}\rfloor$.
 
@@
 For 2-input sequential NB multipliers, abstraction is utilized to implement following algorithm:
-\vspace{-0.1in}
+%\vspace{-0.1in}
 \IncMargin{1em}
@@
 \DecMargin{1em}
-\vspace{-0.1in}
+%\vspace{-0.1in}
 

### 2016-11-21 dissertation/prelim_GF.tex
dissertation v2_finish prelims

@@
 examples of such rings are the set of integers, $\mathbb{Z}$, and the set of 
-rational numbers, $\mathbb{Q}$. Note that while both of these examples are
+rational numbers, $\mathbb{Q}$. While both of these examples are
 rings with an infinite number of elements, the number of elements in a ring 
-can also be finite.
+can also be finite, such as the ring of integers modulo $n$ ($\Z_n$).
 
@@
 
-The main concept of field theory is {\bf Field Extensions}. The idea behind a
+An important concept in field theory is {\bf field extension}. The idea behind a
 field extension is to take a base field and construct a larger field which 
 contains the base field as well as satisfies additional properties. For example,
-the set of real numbers $\mathbb{R}$ forms a field; one common extension of 
+the set of real numbers $\mathbb{R}$ forms a field; one extension of 
 $\mathbb{R}$ is the set of complex numbers $\mathbb{C}=\mathbb{R}(i)$. Every
@@
 elements. 
-In this dissertation we focus on finite fields, also known as Galois fields, and 
-the construction of their field extensions.
+In this dissertation we focus on finite fields -- also known as Galois fields, 
+the construction of their field extensions, and their applications on circuit verification and abstraction techniques.
 
-\subsection{Galois Field}
-Galois fields, also known as finite fields, find widespread applications in 
+\subsection{Finite Field}
+Finite fields find widespread applications in 
 many areas of electrical engineering and computer science such as error-
@@
 In this dissertation, we specifically focus on their application to 
-Elliptic Curve Cryptography as Galois field arithmetic circuits.
+the FSM traversal of sequential Galois field circuits as well as abstraction refinement based on UNSAT core extraction.
 This section describes the relevant Galois field concepts
@@
 
-Fields in the form $\F_{p^{k}}$ are called Galois extension fields.
+Fields of the form $\F_{p^{k}}$ are called Galois extension fields.
 We are specifically interested in extension fields of type 
@@
 \begin{Definition}
-A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$ is {\bf irreducible} 
+A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$, i.e. the set of all 
+polynomials in $x$ with coefficients in $\F_2$, is {\bf irreducible} 
 if $P(x)$ is non-constant with degree $k$ and cannot be 
@@
 of real numbers $\R$ with an additional element $i=\sqrt{-1}$, which is an imaginary
-root in $\R$.
+root in the algebraic closure of $\R$ -- the closure is known as the field of complex numbers $\C$.
 Thus $i \notin \R$, rather $i \in \C$.
@@
 such that \{$0$, $1(=\alpha^{{2^k}-1})$, $\alpha$, $\alpha^2$, $\cdots$, $\alpha^{2^k-2}$\} is the set of 
-all elements in $\mathbb{F}_{2^k}$, 
-where $\alpha$ is a {\bf primitive element} of $\mathbb{F}_{2^k}$. 
+all elements in $\mathbb{F}_{2^k}$.
+Here $\alpha$ is called a {\bf primitive element} of $\mathbb{F}_{2^k}$. 
 \end{Definition}
@@
 A primitive polynomial is guaranteed to generate all distinct elements 
-of a finite field $\mathbb{F}_{2^k}$ while an irreducible polynomial
+of a finite field $\mathbb{F}_{2^k}$ while an arbitrary irreducible polynomial
 has no such guarantee.
@@
 That is, $\F_q \subset \F_{q^k}$.
-For example, Fig \ref{fig:contain2_4_16} shows the containment of the fields 
-$\F_2 \subset \F_4 \subset \F_{16}$. It's easy to see that since $\F_4=\F_{2^2}$, it
-contains $\F_2$. Likewise $\F_{16}=\F_{4^2}=\F_{2^4}$ contains $\F_4$ and $\F_2$.
+For example, the containment relation of the fields 
+$\F_2 \subset \Fkk$ is usually used to represent bit-level Boolean variables as field elements
+in larger finite field which allows projection of $k$-bit word-level variables. 
+Concretely, $\F_{16}=\F_{4^2}=\F_{2^4}$ contains $\F_4$ and $\F_2$.
 The elements $\{0,1,\alpha,\dots,\alpha^{14}\}$
@@
 \end{Definition}
+
+\section{Normal Basis Multiplier over Galois Field}
+From algebraic view, a field is a space, and field elements are points in the space. Those elements can be 
+represented with unique coordinates, which requires the pre-definition of a basis vector. In this
+section, we discuss a special basis called normal basis, as well as the advantages adopting it in 
+GF operations esp. multiplication.
+\subsection{Normal Basis}
+Given a Galois field (GF) $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
+Its elements can be written in polynomials of $\alpha$, when there is an irreducible polynomial $p(\alpha)$
+defined.
+
+If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
+to binary bit-vector representations by recording the coefficients. For example,
+
+\begin{table}[H]
+\centering
+\caption{Bit-vector, Exponential and Polynomial representation of
+elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
+\pmod{x^4+x^3+1}$}
+\begin{tabular}{|c|c||c|c|} 
+\hline
+$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
+\hline
+$0000$        & $0$           & $1000$  &$\alpha^3$\\
+\hline
+$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
+\hline
+$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
+\hline
+$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
+\hline
+$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
+\hline
+$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
+\hline
+$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
+\hline
+$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
+\hline
+\end{tabular}
+\label{table:booltogalois}  
+\end{table}
+
+Basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$ is called {\bf standard basis} (StdB), which results in
+a straightforward representation for elements, and operations of elements such as addition and subtraction.
+The addition/subtraction of GF elements in StdB follows the rules of polynomial addition/subtraction
+where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} (XOR) in
+Boolean algebra, element $A$ add/subtract by element $B$ in StdB is defined as
+\begin{align}\label{eqn:StdB}
+A+B = A-B &= (a_0,a_1,\dots,a_{k-1})_{StdB} \xor (b_0,b_1,\dots,b_{k-1})_{StdB} \nonumber\\
+&=(a_0\oplus b_0, a_1\oplus b_1,\dots,a_{k-1}\oplus b_{k-1})_{StdB} 
+\end{align}
+
+\subsection{Multiplication using Normal Basis}
+Besides addition/subtraction, multiplication is also very common in arithmetic circuit design.
+The multiplication of GF elements in $\Fkk$ in StdB follows the rule of polynomial multiplication.
+However, it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
+in bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
+the area and delay of circuit will be costly.
+
+In order to lower down the complexity of arithmetic circuit design, Massey and Omura \cite{MasseyOmura} % ref 7 in RH paper
+use a new basis to represent GF elements, which is called {\bf normal basis} (NB).
+A normal basis over $\Fkk$ is written in the form of
+\begin{equation*}
+N.B. ~~~ \N = \{\beta,\beta^2,\beta^4,\beta^8,\dots,\beta^{2^{k-1}}\}
+\end{equation*}
+
+{\bf Normal element} $\beta$ is an element from the field which is used to construct the normal basis,
+and can be represent as a power of primitive element $\alpha$: 
+\begin{equation*}
+\beta = \alpha^t, ~~~ 1\leq t<2^k
+\end{equation*}
+Exponent $t$ takes value in the given range when $\N$ fulfills the definition of a basis.
+
+Respectively, a field element in NB representation is actually
+\begin{align*}
+A &= (a_0,a_1,\dots,a_{k-1})_{NB} \\
+  &= a_0\beta+a_1\beta^2+\cdots+a_{k-1}\beta^{2^{k-1}} \\
+  &= \sum_{i=0}^{k-1} a_i\beta^{2^i}
+\end{align*}
+
+According to the definition, a normal basis is a vector where the next entry is the square of the former one.
+We note that the vector is cyclic, i.e. $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
+
+The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
+However, what makes NB powerful is its ease of implementation when doing multiplications and exponentiations.
+The following lemmas and examples illustrate this fabulous property very well.
+\begin{Lemma}[Square of NB]
+\label{lem:squareNB}
+In $\Fkk$, 
+\begin{equation*}
+(a+b)^2 = a^2 + b^2
+\end{equation*}
+According to the \textbf{binomial theorem}, it can be extended to
+\begin{align*}
+\beta^2 =&(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}})^2 \\
+=& b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
+=& b_{k-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}}
+\end{align*}
+\end{Lemma}
+This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
+Obviously, StdB representation does not have this benefit.
+
+\begin{Example}[Square of NB]
+In GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $x^3 + x + 1$, the standard basis is denoted as 
+$\{ 1, \alpha, \alpha^2\}$ where $\alpha^3+\alpha+1=0$.
+Let $\beta = \alpha^3$, then $\N = \{ \beta, \beta^2, \beta^4\}$ forms a normal basis. 
+Write down element $E$ using both representations:
+\begin{align*}
+E &= (a_0,a_1,a_2)_{StdB} = (b_0,b_1,b_2)_{NB} \\
+  &= a_0 + a_1\alpha + a_2\alpha^2 = b_0\beta + b_1\beta^2 + b_2\beta^4
+\end{align*}
+Compute the square of $E$ in StdB first:
+\begin{align*}
+E^2 &= a_0 + a_1\alpha^2 + a_2\alpha^4 \\
+    &= a_0 + a_2\alpha + (a_1 + a_2)\alpha^2 \\
+    &= (a_0,a_2,a_1+a_2)_{StdB}
+\end{align*}
+When it is computed in NB, we can make it very simple:
+\begin{align*}
+E^2 &= \overset{\xrightarrow{Cyclic~~shift}}{(b_0,b_1,b_2)}_{NB} \\
+	&= (b_2,b_0,b_1)_{NB}
+\end{align*}
+\end{Example}
+
+This example shows that convenience to use NB when computing $2^k$ power of an element.
+Multiplication is a bit complicated than squaring; but when it is decomposed as bit-wise
+operations, the property in lemma \ref{lem:squareNB} can be well utilized.
+
+\begin{Example}[Bit-wise NB multiplication]
+Assume there are 2 binary vectors representing 2 operands in NB over $\Fkk$: 
+$A = (a_0, a_1, \dots, a_{k-1}), B = (b_0, b_1, \dots, b_{k-1})$. Note that in this example, 
+by default we use normal basis representation so subscript ``NB" is skipped. Their product can also be written
+as: $$C = A\times B = (c_0, c_1, \dots, c_{k-1})$$
+
+Assume the most significant bit (MSB) of the product can be represented by a function $f_{mult}$: 
+\begin{equation}
+\label{eqn:multiMSB}
+c_{k-1} = f_{mult}(a_0, a_1, \dots, a_{k-1}; b_0, b_1, \dots, b_{k-1})
+\end{equation}
+Before discussing the details of the function $f_{mult}$, we can take 
+a square on both side of equation \ref{eqn:multiMSB}, i.e. $C^2 = A^2\times B^2$.
+Obviously, using the property in lemma \ref{lem:squareNB}, the original second most significant bit 
+becomes the new MSB because of right-cyclic shifting. 
+Concretely, 
+$$(c_{k-1},c_0,c_1,\dots,c_{k-2}) = (a_{k-1},a_0,a_1,\dots,a_{k-2})\times(b_{k-1},b_0,b_1,\dots,b_{k-2})$$
+Note $A^2, B^2$ and $C^2$ still belong to $\Fkk$, thus as a universal function implementing MSB multiplication
+over $\Fkk$, $f_{mult}$ still remains the same. As a result, the new MSB can be written as 
+\begin{equation}
+\label{eqn:shiftMSB}
+c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
+\dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})
+\end{equation}
+Similarly, if we take a square again on the new equation, we can get $c_{k-3}$.
+Successively we can derive all bits of product $C$ using the same function $f_{mult}$, and the only
+adjustment we need to make is to right-cyclic shift 2 operands by 1 bit each time.
+\end{Example}
+
+From above example, it is known that a universal structure that implements $f_{mult}$ can be reused
+$k$ times in NB multiplication over $\Fkk$. Compared to StdB, which requires a distinct design 
+for every bit of multiplication, NB is less costly -- as long as we can prove $f_{mult}$ implies 
+a structure with $o(k^2)$ complexity (symbol $o$ denotes ``strictly lower than bound"). 
+% So our next mission is to explore the details of $f_{mult}$
+% to prove it will be a relatively simple design with complexity lower than $O(k^2)$.
+
+If we want to make the complexity of $f_{mult}$ lower than $O(k^2)$, then the best choice is to try out 
+linear functions. As we know, matrix multiplication can simulate all possible combinations of 
+linear functions. 
+% (which is also the reason it is used as basic model to simulate the behavior of a neuron
+% in neural network machine learning algorithms).
+Imagine $A$ is a $k$-bit row vector and $B$ is a $k$-bit 
+column vector, then the single bit product can be written as the product of matrix multiplication
+$$c_{l} = A\times M\times B$$
+where 
+\begin{align*}
+A &= (a_0,a_1,\dots,a_{k-1})\\
+B &= \begin{pmatrix}
+b_0 \\
+b_1 \\
+\vdots \\
+b_{k-1}
+\end{pmatrix}
+\end{align*}
+Moreover, $M$ is a $k\times k$ square matrix. If we can find $M$, we obtain the design of the multiplier.
+
+\begin{Definition}[$\lambda$-Matrix]
+A binary $k\times k$ matrix $M$ is used to describe the bit-wise normal basis multiplication function $f_{mult}$ where
+\begin{equation}
+\label{eqn:def_lambda}
+c_{l} = f_{mult}(A, B) = A \times M \times B^T
+\end{equation}
+Symbol $B^T$ denotes vector transposition. Matrix $M$ is called $\lambda$-Matrix of
+$k$-bit NB multiplication over $\Fkk$.
+\end{Definition}
+
+When taking different bits $l$ of the product in equation \ref{eqn:def_lambda}, 
+we obtain a series of conjugate matrices of $M$. 
+Which means instead of shifting operands $A$ and $B$, we can also shift the 
+matrix.
+
+More specifically, we denote the matrix by \emph{$l$-th $\lambda$-Matrix} as 
+$$c_l = A \times M^{(l)} \cdot B^T$$
+Meanwhile, the operator shifting rule in equation \ref{eqn:shiftMSB} still holds. Then we have relation 
+$$c_{l-1} = A \cdot M^{(l-1)} \cdot B^T = shift(A) \cdot M^{(l)} \cdot shift(B)^T$$ which means
+by right and down cyclically shifting $M^{(l-1)}$, we can get $M^{(l)}$.
+
+\begin{Example}[NB multiplication using $\lambda$-Matrix]
+Over GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $\alpha^3 + \alpha + 1$, let normal element $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
+forms a normal basis. Corresponding $0$-th $\lambda$-Matrix is
+\begin{equation*}
+M^{(0)} = \left(
+\begin{array} {lcr}
+0 & 1 & 0\\
+1 & 0 & 1\\
+0 & 1 & 1
+\end{array} \right).
+\end{equation*}
+i.e.,
+\begin{equation*}
+c_0 = (a_0\  a_1\  a_2)\left(
+\begin{array} {lcr}
+0 & 1 & 0\\
+1 & 0 & 1\\
+0 & 1 & 1
+\end{array} \right)\left(
+\begin{array} {lcr}
+b_0\\
+b_1\\
+b_2
+\end{array} \right)
+\end{equation*}
+From $0$-th $\lambda$-Matrix we can directly write down all remaining $\lambda$-Matrices:
+\begin{equation*}
+M^{(1)} = \left(
+\begin{array} {lcr}
+1 & 0 & 1\\
+0 & 0 & 1\\
+1 & 1 & 0
+\end{array} \right)~~~~~~
+M^{(2)} = \left(
+\begin{array} {lcr}
+0 & 1 & 1\\
+1 & 1 & 0\\
+1 & 0 & 0
+\end{array} \right)
+\end{equation*}
+\end{Example}
+
+If we generalize the definition and explore the nature of $\lambda$-Matrix, it is defined as cross-product terms from multiplication, which is 
+\begin{equation}
+Product~vector~C = (\sum_{i=0}^{k-1}a_i\beta^{2^i})(\sum_{j=0}^{k-1}b_j\beta^{2^j}) = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}
+\end{equation}
+The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
+NB, i.e.
+\begin{equation}
+\beta^{2^i}\beta^{2^j} = \sum_{l=0}^{k-1}\lambda_{ij}^{(l)}\beta^{2^l}, \ \ \lambda_{ij}^{(l)} \in \mathbb F_2.
+\end{equation}
+Substitution yields, result is an expression for l-th digit of product as showed in equation \ref{eqn:multiMSB}:
+\begin{equation}
+c_l = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(l)}a_ib_j
+\end{equation}
+$\lambda_{ij}^{(l)}$ is the entry with coordinate $(i,j)$ in $l$-th $\lambda$-Matrix.
+
+The $\lambda$-Matrix can be implemented with XOR and AND gates in circuit design.
+The very naive implementation requires $O(C_N)$ gates, where $C_N$ is the number of
+nonzero entries in $\lambda$-Matrix.
+There usually exists multiple NBs in $\Fkk, k>3$. If we employ a random NB, there is no mathematical
+guarantee that $C_N$ has bound $o(k)$. However,
+Mullin et al. \cite{mullinONB} % This citation is valid
+proves that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
+its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
+called optimal normal basis (ONB), details are introduced in appendix \ref{append:ONB}.
+
+In practice, large size NB multipliers are usually designed in $\Fkk$ when ONB exists
+to minimize the number of gates. So in the following part of this chapter and our experiments,
+we only focus on ONB multipliers instead of general NB multipliers.
+
+% After fixing this, add a whole-piece StdB vs NB example, list their cost as the conclusion
+\subsection{Comparison between Standard Basis and Normal Basis}
+At the end of this section, a detailed example is used to make a comparison between StdB multiplication
+and NB multiplication.
+\begin{Example}[Rijndael's finite field]
+Rijndael uses a characteristic 2 finite field with 256 elements, which can also be called the GF $\mathbb F_{2^8}$.
+Let us define the primitive element $\alpha$ using irreducible polynomial 
+$\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$. Coincidently, $\alpha$ is also a normal element,
+i.e. $\beta = \alpha$ can construct a NB $\{\alpha,\alpha^2,\alpha^4,\alpha^8,\alpha^{16},\alpha^{32},\alpha^{64},
+\alpha^{128}\}$.
+
+We pick a pair of elements from the Rijndael's field: $A=(0100~1011)_{StdB} = (4B)_{StdB},~B=(1100~1010)_{StdB} = (CA)_{StdB}$.
+First let us compute their product in StdB, the rule follows ordinary polynomial multiplication.
+
+\begin{align*}
+A\cdot B &= (\alpha^6+\alpha^3+\alpha+1)(\alpha^7+\alpha^6+\alpha^3+\alpha)\\
+&= (\alpha^{13}+\alpha^{10}+\alpha^8+\alpha^7)+(\alpha^{12}+\alpha^9+\alpha^7+\alpha^6)+
+	(\alpha^9+\alpha^6+\alpha^4+\alpha^3)\\
+	&~~~+(\alpha^7+\alpha^4+\alpha^2+\alpha) \\
+&= \alpha^{13}+\alpha^{12}+\alpha^{10}+\alpha^8+\alpha^7+\alpha^3+\alpha^2+\alpha
+\end{align*}
+Note that this polynomial is not the final form of the product because it needs to be
+reduced modulo irreducible polynomial $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$.
+This can be done using base-2 long division. Note the dividend and divisor are written in pseudo Boolean
+vectors, not real Boolean vectors in any kind of bases.
+
+% The folowing part is customized latex code mimicing binary long division
+% Note \divrule coordinates may not reflect the intuitive column # in tabular. Adjust by urself!!!
+
+\vspace{1cm}
+
+\newdimen\digitwidth
+\settowidth\digitwidth{0}
+\def~{\hspace{\digitwidth}}
+
+\def\divrule#1#2{%
+\noalign{\moveright#1\digitwidth%
+\vbox{\hrule width#2\digitwidth}}}
+111010111\,\begin{tabular}[b]{@{}r@{}}
+101001 \\ \hline
+\big)\begin{tabular}[t]{@{}l@{}}
+11010110001110 \\
+111010111 \\ \divrule{0}{14}
+~~111101101 \\
+~~111010111 \\ \divrule{2}{12}
+~~~~~111010110 \\
+~~~~~111010111 \\ \divrule{5}{9}
+~~~~~~~~~~~~~1
+\end{tabular}
+\end{tabular}
+\vspace{0.5cm}
+
+The final remainder is $1$, i.e. the product equals to 1 in StdB.
+
+On the other hand, operands $A$ and $B$ can be written in NB as
+$$A = (0010~1001)_{NB},~~B = (0100~0010)_{NB}$$
+The $\lambda$-Matrix for $\mathbb F_{2}[x] \pmod{x^8+x^7+x^6+x^4+x^2+x+1}$
+is
+(Computation of $\lambda$-Matrix refers to appendix \ref{append:NB})
+\begin{equation*}
+M^{(0)} = \left(\begin{array}{lccccccr}
+0 &0 &0 &0 &1 &0 &1 &1 \\
+0 &0 &1 &1 &1 &1 &0 &0 \\
+0 &1 &0 &0 &0 &0 &1 &0 \\
+0 &1 &0 &0 &1 &1 &0 &1 \\
+1 &1 &0 &1 &0 &1 &0 &0 \\
+0 &1 &0 &1 &1 &0 &0 &1 \\
+1 &0 &1 &0 &0 &0 &0 &0 \\
+1 &0 &0 &1 &0 &1 &0 &1
+\end{array}\right)
+\end{equation*}
+Taking matrix multiplication $c_0 = A\times M^{(0)}\times B^T$,
+the result is $c_0 = 1$. Then by cyclic shifting $A$ and $B$
+(or shifting $M^{(0)}$, either is applicable), we can successively
+obtain other bits of product. The final answer is
+$$C = (0000~0001)_{NB}$$
+It is equivalent to the result in StdB.
+\end{Example}
+
+% From the intuition of humans, StdB multiplication is straightforward and easier to understand
+% while NB is difficult to comprehend. However, if we implement both multiplications to 
+% hardware multipliers, it will be clear which side a circuit designer prefers.
+
+% May consider providing details?
+
+Mastrovito multiplier \cite{mastro:1989} and Montgomery multiplier \cite{PT:1985} are 2 common designs
+of GF multipliers using StdB. As a naive implementation of GF multiplication,
+Mastrovito multiplier uses most number of gates:
+$k^2$ AND gates and less than $k^2$ XOR gates \cite{Mastrovito}. Montgomery multiplier 
+applies lazy reduction techniques and results in a better latency performance, while the number of gates are about
+the same with Mastrovito multiplier:
+$k^2$ AND gates and $k^2-k/2$ XOR gates \cite{wu:2002}.
+
+For an 8-bit ($\mathbb F_{2^8}$) multiplier, typical design of Mastrovito multiplier consists of 218 logic gates, while 
+Montgomery multiplier needs 198 gates. However, the NB multiplier reuses the $\lambda$-Matrix 
+logic, so this component will only need to be implemented for once. 
+Consider the definition of matrix multiplication, it needs $C_N$ AND gates to apply 
+bit-wise multiplication and $C_N-1$ XOR gates to sum the intermediate products up. The number of nonzero entries
+in the $\lambda$-Matrix can be counted: $C_N = 27$.
+As a result, the most naive NB multiplier design (or Massey-Omura multiplier \cite{MasseyOmura})
+contains 53 gates in total, which is a great saving in area cost comparing to StdB multipliers.
+
+\section{Design a Normal Basis Multiplier on Gate Level}
+The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
+we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
+very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
+test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
+routine of several kinds of modern NB multipliers.
+\subsection{Sequential Multiplier with Parallel Outputs}
+The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
+the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
+\caption{A typical SMSO structure of Massey-Omura multiplier}
+\label{fig:MasseyOmura}}
+\end{figure}
+
+Figure \ref{fig:MasseyOmura} shows the basic architecture of a Massey-Omura multiplier. The operands
+$A$ and $B$ are 2 arrays of flip-flops which allow 1-bit right-cyclic shift every clock cycle.
+The logic gates in the boxes implements the matrix multiplication with $\lambda$-Matrix $M^{(0)}$, 
+while each AND gate corresponds to term $a_ib_j$ and each XOR gate corresponds to addition $a_ib_j+a_{i'}b_{j'}$. 
+The XOR layer has only 1 output, giving out 1 bit of product $C$ every clock cycle.
+
+The behavior of Massey-Omura multiplier can be concluded as: pre-load operands $A,B$ and reset $C$ to 0, after 
+executing for $k$ clock cycles, the data stored in flip-flop array $C$ is the product $A\times B$.
+We note that there is only one output giving 1 bit of the product each clock cycle, which matches the 
+definition of serial output to communication channel. Therefore this type of design is named as 
+sequential multiplier with serial output (SMSO).
+The SMSO architecture need $C_N$ AND gates and $C_N-1$ XOR gates, which equals to $2k-1$ AND gates and $2k-2$ XOR gates
+if it is designed using ONB. In fact, the number of gates can be reduced if the multiplication is 
+implemented using a conjugate of SMSO.
+
+The gate-level logic boxes are implementing following function:
+\begin{equation}
+\label{eqn:SMPOterms}
+c_l = row_1(A\times M^{(l)}) \times B + row_2(A\times M^{(l)}) \times B + \cdots + row_{k}(A\times M^{(l)}) \times B
+\end{equation}
+It can be decomposed into $k$ terms. If we only compute one term for each $c_l,~0\leq l\leq k-1$ in one 
+clock cycle, make $k$ outputs and add them up using shift register after $k$ clock cycles, 
+it will generate the same result with SMSO. This kind of architecture is named as 
+sequential multiplier with parallel outputs (SMPO). The basic SMPO, as a conjugate of Massey-Omura multiplier,
+is invented by Agnew et al. \cite{agnew1991implementation}.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/mySMPO.eps}
+\caption{5-bit Agnew's SMPO.
+Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, i.e.
+if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
+Index $w = 2i\pmod 5$}
+\label{fig:SMPO}}
+\end{figure}
+
+\begin{Example}[5-bit Agnew's SMPO]
+Given GF $\mathbb F_{2^5}$ and primitive element $\alpha$ defined by irreducible polynomial 
+$\alpha^5+\alpha^2+1=0$, normal element $\beta = \alpha^5$ constructs an ONB $\{\beta,\beta^2,\beta^4,\beta^8,\beta^{16}\}$.
+The $0$-th $\lambda$-Matrix for this ONB is
+\begin{equation*}
+M^{(0)} = \left(\begin{array}{lcccr}
+0 &1 &0 &0 &0 \\
+1 &0 &0 &1 &0 \\
+0 &0 &0 &1 &1 \\
+0 &1 &1 &0 &0 \\
+0 &0 &1 &0 &1
+\end{array}\right)
+\end{equation*}
+Then a typical design of 5-bit Agnew's SMPO is depicted in figure \ref{fig:SMPO}.
+
+The operands part of this circuit is the same with Massey-Omura multiplier. The differences are on 
+the matrix multiplication part, while it is implemented as separate logic blocks for 5 outputs,
+and the 5 blocks are connected in a shift register fashion. By analyzing the detailed function of 
+logic blocks, we can reveal the mechanism of Agnew's SMPO.
+
+Suppose we implement $M^{(0)}$ as the logic block in SMSO. In the first clock cycle, the output is
+\begin{equation}
+\label{eqn:5bitSMPO_c0}
+c_0 = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4
+\end{equation}
+Note it is written in the form of equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
+remaining bits of the product, which can be written in following general form polynomial:
+\begin{align}
+c_i =& b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
+&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
+(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
+\end{align}
+Note all index calculations are reduced modulo 5.
+
+Now let us observe the behavior of 5-bit Agnew's SMPO. Initially all DFFs are reset to 0. 
+In the first clock cycle,
+signal sent to the flip-flop in block $R_0$ denotes function:
+$$R_0^{(1)} = a_1b_0$$
+It equals to the first term of equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
+this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
+from operands (shifted by 1 bit), generating signal $a_u,a_v$ and $b_w$. Concretely,
+signal sent to flip-flop in block $R_1$ is:
+$$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
+which forms first 2 terms of equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
+on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
+$$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
+which equals to $c_0$ in equation \ref{eqn:5bitSMPO_c0}.
+After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
+the result of $c_0$ is computed after 5 clock cycles and given on $r_0$.
+
+If we track $R_1\to R_2\to R_3 \to \cdots \to R_0$, we can obtain $c_1$ respectively.
+Thus we conclude that Agnew's SMPO functions the same with Massey-Omura multiplier.
+\end{Example}
+
+The design of Agnew's SMPO guarantees that there is only one AND gate in each $R_i$ block.
+For ONB, adopting Agnew's SMPO will reduce the number of AND gates from $2k-1$ to $k$.
+
+\subsection{Multiplier not based on $\lambda$-Matrix}
+Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
+which means that they will be identical if unrolled to full combinational circuits. 
+After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
+some of them jump out of the circle and are independent from $\lambda$-Matrix.
+One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
+\cite{RHmulti}, which is therefore called RH-SMPO.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
+\caption{5-bit RH-SMPO}
+\label{fig:5bitRH}}
+\end{figure}
+
+Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
+in figure \ref{fig:SMPO}. A brief proof is as follows:
+
+\begin{Proof}
+First, we define an auxiliary function for $i$-th bit 
+\begin{equation}
+\label{eqn:aux}
+F_i(A,B) = a_ib_i\beta + \sum_{j=1}^v d_{i,j}\beta^{1+2^j}
+\end{equation}
+where $0\leq i\leq k-1, v = \lfloor k/2\rfloor, 1\leq j \leq v$.
+The $d$-layer index $d_{i,j}$ is defined as
+\begin{equation}
+\label{eqn:auxDC}
+d_{i,j} = c_{a,i} c_{b,i} = (a_i+a_{i+j})(b_i+b_{i+j}),~~1\leq j\leq v
+\end{equation}
+$i+j$ here is the result reduced modulo $k$. Note that there is a special boundary case when
+$k$ is an even number ($v = \frac{k}{2}$):
+$$d_{i,v} = (a_i+a_{i+v})b_i$$
+With the auxiliary function, we can utilize following theorem (proof refers to \cite{RHmulti}):
+\begin{Theorem}
+Consider three elements $A,B$ and $R$ such that $R=A\times B$ over $\Fkk$. Then,
+$$R=(((F_{k-1}^2+F_{k-2})^2+F_{k-3})^2+\cdots+F_1)^2+F_0$$
+\end{Theorem}
+This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
+$R_i$ flip-flops. Concretely, the multiplier behavior is an implementation of 
+following algorithm:
+
+\begin{algorithm}[H]
+\SetAlgoNoLine
+
+ \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
+ \KwOut{$R=A\times B$}
+%%%%%%%%%%%%%%%%%%%%
+  Initialize $A,B$ and aux var $X$ to 0\;
+  \For { ($i=0$;   $i < k$; ++i ) }
+  {
+	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Eq.\ref{eqn:aux}*/\;}
+	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
+  }
+  $R\gets X$
+\caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
+\end{algorithm}
+
+In this algorithm, we use a fixed auxiliary function $F_{k-1}$ inside the loop.
+This is because of equation
+$$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
+So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
+$F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
+\end{Proof}
+
+To better understand the mechanism of RH-SMPO, we will use this 5-bit 
+RH-SMPO as an example and introduce the details on how to design it.
+\begin{Example}[Designing a 5-bit RH-SMPO]
+From equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
+and XOR gates in $c$-layer according to equation \ref{eqn:auxDC}. Concretely, as algorithm \ref{alg:RHmulti}
+describes, we implement auxiliary function $F_{k-1}$ in the logic:
+$$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
+\begin{equation}
+\label{eqn:5bitRHaux}
+F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
+\end{equation}
+Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
+$$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
+$$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
+The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
+As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
+$F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
+turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
+$i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.
+
+In order to obtain information 
+indicating interconnections between $d$-layer and $e$-layer, we need to interpret $\beta^{1+2^j}$ 
+to NB representation.
+There is a concept called {\bf multiplication table} (M-table) which can assist this interpretation. It is defined as 
+a $k\times k$ matrix $T$ over $\mathbb F_2$:
+\begin{equation}
+\label{eqn:multitable}
+\begin{bmatrix}
+\beta^{1+2^0} \\ \beta^{1+2^1} \\ \beta^{1+2^2} \\ \vdots \\ \beta^{1+2^{k-1}}
+\end{bmatrix}
+= \beta
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+\end{bmatrix}
+=
+\begin{bmatrix}
+T_{0,0}      &   T_{0,1}        & \dots & T_{0,k-1}\\
+T_{1,0}    &   T_{1,1}           & \dots & T_{1,k-1}\\
+T_{2,0}    &   T_{2,1}           & \dots & T_{2,k-1}\\
+\vdots & \vdots              & \ddots     & \vdots \\
+T_{k-1,0}    &   T_{k-1,1}           & \dots & T_{k-1,k-1}
+\end{bmatrix}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+\end{bmatrix}
+= {\bf T}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
+\end{bmatrix}
+\end{equation}
+
+It is a known fact that M-table $T$ can be converted from $\lambda$-Matrix $M$:
+$$M_{i,j}^{(0)} = T_{j-i,-i}$$
+with indices reduced modulo $k$ (proof given in appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=3in]{newfig/multitable.pdf}
+\caption{A $5\times 5$ multiplication table}
+\label{fig:multitable}}
+\end{figure}
+
+Note that we only use row 1 and row 2 from the M-table since range $1\leq j \leq 2$.
+All nonzero entries in these 2 rows corresponds to the interconnections between $d$-layer and 
+$e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
+between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in equation \ref{eqn:multitable}:
+\begin{equation*}
+\beta\cdot\beta^2 = 
+\begin{bmatrix}
+1 & 0 & 0 & 1 & 0
+\end{bmatrix}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
+\end{bmatrix}
+= \beta + \beta^{2^3}
+\end{equation*}
+
+Similarly, from row 2 of M-table we derive that $d_2$ has fanouts $e_3,e_4$:
+\begin{equation*}
+\beta\cdot\beta^{2^2} = 
+\begin{bmatrix}
+0 & 0 & 0 & 1 & 1
+\end{bmatrix}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
+\end{bmatrix}
+= \beta^{2^3} + \beta^{2^4}
+\end{equation*}
+
+Let us look back at equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
+The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
+After gathering all interconnection information, we can translate it to gate-level circuit implementation:
+$$e_0 = d_0+d_1,~e_3=d_1+d_2,~e_4=d_2$$
+
+Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
+$R_{i-1}$ in last clock cycle, we can connect using relation 
+$$R_i = r_{i-1} + e_i$$
+In this example, according to the M-table in figure \ref{fig:multitable}, columns $e_1,e_2$
+have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.
+
+This finishes the full design procedure for a 5-bit RH-SMPO.
+\end{Example}
+
+The area cost of RH-SMPO is even smaller than Agnew's SMPO. XOR gates corresponds to all nonzero entries 
+in M-table, which is with the same number of nonzero entries in $\lambda$-Matrix ($C_N$). The number of AND gates 
+equals to $v$ plus 1 (for gate $d_0$). When using ONB ($C_N = 2k-1$), the total number of gates 
+is $2k+\lfloor \frac{k}{2}\rfloor$.
+
+\section{Concluding Remarks}
+In this section, we introduce basic concepts such as the definition of finite fields and the construction 
+of finite fields. Moreover, we describes a special kind of basis in finite field, and its 
+application on Galois field hardware design. The sequential Galois field multipliers working based on these 
+principles are used as good candidates for applying our functional verification approach in subsequent chapters.

### 2016-11-21 dissertation/prelim_basis.tex
dissertation v2_finish prelims

--- 2016-09-21 dissertation/algprelim.tex
@@
-\chapter{Computer Algebra Fundamentals} \label{ch:ideals}
+\chapter{Gr\"ober Basis and Algebraic Geometry} \label{ch:ideals}
 This chapter reviews fundamental concepts of commutative and 
@@
 mostly referred from the textbooks \cite{ideals:book} \cite{gb_book} and 
-previous work by {\it Lv} \cite{lv:phd}.
-
-\section{Monomials, Polynomials, and Term Orderings}
-
+previous work by {\it Lv} \cite{lv:phd} as well as {\it Pruss} \cite{pruss:tcad15}
+
+\section{Algebraic Geometry Fundamentals}
+\label{sec:geo}
+\subsection{Monomials, Polynomials and Polynomial Arithmetics}
 \begin{Definition} \label{def:mono}
@@
 A {\bf multivariate polynomial} $f$ in variables $x_1, x_2, \ldots, x_d$ 
-with coefficients in any given field $\K$ is a finite linear 
-combination of monomials with coefficients in $\K$: 
-\begin{equation}
-	f=\sum_{\alpha}a_{\alpha}\cdot x^{\alpha}, ~~a_{\alpha}\in \K \nonumber
+with coefficients in any given field $\F$ is a finite linear 
+combination of monomials with coefficients in $\F$: 
+\begin{equation}
+	f=\sum_{\alpha}a_{\alpha}\cdot x^{\alpha}, ~~a_{\alpha}\in \F \nonumber
 \end{equation}
@@
 The set of all polynomials in $x_1, x_2, \ldots, x_d$ with coefficients in 
-field $\K$ is denoted by $\K[x_1, x_2, \ldots, x_d]$. 
-Thus, $f \in \K[x_1, x_2, \ldots, x_d]$
+field $\F$ is denoted by $\F[x_1, x_2, \ldots, x_d]$. 
+Thus, $f \in \F[x_1, x_2, \ldots, x_d]$
 
 \begin{enumerate}
-\item We refer to the constant $a_{\alpha} \in \K$ as the 
+\item We refer to the constant $a_{\alpha} \in \F$ as the 
 {\bf coefficient} of the monomial $a_{\alpha} x^{\alpha}$.
@@
 \begin{Definition}
-Let $\K$ be a field and let $f, g \in \K[x_1, x_2, \ldots, x_d]$ be polynomials
+Let $\F$ be a field and let $f, g \in \F[x_1, x_2, \ldots, x_d]$ be polynomials
 over the field. {\bf Polynomial division} of $f$ by $g$ is computes following:
@@
 \end{eqnarray}
-Notice that the division cancels the leading term of $f$.
-\end{Example}
-
-\section{Varieties and Ideals}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In computer-algebra based formal verification,
-it is often necessary to analyze the presence 
-or absence of solutions to a given system of constraints.
-In our applications, these constraints are polynomials and their solutions 
-are modeled as {\bf varieties}.
-
-\begin{Definition}
-Let $\K$ be a field, and let $f_1, \ldots, f_s 
-\in \K[x_1, x_2, \ldots, x_d]$. 
-We call $V(f_1, \dots, f_s)$ the {\bf affine variety} 
-defined by $f_1, \dots, f_s$ as:
-\begin{equation}
-V(f_1, \ldots, f_s)= \{(a_1, \ldots, a_{d})\in \K^d:f_i(a_1, \ldots, a_d)=0, \forall{i},1\le i \le s\}
-\end{equation}
-\end{Definition}
-
-$V(f_1, \dots, f_s)\in \K^d$ is {\bf the set of all solutions} in $\K^d$ 
-of the system of equations: 
-$f_1(x_1,\ldots,x_d)=\dots=f_s(x_1,\dots,x_d)=0$.
-
-\begin{Example}
-Given $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2)$ is the set of all elements
-that satisfy $x^2+y^2=0$ over $\mathbb{R}^2$. So $V(x^2+y^2)=\{(0,0)\}$. 
-Similarly, in $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2-1)=\{all\  points\  on\ the\ circle: x^2+y^2-1=0\}$.
-Note that varieties depend on which field we are operating on. 
-For the same polynomial $x^2+1$, we have:
-\begin{itemize}
-\item In $\mathbb{R}[x]$, $V(x^2+1)=\emptyset$.
-\item In $\mathbb{C}[x]$, $V(x^2+1)=\{(\pm i)\}$.
-\end{itemize}
-\end{Example}
-
-The above example shows the variety can be infinite, finite (non-empty set) 
-or empty. It is interesting to note that since we will be operating over
-finite fields $\F_{q}$, and any finite set of points is a variety. Likewise,
-any variety over $\F_{q}$ is finite (or empty).
-Consider the points $\{(a_1,\dots, a_d): a_1, \dots, a_d \in \F_q\}$
-in $\F_q^d$. Any single point is a variety of some polynomial system:
-e.g. $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
-x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
-varieties are also varieties.
-
-\begin{Example}
-Let $U = V(f_1, \dots, f_s)$ and $W =
-V(g_1, \dots, g_t)$ in $\F_{q}$. Then:  
-\begin{itemize}
-\item $U \cap W = V(f_1, \dots, f_s, g_1, \dots, g_t)$
-\item $U \cup W = V(f_i g_j: 1 \leq i \leq s, 1 \leq j \leq t)$
-\end{itemize}
-\end{Example}
-
-One important distinction we need to make about varieties is that a
-variety depends not just on the given system of polynomial 
-equations, but rather on the {\bf ideal} generated by the polynomials.
-
-%%%%%%%%%%%%%%%%%%%%%  ideal %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{Definition} 
-A subset $I \subset \K[x_1, x_2, \ldots, x_d]$ is an {
-\bf ideal} if it satisfies:
-\begin{itemize}
-\item $0 \in I$
-\item $I$ is closed under addition: $x, y \in I \Rightarrow x+y \in I$
-\item If $x \in \K[x_1, x_2, \ldots, x_d]$ and $y \in I$, then $x\cdot y \in I$ and $y\cdot x \in  I$.
-\end{itemize}
-\end{Definition}
-
-An ideal is generated by its {\it basis} or {\it generators}.
-
-%%%%%%%%%%%%%%%%%%%%%  ideal basis%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{Definition}
-Let $f_1, f_2, \ldots, f_s$ be 
-polynomials of the ring $\K[x_1, x_2, \ldots, x_d]$. 
-Let $I$ be an ideal generated by $f_1, f_2, \ldots, f_s$. Then:
-\begin{equation}
-I = \langle f_1,\dots,f_s \rangle=\{h_1 f_1 + h_2 f_2 + \ldots + h_s f_s : h_1,\dots,h_s\in\K[x_1, \dots, x_d]\} \nonumber
-\end{equation}
-then, $f_1, \ldots, f_s$ are called the {\bf basis (or generators)} of the 
-ideal $I$ and correspondingly $I$ is denoted as $I = \langle f_1, f_2, 
-\ldots, f_s \rangle$. 
-\end{Definition}
-
-\begin{Example}
-The set of even integers, which is a subset of the ring of integers
-$Z$, forms an ideal of $Z$. This can be seen from the following;
-\begin{itemize}
-\item $0$ belongs to the set of even integers.
-\item The sum of two even integers $x$ and $y$ is always an even
-  integer.
-\item The product of any integer $x$ with an even integer $y$ is
-  always an even integer.
-\end{itemize}
-\end{Example}
-
-\begin{Example}
-Given $\mathbb{R}\left[x,y\right]$, $I = \langle x, y \rangle$ is an 
-ideal containing all polynomials generated by $x$ and $y$, 
-such as $x^2+y$ and $x+x\cdot y$.
-$J = \langle x^2, y^2 \rangle$ is an ideal containing all polynomials 
-generated by $x^2$ and $y^2$, such as $x^2+y^3$ and $x^{10}+x^2\cdot y^2$. 
-Notice that $J\subset I$ because every polynomial generated by $J$ can
-be generated by $I$. 
-But $I\neq J$ because $x+y$ can only be generated by $I$.
-\end{Example}
-
-The same ideal may have many different bases.
-For instance, it is possible to have different sets of polynomials
-$\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
-ideal, i.e., 
-$\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
-variety depends on the ideal, these sets of polynomials have the same 
-solutions.
-
-\begin{Proposition}
-If $f_1,\dots,f_{s}$ and $g_{1},\dots,g_{t}$ are bases of the same ideal 
-in $\mathbb{F}[x_{1},\dots,x_{d}]$,
-so that $\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$, 
-then $V(f_{1},\dots,f_{s})=V(g_{1},\dots,g_{t})$.
-\end{Proposition}
-
-\begin{Example}
-	Consider the two bases $F_{1}=\{(2x^{2}+3y^{2}-11,x^{2}-y^{2}-3\}$ and $F_{2}=\{x^{2}-4,y^{2}-1\}$.
-	These two bases generate the same ideal, i.e., $\langle F_{1}\rangle= \langle F_{2} \rangle$.{}
-	Therefore, they represent the same variety, i.e., 
-	\begin{equation}
-		V(F_{1})= V( F_{2})=\{\pm 2, \pm 1\}.
-	\end{equation}
-\end{Example}
-
-Ideals and their varieties are a key part of computer-algebra based formal
-verification. A given hardware design can be transformed into a set
-of polynomials over a field, $f_1, \ldots, f_s \in F$ 
-(we showed how this is done for Galois field arithmetic 
-circuits in the previous chapter).
-This set of polynomials gives the system of equations:
-\begin{eqnarray}
-f_1 = 0 \nonumber \\
-\vdots \nonumber \\
-f_s = 0 \nonumber
-\end{eqnarray}
-Using algebra, it is possible to derive new equations from the original 
-system.
-The ideal $\langle f_1,\ldots, f_s \rangle$ provides a way of analyzing 
-such {\it consequences} of a system of polynomials.
-
-\begin{Example}
-Given two equations in $\mathbb{R}[x,y,z]$:
-\begin{eqnarray}
-x=z+1 \nonumber \\
-y=x^2+1 \nonumber 
-\end{eqnarray}
-we can eliminate $x$ to obtain a new equation:
-\begin{equation}
-y=(z+1)^2+1=z^2+2z+2 \nonumber 
-\end{equation}
-Let $f_1, f_2, h \in \mathbb{R}[x,y,z]$ be polynomials based on these 
-equations:
-\begin{eqnarray}
-f_1 = x-z-1 &= 0 \nonumber \\
-f_2 = y-x^2-1 &= 0 \nonumber \\
-h   = y-z^2-2z-2 &= 0 \nonumber
-\end{eqnarray}
-If $I$ is the ideal generated by $f_1$ and $f_2$, i.e. 
-$I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
-\begin{eqnarray}
-g_1 = x+z+1 \nonumber \\
-g_2 = 1     \nonumber \\
-h = g_1\cdot f_1+g_2\cdot f_2  = y-z^2-2z-2 \nonumber
-\end{eqnarray}
-where $g_1, g_2 \in \mathbb{R}[x,y,z]$.
-Thus, we call $h$ a {\bf member of the ideal} $I$.
-\end{Example}
-
-%\subsection{Ideals of Varieties}
-
-Let $\mathbb{K}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{K}^d$ be a point, and $f \in
-\mathbb{K}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ if $f(\mathbf{a}) = 0$, i.e.,
-$\mathbf{a}$ is in the variety of $f$.
-
-\begin{Definition}
-For any variety $V$ of $\mathbb{K}^d$, the ideal of polynomials that vanish on $V$,
-called the {\it vanishing ideal of $V$}, is defined as $I(V) = \{f\in
-\mathbb{F}[x_1,\dots, x_d]: \forall \mathbf{a} \in V, f(\mathbf{a}) =
-0\}$. 
-\end{Definition}
-
-\begin{Proposition}\label{pro:iofv}
-	If a polynomial $f$ vanishes on a variety $V$, then $f \in I(V)$. 
-\end{Proposition}
-
-\begin{Example}
-	Let ideal $J=\langle x^{2},y^{2}\rangle$. Then $V(J)=\{(0,0)\}$.
-	All polynomials in $J$ will obviously agree with the solution and vanish on this variety.
-	However, the polynomials $x,y$ are not in $J$ but they also vanish on this variety. 
-	Therefore, $I(V(J))$ is the set of all polynomials that vanish on $V(J)$, and the polynomials
-	$x,y$ are members of $I(V(J))$.
-\end{Example}
-
-\begin{Definition}\label{def:radical}
-Let $J \subset \mathbb{K}[x_1,\dots, x_d]$ be an ideal. The {\it radical of $J$} is defined as $\sqrt{J} = \{f \in
-\mathbb{K}[x_1,\dots, x_d]: \exists m \in \mathbb{N}, f^m \in J\}$. 
-\end{Definition}
-
-\begin{Example}
-Let $J=\langle x^2,y^2\rangle \subset \mathbb{K}\left[x,y\right]$.
-Note neither $x$ nor $y$ belongs to $J$, but they belong to $\sqrt J$.
-Similarly, $x\cdot y \notin J$, but since $(x \cdot y)^{2}=x^{2}\cdot y^{2}\in J$, therefore,
-$x\cdot y \in \sqrt J$. 
-\end{Example}
-
-When $J = \sqrt J$, then $J$ is said to be a 
-{\it radical ideal}. Moreover, $I(V)$ is a radical ideal.
-By analyzing the ideal $J$, generated by a system of polynomials derived 
-from a hardware design, its variety $V(J)$, and the ideal of 
-polynomials that vanish over this variety, $I(V(J))$, we can reason about the 
-existence of certain properties of the design. To check for the existence of 
-a property, we formulate the 
-property as a polynomial and then perform an {\bf ideal membership test} to 
-determine if this polynomial is contained within the ideal $I(V(J))$. 
-A {\bf \Grobner basis} provides a decision procedure for performing this 
-test, which is described in the following section. 
-A future section focuses on {\bf Hilbert's Nullstellensatz}, 
-which describes the properties of the ideal of a variety, $I(V(J))$.
-
-%%%%%%%%%%%%%%%%%%%%	Grobner bases	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{\Grobner Bases}
-
-As mentioned earlier, different polynomial sets may generate the same 
-ideal. Some of these generating sets may be a better representation of 
-the ideal, and thus provide more information and insight into the properties 
-of ideal. One such ideal representation is a {\bf \Grobner basis}, which has
-a number of important properties that can solve numerous polynomial 
-decision questions:
-
-\begin{itemize}
-\item Presence or absence of solutions (varieties)
-\item Dimension of the varieties
-\item Ideal membership of a polynomial
-\end{itemize}
-
-In essence, a \Grobner basis is a canonical representation of an ideal.
-There are many equivalent definitions of \Grobner bases, so we start with 
-the definition that best describes their properties:
-
-\begin{Definition}
-A set of non-zero polynomials $G=\{g_1,\dots,g_t\}$ which generate the 
-ideal $I=\langle g_1,\dots,g_t\rangle$, is called a 
-{\bf Gr\"obner basis} for $I$ if and only if 
-for all $f \in I$ where $f \neq 0$, there exists a $g_i \in G$ such that $lm(g_i)$ divides $lm(f)$.
-\begin{eqnarray}
-G = \text{Gr\"obner{Basis}} (I) \iff 
-\forall f \in I: f \neq 0, \exists g_i \in G: lm(g_i)\ |\ lm(f)
-        \label{eqn:groebnermin}
-    \end{eqnarray}
-
-\end{Definition}
-
-The foundation for computing the \Grobner basis of an ideal was laid out 
-by Buchberger\cite{buchberger_thesis}.
-Given a set of polynomials $F=\{f_{1},\dots,f_{s}\}$ that generate ideal $I=
-\langle f_{1},\dots,f_{s} \rangle$, 
-Buchberger gives an algorithm to compute a Gr\"obner basis $G=\langle g_{1},
-\dots,g_{t}\rangle$. This algorithm relies on the notions of $S$-polynomials 
-and polynomial reduction.
-
-\begin{Definition}
-For $f, g \in \K[x_1,\dots,x_d]$, an {\bf S-polynomial} $Spoly(f,g)$ is 
-defined as:
-\begin{equation}
-    Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g
-    \label{eqn:spoly}
-\end{equation}
-\begin{equation}
-\text{where }L = lcm\left(lt(f), lt(g)\right) \nonumber
-\end{equation}
-Note $lcm$ denotes least common multiple.
-\end{Definition}
-
+Notice that the division $f\stackrel{g}{\textstyle\longrightarrow}r$ cancels the leading term of $f$.
+\end{Example}
+
+Similarly, we can also define when a polynomial is divided (reduced) by a set of polynomials.
 \begin{Definition}
@@
     \begin{equation*}
-        f\stackrel{g}{\textstyle\longrightarrow}r
+        f\stackrel{g}{\textstyle\longrightarrow}_{+}r
     \end{equation*}
+    which is transitive and reflective closure of relation $f\stackrel{g}{\textstyle\longrightarrow}r$. 
     Reduction is carried out using multivariate, polynomial long division. 
@@
     set of non-zero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
-    $r = 0$  or no term in $r$ is divisible  by a $lm(f_i), \forall f_i \in F$.
+    $r = 0$  or no term in $r$ is divisible  by $lm(f_i), \forall f_i \in F$.
 \end{Definition}
@@
 \end{Example}
+
+\subsection{Varieties and Ideals}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In computer-algebra based formal verification,
+it is often necessary to analyze the presence 
+or absence of solutions to a given system of constraints.
+In our applications, these constraints are polynomials and their solutions 
+are modeled as {\bf varieties}.
+
+\begin{Definition}
+Let $\F$ be a field, and let $f_1, \ldots, f_s 
+\in \F[x_1, x_2, \ldots, x_d]$. 
+We call $V(f_1, \dots, f_s)$ the {\bf affine variety} 
+defined by $f_1, \dots, f_s$ as:
+\begin{equation}
+V(f_1, \ldots, f_s)= \{(a_1, \ldots, a_{d})\in \F^d:f_i(a_1, \ldots, a_d)=0, \forall{i},1\le i \le s\}
+\end{equation}
+\end{Definition}
+
+$V(f_1, \dots, f_s)\in \F^d$ is {\bf the set of all solutions} in $\F^d$ 
+of the system of equations: 
+$f_1(x_1,\ldots,x_d)=\dots=f_s(x_1,\dots,x_d)=0$.
+
+\begin{Example}
+Given $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2)$ is the set of all elements
+that satisfy $x^2+y^2=0$ over $\mathbb{R}^2$. So $V(x^2+y^2)=\{(0,0)\}$. 
+Similarly, in $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2-1)=\{all\  points\  on\ the\ circle: x^2+y^2-1=0\}$.
+Note that varieties depend on which field we are operating on. 
+For the same polynomial $x^2+1$, we have:
+\begin{itemize}
+\item In $\mathbb{R}[x]$, $V(x^2+1)=\emptyset$.
+\item In $\mathbb{C}[x]$, $V(x^2+1)=\{(\pm i)\}$.
+\end{itemize}
+\end{Example}
+
+The above example shows the variety can be infinite, finite (non-empty set) 
+or empty. It is interesting to note that since we will be operating over
+finite fields $\F_{q}$, and any finite set of points is a variety. Likewise,
+any variety over $\F_{q}$ is finite (or empty).
+Consider the points $\{(a_1,\dots, a_d): a_1, \dots, a_d \in \F_q\}$
+in $\F_q^d$. Any single point is a variety of some polynomial system:
+e.g. $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
+x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
+varieties are also varieties.
+
+\begin{Example}
+Let $U = V(f_1, \dots, f_s)$ and $W =
+V(g_1, \dots, g_t)$ in $\F_{q}$. Then:  
+\begin{itemize}
+\item $U \cap W = V(f_1, \dots, f_s, g_1, \dots, g_t)$
+\item $U \cup W = V(f_i g_j: 1 \leq i \leq s, 1 \leq j \leq t)$
+\end{itemize}
+\end{Example}
+
+One important distinction we need to make about varieties is that a
+variety depends not just on the given system of polynomial 
+equations, but rather on the {\bf ideal} generated by the polynomials.
+
+%%%%%%%%%%%%%%%%%%%%%  ideal %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{Definition} 
+A subset $I \subset \F[x_1, x_2, \ldots, x_d]$ is an {
+\bf ideal} if it satisfies:
+\begin{itemize}
+\item $0 \in I$
+\item $I$ is closed under addition: $x, y \in I \Rightarrow x+y \in I$
+\item If $x \in \F[x_1, x_2, \ldots, x_d]$ and $y \in I$, then $x\cdot y \in I$ and $y\cdot x \in  I$.
+\end{itemize}
+\end{Definition}
+
+An ideal is generated by its {\it basis} or {\it generators}.
+
+%%%%%%%%%%%%%%%%%%%%%  ideal basis%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{Definition}
+Let $f_1, f_2, \ldots, f_s$ be 
+polynomials of the ring $\F[x_1, x_2, \ldots, x_d]$. 
+Let $I$ be an ideal generated by $f_1, f_2, \ldots, f_s$. Then:
+\begin{equation}
+I = \langle f_1,\dots,f_s \rangle=\{h_1 f_1 + h_2 f_2 + \ldots + h_s f_s : h_1,\dots,h_s\in\F[x_1, \dots, x_d]\} \nonumber
+\end{equation}
+then, $f_1, \ldots, f_s$ are called the {\bf basis (or generators)} of the 
+ideal $I$ and correspondingly $I$ is denoted as $I = \langle f_1, f_2, 
+\ldots, f_s \rangle$. 
+\end{Definition}
+
+\begin{Example}
+The set of even integers, which is a subset of the ring of integers
+$Z$, forms an ideal of $Z$. This can be seen from the following;
+\begin{itemize}
+\item $0$ belongs to the set of even integers.
+\item The sum of two even integers $x$ and $y$ is always an even
+  integer.
+\item The product of any integer $x$ with an even integer $y$ is
+  always an even integer.
+\end{itemize}
+\end{Example}
+
+\begin{Example}
+Given $\mathbb{R}\left[x,y\right]$, $I = \langle x, y \rangle$ is an 
+ideal containing all polynomials generated by $x$ and $y$, 
+such as $x^2+y$ and $x+x\cdot y$. The ideal
+$J = \langle x^2, y^2 \rangle$ is an ideal containing all polynomials 
+generated by $x^2$ and $y^2$, such as $x^2+y^3$ and $x^{10}+x^2\cdot y^2$. 
+Notice that $J\subset I$ because every polynomial generated by $J$ can
+be generated by $I$. 
+But $I\neq J$ because $x+y$ can only be generated by $I$.
+\end{Example}
+
+The same ideal may have many different bases.
+For instance, it is possible to have different sets of polynomials
+$\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
+ideal, i.e., 
+$\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
+variety depends on the ideal, these sets of polynomials have the same 
+solutions.
+
+\begin{Proposition}
+If $f_1,\dots,f_{s}$ and $g_{1},\dots,g_{t}$ are bases of the same ideal 
+in $\mathbb{F}[x_{1},\dots,x_{d}]$,
+so that $\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$, 
+then $V(f_{1},\dots,f_{s})=V(g_{1},\dots,g_{t})$.
+\end{Proposition}
+
+\begin{Example}
+	Consider the two bases $F_{1}=\{(2x^{2}+3y^{2}-11,x^{2}-y^{2}-3\}$ and $F_{2}=\{x^{2}-4,y^{2}-1\}$.
+	These two bases generate the same ideal, i.e., $\langle F_{1}\rangle= \langle F_{2} \rangle$.{}
+	Therefore, they represent the same variety, i.e., 
+	\begin{equation}
+		V(F_{1})= V( F_{2})=\{\pm 2, \pm 1\}.
+	\end{equation}
+\end{Example}
+
+Ideals and their varieties are a key part of computer-algebra based formal
+verification. A given hardware design can be transformed into a set
+of polynomials over a field, $f_1, \ldots, f_s \in F$.
+This set of polynomials gives the system of equations:
+\begin{eqnarray}
+f_1 = 0 \nonumber \\
+\vdots \nonumber \\
+f_s = 0 \nonumber
+\end{eqnarray}
+Using algebra, it is possible to derive new equations from the original 
+system.
+The ideal $\langle f_1,\ldots, f_s \rangle$ provides a way of analyzing 
+such {\it consequences} of a system of polynomials.
+
+\begin{Example}
+Given two equations in $\mathbb{R}[x,y,z]$:
+\begin{eqnarray}
+x=z+1 \nonumber \\
+y=x^2+1 \nonumber 
+\end{eqnarray}
+we can eliminate $x$ to obtain a new equation:
+\begin{equation}
+y=(z+1)^2+1=z^2+2z+2 \nonumber 
+\end{equation}
+Let $f_1, f_2, h \in \mathbb{R}[x,y,z]$ be polynomials based on these 
+equations:
+\begin{eqnarray}
+f_1 &= x-z-1 &= 0 \nonumber \\
+f_2 &= y-x^2-1 &= 0 \nonumber \\
+h   &= y-z^2-2z-2 &= 0 \nonumber
+\end{eqnarray}
+If $I$ is the ideal generated by $f_1$ and $f_2$, i.e. 
+$I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
+\begin{eqnarray}
+g_1 &= x+z+1 \nonumber \\
+g_2 &= 1     \nonumber \\
+h &= g_1\cdot f_1+g_2\cdot f_2  = y-z^2-2z-2 \nonumber
+\end{eqnarray}
+where $g_1, g_2 \in \mathbb{R}[x,y,z]$.
+Thus, we call $h$ a {\bf member of the ideal} $I$.
+\end{Example}
+
+%\subsection{Ideals of Varieties}
+
+Let $\mathbb{F}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{F}^d$ be a point, and $f \in
+\mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ if $f(\mathbf{a}) = 0$, i.e.,
+$\mathbf{a}$ is in the variety of $f$.
+
+\begin{Definition}
+For any variety $V$ of $\mathbb{F}^d$, the ideal of polynomials that vanish on $V$,
+called the {\it vanishing ideal of $V$}, is defined as {\bf ideal of variety}:
+$$I(V) = \{f\in
+\mathbb{F}[x_1,\dots, x_d]: \forall \mathbf{a} \in V, f(\mathbf{a}) =
+0\}$$ 
+\end{Definition}
+
+\begin{Proposition}\label{pro:iofv}
+	If a polynomial $f$ vanishes on a variety $V$, then $f \in I(V)$. 
+\end{Proposition}
+
+\begin{Example}
+	Let ideal $J=\langle x^{2},y^{2}\rangle$. Then $V(J)=\{(0,0)\}$.
+	All polynomials in $J$ will obviously agree with the solution and vanish on this variety.
+	However, the polynomials $x,y$ are not in $J$ but they also vanish on this variety. 
+	Therefore, $I(V(J))$ is the set of all polynomials that vanish on $V(J)$, and the polynomials
+	$x,y$ are members of $I(V(J))$.
+\end{Example}
+
+\begin{Definition}\label{def:radical}
+Let $J \subset \mathbb{F}[x_1,\dots, x_d]$ be an ideal. The {\bf radical of $J$} is defined as $\sqrt{J} = \{f \in
+\mathbb{F}[x_1,\dots, x_d]: \exists m \in \mathbb{N}, f^m \in J\}$. 
+\end{Definition}
+
+\begin{Example}
+Let $J=\langle x^2,y^2\rangle \subset \mathbb{F}\left[x,y\right]$.
+Neither $x$ nor $y$ belongs to $J$, but they belong to $\sqrt J$.
+Similarly, $x\cdot y \notin J$, but since $(x \cdot y)^{2}=x^{2}\cdot y^{2}\in J$, therefore,
+$x\cdot y \in \sqrt J$. 
+\end{Example}
+
+When $J = \sqrt J$, then $J$ is said to be a 
+{\it radical ideal}. Moreover, $I(V)$ is a radical ideal.
+By analyzing the ideal $J$, generated by a system of polynomials derived 
+from a hardware design, its variety $V(J)$, and the ideal of 
+polynomials that vanish over this variety, $I(V(J))$, we can reason about the 
+existence of certain properties of the design. To check for the validity of 
+a property, we formulate the 
+property as a polynomial and then perform an {\bf ideal membership test} to 
+determine if this polynomial is contained within the ideal $I(V(J))$. 
+A {\bf \Grobner basis} provides a decision procedure for performing this 
+test, which is described in the following part.
+
+\subsection{\Grobner Bases}
+
+As mentioned earlier, different polynomial sets may generate the same 
+ideal. Some of these generating sets may be a better representation of 
+the ideal, and thus provide more information and insight into the properties 
+of the ideal. One such ideal representation is a {\bf \Grobner basis}, which has
+a number of important properties that can solve numerous polynomial 
+decision questions:
+
+\begin{itemize}
+\item Presence or absence of solutions (varieties)
+\item Dimension of the varieties
+\item Ideal membership of a polynomial
+\item $\dots \dots$
+\end{itemize}
+
+In essence, a \Grobner basis is a canonical representation of an ideal.
+There are many equivalent definitions of \Grobner bases, so we start with 
+the definition that best describes their properties:
+
+\begin{Definition}
+A set of non-zero polynomials $G=\{g_1,\dots,g_t\}$ which generate the 
+ideal $I=\langle g_1,\dots,g_t\rangle$, is called a 
+{\bf Gr\"obner basis} for $I$ if and only if 
+for all $f \in I$ where $f \neq 0$, there exists a $g_i \in G$ such that $lm(g_i)$ divides $lm(f)$.
+\begin{eqnarray}
+G = \text{Gr\"obner{Basis}} (I) \iff 
+\forall f \in I: f \neq 0, \exists g_i \in G: lm(g_i)\ |\ lm(f)
+        \label{eqn:groebnermin}
+    \end{eqnarray}
+
+\end{Definition}
+
+\Grobner basis has an important property, and therefore can be used to perform ideal membership test.
+Formally speaking, a \Grobner basis gives a decision procedure to test 
+for polynomial membership in an ideal. This is explained in the following 
+Theorem.
+
+\begin{Theorem}\label{the:membership}
+	{\bf Ideal Membership Test}
+ Let $G = \{g_1,\cdots,g_t \}$ be a Gr\"obner basis for an ideal $I \subset \mathbb{K}[x_1,\cdots,x_d ]$
+	and let $f \in \mathbb{K}[x_{1},\dots, x_{d}]$. Then $f \in I$ if and only if the remainder on division of $f$ by
+	$G$ is zero.
+\end{Theorem}
+In other words, 
+\begin{equation}
+f \in I \iff f \stackrel{G}{\textstyle\longrightarrow}_+0
+\end{equation}
+
+\begin{Example}
+Consider Example \ref{exp:gbsimple}. Let $f = y^2x - x$ be another
+polynomial. Note that $f = yf_1 + f_2$, so $f \in I$. If we divide $f$
+by $f_1$ first and then by $f_2$, we will obtain a zero
+remainder. However, since the set $\{f_1, f_2\}$ is not a Gr\"{o}bner
+basis, we find that the reduction $f
+\stackrel{f_2}{\textstyle\longrightarrow} x^2 - x
+\stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
+i.e. dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
+zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
+$I$, $G = \{x^2 - x, yx - y, y^2 - x\}$, dividing $f$ by polynomials
+in $G$ in any order will always lead to the zero remainder. Therefore,
+one can decide ideal membership unequivocally using the Gr\"{o}bner
+basis. 
+\end{Example}
+
+The foundation for computing the \Grobner basis of an ideal was laid out 
+by Buchberger\cite{buchberger_thesis}.
+Given a set of polynomials $F=\{f_{1},\dots,f_{s}\}$ that generate ideal $I=
+\langle f_{1},\dots,f_{s} \rangle$, 
+Buchberger gives an algorithm to compute a Gr\"obner basis $G=\langle g_{1},
+\dots,g_{t}\rangle$. This algorithm relies on the notions of $S$-polynomials 
+and polynomial reduction.
+
+\begin{Definition}
+For $f, g \in \F[x_1,\dots,x_d]$, an {\bf S-polynomial} $Spoly(f,g)$ is 
+defined as:
+\begin{equation}
+    Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g
+    \label{eqn:spoly}
+\end{equation}
+\begin{equation}
+\text{where }L = lcm\left(lt(f), lt(g)\right) \nonumber
+\end{equation}
+Note $lcm$ denotes least common multiple.
+\end{Definition}
 
@@
 and $lt(f_j)$ have no common variables, the S-poly reduction step in 
-Buchberger's algorithm,
-$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$,
+Buchberger's algorithm will reduce to 0.
+\begin{Lemma}
+In Buchberger's algorithm, when $gcd(lt(f_i),lt(f_j)) = 0$, the S-poly reduction
+$$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$$
 will produce $r=0$.
+\end{Lemma}
 
@@
 \end{Proof}
-
-As mentioned previously, a \Grobner basis gives a decision procedure to test 
-for polynomial membership in an ideal. This is explained in the following 
-Theorem.
-
-\begin{Theorem}\label{the:membership}
-	{\bf Ideal Membership Test}
- Let $G = \{g_1,\cdots,g_t \}$ be a Gr\"obner basis for an ideal $I \subset \mathbb{K}[x_1,\cdots,x_d ]$
-	and let $f \in \mathbb{K}[x_{1},\dots, x_{d}]$. Then $f \in I$ if and only if the remainder on division of $f$ by
-	$G$ is zero.
-\end{Theorem}
-In other words, 
-\begin{equation}
-f \in I \iff f \stackrel{G}{\textstyle\longrightarrow}_+0
-\end{equation}
-
-\begin{Example}
-Consider Example \ref{exp:gbsimple}. Let $f = y^2x - x$ be another
-polynomial. Note that $f = yf_1 + f_2$, so $f \in I$. If we divide $f$
-by $f_1$ first and then by $f_2$, we will obtain a zero
-remainder. However, since the set $\{f_1, f_2\}$ is not a Gr\"{o}bner
-basis, we find that the reduction $f
-\stackrel{f_2}{\textstyle\longrightarrow} x^2 - x
-\stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
-i.e. dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
-zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
-$I$, $G = \{x^2 - x, yx - y, y^2 - x\}$, dividing $f$ by polynomials
-in $G$ in any order will always lead to the zero remainder. Therefore,
-one can decide ideal membership unequivocally using the Gr\"{o}bner
-basis. 
-\end{Example}
 
@@
 
+% TODO discuss complexity
+
+The high computational complexity if Gr\"obner basis computation is treated as an issue
+because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
+the worst case computation time/space cost of its Gr\"obner basis is doubly exponential bounded -- $q^{q^{O(|\phi|)}}$
+\cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
+the polynomial is well restricted rather than arbitrary.
+Gao et al. \cite{gao:qe-gf-gb} proves that when the number of variables and degree of polynomials 
+are restricted, the complexity reduces to single exponential $q^O(|\phi|)$.
+This provides the possibility to make use of Gr\"obner basis under restricted situations,
+which is the theoretic basis of this dissertation.
+
 \Grobner basis computation depends on the $Spoly$ computation, which in turn 
@@
 
-\section{Elimination Theory}
-
-Elimination theory uses {\bf elimination ordering} 
-to systematically eliminate variables from a system of polynomial
-equations.
-
-\begin{Definition}
-Let $I$ be an ideal in  $\K[x_1,\dots,x_k]$. The $i$-th 
-{\bf elimination ideal} $I_i$ is the ideal of $\K[x_{i+1},\dots,x_k]$ defined
-by
-\begin{equation}
-I_k = I \cap \K[x_{i+1},\dots,x_k]
-\end{equation}
-\end{Definition}
-
-The elimination ideal $I_i$ has eliminated all the variables 
-$x_1,\dots,x_i$, i.e. it only contains polynomials with variables in
-$x_{i+1},\dots,x_k$. 
-We can generate elimination ideals by computing
-\Grobner bases using elimination orderings.
-
-\begin{Theorem}
-$\left[\bf{Elimination\  Theorem}\right]$
-Let $I$ be an ideal in $\K[x_1,\dots,x_k]$ and let $G$ be the \Grobner 
-basis of $I$ with respect to the lex order (elimination order) 
-$x_1>x_2>\dots>x_k$. Then, for every $0\leq i\leq k$,
-\begin{equation}
-G_k=G\cap \K[x_{i+1},\dots,x_k]
-\end{equation}
-is a \Grobner basis of the $i$-th elimination ideal $I_i$.
-\label{thm:elimth}
-\end{Theorem}
-
-This can be better visualized using the following example.
-
-\begin{Example}
-Given the following equations in $\R[x,y,z]$
-\begin{eqnarray}
-x^2+y+z&=1 \nonumber \\
-x+y^2+z&=1 \nonumber \\
-x+y+z^2&=1 \nonumber
-\end{eqnarray}
-let $I$ be the ideal generated by these equations:
-\begin{equation}
-I=\langle x^2+y+z-1, x+y^2+z-1, x+y+z^2-1\rangle \nonumber
-\end{equation}
-The \Grobner basis for $I$ with respect to lex order $x>y>z$ is 
-found to be $G=\{g_1,g_2,g_3,g_4\}$ where
-\begin{eqnarray}
-g_1&=&x+y+z^2-1 \nonumber \\
-g_2&=&y^2-y-z^2+z \nonumber \\
-g_3&=&2yz^2+z^4-z^2 \nonumber \\
-g_4&=&z^6-4z^4+4z^3-z^2 \nonumber
-\end{eqnarray}
-
-Notice that while $g_1$ has variables in $\R[x,y,z]$, $g_2$ and $g_3$ only 
-have variables in $\R[y,z]$ and $g_4$ only has variables in $\R[z]$. Thus, 
-$G_1=G\cap \R[y,z]=\{g_2,g_3,g_4\}$ and $G_2=G\cap \R[z]=\{g_4\}$
-
-Also notice that since $g_4$ only contains variable $z$, and since $g_4=0$, 
-a solution for $z$ can be obtained. This solution can then be applied to 
-$g_2$ and $g_3$ to obtain solutions for $y$, and so on.
-\end{Example}
-
-Elimination theory provides the basis for our abstraction approach.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Hilbert's Nullstellensatz}
@@
 \begin{Definition}\label{def:acf}
-A field $\overline {\K}$ is an {\bf algebraically closed} field if every  
+A field $\overline {\F}$ is an {\bf algebraically closed} field if every  
 polynomial in one variable with degree at least $1$, with coefficients 
-in $\overline {\K}$, has a root in $\overline {\K}$. 
+in $\overline {\F}$, has a root in $\overline {\F}$. 
 \end{Definition}
 In other words, any non-constant polynomial equation over 
-$\overline {\K}\left[x\right]$ always has at least one root 
-in $\overline {\K}$. Every field $\K$ is contained in an algebraically 
-closed one $\overline {\K}$. 
+$\overline {\F}\left[x\right]$ always has at least one root 
+in $\overline {\F}$. Every field $\F$ is contained in an algebraically 
+closed one $\overline {\F}$. 
 For example, the field of real numbers $\mathbb{R}$ is not an algebraically closed 
@@
 
-%%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
-%\begin{Theorem}
-%$\left[\bf{Weak\  Nullstellensatz}\right]$ Let 
-%$I \subset \overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$ 
-%be an ideal satisfying $V(I)=\emptyset$. 
-%Then $I=\overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
-%\begin{equation}
-%V(I)=\emptyset\ \iff\ I=\overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
-%\end{equation}
-%\end{Theorem}
-%
-%\begin{Corollary}
-%	Let $I=\langle f_{1},\dots,f_{s} \rangle \subset \overline {\mathbb{K}}[x_1, x_2, \cdots, x_d]$. 
-%	Let $G$ be the reduced Gr\"obner basis of $I$. Then $V(I)=0 \iff G=\{1\}$.
-%\end{Corollary}
-
-%{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
-%system of multivariate polynomial equations (ideal $I$) has common solutions 
-%in ${\overline {\mathbb{K}}}^d$. For this purpose, we only need to check if 
-%the ideal is generated by the unit element, i.e., $1\in I$. 
-%This approach can be used to evaluate the feasibility of constraints in 
-%verification problems.
+%%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
+\begin{Theorem}{Weak Nullstellensatz}
+%$\left[\bf{Weak\  Nullstellensatz}\right]$ 
+Let $J \subset \overline {\F}[x_1, x_2, \cdots, x_d]$ 
+be an ideal satisfying $V(J)=\emptyset$. 
+Then $J=\overline {\F}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
+\begin{equation}
+V(J)=\emptyset\ \iff\ J=\overline {\F}[x_1, x_2, \cdots, x_d]=\langle 1 \rangle 
+\end{equation}
+\end{Theorem}
+
+\begin{Corollary}
+	Let $J=\langle f_{1},\dots,f_{s} \rangle \subset \overline {\F}[x_1, x_2, \cdots, x_d]$. 
+	Let $G$ be the reduced Gr\"obner basis of $J$. Then $V(J)=0 \iff G=\{1\}$.
+\end{Corollary}
+
+{\bf Weak Nullstellensatz} offers a way to evaluate whether or not the 
+system of multivariate polynomial equations (ideal $J$) has common solutions 
+in ${\overline {\F}}^d$. For this purpose, we only need to check if 
+the ideal is generated by the unit element, i.e., $1\in J$. 
+This approach can be used to evaluate the feasibility of constraints in 
+verification problems.
 %%%%%%%%%%%%%%%%%%%%%%%%%strong Nullstellensatz%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 ({\it The Strong Nullstellensatz} \cite{gb_book}) 
-Let $\overline{\mathbb{K}}$ be an algebraically closed field, and let $J$
-be an ideal in $\overline{\mathbb{K}}[x_1,\dots, x_d]$. 
-Then we have $I(V_{\overline{\mathbb{K}}}(J)) =\sqrt{J}$. 
+Let $\overline{\F}$ be an algebraically closed field, and let $J$
+be an ideal in $\overline{\F}[x_1,\dots, x_d]$. 
+Then we have $I(V_{\overline{\F}}(J)) =\sqrt{J}$. 
 \end{Theorem}
@@
 
-%% 
-%% Using this result, Weak Nullstellensatz can be 
-%% modified to be applicable over finite fields $\Fq$.
-%% %%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
-%% \begin{Theorem}\label{wnull:ff}
-%% $[\bf{Weak~Nullstellensatz~in~\Fq}]$\cite{null:1890}\\
-%% Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
-%% Let $J=\langle f_1,f_2,\cdots,f_s\rangle \subset \Fq[x_1,
-%% x_2, \cdots, x_d]$ be an ideal. Let $J_0 = \langle 
-%% x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
-%% of vanishing polynomials in $\Fq$. Then
-%% $V_{\Fq}(J) = V_{\overline {\Fq}}(J +
-%% J_0)=\emptyset$,  if and only if the reduced
-%% Gr\"obnerBasis$(J+J_{0})=\{1\}$. 
-%% \end{Theorem}
-
-%% The proof is given in \cite{null:1890}. Here, we provide a proof outline.
-
-%% \begin{Proof}
-%% The variety of $J$ over $\Fq[x_1,x_2,\cdots,x_d]$ 
-%% is equivalent to the variety over the algebraic closure of $\Fq$ 
-%% intersected by the entire field $\Fq$. That is, $V_{\Fq}(J)=V_{\overline 
-%% {\Fq}}(J) \cap \Fq$.
-
-%% Let $J_0 = \langle 
-%% x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
-%% generated by all vanishing polynomials in $\Fq[x_1,x_2,\cdots,x_d]$.
-%% Then $V_{\overline{\Fq}}(J_0)=\Fq$.
-
-%% Thus, $V_{\Fq}(J) = V_{\overline{\Fq}}(J)\cap V_{\overline{\Fq}}(J_0)
-%% = V_{\overline{\Fq}}(J+J_0)$.
-%% \end{Proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Using this result, Weak Nullstellensatz can be 
+modified to be applicable over finite fields $\Fq$.
+%%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
+\begin{Theorem}\label{wnull:ff}
+$[\bf{Weak~Nullstellensatz~in~\Fq}]$\cite{null:1890}\\
+Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
+Let $J=\langle f_1,f_2,\cdots,f_s\rangle \subset \Fq[x_1,
+x_2, \cdots, x_d]$ be an ideal. Let $J_0 = \langle 
+x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
+of vanishing polynomials in $\Fq$. Then
+$V_{\Fq}(J) = V_{\overline {\Fq}}(J +
+J_0)=\emptyset$,  if and only if the reduced
+Gr\"obnerBasis$(J+J_{0})=\{1\}$. 
+\end{Theorem}
+
+The proof is given in \cite{null:1890}. Here, we provide a proof outline.
+
+\begin{Proof}
+The variety of $J$ over $\Fq[x_1,x_2,\cdots,x_d]$ 
+is equivalent to the variety over the algebraic closure of $\Fq$ 
+intersected by the entire field $\Fq$. That is, $V_{\Fq}(J)=V_{\overline 
+{\Fq}}(J) \cap \Fq$.
+
+Let $J_0 = \langle 
+x_1^{2^k}-x_1,x_2^{2^k}-x_2,\cdots,x_d^{2^k}-x_d \rangle$ be the ideal
+generated by all vanishing polynomials in $\Fq[x_1,x_2,\cdots,x_d]$.
+Then $V_{\overline{\Fq}}(J_0)=\Fq$.
+
+Thus, $V_{\Fq}(J) = V_{\overline{\Fq}}(J)\cap V_{\overline{\Fq}}(J_0)
+= V_{\overline{\Fq}}(J+J_0)$.
+\end{Proof}
+
+\section{Elimination and Abstraction Theory}
+\label{sec:abstraction}
+Elimination of certain variables in a system of polynomials is a common 
+operation when some variables is not needed in modeling and analysis. 
+In this section, eliminating variables targets a tight-bound over-approximation 
+which is equivalent to existential quantifier elimination in first-order logic,
+or variable smoothing in Boolean operations. We introduce a eliminating method 
+based on algebraic geometry concepts, and use it as the fundamentals of 
+abstraction of a circuit.
+
+\subsection{Elimination Theory}
+Assume we are given a set of polynomials $f_1,\dots,f_s$ belongs to ring $\Fq[x_1,\dots,x_l,\dots,x_d]$.
+First, we show that eliminating the $x_1,\dots,x_l$ variables is equivalent to 
+{\it projecting} the variety $V(\langle f_1,\dots,f_s\rangle)$ from 
+$\Fq^d$ to $\Fq^{d-l}$. Figure \ref{fig:projection} is an example of projection in space of varieties
+from $\Fq^3$ to $\Fq^2$, corresponding to eliminating variable $x_1$ from 
+a system of polynomials belonging to $\Fq[x_1,x_2,x_3]$.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/projection.pdf}
+\caption{An example of projection from $\Fq^3$ to $\Fq^2$}
+\label{fig:projection}}
+\end{figure}
+
+Formally, we define concepts of projection and elimination ideal:
+\begin{Definition}
+The {\bf $l$-th projection} mapping is defined as:
+$$\pi_l:\Fq^d\to\Fq^{d-l},~~\pi_l((x_1,\dots,x_d)) = (x_{l+1}.\dots,x_d)$$
+where $l<d$. For any set $A \subseteq \Fq^d$, we write 
+$$\pi_l(A) = \{\pi_i(x):x\in A\}\subseteq \Fq^{d-l}$$
+\end{Definition}
+
+\begin{Definition}
+Let $J$ be an ideal in  $\Fq[x_1,\dots,x_d]$. The $l$-th 
+{\bf elimination ideal} $J_l$ is the ideal of $\Fq[x_{l+1},\dots,x_d]$ defined
+by
+\begin{equation}
+J_l = J \cap \Fq[x_{l+1},\dots,x_d]
+\end{equation}
+\end{Definition}
+
+The elimination ideal $J_l$ has eliminated all the variables 
+$x_1,\dots,x_l$, i.e. it only contains polynomials with variables in
+$x_{l+1},\dots,x_d$. 
+This shows that in finite fields, projection of the variety $V_{d}(\langle f_1,\dots,f_s\rangle)$
+from $\Fq^d$ to $\Fq^{d-l}$, is exactly the variety $V_{d-l}(\langle f_1,\dots,f_s\rangle\cap \Fq[x_{l+1},\dots,x_d])$.
+
+Elimination theory uses {\bf elimination ordering} 
+to systematically eliminate variables from a system of polynomial
+equations.
+We can generate elimination ideals by computing
+\Grobner bases using elimination orderings.
+
+\begin{Theorem}{Elimination Theorem}
+%$\left[\bf{Elimination\  Theorem}\right]$
+Let $J$ be an ideal in $\F[x_1,\dots,x_d]$ and let $G$ be the \Grobner 
+basis of $J$ with respect to the lex order (elimination order) 
+$x_1>x_2>\dots>x_d$. Then, for every $0\leq l\leq d$,
+\begin{equation}
+G_l=G\cap \F[x_{l+1},\dots,x_d]
+\end{equation}
+is a \Grobner basis of the $l$-th elimination ideal $J_l$.
+\label{thm:elimth}
+\end{Theorem}
+
+This can be better understood using the following example.
+
+\begin{Example}
+Given the following equations in $\R[x,y,z]$
+\begin{eqnarray}
+x^2+y+z&=1 \nonumber \\
+x+y^2+z&=1 \nonumber \\
+x+y+z^2&=1 \nonumber
+\end{eqnarray}
+let $I$ be the ideal generated by these equations:
+\begin{equation}
+I=\langle x^2+y+z-1, x+y^2+z-1, x+y+z^2-1\rangle \nonumber
+\end{equation}
+The \Grobner basis for $I$ with respect to lex order $x>y>z$ is 
+found to be $G=\{g_1,g_2,g_3,g_4\}$ where
+\begin{eqnarray}
+g_1&=&x+y+z^2-1 \nonumber \\
+g_2&=&y^2-y-z^2+z \nonumber \\
+g_3&=&2yz^2+z^4-z^2 \nonumber \\
+g_4&=&z^6-4z^4+4z^3-z^2 \nonumber
+\end{eqnarray}
+
+Notice that while $g_1$ has variables in $\R[x,y,z]$, $g_2$ and $g_3$ only 
+have variables in $\R[y,z]$ and $g_4$ only has variables in $\R[z]$. Thus, 
+$G_1=G\cap \R[y,z]=\{g_2,g_3,g_4\}$ and $G_2=G\cap \R[z]=\{g_4\}$
+
+Also notice that since $g_4$ only contains variable $z$, and since $g_4=0$, 
+a solution for $z$ can be obtained. This solution can then be applied to 
+$g_2$ and $g_3$ to obtain solutions for $y$, and so on.
+\end{Example}
+
+Elimination theory provides the basis for following abstraction approach.
+\subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/abs_setup.pdf}
+\caption{Galois field arithmetic circuit model for abstraction}
+\label{fig:abs_setup}}
+\end{figure}
+
+{\bf Problem setup:} Let $S$ be the system of polynomials, 
+$\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
+derived from the hardware
+implementation of the Galois field arithmetic circuit over $\Fkk$, as figure \ref{fig:abs_setup} shows.
+This circuit performs some unknown function $f$ over 
+$\Fkk$ in the form of $Z=\Func(A_1,\dots,A_n)$, where $Z$ is the $k$-bit 
+output and $A_1,\dots,A_n$ are the $k$-bit inputs.
+The polynomial representation of $\Func$ over $\Fkk$ is thus:
+
+\begin{equation}
+f_{\Func}: Z+\Func(A_1,\dots,A_n) \nonumber
+\end{equation}
+
+Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
+it agrees with the solution to the system of polynomials $\{S\}=0$, i.e.:
+\begin{equation}
+f_1=\dots=f_s=f_{A_1}=\dots=f_{A_n}=f_{Z}=0 \nonumber
+\end{equation}
+Thus, if we let $J=\langle f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\rangle$ 
+be the ideal generated by $S$, 
+$f_{\Func}$ {\bf vanishes} on the variety $V_{\Fkk}(J)$. 
+Therefore, due to Proposition \ref{pro:iofv}, 
+$f_{\Func}$ must be contained in the ideal of polynomials that vanish on
+this variety, $f_{\Func} \in I(V_{\Fkk}(J))$.
+
+By applying Strong Nullstellensatz over $\Fkk$ (Theorem \ref{thm:sns}), 
+$I(V_{\Fkk}(J))=J+J_0$ where
+$J_0$ is the ideal generated by all vanishing polynomials in $\Fkk$. 
+Recall that a vanishing polynomial in $\Fkk[x]$ is $x^q-x=x^q+x$. 
+In our case, 
+$\{x_1,\dots,x_d\} \in \F_2$ and $\{A_1,\dots,A_n,Z\} \in \Fkk$.
+Thus, for $\Fkk[x_1,\dots,x_d,A_1,\dots,A_n,Z]$: 
+\begin{equation}
+J_0 = \langle x_1^2+x_1,\dots,x_d^2+x_d, A_1^{2^k}+A_1,\dots,A_n^{2^k}+A_n,Z^{2^k}+Z\rangle \nonumber
+\end{equation}
+
+The generators of the ideal sum $J+J_0$ are simply the combination of the 
+generators of $J$ and the generators $J_0$.
+
+The variety $V_{\Fq}(J)$ is
+the set of all consistent assignments to the nets (signals) in the
+circuit $C$. If we {\it project this variety on the word-level input and
+output variables of the circuit $C$, we essentially generate the
+function $\Func$ implemented by the circuit.} Projection of varieties from
+$d$-dimensional space $\Fq^d$ onto a lower dimensional subspace
+$\Fq^{d-l}$ is equivalent to {\it eliminating $l$ variables} from the
+corresponding ideal. This can be done by computing a \Grobner basis
+of the ideal with elimination ordering, as described in the 
+Elimination Theorem (Theorem \ref{thm:elimth}).
+Thus, we can find the polynomial 
+$f_\Func:Z+\Func(A_1,\dots,A_n)$ by computing the \Grobner
+basis of $J+J_0$ using the proper elimination ordering.
+
+The proposed elimination order for 
+abstraction is defined as the {\bf abstraction term order}. 
+\begin{Definition}
+\label{def:ato}
+Given a circuit $C$,
+let $x_1, \dots, x_d$ denote all the bit-level variables, 
+let $A_1,\dots,A_n$ denote the $k$-bit word-level inputs, 
+and let $Z$ denote the $k$-bit word-level output. 
+Using any refinement of the partial variable order 
+$\{x_1, \dots, x_d\} > Z > \{A_1, \dots, A_n\}$,
+% where any refinement of the order will do,
+impose a lex term order $>$ on the polynomial ring 
+$R = \Fq[x_1, \dots, x_d, Z, A_1, \dots, A_n]$. 
+This elimination term order $>$ is defined as
+the {\bf Abstraction Term Order (ATO)}. The relative ordering among $x_1,
+\dots, x_d$ is not important and can be chosen arbitrarily. Likewise,
+the relative ordering among $A_1, \dots, A_n$ is also unimportant.
+\end{Definition}
+
+\begin{Theorem}
+{\bf Abstraction Theorem:} Using the notations from problem setup at the beginning of this subsection,
+we compute a \Grobner basis $G$ of ideal $(J+J_0)$ using the abstraction term 
+order $>$. Then: \\
+(i) For every word-level input $A_i$, $G$ must contain the vanishing 
+polynomial $A_i^q - A_i$ as the only polynomial with $A_i$ as its only 
+variable;\\
+(ii) $G$ must contain a polynomial of the form 
+$Z + \mathcal{G}(A_1,\dots,A_n)$; and\\ 
+(iii) $Z + \mathcal{G}(A_1,\dots,A_n)$ is such that 
+$\Func(A_1,\dots,A_n) = {\mathcal{G}}(A_1,\dots,A_n),
+\forall A_1,\dots,A_n \in \Fq$. 
+In other words, ${\mathcal{G}}(A_1,\dots,A_n)$ and $\Func(A_1,\dots,A_n)$ are
+equal as polynomial functions over $\Fq$.
+\label{thm:abs}
+\end{Theorem}
+
+\begin{Corollary}
+By computing a {\bf reduced} \Grobner basis $G_r$ of $J + J_0$, 
+$G_r$ will contain one and only one polynomial in of the form 
+$Z + \G(A_1,\dots,A_n)$, such that $Z = \G(A_1,\dots,A_n)$ is the 
+{\bf unique, minimal, canonical} representation of the function 
+$\Func$ implemented by the circuit.  
+\end{Corollary}
+
+\begin{Example}
+\label{exp:finalMul2Bit}
+Consider a 2-bit multiplier over ${\mathbb{F}}_{2^2}$ with
+$P(x)=x^{2}+x+1$, given in Fig. \ref{fig:2bitmul}. Variables $a_0,
+a_1, b_0, b_1$ are primary inputs, $z_0, z_1$ are primary outputs, and
+$c_0, c_1, c_2, c_3, r_0$ are intermediate variables. 
+%The gate $\otimes$
+%corresponds to AND-gate, i.e. bit-level multiplication modulo 2. The
+%gate $\oplus$ corresponds to XOR-gate, i.e. addition modulo 2.
+
+\begin{figure}[H]
+\centerline{
+\includegraphics[scale=0.5]{./figures/2bitmasmult.eps}
+}
+\caption{ A 2-bit multiplier over ${\mathbb{F}}_{2^2}$.}
+\label{fig:2bitmul}
+\end{figure}
+
+Polynomials extracted from the circuit implementation represent the
+ideal $J$. Along with the ideal of vanishing polynomials $J_0$, 
+the following polynomials represent the generators 
+of $J+J_0$ for the multiplier circuit.
+
+\begin{eqnarray}
+ \left .  
+	\begin{aligned}
+		f_1:c_0+a_0 \cdot b_0  \\
+		f_2:c_1+a_0 \cdot b_1  \\
+		f_3:c_2+a_1 \cdot b_0  \\
+		f_4:c_3+a_1 \cdot b_1  \\
+		f_5:r_0+c_1 + c_2		\\
+		f_6:z_0+c_0 + c_3		\\
+		f_7:z_1+r_0 + c_3		
+	\end{aligned} 
+ \ \right\}
+ &\qquad&  {\it  \text{Bit-level circuit constraints} ~(\subset J)} \nonumber \\
+ \left . 
+	\begin{aligned}
+		f_{A}:A+a_0+a_1\cdot \alpha   \\ 
+		f_{B}:B+b_0+b_1\cdot \alpha  \\ 
+		f_{Z}:Z+z_0+z_1\cdot \alpha   
+	\end{aligned} 
+ \right\}
+ &\qquad&  {\it  \text{Word-level designation} ~(\subset J)} \nonumber \\
+  \left . 
+	\begin{aligned}
+		a_0^2-a_0, ~a_1^2-a_1,~b_0^2-b_0, ~b_1^2-b_1   \\ 
+		c_0^2-c_0, ~c_1^2-c_1,~c_2^2-c_2, ~c_3^2-c_3  \\ 
+		r_0^2-r_0, ~z_0^2-z_0,~z_1^2-z_1    \\ 
+		A^4-A, ~B^4-B ,~Z^4-Z		  
+	\end{aligned} 
+ \right\}
+ &\qquad&  {\it \text{vanishing polynomials} (J_0)} \nonumber
+\end{eqnarray}
+We apply abstraction term order $>$, i.e a lex order with
+"bit-level variables" $>$ "Output Z" $>$ "Inputs A, B".
+
+When we compute the reduced \Grobner basis, $G_r$, of \{$J + J_0$\} with 
+respect to this ordering, $G_r$ = \{$g_1, \dots, g_{14}\}:$
+\begin{align*}
+g_1: B^4+B; 
+~~g_2: b_0+b_1 \alpha + B; 
+~~g_3: a_0+a_1 \alpha + A;  \nonumber \\
+~~g_4: c_0+c_1 \alpha + c_2 \alpha + c_3(\alpha+1)+Z;
+g_5: r_0+c_1+c_2; 
+~~g_6: z_0+c_0+c_3; \nonumber \\
+~~g_7: z_1+r_0+c_3; 
+~~{\bf g_8: Z+A\cdot B};
+~~g_9: b_1+B^2+B; 
+~~g_{10}: a_1+A^2+A; \nonumber \\
+~~g_{11}: c_3+a_1\cdot b_1
+g_{12}: c_2+a_1\cdot b_1 \alpha + a_1 \cdot B; 
+~~g_{13}: c_1+a_1\cdot b_1 \alpha +b_1 A; 
+~~g_{14}: A^4+A  \nonumber
+\end{align*}
+
+$g_8=Z+A\cdot B$ is the {\bf canonical, word-level polynomial } 
+representing the function performed by the multiplier $Z=A\cdot B$.
+\end{Example}
+
 \section{Concluding Remarks}
-
 Our approach to word-level abstraction of Galois field arithmetic 
 circuits applies concepts of polynomial ideals, varieties, \Grobner basis, 
-and elimination theory to abstract a word-level representation of the 
-circuit. This approach is described in the next chapter. However, a \Grobner
-basis computation is prohibitively expensive; thus we propose improvements 
-to our original approach in a subsequent chapter.
-
-%Once we have a word-level representation, we can apply these representations
-%to perform equivalence checking of Galois field arithmetic circuits.
-%The verification problem is formulated using Weak
-%Nullstellensatz, as it applies over Galois fields, and subsequently solved 
-%using a \Grobner basis approach. This approach is described in a future 
-%chapter.
+and abstraction theory to implement verifications on sequential circuits. 
+These approaches are described in the following chapters.

### 2016-11-23 dissertation/normal.tex
dissertation v2_part reacha update 1

@@
 a sort of suitable circuit benchmarks which is easy to compute its
-Gr\"obner basis (GB). From the work of Lv et al. \cite{lv_dissertation},
+Gr\"obner basis (GB). From the work of {\it Lv} \cite{lv:phd},
 we learn that arithmetic circuits in Galois field (GF) is
@@
 ideal generators form a GB themselves when applying reverse topological
-term order. Furthermore, according to the work of Pruss et al.
-\cite{tim_dissertation}, with a limited computation complexity,
+term order. Furthermore, according to the work of {\it Pruss et al.}
+\cite{pruss:tcad15}, with a limited computation complexity,
 we can abstract the word-level signature of an arithmetic 
@@
 field circuits. In each frame, we can use the techniques 
-from \cite{tim_dissertation} to abstract the word-level
+from \cite{pruss:tcad15} to abstract the word-level
 signature of the combinational logic, which corresponds
@@
 \label{sec:normal_motiv}
-From the preliminaries (Chapter \ref{ch:prelim}) about FSMs, we learn that the
+From the preliminaries about FSMs in section \ref{sec:reach}, we learn that the
 Moore machine does not rely on inputs for state transitions. 

### 2016-11-23 dissertation/prelim_basis.tex
dissertation v2_part reacha update 1

@@
 mostly referred from the textbooks \cite{ideals:book} \cite{gb_book} and 
-previous work by {\it Lv} \cite{lv:phd} as well as {\it Pruss} \cite{pruss:tcad15}
+previous work by {\it Lv} \cite{lv:phd} as well as {\it Pruss} \cite{pruss:tcad15}.
 
@@
 have continuously eliminated variables in the ordering. This is the topic of 
-elimination theory, which is described in the following section.
+elimination theory, which is described in the following sections as well as 
+its theoretic basis -- the Nullstellensatz theory.
 

### 2016-11-26 dissertation/normal.tex
dissertation v2_part normal complete

@@
 \label{sec:normal_motiv}
-From the preliminaries about FSMs in section \ref{sec:reach}, we learn that the
+From the preliminaries about FSMs in Section \ref{sec:reach}, we learn that the
 Moore machine does not rely on inputs for state transitions. 
@@
 and store the results in registers after $k$ clock cycles. Its behavior can be described using
-STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, $R_init$ is the initial
+STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, $R_{init}$ is the initial
 state (usually reset to all zeros), $R_1$ to $R_{k-1}$ are intermediate results stored as SO of current state and SI
@@
 reusing a smaller combinational logic component such that the area cost is greatly optimized.
+An example of these designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
 However, it also brings difficulties in verifying the the circuit functions.
@@
 
-% \section{Normal Basis Multiplier over Galois Field}
-% From algebraic view, a field is a space, and field elements are dots in the space. Those elements can be 
-% represented with unique coordinates, which requires the pre-definition of a basis vector. In this
-% section, we discuss a special basis called normal basis, as well as the advantages adopting it in 
-% GF operations esp. multiplication.
-% \subsection{Normal Basis}
-% Given a Galois field (GF) $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
-% Its elements can be written in polynomials of $\alpha$, when there is an irreducible polynomial $p(\alpha)$
-% defined.
-% 
-% If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
-% to binary bit-vector representations by recording the coefficients. For example,
-% 
-% \begin{table}[H]
-% \centering
-% \caption{Bit-vector, Exponential and Polynomial representation of
-% elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
-% \pmod{x^4+x^3+1}$}
-% \begin{tabular}{|c|c||c|c|} 
-% \hline
-% $a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
-% \hline
-% $0000$        & $0$           & $1000$  &$\alpha^3$\\
-% \hline
-% $0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
-% \hline
-% $0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
-% \hline
-% $0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
-% \hline
-% $0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
-% \hline
-% $0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
-% \hline
-% $0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
-% \hline
-% $0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
-% \hline
-% \end{tabular}
-% \label{table:booltogalois}  
-% \end{table}
-% 
-% Basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$ is called {\bf standard basis} (StdB), which results in
-% a straightforward representation for elements, and operations of elements such as addition and subtraction.
-% The addition/subtraction of GF elements in StdB follows the rules of polynomial addition/subtraction
-% where coefficients belong to $\mathbb F_2$. In other words, using the definition of {\it exclusive or} (XOR) in
-% Boolean algebra, element $A$ add/subtract by element $B$ in StdB is defined as
-% \begin{align}\label{eqn:StdB}
-% A+B = A-B &= (a_0,a_1,\dots,a_{k-1})_{StdB} \xor (b_0,b_1,\dots,b_{k-1})_{StdB} \nonumber\\
-% &=(a_0\oplus b_0, a_1\oplus b_1,\dots,a_{k-1}\oplus b_{k-1})_{StdB} 
-% \end{align}
-% 
-% \subsection{Multiplication using Normal Basis}
-% Besides addition/subtraction, multiplication is also very common in arithmetic circuit design.
-% The multiplication of GF elements in $\Fkk$ in StdB follows the rule of polynomial multiplication.
-% However, it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
-% in bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
-% the area and delay of circuit will be costly.
-% 
-% In order to lower down the complexity of arithmetic circuit design, Massey and Omura \cite{MasseyOmura} % ref 7 in RH paper
-% use a new basis to represent GF elements, which is called {\bf normal basis} (NB).
-% A normal basis over $\Fkk$ is written in the form of
-% \begin{equation*}
-% N.B. ~~~ \N = \{\beta,\beta^2,\beta^4,\beta^8,\dots,\beta^{2^{k-1}}\}
-% \end{equation*}
-% Respectively, a field element in NB representation is actually
-% \begin{align*}
-% A &= (a_0,a_1,\dots,a_{k-1})_{NB} \\
-%   &= a_0\beta+a_1\beta^2+\cdots+a_{k-1}\beta^{2^{k-1}} \\
-%   &= \sum_{i=0}^{k-1} a_i\beta^{2^i}
-% \end{align*}
-% 
-% According to the definition, a normal basis is a vector where the next entry is the square of the former one.
-% We note that the vector is cyclic, i.e. $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
-% {\bf Normal element} $\beta$ is an element from the field which is used to construct the normal basis,
-% and can be represent as a power of primitive element $\alpha$: 
-% \begin{equation*}
-% \beta = \alpha^t, ~~~ 1\leq t<2^k
-% \end{equation*}
-% 
-% The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
-% However, what makes NB powerful is its property when doing multiplications and exponentiations.
-% The following lemmas and examples illustrate this fabulous property very well.
-% \begin{Lemma}[Square of NB]
-% \label{lem:squareNB}
-% In $\Fkk$, equation 
-% \begin{equation*}
-% (a+b)^2 = a^2 + b^2
-% \end{equation*}
-% has been proved. According to the \textbf{binomial theorem}, it can be extended to
-% \begin{align*}
-% &(b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{k-1}\beta^{2^{k-1}})^2 \\
-% &= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{k-1}^2\beta \\
-% &= b_{k-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{k-2}\beta^{2^{k-1}}
-% \end{align*}
-% \end{Lemma}
-% This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
-% Obviously, StdB representation does not have this benefit.
-% 
-% \begin{Example}[Square of NB]
-% In GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $x^3 + x + 1$, the standard basis is denoted as 
-% $\{ 1, \alpha, \alpha^2\}$ where $\alpha^3+\alpha+1=0$.
-% Let $\beta = \alpha^3$, then $\N = \{ \beta, \beta^2, \beta^4\}$ forms a normal basis. 
-% Write down element $E$ using both representations:
-% \begin{align*}
-% E &= (a_0,a_1,a_2)_{StdB} = (b_0,b_1,b_2)_{NB} \\
-%   &= a_0 + a_1\alpha + a_2\alpha^2 = b_0\beta + b_1\beta^2 + b_2\beta^4
-% \end{align*}
-% Compute the square of $E$ in StdB first:
-% \begin{align*}
-% E^2 &= a_0 + a_1\alpha^2 + a_2\alpha^4 \\
-%     &= a_0 + a_2\alpha + (a_1 + a_2)\alpha^2 \\
-%     &= (a_0,a_2,a_1+a_2)_{StdB}
-% \end{align*}
-% When it is computed in NB, we can make it very simple:
-% \begin{align*}
-% E^2 &= \overset{\xrightarrow{Cyclic~~shift}}{(b_0,b_1,b_2)}_{NB} \\
-% 	&= (b_2,b_0,b_1)_{NB}
-% \end{align*}
-% \end{Example}
-% 
-% This example shows that convenience to use NB when computing $2^k$ power of an element.
-% Multiplication is a bit complicated than squaring; but when it is decomposed as bit-wise
-% operations, the property in lemma \ref{lem:squareNB} can be well utilized.
-% 
-% \begin{Example}[Bit-wise NB multiplication]
-% Assume there are 2 binary vectors representing 2 operands in NB over $\Fkk$: 
-% $A = (a_0, a_1, \dots, a_{k-1}), B = (b_0, b_1, \dots, b_{k-1})$. Note that in this example, 
-% by default we use normal basis representation so subscript ``NB" is skipped. Their product can also be written
-% as: $$C = A\times B = (c_0, c_1, \dots, c_{k-1})$$
-% 
-% Assume the most significant bit (MSB) of the product can be represented by a function $f_{mult}$: 
-% \begin{equation}
-% \label{eqn:multiMSB}
-% c_{n-1} = f_{mult}(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})
-% \end{equation}
-% Before discussing the details of the function $f_{mult}$, we can take 
-% a square on both side of equation \ref{eqn:multiMSB}, i.e. $C^2 = A^2\times B^2$.
-% Obviously, using the property in lemma \ref{lem:squareNB}, the original second most significant bit 
-% becomes the new MSB because of right-cyclic shifting. 
-% Concretely, 
-% $$(c_{k-1},c_0,c_1,\dots,c_{k-2}) = (a_{k-1},a_0,a_1,\dots,a_{k-2})\times(b_{k-1},b_0,b_1,\dots,b_{k-2})$$
-% Note $A^2, B^2$ and $C^2$ still belong to $\Fkk$, thus as a universal function implementing MSB multiplication
-% over $\Fkk$, $f_{mult}$ still keeps the same. As a result, the new MSB can be written as 
-% \begin{equation}
-% \label{eqn:shiftMSB}
-% c_{k-2} = f_{mult}(a_{k-1}, a_0, a_1, 
-% \dots, a_{k-2}; b_{k-1}, b_0, b_1, \dots, b_{k-2})
-% \end{equation}
-% Similarly, if we take a square again on the new equation, we can get $c_{k-3}$.
-% Successively we can derive all bits of product $C$ using the same function $f_{mult}$, and the only
-% adjustment we need to make is to right-cyclic shift 2 operands by 1 bit each time.
-% \end{Example}
-% 
-% From above example, it is proved that a universal structure that implements $f_{mult}$ can be reused
-% for $k$ times in NB multiplication over $\Fkk$. Comparing to StdB, which requires distinct design 
-% for every bit of multiplication, NB is less costly if we can prove $f_{mult}$ will not result in 
-% a structure with $O(k^2)$ complexity. So our next mission is to explore the details of $f_{mult}$
-% to prove it will be a relatively simple design with complexity lower than $O(k^2)$.
-% 
-% If we want to make the complexity of $f_{mult}$ lower than $O(k^2)$, then the best choice is to try out 
-% linear functions. As we know, matrix multiplication can simulate all possible combinations of 
-% linear functions (which is also the reason it is used as basic model to simulate the behavior of a neuron
-% in neural network machine learning algorithms). Imagine $A$ is a $k$-bit row vector and $B$ is a $k$-bit 
-% column vector, then the single bit product can be written as the product of matrix multiplication
-% $$c_{l} = A\times C\times B$$
-% where $C$ is a $k\times k$ square matrix.
-% 
-% \begin{Definition}[$\lambda$-Matrix]
-% A binary $k\times k$ matrix $M$ is used to describe the bit-wise normal basis multiplication function $f_{mult}$ where
-% \begin{equation}
-% \label{eqn:def_lambda}
-% c_{l} = f_{mult}(A, B) = A \times M \times B^T
-% \end{equation}
-% $B^T$ denotes vector transposition. Matrix $M$ is called $\lambda$-Matrix of
-% $k$-bit NB multiplication over $\Fkk$.
-% \end{Definition} 
-% 
-% When taking different bits $l$ of the product in equation \ref{eqn:def_lambda}, 
-% we obtain a series of conjugate matrices of $M$. 
-% Which means instead of shifting operands $A$ and $B$, we can also shift the 
-% matrix.
-% 
-% More specifically, we denote the matrix by \emph{$l$-th $\lambda$-Matrix} as 
-% $$c_l = A \times M^{(l)} \cdot B^T$$
-% Meanwhile, the operator shifting rule in equation \ref{eqn:shiftMSB} still holds. Then we have relation 
-% $$c_{l-1} = A \cdot M^{(l-1)} \cdot B^T = shift(A) \cdot M^{(l)} \cdot shift(B)^T$$ which means
-% by right and down cyclically shifting $M^{(l-1)}$, we can get $M^{(l)}$.
-% 
-% \begin{Example}[NB multiplication using $\lambda$-Matrix]
-% Over GF $\mathbb F_{2^3}$ constructed by irreducible polynomial $\alpha^3 + \alpha + 1$, let normal element $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
-% forms a normal basis. Corresponding $0$-th $\lambda$-Matrix is
-% \begin{equation*}
-% M^{(0)} = \left(
-% \begin{array} {lcr}
-% 0 & 1 & 0\\
-% 1 & 0 & 1\\
-% 0 & 1 & 1
-% \end{array} \right).
-% \end{equation*}
-% i.e.,
-% \begin{equation*}
-% c_0 = (a_0\  a_1\  a_2)\left(
-% \begin{array} {lcr}
-% 0 & 1 & 0\\
-% 1 & 0 & 1\\
-% 0 & 1 & 1
-% \end{array} \right)\left(
-% \begin{array} {lcr}
-% b_0\\
-% b_1\\
-% b_2
-% \end{array} \right)
-% \end{equation*}
-% From $0$-th $\lambda$-Matrix we can directly write down all remaining $\lambda$-Matrices:
-% \begin{equation*}
-% M^{(1)} = \left(
-% \begin{array} {lcr}
-% 1 & 0 & 1\\
-% 0 & 0 & 1\\
-% 1 & 1 & 0
-% \end{array} \right)~~~~~~
-% M^{(2)} = \left(
-% \begin{array} {lcr}
-% 0 & 1 & 1\\
-% 1 & 1 & 0\\
-% 1 & 0 & 0
-% \end{array} \right)
-% \end{equation*}
-% \end{Example}
-% 
-% If we generalize the definition and explore the nature of $\lambda$-Matrix, it is defined as cross-product terms from multiplication, which is 
-% \begin{equation}
-% Product~vector~C = (\sum_{i=0}^{k-1}a_i\beta^{2^i})(\sum_{j=0}^{k-1}b_j\beta^{2^j}) = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}a_ib_j\beta^{2^i}\beta^{2^j}
-% \end{equation}
-% The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
-% NB, i.e.
-% \begin{equation}
-% \beta^{2^i}\beta^{2^j} = \sum_{l=0}^{k-1}\lambda_{ij}^{(l)}\beta^{2^l}, \ \ \lambda_{ij}^{(l)} \in \mathbb F_2.
-% \end{equation}
-% Substitution yields, result is an expression for l-th digit of product as showed in equation \ref{eqn:multiMSB}:
-% \begin{equation}
-% c_l = \sum_{i=0}^{k-1}\sum_{j=0}^{k-1}\lambda_{ij}^{(l)}a_ib_j
-% \end{equation}
-% $\lambda_{ij}^{(l)}$ is the entry with coordinate $(i,j)$ in $l$-th $\lambda$-Matrix.
-% 
-% The $\lambda$-Matrix can be implemented with XOR and AND gates in circuit design.
-% The very naive implementation requires $O(C_N)$ gates, where $C_N$ is the number of
-% nonzero entries in $\lambda$-Matrix.
-% There usually exists multiple NBs in $\Fkk, k>3$. If we employ a random NB, there is no mathematical
-% guarantee that $C_N \sim o(k)$ (symbol $o$ denotes ``strictly lower than bound"). However,
-% Mullin et al. \cite{mullinONB} % This citation is valid
-% proves that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
-% its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
-% called optimal normal basis (ONB), details are introduced in appendix \ref{append:ONB}.
-% 
-% In practice, large size NB multipliers are usually designed in $\Fkk$ when ONB exists
-% to minimized the number of gates. So in the following part of this chapter and our experiments,
-% we only focus on ONB multipliers instead of general NB multipliers.
-% 
-% 
-% % After fixing this, add a whole-piece StdB vs NB example, list their cost as the conclusion
-% \subsection{Comparison between Standard Basis and Normal Basis}
-% At the end of this section, a detailed example is used to make a comparison between StdB multiplication
-% and NB multiplication.
-% \begin{Example}[Rijndael's finite field]
-% Rijndael uses a characteristic 2 finite field with 256 elements, which can also be called the GF $\mathbb F_{2^8}$.
-% Let us define the primitive element $\alpha$ using irreducible polynomial 
-% $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$. Coincidently, $\alpha$ is also a normal element,
-% i.e. $\beta = \alpha$ can construct a NB $\{\alpha,\alpha^2,\alpha^4,\alpha^8,\alpha^{16},\alpha^{32},\alpha^{64},
-% \alpha^{128}\}$.
-% 
-% We pick a pair of elements from the Rijndael's field: $A=(0100~1011)_{StdB} = (4B)_{StdB},~B=(1100~1010)_{StdB} = (CA)_{StdB}$.
-% First let us compute their product in StdB, the rule follows ordinary polynomial multiplication.
-% 
-% \begin{align*}
-% A\cdot B &= (\alpha^6+\alpha^3+\alpha+1)(\alpha^7+\alpha^6+\alpha^3+\alpha)\\
-% &= (\alpha^{13}+\alpha^{10}+\alpha^8+\alpha^7)+(\alpha^{12}+\alpha^9+\alpha^7+\alpha^6)+
-% 	(\alpha^9+\alpha^6+\alpha^4+\alpha^3)\\
-% 	&~~~+(\alpha^7+\alpha^4+\alpha^2+\alpha) \\
-% &= \alpha^{13}+\alpha^{12}+\alpha^{10}+\alpha^8+\alpha^7+\alpha^3+\alpha^2+\alpha
-% \end{align*}
-% Note that this polynomial is not the final form of the product because it needs to be
-% reduced modulo irreducible polynomial $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$.
-% This can be done using base-2 long division. Note the dividend and divisor are written in pseudo Boolean
-% vectors, not real Boolean vectors in any kind of bases.
-% 
-% % The folowing part is customized latex code mimicing binary long division
-% % Note \divrule coordinates may not reflect the intuitive column # in tabular. Adjust by urself!!!
-% 
-% \vspace{1cm}
-% 
-% \newdimen\digitwidth
-% \settowidth\digitwidth{0}
-% \def~{\hspace{\digitwidth}}
-% 
-% 
-% \def\divrule#1#2{%
-% \noalign{\moveright#1\digitwidth%
-% \vbox{\hrule width#2\digitwidth}}}
-% 111010111\,\begin{tabular}[b]{@{}r@{}}
-% 101001 \\ \hline
-% \big)\begin{tabular}[t]{@{}l@{}}
-% 11010110001110 \\
-% 111010111 \\ \divrule{0}{14}
-% ~~111101101 \\
-% ~~111010111 \\ \divrule{2}{12}
-% ~~~~~111010110 \\
-% ~~~~~111010111 \\ \divrule{5}{9}
-% ~~~~~~~~~~~~~1
-% \end{tabular}
-% \end{tabular}
-% \vspace{0.5cm}
-% 
-% The final remainder is $1$, i.e. the product equals to 1 in StdB.
-% 
-% On the other hand, operands $A$ and $B$ can be written in NB as
-% $$A = (0010~1001)_{NB},~~B = (0100~0010)_{NB}$$
-% The $\lambda$-Matrix for $\mathbb F_{2}[x] \pmod{x^8+x^7+x^6+x^4+x^2+x+1}$
-% is
-% (Computation of $\lambda$-Matrix refers to appendix \ref{append:NB})
-% \begin{equation*}
-% M^{(0)} = \left(\begin{array}{lccccccr}
-% 0 &0 &0 &0 &1 &0 &1 &1 \\
-% 0 &0 &1 &1 &1 &1 &0 &0 \\
-% 0 &1 &0 &0 &0 &0 &1 &0 \\
-% 0 &1 &0 &0 &1 &1 &0 &1 \\
-% 1 &1 &0 &1 &0 &1 &0 &0 \\
-% 0 &1 &0 &1 &1 &0 &0 &1 \\
-% 1 &0 &1 &0 &0 &0 &0 &0 \\
-% 1 &0 &0 &1 &0 &1 &0 &1
-% \end{array}\right)
-% \end{equation*}
-% Taking matrix multiplication $c_0 = A\times M^{(0)}\times B^T$,
-% the result is $c_0 = 1$. Then by cyclic shifting $A$ and $B$
-% (or shifting $M^{(0)}$, either is applicable), we can successively
-% obtain other bits of product. The final answer is
-% $$C = (0000~0001)_{NB}$$
-% It is equivalent to the result in StdB.
-% \end{Example}
-% 
-% From the intuition of humans, StdB multiplication is straightforward and easier to understand
-% while NB is difficult to comprehend. However, if we implement both multiplications to 
-% hardware multipliers, it will be clear which side a circuit designer prefers.
-% 
-% % May consider providing details?
-% 
-% Mastrovito multiplier and Montgomery multiplier are 2 common designs
-% of GF multipliers using StdB. As a naive implementation of GF multiplication,
-% Mastrovito multiplier uses most number of gates:
-% $k^2$ AND gates plus $k^2-\Delta$ XOR gates \cite{Mastrovito}. Montgomery multiplier 
-% applies lazy reduction techniques and results in a better latency performance, while the number of gates are about
-% the same with Mastrovito multiplier:
-% $k^2$ AND gates plus $k^2-k/2$ XOR gates \cite{Montgomery}. 
-% Concretely, typical design of Mastrovito multiplier consists of 218 logic gates, while 
-% Montgomery multiplier needs 198 gates. However, the NB multiplier reuses the $\lambda$-Matrix 
-% logic, so this component will only need to be implemented for once. 
-% Consider the definition of matrix multiplication, it needs $C_N$ AND gates to apply 
-% bit-wise multiplication and $C_N-1$ XOR gates to sum the intermediate products up. The number of nonzero entries
-% in the $\lambda$-Matrix can be counted: $C_N = 27$.
-% As a result, the most naive NB multiplier design (or Massey-Omura multiplier \cite{MasseyOmura})
-% contains 53 gates in total, which is a great saving in area cost comparing to StdB multipliers.
-% 
-% \section{Design a Normal Basis Multiplier on Gate Level}
-% The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
-% we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
-% very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
-% test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
-% routine of several kinds of modern NB multipliers.
-% \subsection{Sequential Multiplier with Parallel Outputs}
-% The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
-% the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.
-% 
-% \begin{figure}[H]
-% \centering{
-% \includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
-% \caption{A typical SMSO structure of Massey-Omura multiplier}
-% \label{fig:MasseyOmura}}
-% \end{figure}
-% 
-% Figure \ref{fig:MasseyOmura} shows the basic architecture of a Massey-Omura multiplier. The operands
-% $A$ and $B$ are 2 arrays of flip-flops which allow 1-bit right-cyclic shift every clock cycle.
-% The logic gates in the boxes implements the matrix multiplication with $\lambda$-Matrix $M^{(0)}$, 
-% while each AND gate corresponds to term $a_ib_j$ and each XOR gate corresponds to addition $a_ib_j+a_{i'}b_{j'}$. 
-% The XOR layer has only 1 output, giving out 1 bit of product $C$ every clock cycle.
-% 
-% The behavior of Massey-Omura multiplier can be concluded as: pre-load operands $A,B$ and reset $C$ to 0, after 
-% executing for $k$ clock cycles, the data stored in flip-flop array $C$ is the product $A\times B$.
-% We note that there is only one output giving 1 bit of the product each clock cycle, which matches the 
-% definition of serial output to communication channel. Therefore this type of design is named as 
-% sequential multiplier with serial output (SMSO).
-% The SMSO architecture need $C_N$ AND gates and $C_N-1$ XOR gates, which equals to $2k-1$ AND gates and $2k-2$ XOR gates
-% if it is designed using ONB. In fact, the number of gates can be reduced if the multiplication is 
-% implemented using a conjugate of SMSO.
-% 
-% The gate-level logic boxes are implementing following function:
-% \begin{equation}
-% \label{eqn:SMPOterms}
-% c_l = row_1(A\times M^{(l)}) \times B + row_2(A\times M^{(l)}) \times B + \cdots + row_{k}(A\times M^{(l)}) \times B
-% \end{equation}
-% It can be decomposed into $k$ terms. If we only compute one term for each $c_l,~0\leq l\leq k-1$ in one 
-% clock cycle, make $k$ outputs and add them up using shift register after $k$ clock cycles, 
-% it will generate the same result with SMSO. This kind of architecture is named as 
-% sequential multiplier with parallel outputs (SMPO). The basic SMPO, as a conjugate of Massey-Omura multiplier,
-% is invented by Agnew et al. \cite{agnew1991implementation}.
-% 
-% \begin{figure}[H]
-% \centering{
-% \includegraphics[width=\textwidth]{newfig/mySMPO.eps}
-% \caption{5-bit Agnew's SMPO.
-Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, i.e.
-if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
-Index $w = 2i\pmod 5$}
-% \label{fig:SMPO}}
-% \end{figure}
-% 
-% \begin{Example}[5-bit Agnew's SMPO]
-% Given GF $\mathbb F_{2^5}$ and primitive element $\alpha$ defined by irreducible polynomial 
-% $\alpha^5+\alpha^2+1=0$, normal element $\beta = \alpha^5$ constructs an ONB $\{\beta,\beta^2,\beta^4,\beta^8,\beta^{16}\}$.
-% The $0$-th $\lambda$-Matrix for this ONB is
-% \begin{equation*}
-% M^{(0)} = \left(\begin{array}{lcccr}
-% 0 &1 &0 &0 &0 \\
-% 1 &0 &0 &1 &0 \\
-% 0 &0 &0 &1 &1 \\
-% 0 &1 &1 &0 &0 \\
-% 0 &0 &1 &0 &1
-% \end{array}\right)
-% \end{equation*}
-% Then a typical design of 5-bit Agnew's SMPO is depicted in figure \ref{fig:SMPO}.
-% 
-% The operands part of this circuit is the same with Massey-Omura multiplier. The differences are on 
-% the matrix multiplication part, while it is implemented as separate logic blocks for 5 outputs,
-% and the 5 blocks are connected in a shift register fashion. By analyzing the detailed function of 
-% logic blocks, we can reveal the mechanism of Agnew's SMPO.
-% 
-% Suppose we implement $M^{(0)}$ as the logic block in SMSO. In the first clock cycle, the output is
-% \begin{equation}
-% \label{eqn:5bitSMPO_c0}
-% c_0 = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4
-% \end{equation}
-% Note it is written in the form of equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
-% remaining bits of the product, which can be written in following general form polynomial:
-% \begin{align}
-% c_i =& b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
-% &+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
-% (a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
-% \end{align}
-% Note all index calculations are reduced modulo 5.
-% 
-% Now let us observe the behavior of 5-bit Agnew's SMPO. Initially all DFFs are reset to 0. 
-% In the first clock cycle,
-% signal sent to the flip-flop in block $R_0$ denotes function:
-% $$R_0^{(1)} = a_1b_0$$
-% It equals to the first term of equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
-% this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
-% from operands (shifted by 1 bit), generating signal $a_u,a_v$ and $b_w$. Concretely,
-% signal sent to flip-flop in block $R_1$ is:
-% $$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
-% which forms first 2 terms of equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
-% on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
-% $$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
-% which equals to $c_0$ in equation \ref{eqn:5bitSMPO_c0}.
-% After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
-% the result of $c_0$ is computed after 5 clock cycles and given on $r_0$.
-% 
-% If we track $R_1\to R_2\to R_3 \to \cdots \to R_0$, we can obtain $c_1$ respectively.
-% Thus we conclude that Agnew's SMPO functions the same with Massey-Omura multiplier.
-% \end{Example}
-% 
-% The design of Agnew's SMPO guarantees that there is only one AND gate in each $R_i$ block.
-% For ONB, adopting Agnew's SMPO will reduce the number of AND gates from $2k-1$ to $k$.
-% 
-% \subsection{Multiplier not based on $\lambda$-Matrix}
-% Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
-% which means that they will be identical if unrolled to full combinational circuits. 
-% After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
-% some of them jump out of the circle and are independent from $\lambda$-Matrix.
-% One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
-% \cite{RHmulti}, which is therefore called RH-SMPO.
-% 
-% \begin{figure}[H]
-% \centering{
-% \includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
-% \caption{5-bit RH-SMPO}
-% \label{fig:5bitRH}}
-% \end{figure}
-% 
-% Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
-% in figure \ref{fig:SMPO}. A brief proof is as follows: 
-% 
-% \begin{Proof}
-% First, we define an auxiliary function for $i$-th bit 
-% \begin{equation}
-% \label{eqn:aux}
-% F_i(A,B) = a_ib_i\beta + \sum_{j=1}^v d_{i,j}\beta^{1+2^j}
-% \end{equation}
-% where $0\leq i\leq k-1, v = \lfloor k/2\rfloor, 1\leq j \leq v$.
-% The $d$-layer index $d_{i,j}$ is defined as
-% \begin{equation}
-% \label{eqn:auxDC}
-% d_{i,j} = c_{a,i} c_{b,i} = (a_i+a_{i+j})(b_i+b_{i+j}),~~1\leq j\leq v
-% \end{equation}
-% $i+j$ here is the result reduced modulo $k$. Note that there is a special boundary case when
-% $k$ is an even number ($v = \frac{k}{2}$):
-% $$d_{i,v} = (a_i+a_{i+v})b_i$$
-% With the auxiliary function, we can utilize following theorem (proof refers to \cite{RHmulti}):
-% \begin{Theorem}
-% Consider three elements $A,B$ and $R$ such that $R=A\times B$ over $\Fkk$. Then,
-% $$R=(((F_{k-1}^2+F_{k-2})^2+F_{k-3})^2+\cdots+F_1)^2+F_0$$
-% \end{Theorem}
-% This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
-% $R_i$ flip-flops. Concretely, the multiplier behavior is an implementation of 
-% following algorithm:
-% 
-% \begin{algorithm}[H]
-% \SetAlgoNoLine
-% 
-%  \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
-%  \KwOut{$R=A\times B$}
-% %%%%%%%%%%%%%%%%%%%%
-%   Initialize $A,B$ and aux var $X$ to 0\;
-%   \For { ($i=0$;   $i < k$; ++i ) }
-%   {
-% 	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Eq.\ref{eqn:aux}*/\;}
-% 	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
-%   }
-%   $R\gets X$
-% \caption{NB Multiplication Algorithm in RH-SMPO \cite{RHmulti}}\label{alg:RHmulti}
-% \end{algorithm}
-% 
-% In this algorithm, we use a fixed auxiliary function $F_{k-1}$ inside the loop.
-% This is because of equation
-% $$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
-% So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
-% $F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
-% \end{Proof}
-% 
-% To better understand the mechanism of RH-SMPO, we will use this 5-bit 
-% RH-SMPO as an example and introduce the details on how to design it.
-% \begin{Example}[Designing a 5-bit RH-SMPO]
-% From equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
-% and XOR gates in $c$-layer according to equation \ref{eqn:auxDC}. Concretely, as algorithm \ref{alg:RHmulti}
-% describes, we implement auxiliary function $F_{k-1}$ in the logic:
-% $$i = k-1 = 4;~~v=\lfloor 5/2 \rfloor = 2$$
-% \begin{equation}
-% \label{eqn:5bitRHaux}
-% F_{4}(A,B) = a_4b_4\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j} = d_0\beta+\sum_{j=1}^2 d_{4,j}\beta^{1+2^j}
-% \end{equation}
-% Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
-% $$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$
-% $$c_3 = a_1+a_4,~c_4 = b_1+b_4,~d_2=d_{4,2}= c_3c_4 = (a_4+a_1)(b_4+b_1)$$
-% The difficult part of the whole design is to deploy XOR gates in $e$-layer. 
-% As the logic layer closest to the outputs $R_i$, $e$-layer actually finishes the implementation of 
-% $F_{k-1}(A,B)$. But it is not a simple addition; the reason is before bit-wise adding to $X^2$, it is necessary to 
-% turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
-% $i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.
-% 
-% In order to obtain information 
-% indicating interconnections between $d$-layer and $e$-layer, we need to interpret $\beta^{1+2^j}$ 
-% to NB representation.
-% There is a concept called {\bf multiplication table} (M-table) which can assist this interpretation. It is defined as 
-% a $k\times k$ matrix $T$ over $\mathbb F_2$:
-% \begin{equation}
-% \label{eqn:multitable}
-% \begin{bmatrix}
-% \beta^{1+2^0} \\ \beta^{1+2^1} \\ \beta^{1+2^2} \\ \vdots \\ \beta^{1+2^{k-1}}
-% \end{bmatrix}
-% = \beta
-% \begin{bmatrix}
-% \beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
-% \end{bmatrix}
-% =
-% \begin{bmatrix}
-% T_{0,0}      &   T_{0,1}        & \dots & T_{0,k-1}\\
-% T_{1,0}    &   T_{1,1}           & \dots & T_{1,k-1}\\
-% T_{2,0}    &   T_{2,1}           & \dots & T_{2,k-1}\\
-% \vdots & \vdots              & \ddots     & \vdots \\
-% T_{k-1,0}    &   T_{k-1,1}           & \dots & T_{k-1,k-1}
-% \end{bmatrix}
-% \begin{bmatrix}
-% \beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
-% \end{bmatrix}
-% = {\bf T}
-% \begin{bmatrix}
-% \beta \\ \beta^2 \\ \beta^4 \\ \vdots \\ \beta^{2^{k-1}}
-% \end{bmatrix}
-% \end{equation}
-% 
-% It is a known fact that M-table $T$ can be converted from $\lambda$-Matrix $M$:
-% $$M_{i,j}^{(0)} = T_{j-i,-i}$$
-% with indices reduced modulo $k$ (proof refers to appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
-% 
-% \begin{figure}[H]
-% \centering{
-% \includegraphics[width=3in]{newfig/multitable.pdf}
-% \caption{A $5\times 5$ multiplication table}
-% \label{fig:multitable}}
-% \end{figure}
-% 
-% Note that we only use row 1 and row 2 from the M-table since range $1\leq j \leq 2$.
-% All nonzero entries in these 2 rows corresponds to the interconnections between $d$-layer and 
-% $e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
-% between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in equation \ref{eqn:multitable}:
-% \begin{equation*}
-% \beta\cdot\beta^2 = 
-% \begin{bmatrix}
-% 1 & 0 & 0 & 1 & 0
-% \end{bmatrix}
-% \begin{bmatrix}
-% \beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
-% \end{bmatrix}
-% = \beta + \beta^{2^3}
-% \end{equation*}
-% 
-% Similarly, from row 2 of M-table we derive that $d_2$ has fanouts $e_3,e_4$:
-% \begin{equation*}
-% \beta\cdot\beta^{2^2} = 
-% \begin{bmatrix}
-% 0 & 0 & 0 & 1 & 1
-% \end{bmatrix}
-% \begin{bmatrix}
-% \beta \\ \beta^2 \\ \beta^4 \\ \beta^8 \\ \beta^{16}
-% \end{bmatrix}
-% = \beta^{2^3} + \beta^{2^4}
-% \end{equation*}
-% 
-% Let us look back at equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
-% The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
-% After gathering all interconnection information, we can translate it to gate-level circuit implementation:
-% $$e_0 = d_0+d_1,~e_3=d_1+d_2,~e_4=d_2$$
-% 
-% Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
-% $R_{i-1}$ in last clock cycle, we can connect using relation 
-% $$R_i = r_{i-1} + e_i$$
-% In this example, according to the M-table in figure \ref{fig:multitable}, columns $e_1,e_2$
-% have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.
-% 
-% This finishes the full design procedure for a 5-bit RH-SMPO.
-% \end{Example}
-% 
-% The area cost of RH-SMPO is even smaller than Agnew's SMPO. XOR gates corresponds to all nonzero entries 
-% in M-table, which is with the same number of nonzero entries in $\lambda$-Matrix ($C_N$). The number of AND gates 
-% equals to $v$ plus 1 (for gate $d_0$). When using ONB ($C_N = 2k-1$), the total number of gates 
-% is $2k+\lfloor \frac{k}{2}\rfloor$.
-
-\section{Full-Blown Verification Procedure for Normal Basis Multiplier Functional Correctness Checking}
-Once a gate-level design of a NB multiplier is generated, it is ready to be verified using 
-similar approach appear in section \ref{sec:recha_approach}. The following part borrows contents
-from my own conference paper \cite{myDATE}.
+\section{Formal Verification of Normal Basis Multiplier using Gr\"obner Basis}
+The gate-level design of a NB multiplier can be generated using appoaches introduced in 
+Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
+an approach similar to that in Chapter \ref{ch:reacha}. 
+First we introduce the sketch of our approach using abstraction term order (ATO)
+mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
+the concept ``RATO", which is previously used in Section \ref{sec:improve}.
+% The following part borrows contents
+% from my own conference paper \cite{myDATE}.
 \subsection{Implicit Unrolling based on Abstraction with ATO}
-In preliminaries we talk about the abstraction basics. If we use elimination term order 
-$intermediate~variables~>~R~>~A,B$, the function of the combinational logic component 
-can be abstracted as 
+% fig:comb_RAB
+
+The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
+$intermediate~variables~>~R~>~A,B$ for the circuit in Figure \ref{fig:comb_RAB} to compute Gr\"obner basis, 
+the function of the combinational logic component can be abstracted as 
 $$R = \Func(A,B)$$
-If we are verifying the functional correctness of a combinational NB multiplier (e.g. Mastrovito multiplier
-or Montgomery multiplier), the function given by abstraction will be $R=AB$.
-While in the sequential case, the function of combinational logic only fulfills a part of the multiplication,
-such as $F_{k-1}$ in equation \ref{eqn:aux}. Nevertheless, the abstraction still provides a word-level 
-representation which works definitely better on unrolling than bit-level expressions.
+To verify the functional correctness of a combinational NB multiplier (e.g. Mastrovito multiplier
+or Montgomery multiplier), the function given by abstraction will be computed as
+$$R=A\cdot B$$
+While in the sequential case, the function of combinational logic only fulfills a part of the multiplication.
+For example, in the RH-SMPO introduced in Section \ref{sec:nbdesign}, 
+the combinational logic actually implements  $F_{k-1}$ in Equation \ref{eqn:aux},
+while computing all of $F_i$ requires $k$-cycle unrolling of the circuit.
+Nevertheless, the abstraction still provides a word-level 
+representation which can be more efficient for unrolling than bit-level expressions.
 In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
@@
  \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
-   state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}
-
+   state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
+\KwOut{Circuit function $R_{final}$ after $k$ cycles of $A_{init}, B_{init}$}
   $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
+  \tcc{$A, B$ are word-level variables, solutions to polynomial equations 
+  $\mathcal{G},\mathcal{H}$ denote the initial values of preloaded operands $A_{init},B_{init}$}
   $i = 0$\;
@@
   	$i \gets i + 1$\;
-%	$to^i(R',A',B') \gets$  $GB( \langle J_{ckt}, J_0,
-%    from^{i-1}(R,A,B)\rangle )$ with abstraction term order\;
 	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
@@
 	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
+	\tcc{Using projections of varieties from abstraction theory, we get NS variables $R',A',B'$
+	in terms of PS $A,B$}
 	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
+	\tcc{By replacing NS variables with PS variables we push it to next time-frame}
   }
@@
 \DecMargin{1em}
-%\vspace{-0.1in}
+
+Notations in the following figure as well as in next example comply with 
+the notations in this algorithm.
 
@@
 We follow  the sequential GF circuit model of
-figure \ref{fig:sequential}, with word-level variables $A, B, R$
-denoting {\it present states (PS)} and $A', B', R'$ denoting {\it next
+Figure \ref{fig:sequential}, with word-level variables $A, B, R$
+denoting the {\it present states (PS)} and $A', B', R'$ denoting the {\it next
   states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
@@
 unroll the machine, and provide word-level algebraic $k$-cycle
-abstraction for $R'$ as $R' = \F(A_{init}, B_{init})$.
+abstraction for $R'$ as $R' = \Func(A_{init}, B_{init})$.
 
@@
 reachable states up to iteration $i$ in terms of the elimination ideal. 
-These computations are analogous to those of image computations performed in FSM reachability. 
+These computations are analogous to those of image computations performed in 
+the FSM reachability algorithm (given in Chapter \ref{ch:reacha}).
+
+In the practice of sequential GF multiplier verification, the combinational 
+logic actually implements function not only related to current operands $A$ and $B$,
+but also involved with PS variable (i.e. temporary product) $R$. Which can be 
+obtained using the abstraction:
+$$R' = \Func(A,B,R)$$
+
+Using ATO, if we put $R$ ahead of $R',A,B$ in term ordering, $R$ is thus eliminated
+and the result in single iteration will be $R' = \Func(A,B)$. NS operands 
+$A'$ and $B'$ are right-cyclic shift of $A$ and $B$, which can be directly written.
+If initial values $A_{init}, B_{init}$ are treated as parameters, the NS ideal 
+contains polynomials with $A', B'$ and $R' = \Func(A_{init},B_{init})$. This 
+is also shown in the following example.
+
 \begin{Example}[Functional verification of 5-bit RH-SMPO]
@@
 operands $A,B$ is doing cyclic shift, while transition function for $R$ has to be computed through Gr\"obner basis
-abstraction approach. Following ideal $J_{ckt}$ from line 5 in algorithm \ref{alg:modified} is the ideal for 
+abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
 all gates in combinational logic block and definition of word-level variables.
-\begin{align}
-J_{ckt} = & d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4,\nonumber\\
-& c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2, \nonumber\\
-& R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4,\nonumber\\
-& A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},\nonumber\\
-		 & B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18},\nonumber\\
-		 & R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18},\nonumber\\
-		 & R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18};\nonumber
-\end{align}
+$$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$
+
+\begin{eqnarray*}
+ \left. 
+\begin{aligned}
+d_0+a_4b_4, c_1+a_0+a_4, c_2+b_0+b_4, d_1+c_1c_2, c_3+a_1a_4\\
+c_4+b_1b_4, d_2+c_3c_4, e_0+d_0+d_1, e_3+d_1+d_2, e_4+d_2 \\
+R_0+r_4+e_0, R_1+r_0, R_2+r_1, R_3+r_2+e_3, R_4+r_3+e_4
+\end{aligned}\right\} ~~~\{f_1\dots f_{15}\}\\
+ \left.
+\begin{aligned}
+ A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}\\
+		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\\
+		  R'+r_0'\alpha^5+r_1'\alpha^{10}+r_2'\alpha^{20}+r_3'\alpha^9+r_4'\alpha^{18}\\
+		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18}
+		 \end{aligned}\right\} ~~~\{f_{16}\dots f_{19}\}
+\end{eqnarray*}
+
 In our implementation here, since we only focus on the output variable $R$, evaluations of intermediate input 
-operands $A, B$ are unnecessary. Polynomials about $A$ and $B$ can be removed from $J_{ckt}$, and $R$ is directly
-evaluated by initial operands $A_{init}$ and $B_{init}$, which are associated with present state bit-level inputs
-$a_0,a_1,\dots,a_4$ and $b_0,b_1,\dots,b_4$ by polynomials in $from^i$.
-
-According to line 5 of algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
-Gr\"obner basis with abstraction term order (copy details here). There is a polynomial in form of $R'+\Func(A_{init},B_{init})$,
+operands $A, B$ are replaced by evaluations with parameters $A_{init},B_{init}$. 
+Thus polynomials describing $A$ and $B$ ($f_{16}$ and $f_{17}$) can be removed from $J_{ckt}$, and $R$ is directly
+evaluated by initial operands $A_{init}$ and $B_{init}$. Those two parameters are associated with PS bit-level inputs
+$a_0,a_1,\dots,a_4$ and $b_0,b_1,\dots,b_4$ by polynomials given in $from^i$.
+
+According to line 5 of Algorithm \ref{alg:modified}, we merge $J_{ckt}$, $J_0$ and $from^i$, then compute its
+Gr\"obner basis with abstraction term order. 
+Concretely, in the first iteration $from^0$ contains three generators. The first one describes PS variable $R$ --
+temporary (or initial, in first iteration) product, which equals to 0 according to the mechanism of sequential 
+GF multiplier. The polynomial is written as 
+$$R$$
+The other two polynomials describes PS variable $A,B$ -- current multiplication operands, and we write them 
+in parameters $A_{init}, B_{init}$. Since this is the first iteration, they have the same form with $f_{16}$ and $f_{17}$:
+\begin{align*}
+& A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18} \\
+& B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}
+\end{align*}
+After computing the Gr\"obner basis of $J_{ckt}+J_0+from^0$ using ATO 
+$$\text{all other bit-level variables} > R > \{R',A_{init},B_{init}\}$$
+there is a polynomial in form of $R'+\Func(A_{init},B_{init})$,
 which should be included by $to^{i+1}$. $to^{i+1}$ also exclude next state variable $A'$ and $B'$, instead we 
@@
 
-Line 7 in algorithm \ref{alg:modified} is implemented by replacing $R'$ with $R$, $\{a_i', b_j'\}$ with $\{a_i,b_j\}$.
+Line 7 in Algorithm \ref{alg:modified} is implemented by replacing $R'$ with $R$, $\{a_i', b_j'\}$ with $\{a_i,b_j\}$.
 
@@
 \begin{itemize}
-\item Clock 1: ${\bf from^0} = \{R, A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
+\item Clock-cycle 1: ${\bf from^0} = \{R, A_{init}+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
 B_{init}+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18}\}$, \\
@@
 B_{init}+b_4'\alpha^5+b_0'\alpha^{10}+b_1'\alpha^{20}+b_2'\alpha^9+b_3'\alpha^{18}\}$
-\item Clock 2: ${\bf from^1} = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^2) A_{init}^8 B_{init}+(\alpha^4+\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^8+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha^2+\alpha) A_{init}^2 B_{init}^2+(\alpha^4+\alpha) A_{init}^2 B_{init}+(\alpha^4+\alpha^3+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init}^2+(\alpha^3+\alpha+1) A_{init} B_{init},
+\item Clock-cycle 2: ${\bf from^1} = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^2) A_{init}^8 B_{init}+(\alpha^4+\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^8+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha^2+\alpha) A_{init}^2 B_{init}^2+(\alpha^4+\alpha) A_{init}^2 B_{init}+(\alpha^4+\alpha^3+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^2) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init}^2+(\alpha^3+\alpha+1) A_{init} B_{init},
 	A_{init}+a_4\alpha^5+a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^9+a_3\alpha^{18},
@@
 B_{init}+b_3'\alpha^5+b_4'\alpha^{10}+b_0'\alpha^{20}+b_1'\alpha^9+b_2'\alpha^{18}\}$
-\item Clock 3: ${\bf from^2} = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^8+(\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^2+(\alpha^3+1) A_{init}^8 B_{init}+(\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4) A_{init}^4 B_{init}^4+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^2) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}+(\alpha^3+1) A_{init} B_{init}^8+(\alpha) A_{init} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init},
+\item Clock-cycle 3: ${\bf from^2} = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^8+(\alpha^2) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^{16}+(\alpha^4+\alpha^2) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^4+\alpha^3+1) A_{init}^8 B_{init}^2+(\alpha^3+1) A_{init}^8 B_{init}+(\alpha^2) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4) A_{init}^4 B_{init}^4+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^4+\alpha^3+1) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^2) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}+(\alpha^3+1) A_{init} B_{init}^8+(\alpha) A_{init} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init},
 A_{init}+a_3\alpha^5+a_4\alpha^{10}+a_0\alpha^{20}+a_1\alpha^9+a_2\alpha^{18},
@@
 B_{init}+b_2'\alpha^5+b_3'\alpha^{10}+b_4'\alpha^{20}+b_0'\alpha^9+b_1'\alpha^{18}\}$
-\item Clock 4: ${\bf from^3} = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}+(\alpha) A_{init}^8 B_{init}^{16}+(\alpha+1) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init},
+\item Clock-cycle 4: ${\bf from^3} = \{R+(\alpha^4+\alpha^3+1) A_{init}^{16} B_{init}^{16}+(\alpha) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}^4+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^{16} B_{init}+(\alpha) A_{init}^8 B_{init}^{16}+(\alpha+1) A_{init}^8 B_{init}^8+(\alpha^4) A_{init}^8 B_{init}^4+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}^2+(\alpha^4+\alpha^3+\alpha+1) A_{init}^8 B_{init}+(\alpha^4+\alpha^3+\alpha^2+1) A_{init}^4 B_{init}^{16}+(\alpha^4) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^3+\alpha+1) A_{init}^4 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^4 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^4 B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^2 B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}^8+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^4+(\alpha^3+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^4+\alpha^3+\alpha^2+1) A_{init} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha+1) A_{init} B_{init}^8+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init} B_{init}^4+(\alpha^4+\alpha) A_{init} B_{init},
 A_{init}+a_2\alpha^5+a_3\alpha^{10}+a_4\alpha^{20}+a_0\alpha^9+a_1\alpha^{18},
@@
 B_{init}+b_1'\alpha^5+b_2'\alpha^{10}+b_3'\alpha^{20}+b_4'\alpha^9+b_0'\alpha^{18}\}$
-\item Clock 5: ${\bf from^4} = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^8 B_{init}^{16}+(\alpha^3+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^8 B_{init}^4+(\alpha^2+\alpha) A_{init}^8 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}+(\alpha^4+\alpha) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^2+\alpha) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^2) A_{init}^2 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}+(\alpha^3+\alpha+1) A_{init} B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init} B_{init}^8+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^3+\alpha^2+1) A_{init} B_{init}^2+(\alpha) A_{init} B_{init},
+\item Clock-cycle 5: ${\bf from^4} = \{R+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}^{16}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^{16} B_{init}^8+(\alpha^4+\alpha) A_{init}^{16} B_{init}^4+(\alpha^3+1) A_{init}^{16} B_{init}^2+(\alpha^3+\alpha+1) A_{init}^{16} B_{init}+(\alpha^4+\alpha^3+\alpha^2+\alpha+1) A_{init}^8 B_{init}^{16}+(\alpha^3+1) A_{init}^8 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^8 B_{init}^4+(\alpha^2+\alpha) A_{init}^8 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^8 B_{init}+(\alpha^4+\alpha) A_{init}^4 B_{init}^{16}+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^8+(\alpha^4+\alpha^2+\alpha) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^3+1) A_{init}^2 B_{init}^{16}+(\alpha^2+\alpha) A_{init}^2 B_{init}^8+(\alpha^4+\alpha^2) A_{init}^2 B_{init}^2+(\alpha^3+\alpha^2+1) A_{init}^2 B_{init}+(\alpha^3+\alpha+1) A_{init} B_{init}^{16}+(\alpha^3+\alpha^2+1) A_{init} B_{init}^8+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^3+\alpha^2+1) A_{init} B_{init}^2+(\alpha) A_{init} B_{init},
 A_{init}+a_1\alpha^5+a_2\alpha^{10}+a_3\alpha^{20}+a_4\alpha^9+a_0\alpha^{18},
@@
 \end{itemize}
-The final result is
-$from^5(R_{final}) = R_{final}+A_{init}\cdot B_{init}$
+The algorithm returns
+$$from^5(R_{final}) = R_{final}+A_{init} B_{init}$$
+which is the function of the multiplier: $R_{final} = A_{init}\cdot B_{init}$
 \end{Example}
+
 \subsection{Overcome Computational Complexity using RATO}
-Similar to our improvements in section \ref{sec:reachaRATO}, RATO \cite{TimDAC} is also available to 
-accelerate the GB computation here. More specifically, we obviate GB computation by turning it 
-into a single-step multivariate polynomial division: first, in ideal generated by
-gates information polynomials and word-level variable definition polynomials, find the unique pair of polynomial
-generators with leading monomials not relatively prime to each other; then, compute their specification polynomial
-using definition 
-$$Spoly(f_w,f_g) = \frac{LCM}{lt(f_w)}\cdot f_w - \frac{LCM}{lt(f_g)}\cdot f_g$$
-where $LCM$ is least common multiple of $lm(f_w)$ and $lm(f_g)$, and $lt$ denotes the leading term;
-last, reduce $Spoly$ with ideal $J_{ckt}+J_0$, it is possible that the remainder will be canonical polynomial
-function of the circuit. We will illustrate the whole improved procedure by applying RATO on 
-5-bit RH-SMPO in figure \ref{fig:5bitRH}.
+% although commented, but I insist refer to previous chapter
+% TODO experiment on naive slimgb on SMPO
+We implement Algorithm \ref{alg:modified} in \textsc{Singular}, with the same setup and environment 
+with the experiment in Section \ref{sec:exp_reacha}. The result on verifying Agnew's SMPO is shown 
+in following table:
+
+\begin{table}[H]
+\label{tab:slimgb}
+	\begin{center}
+	    \caption{(Dummy)Runtime of Gr\"obner Basis Computation of Mastrovito Multipliers in Singular using ATO $>$.}\label{tab:ATO}
+	    \begin{tabular}{|c|c||c|} 
+	        \hline
+		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (minutes)   \\
+		\hline
+	        $16$	&  $1,871$  & $2.4$ \\
+		$24$	&  $3,135$  & $12$  \\
+	        $32$	&  $5,549$  & $22.6$ \\
+	        $40$	&  $8,587$  & $266$ \\
+		$48$	& $12,327$  & NA (Out of Memory) \\
+	        \hline
+	    \end{tabular}
+	\end{center} 
+\end{table}
+
+It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 40(dummy).
+Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} is also available to 
+accelerate the GB computation here. More specifically, we obviate the GB computation by turning the Buchberger's algorithm 
+into a single-step multivariate polynomial division.
+
+\begin{itemize}
+\item First, a set of polynomials is constructed by polynomials translated from each logic gates, and they generate 
+ideal $J_{gates}$. Similarly, ideal $J_{W-B}$ is generated by polynomials representing the word-bit correspondences.
+They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{W-B}$.
+\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using reverse topological order
+in circuit structure, followed by word-level NS output (e.g. $R'$ in Figure \ref{fig:sequential}), 
+then word-level PS inputs (e.g. $A,B,R$ in Figure \ref{fig:sequential}).
+\item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
+RATO, there only exist one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
+\item Then we only need to compute the S-poly for $(f_w,f_g)$, the reduce the S-poly by $J_{ckt}+J_0$.
+\end{itemize}
+
+% first, in ideal generated by
+% gates information polynomials and word-level variable definition polynomials, find the unique pair of polynomial
+% generators with leading monomials not relatively prime to each other; then, compute their specification polynomial
+% using definition 
+% $$Spoly(f_w,f_g) = \frac{LCM}{lt(f_w)}\cdot f_w - \frac{LCM}{lt(f_g)}\cdot f_g$$
+% where $LCM$ is least common multiple of $lm(f_w)$ and $lm(f_g)$, and $lt$ denotes the leading term;
+% last, reduce $Spoly$ with ideal $J_{ckt}+J_0$, 
+After executing the reduction, the remainder only contains a limited number of variables. It can be further 
+transformed to a canonical polynomial
+function of the circuit. We illustrate the whole improved procedure by applying RATO on 
+5-bit RH-SMPO described in Example \ref{ex:RHSMPO} and Figure \ref{fig:5bitRH}.
 
@@
 \label{ex:newRATO}
-Variable order under RATO is:
+From the circuit topological structure in Figure \ref{fig:5bitRH}, the term order under RATO is:
 \begin{align}
@@
 \end{align}
-Search among all generators of $J_{ckt}$ from Ex.\ref{ex:RHSMPO} using RATO, we find a pair of polynomials whose leading 
+The variables in braces are at the same topological level, they are arranged in LEX order. 
+We search among all generators of $J_{ckt}$ from Example \ref{ex:RHSMPO} using RATO, 
+and find a pair of polynomials whose leading 
 monomials are not relatively prime:
@@
 \end{Example}
-Above example indicates that RATO based abstraction on 5-bit RH-SMPO will result a remainder contains both
-bit-level variables and word-level variables, and the number of remaining variables is still large such that
+Above example indicates that RATO based abstraction on 5-bit RH-SMPO will result a remainder containing both
+bit-level variables and word-level variables. Moreover, the number of variables is still large such that
 Gr\"obner basis computation will be inefficient.
@@
 Since the remainder from \emph{Spoly} reduction contains some bit-level variables, our objective is to compute
-a polynomial contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
-this problem is to replace bit-level variable monomials by equivalent polynomials that only contain word-level
-variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination-fashion 
+a polynomial that contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
+this problem is to replace bit-level variables by equivalent polynomials that only contain word-level
+variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
 approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.
@@
 
-Use Gaussian elimination on coefficients matrix, for example  
-\begin{align}
-Row\ 2 =& Row\ 1\times \alpha^5 + Row\ 2: \nonumber\\
-&a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
-=&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
-\end{align}
-Recursively eliminate $a_1$ from third row, $a_2$ from fourth row, etc. The final solution to this system of
+Then we use Gaussian elimination on coefficients matrix 
+% \begin{align}
+% Row\ 2 =& Row\ 1\times \alpha^5 + Row\ 2: \nonumber\\
+% &a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
+% =&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
+% \end{align}
+to recursively eliminate $a_1$ from third row, $a_2$ from fourth row, etc. The final solution to this system of
 equations is
@@
 \end{Example}
-Similarly we can compute equivalent polynomials $\mathcal{H}_i(R)$ for $r_i$, $\mathcal{T}_j(B)$ for $b_j$.
+Similarly we can compute equivalent polynomials $\mathcal{H}_i(R)$ for $r_i$ and $\mathcal{T}_j(B)$ for $b_j$, respectively.
 Using those polynomial equations, it is sufficient to translate all bit-level inputs in the remainder polynomial
 because of following lemma:
-\begin{Lemma}
+\begin{Proposition}
 \label{lem:bitlevelremainder}
@@
 only one term containing word-level output whose monomial is word-level output itself rather than higher order form.
-\end{Lemma}
+\end{Proposition}
 
@@
 
-\subsection{Solving Linear System for Bit-to-Word Substitution}
-In example \ref{ex:BLVS} we use a Gaussian-elimination-fashion method to solve the system of polynomial
-equations. There is another formalized method to solve following system of equations
+\subsection{Solving a Linear System for Bit-to-Word Substitution}
+In Example \ref{ex:BLVS} we use a Gaussian-elimination method to solve the system of polynomial
+equations. We describe another formal method to solve following system of equations
 \begin{align}
@@
 Let ${\bf s}$ be a vector of $k$ unknowns $s_0,\dots,s_{k-1}$,
-then equation \ref{eqn:betamat} can be solved by using Cramer's rule:
+then Equation \ref{eqn:betamat} can be solved by using Cramer's rule:
 \begin{equation}
@@
 Notice that ${\bf M}$ is constructed by squaring a row and assigning it to next row,
-therefore its determinant belongs to a special sort of determinants:
+therefore its determinant exhibits certain special properties:
 \begin{Definition}
@@
 \label{eqn:Moore}
-\det{M(\alpha_0,\dots,\alpha_{k-1})} = \alpha_0 \prod_{i=1}^{k-1} \prod_{c_0,\dots,c_{i-1}\in \Fp} (\alpha_{i} - \sum_{j=0}^{i-1}c_j\alpha_j)
+\det{M(\alpha_0,\dots,\alpha_{k-1})} = \alpha_0 \prod_{i=1}^{k-1}\left( \prod_{c_0,\dots,c_{i-1}\in \Fp} (\alpha_{i} - \sum_{j=0}^{i-1}c_j\alpha_j)\right)
 \end{equation}
-We use an example to help understanding the notations in Eqn.\ref{eqn:Moore}:
+We use an example to help understanding the notations in Equation \ref{eqn:Moore}:
 \begin{Example}
@@
 Then let $i=2$, we obtain ${c_0,c_1}\in \mathbb F_2$, they can take value from $\{0,0\},\{0,1\},\{1,0\}$ and $\{1,1\}$.
-We add 4 more product terms $\alpha_2,(\alpha_2-\alpha_1),(\alpha_2-\alpha_0),(\alpha_2-\alpha_0-\alpha_1)$ respectively.
+We add 4 more product terms $\alpha_2,(\alpha_2-\alpha_1),(\alpha_2-\alpha_0),(\alpha_2-\alpha_0-\alpha_1)$, respectively.
 
@@
 \end{Lemma}
-\begin{proof}
-According to the definition Eqn.\ref{eqn:Moore}, the Moore determinant consists of all possible linear combinations of 
+\begin{Proof}
+According to the definition Equation \ref{eqn:Moore}, the Moore determinant consists of all possible linear combinations of 
 $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ with coefficients over $\Fp$. If $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$
@@
 of all elements of a field equals to 1, the Moore determinant $\M=1$.
-\end{proof}
-Applying Lemma \ref{lemma:Moore} to Eqn.\ref{eqn:Cramer} gives
+\end{Proof}
+
+Applying Lemma \ref{lemma:Moore} to Equation \ref{eqn:Cramer} gives
 \begin{equation}
@@
 
-However, we observe a fact that solution, if written as the inversion of matrix $M$, is a circulant matrix itself.
-We have following proposition:
-$$\dots$$
-Therefore, for a certain type of ONB, we can directly write down the inversion of corresponding matrix $M$
-as 
-$$\dots$$
-% TODO to be continued...
+% However, we observe a fact that solution, if written as the inversion of matrix $M$, is a circulant matrix itself.
+% We have following proposition:
+% $$\dots$$
+% Therefore, for a certain type of ONB, we can directly write down the inversion of corresponding matrix $M$
+% as 
+% $$\dots$$
+% TODO Since our finding only works on type-II ONB, too complicated to explain, so omit
+
+\subsection{The Overall Verification Approach}
+Based on above concepts and improvements, the functional verification for sequential GF multiplier 
+on word-level with $k$-bit input operands $A,B$ and $k$-bit output $R$ is described as follows:
+\begin{enumerate}[{1)}]
+\item Given a sequential GF multiplier $S$, with word-level $k$-bit inputs $A,B$ and output $R$.
+\item Choose a primitive polynomial $P(x)$ of degree $k$ and construct $\Fkk$, and let $P(\alpha)=0$.
+\item Perform a reverse-topological traversal of $S$ to derive RATO: LEX with $\{x_1>x_2>\cdots>x_d>R'>R>A>B\}$,
+where $\{x_1,\dots,x_d\}$ are bit-level variables of $S$, $R'$ is the NS output value.
+\item Derive the set of bit-level polynomials $\{f_1,\dots,f_s\}$ from each gate in $S$, and represent them 
+using RATO. These polynomials are in the form $x_i+tail(f_i)$ where $x_i$ is the output of corresponding gate.
+\item Compose the bit-level to word-level correspondences polynomials such as $r_0\beta+r_1\beta^2+\cdots+r_{k-1}\beta^{2^{k-1}}+R$.
+Add them to set $\{f_1,\dots,f_s\}$ and generate ideal $J_{ckt}$. Compose the ideal of vanishing polynomials
+$J_0 = \langle x_i^2+x_i, \dots, R^{2^k}+R,\dots\rangle$.
+\item Select the only critical pair $(f_w,f_g)$ that does not have relatively prime leading terms. Compute
+$Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}r$.
+\item Construct matrices $M_0,\dots,M_{k-1}$, where $M_i$ is $M$ with the $i$-column replaced by vector 
+$[R R^2 \cdots R^{2^{k-1}}]^T$, and $M$ is the reverse-circulant matrix generated by vector $(\beta,\beta^2,\dots,\beta^{2^{k-1}})$.
+\item Symbolically compute the determinants $|M_i|$ to find set $F_R$, where $f_{R_i}:R_i+|M_i|$, for $0\leq i \leq k-1$.
+Obtain $F_A, F_B$ from $F_R$ since they have the same form.
+\item Compute $r \xrightarrow{F_A\cup F_B \cup F_R}_{+}r_w$, and iterate $r_w$ as $to^i$ in Algorithm \ref{alg:modified}.
+{\bf Then $r_w$ after iteration $k$ is of the form $R'+\Func(A,B)$ and it is the unique, canonical word-level abstraction
+of $S$ over $\Fkk$ after $k$ clock-cycles}.
+\end{enumerate}
 
 \section{Software Implementation of Implicit Unrolling Approach}
-% TODO this can be divided into Singular part and C++ part, with detailed example on the trick
 Our experiment on different size of SMPO designs is performed with both \textsc{Singular} \cite{DGPS}
-symbolic algebra computation system and our customized toolset deployed using C++.
+symbolic algebra computation system and our customized toolset developed using C++.
 The SMPO designs are given as gate-level netlists with registers, then translated to polynomials to compose
@@
 3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and running 64-bit Linux OS.
-\subsection{Architecture in Singular}
-The Singular tool can read in scripts written in its own format similar to ANSI-C. For SMPO experiment, the main 
-loop of our script file performs the same function as algorithm \ref{alg:modified} describes, while Gr\"obner basis
+
+The \textsc{Singular} tool can read in scripts written in its own format similar to ANSI-C. 
+The input file format for our C++ tool is also designed to compatible with \text{Singular} 
+ring and polynomial definitions. For SMPO experiments, the main 
+loop of our script file performs the same function as Algorithm \ref{alg:modified} describes, while Gr\"obner basis
 computation in main loop can be divided into 4 different function parts:
 
-(i) Pre-process:
+1) Pre-process:
 
 This step is executed only once before main loop starts. The function of pre-process is to compute following system
-of equations for bit-level inputs $a_0 \sim a_{k-1}$:
+of equations for bit-level inputs $a_0 \dots a_{k-1}$:
 \begin{displaymath}
@@
 \end{displaymath}
-The methodology has been discussed in section \ref{sec:blvs}. For 5-bit SMPO example, we start from word-level
-expression polynomial
+The methodology has been discussed in this chapter. For 5-bit SMPO example, we start from bit-word correspondence polynomial
 \begin{displaymath}
@@
 \end{displaymath}
-By replacing bit-level variable $a_i$ with $b_i, r_i$ or $R_i$, and word-level variable $A$ with $B, r, R$ respectively,
+By replacing bit-level variable $a_i$ with $b_i, r_i$ or $r_i'$, and word-level variable $A$ with $B, R, R'$ respectively,
 we can directly get bit-word relation functions for another operand input, pseudo input and pseudo output.
 
-One limitation to Singular tool is the exponential cannot exceed $2^{63}$, so when doing experiments for SMPO larger than
-62 bits, we use a little trick (the feasibility of this trick can also be verified in following steps). Since the BLVS method
-only requires squaring of equations each time, the exponential of word $A$ can only be in the form $2^{i-1}$, i.e. 
-$A^{2^0},A^{2^1},\dots,A^{2^{k-1}}$. To minimize the exponential presenting in Singular tool, we rewrite $2^{i-1}$ to $i$, i.e.
+One limitation to \textsc{Singular} tool is the degree of a term cannot exceed $2^{63}$, so when conducting experiments 
+for SMPO circuits larger than
+62 bits, we rewrite the degree using smaller integers (the feasibility of this rewriting can also be verified 
+in following steps). Our C++ tool does not have data size limit, but it accumulates exponent when computing 
+the power of a variable. Thus if the exponent is as large as $2^{63}$, its runtime is very long. As a 
+result, our C++ tool also adopts and benefits from this rewriting technique.
+Since the bit-to-word substitution procedure
+only requires squaring of equations each time, the exponent of word $A$ can only be in the form $2^{i-1}$, i.e. 
+$A^{2^0},A^{2^1},\dots,A^{2^{k-1}}$. To minimize the exponents presenting in both \textsc{Singular} and our tool, 
+we rewrite exponent $2^{i-1}$ to exponent $i$, i.e.
 $(A^{2^0},A^{2^1},\dots,A^{2^{k-1}}) \to (A, A^2, \dots, A^k)$. In this way result is rewritten to be
 $$a_0 = (\alpha+1)A^5+(\alpha^4+\alpha^3+\alpha)A^4+(\alpha^3+\alpha^2)A^3+(\alpha^4+1)A^2+(\alpha^2+1)A$$
-Thus the exponential will not exceed the Singular data size limit.
-
-This step requires limited substitution operations in Singular, so although we use the naive Gaussian elimination
-method (whose time complexity is $O(k^3)$), the time cost is trivial comparing to following steps. For 33 bits experiment,
-pre-process execution time is 2.7 sec; while for 100 bits experiment time cost is 36 sec.
-
-(ii) Spoly reduction:
-
-First, Spoly is calculated based on RATO, then reduced with the ideal composed by circuit description polynomials ($J$).
+Thus the exponents do not exceed the \textsc{Singular} data size limit.
+
+This step requires limited substitution operations, so although we use the naive Gaussian elimination
+method (whose time complexity is $O(k^3)$), the time cost is trivial comparing to following steps. 
+% For 33 bits experiment,
+% pre-process execution time is 2.7 sec; while for 100 bits experiment time cost is 36 sec.
+
+2) S-poly reduction:
+
+First, S-poly is calculated based on RATO, then reduced with the ideal composed by circuit description polynomials ($J_{ckt}$).
 For already finished experiments, naive reduction (multi-division) is adopted, and this step takes largest portion of total 
@@
 For SMPO experiments, reduced Spoly has following generic form (all coefficients are omitted):
-$$redSpoly = \sum r_i + \sum a_ib_i + \sum a_iB + \sum b_iA + R + r$$
-Note there is no cross-term for bit-level or word-level variables from same side such as $a_ia_j, a_iA$, etc. Consider the necessary
+$$redSpoly = R'+\Func(a_i,b_j,r_l,A,B,R) = \sum r_i + \sum a_ib_i + \sum a_iB + \sum b_iA + R' + R$$
+From observation, there is no cross-product terms of bit-by-bit or bit-by-word variables from the same input/output
+such as $a_ia_j, a_iA$, etc. Consider the necessary
 condition of our trick, this property of reduced Spoly guarantees the word level variable can only exist in the form $A^{2^{i-1}}$,
@@
 
-(iii) Substitute bit-level variables in reduced Spoly:
+3) Substitute bit-level variables in reduced S-poly, i.e. 
+$$\Func(a_i,b_j,r_l,A,B,R) \xrightarrow{a_i+\mathcal{G}_i(A),b_j+\mathcal{G}_j(B),r_l+\mathcal{G}_l(R)} \mathcal{H}(A,B,R)$$
 
 Use the result from pre-process, get rid of $r_i, a_i$ and $b_i$ through substitution. This step yields following polynomial (consider the trick we used):
-$$R + \sum r^i + \sum A^iB^j$$
+\begin{equation}
+\label{eqn:HABR}
+R'+ \mathcal{H}(A,B,R) = R' + \sum R^i + \sum A^iB^j
+\end{equation}
 all coefficients omitted.
 
-(iv) Substitute present state word-level variable $r$ with inputs $A$ and $B$:
-
-According to section \ref{sec:SMPOexperiment}, there is still a polynomial $r_{in}$ in the ideal we want to compute Gr\"obner basis.
-This polynomial has form $r+f'(A,B)$, which is last clock cycle's output ($R+f'(A,B)$) with only leading term replaced in step "$from^i\gets to^i$" in 
-algorithm \ref{alg:modified}. Basically this step has nothing different from last one, however, it must be taken good care of when using our
-trick. There is power of $r$, $r^m$ is originally $r^{2^{m-1}}$; so if $r+f'(A,B)$ contains terms $A^iB^j$, the correct result after doing power
+4) Substitute PS word-level variable $R$ with inputs $A$ and $B$, i.e. 
+$$\Func(A,B,R) \xrightarrow{R+\Func'(A,B)} \Func(A,B)$$
+
+In Equation \ref{eqn:HABR}, there are still terms containing $R$ in the ideal we want to compute Gr\"obner basis.
+For PS variable $R$ we can use relation $R+\Func'(A,B)$, which is the last clock-cycle's output ($R'+\Func'(A,B)$) 
+with only leading term replaced in step "$from^i\gets to^i$" in 
+Algorithm \ref{alg:modified}. Basically this step has nothing different from the last one, however, the degree
+of its terms are not actual degree because of the exponent rewriting technique. For example,
+a power of $r$ -- $r^m$ is originally $r^{2^{m-1}}$. So if $R+\Func'(A,B)$ contains terms $A^iB^j$, the result 
+after exponent rewriting
 is $$(A^{2^{i-1}}B^{2^{j-1}})^{2^{m-1}} = A^{2^{((i+m-2)\bmod k)+1}}B^{2^{((j+m-2)\bmod k)+1}}$$
-So the correct exponential for $A$ and $B$ in $(A^iB^j)^m$ should be $((i+m-2)\bmod k)+1$ and $((j+m-2)\bmod k)+1$, respectively.
-
-Within one main loop, after finishing steps (ii) to (iv), the output should be intermediate multiplication result $R+f(A,B)$. After $k$ loops,
-the output is $R+A\cdot B$ when SMPO is bug-free.
-
-\subsection{Architecture in Customized C++ Toolset}
+Thus the actual exponent for $A$ and $B$ in $(A^iB^j)^m$ should be $((i+m-2)\bmod k)+1$ and $((j+m-2)\bmod k)+1$, respectively.
+
+Within one iteration, after finishing steps 2) to 4), the output should be intermediate multiplication result (temporary product) 
+$R'+\Func(A,B)$. After $k$ iterations,
+the output is $R+A\cdot B$ when SMPO circuit is bug-free.
+
+% \subsection{Architecture in Customized C++ Toolset}
 % what kind of package, what kind of data structure, what kind of algorithm we use (refer to preliminaries part)
@@
 We have implemented our approach within the \textsc{Singular} symbolic
-algebra computation system [v. 3-1-6] \cite{DGPS}. Using our
+algebra computation system [v. 3-1-6] \cite{DGPS} as well as C++/GCC. Using our
 implementation, we have performed experiments to verify two SMPO
@@
 {\it Evaluation of SAT/ABC/BDD based methods:} To verify circuit $S$
-against the polynomial $\F$, we unroll the SMPO over $k$ time-frames,
-and construct a miter against a combinational implementation of $\F$.
+against the polynomial $\Func$, we unroll the SMPO over $k$ time-frames,
+and construct a miter against a combinational implementation of $\Func$.
 A (pre-verified) $\Fkk$ Mastrovito multiplier \cite{mastro:1989} is

### 2016-11-26 dissertation/prelim_GF.tex
dissertation v2_part reacha update 1
dissertation v2_part normal complete

@@
 \chapter{Galois Field and Sequential Arithmetic Circuits}
+\label{ch:prelim_GF}
 This chapter provides a mathematical background for understanding 
@@
 If we use a basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$, we can easily transform polynomial representations
-to binary bit-vector representations by recording the coefficients. For example,
-
-\begin{table}[H]
-\centering
-\caption{Bit-vector, Exponential and Polynomial representation of
-elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
-\pmod{x^4+x^3+1}$}
-\begin{tabular}{|c|c||c|c|} 
-\hline
-$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
-\hline
-$0000$        & $0$           & $1000$  &$\alpha^3$\\
-\hline
-$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
-\hline
-$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
-\hline
-$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
-\hline
-$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
-\hline
-$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
-\hline
-$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
-\hline
-$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
-\hline
-\end{tabular}
-\label{table:booltogalois}  
-\end{table}
+to binary bit-vector representations by recording the coefficients. For example, for elements in
+$\F_{2^4}$, the results are shown in table \ref{tab:gfelement}, column ``Polynomial".
+
+% \begin{table}[H]
+% \centering
+% \caption{Bit-vector, Exponential and Polynomial representation of
+% elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
+% \pmod{x^4+x^3+1}$}
+% \begin{tabular}{|c|c||c|c|} 
+% \hline
+% $a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
+% \hline
+% $0000$        & $0$           & $1000$  &$\alpha^3$\\
+% \hline
+% $0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
+% \hline
+% $0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
+% \hline
+% $0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
+% \hline
+% $0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
+% \hline
+% $0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
+% \hline
+% $0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
+% \hline
+% $0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
+% \hline
+% \end{tabular}
+% \label{table:booltogalois}  
+% \end{table}
 
@@
 0 & 1 & 1
-\end{array} \right).
+\end{array} \right)
 \end{equation*}
@@
 
+\begin{center}
 \def\divrule#1#2{%
@@
 \end{tabular}
+\end{center}
 \vspace{0.5cm}
@@
 is
-(Computation of $\lambda$-Matrix refers to appendix \ref{append:NB})
+(Computation of $\lambda$-Matrix refers to Appendix \ref{append:NB})
 \begin{equation*}
@@
 \section{Design a Normal Basis Multiplier on Gate Level}
+\label{sec:nbdesign}
 The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
@@
 \end{equation*}
-Then a typical design of 5-bit Agnew's SMPO is depicted in figure \ref{fig:SMPO}.
+Then a typical design of 5-bit Agnew's SMPO is depicted in Figure \ref{fig:SMPO}.
 
@@
 \end{equation}
-Note it is written in the form of equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
+Note it is written in the form of Equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
 remaining bits of the product, which can be written in following general form polynomial:
@@
 $$R_0^{(1)} = a_1b_0$$
-It equals to the first term of equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
+It equals to the first term of Equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
 this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
@@
 $$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
-which forms first 2 terms of equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
+which forms first 2 terms of Equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
 on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
 $$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
-which equals to $c_0$ in equation \ref{eqn:5bitSMPO_c0}.
+which equals to $c_0$ in Equation \ref{eqn:5bitSMPO_c0}.
 After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
@@
 Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
-in figure \ref{fig:SMPO}. A brief proof is as follows:
+in Figure \ref{fig:SMPO}. A brief proof is as follows:
 
@@
   {
-	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Eq.\ref{eqn:aux}*/\;}
+	$X \gets X^2+F_{k-1}(A,B)$ \CommentSty{/*use aux-func from Equation \ref{eqn:aux}*/\;}
 	$A\gets A^2,~B\gets B^2$ \CommentSty{/*Right-cyclic shift $A$ and $B$*/\;}
@@
 \begin{Example}[Designing a 5-bit RH-SMPO]
-From equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
-and XOR gates in $c$-layer according to equation \ref{eqn:auxDC}. Concretely, as algorithm \ref{alg:RHmulti}
+From Equation \ref{eqn:aux} we can deploy AND gates in $d$-layer according to $d_{i,j}$,
+and XOR gates in $c$-layer according to Equation \ref{eqn:auxDC}. Concretely, as Algorithm \ref{alg:RHmulti}
 describes, we implement auxiliary function $F_{k-1}$ in the logic:
@@
 $$M_{i,j}^{(0)} = T_{j-i,-i}$$
-with indices reduced modulo $k$ (proof given in appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
+with indices reduced modulo $k$ (proof given in Appendix \ref{append:ONB}). Thus we can write down the M-table of $\mathbb F_{2^5}$ with current NB $N$:
 
@@
 $e$-layer. For example, row 1 has two nonzero entries at column 0 and column 3, which corresponds to interconnections 
-between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in equation \ref{eqn:multitable}:
+between $d_1$ and $e_0,e_3$. This conclusion comes from row 1 in Equation \ref{eqn:multitable}:
 \begin{equation*}
@@
 
-Let us look back at equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
+Let us look back at Equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
 The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
@@
 $$R_i = r_{i-1} + e_i$$
-In this example, according to the M-table in figure \ref{fig:multitable}, columns $e_1,e_2$
+In this example, according to the M-table in Figure \ref{fig:multitable}, columns $e_1,e_2$
 have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.

### 2016-11-26 dissertation/prelim_basis.tex
dissertation v2_part normal complete

@@
 derived from the hardware
-implementation of the Galois field arithmetic circuit over $\Fkk$, as figure \ref{fig:abs_setup} shows.
+implementation of the Galois field arithmetic circuit over $\Fkk$, as Figure \ref{fig:abs_setup} shows.
 This circuit performs some unknown function $f$ over 
@@
 Consider a 2-bit multiplier over ${\mathbb{F}}_{2^2}$ with
-$P(x)=x^{2}+x+1$, given in Fig. \ref{fig:2bitmul}. Variables $a_0,
+$P(x)=x^{2}+x+1$, given in Figure \ref{fig:2bitmul}. Variables $a_0,
 a_1, b_0, b_1$ are primary inputs, $z_0, z_1$ are primary outputs, and

### 2016-11-26 dissertation/recha.tex
dissertation v2_part normal complete

@@
 \section{Motivation}
+The motivation of this part of my research consists of many inspirations and observations.
+In this dissertation, we will summarize those major ones into 3 topics. First, we start from 
+a sequential circuit partial scan paper to discuss the inspirations of our work as well as some 
+difficulties to overcome. Secondly, we state the observation that word-level description is 
+increasingly important and common in characterizing the data flow on modern large-size datapaths. 
+Last but not least, we give instances that some prerequisites for supporting word-level techniques 
+are already available. Thus we cover both the necessity and sufficiency of developing such 
+a word-level traversal technique in our work.
+
 \subsection{Limitations of Full Scan Algorithms}
@@
 procedure, this unrolling is called {\bf explicit unrolling}. In the SIMPSON algorithm, the states cannot be 
-directly read in the execution; instead, they are implicitly represented using a conjuction of several 
+directly read in the execution; instead, they are implicitly represented using a conjunction of several 
 Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.
 
-However, the SIMPSON algorithm proposed by the author is usually not practical. The conjuctions of 
+However, the SIMPSON algorithm proposed by the author is usually not practical. The conjunctions of 
 Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
@@
 
-\subsection{Word-level Data Flow on Modern Datapaths}
+\subsection{Word-level Data Flow on Modern Datapath Designs}
+The integrity of modern digital circuit designs is very high. For example, processor A10 designed by Apple
+integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such high integrity make the silicon 
+implementation of large datapaths possible. In recent decades, leading by top-notch central processing units (CPUs)
+and high bandwidth memory (HBM), 64-bit or even larger datapaths frequently appear in digital ICs.
+Meanwhile, with the development of electrical design automation (EDA) tools, data flow on those datapaths
+are directly described by the designer as word-level specifications. Therefore, it will be straightforward 
+and beneficial for users if formal verification tools can work on word-level. Moreover, adopting word-level 
+techniques will greatly reduce the state space and make verification executes faster by orders of magnitude.
+
+In order to throw light on the advantages using word-level techniques, we pick a typical 
+digital circuit component in modern 64-bit MIPS processor as an example.
+
+\begin{Example}[Verification of sequential multiplication hardware]
+
+% fig:MIPS
+
+Figure \ref{fig:MIPS} depicts a sequential multiplication hardware implementation within a 64-bit 
+MIPS. Initially, one multiplicand is preloaded to the lower 64 bits of the product registers. 
+Iteratively, the last significant bit (LSB) of current (temporary) product is used as flag to activate the ALU 
+to add on the other multiplicand. For each iteration the data in product registers shift right by 1 bit.
+Finally when the most significant bit (MSB) of preloaded multiplicand arrives the MSB of product registers,
+the registers contains the result -- 128 bits product. The behavior can be concluded as following algorithm:
+
+\vspace{0.2in}
+
+\IncMargin{1em}
+\begin{algorithm}[H]
+\SetAlgoNoLine
+\Indm
+ \KwIn{Multiplicand $A,B$}
+ \KwOut{Product $C$}
+\Indp
+
+  Preload $B$ into lower 64-bit of Product Register $P$\;
+  \Repeat{64 Repetitions}
+  {
+  	\If{Last Bit of Product Register $LSB(P)==1$}
+  	{
+		$P = P_{1/2}+B$\;
+	}
+	Right shift $P$\;
+  }
+\Return{$C=P$}
+\caption {Sequential multiplication hardware in 64-bit MIPS}\label{alg:MIPS}
+\end{algorithm}
+\DecMargin{1em}
+
+\vspace{0.2in}
+
+% List the detailed bit-level property need to check P==P+B
+Traditionally, to verify the functional correctness of this multiplier, satisfiability (SAT) based or BDD based 
+model checking is applied on basic function units. For example, as a part of functional verification, 
+we would like to check ``$P=P_{1/2}+B$" is correctly executed. Then in a model checker we need to add following 
+specifications:
+
+\begin{align*}
+& en\_ALU \\
+\land & s_0 = a_0 \oplus p_0 \\
+\land & s_1 = a_1 \oplus p_1 \oplus (a_0 \land p_0) \\
+\land & s_2 = a_2 \oplus p_2 \oplus ((a_1 \oplus p_1) \land a_0 \land p_0 \lor a_0 \land p_0) \\
+\land & \cdots \\
+\land & s_{63} = a_{63} \oplus p_{63} \oplus (c_{63})
+\end{align*}
+
+We can see that when checking a single part of the whole structure, the number of clauses 
+needed will increase to $k+1$ when using $k$-bit datapath. Considering the formula representing 
+carry-in will become longer and longer, the final conjunction of all clauses will contain 
+$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables
+\begin{equation}
+\label{eqn:wordspec}
+S=P_{1/2}+B
+\end{equation}
+The state space will be greatly reduced.
+\end{Example}
+
+On the other hand, the user of conventional model checker will have to write all these clauses.
+These clauses contain cross-literals, e.g. $s_2$ may associate with $a_1, p_0$, etc. If the user 
+is not familiar with the implementation of this adder, those cross-literals will bring inconveniences
+and confusions. However, if word-level techniques allow specification like equation \ref{eqn:wordspec},
+the verification tool will be very user-friendly and straightforward.
 
 \subsection{Prerequisites of Word-level Techniques}
+When given a bit-level netlist, the prerequisites to use word-level techniques are to convert bit-level 
+to word-level first. This conversion is usually completed by abstraction techniques.
+
+An old but universally effective abstraction method is {\bf Lagrange's interpolation}. Instead of using 
+it in real algebra, it can also be extended to Galois field. Here we use an example to illustrate the 
+conversion in Galois field using Lagrange's interpolation.
+
+\begin{Example}[Lagrange's interpolation]
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=4in]{newfig/Lagrange.pdf}
+\caption{Gate-level netlist for Lagrange's interpolation example}
+\label{fig:Lagrange}}
+\end{figure}
+
+Assume we are given gate-level netlist shown in figure \ref{fig:Lagrange}. It can be written as 3 
+Boolean equations:
+\begin{align*}
+z_0 &= a_0 \lor a_1 \lor a_2 \\
+z_1 &= a_1 \land \neg a_2 \lor a_0 \land \neg a_1 \land a_2 \\
+z_2 &= a_1 \lor \neg a_0 \land a_2
+\end{align*}
+Define 2 word-level variables $A,Z$ as input and output:
+$$A = \{a_2a_1a_0\},~Z = \{z_2z_1z_0\}$$
+To convert bit-level to word-level, we need to find mapping $\mathbb{B}^3 \rightarrow \mathbb{B}^3$,
+or $\F_{2^3} \rightarrow \F_{2^3}$. The latter one, as mentioned in preliminaries, is a polynomial
+function in $\F_{2^3}$.
+
+For each element in $\F_{2^3}$, we write down the truth table as follows:
+\begin{table}[H]
+\label{tab:truthtable}
+\centering{\small
+\begin{tabular}{|c|ccc|c|}
+\hline
+$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
+\hline
+000  &0 &$\rightarrow$&000 & 0 \\
+001  &1 &$\rightarrow$&001 & 1 \\
+010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
+011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
+100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
+101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
+110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
+111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
+\hline
+\end {tabular}
+}
+\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
+\end{table}
+Now our objective is to abstract a function over GF $\F_{2^3}$ about word-level variables, i.e. $Z = \Func(A)$.
+Recall the definition of Lagrange's interpolation:
+\begin{equation}
+\label{eqn:Lagrange}
+\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
+\end{equation}
+The geometry meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
+they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
+written in the form of equation \ref{eqn:Lagrange}. In this example, although defined in Galois field instead of 
+real number field, the essential concept of Lagrange's interpolation keeps the same. We can get 8 ``dots":
+\begin{align*}
+Uni-form&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
+Dot~1&: (0, 0) \gets (000,000) \\
+Dot~2&: (1,1) \gets (001,001) \\
+Dot~3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
+Dot~4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
+Dot~5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
+Dot~6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
+Dot~7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
+Dot~8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
+\end{align*}
+Substitute 8 $(x_i,y_i)$ pairs in equation \ref{eqn:Lagrange} with these 8 ``dots" in $\F_{2^3}$.
+The result is a polynomial function with degree no greater than 7:
+\begin{align*}
+Z =& \Func(A) \\
+=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
+&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
+\end{align*}
+
+\end{Example}
+
+The Lagrange's interpolation theorem also proves the existence of a word-level abstraction for a bit-level netlist.
+In practice, Lagrange's interpolation is too weak on scalability. Our approach uses abstraction based on 
+Gr\"obner basis with abstraction term order (ATO), which is briefly introduced in preliminaries.
 
@@
 
+\subsection{Conventional Traversal Method}
 Conceptually, the state-space of a FSM is traversed in a breadth-first
@@
 Let us describe the application of the algorithm on the  FSM circuit
-of Fig. \ref{fig:fsm}. {\it We will first describe its operation at the
+of figure \ref{fig:fsm}. {\it We will first describe its operation at the
 Boolean level, and then describe how this algorithm can be implemented
@@
 In Line 1 of the BFS algorithm, assume that the initial state
-is $S_3$ in Fig.\ref{fig:fsm}(b), which is encoded as 
+is $S_3$ in figure \ref{fig:fsm}(b), which is encoded as 
 $S_3 = \{11\}$. Using Boolean variables $s_0, s_1$ for the present
@@
 \begin{Example}
-For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
+For the circuit in figure \ref{fig:fsm} (b), we have the transition
 functions of the machine as:
@@
 
-{\bf FSM Traversal at word-level over $\Fkk$:} 
-The state transition graph (STG) shown in Fig.\ref{fig:fsm}(a) uses a
+\subsection{FSM Traversal at word-level over $\Fkk$} 
+The state transition graph (STG) shown in figure \ref{fig:fsm}(a) uses a
 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
@@
 
-{\it Initial state:} In Line 1 of Alg.\ref{alg:BFS}, the initial state is
+{\it Initial state:} In Line 1 of algorithm \ref{alg:BFS}, the initial state is
 specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
@@
 {\bf Set operations:} In Lines 5 and 6 of
-Alg. \ref{alg:BFS}, we need  \textbf{union}, 
+algorithm  \ref{alg:BFS}, we need  \textbf{union}, 
 \textbf{intersection} and \textbf{complement} of varieties over
@@
 
-In Line 5 of Alg. \ref{alg:BFS}, we need to compute the complement of a
+In Line 5 of algorithm  \ref{alg:BFS}, we need to compute the complement of a
 set of states. Assume that $J$ denotes a polynomial ideal whose
@@
 
-\begin{proof}
+\begin{Proof}
 Let $d = gcd(f, g)$. So, $f = df_1 , g = dg_1$ with $gcd(f_1 , g_1 ) =
@@
 \langle f \rangle$. Therefore, $h \in \langle f : g\rangle$.
-\end{proof}
+\end{Proof}
 
@@
 
-\begin{proof}
+\begin{Proof}
 Since $\Fkk[x]$ is a principal ideal domain, $ J = \langle g\rangle$
@@
 So, $x \notin V(g)$ and this concludes the proof.
-\end{proof}
+\end{Proof}
 
@@
 concept of word-level FSM traversal over $\Fkk$ using algebraic
-geometry. The algorithm is given in Alg. \ref{alg:univa}. Note that in
+geometry. The algorithm is given in algorithm  \ref{alg:univa}. Note that in
 the algorithm, $from^i, to^i, new^i$ are {\it univariate polynomials in variables $S$
@@
 variables are abstracted and quantified away.
-
-%% Assume $I = \langle f\rangle  = \langle T^2 +
-%% (1+\alpha)\cdot T+\alpha\rangle $,  we can add vanishing polynomial
-%% $T^4 - T$ to this ideal such that its variety  $V(I) = \{a\ |\ a \in
-%% \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$. For complement set
-%% of variety $\{1, \alpha\}$, the universal set is the variety of ideal
-%% of vanishing polynomial $V(\langle T^4-T\rangle ) =
-%% \{0,1,\alpha,1+\alpha\}$, 
-%% so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
-%% which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
-%% the result is $\{0,1+\alpha\}$.
 
@@
 
+\subsection{Word-level FSM Traversal Example}
+
 \begin{Example}
@@
 We apply Algorithm \ref{alg:univa} to the example shown in
-Fig. \ref{fig:fsm} to execute the FSM traversal. Let the initial state
-$from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F _4$. Polynomially, it is
+figure \ref{fig:fsm} to execute the FSM traversal. Let the initial state
+$from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F_4$. Polynomially, it is
 written as $from^0 = S - 0$. In the first iteration, we compose
@@
 polynomial in $T$, quantifying away the rest of the variables.
+
+\section{Improving our Approach}
+\label{sec:improve}
+% TODO Revised according to comments in next chapter
+
+Using ATO on computing GB for large set of polynomial is time-consuming, and usually 
+intractable. The reason is the exponential computational complexity \cite{gao:gf-gb-ms}:
+\begin{Theorem}
+Let $J+J_0 = \langle f_1, \dots, f_s, ~x_1^q - x_1, \dots, x_d^q -
+x_d\rangle \subset \Fq [x_1, \dots, x_d]$ be an ideal. The time and
+space complexity of Buchberger's algorithm to compute a \Grobner
+basis of $J+J_0$ is bounded by $q^{O(d)}$.
+\end{Theorem}
+
+In our case $q = 2^k$, and when $k$ and $d$ are large, this complexity 
+makes abstraction infeasible.
+
+Buchberger's algorithm consists of two major operations: one is finding 
+S-poly pairs, the other is dividing the S-poly by the set of polynomials.
+Its actual cost is very sensitive to the term ordering: if there exists 
+a term order making most S-poly computation unnecessary, then the cost to 
+compute S-poly is saved. Moreover, it prevents new polynomials generating 
+from unnecessary S-poly reductions, which further reduces the cost
+because there are less S-poly pairs as well as candidate divisors.
+Besides tuning the term ordering, directly cutting down the number of 
+polynomials in the set is also effective.
+
+In order to make our approach scalable, we propose improvements from 
+two aspects: on the one hand, using another term ordering which can lower down the 
+computational complexity of obtaining GB; on the other hand, reducing the 
+polynomials by collecting bit-level primary inputs (PIs) and integrating them as word-level 
+variables which are compatible with our working GF.
+\subsection{Simplifying the Gr\"obner Basis Computation}
+In algorithm \ref{alg:univa}, a Gr\"obner basis is computed for each
+iteration to attain the word-level polynomial representation of the next states. In practice, for 
+a sequential circuit with complicated structure and large size, Gr\"obner basis computation
+is intractable. To overcome the high computational complexity of computing a GB, 
+we describe a method that computes a GB of a smaller subset of polynomials.
+The approach draws inspirations from \cite{pruss:tcad}. 
+According to proposition 2 in \cite{pruss:tcad}, if the GB is computed using {\it refined abstraction term
+order} (RATO), there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
+their leading monomials are not relatively prime, i.e. 
+$$gcd(LM(f_w), LM(f_g)) \neq 1$$
+As a well-known fact from Buchberger's algorithm, the S-polynomial 
+(Spoly) pairs with relatively prime 
+leading monomials will always reduce to 0 modulo the basis and have no contribution to 
+the Gr\"obner basis computation.
+Therefore, by removing the relatively prime polynomials from $J_{ckt}$, 
+the Gr\"obner basis computation is transformed to the reduction of 
+$Spoly(f_w,f_g)$ modulo $J_{ckt}$. More specifically, we turn the
+GB computation into one-step multivariate polynomial division, and the obtained
+remainder $r$ will only contain bit-level inputs and word-level output.
+
+\begin{Example}
+After adding intermediate bit-level signal $a,b,c,d$, the elimination ideal for example circuit 
+(example \ref{ex:SMPO}) can be rewritten LEX order with RATO:
+\begin{align}
+&(t_0,t_1)>(a,b,c,d)\nonumber\\&>T>(x,s_0,s_1)\nonumber
+\end{align}
+We can write down all polynomial generators of $J_{ckt}$:
+\begin{equation*}
+f_1: a+xs_0s_1+xs_0+xs_1+x+s_0s_1+s_0+s_1+1
+\end{equation*}
+\vspace{-0.8cm}
+\begin{alignat*}{3}
+& f_2: b+s_0s_1 &&~~~ f_3: c+x+xs_0 \\
+& f_4: d+s_0s_1+s_0 &&~~~ f_5: t_0+ab+a+1 \\
+& f_6: t_1+cd+c+d &&~~~ f_7: t_0+t_1\alpha+T 
+\end{alignat*}
+
+From observation, the only pair which is not relatively prime is
+$(f_5,f_7)$, thus the critical candidate polynomial pair is
+$(f_w,f_g)$, where $$f_w = t_0+a\cdot b+a+b, ~~f_g =t_0+t_1\alpha + T$$
+Result after reduction is:
+\begin{align}
+&Spoly(f_w,f_g) \xrightarrow{J+J_0}_{+}T + s_0 s_1 x+\alpha s_0 s_1 \nonumber\\
+&+(1+\alpha)s_0 x+(1+\alpha) s_0+s_1 x+s_1+(1+\alpha) x+1\nonumber
+\end{align}
+The remainder contains only bit-level inputs $(x,s_0,s_1)$ and word-level output $T$.
+\end{Example}
+
+The remainder from \emph{Spoly} reduction contains bit-level PS variables, and our objective is to get a polynomial 
+containing only word-level PS variables. One possible method is 
+to rewrite bit-level variables in term of word-level variables, i.e.
+\begin{equation}
+\label{eqn:BLVS_reacha}
+s_i = \mathcal{G}(S)
+\end{equation}
+
+Then we can substitute all bit-level variables with the word-level variable
+and obtain a word-level expression.
+The authors of \cite{pruss:tcad} propose a method to construct 
+a system of equations and solution to the system consists of equation \ref{eqn:BLVS_reacha}.
+One way to obtain  is to apply a Gaussian-elimination-fashion approach, which could 
+compute corresponding $\mathcal{G}(S)$ efficiently with time complexity $O(k^3)$.
+
+\begin{Example}
+{\bf Objective}:\ Abstract polynomial $s_i + \mathcal{G}_i(S)$ from $f_0: s_0+s_1\alpha+S$.
+
+First, compute $f_0^2: s_0+s_1\alpha^2+S^2$. Apparently variable $s_0$ can be
+eliminated by operation 
+\begin{align}
+f_1 =& f_0 + f_0^2 \nonumber\\
+=&(\alpha^2+\alpha)s_1+S^2+S\nonumber
+\end{align}
+Now we can solve univariate polynomial equation $f_1 = 0$ and get solution
+$$s_1 = S^2 + S$$
+Using this solution we can easily solve equation $f_0 = 0$. The result is
+$$s_0 = \alpha S^2+(1+\alpha)S$$
+\end{Example}
+
+More formally, polynomial expressions for $s_i$ in terms of $S$ can be
+obtained by setting up and solving the following system of equations:
+\begin{align}
+\label{eqn:betamat}
+\begin{bmatrix}
+S \\
+S^2 \\
+S^{2^2} \\
+\vdots \\
+S^{2^{k-1}}
+\end{bmatrix}
+&=
+\begin{bmatrix}
+1 & \alpha & \alpha^{2} & \cdots & \alpha^{{k-1}}\\
+1 & \alpha^{2} & \alpha^{4} & \cdots & \alpha^{2(k-1)} \\
+1 & \alpha^{4} & \alpha^{8} & \cdots & \alpha^{4(k-1)}\\
+\vdots & \vdots & \vdots & \ddots & \vdots \\
+1 & \alpha^{2^{k-1}} & \alpha^{2\cdot 2^{k-1}} & \cdots & \alpha^{(k-1)\cdot 2^{k-1}}
+\end{bmatrix}
+\begin{bmatrix}
+s_0\\
+s_1\\
+s_2\\
+\vdots\\
+s_{k-1}
+\end{bmatrix}
+\end{align}
+Treat ${\bf s}$ as a vector of $k$ unknowns $s_0,\dots,s_{k-1}$,
+then equation \ref{eqn:betamat} can be solved by using Cramer's rule or Gaussian elimination.
+In other words, we can obtain $s_i=\mathcal G(S)$ by solving equation \ref{eqn:betamat} symbolically.
+
+In this approach we get word-level variable representation for each bit-level PS variables. 
+By substitution, a new polynomial in word-level PS/NS variables could be obtained.
+
+After processing with RATO and bit-to-word conversions, we get a polynomial in the 
+form of $f_T = T+\mathcal{F}(S,x)$
+denoting the {\bf transition function}. We include a polynomial in $S$ to define the present states $f_S$,
+as well as the set of vanishing polynomials for primary inputs 
+$J_0^{PI} = \langle x_1^2-x_1,\dots,x_n^2-x_n\rangle$. Using elimination term order with $S>x_i>T$,
+we can compute a GB of the elimination ideal $\langle f_T,f_S\rangle + J_0^{PI}$. This GB contains a univariate
+polynomial denoting next states. The improved algorithm is depicted in Alg. \ref{alg:refined}.
+
+\IncMargin{1em}
+\begin{algorithm}[hbt]
+\SetAlgoNoLine
+\Indm
+ \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\F(S)$}
+\Indp
+
+  $from^0 = reached = \F(S)$\;
+  $f_T = $Reduce($Spoly(f_w,f_g), J_{ckt})$\;
+  \tcc{Compute S-poly for the critical pair, then reduce it with circuit ideal under RATO}
+	Eliminate bit-level variables in $f_T$\;
+  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
+  {
+  	$i \gets i + 1$\;
+  	$G \gets$GB($\langle f_T , from^{i-1}\rangle+J_0^{PI}$)\;
+  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest; $J_0^{PI}$ covers all possible inputs from PIs}
+	$to^i \gets G\cup \Fkk[T]$\;
+	\tcc{There will be a univariate polynomial in $G$ denoting next state in word-level variable $T$}
+	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
+	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
+  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
+  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
+	$from^i \gets new^i(S\setminus T)$\;
+	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
+  }
+  \tcc{Loop until a fixpoint reached: newly reached state is empty}
+\Return{$\langle reached\rangle$}
+\caption {Refined Algebraic Geometry based FSM Traversal}\label{alg:refined}
+\end{algorithm}
+\DecMargin{1em}
+
+\subsection{PI Partition}
+Using above techniques we can get a remainder polynomial with only word-level PS/NS variables. However in most 
+cases, the number of bit-level PIs will be too large for the last-step Gr\"obner basis computation. 
+Therefore it is necessary to convert bit-level PIs to word-level PI variables.
+
+Assume we have $k$-bit datapath and $n$-bit PIs. In Galois field, we need to carefully partition $n$ PIs
+such that states of each partition can be covered by a univariate polynomial respectively.
+
+\begin{Proposition}
+Divide $n$-bit PIs to partitions $n_1,n_2,\dots, n_s$ and let $n_1$-bit, $n_2$-bit, $\dots$, $n_s$-bit word-level variables
+represent their evaluations in $\mathbb F_{2^k}$ respectively. Then $n_1,n_2,\dots,n_s~|~k$.
+\end{Proposition}
+
+Again, assume a partition $n_i~|~k$ and corresponding word-level variable is $P$. Then we can use polynomial
+$P^{2^{n_i}}-P$ to represent all signals at free-end PIs, according to following theorem about {\bf composite
+field}:
+\begin{Theorem}
+\label{thm:composite}
+Let $k = m\cdot n_i$, such that $\mathbb F_{2^k} = \mathbb F_{(2^{n_i})^m}$. Let $\alpha$ be primitive root of 
+$\mathbb F_{2^k}$, $\beta$ be primitive root of the ground field $\mathbb F_{2^{n_i}}$. Then
+$$\beta = \alpha^\omega,~~\text{where}~~\omega = \frac{2^k-1}{2^{n_i}-1}$$
+\end{Theorem}
+\begin{Example}
+\label{ex:PI}
+In a sequential circuit, PS/NS inputs/outputs are 4-bit signals, which means we will use $\mathbb F_{2^4}$
+as working field. PIs are partitioned to 2-bit vectors, which means the ground field is $\mathbb F_{2^2}$.
+In ground field we can represent all possible evaluations of this PI partition $\{p_0,p_1\}$ with
+$$P^4+P,~~\text{where}~~P=p_0+p_1\cdot\beta$$
+Using theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
+$$P = p_0 + p_1\cdot\alpha^5$$
+\end{Example}
+Using this method we can efficiently partition large size PIs to small number of word-level PI variables.
+One limitation of this approach is PIs cannot be partitioned when $k$ is prime.
+
+\section{Implementation of Word-level FSM Traversal Algorithm}
+In this section, we will introduce the tool implementation of our approach.
+The architecture of our tool consists of 4 functional components.
+First, given circuit is synthesized to gate-level using state-of-art synthesizers.
+Secondly, the synthesized netlist is translated to polynomial form and variables are sorted in 
+RATO. This part is implemented using scripting language such as \emph{Perl}. Thirdly, 
+the polynomial reduction is executed using our customized reduction engine, which is dumped in C++.
+Finally, we utilize symbolic computation engine -- \textsc{Singular} \cite{DGPS} to code algorithm 
+\ref{alg:refined} and execute the BFS traversal. The result is given in a univariate polynomial 
+about NS word $T$, denoting the set of reachable states from given initial state.
+
+The following example illustrates the full-blown reachability analysis approach
+based on C++ and \textsc{Singular} implementation.
+
+\begin{Example}[Full Blown Reachability Analysis Walk-through]
+Usually FSM designs can be described in behavior/structural hybrid languages.
+One of these languages is Berkeley logic interchange format (BLIF) \cite{BLIF},
+it allows state behavior representation and logic component representation.
+
+In this example, we use benchmark FSM ``lion9" from MCNC benchmark library.
+Figure \ref{fig:BLIF} is the truth table based structural BLIF representation of this FSM.
+
+\begin{figure}[hbt]
+\centering{
+\includegraphics[width=\textwidth]{newfig/BLIF.pdf}
+\caption{Sample input BLIF file}
+\label{fig:BLIF}}
+\end{figure}
+
+In order to compose the polynomial set in elimination ideal, we need to synthesize it to 
+gate-level netlist. Modern synthesizers including ABC \cite{brayton2010abc} and SIS \cite{SIS} 
+can complete this work. Figure \ref{fig:BLIF_syn} shows a synthesized FSM given in 
+BLIF format. Note that for simplifying purposes in this example, we only use 2-input logic gates.
+
+\begin{figure}[hbt]
+\centering{
+\includegraphics[width=\textwidth]{newfig/BLIF_syn.pdf}
+\caption{Synthesized BLIF file}
+\label{fig:BLIF_syn}}
+\end{figure}
+
+Using our interpreter, the synthesized BLIF file turns to a polynomial file customized for our 
+polynomial reduction engine as figure \ref{fig:ALG} depicts. Note the first line includes all variables in RATO, the second line 
+denotes the word-level variable. The third line is the irreducible polynomial of current GF $(\F_{2^4})$, the fourth line is the Spoly need to be reduced. The rest part has one polynomial in 
+the elimination ideal in each line.
+
+\begin{figure}[hbt]
+\centering{
+\includegraphics[width=\textwidth]{newfig/ALG.pdf}
+\caption{Polynomial file prepared for polynomial reduction engine}
+\label{fig:ALG}}
+\end{figure}
+
+The result given by our reduction engine is in the format 
+$$T+\Func(s_0,\dots,s_{k-1},x_0,\dots,x_{n-1})$$ 
+where $s_i$ and $x_j$ denotes bit-level PS variables and PIs. Concretely in this example, the result is
+\begin{align*}
+T&+(\alpha^2+\alpha+1) x_0 x_1 s_1 s_3+(\alpha^3+\alpha) x_0 x_1 s_1+(\alpha+1) x_0 x_1 s_3 \\
+&+(\alpha^3+1) x_0 s_1 s_3+(\alpha^3+\alpha) x_0 s_1 +(\alpha+1) x_0 s_3+(\alpha^2) x_0 \\
+&+(\alpha^2+\alpha+1) x_1 s_1 s_3+(\alpha^3+\alpha) x_1 s_1+(\alpha^2+\alpha+1) x_1 s_3\\
+&+(\alpha) x_1+(\alpha^3+1) s_1 s_3+(\alpha^3+\alpha) s_1+(\alpha^3+1) s_3+\alpha^2
+\end{align*}
+This is the transition function of this FSM. We utilized \textsc{Singular} to integrat both 
+bit-word substitution and tranversal algorithm. In figure \ref{fig:SING}, ``tran" is the transition
+function we just obtained. ``init\_S" is the initial state, note it equals to ``0100" register 
+reset values in figure \ref{fig:BLIF}. Moreover, 2 bits PIs $x_0,x_1$ are combined to a word-level 
+PI variable $P$ using our conclusion in example \ref{ex:PI}: ``def\_X" is the definition of 2-bit 
+word, and ``red\_X" denotes the vanishing polynomial for word $P$.
+
+\begin{figure}[hbt]
+\centering{
+\includegraphics[width=\textwidth]{newfig/SING.pdf}
+\caption{Singular script for executing bit-to-word substitution and traversal loop}
+\label{fig:SING}}
+\end{figure}
+
+After the script is executed, the trversal finishes after 4 
+transition iterations, which denotes BFS depth equals to 4. The final reachable states 
+is a degree-9 polynomial about $T$, indicating final reachable states set contains 9 states.
+And state encodings can be obtained by solving this polynomial equation.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/Recha_result.pdf}
+\caption{The output given by our traversal tool}
+\label{fig:Recha_result}}
+\end{figure}
+
+\end{Example}
+
+\section{Experiment Results}
+\label{sec:exp_reacha}
+We have implemented our traversal algorithm in 3 parts:
+the first part implements polynomial reductions (division) of the
+Gr\"obner basis computations, under the term order derived from the
+circuit as Line 2 in Alg. \ref{alg:refined}. This is implemented 
+with our customized data structure in
+C++. The second part implements the bit-level to word-level
+abstraction to attain transition functions at the word-level 
+%$T+\mathcal{F}(S,PIs)$ 
+using the \textsc{Singular} symbolic
+algebra computation system [v. 3-1-6] \cite{DGPS}, as Line 3 in 
+Alg. \ref{alg:refined}; and the third part
+executes the reachability checking iterations using 
+\textsc{Singular} as well. With our
+tool implementation, we have performed experiments to analyze reachability
+of several FSMs. Our experiments run on a desktop with
+3.5GHz Intel $\text{Core}^\text{TM}$ i7-4770K Quad-core CPU, 32 GB RAM and
+64-bit Ubuntu Linux OS. The experiments are shown in Table \ref{tab:recha_result}.
+
+There are 2 bottlenecks which restricts the performance of our tool:
+one bottleneck is that the polynomial reduction engine is slow when the number of
+gates (especially OR gates) is large; the other one is the high computational complexity
+of Gr\"obner basis engine in general. Therefore, we pick 10 FSM
+benchmarks of reasonable size for testing our tool. Among them ``b01,
+b02, b06" come from ITC'99 benchmarks, %\cite{ITC99}, 
+``s27, s208, s386" are from  ISCAS'89 benchmarks %\cite{yang1991logic}
+ and ``bbara, beecount, dk14, donfile" are from MCNC benchmarks. %\cite{yang1991logic}.
+ISCAS benchmarks are given as {\it bench} format so we can directly read gate information,
+where ITC/MCNC FSMs are given in unsynthesized {\it blif} format so we first turn them
+into gate-level netlists using AIG based synthesizer ABC. % \cite{brayton2010abc}.
+Since the number of primary inputs $(m)$ is relatively small, in our experiments we partition
+primary inputs as $m$ single bit-level variables. To verify the
+correctness of our techniques and implementations, we compare the
+number of reachable states obtained from our tool against the results
+obtained from the VIS tool \cite{brayton1996vis}.
+
+\begin{table}[H]
+\centering
+\caption{Results of running benchmarks using our tool. 
+\small{Parts I to III denote the time taken by polynomial divisions,
+  bit-level to word-level abstraction and iterative reachability
+  convergence checking part of our approach, respectively.}}
+{\small 
+\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|}
+\hline
+\multirow{3}{*}{\centering Benchmark} 
+& \multirow{3}{0.9cm}{\centering \# Gates} 
+& \multirow{3}{1.1cm}{\centering \# Latches} 
+& \multirow{3}{*}{\centering \# PIs}
+ & \multirow{3}{0.9cm}{\centering \# States}
+ & \multirow{3}{1.3cm}{\centering \# iterations}
+ & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
+ & \multirow{3}{1.8cm}{\centering Runtime of VIS (sec)} \\
+  & & & & & &\multicolumn{3}{c|}{}& \\
+  \cline{7-9}
+    & & & & & & I & II & III & \\
+\hline
+\hline
+b01 & 39  & 5  & 2 & 18  & 5  & $<0.01$ & 0.01 & 0.02 & $<0.01$\\
+b02 & 24  & 4  & 1 & 8 & 5 & $<0.01$  & 0.01 & $<0.01$ & $<0.01$ \\
+b06 & 49  & 9  & 2 & 13 & 4 & $<0.01$ & 0.07 & 5.0 & $<0.01$ \\
+s27 & 10 & 3 & 4 & 6 & 2 & $<0.01$ & 0.01 & 0.02  & $<0.01$  \\
+s208 & 61 & 8 & 11 & 16 & 16 & $<0.01$ & 0.32 & 2.4 & $<0.01$ \\
+s386 & 118 & 6 & 13 & 13  & 3 & 1.0 & 7.6 & 8.2  & $<0.01$ \\
+bbara & 82 & 4 & 4 & 10 & 6 & 0.04 & 0.01 & 0.04  & $<0.01$ \\
+beecount & 48  & 3  & 3 & 7  & 3  &$<0.01$ & 0.01 & 0.01 & $<0.01$ \\
+dk14 & 120  & 3  & 3 & 7  & 2  & 45 & $<0.01$ & 0.08 & $<0.01$\\
+donfile & 205  & 5  & 2 & 24 & 3  & 12316 & 0.02 & 1.7  & $<0.01$\\
+\hline
+\end{tabular}
+}
+\label{tab:recha_result}  
+\end{table}
+
+In Table \ref{tab:recha_result}, \# States denotes the final reachable states
+starting from given reset state, which given by our tool is the same with
+the return value of {\it compute\_reach} in VIS. Meanwhile, from observation
+of the experiment run-times, we find the reduction runtime increases
+as the number of gates grows. Also, iterative reachability convergence
+check's runtime reflects both the size of present state/next state
+words ($k$) and the number of final reached states, which corresponds
+to the degree of polynomial {\it reached} in Alg.\ref{alg:univa}. 
+Although the efficiency of our initial implementation fails to compete
+with the BDD based FSM analyzer VIS, the experiment
+demonstrates the power of abstraction of algebraic geometry techniques
+for reachability analysis applications. Currently, we are
+investigating techniques that
+can help us overcome the complexity of the GB computation with elimination
+orders and speed-up our approach.
+
+%\vspace{-0.1in}
+
+\section{Conclusion} \label{sec:conc}
+This paper has presented a new approach to perform reachability
+analysis of finite state machines at the word-level. This is achieved
+by modeling the transition relations and sets of states by way of
+polynomials over finite fields $\Fkk$, where $k$ represents the size
+of the state register bits. Subsequently using the concepts of
+elimination ideals, Gr\"obner bases, and quotients of ideals, we show
+that the set of reachable states can be encoded, canonically, as the
+variety of a univariate polynomial. This polynomial is computed using
+the Gr\"obner basis algorithm w.r.t. an elimination term
+order. Experiments are conducted with a few FSMs that validate the
+concept of word-level FSM traversal using algebraic geometry.

### 2016-11-27 dissertation/recha.tex
dissertation v2:part recha

@@
 \label{ch:reacha}
-Reachability analysis is a basic component of sequential circuit verification, esp. for 
+Reachability analysis is a basic component of sequential circuit verification, especially for 
 formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
 tools, in order to improve various performance indicators such as latency, clock skew or power
-density, some major refactoring and attachments are made on original designs. 
-Those modifications may introduce malfunctions or glitches to the whole logic. 
-In localized simulation or formal verification (esp. equivalence checking), the modifications may be 
-denied since the malfunctions or glitches are considered as ``faults" in this circuit.
-However, if the circuit behavior is carefully investigated, it may come to a verdict that 
+density, sequential optimization techniques such as retiming \cite{retiming}, scan logic \cite{scan},
+sequential power optimization \cite{mathur2009power} and clock-gating techniques \cite{clockgating}.
+% some major refactoring and attachments are made on original designs. 
+These modifications may introduce malfunctions to the original logic and cause problems. 
+Based on traditional localized simulation or formal verification method (e.g. equivalence checking), 
+designers are reluctant to make aggressive optimization
+since the malfunctions are considered as ``faults" in this circuit.
+However, if the circuit behavior is carefully investigated, it may become evident that 
 those ``faults" will never be activated/excitated during a restricted execution starting 
 from legal initial states and with legal inputs. Thus we will call those ``faults" as 
-{\bf spurious faults}, since they will not affect the circuit's normal behavior.
+{\bf spurious faults} (false negatives), since they will not affect the circuit's normal behavior.
 
 Almost all practical sequential logic components can be modeled as finite state machines (FSMs). 
-If we apply constrains upon the machine to make it start from designated initial states and 
-take in specific legal inputs, a set of reachable states can be derived. 
+If we apply constraints upon the machine to make it start from designated initial states, and 
+take specific legal inputs, a set of reachable states can be derived. 
 As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
-spurious faults by checking if they sit in the reachable states. From the spurious fault validation 
+spurious faults by checking if they belong to the reachable states. From the spurious fault validation 
 perspective, reachability analysis is a must when developing full set of sequential circuit verification
@@
 binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
-which bring critical advantages. In this chapter the approach will be described and discussed in depth, 
+which brings critical advantages. In this chapter the approach will be described and discussed in depth, 
 with examples and experiments showing its feasibility when applied on general circuit benchmarks.
@@
 \section{Motivation}
-The motivation of this part of my research consists of many inspirations and observations.
-In this dissertation, we will summarize those major ones into 3 topics. First, we start from 
-a sequential circuit partial scan paper to discuss the inspirations of our work as well as some 
-difficulties to overcome. Secondly, we state the observation that word-level description is 
-increasingly important and common in characterizing the data flow on modern large-size datapaths. 
-Last but not least, we give instances that some prerequisites for supporting word-level techniques 
-are already available. Thus we cover both the necessity and sufficiency of developing such 
-a word-level traversal technique in our work.
-
-\subsection{Limitations of Full Scan Algorithms}
-The inspiration of this research mainly comes from a journal paper \cite{KallaPartialScan}. 
-In that paper, the author proposed an traversal algorithm using concept ``implicit state enumeration".
+Before introducing the details of our approach, there are a couple of questions to ask: why is it
+necessary to execute word-level FSM traversal? Why does algebraic geometry make this happen?
+The answers can be found in this section, as statement of the motivation of our research.
+
+\subsection{FSM Traversal Algorithms}
+Sequential circuits are modeled as FSMs, which can be implemented as graph. Thus a 
+graph-traversal based algorithm is created to analysis the reachable states \cite{coudert2003unified}.
+A traversal algorithm using concept ``implicit state enumeration" is proposed \cite{cho1993redundancy}.
 Concretely, the algorithm is written as follows:
@@
 \LinesNumbered
- \KwIn{Sequential circuit, number of registers to scan}
+ \KwIn{Sequential circuit with given number of registers}
  \KwOut{Scan registers listed in decreasing order of their non-controllability}
@@
   }\
-  \tcc{\ \ \ \ \ \ \ \ \ \ \ \ \ \ FSM traversal completed}\
-  \For{each state variable $r_j$}
-  {
-	\If{missing transition for $r_j$}
-	{
-		scan state variable $r_j$\;
-	}
-  }
-  $illegal\_states = $bdd\_complement$(reached)$\;
-  \For{each state variable $r_j$}
-  {
-	compute\_degree\_of\_unateness$(r_j,illegal\_states)$\;
-	non-controllability$(r_j)=$degree\_of\_unsettability$(r_j)+$degree\_of\_unateness$(r_j)$\;
-  }
-  order state variables in terms of their non-controllabilities;\tcc{\ Sorting}\
-  output the required scan registers\;
-%\Return{$from_k(R_{final})$}
-\caption {SIMPSON: Scan Register Selection using Implicit State Enumeration\cite{KallaPartialScan}}
+\Return{$reached$}
+\caption {FSM Traversal using Implicit State Enumeration\cite{KallaPartialScan}}
 \label{alg:SIMPSON}
@@
 
-Note that the former part of this algorithm describes a breath-first-search (BFS) traversal in
-state space. The traversal algorithm is a simple variation of BFS algorithm in graph theory where 
+Above algorithm describes a breath-first-search (BFS) traversal in
+state space. The traversal algorithm is a simple variation of BFS algorithm where 
 states are nodes and transitions are arcs. Each state is uniquely encoded by a combination
@@
 
-Since a typical sequential circuit usually contains an functionally independent combinational logic
-component, the traversal algorithm analyzes the combinational logic and derive the transition 
-function for one-step reachability within current time-frame, and extend the result to complete 
+Since a typical sequential circuit usually contains a combinational logic
+component, the traversal algorithm analyzes the combinational logic and derives the transition 
+function for one-step reachability within current time-frame, and extends the result to complete 
 execution paths through unrolling. If each state encoding (i.e. exact values in the selected registers) is
-explicitly exposed and counted during the unrolling 
-procedure, this unrolling is called {\bf explicit unrolling}. In the SIMPSON algorithm, the states cannot be 
+explicitly analyzed and counted during the unrolling 
+procedure, this unrolling is called {\bf explicit unrolling}. In the BFS traversal algorithm, the states cannot be 
 directly read in the execution; instead, they are implicitly represented using a conjunction of several 
@@
 
-However, the SIMPSON algorithm proposed by the author is usually not practical. The conjunctions of 
+However, the BFS traversal algorithm proposed by the author is usually not practical. The conjunctions of 
 Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
-the size of BDD is easy to explode when the formulas become too long and too complicated. In the paper,
-the author make a compromise between accuracy and cost, and turn to a partial scan technique rather than 
-a full scan. In my research work, the aim is to explore a word-level technique which can make a full scan 
-available.
+the size of BDD explodes when the formulas become too long and too complicated. In \cite{cho1993redundancy},
+the authors make a compromise between accuracy and cost, and turn to approximate reachability. 
+In this research work, the aim is to explore a word-level technique which can make a accurate reachability
+analysis available.
 
 \subsection{Word-level Data Flow on Modern Datapath Designs}
-The integrity of modern digital circuit designs is very high. For example, processor A10 designed by Apple
-integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such high integrity make the silicon 
-implementation of large datapaths possible. In recent decades, leading by top-notch central processing units (CPUs)
-and high bandwidth memory (HBM), 64-bit or even larger datapaths frequently appear in digital ICs.
-Meanwhile, with the development of electrical design automation (EDA) tools, data flow on those datapaths
-are directly described by the designer as word-level specifications. Therefore, it will be straightforward 
+The level of integration of modern digital circuit designs is very high. For example, processor A10 designed by Apple
+integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such a high density makes the silicon 
+implementation of large datapaths possible. In recent decades, 64-bit or even larger datapaths frequently appear in 
+modern digital IC designs such as powerful central processing units (CPUs) and high bandwidth memory (HBM).
+Meanwhile, with the development of electronic design automation (EDA) tools, data flow 
+is described by the designer as word-level specifications. Therefore, it will be straightforward 
 and beneficial for users if formal verification tools can work on word-level. Moreover, adopting word-level 
-techniques will greatly reduce the state space and make verification executes faster by orders of magnitude.
+techniques will greatly reduce the state space and make verification more efficient.
 
@@
 
-\begin{Example}[Verification of sequential multiplication hardware]
-
-% fig:MIPS
+\begin{Example}
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/MIPS.pdf}
+\caption{A 64-bit sequential multiplication hardware design}
+\label{fig:MIPS}}
+\end{figure}
 
@@
 Iteratively, the last significant bit (LSB) of current (temporary) product is used as flag to activate the ALU 
-to add on the other multiplicand. For each iteration the data in product registers shift right by 1 bit.
-Finally when the most significant bit (MSB) of preloaded multiplicand arrives the MSB of product registers,
-the registers contains the result -- 128 bits product. The behavior can be concluded as following algorithm:
-
-\vspace{0.2in}
+to add on the other multiplicand. For each iteration the data in product registers shifts right by 1 bit.
+Finally when the most significant bit (MSB) of preloaded multiplicand arrives at the MSB of product registers,
+the registers contains the result -- 128 bits product. The behavior can be described by the following algorithm:
+
+% \vspace{0.2in}
 
@@
 \end{equation}
-The state space will be greatly reduced.
+
+Since the abstraction to word-level reduces symbolic storage and execution cost, the complexity 
+to traverse the state space is greatly reduced.
 \end{Example}
 
-On the other hand, the user of conventional model checker will have to write all these clauses.
-These clauses contain cross-literals, e.g. $s_2$ may associate with $a_1, p_0$, etc. If the user 
-is not familiar with the implementation of this adder, those cross-literals will bring inconveniences
-and confusions. However, if word-level techniques allow specification like equation \ref{eqn:wordspec},
-the verification tool will be very user-friendly and straightforward.
-
-\subsection{Prerequisites of Word-level Techniques}
+On the other hand, when implementation details of the datapath are not available, it is not convenient 
+any more for users of conventional model checker. The reason is that the user has to write all clauses 
+for the implementation, which contains cross-literals, e.g. $s_2$ may associate with $a_1, p_0$, etc. If the user 
+is not familiar with the implementation of this adder, those cross-literals will bring
+confusions. However, if word-level techniques allow specification like Equation \ref{eqn:wordspec},
+the verification tool will be very user-friendly and straightforward even if the implementation details
+are in a black box.
+
+\subsection{On the Existence of Word-level Abstraction}
 When given a bit-level netlist, the prerequisites to use word-level techniques are to convert bit-level 
@@
 
-An old but universally effective abstraction method is {\bf Lagrange's interpolation}. Instead of using 
+An old but universally effective abstraction method is {\it Lagrange's interpolation}. Instead of using 
 it in real algebra, it can also be extended to Galois field. Here we use an example to illustrate the 
@@
 \end{equation}
-The geometry meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
+The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
 they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
-written in the form of equation \ref{eqn:Lagrange}. In this example, although defined in Galois field instead of 
-real number field, the essential concept of Lagrange's interpolation keeps the same. We can get 8 ``dots":
+written in the form of equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
+the real number field, the essential concept of Lagrange's interpolation remains the same. 
+We can get 8 points in the affine space:
 \begin{align*}
-Uni-form&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
-Dot~1&: (0, 0) \gets (000,000) \\
-Dot~2&: (1,1) \gets (001,001) \\
-Dot~3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
-Dot~4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
-Dot~5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
-Dot~6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
-Dot~7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
-Dot~8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
+\text{Generic form}&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
+\text{Point }1&: (0, 0) \gets (000,000) \\
+\text{Point }2&: (1,1) \gets (001,001) \\
+\text{Point }3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
+\text{Point }4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
+\text{Point }5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
+\text{Point }6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
+\text{Point }7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
+\text{Point }8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
 \end{align*}
-Substitute 8 $(x_i,y_i)$ pairs in equation \ref{eqn:Lagrange} with these 8 ``dots" in $\F_{2^3}$.
+Substitute 8 $(x_i,y_i)$ pairs in Equation \ref{eqn:Lagrange} with these 8 points in $\F_{2^3}$.
 The result is a polynomial function with degree no greater than 7:
@@
 The Lagrange's interpolation theorem also proves the existence of a word-level abstraction for a bit-level netlist.
-In practice, Lagrange's interpolation is too weak on scalability. Our approach uses abstraction based on 
-Gr\"obner basis with abstraction term order (ATO), which is briefly introduced in preliminaries.
+In practice, Lagrange's interpolation is not scalable. 
+The reason is that it needs entire function (state space), but usually we only have the circuit representation
+of the FSM. Considering this fact, a symbolic method is needed.
+Our approach in this section uses abstraction based on 
+Gr\"obner basis with abstraction term order (ATO), which is briefly introduced in Section \ref{sec:abstraction}.
+
+\subsection{Significance of Developing Word-level FSM Traversal}
+At the end of this section, we summarize each subsection and conclude the motivation of our research.
+First, we start from the basic FSM traversal algorithms and exhibit their
+difficulties to overcome. Secondly, we state the observation that word-level description is 
+increasingly important and common in characterizing the data flow on modern large-size datapaths. 
+Last but not least, we give instances that some prerequisites for supporting word-level techniques 
+are already available.
+
+To conclude, first we state the importance of performing FSM traversal in sequential circuit reachability analysis.
+To overcome the vast cost brought by searching in a large space, we propose to use word-level 
+polynomials to represent the states and transition relations. As a result, states are categorized into 
+sets represented by a small number of word-level polynomials (more specifically, the varieties to the polynomial ideals),
+and multiple transition relations are therefore merged together. All of these efforts reduces the space cost of 
+state space, meanwhile lower the time complexity to traverse such a state space. By using Lagrange's interpolation,
+we prove the feasibility of word-level techniques on general circuits' verification.
+Thus we cover both the necessity and sufficiency of developing such 
+a word-level traversal technique in our work.
+They are also answers to the questions at the beginning of this section.
 
@@
 computing a GB over $\Fkk$ of the polynomials of the circuit using an
-elimination term order based on Theorem \ref{thm:elim}.
+elimination term order based on Theorem \ref{thm:elimth}.
+
+\subsection{FSM model for sequential circuits}
+A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
+circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
+the formal definition is as follows:
+\begin{Definition}
+A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
+\begin{itemize}
+\item $\Sigma$ is the input label, $O$ is the output label;
+\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
+\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
+\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
+\end{itemize}
+\end{Definition}
+The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
+its primary outputs only depend on the present states, i.e. the output function is defined as
+$$\Lambda:\ S \to O$$
+Typical sequential circuits can be depicted as Figure \ref{fig:seqmodel}(a). Primary inputs
+$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
+are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
+We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
+in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
+at bit level are defined as $\Delta_i$: 
+$$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$
+\begin{figure}[H]
+\centering{
+%\begin{minipage}{12cm}
+\includegraphics[width=5.5in]{newfig/seqmodel.eps}
+% \vspace{-0.2in}
+\caption{FSM models of sequential circuits}
+%\end{minipage}
+\label{fig:seqmodel}}
+\end{figure}
+In some cases, arithmetic computations are implemented as Moore machines where input operands
+are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
+We can simplify them to the model in Figure \ref{fig:seqmodel}(b).
 
@@
 \begin{align}
-\label{img}
+\label{eqn:img}
 to = \text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
@@
 Here, $\exists x (f)$ represents the {\it existential quantification
-  of $f$ w.r.t. variable $x$}.
+  of $f$ w.r.t. variable $x$}. In Boolean logic, this operator is implemented as
+  $$\exists x (f) = f_x\lor f_{\overline{x}}$$
 
@@
 Let us describe the application of the algorithm on the  FSM circuit
-of figure \ref{fig:fsm}. {\it We will first describe its operation at the
+of Figure \ref{fig:fsm}. {\it We will first describe its operation at the
 Boolean level, and then describe how this algorithm can be implemented
@@
 In Line 1 of the BFS algorithm, assume that the initial state
-is $S_3$ in figure \ref{fig:fsm}(b), which is encoded as 
+is $S_3$ in Figure \ref{fig:fsm}(b), which is encoded as 
 $S_3 = \{11\}$. Using Boolean variables $s_0, s_1$ for the present
@@
 \begin{Example}
-For the circuit in figure \ref{fig:fsm} (b), we have the transition
+For the circuit in Figure \ref{fig:fsm} (b), we have the transition
 functions of the machine as:
@@
 
-When the formula of Eqn. (\ref{img}) is applied to compute 1-step
+When the formula of Equation \ref{eqn:img} is applied to compute 1-step
 reachability, $to = \exists _{s_0, s_1, x} (\Delta_1 \cdot \Delta_2
@@
 \subsection{FSM Traversal at word-level over $\Fkk$} 
-The state transition graph (STG) shown in figure \ref{fig:fsm}(a) uses a
+The state transition graph (STG) shown in Figure \ref{fig:fsm}(a) uses a
 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
@@
 
-{\it Initial state:} In Line 1 of algorithm \ref{alg:BFS}, the initial state is
+{\it Initial state:} In Line 1 of Algorithm \ref{alg:BFS}, the initial state is
 specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
@@
 {\bf Set operations:} In Lines 5 and 6 of
-algorithm  \ref{alg:BFS}, we need  \textbf{union}, 
+Algorithm  \ref{alg:BFS}, we need  \textbf{union}, 
 \textbf{intersection} and \textbf{complement} of varieties over
@@
 
-In Line 5 of algorithm  \ref{alg:BFS}, we need to compute the complement of a
+In Line 5 of Algorithm  \ref{alg:BFS}, we need to compute the complement of a
 set of states. Assume that $J$ denotes a polynomial ideal whose
@@
 Take $h \in \langle f : g\rangle$. According to the
-Def. \ref{def:quo}, $hg \in \langle f \rangle$, which means $hg = f
+Definition \ref{def:quo}, $hg \in \langle f \rangle$, which means $hg = f
 \cdot r$ with $r \in \Fkk[x]$. Therefore, $hdg_1 = df_1 r$ and $hg_1 =
@@
 concept of word-level FSM traversal over $\Fkk$ using algebraic
-geometry. The algorithm is given in algorithm  \ref{alg:univa}. Note that in
+geometry. The algorithm is given in Algorithm  \ref{alg:univa}. Note that in
 the algorithm, $from^i, to^i, new^i$ are {\it univariate polynomials in variables $S$
@@
 We apply Algorithm \ref{alg:univa} to the example shown in
-figure \ref{fig:fsm} to execute the FSM traversal. Let the initial state
+Figure \ref{fig:fsm} to execute the FSM traversal. Let the initial state
 $from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F_4$. Polynomially, it is
@@
 
-{\it Significance of using GB:} A reduced GB is a unique, minimal and
-{\it canonical} representation of the circuit's function. Starting
-from a certain initial state and using a reduced GB to represent the
-transition function, reachable states can be computed and represented
-canonically. Then it becomes possible to identify  when a fixpoint
-is reached (termination of the algorithm) by performing an equality
-check of polynomial ideals. Moreover, the GB computation is also used
-as a quantification procedure. As the GB is computed w.r.t. an
-elimination term order with ``bit-level variables'' $>$
-``present-state word'' $S$ $>$ ``next-state word'' $T$, the set of
-reachable states are encoded, canonically, using a univariate
-polynomial in $T$, quantifying away the rest of the variables.
+% {\it Significance of using GB:} A reduced GB is a unique, minimal and
+% {\it canonical} representation of the circuit's function. Starting
+% from a certain initial state and using a reduced GB to represent the
+% transition function, reachable states can be computed and represented
+% canonically. Then it becomes possible to identify  when a fixpoint
+% is reached (termination of the algorithm) by performing an equality
+% check of polynomial ideals. Moreover, the GB computation is also used
+% as a quantification procedure. As the GB is computed w.r.t. an
+% elimination term order with ``bit-level variables'' $>$
+% ``present-state word'' $S$ $>$ ``next-state word'' $T$, the set of
+% reachable states are encoded, canonically, using a univariate
+% polynomial in $T$, quantifying away the rest of the variables.
+
+\subsection{Significance of using Algebraic Geometry and Gr\"obner Basis}
+The essences of out approach consist of utilizations of algebraic geometry and Gr\"obner basis concepts.
+One the one hand, we use GB with elimination term ordering as an analog of image function.
+As mentioned in Section \ref{sec:abstraction}, we construct the elimination ideal $J+J_0$ to 
+describe the circuit in Figure \ref{fig:proj_reacha} using algebraic geometry.
+When given the present states, the next states are implicitly represented in the variety of 
+the elimination ideal. Imposing elimination term order and computing GB, is actually 
+projecting the variety on NS output $T$ by eliminating all other variables.
+This projection gives us the canonical representation of NS in a polynomial with $T$.
+
+% fig:proj_reacha
+
+On the other hand, we use ideal arithmetic to implement set operations in Boolean domain.
+In algebraic geometry, manipulating ideals provides a high-level solution to 
+operate on their varieties without actually solving the system. The intersection, union and 
+complement of varieties are mapped to varieties of the sum, product and quotient of ideals, respectively.
+Furthermore, the varieties are equivalent to set of states because of correspondence of 
+affine space and state space.
+
+As a result, we find an analog of the FSM traversal algorithm in algebraic geometry which is 
+compatible with word-level variables (e.g. $S,T$). We show it is effective to perform the 
+reachability analysis.
 
@@
 
-Using ATO on computing GB for large set of polynomial is time-consuming, and usually 
+Using elimination term ordering on computing GB for large set of polynomial is time-consuming, and usually 
 intractable. The reason is the exponential computational complexity \cite{gao:gf-gb-ms}:
@@
 variables which are compatible with our working GF.
+
 \subsection{Simplifying the Gr\"obner Basis Computation}
-In algorithm \ref{alg:univa}, a Gr\"obner basis is computed for each
+In Algorithm \ref{alg:univa}, a Gr\"obner basis is computed for each
 iteration to attain the word-level polynomial representation of the next states. In practice, for 
@@
 and obtain a word-level expression.
-The authors of \cite{pruss:tcad} propose a method to construct 
-a system of equations and solution to the system consists of equation \ref{eqn:BLVS_reacha}.
-One way to obtain  is to apply a Gaussian-elimination-fashion approach, which could 
+The authors of \cite{pruss:tcad15} propose a method to construct 
+a system of equations and solution to the system consists of Equation \ref{eqn:BLVS_reacha}.
+This can be obtained by Gaussian elimination, which could 
 compute corresponding $\mathcal{G}(S)$ efficiently with time complexity $O(k^3)$.
@@
 \begin{align}
-\label{eqn:betamat}
+\label{eqn:alphamat}
 \begin{bmatrix}
@@
 \end{align}
-Treat ${\bf s}$ as a vector of $k$ unknowns $s_0,\dots,s_{k-1}$,
-then equation \ref{eqn:betamat} can be solved by using Cramer's rule or Gaussian elimination.
-In other words, we can obtain $s_i=\mathcal G(S)$ by solving equation \ref{eqn:betamat} symbolically.
+Let $\vec{S}$ be a vector of $k$ unknowns $(s_0,\dots,s_{k-1})$,
+then Equation \ref{eqn:alphamat} can be solved by using Cramer's rule or Gaussian elimination.
+In other words, we can obtain $s_i=\mathcal G(S)$ by solving Equation \ref{eqn:alphamat} symbolically.
 
@@
 as well as the set of vanishing polynomials for primary inputs 
-$J_0^{PI} = \langle x_1^2-x_1,\dots,x_n^2-x_n\rangle$. Using elimination term order with $S>x_i>T$,
+$J_0^{PI} = \langle x_1^2-x_1,\dots,x_d^2-x_d\rangle$. Using elimination term order with $S>x_i>T$,
 we can compute a GB of the elimination ideal $\langle f_T,f_S\rangle + J_0^{PI}$. This GB contains a univariate
-polynomial denoting next states. The improved algorithm is depicted in Alg. \ref{alg:refined}.
+polynomial denoting next states. The improved algorithm is depicted in Algorithm \ref{alg:refined}.
 
@@
 \Indm
- \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\F(S)$}
+ \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\Func(S)$}
+ \KwOut{Final reachable states represented by polynomial $\mathcal G(T)$}
 \Indp
 
-  $from^0 = reached = \F(S)$\;
+  $from^0 = reached = \Func(S)$\;
   $f_T = $Reduce($Spoly(f_w,f_g), J_{ckt})$\;
@@
   	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
-  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
+  	\tcc{Use product of ideals to attain a union of new reached states and formerly reached states}
 	$from^i \gets new^i(S\setminus T)$\;
-	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
+	\tcc{Start a new iteration by replacing variable $T$ in new reached states with current state variable $S$}
   }
@@
 
-Assume we have $k$-bit datapath and $n$-bit PIs. In Galois field, we need to carefully partition $n$ PIs
+% fig:PI
+
+Assume we have $k$-bit datapath and $n$-bit PIs. In finite field, we need to carefully partition $n$ PIs
 such that states of each partition can be covered by a univariate polynomial respectively.
@@
 \begin{Proposition}
-Divide $n$-bit PIs to partitions $n_1,n_2,\dots, n_s$ and let $n_1$-bit, $n_2$-bit, $\dots$, $n_s$-bit word-level variables
-represent their evaluations in $\mathbb F_{2^k}$ respectively. Then $n_1,n_2,\dots,n_s~|~k$.
+Divide $n$-bit PIs into partitions $n_1,n_2,\dots, n_s$ where each $n_i|k$. Then let $n_1$-bit, $n_2$-bit, $\dots$, $n_s$-bit word-level variables
+represent their evaluations in $\mathbb F_{2^k}$ as $\F_{2^{n_i}} \subset \Fkk$.
 \end{Proposition}

### 2016-11-28 dissertation/normal.tex
dissertation v2:part unsat update 1

@@
 \subsection{Implicit Unrolling based on Abstraction with ATO}
-% fig:comb_RAB
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
+\caption{Architecture of a combinational GF multiplier}
+\label{fig:comb_RAB}}
+\end{figure}
 
@@
 Analysis on experiment results shows major time consumption is on ``multi-division" part.
-A matrix-based technique named as "F-4 style reduction" \cite{F4reduce} can speed up the procedure
+A matrix-based technique named as "F-4 style reduction" \cite{f4} can speed up the procedure
 dividing a low-degree polynomial with term-sparse polynomial ideal.

### 2016-11-28 dissertation/recha.tex
dissertation v2:part unsat update 1

@@
 
-Assume we are given gate-level netlist shown in figure \ref{fig:Lagrange}. It can be written as 3 
+Assume we are given gate-level netlist shown in Figure \ref{fig:Lagrange}. It can be written as 3 
 Boolean equations:
@@
 they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
-written in the form of equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
+written in the form of Equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
 the real number field, the essential concept of Lagrange's interpolation remains the same. 
@@
 state $from^i = S^0$, the algorithm computes the states reachable in
-1-step from $from^i$ in each iteration. In line 4 of algorithm
+1-step from $from^i$ in each iteration. In line 4 of Algorithm
 \ref{alg:BFS}, the {\it image computation} is 
@@
 \begin{Example}
-For the circuit in Figure \ref{fig:fsm} (b), we have the transition
+For the circuit in Figure \ref{fig:fsm}(b), we have the transition
 functions of the machine as:
@@
 
-% fig:proj_reacha
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/proj_reacha.pdf}
+\caption{Projection of the variety of circuit description ideal}
+\label{fig:proj_reacha}}
+\end{figure}
 
@@
 we describe a method that computes a GB of a smaller subset of polynomials.
-The approach draws inspirations from \cite{pruss:tcad}. 
-According to proposition 2 in \cite{pruss:tcad}, if the GB is computed using {\it refined abstraction term
+The approach draws inspirations from \cite{pruss:tcad15}. 
+According to proposition 2 in \cite{pruss:tcad15}, if the GB is computed using {\it refined abstraction term
 order} (RATO), there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
@@
 After adding intermediate bit-level signal $a,b,c,d$, the elimination ideal for example circuit 
-(example \ref{ex:SMPO}) can be rewritten LEX order with RATO:
+(Example \ref{ex:SMPO}) can be rewritten LEX order with RATO:
 \begin{align}
@@
 
-% fig:PI
+\begin{figure}[hbt]
+\centering{
+\includegraphics[width=\textwidth]{newfig/PI.pdf}
+\caption{PI partition of a sequential circuit}
+\label{fig:PI}}
+\end{figure}
 
@@
 $$P^4+P,~~\text{where}~~P=p_0+p_1\cdot\beta$$
-Using theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
+Using Theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
 $$P = p_0 + p_1\cdot\alpha^5$$
@@
 \section{Implementation of Word-level FSM Traversal Algorithm}
-In this section, we will introduce the tool implementation of our approach.
-The architecture of our tool consists of 4 functional components.
-First, given circuit is synthesized to gate-level using state-of-art synthesizers.
-Secondly, the synthesized netlist is translated to polynomial form and variables are sorted in 
-RATO. This part is implemented using scripting language such as \emph{Perl}. Thirdly, 
-the polynomial reduction is executed using our customized reduction engine, which is dumped in C++.
-Finally, we utilize symbolic computation engine -- \textsc{Singular} \cite{DGPS} to code algorithm 
-\ref{alg:refined} and execute the BFS traversal. The result is given in a univariate polynomial 
-about NS word $T$, denoting the set of reachable states from given initial state.
-
-The following example illustrates the full-blown reachability analysis approach
+In this section, we described the architecture of our tool which can perform word-level FSM traversal
+on FSM benchmark circuits.
+Briefly speaking, our tool consists of 3 functional components.
+% First, the given circuit is synthesized to gate-level using state-of-art synthesizers.
+First, the gate-level netlist of circuit is translated to polynomial form and variables are sorted in 
+RATO. This part is implemented using scripting language such as \emph{Perl}. If 
+the given benchmark is a structural/behavior hybrid description, we perform a pre-process on it to get a synthesized netlist.
+Secondly, the polynomial reduction is executed using our customized reduction engine, which is written in C++.
+Finally, we utilize symbolic computation engine -- \textsc{Singular} \cite{DGPS} to code Algorithm 
+\ref{alg:refined} and execute the BFS traversal. The tool outputs a univariate polynomial 
+for NS word $T$, denoting the set of reachable states from given initial state.
+
+Figure \ref{fig:flowchart} illustrates the execution of reachability analysis approach
 based on C++ and \textsc{Singular} implementation.
 
-\begin{Example}[Full Blown Reachability Analysis Walk-through]
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/flowchart.pdf}
+\caption{Execution process of word-level FSM traversal tool}
+\label{fig:flowchart}}
+\end{figure}
+
 Usually FSM designs can be described in behavior/structural hybrid languages.
@@
 
+\begin{Example}
 In this example, we use benchmark FSM ``lion9" from MCNC benchmark library.
-Figure \ref{fig:BLIF} is the truth table based structural BLIF representation of this FSM.
-
-\begin{figure}[hbt]
-\centering{
-\includegraphics[width=\textwidth]{newfig/BLIF.pdf}
-\caption{Sample input BLIF file}
-\label{fig:BLIF}}
-\end{figure}
+This benchmark circuit is given as truth-table based structural BLIF representation of a FSM.
+
+% \begin{figure}[hbt]
+% \centering{
+% \includegraphics[width=\textwidth]{newfig/BLIF.pdf}
+% \caption{Sample input BLIF file}
+% \label{fig:BLIF}}
+% \end{figure}
 
@@
 gate-level netlist. Modern synthesizers including ABC \cite{brayton2010abc} and SIS \cite{SIS} 
-can complete this work. Figure \ref{fig:BLIF_syn} shows a synthesized FSM given in 
-BLIF format. Note that for simplifying purposes in this example, we only use 2-input logic gates.
-
-\begin{figure}[hbt]
-\centering{
-\includegraphics[width=\textwidth]{newfig/BLIF_syn.pdf}
-\caption{Synthesized BLIF file}
-\label{fig:BLIF_syn}}
-\end{figure}
+can complete this work. In this example we use a synthesis library containing only 2-input 
+AND, NAND, OR, NOR, XOR, XNOR gates as well as the inverter, the output synthesized FSM is also given in 
+BLIF format.
+
+% \begin{figure}[hbt]
+% \centering{
+% \includegraphics[width=\textwidth]{newfig/BLIF_syn.pdf}
+% \caption{Synthesized BLIF file}
+% \label{fig:BLIF_syn}}
+% \end{figure}
 
 Using our interpreter, the synthesized BLIF file turns to a polynomial file customized for our 
-polynomial reduction engine as figure \ref{fig:ALG} depicts. Note the first line includes all variables in RATO, the second line 
-denotes the word-level variable. The third line is the irreducible polynomial of current GF $(\F_{2^4})$, the fourth line is the Spoly need to be reduced. The rest part has one polynomial in 
-the elimination ideal in each line.
-
-\begin{figure}[hbt]
-\centering{
-\includegraphics[width=\textwidth]{newfig/ALG.pdf}
-\caption{Polynomial file prepared for polynomial reduction engine}
-\label{fig:ALG}}
-\end{figure}
+polynomial reduction engine. 
+The input file format includes all 
+variables in RATO, along with the S-poly need to be reduced and polynomials in $J_{ckt}$.
+% 
+% \begin{figure}[hbt]
+% \centering{
+% \includegraphics[width=\textwidth]{newfig/ALG.pdf}
+% \caption{Polynomial file prepared for polynomial reduction engine}
+% \label{fig:ALG}}
+% \end{figure}
 
@@
 This is the transition function of this FSM. We utilized \textsc{Singular} to integrat both 
-bit-word substitution and tranversal algorithm. In figure \ref{fig:SING}, ``tran" is the transition
+bit-word substitution and tranversal algorithm. In Figure \ref{fig:SING}, ``tran" is the transition
 function we just obtained. ``init\_S" is the initial state, note it equals to ``0100" register 
-reset values in figure \ref{fig:BLIF}. Moreover, 2 bits PIs $x_0,x_1$ are combined to a word-level 
-PI variable $P$ using our conclusion in example \ref{ex:PI}: ``def\_X" is the definition of 2-bit 
+reset values preloaded in the BLIF file. Moreover, 2 bits PIs $x_0,x_1$ are combined to a word-level 
+PI variable $P$ using our conclusion in Example \ref{ex:PI}: ``def\_X" is the definition of 2-bit 
 word, and ``red\_X" denotes the vanishing polynomial for word $P$.
 
-\begin{figure}[hbt]
+After the script is executed, the trversal finishes after 4 
+transition iterations, which denotes BFS depth equals to 4. The final reachable states 
+is a degree-9 polynomial about $T$, indicating final reachable states set contains 9 states.
+And state encodings can be obtained by solving this polynomial equation.
+
+\begin{figure}[H]
 \centering{
@@
 \end{figure}
-
-After the script is executed, the trversal finishes after 4 
-transition iterations, which denotes BFS depth equals to 4. The final reachable states 
-is a degree-9 polynomial about $T$, indicating final reachable states set contains 9 states.
-And state encodings can be obtained by solving this polynomial equation.
 
@@
 Gr\"obner basis computations, under the term order derived from the
-circuit as Line 2 in Alg. \ref{alg:refined}. This is implemented 
+circuit as Line 2 in Algorithm \ref{alg:refined}. This is implemented 
 with our customized data structure in
@@
 algebra computation system [v. 3-1-6] \cite{DGPS}, as Line 3 in 
-Alg. \ref{alg:refined}; and the third part
+Algorithm \ref{alg:refined}; and the third part
 executes the reachability checking iterations using 
@@
 words ($k$) and the number of final reached states, which corresponds
-to the degree of polynomial {\it reached} in Alg.\ref{alg:univa}. 
+to the degree of polynomial {\it reached} in Algorithm \ref{alg:univa}. 
 Although the efficiency of our initial implementation fails to compete
@@
 \section{Conclusion} \label{sec:conc}
-This paper has presented a new approach to perform reachability
+This dissertation has presented a new approach to perform reachability
 analysis of finite state machines at the word-level. This is achieved

### 2016-11-28 dissertation/unsat.tex
dissertation v2:part unsat update 1

@@
-\chapter{Finding Unsatisfiable Cores Extraction for a Set of Polynomials using the Gr\"obner Basis Algorithm}
+\chapter{Finding Unsatisfiable Cores of a Set of Polynomials using the Gr\"obner Basis Algorithm}
 \label{ch:UNSAT}
-Besides elimination ideal based abstraction, Gr\"obner basis can also be applied to a vast field
-about other techniques in formal verification. Satisfiability (SAT) problem is the basis of modern 
-computation theories, as well as the origin of most formal verification problem. 
-In this chapter, we will discuss a sort of problems branching out from SAT. They are about a situation 
+Gr\"obner basis can find further applications to a vast field
+of other techniques in formal verification. Satisfiability (SAT) problem is the basis of 
+most decidable decision problems, as well as the basis of many formal verification techniques. 
+In this chapter, we discuss a special topic branching out from SAT theory. They are about a situation 
 when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
 a set of constrains (e.g. clauses, formulas or polynomials) which is unsatisfiable, sometimes 
-it is worthy to explore a smaller subset (core) that keeps UNSAT. From the execution of GB algorithm, 
-an auxiliary structure can be obtained to help UNSAT core extraction. This chapter will introduce the 
+it is worthy to explore the reasons for UNSAT. From the execution of the GB algorithm, 
+an auxiliary structure can be obtained to help explore the reasons causing UNSAT. This chapter introduces the 
 details about the motivation, mechanism and implementation.
@@
 \section{Motivation}
+In this section, we introduce the motivation of our UNSAT core extraction research. We
+start with defining an UNSAT problem, then review the previous work and applications 
+to abstraction refinement. Moreover, we then present the need of finding a word-level 
+analog to contemporary UNSAT reasoning techniques.
+
+\subsection{Preliminaries of SAT/UNSAT Theory}
+Boolean satisfiability (SAT) problem is one of most basic problems in computer science.
+In the following part we define the terminology related to SAT problem \cite{anaICCAD}.
+
+\begin{Definition}
+A {\bf literal} $l$ is defined as a variable $v$, or its negation $\overline{v}$. A disjunction (OR relation)
+of literals forms a {\bf clause}, i.e. $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
+A Boolean formula can always be written as {\bf conjunctive normal form} (CNF),
+which is the conjunction (AND relation) of clauses: $F = c_1\land c_2\land\cdots\land c_k$.
+\end{Definition}
+
+Using above concepts, the SAT/UNSAT problem can also be formally defined:
+\begin{Definition}
+A {\bf satisfiability (SAT) problem} is a decision problem that takes a CNF formula and returns
+that the formula is satisfiable (SAT), when there is an assignment of the variables making the 
+formula evaluate to true. Otherwise, the formula is {\bf unsatisfiable (UNSAT)}.
+\end{Definition}
+
+Figure \ref{fig:SAT} shows a simple example of SAT problem on circuits verification:
+we need to verify whether subcircuit $A$ and subcircuit $B$ has the same function, so
+we build a miter circuit for their outputs $X$ and $Y$, and the equivalence checking 
+problem is turned into a SAT problem as follows:
+
+\begin{center}
+\begin{align*}
+&\textit{Is\ subcircuit\ A\ functionally\ equivalent\ to\ subcircuit\ B?}\\
+&\Longleftrightarrow
+\textit{Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ primary\ inputs\ a,b,c\ such\ that\ Z=1?}
+\end{align*}
+\end{center}
+
+\begin{figure}[H]
+\centering{
+\includegraphics[scale=0.3]{newfig/fig_SAT.eps}
+\caption{An example of Boolean satisfiability problem on circuits}
+\label{fig:SAT}}
+\end{figure}
+
+For an UNSAT problem, the cause of UNSAT may include a part of clauses.
+\begin{Definition}
+Assume a CNF formula $F$ is UNSAT. A subformula $M \subseteq F$ is an {\bf UNSAT core} 
+if $M$ is also UNSAT. Further, if $\forall c\in M$, $M\setminus\{c\}$ is SAT, then 
+$M$ is called a {\bf minimal UNSAT core} of $F$.
+\end{Definition}
+
+\subsection{Previous Work on CNF-UNSAT}
+In most cases, SAT problems are modeled as conjunctive normal form (CNF) formulas, and solved by procedures based on 
+Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of DPLL algorithm
+is recursive branching and backtrack searching. To improve the efficiency of DPLL-based
+SAT solver, efforts are made on minimizing number of branches, accelerating unit propagation
+and modifying backtrack algorithm. Recently state-of-art SAT solvers developed conflict-driven
+clause learning (CDCL) technique to prune the search space, which is effective to reduce
+result search time.
+
+When a SAT solver fails to give a SAT assignment, it will provide a {\it UNSAT proof} or {\it refutation proof}
+to prove the problem is UNSAT. By analyzing clauses involved to the UNSAT proof, we can 
+generate a subformula which remains to be UNSAT. A n\"aive method is to collect all leaf 
+clauses in an UNSAT proof as an UNSAT core. In practice, a minimal UNSAT core is more valuable.
+There are mainly 2 kinds of method to find minimal UNSAT cores.
+The one is insertion-based method, which is achieved by adding clauses to the smallest subset 
+until the subset turns to be UNSAT. The other method is deletion-based method,
+which is realized by deleting clauses from a larger subset until the subset turns to be SAT.
+Recently heuristics such as clause-set refinement \cite{modelrotation} and model rotation \cite{belov2011accelerating}
+are applied as improvements on the deletion/insertion based methods.
+These methods are also expanded to satisfiability modulo theories (SMT) \cite{cimatti2007simple}.
+
+On the other hand, researchers are seeking an alternative solution for SAT problem using a totally
+different method from ``old-fashioned" DPLL algorithm. One promising option is polynomial calculus (PC),
+mapping Boolean variables and connectors in CNF formulas to variables and operators in 
+Galois fields. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
+algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
+valid assignments or proof of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
+from computer algebra.
+
+The inspiration that use PC to solve SAT problems first comes from \cite{ceiSTOC96}.
+Using PC and its variations, researchers develop many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
+Besides, researchers borrow concepts from PC, combine them with traditional DPLL and clause learning techniques
+to build hybrid SAT solvers \cite{condratTACAS07,Zengler2010}.
+
 \subsection{Exploiting UNSAT cores for abstraction refinement}
-By exploring previous work, we learn that most state-of-art abstraction refinement techniques require 
+
+UNSAT cores can find broad applications in circuit verification. Many abstraction refinement techniques require 
 information mining from UNSAT proofs of intermediate abstractions. 
@@
 utilized in such techniques.
+
+% TODO explain BMC
 

### 2016-11-29 dissertation/unsat.tex
dissertation v2:major chapters completed

@@
 
+\begin{table}[htb]
+\centering
+\begin{tabular}{|c|c|} 
+\hline
+ Boolean operator & Function over $\Fkk$ \\
+\hline
+\hline
+ $a\land b$ & $a\cdot b$ \\
+\hline
+ $a\oplus b$ & $a+b$ \\
+\hline
+ $\bar{a}$ & $1+a$ \\
+\hline
+ $a\lor b$ & $a+b+a\cdot b$\\
+\hline
+\end{tabular}
+\caption{Mapping Boolean operators to functions over $\Fkk$}
+\label{tab:booltof4}  
+\end{table}
+
 The inspiration that use PC to solve SAT problems first comes from \cite{ceiSTOC96}.
@@
 \subsection{Exploiting UNSAT cores for abstraction refinement}
-
-UNSAT cores can find broad applications in circuit verification. Many abstraction refinement techniques require 
+Finding small UNSAT cores attracts interest for decades because of the need for all kinds of applications.
+In solving MaxSAT problems, small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
+It is applied to solving Boolean algebra problems, such as Boolean function decomposition \cite{lee2008bi}. 
+Small explanation generation in general constraint programming problems also rely on small UNSAT cores \cite{cambazard2008reformulating}.
+
+UNSAT cores can find wide range of applications in circuit verification as well. Many abstraction refinement techniques require 
 information mining from UNSAT proofs of intermediate abstractions. 
-Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain that how an UNSAT proof is
+Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain how an UNSAT proof is
 utilized in such techniques.
 
-% TODO explain BMC
+Bounded model checking (BMC) is a model checking technique which set upper bound to the length of all paths. 
+It can be solved by solving a SAT problem. Given a model $M$, property $p$ and bound $k$, 
+The $k$-BMC unrolls $M$ by $k$ clock-cycles and generates Boolean formula $F$ including violation check of $p$.
+Then $F$ is fed to a SAT solver, if the SAT solver returns SAT, then $p$ is violated in 
+some paths shorter than $k$. Otherwise, UNSAT indicates $p$ is not violated in all paths 
+shorter than $k$. Algorithm \ref{alg:absrefine} makes an improvement on BMC using 
+abstraction refinement.
 
@@
 \end{figure}
-Assume that we are given a sequential circuit with $n$ latches as shown in Fig.\ref{fig:refine}(a). 
+Assume that we are given a sequential circuit with $n$ latches as shown in Figure \ref{fig:refine}(a). 
 This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
@@
 $$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
-is $SAT$ or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
-$SAT$, then a counterexample is found that violates property $p$. If the result is $UNSAT$, we cannot
+is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
+SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
 assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
-without any fixpoint detected.
-In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is $UNSAT$.
-If there are some latch variables not included in this UNSAT core (denoted by $L_{abs}$), then we can assert that the evaluations of 
+without any fixpoint detection.
+In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
+If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
 these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
-In practice, we turn these latches into primary inputs/outputs as shown in Fig.\ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).
-
-The second ``if-else" branch means: if we do an ordinary model checking on the abstracted machine $M'$ and find no
+In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).
+
+The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
 error trace, we can assert that property $p$ also holds on the original machine $M$. The reason of this assertion is that
-the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an over-approximation
-of $M$, such that $(M'\implies p) \implies (M\implies p)$. If we find a violation on abstracted machine, then this
+the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf over-approximation}
+of $M$, such that 
+$$(M'\implies p) \implies (M\implies p)$$
+
+If we find a violation on abstracted machine, then this
 abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.
@@
 It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
-the UNSAT core is extracted using a conventional CNF-SAT solver, which will encounter the ``bit-blasting" problem
-when the size of datapath (number of latches in Fig.\ref{fig:refine}) is very large. \textbf{Here we propose a totally new 
+the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
+when the size of datapath (number of latches in Figure \ref{fig:refine}) is very large. \textbf{Here we propose an altogether new 
 method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
-method according to the following observation based on our experience:}
+method according to the following observation:}
 
 % {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
-{\it While computing GB over finite fields is exponential in the number of variables, GB  computation 
-is observed to be more efficient for UNSAT problems.} 
-The reason is discussed as follows:
+{\it While the complexity of computing a GB over finite fields is exponential in the number of variables,
+the GB  computation is observed to be more efficient for UNSAT problems.} 
+The reason is as follows:
 \begin{Theorem}
@@
 of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
-$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow reducedGB(J) = \{1\}$$
+$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow \text{reduced GB}(J) = \{1\}$$
 \end{Theorem}
-It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high
-time complexity and is usually not practicable. One reason is that the size of GB may explode even if the term
-ordering is carefully chosen. However if the reduced GB is 1, which means every term in the original polynomials
+It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high space and
+time complexity and is usually not practical. One reason is that the size of the GB may explode even if the term
+ordering is carefully chosen. However if the reduced GB is $\{1\}$, which means every term in the original polynomials
 will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
@@
 
-\subsection{A Demonstration of Motivating Example}
-One  important research topic about UNSAT problems is to identify UNSAT cores efficiently.
-An UNSAT core in a CNF formula denotes a subset of clauses which is still unsatisfiable. Here
-we re-define this concept in algebraic geometry:
-\begin{Definition}
-Assume a set of polynomials $F$ and its subset $F_s\subset F$. If ${\bf V}(\langle F\rangle) = {\bf V}
-(\langle F_s\rangle) = \emptyset$, we call $F_s$ an {\bf UNSAT core} of $F$. Additionally, if
-$F_s$ contains no other UNSAT core, we call it a {\bf minimal} UNSAT core.
-\end{Definition}
-
-We  conjecture that based on observation of Buchberger's algorithm's execution, an UNSAT core can be identified.
-\begin{Proposition}
+Moreover, conventional techniques are limited to bit-level variables (literals). Algebraic geometry methods
+allow the use of word-level variables, which provides a strong potential for all applications which can be
+modeled as polynomials in finite fields extensions.
+
+\section{UNSAT Cores of Polynomial Ideals}
+\label{sec:core}
+In this section we provide a solution to the problem that how to use 
+GB to extract the UNSAT core.
+
+\textbf{Problem statement: }
+Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
+the ring $R = \F[x_1,\dots,x_d]$ that generate ideal $J = \langle
+f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
+\emptyset$, or it is determined to be so by applying the Gr\"obner
+basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
+J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
+too. Borrowing the terminology from the Boolean SAT domain, we
+call $F_c$ the infeasible core or the unsat core of $F$.
+
+%% Any set of polynomials $F$ with empty variety is an unsat core in
+%% itself. There may be more than one unsat cores in $F$, some
+%% polynomials may be common to multiple cores (i.e. the
+%% cores may be non-disjoint), and these cores may have
+%% different cardinalities $|F_c|$. The unsat core problems
+%% are therefore further classified as: 
+%% \bi
+%% \item Identify a {\it minimum} core, i.e. a core of minimum
+%%   cardinality. 
+%% \item Identify a {\it minimal} core. An unsat core $F_{c}$ is minimal
+%%   with respect to the property that while the variety of $F_{c}$ is
+%%   empty, the variety of {\it any} subset of $F_c$ is {\it non-empty},
+%%   i.e. $V( F_{c} - \{f_i\} ) \neq \emptyset$ for any $f_i \in
+%%   F_c$. 
+%% \item Identify any unsat core disregarding minimality; i.e. find any
+%%   subset $F_c$ of $F$.
+%% \ei
+
+%% Let us first consider the problem of finding any $F_c \subset F$,
+%% disregarding minimality. 
+It is not hard to motivate that an unsat core should be identifiable
+using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
+$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
+the ideal generated by $(F - \{f_j\})$, i.e. $f_j \in \langle F -
+\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
+polynomials of $F_c$, so $f_j$ is  not a part of the unsat core, and
+it can be safely discarded from $F_c$. This can be identified by means
+of the GB algorithm for this ideal membership test.% (Definition \ref{def:gb}(ii)).
+
+A n\"aive way (and inefficient way) to identify {\it a minimal core}
+using the GB computation is as follows:
+select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
+(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
+from the core; otherwise retain $f_i$ in $F_c$. Select a different
+$f_i$ and continue until all polynomials $f_i$ are visited for
+inclusion in $F_c$. This approach will produce a minimal core, as we
+would have tested each polynomial $f_i$ for inclusion in the
+core. This requires $O(|F|)$ calls to the GB engine, which is really
+impractical.
+
+\subsection{An Example Motivating our Approach}
+
 Buchberger's algorithm picks pairs of polynomials from a given set, computes their S-poly, then reduces this S-poly
@@
  execution of Buchberger's algorithm.
-\end{Proposition}
+
 \begin{Example}
@@
 
-% \begin{align*}
-% &\bar{a}\lor\bar{b}\\
-% &a\lor\bar{b}\\
-% &\bar{a}\lor b\\
-% &a\lor b\\
-% &x\lor y\\
-% &y\lor z\\
-% &b\lor \neg y\\
-% &a\lor x\lor \neg z
-% \end{align*}
+\vspace{0.1in}
 Using Boolean to polynomial mappings given in Table \ref{tab:booltof4}, we can transform them to a set of
-polynomials $F$ over ring $\mathbb F_2[a,b,x,y,z]$:
+polynomials $F=\{f_1,\dots,f_8\}$ over ring $\mathbb F_2[a,b,x,y,z]$:
 
@@
 \end{minipage}
-% \begin{align*}
-% &f_1:ab\\
-% &f_2:ab+a\\
-% &f_3:ab+b\\
-% &f_4:ab+a+b+1\\
-% &f_5:xy+y+x+1\\
-% &f_6:yz+y+z+1\\
-% &f_7:by+y\\
-% &f_8:axz+az+xz+z
-% \end{align*}
+
+\vspace{0.1in}
 We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
 Since this problem is UNSAT, we will stop when ``1" is added to GB.
-\begin{enumerate}
+\begin{enumerate}[{1)}]
 \item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
@@
 \item Update $F=F\cup r_2$;
-\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Fig.\ref{fig:UNSAT}(a) shows;
+\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:UNSAT}(a) shows;
 \item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
@@
 terminate the Buchberger's algorithm;
-\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Fig.\ref{fig:UNSAT}(b). 
+\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:UNSAT}(b). 
 From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
-as these polynomials are the ``causes" of unsatisfiability of original set of clauses.
+as these polynomials are the ``causes" of unsatisfiability of original set of polynomials.
 \end{enumerate}
@@
 
-We conclude our approach as a conjecture algorithm (Algorithm \ref{alg:UNSAT}).
+We conclude our approach as a conjecture Algorithm \ref{alg:UNSAT}.
 
@@
 \end{algorithm}
-
-\section{Formalize the Buchberger's Algorithm based UNSAT Core Identification}
-\label{sec:core}
-
-\textbf{Problem: }
-Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
-the ring $R = \F[x_1,\dots,x_d]$ that generate ideal $J = \langle
-f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
-\emptyset$, or it is determined to be so by applying the Gr\"obner
-basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
-J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
-too. Borrowing the terminology from the Boolean SAT domain, we
-call $F_c$ the infeasible core or the unsat core of $F$.
-
-%% Any set of polynomials $F$ with empty variety is an unsat core in
-%% itself. There may be more than one unsat cores in $F$, some
-%% polynomials may be common to multiple cores (i.e. the
-%% cores may be non-disjoint), and these cores may have
-%% different cardinalities $|F_c|$. The unsat core problems
-%% are therefore further classified as: 
-%% \bi
-%% \item Identify a {\it minimum} core, i.e. a core of minimum
-%%   cardinality. 
-%% \item Identify a {\it minimal} core. An unsat core $F_{c}$ is minimal
-%%   with respect to the property that while the variety of $F_{c}$ is
-%%   empty, the variety of {\it any} subset of $F_c$ is {\it non-empty},
-%%   i.e. $V( F_{c} - \{f_i\} ) \neq \emptyset$ for any $f_i \in
-%%   F_c$. 
-%% \item Identify any unsat core disregarding minimality; i.e. find any
-%%   subset $F_c$ of $F$.
-%% \ei
-
-%% Let us first consider the problem of finding any $F_c \subset F$,
-%% disregarding minimality. 
-It is not hard to motivate that an unsat core should be identifiable
-using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
-$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
-the ideal generated by $(F - \{f_j\})$, i.e. $f_j \in \langle F -
-\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
-polynomials of $F_c$, so $f_j$ is  not a part of the unsat core, and
-it can be safely discarded from $F_c$. This can be identified by means
-of the GB algorithm for this ideal membership test.% (Def. \ref{def:gb}(ii)).
-
-A n\"aive way (and inefficient way) to identify {\it a minimal core}
-using the GB computation is as follows:
-select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
-(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
-from the core; otherwise retain $f_i$ in $F_c$. Select a different
-$f_i$ and continue until all polynomials $f_i$ are visited for
-inclusion in $F_c$. This approach will produce a minimal core, as we
-would have tested each polynomial $f_i$ for inclusion in the
-core. This requires $O(|F|)$ calls to the GB engine, which is really
-impractical.
 
@@
 corresponding to this sequence of reductions is depicted by means of a
-graph in Fig. \ref{fig:refute}. We call this graph a {\it refutation tree}.
+graph in Figure \ref{fig:refute}. We call this graph a {\it refutation tree}.
 
@@
 \centering
-\includegraphics[width=\textwidth]{newfig/refutation_tree.eps}
+\includegraphics[width=4.5in]{newfig/refutation_tree.eps}
 \caption{Generating refutation trees to record unsat cores}
@@
 division by $f_2$ resulted in the quotient $a$. The refutation tree of
-Fig. \ref{fig:refute} shows further that
+Figure  \ref{fig:refute} shows further that
 $Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
@@
 the transitive fanin of ``1'' can be safely discarded from $F_c$. From
-Fig. \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
+Figure \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
 as an unsat core of $F$.
@@
 %how the size of this core could be reduced further.  
-For this purpose, we will have to perform a more 
+For this purpose, we perform a more 
 systematic book-keeping of the data generated during the execution of
@@
 
-In Eqns. (\ref{eqn1}) and (\ref{data1}), $g_{ij}$ denotes the
+In Equation \ref{eqn1} and Equation \ref{data1}, $g_{ij}$ denotes the
 remainder of the $S$-polynomial $Spoly(f_i,f_j)$ modulo the current system
@@
 the coefficients of $f_i$, respectively $f_j$, in the $S$-polynomial
-$Spoly(f_i,f_j)$. Furthermore, in Eqn. (\ref{data1}), $(c_{k1}, \dots
+$Spoly(f_i,f_j)$. Furthermore, in Equation \ref{data1}, $(c_{k1}, \dots
 c_{kl})$ are the respective quotients of division by
@@
 \begin{Example}
-Revisiting Ex. \ref{ex:1}, and Fig. \ref{fig:refute}, the data
+Revisiting Example \ref{ex:1}, and Figure \ref{fig:refute}, the data
 corresponding to $Spoly(f_1,f_2)$\\$\xrightarrow{F}_+  g_{12} = f_{10}$
@@
 $$f_{10}=g_{12}=f_1+acf_2+af_2+f_3+cf_2+f_2$$ is obtained.
-The data is recorded according to Eqn. (\ref{data1}):
+The data is recorded according to Equation \ref{data1}:
 
@@
 f_s\}$. Moreover, this matrix $M$ is constructed precisely using the
-data that is recorded in the form of Eqn. (\ref{data1}). We now give a condition
+data that is recorded in the form of Equation \ref{data1}. We now give a condition
 when the matrix $M$ may identify some redundant elements.
@@
 Corresponding to Example \ref{ex:1} and the refutation tree shown in
-Fig. \ref{fig:refute}, we discover that the polynomial $f_3$ is used
+Figure \ref{fig:refute}, we discover that the polynomial $f_3$ is used
 only twice in the division process. In both occasions, the quotient of
-the division is 1. From Fig. \ref{fig:refute}, it follows that:
+the division is 1. From Figure \ref{fig:refute}, it follows that:
 \begin{equation}
@@
 $f_{m+1}$ and add it to the current set of generators $G$. We
-also record the data as in Eqn. (\ref{data1}): 
+also record the data as in Equation \ref{data1}: 
 \begin{displaymath}
@@
 
-{\bf Recording the coefficients:} In Eqn. (\ref{eqn:red}) we obtain a
+{\bf Recording the coefficients:} In Equation \ref{eqn:red} we obtain a
 vector of polynomial coefficients $c_k$ where $k>s$. These
@@
 corresponding data is stored in a data-structure $D$ corresponding to
-Eqn. (\ref{data1}). The matrix $M$ is also constructed. From this
+Equation \ref{data1}. The matrix $M$ is also constructed. From this
 recorded data the refutation tree can be easily derived.
@@
 
-Similarly, the motivation for using the \emph{frequency of occurrence}
+Similarly, we can define the concept of another heuristic:
+\begin{Definition}[Frequency of Occurrence]
+Frequency of occurrence of a polynomial $f_i$ in a refutation tree
+corresponds to the number it appears in the refutation tree. 
+\end{Definition}
+
+The motivation for using the \emph{frequency of occurrence}
 of $f_i$ in the refutation tree is as follows: polynomials that appear
@@
 identified as $\{f_1, f_2,f_3,f_4\}$, and we obtain a
-refutation tree as shown in Fig.\ref{fig:refine}(a).  
+refutation tree as shown in Figure \ref{fig:refine}(a).  
 \begin{figure}[hbt]
@@
 The refutation distance corresponding to polynomial $f_2$ is equal to
-2 levels. Note that while three leaf nodes in Fig. \ref{fig:refine} (a)
+2 levels. Note that while three leaf nodes in Figure \ref{fig:refine} (a)
 correspond to $f_2$, the shortest distance from ``1'' to any $f_2$
@@
 $\{f_1',f_2',f_3',f_4'\}$. The result is shown in
-Fig.\ref{fig:refine}(b) with the core identified as $\{f_1', f_3',
+Figure \ref{fig:refine}(b) with the core identified as $\{f_1', f_3',
 f_4'\} = \{f_2,f_3,f_4\}$. Further iterations do not refine the core
@@
 For each $Spoly(f_i,f_j)\xrightarrow{F}_+0$ reduction, we record the
-information on corresponding syzygies as in Eqn. (\ref{eqn:syz}), also
-represented in matrix form in Eqn. (\ref{mat:syzygy}):
+information on corresponding syzygies as in Equation \ref{eqn:syz}, also
+represented in matrix form in Equation \ref{mat:syzygy}:
 
@@
  r_2c_j^2 + \cdots + r_mc_j^m = \sum_{i=1}^m r_ic_j^i.$ 
- If we multiply each row $i$ in the matrix of Eqn. (\ref{mat:syzygy})
+ If we multiply each row $i$ in the matrix of Equation \ref{mat:syzygy}
  with $r_i$, and sum up all the rows, we will obtain the
@@
  \begin{Example}
- Revisiting Example\ref{ex:1}, execute the GB-core algorithm and
+ Revisiting Example \ref{ex:1}, execute the GB-core algorithm and
  record the syzygies on $f_1,\dots,f_s$ corresponding to the
@@
 Usually, we need to generate extra columns compared to the syzygy
-matrix of Eqn. (\ref{mat:syzygy}). In this example, we need to add an
+matrix of Equation \ref{mat:syzygy}. In this example, we need to add an
 extra column for the coefficient of $f_{10}$. This is because $f_{10}$
@@
 
- By adding this row vector (Eqn. (\ref{sr})) to the rows in
- Eqn. (\ref{eqn:sm}) corresponding to the non-zero entries in the
+ By adding this row vector (Equation \ref{sr}) to the rows in
+ Equation \ref{eqn:sm} corresponding to the non-zero entries in the
  column for $f_{10}$, we obtain the syzygy matrix only for the
@@
 
+\section{Application to Abstraction Refinement}
+In this section we apply our UNSAT core extraction approach to the 
+$k$-BMC in Algorithm \ref{alg:refined}. This algorithm utilize UNSAT core to 
+remove irrelevant latches (state variables) to
+reduce the state space. Since those state variables contribute nothing to the violation of 
+property $p$, the abstracted model is a reasonable over-approximation for checking 
+$p$ without loss of accuracy. In the following example, we apply the UNSAT core extraction 
+approach to a FSM which is complicated for bounded model checking and show the power of 
+abstraction refinement to reduce the state space for refined $k$-BMC.
+
+\begin{Example}
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=\textwidth]{newfig/s27.pdf}
+\caption{Gate-level Schematic of Example Circuit}
+\label{fig:s27}}
+\end{figure}
+
+Figure \ref{fig:s27} shows a sequential circuit (``s27'' from the
+ISCAS benchmark set) with 3-bit state registers
+PS = $\{G7,G6,G5\}$ and NS = $\{G13,G11,G10\}$. 
+Its underlying FSM contains 8 states.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=4.5in]{newfig/s27_stg.pdf}
+\caption{State transition graph of example circuit}
+\label{fig:s27_stg}}
+\end{figure}
+
+On this FSM, define a linear temporal logic (LTL) property 
+$$p = {\bf AG}( (\neg G13) {\bf U} (\neg Z) )$$ 
+with given initial state $\{000\}$. Model checking on this machine requires
+traversal on the STG to search for an accepting trace. If we use $k$-BMC
+without abstraction refinement, we need to unroll the machine to
+iteration  $k=3$. We show how abstraction will be applied using our
+setup.
+
+Because the circuit has 3 latches, we model the problem over the field
+${\mathbb{F}}_{2^3}$. First, let the bound $k = 0$. Generate the
+polynomial constraints for initial state ideal $I$ and check $SAT(I\land
+\neg p)$ using \Grobner bases. The ideal
+
+\begin{align*}
+I = \langle & G14+1+G0, G8+G14\cdot G6, G15+G12+G8+G12\cdot G8,\\ 
+	& G16+G3+G8+G3\cdot G8, G9+1+G16\cdot G15, \\ 
+	& G10+1+G14+G11+G14\cdot G11, G11+1+G5+G9+G5\cdot G9, \\
+	& G12+1+G1+G7+G1\cdot G7, G13+1+G2+G12+G2\cdot G12, \\
+	& Z+1+G1, \\
+	& \text{(Initial state 000 )} G5, G6, G7 \rangle;
+\end{align*}
+
+Property $\neg p$ is also written as a polynomial in the first
+time-frame: 
+\begin{equation*}
+\neg p = Z\cdot G13 + 1
+\end{equation*}
+
+As $I\land \neg p$ is UNSAT by the weak
+Nullstellensatz, we extract an UNSAT core using our approach:
+\begin{align*}
+Core(I\land \neg p) =& G12+1+G1+G7+G1\cdot G7, G13+1+G2+G12+G2\cdot G12, \\
+& Z+1+G1, G7;
+\end{align*}
+
+The result shows that state variables $\{G5,G6, G10,G11\}$ are
+irrelevant when considering the violation of $p$. Thus, we can remove
+the latches $G10, G11$, and make $G5,G6$ primary inputs. In this way,
+the number of state variables is reduced to 1, such that the new
+machine only contains 2 states.
+
+\begin{figure}[H]
+\centering{
+\includegraphics[width=4.5in]{newfig/s27_stg_refined.pdf}
+\caption{State transition graph of abstracted machine}
+\label{fig:s27_stg_refined}}
+\end{figure}
+
+Since the state space is greatly
+reduced, we can execute unbounded model checking on this  abstracted
+machine with less cost. As a result, property $p$ is not violated on
+the abstracted machine. Therefore, $p$ is also a passing property of
+the original machine and  the refined $k$-BMC algorithm terminates!
+
+\end{Example}
+
 \section{Experiment results}
@@
 industry-size CNF-SAT benchmarks, as the translation from CNF introduces
-too many variables and clauses for GB engines to handle. In order to
-validate our approach, we use a somewhat customized  benchmark
-library: i) "aim-100" is a modified version  of the random 3-SAT
-benchmark "aim-50/100", modified by adding some redundant clauses; ii)
-The "subset" series are generated for random subset sum problems; iii)
+too many variables and clauses for GB engines to handle. 
+On the other hand, although our approach is totally compatible with 
+any constraints which can be written as polynomials in GF extensions,
+there is no such benchmark libraries clearly identifying a minimal core within 
+to test our tool.
+
+In order to
+validate our approach, we make a compromise and create a somewhat customized  benchmark
+library by modifying SAT benchmarks and translating from circuit benchmarks: 
+1) "aim-100" is a modified version  of the random 3-SAT
+benchmark "aim-50/100", modified by adding some redundant clauses; 2)
+The "subset" series are generated for random subset sum problems; 3)
 "cocktail" is similarly revised from a combination of factorization
-and a random 3-SAT benchmark; iv) and "phole4/5" are generated by
-adding redundant clauses to pigeon hole benchmarks; v) Moreover,
-SMPO and RH benchmarks correspond to hardware equivalence checking
-instances of sequential Galois field normal basis modulo multiplier
-circuits \cite{SMPO,RHmulti}, compared against a golden model
+and a random 3-SAT benchmark; 4) and "phole4/5" are generated by
+adding redundant clauses to pigeon hole benchmarks; 5) Moreover,
+``SMPO" and ``RH" benchmarks correspond to hardware equivalence checking
+instances of Agnew's SMPO and RH-SMPO
+circuits \cite{agnew1991implementation,RHmulti}, compared against a golden model
 {\it spec}. Similarly, the "MasVMon" benchmarks are the equivalence

### 2016-11-29 dissertation/prelim_basis.tex
dissertation v2:minor amendment

@@
 in subsequent chapters. The material of this chapter is
-mostly referred from the textbooks \cite{ideals:book} \cite{gb_book} and 
+mostly referred from the textbooks \cite{ideals:book,gb_book} and 
 previous work by {\it Lv} \cite{lv:phd} as well as {\it Pruss} \cite{pruss:tcad15}.
@@
 Gao et al. \cite{gao:qe-gf-gb} proves that when the number of variables and degree of polynomials 
-are restricted, the complexity reduces to single exponential $q^O(|\phi|)$.
+are restricted, the complexity reduces to single exponential $q^{O(|\phi|)}$.
 This provides the possibility to make use of Gr\"obner basis under restricted situations,
@@
 \begin{align*}
-g_1: B^4+B; 
+&g_1: B^4+B; 
 ~~g_2: b_0+b_1 \alpha + B; 
 ~~g_3: a_0+a_1 \alpha + A;  \nonumber \\
-~~g_4: c_0+c_1 \alpha + c_2 \alpha + c_3(\alpha+1)+Z;
+&g_4: c_0+c_1 \alpha + c_2 \alpha + c_3(\alpha+1)+Z;
 g_5: r_0+c_1+c_2; 
 ~~g_6: z_0+c_0+c_3; \nonumber \\
-~~g_7: z_1+r_0+c_3; 
+&g_7: z_1+r_0+c_3; 
 ~~{\bf g_8: Z+A\cdot B};
@@
 ~~g_{10}: a_1+A^2+A; \nonumber \\
-~~g_{11}: c_3+a_1\cdot b_1
+&g_{11}: c_3+a_1\cdot b_1
 g_{12}: c_2+a_1\cdot b_1 \alpha + a_1 \cdot B; 

### 2016-12-01 dissertation/intro.tex
dissertation v2_chap 3 4_update1
dissertation v2: part intro update 1

--- 2016-09-21 dissertation/intro.tex
@@
-\chapter{Introduction} \label{ch:intro}
-There is an ever-increasing need for secure communication within information 
-technology. Security of sensitive information relies more and more heavily 
-on encryption methodologies implemented in hardware by cryptographic 
-circuits. One of the most prominent of these methodologies is Elliptical 
-Curve Cryptography (\emph{ECC}), which provides more strength per encryption bit 
-than other encryption methodologies.
-The main building blocks of ECC hardware implementations 
-are fast, \emph{custom-built Galois field arithmetic circuits}.
-These circuits are notoriously hard to verify, yet their correctness is 
-vitally important in critical applications. In \cite{crypto:bug_attacks}, 
-for example, it is shown that a bug in the hardware could lead to the full 
-leakage of the secret cryptographic key, which could compromise the entire
-system. Thus, formal verification is imperative in Electronic Design 
-Automation (\emph{EDA}) when dealing with cryptographic circuits.
+\chapter{Introduction}
+\label{ch:intro}
+With the increasing size of integrated
+circuits, sequential circuit designers face complicated problems of
+design errors in specification models and implementations. These
+errors are usually modeled as ``bad" states, and the
+circuits/functional components are modeled as finite state machines
+(FSMs). Once state reachability is analyzed, the existence of errors
+can be identified by checking whether ``bad" states are {\it
+  reachable} from certain initial states. Temporal logic {\it model
+  checking} (MC) formulations and solvers are often used for this
+purpose. Once the designs and specification models are validated using
+model checking, optimized implementations of sequential circuits are
+synthesized. A subsequent problem then needs to be solved to prove
+that the sequential circuit implementations are equivalent to the
+original specification models, i.e. {\it Sequential Equivalence
+  Checking   (SEC)}.
 
-To facilitate this verification, it is highly desirable to obtain a {\it word-level 
-representation of the datapath of the ECC arithmetic block} from its 
-bit-level implementation. Ideally, this abstraction should be {\it canonical},
-as this allows the it to be directly applicable to equivalence checking. 
-Such a canonical, word-level abstraction of the 
-Galois field arithmetic block would not only make it easier to verify and 
-reason 
-about the cryptographic system as a whole, but also enable the use of 
-higher level abstraction and synthesis tools. 
-As arithmetic circuits are custom-designed, often modularly, using Galois field arithmetic blocks, 
-the abstraction should also exploit the hierarchical nature of the circuitry.
-Due to the modular circuit structure, abstraction of each arithmetic block
-becomes the key in verification of the full circuit. 
-Practical 
-applications of ECC dictate a datapath of a minimum of $163$-bits, up to 
-$571$-bits, as 
-designated by the National Institute for Standards and Technology (NIST).
-However abstraction of 
-Galois field arithmetic circuits has been infeasible for data-paths beyond 
-$16$ bits.
+Reachability analysis forms the backbone of most sequential
+verification techniques. As the state-space of FSMs increases,
+reachability analysis forms a fundamental bottleneck in sequential
+verification. Contemporary approaches employ various techniques to
+overcome this state-explosion problem: i) Bounded model checking
+\cite{bitlevel1} traverses the FSMs for a fixed number of steps $k$
+($k$-BMC) to check whether a property violation can occur in $k$ or
+fewer steps. ii) Analyze over-approximations (or {\it abstractions})
+of the state-space. Abstraction proves properties on the system by
+first simplifying it, and when the abstraction does not satisfy the
+same properties as the original one, a process of refinement is
+needed. {\it Craig interpolants} \cite{craig-interpolate}
+\cite{pudlak:ci} are used as a means of abstraction and reachability
+analysis for linear temporal logic (LTL) model checking
+\cite{mcmillan:interpolation}. Similarly, counterexample guided
+abstraction refinement (CEGAR) \cite{cegar-journal} uses proofs of
+unsatisfiability (UNSAT cores) to refine the abstractions. iii) the
+recent breakthrough method of \cite{bradley:pdr} where the set of
+over-approximations to forward reachable states are refined with
+inductive constraints -- property directed reachability (PDR).
 
-This dissertation proposes an algebraic geometry based approach to abstract canonical, 
-word-level representations of bit-level Galois field arithmetic circuits.
-The approach is able to abstract representations
-for circuits up to 571 bits in size, which is the largest NIST
-standard for datapath size in ECC. 
-Verification of circuits for which this abstraction has been computed is shown to 
-be trivial; thus, the focus is on deriving the abstraction quickly and efficiently.
-%Furthermore, we show how we can use these abstract representations to 
-%perform improved formal verification of these circuits.
+While the above techniques have made significant strides in sequential
+verification, numerous practical instances remain unsolved. One issue
+with all of the above techniques is that they mostly use bit-level
+constraints to model the transition relations and sets of 
+states. We have access to instances from our industrial
+collaborators (Calypto Design Systems, now Mentor Graphics) where
+the designs are expressed at the level of bit-vector words
+(e.g. Matlab code, Verilog), and these word-level abstractions are
+rarely exploited in verification. The problem is further exacerbated
+when there are arithmetic operators on word-level operands embedded in
+the control logic. While attempts have been made towards word-level
+predicate abstraction \cite{Jain05wordlevel} \cite{mcmillan:cav06}
+\cite{mcmillan:impact}, {\it using a purely word-level representation
+  of the state-space, the properties and their abstractions have not
+  been fully explored as another dimension in improving sequential
+  verification.}
 
-\section{Hardware Design and Verification Overview}
+There are several advantages in exploiting word-level information for
+verification. A number of designs have  their
+datapaths and/or system-level models described as word-level RTL
+models.  Exploiting word-level instead of bit-level information is one
+way of abstraction -- a key technique to reduce the state space of a
+FSM. It has the effect of combining sets of states with similar 
+properties. During reachability analysis, if we use bit-level
+variables to  represent the states, the representations may become too
+large to handle. However, when a ``bundle" of bit-level variables are
+represented as {\it only one word-level variable}, the set of
+reachable states can be represented by a  word-level constraint
+expression; which may lower verification complexity.
+
+{\bf Overall Research Objective:} In this proposal we investigate a
+set of new, promising approaches for {\it word-level representation,
+reachability analysis and abstraction} with applications to SEC,
+$k$-BMC and abstraction-based unbounded LTL model checking
+(A-UMC). Our techniques operate at the word-level and they are based
+largely on the concepts from {\it algebraic geometry}.
+
+For word-level SEC, we are given two designs, or their corresponding
+Mealy/Moore FSMs ${\mathcal{M}}_1,{\mathcal{M}}_2$, along with their
+initial starting states $S_0^1,S_0^2$. We wish to prove the absence of
+a sequence of inputs (string) that distinguishes the initial
+states \cite{coudert:iccad90}
+\cite{coudert1990verification}. Fundamentally, this requires 
+the construction of a product machine; and the main research problem
+relates to that of performing {\it FSM traversal \cite{touati:iccad90}
+but  at the word-level}. Analogously, in the case of MC, the problem
+is setup w.r.t. a FSM $\mathcal{M}$, a set of initial states $S_0$ and
+a set of property states $p$. Techniques are to be researched that
+verify that there exist no sequence of transitions from an initial
+state to a non-property state (``bad'' state). These problems have to
+be solved in the context of word-level verification -- i.e. data
+representation, abstraction (Craig interpolation, UNSAT cores) and
+algorithm execution has to be carried out at word-level.
+
+{\bf The Algebraic Geometry Model and Rationale:} We will represent
+the FSMs -- the transition relations -- by means of a set of
+multi-variate polynomials with coefficients from the finite (Galois)
+field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
+$\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
+Boolean assignment to a set of $k$-bit state register variables
+$S=\{s_0,\dots,s_{k-1}\}$. Therefore, we can consider each ($k$-bit)
+state as a word-level element $S$ of the finite field
+$\Fkk$. Algorithms can directly operate on polynomials in word-level
+variable $S$.
+
+Boolean functions with $k$-bit inputs and $k$-bit outputs 
+$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
+$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
+field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
+\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
+there exists a unique canonical polynomial $\FF$ that describes $f$.
+This implies that one can derive a {\it  canonical, polynomial
+  abstraction} of the function as $Z = \FF(A)$ where $Z, A$ are
+word-level symbols representing $k$-bit operands. The concept also
+generalizes to functions with different input/output bit-vector sizes,
+i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
+polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
+{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
+This implies that the FSM's transition relations can be
+represented as polynomial functions (ideals) in $\Fkk$, and values of
+state variables can be represented as solutions to these polynomials
+(variety of the ideal). Subsequently, the ideal-variety correspondences
+in algebraic geometry can be applied to implement symbolic reasoning
+about state reachability. Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
+our model provides a single, unified and bit-precise representation
+for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
+constraints.
+
+The decision and abstraction procedures in our setting will rely on
+the theory and technology of {\it \Grobner bases (GB)}. GB-based
+algebraic reasoning is very powerful; in fact it is known to be
+strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
+light of the above discussion, using concepts from algebraic geometry
+and \Grobner bases over $\Fkk$, we can introduce another dimension of
+word-level abstraction to the techniques in sequential verification.
+
+{\bf The core focus of this research:} The use of algebraic geometry
+for sequential circuit verification and symbolic model checking has
+been presented before. Avrunin presented the
+concept of symbolic MC using algebraic geometry in
+\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
+GB-algorithms for CTL, LTL and 
+bounded MC over Boolean rings. However, these approaches are a
+straight-forward transformation of the problem to {\it bit-level}
+Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
+the concepts of word-level reachability, abstraction-refinement using
+interpolation or UNSAT cores, etc., that we desire was not the focus of
+\cite{Avrunin:CAV} \cite{vardi-iasted07}.
+
+In this research, we will focus on three specific research problems: 
+\begin{enumerate}
+\item How to perform reachability analysis at the word-level? This
+  of course forms the foundation for word-level SEC and MC.
+\item What is the analog of Craig interpolation in an algebraic
+  geometry setting (theoretical challenge), and how can we compute
+  various interpolants using GB-engines efficiently (practical
+  implementation)? This will enable word-level abstraction of the
+  state-space. 
+\item Explore the concept, and the computation, of unsatisfiable (UNSAT)
+  cores of  a set of polynomials. Many abstraction-refinement
+  techniques make use of UNSAT cores of CNF-formulas in MC. Therefore,
+  our investigation of UNSAT cores of polynomial ideals will enable
+  similar applications at word-level. 
+\end{enumerate}
+
+\section{Design, Test and Verification of Modern Sequential Circuit Components}\section{Hardware Design and Verification Overview}
 The typical design flow of a hardware system, 
@@
 
-\section{Dissertation Objective, Motivation, and Contributions}
-This dissertation focuses on abstracting a canonical, 
-word-level representation of hardware (bit-level) implementations of 
-combinational circuits. The proposed technique is a full abstraction solution which can be 
-applied to any arbitrary acyclic combinational circuit. 
-It is particularly efficient when applied to Galois field arithmetic circuits.
-Using this technique, if the abstraction of the circuit's implementation and its
-specification are found, they can be easily compared to determine equivalence.
-Implementation of a custom software tool, developed to compute the abstractions, is
-also described.
-
-\subsection{Motivating Application}
-The motivation for this work comes from applications of Galois field 
-arithmetic circuits in elliptical curve cryptography ({\it ECC}) hardware systems.
-The main operations of encryption, decryption, and 
-authentication in ECC rely on operations performed on elliptic curves, which 
-are implemented in hardware as polynomial functions over Galois fields. 
-To be applicable in real-world situations, ECC data-paths
-should be a minimum of $163$-bits wide, which is the minimum NIST standard, 
-up to a recommended size of $571$-bit operand widths. Many non-ECC cryptosystems
-have datapaths on the order of $1000$-bits.
-
-A Galois field arithmetic circuit with a datapath size of $k$ 
-is built as a Boolean function: $\mathbb{B}^k \rightarrow \mathbb{B}^k$. 
-This function is mapped to an operation 
-$f:\mathbb{F}_{2^k} \rightarrow \mathbb{F}_{2^k}$ 
-over the Galois field $\mathbb{F}_{2^k}$. 
-These circuits are custom-built, modular
-systems which cannot be synthesized due to their complex nature. Thus, 
-formal verification is needed to ensure they operate correctly.
-
-Recent computer-algebra based formal verification techniques have been
-able to perform verification of Galois field arithmetic circuits with
-a datapath size up to $163$-bits \cite{lv:phd}. Word-level abstractions 
-of Galois
-field arithmetic circuits could be used to further improve these formal 
-verification techniques to allow for verification of larger circuits, as
-well as provide the other benefits of word-level abstraction.
-However, there is currently no technique for computing word-level 
-abstractions of Galois field circuits of any practical size.
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.5\textwidth]{./figures/interpolate}
-}
-\caption{Circuit with $k$-bit input $A$ and $k$-bit output $Z$. 
-Abstraction to be derived as $Z=\Func(A)$.}
-\label{fig:abstractA_Z}
-\end{figure}
-}
-
-While the motivation comes from the need to verify Galois field arithmetic
-circuits, the presented approach can be generalized to be applicable to any
-combinational acyclic circuit.
-Any such circuit with a $k$-bit input $A$ and a $k$-bit output $Z$, such
-as the one shown in Fig. \ref{fig:abstractA_Z}, computes 
-$f: \mathbb{B}^k \rightarrow \mathbb{B}^k$ and can thus be analyzed as the
-function $f: \Fkk \rightarrow \Fkk$. Over $\Fkk$ this function can be 
-represented as the polynomial $Z=\Func(A)$. This is trivially generalized
-when there are multiple $k$-bit inputs $A_1,A_2,\dots,A_i$, i.e. $Z=\Func(A_1,\dots,A_i)$.
-Now assume the word-size of the input differs from the output, that is the circuit
-computes $f: \mathbb{B}^m \rightarrow \mathbb{B}^n$ for $m \neq n$. This can 
-be represented as a function over Galois fields as $f: \F_{2^m} \rightarrow \F_{2^n}$.
-This function can be analyzed over the field $\Fkk$ such that $\Fkk \supset \F_{2^m}$ 
-and $\Fkk \supset \F_{2^m}$, where $k=LCM(m,n)$.
-
-\subsection{Dissertation Contributions}
-To solve the problem of word-level abstraction, this dissertation proposes 
-a full solution consisting of three main contributions.
-
-\begin{enumerate}
-\item A theory for finding the word-level abstraction from a bit-level circuit over Galois fields is created.
-The given bit-level circuit implementation is modelled as a system of
-polynomials over the field.
-This theory is derived using techniques from computer-algebra, notably the theory of
-\Grobner basis \cite{pruss:iwls13}.
-\item Using this theory, new algorithms based on symbolic computation are developed to 
-derive the word-level abstraction. The algorithms are designed to be applicable to 
-industry-size arithmetic circuits over Galois fields \cite{pruss:dac14}\cite{sun:date15}.
-A complexity analysis of the algorithmic approach is also presented. Furthermore, the
-approach is also generalized to make it applicable to arbitrary combinational circuits.
-Finally, we show how the approach can be used to exploit the hierarchical structure of
-large Galois field multipliers designed over composite fields.
-\item A custom software tool implementation of the algorithmic approach is described, including
-an analysis of efficient data structures designed for this purpose \cite{pruss:tcad15}.
-\end{enumerate}
-
-Experiments show that the proposed solution can abstract canonical, word-level, 
-polynomial representations of Galois field arithmetic circuits up to $1024$-bits in
-size, while other contemporary approaches are infeasible beyond a $32$-bit designs.
-
-%an 
-%approach based on computer-algebra, algebraic geometry - notably the theory of
-%{\it Gr\"obner bases} \cite{gb_book} \cite{buchberger_thesis}.
-%and {\it elimination ideals} \cite{ideals:book}, as our approach for 
-%abstracting canonical word-level representations from bit-level Galois 
-%field arithmetic circuits. 
-%Computer-algebra techniques easily integrate with 
-%Galois field theory and allow for polynomial abstraction of the circuit over 
-%the field itself. Thus, if the circuit computes a function over some field $
-%\mathbb{F}_{2^k}$, the resulting abstraction is a word-level polynomial 
-%function over the same field $\mathbb{F}_{2^k}$.
-
-%The given bit-level circuit implementation is first modeled as a system of
-%polynomials over the field. Using the theory of elimination ideals, we 
-%derive an elimination term ordering and prove that by using this ordering we 
-%can obtain a canonical word-level representation of a circuit by 
-%computing a Gr\"obner basis of the polynomials. However, complexity 
-%of the computation of a Gr\"obner basis proves to be prohibitive for circuits
-%of practical size. Therefore, we employ a select sub-set of computations 
-%from the Gr\"obner basis theory to overcome this complexity and obtain the
-%abstraction. We prove that this simpler computation can be 
-%performed via a polynomial reduction process, and engineer a custom 
-%verification tool for this purpose. Using this approach, we are able to 
-%successfully 
-%abstract word-level representations of Galois field arithmetic circuits up 
-%to $571$-bits, which is the {\it largest recommended NIST standard} for ECC.
+\section{Objective, Motivation and Contribution of this Dissertation}
+\subsection{Word-level Reachability and FSM Traversal}
+\subsection{Application to Galois Field Multipliers}
+\subsection{Abstractions for Sequential Circuits}
+\subsection{Other Applications}
 
 \section{Dissertation Organization}
-The rest of this dissertation is organized as follows. Chapter
-\ref{ch:prev} reviews previous applicable work and highlights their
-drawbacks with respect to the canonical, word-level abstraction problem. 
-Chapter \ref{ch:prelim} describes the properties of Galois fields, 
-$\mathbb{F}_{2^k}$, and explains the process of constructing them.
-It also describes how to design arithmetic circuits over such fields, their 
-complexities, and the role of these circuits in Elliptic Curve Cryptography.
-Chapter \ref{ch:ideals} provides a theoretical background of 
-computer-algebra and Gr\"obner bases and explains their application
-to Galois fields. 
-Chapter \ref{ch:abstract} describes an approach to abstract 
-word-level polynomial representations of combinational circuits using a 
-Gr\"obner basis computation.
-Chapter \ref{ch:improv} improves on this word-level abstraction approach to 
-make it applicable to much larger circuits.
-Chapter \ref{ch:generalize} generalizes the abstraction approach to make
-it applicable to circuits with varying operand word-lengths. It also describes how the 
-approach can take advantage of the hierarchy of arithmetic circuits designed over
-composite fields.
-Chapter \ref{ch:implement} describes the implementation details of a 
-custom abstraction tool and gives experimental results of
-abstracting large Galois field multiplier circuits.
-Chapter \ref{ch:concl} concludes the dissertation and outlines potential future 
-research for continuation of this work.
+{\bf Proposal Organization:} The rest of the proposal is organized as
+follows: The following Subsection describes the intellectual merit of
+the proposed work.  Section \ref{sec:prelim} covers preliminary
+concepts and notation. Section \ref{sec:reach} describes the basic
+concept of word-level FSM traversal and the research problems to
+investigate. Section \ref{sec:craig} explores the algebraic geometry
+analog of Craig Interpolants. Section \ref{sec:core} describes
+algorithmic techniques to derive UNSAT cores of polynomial
+ideals. Section \ref{sec:plan} puts the whole research in perpective,
+by demonstrating  with the help of an example how abstraction via
+unsat cores in algebraic geometry can simplify BMC. 
+Section \ref{sec:impact} addresses the broader impact of the
+proposed activity and concludes the proposal.

### 2016-12-02 dissertation/intro.tex
dissertation v2:part intro completed

@@
 original specification models, i.e. {\it Sequential Equivalence
-  Checking   (SEC)}.
+  Checking   (SEC)}. When the specification is given as an arithmetic function
+which canonically represents the circuit, then the problem 
+becomes {\it functional verification} of sequential arithmetic circuits.
 
@@
 verification. Contemporary approaches employ various techniques to
-overcome this state-explosion problem: i) Bounded model checking
+overcome this state-explosion problem:
+
+\begin{enumerate}[{1)}]
+\item Bounded model checking
 \cite{bitlevel1} traverses the FSMs for a fixed number of steps $k$
 ($k$-BMC) to check whether a property violation can occur in $k$ or
-fewer steps. ii) Analyze over-approximations (or {\it abstractions})
+fewer steps.  
+\item Analyze over-approximations (or {\it abstractions})
 of the state-space. Abstraction proves properties on the system by
@@
 same properties as the original one, a process of refinement is
-needed. {\it Craig interpolants} \cite{craig-interpolate}
-\cite{pudlak:ci} are used as a means of abstraction and reachability
-analysis for linear temporal logic (LTL) model checking
-\cite{mcmillan:interpolation}. Similarly, counterexample guided
+needed. For example, counterexample guided
 abstraction refinement (CEGAR) \cite{cegar-journal} uses proofs of
-unsatisfiability (UNSAT cores) to refine the abstractions. iii) the
-recent breakthrough method of \cite{bradley:pdr} where the set of
+unsatisfiability (UNSAT cores) to refine the abstractions.
+\item The recent breakthrough method of \cite{bradley:pdr} where the set of
 over-approximations to forward reachable states are refined with
 inductive constraints -- property directed reachability (PDR). 
+\end{enumerate}
 
@@
 
-There are several advantages in exploiting word-level information for
-verification. A number of designs have  their
-datapaths and/or system-level models described as word-level RTL
-models.  Exploiting word-level instead of bit-level information is one
-way of abstraction -- a key technique to reduce the state space of a
-FSM. It has the effect of combining sets of states with similar 
-properties. During reachability analysis, if we use bit-level
-variables to  represent the states, the representations may become too
-large to handle. However, when a ``bundle" of bit-level variables are
-represented as {\it only one word-level variable}, the set of
-reachable states can be represented by a  word-level constraint
-expression; which may lower verification complexity.
-
-{\bf Overall Research Objective:} In this proposal we investigate a
-set of new, promising approaches for {\it word-level representation,
-reachability analysis and abstraction} with applications to SEC,
-$k$-BMC and abstraction-based unbounded LTL model checking
-(A-UMC). Our techniques operate at the word-level and they are based
-largely on the concepts from {\it algebraic geometry}.
-
 For word-level SEC, we are given two designs, or their corresponding
@@
 be solved in the context of word-level verification -- i.e. data
-representation, abstraction (Craig interpolation, UNSAT cores) and
+representation, abstraction using UNSAT cores and
 algorithm execution has to be carried out at word-level.
 
-{\bf The Algebraic Geometry Model and Rationale:} We will represent
-the FSMs -- the transition relations -- by means of a set of
-multi-variate polynomials with coefficients from the finite (Galois)
-field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
-$\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
-Boolean assignment to a set of $k$-bit state register variables
-$S=\{s_0,\dots,s_{k-1}\}$. Therefore, we can consider each ($k$-bit)
-state as a word-level element $S$ of the finite field
-$\Fkk$. Algorithms can directly operate on polynomials in word-level
-variable $S$.
+\section{Hardware Design and Verification Overview}
+Following the Moore's law, the level of integration on modern VLSI becomes very high, 
+such that a single digital VLSI design can implement a true system-level component.
+The design process also evolves from manual design with little validation, to 
+a formal 3-step procudure which needs collaboration of team with large number of 
+engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
+2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
+tools; 3) Verify, which is to verify the correctness of design and implementation.
 
-Boolean functions with $k$-bit inputs and $k$-bit outputs 
-$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
-$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
-field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
-\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
-there exists a unique canonical polynomial $\FF$ that describes $f$.
-This implies that one can derive a {\it  canonical, polynomial
-  abstraction} of the function as $Z = \FF(A)$ where $Z, A$ are
-word-level symbols representing $k$-bit operands. The concept also
-generalizes to functions with different input/output bit-vector sizes,
-i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
-polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
-{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
-This implies that the FSM's transition relations can be
-represented as polynomial functions (ideals) in $\Fkk$, and values of
-state variables can be represented as solutions to these polynomials
-(variety of the ideal). Subsequently, the ideal-variety correspondences
-in algebraic geometry can be applied to implement symbolic reasoning
-about state reachability. Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
-our model provides a single, unified and bit-precise representation
-for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
-constraints.
+Nowadays the verify step is usually completed by a team with specialities on test, verification and validation of 
+circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
+Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procudures starting from 
+register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
+physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
+to ensure the implementation is consistent with the original design intent. Verification is 
+an important quality control measure before sending design layout to the VLSI foundaries.
+Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
+waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
+the stability and accuracy of digital VLSI circuits, even small flaws or short-time failures can cause 
+huge loss, especially in medical applications, military facilities and financial systems.
+Therefore, it is of utmost importance to verify the correctness of VLSI designs.
 
-The decision and abstraction procedures in our setting will rely on
-the theory and technology of {\it \Grobner bases (GB)}. GB-based
-algebraic reasoning is very powerful; in fact it is known to be
-strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
-light of the above discussion, using concepts from algebraic geometry
-and \Grobner bases over $\Fkk$, we can introduce another dimension of
-word-level abstraction to the techniques in sequential verification.
-
-{\bf The core focus of this research:} The use of algebraic geometry
-for sequential circuit verification and symbolic model checking has
-been presented before. Avrunin presented the
-concept of symbolic MC using algebraic geometry in
-\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
-GB-algorithms for CTL, LTL and 
-bounded MC over Boolean rings. However, these approaches are a
-straight-forward transformation of the problem to {\it bit-level}
-Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
-the concepts of word-level reachability, abstraction-refinement using
-interpolation or UNSAT cores, etc., that we desire was not the focus of
-\cite{Avrunin:CAV} \cite{vardi-iasted07}.
-
-In this research, we will focus on three specific research problems: 
-\begin{enumerate}
-\item How to perform reachability analysis at the word-level? This
-  of course forms the foundation for word-level SEC and MC.
-\item What is the analog of Craig interpolation in an algebraic
-  geometry setting (theoretical challenge), and how can we compute
-  various interpolants using GB-engines efficiently (practical
-  implementation)? This will enable word-level abstraction of the
-  state-space. 
-\item Explore the concept, and the computation, of unsatisfiable (UNSAT)
-  cores of  a set of polynomials. Many abstraction-refinement
-  techniques make use of UNSAT cores of CNF-formulas in MC. Therefore,
-  our investigation of UNSAT cores of polynomial ideals will enable
-  similar applications at word-level. 
-\end{enumerate}
-
-\section{Design, Test and Verification of Modern Sequential Circuit Components}\section{Hardware Design and Verification Overview}
-The typical design flow of a hardware system, 
-as shown in Fig. \ref{fig:cadflow}, starts with a hardware system 
-specification, which describes the necessary functions and parameters that 
-the system must perform and adhere to. The specification is typically
-modelled using a transaction-level model (\emph{TLM}), which describes 
-communication details between large circuit modules. 
-The TLM is then translated 
-into a register-transfer-level (\emph{RTL}) description, which is composed of abstracted, 
-interconnected circuit blocks that compose the entire system. RTL is 
-typically implemented 
-in hardware description languages (\emph{HDL}) such as Verilog and VHDL, 
-which are the most popular choices in the industry.
-Next, the RTL is optimized and converted into a \emph{netlist}, i.e. a large 
-collection of small physical blocks (MOSFET, Boolean logic gates, etc.) and 
-the inter-connections (wires) between them. 
-Lastly, the netlist is further optimized and then mapped onto a physical space 
-on a chip, which is then sent off for fabrication. This entire design flow
-is automated by Computer-Aided Design (\emph{CAD}) tools.
+One way to perform the verification is called {\it simulation}. It is defined as the collage of all circuit validation 
+methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
+However, simulation is not a complete solution to circuit verification problems. In modern designs with 
+large number of logic components and complicated architecures, it is impractical to simulate all possible 
+test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
+cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
+is a good example when simulation fails. Failure occurs when only 5 entries in an $84\times 16$ look-up table
+(LUT) are activated,  which is rarely used in most divisions. Because of the limitation of simulation,  
+test engineers from Intel failed to detect the bug,  which brought $\$475$ million dollors recalling bill
+for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.
 
@@
 \centerline{
-\includegraphics[width=0.5\textwidth]{./figures/designFlow}
+\includegraphics[width=0.7\textwidth]{newfig/designflow.eps}
 }
-\caption{Typical hardware design flow.}
-\label{fig:cadflow}
+\caption{Typical hardware design flow}
+\label{fig:designflow}
 \end{figure}
@@
 
-%\subsection{The Hardware Verification Imperative}
-When moving from one abstraction level of the hardware design process to the 
-next, an important issue arises: how can one ensure that the functionality of 
-the optimized design matches original spec? 
-Bugs in hardware design which are 
-not caught early can have costly effects later, such as the need for a 
-redesign. Bugs in arithmetic circuits can be especially catastrophic.
-One infamous example is the 1994 floating point division (FDIV)
-bug that affected the Intel Pentium chip \cite{nicely:FDIV}, 
-and subsequently cost the company \$475 million because it was  
-discovered after the chip's release. In another more fatal case, during the
-Gulf war, an American Patriot Missile battery failed to intercept an incoming
-enemy missile due to an arithmetic error \cite{arnold:patriot}.
-Since hardware bugs can have significant consequences, 
-there has been extensive work in field of hardware verification to find and
-eliminate bugs prior to fabrication.
-
-The two main methodologies used in hardware verification are simulation and 
-formal verification. \emph{Simulation} checks correctness by applying exhaustive 
-assignments to the circuit inputs and verifying correctness of the output. 
-This ensures that the circuit performs as designed under all possible 
-inputs. Such exhaustive testing is quite effective for smaller circuits. 
-However, as the size of the circuit increases, it becomes 
-computationally infeasible to simulate all possible test vectors. This is the 
-case with Galois field arithmetic circuits, which are commonly very large in 
-real-world applications. Often for such large circuits, simulations of a smaller and more 
-manageable subset of test vectors are employed to catch bugs. While these tests
-can increase confidence in the correctness of the design,  
-{\it they do not guarantee correctness} since every data-flow of the design hasn't
-been analyzed.
-%Such is the case with large Galois field arithmetic circuits, so we instead 
-%focus on formal verification.
-
 \section{Formal Verification}
-Instead of simulating input vectors, \emph{formal verification} utilizes 
+% TODO this part require less modifications
+Another method developed besides simulation is \emph{formal verification}, it utilizes 
 mathematical theory to reason about the correctness of hardware designs.
-%which overcomes some limitations of simulation. 
+Formal verification can provide $100\%$ fault coverage from 2 aspects. On the one hand,  it formalize properties
+for the circuit model which are irrelevant to specific input signals,  and prove them mathematically.
+On the other hand,  it adopts formal languages to strictly describe the design intents and detailed implementation, 
+and deduces circuit function from the implementation. These descriptions are named as {\it specifications}.
 Formal verification has two main forms: property checking and equivalence 
@@
 of the circuits they verify.
-Abstraction is defined as state-space reduction, i.e{\text . }abstraction
-reduces state-space by mapping the set of states of a system to a smaller 
-set of states. Because the new representation contains fewer states, it
-is easier to comprehend and thus easier to use. 
-Word-level abstraction focuses specifically on abstracting a word-level
-representation of a circuit out of a bit-level representation. For example,
-a bit-level representation of an integer multiplier is represented by a
-collection of Boolean inputs and outputs, whereas a word-level
-abstraction hides the underlying logic and represents the circuit as two 
-integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
-multiplier increases, the logical implementation of the multiplier grows (typically
-exponentially) while the word-level abstraction stays the same.
+There are several advantages in exploiting word-level information for
+verification. A number of designs have  their
+datapaths and/or system-level models described as word-level RTL
+models.  Exploiting word-level instead of bit-level information is one
+way of abstraction -- a key technique to reduce the state space of a
+FSM. It has the effect of combining sets of states with similar 
+properties. During reachability analysis, if we use bit-level
+variables to  represent the states, the representations may become too
+large to handle. However, when a ``bundle" of bit-level variables are
+represented as {\it only one word-level variable}, the set of
+reachable states can be represented by a  word-level constraint
+expression; which may lower verification complexity.
+
+% Abstraction is defined as state-space reduction, i.e{\text . }abstraction
+% reduces state-space by mapping the set of states of a system to a smaller 
+% set of states. Because the new representation contains fewer states, it
+% is easier to comprehend and thus easier to use. 
+% Word-level abstraction focuses specifically on abstracting a word-level
+% representation of a circuit out of a bit-level representation. For example,
+% a bit-level representation of an integer multiplier is represented by a
+% collection of Boolean inputs and outputs, whereas a word-level
+% abstraction hides the underlying logic and represents the circuit as two 
+% integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
+% multiplier increases, the logical implementation of the multiplier grows (typically
+% exponentially) while the word-level abstraction stays the same.
 
@@
 
-Finally, abstractions can also be applied to detect malicious 
-modifications to a circuit, potentially inserted as a hardware trojan horse.
-Hardware trojans, a relatively new security concern in the hardware 
-industry, use certain techniques to add incorrect behavior to a 
-design. 
-This behavior is only activated under certain rare circumstances that only 
-the mal-intent designer has knowledge of.
-The behavior is purposely hidden and is very difficult to encounter during 
-simulation of the design. A manufactured chip with a subsystem 
-that contains a hardware trojan could compromise the entire system in which 
-it is used.
-In some hardware trojan cases, formal verification techniques may be applied 
-to catch a bug in a design and provide a counter-example which exercises it. 
-However, it can be difficult to tell whether the bug in the design was 
-introduced intentionally of not. On the other hand, word-level abstractions 
-of bit-level circuits {\it effectively reverse-engineer the true function 
-implemented by the circuit}, which could be used to determine the designer's 
-true intention.
+% Finally, abstractions can also be applied to detect malicious 
+% modifications to a circuit, potentially inserted as a hardware trojan horse.
+% Hardware trojans, a relatively new security concern in the hardware 
+% industry, use certain techniques to add incorrect behavior to a 
+% design. 
+% This behavior is only activated under certain rare circumstances that only 
+% the mal-intent designer has knowledge of.
+% The behavior is purposely hidden and is very difficult to encounter during 
+% simulation of the design. A manufactured chip with a subsystem 
+% that contains a hardware trojan could compromise the entire system in which 
+% it is used.
+% In some hardware trojan cases, formal verification techniques may be applied 
+% to catch a bug in a design and provide a counter-example which exercises it. 
+% However, it can be difficult to tell whether the bug in the design was 
+% introduced intentionally of not. On the other hand, word-level abstractions 
+% of bit-level circuits {\it effectively reverse-engineer the true function 
+% implemented by the circuit}, which could be used to determine the designer's 
+% true intention.
 
 \section{Objective, Motivation and Contribution of this Dissertation}
-\subsection{Word-level Reachability and FSM Traversal}
-\subsection{Application to Galois Field Multipliers}
-\subsection{Abstractions for Sequential Circuits}
-\subsection{Other Applications}
+This research proposes a
+set of new, promising approaches for {\it word-level representation,
+reachability analysis and abstraction} with applications to SEC, sequential GF
+arithmetic functional verification and 
+$k$-BMC with abstraction refinement. Our techniques operate at the word-level and they are based
+largely on the concepts from {\it algebraic geometry}.
+
+\subsection{Motivating Application}
+
+The use of algebraic geometry
+for sequential circuit verification and symbolic model checking has
+been presented before. Avrunin presented the
+concept of symbolic MC using algebraic geometry in
+\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
+GB-algorithms for CTL, LTL and 
+bounded MC over Boolean rings. However, these approaches are a
+straight-forward transformation of the problem to {\it bit-level}
+Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
+the concepts of word-level reachability, abstraction-refinement using
+interpolation or UNSAT cores, etc., that we desire was not the focus of
+\cite{Avrunin:CAV} \cite{vardi-iasted07}.
+
+{\bf The Algebraic Geometry Model and Rationale:} We will represent
+the FSMs -- the transition relations -- by means of a set of
+multi-variate polynomials with coefficients from the finite (Galois)
+field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
+$\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
+Boolean assignment to a set of $k$-bit state register variables
+$S=\{s_0,\dots,s_{k-1}\}$. Therefore, we can consider each ($k$-bit)
+state as a word-level element $S$ of the finite field
+$\Fkk$. Algorithms can directly operate on polynomials in word-level
+variable $S$.
+
+Boolean functions with $k$-bit inputs and $k$-bit outputs 
+$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
+$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
+field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
+\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
+there exists a unique canonical polynomial $\Func$ that describes $f$.
+This implies that one can derive a {\it  canonical, polynomial
+  abstraction} of the function as $Z = \Func(A)$ where $Z, A$ are
+word-level symbols representing $k$-bit operands. The concept also
+generalizes to functions with different input/output bit-vector sizes,
+i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
+polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
+{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
+This implies that the FSM's transition relations can be
+represented as polynomial functions (ideals) in $\Fkk$, and values of
+state variables can be represented as solutions to these polynomials
+(variety of the ideal). Subsequently, the ideal-variety correspondences
+in algebraic geometry can be applied to implement symbolic reasoning
+about state reachability. Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
+our model provides a single, unified and bit-precise representation
+for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
+constraints.
+
+The decision and abstraction procedures in our setting will rely on
+the theory and technology of {\it \Grobner bases (GB)}. GB-based
+algebraic reasoning is very powerful; in fact it is known to be
+strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
+light of the above discussion, using concepts from algebraic geometry
+and \Grobner bases over $\Fkk$, we can introduce another dimension of
+word-level abstraction to the techniques in sequential verification.
+
+\subsection{Dissertation Contributions}
+% TODO revise this to contributions
+In this dissertation, we focus on sequential circuit verification problems and 
+propose solutions consisting of three main contributions: 
+\begin{enumerate}[{1)}]
+\item A method to perform reachability analysis at the word-level. 
+	The given FSM is modeled as a system of polynomials over finite field,
+	where the state space is mapped to its solution space.
+	Our proposed algorithm utilize concepts in algebraic geometry including ideals and varieties.
+	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
+  state-space.
+\item Using the theory of FSM traversal, we apply the algorithm to verify the 
+function of sequential GF arithmetic circuits. Our proposed approach uses GB-engines efficiently 
+and can verify sequential multipliers with 162-bit datapaths.
+\item We explore the concept, and the computation, of unsatisfiable (UNSAT)
+  cores of  a set of polynomials using Gr\"obner bases calculation. We apply the UNSAT core extraction 
+  to abstraction-refinement
+  techniques such as bounded model checking (BMC). 
+\end{enumerate}
 
 \section{Dissertation Organization}
-{\bf Proposal Organization:} The rest of the proposal is organized as
-follows: The following Subsection describes the intellectual merit of
-the proposed work.  Section \ref{sec:prelim} covers preliminary
-concepts and notation. Section \ref{sec:reach} describes the basic
-concept of word-level FSM traversal and the research problems to
-investigate. Section \ref{sec:craig} explores the algebraic geometry
-analog of Craig Interpolants. Section \ref{sec:core} describes
+The rest of the dissertation is organized as
+follows: Chapter \ref{ch:prev} reviews previous work, and analyzes their drawbacks with respect to 
+the word-level sequential verification problem.
+Chapter \ref{ch:prelim_GF} covers preliminary
+concepts and notation on finite fields, and the methodology about design of multipliers in finite fields.
+Chapter \ref{ch:ideals} provides a theoretical background of algebraic geometry and Gr\"obner bases.
+Chapter \ref{ch:reacha} describes the basic
+concept of word-level FSM traversal and introduces our proposed word-level FSM traversal algorithm. 
+Chapter \ref{ch:normal} explores the application of FSM traversal algorithm on 
+functional verification of sequential GF arithmetic circuits. Chapter \ref{ch:UNSAT} describes
 algorithmic techniques to derive UNSAT cores of polynomial
-ideals. Section \ref{sec:plan} puts the whole research in perpective,
-by demonstrating  with the help of an example how abstraction via
-unsat cores in algebraic geometry can simplify BMC. 
-Section \ref{sec:impact} addresses the broader impact of the
-proposed activity and concludes the proposal.
+ideals. It also demonstrates with the help of an example how abstraction via
+UNSAT cores in algebraic geometry can simplify BMC. 
+Chapter \ref{ch:conclude} concludes the dissertation and outlines potential future research for
+continuation of this work.

### 2016-12-02 dissertation/intro_xiaojun.tex
dissertation v2:part intro completed

--- 2016-10-24 dissertation/intro.tex
@@
-\chapter{Introduction}
-\label{ch:intro}
-\section{Design, Test and Verification of Modern Sequential Circuit Components}
+Following the Moore's law, the level of integration on modern VLSI becomes very high, 
+such that a single digital VLSI design can implement a true system-level component.
+The design process also evolves from manual design with little validation, to 
+a formal 3-step procudure which needs collaboration of team with large number of 
+engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
+2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
+tools; 3) Verify, which is to verify the correctness of design and implementation.
 
-\section{Formal Verification Techniques}
-\subsection{Conventional Formal Verification}
-\subsection{Word-level Verification Techniques}
+Nowadays the verify step is usually completed by a team with specialities on test, verification and validation of 
+circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
+Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procudures starting from 
+register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
+physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
+to ensure the implementation is consistent with the original design intent. Verification is 
+an important quality control measure before sending design layout to the VLSI foundaries.
+Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
+waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
+the stability and accuracy of digital VLSI circuits, even small flaws or short-time failures can cause 
+huge loss, especially in medical applications, military facilities and financial systems.
+Therefore, it is of utmost importance to verify the correctness of VLSI designs.
 
-\section{Objective, Motivation and Contribution of this Dissertation}
+One way to perform the verification is called {\it simulation}. It is defined as the collage of all circuit validation 
+methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
+However, simulation is not a complete solution to circuit verification problems. In modern designs with 
+large number of logic components and complicated architecures, it is impractical to simulate all possible 
+test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
+cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
+is a good example when simulation fails. Failure occurs when only 5 entries in an $84\times 16$ look-up table
+(LUT) are activated,  which is rarely used in most divisions. Because of the limitation of simulation,  
+test engineers from Intel failed to detect the bug,  which brought $\$475$ million dollors recalling bill
+for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.
 
-\section{Dissertation Organization}
+Formal verification can provide $100\%$ fault coverage from 2 aspects. On the one hand,  it formalize properties
+for the circuit model which are irrelevant to specific input signals,  and prove them mathematically.
+On the other hand,  it adopts formal languages to strictly describe the design intents and detailed implementation, 
+and deduces circuit function from the implementation. These descriptions are named as {\it specifications}.

### 2016-12-04 dissertation/prev.tex
dissertation v2: part prev update 1

@@
 \label{ch:prev}
-\section{SAT,DD,AIG-based Sequential Circuit Verification}
+\section{Sequential Circuit Verification based on Unrolling}
+The most significant difference of sequential circuits from combinational circuits is 
+that the outputs of the circuit depend not only 
+on the primary inputs, but also on current state. 
+The behavioral difference reflect on the structual design of circuits and result in 
+the existance of memory components such as latches and flip-flops.
+In order to test certain properties on some signals accross multiple clock-cycles,
+the most straightforward method is to propogate those signals throughout 
+all clock-cycles. Moreover, for formal verification, all signals on all paths from the circuit
+need to be propogated through multiple clock-cycles.
+This indicates a time-to-space conversion, where 
+a cut is set on the latches/flip-flops and the combinational part of circuit
+is copied over several time-frames then connected together.
+The procedure is called {\it unrolling} of sequential circuit, as 
+Figure \ref{fig:unrolling} shows.
 
-\section{Bounded Model Checking}
+% fig:unrolling
 
-\section{Model Checking with Abstraction Refinement}
+Unrolling provides a way to transform a sequential circuit to combinational 
+circuit. Therefore,  methods which can be applied to combinational circuit 
+verification is also suitable for unrolled sequential circuits. The canonical 
+graphical representation of the combinational circuit after unrolling is also 
+the canonical representation of the original sequential circuit. For the sequential 
+equivlance checking problem, we can also unroll the circuit to be verified and 
+specification to combinational ones, then perform combinational equivlance checking
+techniques. In the following part we review research and techniques which 
+can be applied to unrolled sequential circuits.
 
-\section{Word-level Techniques in Sequential Circuit Verification}
-Term rewriting?
+\subsection{Canonical Decision Diagrams}
+Canonical representations of Boolean functions have been the subject
+of extensive investigation for logic synthesis and design
+verification. The Reduced Ordered Binary Decision Diagram (ROBBD)
+\cite{BRYA86} was the first significant contribution in this area. 
+Efficient implementation of ROBDDs as a software package \cite{brace}
+allowed for efficient formal verification of combinational and
+sequential circuits.  ROBDDs represent a Boolean function as an
+implicit set of points on a canonical directed acyclic graph
+(DAG). Manipulation of Boolean functions can then be carried out as
+composition operations on their respective DAGs.  The decomposition
+principle behind BDDs is one of Shannon's expansion, i.e.
+\begin{equation}
+f(x, y, \dots) = x f_x + x' f_{x'}
+\end{equation}
 
-\section{Verification of Galois Field Circuits}
+where $f_x = f(x = 1)$ and $f_{x'} = f(x = 0)$ denote the positive and
+negative co-factors of $f$ w.r.t. $x$, respectively. Motivated by the
+success of BDDs,  variants of the Shannon's decomposition principle
+(Davio, Reed-Muller, etc.) were explored to develop other functional
+decision diagrams. For example, the AND-OR-NOT logic based Shannon's
+expansion is transformed into an AND-XOR logic based decomposition,
+termed as the Davio's decomposition:
 
-\section{Conventional UNSAT Core Extraction}
+\begin{eqnarray}
+f(x, y, \dots) &=& x f_x + x' f_{x'}\\
+& = & x f_x \oplus x' f_{x'}\\
+& = & x f_x \oplus (1 \oplus x) f_{x'}\\
+& = & f_{x'} \oplus x (f_x \oplus f_{x'})
+\end{eqnarray}
 
-\section{Conclusion Remarks}
+Decision diagrams based on such decompositions include FDDs
+\cite{okfdd}, ADDs \cite{add}, MTBDDs \cite{mtbdd}, and their hybrid 
+edge-valued counterparts, HDDs \cite{hdd} and EVBDDs \cite{evbdd}. 
+While these are referred to as {\it Word-Level Decision Diagrams}
+\cite{WLS}, the decomposition is still point-wise, binary, 
+w.r.t. each Boolean variable. These representations do not
+serve the purpose of word-level abstraction from bit-level
+representations.
+
+Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivatives K*BMDs
+\cite{kbmd} and *PHDDs \cite{phdd}, depart from the Boolean
+decomposition and perform the decomposition of a {\it linear} function
+based on its two moments. BMDs provide a compact representation for
+integer arithmetic circuits such as multipliers and squarers. However,
+these are inapplicable to word-level abstraction of modulo-arithmetic
+circuits over Galois fields.
+
+Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp} are a
+word-level canonical representation of a {\it polynomial expression},
+based on the Taylor's series expansion of a polynomial. However, they do
+not represent a {\it polynomial function} canonically. For example,
+$f_1 = 0$ and $f_2 = 2x^2 - 2x \pmod{ 4}$ are two different polynomial
+representations of the zero function over $\Z_4$; but they are
+symbolically different polynomials and they have non-isomorphic TED
+DAGs.  While \cite{namrata:phd}  and \cite{alizadeh:tcad2010} provide 
+canonical representations of polynomial functions, they do so over
+finite integer rings $\Z_{2^k}$ and not over Galois fields $\Fkk$.
+
+MODDs \cite{modd} \cite{modd_tcomp} are a DAG representation of the
+characteristic function of a circuit over Galois fields $\Fkk$. MODDs
+come very close to satisfying our requirements as a canonical
+word-level representation that can be employed over Galois fields, as
+it essentially interpolates a polynomial from the characteristic 
+function. However, MODDs do not scale very well for large circuits ---
+this is because every node in the DAG can have up to $k$ children and
+the normalization operations are very complicated for MODDs. They also
+suffer from the size explosion problem during intermediate
+computations. They are known to be infeasible in representing
+functions over $32$-bit operand words.
+
+% TODO find good shifts on this paragraph
+
+DD-based model checking is very efficient as long as the DDs of sequential 
+circuit can be setup. The existence of violating states in constructed DDs 
+immediately deduces the violation of property. However, when the design gets
+larger and larger, the time and space cost of building and storing the diagram 
+increases rapidly. In our experiment of verifying a $k$-bit arithmetic circuit 
+using ZDDs, when $k$ is larger than 100, the construction of ZDDs occupies 
+over $99\%$ runtime of the whole procedure.
+\subsection{Combinational Equivalence Checking Techniques}
+The verification problem addressed in this dissertation is a
+manifestation of the combinational equivalence checking (CEC) problem,
+where the specification (polynomial) and  the implementation (circuit)
+are custom-designed, structurally very dissimilar circuits. To make
+use of contemporary gate-level CEC tools, we can take the 
+specification circuit (``golden model'') and check its equivalence
+against the implementation circuit. Canonical decision diagrams (BDDs
+\cite{BRYA86} and their word-level variants \cite{WLS}),
+And-Invert-Graph (AIG) based reductions \cite{AIG:2002}
+\cite{alanmi:cec:iccad2006}, circuit-SAT solvers \cite{csat}, etc.,
+are among the many techniques that can be employed for this CEC.  
+When one circuit is synthesized from the 
+other, this problem can be efficiently solved using AIG-based
+reductions (e.g. the ABC tool \cite{abc}) and circuit-SAT solvers
+(e.g., CSAT \cite{csat}). 
+Synthesized circuits generally contain many
+sub-circuit equivalences which AIG and CSAT based tools can identify
+and exploit for verification. However, when the circuits are
+functionally equivalent but structurally very dissimilar (e.g.,
+Mastrovito \cite{mastro:1989} versus Montgomery implementations
+\cite{acar:1998} of Galois field circuits), none of the 
+  contemporary techniques, including ABC and CSAT, offer a practical
+  solution. Automatic formal verification of large {\it
+    custom-designed modulo-arithmetic circuits} largely remains
+  unsolved today.
+
+This verification problem is very hard for SAT solvers and also for
+quantifier-free bit-vector (QF-BV) theory based SMT-solvers, due to
+the large circuit size, and the presence of AND-XOR-SHIFT structures. 
+Similarly, the Cryptol tool-set \cite{Cryptol:fmcad09} also employs
+AIG-based reductions (SAT-sweeping) and SAT/SMT-solvers for
+verification of crypto-protocols. For applications where
+AIGs/SAT/SMT-techniques fail, the {\it Cryptol} tool-set also does not
+deliver. As shown in \cite{lv:phd}, {\it none of BDDs, SAT, SMT
+  solvers, nor the  ABC tool can prove design equivalence beyond
+  $16$-bit circuits.}
+
+%TODO add what I wrote for rolf & cunxi (refer to the new paper)%%%%%%%
+In \cite{ciesielski:flow}, the authors present a method
+for verification of integer multipliers using a data-flow
+approach. This work abstracts a polynomial function
+of the given multiplier and then solves the network flow problem 
+using algebraic techniques. However, the abstraction is 
+solely bit-level, and is thus not applicable to deriving 
+a word-level representation of a given design.
+
+\section{Word-level Techniques applied to Sequential Circuit Synthesis and Validation}
+% TODO Talk bit-level and bit-blast
+
+Other attempts to derive high-level representations of functions,
+along with associated decision procedures, can be found in
+the rich domain of formal model checking \cite{BHEL96} \cite{SMV},
+theorem proving \cite{arditi:bmd}, bit-vector SMT-solvers
+\cite{boolector} \cite{cvc3} \cite{z3} \cite{bitvector98}, automated
+decision procedures for Presburger arithmetic \cite{presburger}
+\cite{bultan:mixed_verification}, algebraic manipulation techniques 
+\cite{devadas:algebraic_manipulation_iccd91}, or the ones based on
+term re-writing \cite{AST}, etc. % TODO expand term rewriting
+
+Polynomial, integer and other non-linear representations have also
+been researched: Difference Decision Diagrams (DDDs) \cite{ddd-csl99} \cite{ddd-mt-98}, interval
+diagrams \cite{interval_dd}, interval analysis using polynomials
+\cite{polynomial_sanchez99}, etc. Most of these have found 
+application in constraint satisfaction for simulation-based
+validation:  \cite{Ritter99} \cite{hsat} \cite{lpsat}
+\cite{brinkmann:asp-dac} \cite{Huang:tcad01} \cite{bitvector98}. Among
+these, \cite{brinkmann:asp-dac} \cite{Huang:tcad01} \cite{bitvector98}
+have been used to {\it solve} integer modular arithmetic on linear
+expressions - a different application from {\it representing}
+finite field modulo-arithmetic on polynomials in a canonical form.
+
+% TODO Uninterpreted function abstraction
+
+\section{Abstraction Refinement Techniques}
+
+\section{Verification using Algebraic Geometry}
+
+\section{Concluding Remarks}
+For the problem of word-level, canonical, polynomial abstractions
+of Galois field arithmetic circuits over $\Fkk$, previous related work 
+is either inapplicable or only applicable to circuits no larger than
+$32$-bits in size.
+Therefore, we propose a {\it symbolic approach} to
+polynomial interpolation from a circuit using the Gr\"obner basis
+computation. However, the complexity of a \Grobner basis computation is 
+prohibitively expensive; thus, we propose further improvements to this 
+approach by deriving a smaller subset of computations based on a \Grobner
+basis analysis. These improvements allow for abstractions of 
+flattened Galois field circuits up to $571$-bits, which is the largest NIST 
+standard for ECC, or up to $1024$-bits when a hierarchy is given. 
+Furthermore, we propose applications of this 
+approach to allow for formal verification of flattened Galois field circuits up to 
+$1024$-bits, where current techniques are only applicable for circuits up 
+to $163$-bits.

### 2016-12-05 dissertation/intro.tex
dissertation v2:part prev finished

@@
 unsatisfiability (UNSAT cores) to refine the abstractions.
-\item The recent breakthrough method of \cite{bradley:pdr} where the set of
+\item The recent breakthrough method of \cite{bradley2011sat} where the set of
 over-approximations to forward reachable states are refined with
@@
 the control logic. While attempts have been made towards word-level
-predicate abstraction \cite{Jain05wordlevel} \cite{mcmillan:cav06}
-\cite{mcmillan:impact}, {\it using a purely word-level representation
+predicate abstraction \cite{jain2005word} \cite{mcmillan:cav06}
+\cite{mcmillan2010lazy}, {\it using a purely word-level representation
   of the state-space, the properties and their abstractions have not
@@
 the construction of a product machine; and the main research problem
-relates to that of performing {\it FSM traversal \cite{touati:iccad90}
-but  at the word-level}. Analogously, in the case of MC, the problem
+relates to that of performing {\it FSM traversal} \cite{touati1990implicit}
+but  at the word-level. Analogously, in the case of MC, the problem
 is setup w.r.t. a FSM $\mathcal{M}$, a set of initial states $S_0$ and

### 2016-12-05 dissertation/prev.tex
dissertation v2:part prev finished

--- 2016-10-24 dissertation/prev.tex
@@
 \label{ch:prev}
-\section{SAT,DD,AIG-based Sequential Circuit Verification}
+\section{Sequential Equivalence Checking}
+As an important component of formal verification for sequential circuits, SEC techniques 
+have been developed over decades and widely utilized in both academia and industry. 
+The specification of a sequential circuit can also be modeled as a golden state machine,
+SEC is performed to compare between the circuit for test and the golden one.
+The n\"aive method to implement SEC is: preload both circuits to the same initial states, 
+and assign their primary inputs to the same values during all clock-cycles.
+This method needs to be operated along with state space traversal,  therefore it is 
+less efficient. Moreover,  most SEC only check the primary outputs/inputs consistency 
+and does not require the 1:1 state correspondence,  so state space traversal is not always necessary.
 
-\section{Bounded Model Checking}
+Researchers proposed improvements by using Boolean functions represent 
+a set of states/transitions \cite{coudert2003unified, coudert1990verification},  or by dividing the sequential circuit to 
+smaller subcircuit and re-model the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
+interfaces that focuses on only designated initial states and remove redundancies in state space \cite{baumgartner2007scalable}.
 
-\section{Model Checking with Abstraction Refinement}
+Another direction to improve SEC algorithms is to avoid using state space traversal. 
+The forward retiming method \cite{van1998sequential} and time-frame merging \cite{stoffel1997record} 
+all work on an array of time-frames,  with the assistance of combinational equivalence checking (CEC)
+techniques. These techniques require structural similarities between the two circuits.
 
-\section{Word-level Techniques in Sequential Circuit Verification}
-Term rewriting?
+The most significant difference of sequential circuits from combinational circuits is 
+that the outputs of the circuit depend not only 
+on the primary inputs, but also on current state. 
+The behavioral difference reflect on the structual design of circuits and result in 
+the existance of memory components such as latches and flip-flops.
+In order to test certain properties on some signals accross multiple clock-cycles,
+the most straightforward method is to propogate those signals throughout 
+all clock-cycles. Moreover, for formal verification, all signals on all paths from the circuit
+need to be propogated through multiple clock-cycles.
+This indicates a time-to-space conversion, where 
+a cut is set on the latches/flip-flops and the combinational part of circuit
+is copied over several time-frames then connected together.
+The procedure is called {\it unrolling} of sequential circuit, as 
+Figure \ref{fig:unrolling} shows.
 
-\section{Verification of Galois Field Circuits}
+% fig:unrolling
 
-\section{Conventional UNSAT Core Extraction}
+Unrolling provides a way to transform a sequential circuit to combinational 
+circuit. Therefore,  methods which can be applied to combinational circuit 
+verification is also suitable for unrolled sequential circuits. The canonical 
+graphical representation of the combinational circuit after unrolling is also 
+the canonical representation of the original sequential circuit. For the sequential 
+equivlance checking problem, we can also unroll the circuit to be verified and 
+specification to combinational ones, then perform combinational equivlance checking
+techniques. In the following part we review research and techniques which 
+can be applied to unrolled sequential circuits.
 
-\section{Conclusion Remarks}
+\subsection{Canonical Decision Diagrams}
+The decision diagrams (DDs) are invented by inspirations of 
+using optimized data structures to accelerate formal verification.
+The most fundamental DD is the binary DD (BDD), which origins from 
+Shannon's expansion:
+\begin{equation}
+f(x, y, \dots) = x f_x + x' f_{x'}
+\end{equation}
+where $f_x = f(x = 1)$ and $f_{x'} = f(x = 0)$ denote the positive and
+negative co-factors of $f$ w.r.t. $x$, respectively.
+A BDD is usually represented as a binary tree.
+Its ordered and reduced form -- the Reduced Ordered Binary Decision Diagram (ROBBD)
+\cite{BRYA86}, was the first significant contribution because of its canonicality.  
+ ROBDDs represent a Boolean function as an
+implicit set of points on a canonical directed acyclic graph
+(DAG). Manipulation of Boolean functions can then be carried out as
+composition operations on their respective DAGs.
+
+Following BDDs,  variants of the Shannon's decomposition principle
+were explored to develop other functional decision diagrams such as
+ FDDs \cite{okfdd}, ADDs \cite{add}, MTBDDs \cite{mtbdd}, and their hybrid 
+edge-valued counterparts, HDDs \cite{hdd} and EVBDDs \cite{evbdd}. 
+Zero-suppressed BDDs (ZDDs) \cite{minato1993zero,minato1994calculation} use the if-then-else branches
+to represent the existence of variables in a cube, and result in lower 
+space complexity. It can be used to represent polynomials with integer coefficients.
+
+% fig: ROBDD vs ZDD
+
+DDs above are all based on bit-level operations. Even in the {\it Word-Level Decision Diagrams}
+\cite{WLS}, the decomposition is still point-wise, binary, 
+w.r.t. each Boolean variable. These representations do not
+serve the purpose of word-level abstraction from bit-level
+representations.
+
+Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivatives K*BMDs
+\cite{kbmd} and *PHDDs \cite{phdd}, perform the decomposition of a {\it linear} function
+based on its two moments instead of relying on Boolean decomposition. 
+MODDs \cite{modd} \cite{modd_tcomp} are a DAG representation of the
+characteristic function of a circuit over Galois fields $\Fkk$. 
+However, MODDs fails to work large circuits because of its poor scalability.
+
+Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp} are a
+word-level canonical representation of a {\it polynomial expression},
+based on the Taylor's series expansion of a polynomial. However, they do
+not represent a {\it polynomial function} canonically.
+
+The use of DDs in traditional formal verification has a lot of advantages. 
+For example, DD-based model checking is very efficient as long as the DDs of sequential 
+circuit can be setup. The existence of violating states in constructed DDs 
+immediately deduces the violation of property. However, when the design gets
+larger and larger, the time and space cost of building and storing the diagram 
+increases rapidly. In our experiment of verifying a $k$-bit arithmetic circuit 
+using ZDDs, when $k$ is larger than 100, the construction of ZDDs occupies 
+over $99\%$ runtime of the whole procedure.
+\subsection{Combinational Equivalence Checking Techniques}
+The CEC problem can be solved using methods from many aspects.
+Besides using canonical DDs (BDDs
+\cite{BRYA86} and their word-level variants \cite{WLS}),
+non-canonical representations And-Invert-Graph (AIG) based reductions \cite{AIG:2002}
+\cite{alanmi:cec:iccad2006} is also very effective. 
+Solvers for satisfiability problems (SAT) are good candidates to solve CEC problems,
+as long as the miter circuit of two circuits can be described using conjunctive normal form (CNF)
+formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, etc.,
+If the circuits being compared are structurally highly similar, AIG and circuit-SAT 
+based approaches are proved to be efficient.
+However, when the circuits are functionally equivalent but structurally very dissimilar (e.g.,
+Mastrovito \cite{mastro:1989} versus Montgomery implementations
+\cite{acar:1998} of Galois field circuits), none of the 
+  contemporary techniques including quantifier-free bit-vector 
+  (QF-BV) theory based SMT-solvers \cite{Cryptol:fmcad09}
+  offers a practical solution. As shown in \cite{lv:phd}, {\it none of BDDs, SAT, SMT
+  solvers, nor the  ABC tool can prove design equivalence beyond
+  $16$-bit circuits.}
+
+Recently integer polynomial based techniques \cite{ciesielski2014function,rolf:date16} are proposed  to verify the functional 
+correctness of integer arithmetic circuits. Their approach formulates the output signature as a polynomial function 
+with binary variables and integer coefficients, then rewrite the polynomial by substituting gate output with gate inputs
+variables. After going through the backward rewriting procedure using reverse topological variable ordering, the polynomial
+will be composed by only input variables. Then the polynomial is converted to a canonical representation, and compared
+with designated input signature. If they are equivalent, then the arithmetic circuit is successfully verified.
+This approach has a risk of polynomial-size-explosion during the backward rewriting. The authors proposed a heuristic
+to levelize arithmetic circuit, and substitute several gates' variables at the same time to minimize the risk. However, 
+the heuristic is proved to be less effective when the inner symmetry of circuit structure is missing.
+
+To conclude, automatic formal verification of large {\it
+    custom-designed modulo-arithmetic circuits} largely remains
+  unsolved today.
+
+\section{Symbolic Model Checking and Abstraction Refinement}
+Model checking is a way to verify certain safety and liveness properties 
+in sequential circuits. Symbolic model checking which avoids using explicit state encodings
+provides more flexibilities to reduce the state space and enhance the 
+efficiency of model checkers. The implementations of symbolic model checking 
+requires canonical DDs or SAT solvers \cite{burch1990sequential,burch1991representing,biere1999symbolic,biere1999symbolic}.
+
+Abstraction is a technique to minimize the state space representation by combining states with certain 
+similarities. Sometimes it can effectively lower the number of states that require analysis by orders of magnitude,
+without affecting the properties we need to verify. There is a kind of model checkers utilize abstracted models 
+with interpolation \cite{mcmillan2003interpolation,mcmillan:cav06}.
+At first, abstraction was done manually by designers. Clarke {\it et al.} \cite{clarke2000counterexample}
+proposed an BDD-based automated abstraction by removing spurious paths from analysis of counterexamples. 
+Zhang {\it et al.} \cite{zhang2005design} proposed another abstraction method based on CNF-SAT.
+It implemented latch abstraction by removing "irrelevant" latches from the analysis
+according to SAT/UNSAT analysis from the $k$-BMC. H. Jain {\it et al.} \cite{HimanshuDAC2005} improved the abstraction refinement technique of \cite{clarke2000counterexample},
+where they use CNF-SAT to perform the refinement instead of using BDDs. The new approach is applied to verify RTL Verilog
+and was known to be successful.
+
+$k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
+on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
+\cite{bradley2011sat} \cite{bradley2011incremental}. IC3 works incrementally to find out inductive subclauses
+of negations of reached states, meanwhile it is monolithic when computing over-approximations to sets of reachable
+states within $1,2,\dots,k$ steps. It is proved to be more efficient than interpolation based model checking
+although using similar mechanisms.
+
+Above techniques have limitations: they are all relying on bit-level informations from 
+the circuit, which prevent them from being applied to circuit with large datapaths.
+Meanwhile, their implementation replies on SAT/BDDs, which is an extension of Boolean 
+functions and not compatible with other forms of constraints.
+
+\section{Word-level Techniques applied to Sequential Circuit Synthesis and Validation}
+To better verify word-level designs, word-level verification techniques are 
+explored in recent years. Directly translating bit-vectors problem to bit-level 
+problems is called {\it bit-blasting}, and usually brings high computational complexity issues.
+Attempts to develop pure word-level techniques can be found in
+the rich domain of formal model checking \cite{BHEL96} \cite{SMV},
+theorem proving \cite{arditi:bmd}, bit-vector SMT-solvers
+\cite{boolector} \cite{cvc3} \cite{z3} \cite{bitvector98}, automated
+decision procedures for Presburger arithmetic \cite{presburger}
+\cite{bultan:mixed_verification}, algebraic manipulation techniques 
+\cite{devadas:algebraic_manipulation_iccd91}, or the ones based on
+term re-writing \cite{AST}, etc. % TODO expand term rewriting
+
+Polynomial, integer and other non-linear representations have also
+been researched: Difference Decision Diagrams (DDDs) \cite{ddd-csl99} \cite{ddd-mt-98}, interval
+diagrams \cite{interval_dd}, interval analysis using polynomials
+\cite{polynomial_sanchez99}, etc. Most of these have found 
+application in constraint satisfaction for simulation-based
+validation:  \cite{Ritter99} \cite{hsat} \cite{lpsat}
+\cite{brinkmann:asp-dac} \cite{Huang:tcad01} \cite{bitvector98}. Among
+these, \cite{brinkmann:asp-dac} \cite{Huang:tcad01} \cite{bitvector98}
+have been used to {\it solve} integer modular arithmetic on linear
+expressions -- a different application from {\it representing}
+finite field modulo-arithmetic on polynomials in a canonical form.
+
+Uninterpreted function abstraction is also an important category of 
+word-level techniques which facilitates word-level model checking.
+For uninterpreted functions, researchers manage to constrain them 
+using functional consistency between word variables' evaluations
+\cite{UF1,UF2,UF3}.
+
+\section{Verification using Algebraic Geometry}
+
+Symbolic computer algebra techniques have been employed for formal
+verification of circuits over $\Z_{2^k}$ and also over
+Galois fields $\Fkk$. 
+General verification techniques using Gr\"obner bases
+\cite{Avrunin:CAV} \cite{gbverify:2007} \cite{manna:program} are propose,
+but they did not address the problems of high computational complexity to
+compute Gr\"obner bases.
+
+Verification of a combinational GF arithmetic circuit $C$ against a
+polynomial specification $\F$ has been previously addressed \cite{ibm:blueveri}
+\cite{lv:tcad2013} \cite{pruss:dac14}. Verification problems in
+\cite{ibm:blueveri} \cite{lv:tcad2013} are formulated using
+Nullstellensatz and decided using the \Grobner basis algorithm.
+
+The paper 
+\cite{pruss:dac14} performs verification by deriving a canonical
+word-level polynomial representation $\F$ from the circuit $C$. Their
+approach views any arbitrary Boolean function (circuit) $f: \B^k
+\rightarrow \B^k$ as a polynomial function $f: \Fkk \rightarrow \Fkk$,
+and derives a canonical polynomial representation $\F$ over
+$\Fkk$. They show that this can be achieved by computing a reduced 
+\Grobner basis w.r.t. an {\it abstraction term order} derived from the
+circuit. Subsequently, they propose a \underline{r}efinement of this
+\underline{a}bstraction \underline{t}erm \underline{o}rder (called
+RATO), that enables to compute the \Grobner basis of a smaller subset
+of polynomials. The authors show that their approach can prove
+correctness of up to 571-bit combinational GF multipliers.
+
+\section{Concluding Remarks}
+From the investigation of previous work, techniques are to be researched that 
+can operate the FSM traversal on word-level to verify a property excluding spurious 
+faults. We propose to solve this problem in the context of word-level verification,
+with data representation, abstraction and algorithm execution all carried out 
+on word-level.

### 2016-12-06 dissertation/conclude.tex
dissertation v2:part future finished

@@
 \label{ch:conclude}
+This dissertation first present a new approach to perform reachability
+analysis of FSMs at the word-level. It is facilitated by reproducing 
+the implicit state enumeration algorithm in computer algebra and algebraic
+geometry domain. The image function part is mapped to variety projection on 
+next state variables. This projection is implemented by computing the 
+Gr\"obner basis of the elimination ideal with abstraction term orders.
+Moreover, the set operations in the state space is mapped to 
+the arithmetic of ideals, because algebraic geometry provides a 
+way to reason about the variety (solution) without actually solving the 
+system of polynomials. RATO is applied to improve the Gr\"obner bases 
+computation process, and a tool is developed to implement our 
+word-level FSM traversal algorithm. Experiments are performed to 
+analyze the reachability for circuit benchmarks and gain success.
+
+Next, we describe a method to execute functional verification on sequential Galois field
+multipliers over $\Fkk$.  The core algorithm is based on word-level unrolling of 
+a Moore machine, which borrows ideas from our word-level FSM traversal 
+algorithm. As a result, bit-level transition relations are represented 
+as a polynomial with word-level inputs/outputs.
+We implement our algorithm with both \textsc{Singular} platform and C++ toolset 
+and perform experiments on Agnew's SMPO and RH-SMPO.
+Our approach is able to verify up to 162-bit sequential
+circuits, whereas contemporary techniques fail beyond 23-bit
+datapaths.
+
+At last we explore abstraction refinement techniques to boost sequential circuit verification.
+UNSAT core extraction, as an indispensable component of many refinement algorithms 
+which rely on UNSAT informations, is thus proposed in polynomial rings.
+We use weak Nullstellensatz as essential theory of UNSAT core extraction, 
+then develop heuristics depending on the structural information of the refutation.
+An algorithm was implemented within the \textsc{Singular} computer algebra tool
+and experiments were conducted to validate the approach.
+
+Our approaches still have limitations: for word-level FSM traversal and 
+sequential GF multipliers verification, our method are more efficient
+on XOR-rich circuits, while most industrial designs are AND-OR gates dominant.
+For UNSAT core extractions, the abstraction refinement approach such as $k$-BMC 
+has only limited application on certain model checking problems. To overcome these limitations,
+further explorations are needed for my research.
 
 \section{Future Work}
+In this section, we point out some research problems that deserve further study.
 \subsection{Multivariate polynomial ideal based FSM Traversal}
-\subsection{New Diagram Structure accelerating Polynomial Reduction}
+In Chapter \ref{ch:reacha}, we always use a single word-level variable $T$ to denote 
+next state. However, some verification requires the bit-level outputs to be classified to 
+multiple sets. While the elimination and set union/intersection are compatible with multiple NS variables,
+the set complement requires an extension to Theorem \ref{thm:quotient} on ideals with 
+multivariate polynomial generators. We make a conjecture in details:
+
+\begin{Conjecture}
+Assume we have 2 ring variables $x,y$, and an ideal $J$ composed by 2 generators:
+$J = \langle f_y, v_x\rangle$. Here $f_y$ has the form $f_y = \mathcal F(x,y)$, where $\mathcal F$
+is a 2-argument function, for example $f_y = y^2+(1+x)y+x$. And $v_x$ is "vanishing polynomial" about
+$x$: $v_x = x^{q_1} - x$. Now we have another generator serving as "universal set"
+$J_0 = \langle v_x, v_y\rangle$, $v_y = y^{q_2} - y$.
+We conjecture that
+$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J).$$
+And we believe this conjecture is scalable such that it is also valid for even more variables and generators like 
+$J = \langle f_z, v_x,v_y,v_u,v_w\rangle$, $f_z =\mathcal F (z,x,y,u,w)$.
+\end{Conjecture}
+
+The following example illustrates a different operation from Example \ref{ex:SMPO}.
+\begin{Example}
+Recall that 
+$\{s_0, s_1\}$ are state/pseudo inputs, $\{t_0,t_1\}$ are state/pseudo outputs, and there is a primary input (1-bit) 
+$x$. We propose a new algorithm by modifying Algorithm \ref{alg:univa}
+
+\begin{algorithm}[hbt]
+\SetAlgoNoLine
+ \KwIn{Transition polynomial $f_t = T + \mathcal F (S,x)$, 
+	initial state ideal $from^0 = \langle S+\mathcal G(x), x^{q_1} - x\rangle$}
+
+  $reached = from^0(T\setminus S)$\;
+  \Repeat{$GB(new^i) == 1$}
+  {
+  	$i \gets i + 1$\;
+	$to^i \gets$GB$(\langle f_t, from^{i-1}\rangle) \setminus \mathcal H(S)$\;
+	$\overline{reached} = \langle T^{q_2}-T, x^{q_1} - x \rangle : reached$\;
+	$new^i \gets $GB$(to^i + \overline{reached})$\;
+  	$reached \gets $GB$( reached \cdot new^i)$\;
+	$from^i \gets new^i(S\setminus T)$\;
+  }
+\Return{$reached$}
+\caption {Algebraic Geometry based Traversal Algorithm (multivariate-generator ideals)}\label{alg:multi}
+\end{algorithm}
+
+Inputs of this algorithm are: transition polynomial, which is the result of word-level abstraction; initial 
+states description ideal, contains 2 generators defining PI could be any inputs and state inputs $S$ is a specific
+value. (Ex: $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
+$\langle S+x\cdot\alpha, x^2+x\rangle$ means initial states$=\{00,10\}$).
+
+Transition polynomial calculation uses Tim's abstraction and bit-level variable substitution. After building
+an elimination ideal, use RATO such that \emph{reverse\ topo\ order\ ckt\ variables }$> T > S > x$, the reduction
+remainder has the form $T+\mathcal F(s_0,s_1,x)$. From word definition $S+s_0+s_1\alpha$ we get
+$$s_0 = \alpha S^2+ (1+\alpha)S, s_1 = S^2+S$$
+Do substitution, the transition polynomial of example ckt is 
+$$f_T = T+S^3\cdot x+\alpha S^3+(1+\alpha)S^2\cdot x+S^2+S\cdot x+(1+\alpha)x+1$$
+
+Assume we start from state $\{11\}$. First iteration, it will reach state $\{01\}$. Line 4 is to compose an
+ideal with 2 generators from $from^0$ and transition polynomial $f_T$, compute its Gr\"obner basis. Note this ideal
+has the form
+\begin{equation}
+I_{tran} = \left\{
+             \begin{array}{c}
+             T+\mathcal F(S,x) \\
+             S + \mathcal G(x) \\
+             v_x
+             \end{array}  
+        \right.
+\end {equation}
+$v_x$ is a polynomial containing only $x$, initially it should be vanishing polynomial $x^{q_1}-x$, with the program executing
+it may be factorized.
+
+Consider Buchberger's algorithm, all generators' leading terms are relatively prime, so it is a GB itself. 
+Then we need to reduce this GB, we will find out $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
+and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. So, at last we will get a polynomial
+$T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
+$S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$
+
+Line 5 is the ideal quotient of universal set and reached states. In the first iteration, 
+$reached$ is the initial state $\langle T+1+\alpha, x^2+x \rangle$. Result of ideal quotient
+is $\langle T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle$ represents $\{00,01,10\}$.
+
+Line 6 is ideals' sum (intersection of their varieties), it is done by combining all generators
+from 2 ideals and compute GB. For first iteration result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.
+
+Line 7 is ideals' product (union of their varieties), it is done by multiplying all pairs of
+generators from both ideal. For first iteration result is $GB(\langle (T+1)(T+1+\alpha),
+(T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$ representing $\{01,11\}$.
+
+The traversal will run 3 iterations and terminate at 4th iteration. I list all intermediate 
+results below:
+\begin{itemize}
+\item Iteration 1: $from^0 = \langle S+1+\alpha, x^2+x\rangle, to^1 = \langle T+1, x^2+x\rangle,
+ reached = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$
+\item Iteration 2: $from^1 = \langle S+1, x^2+x\rangle, to^2= \langle T+\alpha, x^2+x\rangle,
+reached = \langle T^3+1, x^2+x\rangle$
+\item Iteration 3: $from^2 = \langle S+\alpha, x^2+x\rangle, to^3 = \langle T+\alpha x, x^2+x
+\rangle, reached = \langle T^3\cdot x+x, T^4+T, x^2+x\rangle$
+\item Iteration 4: $from^3 = \langle S, x\rangle, to^4 = \langle T+1, x\rangle, new = \langle1\rangle$
+\end{itemize}
+The final reachable states are represented by ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
+which means $\{00,01,10,11\}$.
+\end{Example}
+
+\subsection{Use F-4 and ZDDs to Accelerate Polynomial Reduction}
+In Chapter \ref{ch:reacha} and Chapter \ref{ch:normal}, by using RATO we transform the GB 
+computation to a multivariate polynomial division. However, this division (reduction)
+is still with exponential complexity at worst case. In a situation where 
+a chain of OR gate exists, the size of polynomial will explode. For example, a chain of OR gates 
+which can be written as Boolean function 
+$$f = ((a\lor b) \lor c) \lor d$$
+equals to following polynomial function in $\F_2$:
+$$f = abcd+abc+abd+ab+acd+ac+ad+a+bcd+bc+bd+b+cd+c+d$$
+Notice $f$ contains $2^4-1 = 15$ terms, which is exponential to number of variables.
+The polynomial size explosion is a major factor affect the efficiency of polynomial division.
+
+One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
+Analysis on experiment results shows major time consumption is on multivariate polynomial division procedure.
+A matrix-based technique named as "F-4 style reduction" \cite{f4} can speed up the procedure
+dividing a low-degree polynomial with term-sparse polynomial ideal.
+
+Another way is to utilize DDs, {\it e.g.} ZDDs. ZDDs can represent unate covers of Boolean formulas.
+We can represent $f$ using a ZDD, while the dotted edges denote the corresponding literal does not 
+exist in a minterm.
+
+% fig:ZDD
+
+From Figure \ref{fig:ZDD} we find the size of the ZDD is $2\times4 -1 = 7$, which is linear to the number of variables.
+The reduction process using ZDDs can be executed as in \cite{polybori:2009}. The graphical illustration 
+of reduction in ZDDs is shown in Figure \ref{fig:reduce_steps}.
+
+% fig:reduce_steps
+
 \subsection{Interpolation extraction using GB Algorithm}
-\subsection{Verification of Integer Arithmetic Circuits}
+The concept of Craig Interpolants (CI) and their existence comes from
+symbolic logic \cite{craig-interpolate}; later, algorithms were
+presented to find the CI for Boolean formulae
+\cite{pudlak:ci} \cite{mcmillan:interpolation}. Assume that Boolean
+formulae are represented in Clause Normal Form (CNF) as 
+$f = C_1 \wedge C_2 \wedge \dots \wedge C_m$ where: 1) Each clause
+$C_i$ is a disjunction (Boolean OR, denoted $\vee$) of literals; 2)
+Each literal is a Boolean variable $x_i$ or its complement
+$\overline{x_i}$. 
+The Boolean satisfiability (SAT) problem requires that we find an
+assignment to the variables such that the formula $f$ is satisfied
+(SAT), or otherwise prove that no such assignment exists (UNSAT). 
+A CI is related to an UNSAT formula.
+
+\begin{Definition}
+(From \cite{mcmillan:interpolation}) Let $A$ and $B$ be Boolean
+  formulae given as sets of clauses such that $A \wedge B$ is
+  unsatisfiable (UNSAT). Then there exists a formula $P$ such that: 
+1) $A$ implies $P$ (or $A \subseteq P$); ~2) $P \wedge B$ is UNSAT; ~iii)
+$P$ refers to only the common variables of $A$ and $B$. The formula
+$P$ represents an {\bf interpolant} of $A$ and $B$.
+
+Given  the pair $(A, B)$ and their refutation proof, a procedure
+called interpolation system constructs an interpolant in linear time
+and space in the size of the proof \cite{mcmillan:interpolation}
+\cite{pudlak:ci}. 
+\end{Definition}
+
+\begin{Example}\label{ex1}
+Let $f = (\overline{d})(\overline{c})(\overline{a}\vee d)(a
+\vee b \vee c)(\overline{b})$ be a CNF formula. Let $f = A \wedge B =
+\emptyset$, where $A = (\overline{d})(\overline{c})(\overline{a}\vee
+d)$ and  $B = (a \vee b \vee c)(\overline{b})$. Then $P = \overline{a}
+\wedge \overline{c}$ is an interpolant of $(A, B)$. 
+\end{Example}
+
+CIs are used to derive abstractions to produce over-approximate image
+operators in model checking \cite{mcmillan:interpolation}. Since $A
+\implies P$, $P$ contains $A$ and is an {\it abstraction} of $A$. It
+also has fewer variables, so checking  invariants on $P \wedge B$ is
+easier. The interpolant is derived through a resolution proof of the
+SAT problem. There can be many interpolants $P_i$ for a pair $(A,B)$; 
+however, it is not feasible to explore a few or all of these
+interpolants by means of the resolution proof.
+
+\begin{Problem}
+{\it As we wish to perform MC with abstraction using algebraic
+  geometry, we introduce the algebraic geometry analog of CI
+  and explore algorithms to compute them. We conjecture that the
+  concept of CI should be related to elimination ideals, so our line
+  of investigation will focus on \Grobner basis computations with
+  elimination term orders for their computation.
+}
+\end{Problem}
+
+\begin{Definition}\label{ci}
+Let $F = \{f_1, \dots, f_s\}$ be a set of polynomials in the ring 
+$R = \Fq[x_1,\dots,x_n]$. Let $F = F_A \cup F_B$ and ideals $J =
+\langle F \rangle, J_A = \langle F_A \rangle, J_B = \langle F_B
+\rangle$ be corresponding ideals in $R$ such that $J = J_A + J_B$. Let
+it be known (say, due to application of Weak Nullstellensatz and
+Gr\"obner basis) that the varieties $V_{\Fq}(J) = V_{\Fq}(J_A) \cap
+V_{\Fq}(J_B) =V_{\Fq}(J_A+J_B)=\emptyset$. Also, let the set of
+variables $X = \{x_1,\dots,x_n\} = X_A \cup X_c \cup X_B$ where $X_A,
+X_B$ are the set of variables present exclusively in the sets of
+polynomials $F_A, F_B$ respectively. Only $X_c$ is the set of
+variables that are common to both sets of polynomials $F_A,
+F_B$. Then, there exists a set of polynomials  $F_P$ and ideal $J_P =
+\langle F_P \rangle$ such that 
+\bi
+\item $V_{\Fq}(J_A) \subseteq V_{\Fq}(J_P)$
+\item $V_{\Fq}(J_P) \cap V_{\Fq}(J_B) = \emptyset$
+\item Polynomials of $F_P$ contain the common variables ($X_C$) of
+  $F_A, F_B$. 
+\ei
+
+We call the ideal $J_P = \langle F_P\rangle$ the {\bf algebraic
+  interpolant} of  $J_A+J_B$. 
+\end{Definition}
+
+In the above definition $V_{\Fq}(J_A+J_B)=\emptyset$ implies that the
+ideal $J_A+J_B$ has an empty variety (UNSAT) problem.
+
+\begin{Example} \label{ex2}
+Based on Example \ref{ex1}, we translated the system over
+$\mathbb{F}_2[a, b, c, d]$. Let $F_A = \{f_1, f_2, f_3\}$ and $F_B =
+\{f_4, f_5\}$ where: $f_1: d; ~~~f_2: c; ~~~f_3: a + da; f_4:  abc +
+ab + ac + bc + a + b + c + 1; ~~~f_5: b.$ The Boolean interpolant
+$\overline{a}\wedge\overline{c}$ from Example \ref{ex1} translates to
+${\mathbb{F}}_2$ as the polynomial $f_p = ac + a + c$, with its
+variety $V(f_p) = \{a=0, c=0\}$.   
+\end{Example}
+
+\begin{Problem}
+{\it How can this algebraic interpolant $f_p$ (or in the general case,
+  the set of polynomials $F_P$) be computed? As the interpolant $F_P$ 
+contains only variables $X_C$ that are common to the polynomial sets
+$F_A, F_B$, does this imply that the interpolants can be computed by
+means of \Grobner bases with an elimination term order $X_A>X_B>X_C$
+with $X_A, X_B$ eliminated from the problem? }
+\end{Problem}
+
+We believe that algebraic interpolation is strongly related
+to the GB computation with the elimination order $X_A>X_B>X_C$, and
+this relationship needs to be formally derived.
+
+\begin{Conjecture} \label{con:ci}
+Computations of algebraic interpolants: Let $J_0$ denote the ideal of
+all vanishing polynomials in $\Fkk[x_1,\dots,x_n]$.
+\bi
+\item  Compute a Gr\"obner basis $G_1 = GB(J_A + J_0)$ 
+with the elimination order $X_A > X_B > X_C$, and select
+$F_{P1}=G_1\cap\Fkk[X_C]$. We conjecture that the \Grobner
+basis $F_{P1}$ of the elimination ideal corresponds to an algebraic
+interpolant.
+
+\item Analogously, compute a Gr\"obner basis $G_2 = GB(J_B + J_0)$ 
+with the elimination order $X_B > X_A > X_C$, and select
+$F_{P2}=G_2\cap\Fkk[X_C]$. Find an ideal $F'_{P2}$ such that
+the variety $V(F'_{P2})$ is the complement of the variety
+$V(F_{P2})$. Then the set $F'_{P2}$ gives another interpolant.
+\ei
+\end{Conjecture}
+
+\begin{Example}\label{ex3}
+Consider the polynomials $\{f_1, \dots,f_5\}$ from Example
+\ref{ex2}. Computing $F_{P1}$ as described in Conjecture \ref{con:ci} 
+produces $F_{P1} = \{a, c\}$, correctly giving us the desired variety
+$V(a = 0, c = 0)$. Similarly, when we compute $F_{P2}$, we find that
+the interpolant is $ac + a + c + 1$. Notice that the variety 
+$V(ac+a+c+1)=\{(0,1), (1,0), (1,1)\}$, which is exactly the complement
+of $V(F_{P1})$.  
+\end{Example}
+
+The aforementioned experiments in Example \ref{ex3} do not invalidate
+our conjectures, therefore we will attempt to prove them. If we are
+able to prove our conjectures, then we will further study the proof to
+derive efficient algorithms to compute the algebraic
+interpolants. Moreover, the above experiment shows that there can be
+multiple ways of computing the interpolants. What can these \Grobner
+basis computations tell us about the number of valid algebraic
+interpolants for any given problem? Can they be classified as {\it weak
+  or strong interpolants} based on the sparsity of the polynomials
+(power of abstraction)?
+
+\begin{Problem}
+As there can be many interpolants for an ideal-pair $(J_A, J_B)$, we
+will further resolve these questions:
+\bi
+\item Find the minimal interpolant: i.e. find the interpolant $F_P$
+  such that $V(J_P)$ is the smallest variety larger than $V(J_A)$ such
+  that $V(J_P) \cap V(J_B) = \emptyset$.
+\item Analogously, find the maximal interpolant. 
+\item Over finite fields, the variety of an elimination ideal is
+  exactly equal to the projection of the variety on the un-eliminated
+  variables. Consider Fig. \ref{fig:projection}, where variety of the
+  ideals $J_A, J_B$ are respectively projected on the common variables
+  $X_C = \{a, c\}$. Then, does computing $F_{P1}$ (resp. $F_{P2}$) 
+  deliver the minimal (resp. maximal) interpolant?
+\ei
+\end{Problem}
+
+\begin{figure}
+\centering{
+\includegraphics[width=3in]{newfig/project.eps}
+\caption{Algebraic interpolant: Projection of varieties on common variables? }
+\label{fig:projection}}
+\end{figure}
+
+Once these problems are understood and algorithms for the computation
+of algebraic interpolants derived, then we will integrate them 
+with word-level reachability analysis for abstraction-based model
+checking.
+
+\subsection{Technology Mapping for Word-level Functional Blocks}
+Technology mapping is an important problem in digital circuit synthesis.
+Designers are given a library of well-designed functional blocks (including IP cores) and a raw netlist, 
+technology mapping's objective is to map as many as blocks to the raw netlist and 
+keep the functional equivalence. Contemporary techniques rely on bit-wise 
+analysis on the signals to deduce the boundary of mapped blocks.
+It is possible to use the equivalence checking techniques proposed in this dissertation 
+as a alternative way to perform technology mapping, especially on word-level when 
+given blocks represent word-level functions.
+
+\begin{Problem}
+The objective of our approach is to map the macro blocks without boundary information.
+Mapping is an essential technique used in synthesis and verification. In synthesis, we can map the 
+macro functions with smaller and faster implementations to optimize the timing and area; in simulation,
+we can map a complicated function to a simple execution to accelerate simulation speed.
+
+Given a gate-level design $D$ and several word-level macro blocks $\{B_i\}$, we need to map macro
+blocks $B_i$ into design $D$, and write out the mapped design $D'$ which is equivalent to original
+design $D$. The objective is to generate a mapped design $D$ with as many of $B_i$ such that the area
+and timing is optimal.
+
+\begin{figure}[hbt]
+	\begin{center}
+	\includegraphics[scale=0.3]{macro.png}
+	\end{center}
+	\caption{The outline and flow of mapping for macro blocks}
+	\label{fig:macro}
+\end{figure}
+
+\end{Problem}
+
+This following part describes the sketch of our proposed approach base on loop invariant constraints \cite{sankaranarayanan2004non}.
+First, a transition system is modeled by algebraic assertions; then \emph{ideal membership test}
+is applied on the set of assertions to help abstract the loop invariant. "Template" is the concept
+I borrowed from this paper to apply on my own approach.
+
+1) {\bf Template and state constraints:} To effectively and efficiently utilize the ideal membership test technique, a heuristic to
+generate $\F$ is needed. One heuristic can provide better coverage for loop invariant abstraction as well as
+relatively small size is \emph{generic quadratic form}.
+
+For example, a pair of state variables $\{x,y\}$'s generic quadratic form is
+$$\F = a_0x^2+a_1x+a_2xy+a_3y+a_4y^2+a_5$$
+It covers all possible terms with degree at most 2. $a_0\sim a_5$ are usually real number parameters,
+a certain assignment about them can turn $\F$ into desired invariant. This parameterized constrain
+polynomial covers all combinations of state variables can also be called a \emph{template}.
+
+Finding a proper assignment is the major part of original paper and has potential to be even further explored
+beyond that paper, and/or to be applied on different research field.
+
+2) {\bf Initial state constraints:} For initial state, the constrains are explicit. A template is adopted and refined by Gr\"obner basis 
+generated by original constrains, by equaling the remainder to 0 we can get constrains on
+parameters from the template.
+
+An example is Fig.\ref{fig:loopinv}. The template is generic quadratic form of $\{s,i,j,j_0\}$, which is
+\begin{figure}[hbt]
+	\begin{center}
+	\includegraphics[scale=0.3]{../../../Pictures/Selection_006.png}
+	\end{center}
+	\caption{An example program of 2 Natural numbers' multiplication (Fig.1 from original paper)}
+	\label{fig:loopinv}
+\end{figure}
+\begin{align}
+\F =& a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
+&a_5i^2+a_6i+a_7ij+a_8ij_0+a_9j^2\nonumber\\
+&a_{10}j+a_{11}jj_0+a_{12}j_0^2+a_{13}j_0+a_{14}\nonumber
+\end{align}
+
+Constrains of initial state: $s=0\land j=j_0$ can be interpreted to polynomials:
+$\{s, j-j_0\}$. Check if it is Gr\"obner basis using \emph{Buchberger's algorithm}. Since
+their leading terms are relatively prime, we can write Gr\"obner basis $G=\{s, j-j_0\}$. Its ideal
+$J=\langle s,j-j_0\rangle$.
+
+Compute multi-division $\F \xrightarrow{G}_+ r$, the remainder is
+$r = a_5i^2+a_6i+(a_7+a_8)ij_0+(a_9+a_{11}+a_{12})j_0^2+(a_{10}+a_{13})j_0+a_{14}$.
+Let it equal to 0, each coefficient will generate a constrain, solution to the system is candidate
+assignment to generate invariant.
+\begin{equation}
+\left\{
+\begin{array}{l}
+a_5=a_6=a_{14}=0\\
+a_7+a_8=0\\
+a_9+a_{11}+a_{12}=0\\
+a_{10}+a_{13}=0
+\end{array}\right.
+\nonumber
+\end{equation}
+
+3) {\bf Modeling state transitions:}
+2 states get involved in a state transition. The technique requires to model the 2 states individually,
+do multi-division separately to get remainder $r_1$ and $r_2$. Suppose a constrain polynomial for
+this transition is $r_t$, we requires that when invariant for state 1 holds and transition $1\to 2$ stands,
+invariant for state 2 should also holds, which is
+$$(r_1 = 0)\land (r_t = 0) \implies (r_2 = 0)$$
+One reasonable conjecture is
+$$r_t = r_1 - \lambda r_2$$
+Theoretically $\lambda$ could be any polynomial. Consider the system to be easy to solve, we keep $\lambda$
+only taking value from real numbers.
+
+Still take Fig.\ref{fig:loopinv} as example (also refer to Example 10 in original paper). State 1 is initial state we 
+just characterized $\F = f(s,i,j,j_0)$, and the post state (state 2) have exactly the same form constrain:
+$\F' = f(s',i',j',j_0')$. Considering transition relation, substitute $s'$ with $s+i$, replace $j'$
+with $j-1$, and $i'$ for $i$, $j_0'$ for $j_0$, the template for state 2 is expression $f'$ in Example 10 from
+original paper.
+
+Do multi-division with Gr\"obner basis generated by transition relation, record its remainder $r_2$.
+From equality $r_2 = \lambda r_1$, we get constrains for parameters. Solve the system using covering-problem-like
+techniques (Part. \emph{Elimination by Splitting} in original paper). One branching result is:
+$a_0,a_2\sim a_6, a_9\sim a_{14}$ are all zero; $a_1=a_7=-a_8$. Then reduced remainder 
+$$r_2 = a_1s +(a_1-a_7)i+a_7ij+a_8ij_0 = a_1(s+ij-ij_0) = 0$$
+Desired invariant is 
+$$s = i(j_0- j)$$
+We knew the program calculates $i\times j_0$. Initially $s=0$, $j_0-j=0$, invariant holds;
+during each cycle, $s'=s+i, (j_0'-j') = (j_0-j)+1$, the invariant also holds. In conclusion, this is a
+loop invariant for program in Fig.\ref{fig:loopinv}.
+
+{\bf Our proposed approach on technology mapping:}
+Our approach borrows inspiration of "templates" from loop invariant paper. The difference is,
+we are not using template polynomial to describe the system, instead we add some templates in the ideal
+of macro blocks we want to map to simulate all possibilities.
+1) {\bf System abstraction:}
+The polynomial we use to test ideal membership should include all information of a circuit partition,
+this requires us to abstract information from the system and write it into only 1 polynomial.
+Usually this polynomial has the form like:
+$$Z + f(i_1,i_2,\dots,i_n)$$
+Here $Z$ is the output. When there is only one output, $Z$ will be a bit-level variable. However in most
+cases we have multiple outputs, this asks us to write $Z$ as a word-level variable. $f$ is a Boolean
+function about all inputs, and $i_1,i_2,\dots,i_n$ are all bit-level inputs.
+
+\begin{figure}[hbt]
+	\begin{center}
+	\includegraphics[scale=0.2]{tobemapped.png}
+	\end{center}
+	\caption{An example gate-level netlist to be mapped}
+	\label{fig:tobemapped}
+\end{figure}
+
+Fig.\ref{fig:tobemapped} shows an example circuit partition with $a1,b1,a0,b0$ as inputs and
+$Z = \{z0,z1\}$ as outputs. If we use elements from Galois field $F_{2^2}$ to represent word $Z$,
+we have $Z = z_0 + \alpha\cdot z_1$.
+
+The abstraction also uses the property of Gr\"obner basis. If we arrange a term ordering as
+$$Other\ circuit\ variables > output\ word\ Z > all\ bit\ level\ inputs$$
+and include all gates information, word level variable definition and vanishing polynomials,
+the reduced Gr\"obner basis will have a polynomial (generator) in the form of $Z + f(a1,b1,a0,b0)$,
+this is an application of "elimination using GB", where we eliminate all variables except output
+word variable $Z$ and inputs $a1,a0,b1,b0$.
+
+In this example:
+
+\begin{itemize}
+\item Term ordering: $z0,z1,t,g1,p1,ci,t1,t2,Z,a0,b0,a1,b1$
+\item Gate description (from table \ref{table:booltogalois_op}): $z0+t+g1+t\cdot g1, t+t1+t2+t1\cdot t2, g1+a1\cdot b1,
+			t1+a1\cdot ci, t2+b1\cdot ci, ci+a0\cdot b0, z1+p1+ci, p1+a1+b1$
+\item Word definition: $Z+z0+z1\cdot \alpha$
+\item Vanishing polynomials $(J_0)$: $z0^2+z0, z1^2+z1, t^2+t, g1^2+g1, t1^2+t1, t2^2+t2, p1^2+p1, ci^2+ci,
+			a0^2+a0, b0^2+b0, a1^2+a1, b1^2+b1, Z^4+Z$(note Z is 2-bit word)
+\end{itemize}
+
+The result is one polynomial in Gr\"obner basis with leading term $Z$:
+$$Z+a0\cdot b0\cdot a1+a0\cdot b0\cdot b1+\alpha\cdot a0\cdot b0+a1\cdot b1+\alpha\cdot a1+\alpha\cdot b1$$
+
+\begin{table}
+\centering
+\begin{tabular}{|c|c|} \hline
+Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
+$\overline{a}$ & $1 + a$\\ \hline
+$a\ and\ b$ & $ab$\\ \hline
+$a\ or\ b$ & $a + b + ab$\\ \hline
+$a \oplus b$ & $a + b$\\
+\hline\end{tabular}
+\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
+\label{table:booltogalois_op}
+\end{table}
+
+2) {\bf Templates on Boundary Information:}
+\begin{figure}[hbt]
+	\begin{center}
+	\includegraphics[scale=0.2]{template.png}
+	\end{center}
+	\caption{Standard 2-bit adder with input/output mapping}
+	\label{fig:template}
+\end{figure}
+
+Fig.\ref{fig:template} shows a 2-bit standard cell. It has 3 outputs mapping to 2 output pins,
+and 4 inputs mapped to 3 input pins (the rest pin is mapped to fixed 0/1 signal). Dashed connection
+lines show one possible mapping, to find out this kind of feasible mapping, we need to simulate
+all possible mappings, where the concept "template" can be used.\vspace{5mm}
+
+{\it Output template}: $z0+ct0_{z0}\cdot t0+ct1_{z0}\cdot t1+c0_{CO}\cdot CO+cn0$
+
+When $cn0 = 0$, any one of $ct0_{z0},ct1_{z0},c0_{CO}$ could be 1 means mapping to corresponding output
+pin. If $cn0 = 1$, mapping to negation of corresponding output pin.
+
+{\it Input template}: $a0+pa0+c0_{A0}\cdot s0+c1_{A0}\cdot s1+c2_{A0}\cdot s2$
+
+When $pa0 = 0$ and any one of $c0_{A0},c1_{A0},c2_{A0}$ equal to 1 means mapping to corresponding input
+pin; if all $c0_{A0},c1_{A0},c2_{A0}$ equal to 0 then means mapping to fixed signal "0" (when $pa0 = 1$ then
+mapping to fixed signal "1").

### 2016-12-06 dissertation/normal.tex
dissertation v2:part future finished

@@
 further explorations are needed for my research.
-
-One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
-Analysis on experiment results shows major time consumption is on ``multi-division" part.
-A matrix-based technique named as "F-4 style reduction" \cite{f4} can speed up the procedure
-dividing a low-degree polynomial with term-sparse polynomial ideal.

### 2016-12-07 dissertation/ONB.tex
dissertation v2:Appendix update 1

% LaTeX file for a 1 page document
\documentclass[12pt]{article}
%\usepackage{e-jc}

\title{Theory about Normal Basis on $F(2^k)$}

\author{Xiaojun Sun \ \ \ \ Prof. Priyank Kalla\\
\small Department of Electrical \& Computer Engineering\\[-0.8ex]
\small University of Utah, Salt Lake City\\
\small \texttt{\{xiaojun.sun,kalla\}@ece.utah.edu}
}

\date{Last Update: Apr 17, 2013}

\begin{document}
\maketitle

\begin{abstract}
N/A
\end{abstract}

\section{Introduction}

a) Why use NB?\\
b) Relations between NB \& StB?\\
c) What's ONB?\\
{\bf Simple stuff we've discussed.}

\section{Characterization of Normal Basis}
\textbf{Conceptions from Linear Algebra}

%\hspace{8mm}\\

\textit{\underline{Frobenius Map:} \\
$\sigma : x \rightarrow x^q, x \in F_{q^n}$ \\
Linear transformation of $F_{q^n}$ over $F_q$. \\
}

\textit{\underline{T-invariant subspace / cyclic vector:} \\
A subspace $W \subset V$ is called T-invariant when $Tu \in W \forall vector u \in W$\\
Subspace $Z(u,T) = <u,Tu,T^2u,...>$ is called T-cylic subspace of V. \\
If $Z(u,T) = V$, then u is called a cyclic vector of V for T.\\
}

\textit{\underline{Nullspace of polynomial:} \\
For any polynomial $g(x) \in F[x]$, the null space of g(T) consists of all vectors u such that
$g(T)u = 0$.\\
}

\textit{\underline{T-Order / minimal polynomial:} \\
For any vector $u \in V$, the monic polynomial $g(x) \in F[x]$ with smallest degree such that
$g(T)u = 0$ is called the T-Order of u or minimal polynomial of u.\\
That is, for an arbitrary element $\theta$ in $F_{q^n}$, find least positive integer k such that
$\sigma^k\theta = \sum_{i=0}^{k-1} c_i\sigma^i\theta$, then the $\sigma$-Order of $\theta$ can be 
written by $Ord_{\theta,\sigma}(x) = x^k - \sum_{i=0}^{k-1} c_ix^i$. \\
}

\textbf{Lemmas \& Theorems from linear algebra}
\begin{itemize}
\item[-] \textbf{Lemma 1}\ \ \ \ $g(x) \in F[x]$ and W is its null space. Let $d(x) = gcd(f(x),g(x)), e(x) = f(x)/d(x)$. Then $dim(W) = deg(d(x))$ and $W = \{e(T)u | u \in V\}$.
\item[-] \textbf{Lemma 2}\ \ \ \ Factorize f(x): $f(x) = \prod_{i=1}^{r} f_{i}^{d_i} (x)$, each $f_i(x)$ is prime to others. Assume $V_i$ be null space of $f_{i}^{d_i} (x)$, then $V = V_1\oplus V_2 \oplus ...\oplus V_r$.
Furthermore, define $\Psi_i(x) = f(x)/f_{i}^{d_i} (x)$. $\forall u_j \in V_j, u_j \neq 0, \Psi_i(T)u_j \neq 0$, only if $i = j$.
\item[-] \textbf{Lemma 3}\ \ \ \ Minimal and characteristic polynomial for $\sigma$ are both $x^n - 1$.
\item[-] \textbf{Corollary 1}\ \ \ \ An element $\alpha \in F_{q^n}$ is normal element if and only if $Ord_{\alpha,\sigma}(x) = x^n - 1$.
\item[-] \textbf{Theorem 1}\ \ \ \ Consider we are dealing with $F_{2^n}$, then field characteristic $p = 2$.
				Define $t = p^e$ where $n = kp^e, gcd(k,p) = 1$, so here t=1 if n is odd. Then 
			$x^n - 1$ can be factorized as $(\varphi_1(x)\varphi_2(x)\cdot\cdot\cdot\varphi_r(x))^t$.
			Additionally define $\Phi_i(x) = (x^n - 1)/\varphi_i(x)$. We get: \\ 
			An element $\alpha \in 	F_{q^n}$ is normal element if and only if $\Phi_i(\sigma)\alpha \neq 0, i = 1,2,...,r$.
\item[-] \textbf{Theorem 2}\ \ \ \ Let $W_i$ be the null space of $\varphi_{i}^{t} (x)$ and $\widetilde{W_i}$ the
				 null space of $\varphi_{i}^{t-1} (x)$. Let $\overline{W_i}$ be any subspace of
			 $W_i$ such that $W_i = \overline{W_i}\oplus \widetilde{W_i}$. Then\\
			$F_{q^n} = \displaystyle\sum_{i=1}^{r} \overline{W_i}\oplus \widetilde{W_i}$ \\
			is a direct sum where $dim(\overline{W_i}) = d_i$ and $dim(\widetilde{W_i}) = (t-1)d_i$.\\
			Furthermore, an element $\alpha \in F_{q^n}$ with $\alpha = \sum_{i=1}^{r} (\overline{\alpha_i} + \widetilde{\alpha_i}), \overline{\alpha_i} \in \overline{W_i}, \widetilde{\alpha_i} \in \widetilde{W_i}$,
			is a normal element if and only if $\overline{\alpha_i} \neq 0 \forall i = 1,2,...,r$.
\item[-] \textbf{Normal Basis Theorem for Finite Fields}\ \ \ \ There always exists a normal basis of $F_{q^n}$ over $F_q$.
\end{itemize}

\section{Algorithms for Normal Basis Construction}
\textbf{L\"uneburg's Algorithm}\\
Step 1: For each i = 0,1,...,n-1, compute $\sigma$-Order $f_i = Ord_{\alpha^i}(x)$. Here $x^n - 1 = lcm(f_0,f_1,...,f_{n-1})$.\\
Step 2: Apply factor refinement to $\{f_i\}$ and get $f_i = \prod_{1\leq j\leq r} g_{j}^{e_{ij}}, i = 0,1,...,n-1$.\\
Step 3: For each j, find an index $i_j$ (denote as i(j)) so that $e_{ij}$ is max in this j-th column.\\
Step 4: Let $h_j = f_{i(j)}/g_{j}^{e_{i(j)j}}$, take $\beta_j = h_j(\sigma)\alpha^{i(j)}$. Then\\
\indent\indent\indent\indent\indent\indent\indent\indent\indent$\beta = \displaystyle\sum_{j=1}^{r} \beta_j$\\
is the normal element.\\

\textbf{Preliminary to Lenstra's Algorithm}
\begin{itemize}
\item[-] \textbf{Lemma 4}\ \ \ \ For an arbitrary element $\theta \in F_{q^n}$ that $Ord_\theta(x) \neq x^n - 1$,
let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.\
\item[-] \textbf{Lemma 5}\ \ \ \ Same $\theta$ and $g(x)$ defined as last lemma. Assume there exists a solution 
$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Then there exists a non-zero element $\eta$ such that\\
$g(\sigma)\eta = 0$, and\\
$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$.

\end{itemize}

Suppose we take only one polynomial only contains T(next state), like\\
\hspace{8mm}\par
$ideal\ G = T^2 + (1+X) \cdot T + X$\\
\hspace{8mm}\par
Since we get G from GB based image function, so G is already a Groebner Basis itself.
Considering Theorem 1 $\sim$ 3, it's easy to do ideal operation if its generator is only one polynomial.
So using $G+G'$ we can get intersection, using $G \cdot G'$ we can get union. If we take V = Universe = $<vanishing\ polynomial>$, we can get ideal quotient $I(V):I(W)$ as complementary set for
specific varieties.

\hspace{8mm}\par
\textbf{More about Ideal Quotient}
\begin{itemize}
\item[-] \textbf{Definition 1}\ \ \ \ If I,J are ideals in $k[x_1,...,x_n]$, then $I:J$ is the set \\ \indent $\{f\in k[x_1,...,x_n]:fg\in I\ for \ all\ g\in J\}$
\end{itemize} \par
Why we can get complementary set $\overline {reached}$ through this?\par
First, we only care about varieties. Say we redefine the "equal" as $V(I)==V(J)\Leftrightarrow I == J$.\par
Second, we only care about the ideal/GB {\bf G} contains only 1 generator f. Say this polynomial f is a function about next state (word level) T, then $f(T=V(\bf{G}))==0$. To ensure then unique of generator f, we'll reduce f (or say {\bf G}) every time.\par
Third, for ideal quotient ${\bf U:G}$ here, we need to find f from $g~f==vanish$. This means $<f>\bigcup <g>==<vanish>$. How can we develop a division algorithm to get exactly a factor without any remainders? Let's see 3 examples following:\\
\hspace{8mm}\par
\indent $T^2+(1+X)\cdot T+X$\\
\hspace{8mm}\par
\indent $T^2+T$\\
\hspace{8mm}\par
\indent $T^2+1$\\

\end{document}

### 2016-12-07 dissertation/appendix.tex
dissertation v2:Appendix update 1

@@
 \label{append:NB}
+In Chapter \ref{ch:prelim_GF} we briefly introduced the concept of normal basis (NB), and 
+the benefits of using NB. In the section we describe more details about the normal basis theory by 
+characterizing NB from linear algebra perspective, constructing a general NB in an 
+arbitrary field and converting between NB and StdB. All theorems and lemmas refer to the 
+dissertation of S. Gao \cite{gao:phd_normal_basis}, and we deduce all the proofs for them.
 \subsection{Characterization of Normal Basis}
+In order to depict the characterization of NB, we need the introduce some concepts from linear algebra domain.
+\begin{Definition}[Frobenius Map]
+Define map $\sigma:x\to x^p,~x\in \F_{p^n}$. This map denoting the linear map of field extension $\F_{p^n}$
+over $\F_p$.
+\end{Definition}
+Additionally we can define a subspace based on linear map $T$:
+\begin{Definition}
+A subspace $W\subset V$ is called {\bf $T$-invariant} when 
+$$Tu \in W, ~~ \forall \text{vector }u\in W$$
+
+Subspace $Z(u,T) = \langle u,Tu,T^2u,\dots\rangle$ is called {\bf $T$-cyclic subspace} of $V$.
+If $Z(u,T) = V$ holds, then $u$ is called a {\bf cyclic vector} of $V$ for $T$.
+\end{Definition}
+Field elements $x$ constitutes polynomials in ring $\F[x]$. Subsequently we can define the 
+nullspace of polynomial:
+\begin{Definition}[Nullspace of polynomial]
+\label{def:nullspace}
+For any polynomial $g(x) \in \F[x]$, the null space of $g(T)$ consists of all vectors $u$ such that
+$g(T)u = 0$.
+\end{Definition}
+Finally we get to the most important concept: $T$-Order. Meanwhile it derives the construction of a ring extended by a field requires
+a minimal polynomial (while we use stronger irreducible polynomial $P(x)$ in Chapter \ref{ch:prelim_GF}).
+\begin{Definition}
+For any vector $u \in V$, the monic polynomial $g(x) \in \F[x]$ with smallest degree such that
+$g(T)u = 0$ is called the {\bf $T$-Order} of $u$ or {\bf minimal polynomial} of $u$.
+
+That is, for an arbitrary element $\theta$ in $F_{q^n}$, find least positive integer $k$ such that
+$\sigma^k\theta = \sum_{i=0}^{k-1} c_i\sigma^i\theta$, then the $\sigma$-Order of $\theta$ can be 
+written as $$Ord_{\theta,\sigma}(x) = x^k - \sum_{i=0}^{k-1} c_ix^i$$
+\end{Definition}
+
+Using the concepts introduced above along with other basic concepts in linear algebra and finite field theory,
+we can derive following theorems.
+
+\begin{Lemma}
+Given $g(x) \in \F[x]$ and $W$ is its nullspace. Let $d(x) = gcd(f(x),g(x)), e(x) = f(x)/d(x)$. 
+Then $dim(W) = deg(d(x))$ and $W = \{e(T)u~|~u \in V\}$.
+\end{Lemma}
+\begin{Proof}
+Assume $f(x)$ is the minimal and characteristic polynomial for $T$. Then according to Definition \ref{def:nullspace}
+we obtain 
+$$W = \{u\in V~|~g(T)u = 0\}$$
+
+Let $k$ be a polynomial whose degree is larger than $f(x)$, i.e. $deg(k(x))>deg(f(x))$, then
+$$f(x)|k(x)~~\text{iff}~k(T)=0$$
+
+Since the construction of $f(x)$ relies on $T$, $f(T) = 0$, we get
+$$\forall u, f(T)u = 0$$
+
+Consider $W(u,T)$ subsequently relies on $g(T)$, we deduce 
+$$dim(W) = deg(d(x))$$
+
+According to the definition of $W$, we have
+\begin{align*}
+\forall u,~~& e(T)u \in W \Longleftrightarrow \\
+& g(T)e(T)u = g(T)\frac{f(T)}{d(T)} u = h(T)\cdot f(T) u = 0
+\end{align*}
+\end{Proof}
 \subsection{Construction of General Normal Basis}

### 2016-12-07 dissertation/conclude.tex
dissertation v2:Appendix update 1

@@
 In Chapter \ref{ch:reacha}, we always use a single word-level variable $T$ to denote 
-next state. However, some verification requires the bit-level outputs to be classified to 
-multiple sets. While the elimination and set union/intersection are compatible with multiple NS variables,
+next state. However, in some situations we need to keep recording the relations between $T$ and
+inputs, {\it i.e.} the reached states contain multivariate polynomials in $T,S$ and PI $x$. 
+While the elimination and set union/intersection are compatible with multiple NS variables,
 the set complement requires an extension to Theorem \ref{thm:quotient} on ideals with 
-multivariate polynomial generators. We make a conjecture in details:
+multivariate polynomial generators. We make a conjecture as follows:
 
 \begin{Conjecture}
-Assume we have 2 ring variables $x,y$, and an ideal $J$ composed by 2 generators:
+Assume we are given 2 ring variables $x,y$, and an ideal $J$ composed by 2 generators:
 $J = \langle f_y, v_x\rangle$. Here $f_y$ has the form $f_y = \mathcal F(x,y)$, where $\mathcal F$
-is a 2-argument function, for example $f_y = y^2+(1+x)y+x$. And $v_x$ is "vanishing polynomial" about
-$x$: $v_x = x^{q_1} - x$. Now we have another generator serving as "universal set"
+is a 2-argument function, e.g. $f_y = y^2+(1+x)y+x$. And $v_x$ is the vanishing polynomial for variable
+$x$: $v_x = x^{q_1} - x$. Now we have another generator serving as the universal set
 $J_0 = \langle v_x, v_y\rangle$, $v_y = y^{q_2} - y$.
 We conjecture that
-$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J).$$
+$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J)$$
 And we believe this conjecture is scalable such that it is also valid for even more variables and generators like 
@@
 \begin{Example}
-Recall that 
+In Figure \ref{fig:fsm}
 $\{s_0, s_1\}$ are state/pseudo inputs, $\{t_0,t_1\}$ are state/pseudo outputs, and there is a primary input (1-bit) 
-$x$. We propose a new algorithm by modifying Algorithm \ref{alg:univa}
+$x$. We propose a new algorithm by modifying Algorithm \ref{alg:univa}:
 
+\IncMargin{1em}
 \begin{algorithm}[hbt]
 \SetAlgoNoLine
+\LinesNumbered
  \KwIn{Transition polynomial $f_t = T + \mathcal F (S,x)$, 
 	initial state ideal $from^0 = \langle S+\mathcal G(x), x^{q_1} - x\rangle$}
-
+ \KwOut{Reachable states}
   $reached = from^0(T\setminus S)$\;
@@
 \end{algorithm}
+\DecMargin{1em}
 
-Inputs of this algorithm are: transition polynomial, which is the result of word-level abstraction; initial 
-states description ideal, contains 2 generators defining PI could be any inputs and state inputs $S$ is a specific
-value. (Ex: $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
+The inputs of this algorithm includes transition polynomial (result of word-level abstraction) and initial 
+states description ideal, which contains 2 generators corresponds to constraints of PI and combinational input $S$. 
+For example, $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
 $\langle S+x\cdot\alpha, x^2+x\rangle$ means initial states$=\{00,10\}$).
 
-Transition polynomial calculation uses Tim's abstraction and bit-level variable substitution. After building
-an elimination ideal, use RATO such that \emph{reverse\ topo\ order\ ckt\ variables }$> T > S > x$, the reduction
+Transition polynomial calculation uses the abstraction and bit-to-level conversion method from 
+Chapter \ref{ch:reacha}. After Constructing
+an elimination ideal, we impose RATO such that \emph{reverse\ topo\ order\ ckt\ variables }$> T > S > x$, the reduction
 remainder has the form $T+\mathcal F(s_0,s_1,x)$. From word definition $S+s_0+s_1\alpha$ we get
@@
 
-Assume we start from state $\{11\}$. First iteration, it will reach state $\{01\}$. Line 4 is to compose an
+Assume we start from state $\{11\}$. In the first iteration, the reached state is $\{01\}$. Line 4 is to compose an
 ideal with 2 generators from $from^0$ and transition polynomial $f_T$, compute its Gr\"obner basis. Note this ideal
@@
 \end {equation}
-$v_x$ is a polynomial containing only $x$, initially it should be vanishing polynomial $x^{q_1}-x$, with the program executing
-it may be factorized.
+$v_x$ is a polynomial containing only $q$-bit PI $x$, initially it should be vanishing polynomial $x^{q_1}-x$, 
+with the program executing it may be factorized.
 
@@
 $T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
-$S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$
+$S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. 
+In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$
 
@@
 \end{itemize}
-The final reachable states are represented by ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
-which means $\{00,01,10,11\}$.
+The final reachable states are represented by multivariate polynomial ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
+which denotes $\{00,01,10,11\}$.
 \end{Example}
@@
 which can be written as Boolean function 
-$$f = ((a\lor b) \lor c) \lor d$$
+\begin{equation}
+\label{eqn:chainOR}
+f = ((a\lor b) \lor c) \lor d
+\end{equation}
 equals to following polynomial function in $\F_2$:
@@
 
-% fig:ZDD
+\begin{figure}[h]
+\centerline{
+\includegraphics[width=\textwidth]{newfig/red_steps}
+}
+\caption{The reduction process of chain of OR gates represent using ZDDs with order $d>c>b>a$}
+\label{fig:ZDD}
+\end{figure}
 
-From Figure \ref{fig:ZDD} we find the size of the ZDD is $2\times4 -1 = 7$, which is linear to the number of variables.
+From Figure \ref{fig:ZDD} we find the size of the ZDD on the right-most side representing 
+Equation \ref{eqn:chainOR} is $2\times4 -1 = 7$, which is linear to the number of variables.
 The reduction process using ZDDs can be executed as in \cite{polybori:2009}. The graphical illustration 
-of reduction in ZDDs is shown in Figure \ref{fig:reduce_steps}.
-
-% fig:reduce_steps
+of reduction in ZDDs is also shown in Figure \ref{fig:ZDD}.
 
@@
 presented to find the CI for Boolean formulae
-\cite{pudlak:ci} \cite{mcmillan:interpolation}. Assume that Boolean
+\cite{pudlak:ci} \cite{mcmillan2003interpolation}. Assume that Boolean
 formulae are represented in Clause Normal Form (CNF) as 
@@
 \begin{Definition}
-(From \cite{mcmillan:interpolation}) Let $A$ and $B$ be Boolean
+(From \cite{mcmillan2003interpolation}) Let $A$ and $B$ be Boolean
   formulae given as sets of clauses such that $A \wedge B$ is
@@
 called interpolation system constructs an interpolant in linear time
-and space in the size of the proof \cite{mcmillan:interpolation}
+and space in the size of the proof \cite{mcmillan2003interpolation}
 \cite{pudlak:ci}. 
@@
 CIs are used to derive abstractions to produce over-approximate image
-operators in model checking \cite{mcmillan:interpolation}. Since $A
+operators in model checking \cite{mcmillan2003interpolation}. Since $A
 \implies P$, $P$ contains $A$ and is an {\it abstraction} of $A$. It
@@
 
-\begin{Problem}
+{\bf Research problem 1:}
 {\it As we wish to perform MC with abstraction using algebraic
@@
 }
-\end{Problem}
 
@@
 
-\begin{Problem}
+{\bf Research problem 2:}
 {\it How can this algebraic interpolant $f_p$ (or in the general case,
@@
 with $X_A, X_B$ eliminated from the problem? }
-\end{Problem}
 
@@
 
-\begin{Problem}
+{\bf Research problem 3:}
 As there can be many interpolants for an ideal-pair $(J_A, J_B)$, we
@@
 \ei
-\end{Problem}
 
-\begin{figure}
+\begin{figure}[h]
 \centering{
-\includegraphics[width=3in]{newfig/project.eps}
+\includegraphics[width=4.5in]{newfig/project.eps}
 \caption{Algebraic interpolant: Projection of varieties on common variables? }
@@
 Technology mapping is an important problem in digital circuit synthesis.
-Designers are given a library of well-designed functional blocks (including IP cores) and a raw netlist, 
+Designers are given a library of well-designed functional/macro blocks (including IP cores) and a raw netlist, 
 technology mapping's objective is to map as many as blocks to the raw netlist and 
@@
 design $D$. The objective is to generate a mapped design $D$ with as many of $B_i$ such that the area
-and timing is optimal.
+and timing is optimal. The procedure is also illustrated in Figure \ref{fig:macro}.
 
-\begin{figure}[hbt]
+\begin{figure}[h]
 	\begin{center}
-	\includegraphics[scale=0.3]{macro.png}
+	\includegraphics[width=\textwidth]{newfig/macro.pdf}
 	\end{center}
-	\caption{The outline and flow of mapping for macro blocks}
+	\caption{The outline and flow of technology mapping of macro blocks}
 	\label{fig:macro}
@@
 
-This following part describes the sketch of our proposed approach base on loop invariant constraints \cite{sankaranarayanan2004non}.
-First, a transition system is modeled by algebraic assertions; then \emph{ideal membership test}
-is applied on the set of assertions to help abstract the loop invariant. "Template" is the concept
-I borrowed from this paper to apply on my own approach.
+The following part describes the sketch of our proposed approach base on loop invariant constraints \cite{sankaranarayanan2004non}.
+First, a transition system is modeled by algebraic assertions; then \emph{ideal membership test} \cite{lv:phd}
+is applied on the set of assertions to help abstract the loop invariant. We introduce concept \emph{template}
+from \cite{sankaranarayanan2004non} to apply on our proposed approach.
 
-1) {\bf Template and state constraints:} To effectively and efficiently utilize the ideal membership test technique, a heuristic to
-generate $\F$ is needed. One heuristic can provide better coverage for loop invariant abstraction as well as
-relatively small size is \emph{generic quadratic form}.
+1) {\bf Template and state constraints:}  Ideal membership test technique requires a setup of ideal $J_{loop}$
+for loop invariants. 
+One heuristic can provide better coverage for loop invariant abstraction as well as
+relatively small size is \emph{generic quadratic form} (GQF).
 
-For example, a pair of state variables $\{x,y\}$'s generic quadratic form is
-$$\F = a_0x^2+a_1x+a_2xy+a_3y+a_4y^2+a_5$$
-It covers all possible terms with degree at most 2. $a_0\sim a_5$ are usually real number parameters,
-a certain assignment about them can turn $\F$ into desired invariant. This parameterized constrain
+For example, the GQF of a pair of state variables $\{x,y\}$ is
+$$\Func = a_0x^2+a_1x+a_2xy+a_3y+a_4y^2+a_5$$
+It covers all possible terms with degree no more than 2. $a_0,a_1,\dots, a_5$ are usually real number parameters,
+some of their assignments can turn $\Func$ into desired invariant. This parameterized constrain
 polynomial covers all combinations of state variables can also be called a \emph{template}.
+Subsequently, finding a proper assignment is the essential part of in our proposed research.
 
-Finding a proper assignment is the major part of original paper and has potential to be even further explored
-beyond that paper, and/or to be applied on different research field.
-
-2) {\bf Initial state constraints:} For initial state, the constrains are explicit. A template is adopted and refined by Gr\"obner basis 
+2) {\bf Initial state constraints:} For initial states, the constrains are explicit. A template is adopted and refined by Gr\"obner basis 
 generated by original constrains, by equaling the remainder to 0 we can get constrains on
@@
 
-An example is Fig.\ref{fig:loopinv}. The template is generic quadratic form of $\{s,i,j,j_0\}$, which is
-\begin{figure}[hbt]
-	\begin{center}
-	\includegraphics[scale=0.3]{../../../Pictures/Selection_006.png}
-	\end{center}
-	\caption{An example program of 2 Natural numbers' multiplication (Fig.1 from original paper)}
-	\label{fig:loopinv}
-\end{figure}
+An example is shown in the following 3-line algorithm multiplying two natural numbers. 
 \begin{align}
-\F =& a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
+\label{eqn:loopinv}
+& {\bf integer}~~i,j,s~~{\bf where}(s=0 \land j=j_0) \nonumber\\
+& l_0: {\bf while}~~(\cdots)~~{\bf do} \nonumber\\
+& ~~~~ l_1:~(s,j)\gets(s+i,j-1) 
+\end{align}
+
+Its template is generic quadratic form of $\{s,i,j,j_0\}$, which is
+\begin{align}
+\Func = & a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
 &a_5i^2+a_6i+a_7ij+a_8ij_0+a_9j^2\nonumber\\
@@
 \end{align}
+with parameters $a_0,\dots,a_{14}$.
 
-Constrains of initial state: $s=0\land j=j_0$ can be interpreted to polynomials:
-$\{s, j-j_0\}$. Check if it is Gr\"obner basis using \emph{Buchberger's algorithm}. Since
-their leading terms are relatively prime, we can write Gr\"obner basis $G=\{s, j-j_0\}$. Its ideal
+Constrains of initial state $s=0\land j=j_0$ can be interpreted to polynomials:
+$$\{s, j-j_0\}$$
+Since their leading terms are relatively prime, the polynomial set itself constitutes a Gr\"obner basis 
+$$G=\{s, j-j_0\}$$ 
+Therefore its ideal can be written as
 $J=\langle s,j-j_0\rangle$.
 
-Compute multi-division $\F \xrightarrow{G}_+ r$, the remainder is
-$r = a_5i^2+a_6i+(a_7+a_8)ij_0+(a_9+a_{11}+a_{12})j_0^2+(a_{10}+a_{13})j_0+a_{14}$.
-Let it equal to 0, each coefficient will generate a constrain, solution to the system is candidate
-assignment to generate invariant.
+In the next step, we reduce the template with the Gr\"obner basis $G$: $\Func \xrightarrow{G}_+ r$, the remainder equals to
+$$r = a_5i^2+a_6i+(a_7+a_8)ij_0+(a_9+a_{11}+a_{12})j_0^2+(a_{10}+a_{13})j_0+a_{14}$$
+Let it equal to 0, then each coefficient will generate a constraint. The solution to the system forms the candidate
+assignment to generate loop invariants.
 \begin{equation}
@@
 3) {\bf Modeling state transitions:}
-2 states get involved in a state transition. The technique requires to model the 2 states individually,
-do multi-division separately to get remainder $r_1$ and $r_2$. Suppose a constrain polynomial for
-this transition is $r_t$, we requires that when invariant for state 1 holds and transition $1\to 2$ stands,
-invariant for state 2 should also holds, which is
+A typical state transition starts from the previous state (PS) and ends at next state (NS). 
+Our proposed approach models the 2 states individually,
+{i.e.} performs polynomial reduction separately and obtains remainder $r_1$ and $r_2$. 
+Assume a constraint polynomial describing
+the state transition is $r_t$, then we require that when invariant of PS holds and transition $PS \to NS$ stands,
+invariant of NS should also be satisfied:
 $$(r_1 = 0)\land (r_t = 0) \implies (r_2 = 0)$$
@@
 $$r_t = r_1 - \lambda r_2$$
-Theoretically $\lambda$ could be any polynomial. Consider the system to be easy to solve, we keep $\lambda$
-only taking value from real numbers.
+Theoretically $\lambda$ can be any polynomial in arbitrary rings. To make it practical solving the system, 
+we limit $\lambda$ to the polynomial ring with only real numbers.
 
-Still take Fig.\ref{fig:loopinv} as example (also refer to Example 10 in original paper). State 1 is initial state we 
-just characterized $\F = f(s,i,j,j_0)$, and the post state (state 2) have exactly the same form constrain:
-$\F' = f(s',i',j',j_0')$. Considering transition relation, substitute $s'$ with $s+i$, replace $j'$
-with $j-1$, and $i'$ for $i$, $j_0'$ for $j_0$, the template for state 2 is expression $f'$ in Example 10 from
-original paper.
+We take Equation \ref{eqn:loopinv} as the example (which refers to Example 10 in \cite{sankaranarayanan2004non}). 
+PS is initial state we just characterized 
+$$\Func = f(s,i,j,j_0)$$ 
+and NS has exactly the same form of constraint:
+$$\Func' = f(s',i',j',j_0')$$
+Considering the transition relation, we substitute $s'$ with $s+i$ and replace $j'$
+with $j-1$, $i'$ for $i$, $j_0'$ for $j_0$, respectively. 
+The template for NS is polynomial $f'$ in Example 10 in \cite{sankaranarayanan2004non}.
 
-Do multi-division with Gr\"obner basis generated by transition relation, record its remainder $r_2$.
-From equality $r_2 = \lambda r_1$, we get constrains for parameters. Solve the system using covering-problem-like
-techniques (Part. \emph{Elimination by Splitting} in original paper). One branching result is:
-$a_0,a_2\sim a_6, a_9\sim a_{14}$ are all zero; $a_1=a_7=-a_8$. Then reduced remainder 
+We then perform reduction with the Gr\"obner basis generated from the transition relation modeling, 
+and record its remainder $r_2$.
+From equality $r_2 = \lambda r_1$, we get constraints for the parameters. Solve the system using 
+similar algorithm to solve binate covering problems
+which is also described in Section 4 as \emph{elimination by splitting} technique in \cite{sankaranarayanan2004non}. 
+One branching result is:
+\begin{align*}
+&a_0,a_2,\dots, a_6, a_9,\dots,a_{14}=  0\\ 
+&a_1=a_7=-a_8
+\end{align*}
+The reduced remainder is: 
 $$r_2 = a_1s +(a_1-a_7)i+a_7ij+a_8ij_0 = a_1(s+ij-ij_0) = 0$$
-Desired invariant is 
+Thus the invariant of the program in Equation \ref{eqn:loopinv} is:
 $$s = i(j_0- j)$$
-We knew the program calculates $i\times j_0$. Initially $s=0$, $j_0-j=0$, invariant holds;
+We can verify the invariant by executing the program which calculates $i\times j_0$. 
+Initially $s=0$, $j_0-j=0$, invariant holds;
 during each cycle, $s'=s+i, (j_0'-j') = (j_0-j)+1$, the invariant also holds. In conclusion, this is a
-loop invariant for program in Fig.\ref{fig:loopinv}.
+loop invariant for program in Equation \ref{eqn:loopinv}.
 
 {\bf Our proposed approach on technology mapping:}
-Our approach borrows inspiration of "templates" from loop invariant paper. The difference is,
-we are not using template polynomial to describe the system, instead we add some templates in the ideal
-of macro blocks we want to map to simulate all possibilities.
+Our approach borrows inspiration of templates in \cite{sankaranarayanan2004non}. However, we make an improvement
+on the original approach:
+instead of using template polynomial to describe the system, we add some templates into the ideal
+of macro blocks which serve as technology mapping candidates. In this way we can cover all possibilities
+of boundary cutting (circuit partition).
+
 1) {\bf System abstraction:}
 The polynomial we use to test ideal membership should include all information of a circuit partition,
-this requires us to abstract information from the system and write it into only 1 polynomial.
-Usually this polynomial has the form like:
+which requires us to abstract information from the system and write it into a single polynomial.
+Usually this polynomial has the following generic form:
 $$Z + f(i_1,i_2,\dots,i_n)$$
-Here $Z$ is the output. When there is only one output, $Z$ will be a bit-level variable. However in most
-cases we have multiple outputs, this asks us to write $Z$ as a word-level variable. $f$ is a Boolean
-function about all inputs, and $i_1,i_2,\dots,i_n$ are all bit-level inputs.
+where $Z$ is the output. When there is only one output, $Z$ collapse to a bit-level variable. However in most
+cases there are multiple outputs on the cut, indicating $Z$ as a word-level variable. Boolean function $\Func$ covers
+all inputs, and $i_1,i_2,\dots,i_n$ are all bit-level inputs.
 
-\begin{figure}[hbt]
+\begin{figure}[h]
 	\begin{center}
-	\includegraphics[scale=0.2]{tobemapped.png}
+	\includegraphics[width=\textwidth]{newfig/tobemapped.pdf}
 	\end{center}
@@
 
-Fig.\ref{fig:tobemapped} shows an example circuit partition with $a1,b1,a0,b0$ as inputs and
-$Z = \{z0,z1\}$ as outputs. If we use elements from Galois field $F_{2^2}$ to represent word $Z$,
+Figure \ref{fig:tobemapped} shows an example circuit partition with $a_1,b_1,a_0,b_0$ as inputs and
+$Z = \{z_0,z_1\}$ as outputs. If we use elements from Galois field $\F_{2^2}$ to represent word $Z$,
 we have $Z = z_0 + \alpha\cdot z_1$.
 
-The abstraction also uses the property of Gr\"obner basis. If we arrange a term ordering as
+After imposing ATO of LEX with
 $$Other\ circuit\ variables > output\ word\ Z > all\ bit\ level\ inputs$$
-and include all gates information, word level variable definition and vanishing polynomials,
-the reduced Gr\"obner basis will have a polynomial (generator) in the form of $Z + f(a1,b1,a0,b0)$,
-this is an application of "elimination using GB", where we eliminate all variables except output
-word variable $Z$ and inputs $a1,a0,b1,b0$.
-
+on the ideal describing the partitioned circuit as well as ideal with vanishing polynomials,
+the reduced Gr\"obner basis has a single polynomial generator in the form of $Z + \Func(a_1,b_1,a_0,b_0)$.
 In this example:
@@
 \begin{itemize}
-\item Term ordering: $z0,z1,t,g1,p1,ci,t1,t2,Z,a0,b0,a1,b1$
-\item Gate description (from table \ref{table:booltogalois_op}): $z0+t+g1+t\cdot g1, t+t1+t2+t1\cdot t2, g1+a1\cdot b1,
-			t1+a1\cdot ci, t2+b1\cdot ci, ci+a0\cdot b0, z1+p1+ci, p1+a1+b1$
-\item Word definition: $Z+z0+z1\cdot \alpha$
-\item Vanishing polynomials $(J_0)$: $z0^2+z0, z1^2+z1, t^2+t, g1^2+g1, t1^2+t1, t2^2+t2, p1^2+p1, ci^2+ci,
-			a0^2+a0, b0^2+b0, a1^2+a1, b1^2+b1, Z^4+Z$(note Z is 2-bit word)
+\item Term ordering: $\{z_0,z_1,t,g_1,p_1,ci,t_1,t_2\}>Z>\{a_0,b_0,a_1,b_1$
+\item Gate description: $z_0+t+g_1+t\cdot g_1, t+t_1+t_2+t_1\cdot t_2, g_1+a_1\cdot b_1,
+			t_1+a_1\cdot ci, t_2+b_1\cdot ci, ci+a_0\cdot b_0, z_1+p_1+ci, p_1+a_1+b_1$
+\item Word definition: $Z+z_0+z_1\cdot \alpha$
+\item Vanishing polynomial ideal $(J_0)$: $z_0^2+z_0, z_1^2+z_1, t^2+t, g_1^2+g_1, t_1^2+t_1, t_2^2+t_2, p_1^2+p_1, ci^2+ci,
+			a_0^2+a_0, b_0^2+b_0, a_1^2+a_1, b_1^2+b_1, Z^4+Z$(since Z is a 2-bit word)
 \end{itemize}
 
-The result is one polynomial in Gr\"obner basis with leading term $Z$:
-$$Z+a0\cdot b0\cdot a1+a0\cdot b0\cdot b1+\alpha\cdot a0\cdot b0+a1\cdot b1+\alpha\cdot a1+\alpha\cdot b1$$
+The result is a Gr\"obner basis with single polynomial generator. The polynomial has leading term $Z$:
+$$Z+a_0\cdot b_0\cdot a_1+a_0\cdot b_0\cdot b_1+\alpha\cdot a_0\cdot b_0+a_1\cdot b_1+\alpha\cdot a_1+\alpha\cdot b_1$$
 
-\begin{table}
-\centering
-\begin{tabular}{|c|c|} \hline
-Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
-$\overline{a}$ & $1 + a$\\ \hline
-$a\ and\ b$ & $ab$\\ \hline
-$a\ or\ b$ & $a + b + ab$\\ \hline
-$a \oplus b$ & $a + b$\\
-\hline\end{tabular}
-\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
-\label{table:booltogalois_op}
-\end{table}
+% \begin{table}
+% \centering
+% \begin{tabular}{|c|c|} \hline
+% Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
+% $\overline{a}$ & $1 + a$\\ \hline
+% $a\ and\ b$ & $ab$\\ \hline
+% $a\ or\ b$ & $a + b + ab$\\ \hline
+% $a \oplus b$ & $a + b$\\
+% \hline\end{tabular}
+% \caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
+% \label{table:booltogalois_op}
+% \end{table}
 
 2) {\bf Templates on Boundary Information:}
-\begin{figure}[hbt]
+\begin{figure}[h]
 	\begin{center}
-	\includegraphics[scale=0.2]{template.png}
+	\includegraphics[width=\textwidth]{newfig/template.pdf}
 	\end{center}
@@
 
-Fig.\ref{fig:template} shows a 2-bit standard cell. It has 3 outputs mapping to 2 output pins,
-and 4 inputs mapped to 3 input pins (the rest pin is mapped to fixed 0/1 signal). Dashed connection
+Figure \ref{fig:template} shows a 2-bit standard cell. It has 3 outputs mapped to 2 output pins,
+and 4 inputs mapped to 3 input pins (the rest pins are assigned to fixed 0/1 signal). Dashed connection
 lines show one possible mapping, to find out this kind of feasible mapping, we need to simulate
-all possible mappings, where the concept "template" can be used.\vspace{5mm}
+all possible mappings, where the concept "template" can be used.
 
-{\it Output template}: $z0+ct0_{z0}\cdot t0+ct1_{z0}\cdot t1+c0_{CO}\cdot CO+cn0$
+{\it Output template}: $z_0+c_{{t_0}{z_0}}\cdot t_0+c_{t_1z_0}\cdot t_1+c0_{CO}\cdot CO+c_{n_0}$
 
-When $cn0 = 0$, any one of $ct0_{z0},ct1_{z0},c0_{CO}$ could be 1 means mapping to corresponding output
-pin. If $cn0 = 1$, mapping to negation of corresponding output pin.
+When $c_{n_0} = 0$, then $\{c_{t_0z_0},c_{t_1z_0},c0_{CO}\}$ equals to 1 denotes a valid mapping to corresponding output pin. 
+Conversely if $cn0 = 1$, the ``1" evaluation denotes mapping to negation of corresponding output pin.
 
-{\it Input template}: $a0+pa0+c0_{A0}\cdot s0+c1_{A0}\cdot s1+c2_{A0}\cdot s2$
+{\it Input template}: $a_0+p_{a_0}+c0_{A_0}\cdot s_0+c1_{A_0}\cdot s_1+c2_{A_0}\cdot s_2$
 
-When $pa0 = 0$ and any one of $c0_{A0},c1_{A0},c2_{A0}$ equal to 1 means mapping to corresponding input
-pin; if all $c0_{A0},c1_{A0},c2_{A0}$ equal to 0 then means mapping to fixed signal "0" (when $pa0 = 1$ then
-mapping to fixed signal "1").
+When $p_{a_0} = 0$, then $\{c0_{A_0},c1_{A_0},c2_{A_0}\}$ equal to 1 denotes a valid mapping to corresponding input
+pin. Conversely if every variable in set $\{c0_{A_0},c1_{A_0},c2_{A_0}\}$ equals to 0, it denotes mapping to fixed signal ``0" (when $p_{a_0} = 1$ then
+mapping to fixed signal ``1").
+
+With these settings, we propose an approach to perform technology mapping with boundary information:
+\begin{itemize}
+\item First, choose a set of independent wires as input pins;
+
+\item Second, push forward input signal for a certain number of gates (depth), choose a cut set
+which fully dependent on these inputs as output pins;
+
+\item Third, abstract a description polynomial using GB based abstraction;
+
+\item Fourth, reduce the description polynomial with the GB we computed associated to macro block 
+and templates, obtain a remainder polynomial;
+
+\item Last but not least, abstract all coefficients from the remainder polynomial and set up a system of equations.
+If we find a solution to this system, then the solution is a feasible mapping; no solution means we cannot make a
+feasible mapping.
+\end{itemize}

### 2016-12-08 dissertation/abstract.tex
dissertation v2:reconstructing part intro

@@
 
-Verification of sequential circuits remains to be a topic of all time. With increasing size of new
-integrated circuits, sequential circuit designers may face much more complicated problems on 
-logic errors, unexpected delays and failures on critical path. 
-This dissertation addresses the problem of sequential circuit verification at the word-level
-and is based on concepts derived from algebraic geometry.
+% Verification of sequential circuits remains to be a topic of all time. With increasing size of new
+% integrated circuits, sequential circuit designers may face much more complicated problems on 
+% logic errors, unexpected delays and failures on critical path. 
+% This dissertation addresses the problem of sequential circuit verification at the word-level
+% and is based on concepts derived from algebraic geometry. 
+% 
+% Analyzing the sequential circuits at word level is an efficient way
+% of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
+% state-space. We manage to model the verification properties and the gate-level sequential circuit 
+% implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
+% their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
+% techniques from algebraic geometry can be used to reason about the state-space on high-level.
+% In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
+% polynomial ideals. Without actually solving the system, the states in state space can be investigated 
+% by manipulating these ideals. 
+% 
+% We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
+% analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
+% verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
+% abstraction-refinement techniques for verification, we propose to investigate a word-level 
+% analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
+% not only derive new algorithms and techniques, but will also consider efficient CAD 
+% implementations to target sequential equivalence and model checking problems. 
+% % \end{abstract}
 
-Analyzing the sequential circuits at word level is an efficient way
-of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
-state-space. We manage to model the verification properties and the gate-level sequential circuit 
-implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
-their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
-techniques from algebraic geometry can be used to reason about the state-space on high-level.
-In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
-polynomial ideals. Without actually solving the system, the states in state space can be investigated 
-by manipulating these ideals.
-
-We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
-analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
-verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
-abstraction-refinement techniques for verification, we propose to investigate a word-level 
-analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
-not only derive new algorithms and techniques, but will also consider efficient CAD 
-implementations to target sequential equivalence and model checking problems. 
-% \end{abstract}
+This dissertation presents work on a new class of terahertz (THz) waveguides based on structured 
+metal geometries. The waveguides are designed with the core idea that adoption of planar layout in 
+fabrication can lead to exponential growth in device capabilities, analogous to the growth in device 
+capabilities based in electronics. From a functional point of view, the waveguides rely upon 
+propagation of surface waves along the surface of metals. This approach is preferred, since 
+dielectrics tend to be lossy at THz and the loss parameters scales almost quadratically with 
+frequency for most dielectrics. The loss in propagating wave is minimized by utilizing metals, 
+which are highly conducting at THz frequencies. Structuring the metal surface with periodic array 
+of apertures of sub-wavelength dimension allows bound surface wave to propagate as the wave can 
+evanescently decay into the metal. This phenomenon is referred as the coupling of propagating wave 
+to surface plasmon polariton (SPP) like mode at the interface of structured metal surface and air. 
+Thus, these propagating THz waves are simply surface plasmon-polaritons (SPPs) . Similarly, 
+complimentary structures that do not perforate the metal, but rather stand on the metal surface 
+also support SPPs. The wavelength of SPPs can be controlled by changing the dimension of these 
+apertures/structures, since the dispersion relationship of the medium depends on the geometrical size. 
+This engineering capability has been exploited in creating all the waveguides presented in this thesis. 
+The devices presented are categorized based on the fabrication technique. Each technique is unique 
+in its own regard and can be selected based on functional needs. A commonly adopted process of laser 
+ablation covers a wide set of waveguides presented here. In one of the waveguides fabricated using 
+ablation technique, the role of disorder is discussed. The waveguide with introduction leads to 
+observance of localized mode with spectral and spatial feature like Anderson localized modes of photons. 
+It is the first report of localized mode at THz frequency. 3D rapid prototyping involving 3D printer is 
+used to create waveguides with complex layout that can allow for multi-plane signal routing. 
+This also is the first demonstration of 3D printing in the development of THz devices. In another 
+approach a unique fabrication technique had to be developed to create waveguides mediums with negative 
+index of refraction (NIM) as they require feature sizes which cannot easily be attained using conventional 
+clean room techniques and 3D printing. This new fabrication approach uses a sacrificial layer technique 
+that is used create an effective medium with negative index of refraction and length on the order of tens 
+of wavelength. Again, this happens to be the first demonstration of a waveguide with NIM capability at 
+THz length scales, as well as a figure of merit (FOM) defined in terms of loss parameter that is an order 
+of magnitude better than the previous reported NIM’s in the THz spectral range.

### 2016-12-08 dissertation/appendix.tex
dissertation v2:reconstructing part intro

@@
 \begin{Lemma}
+\label{lem:app1}
 Given $g(x) \in \F[x]$ and $W$ is its nullspace. Let $d(x) = gcd(f(x),g(x)), e(x) = f(x)/d(x)$. 
@@
 \end{Proof}
+
+The $T$-Order and corresponding nullspace also affect the factorization of polynomials:
+\begin{Lemma}[Factorization of $f(x)$]
+Factorization $f(x) = \prod_{i=1}^{r} f_{i}^{d_i} (x)$, where each $f_i(x)$ is prime to others. 
+Assume $V_i$ be nullspace of $f_{i}^{d_i} (x)$, then $V = V_1\oplus V_2 \oplus \dots\oplus V_r$.
+
+Furthermore, we can define a polynomial $\Psi_i(x) = f(x)/f_{i}^{d_i} (x)$, where 
+$$\forall u_j \in V_j, u_j \neq 0, \Psi_i(T)u_j \neq 0$$
+only if $i = j$.
+\end{Lemma}
+\begin{Proof}
+We can use Lemma \ref{lem:app1}:
+$$gcd(f(x),f_{i}^{d_i} (x)) = f_{i}^{d_i} (x),~~dim(V_i) = deg(f_{i}^{d_i} (x))$$
+
+Which implies 
+$$dim(V_i) = deg\left(\prod_i f_{i}^{d_i} (x)\right) = \prod_i dim(V_i) \implies V = \bigoplus_i V_i$$
+
+Assume $i\neq j$, $\Psi_i(x) = \frac{f(x)}{f_{i}^{d_i} (x)} = h(x)f_{j}^{d_j} (x)$. Then
+\begin{align*}
+&\forall u_j\in V_j, f_j(T)u_j = 0 \implies h(x)f_{j}^{d_j} (T)u_j = 0 \\
+& \implies \Psi_i(x)u_j = 0
+\end{align*}
+
+Conversely, if $i=j$, then
+\begin{align*}
+& \Psi_i(x) = \frac{f(x)}{f_{i}^{d_i} (x)} \perp f_i(x) \\
+& \implies \Psi_i(x)u_j \neq 0
+\end{align*}
+\end{Proof}
+
+For a given Frobenius map $\sigma$, corresponding minimal polynomial is restricted:
+\begin{Lemma}
+The minimal (and characteristic) polynomial for $\sigma$ is $x^n-1$.
+\end{Lemma}
+\begin{Proof}
+Consider Fermat's little theorem in $\F_{p^n}$. Let $\beta$ be an element in the field, then
+$$\sigma^n\beta = \beta^{p^n} = \beta \implies \sigma^n-I = 0$$
+where $I$ is the identity map. For characteristic polynomial, assume $\exists f(x) = \sum_i f_ix^i \in \F_p[x]$, 
+such that the degree of characteristic polynomial is lower than $n$:
+$$\sum_if_i\sigma^i = 0~~~\text{and}~~~deg(f(x))<n$$
+
+Then $\forall \beta\in\F_{p^n}$, 
+$$\left(\sum_if_i\sigma^i\right)\beta = \sum_if_i\beta^{p^i} = 0$$
+
+This equation denotes that $\beta$ is one of $p^n$ roots of polynomial $\Func(x) = \sum_if_i(x^{p^i})$.
+However, the maximum number of roots allowed equals to the degree of $\Func(x)$, which is 
+$p^{n-1}<p^n$, we find a contradiction. Thus we conclude that
+$$\textit{Both characteristic and minimal polynomial is }x^n-1$$
+\end{Proof}
+
+From lemmas above we can deduce the following corollary:
+\begin{Corollary}
+An element $\alpha \in \F_{p^n}$ is a {\bf normal element} if and only if $Ord_{\alpha,\sigma}(x) = x^n-1$.
+\end{Corollary}
 \subsection{Construction of General Normal Basis}

### 2016-12-08 dissertation/intro.tex
dissertation v2:Appendix update 1
dissertation v2:reconstructing part intro

@@
 \label{ch:intro}
+
+\section{Hardware Design and Verification Overview}
+Because of the Moore's law, the level of integration in modern VLSI systems is becoming higher and higher. 
+As a result, an entire system with billions of transistors can be built upon a single chip.
+The design process also evolves from manual design with little validation, to 
+a formal 3-step procudure which needs collaboration of teams with large number of 
+engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
+2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
+tools; 3) Verify, which is to verify the correctness of design and implementation.
+
+Nowadays the verification step is usually completed by a team that specializes on test, verification and validation of 
+circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
+Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procedures starting from 
+register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
+physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
+to ensure the implementation is consistent with the original design intent. Verification is 
+an important quality control measure before sending design layout to the VLSI foundaries.
+Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
+waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
+the stability and accuracy of digital VLSI circuits; even small flaws or short-time failures can cause 
+huge loss, especially in medical applications, military facilities and financial systems.
+Therefore, it is of utmost importance to verify the correctness of VLSI designs.
+
+One way to perform verification is by {\it simulation}. It is defined as the collage of all circuit validation 
+methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
+However, simulation is not a complete solution to circuit verification problems. In modern designs with 
+large number of logic components and complicated architecures, it is impractical to simulate all possible 
+test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
+cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
+is a good example where simulation failed. Failure occurs with input assignments which are rarely used in most divisions. Because of the limitation of simulation,  
+test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollors recalling bill
+for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.
+
+{
+\begin{figure}[h]
+\centerline{
+\includegraphics[width=0.7\textwidth]{newfig/designflow.eps}
+}
+\caption{Typical hardware design flow}
+\label{fig:designflow}
+\end{figure}
+}
+
+Another method developed besides simulation is \emph{formal verification}, it utilizes 
+mathematical theory to reason about the correctness of hardware designs.
+Formal verification can provide $100\%$ fault coverage from two aspects. On the one hand,  
+it adopts formal languages to strictly describe the design intent and detailed implementation, 
+and deduces circuit function from the implementation.
+On the other hand,  it formalizes properties
+for the circuit model which are relevant only to specific input signals,  and prove the properties mathematically. 
+These descriptions are named as {\it specifications}.
+Formal verification has two main forms: property checking and equivalence 
+checking.
+
+\section{Formal Verification: Property and Equilavence Checking}
+
+{\it Property checking} (or property verification) verifies
+that a design satisfies certain given properties. Property checking is done mainly 
+in the form of theorem proving (TP), model checking (MC), or TP/MC hybrid approaches.
+\begin{enumerate}[{1)}]
+\item \emph{Theorem proving} \cite{theoremproving:91} 
+is a method of reasoning and mathematical logic dealing with proving 
+mathematical theorems. In the application to circuit property checking, 
+the specification as well as the circuit implementation
+is described as theorems in mathematical logic. Subsequently logic rules
+are employed to deduce new objective theorem. In practice, the tool can reduce
+a proof goal to simpler sub-goals for automatic verification.
+
+\item \emph{Model checking} \cite{modelcheck:99} is a technique 
+for verifying if the specification properties are violated in finite-state systems. In circuit verification 
+domain, both the specification and the circuit implementation are modeled as a system of 
+logic formulas. The finite-state system is then traversed to check if the 
+properties are violated. If violation occurs, 
+a counter-example is then generated as a transition firing path corresponds to the
+false behavior in the design. 
+Modern model checking techniques use the error-trace to automatically refine
+the system and perform further checkings.
+\end{enumerate}
+
+{\it Equivalence Checking} verifies that two different representations of
+a circuit design have equivalent functionality. It can be applied to 
+multiple steps in the hardware design flow in Figure \ref{fig:designflow},
+such as checking functional equivalence between HDL description and RTL,
+checking RTL equivalence between RTL and synthesized/optimized netlist
+,and checking layout verification between netlist and layout for fabrication.
+
+There are three major
+equivalence checking techniques: graph-based,
+satisfiability-based (SAT-based) and induction-based.
+\begin{enumerate}[{1)}]
+\item \emph{Graph-based} techniques compare between 2 circuit implementations 
+by representing them using canonical graphs. 
+The earliest invented canonical graph is the \emph{Binary Decision Diagram} (BDD) \cite{BRYA86}.
+Many variants branch out from BDD, some widely used variants include 
+ZDD \cite{minato1993zero}, BMD \cite{bmd}, FDD \cite{okfdd}, etc. The comparison algorithms can 
+determine whether the two graphs are isomorphic. The canonity of the graph representation guarantees that
+the graphs correspond to the two circuits will be 
+equivalent if and only if the circuits are implementing the same function.
+\item \emph{Satisfiability} (SAT) techniques utilize the satisfiability theory.
+In circuit equivalence checking, a miter of the two circuits is created.
+A \emph{miter} is a combination of the two circuits with one bit-level output, which 
+gives output ``1" only when the outputs of the circuits differ with 
+the same inputs, {\it e.g. }inputs $a,b,c$ shown in Figure \ref{fig:miter}. 
+A SAT solver \cite{csat,mishchenko2006improvements} is then employed to simplify the problem 
+and find a satisfying assignment to the inputs for which the 
+miter output is ``1". If such an assignment exists, this solution acts as a 
+counter-example of when the circuit outputs differ; otherwise the circuits
+are functionally equivalent.
+\item \emph{Induction-based} techniques are developed and applied to verify the 
+equivalence between sequential circuits, which is called the {\it sequential equivalence checking}
+(SEC) problem. A miter model can also be built with two sequential circuits. Through the 
+miter model, SEC problem is transformed into a sequential backward justification problem.
+Equivalence of states and transitions between states can be proved using induction-based 
+proof and fix-point calculation \cite{bjesse2000sat,stoffel1997record}.
+\end{enumerate}
+
+\begin{figure}[h]
+\centering{
+\includegraphics[scale=0.3]{newfig/fig_SAT.eps}
+\caption{An example of equivalence checking on miter of circuit $A$ and $B$ using SAT}
+\label{fig:miter}}
+\end{figure}
+
+A part of formal verification techniques adopt concepts and algorithms from 
+\emph{computer-algebra} and \emph{algebraic geometry}.
+Algebraic geometry provide a way to reason about the presence or absence of solutions 
+without actually solving the system of constraints.
+Using methods in \cite{Avrunin:CAV,condrat-tacas07,gbverify:2007,jinpeng,pruss:tcad15}, 
+the circuit design can be transformed into a polynomial system. Subsequently, this system
+of polynomials is canonicalized by computing a Gr\"obner basis (GB) \cite{gb_book}. 
+Computation of GB allows for 
+a straightforward proof of important properties of a polynomial system, 
+such as the presence or absence of 
+solutions. These properties can also be leveraged for 
+verification. The disadvantage of the GB computation method is that its complexity can be doubly 
+exponential in the worst case \cite{dube1986complexity}, and thus methods directly performing GB computation 
+from an arbitrary setup is not practical for industry-level applications. However, recent
+breakthroughs in computer-algebra hardware verification have shown
+that it is possible to overcome the complexity of this computation while
+still utilizing the beneficial properties of GB
+\cite{lv:phd,tim:phd}.
+
+\section{Importance of Word-level Abstraction}
+Most formal verification techniques can benefit from word-level abstractions 
+of the circuits they verify.
+There are several advantages in exploiting word-level information for
+verification. A number of designs have  their
+datapaths and/or system-level models described as word-level RTL
+models.  Exploiting word-level instead of bit-level information is one
+way of abstraction -- a key technique to reduce the state space of a
+FSM. It has the effect of combining sets of states with similar 
+properties. During reachability analysis, if we use bit-level
+variables to  represent the states, the representations may become too
+large to handle. However, when a ``bundle" of bit-level variables are
+represented as {\it only one word-level variable}, the set of
+reachable states can be represented by a  word-level constraint
+expression; which may lower verification complexity.
+
+% Abstraction is defined as state-space reduction, i.e{\text . }abstraction
+% reduces state-space by mapping the set of states of a system to a smaller 
+% set of states. Because the new representation contains fewer states, it
+% is easier to comprehend and thus easier to use. 
+% Word-level abstraction focuses specifically on abstracting a word-level
+% representation of a circuit out of a bit-level representation. For example,
+% a bit-level representation of an integer multiplier is represented by a
+% collection of Boolean inputs and outputs, whereas a word-level
+% abstraction hides the underlying logic and represents the circuit as two 
+% integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
+% multiplier increases, the logical implementation of the multiplier grows (typically
+% exponentially) while the word-level abstraction stays the same.
+
+Word-level abstractions have a wide variety of applications in formal 
+verification. For example, it can work as automatic decision and canonical reduction engine in theorem proving;
+for RTL composed of macro blocks, abstractions of these blocks are also benefit RTL verification.
+Concretely, MC and equivalence checking with abstractions can be classified as:
+\begin{itemize}  
+\item Model checking with abstractions \cite{kroening:model}, 
+where an over-pproximation of RTL blocks is abstracted and used to 
+refine the RTL model.
+\item Graph-based equivalence 
+checking with abstraction \cite{WLS,arditi:bmd}, where abstraction 
+generates a word-level canonical graph representation of the circuit.
+\item SAT-based equivalence checking with abstraction \cite{lpsat}, where 
+abstractions are used to analyze structural symmetries and similarities such that
+the Boolean formulas fed to SAT solver is minimized.
+\end{itemize}
+
+Other equivalence checking techniques that employ abstractions 
+include {\it satisfiability modulo theory} (SMT) solvers \cite{boolector,bryant:tacas07}.
+as well as constraint programming (CP) techniques \cite{ms:research,tew:iccad08}.
+
+Word-level abstractions also find applications in RTL and datapath 
+synthesis \cite{demicheli:iccad_98,demicheli:dac_99,demicheli:tcad_03}. 
+Since modern datapath design specifications are mostly word-level, synthesis tools with abstractions
+can make use of larger macro blocks to generate and optimize the
+datapaths. Moreover, 
+word-level abstractions facilitates uninterpreted functions (UFs) \cite{UF3}, which 
+can be transformed into proposition formulas with word-level information and verified using 
+word-level theorem provers and model checkers.
+
+\section{Abstractions in Sequential Design Verification}
 With the increasing size of integrated
@@
 that the sequential circuit implementations are equivalent to the
-original specification models, i.e. {\it Sequential Equivalence
-  Checking   (SEC)}. When the specification is given as an arithmetic function
+original specification models, {\it i.e.} {\it Sequential Equivalence
+  Checking }(SEC). When the specification is given as an arithmetic function
 which canonically represents the circuit, then the problem 
@@
 but  at the word-level. Analogously, in the case of MC, the problem
-is setup w.r.t. a FSM $\mathcal{M}$, a set of initial states $S_0$ and
+is setup {\it w.r.t.} a FSM $\mathcal{M}$, a set of initial states $S_0$ and
 a set of property states $p$. Techniques are to be researched that
@@
 state to a non-property state (``bad'' state). These problems have to
-be solved in the context of word-level verification -- i.e. data
+be solved in the context of word-level verification -- {\it i.e.} data
 representation, abstraction using UNSAT cores and
 algorithm execution has to be carried out at word-level.
-
-\section{Hardware Design and Verification Overview}
-Following the Moore's law, the level of integration on modern VLSI becomes very high, 
-such that a single digital VLSI design can implement a true system-level component.
-The design process also evolves from manual design with little validation, to 
-a formal 3-step procudure which needs collaboration of team with large number of 
-engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
-2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
-tools; 3) Verify, which is to verify the correctness of design and implementation.
-
-Nowadays the verify step is usually completed by a team with specialities on test, verification and validation of 
-circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
-Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procudures starting from 
-register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
-physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
-to ensure the implementation is consistent with the original design intent. Verification is 
-an important quality control measure before sending design layout to the VLSI foundaries.
-Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
-waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
-the stability and accuracy of digital VLSI circuits, even small flaws or short-time failures can cause 
-huge loss, especially in medical applications, military facilities and financial systems.
-Therefore, it is of utmost importance to verify the correctness of VLSI designs.
-
-One way to perform the verification is called {\it simulation}. It is defined as the collage of all circuit validation 
-methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
-However, simulation is not a complete solution to circuit verification problems. In modern designs with 
-large number of logic components and complicated architecures, it is impractical to simulate all possible 
-test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
-cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
-is a good example when simulation fails. Failure occurs when only 5 entries in an $84\times 16$ look-up table
-(LUT) are activated,  which is rarely used in most divisions. Because of the limitation of simulation,  
-test engineers from Intel failed to detect the bug,  which brought $\$475$ million dollors recalling bill
-for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.7\textwidth]{newfig/designflow.eps}
-}
-\caption{Typical hardware design flow}
-\label{fig:designflow}
-\end{figure}
-}
-
-\section{Formal Verification}
-% TODO this part require less modifications
-Another method developed besides simulation is \emph{formal verification}, it utilizes 
-mathematical theory to reason about the correctness of hardware designs.
-Formal verification can provide $100\%$ fault coverage from 2 aspects. On the one hand,  it formalize properties
-for the circuit model which are irrelevant to specific input signals,  and prove them mathematically.
-On the other hand,  it adopts formal languages to strictly describe the design intents and detailed implementation, 
-and deduces circuit function from the implementation. These descriptions are named as {\it specifications}.
-Formal verification has two main forms: property checking and equivalence 
-checking.
-
-{\it Property checking} (or property verification) verifies
-that a design satisfies certain given properties. Property checking is done mainly 
-in the form of theorem proving, model checking, or approaches which 
-combine the two.
-\begin{enumerate}
-\item \emph{Theorem proving} \cite{theoremproving:91} requires the existence of
-mathematical descriptors of the specification and implementation of the 
-circuit. Theorem provers apply mathematical rules to these descriptors to
-derive new properties of the specification. In this way, the tool can reduce
-a proof goal to simpler sub-goals, which can be automatically verified.
-However, generating the initial proof-goal requires extensive guidance from
-the user, so there is an overall lack of automation in theorem 
-proving.
-\item \emph{Model checking} \cite{modelcheck:99} is an approach
-to verifying finite-state systems where specification 
-properties are modeled as a system of 
-logic formulas. The design is then traversed to check if the 
-properties hold. If the design is found to violate a
-particular property, a counter-example is generated which exercises the
-incorrect behavior in the design. Such counter-examples allow the designer
-to trace the behavior and find where the error in the design lies.
-Modern model checking techniques use the result to automatically refine
-the system and perform further checking.
-These tools are typically automated, and thus have found widespread 
-use in CAD tool suites.
-\end{enumerate}
-
-{\it Equivalence Checking} verifies that two different representations of
-a circuit design have equivalent functionality. An example of equivalence
-checking as it applies to the hardware design flow is shown in
-Fig. \ref{fig:equivflow}.
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.5\textwidth]{./figures/designVerification}
-}
-\caption{Equivalence checking as applied to the hardware design flow.}
-\label{fig:equivflow}
-\end{figure}
-}
-
-There are two major
-equivalence checking techniques: graph-based
-and satisfiability-based.
-\begin{enumerate}
-\item \emph{Graph-based} techniques construct a canonical graph 
-representation, such as a Binary Decision Diagram (\emph{BDD}) or one of
-its many variants, of each circuit. A linear comparison is then conducted to 
-determine whether the two graphs are isomorphic. Since the graph 
-representation is canonical, the graphs of the two circuits will be 
-equivalent if and only if the circuits perform the same function.
-\item \emph{Satisfiability} techniques construct a miter of the two circuits,
-typically in a graph such as an And-Inverter graph (\emph{AIG}). A
-\emph{miter} is a combination of the two circuits with one bit-level output, which 
-is only in a "1" state when the outputs of the circuits differ given 
-the same given 
-input, as shown in Fig. \ref{fig:miter}. 
-A satisfiability (\emph{SAT}) tool \cite{csat} 
-is then employed to simplify the graph and find a solution to the miter, 
-i.e. find an input for which the 
-miter output is "1". If a solution is found, this solution acts as a 
-counter-example of when the circuit outputs differ; otherwise the circuits
-are functionally equivalent.
-\end{enumerate}
-
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.8\textwidth]{./figures/betterMiter}
-}
-\caption{A miter of two circuits.}
-\label{fig:miter}
-\end{figure}
-}
-
-%\subsection{Computer-Algebra Based Formal Verification}
-Certain formal verification methods use \emph{computer-algebra} and \emph{algebraic
-geometry} techniques based on mathematical theories.
-Unlike SAT-based verification, modern algebraic geometry 
-techniques do not explicitly solve the constraints to find a solution; 
-rather, they reason about the presence or absence of solutions, or explore
-the geometry of the solutions.
-These methods \cite{Avrunin:CAV} \cite{condrat-tacas07} \cite{gbverify:2007} 
-transform the circuit design into a polynomial system. Typically, this system
-of polynomials is then used to compute a Gr\"obner basis \cite{gb_book}. 
-Computation of Gr\"obner bases allows for 
-the easy deduction of important properties of a polynomial system, 
-such as the presence or absence of 
-solutions. These properties are then leveraged to perform 
-verification. Unfortunately, such a computation 
-has been shown to be doubly 
-exponential in the worst case, and thus these methods have not been 
-practical for real-world applications. However, recent
-breakthroughs in computer-algebra hardware verification have shown
-that it is possible to overcome the complexity of this computation while
-still utilizing the beneficial properties of a Gr\"obner bases
-\cite{lv:phd}.
-
-\section{Importance of Word-level Abstraction}
-Most formal verification techniques can benefit from word-level abstractions 
-of the circuits they verify.
-There are several advantages in exploiting word-level information for
-verification. A number of designs have  their
-datapaths and/or system-level models described as word-level RTL
-models.  Exploiting word-level instead of bit-level information is one
-way of abstraction -- a key technique to reduce the state space of a
-FSM. It has the effect of combining sets of states with similar 
-properties. During reachability analysis, if we use bit-level
-variables to  represent the states, the representations may become too
-large to handle. However, when a ``bundle" of bit-level variables are
-represented as {\it only one word-level variable}, the set of
-reachable states can be represented by a  word-level constraint
-expression; which may lower verification complexity.
-
-% Abstraction is defined as state-space reduction, i.e{\text . }abstraction
-% reduces state-space by mapping the set of states of a system to a smaller 
-% set of states. Because the new representation contains fewer states, it
-% is easier to comprehend and thus easier to use. 
-% Word-level abstraction focuses specifically on abstracting a word-level
-% representation of a circuit out of a bit-level representation. For example,
-% a bit-level representation of an integer multiplier is represented by a
-% collection of Boolean inputs and outputs, whereas a word-level
-% abstraction hides the underlying logic and represents the circuit as two 
-% integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
-% multiplier increases, the logical implementation of the multiplier grows (typically
-% exponentially) while the word-level abstraction stays the same.
-
-Word-level abstractions have a wide variety applications in formal 
-verification. Theorem proving techniques can leverage abstraction as an 
-automatic decision procedure or as a canonical reduction 
-engine. For example, since RTL is composed of circuit blocks that represent 
-the underlying circuit, RTL verification methods can exploit 
-abstractions of these blocks.
-This is seen in the following RTL verification methods:
-\begin{itemize}  
-\item Model checking \cite{kroening:model}, 
-where an approximation abstraction of RTL blocks is generated and then 
-refined.
-\item Graph-based equivalence 
-checking \cite{WLS} \cite{arditi:bmd}, where abstraction methods are used
-to generate a canonical word-level graph representation of the circuit.
-\item Satisfiability-based equivalence checking \cite{lpsat}, where 
-abstractions are used identify symmetrics and similarities in order to 
-minimize the amount of logic that is sent to the 
-SAT tool. 
-\end{itemize}
-
-Other equivalence checking techniques that employ abstractions 
-include satisfiability modulo theory ({\it SMT}) techniques \cite{boolector} \cite{bryant:tacas07}, 
-which are similar to SAT except they operate on higher-level data
-structures (integers, reals, bit vectors, etc.), as well as 
-constraint solving techniques \cite {ms:research} \cite{tew:iccad08}.
-In general, RTL equivalence checking approaches would ideally maintain a 
-high-level of abstraction while still retaining sufficient lower-level 
-functional details  (such as bit-vector size, precision, etc) 
-\cite{gupta_survey}.
-
-Word-level hardware abstractions also have applications in RTL and datapath 
-synthesis \cite{demicheli:iccad_98} \cite{demicheli:dac_99}
-\cite{demicheli:tcad_03}. 
-Abstractions of circuits allow for design reuse, which allows for tool-automated 
-synthesis of larger circuit blocks.
-Since hardware design specifications tend to be word-level, synthesis tools 
-can use these larger circuit blocks to generate and optimize the
-datapaths and create the RTL of the system. Thus, in order for a circuit to 
-be used by these automated synthesis tools, its word-level abstraction must
-be known.
 
@@
 
-\section{Objective, Motivation and Contribution of this Dissertation}
+\section{Objective and Contribution of this Dissertation}
 This research proposes a
 set of new, promising approaches for {\it word-level representation,
-reachability analysis and abstraction} with applications to SEC, sequential GF
-arithmetic functional verification and 
-$k$-BMC with abstraction refinement. Our techniques operate at the word-level and they are based
+reachability analysis and abstraction} for sequential design verification techniques. 
+Our approaches operate at the word-level and are based
 largely on the concepts from {\it algebraic geometry}.
@@
 
-The use of algebraic geometry
-for sequential circuit verification and symbolic model checking has
-been presented before. Avrunin presented the
-concept of symbolic MC using algebraic geometry in
-\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
-GB-algorithms for CTL, LTL and 
-bounded MC over Boolean rings. However, these approaches are a
-straight-forward transformation of the problem to {\it bit-level}
-Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
-the concepts of word-level reachability, abstraction-refinement using
-interpolation or UNSAT cores, etc., that we desire was not the focus of
-\cite{Avrunin:CAV} \cite{vardi-iasted07}.
+% TODO move to prev
+% The use of algebraic geometry
+% for sequential circuit verification and symbolic model checking has
+% been presented before. Avrunin presented the
+% concept of symbolic MC using algebraic geometry in
+% \cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
+% GB-algorithms for CTL, LTL and 
+% bounded MC over Boolean rings. However, these approaches are a
+% straight-forward transformation of the problem to {\it bit-level}
+% Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
+% the concepts of word-level reachability, abstraction-refinement using
+% interpolation or UNSAT cores, etc., that we desire was not the focus of
+% \cite{Avrunin:CAV} \cite{vardi-iasted07}.
 
-{\bf The Algebraic Geometry Model and Rationale:} We will represent
+\subsection{Word-level Reachability Analysis}
+In this dissertation, we propose a method to perform reachability analysis at the word-level. 
+	The given FSM is modeled as a system of polynomials over a finite field,
+	where the state space is mapped to the solutions of polynomial system.
+	Our proposed algorithm utilizes concepts of ideal-variety correspondence in algebraic geometry.
+	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
+  state-space.
+
+In this dissertation we represent
 the FSMs -- the transition relations -- by means of a set of
@@
 in algebraic geometry can be applied to implement symbolic reasoning
-about state reachability. Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
-our model provides a single, unified and bit-precise representation
-for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
-constraints.
+about state reachability. 
+% Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
+% our model provides a single, unified and bit-precise representation
+% for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
+% constraints.
 
 The decision and abstraction procedures in our setting will rely on
-the theory and technology of {\it \Grobner bases (GB)}. GB-based
+the theory and technology of {\it \Grobner bases}. GB-based
 algebraic reasoning is very powerful; in fact it is known to be
@@
 
-\subsection{Dissertation Contributions}
-% TODO revise this to contributions
-In this dissertation, we focus on sequential circuit verification problems and 
-propose solutions consisting of three main contributions: 
-\begin{enumerate}[{1)}]
-\item A method to perform reachability analysis at the word-level. 
-	The given FSM is modeled as a system of polynomials over finite field,
-	where the state space is mapped to its solution space.
-	Our proposed algorithm utilize concepts in algebraic geometry including ideals and varieties.
-	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
-  state-space.
-\item Using the theory of FSM traversal, we apply the algorithm to verify the 
-function of sequential GF arithmetic circuits. Our proposed approach uses GB-engines efficiently 
-and can verify sequential multipliers with 162-bit datapaths.
-\item We explore the concept, and the computation, of unsatisfiable (UNSAT)
-  cores of  a set of polynomials using Gr\"obner bases calculation. We apply the UNSAT core extraction 
-  to abstraction-refinement
-  techniques such as bounded model checking (BMC). 
-\end{enumerate}
+\subsection{Application to Sequential Galois Field Arithmetic Circuits Verification}
+Sequential Galois field (GF) arithmetic circuits can be modeled as a special type of 
+FSM, where the pre-loaded operands are mapped to initial states and pseudo outputs
+after $k$ clock-cycles are recognized as the final reached state after $k$ transitions.
+Therefore, the word-level FSM traversal algorithm can be applied to 
+verify the correctness of final reached state, {\it i.e.} the functional correctness of
+a sequential arithmetic circuit.
+
+In our proposed approach, word-level abstraction is employed to generate the 
+word-level signature of the combinational logic component of the sequential 
+arithmetic circuit. This abstraction requires GB computation, which usually has high time/space 
+complexity. We propose several improvements to simplify the GB computation procedure
+and make the entire algorithm execution practical. As a result, we successfully verify sequential 
+multipliers with 162-bit datapaths.
+
+\subsection{UNSAT Cores in Algebraic Geometry for Abstraction Refinement}
+Abstraction is an effective method to lower the cost to traverse the state space.
+In modern model checkers, abstraction is used to refine the model 
+during the iterative execution of the tool. UNSAT core is widely used 
+as an important component of abstraction refinement. The reason is that UNSAT cores can provide 
+information to indicate the redundancy in the state variables, and that information is necessary for the refinement 
+process.
+
+In this dissertation, we explore the concept, and the computation, of unsatisfiable (UNSAT)
+  cores of  a set of polynomials using Gr\"obner bases calculation. We also propose
+a number of heuristics by cracking the Buchberger's algorithm to reduce the size of UNSAT core. 
+  We apply the UNSAT core extraction 
+  to a bounded model checker (BMC) with abstraction refinement.
+
+% \subsection{Dissertation Contributions}
+% In this dissertation, we focus on sequential circuit verification problems and 
+% propose solutions consisting of three main contributions: 
+% \begin{enumerate}[{1)}]
+% \item A method to perform reachability analysis at the word-level. 
+% 	The given FSM is modeled as a system of polynomials over finite field,
+% 	where the state space is mapped to its solution space.
+% 	Our proposed algorithm utilize concepts in algebraic geometry including ideals and varieties.
+% 	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
+%   state-space.
+% \item Using the theory of FSM traversal, we apply the algorithm to verify the 
+% function of sequential GF arithmetic circuits. Our proposed approach uses GB-engines efficiently 
+% and can verify sequential multipliers with 162-bit datapaths.
+% \item We explore the concept, and the computation, of unsatisfiable (UNSAT)
+%   cores of  a set of polynomials using Gr\"obner bases calculation. We apply the UNSAT core extraction 
+%   to abstraction-refinement
+%   techniques such as bounded model checking (BMC). 
+% \end{enumerate}
 
@@
 Chapter \ref{ch:prelim_GF} covers preliminary
-concepts and notation on finite fields, and the methodology about design of multipliers in finite fields.
-Chapter \ref{ch:ideals} provides a theoretical background of algebraic geometry and Gr\"obner bases.
+concepts and notation on finite fields, and the methodology about design of arithmetic circuits in finite fields.
+Chapter \ref{ch:ideals} provides a theoretical background on algebraic geometry and Gr\"obner bases.
 Chapter \ref{ch:reacha} describes the basic
@@
 Chapter \ref{ch:conclude} concludes the dissertation and outlines potential future research for
-continuation of this work.
+continuation of this work. Appendix provides theory and methodology on characterization of normal basis,
+as well as the construction of optimal normal basis and application to normal basis multiplier design.

### 2016-12-09 dissertation/ONB.tex
dissertation v2:abstract & appendix update 2

@@
 %\usepackage{e-jc}
+\usepackage{amsfonts}
 
@@
 
-\date{Last Update: Apr 17, 2013}
+\date{Last Update: May 14, 2013}
 
@@
 $g(\sigma)\eta = 0$, and\\
-$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$.
+$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$.\\
 
@@
 
-Suppose we take only one polynomial only contains T(next state), like\\
+\textbf{Lenstra's Algorithm}\\
+Step 1: Take an arbitrary element $\theta \in F_{q^n}$, determine $Ord_\theta(x)$.\\
+Step 2: If $Ord_\theta(x) = x^n - 1$ then algorithm ends.\\
+Step 3: Calculate $g(x) = (x^n - 1)/Ord_\theta(x)$, and solve $\beta$ from $g(\sigma)\beta = \theta$.\\
+Step 4: Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to step 2;
+	otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a non-zero element $\eta$ such that $g(\sigma)\eta = 0$,
+	replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, then go to step 2.\\
+
+\section{Optimal Normal Basis Characterization}
+\textbf{Definition of Optimal Normal Basis}\\
+To define optimal normal basis, it is necessary to find out the minimum number
+of terms in multiplication table(defined below).
+\textit{Theorem 3}\ \ If N is a normal basis for $F_{p^n}$ with multiplication table $M_T$, then non-zero terms in 
+matrix $C_N\leq 2n-1$.\\
+\textit{Proof}\ \ Let $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}$, denote $\beta^{p^i}$ by $\beta_i$. 
+Then $\sum_{i=0}^{n-1} \beta_i = trace \beta$. \\
+Denote $trace \beta$ with $b$, consider $n\times n$ matrix (multiplication table for lowest digit) $M_0$. Then\\
+$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$.\\
+Therefore, the sum of all rows in $M_0$ is an n-tuple with $b$ as the first element and zeros elsewhere.
+So the first column must contain at least 1 nonzero element, and other columns must get a sum of zero 
+while keeping every row linearly independent so there should be at least 2 non-zero elements each column.\\
+For $F_{2^n}$, easy to figure out the first row must contain a "1" because $\beta \beta^{2^0} = \beta^{2^1}$.
+for the rest rows $\beta \beta^{2^i} \neq \beta^{2^{i+1}}$,
+so there must be at least 2 "1"s to fulfil the equation.\\
+\textbf{Multiplication table}\\
+$\beta \left( \begin{array} {lcr}
+ \beta \\ 
+\beta^2 \\ 
+\cdot \\ 
+\cdot \\ 
+\cdot \\ 
+\beta^{2^{n-1}} 
+\end{array} \right) = M_T \left( \begin{array} {lcr}
+ \beta \\ 
+\beta^2 \\ 
+\cdot \\ 
+\cdot \\ 
+\cdot \\ 
+\beta^{2^{n-1}} 
+\end{array} \right)$. \\
+
+Complexity $C_N \geq 2n - 1$.\\
+
+\textbf{ONB Existance Theory}\\
+Key words: k-th primitive root of unity, Euler's criterion, quadratic residues\\
 \hspace{8mm}\par
-$ideal\ G = T^2 + (1+X) \cdot T + X$\\
+Type I ONB: n+1 is prime and q is primitive in $\mathbb{Z}_{n+1}$, then the n nonunit (n+1)th roots of unity form ONB.\\
 \hspace{8mm}\par
-Since we get G from GB based image function, so G is already a Groebner Basis itself.
-Considering Theorem 1 $\sim$ 3, it's easy to do ideal operation if its generator is only one polynomial.
-So using $G+G'$ we can get intersection, using $G \cdot G'$ we can get union. If we take V = Universe = $<vanishing\ polynomial>$, we can get ideal quotient $I(V):I(W)$ as complementary set for
-specific varieties.
+Type II ONB: 2n+1 is prime, if (1) 2 is primitive in $\mathbb{Z}_{2n+1}$, OR \\
+(2) $2n + 1 \equiv 3(mod 4)$ and 2 generates the quadratic residues in $\mathbb{Z}_{2n+1}$.\\
+Then $\alpha = \gamma + \gamma^{-1}$ generates ONB, where $\gamma$ is a primitive (2n+1)th root if unity.\\
 
+\textbf{Low-complexity Normal Basis design}\\
+N/A.\\
+
+\section{Other NB Properties \& Problems}
+\textbf{N-poly}\\
+N-poly is irreducible polynomial with linearly independent roots. Normal basis is a set of roots of N-poly.
+\begin{itemize}
+\item[-] \textbf{Corollary 2}\ \ \ \ For irreducible polynomial $f(x) = x^n + a_1x^{n-1} + ... + a_n \in F_{2^n}[x]$, 
+		that is an N-poly if and only if $a_1 \neq 0$.
+\end{itemize}
+
+\textbf{Problems and Discussion}\\
 \hspace{8mm}\par
-\textbf{More about Ideal Quotient}
-\begin{itemize}
-\item[-] \textbf{Definition 1}\ \ \ \ If I,J are ideals in $k[x_1,...,x_n]$, then $I:J$ is the set \\ \indent $\{f\in k[x_1,...,x_n]:fg\in I\ for \ all\ g\in J\}$
-\end{itemize} \par
-Why we can get complementary set $\overline {reached}$ through this?\par
-First, we only care about varieties. Say we redefine the "equal" as $V(I)==V(J)\Leftrightarrow I == J$.\par
-Second, we only care about the ideal/GB {\bf G} contains only 1 generator f.
-Say this polynomial f is a function about next state (word level) T, then $f(T=V(\bf{G}))==0$.
-To ensure then unique of generator f, we'll reduce f (or say {\bf G}) every time.\par
-Third, for ideal quotient ${\bf U:G}$ here, we need to find f from $g~f==vanish$.
-This means $<f>\bigcup <g>==<vanish>$.
-How can we develop a division algorithm to get exactly a factor without any remainders?
-Let's see 3 examples following:\\
+Prob1: how to prove $\lambda$-Matrix is multiplication table?\\
+\indent I think this is a good way to understand the essence of normal basis 
+theory. This problem can be rewrite like this:\\
+Why there exist a rotating symmetry in $\lambda$-Matrix that
+ $\lambda_{0j}^{(k)} = \lambda_{jk}^{(0)}$? Andrew guess it's a Frobenius
+ element matrix rotating symmetry involves with field automorphism. Need 
+further consultation on this part. How to get this property? Check out the
+ last part of my note.\\
 \hspace{8mm}\par
-\indent $T^2+(1+X)\cdot T+X$\\
+Prob2: Proof of Lemma 5.\\
+(Already proved. Please check out the note.)\\
 \hspace{8mm}\par
-\indent $T^2+T$\\
-\hspace{8mm}\par
-\indent $T^2+1$\\
+Prob3: Fast algorithm for determining $\sigma$-Order.\\
+(Do not need FAST algorithms. Just do a $(O(n^2))$ scan.)\\
 

### 2016-12-09 dissertation/abstract.tex
dissertation v2:abstract & appendix update 2

--- 2016-10-24 dissertation/abstract.tex
@@
 \begin{center}{\bf ABSTRACT}\end{center}
+Sequential equivalence and model checking techniques are widely
+employed in hardware and software verification. With the increasing
+size and complexity of hardware/software systems, abstraction has
+become a key component to address scalability.
+This dissertation investigates new abstraction techniques at word-level and their applications 
+on sequential circuits verification. By bundling together $k$ bit-level state-variables into
+one word-level constraint expression, the state-space can be construed
+as a polynomial function over the finite (Galois) field of $2^k$
+elements. As long as the state-space are modeled as polynomial ideals
+over word-level variables, algebraic geometry techniques including ideal manipulations,
+Gr\"obner bases reasoning and finite field algebra are researched to 
+facilitate new word-level abstractions.
 
-Verification of sequential circuits remains to be a topic of all time. With increasing size of new
-integrated circuits, sequential circuit designers may face much more complicated problems on 
-logic errors, unexpected delays and failures on critical path. 
-This dissertation addresses the problem of sequential circuit verification at the word-level
-and is based on concepts derived from algebraic geometry.
+Algebraic geometry offers a very powerful set of tools to reason about
+the properties of the solutions to a system of polynomial
+constraints. Moreover, the algebraic model inherently provides a
+framework for abstraction. 
+The algebraic geometry algorithms (Groebner basis computations) exhibit
+high complexity which prevent it from being widely used in conventional techniques. 
+Interestingly, recent experience has shown that
+by analyzing the topology of circuits in conjunction with Gr\"obner
+basis computations helps to identify the structure and symmetry
+inherent in the problem, this lower the computational complexity.
 
-Analyzing the sequential circuits at word level is an efficient way
-of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
-state-space. We manage to model the verification properties and the gate-level sequential circuit 
-implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
-their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
-techniques from algebraic geometry can be used to reason about the state-space on high-level.
-In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
-polynomial ideals. Without actually solving the system, the states in state space can be investigated 
-by manipulating these ideals.
-
-We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
-analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
-verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
-abstraction-refinement techniques for verification, we propose to investigate a word-level 
-analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
-not only derive new algorithms and techniques, but will also consider efficient CAD 
-implementations to target sequential equivalence and model checking problems. 
-% \end{abstract}
+Formal verification of sequential circuits requires the analysis of
+the underlying finite state machine. Reachability analysis becomes a
+fundamental tool for this purpose. We devised traditional traversal algorithms 
+using the new abstraction method to perform
+the reachability analysis at word-level. 
+The abstractions are also exploited on functional
+verification of sequential Galois field arithmetic circuits, where conventional 
+techniques fail to provide scalability. At last, we propose a new method to extract 
+unsatisfiable cores and utilize the information about the cores to 
+perform model checking with abstraction refinement.

### 2016-12-09 dissertation/appendix.tex
dissertation v2:abstract & appendix update 2

@@
 \begin{Lemma}
+\label{lem:3}
 The minimal (and characteristic) polynomial for $\sigma$ is $x^n-1$.
@@
 \begin{Corollary}
+\label{corol:1}
 An element $\alpha \in \F_{p^n}$ is a {\bf normal element} if and only if $Ord_{\alpha,\sigma}(x) = x^n-1$.
 \end{Corollary}
-\subsection{Construction of General Normal Basis}
+\begin{Proof}
+Normal bases require $\{\beta,\beta^2,\dots,\beta^{2^{n-1}}\}$ to be linearly independent,
+which is equivalent to the following expression:
+$$\forall f(x)\in \F_p[x],~deg(f(x))<n$$
+Furthermore, because there is no annihilators in the $T$-Order, this implies that
+$$Ord_{\alpha,\sigma}(x) = x^n-1$$
+\end{Proof}
+
+From corollary above we can deduce another form of criterion of normal element:
+\begin{Theorem}
+\label{thm:1}
+Given finite field $\F_{2^n}$, the field characteristic $p = 2$.
+Define $t = p^e$ such that $n = kp^e, gcd(k,p) = 1$, so $t=1$ if $n$ is an odd integer. Then 
+$x^n - 1$ can be factorized as 
+$$(\varphi_1(x)\varphi_2(x)\cdot\cdot\cdot\varphi_r(x))^t$$
+Additionally, define $\Phi_i(x) = (x^n - 1)/\varphi_i(x)$. We assert that
+
+An element $\alpha \in 	F_{p^n}$ is a normal element if and only if $\Phi_i(\sigma)\alpha \neq 0, i = 1,2,\dots,r$.
+\end{Theorem}
+\begin{Proof}
+Let us analyze the auxiliary polynomial $\Phi_i(x) = \frac{x^n - 1}{\varphi_i(x)}$ first.
+$$\Phi_i(\sigma)\alpha \neq 0 \Leftrightarrow \text{No factor in }x^n-1\text{ annihilates }\alpha$$
+On the other hand, according to Lemma \ref{lem:3}, it is a known 
+fact that \emph{any annihilator always divide} $x^n-1$. As a result, the only possible situation is
+$$Ord_{\alpha,\sigma}(x) = x^n-1$$
+Using Corollary \ref{corol:1} we deduce that $\alpha$ is a normal element.
+\end{Proof}
+
+The normal element identification can also be described from the nullspace perspective:
+\begin{Theorem}
+\label{thm:2}
+Let $W_i$ be the nullspace of $\varphi_{i}^{t} (x)$ and $\widetilde{W_i}$ the
+nullspace of $\varphi_{i}^{t-1} (x)$. Let $\overline{W_i}$ be any subspace of
+$W_i$ such that $W_i = \overline{W_i}\oplus \widetilde{W_i}$. Then
+$$\F_{p^n} = \displaystyle\sum_{i=1}^{r} \overline{W_i}\oplus \widetilde{W_i}$$
+is a direct sum where $dim(\overline{W_i}) = d_i$ and $dim(\widetilde{W_i}) = (t-1)d_i$.
+
+Using above setup we define an element $\alpha \in \F_{p^n}$ with 
+$\alpha = \sum_{i=1}^{r} (\overline{\alpha_i} + \widetilde{\alpha_i}), \overline{\alpha_i} \in \overline{W_i}, \widetilde{\alpha_i} \in \widetilde{W_i}$,
+as a normal element if and only if $\overline{\alpha_i} \neq 0,~ \forall i = 1,2,\dots,r$.
+\end{Theorem}
+
+Ultimately, since there always exist at least one element fulfilling requirements in Theorem \ref{thm:1}
+or Theorem \ref{thm:2}, we obtain following theorem:
+\begin{Theorem}[Normal Basis Theorem over Finite Fields]
+There always exists a normal basis of $\F_{p^n}$ over $\F_p$.
+\end{Theorem}
+\subsection{Construction of General Normal Bases}
+After proving the existence of normal bases, the upcoming question is that how to find such a normal basis/element.
+For general normal basis identification, there are two methods widely used: the L\"uneburg's algorithm
+and Lenstra's algorithm.\\
+\textbf{L\"uneburg's Algorithm:}
+\begin{enumerate}[{1)}]
+\item Randomly pick an element $\alpha$ from $\F_{p^n}$. For each $i = 0,1,\dots,n-1$, 
+compute $\sigma$-Order $f_i = Ord_{\alpha^i}(x)$. Then $x^n - 1 = lcm(f_0,f_1,...,f_{n-1})$.
+\item Apply factor refinement to set $\{f_0,\dots,f_{n-1}\}$ and obtain $f_i = \prod_{1\leq j\leq r} g_{j}^{e_{ij}}, i = 0,1,\dots,n-1$.
+We can write the result as an $i\times j$ matrix.
+\item For each $j$, find an index $i_j$ (denote as $i(j)$) such that $e_{ij}$ is maximum in the $j$-th column.
+\item Let $h_j = f_{i(j)}/g_{j}^{e_{i(j)j}}$, take $\beta_j = h_j(\sigma)\alpha^{i(j)}$. Then
+$$\beta = \displaystyle\sum_{j=1}^{r} \beta_j$$
+is a normal element.
+\end{enumerate}
+
+L\"uneburg's algorithm starts with a random element and ends with a normal element. The justification of 
+its soundness is as follows:
+\begin{Proposition}
+L\"uneburg's algorithm always generate a normal element over field $\F_{p^n}$.
+\end{Proposition}
+
+\begin{Proof}
+First, according to the definition of $T$-Order:
+$$f_i = Ord_{\alpha^i} \Leftrightarrow f_i(\sigma)\alpha^i = 0$$
+
+Use Lemma \ref{lem:3}, the minimal/characteristic polynomial for $\sigma$ is $x^n-1$ implies that
+any annihilator of $\alpha^i$ ({\it i.e.} $f_i(\sigma)$) divides $x^n-1$. Meanwhile, $\{\alpha^i~|~0<i<n-1\}$ forms a (standard) basis
+of field $\F_{p^n}$, this means $\alpha_i$ are linearly independent with each other ($i=0,1,\dots,n-1$).
+Since linear map $f(\sigma)$ guarantees that all elements in the field 
+have order $lcm(f_0,\dots,f_{n-1})$, i.e. $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
+By contradiction we can prove that there is no factor of $x^n-1$ that can be divided by the product 
+of elements in set $\{f_0,f_1,\dots,f_{n-1}\}$. This actually corresponds to the assertion in the first step of the algorithm:
+$$x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$$
+
+In the second step, after factorization of set $\{f_i\}$, we transform $f_i(\sigma)\alpha^i=0$ to
+$$h_j(\sigma)\cdot g_{j}^{e_{i(j)j}}(\sigma) \cdot \alpha^{i(j)} = 0$$
+
+Furthermore we have 
+$$g_{j}^{e_{i(j)j}}(\sigma)\beta_j = 0 \Leftrightarrow Ord_{\sigma,\beta_j}(x) = g_{j}^{e_{i(j)j}}(x)$$
+
+Consider the facts: 1) elements in set $\{g_j\}$ are relatively prime; 2) $g_{j}^{e_{i(j)j}}$ is the maximum factor; 
+3) $x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$, we deduce that 
+$$x^n-1 = \prod_j g_{j}^{e_{i(j)j}}(x) = \prod_jOrd_{\sigma,\beta_j}(x)$$
+
+As a result
+$$x^n-1 = Ord_{\sigma,\beta}(x) \implies \beta\text{ is a normal element.}$$
+\end{Proof}
+
+L\"uneburg's algorithm is an analytic method, which requires relatively high computational complexity 
+mainly because of the factorization in its second step. To overcome the high cost, an inductive method 
+is proposed, which is the Lenstra's algorithm. It allows for setup of heuristics to accelerate the procedure.
+Before introducing the details of the algorithm, we demonstrate two preliminary lemmas.
+\begin{Lemma}
+\label{lem:4}
+For an arbitrary element $\theta \in \F_{p^n}$ that $Ord_\theta(x) \neq x^n - 1$,
+let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.
+\end{Lemma}
+\begin{Proof}
+Assume $\gamma$ is the desired normal element. From the definition of normal element we derive 
+$$\exists f(x)\in \F_{p^n}[x],~~f(\sigma)\gamma = \theta$$
+
+Then using the definition of $T$-Order, we derive
+$$Ord_{\sigma}\theta = 0 \implies (Ord_{\sigma}f(\sigma))\gamma = 0$$
+
+Since $\gamma$ is the normal element, $Ord_\gamma(x) = x^n-1\implies x^n-1~|~(Ord_\theta(x)f(x))$.
+Furthermore,
+$$g(x) = \frac{x^n-1}{Ord_\theta(x)} \implies x^n-1\Bigm|\left(\frac{x^n-1}{g(x)}\cdot f(x)\right)\implies g(x)~|~f(x)$$
+
+Let $f(x)=h(x)g(x)$. Then
+$$g(\sigma)(h(\sigma)\gamma) = \theta$$
+
+Therefore, $\exists \beta$ {\it i.e. }$g(\sigma)\beta = \theta$. Concretely, $\beta = h(\sigma)\gamma$.
+\end{Proof}
+
+\begin{Lemma}
+\label{lem:5}
+Define $\theta$ and $g(x)$ as in Lemma \ref{lem:4}. Assume there exists a solution 
+$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Respectively there exists a non-zero element $\eta$ such that
+$$g(\sigma)\eta = 0$$
+and
+$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
+\end{Lemma}
+\begin{Proof}
+Assume $\gamma$ is the desired normal element. Similarly,
+$$\exists \eta = Ord_\theta(\sigma)\gamma \neq 0,~~g(\sigma)\eta = 0$$
+
+Follow the setup in Lemma \ref{lem:4}, we have
+$$g(\sigma)\beta = \theta,~~\frac{x^n-1}{Ord_\theta(x)}\Bigm|_\sigma\cdot \beta = \theta$$
+
+Subsequently,
+$$Ord_\theta(x)\cdot \frac{x^n-1}{Ord_\theta(x)}\Bigm|_\sigma\beta = 0 = Ord_\beta(x)\beta$$
+which implies
+$$Ord_\theta(x)\bigm|Ord_\beta(x),~~deg(Ord_\theta(x)) \leq deg(Ord_\beta(x))$$
+Combining with the assumption in the lemma: $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$, we derive
+$$deg(Ord_\beta(x)) = deg(Ord_\theta(x))\implies Ord_\theta(x) =Ord_\beta(x)$$
+
+In the next part we need to prove $g(x)\perp Ord_\theta(x)$ by contradiction. Assume 
+$h(x) = gcd(g(x),Ord_\theta(x))\neq 1$, and
+$$g(\sigma)\beta = a(\sigma)h(\sigma)\beta = \theta,~~Ord_\theta(\sigma)\theta = a(\sigma)b(\sigma)h^2(\sigma)\beta$$
+However, $Ord_\theta(x) =Ord_\beta(x)$ indicates that $Ord_\beta(x)\beta = b(x)h(x)a(x)\beta = 0$. Thus,
+$Ord_\theta(x)=b(x)$. This is true if and only if $h(x)=1$. As a result, we assert that
+$$g(x) \perp Ord_\theta(x)$$
+
+Consider $g(\sigma)\eta = 0 \implies Ord_\eta(x)\bigm|g(x)$. It further implies that 
+$$Ord_\theta(x)\perp Ord_\eta(x) \implies Ord_{\theta+\eta}(x) = Ord_\theta(x)\cdot Ord_\eta(x)$$
+
+Since $\eta \neq 0$, we derive the result
+$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
+\end{Proof}
+
+Based on Lemmas \ref{lem:4} and \ref{lem:5}, the Lenstra's algorithm is described below:\\
+\textbf{Lenstra's Algorithm}
+\begin{enumerate}[{1)}]
+\item Take an arbitrary element $\theta \in \F_{p^n}$, determine $Ord_\theta(x)$.
+\item If $Ord_\theta(x) = x^n - 1$ then algorithm terminates and return $\theta$ as a normal element.
+\item Otherwise, compute $g(x) = (x^n - 1)/Ord_\theta(x)$, and solve $\beta$ from $g(\sigma)\beta = \theta$.
+\item Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to 2);
+otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a non-zero element $\eta$ such that $g(\sigma)\eta = 0$,
+replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, and go to 2).
+\end{enumerate}
+
+Lenstra's algorithm is an approximation algorithm in nature. For each iteration $Ord_\theta(x)$ monotonically increase,
+so it will finally reach termination condition $Ord_\theta(x) = x^n-1$.
 \subsection{Bases Conversion}
@@
 % Multi table <=> \lambda matrix
+In Section \ref{sec:nbdesign}, we introduce the concept of multiplication table and explained how its 
+entries are transformed from the $0$-th $\lambda$-Matrix. Actually we can mathematically prove the relationship 
+between them:
+\begin{Theorem}
+Multiplication table $T$ is actually a conjugate of the $\lambda$-Matrix $M$. 
+We can use an equation $M_{i,j}^{(0)} = T_{j-i,-i}$ to demonstrate the relation between their entries.
+\end{Theorem}
+\begin{Proof}
+Recall the definition of $\lambda$-Matrix over field $\F_{2^k}$:
+$$C = A\times B = \left(\sum_i a_i\beta^{2^i}\right)\left(\sum_jb_j\beta^{2^j}\right) = \sum_i\sum_ja_ib_j\beta^{2^i}\beta^{2^j}$$
+Then there always exist $\lambda_{ij}^{(l)}$, such that 
+$$\beta^{2^i}\beta^{2^j} = \sum_l\lambda_{ij}^{(l)}\beta^{2^l}$$
+We call it the cross-product term. The $l$-th bit of the product $C$ is
+$$c_l = \sum_i\sum_ja_ib_j\lambda_{ij}^{(l)}$$
+For the sake of simplification, we let $i=0$, thus 
+$$\beta\cdot\beta^{2^j} = \sum_l\lambda_{0j}^{(l)}\beta^{2^l}$$
+Since the multiplication table is defined as
+\begin{equation}
+\beta
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
+\end{bmatrix}
+= {\bf T}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
+\end{bmatrix}
+\end{equation}
+The $j$-th row of $T$ can be written as
+\begin{align}
+\beta\cdot\beta^{2^j} =& \sum_l\lambda_{0j}^{(l)}\beta^{2^l} & \text{(multi-table definition)} \nonumber\\
+					=& \sum_l\lambda_{jl}^{(0)}\beta^{2^l} & \text{(cross-product term)} \label{eqn:mteq}
+\end{align}
+Notice that $\lambda$ corresponds to $\lambda$-Matrix entries. In Equation \ref{eqn:mteq} we assign $0$ to the 
+row index $i$, the proof can be extended to all row indices $i<n$ but fix $l$ to 0 because of the 
+conjugation generated by right-down cyclic-shift of $\lambda$-Matrix. Therefore we have 
+$$M_{i,j}^{(0)} = \lambda_{ij}^{(0)} = \lambda_{j,0}^{(i)} = T_{j-i,0-i}$$
+\end{Proof}

### 2016-12-09 dissertation/abstract.tex
dissertation v2:add more to appendix update 3

--- 2016-10-24 dissertation/abstract.tex
@@
 \begin{center}{\bf ABSTRACT}\end{center}
+Formal verification of hardware designs has become an essential
+component of the overall system design flow. The designs are generally
+modeled as finite state machines, on which property and equivalence
+checking problems are solved for verification. Reachability analysis
+forms the core of these techniques. However, increasing
+size and complexity of the circuits causes the state explosion
+problem. Abstraction is key to tackle the scalability challenge.
 
-Verification of sequential circuits remains to be a topic of all time. With increasing size of new
-integrated circuits, sequential circuit designers may face much more complicated problems on 
-logic errors, unexpected delays and failures on critical path. 
-This dissertation addresses the problem of sequential circuit verification at the word-level
-and is based on concepts derived from algebraic geometry.
+This dissertation presents new techniques for word-level abstraction
+with applications in sequential design verification. By bundling
+together $k$ bit-level state-variables into one word-level constraint 
+expression, the state-space is construed as solutions (variety) to
+a set of polynomial constraints (ideal), modeled over the finite
+(Galois) field of $2^k$ elements. Subsequently, techniques from
+algebraic geometry -- notable, Gr\"obner basis theory and technology
+-- are researched to perform reachability analysis and
+verification of sequential circuits. This approach adds a ``word-level
+dimension'' to state-space abstraction and verification to make the
+process more efficient.
 
-Analyzing the sequential circuits at word level is an efficient way
-of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
-state-space. We manage to model the verification properties and the gate-level sequential circuit 
-implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
-their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
-techniques from algebraic geometry can be used to reason about the state-space on high-level.
-In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
-polynomial ideals. Without actually solving the system, the states in state space can be investigated 
-by manipulating these ideals.
-
-We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
-analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
-verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
-abstraction-refinement techniques for verification, we propose to investigate a word-level 
-analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
-not only derive new algorithms and techniques, but will also consider efficient CAD 
-implementations to target sequential equivalence and model checking problems. 
-% \end{abstract}
+While algebraic geometry provides powerful abstraction and reasoning
+capabilities, the algorithms exhibit high computational complexity. In
+the dissertation, we show that by analyzing the constraints, it is
+possible to obtain more insights about the polynomial ideals, which can be
+exploited to overcome the complexity. Using our algorithm design and
+implementations, we demonstrate how to perform reachability
+analysis of finite-state machines purely at the word-level. Using this
+concept, we perform scalable verification of sequential arithmetic
+circuits. As contemporary approaches make use of resolution proofs and
+unsatisfiable cores for state-space abstraction, we introduce the
+algebraic geometry analog of unsatisfiable cores, and present
+algorithms to extract and refine unsatisfiable cores of polynomial
+ideals. Experiments are performed to demonstrate the efficacy of our
+approaches.

### 2016-12-09 dissertation/appendix.tex
dissertation v2:add more to appendix update 3

@@
 \label{append:ONB}
+The number of non-zero entries in $\lambda$-Matrix or multiplication table is known as \emph{Complexity $(C_N)$}.
+To define optimal normal basis, it is necessary to find the lower bound of $C_N$.
+\begin{Theorem}
+If N is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
+\end{Theorem}
+\begin{Proof}
+Let basis $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
+Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. 
+Denote $trace\ \beta$ by $b$, consider a $n\times n$ matrix $M^{(0)}$. Then
+$$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$$
+
+Therefore, the sum of all rows in $M^{(0)}$ is an n-tuple with $b$ as the first element and zeros elsewhere.
+So the first column always includes at least one nonzero element. 
+Since the sum of enries in each of other columns should equal to zero (modulo 2), they 
+need to include an even number of ``1"s.
+in order to maintain the linear independence among each row, they cannot be purely ``0"s. 
+Therefore, there are always at least two non-zero elements in each column.
+
+As a result, $C_N \geq 2(n-1)+1  = 2n-1$.
+\end{Proof}
+
+If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as \emph{Optimal Normal Basis}.
+
+\begin{Example}
+In $\F_{2^4}$ constructed with $p(\alpha) = \alpha^4 + \alpha + 1$, two sets of normal bases can be found:
+$\beta = \alpha^3, N = \{ \alpha^3, \alpha^6, \alpha^{12}, \alpha^9\}$ and 
+$\beta = \alpha^7, N = \{\alpha^7, \alpha^{14}, \alpha^{13}, \alpha^{11}\}$. Multiplication table for $\beta = \alpha^3$:
+\begin{equation}
+T_1 = \left(
+\begin{array}{lccr}
+0 & 1 & 0 & 0\\
+0 & 0 & 0 & 1\\
+1 & 1 & 1 & 1\\
+0 & 0 & 1 & 0
+\end{array} \right)
+\end{equation}
+Complexity $C_N$ = 7, this is optimal. For $\beta = \alpha^7$:
+\begin{equation}
+T_2 = \left(
+\begin{array}{lccr}
+0 & 1 & 0 & 0\\
+1 & 1 & 0 & 1\\
+1 & 0 & 1 & 0\\
+1 & 0 & 1 & 1
+\end{array} \right)
+\end{equation}
+Complexity $C_N$ = 9, this is NOT optimal.
+\end{Example}
+
+There should be an efficient method to construct optimal normal basis. The following  subsection refers to \cite{rosingbook}.
 \subsection{Construction of Optimal Normal Basis}

### 2016-12-10 dissertation/appendix.tex
dissertation v2:milestone

@@
 so it will finally reach termination condition $Ord_\theta(x) = x^n-1$.
-\subsection{Bases Conversion}
+\subsection{Bases Conversion and $\lambda$-Matrix Construction}
 % General \lambda matrix computation
+Given a normal basis (NB), we can easily convert it to a standard basis (StdB) by reducing $\alpha^t$
+with irreducible polynomial $p(\alpha)$ since 
+normal element is already given as $\beta = \alpha^t$. The reverse conversion requires 
+solving a system of equations. In Example \ref{ex:Rijndael}, the system of equations
+is as follows:
 
-\section{Optimal Normal Basis}
-\label{append:ONB}
-The number of non-zero entries in $\lambda$-Matrix or multiplication table is known as \emph{Complexity $(C_N)$}.
-To define optimal normal basis, it is necessary to find the lower bound of $C_N$.
-\begin{Theorem}
-If N is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
-\end{Theorem}
-\begin{Proof}
-Let basis $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
-Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. 
-Denote $trace\ \beta$ by $b$, consider a $n\times n$ matrix $M^{(0)}$. Then
-$$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$$
+\begin{equation}
+\begin{cases}
+\alpha = \alpha \\
+\alpha^2 = \alpha^2 \\
+\alpha^4 = \alpha^4 \\
+\alpha^8 = \alpha^7 + \alpha^6+\alpha^4+\alpha^2+\alpha+1\\
+\alpha^{16} = \alpha^7+\alpha^6+\alpha^5+\alpha^3+\alpha+1\\
+\alpha^{32} = \alpha^7+\alpha^4+\alpha^3+\alpha^2+\alpha\\
+\alpha^{64} = \alpha^7+\alpha^6+\alpha^3+\alpha^2\\
+\alpha^{128} = \alpha^6+\alpha^5+\alpha^4+\alpha^3+1
+\end{cases}
+\end{equation}
 
-Therefore, the sum of all rows in $M^{(0)}$ is an n-tuple with $b$ as the first element and zeros elsewhere.
-So the first column always includes at least one nonzero element. 
-Since the sum of enries in each of other columns should equal to zero (modulo 2), they 
-need to include an even number of ``1"s.
-in order to maintain the linear independence among each row, they cannot be purely ``0"s. 
-Therefore, there are always at least two non-zero elements in each column.
+We treat StdB $\{1,\alpha,\dots,\alpha^7$ as unknowns and NB $\{\alpha,\alpha^2,\dots,\alpha^{128}\}$
+as parameters. Then we can obtain a solution to this system denoting mapping from StdB to NB:
 
-As a result, $C_N \geq 2(n-1)+1  = 2n-1$.
-\end{Proof}
+\begin{equation}
+\begin{cases}
+1 = \alpha^{128}+\alpha^{64}+\alpha^{32}+\alpha^{16}+\alpha^8+\alpha^4+\alpha^2+\alpha \\
+\alpha = \alpha \\
+\alpha^2 = \alpha^2 \\
+\alpha^3 = \alpha^{128} + \alpha^{32}+\alpha^{16}+\alpha^2\\
+\alpha^4 = \alpha^4\\
+\alpha^5 = \alpha^{128} + \alpha^{32}+\alpha^{8}+\alpha^4\\
+\alpha^6 = \alpha^{64} + \alpha^{32}+\alpha^{4}+\alpha\\
+\alpha^7 = \alpha^{128} + \alpha^{16}+\alpha^{4}+\alpha
+\end{cases}
+\end{equation}
 
-If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as \emph{Optimal Normal Basis}.
+Directly constructing $\lambda$-Matrix is difficult for general NBs. 
+In Section \ref{sec:nbdesign}, we introduce the concept of multiplication table (M-table) and explained how its 
+entries are transformed from the $0$-th $\lambda$-Matrix. 
+Recall the M-table is defined as
+\begin{equation}
+\label{eqn:appmt}
+\beta
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
+\end{bmatrix}
+= {\bf T}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
+\end{bmatrix}
+\end{equation}
+Therefore we can compute M-table and then covert it to $\lambda$-Matrix.
+Using Equation \ref{eqn:appmt}, we derive entry $T_{i,j}$ in the M-table equals to
+the $j$-th bit of the NB representation of element $\beta\cdot\beta^{2^{i}}$.
+For example, the $i$-th row of the M-table for NB in Example \ref{ex:Rijndael} satisfies
+$$\beta\cdot\beta^{2^i} = \sum_{j=0}^{7} T_{i,j}\beta^{2^j}$$
+Thus, the M-table can be written as
+\begin{equation*}
+T = \left(\begin{array}{lccccccr}
+0 &1 &0 &0 &0 &0 &0 &1 \\
+0 &1 &0 &0 &1 &1 &0 &1 \\
+0 &0 &1 &0 &0 &1 &0 &1 \\
+0 &0 &0 &0 &0 &0 &0 &1 \\
+1 &1 &1 &1 &1 &1 &1 &1 \\
+0 &0 &0 &0 &1 &0 &0 &0 \\
+1 &1 &0 &1 &0 &1 &0 &0 \\
+1 &0 &0 &1 &1 &0 &1 &0
+\end{array}\right)
+\end{equation*}
+Use equation $M_{i,j}^{(0)} = T_{j-i,-i}$, we can obtain the
+$\lambda$-Matrix in Equation \ref{eqn:mat_R}.
 
-\begin{Example}
-In $\F_{2^4}$ constructed with $p(\alpha) = \alpha^4 + \alpha + 1$, two sets of normal bases can be found:
-$\beta = \alpha^3, N = \{ \alpha^3, \alpha^6, \alpha^{12}, \alpha^9\}$ and 
-$\beta = \alpha^7, N = \{\alpha^7, \alpha^{14}, \alpha^{13}, \alpha^{11}\}$. Multiplication table for $\beta = \alpha^3$:
-\begin{equation}
-T_1 = \left(
-\begin{array}{lccr}
-0 & 1 & 0 & 0\\
-0 & 0 & 0 & 1\\
-1 & 1 & 1 & 1\\
-0 & 0 & 1 & 0
-\end{array} \right)
-\end{equation}
-Complexity $C_N$ = 7, this is optimal. For $\beta = \alpha^7$:
-\begin{equation}
-T_2 = \left(
-\begin{array}{lccr}
-0 & 1 & 0 & 0\\
-1 & 1 & 0 & 1\\
-1 & 0 & 1 & 0\\
-1 & 0 & 1 & 1
-\end{array} \right)
-\end{equation}
-Complexity $C_N$ = 9, this is NOT optimal.
-\end{Example}
-
-There should be an efficient method to construct optimal normal basis. The following  subsection refers to \cite{rosingbook}.
-\subsection{Construction of Optimal Normal Basis}
-\subsection{Optimal Normal Basis Multiplier Design}
-% Customized Galois Field IEEE1363-2000
-% Multi table <=> \lambda matrix
-In Section \ref{sec:nbdesign}, we introduce the concept of multiplication table and explained how its 
-entries are transformed from the $0$-th $\lambda$-Matrix. Actually we can mathematically prove the relationship 
+Actually we can mathematically prove the relationship 
 between them:
 \begin{Theorem}
-Multiplication table $T$ is actually a conjugate of the $\lambda$-Matrix $M$. 
+M-table $T$ is actually a conjugate of the $\lambda$-Matrix $M$. 
 We can use an equation $M_{i,j}^{(0)} = T_{j-i,-i}$ to demonstrate the relation between their entries.
@@
 \begin{align}
-\beta\cdot\beta^{2^j} =& \sum_l\lambda_{0j}^{(l)}\beta^{2^l} & \text{(multi-table definition)} \nonumber\\
+\beta\cdot\beta^{2^j} =& \sum_l\lambda_{0j}^{(l)}\beta^{2^l} & \text{(M-table definition)} \nonumber\\
 					=& \sum_l\lambda_{jl}^{(0)}\beta^{2^l} & \text{(cross-product term)} \label{eqn:mteq}
@@
 \end{Proof}
+
+\section{Optimal Normal Basis}
+\label{append:ONB}
+The number of non-zero entries in $\lambda$-Matrix or multiplication table (M-table) is known as \emph{Complexity $(C_N)$}.
+To define optimal normal basis, it is necessary to find the lower bound of $C_N$.
+\begin{Theorem}
+If N is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
+\end{Theorem}
+\begin{Proof}
+Let basis $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
+Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. 
+Denote $trace\ \beta$ by $b$, consider a $n\times n$ matrix $M^{(0)}$. Then
+$$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$$
+
+Therefore, the sum of all rows in $M^{(0)}$ is an n-tuple with $b$ as the first element and zeros elsewhere.
+So the first column always includes at least one nonzero element. 
+Since the sum of enries in each of other columns should equal to zero (modulo 2), they 
+need to include an even number of ``1"s.
+Meanwhile, in order to maintain the linear independence among each row, they cannot be all ``0"s. 
+Therefore, there are always at least two non-zero elements in each column.
+
+As a result, $C_N \geq 2(n-1)+1  = 2n-1$.
+\end{Proof}
+
+If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as 
+\emph{Optimal Normal Basis} (ONB).
+
+\begin{Example}
+\label{ex:app4}
+In $\F_{2^4}$ constructed with $p(\alpha) = \alpha^4 + \alpha + 1$, two normal bases can be found:
+$\beta = \alpha^3, N_1 = \{ \alpha^3, \alpha^6, \alpha^{12}, \alpha^9\}$ and 
+$\beta = \alpha^7, N_2 = \{\alpha^7, \alpha^{14}, \alpha^{13}, \alpha^{11}\}$. 
+Their multiplication tables are listed below. For basis $N_1$:
+\begin{equation}
+T_1 = \left(
+\begin{array}{lccr}
+0 & 1 & 0 & 0\\
+0 & 0 & 0 & 1\\
+1 & 1 & 1 & 1\\
+0 & 0 & 1 & 0
+\end{array} \right)
+\end{equation}
+Complexity $C_N$ = 7, so $N_1$ is an ONB. For basis $N_2$:
+\begin{equation}
+T_2 = \left(
+\begin{array}{lccr}
+0 & 1 & 0 & 0\\
+1 & 1 & 0 & 1\\
+1 & 0 & 1 & 0\\
+1 & 0 & 1 & 1
+\end{array} \right)
+\end{equation}
+Complexity $C_N$ = 9, so basis $N_2$ is not optimal.
+\end{Example}
+
+ONBs are widely used in finite field circuit design, not only because its low complexity, 
+but also because of the convenience to construct the $\lambda$-Matrix and M-table for them. 
+To explain the convenience and the reason behind it, we introduce two types of ONBs and the method 
+to construct them.
+The following subsection refers to \cite{rosingbook}.
+\subsection{Construction of Optimal Normal Basis}
+{\bf Type-I ONB} over $F_{2^n}$ satisfies following criteria:
+\begin{itemize}
+\item $n+1$ must be prime.
+\item 2 must be primitive in $\mathbb{Z}_{n+1}$.
+\end{itemize}
+The second criterion indicates that powers of 2 (exponent from 0 to $n-1$) modulo $n+1$ must cover all integers
+from 1 to $n$.
+
+In following part we derive a simple way to construct the $\lambda$-Matrix of type-I ONB 
+from the criteria above. 
+Assume $\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ from k-th $\lambda$-Matrix. Then the crossproduct
+term can be written as
+\begin{equation}
+\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1} \lambda_{ij}^{(k)} \beta^{2^k}
+\end{equation}
+Suppose we only care about k=0. So simplified to following equations:
+
+$$
+\begin{cases}
+&\beta^{2^i}\beta^{2^j} = \beta\\
+&\beta^{2^i}\beta^{2^j} = 1 ~~~(\text{ if }~  2^i = 2^j\pmod{n+1})
+\end{cases}$$
+
+Solution $(i,j)$ implies location of entries that equal to ``1" in $\lambda$-Matrix. 
+Let $\beta$ be an optimal normal element, 
+$\{\beta^{2^i}~|~0\leq i<n\}$ cover all powers of $\beta$ and generates the basis. Thus by solving
+$$\begin{cases}
+2^i + 2^j = 1 \pmod{n+1}\\
+2^i + 2^j = 0 \pmod{n+1}
+\end{cases}$$
+we obtain $\lambda$-Matrix $M^{(0)}$. This construction method does not require information about 
+primitive polynomial nor normal element, so it is very convenient for circuit designers.
+Basis $N_1$ in Example \ref{ex:app4} is a type-I ONB over $\F_{2^4}$.
+
+{\bf Type-II ONB} over $\F_{2^n}$ satisfies following criteria:
+\begin{itemize}
+\item $2n+1$ must be prime. And either
+\item 2 is primitive in $\mathbb{Z}_{2n+1}$, or
+\item $2n+1 = 3 \pmod{4}$ and 2 generates the quadratic residues in $\mathbb{Z}_{2n+1}$
+\end{itemize}
+The last criterion means that that $2n+1$ should be congruent to $1$ or $3$ modulo 4. 
+To generate a Type-II ONB, we first pick an element $\gamma$ with order $2n+1$ in $\F_{2^{2n}}$, 
+such that the corresponding normal element $\beta$
+from $F_{2^n}$ can be written as $\beta = \gamma + \gamma^{-1}$. Thus the cross-product terms will be:
+\begin{equation}
+\begin{split}
+\beta^{2^i}\beta^{2^j} &= (\gamma^{2^i} + \gamma^{-2^i})(\gamma^{2^j} + \gamma^{-2^j})\\ &= 
+(\gamma^{2^i+2^j} + \gamma^{-(2^i+2^j)}) + (\gamma^{2^i-2^j} + \gamma^{-(2^i-2^j)})\\ &= 
+\begin{cases}
+\beta^{2^k} + \beta^{2^{k\prime}}~~~\text{if }~ 2^i \neq 2^j \pmod{2n+1}\\
+\beta^{2^k}~~~~~~~~~~~~~~\text{if }~ 2^i = 2^j \pmod{2n+1}
+\end{cases}
+\end{split}
+\end{equation}
+$k$ and $k\prime$ are the 2 possible solutions to multiplication of any 2 basis elements. 
+This guarantees the optimum of the basis since it has the minimum number of possible terms. 
+In the case of $2^i \neq 2^j \pmod{2n+1}$, at least one of following
+equations
+\begin{equation}
+\label{eqn:app1st}
+\begin{cases}
+2^i + 2^j = 2^k ~~~\pmod{2n+1}\\
+2^i + 2^j = -2^k \pmod{2n+1}
+\end{cases}
+\end{equation}
+has a solution, meanwhile at least one of following equations
+\begin{equation}
+\label{eqn:app2nd}
+\begin{cases}
+2^i - 2^j = 2^{k\prime} ~~~\pmod{2n+1}\\
+2^i - 2^j = -2^{k\prime} \pmod{2n+1}
+\end{cases}
+\end{equation}
+has a solution as well.
+
+In another case that $2^i = \pm 2^j \pmod{2n+1}$, at least one of the following 4 equations has a solution
+\begin{equation}
+\label{eqn:app3rd}
+\begin{cases}
+2^i + 2^j = 2^k ~~~\pmod{2n+1}\\
+2^i + 2^j = -2^k \pmod{2n+1}\\
+2^i - 2^j = 2^k ~~~\pmod{2n+1}\\
+2^i - 2^j = -2^k \pmod{2n+1}
+\end{cases}
+\end{equation}
+In set of Equations \ref{eqn:app1st} and \ref{eqn:app2nd}, there are two possible solutions in total. 
+In set of Equations \ref{eqn:app3rd},
+there is only one possible solution. Since these equations are all similar, instead of 
+working with two different sets we can combine them together and solve a system of 4 equations. 
+As a result, to construct 
+the $\lambda$-Matrix $M^{(0)}$, we set $k=0$ and find solutions to:
+\begin{equation}
+\label{eqn:app4th}
+\begin{cases}
+2^i + 2^j = 1 ~~~\pmod{2n+1}\\
+2^i + 2^j = -1 \pmod{2n+1}\\
+2^i - 2^j = 1 ~~~\pmod{2n+1}\\
+2^i - 2^j = -1 \pmod{2n+1}
+\end{cases}
+\end{equation}
+
+\begin{Example} 
+By solving the system of Equations \ref{eqn:app4th} with $n=5$, we find 9 pairs of
+indices $(i,j)$ such that $0\leq i,j <5$. Assign ``1" to corresponding entries in a $5\times5$ matrix,
+the result is $M^{(0)}$ for type-II ONB over $\F_{2^5}$:
+\begin{equation}
+M^{(0)} = \left(
+\begin{array}{lcccr}
+0 & 1 & 0 & 0 & 0\\
+1 & 0 & 0 & 1 & 0\\
+0 & 0 & 0 & 1 & 1\\
+0 & 1 & 1 & 0 & 0\\
+0 & 0 & 1 & 0 & 1
+\end{array} \right)
+\end{equation}
+\end{Example}
+
+\subsection{Optimal Normal Basis Multiplier Design}
+% Customized Galois Field IEEE1363-2000
+Designers can easily generate the $\lambda$-Matrix for GF multipliers, which is sufficient to derive the 
+structure of the circuit. However, our GB based approach requires specifying the exact normal element $\beta$,
+{\it i.e. } it is necessary to obtain $t$ as $\beta = \alpha^t$ before executing our technique.
+L\"uneburg's algorithm and Lenstra's algorithm do not guarantee the output normal element is optimal 
+normal element, and usually result in high computation cost. Actually if the type of ONB within the design 
+is known, instead of looking up the optimal normal element, we can construct a special irreducible $p(\alpha)$
+such that optimal normal element equals to the primitive element: $\beta=\alpha$. The concepts and algorithms 
+refer to IEEE standard 1363-2000 \cite{IEEE1363}.
+
+For type-I ONB over $\F_{2^k}$, the irreducible polynomial $p(\alpha) = \alpha^k+\alpha^{k-1}+\cdots+\alpha+1
+= \sum_{i=0}^k \alpha^i$. For type-II ONB, the following iterative algorithm is required to generate desired 
+irreducible polynomial:
+
+\begin{algorithm}[H]
+\SetAlgoNoLine
+\LinesNumbered
+ \KwIn{Field $\Fkk$ which contains a type-II ONB}
+ \KwOut{The irreducible polynomial $p(\alpha)$ for the ONB}
+
+  $f(\alpha)\gets 1$\;
+  $p(\alpha)\gets \alpha+1$\;
+  \For{$i=1\dots k-1$}
+  {
+  	$g(\alpha)\gets f(\alpha)$\;
+	$f(\alpha)\gets p(\alpha)$\;
+	$p(\alpha)\gets \alpha f(\alpha)+g(\alpha)$\;
+  }\
+\Return{$p(\alpha)$}
+\caption {Generating irreducible polynomial for type-II ONB over $\Fkk$}
+\label{alg:ieee1363}
+\end{algorithm}
+\DecMargin{1em}
+\vspace{0.2in}

### 2016-12-10 dissertation/conclude.tex
dissertation v3:update 1

--- 2016-10-24 dissertation/conclude.tex
@@
 \label{ch:conclude}
+This dissertation presents a new approach to perform reachability
+analysis of FSMs at the word-level. It is facilitated by investigating the analog of 
+implicit state enumeration algorithm in computer algebra and algebraic
+geometry domains. The image function part of the computation is mapped to a variety projection on 
+next state variables. This projection is implemented by computing the 
+Gr\"obner basis of an elimination ideal with abstraction term orders.
+Moreover, the set operations in the state space are mapped to 
+the arithmetic of ideals, as algebraic geometry provides a 
+way to reason about the variety (solutions) by manipulating the ideals.
+A special term order (RATO) is utilized to improve the Gr\"obner bases 
+computation, and a tool is developed to implement our 
+word-level FSM traversal algorithm. Experiments are performed to 
+analyze the reachability for ISCAS'89 and ITC'99 circuit benchmarks.
+
+Next, we describe a method to execute functional verification on sequential Galois field
+multipliers over $\Fkk$.  The core algorithm is based on word-level unrolling of 
+a Moore machine, which applies concepts from the word-level FSM traversal 
+algorithm. As a result, transition relations are represented 
+as a polynomial with word-level inputs/outputs.
+We implement our algorithm with both the \textsc{Singular} platform and a custom C++ toolset 
+and perform experiments on two classes of circuits.
+Our approach is able to verify up to 162-bit sequential
+circuits, whereas contemporary techniques fail beyond 23-bit
+datapaths.
+
+At last we explore abstraction UNSAT cores in algebraic geometry -- the foundation for refinement techniques
+to boost sequential circuit verification.
+% UNSAT core extraction, as an indispensable component of many refinement algorithms 
+% which rely on UNSAT informations, is thus proposed in polynomial rings.
+We use the Weak Nullstellensatz as the essential theory of UNSAT core extraction, 
+then develop heuristics to improve the core that exploit the structure of the refutation proof.
+An algorithm was implemented within the \textsc{Singular} computer algebra tool
+and experiments were conducted to validate the approach.
+
+Our approaches still have limitations: for word-level FSM traversal and 
+sequential GF multiplier verification, our methods are more efficient
+for XOR-rich circuits, while most industrial designs are AND-OR gate dominant.
+For UNSAT core extraction, the abstraction refinement approach such as $k$-BMC 
+has only limited application on certain model checking problems. To overcome these limitations,
+the following further explorations are worthy of investigation.
 
 \section{Future Work}
-\subsection{Multivariate polynomial ideal based FSM Traversal}
-\subsection{New Diagram Structure accelerating Polynomial Reduction}
-\subsection{Interpolation extraction using GB Algorithm}
-\subsection{Verification of Integer Arithmetic Circuits}
+In this section, we highlight some research problems that deserve further study.
+\subsection{Multivariate polynomial ideal based FSM traversal}
+In Chapter \ref{ch:reacha}, we always use a single word-level variable $T$ to denote the
+next state. However, in some situations we need to keep recording the relations between $T$ and
+inputs, {\it i.e.} the reached states contain multivariate polynomials in $T,S$ and PI $x$. 
+While the elimination and set union/intersection are compatible with multiple NS variables,
+the set complement requires an extension of Theorem \ref{thm:quotient} on ideals with 
+multivariate polynomial generators. We conjecture as follows:
+
+\begin{Conjecture}
+Assume we are given ring variables $x_i$, and an ideal $J$ composed by $s$ generators:
+$J = \langle f_1,\dots, f_s\rangle$.  Additionally $J_0$ is the vanishing polynomial for variables
+$x_i$: $ x_i^{q} - x_i$ where $q$ is the size of signal $x_i$ represents.
+We conjecture that
+$$V(J') = V(J_0:J) = V(J_0)\setminus V(J) = \overline{V(J)}$$
+% And we believe this conjecture is scalable such that it is also valid for even more variables and generators like 
+% $J = \langle f_z, v_x,v_y,v_u,v_w\rangle$, $f_z =\mathcal F (z,x,y,u,w)$.
+\end{Conjecture}
+
+In univariate case, $\F[x]$ is the principle ideal domain thus Theorem \ref{thm:quotient} can be proved.
+However over multivariate case, the proof of this conjecture is not available.
+% 
+% The following example illustrates a different operation from Example \ref{ex:SMPO}.
+% \begin{Example}
+% In Figure \ref{fig:fsm}
+% $\{s_0, s_1\}$ are state/pseudo inputs, $\{t_0,t_1\}$ are state/pseudo outputs, and there is a primary input (1-bit) 
+% $x$. We propose a new algorithm by modifying Algorithm \ref{alg:univa}:
+% 
+% \IncMargin{1em}
+% \begin{algorithm}[hbt]
+% \SetAlgoNoLine
+% \LinesNumbered
+%  \KwIn{Transition polynomial $f_t = T + \mathcal F (S,x)$, 
+% 	initial state ideal $from^0 = \langle S+\mathcal G(x), x^{q_1} - x\rangle$}
+%  \KwOut{Reachable states}
+%   $reached = from^0(T\setminus S)$\;
+%   \Repeat{$GB(new^i) == 1$}
+%   {
+%   	$i \gets i + 1$\;
+% 	$to^i \gets$GB$(\langle f_t, from^{i-1}\rangle) \setminus \mathcal H(S)$\;
+% 	$\overline{reached} = \langle T^{q_2}-T, x^{q_1} - x \rangle : reached$\;
+% 	$new^i \gets $GB$(to^i + \overline{reached})$\;
+%   	$reached \gets $GB$( reached \cdot new^i)$\;
+% 	$from^i \gets new^i(S\setminus T)$\;
+%   }
+% \Return{$reached$}
+% \caption {Algebraic Geometry based Traversal Algorithm (multivariate-generator ideals)}\label{alg:multi}
+% \end{algorithm}
+% \DecMargin{1em}
+% 
+% The inputs of this algorithm includes transition polynomial (result of word-level abstraction) and initial 
+% states description ideal, which contains 2 generators corresponds to constraints of PI and combinational input $S$. 
+% For example, $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
+% $\langle S+x\cdot\alpha, x^2+x\rangle$ means initial states$=\{00,10\}$).
+% 
+% Transition polynomial calculation uses the abstraction and bit-to-level conversion method from 
+% Chapter \ref{ch:reacha}. After Constructing
+% an elimination ideal, we impose RATO such that \emph{reverse\ topo\ order\ ckt\ variables }$> T > S > x$, the reduction
+% remainder has the form $T+\mathcal F(s_0,s_1,x)$. From word definition $S+s_0+s_1\alpha$ we get
+% $$s_0 = \alpha S^2+ (1+\alpha)S, s_1 = S^2+S$$
+% Do substitution, the transition polynomial of example ckt is 
+% $$f_T = T+S^3\cdot x+\alpha S^3+(1+\alpha)S^2\cdot x+S^2+S\cdot x+(1+\alpha)x+1$$
+% 
+% Assume we start from state $\{11\}$. In the first iteration, the reached state is $\{01\}$. Line 4 is to compose an
+% ideal with 2 generators from $from^0$ and transition polynomial $f_T$, compute its Gr\"obner basis. Note this ideal
+% has the form
+% \begin{equation}
+% I_{tran} = \left\{
+%              \begin{array}{c}
+%              T+\mathcal F(S,x) \\
+%              S + \mathcal G(x) \\
+%              v_x
+%              \end{array}  
+%         \right.
+% \end {equation}
+% $v_x$ is a polynomial containing only $q$-bit PI $x$, initially it should be vanishing polynomial $x^{q_1}-x$, 
+% with the program executing it may be factorized.
+% 
+% Consider Buchberger's algorithm, all generators' leading terms are relatively prime, so it is a GB itself. 
+% Then we need to reduce this GB, we will find out $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
+% and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. So, at last we will get a polynomial
+% $T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
+% $S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. 
+% In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$
+% 
+% Line 5 is the ideal quotient of universal set and reached states. In the first iteration, 
+% $reached$ is the initial state $\langle T+1+\alpha, x^2+x \rangle$. Result of ideal quotient
+% is $\langle T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle$ represents $\{00,01,10\}$.
+% 
+% Line 6 is ideals' sum (intersection of their varieties), it is done by combining all generators
+% from 2 ideals and compute GB. For first iteration result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.
+% 
+% Line 7 is ideals' product (union of their varieties), it is done by multiplying all pairs of
+% generators from both ideal. For first iteration result is $GB(\langle (T+1)(T+1+\alpha),
+% (T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$ representing $\{01,11\}$.
+% 
+% The traversal will run 3 iterations and terminate at 4th iteration. I list all intermediate 
+% results below:
+% \begin{itemize}
+% \item Iteration 1: $from^0 = \langle S+1+\alpha, x^2+x\rangle, to^1 = \langle T+1, x^2+x\rangle,
+%  reached = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$
+% \item Iteration 2: $from^1 = \langle S+1, x^2+x\rangle, to^2= \langle T+\alpha, x^2+x\rangle,
+% reached = \langle T^3+1, x^2+x\rangle$
+% \item Iteration 3: $from^2 = \langle S+\alpha, x^2+x\rangle, to^3 = \langle T+\alpha x, x^2+x
+% \rangle, reached = \langle T^3\cdot x+x, T^4+T, x^2+x\rangle$
+% \item Iteration 4: $from^3 = \langle S, x\rangle, to^4 = \langle T+1, x\rangle, new = \langle1\rangle$
+% \end{itemize}
+% The final reachable states are represented by multivariate polynomial ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
+% which denotes $\{00,01,10,11\}$.
+% \end{Example}
+
+\subsection{Use of the $F_4$ algorithm and ZDDs to Accelerate GB Reduction}
+In Chapters \ref{ch:reacha} and \ref{ch:normal}, by using RATO we transform the GB-based
+computation to that of a multivariate polynomial division. However, this division (reduction)
+still incurs exponential complexity in the worst case. In a situation where division is to be performed modulo
+a chain of OR gates, the size of remainder polynomial will explode. For example, a chain of OR gates 
+can be written as the Boolean function 
+\begin{equation}
+\label{eqn:chainOR}
+f = ((a\lor b) \lor c) \lor d
+\end{equation}
+which equals to the following polynomial function in $\F_2$:
+$$f = abcd+abc+abd+ab+acd+ac+ad+a+bcd+bc+bd+b+cd+c+d$$
+Notice, $f$ contains $2^4-1 = 15$ terms, which is exponential in the number of variables.
+The size explosion is a major factor affecting the efficiency of polynomial division.
+
+One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
+Analysis of experimental results shows that multivariate polynomial division procedure
+consumes most of the verification time.
+A matrix-based technique named as "$F_4$ style reduction" \cite{f4} can speed up the procedure of
+dividing a low-degree polynomial with a term-sparse polynomial ideal.
+
+Another way is to utilize DDs, {\it e.g.} ZDDs. ZDDs can represent unate covers of Boolean formulas.
+We can represent $f$ using a ZDD, where every path in the ZDD from a root to a terminal corresponds
+to a monomial term.
+
+\begin{figure}[h]
+\centerline{
+\includegraphics[width=0.35\textwidth]{newfig/ZDD.eps}
+}
+\caption{A ZDD representing remainder polynomial reducing by a chain of OR gates with order $d>c>b>a$}
+\label{fig:ZDD}
+\end{figure}
+
+Figure \ref{fig:ZDD} shows a ZDD representing 
+Equation \ref{eqn:chainOR} with size $2\times4 -1 = 7$, which is linear in the number of variables.
+The reduction process using ZDDs can be executed as in \cite{polybori:2009}. The graphical illustration 
+of the remainder in ZDDs is also shown in Figure \ref{fig:ZDD}.
+
+\subsection{Craig Interpolants in Algebraic Geometry}
+The concept of Craig Interpolants (CI) and their existence comes from
+symbolic logic \cite{craig-interpolate}; later, algorithms were
+presented to find the CI for Boolean formulae
+\cite{pudlak:ci} \cite{mcmillan2003interpolation}. Assume that Boolean
+formulae are represented in Clause Normal Form (CNF) as 
+$f = C_1 \wedge C_2 \wedge \dots \wedge C_m$ where: 1) Each clause
+$C_i$ is a disjunction (Boolean OR, denoted $\vee$) of literals; 2)
+Each literal is a Boolean variable $x_i$ or its complement
+$\overline{x_i}$. 
+The Boolean satisfiability (SAT) problem requires that we find an
+assignment to the variables such that the formula $f$ is satisfied
+(SAT), or otherwise prove that no such assignment exists (UNSAT). 
+A CI is related to an UNSAT formula.
+
+\begin{Definition}
+(From \cite{mcmillan2003interpolation}) Let $A$ and $B$ be Boolean
+  formulae given as sets of clauses such that $A \wedge B$ is
+  unsatisfiable (UNSAT). Then there exists a formula $P$ such that: 
+1) $A$ implies $P$ (or $A \subseteq P$); ~2) $P \wedge B$ is UNSAT; ~3)
+$P$ refers to only the common variables of $A$ and $B$. The formula
+$P$ represents an {\bf interpolant} of $A$ and $B$.
+
+Given  the pair $(A, B)$ and their refutation proof, a procedure
+called interpolation system constructs an interpolant in linear time
+and space in the size of the proof \cite{mcmillan2003interpolation}
+\cite{pudlak:ci}. 
+\end{Definition}
+
+\begin{Example}\label{ex1}
+Let $f = (\overline{d})(\overline{c})(\overline{a}\vee d)(a
+\vee b \vee c)(\overline{b})$ be a CNF formula. Let $f = A \wedge B =
+\emptyset$, where $A = (\overline{d})(\overline{c})(\overline{a}\vee
+d)$ and  $B = (a \vee b \vee c)(\overline{b})$. Then $P = \overline{a}
+\wedge \overline{c}$ is an interpolant of $(A, B)$. 
+\end{Example}
+
+CIs are used to derive abstractions to produce over-approximate image
+operators in model checking \cite{mcmillan2003interpolation}. Since $A
+\implies P$, $P$ contains $A$ and is an {\it abstraction} of $A$. It
+also has fewer variables, so checking  invariants on $P \wedge B$ is
+easier. The interpolant is derived through a resolution proof of the
+SAT problem. There can be many interpolants $P_i$ for a pair $(A,B)$; 
+however, it is not feasible to explore a few or all of these
+interpolants by means of the resolution proof.
+
+We introduce the algebraic geometry analog of CI.
+We conjecture that the
+  concept of CI should be related to elimination ideals, so future lines
+  of investigation should focus on \Grobner basis computations with
+  elimination term orders for their computation.
+
+\begin{Definition}\label{ci}
+Let $F = \{f_1, \dots, f_s\}$ be a set of polynomials in the ring 
+$R = \Fq[x_1,\dots,x_n]$. Let $F = F_A \cup F_B$ and ideals $J =
+\langle F \rangle, J_A = \langle F_A \rangle, J_B = \langle F_B
+\rangle$ be corresponding ideals in $R$ such that $J = J_A + J_B$. Let
+it be known (say, due to application of Weak Nullstellensatz and
+Gr\"obner basis) that the varieties $V_{\Fq}(J) = V_{\Fq}(J_A) \cap
+V_{\Fq}(J_B) =V_{\Fq}(J_A+J_B)=\emptyset$. Also, let the set of
+variables $X = \{x_1,\dots,x_n\} = X_A \cup X_c \cup X_B$ where $X_A,
+X_B$ are the set of variables present exclusively in the sets of
+polynomials $F_A, F_B$ respectively. Only $X_c$ is the set of
+variables that are common to both sets of polynomials $F_A,
+F_B$. Then, there exists a set of polynomials  $F_P$ and ideal $J_P =
+\langle F_P \rangle$ such that 
+\bi
+\item $V_{\Fq}(J_A) \subseteq V_{\Fq}(J_P)$
+\item $V_{\Fq}(J_P) \cap V_{\Fq}(J_B) = \emptyset$
+\item Polynomials of $F_P$ contain the common variables ($X_C$) of
+  $F_A, F_B$. 
+\ei
+
+We call the ideal $J_P = \langle F_P\rangle$ the {\bf algebraic
+  interpolant} of  $J_A+J_B$. 
+\end{Definition}
+
+Existence of algebraic interpolant comes from \cite{craig-interpolate}.
+The question for exploring is that how do we compute it in algebraic geometry?
+
+\begin{Example} \label{ex2}
+Based on Example \ref{ex1}, we translated the system over
+$\mathbb{F}_2[a, b, c, d]$. Let $F_A = \{f_1, f_2, f_3\}$ and $F_B =
+\{f_4, f_5\}$ where: $f_1: d; ~~~f_2: c; ~~~f_3: a + da; f_4:  abc +
+ab + ac + bc + a + b + c + 1; ~~~f_5: b.$ The Boolean interpolant
+$\overline{a}\wedge\overline{c}$ from Example \ref{ex1} translates to
+${\mathbb{F}}_2$ as the polynomial $f_p = ac + a + c$, with its
+variety $V(f_p) = \{a=0, c=0\}$.   
+\end{Example}
+
+% {\bf Research problem 2:}
+% {\it How can this algebraic interpolant $f_p$ (or in the general case,
+%   the set of polynomials $F_P$) be computed? As the interpolant $F_P$ 
+% contains only variables $X_C$ that are common to the polynomial sets
+% $F_A, F_B$, does this imply that the interpolants can be computed by
+% means of \Grobner bases with an elimination term order $X_A>X_B>X_C$
+% with $X_A, X_B$ eliminated from the problem? }
+
+Algebraic interpolation is strongly related
+to the GB computation with the elimination order $X_A>X_B>X_C$, and
+this relationship needs to be formally derived.
+
+\begin{Conjecture} \label{con:ci}
+Computations of algebraic interpolants: Let $J_0$ denote the ideal of
+all vanishing polynomials in $\Fkk[x_1,\dots,x_n]$.
+\bi
+\item  Compute a Gr\"obner basis $G_1 = GB(J_A + J_0)$ 
+with the elimination order $X_A > X_B > X_C$, and select
+$F_{P1}=G_1\cap\Fkk[X_C]$. We conjecture that the \Grobner
+basis $F_{P1}$ of the elimination ideal corresponds to an algebraic
+interpolant.
+
+\item Analogously, compute a Gr\"obner basis $G_2 = GB(J_B + J_0)$ 
+with the elimination order $X_B > X_A > X_C$, and select
+$F_{P2}=G_2\cap\Fkk[X_C]$. Find an ideal $F'_{P2}$ such that
+the variety $V(F'_{P2})$ is the complement of the variety
+$V(F_{P2})$. Then the set $F'_{P2}$ gives another interpolant.
+\ei
+\end{Conjecture}
+
+\begin{Example}\label{ex3}
+Consider the polynomials $\{f_1, \dots,f_5\}$ from Example
+\ref{ex2}. Computing $F_{P1}$ as described in Conjecture \ref{con:ci} 
+produces $F_{P1} = \{a, c\}$, correctly giving us the desired variety
+$V(a = 0, c = 0)$. Similarly, when we compute $F_{P2}$, we find that
+the interpolant is $ac + a + c + 1$. Notice that the variety 
+$V(ac+a+c+1)=\{(0,1), (1,0), (1,1)\}$, which is exactly the complement
+of $V(F_{P1})$.  
+\end{Example}
+
+The aforementioned experiments in Example \ref{ex3} do not invalidate
+our conjectures. Moreover, the above experiment shows that there can be
+multiple ways of computing the interpolants. What can these \Grobner
+basis computations tell us about the number of valid algebraic
+interpolants for any given problem? Can they be classified as {\it weak
+  or strong interpolants} based on the sparsity of the polynomials
+(power of abstraction)?
+
+As there can be many interpolants for an ideal-pair $(J_A, J_B)$, 
+the following questions should also be investigated in the future:
+\bi
+\item Find the minimal interpolant: i.e. find the interpolant $F_P$
+  such that $V(J_P)$ is the smallest variety larger than $V(J_A)$ such
+  that $V(J_P) \cap V(J_B) = \emptyset$.
+\item Analogously, find the maximal interpolant. 
+\item Over finite fields, the variety of an elimination ideal is
+  exactly equal to the projection of the variety on the un-eliminated
+  variables. Consider Fig. \ref{fig:projection}, where variety of the
+  ideals $J_A, J_B$ are respectively projected on the common variables
+  $X_C = \{a, c\}$. Then, does computing $F_{P1}$ (resp. $F_{P2}$) 
+  deliver the minimal (resp. maximal) interpolant?
+\ei
+
+\begin{figure}[h]
+\centering{
+\includegraphics[width=4.5in]{newfig/project.eps}
+\caption{Algebraic interpolant: Projection of varieties on common variables? }
+\label{fig:projection}}
+\end{figure}
+
+% 
+% Once these problems are understood and algorithms for the computation
+% of algebraic interpolants derived, then we will integrate them 
+% with word-level reachability analysis for abstraction-based model
+% checking.
+
+% \subsection{Technology Mapping for Word-level Functional Blocks}
+% Technology mapping is an important problem in digital circuit synthesis.
+% Designers are given a library of well-designed functional/macro blocks (including IP cores) and a raw netlist, 
+% technology mapping's objective is to map as many as blocks to the raw netlist and 
+% keep the functional equivalence. Contemporary techniques rely on bit-wise 
+% analysis on the signals to deduce the boundary of mapped blocks.
+% It is possible to use the equivalence checking techniques proposed in this dissertation 
+% as a alternative way to perform technology mapping, especially on word-level when 
+% given blocks represent word-level functions.
+% 
+% \begin{Problem}
+% The objective of our approach is to map the macro blocks without boundary information.
+% Mapping is an essential technique used in synthesis and verification. In synthesis, we can map the 
+% macro functions with smaller and faster implementations to optimize the timing and area; in simulation,
+% we can map a complicated function to a simple execution to accelerate simulation speed.
+% 
+% Given a gate-level design $D$ and several word-level macro blocks $\{B_i\}$, we need to map macro
+% blocks $B_i$ into design $D$, and write out the mapped design $D'$ which is equivalent to original
+% design $D$. The objective is to generate a mapped design $D$ with as many of $B_i$ such that the area
+% and timing is optimal. The procedure is also illustrated in Figure \ref{fig:macro}.
+% 
+% \begin{figure}[h]
+% 	\begin{center}
+% 	\includegraphics[width=\textwidth]{newfig/macro.pdf}
+% 	\end{center}
+% 	\caption{The outline and flow of technology mapping of macro blocks}
+% 	\label{fig:macro}
+% \end{figure}
+% 
+% \end{Problem}
+% 
+% The following part describes the sketch of our proposed approach base on loop invariant constraints \cite{sankaranarayanan2004non}.
+% First, a transition system is modeled by algebraic assertions; then \emph{ideal membership test} \cite{lv:phd}
+% is applied on the set of assertions to help abstract the loop invariant. We introduce concept \emph{template}
+% from \cite{sankaranarayanan2004non} to apply on our proposed approach.
+% 
+% 1) {\bf Template and state constraints:}  Ideal membership test technique requires a setup of ideal $J_{loop}$
+% for loop invariants. 
+% One heuristic can provide better coverage for loop invariant abstraction as well as
+% relatively small size is \emph{generic quadratic form} (GQF).
+% 
+% For example, the GQF of a pair of state variables $\{x,y\}$ is
+% $$\Func = a_0x^2+a_1x+a_2xy+a_3y+a_4y^2+a_5$$
+% It covers all possible terms with degree no more than 2. $a_0,a_1,\dots, a_5$ are usually real number parameters,
+% some of their assignments can turn $\Func$ into desired invariant. This parameterized constrain
+% polynomial covers all combinations of state variables can also be called a \emph{template}.
+% Subsequently, finding a proper assignment is the essential part of in our proposed research.
+% 
+% 2) {\bf Initial state constraints:} For initial states, the constrains are explicit. A template is adopted and refined by Gr\"obner basis 
+% generated by original constrains, by equaling the remainder to 0 we can get constrains on
+% parameters from the template.
+% 
+% An example is shown in the following 3-line algorithm multiplying two natural numbers. 
+% \begin{align}
+% \label{eqn:loopinv}
+% & {\bf integer}~~i,j,s~~{\bf where}(s=0 \land j=j_0) \nonumber\\
+% & l_0: {\bf while}~~(\cdots)~~{\bf do} \nonumber\\
+% & ~~~~ l_1:~(s,j)\gets(s+i,j-1) 
+% \end{align}
+% 
+% Its template is generic quadratic form of $\{s,i,j,j_0\}$, which is
+% \begin{align}
+% \Func = & a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
+% &a_5i^2+a_6i+a_7ij+a_8ij_0+a_9j^2\nonumber\\
+% &a_{10}j+a_{11}jj_0+a_{12}j_0^2+a_{13}j_0+a_{14}\nonumber
+% \end{align}
+% with parameters $a_0,\dots,a_{14}$.
+% 
+% Constrains of initial state $s=0\land j=j_0$ can be interpreted to polynomials:
+% $$\{s, j-j_0\}$$
+% Since their leading terms are relatively prime, the polynomial set itself constitutes a Gr\"obner basis 
+% $$G=\{s, j-j_0\}$$ 
+% Therefore its ideal can be written as
+% $J=\langle s,j-j_0\rangle$.
+% 
+% In the next step, we reduce the template with the Gr\"obner basis $G$: $\Func \xrightarrow{G}_+ r$, the remainder equals to
+% $$r = a_5i^2+a_6i+(a_7+a_8)ij_0+(a_9+a_{11}+a_{12})j_0^2+(a_{10}+a_{13})j_0+a_{14}$$
+% Let it equal to 0, then each coefficient will generate a constraint. The solution to the system forms the candidate
+% assignment to generate loop invariants.
+% \begin{equation}
+% \left\{
+% \begin{array}{l}
+% a_5=a_6=a_{14}=0\\
+% a_7+a_8=0\\
+% a_9+a_{11}+a_{12}=0\\
+% a_{10}+a_{13}=0
+% \end{array}\right.
+% \nonumber
+% \end{equation}
+% 
+% 3) {\bf Modeling state transitions:}
+% A typical state transition starts from the previous state (PS) and ends at next state (NS). 
+% Our proposed approach models the 2 states individually,
+% {i.e.} performs polynomial reduction separately and obtains remainder $r_1$ and $r_2$. 
+% Assume a constraint polynomial describing
+% the state transition is $r_t$, then we require that when invariant of PS holds and transition $PS \to NS$ stands,
+% invariant of NS should also be satisfied:
+% $$(r_1 = 0)\land (r_t = 0) \implies (r_2 = 0)$$
+% One reasonable conjecture is
+% $$r_t = r_1 - \lambda r_2$$
+% Theoretically $\lambda$ can be any polynomial in arbitrary rings. To make it practical solving the system, 
+% we limit $\lambda$ to the polynomial ring with only real numbers.
+% 
+% We take Equation \ref{eqn:loopinv} as the example (which refers to Example 10 in \cite{sankaranarayanan2004non}). 
+% PS is initial state we just characterized 
+% $$\Func = f(s,i,j,j_0)$$ 
+% and NS has exactly the same form of constraint:
+% $$\Func' = f(s',i',j',j_0')$$
+% Considering the transition relation, we substitute $s'$ with $s+i$ and replace $j'$
+% with $j-1$, $i'$ for $i$, $j_0'$ for $j_0$, respectively. 
+% The template for NS is polynomial $f'$ in Example 10 in \cite{sankaranarayanan2004non}.
+% 
+% We then perform reduction with the Gr\"obner basis generated from the transition relation modeling, 
+% and record its remainder $r_2$.
+% From equality $r_2 = \lambda r_1$, we get constraints for the parameters. Solve the system using 
+% similar algorithm to solve binate covering problems
+% which is also described in Section 4 as \emph{elimination by splitting} technique in \cite{sankaranarayanan2004non}. 
+% One branching result is:
+% \begin{align*}
+% &a_0,a_2,\dots, a_6, a_9,\dots,a_{14}=  0\\ 
+% &a_1=a_7=-a_8
+% \end{align*}
+% The reduced remainder is: 
+% $$r_2 = a_1s +(a_1-a_7)i+a_7ij+a_8ij_0 = a_1(s+ij-ij_0) = 0$$
+% Thus the invariant of the program in Equation \ref{eqn:loopinv} is:
+% $$s = i(j_0- j)$$
+% We can verify the invariant by executing the program which calculates $i\times j_0$. 
+% Initially $s=0$, $j_0-j=0$, invariant holds;
+% during each cycle, $s'=s+i, (j_0'-j') = (j_0-j)+1$, the invariant also holds. In conclusion, this is a
+% loop invariant for program in Equation \ref{eqn:loopinv}.
+% 
+% {\bf Our proposed approach on technology mapping:}
+% Our approach borrows inspiration of templates in \cite{sankaranarayanan2004non}. However, we make an improvement
+% on the original approach:
+% instead of using template polynomial to describe the system, we add some templates into the ideal
+% of macro blocks which serve as technology mapping candidates. In this way we can cover all possibilities
+% of boundary cutting (circuit partition).
+% 
+% 1) {\bf System abstraction:}
+% The polynomial we use to test ideal membership should include all information of a circuit partition,
+% which requires us to abstract information from the system and write it into a single polynomial.
+% Usually this polynomial has the following generic form:
+% $$Z + f(i_1,i_2,\dots,i_n)$$
+% where $Z$ is the output. When there is only one output, $Z$ collapse to a bit-level variable. However in most
+% cases there are multiple outputs on the cut, indicating $Z$ as a word-level variable. Boolean function $\Func$ covers
+% all inputs, and $i_1,i_2,\dots,i_n$ are all bit-level inputs.
+% 
+% \begin{figure}[h]
+% 	\begin{center}
+% 	\includegraphics[width=\textwidth]{newfig/tobemapped.pdf}
+% 	\end{center}
+% 	\caption{An example gate-level netlist to be mapped}
+% 	\label{fig:tobemapped}
+% \end{figure}
+% 
+% Figure \ref{fig:tobemapped} shows an example circuit partition with $a_1,b_1,a_0,b_0$ as inputs and
+% $Z = \{z_0,z_1\}$ as outputs. If we use elements from Galois field $\F_{2^2}$ to represent word $Z$,
+% we have $Z = z_0 + \alpha\cdot z_1$.
+% 
+% After imposing ATO of LEX with
+% $$Other\ circuit\ variables > output\ word\ Z > all\ bit\ level\ inputs$$
+% on the ideal describing the partitioned circuit as well as ideal with vanishing polynomials,
+% the reduced Gr\"obner basis has a single polynomial generator in the form of $Z + \Func(a_1,b_1,a_0,b_0)$.
+% In this example:
+% 
+% \begin{itemize}
+% \item Term ordering: $\{z_0,z_1,t,g_1,p_1,ci,t_1,t_2\}>Z>\{a_0,b_0,a_1,b_1$
+% \item Gate description: $z_0+t+g_1+t\cdot g_1, t+t_1+t_2+t_1\cdot t_2, g_1+a_1\cdot b_1,
+% 			t_1+a_1\cdot ci, t_2+b_1\cdot ci, ci+a_0\cdot b_0, z_1+p_1+ci, p_1+a_1+b_1$
+% \item Word definition: $Z+z_0+z_1\cdot \alpha$
+% \item Vanishing polynomial ideal $(J_0)$: $z_0^2+z_0, z_1^2+z_1, t^2+t, g_1^2+g_1, t_1^2+t_1, t_2^2+t_2, p_1^2+p_1, ci^2+ci,
+% 			a_0^2+a_0, b_0^2+b_0, a_1^2+a_1, b_1^2+b_1, Z^4+Z$(since Z is a 2-bit word)
+% \end{itemize}
+% 
+% The result is a Gr\"obner basis with single polynomial generator. The polynomial has leading term $Z$:
+% $$Z+a_0\cdot b_0\cdot a_1+a_0\cdot b_0\cdot b_1+\alpha\cdot a_0\cdot b_0+a_1\cdot b_1+\alpha\cdot a_1+\alpha\cdot b_1$$
+% 
+% % \begin{table}
+% % \centering
+% % \begin{tabular}{|c|c|} \hline
+% % Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
+% % $\overline{a}$ & $1 + a$\\ \hline
+% % $a\ and\ b$ & $ab$\\ \hline
+% % $a\ or\ b$ & $a + b + ab$\\ \hline
+% % $a \oplus b$ & $a + b$\\
+% % \hline\end{tabular}
+% % \caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
+% % \label{table:booltogalois_op}
+% % \end{table}
+% 
+% 2) {\bf Templates on Boundary Information:}
+% \begin{figure}[h]
+% 	\begin{center}
+% 	\includegraphics[width=\textwidth]{newfig/template.pdf}
+% 	\end{center}
+% 	\caption{Standard 2-bit adder with input/output mapping}
+% 	\label{fig:template}
+% \end{figure}
+% 
+% Figure \ref{fig:template} shows a 2-bit standard cell. It has 3 outputs mapped to 2 output pins,
+% and 4 inputs mapped to 3 input pins (the rest pins are assigned to fixed 0/1 signal). Dashed connection
+% lines show one possible mapping, to find out this kind of feasible mapping, we need to simulate
+% all possible mappings, where the concept "template" can be used.
+% 
+% {\it Output template}: $z_0+c_{{t_0}{z_0}}\cdot t_0+c_{t_1z_0}\cdot t_1+c0_{CO}\cdot CO+c_{n_0}$
+% 
+% When $c_{n_0} = 0$, then $\{c_{t_0z_0},c_{t_1z_0},c0_{CO}\}$ equals to 1 denotes a valid mapping to corresponding output pin. 
+% Conversely if $cn0 = 1$, the ``1" evaluation denotes mapping to negation of corresponding output pin.
+% 
+% {\it Input template}: $a_0+p_{a_0}+c0_{A_0}\cdot s_0+c1_{A_0}\cdot s_1+c2_{A_0}\cdot s_2$
+% 
+% When $p_{a_0} = 0$, then $\{c0_{A_0},c1_{A_0},c2_{A_0}\}$ equal to 1 denotes a valid mapping to corresponding input
+% pin. Conversely if every variable in set $\{c0_{A_0},c1_{A_0},c2_{A_0}\}$ equals to 0, it denotes mapping to fixed signal ``0" (when $p_{a_0} = 1$ then
+% mapping to fixed signal ``1").
+% 
+% With these settings, we propose an approach to perform technology mapping with boundary information:
+% \begin{itemize}
+% \item First, choose a set of independent wires as input pins;
+% 
+% \item Second, push forward input signal for a certain number of gates (depth), choose a cut set
+% which fully dependent on these inputs as output pins;
+% 
+% \item Third, abstract a description polynomial using GB based abstraction;
+% 
+% \item Fourth, reduce the description polynomial with the GB we computed associated to macro block 
+% and templates, obtain a remainder polynomial;
+% 
+% \item Last but not least, abstract all coefficients from the remainder polynomial and set up a system of equations.
+% If we find a solution to this system, then the solution is a feasible mapping; no solution means we cannot make a
+% feasible mapping.
+% \end{itemize}

### 2016-12-10 dissertation/intro.tex
dissertation v3:update 1

@@
 \section{Hardware Design and Verification Overview}
-Because of the Moore's law, the level of integration in modern VLSI systems is becoming higher and higher. 
+During the last decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
 As a result, an entire system with billions of transistors can be built upon a single chip.
 The design process also evolves from manual design with little validation, to 
-a formal 3-step procudure which needs collaboration of teams with large number of 
+a formal 3-step procudure which requires collaboration of teams with large number of 
 engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
-2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
+2) Implement, which is to refine the design through various abstraction levels with the assistance of Computer-Aided-Design (CAD)
 tools; 3) Verify, which is to verify the correctness of design and implementation.
@@
 Nowadays the verification step is usually completed by a team that specializes on test, verification and validation of 
-circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
-Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procedures starting from 
+circuits. This step is also automated as an indispensable part of the CAD flow, when circuit synthesis is performed. 
+Figure \ref{fig:designflow} shows the typical synthesis flow, which covers procedures starting from the 
 register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
@@
 an important quality control measure before sending design layout to the VLSI foundaries.
-Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
+Considering the high cost of fabrication, faults and errors in the design will bring considerable 
 waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
@@
 One way to perform verification is by {\it simulation}. It is defined as the collage of all circuit validation 
-methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
+methods which apply stimulus on the inputs of circuit model and verify correctness of the outputs.
 However, simulation is not a complete solution to circuit verification problems. In modern designs with 
@@
 cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
-is a good example where simulation failed. Failure occurs with input assignments which are rarely used in most divisions. Because of the limitation of simulation,  
+is a good example where simulation failed. Failure occurred with input assignments which were rarely used in most divisions. Because of the limitation of simulation,  
 test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollors recalling bill
@@
 is described as theorems in mathematical logic. Subsequently logic rules
-are employed to deduce new objective theorem. In practice, the tool can reduce
+are employed to deduce new objective theorems. In practice, the tool can reduce
 a proof goal to simpler sub-goals for automatic verification.
@@
 \item \emph{Model checking} \cite{modelcheck:99} is a technique 
-for verifying if the specification properties are violated in finite-state systems. In circuit verification 
+for verifying if the specification properties are violated in finite-state systems. In the circuit verification 
 domain, both the specification and the circuit implementation are modeled as a system of 
@@
 properties are violated. If violation occurs, 
-a counter-example is then generated as a transition firing path corresponds to the
+a counter-example is then generated as a transition firing path that corresponds to the
 false behavior in the design. 
@@
 such as checking functional equivalence between HDL description and RTL,
-checking RTL equivalence between RTL and synthesized/optimized netlist
-,and checking layout verification between netlist and layout for fabrication.
+checking RTL equivalence between RTL and synthesized/optimized netlist, 
+and checking layout verification between netlist and layout for fabrication.
 
@@
 \begin{enumerate}[{1)}]
-\item \emph{Graph-based} techniques compare between 2 circuit implementations 
+\item \emph{Graph-based} techniques compare two circuit implementations 
 by representing them using canonical graphs. 
 The earliest invented canonical graph is the \emph{Binary Decision Diagram} (BDD) \cite{BRYA86}.
-Many variants branch out from BDD, some widely used variants include 
+Many variants branch out from BDDs; some widely used variants include 
 ZDD \cite{minato1993zero}, BMD \cite{bmd}, FDD \cite{okfdd}, etc. The comparison algorithms can 
@@
 the graphs correspond to the two circuits will be 
-equivalent if and only if the circuits are implementing the same function.
+equivalent if and only if the circuits implement the same function.
 \item \emph{Satisfiability} (SAT) techniques utilize the satisfiability theory.
@@
 miter output is ``1". If such an assignment exists, this solution acts as a 
-counter-example of when the circuit outputs differ; otherwise the circuits
+counter-example to equivalence; otherwise the circuits
 are functionally equivalent.
@@
 \emph{computer-algebra} and \emph{algebraic geometry}.
-Algebraic geometry provide a way to reason about the presence or absence of solutions 
+Algebraic geometry provides a way to reason about the presence or absence of solutions 
 without actually solving the system of constraints.
@@
 Computation of GB allows for 
-a straightforward proof of important properties of a polynomial system, 
+a straightforward proof of important properties of the polynomial system, 
 such as the presence or absence of 
@@
 verification. The disadvantage of the GB computation method is that its complexity can be doubly 
-exponential in the worst case \cite{dube1986complexity}, and thus methods directly performing GB computation 
-from an arbitrary setup is not practical for industry-level applications. However, recent
+exponential in the worst case \cite{dube1986complexity}. Thus, directly performing GB computation 
+over arbitrary setup is not practical for industry-level applications. However, recent
 breakthroughs in computer-algebra hardware verification have shown
@@
 way of abstraction -- a key technique to reduce the state space of a
-FSM. It has the effect of combining sets of states with similar 
+sequential circuit. It has the effect of combining sets of states with similar 
 properties. During reachability analysis, if we use bit-level
@@
 verification. For example, it can work as automatic decision and canonical reduction engine in theorem proving;
-for RTL composed of macro blocks, abstractions of these blocks are also benefit RTL verification.
+for RTL composed of macro blocks, abstractions of these blocks also benefit RTL verification.
 Concretely, MC and equivalence checking with abstractions can be classified as:
@@
 \item Model checking with abstractions \cite{kroening:model}, 
-where an over-pproximation of RTL blocks is abstracted and used to 
-refine the RTL model.
+where an over-approximation of RTL blocks is abstracted and used for property checking on a simplified model.
 \item Graph-based equivalence 
@@
 abstractions are used to analyze structural symmetries and similarities such that
-the Boolean formulas fed to SAT solver is minimized.
+the Boolean formulas fed to SAT solver is simplified.
 \end{itemize}
@@
 datapaths. Moreover, 
-word-level abstractions facilitates uninterpreted functions (UFs) \cite{UF3}, which 
+word-level abstractions facilitates the use of uninterpreted functions (UFs) \cite{UF3}, which 
 can be transformed into proposition formulas with word-level information and verified using 
@@
 can be identified by checking whether ``bad" states are {\it
-  reachable} from certain initial states. Temporal logic {\it model
-  checking} (MC) formulations and solvers are often used for this
+  reachable} from certain initial states. Temporal logic model
+  checking formulations and solvers are often used for this
 purpose. Once the designs and specification models are validated using
@@
 constraints to model the transition relations and sets of 
-states. We have access to instances from our industrial
-collaborators (Calypto Design Systems, now Mentor Graphics) where
+states. Often,
 the designs are expressed at the level of bit-vector words
@@
 the control logic. While attempts have been made towards word-level
-predicate abstraction \cite{jain2005word} \cite{mcmillan:cav06}
-\cite{mcmillan2010lazy}, {\it using a purely word-level representation
+predicate abstraction \cite{jain2005word,mcmillan:cav06,mcmillan2010lazy}, 
+{\it using a purely word-level representation
   of the state-space, the properties and their abstractions have not
@@
   verification.}
-
-For word-level SEC, we are given two designs, or their corresponding
-Mealy/Moore FSMs ${\mathcal{M}}_1,{\mathcal{M}}_2$, along with their
-initial starting states $S_0^1,S_0^2$. We wish to prove the absence of
-a sequence of inputs (string) that distinguishes the initial
-states \cite{coudert:iccad90}
-\cite{coudert1990verification}. Fundamentally, this requires 
-the construction of a product machine; and the main research problem
-relates to that of performing {\it FSM traversal} \cite{touati1990implicit}
-but  at the word-level. Analogously, in the case of MC, the problem
-is setup {\it w.r.t.} a FSM $\mathcal{M}$, a set of initial states $S_0$ and
-a set of property states $p$. Techniques are to be researched that
-verify that there exist no sequence of transitions from an initial
-state to a non-property state (``bad'' state). These problems have to
-be solved in the context of word-level verification -- {\it i.e.} data
-representation, abstraction using UNSAT cores and
-algorithm execution has to be carried out at word-level.
 
@@
 
-\subsection{Motivating Application}
+For word-level SEC, we are given two designs, or their corresponding
+Mealy/Moore FSMs ${\mathcal{M}}_1,{\mathcal{M}}_2$, along with their
+initial starting states $S_0^1,S_0^2$. We wish to prove the absence of
+a sequence of inputs (string) that distinguishes the initial
+states \cite{coudert:iccad90,coudert1990verification}. Fundamentally, this requires 
+the construction of a product machine; and the main research problem
+relates to that of performing {\it FSM traversal} \cite{touati1990implicit}
+but  at the word-level. Analogously, in the case of MC, the problem
+is setup {\it w.r.t.} a FSM $\mathcal{M}$, a set of initial states $S_0$ and
+a set of property states $p$. Techniques are to be researched that
+verify that there exist no sequence of transitions from an initial
+state to a non-property state (``bad'' state). These problems have to
+be solved in the context of word-level verification -- {\it i.e.} data
+representation, abstraction using UNSAT cores and
+algorithm execution has to be carried out at word-level.
 
-% TODO move to prev
-% The use of algebraic geometry
-% for sequential circuit verification and symbolic model checking has
-% been presented before. Avrunin presented the
-% concept of symbolic MC using algebraic geometry in
-% \cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
-% GB-algorithms for CTL, LTL and 
-% bounded MC over Boolean rings. However, these approaches are a
-% straight-forward transformation of the problem to {\it bit-level}
-% Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
-% the concepts of word-level reachability, abstraction-refinement using
-% interpolation or UNSAT cores, etc., that we desire was not the focus of
-% \cite{Avrunin:CAV} \cite{vardi-iasted07}.
-
-\subsection{Word-level Reachability Analysis}
+\subsection{Word-level Reachability Analysis of FSMs}
 In this dissertation, we propose a method to perform reachability analysis at the word-level. 
 	The given FSM is modeled as a system of polynomials over a finite field,
-	where the state space is mapped to the solutions of polynomial system.
-	Our proposed algorithm utilizes concepts of ideal-variety correspondence in algebraic geometry.
-	It also forms the foundation for word-level SEC and MC by enabling word-level abstraction of the
+	where the state space is mapped to the solutions of the polynomial system.
+	Our proposed algorithm utilizes ideal-variety correspondences in algebraic geometry.
+	It also forms the foundation for word-level verification by enabling word-level abstraction of the
   state-space.
@@
 $f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
-field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
+field ($\Fq$) of  $q$ elements, every function $f: \Fq
 \rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
 there exists a unique canonical polynomial $\Func$ that describes $f$.
-This implies that one can derive a {\it  canonical, polynomial
-  abstraction} of the function as $Z = \Func(A)$ where $Z, A$ are
+This implies that one can derive a canonical, polynomial
+  abstraction of the function as $Z = \Func(A)$ where $Z, A$ are
 word-level symbols representing $k$-bit operands. The concept also
@@
 {\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
-This implies that the FSM's transition relations can be
+{\it This implies that the FSM's transition relations can be
 represented as polynomial functions (ideals) in $\Fkk$, and values of
-state variables can be represented as solutions to these polynomials
+state variables can be represented as solutions to these polynomials}
 (variety of the ideal). Subsequently, the ideal-variety correspondences
@@
 word-level abstraction to the techniques in sequential verification. 
+This work was published at \cite{myHLDVT}.
 
-\subsection{Application to Sequential Galois Field Arithmetic Circuits Verification}
+\subsection{Application to Sequential Galois Field Arithmetic Circuit Verification}
 Sequential Galois field (GF) arithmetic circuits can be modeled as a special type of 
-FSM, where the pre-loaded operands are mapped to initial states and pseudo outputs
+FSM, where the pre-loaded operands are mapped to initial states, and pseudo outputs
 after $k$ clock-cycles are recognized as the final reached state after $k$ transitions.
@@
 
-In our proposed approach, word-level abstraction is employed to generate the 
+In our proposed approach, word-level abstraction is employed to generate, in every time-frame, the 
 word-level signature of the combinational logic component of the sequential 
-arithmetic circuit. This abstraction requires GB computation, which usually has high time/space 
+arithmetic circuit. This abstraction requires a GB computation, which usually has high time/space 
 complexity. We propose several improvements to simplify the GB computation procedure
 and make the entire algorithm execution practical. As a result, we successfully verify sequential 
-multipliers with 162-bit datapaths.
+multipliers with 162-bit datapaths. This work was published at \cite{myDATE} and a journal paper
+is under preparation.
 
-\subsection{UNSAT Cores in Algebraic Geometry for Abstraction Refinement}
+\subsection{UNSAT Cores in Algebraic Geometry}
 Abstraction is an effective method to lower the cost to traverse the state space.
-In modern model checkers, abstraction is used to refine the model 
-during the iterative execution of the tool. UNSAT core is widely used 
+In modern model checkers, abstraction is used to simplify and refine the model 
+during the iterative execution of the tool. An UNSAT core is widely used 
 as an important component of abstraction refinement. The reason is that UNSAT cores can provide 
-information to indicate the redundancy in the state variables, and that information is necessary for the refinement 
-process.
+information about the state variables that truly affect the property, 
+and that information is necessary for the refinement process.
 
 In this dissertation, we explore the concept, and the computation, of unsatisfiable (UNSAT)
-  cores of  a set of polynomials using Gr\"obner bases calculation. We also propose
-a number of heuristics by cracking the Buchberger's algorithm to reduce the size of UNSAT core. 
-  We apply the UNSAT core extraction 
-  to a bounded model checker (BMC) with abstraction refinement.
+  cores of  a set of polynomials using the Gr\"obner bases algorithm. We also propose
+a number of heuristics that extend the Buchberger's algorithm to reduce the size of UNSAT core. 
+  We demonstrate the use of UNSAT core extraction 
+  to a bounded model checking instance with abstraction refinement.
+ This work was published at \cite{myCP}.
 
@@
 UNSAT cores in algebraic geometry can simplify BMC. 
-Chapter \ref{ch:conclude} concludes the dissertation and outlines potential future research for
-continuation of this work. Appendix provides theory and methodology on characterization of normal basis,
+Chapter \ref{ch:conclude} outlines potential future research for
+continuation of this work and concludes the dissertation. 
+An appendix provides theory and methodology on the characterization of finite field normal basis,
 as well as the construction of optimal normal basis and application to normal basis multiplier design.

### 2016-12-10 dissertation/prev.tex
dissertation v2:Appendix update 1
dissertation v3:update 1

--- 2016-10-24 dissertation/prev.tex
@@
 \label{ch:prev}
-\section{SAT,DD,AIG-based Sequential Circuit Verification}
+\section{Sequential Equivalence Checking}
+As an important component of formal verification for sequential circuits, SEC techniques 
+have been developed over decades and widely utilized in both academia and industry. 
+The specification of a sequential circuit can be modeled as a (golden model) state machine,
+SEC is performed to compare between the circuit for test and the golden one.
+% The n\"aive method to implement SEC is: preload both circuits to the same initial states, 
+% and assign their primary inputs to the same values during all clock-cycles.
+% This method needs to be operated along with state space traversal,  therefore it is 
+% less efficient. Moreover,  most SEC only check the primary outputs/inputs consistency 
+% and does not require the 1:1 state correspondence,  so state space traversal is not always necessary.
+One way to implement SEC is to create miter with two circuits to be verified, then 
+prove that there exists no sequence of inputs that generates different outputs.
 
-\section{Bounded Model Checking}
+Researchers proposed improvements by using Boolean functions to represent 
+a set of states/transitions \cite{coudert2003unified, coudert1990verification},  or by dividing the sequential circuit to 
+smaller subcircuit and re-model the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
+interfaces that focuses on only designated initial states and remove redundancies in state space \cite{baumgartner2007scalable}.
 
-\section{Model Checking with Abstraction Refinement}
+Another direction to improve SEC algorithms is to avoid using state space traversal. 
+The forward retiming method \cite{van1998sequential} and time-frame merging \cite{stoffel1997record} 
+all work on an array of time-frames,  with the assistance of combinational equivalence checking (CEC)
+techniques. These techniques require structural similarities between the two circuits.
 
-\section{Word-level Techniques in Sequential Circuit Verification}
-Term rewriting?
+The most significant difference of sequential circuits from combinational circuits is 
+that the outputs of the circuit depend not only 
+on the primary inputs, but also on current state. 
+The behavioral difference reflects on the structural design of circuits and in 
+the existence of memory components such as latches and flip-flops.
+In order to test certain properties on some signals across multiple clock-cycles,
+the most straightforward method is to propagate those signals throughout 
+all clock-cycles. Moreover, for formal verification, all signals on all paths from the circuit
+need to be propagated through multiple clock-cycles.
+This indicates a time-to-space conversion, where 
+ the combinational part of circuit
+is copied over several time-frames then connected together.
+The procedure is called {\it unrolling} of sequential circuit, as 
+Figure \ref{fig:unrolling} shows.
 
-\section{Verification of Galois Field Circuits}
+\begin{figure}[h]
+\centerline{
+\includegraphics[width=\textwidth]{newfig/unroll.pdf}
+}
+\caption{The unrolling of a sequential circuit}
+\label{fig:unrolling}
+\end{figure}
 
-\section{Conventional UNSAT Core Extraction}
+Unrolling provides a way to transform a sequential circuit to combinational 
+circuit. Therefore,  methods which can be applied to combinational circuit 
+verification are also suitable for unrolled sequential circuits. The canonical 
+graphical representation of the combinational circuit after unrolling is also 
+the canonical representation of the original sequential circuit. For the sequential 
+equivalence checking problem, we can also unroll the circuit to be verified and the 
+specification to combinational ones, and then perform combinational equivalence checking
+techniques \cite{avoj2010combinational}. In the following part we review research and techniques which 
+can be applied to unrolled sequential circuits.
 
-\section{Conclusion Remarks}
+\subsection{Canonical Decision Diagrams}
+The decision diagrams (DDs) are invented by inspirations of 
+using optimized data structures to accelerate formal verification.
+The most fundamental DD is the Binary DD (BDD), which originates from the 
+Shannon's expansion:
+\begin{equation}
+f(x, y, \dots) = x f_x + x' f_{x'}
+\end{equation}
+where $f_x = f(x = 1)$ and $f_{x'} = f(x = 0)$ denote the positive and
+negative co-factors of $f$ w.r.t. $x$, respectively.
+A BDD is usually represented as a binary tree.
+Its ordered and reduced form -- the Reduced Ordered Binary Decision Diagram (ROBBD)
+\cite{BRYA86}, was the first significant contribution because of its canonicity.  
+ ROBDDs represent a Boolean function as an
+implicit set of points on a canonical directed acyclic graph
+(DAG). Manipulation of Boolean functions can then be carried out as
+composition operations on their respective DAGs.
+
+Following BDDs,  variants of the Shannon's decomposition principle
+were explored to develop other functional decision diagrams such as
+ FDDs \cite{okfdd}, ADDs \cite{add}, MTBDDs \cite{mtbdd}, and their hybrid 
+edge-valued counterparts, HDDs \cite{hdd} and EVBDDs \cite{evbdd}. 
+Zero-suppressed BDDs (ZDDs) \cite{minato1993zero,minato1994calculation} use the if-then-else branches
+to represent the existence of variables in a cube, and result in lower 
+space complexity. It can be used to represent polynomials with integer coefficients.
+
+\begin{figure}[h]
+\centerline{
+\includegraphics[width=0.35\textwidth]{newfig/BDD.pdf}
+}
+\caption{ROBDD representing Boolean function $\neg a \land b \land (c\lor d)$ with order $a>b>c>d$}
+\label{fig:BDD}
+\end{figure}
+
+The DDs above are all based on bit-level operations. Even in the {\it Word-Level Decision Diagrams}
+\cite{WLS}, the decomposition is still point-wise, binary, 
+w.r.t. each Boolean variable. These representations do not
+serve the purpose of word-level abstraction from bit-level
+representations.
+
+Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivatives K*BMDs
+\cite{kbmd} and *PHDDs \cite{phdd}, perform the decomposition of a {\it linear} function
+based on its two moments instead of relying on Boolean decomposition. 
+MODDs \cite{modd,modd_tcomp} are a DAG representation of the
+characteristic function of a circuit over Galois fields $\Fkk$. 
+However, MODDs fails to compactly represent large circuits.
+
+Taylor Expansion Diagrams (TEDs) \cite{ted_tcomp} are a
+word-level canonical representation of a {\it polynomial expression},
+based on the Taylor's series expansion of a polynomial. However, they do
+not represent a {\it polynomial function} canonically.
+
+The use of DDs in traditional formal verification has a lot of advantages. 
+For example, DD-based model checking is very efficient as long as the DDs of sequential 
+circuit can be setup. The existence of violating states in constructed DDs 
+immediately deduces the violation of property. However, when the design gets
+larger and larger, the time and space cost of building and storing the diagram 
+increases rapidly. In our experiment of verifying a $k$-bit arithmetic circuit 
+using ZDDs, when $k$ is larger than 100, the construction of ZDDs occupies 
+over $99\%$ runtime of the whole procedure.
+\subsection{Combinational Equivalence Checking Techniques}
+The CEC problem can be solved using various methods.
+Besides using canonical DDs (BDDs
+\cite{BRYA86} and their word-level variants \cite{WLS}),
+non-canonical representations such as And-Invert-Graph (AIG) based reductions 
+\cite{AIG:2002,alanmi:cec:iccad2006} are also very effective. 
+Solvers for satisfiability problems (SAT) are good candidates to solve CEC problems,
+as long as the miter of two circuits can be described using conjunctive normal form (CNF)
+formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, {\it etc.}
+If the circuits being compared are structurally highly similar, AIG and circuit-SAT 
+based approaches are proved to be efficient.
+However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
+  contemporary techniques including quantifier-free bit-vector 
+  (QF-BV) theory based SMT-solvers \cite{Cryptol:fmcad09},
+  offer a practical solution.
+
+Recently integer polynomial based techniques \cite{ciesielski2014function,rolf:date16} are proposed  to verify the functional 
+correctness of integer arithmetic circuits. Their approach formulates the output signature as a polynomial function 
+with binary variables and integer coefficients, then rewrite the polynomial by substituting gate output with gate inputs
+variables. After going through the backward rewriting procedure,  the polynomial
+will be composed by only input variables. Then the polynomial is converted to a canonical representation, and compared
+with designated input signature. If they are equivalent, then the arithmetic circuit is successfully verified.
+This approach incurs of polynomial term explosion during the backward rewriting. The authors proposed a heuristic
+to levelize the arithmetic circuit, and substitute several gates' variables at the same time to minimize the risk. However, 
+the heuristic is proved to be less effective when the inner symmetry of circuit structure is missing.
+
+% To conclude, automatic formal verification of large {\it
+%     custom-designed modulo-arithmetic circuits} largely remains
+%   unsolved today.
+
+\section{Symbolic Model Checking and Abstraction Refinement}
+Model checking is a way to verify certain safety and liveness properties 
+in sequential circuits. Symbolic model checking which avoids using explicit state encoding
+provides more flexibility to reduce the state space and enhance the 
+efficiency of model checkers. The implementations of symbolic model checking 
+require canonical DDs or SAT solvers \cite{burch1990sequential,burch1991representing,biere1999symbolic}.
+
+Abstraction is a technique to reduce the state space representation by combining states with similar 
+characteristics. Sometimes it can effectively lower the number of states that require analysis by orders of magnitude,
+without affecting the properties we need to verify. Model checkers then utilize abstracted models 
+with interpolation \cite{mcmillan2003interpolation,mcmillan:cav06}.
+At first, abstraction was done manually by designers. Clarke {\it et al.} \cite{clarke2000counterexample}
+proposed a BDD-based automated abstraction by removing spurious paths from analysis of counterexamples. 
+Zhang {\it et al.} \cite{zhang2005design} proposed another abstraction method based on CNF-SAT.
+It implemented latch abstraction by removing "irrelevant" latches by analyzing the 
+UNSAT core from the $k$-BMC. Jain {\it et al.} \cite{HimanshuDAC2005} improved the abstraction refinement technique of \cite{clarke2000counterexample},
+where they use CNF-SAT to perform the refinement instead of using BDDs. The new approach is applied to verify RTL Verilog
+and was known to be successful.
+
+$k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
+on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
+\cite{bradley2011sat,bradley2011incremental}. IC3 works incrementally to find inductive subclauses
+of negations of reached states, meanwhile it is monolithic when computing over-approximations to sets of reachable
+states within $1,2,\dots,k$ steps. It is proved to be more efficient than interpolation based model checking
+although using similar mechanisms.
+
+Above techniques have limitations: they all rely on bit-level information from 
+the circuit, which prevents them from being applied to circuit with large datapaths.
+Meanwhile, their implementation relies on SAT/BDDs, which is an extension of Boolean 
+functions and not compatible with other forms of constraints.
+
+\section{Word-level Techniques applied to Sequential Circuit Synthesis and Validation}
+To better verify word-level designs, word-level verification techniques have been 
+explored in recent years. Directly translating bit-vector problems to bit-level 
+problems is called {\it bit-blasting}, and usually brings high redundancy and computational complexity in verification.
+Attempts to develop pure word-level techniques can be found in
+the rich domain of 
+theorem proving \cite{arditi:bmd} and bit-vector SMT-solvers
+\cite{boolector,cvc3,z3,bitvector98}, automated
+decision procedures for Presburger arithmetic \cite{presburger,bultan:mixed_verification}, 
+algebraic manipulation techniques 
+\cite{devadas:algebraic_manipulation_iccd91}, or the ones based on
+term re-writing \cite{AST}, etc.
+
+Polynomial, integer and other non-linear representations have also
+been researched: Difference Decision Diagrams (DDDs) \cite{ddd-csl99,ddd-mt-98}, interval
+diagrams \cite{interval_dd}, interval analysis using polynomials
+\cite{polynomial_sanchez99}, etc. Most of these have found 
+application in constraint satisfaction for simulation-based
+validation:  \cite{Ritter99,hsat,lpsat,brinkmann:asp-dac,Huang:tcad01,bitvector98}. Among
+these, \cite{brinkmann:asp-dac,Huang:tcad01,bitvector98}
+have been used to {\it solve} integer modular arithmetic on linear
+expressions -- a different application from {\it representing}
+finite field modulo-arithmetic on polynomials in a canonical form.
+
+Uninterpreted function abstraction is also an important category of 
+word-level techniques which facilitates word-level model checking.
+Usually uninterpreted symbols have no notion of bit-vector-precision. However, these techniques
+constrain them 
+using functional consistency among the evaluations of word variables
+\cite{UF1,UF2,UF3}.
+
+\section{Verification using Algebraic Geometry}
+
+Symbolic computer algebra techniques have been employed for formal
+verification of circuits over $\Z_{2^k}$ and also over
+Galois fields $\Fkk$. 
+Verification techniques using Gr\"obner bases
+\cite{Avrunin:CAV,gbverify:2007,manna:program} are proposed,
+but they do not address the problem of high computational complexity to
+compute Gr\"obner bases.
+
+Verification of a combinational Galois field arithmetic circuit $C$ against a
+polynomial specification $\Func$ has been previously addressed 
+\cite{ibm:blueveri,lv:tcad2013,pruss:dac14}. Verification problems in
+\cite{ibm:blueveri,lv:tcad2013} are formulated using
+Nullstellensatz and decided using the \Grobner basis algorithm.
+
+The paper 
+\cite{pruss:dac14} performs verification by deriving a canonical
+word-level polynomial representation $\F$ from the circuit $C$. Their
+approach views any arbitrary Boolean function (circuit) $f: \B^k
+\rightarrow \B^k$ as a polynomial function $f: \Fkk \rightarrow \Fkk$,
+and derives a canonical polynomial representation $\F$ over
+$\Fkk$. They show that this can be achieved by computing a reduced 
+\Grobner basis w.r.t. an {\it abstraction term order} derived from the
+circuit. Subsequently, they propose a \underline{r}efinement of this
+\underline{a}bstraction \underline{t}erm \underline{o}rder (called
+RATO), that enables to compute the \Grobner basis of a smaller subset
+of polynomials. The authors show that their approach can prove
+correctness of up to 571-bit combinational GF multipliers.
+
+IBM proposed a method to apply algebraic geometry techniques 
+on verifying error coding circuits \cite{BLUEVERI}.
+Recent papers \cite{rolf:date16,rolf:FMCAD16} provide a way to utilize 
+algebraic geometry and GB-based symbolic computing and perform 
+equivalence checking on integer arithmetic circuits and floating-point 
+arithmetic circuits, respectively.
+
+The use of algebraic geometry
+for sequential circuit verification and symbolic model checking has
+been presented before. Avrunin presented the
+concept of symbolic MC using algebraic geometry in
+\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
+GB-algorithms for CTL, LTL and 
+bounded MC over Boolean rings. However, these approaches are a
+straight-forward transformation of the problem to {\it bit-level}
+Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
+the concepts of word-level reachability, abstraction-refinement using
+interpolation or UNSAT cores, etc., that we desire was not the focus of
+\cite{Avrunin:CAV,vardi-iasted07}.
+
+\section{Concluding Remarks}
+From the investigation of previous work, techniques are to be researched that 
+can operate the FSM traversal on word-level to verify a property excluding spurious 
+faults. Meanwhile, many abstraction refinement techniques utilize information from UNSAT cores.
+We propose to solve these problems in the context of word-level verification,
+with data representation, abstraction and algorithm execution all carried out 
+on word-level.
+
+In this dissertation, we propose a purely word-level reachability analysis approach which is never done before.
+We achieve it by modeling the transition relations, states and the traversal algorithm at word-level.
+We borrow inspirations from \cite{tim:phd,gao:qe-gf-gb} to perform state space abstraction.
+Moreover, we find applications of our proposed approach to sequential arithmetic verification which is 
+not done before, either.

### 2016-12-11 dissertation/appendix.tex
dissertation v3:milestone

@@
 
-Let $k$ be a polynomial whose degree is larger than $f(x)$, i.e. $deg(k(x))>deg(f(x))$, then
+Let $k$ be a polynomial whose degree is larger than $f(x)$, {\it i.e.} $deg(k(x))>deg(f(x))$, then
 $$f(x)|k(x)~~\text{iff}~k(T)=0$$
@@
 Since linear map $f(\sigma)$ guarantees that all elements in the field 
-have order $lcm(f_0,\dots,f_{n-1})$, i.e. $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
+have order $lcm(f_0,\dots,f_{n-1})$, {\it i.e.} $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
 By contradiction we can prove that there is no factor of $x^n-1$ that can be divided by the product 
@@
 \begin{Theorem}
-If N is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
+If $\mathcal{N}$ is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
 \end{Theorem}
 \begin{Proof}
-Let basis $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
+Let basis $\mathcal{N} = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
 Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. 
@@
 So the first column always includes at least one nonzero element. 
-Since the sum of enries in each of other columns should equal to zero (modulo 2), they 
+Since the sum of entries in each of other columns should equal to zero (modulo 2), they 
 need to include an even number of ``1"s.
@@
 In $\F_{2^4}$ constructed with $p(\alpha) = \alpha^4 + \alpha + 1$, two normal bases can be found:
-$\beta = \alpha^3, N_1 = \{ \alpha^3, \alpha^6, \alpha^{12}, \alpha^9\}$ and 
-$\beta = \alpha^7, N_2 = \{\alpha^7, \alpha^{14}, \alpha^{13}, \alpha^{11}\}$. 
-Their multiplication tables are listed below. For basis $N_1$:
+$\beta = \alpha^3, \mathcal N_1 = \{ \alpha^3, \alpha^6, \alpha^{12}, \alpha^9\}$ and 
+$\beta = \alpha^7, \mathcal N_2 = \{\alpha^7, \alpha^{14}, \alpha^{13}, \alpha^{11}\}$. 
+Their multiplication tables are listed below. For basis $\mathcal N_1$:
 \begin{equation}
@@
 \end{equation}
-Complexity $C_N$ = 7, so $N_1$ is an ONB. For basis $N_2$:
+Complexity $C_N$ = 7, so $\mathcal N_1$ is an ONB. For basis $\mathcal N_2$:
 \begin{equation}
@@
 \end{equation}
-Complexity $C_N$ = 9, so basis $N_2$ is not optimal.
+Complexity $C_N$ = 9, so basis $\mathcal N_2$ is not optimal.
 \end{Example}
@@
 primitive polynomial nor normal element, so it is very convenient for circuit designers.
-Basis $N_1$ in Example \ref{ex:app4} is a type-I ONB over $\F_{2^4}$.
+Basis $\mathcal N_1$ in Example \ref{ex:app4} is a type-I ONB over $\F_{2^4}$.
 

### 2016-12-11 dissertation/conclude.tex
dissertation v3:milestone

@@
 In this section, we highlight some research problems that deserve further study.
-\subsection{Multivariate polynomial ideal based FSM traversal}
+\subsection{Multivariate Polynomial Ideal based FSM Traversal}
 In Chapter \ref{ch:reacha}, we always use a single word-level variable $T$ to denote the
@@
 
-\subsection{Use of the $F_4$ algorithm and ZDDs to Accelerate GB Reduction}
+\subsection{Use of the $F_4$ Algorithm and ZDDs to Accelerate GB Reduction}
 In Chapters \ref{ch:reacha} and \ref{ch:normal}, by using RATO we transform the GB-based
@@
 (From \cite{mcmillan2003interpolation}) Let $A$ and $B$ be Boolean
-  formulae given as sets of clauses such that $A \wedge B$ is
+  formulas given as sets of clauses such that $A \wedge B$ is
   unsatisfiable (UNSAT). Then there exists a formula $P$ such that: 
@@
 \bi
-\item Find the minimal interpolant: i.e. find the interpolant $F_P$
+\item Find the minimal interpolant: {\it i.e.} find the interpolant $F_P$
   such that $V(J_P)$ is the smallest variety larger than $V(J_A)$ such
@@
 \item Over finite fields, the variety of an elimination ideal is
-  exactly equal to the projection of the variety on the un-eliminated
-  variables. Consider Fig. \ref{fig:projection}, where variety of the
+  exactly equal to the projection of the variety on the remaining
+  variables. Consider Figure \ref{fig:projection}, where variety of the
   ideals $J_A, J_B$ are respectively projected on the common variables

### 2016-12-11 dissertation/intro.tex
dissertation v3:milestone

@@
 The design process also evolves from manual design with little validation, to 
-a formal 3-step procudure which requires collaboration of teams with large number of 
+a formal 3-step procedure which requires collaboration of teams with large number of 
 engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
@@
 Figure \ref{fig:designflow} shows the typical synthesis flow, which covers procedures starting from the 
-register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
+register-transfer-level (RTL) description (using hardware design languages, {\it i.e.} HDL) to  the 
 physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
 to ensure the implementation is consistent with the original design intent. Verification is 
-an important quality control measure before sending design layout to the VLSI foundaries.
+an important quality control measure before sending design layout to the VLSI foundries.
 Considering the high cost of fabrication, faults and errors in the design will bring considerable 
@@
 However, simulation is not a complete solution to circuit verification problems. In modern designs with 
-large number of logic components and complicated architecures, it is impractical to simulate all possible 
+large number of logic components and complicated architectures, it is impractical to simulate all possible 
 test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
@@
 is a good example where simulation failed. Failure occurred with input assignments which were rarely used in most divisions. Because of the limitation of simulation,  
-test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollors recalling bill
+test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollars recalling bill
 for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.
@@
 
-\section{Formal Verification: Property and Equilavence Checking}
+\section{Formal Verification: Property and Equivalence Checking}
 
@@
 Modern model checking techniques use the error-trace to automatically refine
-the system and perform further checkings.
+the system and perform further checking.
 \end{enumerate}
@@
 Many variants branch out from BDDs; some widely used variants include 
-ZDD \cite{minato1993zero}, BMD \cite{bmd}, FDD \cite{okfdd}, etc. The comparison algorithms can 
-determine whether the two graphs are isomorphic. The canonity of the graph representation guarantees that
+ZDD \cite{minato1993zero}, BMD \cite{bmd}, FDD \cite{okfdd}, {\it etc.} The comparison algorithms can 
+determine whether the two graphs are isomorphic. The canonicity of the graph representation guarantees that
 the graphs correspond to the two circuits will be 
@@
 the designs are expressed at the level of bit-vector words
-(e.g. Matlab code, Verilog), and these word-level abstractions are
+({\it e.g.} Matlab code, Verilog), and these word-level abstractions are
 rarely exploited in verification. The problem is further exacerbated
@@
 multi-variate polynomials with coefficients from the finite (Galois)
-field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
+field $\Fkk$ of $2^k$ elements, {\it i.e.} polynomials in the ring
 $\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
@@
 generalizes to functions with different input/output bit-vector sizes,
-i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
+{\it i.e.} functions of the type $f: \B^n \rightarrow \B^m$, modeled as
 polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow

### 2016-12-11 dissertation/normal.tex
dissertation v3:milestone

@@
-\chapter{Functional Verification of Sequential Normal Basis Multiplier}
+\chapter{Functional Verification of Sequential Normal Basis Multipliers}
 \label{ch:normal}
-In order to utilize our traversal algorithm, it is necessary to find out
-a sort of suitable circuit benchmarks which is easy to compute its
-Gr\"obner basis (GB). From the work of {\it Lv} \cite{lv:phd},
-we learn that arithmetic circuits in Galois field (GF) is
-convertible to an ideal of circuit polynomials, and the 
-ideal generators form a GB themselves when applying reverse topological
-term order. Furthermore, according to the work of {\it Pruss et al.}
-\cite{pruss:tcad15}, with a limited computation complexity,
-we can abstract the word-level signature of an arithmetic 
-component working in GF. Thus, we consider the possibility 
-of applying our traversal algorithm on sequential Galois
-field circuits. In each frame, we can use the techniques 
+With the size of sequential arithmetic circuits increasing rapidly, 
+conventional methods become inefficient on verifying the correctness of these circuits.
+In Chapter \ref{ch:reacha} we proposed a FSM traversal approach at word-level 
+based on abstractions. Since the sequential arithmetic circuits can be modeled 
+as FSMs, it is possible to apply our word-level traversal algorithm to solving 
+the functional verification problem on these circuits.
+From the work of Lv \cite{lv:phd},
+we learn that arithmetic circuits in Galois field (GF) can be
+represented using an ideal of polynomials, and the 
+ideal generators themselves form a GB under a specialized term order
+obtained by a reverse topological traversal of the circuit.
+Furthermore, according to the work of Pruss {\it et al.}
+\cite{pruss:tcad15}, we can also abstract the word-level signature of an arithmetic 
+component working in GF. Thus, we consider exploiting the concepts of \cite{lv:phd} and \cite{pruss:tcad15}
+to efficiently perform abstraction of the state-space to specific sequential GF circuits.
+In each time-frame, we can use the techniques 
 from \cite{pruss:tcad15} to abstract the word-level
@@
 to the transition function in our traversal algorithm.
-As a result, we manage to find a type of sequential GF multiplier
+We show how this approach can be applied to verify sequential GF multipliers
 which we can apply our traversal algorithm to actually 
@@
 \label{sec:normal_motiv}
-From the preliminaries about FSMs in Section \ref{sec:reach}, we learn that the
-Moore machine does not rely on inputs for state transitions. 
+% From the preliminaries about FSMs in Section \ref{sec:reach}, we learn that the
+% Moore machine does not rely on inputs for state transitions. 
 As depicted in Figure \ref{fig:Moore}(a), a typical Moore machine implementation
@@
 $r_0,\dots,r_k$ are present state (PS) variables 
-standing for state inputs (SI), and $r_0',\dots,r_k'$ are next state (NS) variables standing for
-state outputs (SO). Figure \ref{fig:Moore}(b) shows the state transition graph (STG) of 
+standing for pseudo inputs, and $r_0',\dots,r_k'$ are next state (NS) variables standing for
+pseudo outputs. Figure \ref{fig:Moore}(b) shows the state transition graph (STG) of a
+GF arithmetic circuit implementation of 
 a Moore machine with $k+1$ distinct states. We notice that it forms a simple chain,
@@
 then the closed circuit executes without taking any additional information from outside,
-and store the results in registers after $k$ clock cycles. Its behavior can be described using
-STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, $R_{init}$ is the initial
-state (usually reset to all zeros), $R_1$ to $R_{k-1}$ are intermediate results stored as SO of current state and SI
-for next state, and $R_k$ (or $R_{final}$) is the final result given by arithmetic circuits (and equals to the answer
-to arithmetic function when circuit is working functional correctly).
+and stores the results in registers after $k$ clock cycles. Its behavior can be described using the 
+STG in Figure \ref{fig:Moore}(b): state $R$ denotes the bits stored in registers. Concretely, 
+$R_{init}$ is the initial state, $R_1$ to $R_{k-1}$ are intermediate results stored as NS variables 
+of current time-frame and PS variables for the next time-frame, and $R_k$ (or $R_{final}$) 
+is the final result given by arithmetic circuits (and equals to the
+arithmetic function when circuit is working correctly).
 This kind of design results in 
 reusing a smaller combinational logic component such that the area cost is greatly optimized.
-An example of these designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
+An example of such designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
 However, it also brings difficulties in verifying the the circuit functions.
@@
 $k$ time-frames, and performing an equivalence checking between the unrolled machine and
-the specification function. However, the number of gates will grow fast when doing unrolling
+the specification function. However, the number of gates will grow fast when unrolling
 on bit-level. Meanwhile the structural similarity based equivalence checking techniques 
 will fail when the sequential circuit is highly customized and optimized from the naive specification 
-function. As a result, conventional techniques is grossly inefficient for large circuits.
-Therefore, a new method based on our proposed word-level FSM traversal technique is worthy to be explored.
+function. As a result, conventional techniques are grossly inefficient for large circuits.
+Therefore, a new method based on our proposed word-level FSM traversal technique is worthy of exploration.
 
-\section{Formal Verification of Normal Basis Multiplier using Gr\"obner Basis}
-The gate-level design of a NB multiplier can be generated using appoaches introduced in 
+\section{Formal Verification of Normal Basis Multipliers using Gr\"obner Bases}
+The gate-level design of a NB multiplier can be generated using approaches introduced in 
 Section \ref{sec:nbdesign}. The gate-level netlist is ready to be verified using 
@@
 mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
-the concept ``RATO", which is previously used in Section \ref{sec:improve}.
+the concept of ``RATO", which is previously used in Section \ref{sec:improve}.
 % The following part borrows contents
@@
 $$R = \Func(A,B)$$
-To verify the functional correctness of a combinational NB multiplier (e.g. Mastrovito multiplier
+To verify the functional correctness of a combinational NB multiplier ({\it e.g.} Mastrovito multiplier
 or Montgomery multiplier), the function given by abstraction will be computed as
@@
 \LinesNumbered
- \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial
+ \KwIn{Circuit polynomial ideal $J_{ckt}$, vanishing ideal $J_0$, initial
    state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
@@
   	$i \gets i + 1$\;
-	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle
+	$G \gets$GB$( \langle J_{ckt} + J_0+ from_{i-1}(R,A,B) \rangle
     )$ with ATO\;
@@
 
-We follow  the sequential GF circuit model of
-Figure \ref{fig:sequential}, with word-level variables $A, B, R$
+Given a sequential GF circuit in Figure \ref{fig:sequential} over $\Fkk=\F_2[x]$, irreducible polynomial $p(\alpha) = 0$
+defines the primitive element $\alpha$, $\beta = \alpha^t$ is the normal element.
+The circuit includes word-level variables $A, B, R$
 denoting the {\it present states (PS)} and $A', B', R'$ denoting the {\it next
@@
 \beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
-store the result, and $A, B\ (A', B')$ store input operands. E.g.,
+store the result, and $A, B\ (A', B')$ store input operands. For example,
 for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
@@
 \sum_{i=0}^{k-1} a_i \beta^{2^i}$). We denote this set of polynomials
-as ideal $J = \langle 
-f_1, \dots, f_s \rangle \subset \Fkk[x_1, \dots, x_d, R, R', A, A', B,
+as ideal $J_{ckt} = \langle 
+f_1, \dots, f_s \rangle \subset \F_2[x_1, \dots, x_d, R, R', A, A', B,
   B']$, where $x_1, \dots, x_d$ denote the bit-level (Boolean) variables
-  of the circuit. The ideal of vanishing polynomials $J_0$ is also included, and
-then the implicit FSM unrolling problem is setup for abstraction.
+  of the circuit. The ideal of vanishing polynomials 
+$J_0 = \langle x_i^2-x_i,R_j^{2^k}-R_j\rangle,~\forall \text{ bit-level variables }x_i\text{ and word-level variables }R_j$, is also included. 
+Then the implicit FSM unrolling problem is setup for abstraction.
 
@@
 Projection of varieties (geometry) corresponds to elimination ideals
-(algebra), and can be analyzed via \Grobner bases. Therefore, we
+(algebra), and can be analyzed via \Grobner bases as shown in Section \ref{sec:abstraction}. Therefore, we
 employ a \Grobner basis computation with ATO: we use a {\it lex term
@@
 Consequently, $k$-successive \Grobner basis computations implicitly
-unroll the machine, and provide word-level algebraic $k$-cycle
+unroll the machine, and provide a word-level algebraic $k$-cycle
 abstraction for $R'$ as $R' = \Func(A_{init}, B_{init})$.
@@
 logic actually implements function not only related to current operands $A$ and $B$,
-but also involved with PS variable (i.e. temporary product) $R$. Which can be 
+but also involved with PS variable ({\it i.e.} temporary product) $R$. Which can be 
 obtained using the abstraction:
@@
 \label{ex:RHSMPO}
-Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO. The transition function for
-operands $A,B$ is doing cyclic shift, while transition function for $R$ has to be computed through Gr\"obner basis
+Figure \ref{fig:5bitRH} shows the detailed structure of a 5-bit RH-SMPO, with normal element $\beta = \alpha^5$. The transition function for
+operands $A,B$ perform cyclic shift, while the transition function for $R$ has to be computed through the Gr\"obner basis
 abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
@@
 After computing the Gr\"obner basis of $J_{ckt}+J_0+from^0$ using ATO 
-$$\text{all other bit-level variables} > R > \{R',A_{init},B_{init}\}$$
-there is a polynomial in form of $R'+\Func(A_{init},B_{init})$,
-which should be included by $to^{i+1}$. $to^{i+1}$ also exclude next state variable $A'$ and $B'$, instead we 
+$$\text{all bit-level variables} > R > \{R',A_{init},B_{init}\}$$
+there is a polynomial in the form of $R'+\Func(A_{init},B_{init})$,
+which should be included in $to^{i+1}$. $to^{i+1}$ also excludes next state variable $A'$ and $B'$, instead we 
 redefine $A_{init}$ and $B_{init}$ using next state bit-level variables $\{a_i', b_j'\}$. Next state Bit-level variables
@@
 \subsection{Overcome Computational Complexity using RATO}
-% although commented, but I insist refer to previous chapter
-% TODO experiment on naive slimgb on SMPO
-We implement Algorithm \ref{alg:modified} in \textsc{Singular}, with the same setup and environment 
-with the experiment in Section \ref{sec:exp_reacha}. The result on verifying Agnew's SMPO is shown 
+We implemented Algorithm \ref{alg:modified} in \textsc{Singular}, with similar setup and environments as
+with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
 in following table:
@@
 	\begin{center}
-	    \caption{(Dummy)Runtime of Gr\"obner Basis Computation of Mastrovito Multipliers in Singular using ATO $>$.}\label{tab:ATO}
+	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:ATO}
 	    \begin{tabular}{|c|c||c|} 
 	        \hline
-		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (minutes)   \\
+		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (seconds)   \\
 		\hline
-	        $16$	&  $1,871$  & $2.4$ \\
-		$24$	&  $3,135$  & $12$  \\
-	        $32$	&  $5,549$  & $22.6$ \\
-	        $40$	&  $8,587$  & $266$ \\
-		$48$	& $12,327$  & NA (Out of Memory) \\
+	        $5$	&  $65$  & $20$ \\
+		$11$	&  $275$  & $>35,000$  \\
 	        \hline
@@
 
-It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 40(dummy).
-Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} is also available to 
-accelerate the GB computation here. More specifically, we obviate the GB computation by turning the Buchberger's algorithm 
-into a single-step multivariate polynomial division.
+It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 11 bits operands.
+Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} can also be applied to 
+accelerate the GB computation here. More specifically, we circumvent the GB computation 
+of Line 5 in Algorithm \ref{alg:modified} by turning the Buchberger's algorithm 
+into a single-step of multivariate polynomial division.
 
 \begin{itemize}
-\item First, a set of polynomials is constructed by polynomials translated from each logic gates, and they generate 
-ideal $J_{gates}$. Similarly, ideal $J_{W-B}$ is generated by polynomials representing the word-bit correspondences.
-They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{W-B}$.
-\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using reverse topological order
-in circuit structure, followed by word-level NS output (e.g. $R'$ in Figure \ref{fig:sequential}), 
-then word-level PS inputs (e.g. $A,B,R$ in Figure \ref{fig:sequential}).
+\item First, a set of polynomials is constructed from each logic gates in the circuit, and they generate 
+ideal $J_{gates}$. Similarly, ideal $J_{WB}$ is generated by polynomials representing the word-to-bit correspondences.
+They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{WB}$.
+\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using a reverse topological order
+derived from the circuit structure, followed by word-level NS outputs ({\it e.g.} $R'$ in Figure \ref{fig:sequential}), 
+then word-level PS inputs ({\it e.g.} $A,B,R$ in Figure \ref{fig:sequential}).
 \item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
-RATO, there only exist one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
-\item Then we only need to compute the S-poly for $(f_w,f_g)$, the reduce the S-poly by $J_{ckt}+J_0$.
+RATO, there exists only one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
+\item Then we only need to compute the $Spoly$ for $(f_w,f_g)$, and reduce the $Spoly$ by $J_{ckt}+J_0$.
 \end{itemize}
@@
 \begin{align}
-&Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
+&S\text{-}poly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}\nonumber\\
 &(\alpha^3+\alpha^2+\alpha) r_1+(\alpha^4+\alpha^3+\alpha^2) r_2+(\alpha^2+\alpha) r_3+(\alpha) r_4\nonumber\\
@@
 \end{Example}
-Above example indicates that RATO based abstraction on 5-bit RH-SMPO will result a remainder containing both
+The above example indicates that RATO based abstraction on 5-bit RH-SMPO will result in a remainder containing both
 bit-level variables and word-level variables. Moreover, the number of variables is still large such that
@@
 a polynomial that contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
-this problem is to replace bit-level variables by equivalent polynomials that only contain word-level
-variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
+this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
+variables, {\it e.g.} $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
 approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.
@@
 \label{ex:BLVS}
-{\bf Objective}:\ Compute polynomial $a_i + \mathcal{G}_i(A)$ from $f_0 = a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A
+{\bf Objective}:\ In $\F_{2^5}$ with normal element $\beta = \alpha^5$, compute polynomial 
+$a_i + \mathcal{G}_i(A)$ from $f_0 = a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A
  = g_0 + A$.
@@
 First, compute $f_0^2= a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2 = g_0^2 + A^2$;
-then $f_0^4, f_0^8, f_0^{16}$. By repeating squaring we get a system of equations:
+then $f_0^4, f_0^8, f_0^{16}$. Use Lemma \ref{lem:Fermat}, by repeating squaring we get a system of equations:
 \begin{displaymath}
@@
 \end{displaymath}
-Following is the coefficients matrix form of this system of equations:
+Following is the coefficient matrix form of this system of equations:
 \begin{displaymath}
@@
 Using those polynomial equations, it is sufficient to translate all bit-level inputs in the remainder polynomial
-because of following lemma:
+because of following proposition:
 \begin{Proposition}
 \label{lem:bitlevelremainder}
-Remainder of S-poly reduction will only contain primary inputs (bit-level) and word-level output; furthermore, there will be one and
-only one term containing word-level output whose monomial is word-level output itself rather than higher order form.
+Due to RATO, remainder of $Spoly$ reduction will only contain primary inputs (bit-level) and word-level output; 
+furthermore, there will be one and
+only one term containing word-level output with degree equals to 1.
 \end{Proposition}
@@
 single word-level output variable (say it is $R$) and this term is the last term under RATO, which means there is only one term with
-$R$ in Spoly. Meanwhile in other polynomials from $J_{ckt}+J_0$ there is no such term containing $R$, so this term will be
-kept to remainder $r$, with exponent equals to 1.
+$R$ in $Spoly$. Meanwhile in other polynomials from $J_{ckt}+J_0$ there is no such term containing $R$, so this term 
+will appear in remainder $r$, with exponent equals to 1.
 \end{Proof}
@@
 \subsection{Solving a Linear System for Bit-to-Word Substitution}
-In Example \ref{ex:BLVS} we use a Gaussian-elimination method to solve the system of polynomial
+In Example \ref{ex:BLVS} we show a Gaussian-elimination method to solve the system of polynomial
 equations. We describe another formal method to solve following system of equations
@@
 
-Notice that ${\bf M}$ is constructed by squaring a row and assigning it to next row,
+Notice that ${\bf M}$ is constructed by squaring a row and assigning it to the next row,
 therefore its determinant exhibits certain special properties:
@@
 \begin{Proof}
-According to the definition Equation \ref{eqn:Moore}, the Moore determinant consists of all possible linear combinations of 
+According to Equation \ref{eqn:Moore}, the Moore determinant consists of all possible linear combinations of 
 $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$ with coefficients over $\Fp$. If $\{\alpha_0,\alpha_1,\dots,\alpha_{k-1}\}$
@@
 \end{equation}
-where $\Mi$ can be easily computed using Laplace expansion method, with complexity $O(k!)$.
+where $\Mi$ can be computed using Laplace expansion method, with complexity $O(k!)$.
 
@@
 \subsection{The Overall Verification Approach}
-Based on above concepts and improvements, the functional verification for sequential GF multiplier 
+Based on the above concepts and improvements, the functional verification for sequential GF multiplier 
 on word-level with $k$-bit input operands $A,B$ and $k$-bit output $R$ is described as follows:
@@
 \item Given a sequential GF multiplier $S$, with word-level $k$-bit inputs $A,B$ and output $R$.
-\item Choose a primitive polynomial $P(x)$ of degree $k$ and construct $\Fkk$, and let $P(\alpha)=0$.
+\item Given a primitive polynomial $P(x)$ of degree $k$ and construct $\Fkk$, and let $P(\alpha)=0$.
 \item Perform a reverse-topological traversal of $S$ to derive RATO: LEX with $\{x_1>x_2>\cdots>x_d>R'>R>A>B\}$,
-where $\{x_1,\dots,x_d\}$ are bit-level variables of $S$, $R'$ is the NS output value.
+where $\{x_1,\dots,x_d\}$ are bit-level variables of $S$, $R'$ is the NS output value, and 
+$R,A,B$ are word-level PS variables.
 \item Derive the set of bit-level polynomials $\{f_1,\dots,f_s\}$ from each gate in $S$, and represent them 
-using RATO. These polynomials are in the form $x_i+tail(f_i)$ where $x_i$ is the output of corresponding gate.
+using RATO. These polynomials are in the form $x_i+tail(f_i)$ where $x_i$ is the output of the corresponding logic gate.
 \item Compose the bit-level to word-level correspondences polynomials such as $r_0\beta+r_1\beta^2+\cdots+r_{k-1}\beta^{2^{k-1}}+R$.
@@
 \item Construct matrices $M_0,\dots,M_{k-1}$, where $M_i$ is $M$ with the $i$-column replaced by vector 
-$[R R^2 \cdots R^{2^{k-1}}]^T$, and $M$ is the reverse-circulant matrix generated by vector $(\beta,\beta^2,\dots,\beta^{2^{k-1}})$.
+$[R~~R^2~~\cdots~~R^{2^{k-1}}]^T$, and $M$ is the reverse-circulant matrix generated by vector $(\beta,\beta^2,\dots,\beta^{2^{k-1}})$.
 \item Symbolically compute the determinants $|M_i|$ to find set $F_R$, where $f_{R_i}:R_i+|M_i|$, for $0\leq i \leq k-1$.
@@
 \section{Software Implementation of Implicit Unrolling Approach}
-Our experiment on different size of SMPO designs is performed with both \textsc{Singular} \cite{DGPS}
+Our experiment on different sizes of SMPO designs is performed with both \textsc{Singular} \cite{DGPS}
 symbolic algebra computation system and our customized toolset developed using C++.
 The SMPO designs are given as gate-level netlists with registers, then translated to polynomials to compose
-elimination ideal for Gr\"obner basis calculation. The experiment is conducted on desktop with
+elimination ideals for Gr\"obner basis calculation. The experiment is conducted on a desktop with
 3.5GHz Intel $\text{Core}^\text{TM}$ i7 Quad-core CPU, 16 GB RAM and running 64-bit Linux OS.
 
-The \textsc{Singular} tool can read in scripts written in its own format similar to ANSI-C. 
-The input file format for our C++ tool is also designed to compatible with \text{Singular} 
+The \textsc{Singular} tool can read scripts written in its own format similar to ANSI-C. 
+The input file format for our C++ tool is also designed to be compatible with \textsc{Singular} 
 ring and polynomial definitions. For SMPO experiments, the main 
-loop of our script file performs the same function as Algorithm \ref{alg:modified} describes, while Gr\"obner basis
-computation in main loop can be divided into 4 different function parts:
+loop of our script file performs the same function as Algorithm \ref{alg:modified} describes, while the Gr\"obner basis
+computation in main loop is divided into 4 different functional parts:
 
-1) Pre-process:
+1) Pre-processing:
 
-This step is executed only once before main loop starts. The function of pre-process is to compute following system
+This step is executed only once before the main loop starts. The function of pre-processing is to compute following system
 of equations for bit-level inputs $a_0 \dots a_{k-1}$:
@@
 \end{displaymath}
-The methodology has been discussed in this chapter. For 5-bit SMPO example, we start from bit-word correspondence polynomial
+The methodology has been discussed in this chapter. For 5-bit SMPO example, we start from bit-to-word correspondence polynomial
 \begin{displaymath}
@@
 By replacing bit-level variable $a_i$ with $b_i, r_i$ or $r_i'$, and word-level variable $A$ with $B, R, R'$ respectively,
-we can directly get bit-word relation functions for another operand input, pseudo input and pseudo output.
+we can directly get bit-to-word relation functions for another operand input, pseudo input and pseudo output.
 
-One limitation to \textsc{Singular} tool is the degree of a term cannot exceed $2^{63}$, so when conducting experiments 
+One limitation of the \textsc{Singular} tool is the degree of a term cannot exceed $2^{63}$, so when conducting experiments 
 for SMPO circuits larger than
@@
 Since the bit-to-word substitution procedure
-only requires squaring of equations each time, the exponent of word $A$ can only be in the form $2^{i-1}$, i.e. 
+only requires squaring of equations each time, the exponent of word $A$ can only be in the form $2^{i-1}$, {\it i.e.} 
 $A^{2^0},A^{2^1},\dots,A^{2^{k-1}}$. To minimize the exponents presenting in both \textsc{Singular} and our tool, 
-we rewrite exponent $2^{i-1}$ to exponent $i$, i.e.
+we rewrite exponent $2^{i-1}$ to exponent $i$, {\it i.e.}
 $(A^{2^0},A^{2^1},\dots,A^{2^{k-1}}) \to (A, A^2, \dots, A^k)$. In this way result is rewritten to be
@@
 
-2) S-poly reduction:
+2) $Spoly$ reduction:
 
-First, S-poly is calculated based on RATO, then reduced with the ideal composed by circuit description polynomials ($J_{ckt}$).
-For already finished experiments, naive reduction (multi-division) is adopted, and this step takes largest portion of total 
-time consumption.
+First, $Spoly$ is calculated based on RATO, then reduced with the ideal composed by the circuit description polynomials ($J_{ckt}$).
+% For already finished experiments, naive reduction (multi-division) is adopted, and this step takes largest portion of total 
+% time consumption.
 
-For SMPO experiments, reduced Spoly has following generic form (all coefficients are omitted):
-$$redSpoly = R'+\Func(a_i,b_j,r_l,A,B,R) = \sum r_i + \sum a_ib_i + \sum a_iB + \sum b_iA + R' + R$$
+For SMPO experiments, reduced $Spoly$ has following generic form (all coefficients are omitted):
+\begin{equation}
+\label{eqn:redSpoly}
+redSpoly = R'+\Func(a_i,b_j,r_l,A,B,R) = \sum r_i + \sum a_ib_i + \sum a_iB + \sum b_iA + R' + R
+\end{equation}
 From observation, there is no cross-product terms of bit-by-bit or bit-by-word variables from the same input/output
-such as $a_ia_j, a_iA$, etc. Consider the necessary
-condition of our trick, this property of reduced Spoly guarantees the word level variable can only exist in the form $A^{2^{i-1}}$,
-after substituting bit-level variables with corresponding word-level variable.
+such as $a_ia_j, a_iA$, {\it etc.} 
+For our exponent rewriting approach to give a faithful result, it has to satisfy an important condition:
+there exists no cross-product terms simultaneously containing elements within the same set, {\it i.e. } sets
+$\{A,a_i\},\{B,b_j\}$ and $\{R,r_i\}$.
+We now show that $redSpoly$ actually satisfies this condition.
+This property of $redSpoly$ guarantees the word level variable can only exist in the form $A^{2^{i-1}}$,
+after substituting a bit-level variables with corresponding word-level variable.
 
-3) Substitute bit-level variables in reduced S-poly, i.e. 
+3) Substitute bit-level variables in reduced $Spoly$, {\it i.e.} 
 $$\Func(a_i,b_j,r_l,A,B,R) \xrightarrow{a_i+\mathcal{G}_i(A),b_j+\mathcal{G}_j(B),r_l+\mathcal{G}_l(R)} \mathcal{H}(A,B,R)$$
 
-Use the result from pre-process, get rid of $r_i, a_i$ and $b_i$ through substitution. This step yields following polynomial (consider the trick we used):
+Use the result from pre-processing, get rid of $r_i, a_i$ and $b_i$ through substitution. This step yields following polynomial:
 \begin{equation}
@@
 
-4) Substitute PS word-level variable $R$ with inputs $A$ and $B$, i.e. 
+4) Substitute PS word-level variable $R$ with inputs $A$ and $B$, {\it i.e.} 
 $$\Func(A,B,R) \xrightarrow{R+\Func'(A,B)} \Func(A,B)$$
 
-In Equation \ref{eqn:HABR}, there are still terms containing $R$ in the ideal we want to compute Gr\"obner basis.
+In Equation \ref{eqn:HABR}, there are still terms containing $R$ in the ideal for which we want to compute Gr\"obner basis.
 For PS variable $R$ we can use relation $R+\Func'(A,B)$, which is the last clock-cycle's output ($R'+\Func'(A,B)$) 
-with only leading term replaced in step "$from^i\gets to^i$" in 
-Algorithm \ref{alg:modified}. Basically this step has nothing different from the last one, however, the degree
-of its terms are not actual degree because of the exponent rewriting technique. For example,
-a power of $r$ -- $r^m$ is originally $r^{2^{m-1}}$. So if $R+\Func'(A,B)$ contains terms $A^iB^j$, the result 
-after exponent rewriting
-is $$(A^{2^{i-1}}B^{2^{j-1}})^{2^{m-1}} = A^{2^{((i+m-2)\bmod k)+1}}B^{2^{((j+m-2)\bmod k)+1}}$$
-Thus the actual exponent for $A$ and $B$ in $(A^iB^j)^m$ should be $((i+m-2)\bmod k)+1$ and $((j+m-2)\bmod k)+1$, respectively.
+with only leading term replaced in step ``$from^i\gets to^i$" in 
+Algorithm \ref{alg:modified}. This step needs to be specially executed because of the exponent rewriting technique. 
+If $R+\Func'(A,B)$ contains terms $A^iB^j$, it should be transformed back to the original exponent first,
+then compute the exponential of the entire term.
+Because of following equation
+$$(A^{2^{i-1}}B^{2^{j-1}})^{2^{m-1}} = A^{2^{((i+m-2)\bmod k)+1}}B^{2^{((j+m-2)\bmod k)+1}}$$
+the correct exponent for $A$ and $B$ in $(A^iB^j)^m$ should be $((i+m-2)\bmod k)+1$ and $((j+m-2)\bmod k)+1$, respectively.
 
-Within one iteration, after finishing steps 2) to 4), the output should be intermediate multiplication result (temporary product) 
+Within one iteration, after finishing steps 2) to 4), the output should correspond to the intermediate multiplication result (temporary product) 
 $R'+\Func(A,B)$. After $k$ iterations,
@@
 
-\section{Experimental Results}
+\section{Experiment Results}
 \label{sec:result}
-% TODO revise repeated parts and references, add C++ results
 
@@
 clock-cycles, except for the $k^{th}$ one, are also different. These
-designs have no internal logical/structural equivalencies, and
+designs have no internal logical/structural equivalences, and
 verification with SAT/BDDs/ABC is infeasible. Their dissimilarity is
@@
 run-times for verifying bug-free and buggy RH-SMPO and Agnew-SMPO are
-shown in Table \ref{tbl:exp1} and Table \ref{tbl:exp2}, respectively. 
-We can verify, as well as catch bugs in, up to 100-bit
-multipliers. Beyond 100-bit fields, our approach is infeasible --
+shown in Tables \ref{tbl:exp1} and \ref{tbl:exp2} (using \textsc{Singular})
+and Table \ref{tab:Cpp} (using C++ implementation), respectively. 
+We can verify, as well as catch bugs in, up to 100-bit (using \textsc{Singular})
+and 162-bit (using C++ implementation) multipliers.
+
+Beyond 162-bit fields, our approach is infeasible --
 mostly due to the fact that the intermediate 
@@
 \caption{\small Run-time (seconds) for verification of bug-free and
-  buggy RH-SMPO using our approach} 
+  buggy RH-SMPO with \textsc{Singular}} 
 \label{tbl:exp1}  
@@
 \caption{\small Run-time (seconds) for verification of
-  bug-free and buggy Agnew's SMPO our approach}
+  bug-free and buggy Agnew's SMPO with \textsc{Singular}}
 \label{tbl:exp2}
@@
 
-\section{Conclusions and Further Work}\label{sec:concl}
-%TODO some of this should be moved to future work chapter
+\begin{table}[hbt]
+\centering
+\caption{\small Run-time (seconds) for verification of
+  RH-SMPO and Agnew's SMPO using C++ implementation}
+\label{tab:Cpp}
+{\small 
+\begin{tabular}{|c|c||c|c|c|c|c|c|}
+\hline
+\multicolumn{2}{|c||}{\centering Operand size $k$} & 36 & 60 & 81 & 100 & 131 & 162 \\
+\hline
+\multirow{2}{2cm}{\centering RH-SMPO} & \#Polynomials & 4716 & 12960 & 21870 & 35600 & 56592 & 92826 \\
+\cline{2-8}
+ & Runtime & 14.3 & 213.3 & 1343 & 4685 & 26314 & 124194 \\
+\hline
+\multirow{2}{2cm}{\centering Agnew's SMPO} & \#Polynomials & 2700 & 7380 & 13356 & 20300 & 34715 & 52974 \\
+\cline{2-8}
+ & Runtime & 10.2 & 212.0 & 2684 & 4686 & 56568 & 119441 \\
 
-This proposal has described a method to verify sequential Galois field
+ \hline
+
+%  \multirow{3}{2.2cm}{\centering Benchmark} 
+% & \multirow{3}{0.9cm}{\centering \# Polys} 
+% & \multirow{3}{0.8cm}{\centering \# MUS} 
+% & \multicolumn{3}{c|}{\multirow{2}{1.8cm}{\centering Size of core}}
+%  & \multirow{3}{1.6cm}{\centering \# GB-core iterations}
+%  & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
+%  & \multirow{3}{1.8cm}{\centering Runtime of PicoMUS (sec)} \\
+%   & & &\multicolumn{3}{c|}{}& &\multicolumn{3}{c|}{}& \\
+%   \cline{4-6} \cline{8-10}
+%     & & & I & II & III & & I & II & III & \\
+% \hline
+% \hline
+% $5\times 5$ SMPO & 240  & 137  & 169 & 137 & 137  & 8  & 1222 & 1938 & 1698 & $<0.1$\\
+
+\end{tabular}
+}
+\end{table}
+
+\section{Concluding Remarks}
+
+This chapter has described a method to verify sequential Galois field
 multipliers over $\Fkk$ using computer algebra and algebraic geometry
@@
 
-Our approach still has following limitations: first, it can only be applied
-on XOR-rich circuits, while most industrial designs are AND-OR gates dominant;
-second, it only uses naive bit-word abstraction based on functions of arithmetic
-circuits, which will be inefficient when the function does not have a straightforward
-expression (such as an implicit function); last but not least, Gr\"obner basis computation
-in our improved approach still requires a very long time. To overcome these limitations,
-further explorations are needed for my research.
+% Our approach still has following limitations: first, it can only be applied
+% on XOR-rich circuits, while most industrial designs are AND-OR gates dominant;
+% second, it only uses naive bit-word abstraction based on functions of arithmetic
+% circuits, which will be inefficient when the function does not have a straightforward
+% expression (such as an implicit function); last but not least, Gr\"obner basis computation
+% in our improved approach still requires a very long time. To overcome these limitations,
+% further explorations are needed for my research.

### 2016-12-11 dissertation/prelim_GF.tex
dissertation v2:milestone
dissertation v3:milestone

@@
-\chapter{Galois Field and Sequential Arithmetic Circuits}
+\chapter{Finite Fields and Sequential Arithmetic Circuits}
 \label{ch:prelim_GF}
 This chapter provides a mathematical background for understanding 
-finite fields (Galois fields) and explains how to design Galois field arithmetic circuits.
+finite fields (Galois fields) and explains how to design Galois field (GF) arithmetic circuits.
 We first introduce the mathematical concepts of groups, rings, fields, and 
@@
 explain how to map them to a Boolean circuit implementation.
-Additonally, we introduce a special type of sequential arithmetic hardware based on normal basis, as well
+Additionally, we introduce a special type of sequential arithmetic hardware based on normal basis, as well
 as the normal basis theory behind the designing such hardware.
@@
 \cite{mastro:1989, PT:1985, acar:1998, wu:2002, Knezevic:2008} for hardware design over Galois fields 
-and previous work by {\it Lv} \cite{lv:phd}.
-Normal basis theory in this section refers to \cite{normal_book, gao:phd_normal_basis} and sequential
+and previous work by Lv \cite{lv:phd}.
+Normal basis theory in this section is referred from \cite{normal_book, gao:phd_normal_basis} and sequential
 normal basis arithmetic hardware designs come from \cite{mullinONB,MasseyOmura,agnew1991implementation, RHmulti}.
@@
 \label{sec:algebra}
-\subsection{Group, ring and field}
+\subsection{Group, Ring and Field}
 \begin{Definition}
-An {\bf abelian group} is a set $\mathbb{S}$ with a binary operation $'+'$
+An {\bf Abelian group} is a set $\mathbb{S}$ with a binary operation $'+'$
 which satisfies the following properties: 
@@
 
-The set of integers $\mathbb{Z}$ forms an abelian group under the addition operation.
+The set of integers $\mathbb{Z}$ forms an Abelian group under the addition operation.
 
@@
 \begin{itemize}
-\item $\mathbb{R}$ forms an abelian group under the '+' operation with additive identity element $0$.
+\item $\mathbb{R}$ forms an Abelian group under the '+' operation with additive identity element $0$.
 \item {\it Multiplicative Distributive Law}: For all $a, b, c \in$ $\mathbb{R}$, $a\cdot (b + c) = a\cdot b + a\cdot c$.
@@
 A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
-non-zero element in $\mathbb{F}$ has a multiplicative inverse; i.e. $\forall
+non-zero element in $\mathbb{F}$ has a multiplicative inverse; {\it i.e.} $\forall
 a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
@@
 A {\bf Galois field}, denote $\Fq$, is a field with a finite
-number of elements, $q$. The number of elements $q$ of the Galois field is
-a power of a prime integer, i.e. $q = p^k$, where $p$ is a prime
+number of elements, $q$. The number of elements $q$ of the field is
+a power of a prime integer, {\it i.e.} $q = p^k$, where $p$ is a prime
 integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
@@
 \begin{Definition}
-A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$, i.e. the set of all 
+A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$, {\it i.e.} the set of all 
 polynomials in $x$ with coefficients in $\F_2$, is {\bf irreducible} 
@@
 For example, $x^2+x+1$ is an irreducible polynomial over $\mathbb{F}_{2}$
-because it has no solutions in $\mathbb{F}_{2}$, i.e. $(0)^2+(0)+1=1\neq0$ 
+because it has no solutions in $\mathbb{F}_{2}$, {\it i.e.} $(0)^2+(0)+1=1\neq0$ 
 and $(1)^2+(1)+1=1\neq0$ over $\F_2$.
@@
 $\mathbb{F}_{2^k} \equiv \mathbb{F}_2[x] \pmod{P(x)}$.
-Let $\alpha$ be a root of $P(x)$, i.e., $P(\alpha)=0$.
+Let $\alpha$ be a root of $P(x)$, {\it i.e.} $P(\alpha)=0$.
 Since $P(x)$ is irreducible over
@@
 $P(x)=x^4+x^3+1 \in \mathbb{F}_2[x]$ is an irreducible polynomial of degree $k=4$. 
-Let $\alpha$ be the root of $P(x)$, i.e. $P(\alpha)=0$.
+Let $\alpha$ be the root of $P(x)$, {\it i.e.} $P(\alpha)=0$.
 
@@
 \begin{Theorem}
-$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, i.e. if $n$ divides $m$.
+$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, {\it i.e.} if $n$ divides $m$.
 \end{Theorem}
@@
 \section{Normal Basis Multiplier over Galois Field}
-From algebraic view, a field is a space, and field elements are points in the space. Those elements can be 
+From an algebraic perspective, a field is a space, and field elements are points in the space. Those elements can be 
 represented with unique coordinates, which requires the pre-definition of a basis vector. In this
 section, we discuss a special basis called normal basis, as well as the advantages adopting it in 
-GF operations esp. multiplication.
+GF operations, especially multiplication.
 \subsection{Normal Basis}
-Given a Galois field (GF) $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
+Given a Galois field $\Fkk$ is a finite field with  $2^k$ elements and characteristic equals to 2.
 Its elements can be written in polynomials of $\alpha$, when there is an irreducible polynomial $p(\alpha)$
@@
 to binary bit-vector representations by recording the coefficients. For example, for elements in
-$\F_{2^4}$, the results are shown in table \ref{tab:gfelement}, column ``Polynomial".
+$\F_{2^4}$, the results are shown in Table \ref{tab:gfelement}, column ``Polynomial".
 
@@
 
-Basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$ is called {\bf standard basis} (StdB), which results in
+The basis $\{1,\alpha,\alpha^2,\alpha^3,\dots,\alpha^{k-1}\}$ is called a {\bf standard basis} (StdB), which results in
 a straightforward representation for elements, and operations of elements such as addition and subtraction.
@@
 However, it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
-in bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
+in a bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
 the area and delay of circuit will be costly.
 
-In order to lower down the complexity of arithmetic circuit design, Massey and Omura \cite{MasseyOmura} % ref 7 in RH paper
-use a new basis to represent GF elements, which is called {\bf normal basis} (NB).
+In order to lower the complexity of arithmetic circuit design, Massey and Omura \cite{MasseyOmura} % ref 7 in RH paper
+use a new basis to represent GF elements, which is called a {\bf normal basis} (NB).
 A normal basis over $\Fkk$ is written in the form of
@@
 {\bf Normal element} $\beta$ is an element from the field which is used to construct the normal basis,
-and can be represent as a power of primitive element $\alpha$: 
+and can be represented as a power of the primitive element $\alpha$: 
 \begin{equation*}
@@
 \end{equation*}
-Exponent $t$ takes value in the given range when $\N$ fulfills the definition of a basis.
-
-Respectively, a field element in NB representation is actually
+Exponent $t$ takes values in the given range when $\N$ fulfills the definition of a basis.
+
+Correspondingly, a field element in NB representation is actually
 \begin{align*}
@@
 According to the definition, a normal basis is a vector where the next entry is the square of the former one.
-We note that the vector is cyclic, i.e. $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
-
-The addition and subtraction of elements in NB representation are similar to equation \ref{eqn:StdB}.
+We note that the vector is cyclic, {\it i.e.} $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
+
+The addition and subtraction of elements in NB representation are similar to Equation \ref{eqn:StdB}.
 However, what makes NB powerful is its ease of implementation when doing multiplications and exponentiations.
@@
 \end{Lemma}
-This lemma concludes that the square of an element in NB equals to a simple right-cyclic shift of the bit-vector.
+This lemma concludes that the square of an element in NB equals to a simple {\it right-cyclic shift} of the bit-vector.
 Obviously, StdB representation does not have this benefit.
@@
 This example shows that convenience to use NB when computing $2^k$ power of an element.
-Multiplication is a bit complicated than squaring; but when it is decomposed as bit-wise
-operations, the property in lemma \ref{lem:squareNB} can be well utilized.
+Multiplication is more complicated than squaring; but when it is decomposed as bit-wise
+operations, the property in Lemma \ref{lem:squareNB} can be well utilized.
 
@@
 \end{equation}
-Before discussing the details of the function $f_{mult}$, we can take 
-a square on both side of equation \ref{eqn:multiMSB}, i.e. $C^2 = A^2\times B^2$.
-Obviously, using the property in lemma \ref{lem:squareNB}, the original second most significant bit 
+Before discussing the details of the function $f_{mult}$, we can 
+square both sides of Equation \ref{eqn:multiMSB}, {\it i.e.} $C^2 = A^2\times B^2$.
+Obviously, using the property in Lemma \ref{lem:squareNB}, the original second most significant bit 
 becomes the new MSB because of right-cyclic shifting. 
@@
 \end{align*}
-Moreover, $M$ is a $k\times k$ square matrix. If we can find $M$, we obtain the design of the multiplier.
+Moreover, $M$ is a $k\times k$ square binary matrix. If we can find $M$, we obtain the design of the multiplier.
 
@@
 
-When taking different bits $l$ of the product in equation \ref{eqn:def_lambda}, 
+When taking different bits $l$ of the product in Equation \ref{eqn:def_lambda}, 
 we obtain a series of conjugate matrices of $M$. 
@@
 $$c_l = A \times M^{(l)} \cdot B^T$$
-Meanwhile, the operator shifting rule in equation \ref{eqn:shiftMSB} still holds. Then we have relation 
+Meanwhile, the operator shifting rule in Equation \ref{eqn:shiftMSB} still holds. Then we have the relation 
 $$c_{l-1} = A \cdot M^{(l-1)} \cdot B^T = shift(A) \cdot M^{(l)} \cdot shift(B)^T$$ which means
@@
 The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
-NB, i.e.
+NB, {\it i.e.}
 \begin{equation}
@@
 \end{equation}
-Substitution yields, result is an expression for l-th digit of product as showed in equation \ref{eqn:multiMSB}:
+Substitution yields an expression for $l$-th digit of product as showed in Equation \ref{eqn:multiMSB}:
 \begin{equation}
@@
 guarantee that $C_N$ has bound $o(k)$. However,
-Mullin et al. \cite{mullinONB} % This citation is valid
+Mullin {\it et al.} \cite{mullinONB} % This citation is valid
 proves that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
 its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
-called optimal normal basis (ONB), details are introduced in appendix \ref{append:ONB}.
+called optimal normal basis (ONB), details are introduced in Appendix \ref{append:ONB}.
 
@@
 \begin{Example}[Rijndael's finite field]
+\label{ex:Rijndael}
 Rijndael uses a characteristic 2 finite field with 256 elements, which can also be called the GF $\mathbb F_{2^8}$.
@@
 
-The final remainder is $1$, i.e. the product equals to 1 in StdB.
+The final remainder is $1$, {\it i.e.} the product equals to 1 in StdB.
 
@@
 \begin{equation*}
+\label{eqn:mat_R}
 M^{(0)} = \left(\begin{array}{lccccccr}
@@
 \label{sec:nbdesign}
-The NB multiplier design consumes much less gates than ordinary StdB multiplier design, even if 
+The NB multiplier design consists of fewer gates than ordinary StdB multiplier design, even if 
 we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
@@
 
-The behavior of Massey-Omura multiplier can be concluded as: pre-load operands $A,B$ and reset $C$ to 0, after 
+The behavior of Massey-Omura multiplier can be described as: pre-load operands $A,B$ and reset $C$ to 0, after 
 executing for $k$ clock cycles, the data stored in flip-flop array $C$ is the product $A\times B$.
@@
 
-The gate-level logic boxes are implementing following function:
+The gate-level logic boxes implement the following function:
 \begin{equation}
@@
 clock cycle, make $k$ outputs and add them up using shift register after $k$ clock cycles, 
-it will generate the same result with SMSO. This kind of architecture is named as 
+it will generate the same result with SMSO. This kind of architecture is named as a 
 sequential multiplier with parallel outputs (SMPO). The basic SMPO, as a conjugate of Massey-Omura multiplier,
-is invented by Agnew et al. \cite{agnew1991implementation}.
+was invented by Agnew {\it et al.} \cite{agnew1991implementation}.
 
@@
 \caption{5-bit Agnew's SMPO.
-Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, i.e.
-if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
+Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, {\it i.e.} if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
 Index $w = 2i\pmod 5$}
@@
 \end{equation}
-Note it is written in the form of Equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
+It is written in the form of Equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
 remaining bits of the product, which can be written in following general form polynomial:
@@
 After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
-some of them jump out of the circle and are independent from $\lambda$-Matrix.
+some of them jump out of the box and are independent from $\lambda$-Matrix.
 One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
@@
 \end{Theorem}
+where $F$ is as given in Equation \ref{eqn:aux}.
 This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
@@
 
- \KwIn{$A,B\in \Fkk$ given w.r.t. NB $N$}
+ \KwIn{$A,B\in \Fkk$ given {\it w.r.t.} NB $N$}
  \KwOut{$R=A\times B$}
@@
 turn the sum to NB form. In other words, theoretically we need $k$ XOR gates in $e$-layer, the output of 
-$i$-th gate corresponds to the bit multiplying $\beta^{2^i}$.
+$i$-th gate forms the coefficient of $\beta^{2^i}$.
 
 In order to obtain information 
-indicating interconnections between $d$-layer and $e$-layer, we need to interpret $\beta^{1+2^j}$ 
+indicating interconnections between the $d$-layer and $e$-layer, we need to interpret $\beta^{1+2^j}$ 
 to NB representation.
@@
 \section{Concluding Remarks}
-In this section, we introduce basic concepts such as the definition of finite fields and the construction 
-of finite fields. Moreover, we describes a special kind of basis in finite field, and its 
-application on Galois field hardware design. The sequential Galois field multipliers working based on these 
+In this chapter, we introduced basic concepts such as the definition of finite fields and the construction 
+of finite fields. Moreover, we described a special kind of basis in finite field, and its 
+application on Galois field hardware design. The sequential Galois field multipliers based on these 
 principles are used as good candidates for applying our functional verification approach in subsequent chapters.

### 2016-12-11 dissertation/prelim_basis.tex
dissertation v3:milestone

@@
-\chapter{Gr\"ober Basis and Algebraic Geometry} \label{ch:ideals}
+\chapter{Gr\"ober Bases and Algebraic Geometry} \label{ch:ideals}
 This chapter reviews fundamental concepts of commutative and 
@@
 mostly referred from the textbooks \cite{ideals:book,gb_book} and 
-previous work by {\it Lv} \cite{lv:phd} as well as {\it Pruss} \cite{pruss:tcad15}.
+previous work by Lv \cite{lv:phd} as well as Pruss \cite{pruss:tcad15}.
 
@@
 where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
-integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.
+integers $\ge 0$, {\it i.e.} $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.
 
@@
 
-A {\bf lexicographic ordering} (lex) is a total-ordering $>$ such that 
-variables in the terms are lexicographically ordered, i.e. simply based on 
-when the variables appear in the ordering.
+A {\bf lexicographic ordering} (LEX) is a total-ordering $>$ such that 
+variables in the terms are lexicographically ordered, {\it i.e.} simply based on 
+when the variables appear in the ordering. It is also a well-order, where the least element exists.
 Higher variable-degrees take 
-precedence over lower degrees for equivalent variables (e.g. $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
+precedence over lower degrees for equivalent variables ({\it e.g.} $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
 \begin{Definition}
@@
 
-A {\bf degree-lexicographic ordering} (deglex) is a total-ordering $>$ such 
+A {\bf degree-lexicographic ordering} (DEGLEX) is a total-ordering $>$ such 
 that the total degree of a term takes precedence over the lexicographic 
 ordering.  
-A {\bf degree-reverse-lexicographic ordering} (degrevlex) is the same as a
-deglex ordering, however terms are lexed in reverse.
+A {\bf degree-reverse-lexicographic ordering} (DEGREVLEX) is the same as a
+DEGLEX ordering, however terms are LEXed in reverse.
 
@@
 \sum_{i=1}^{d}\alpha_i = \sum_{i=1}^{d} \beta_i  \text{ and }
-x^{\alpha} > x^{\beta} & \text{w.r.t. lex order}
+x^{\alpha} > x^{\beta} & \textit{w.r.t.}\text{ LEX order}
 \end{cases}
@@
 \begin{eqnarray}
-     lt(f)&& \text{--- Leading Term} \\
-     lc(f)&& \text{--- Leading Coefficient} \\
-     lm(f)&& \text{--- Leading Monomial} \\
+     lt(f)&& \text{Leading Term} \\
+     lc(f)&& \text{Leading Coefficient} \\
+     lm(f)&& \text{Leading Monomial} \\
      tail(f)&& f - lt(f)
@@
 If $\frac{lt(f)}{lt(g)}$ is non-zero, then $f$ is considered divisible by $g$, 
-i.e. $g \mid f$.
+{\it i.e.} $g \mid f$.
 \end{Definition}
@@
     \end{equation*}
-    which is transitive and reflective closure of relation $f\stackrel{g}{\textstyle\longrightarrow}r$. 
+    which is the transitive and reflective closure of the relation $f\stackrel{g}{\textstyle\longrightarrow}r$. 
     Reduction is carried out using multivariate, polynomial long division. 
@@
 in $\F_q^d$. Any single point is a variety of some polynomial system:
-e.g. $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
+{\it e.g.} $(a_1,\dots, a_d)$ is a variety of $x_1-a_1 = x_2 - a_2 = \dots =
 x_d-a_d=0$. {\bf Finite unions} and {\bf finite  intersections} of
@@
 $\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
-ideal, i.e., 
+ideal, {\it i.e.} 
 $\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
@@
 verification. A given hardware design can be transformed into a set
-of polynomials over a field, $f_1, \ldots, f_s \in F$.
+of polynomials over a field, $F = \{f_1, \ldots, f_s\}$.
 This set of polynomials gives the system of equations:
@@
 \end{eqnarray}
-If $I$ is the ideal generated by $f_1$ and $f_2$, i.e. 
+If $I$ is the ideal generated by $f_1$ and $f_2$, {\it i.e.} 
 $I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
@@
 Let $\mathbb{F}$ be any field and let $\mathbf{a}=(a_{1},\dots,a_{d}) \in \mathbb{F}^d$ be a point, and $f \in
-\mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ if $f(\mathbf{a}) = 0$, i.e.,
-$\mathbf{a}$ is in the variety of $f$.
+\mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ 
+if $f(\mathbf{a}) = 0$, {\it i.e.} $\mathbf{a}$ is in the variety of $f$.
 
@@
 a number of important properties that can solve numerous polynomial 
-decision questions:
+decision questions; the following are used in this dissertation:
 
 \begin{itemize}
-\item Presence or absence of solutions (varieties)
-\item Dimension of the varieties
-\item Ideal membership of a polynomial
-\item $\dots \dots$
+\item Check for presence or absence of solutions (varieties)
+\item Compute dimension of the varieties
+\item Test ideal membership of a polynomial
+\item Find projection of varieties by eliminating variables in LEX order
 \end{itemize}
@@
 \stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
-i.e. dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
+{\it i.e.} dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
 zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
@@
 \SetAlgoNoLine
- \KwIn{$F = \{f_1, \dots, f_s\}$, such that $I=\langle f_1, \dots, f_s\rangle$}, and term order $>$
+\LinesNumbered
+ \KwIn{$F = \{f_1, \dots, f_s\}$, such that $I=\langle f_1, \dots, f_s\rangle$, and term order $>$}
  \KwOut{$G = \{g_1,\dots ,g_t\}$, a Gr\"{o}bner basis of $I$ }
@@
 When computing a \Grobner basis, it's important to note that if $lt(f_i)$ 
-and $lt(f_j)$ have no common variables, the S-poly reduction step in 
+and $lt(f_j)$ have no common variables, the $Spoly$ reduction step in 
 Buchberger's algorithm will reduce to 0.
 \begin{Lemma}
-In Buchberger's algorithm, when $gcd(lt(f_i),lt(f_j)) = 0$, the S-poly reduction
+In Buchberger's algorithm, when $gcd(lt(f_i),lt(f_j)) = 0$, the $Spoly$ reduction
 $$Spoly(f_{i}, f_{j}) \stackrel{G'}{\textstyle\longrightarrow}_+r$$
@@
 \item Minimize every $g_i \in G$, i.e $g_i=g_i/lc(g_i)$
-\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, i.e.
-remove every polynomial in $G$ whose leading term is divisible by the leading term of some other polynomial in $G$.
+\item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
+{\it i.e.} remove every polynomial in $G$ whose leading term is divisible by the leading term 
+of some other polynomial in $G$.
 \end{enumerate}
@@
 \item $h_1 \stackrel{G_1}{\textstyle\longrightarrow}_+ g_1$, where
-  $g_1$ is reduced w.r.t. $G_1 = \{h_2, \ldots, h_t\}$
+  $g_1$ is reduced {\it w.r.t.} $G_1 = \{h_2, \ldots, h_t\}$
 
 \item $h_2 \stackrel{G_2}{\textstyle\longrightarrow}_+ g_2$, where
-  $g_2$ is reduced w.r.t. $G_2 = \{g_1, h_3, \ldots, h_t\}$
+  $g_2$ is reduced {\it w.r.t.} $G_2 = \{g_1, h_3, \ldots, h_t\}$
 \item $h_3 \stackrel{G_3}{\textstyle\longrightarrow}_+ g_3$, where
-  $g_3$ is reduced w.r.t. $G_3 = \{g_1, g_2, h_4, \ldots, h_t\}$
+  $g_3$ is reduced {\it w.r.t.} $G_3 = \{g_1, g_2, h_4, \ldots, h_t\}$
 
@@
 \item $h_t \stackrel{G_t}{\textstyle\longrightarrow}_+ g_t$, where
-  $g_t$ is reduced w.r.t. $G_t = \{g_1, g_2, g_3, \ldots, g_{t-1}\}$
+  $g_t$ is reduced {\it w.r.t.} $G_t = \{g_1, g_2, g_3, \ldots, g_{t-1}\}$
 \end{itemize}
@@
 
-% TODO discuss complexity
-
-The high computational complexity if Gr\"obner basis computation is treated as an issue
+The high computational complexity of the Gr\"obner basis problem is an important issue
 because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
-the worst case computation time/space cost of its Gr\"obner basis is doubly exponential bounded -- $q^{q^{O(|\phi|)}}$
+the worst case computation time/space cost of its Gr\"obner basis is doubly exponential while the total degree of 
+polynomials in the Gr\"obner basis is bounded by 
+$O(2(\frac{1}{2}d^2+d)^{2^{n-1}})$ where $d$ is the degree of the ideal and $n$ is the number of variables
 \cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
 the polynomial is well restricted rather than arbitrary.
-Gao et al. \cite{gao:qe-gf-gb} proves that when the number of variables and degree of polynomials 
-are restricted, the complexity reduces to single exponential $q^{O(|\phi|)}$.
+Gao {\it et al.} \cite{gao:qe-gf-gb} proves that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
 This provides the possibility to make use of Gr\"obner basis under restricted situations,
@@
 orderings can result in different \Grobner basis computations for the 
-same ideal. Computation using a degrevlex ordering tends to be least 
+same ideal. Computation using a DEGREVLEX ordering tends to be least 
 difficult, while lex ordering tends to be computationally complex. However, 
@@
 %%%%%%%%%%%%%%%weak nullstellensatz%%%%%%%%%%%%%
-\begin{Theorem}{Weak Nullstellensatz}
+\begin{Theorem}[Weak Nullstellensatz]
 %$\left[\bf{Weak\  Nullstellensatz}\right]$ 
@@
 in ${\overline {\F}}^d$. For this purpose, we only need to check if 
-the ideal is generated by the unit element, i.e., $1\in J$. 
+the ideal is generated by the unit element, {\it i.e.} $1\in J$. 
 This approach can be used to evaluate the feasibility of constraints in 
@@
 
-\begin{Theorem}\label{thm:sns}
-({\it The Strong Nullstellensatz} \cite{gb_book}) 
+\begin{Theorem}[Strong Nullstellensatz]
+\label{thm:sns}
+\cite{gb_book} 
 Let $\overline{\F}$ be an algebraically closed field, and let $J$
@@
 polynomial, $V(J_0)=\Fq$. Similarly, over $\Fq[x_1,\dots,x_d]$, $J_0$ is 
-$\langle x_1^q-x_1,\dots,x_d^q-x_d \rangle$ and $V(J_0)=(F_q)^d$.
+$\langle x_1^q-x_1,\dots,x_d^q-x_d \rangle$ and $V(J_0)=(\F_q)^d$.
 
@@
 $I_2=\langle g_1,\dots g_t\rangle$, then the {\bf sum of ideals} 
-$I_1+I_2=\langle f_1,\dots,f_s,g_1,\dots g_t\rangle$
+$I_1+I_2=\langle f_1,\dots,f_s,g_1,\dots g_t\rangle$.
 \end{Definition}
@@
 For any Galois field $\Fq$, let $J \subset \Fq[x_1,\dots,x_d]$ be any ideal 
-and let $J_0 = \langle x_1^q-x_1, x_d^q-x_d \rangle$ be the ideal of all
+and let $J_0 = \langle x_1^q-x_1,\dots, x_d^q-x_d \rangle$ be the ideal of all
 vanishing polynomials. Let $V_{\Fq}(J)$ denote the variety of $J$ over $\Fq$.
@@
 \begin{Proof}
-\begin{enumerate}
-\item $\sqrt{J+J_0} = J+J_0$. That is, $J+J_0$ is a radical ideal.
+\begin{enumerate}[{1)}]
+\item $\sqrt{J+J_0} = J+J_0$. That is, $J+J_0$ is a radical ideal (Lemma 2.1 in \cite{gao:qe-gf-gb}).
 \item $V_{\Fq}(J)=V_{\overline{\Fq}}(J+J_0)$.
-\item Due to (2), $I(V_{\Fq}(J)) = I(V_{\overline{\Fq}}(J+J_0))$. 
+\item Due to 2), $I(V_{\Fq}(J)) = I(V_{\overline{\Fq}}(J+J_0))$. 
 By Strong Nullstellensatz, this is equivalent to $\sqrt{J+J_0}$.
-Finally, due to (1), this is equivalent to $J+J_0$.
+Finally, due to 1), this is equivalent to $J+J_0$.
 \end{enumerate}
@@
 %%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
-\begin{Theorem}\label{wnull:ff}
-$[\bf{Weak~Nullstellensatz~in~\Fq}]$\cite{null:1890}\\
+\begin{Theorem}[Weak Nullstellensatz in $\Fq$]
+\label{wnull:ff}
+\cite{null:1890}
 Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
@@
 J_0)=\emptyset$,  if and only if the reduced
-Gr\"obnerBasis$(J+J_{0})=\{1\}$. 
+Gr\"obner Basis $RGB(J+J_{0})=\{1\}$. 
 \end{Theorem}
@@
 
-\section{Elimination and Abstraction Theory}
+\section{Elimination Theory and Application to Abstraction}
 \label{sec:abstraction}
@@
 operation when some variables is not needed in modeling and analysis. 
-In this section, eliminating variables targets a tight-bound over-approximation 
-which is equivalent to existential quantifier elimination in first-order logic,
-or variable smoothing in Boolean operations. We introduce a eliminating method 
-based on algebraic geometry concepts, and use it as the fundamentals of 
+In this section, eliminating variables targets a tight-bound over-approximation.
+This is equivalent to the existential quantifier elimination in first-order logic,
+or variable smoothing in Boolean operations. We introduce a elimination method 
+based on algebraic geometry concepts, and use it as the fundamental of 
 abstraction of a circuit.
@@
 \subsection{Elimination Theory}
-Assume we are given a set of polynomials $f_1,\dots,f_s$ belongs to ring $\Fq[x_1,\dots,x_l,\dots,x_d]$.
-First, we show that eliminating the $x_1,\dots,x_l$ variables is equivalent to 
+Assume we are given a set of polynomials $f_1,\dots,f_s$ from a ring $\Fq[x_1,\dots,x_l,\dots,x_d]$.
+First, we show that eliminating $x_1,\dots,x_l$ variables related to 
 {\it projecting} the variety $V(\langle f_1,\dots,f_s\rangle)$ from 
@@
 $$\pi_l:\Fq^d\to\Fq^{d-l},~~\pi_l((x_1,\dots,x_d)) = (x_{l+1}.\dots,x_d)$$
-where $l<d$. For any set $A \subseteq \Fq^d$, we write 
-$$\pi_l(A) = \{\pi_i(x):x\in A\}\subseteq \Fq^{d-l}$$
+where $l<d$. For any subset $V \subseteq \Fq^d$, we write 
+$$\pi_l(V) = \{\pi_i(x):x\in A\}\subseteq \Fq^{d-l}$$
 \end{Definition}
@@
 The elimination ideal $J_l$ has eliminated all the variables 
-$x_1,\dots,x_l$, i.e. it only contains polynomials with variables in
+$x_1,\dots,x_l$, {\it i.e.} it only contains polynomials with variables in
 $x_{l+1},\dots,x_d$. 
+In a general setting,
+$\pi_l(V_d(J))$ from $\Fq^d$ to $\Fq^{d-l}$ is a subset of $V(J_l)$. But over $\Fq$, 
+consider ideal $J+J_0$ where $J_0$ is the ideal of vanishing polynomials, it is proved in \cite{gao:qe-gf-gb}
+that 
+$$\pi_l(V(J+J_0)) = V((J+J_0)_l)$$
+
 This shows that in finite fields, projection of the variety $V_{d}(\langle f_1,\dots,f_s\rangle)$
@@
 
-\begin{Theorem}{Elimination Theorem}
+\begin{Theorem}[Elimination Theorem]
 %$\left[\bf{Elimination\  Theorem}\right]$
@@
 
-{\bf Problem setup:} Let $S$ be the system of polynomials, 
+\begin{Problem}
+\label{setup:abs}
+Let $S$ be the system of polynomials, 
 $\{f_1,\dots,f_s,f_{A_1},\dots,f_{A_n},f_{Z}\}\subset \Fkk$, 
 derived from the hardware
-implementation of the Galois field arithmetic circuit over $\Fkk$, as Figure \ref{fig:abs_setup} shows.
-This circuit performs some unknown function $f$ over 
-$\Fkk$ in the form of $Z=\Func(A_1,\dots,A_n)$, where $Z$ is the $k$-bit 
-output and $A_1,\dots,A_n$ are the $k$-bit inputs.
-The polynomial representation of $\Func$ over $\Fkk$ is thus:
+implementation of any circuit over $\Fq, q=2^k$, as Figure \ref{fig:abs_setup} shows.
+In general, this circuit performs some unknown function $f$ over 
+$\Fq$ such that every $f:\Fq\to\Fq$ is represented in polynomial form.
+For any circuit $\mathcal C$, $Z = \Func(A)$. Then how to derive this function $\Func$
+using GB?
+\end{Problem}
+
+The polynomial representation of $\Func$ over $\Fkk$ is:
 
@@
 Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
-it agrees with the solution to the system of polynomials $\{S\}=0$, i.e.:
+it agrees with the solution to the system of polynomials $\{S\}=0$, {\it i.e.}:
 \begin{equation}
@@
 
-\begin{Theorem}
-{\bf Abstraction Theorem:} Using the notations from problem setup at the beginning of this subsection,
+\begin{Theorem}[Abstraction Theorem]
+Using the notations from Problem Setup \ref{setup:abs} at the beginning of this subsection,
 we compute a \Grobner basis $G$ of ideal $(J+J_0)$ using the abstraction term 
 order $>$. Then: \\
-(i) For every word-level input $A_i$, $G$ must contain the vanishing 
+1) For every word-level input $A_i$, $G$ must contain the vanishing 
 polynomial $A_i^q - A_i$ as the only polynomial with $A_i$ as its only 
 variable;\\
-(ii) $G$ must contain a polynomial of the form 
+2) $G$ must contain a polynomial of the form 
 $Z + \mathcal{G}(A_1,\dots,A_n)$; and\\ 
-(iii) $Z + \mathcal{G}(A_1,\dots,A_n)$ is such that 
+3) $Z + \mathcal{G}(A_1,\dots,A_n)$ is such that 
 $\Func(A_1,\dots,A_n) = {\mathcal{G}}(A_1,\dots,A_n),
@@
 
-$g_8=Z+A\cdot B$ is the {\bf canonical, word-level polynomial } 
+Here $g_8=Z+A\cdot B$ is the {\bf canonical, word-level polynomial } 
 representing the function performed by the multiplier $Z=A\cdot B$.
@@
 and abstraction theory to implement verifications on sequential circuits. 
-These approaches are described in the following chapters.
+The application of these approaches to sequential circuit verification is described in the following chapters.

### 2016-12-11 dissertation/prev.tex
dissertation v3:milestone

@@
 where $f_x = f(x = 1)$ and $f_{x'} = f(x = 0)$ denote the positive and
-negative co-factors of $f$ w.r.t. $x$, respectively.
+negative co-factors of $f$ {\it w.r.t.} $x$, respectively.
 A BDD is usually represented as a binary tree.
@@
 \centerline{
-\includegraphics[width=0.35\textwidth]{newfig/BDD.pdf}
+\includegraphics[width=0.45\textwidth]{newfig/BDD.pdf}
 }
@@
 \cite{devadas:algebraic_manipulation_iccd91}, or the ones based on
-term re-writing \cite{AST}, etc.
+term re-writing \cite{AST}, {\it etc.}
 
@@
 diagrams \cite{interval_dd}, interval analysis using polynomials
-\cite{polynomial_sanchez99}, etc. Most of these have found 
+\cite{polynomial_sanchez99}, {\it etc.} Most of these have found 
 application in constraint satisfaction for simulation-based
@@
 $\Fkk$. They show that this can be achieved by computing a reduced 
-\Grobner basis w.r.t. an {\it abstraction term order} derived from the
+\Grobner basis {\it w.r.t.} an {\it abstraction term order} derived from the
 circuit. Subsequently, they propose a \underline{r}efinement of this
@@
 the concepts of word-level reachability, abstraction-refinement using
-interpolation or UNSAT cores, etc., that we desire was not the focus of
+interpolation or UNSAT cores, {\it etc.}, that we desire was not the focus of
 \cite{Avrunin:CAV,vardi-iasted07}.

### 2016-12-11 dissertation/recha.tex
dissertation v3:milestone

@@
-\chapter{Word-level Traversal of Finite State Machines using Algebraic Geometry}
+\chapter{Word-Level Traversal of Finite State Machines using Algebraic Geometry}
 \label{ch:reacha}
@@
 density, sequential optimization techniques such as retiming \cite{retiming}, scan logic \cite{scan},
-sequential power optimization \cite{mathur2009power} and clock-gating techniques \cite{clockgating}.
+sequential power optimization \cite{mathur2009power} and clock-gating techniques \cite{clockgating} are applied.
 % some major refactoring and attachments are made on original designs. 
-These modifications may introduce malfunctions to the original logic and cause problems. 
-Based on traditional localized simulation or formal verification method (e.g. equivalence checking), 
+These modifications may introduce, bugs, errors or malfunctions to the original logic and cause problems. 
+Based on traditional localized simulation or formal verification method ({\it e.g.} equivalence checking), 
 designers are reluctant to make aggressive optimization
-since the malfunctions are considered as ``faults" in this circuit.
+since the malfunctions are considered as ``faults" in the circuits.
 However, if the circuit behavior is carefully investigated, it may become evident that 
-those ``faults" will never be activated/excitated during a restricted execution starting 
+those ``faults" may never be activated during a restricted execution starting 
 from legal initial states and with legal inputs. Thus we will call those ``faults" as 
-{\bf spurious faults} (false negatives), since they will not affect the circuit's normal behavior.
+{\it spurious faults} (false negatives), since they will not affect the circuit's normal behavior.
 
@@
 As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
-spurious faults by checking if they belong to the reachable states. From the spurious fault validation 
+spurious faults by checking if they belong to the unreachable states. From the spurious fault validation 
 perspective, reachability analysis is a must when developing full set of sequential circuit verification
@@
 There are quite a few methods to perform reachability checking on FSMs. One among them is 
-state space traversal. Conventionally the algorithm is based on bit-level techniques such as
+state space traversal. Conventionally, the algorithm is based on bit-level techniques such as
 binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
@@
 \section{Motivation}
-Before introducing the details of our approach, there are a couple of questions to ask: why is it
-necessary to execute word-level FSM traversal? Why does algebraic geometry make this happen?
-The answers can be found in this section, as statement of the motivation of our research.
+Before introducing the details of our approach, there are a few questions to ask: what is 
+the benefit of execute FSM traversal at word-level? How can algebraic geometry make this happen?
+The answers can be found in this section, as a statement of the motivation of our research.
 
 \subsection{FSM Traversal Algorithms}
-Sequential circuits are modeled as FSMs, which can be implemented as graph. Thus a 
+Sequential circuits are modeled as FSMs, which can be implemented as graphs. Thus a 
 graph-traversal based algorithm is created to analysis the reachable states \cite{coudert2003unified}.
-A traversal algorithm using concept ``implicit state enumeration" is proposed \cite{cho1993redundancy}.
-Concretely, the algorithm is written as follows:
-
-\begin{algorithm}[H]
+A traversal algorithm using the concept of implicit state enumeration is proposed \cite{cho1993redundancy}.
+Concretely, the algorithm is given in Algorithm \ref{alg:BFS}:
+
+% \begin{algorithm}[H]
+% \SetAlgoNoLine
+% \LinesNumbered
+%  \KwIn{Sequential circuit with given number of registers}
+%  \KwOut{Scan registers listed in decreasing order of their non-controllability}
+% 
+%   $from^0 = reached = S^0$\;
+%   $i = 0$\;
+%   \While{TRUE}
+%   {
+%   	$i++$\;
+% 	$to^i = $IMAGE$(\Delta,from^{i-1})$\;
+% 	$new^i = to^i \cdot \overline{reached}$\;
+% 	\For{each state variable $r_j$}
+% 	{
+% 		record\_if\_transitions\_present\_or\_missing$(r_j,new^i)$\;
+% 		compute\_degree\_of\_unsettability$(r_j,new^i)$\;
+% 	}
+% 	\If{$new^i == 0$}
+% 	{
+% 		break\;
+% 	}
+% 	$reached = reached + new^i$\;
+% 	$from^i = new^i$\;
+%   }\
+% \Return{$reached$}
+% \caption {FSM Traversal using Implicit State Enumeration\cite{KallaPartialScan}}
+% \label{alg:SIMPSON}
+% \end{algorithm}
+% \DecMargin{1em}
+
+\IncMargin{1em}
+\begin{algorithm}[hbt]
 \SetAlgoNoLine
 \LinesNumbered
- \KwIn{Sequential circuit with given number of registers}
- \KwOut{Scan registers listed in decreasing order of their non-controllability}
+\Indm
+ \KwIn{Transition functions $\Delta$, initial state $S^0$}
+\Indp
 
   $from^0 = reached = S^0$\;
-  $i = 0$\;
-  \While{TRUE}
+  \Repeat{$new^i == 0$}
   {
-  	$i++$\;
-	$to^i = $IMAGE$(\Delta,from^{i-1})$\;
-	$new^i = to^i \cdot \overline{reached}$\;
-	\For{each state variable $r_j$}
-	{
-		record\_if\_transitions\_present\_or\_missing$(r_j,new^i)$\;
-		compute\_degree\_of\_unsettability$(r_j,new^i)$\;
+  	$i \gets i + 1$\;
+	$to^i \gets$Img$(\Delta, from^{i-1})$\;
+	$new^i \gets to^i \cap \overline{reached}$\;
+  	$reached \gets reached \cup new^i$\;
+	$from^i \gets new^i$\;
   }
-	\If{$new^i == 0$}
-	{
-		break\;
-	}
-	$reached = reached + new^i$\;
-	$from^i = new^i$\;
-  }\
 \Return{$reached$}
-\caption {FSM Traversal using Implicit State Enumeration\cite{KallaPartialScan}}
-\label{alg:SIMPSON}
+\caption {BFS Traversal for FSM Reachability}\label{alg:BFS}
 \end{algorithm}
@@
 function for one-step reachability within current time-frame, and extends the result to complete 
-execution paths through unrolling. If each state encoding (i.e. exact values in the selected registers) is
+execution paths through unrolling. If each state encoding ({\it i.e.} exact values in the selected registers) is
 explicitly analyzed and counted during the unrolling 
@@
 
-\subsection{Word-level Data Flow on Modern Datapath Designs}
+\subsection{Word-Level Data Flow on Modern Datapath Designs}
 The level of integration of modern digital circuit designs is very high. For example, processor A10 designed by Apple
@@
 \SetAlgoNoLine
+\LinesNumbered
 \Indm
@@
 carry-in will become longer and longer, the final conjunction of all clauses will contain 
-$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables
+$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables:
 \begin{equation}
@@
 
-Since the abstraction to word-level reduces symbolic storage and execution cost, the complexity 
-to traverse the state space is greatly reduced.
+The abstraction to word-level will reduce symbolic storage and execution cost; the complexity 
+to traverse the state space will be greatly reduced.
 \end{Example}
@@
 any more for users of conventional model checker. The reason is that the user has to write all clauses 
-for the implementation, which contains cross-literals, e.g. $s_2$ may associate with $a_1, p_0$, etc. If the user 
+for the implementation, which contains cross-literals, {\it e.g.} $s_2$ may associate with $a_1, p_0$, {\it etc.} If the user 
 is not familiar with the implementation of this adder, those cross-literals will bring
@@
 
-\subsection{On the Existence of Word-level Abstraction}
+\subsection{On the Existence of Word-Level Abstraction for Arbitrary Circuits}
 When given a bit-level netlist, the prerequisites to use word-level techniques are to convert bit-level 
@@
 
-An old but universally effective abstraction method is {\it Lagrange's interpolation}. Instead of using 
-it in real algebra, it can also be extended to Galois field. Here we use an example to illustrate the 
-conversion in Galois field using Lagrange's interpolation.
+An old but universally effective abstraction method is {\it Lagrange's interpolation}, 
+which can be applied over finite fields. Here we use an example to illustrate the 
+conversion in Galois field using Lagrange's interpolation for an arbitrary circuit.
 
@@
 \end{table}
-Now our objective is to abstract a function over GF $\F_{2^3}$ about word-level variables, i.e. $Z = \Func(A)$.
-Recall the definition of Lagrange's interpolation:
+Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
+$Z = \Func(A)$. Recall Lagrange's interpolation formula:
 \begin{equation}
@@
 In practice, Lagrange's interpolation is not scalable. 
-The reason is that it needs entire function (state space), but usually we only have the circuit representation
+The reason is that it needs the entire function (state space), but usually we only have the circuit representation
 of the FSM. Considering this fact, a symbolic method is needed.
@@
 
-\subsection{Significance of Developing Word-level FSM Traversal}
-At the end of this section, we summarize each subsection and conclude the motivation of our research.
-First, we start from the basic FSM traversal algorithms and exhibit their
-difficulties to overcome. Secondly, we state the observation that word-level description is 
-increasingly important and common in characterizing the data flow on modern large-size datapaths. 
-Last but not least, we give instances that some prerequisites for supporting word-level techniques 
-are already available.
-
-To conclude, first we state the importance of performing FSM traversal in sequential circuit reachability analysis.
-To overcome the vast cost brought by searching in a large space, we propose to use word-level 
+\subsection{Significance of Developing Word-Level Reachability Analysis}
+% At the end of this section, we summarize each subsection and conclude the motivation of our research.
+% First, we start from the basic FSM traversal algorithms and exhibit their
+% difficulties to overcome. Secondly, we state the observation that word-level description is 
+% increasingly important and common in characterizing the data flow on modern large-size datapaths. 
+% Last but not least, we give instances that some prerequisites for supporting word-level techniques 
+% are already available.
+
+Based on the aforementioned discussions, the importance of performing FSM traversal at word-level 
+can bring a dimension of abstraction in sequential circuit reachability analysis.
+To overcome the cost incurred by searching in a large space, we propose to use word-level 
 polynomials to represent the states and transition relations. As a result, states are categorized into 
 sets represented by a small number of word-level polynomials (more specifically, the varieties to the polynomial ideals),
-and multiple transition relations are therefore merged together. All of these efforts reduces the space cost of 
-state space, meanwhile lower the time complexity to traverse such a state space. By using Lagrange's interpolation,
-we prove the feasibility of word-level techniques on general circuits' verification.
-Thus we cover both the necessity and sufficiency of developing such 
-a word-level traversal technique in our work.
-They are also answers to the questions at the beginning of this section.
+and multiple transition relations are therefore merged together. All of these efforts reduce the cost of 
+state space, meanwhile lowering the time complexity to traverse such a state space. As Lagrange's interpolation
+confirms the existence of word-level abstraction for bit-level circuits, the concept
+can be extended to encode the state space of sequential circuit by word-level polynomials in $\Fkk$.
+% we prove the feasibility of word-level techniques on general circuits' verification.
+% Thus we cover both the necessity and sufficiency of developing such 
+% a word-level traversal technique in our work.
+% They are also answers to the questions at the beginning of this section.
 
@@
 
-\subsection{FSM model for sequential circuits}
+\subsection{FSM Model for Sequential Circuits}
 A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
@@
 The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
-its primary outputs only depend on the present states, i.e. the output function is defined as
+its primary outputs only depend on the present states, {\it i.e.} the output function is defined as
 $$\Lambda:\ S \to O$$
@@
 %\begin{minipage}{12cm}
-\includegraphics[width=5.5in]{newfig/seqmodel.eps}
+\includegraphics[width=5.5in]{newfig/seqmodel.pdf}
 % \vspace{-0.2in}
@@
 Here, $\exists x (f)$ represents the {\it existential quantification
-  of $f$ w.r.t. variable $x$}. In Boolean logic, this operator is implemented as
+  of $f$ {\it w.r.t.} variable $x$}. In Boolean logic, this operator is implemented as
   $$\exists x (f) = f_x\lor f_{\overline{x}}$$
-
-\IncMargin{1em}
-\begin{algorithm}[hbt]
-\SetAlgoNoLine
-\Indm
- \KwIn{Transition functions $\Delta$, initial state $S^0$}
-\Indp
-
-  $from^0 = reached = S^0$\;
-  \Repeat{$new^i == 0$}
-  {
-  	$i \gets i + 1$\;
-	$to^i \gets$Img$(\Delta, from^{i-1})$\;
-	$new^i \gets to^i \cap \overline{reached}$\;
-  	$reached \gets reached \cup new^i$\;
-	$from^i \gets new^i$\;
-  }
-\Return{$reached$}
-\caption {BFS Traversal for FSM Reachability}\label{alg:BFS}
-\end{algorithm}
-\DecMargin{1em}
 
@@
 
-\subsection{FSM Traversal at word-level over $\Fkk$} 
+\subsection{FSM Traversal at Word-Level over $\Fkk$} 
 The state transition graph (STG) shown in Figure \ref{fig:fsm}(a) uses a
@@
 
-{\it Initial state:} In Line 1 of Algorithm \ref{alg:BFS}, the initial state is
-specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
+{\it Initial state:} Line 1 of Algorithm \ref{alg:BFS} specifies the initial state.
+In algebraic geometry, it can be specified by means of a corresponding polynomial
+ $f = \mathcal{F}(S) =
 S - 1 - \alpha$. Notice that if we consider the ideal generated by the
@@
 Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$
-i.e. the variety of vanishing ideal contains
+{\it i.e.} the variety of vanishing ideal contains
  all possible valuations of variables, so it constitutes the {\bf
@@
 variables are abstracted and quantified away.
+
+\section{Problem Setup and Formulation}
+\begin{Problem}
+We use the notions from Figure \ref{fig:seqmodel}(a) in this setup. 
+\begin{enumerate}[{1)}] 
+\item The circuit is modeled over $\Fkk = \F_2[x]\pmod{p(x)}, p(\alpha) = 0$, where $k$ is the number of flip-flops,
+	the PS variables $\{s_0,dots,s_{k-1}\}$, NS variables $\{t_0,\dots,t_{k-1}\}$.
+\item Denote $S$ as the PS word-level by the following polynomial:
+$$f_S: S = s_0+s_1\alpha+s_2\alpha^2+\cdots+s_{k-1}\alpha^{k-1}$$
+Similarly we define NS word-level variable $T$ by:
+$$f_T: T = t_0+t_1\alpha+t_2\alpha^2+\cdots+t_{k-1}\alpha^{k-1}$$
+\item Impose ATO for sequential circuits which is 
+$$LEX: \text{ all bit-level variables in any order }>S>T$$
+\item Write polynomials $f_1,\dots,f_s$ for each gate, and construct ideal describing the circuit:
+$$J_{ckt} = \langle f_1,\dots,f_s,f_S,f_T\rangle$$
+as well as the ideal with vanishing polynomials:
+$$J_0 = \langle \dots,x_i^2-x_i,\dots,S^{2^k}-S,T^{2^k}-T\rangle$$
+where $x_i$ corresponds to bit-level variables denoting wires in the circuit.
+\item Compute GB with ATO for $G$. Then obtain the projection of variety only on NS variable $T$, by
+$$G\cap \Fkk[T]$$
+\end{enumerate}
+\end{Problem}
 
@@
 
-\subsection{Word-level FSM Traversal Example}
-
+\subsection{Word-Level FSM Traversal Example}
 \begin{Example}
@@
 \end{align*}
- i.e. states $\{00,01,10\}$. We update the present states
+ {\it i.e.} states $\{00,01,10\}$. We update the present states
 for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha.$$
@@
 $\{00,10\}$. The complement of $reached$ is $$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
-= \langle T + 1+\alpha\rangle$$ (i.e. states $\{11\}$). We compute the newly reached state 
+= \langle T + 1+\alpha\rangle$$ ({\it i.e.} states $\{11\}$). We compute the newly reached state 
 $$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$$
@@
 Since the GB contains the unit ideal, it means the newly reached state
-set is empty, thus a fixpoint has been reached. The algorithm
+set is empty, thus a fix-point has been reached. The algorithm
 terminates and returns $$reached = \langle T^3+(\alpha+1)T^2+\alpha
@@
 
-\subsection{Significance of using Algebraic Geometry and Gr\"obner Basis}
-The essences of out approach consist of utilizations of algebraic geometry and Gr\"obner basis concepts.
+\subsection{Significance of using Algebraic Geometry and Gr\"obner Bases}
+The essence of our approach is based on algebraic geometry and Gr\"obner basis concepts.
 One the one hand, we use GB with elimination term ordering as an analog of image function.
-As mentioned in Section \ref{sec:abstraction}, we construct the elimination ideal $J+J_0$ to 
+As mentioned in Section \ref{sec:abstraction}, we construct the ideal $J+J_0$ to 
 describe the circuit in Figure \ref{fig:proj_reacha} using algebraic geometry.
-When given the present states, the next states are implicitly represented in the variety of 
-the elimination ideal. Imposing elimination term order and computing GB, is actually 
-projecting the variety on NS output $T$ by eliminating all other variables.
-This projection gives us the canonical representation of NS in a polynomial with $T$.
+Given the present states, the next states are implicitly represented in the variety of 
+the elimination ideal obtained by quantify away the remaining variables. 
+This projects the variety on the NS output $T$ by eliminating all other variables.
+This projection gives us the canonical representation of NS in a polynomial in $T$.
 
@@
 
-On the other hand, we use ideal arithmetic to implement set operations in Boolean domain.
-In algebraic geometry, manipulating ideals provides a high-level solution to 
-operate on their varieties without actually solving the system. The intersection, union and 
+On the other hand, we use the algebraic ideals to implement set operations. States are finite set of points,
+which can be mapped to a variety.
+In algebraic geometry, manipulating the ideals provides a mechanism to 
+operate on the varieties without actually solving the system of equations. The intersection, union and 
 complement of varieties are mapped to varieties of the sum, product and quotient of ideals, respectively.
-Furthermore, the varieties are equivalent to set of states because of correspondence of 
-affine space and state space.
-
-As a result, we find an analog of the FSM traversal algorithm in algebraic geometry which is 
-compatible with word-level variables (e.g. $S,T$). We show it is effective to perform the 
+% Furthermore, the varieties are equivalent to set of states because of correspondence of 
+% affine space and state space.
+
+As a result, we create the analog of the FSM traversal algorithm in algebraic geometry which is 
+compatible with word-level variables ({\it e.g.} $S,T$). We show it is effective to perform the 
 reachability analysis.
@@
 \label{sec:improve}
-% TODO Revised according to comments in next chapter
 
@@
 Buchberger's algorithm consists of two major operations: one is finding 
-S-poly pairs, the other is dividing the S-poly by the set of polynomials.
+$Spoly$ pairs, the other is dividing the $Spoly$ by the set of polynomials.
 Its actual cost is very sensitive to the term ordering: if there exists 
-a term order making most S-poly computation unnecessary, then the cost to 
-compute S-poly is saved. Moreover, it prevents new polynomials generating 
-from unnecessary S-poly reductions, which further reduces the cost
-because there are less S-poly pairs as well as candidate divisors.
+a term order making most $Spoly$ computation unnecessary, then the cost to 
+compute $Spoly$ is saved. Moreover, it prevents the generation of new polynomials 
+from unnecessary $Spoly$ reductions, which further reduces the cost
+because there are less $Spoly$ pairs as well as candidate divisors.
 Besides tuning the term ordering, directly cutting down the number of 
@@
 In order to make our approach scalable, we propose improvements from 
-two aspects: on the one hand, using another term ordering which can lower down the 
-computational complexity of obtaining GB; on the other hand, reducing the 
+two aspects: 1) using another term ordering which can lower the 
+computational complexity of obtaining GB; and 2) reducing the number of
 polynomials by collecting bit-level primary inputs (PIs) and integrating them as word-level 
@@
 
-\subsection{Simplifying the Gr\"obner Basis Computation}
+\subsection{Simplifying the Gr\"obner Bases Computation}
 In Algorithm \ref{alg:univa}, a Gr\"obner basis is computed for each
@@
 we describe a method that computes a GB of a smaller subset of polynomials.
-The approach draws inspirations from \cite{pruss:tcad15}. 
-According to proposition 2 in \cite{pruss:tcad15}, if the GB is computed using {\it refined abstraction term
-order} (RATO), there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
-their leading monomials are not relatively prime, i.e. 
+The approach draws inspirations from \cite{pruss:tcad15}, which defined and justified a 
+{\it refined abstraction term order} (RATO). The following definition rephrases Definition 5.1 in \cite{pruss:tcad15}
+with our sequential circuit setup and notations.
+
+% From Tim's journal
+\begin{Definition}[Refined Abstraction Term Order $>_r$]
+Starting from the pseudo outputs (NS variables) of the combinational component of a sequential circuit $C$, 
+perform a reverse topological 
+traversal toward the pseudo inputs (PS variables) and primary inputs. Order each variable of the circuit according to
+its reverse topological order. Impose a LEX term order $>_r$ on $\Fq[x_1,\dots,x_d,T,S]$
+with the ``bit-level variables $x_1,\dots,x_d$ ordered reverse topologically "$ > T>S$,
+this term order is called RATO.
+\end{Definition}
+
+According to proposition 5.1 in \cite{pruss:tcad15}, if the GB is computed using RATO, 
+there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
+their leading monomials are not relatively prime, {\it i.e.} 
 $$gcd(LM(f_w), LM(f_g)) \neq 1$$
 As a well-known fact from Buchberger's algorithm, the S-polynomial 
-(Spoly) pairs with relatively prime 
+($Spoly$) pairs with relatively prime 
 leading monomials will always reduce to 0 modulo the basis and have no contribution to 
 the Gr\"obner basis computation.
-Therefore, by removing the relatively prime polynomials from $J_{ckt}$, 
+Therefore, by removing the polynomials with relatively prime leading terms from $J_{ckt}$, 
 the Gr\"obner basis computation is transformed to the reduction of 
@@
 \begin{Example}
-After adding intermediate bit-level signal $a,b,c,d$, the elimination ideal for example circuit 
-(Example \ref{ex:SMPO}) can be rewritten LEX order with RATO:
+In this example, we impose RATO on the polynomial ideal generated from the circuit in Figure \ref{fig:fsm}(b).
+We start from the outputs $t_0,t_1$, then intermediate bit-level signals $a,b,c,d$ because they are 
+the fanins of the corresponding gates which fanout $t_0,t_1$. Then we ends at the pseudo inputs $s_0,s_1$
+and primary input $x$.
+Thus variables in $J_{ckt}$ can be ordered by LEX with:
 \begin{align}
-&(t_0,t_1)>(a,b,c,d)\nonumber\\&>T>(x,s_0,s_1)\nonumber
+&(t_0,t_1)>(a,b,c,d)>(x,s_0,s_1)\nonumber\\&>T>S\nonumber
 \end{align}
+This is the RATO for circuit in Figure \ref{fig:fsm}(b).
+
 We can write down all polynomial generators of $J_{ckt}$:
@@
 $(f_5,f_7)$, thus the critical candidate polynomial pair is
-$(f_w,f_g)$, where $$f_w = t_0+a\cdot b+a+b, ~~f_g =t_0+t_1\alpha + T$$
+$(f_w,f_g)$, where $$f_w =f_5= t_0+a\cdot b+a+b, ~~f_g =f_7=t_0+t_1\alpha + T$$
 Result after reduction is:
@@
 
-The remainder from \emph{Spoly} reduction contains bit-level PS variables, and our objective is to get a polynomial 
+The remainder from $Spoly$ reduction contains bit-level PS variables, and our objective is to get a polynomial 
 containing only word-level PS variables. One possible method is 
-to rewrite bit-level variables in term of word-level variables, i.e.
+to rewrite bit-level variables in term of word-level variables, {\it i.e.}
 \begin{equation}
@@
 The authors of \cite{pruss:tcad15} propose a method to construct 
-a system of equations and solution to the system consists of Equation \ref{eqn:BLVS_reacha}.
-This can be obtained by Gaussian elimination, which could 
+a system of equations, such that the solution to the system consists of Equation \ref{eqn:BLVS_reacha}.
+It relies on a lemma which can be derived from Fermat's Little Theorem:
+\begin{Lemma}
+\label{lem:Fermat}
+For elements $\alpha_i$ in $\Fkk$, the following equation holds ($n\geq 0$):
+$$(\alpha_1+\alpha_2+\cdots+\alpha_t)^{2^n} = \alpha_1^{2^n}+\alpha_2^{2^n}+\cdots+\alpha_t^{2^n}$$
+\end{Lemma}
+
+Solution to the system of equations can be obtained by Gaussian elimination, which could 
 compute corresponding $\mathcal{G}(S)$ efficiently with time complexity $O(k^3)$.
@@
   $f_T = $Reduce($Spoly(f_w,f_g), J_{ckt})$\;
-  \tcc{Compute S-poly for the critical pair, then reduce it with circuit ideal under RATO}
+  \tcc{Compute $Spoly$ for the critical pair, then reduce it with circuit ideal under RATO}
 	Eliminate bit-level variables in $f_T$\;
@@
   }
-  \tcc{Loop until a fixpoint reached: newly reached state is empty}
+  \tcc{Loop until a fix-point reached: newly reached state is empty}
 \Return{$\langle reached\rangle$}
@@
 
-\subsection{PI Partition}
+\subsection{Primary Inputs Partition}
 Using above techniques we can get a remainder polynomial with only word-level PS/NS variables. However in most 
@@
 $P^{2^{n_i}}-P$ to represent all signals at free-end PIs, according to following theorem about {\bf composite
-field}:
+fields} \cite{ecc:software}:
 \begin{Theorem}
@@
 
-\section{Implementation of Word-level FSM Traversal Algorithm}
-In this section, we described the architecture of our tool which can perform word-level FSM traversal
-on FSM benchmark circuits.
-Briefly speaking, our tool consists of 3 functional components.
+\section{Implementation of Word-Level FSM Traversal Algorithm}
+In this section, we describe the architecture of our tool which can perform word-level FSM traversal
+on FSM benchmark circuits. Our tool consists of 3 functional components.
 % First, the given circuit is synthesized to gate-level using state-of-art synthesizers.
 First, the gate-level netlist of circuit is translated to polynomial form and variables are sorted in 
-RATO. This part is implemented using scripting language such as \emph{Perl}. If 
-the given benchmark is a structural/behavior hybrid description, we perform a pre-process on it to get a synthesized netlist.
+RATO. This part is implemented using a scripting language such as \emph{Perl}. If 
+the given benchmark is a structural/behavior hybrid description, we perform pre-processing on it to get a synthesized netlist.
 Secondly, the polynomial reduction is executed using our customized reduction engine, which is written in C++.
-Finally, we utilize symbolic computation engine -- \textsc{Singular} \cite{DGPS} to code Algorithm 
+Finally, we utilize the symbolic computation engine \textsc{Singular} \cite{DGPS} to code Algorithm 
 \ref{alg:refined} and execute the BFS traversal. The tool outputs a univariate polynomial 
@@
 Usually FSM designs can be described in behavior/structural hybrid languages.
-One of these languages is Berkeley logic interchange format (BLIF) \cite{BLIF},
+One of these languages is the Berkeley logic interchange format (BLIF) \cite{BLIF},
 it allows state behavior representation and logic component representation.
@@
 In this example, we use benchmark FSM ``lion9" from MCNC benchmark library.
-This benchmark circuit is given as truth-table based structural BLIF representation of a FSM.
+This benchmark circuit is given as state table based BLIF representation of a FSM.
 
@@
 
-In order to compose the polynomial set in elimination ideal, we need to synthesize it to 
+In order to compose the polynomial set in elimination ideal, we need to synthesize it to a 
 gate-level netlist. Modern synthesizers including ABC \cite{brayton2010abc} and SIS \cite{SIS} 
-can complete this work. In this example we use a synthesis library containing only 2-input 
-AND, NAND, OR, NOR, XOR, XNOR gates as well as the inverter, the output synthesized FSM is also given in 
+can perform this task. In this example we use a synthesis library containing only 2-input 
+AND, NAND, OR, NOR, XOR, XNOR gates as well as the inverter, the synthesized FSM is also given in 
 BLIF format.
@@
 
-Using our interpreter, the synthesized BLIF file turns to a polynomial file customized for our 
+Using our interpreter, the synthesized BLIF file is translated to a polynomial file customized for our 
 polynomial reduction engine. 
 The input file format includes all 
-variables in RATO, along with the S-poly need to be reduced and polynomials in $J_{ckt}$.
+variables in RATO, along with the $Spoly$ that is need to be reduced, and polynomials in $J_{ckt}$.
 % 
@@
 
-The result given by our reduction engine is in the format 
+The result given by our reduction engine is in the form of the polynomial
 $$T+\Func(s_0,\dots,s_{k-1},x_0,\dots,x_{n-1})$$ 
-where $s_i$ and $x_j$ denotes bit-level PS variables and PIs. Concretely in this example, the result is
+where $s_i$ and $x_j$ denote bit-level PS variables and PIs. Concretely in this example, the result is
 \begin{align*}
@@
 \end{align*}
-This is the transition function of this FSM. We utilized \textsc{Singular} to integrat both 
-bit-word substitution and tranversal algorithm. In Figure \ref{fig:SING}, ``tran" is the transition
+This is the transition function of this FSM. We utilized \textsc{Singular} to integrate both 
+bit-to-word substitution and the traversal algorithm. In Figure \ref{fig:SING}, ``tran" is the transition
 function we just obtained. ``init\_S" is the initial state, note it equals to ``0100" register 
@@
 
-After the script is executed, the trversal finishes after 4 
+After the script is executed, the traversal finishes after 4 
 transition iterations, which denotes BFS depth equals to 4. The final reachable states 
-is a degree-9 polynomial about $T$, indicating final reachable states set contains 9 states.
+is a degree-9 polynomial in $T$, indicating final reachable states set contains 9 states.
 And state encodings can be obtained by solving this polynomial equation.
@@
  and ``bbara, beecount, dk14, donfile" are from MCNC benchmarks. %\cite{yang1991logic}.
-ISCAS benchmarks are given as {\it bench} format so we can directly read gate information,
-where ITC/MCNC FSMs are given in unsynthesized {\it blif} format so we first turn them
+ISCAS benchmarks are given in {\it bench} format so we can directly read gate information,
+where ITC/MCNC FSMs are given in unsynthesized {\it BLIF} format so we first turn them
 into gate-level netlists using AIG based synthesizer ABC. % \cite{brayton2010abc}.
@@
 demonstrates the power of abstraction of algebraic geometry techniques
-for reachability analysis applications. Currently, we are
-investigating techniques that
-can help us overcome the complexity of the GB computation with elimination
-orders and speed-up our approach.
+for reachability analysis applications. 
+% Currently, we are investigating techniques that
+% can help us overcome the complexity of the GB computation with elimination
+% orders and speed-up our approach.
 
@@
 
-\section{Conclusion} \label{sec:conc}
+\section{Concluding Remarks} \label{sec:conc}
 This dissertation has presented a new approach to perform reachability
@@
 variety of a univariate polynomial. This polynomial is computed using
-the Gr\"obner basis algorithm w.r.t. an elimination term
+the Gr\"obner basis algorithm {\it w.r.t.} an elimination term
 order. Experiments are conducted with a few FSMs that validate the

### 2016-12-11 dissertation/unsat.tex
dissertation v3:milestone

@@
 \label{ch:UNSAT}
-Gr\"obner basis can find further applications to a vast field
-of other techniques in formal verification. Satisfiability (SAT) problem is the basis of 
+In previous chapters, we have introduced the concept of word-level abstraction and 
+reachability analysis of the state-space of sequential circuits for equivalence checking.
+Modern property verification techniques employ state-space abstraction using interpolants and UNSAT 
+cores. In this chapter, we explore algebraic geometry analogs of UNSAT cores which can be 
+applied at both bit and word-level, actually over any field.
+
+Boolean satisfiability (SAT) problem is the basis of 
 most decidable decision problems, as well as the basis of many formal verification techniques. 
-In this chapter, we discuss a special topic branching out from SAT theory. They are about a situation 
+In this chapter, we discuss a special topic branching out from SAT theory. It is about a situation 
 when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
-a set of constrains (e.g. clauses, formulas or polynomials) which is unsatisfiable, sometimes 
+a set of constrains ({\it e.g.} clauses, formulas or polynomials) which is unsatisfiable, sometimes 
 it is worthy to explore the reasons for UNSAT. From the execution of the GB algorithm, 
@@
 In this section, we introduce the motivation of our UNSAT core extraction research. We
-start with defining an UNSAT problem, then review the previous work and applications 
-to abstraction refinement. Moreover, we then present the need of finding a word-level 
+start by defining an UNSAT problem, then review the previous work and applications 
+to abstraction refinement. We then present the need for finding a word-level 
 analog to contemporary UNSAT reasoning techniques.
@@
 \subsection{Preliminaries of SAT/UNSAT Theory}
-Boolean satisfiability (SAT) problem is one of most basic problems in computer science.
+Boolean satisfiability (SAT) problem is a fundamental problem in computer science.
 In the following part we define the terminology related to SAT problem \cite{anaICCAD}.
@@
 A {\bf literal} $l$ is defined as a variable $v$, or its negation $\overline{v}$. A disjunction (OR relation)
-of literals forms a {\bf clause}, i.e. $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
+of literals forms a {\bf clause}, {\it i.e}. $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
 A Boolean formula can always be written as {\bf conjunctive normal form} (CNF),
@@
 A {\bf satisfiability (SAT) problem} is a decision problem that takes a CNF formula and returns
-that the formula is satisfiable (SAT), when there is an assignment of the variables making the 
-formula evaluate to true. Otherwise, the formula is {\bf unsatisfiable (UNSAT)}.
+that the formula is SAT, whenever there is an assignment of variables that makes the 
+formula evaluate to true. Otherwise, the formula is unsatisfiable (UNSAT).
 \end{Definition}
 
-Figure \ref{fig:SAT} shows a simple example of SAT problem on circuits verification:
-we need to verify whether subcircuit $A$ and subcircuit $B$ has the same function, so
+Figure \ref{fig:SAT} shows a simple example of SAT problem on circuit verification:
+we need to verify whether subcircuit $A$ and subcircuit $B$ have the same function, so
 we build a miter circuit for their outputs $X$ and $Y$, and the equivalence checking 
@@
 
+\vspace{-0.5in}
 \begin{center}
@@
 &\Longleftrightarrow
-\textit{Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ primary\ inputs\ a,b,c\ such\ that\ Z=1?}
+\textit{Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ PIs\ a,b,c\ exists\ such\ that\ Z=1?}
 \end{align*}
@@
 
-For an UNSAT problem, the cause of UNSAT may include a part of clauses.
+For an UNSAT problem, the cause of UNSAT may include a subset of clauses.
 \begin{Definition}
@@
 clause learning (CDCL) technique to prune the search space, which is effective to reduce
-result search time.
+search time.
 
 When a SAT solver fails to give a SAT assignment, it will provide a {\it UNSAT proof} or {\it refutation proof}
-to prove the problem is UNSAT. By analyzing clauses involved to the UNSAT proof, we can 
-generate a subformula which remains to be UNSAT. A n\"aive method is to collect all leaf 
+to prove the problem is UNSAT. By analyzing clauses involved in the UNSAT proof, we can 
+generate a subformula which remains UNSAT. A n\"aive method is to collect all leaf 
 clauses in an UNSAT proof as an UNSAT core. In practice, a minimal UNSAT core is more valuable.
-There are mainly 2 kinds of method to find minimal UNSAT cores.
-The one is insertion-based method, which is achieved by adding clauses to the smallest subset 
+There are mainly two kinds of methods to find minimal UNSAT cores.
+One is insertion-based method, which is achieved by adding clauses to the smallest subset 
 until the subset turns to be UNSAT. The other method is deletion-based method,
@@
 algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
-valid assignments or proof of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
+valid assignments or proofs of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
 from computer algebra.
@@
 
-The inspiration that use PC to solve SAT problems first comes from \cite{ceiSTOC96}.
-Using PC and its variations, researchers develop many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
+The inspiration to use PC to solve SAT problems first came from \cite{ceiSTOC96}.
+Using PC and its variations, researchers developed many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
 Besides, researchers borrow concepts from PC, combine them with traditional DPLL and clause learning techniques
@@
 
-\subsection{Exploiting UNSAT cores for abstraction refinement}
-Finding small UNSAT cores attracts interest for decades because of the need for all kinds of applications.
-In solving MaxSAT problems, small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
-It is applied to solving Boolean algebra problems, such as Boolean function decomposition \cite{lee2008bi}. 
+\subsection{Exploiting UNSAT Cores for Abstraction Refinement}
+The problem of finding small UNSAT cores has attracted interest for decades because of 
+its applications to various of verification and synthesis problems.
+In solving MaxSAT problems, a small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
+It is applied to solving logic synthesis problems, such as Boolean function decomposition \cite{lee2008bi}. 
 Small explanation generation in general constraint programming problems also rely on small UNSAT cores \cite{cambazard2008reformulating}.
@@
 assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
-without any fixpoint detection.
+without any fix-point detection.
 In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
@@
 The reason is as follows:
-\begin{Theorem}
-{\bf Weak Nullstellensatz:} Given ideal $J\subset \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure
+\begin{Theorem}[Weak Nullstellensatz]
+Given ideal $J\subseteq \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure $\overline{\F}$
 of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
@@
 \label{sec:core}
-In this section we provide a solution to the problem that how to use 
+In this section we provide a solution to the problem of how to use 
 GB to extract the UNSAT core.
@@
 too. Borrowing the terminology from the Boolean SAT domain, we
-call $F_c$ the infeasible core or the unsat core of $F$.
+call $F_c$ the infeasible core or the UNSAT core of $F$.
 
@@
 %% disregarding minimality. 
-It is not hard to motivate that an unsat core should be identifiable
+It is not hard to motivate that an UNSAT core should be identifiable
 using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
 $GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
-the ideal generated by $(F - \{f_j\})$, i.e. $f_j \in \langle F -
+the ideal generated by $(F - \{f_j\})$, {\it i.e.} $f_j \in \langle F -
 \{f_j\}\rangle$. Thus $f_j$ can be composed of the other
-polynomials of $F_c$, so $f_j$ is  not a part of the unsat core, and
+polynomials of $F_c$, so $f_j$ is  not a part of the UNSAT core, and
 it can be safely discarded from $F_c$. This can be identified by means
@@
 select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
-(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
+({\it i.e.} if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
 from the core; otherwise retain $f_i$ in $F_c$. Select a different
@@
 
-Buchberger's algorithm picks pairs of polynomials from a given set, computes their S-poly, then reduces this S-poly
+Buchberger's algorithm picks pairs of polynomials from a given set, computes their $Spoly$, then reduces this $Spoly$
 with the given set of polynomials. If the remainder is non-zero,
 it is added to the set of polynomials.
-By tracking S-poly computations and multivariate divisions that lead to remainder 
+By tracking $Spoly$ computations and multivariate divisions that lead to remainder 
 1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
@@
 \includegraphics[width=\textwidth]{newfig/UNSAT.eps}
-\caption{DAG representing Spoly computations and multivariate divisions}
+\caption{DAG representing $Spoly$ computations and multivariate divisions}
 \label{fig:UNSAT}}
 \end{figure}
-% We conclude our approach as follows: i) Keep track of all $Spoly(f_i,
-% f_j)\xrightarrow{f_1, \dots, f_s}_+ r$; ii) mark $f_i, f_j$ if $r\neq
-% 0$; iii) Mark those $f_l$'s that have leading terms that cancel
-% monomial terms in $Spoly(f_i, f_j)$; iv) Build a DAG recording all reductions including Spoly computations
-% and multivariate divisions; v) Once $1\in J$ is
-% detected, traverse the graph to identify all the marked polynomials that were
-% used in obtaining this unit element. These polynomials constitute the
-% core.
-
-We conclude our approach as a conjecture Algorithm \ref{alg:UNSAT}.
+
+This general idea is depicted in Algorithm \ref{alg:UNSAT}. 
+We now show a formal and efficient implementation.
 
@@
 {
-	Pick a pair $f_i,f_j\in F$ that has never been computed S-poly\;
+	Pick a pair $f_i,f_j\in F$ that has never been computed $Spoly$\;
 	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
@@
 		$F = F\cup r_l$\;
-		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the S-poly, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the S-poly\;
+		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the $Spoly$, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the $Spoly$\;
 	}
@@
 $Spoly(f_i,f_j)\xrightarrow{F}_+ g_{ij}$ reductions in Buchberger's
-algorithm. Since $F$ is itself an unsat core, 
+algorithm. Since $F$ is itself an UNSAT core, 
 definitely {\it there exists  
-a sequence of Spoly reductions in Buchberger's algorithm where
+a sequence of $Spoly$ reductions in Buchberger's algorithm where
 $Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
@@
 the {\it data} generated by Buchberger's algorithm --- namely, the
-critical pairs($f_i,f_j$) used in the Spoly computations,
+critical pairs($f_i,f_j$) used in the $Spoly$ computations,
 and the polynomials from $F$ used to cancel terms in the reduction
@@
 {\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
-$(f_2,f_5)$. It turns out that the Spoly-reductions corresponding to
+$(f_2,f_5)$. It turns out that the $Spoly$-reductions corresponding to
 these 3 pairs lead to the unit ideal. Recording the data
@@
 \includegraphics[width=4.5in]{newfig/refutation_tree.eps}
-\caption{Generating refutation trees to record unsat cores}
+\caption{Generating refutation trees to record UNSAT cores}
 \label{fig:refute}
@@
 leaves. Then, all the leaves in the transitive fanin of ``1''
-constitute an unsat core. The polynomials (nodes) that do not lie in
+constitute an UNSAT core. The polynomials (nodes) that do not lie in
 the transitive fanin of ``1'' can be safely discarded from $F_c$. From
 Figure \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
-as an unsat core of $F$.
-
-\section{Reducing the Size of the Infeasible Core $F_c$}
+as an UNSAT core of $F$.
+
+\section{Reducing the Size of the UNSAT Core $F_c$}
 \label{sec:alg}
@@
 discarded. We will now describe techniques to further reduce the size
-of the unsat core by identifying such redundant elements. 
+of the UNSAT core by identifying such redundant elements. 
 %how the size of this core could be reduced further.  
@@
 
-\subsection{Identifying redundant polynomials from the refutation tree}
+\subsection{Identifying Redundant Polynomials from the Refutation Tree}
 
@@
 coefficients are associated with new elements (remainders) in the
-Gr\"obner basis, that are not a part of the unsat core. 
+Gr\"obner basis, that are not a part of the UNSAT core. 
 %which cannot benefit the UNSAT core extraction. 
@@
  \FOR {{ each pair $(f_i,f_j)\in G$  }}
- 	\STATE  $f_{sp},(f_{i},h_{ij})(f_{j},h_{ji}) \gets$ Spoly($f_i,f_j$) \COMMENT{$f_{sp}$ is the S-polynomial}
+ 	\STATE  $f_{sp},(f_{i},h_{ij})(f_{j},h_{ji}) \gets Spolyf_i,f_j)$ \COMMENT{$f_{sp}$ is the S-polynomial}
  	\STATE{{ $g_{ij}|(c_{k1},f_{k1}),\dots,(c_{kl},f_{kl}) \gets (f_{sp}\xrightarrow{G}_+  g_{ij})$}}
@@
 
-\section{Iterative Refinement of the Unsat Core}
+\section{Iterative Refinement of the UNSAT Core}
 \label{sec:iter}
 
-As with most other unsat core extractors, our algorithm also cannot
+As with most other UNSAT core extractors, our algorithm also cannot
 generate a minimal core in one execution. To obtain a smaller core, we
@@
 by relabeling the polynomial indexes, we can affect their
-chances of being selected in the unsat core. We use two criteria to
-to affect the polynomial selection in the unsat core. One corresponds
+chances of being selected in the UNSAT core. We use two criteria to
+to affect the polynomial selection in the UNSAT core. One corresponds
 to the \emph{refutation distance}, whereas the other corresponds to
@@
 as they can cancel more terms. This may prohibit other (higher-degree)
-polynomials from being present in the unsat core.
+polynomials from being present in the UNSAT core.
 
@@
 \begin{Definition}[Frequency of Occurrence]
-Frequency of occurrence of a polynomial $f_i$ in a refutation tree
+The frequency of occurrence of a polynomial $f_i$ in a refutation tree
 corresponds to the number it appears in the refutation tree. 
@@
 properties (leading terms) that give them a higher likelihood of being
-present in the unsat core.
+present in the UNSAT core.
 
@@
 f_4'\} = \{f_2,f_3,f_4\}$. Further iterations do not refine the core
--- i.e. a fix point is reached. 
+-- {\it i.e.} a fix point is reached. 
 \end{Example}
 
-\section{Refining the Unsat core using Syzygies}
+\section{Refining the UNSAT Core using Syzygies}
 \label{sec:syz}
@@
 
-The unsat core obtained through our GB-core algorithm is by nature a
+The UNSAT core obtained through our GB-core algorithm is by nature a
 refutation polynomial that equals to 1:  
@@
 $f_k \in F$ can be represented using a combination of the rest of the
-polynomials of the core, e.g.:
+polynomials of the core, {\it e.g.}
 $$f_k = \sum_{j\neq k} c_j'f_j.$$
@@
 such polynomials $f_k$ in the generating set $F$ that can be composed
-of the other polynomials in the basis, i.e. 
+of the other polynomials in the basis, {\it i.e.}
 $f_k \in \langle F-\{f_k\} \rangle$. We present an approach targeted
@@
 implies that some polynomial combination of $\{f_1,\dots,f_s\}$
-vanishes: i.e. $c_1f_1 + c_2f_2 + \dots+c_sf_s=0$, for some
+vanishes: {\it i.e.} $c_1f_1 + c_2f_2 + \dots+c_sf_s=0$, for some
 $c_1,\dots,c_s$. These elements ($c_1,\dots,c_s$) form a syzygy on $f_1,\dots,f_s$.
@@
  Here $\{f_1,f_2,\dots,f_s\}$ is the given core.
- Take one column of the syzygy matrix (e.g. the set of polynomials in $j$-th column
+ Take one column of the syzygy matrix ({\it e.g.} the set of polynomials in $j$-th column
  $c_j^1, c_j^2, \dots, c_j^m$)  and compute its reduced Gr\"obner
@@
 discarded from the core. By repeating this procedure, some redundant
-polynomials can be identified and size of unsat core can be reduced
+polynomials can be identified and size of UNSAT core can be reduced
 further.
@@
 
-\textbf{Our Overall Approach for Unsat Core Extraction:} i) Given
+\textbf{Our Overall Approach for UNSAT Core Extraction:} i) Given
 the set $F = \{f_1,\dots,f_s\}$, we apply the GB-core algorithm,
@@
 in $F_c$; all such $f_k$ are discarded from $F_c$. This gives us the
-final unsat core $F_c$.
+final UNSAT core $F_c$.
 
@@
 In this section we apply our UNSAT core extraction approach to the 
-$k$-BMC in Algorithm \ref{alg:refined}. This algorithm utilize UNSAT core to 
+$k$-BMC in Algorithm \ref{alg:refined}. This algorithm utilizes UNSAT core to 
 remove irrelevant latches (state variables) to
@@
 
-\section{Experiment results}
+\section{Experiment Results}
 \label{sec:exp}
@@
 tool implementation, we have performed experiments to extract a
-minimal unsat core from a given set of  
+minimal UNSAT core from a given set of  
 polynomials. Our experiments run on a desktop with
@@
 against  Montgomery multipliers \cite{lv:tcad2013}. Some of these are
-available as CNF formulae, whereas others were available directly as
-polynomials over finite fields. The CNF formulae are translated as
+available as CNF formulas, whereas others were available directly as
+polynomials over finite fields. The CNF formulas are translated as
 polynomial constraints over $\mathbb{F}_2$ (as shown in
@@
 from which a core is to be extracted. \#MUS is the {\it minimal} core either
-extracted by PicoMUS (for CNF benchmarks) or exhaustive deletion method (for non-CNF bencmarks).
+extracted by {\it PicoMUS} (for CNF benchmarks) or exhaustive deletion method (for non-CNF bencmarks).
  \#GB-core iterations corresponds to the
-number of calls to the GB-core engine to arrive at the reduced unsat
+number of calls to the GB-core engine to arrive at the reduced UNSAT
 core. The second last column shows the improvement in the minimal core
@@
 iteratively refined further. 
-We choose PicoMUS as a comparison to our tool because it is a state-of-art MUS extractor,
+We choose {\it PicoMUS} as a comparison to our tool because it is a state-of-art MUS extractor,
 and the results it returned for our set of benchmarks are proved to be minimal. 
@@
 cases, our tool can produce a minimal core. For the subset-3
-benchmark, we obtain another core with even smaller size than the one from PicoMUS. The results
+benchmark, we obtain another core with even smaller size than the one from {\it PicoMUS}. The results
 demonstrate the power of the Gr\"obner basis technique to identify the
@@
 
-\section{Conclusions}\label{sec:conc}
-This paper addresses the problem of identifying an infeasible core of
+\section{Concluding Remarks}\label{sec:conc}
+This chapter addresses the problem of identifying an infeasible core of
 a set of multivariate polynomials, with coefficients from a field,
@@
 the use of GB engines for SAT solving has a rich history, the problem
-of unsat core identification using GB-engines has not been addressed
-by the SAT community. We hope that this paper will kindle some
+of UNSAT core identification using GB-engines has not been addressed
+by the SAT community. We hope that this technique will kindle some
 interest in this topic which is worthy of attention from the SAT

### 2016-12-11 dissertation/abstract.tex
dissertation: committee review version

@@
 (Galois) field of $2^k$ elements. Subsequently, techniques from
-algebraic geometry -- notable, Gr\"obner basis theory and technology
+algebraic geometry -- notably, Gr\"obner basis theory and technology
 -- are researched to perform reachability analysis and

### 2016-12-11 dissertation/intro.tex
dissertation: committee review version

@@
 \section{Hardware Design and Verification Overview}
-During the last decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
+During the past decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
 As a result, an entire system with billions of transistors can be built upon a single chip.
@@
 
-One way to perform verification is by {\it simulation}. It is defined as the collage of all circuit validation 
+One way to perform verification is by {\it simulation}. It is the collage of all circuit validation 
 methods which apply stimulus on the inputs of circuit model and verify correctness of the outputs.
@@
 
-A part of formal verification techniques adopt concepts and algorithms from 
+Many formal verification techniques adopt concepts and algorithms from 
 \emph{computer-algebra} and \emph{algebraic geometry}.
@@
 exponential in the worst case \cite{dube1986complexity}. Thus, directly performing GB computation 
-over arbitrary setup is not practical for industry-level applications. However, recent
+over an arbitrary setup is not practical for industry-level applications. However, recent
 breakthroughs in computer-algebra hardware verification have shown

### 2016-12-11 dissertation/normal.tex
dissertation: committee review version

@@
 \begin{table}[H]
-\label{tab:slimgb}
 \begin{center}
-	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:ATO}
+	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:slimgb}
 	    \begin{tabular}{|c|c||c|} 
@@
 		\hline
-	        $5$	&  $65$  & $20$ \\
-		$11$	&  $275$  & $>35,000$  \\
+	        $5$	&  $65$  & $16$ \\
+	        $6$ &  $90$  & $368$ \\
+	        $9$ &  $189$ & $43602$ \\
+		$11$	&  $275$  & $TO$ ($>48$ hr)  \\
 	        \hline
@@
 \begin{itemize}
-\item First, a set of polynomials is constructed from each logic gates in the circuit, and they generate 
+\item First, a set of polynomials is constructed from each logic gate in the circuit, and they generate 
 ideal $J_{gates}$. Similarly, ideal $J_{WB}$ is generated by polynomials representing the word-to-bit correspondences.
@@
 substitution, and compute reduced \Grobner bases with RATO, then our
-approach does not scale beyond 33-bit datapaths.
+approach does not scale beyond 11-bit datapaths (Table \ref{tab:slimgb}).
 
@@
 the binomial expansion over Galois fields and Gaussian
-elimination. Our approach is able to verify up to 100-bit sequential
-circuits, whereas contemporary techniques fail beyond 23-bit
+elimination. Our approach is able to verify up to 162-bit sequential
+circuits, whereas contemporary techniques ({\it e.g.} BDDs) fail beyond 23-bit
 datapaths.

### 2016-12-11 dissertation/prelim_GF.tex
dissertation: committee review version

@@
 
-\section{Design a Normal Basis Multiplier on Gate Level}
+\section{Design of a Normal Basis Multiplier on Gate Level}
 \label{sec:nbdesign}
@@
 
-\subsection{Multiplier not based on $\lambda$-Matrix}
+\subsection{Multiplier not based on the $\lambda$-Matrix}
 Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
@@
 application on Galois field hardware design. The sequential Galois field multipliers based on these 
-principles are used as good candidates for applying our functional verification approach in subsequent chapters.
+principles are used as candidates for applying our functional verification approach in subsequent chapters.

### 2016-12-11 dissertation/prelim_basis.tex
dissertation: committee review version

@@
 \label{sec:geo}
-\subsection{Monomials, Polynomials and Polynomial Arithmetics}
+\subsection{Monomials, Polynomials and Polynomial Algebra}
 \begin{Definition} \label{def:mono}
@@
 because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
-the worst case computation time/space cost of its Gr\"obner basis is doubly exponential while the total degree of 
+the worst case computation time/space cost of its Gr\"obner basis is doubly exponential, as the total degree of 
 polynomials in the Gr\"obner basis is bounded by 
@@
 This provides the possibility to make use of Gr\"obner basis under restricted situations,
-which is the theoretic basis of this dissertation.
+which is the theoretical basis of this dissertation.
 
@@
 \begin{Theorem}[Abstraction Theorem]
+\cite{tim:phd}
 Using the notations from Problem Setup \ref{setup:abs} at the beginning of this subsection,

### 2016-12-11 dissertation/prev.tex
dissertation: committee review version

@@
 The specification of a sequential circuit can be modeled as a (golden model) state machine,
-SEC is performed to compare between the circuit for test and the golden one.
+SEC is performed to compare the functionality between the circuit for test and the golden one.
 % The n\"aive method to implement SEC is: preload both circuits to the same initial states, 
@@
 % and does not require the 1:1 state correspondence,  so state space traversal is not always necessary.
-One way to implement SEC is to create miter with two circuits to be verified, then 
+One way to implement SEC is to create a miter with two circuits to be verified, then 
 prove that there exists no sequence of inputs that generates different outputs.
@@
 a set of states/transitions \cite{coudert2003unified, coudert1990verification},  or by dividing the sequential circuit to 
-smaller subcircuit and re-model the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
-interfaces that focuses on only designated initial states and remove redundancies in state space \cite{baumgartner2007scalable}.
+into a smaller subcircuit and re-model the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
+interfaces that focuses on only the designated initial states and remove redundancies in state space \cite{baumgartner2007scalable}.
 
@@
 specification to combinational ones, and then perform combinational equivalence checking
-techniques \cite{avoj2010combinational}. In the following part we review research and techniques which 
+techniques \cite{savoj2010combinational}. In the following part we review research and techniques which 
 can be applied to unrolled sequential circuits.
@@
 If the circuits being compared are structurally highly similar, AIG and circuit-SAT 
-based approaches are proved to be efficient.
+based approaches are known to be efficient.
 However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
@@
 correctness of integer arithmetic circuits. Their approach formulates the output signature as a polynomial function 
-with binary variables and integer coefficients, then rewrite the polynomial by substituting gate output with gate inputs
-variables. After going through the backward rewriting procedure,  the polynomial
+with binary variables and integer coefficients, then rewrites the polynomial by substituting gate output with gate 
+inputs. After going through the backward rewriting procedure,  the polynomial
 will be composed by only input variables. Then the polynomial is converted to a canonical representation, and compared
 with designated input signature. If they are equivalent, then the arithmetic circuit is successfully verified.
-This approach incurs of polynomial term explosion during the backward rewriting. The authors proposed a heuristic
+This approach incurs polynomial term explosion during the backward rewriting. The authors proposed a heuristic
 to levelize the arithmetic circuit, and substitute several gates' variables at the same time to minimize the risk. However, 
-the heuristic is proved to be less effective when the inner symmetry of circuit structure is missing.
+the heuristic proved to be less effective when the inner symmetry of circuit structure is missing.
 
@@
 
-$k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
+The $k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
 on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
@@
 From the investigation of previous work, techniques are to be researched that 
-can operate the FSM traversal on word-level to verify a property excluding spurious 
+can perform the FSM traversal on word-level to verify a property excluding spurious 
 faults. Meanwhile, many abstraction refinement techniques utilize information from UNSAT cores.
@@
 
-In this dissertation, we propose a purely word-level reachability analysis approach which is never done before.
-We achieve it by modeling the transition relations, states and the traversal algorithm at word-level.
+In this dissertation, we propose a purely word-level reachability analysis approach which has never been done before.
+We achieve this by modeling the transition relations, states and the traversal algorithm at word-level.
 We borrow inspirations from \cite{tim:phd,gao:qe-gf-gb} to perform state space abstraction.
-Moreover, we find applications of our proposed approach to sequential arithmetic verification which is 
+Moreover, we demonstrate applications of our proposed approach to sequential arithmetic verification which is 
 not done before, either.
+Finally, we show algebraic geometry analogs of UNSAT cores of polynomial ideals, and describe algorithms
+to extract and refine these cores.

### 2016-12-11 dissertation/recha.tex
dissertation: committee review version

@@
 % some major refactoring and attachments are made on original designs. 
-These modifications may introduce, bugs, errors or malfunctions to the original logic and cause problems. 
+These modifications may introduce bugs, errors or malfunctions to the original logic and cause problems. 
 Based on traditional localized simulation or formal verification method ({\it e.g.} equivalence checking), 
@@
 Before introducing the details of our approach, there are a few questions to ask: what is 
-the benefit of execute FSM traversal at word-level? How can algebraic geometry make this happen?
+the benefit of executing FSM traversal at word-level? How can algebraic geometry make this happen?
 The answers can be found in this section, as a statement of the motivation of our research.
@@
 Sequential circuits are modeled as FSMs, which can be implemented as graphs. Thus a 
-graph-traversal based algorithm is created to analysis the reachable states \cite{coudert2003unified}.
+graph-traversal based algorithm is created to analyze the reachable states \cite{coudert2003unified}.
 A traversal algorithm using the concept of implicit state enumeration is proposed \cite{cho1993redundancy}.
@@
 \end{equation}
-The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
+The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ points with coordinates $(x_i,y_i)$,
 they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
@@
 \begin{Proof}
-Since $\Fkk[x]$ is a principal ideal domain, $ J = \langle g\rangle$
-for some polynomial $ g \in \Fkk[x]$. Let $d = gcd(g, x^{2^k} -
-x)$. So, $g = dg_1 , x^{2^k} - x = df_1$, with $gcd(f_1 , g_1 ) =
-1$. Then $J_0 : J = \langle f_1 \rangle$ by Lemma \ref{lem:gcd}.
+Since $\Fkk[x]$ is a principal ideal domain, $ J = \langle g(x)\rangle$
+for some polynomial $ g(x) \in \Fkk[x]$. Let $h(x) = gcd(g(x), x^{2^k} -
+x)$. So, $g(x) = h(x)g_1(x) , x^{2^k} - x = h(x)f_1(x)$, with $gcd(f_1(x) , g_1(x) ) =
+1$. Then $J_0 : J = \langle f_1(x) \rangle$ by Lemma \ref{lem:gcd}.
 
@@
 
-Since $x^{2^k} = x$, we see that either $d(x) = 0$ or $f_1 (x) =
-0$. Considering $g(x) \neq 0$, we can assert that $d(x) \neq 0$. In
+Since $x^{2^k} = x$, we see that either $h(x) = 0$ or $f_1 (x) =
+0$. Considering $g(x) \neq 0$, we can assert that $h(x) \neq 0$. In
 conclusion, $f_1 (x) = 0$ and $x \in V(f_1 )$.
@@
 Now we make an assumption that $x \in V(g)$. Then $g(x) = 0 =
-d(x)g_1(x)$ which means either $d(x) = 0$ or $g_1 (x) = 0$.
-
-If $g_1 (x) = 0$, then since $f_1 (x) = 0$ we get that $f_1 , g_1$
-share a root. This contradicts the fact that $gcd(f_1 , g_1 ) = 1$.
-
-On the other hand, if $d(x) = 0$, then since $f_1 (x) = 0$ and
-$x^{2^k} - x = df_1$, we get that $x^{2^k} - x$ has a double root. 
+d(x)g_1(x)$ which means either $h(x) = 0$ or $g_1 (x) = 0$.
+
+If $g_1 (x) = 0$, then since $f_1 (x) = 0$ we get that $f_1(x) , g_1(x)$
+share a root. This contradicts the fact that $gcd(f_1(x) , g_1(x) ) = 1$.
+
+On the other hand, if $h(x) = 0$, then since $f_1 (x) = 0$ and
+$x^{2^k} - x = d(x)f_1(x)$, we get that $x^{2^k} - x$ has a double root. 
 But this is impossible since the derivative of $x^{2^k} - x$ is $-1$.
 
-So, $x \notin V(g)$ and this concludes the proof.
+So, $x \notin V(g(x))$ and this concludes the proof.
 \end{Proof}
@@
 \item The circuit is modeled over $\Fkk = \F_2[x]\pmod{p(x)}, p(\alpha) = 0$, where $k$ is the number of flip-flops,
-	the PS variables $\{s_0,dots,s_{k-1}\}$, NS variables $\{t_0,\dots,t_{k-1}\}$.
+	the PS variables $\{s_0,\dots,s_{k-1}\}$, NS variables $\{t_0,\dots,t_{k-1}\}$.
 \item Denote $S$ as the PS word-level by the following polynomial:
@@
 
-  $from^0 = reached = \F(S)$\;
+  $from^0 = reached = \Func(S)$\;
   \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
@@
 
-\subsection{Primary Inputs Partition}
+\subsection{Primary Inputs Partitioning}
 Using above techniques we can get a remainder polynomial with only word-level PS/NS variables. However in most 
@@
 
-\section{Concluding Remarks} \label{sec:conc}
-This dissertation has presented a new approach to perform reachability
+\section{Concluding Remarks}
+This chapter has presented a new approach to perform reachability
 analysis of finite state machines at the word-level. This is achieved

### 2016-12-11 dissertation/unsat.tex
dissertation: committee review version

@@
 
-Boolean satisfiability (SAT) problem is the basis of 
+The Boolean satisfiability (SAT) problem is the basis of 
 most decidable decision problems, as well as the basis of many formal verification techniques. 
@@
 \subsection{Preliminaries of SAT/UNSAT Theory}
-Boolean satisfiability (SAT) problem is a fundamental problem in computer science.
+The Boolean satisfiability (SAT) problem is a fundamental problem in computer science.
 In the following part we define the terminology related to SAT problem \cite{anaICCAD}.

### 2016-12-13 dissertation/slides/defense.tex
dissertation: slides v1

% Lump tex of Xiaojun's defense slides

\documentclass[xcolor=dvipsnames]{beamer}
%%\usetheme{default}
\usecolortheme[named=Maroon]{structure}
%\usetheme{Boadilla}
\usetheme{Madrid}
\useoutertheme{default}%[footline=empty]{infolines}

% \usepackage{helvet}
% \usepackage{enumerate}
% \usepackage{amsmath}
% \usepackage{amsfonts}
% \usepackage{graphicx}
% \usepackage{ulem}
\usepackage{multirow}
\usepackage{comment}
\usepackage{xspace}
%\usepackage{columns}

\usepackage[absolute,overlay]{textpos}
\usepackage[ruled,linesnumbered]{../algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\renewcommand{\thealgocf}{}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\RequirePackage{algorithmic}
%\RequirePackage{algorithm}
% \renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
% \renewcommand{\algorithmicensure}{\textbf{Outputs:}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{lemma}
%\newtheorem{corollary}{Corollary}
%\newtheorem{proposition}{Proposition}
%\newtheorem{Q}{Question}
%\newtheorem{Exa}{Example}
%\newtheorem{Definition}{Definition}

\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fp}{{\mathbb{F}}_{p}}
\newcommand{\Fpk}{{\mathbb{F}}_{p^k}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\Zkk}{{\mathbb{Z}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
%\newcommand{\Grobner}{Gr\"{o}bner}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\Func}{{\mathcal{F}}}
\newcommand{\N}{{\mathcal{N}}}
\newcommand{\G}{{\mathcal{G}}}
\newcommand{\F}{{\mathbb{F}}}
\newcommand{\B}{{\mathbb{B}}}
\newcommand{\C}{{\mathbb{C}}}
\newcommand{\R}{{\mathbb{R}}}

\title[PhD Dissertation Defense]{Word-Level Abstractions for Sequential Design Verification using Algebraic Geometry}

\author[Xiaojun Sun]{{\bf Xiaojun Sun}, PhD Candidate}

%\email{rostamian@umbc.edu}
\institute[Univ. of Utah]{
\includegraphics[height=22mm]{./Ulogo_1200px.png}\\
\    \\
\    \\
\    \\
Electrical and Computer Engineering, University of Utah\\
\    \\
Advisor: {\it Priyank Kalla}
}

\date{Dec 16, 2016}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%----------- titlepage ----------------------------------------------%
\begin{frame}[plain]
  \titlepage

\end{frame}

%\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Contributions
\item Motivations
\item Previous Work
\item Preliminaries
	\bi
	\item Finite fields
	\item Polynomial algebra \& Algebraic geometry
	\item Projection
	\ei
\item Projection based abstraction 
	\bi 
	\item Application: \alert{Reachability analysis}
	\item Application: \alert{Sequential arithmetic ckt verification}
	\ei
\item UNSAT core based abstraction
	\bi
	\item \alert{UNSAT core extraction using Gr\"obner basis refutation}
	\item Application: Bounded model checking (BMC) with abstraction refinement
	\ei
\item Conclusion \& Future work
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Contributions}}
\bi
\item Word-level reachability analysis -- analog if implicit state enumeration
	\bi 
	\item {\bf Q:} Why word-level? 
	\pause
	\item {\bf A:} Datapath $\gets$ word-level info 
	\pause
	\item {\bf A:} Simplify representation (abstraction) of state-space $\to$ Efficiency!
	\ei
\pause
\item Apply word-level reachability algo. to sequential arithmetic circuit verification
	\bi
	\item Abstraction $\to$ word-level signature each time-frame
	\item Word-level unrolling
	\ei
\pause
\item UNSAT cores in Algebraic geometry
	\bi
	\item Analyze Buchberger's algorithm $\to$ extract refutation proof $\to$ UNSAT core
	\item Structure of refutation proof $\to$ refine UNSAT core
	\item Apply UNSAT core to $k$-BMC abstraction refinement
	\ei
\pause
\item Implement above algos: C++ \& \textsc{Singular}
	\bi
	\item For sequential GF multipliers: overwhelmingly better than contemporary tools
	\ei
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = linkSTG]{\large{Motivation I: BFS state space traversal}}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{./stg_fig.eps}
\caption{State Transition Graph}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\bi
\item Initial state: $\{00\}$
\item Iteration 1: 
	\bi
	\item Start from $\{00\}$
	\item One-step transition: $\{01,10\}$
	\item Newly reached: $\{01,10\}$
	\ei
\item Iteration 2: 
	\bi
	\item Start from $\{01,10\}$
	\item One-step transition: $\{00,10\}$
	\item Newly reached: $\emptyset$
	\ei
\item All reachable states detected. Final reached states: $\{00,01,10\}$
\pause
\item Still need bit-level Boolean variables to represent states
\ei
\end{column}
\end{columns}
\hyperlink{pptpage2}{\beamergotobutton{Go back to example}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = motiv2]{\large{Motivation II: Sequential arithmetic circuits verification}}
\vspace{-0.1in}
\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{../newfig/convention.eps}
}
\end{figure}
\vspace{-0.2in}
\bi
\item Problem: Verify the function of sequential arithmetic circuit
	\bi
	\item Operands preloaded into registers
	\item After $k$ clock cycles, give desired output
	\ei
\item Conventional: explicitly unroll $k$ time-frames (bit-level) and setup miter
	\bi
	\item Check miter output: SAT, BDDs, AIGs
	\item Bit-blasting!
	\ei
\ei
\hyperlink{Moore}{\beamergotobutton{Go back to Moore machine}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = motiv3]{\large{Motivation III: $k$-BMC with abstraction refinement}}
\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
	\KwIn{$M$ is the original machine, $p$ is the property to check, $k$ is the number of steps unrolling $M$}
  $k = $ InitValue\;
  \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT} (reachable states after $k$ steps violates $p$)}
  {
	\Return{``Found error trace"}
  }
  {
	\alert{Extract UNSAT core $\mathcal P$ of $k$-BMC} \;
	$M' = $ \textit{ABSTRACT}$(M,\mathcal P)$\;
  }
  \eIf{MODEL-CHECK$(M',p)$ returns \textbf{PASS}}
  {
	\Return{``Passing property"}
  }
  {
	Increase bound $k$\;
	goto Line 2\;
  }
\caption{$k$-BMC with Abstraction Refinement (L. Zhang'05)}
\end{algorithm}
\DecMargin{1em}
\hyperlink{refine}{\beamergotobutton{Go back to Abs refine}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Previous work}}
\bi
\item Sequential Equivalence Checking (SEC): bit-blasting, or structural info dependency
	\bi
	\item Sequential miter
	\item Unroll, then use DDs, SAT or AIGs (Combinational)
	\ei
\item Symbolic model checking (counterexample, IC3): SAT/BDDs in nature
\item Word-level techniques (term rewriting, uninterpreted function): ?
\item Algebraic geometry methods
	\bi
	\item Gr\"ober basis in model checking [Avrunin,CAV'96;Vardi,IASTED'07]: Analog of bit-level Boolean functions
	\item Abstract word-level polynomial representation for arbitrary combinational ckt [Pruss,TCAD'16]
	\ei
\pause
\item No purely word-level sequential verification: data/abstraction/algorithm
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Finite/Galois Fields}}
\textbf {Galois field} $\mathbb{F}_q$ is a finite field with $q$
elements, $q = p^k$, $p = $prime
\begin{itemize}
\item 0,1 elements, commutative, associate, distributive laws
\item Closure property: $+,-,\times$, inverse ($\div$)
\end{itemize}

Our interest: $\mathbb{F}_{q} = \Fkk$ ($q = 2^k$)
\begin{itemize}
\item  $\mathbb{F}_{2^k}$: $k$-dimensional extension of  $\mathbb{F}_{2}=\{0,1\}$
	\begin{itemize}
	\item $k$-bit bit-vector, AND/XOR arithmetic
	\end{itemize}
\end{itemize}

To construct $\mathbb{F}_{2^k}$
\begin{itemize}
\item $\mathbb{F}_{2^k} \equiv \mathbb{F}_{2}[x] \pmod {P(x)}$
\item $P(x) \in \mathbb{F}_{2}[x]$, irreducible polynomial of degree
  $k$
\item Operations performed (mod $P(x)$) and coefficient reduced (mod 2)
\bi
\item E.g. $\C   = \R[x] \pmod{ x^2+1}$
\ei
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Preliminaries: Field construction of $\mathbb{F}_8$}
%\ptsize{10}

Consider: $\mathbb{F}_{2^3}= \mathbb{F}_{2}[x] \pmod {x^3 + x +
  1}$\\ $A \in \mathbb{F}_{2}[x]$ \\ 
$A \pmod {x^3 + x + 1} = a_2 x^2 + a_1 x + a_0$. Let $P(\alpha) = 0$:

\begin{itemize}
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 0, 0\rangle = 0$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 0, 1\rangle = 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 1, 0\rangle = \alpha$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 1, 1\rangle = \alpha
  + 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 0, 0\rangle = \alpha^2$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 0, 1\rangle =
  \alpha^2 + 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 1, 0\rangle =
  \alpha^2 + \alpha$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 1, 1\rangle =
  \alpha^2 + \alpha + 1$
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Polynomial function $f:\Fq\to\Fq$}}

\begin{Theorem}[Fermat's Little Theorem over $\Fq$]
Let $\alpha \in \Fq$, then $\alpha^q = \alpha$. Therefore, $x^q -x$
vanishes on all points in $\Fq$. 
\end{Theorem}
\begin{table}[H]
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
\end{table}
\pause
\bi
\item A polynomial function representation always exists in $\Fq[A]$ (Lagrange's interpolation)
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Computer algebra terminology}}
\vspace{-0.2in}
%\ptsize{10}
Let $\mathbb{F}_q = GF(2^k)$, and $\overline{\Fq}$ be its closure
\begin{itemize}
\item $\mathbb{F}_q[x_1, \ldots, x_n]$: ring of all polynomials with
  coefficients in $\mathbb{F}_q$

\item Polynomial $f = c_1X_1 + c_2 X_2 + \dots + c_t X_t$
\bi
\item A monomial ordering is imposed on $f: X_1 > X_2 > \dots > X_t$
\item \alert{Leading term} $lt(f) = c_1 X_1, ~~tail(f) = c_2 X_2 + \dots + c_t X_t$
\item Leading coefficient $lt(f) = c_1$ and leading monomial $lm(f) = X_1$
\ei
\begin{itemize}
\item LEX $x> y> z$: ~~$f = \alert{ -2x^3} + 2x^2yz + 3xy^3$
\item  DEGLEX $x>y>z$:  ~~$f = \alert{ 2x^2yz} + 3xy^3 -2x^3$
\item DEGREVLEX $x>y>z$: ~~$f = \alert{3xy^3} + 2x^2yz - 2x^3$
\end{itemize}
\end{itemize}
\bi
\item Leading terms $lt(f)$ play an important role
\bi
\item Affect division results!
\ei
\ei

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Polynomial division}}
Copy page 12 from FMCAD
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Algebraic geometry terminology (cont.)}}
Let $\mathbb{F}_q = GF(2^k)$:
\begin{itemize}
\item $\mathbb{F}_q[x_1, \ldots, x_n]$: ring of all polynomials with
  coefficients in $\mathbb{F}_q$ 
\item Given a set of polynomials:
\begin{itemize}
\item $f, f_1, f_2, \dots , f_s \in \mathbb{F}_q[x_1, \dots, x_n]$
\item Find solutions to $f_1 = f_2 = \dots = f_s = 0$
\end{itemize}
\item \alert{Variety:} Set of ALL solutions to a given system of polynomial equations: $V(f_1, \dots, f_s)$
	\begin{itemize}
	\item In $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2-1)=\{all\  points\  on\ circle: x^2+y^2-1=0\}$
	\item In $\mathbb{R}[x]$, $V(x^2+1)=\emptyset$
	\item In $\mathbb{C}[x]$, $V(x^2+1)=\{(\pm i)\}$
	\end{itemize}
\item Variety depends on the \alert{ideal} generated by the polynomials.
\item Reason about the Variety by analyzing the Ideals
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Ideals \& Gr\"obner bases}}
\vspace{-0.1in}

\begin{Definition}
{\bf Ideals of Polynomials:} Let $f_1, f_2, \ldots, f_s \in
\mathbb{F}_q[x_1, \dots, x_n]$. Let 
\begin{eqnarray}
J = \langle f_1, f_2 \ldots, f_s\rangle = \{f_1 h_1 + f_2 h_2 + \dots + f_s h_s\} \nonumber 
\end{eqnarray}

$J = \langle f_1, f_2 \ldots, f_s\rangle$ is an ideal generated by
$f_1, \ldots, f_s$ and the polynomials are called the generators. 
\end{Definition}

\vspace{-0.1in}
%\ptsize{10}

\begin{itemize}
\item Different generators can generate the same ideal
\item $\langle f_1,\cdots,f_s \rangle=\cdots=\langle g_1,\cdots,g_t
  \rangle$
\item Some generators are a ``better'' representation of the ideal
\item A {\bf Gr\"obner basis} $G$ is a ``canonical'' representation of an ideal
	\bi
	\item $I = \langle F \rangle = \langle G \rangle$, and $V(F)=V(G)$
	\ei
\item Map: set of states $\to$ \alert{variety} of polynomial ideal
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Buchberger's algorithm}}
 INPUT : $F = \{f_1, \dots, f_s\}$\\
% \ENSURE : $G = \{g_1,\dots ,g_t\}$\\ %, a Gr\"{o}bner basis \\
 OUTPUT : $G = \{g_1,\dots ,g_t\}$\\ %, a Gr\"{o}bner basis \\
  $G:= F$; \\
  REPEAT\\
  \hspace{0.1in} $G' := G$\\
  \hspace{0.1in} For each pair $\{f, g\}, f \neq g$ in $G'$ DO\\
\hspace{0.2in}  $Spoly(f, g) \stackrel{G'}{\textstyle\longrightarrow}_+
r$ \\
\hspace{0.2in}  IF $r \neq 0$ THEN $G:= G \cup \{r\}$ \\
%\hspace{0.2in}  $G(x):=G(x) / x$
UNTIL $G = G'$
%\end{algorithmic}
%\end{algorithm}
\pause
\bi
\item $Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g$\par
$L = \text{LCM}(lm(f), lm(g))$, ~~$lm(f)$: leading monomial of $f$
\item Animation 1...
\pause
\item Inspiration of mapping: set of states $\to$ \alert{variety} of polynomial ideal
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Gr\'obner basis with elimination term order}}
\begin{itemize}
\item Let ideal $I = \langle f_1, f_2, f_3 \rangle$ where
	\begin{itemize}
	\item $f_1 = x^2 + y + z - 1$
	\item $f_2 = x + y^2 + z - 1$
	\item $f_3 = x + y + z^2 - 1$
	\end{itemize}
\item The \Grobner basis of $I$ with \alert{elimination} (LEX) order $(x > y > z)$ is
	\begin{itemize}
	\item $g_1 = x + y + z^2 - 1$
	\item $g_2 = y^2 - y - z^2 + z$
	\item $g_3 = 2yz^2 + z^4 - z^2$
	\item $g_4 = z^6 - 4z^4 + 4z^3 - z^2$
	\end{itemize}
\item Notice that $g_2$ and $g_3$ only contain variables $y$ and $z$
	\begin{itemize}
	\item Eliminates variable $x~\Leftrightarrow  ~\exists_x$ in Boolean formula!
	\end{itemize}
\item Similarly, $g_4$ only contains the variable $z$ and eliminates $x$ and $y$
\pause
\item GB($I_{x,y,z}$)$\cap\Fq[z]\Leftrightarrow$ \alert{projection} on variable $z$!
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Elimination by projection}}
\vspace{-0.1in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/projection.pdf}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/proj_reacha.pdf}
\end{figure}
\end{column}
\end{columns}
\bi
\item Projection of variety from $\F[x_1,x_2,x_3]$ to $\F[x_2,x_3]$
\item Projection of ckt ideal's variety on next state (NS) variables $T$
\pause
\item GB($J_{ckt}+J_0$)$\cap\F_2[T] \implies $Next state polynomial $f(T)$!
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{BFS traversal algorithm}}
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
\alert{	$to^i \gets$Img$(\Delta, from^{i-1})$}\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}
\pause
\bi
\item Image function: $\text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
  \land from ] = \exists _s ~\exists _x ~\bigwedge_
  {i=1}^{n} (t_i
\overline{\oplus } \Delta_i)\land from$
\item In $\mathbb B^k$, image function $\Leftrightarrow \exists$
\item In $\mathbb F_{2^k}$, need to implement \alert{quantifier elimination}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%(copy 17~24 HLDVT)%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Recall: Breadth-First Traversal Algorithm}}
\only<1>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<2>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<3>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\tcp*{\alert{GB of Elim ideal}}
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<4>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\tcp*{\alert{GB of Elim ideal}}
	$new^i \gets to^i \cap \overline{reached}$\tcp*{\alert{ideal quotient \& sum}}
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<5>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\tcp*{\alert{GB of Elim ideal}}
	$new^i \gets to^i \cap \overline{reached}$\tcp*{\alert{ideal quotient \& sum}}
  	$reached \gets reached \cup new^i$\tcp*{\alert{ideal product}}
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Intersection and union in algebraic geometry}}
\begin{Definition}
\label{def:sum}
({\bf Sum/Product of Ideals}) If $I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are 
ideals in $\mathbb F[x_1, \dots, x_n]$, then the {\bf sum} of $I$ and $J$ is defined as
$$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$$ And the {\bf product} of $I$ and $J$ is defined
as
\begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle \nonumber
  \end{equation}
\end{Definition}
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf
  V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) =
{\bf V}(I) \bigcup {\bf V}(J)$. 
\end{Theorem}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Complement set in algebraic geometry}}
\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
  F[x_1, \dots, x_n]$, then $I:J$ is the set
  \begin{equation}
  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}\nonumber
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}
\begin{Theorem}
\label{thm:quotient}
Let $J_0$ be an ideal of vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
$${\bf V}(J_0:J) = {\bf V}(J_0) - {\bf V}(J) = \overline{{\bf V}(J)}$$
\end{Theorem}
\bi
\item Given ideal $J$, compute $J'$ s.t. $V(J') = \overline{V(J)}$
\item $\to J'=J_0:J$
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Our proposed algorithm of BFS traversal based on algebraic geometry}}

\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\Indm
 \KwIn{The circuit's characteristic polynomial ideal $J_{ckt}$,
   initial state polynomial $\F(S)$, and LEX term order: bit-level
   variables $x, s, t$ $>$ PS word $S$ $>$ NS word $T$}
\Indp

  $from^0 = reached = \F(S)$\;
  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
  {
  	$i \gets i + 1$\;
  	$G \gets$GB($\langle J_{ckt},J_0, from^{i-1}\rangle$)\tcp*{This step contains bit-level}
%  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest}
	$\langle to^i\rangle \gets G\cap \Fkk[T]$\tcp*{Only word-level $S,T$ onwards}
%	\tcc{There will be a univariate polynomial in $G$ denoting the
%          set of next states in word-level variable $T$}
	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
%	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
%  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
	$from^i \gets new^i(S\setminus T)$\;
%	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
  }
%  \tcc{Loop until a fixpoint reached: newly reached state is empty}
\Return{$\langle reached\rangle$}
\caption {Algebraic Geometry based FSM Traversal}
\end{algorithm}
\DecMargin{1em}

\hyperlink{pptpage2}{\beamergotobutton{Go to example page 2}}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Initial state $from^0 = S(\{00\})$
\item {\bf Iteration 1:}Compose an elimination ideal $J$ 
\vspace{-0.2in}\\
\begin{align*}
&f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
&f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
&f_3: S - s_0 - s_1\alpha\\
&f_4: T - t_0 - t_1\alpha
\end{align*}
$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$
\begin{align*}
&f_5: x^2-x\\
&f_6: s_0^2-s_0, f_7: s_1^2-s_1\\
&f_8: t_0^2-t_0, f_9: t_1^2-t_1\\
&f_{10}: S^4-S, f_{11}:T^4-T
\end{align*}
$\ \ \ \ \ \ \ \ \ \ \  \ J_0 = \langle f_5,f_6,\dots,f_{11}\rangle$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = pptpage2]{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Elimination term order: 
$$\{x,s_0,s_1,t_0,t_1\}~(all~bits)~>~S~(PS~word)~ >~ T~(NS~word)$$
\item Compute the reduced GB for $J = J_{ckt}+J_0+\langle from^0\rangle$
\item Next state 
$$to^1 = \langle T^2+(\alpha+1)T+\alpha\rangle$$ 
\item Mapping to set of states
$$V(to^1) = \{1,\alpha\} \Leftrightarrow \{01,10\}$$
\item Complement of formerly reached state:
$$\langle T^4-T\rangle:\langle T\rangle = \langle T^3+1\rangle$$ 
\item Mapping to set of states
$$V(\langle T^3+1\rangle) = \{1,\alpha,1+\alpha\} \Leftrightarrow \{01,10,11\}$$
\ei
\hyperlink{linkSTG}{\beamergotobutton{STG}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = pptpage3]{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Newly reached states:
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle~(\{01,10\})$$ 
\item Update current reached states 
$$reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$ 
\item Mapping to set of states
$$V(reached) = \{0,1,\alpha\} \Leftrightarrow \{00,01,10\}$$
\item Update the present states for next iteration
$$from^1 = \langle S^2+(\alpha+1)S+\alpha\rangle$$
\ei
\hyperlink{linkSTG}{\beamergotobutton{STG}}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Iteration 2:
	\bi
	\item Next state: $to^2 = \langle T^2+\alpha T\rangle~(\{00,10\})$
	\item The complement of $reached$:
	$$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
= \langle T + 1+\alpha\rangle ~(\{11\})$$ 
	\item Newly reached state: 
$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle \alert{1}\rangle$$ 
	\ei
\item Algorithm terminates
\item Return value (final reachable states):
$$reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%(End of copy)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Improve complexity using RATO}}
\bi
\item Directly compute GB($J_{ckt}+J_0$) in $\Fq$ is costly ($q^{O(d)}$)
\item GB sensitive to term order$\to$Transform to simpler set for GB?
\pause
\item RATO: reverse topological traverse on ckt structure
\item Benefit: Only one pair of poly with non-relatively-prime leading terms\par
$gcd(lt(f_w),lt(f_g))\neq 1$
\ei
\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{../newfig/new_stg.eps}
\label{fig:fsm}}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Example of using RATO}}
\bi
\item RATO: LEX with $(t_0,t_1)>(a,b,c,d)>(x,s_0,s_1)>T>S$
\ei
\begin{equation*}
f_1: a+xs_0s_1+xs_0+xs_1+x+s_0s_1+s_0+s_1+1
\end{equation*}
\vspace{-0.8cm}
\begin{alignat*}{2}
f_2: b+s_0s_1 &~~~ f_3: c+x+xs_0 &&~~~f_4: d+s_0s_1+s_0 \\
f_5: \alert{t_0}+ab+a+1 &~~~ f_6: t_1+cd+c+d &&~~~ f_7: \alert{t_0}+t_1\alpha+T 
\end{alignat*}
\bi
\item $Spoly$ reduction gives $T+\Func(inputs)$
\ei
\begin{align}
&Spoly(f_5,f_7) \xrightarrow{J_{ckt}+J_0}_{+}T + s_0 s_1 x+\alpha s_0 s_1 \nonumber\\
&+(1+\alpha)s_0 x+(1+\alpha) s_0+s_1 x+s_1+(1+\alpha) x+1\nonumber
\end{align}
\bi
\item {\bf Q:} How to get rid of bit-level inputs?
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Bit-to-word conversion}}
\bi
\item Objective: find $s_i = \mathcal G(S)$
\item Build system of poly eqn by squaring:
\begin{align*}
\label{eqn:alphamat}
\begin{bmatrix}
S \\
S^2 \\
S^{2^2} \\
\vdots \\
S^{2^{k-1}}
\end{bmatrix}
&=
\begin{bmatrix}
1 & \alpha & \alpha^{2} & \cdots & \alpha^{{k-1}}\\
1 & \alpha^{2} & \alpha^{4} & \cdots & \alpha^{2(k-1)} \\
1 & \alpha^{4} & \alpha^{8} & \cdots & \alpha^{4(k-1)}\\
\vdots & \vdots & \vdots & \ddots & \vdots \\
1 & \alpha^{2^{k-1}} & \alpha^{2\cdot 2^{k-1}} & \cdots & \alpha^{(k-1)\cdot 2^{k-1}}
\end{bmatrix}
\begin{bmatrix}
s_0\\
s_1\\
s_2\\
\vdots\\
s_{k-1}
\end{bmatrix}
\end{align*}
\item Transition function $f_T: T+\F(S,x)$
\item Elimination on ideal $\langle f_T,f_S \rangle+J_0'$ using $S,x > T$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment results: word-level traversal}}
\begin{table}[H]
\centering
\caption{Results of running benchmarks using our tool. 
\small{Parts I to III denote the time taken by polynomial divisions,
  bit-level to word-level abstraction and iterative reachability
  convergence checking part of our approach, respectively.}}
{\small 
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
\multirow{3}{*}{\centering Benchmark} 
 & \multirow{3}{0.9cm}{\centering \# States}
 & \multirow{3}{1.3cm}{\centering \# iterations}
 & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
 & \multirow{3}{1.8cm}{\centering Runtime of VIS (sec)} \\
   & & &\multicolumn{3}{c|}{}& \\
  \cline{4-6}
     & & & I & II & III & \\
\hline
\hline
b01 & 18  & 5  & $<0.01$ & 0.01 & 0.02 & $<0.01$\\
b02 & 8 & 5 & $<0.01$  & 0.01 & $<0.01$ & $<0.01$ \\
b06 & 13 & 4 & $<0.01$ & 0.07 & 5.0 & $<0.01$ \\
s27 & 6 & 2 & $<0.01$ & 0.01 & 0.02  & $<0.01$  \\
s208 & 16 & 16 & $<0.01$ & 0.32 & 2.4 & $<0.01$ \\
s386 & 13  & 3 & 1.0 & 7.6 & 8.2  & $<0.01$ \\
bbara & 10 & 6 & 0.04 & 0.01 & 0.04  & $<0.01$ \\
beecount & 7  & 3  &$<0.01$ & 0.01 & 0.01 & $<0.01$ \\
dk14 & 7  & 2  & 45 & $<0.01$ & 0.08 & $<0.01$\\
donfile & 24 & 3  & 12316 & 0.02 & 1.7  & $<0.01$\\
\hline
\end{tabular}
}
\label{tab:result}  
\end{table} 
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = Moore]{\large{Apply FSM traversal to arithmetic ckts}}
\vspace{-0.1in}
\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=4.5in]{../newfig/Moore.eps}
% \vspace{-0.2in}
%\caption{A simple Moore FSM and its state transition graph}
%\end{minipage}
\label{fig:Moore}}
\end{figure}
\vspace{-0.4in}
\bi
\item Model: restricted Moore finite state machine
	\bi
	\item Some sequential arithmetic circuits will give results after running for $k$ clock cycles
	\item The initial operands are preloaded in register files
	\ei
\item State transitions on this model:
$$R_{k} = Tr(R_{k-1}) = Tr(Tr(\cdots Tr(R_{init})\cdots)) = Tr^k(R_{init})$$
\item Word-level unrolling
\ei
\hyperlink{motiv2}{\beamergotobutton{Recall unrolling}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Galois field multiplier}}
SPEC: $R = A_{init}\cdot B_{init} \pmod {P(\alpha)}$ after $k$ clock cycles
\hspace{-0.3in}\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{./new_multi.eps}
% \vspace{-0.2in}
\caption{A 3-bit RH-SMPO and its Moore FSM model}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}
\bi
\item Projection on NS $R',A',B'$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{GF multiplier verification algo.}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	\alert{$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle)$ with ATO}\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit verification}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%(copy from DATE18~21)%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment on 3-bit RH-SMPO}}
\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=1.5in]{./RH3.eps}
% \vspace{-0.2in}
%\caption{Gate-level circuit of 3-bit RH-SMPO}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}
\vspace{-0.2in}
\bi
\item The elimination ideal (first iteration):
\begin{align*}
J = &d_0+b_2\cdot a_2,
c_1+a_0+a_2,
c_2+b_0+b_2,
d_1+c_1\cdot c_2,\\
&e_0+d_0+d_1,
e_2+d_1,
r_0'+r_2+e_0,
r_1'+r_0,
r_2'+r_1+e_2,\\
&A+a_0\beta+a_1\beta^2+a_2\beta^4,
B+b_0\beta+b_1\beta^2+b_2\beta^4,\\
&R+r_0\beta+r_1\beta^2+r_2\beta^4,
R'+r_0'\beta+r_1'\beta^2+r_2'\beta^4;
\end{align*}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment on 3-bit RH-SMPO(2)}}
\bi
\vspace{-3.45cm}
\item $J_0 = \langle x_i^2 - x_i, X^q - X\rangle$
\item $from_0 = \{R, A_{init}+a_0\beta+a_1\beta^2+a_2\beta^4,
B_{init}+b_0\beta+b_1\beta^2+b_2\beta^4\}$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Basic algorithm to verify the function of sequential GF multipliers}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle)$ with ATO\;
	\alert{$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$}\;
	\alert{$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$}\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit verification}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Experiment on 3-bit RH-SMPO(2)}}
\bi
\item $J_0 = \langle x_i^2 - x_i, X^q - X\rangle$
\item $from_0 = \{R, A_{init}+a_0\beta+a_1\beta^2+a_2\beta^4,
B_{init}+b_0\beta+b_1\beta^2+b_2\beta^4\}$ ($\beta = \alpha^3$)
\item $to_1 : 
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
$
\item $from_1 = \{
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
, A_{init}+a_2\alpha^3+a_0\alpha^6+a_1\alpha^{12},
B_{init}+b_2\alpha^3+b_0\alpha^6+b_1\alpha^{12}\}$
\item $\cdots$
\item After 3 iterations: $to_3 = \{ \alert{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Improve using RATO}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	\alert{$f_o \xrightarrow{J + J_0+ from_{i-1}(R,A,B)}_+ f_r$ under RATO} \;
	\alert{$to_i(R',A',B')\gets f_r(\{R',A',B'\}\setminus\{r_0,\dots,r_{k-1},a_0,\dots,a_{k-1},b_0,\dots,b_{k-1}\}$}\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit verification}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment result: sequential GF multiplier verification}}
\begin{table}[H]
\centering
\caption{\small Run-time for verification of bug-free RH-SMPO circuits
  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
\label{tbl:equiv}
\begin{tabular}{|c||c|c|c|c|} 
\hline
& \multicolumn{4}{|c|}{Word size of the operands $k$-bits}  \\
\hline
Solver & 11 & 18 & 23 & 33 \\
\hline
\hline
Lingeling & 593  & \emph{TO}  & \emph{TO}  & \emph{TO}\\
\hline
\hline
ABC & 6.24 & \emph{TO} & \emph{TO} & \emph{TO}\\
\hline
\hline
BDD & 0.1 & 11.7 & 1002.4 & \emph{TO}  \\
\hline
\end{tabular}
\label{table:satbdd}  
\end{table} 
\begin{table}[H]
\centering
\label{tab:Cpp}
{\small 
\begin{tabular}{|c|c||c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c||}{\centering Operand size $k$} & 36 & 60 & 81 & 100 & 131 & 162 \\
\hline
\multirow{2}{1.5cm}{\centering RH-SMPO} & \#Polynomials & 4716 & 12960 & 21870 & 35600 & 56592 & 92826 \\
\cline{2-8}
 & Runtime & 14.3 & 213.3 & 1343 & 4685 & 26314 & 124194 \\
\hline
\multirow{2}{1.5cm}{\centering Agnew's SMPO} & \#Polynomials & 2700 & 7380 & 13356 & 20300 & 34715 & 52974 \\
\cline{2-8}
 & Runtime & 10.2 & 212.0 & 2684 & 4686 & 56568 & 119441 \\

 \hline

\end{tabular}
}
\end{table}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = refine]{\large{Abstraction refinement}}
\begin{figure}[hbt]
\centering{
\includegraphics[width=4.5in]{../newfig/refine.eps}
\caption{Abstraction by reducing latches}
\label{fig:refine}}
\end{figure}
\bi
\item Remove ``irrelevant" latches, reduce state space
\item Provide an over-approximation
\item This algorithm requires \alert{UNSAT core extraction}
\ei
\hyperlink{motiv3}{\beamergotobutton{Recall $k$-BMC}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Refutation within Buchberger's algo.}}
\bi
\item Recall Buchberger's algo....
\pause
\item  What can we learn from the \alert{``1" generated by Buchberger's Algorithm}?
\ei
\begin{Theorem}
\label{thm:weak-ns}
Let $J = \langle f_1,\dots, f_s \rangle $ be
an ideal in the ring $ \mathbb{F}[x_1, \dots, x_d]$ and $V_{\overline{\F}}(J)$ be its
variety over $\overline{\F}$. Then $V_{\overline{\F}}(J) = \emptyset \iff J =
\mathbb{F}[x_1,\dots,x_d] \iff 1 \in J$.  
\end{Theorem}
$V_{\overline{\F}}(J)=\emptyset \iff 1\in J \iff reduced~GB(J) = \{1\}$

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Motivating Example}}
\bi
\item $f_1\sim f_9 \in \mathbb{F}_2[a,b,c,d]$
\item $F = \{f_1,f_2,\dots,f_9\}$
\ei

\begin{minipage}{2in}
\begin{align*}
f_1&: abc + ab + ac + bc \\
   & + a + b + c + 1\\
f_2&: b\\
f_3&: ac\\
f_4&: ac + a
\end{align*}
\end{minipage}
%\hspace{0.1in}
\begin{minipage}{2in}
\begin{align*}
f_5&: bc + c\\
f_6&: abd + ad + bd + d\\
f_7&: cd\\
f_8&: abd + ab + ad + bd + a + b + d + 1\\
f_9&: abd + ab +bd + b
\end{align*}
\end{minipage}
\vspace{1cm}
\bi
\item Find a subset $F_c \subset F$ {\it s.t.} $F_c$ remains UNSAT 
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Motivating Example}}
Following Spoly selection strategy $(f_1,f_2)\to (f_1,f_3) \to (f_2,f_3)\to (f_1,f_4) \to \cdots$,
execute Buchberger's algorithm until adding ``1" to Gr\"obner basis
\bi
\item $Spoly(f_1,f_2)= f_1 - ac\cdot f_2\xrightarrow{a\cdot f_2} ~\xrightarrow{f_3}~
\xrightarrow{c\cdot f_2}  ~\xrightarrow{f_2} f_{10}=a+c+1$
\item $Spoly(f_1,f_3)\xrightarrow{F}_+ 0$
\item \dots
\item $Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$
\item $Spoly(f_2,f_5)\xrightarrow{f_{11}} 1$
\ei
$\{f_1,\dots,f_9,f_{10},f_{11},1\}$ is the Gr\"obner basis generated from Buchberger's algorithm
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Note: this frame contains multiple pages for animation purpose

\begin{frame}{\large{Motivating Example: Refutation Tree}}
\vspace{-0.1in}
\includegraphics[width=3.5in]{./refutation_1.pdf}
\pause
\bi
\item $f_{10}=f_1-acf_2-af_2-f_3-cf_2-f_2 = f_1+acf_2+af_2+f_3+cf_2+f_2$
\item $1 = \F(f_1,f_2,f_3,f_4,f_5)\implies 1 \in \langle f_1,f_2,f_3,f_4,f_5\rangle \implies UNSAT$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size: Redundancy in Refutation Tree}}
\bi
\item Expand: \begin{align*}
		1 &= \F(f_1,f_2,f_3,f_4,f_5)\\
		&= cf_2+f_5+\alert{f_{11}}\\
		&=cf_2+f_5+f_3+f_4+\alert{f_{10}}\\
		&=(cf_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot
  f_3}+ \dots + (f_1 + acf_2)
		\end{align*}
\item UNSAT core reduced to $\{f_1,f_2,f_4,f_5\}$ which is {\bf minimal}
\item \alert{GB-core algorithm}:
	\bi
	\item Execute Buchberger's algorithm
	\item Recording data including Spoly, polynomials for division and remainder
	\item Terminate Buchberger's algorithm after recording remainder ``1"
	\item Build refutation tree and get UNSAT core
	\item Analyze recorded data, remove redundant polynomials from the core
	\ei
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size: Iterative Refinement}}
\bi
\item Spoly pair selection strategy: $(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$
\item High likelihood in minimal core $\to$ Put ahead in Spoly queue $\to$ Faster approaching ``1" in GB-core $\to$ Smaller core
\ei
\vspace{-0.2in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{refutation_2.pdf}
%\caption{Refutation distance \& frequency}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\bi
\item {\bf Refutation distance}: shortest path to leaf
\item Distance$\downarrow \to$ LT Degree$\downarrow \to$ Likelihood in minimal core $\uparrow$
\vspace{0.2in}
\item {\bf Frequency}: number of times $f_i$ appears in refutation tree
\ei
\end{column}
\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% use "alignat" to align to multiple columns

\begin{frame}{\large{Iterative Refinement Example}}
\vspace{-0.1in}
\begin{alignat*}{4}
& f_1: x_1x_3+x_3 && ~~~f_2: x_2 + 1 && ~~~f_3: x_2x_3+x_2 \\
& f_4: x_2x_3 && ~~~f_5: x_2x_3 + x_2 + x_3 + 1 && ~~~f_6 : x_1x_2x_3 +x_1x_3
\end{alignat*}
\vspace{-0.5in}
\begin{columns}[onlytextwidth]
\begin{column}{0.6\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.3]{iterative_1.pdf}
\end{figure}
\end{column}
\begin{column}{0.4\textwidth}
\bi
\item $F=\{f_1,f_2,\dots,f_6\}\in\mathbb{F}_2[x_1,x_2,x_3]$
\item UNSAT core: $f_1,f_2,f_3,f_4$
\item \alert{Refutation distance}: \\
		$f_2~~f_1~~f_3~~f_4$\\
		$2~~~3~~~3~~~3$
\item \alert{Frequency}:\\
		$f_1~~f_3~~f_4$\\
		$1~~~1~~~1$
\item Reorder: $\alert{f_2},f_1,f_3,f_4$
\ei
\end{column}
\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Iterative Refinement Example}}
\vspace{-0.1in}
\begin{alignat*}{4}
& f_1: x_1x_3+x_3 && ~~~f_2: x_2 + 1 && ~~~f_3: x_2x_3+x_2 \\
& f_4: x_2x_3 && ~~~f_5: x_2x_3 + x_2 + x_3 + 1 && ~~~f_6 : x_1x_2x_3 +x_1x_3
\end{alignat*}
\vspace{-0.5in}
\begin{columns}[onlytextwidth]
\begin{column}{0.6\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.3]{iterative_2.pdf}
\end{figure}
\end{column}
\begin{column}{0.4\textwidth}
\bi
\item Initial order: $f_2,f_1,f_3,f_4$
\item Spoly pairs selection:\\
	$(\alert{f_2},f_1)\to(\alert{f_2},f_3)\to(f_1,f_3)\to(\alert{f_2},f_4)\to(f_1,f_4)\to(f_3,f_4)$
\item UNSAT core: $f_2,f_3,f_4$
\item \alert{Fixpoint reached}
\ei
\end{column}
\end{columns}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Reducing Size Further using Syzygy Heuristic}}
\bi
\item Finding interdependencies: $f_i\in\langle F\setminus\{f_i\}\rangle$?
\item Given $F=\{f_1,\dots,f_s\}$, find $f_i = \sum_{j\neq i} h_jf_j$
\item Info lost in Buchberger's algorithm?
\item Inner loop of Buchberger's algorithm: \\
\hspace{0.2in}  $Spoly(f, g) \stackrel{G'}{\textstyle\longrightarrow}_+
r$ \\
\hspace{0.2in}  \alert{IF $r \neq 0$} THEN $G:= G \cup \{r\}$ \\
\hspace{0.2in}  \alert{IF $r = 0$} THEN Discard

\item We collect division info when $r = 0$:
	\bi
	\item Record data for Spoly and polynomial division as in GB-core algorithm
	\ei
\item $Spoly(f_i,f_j)\xrightarrow{F}_+ 0\implies \alert{c_1f_1+c_2f_2+\cdots+c_sf_s = 0}$
\pause
\item $(c_1,c_2,\dots,c_s)$ is a {\bf syzygy} on $(f_1,f_2,\dots,f_s)$
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size Further using Syzygy Heuristic}}
\bi
\item In a single syzygy, $c_i=1 \implies f_i = \sum_{j\neq i} h_jf_j$
\item In general cases, need to analyze all syzygies recorded
\item Collect $m$ syzygies as a system of polynomial equations, or a {\bf Syzygy Matrix}
\ei
\hspace{-0.1in}
\begin{minipage}{2.5in}
\begin{equation*} \label{eqn:syz}
 \begin{cases}
 c_1^1f_1+c_2^1f_2+\cdots+c_s^1f_s = 0\\
 c_1^2f_1+c_2^2f_2+\cdots+c_s^2f_s  = 0\\
 \ \ \ \ \ \  \vdots \\
 c_1^mf_1+c_2^mf_2+\cdots+c_s^mf_s = 0  
 \end{cases}
\end{equation*}
\end{minipage}
\hspace{-0.3in}
\begin{minipage}{2in}
%which can be represented in matrix form as:
\begin{center}
\begin{align*}
\label{mat:syzygy}
   \begin{bmatrix}
           c_1^1 & c_2^1 & \cdots & c_s^1 \\
           c_1^2 & c_2^2 & \cdots & c_s^2 \\
           \vdots & \vdots & \ddots & \vdots \\
           c_1^m & c_2^m & \cdots & c_s^m
         \end{bmatrix}
    \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
         &= 0
  \end{align*}

\end{center}
\end{minipage}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size: Revisiting Motivating Example with Syzygy}}
\vspace{-0.2in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{refutation_tree.pdf}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\bi
\item Recorded in Buchberger's Algo:\\
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$\\
$Spoly(f_3,f_4)\xrightarrow{F}_+ f_{11}$\\
$Spoly(f_2,f_5)\xrightarrow{F}_+ 1$
\item Discarded in Buchberger's Algo:\\
$Spoly(f_1,f_3)\xrightarrow{F}_+ 0$\\
$Spoly(f_2,f_3)\xrightarrow{F}_+ 0$\\
$\dots$\\
$Spoly(f_1,f_5)\xrightarrow{F}_+ 0$
\ei
\end{column}
\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Syzygy Example}}
\bi
\item Syzygy matrix:
\ei
\includegraphics[width=4.5in]{./syzygy_1.pdf}
\pause
\llap{\includegraphics[width=4.5in]{./syzygy_2.pdf}}
\bi
\item $J = \langle f_1,f_2,\dots,f_9\rangle$
\item $f_{10}\in J$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Syzygy Example}}
\bi
\item Considering $f_{10}=f_1-acf_2-af_2-f_3-cf_2-f_2 = f_1+acf_2+af_2+f_3+cf_2+f_2$
\ei
\includegraphics[width=4.5in]{./syzygy_3.pdf}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Syzygy Example}}
\includegraphics[width=4.5in]{./syzygy_4.pdf}
\bi
\item Column $f_3$ contains ``1", means:\\
$f_3 = ac f_2 + a f_5$
\item Generalization of this strategy refer to the paper
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Overall Approach}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{A set of UNSAT polynomials $F$} 
 \KwOut{A subset $F'\subset F$ remains UNSAT}
 $G \gets F$\;
  \Repeat{$G=F'$}
  {
	$F' = G$\;
  	$G \gets $GB-core($F'$ with order $>$)\;
	Update order $>$\;
  }
  $F'\gets$syzygy\_heuristic($G$)\;
\Return{$F'$}
\caption {UNSAT core extraction based on Gr\"obner basis algorithm}
\end{algorithm}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment results: UNSAT core extraction}}
\centering{Table of selected benchmarks}\\
\centering{I:single GB-core; II: Iterative GB-core; III: syzygy}
\includegraphics[width=4.5in]{./table_cp2016.pdf}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Application to abstraction refinement}}
\vspace{-0.1in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/s27_stg.pdf}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/s27_stg_refined.pdf}
\end{figure}
\end{column}
\end{columns}
\bi
\item By removing ``irrelevant" latches, state-space reduced
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Conclusion}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Future work}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Publications \& tools}}
\end{frame}

\end{document}

### 2016-12-14 dissertation/slides/defense.tex
dissertation defense slides v1: milestone

@@
 \usepackage{xspace}
+\usepackage{polynom}
 %\usepackage{columns}
@@
 	\item Polynomial algebra \& Algebraic geometry
-	\item Projection
+	\item Projection of varieties
 	\ei
@@
 \bi
-\item Word-level reachability analysis -- analog if implicit state enumeration
+\item Word-level reachability analysis -- analog of implicit state enumeration
 	\bi 
@@
 	\pause
-	\item {\bf A:} Datapath $\gets$ word-level info 
+	\item {\bf A:} Data $\gets$ word-level info 
 	\pause
@@
 \pause
-\item Apply word-level reachability algo. to sequential arithmetic circuit verification
+\item Apply word-level reachability algorithm to sequential arithmetic circuit verification
 	\bi
 	\item Abstraction $\to$ word-level signature each time-frame
+	\bi
+            \item Word-level abstraction from bit-level ckts [Pruss, 2015]
+        \ei
 	\item Word-level unrolling
@@
 	\item Structure of refutation proof $\to$ refine UNSAT core
-	\item Apply UNSAT core to $k$-BMC abstraction refinement
 	\ei
@@
 	\bi
+	\item Usually based on reachability analysis
 	\item Sequential miter
 	\item Unroll, then use DDs, SAT or AIGs (Combinational)
+	\item Induction-based
 	\ei
 \item Symbolic model checking (counterexample, IC3): SAT/BDDs in nature
-\item Word-level techniques (term rewriting, uninterpreted function): ?
+\item Word-level techniques (term rewriting, uninterpreted function): ?/no encoding
 \item Algebraic geometry methods
@@
 \item No purely word-level sequential verification: data/abstraction/algorithm
+\ei
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{\large{Working field: $\Fkk$}}
+\bi
+\item Our proposed state-space model is based on finite field $\Fkk$
+	\bi
+	\item $\B^k$: bit-vector
+	\item $\Z_{2^k}$, $\R_q$: approaches not compatible
+	\ei
+\item Evaluations in $\B^k \Leftrightarrow$ Elements in $\Fkk$
+\item Functions $\B^k\to\B^k \Leftrightarrow \Func:\Fkk\to\Fkk$
 \ei
@@
 \pause
-\bi
-\item A polynomial function representation always exists in $\Fq[A]$ (Lagrange's interpolation)
-\ei
+\vspace{-0.2in}
+% \bi
+% \item Always possible to generate word-level poly function using Lagrange's interpolation
+% \ei
+\begin{align*}
+Z =& \Func(A) \\
+=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
+&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
+\end{align*}
 \end{frame}
@@
 \begin{frame}{\large{Preliminaries: Polynomial division}}
-Copy page 12 from FMCAD
+Divide $f = x^3-2x^2+2x+8$ ~~by ~~$g =2x^2+3x + 1$\\
+
+\pause
+
+\polylongdiv{x^3-2x^2+2x+8}{2x^2+3x + 1}
+
+\pause
+
+\bi
+\item The key step in division: \alert{$r = f - \frac{lt(f)}{lt(g)}\cdot  g$}, denoted $f\xrightarrow{g} r$
+\item Similarly divide $f$ by a set of polynomials $F = \{f_1,\dots, f_s\}$
+\item Denoted: $f \xrightarrow{f_1, \dots, f_s}_+ r$
+\bi
+\item Remainder $r$ is reduced: no term in $r$ is divisible by $lt(f_i)$
+\ei
+\ei
 \end{frame}
@@
 \begin{itemize}
-\item $\mathbb{F}_q[x_1, \ldots, x_n]$: ring of all polynomials with
-  coefficients in $\mathbb{F}_q$ 
+% \item $\mathbb{F}_q[x_1, \ldots, x_n]$: ring of all polynomials with
+%   coefficients in $\mathbb{F}_q$ 
 \item Given a set of polynomials:
 \begin{itemize}
-\item $f, f_1, f_2, \dots , f_s \in \mathbb{F}_q[x_1, \dots, x_n]$
+\item $f_1, f_2, \dots , f_s \in \mathbb{F}_q[x_1, \dots, x_n]$
 \item Find solutions to $f_1 = f_2 = \dots = f_s = 0$
@@
 \begin{eqnarray}
-J = \langle f_1, f_2 \ldots, f_s\rangle = \{f_1 h_1 + f_2 h_2 + \dots + f_s h_s\} \nonumber 
+J = \langle f_1, f_2 \ldots, f_s\rangle = \{f_1 h_1 + f_2 h_2 + \dots + f_s h_s\},~~h_i\in\Fq[x_1,\dots,x_n] \nonumber 
 \end{eqnarray}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Preliminaries: Buchberger's algorithm}}
+\begin{frame}{\large{Preliminaries: Buchberger's algorithm computes a Gr\"obner basis}}
  INPUT : $F = \{f_1, \dots, f_s\}$\\
@@
 \pause
-\item Inspiration of mapping: set of states $\to$ \alert{variety} of polynomial ideal
+\item GB enables mapping: set of states $\to$ \alert{variety} of polynomial ideal
+	\bi
+	\item Algebraic/reasoning engine: application to elimination
+	\ei
+\item GB enables abstraction
 \ei
@@
 \pause
-\item GB($I_{x,y,z}$)$\cap\Fq[z]\Leftrightarrow$ \alert{projection} on variable $z$!
+\item GB($I_{x,y,z}$)$\cap\Fq[z]$ related to \alert{projection} on variable $z$!
 \end{itemize}
@@
 \bi
-\item Given ideal $J$, compute $J'$ s.t. $V(J') = \overline{V(J)}$
-\item $\to J'=J_0:J$
+\item $V(J)\subseteq \Fkk$ in affine space
+\item Given ideal $J$, compute $J'$ s.t. $V(J') = \overline{V(J)} = \Fkk-V(J)\implies J'=J_0:J$
 \ei
@@
 \item {\bf Iteration 1:}Compose an elimination ideal $J$ 
-\vspace{-0.2in}\\
+\ei
+\vspace{-0.2in}
+\begin{columns}[onlytextwidth]
+\begin{column}{0.5\textwidth}
 \begin{align*}
-&f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
+&f_1: t_0- (xs_0s_1+xs_0\\
+&+xs_1+x+s_0+s_1+1)\\
 &f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
@@
 \end{align*}
-$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$
+\end{column}
+\vspace{-0.2in}
+\begin{column}{0.5\textwidth}
 \begin{align*}
@@
 \end{align*}
-$\ \ \ \ \ \ \ \ \ \ \  \ J_0 = \langle f_5,f_6,\dots,f_{11}\rangle$
-\ei
+\end{column}
+\end{columns}
+\begin{columns}[onlytextwidth]
+\begin{column}{0.7\textwidth}
+\begin{figure}[H]
+\centering{
+\includegraphics[width=3in]{hehe.pdf}
+}
+\end{figure}
+\end{column}
+\begin{column}{0.3\textwidth}
+$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$\\
+$J_0 = \langle f_5,f_6,\dots,f_{11}\rangle$
+\end{column}
+\end{columns}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%(End of copy)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Improve complexity using RATO}}
+\begin{frame}{\large{Improve complexity using RATO [Pruss, '15]}}
 \bi
@@
 \item RATO: reverse topological traverse on ckt structure
-\item Benefit: Only one pair of poly with non-relatively-prime leading terms\par
-$gcd(lt(f_w),lt(f_g))\neq 1$
 \ei
@@
 \end{figure}
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{\large{Benefits of using RATO}}
+\bi
+\item Definition of RATO:\par 
+``bit-level variables ordered reverse topologically "$ > T>S$
+\item Only one pair of poly with non-relatively-prime leading terms
+	\bi
+	\item If $gcd(lt(f_w),lt(f_g)) = 1$:
+	\item $Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_+ 0$
+	\ei
+\item Topology in structure
+	\bi
+	\item Each poly with leading term -- fanout
+	\item Divide with levelization
+	\item Only inputs (primary \& pseudo) left!
+	\ei
+\ei
 \end{frame}
@@
 \bi
-\item $Spoly$ reduction gives $T+\Func(inputs)$
+\item $Spoly$ reduction gives $T+\Func(primary/pseudo~inputs)$
 \ei
@@
 %\begin{minipage}{12cm}
-\includegraphics[width=5.5in]{./new_multi.eps}
-% \vspace{-0.2in}
-\caption{A 3-bit RH-SMPO and its Moore FSM model}
+\includegraphics[width=5in]{./new_multi.eps}
 %\end{minipage}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{GF multiplier verification algo.}}
+\begin{frame}{\large{GF multiplier verification algorithm}}
 \begin{algorithm}[H] % note [hbt] will fail in beamer
@@
 \begin{frame}{\large{Experiment result: sequential GF multiplier verification}}
+\bi
+\item Run-time for verification of bug-free RH-SMPO circuits
+  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs
 \begin{table}[H]
 \centering
-\caption{\small Run-time for verification of bug-free RH-SMPO circuits
-  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
 \label{tbl:equiv}
@@
 \end{table} 
-\begin{table}[H]
+\item Runtime for verification of bug-free Agnew's and RH-SMPO circuits using our approach
+\hspace{-0.3in}\begin{table}[H]
 \centering
@@
 \hline
-\multirow{2}{1.5cm}{\centering RH-SMPO} & \#Polynomials & 4716 & 12960 & 21870 & 35600 & 56592 & 92826 \\
+\multirow{2}{1cm}{\centering RH-SMPO} & \#Polys & 4716 & 12960 & 21870 & 35600 & 56592 & 92826 \\
 \cline{2-8}
@@
 \hline
-\multirow{2}{1.5cm}{\centering Agnew's SMPO} & \#Polynomials & 2700 & 7380 & 13356 & 20300 & 34715 & 52974 \\
+\multirow{2}{1cm}{\centering Agnew's SMPO} & \#Polys & 2700 & 7380 & 13356 & 20300 & 34715 & 52974 \\
 \cline{2-8}
@@
 \end{table}
+\ei
 \end{frame}
@@
 \end{figure}
+\vspace{-0.2in}
 \bi
@@
 \bi
-\item Recall Buchberger's algo....
+\item Recall Buchberger's algorithm
 \pause
@@
 \vspace{-0.1in}
+\begin{figure}
+\centering
+\includegraphics[scale=0.35]{../newfig/s27.pdf}
+\end{figure}
+\vspace{-0.3in}
+\bi
+\item PS = $\{G7,G6,G5\}$, NS = $\{G13,G11,G10\}$: 8 states
+\item Property $p = {\bf AG}( (\neg G13) {\bf U} (\neg Z) )$
+\item $k$-BMC without abstraction refinement: when $k=3$ prove $PASS$
+\ei
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{\large{Application to abstraction refinement}}
+\bi
+\item Circuit ideal when $k=0$:
+\ei
+\begin{align*}
+I = \langle & G14+1+G0, G8+G14\cdot G6, G15+G12+G8+G12\cdot G8,\\ 
+	& G16+G3+G8+G3\cdot G8, G9+1+G16\cdot G15, \\ 
+	& G10+1+G14+G11+G14\cdot G11, G11+1+G5+G9+G5\cdot G9, \\
+	& G12+1+G1+G7+G1\cdot G7, G13+1+G2+G12+G2\cdot G12, \\
+	& Z+1+G1, \\
+	& \text{(Initial state 000 )} G5, G6, G7 \rangle;
+\end{align*}
+\vspace{-0.2in}
+\bi
+\item Property: $\neg p = Z\cdot G13 + 1$
+\item UNSAT core: \begin{align*}
+Core(I\land \neg p) =& G12+1+G1+G7+G1\cdot G7,\\
+&G13+1+G2+G12+G2\cdot G12, \\
+& Z+1+G1, G7;
+\end{align*}
+\ei
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{\large{Application to abstraction refinement}}
+\vspace{-0.1in}
 \begin{columns}[onlytextwidth]
@@
 \bi
-\item By removing ``irrelevant" latches, state-space reduced
+\item $\{G5/G10 ,G6/G11\}$: irrelevant
+\item By removing irrelevant latches, state-space reduced
 \ei
@@
 \begin{frame}{\large{Conclusion}}
+\bi
+\item Word-level abstraction of state-space
+\item Apply to reachability analysis
+\item Prove effectiveness by experiments on ISCAS'89 and ITC'99 benchmarks
+\ei
+\bi
+\item Word-level abstraction of function in a single time-frame
+\item Word-level unrolling
+\item Apply to functional correctness checking of sequential GF multipliers
+\item Succeed to verify 162-bit, while contemporary fails beyond 23 bit 
+\ei
+\bi
+\item UNSAT core extraction for a set of polynomials
+\item Refine the core using refutation proof \& syzygies
+\item UNSAT core info can be applied to abstraction refinement
+\ei
 \end{frame}
@@
 \begin{frame}{\large{Future work}}
+\bi
+\item Multivariate polynomial ideals
+	\bi
+	\item Extend the application of univariate polynomial ideals
+	\ei
+\ei
+\bi
+\item Accelerate GB reduction
+	\bi
+	\item $F_4$ algorithm on term-sparse polynomial ideal (parallel computing)
+	\item ZDDs can represent chain of OR gates logic in linear space complexity (alternative canonical graphic representation)
+	\ei
+\ei
+\bi
+\item Compute Craig's interpolants using algebraic geometry
+	\bi
+	\item Projection of varieties $\implies$ interpolants
+	\ei
+\ei
 \end{frame}
@@
 \begin{frame}{\large{Publications \& tools}}
+\bi
+\item Publications:
+	\bi
+	\item {\it Formal Verification of Sequential Galois Field Arithmetic Circuits using Algebraic Geometry}. {\bf Xiaojun Sun}, Priyank Kalla, Tim Pruss, Florian Enescu. DATE 2015, Grenoble
+	\item {\it Finding Unsatisfiable Cores of a Set of Polynomials using the Groebner Basis Algorithm}. {\bf Xiaojun Sun}, Irina Ilioaea, Priyank Kalla, Florian Enescu. CP 2016, Toulouse
+	\item {\it Word-level Traversal of Finite States Machines using Algebraic Geometry}. {\bf Xiaojun Sun}, Priyank Kalla, Florian Enescu. HLDVT 2016, Santa Cruz
+	\ei
+\item Tools:\\
+My website: {\it ece.utah.edu/$\sim$xiaojuns/code.html}
+\ei
 \end{frame}

### 2016-12-15 dissertation/slides/defense.tex
dissertation: defense slides v2

@@
 \hyperlink{Moore}{\beamergotobutton{Go back to Moore machine}}
+\hyperlink{expSMPO}{\beamergotobutton{Go back to Exp result}}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Preliminaries: Buchberger's algorithm computes a Gr\"obner basis}}
+\begin{frame}[label = Buch]{\large{Preliminaries: Buchberger's algorithm computes a Gr\"obner basis}}
  INPUT : $F = \{f_1, \dots, f_s\}$\\
@@
 \ei
+\hyperlink{linkRATO}{\beamergotobutton{RATO}}
 \end{frame}
@@
 \bi
-\item Directly compute GB($J_{ckt}+J_0$) in $\Fq$ is costly ($q^{O(d)}$)
+\item Directly compute GB($J_{ckt}+J_0$) in $\Fq$ is costly ($q^{O(n)}$)
 \item GB sensitive to term order$\to$Transform to simpler set for GB?
@@
 \end{figure}
+\hyperlink{Buch}{\beamergotobutton{Buchberger}}
 \end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Benefits of using RATO}}
+\begin{frame}[label = linkRATO]{\large{Benefits of using RATO}}
 \bi
@@
 ``bit-level variables ordered reverse topologically "$ > T>S$
+\item Topology in ckt structure
+	\bi
+	\item Output of each gate $= lt(f_i)$
+	\ei
+\item Product criterion: $gcd(lt(f_i),lt(f_j)) = 1 \implies Spoly(f_i,f_j) \xrightarrow{J_{ckt}+J_0}_+ 0$
 \item Only one pair of poly with non-relatively-prime leading terms
-	\bi
-	\item If $gcd(lt(f_w),lt(f_g)) = 1$:
-	\item $Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_+ 0$
-	\ei
-\item Topology in structure
-	\bi
-	\item Each poly with leading term -- fanout
+
+\item $Spoly$ division
+	\bi
 	\item Divide with levelization
@@
 \item Objective: find $s_i = \mathcal G(S)$
-\item Build system of poly eqn by squaring:
+\item $(s_0+s_1\alpha+\cdots+s_{k-1}\alpha^{k-1})^{2^n} = s_0^{2^n}+(s_1\alpha)^{2^n}+\cdots+(s_{k-1}\alpha^{k-1})^{2^n}$
+\item Build system of poly eqns by squaring:
 \begin{align*}
@@
 \end{align*}
-\item Transition function $f_T: T+\F(S,x)$
-\item Elimination on ideal $\langle f_T,f_S \rangle+J_0'$ using $S,x > T$
-\ei
+\item Transition function $f_T: T+\Func(S,x)$
+\item Elimination on ideal $\langle f_T,f_S \rangle+J_0$ using $S,x > T$
+\ei
+\end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{\large{FSM traversal algorithm with RATO}}
+\IncMargin{1em}
+\begin{algorithm}[H]
+\SetAlgoNoLine
+\LinesNumbered
+\Indm
+ \KwIn{Polynomial ideal $J_{ckt}$, initial state polynomial $\Func(S)$}
+ \KwOut{Final reachable states represented by polynomial $\mathcal G(T)$}
+\Indp
+
+  $from^0 = reached = \Func(S)$\;
+  $f_T = $Reduce($Spoly(f_w,f_g), J_{ckt})$\;
+  \tcc{Compute $Spoly$ for the critical pair, then reduce it with circuit ideal under RATO}
+	Eliminate bit-level variables in $f_T$\;
+  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
+  {
+  	$i \gets i + 1$\;
+  	$G \gets$GB($\langle f_T , from^{i-1}\rangle+J_0^{PI}$)\;
+%  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest; $J_0^{PI}$ covers all possible inputs from PIs}
+	$to^i \gets G\cup \Fkk[T]$\;
+%	\tcc{There will be a univariate polynomial in $G$ denoting next state in word-level variable $T$}
+	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
+%	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
+  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
+%  	\tcc{Use product of ideals to attain a union of new reached states and formerly reached states}
+	$from^i \gets new^i(S\setminus T)$\;
+%	\tcc{Start a new iteration by replacing variable $T$ in new reached states with current state variable $S$}
+  }
+%  \tcc{Loop until a fix-point reached: newly reached state is empty}
+\Return{$\langle reached\rangle$}
+\caption {Refined Algebraic Geometry based FSM Traversal}\label{alg:refined}
+\end{algorithm}
+\DecMargin{1em}
+
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Experiment result: sequential GF multiplier verification}}
+\begin{frame}[label = expSMPO]{\large{Experiment result: sequential GF multiplier verification}}
 \bi
@@
 \ei
+\hyperlink{motiv2}{\beamergotobutton{Recall SEC miter}}
 \end{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Refutation within Buchberger's algo.}}
-\bi
-\item Recall Buchberger's algorithm
-\pause
-\item  What can we learn from the \alert{``1" generated by Buchberger's Algorithm}?
-\ei
-\begin{Theorem}
-\label{thm:weak-ns}
+\begin{frame}{\large{UNSAT problems: refutation within Buchberger's algorithm}}
+\begin{Theorem}[Weak Nullstellensatz]
 Let $J = \langle f_1,\dots, f_s \rangle $ be
-an ideal in the ring $ \mathbb{F}[x_1, \dots, x_d]$ and $V_{\overline{\F}}(J)$ be its
+an ideal in the ring $ \mathbb{F}[x_1, \dots, x_n]$ and $V_{\overline{\F}}(J)$ be its
 variety over $\overline{\F}$. Then $V_{\overline{\F}}(J) = \emptyset \iff J =
-\mathbb{F}[x_1,\dots,x_d] \iff 1 \in J$.  
+\mathbb{F}[x_1,\dots,x_n] \iff 1 \in J$.  
 \end{Theorem}
 $V_{\overline{\F}}(J)=\emptyset \iff 1\in J \iff reduced~GB(J) = \{1\}$
-
+\bi
+\item Recall Buchberger's algorithm: UNSAT $\to$ terminates with $1$
+\pause
+\item  Information lies in $Spoly$?
+\item Poly calculus stronger than resolution!
+\ei
+
+\begin{Definition}
+Assume a polynomial set $F$ is UNSAT. A subset $M \subseteq F$ is an {\bf UNSAT core} 
+if $M$ is also UNSAT. Further, if $\forall f\in M$, $M\setminus\{f\}$ is SAT, then 
+$M$ is called a {\bf minimal UNSAT core} of $F$.
+\end{Definition}
+
+Animation 2...
 \end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Motivating Example}}
+\begin{frame}{\large{Motivating example}}
 \bi
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Motivating Example}}
+\begin{frame}{\large{Motivating example}}
 Following Spoly selection strategy $(f_1,f_2)\to (f_1,f_3) \to (f_2,f_3)\to (f_1,f_4) \to \cdots$,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Reducing Size: Redundancy in Refutation Tree}}
+\begin{frame}{\large{Reducing size: redundancy in refutation tree}}
 \bi
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Reducing Size: Iterative Refinement}}
+\begin{frame}{\large{Reducing size: iterative refinement}}
 \bi
@@
 
-\begin{frame}{\large{Iterative Refinement Example}}
+\begin{frame}{\large{Iterative refinement Example}}
 \vspace{-0.1in}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Iterative Refinement Example}}
+\begin{frame}{\large{Iterative refinement: iteration 2}}
 \vspace{-0.1in}
@@
 \bi
-\item Initial order: $f_2,f_1,f_3,f_4$
+\item New order: $f_2,f_1,f_3,f_4$
 \item Spoly pairs selection:\\
@@
 
-\begin{frame}{\large{Reducing Size Further using Syzygy Heuristic}}
+\begin{frame}{\large{Reducing size further using syzygy heuristic}}
 \bi
@@
 \item Given $F=\{f_1,\dots,f_s\}$, find $f_i = \sum_{j\neq i} h_jf_j$
-\item Info lost in Buchberger's algorithm?
+\item Info lost in refutation tree/Buchberger's algorithm?
 \item Inner loop of Buchberger's algorithm: \\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Reducing Size Further using Syzygy Heuristic}}
+\begin{frame}{\large{Reducing size further using syzygy heuristic}}
 \bi
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Reducing Size: Revisiting Motivating Example with Syzygy}}
+\begin{frame}{\large{Reducing size: revisiting motivating example with syzygy}}
 \vspace{-0.2in}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Syzygy Example}}
+\begin{frame}{\large{Syzygy example: setup}}
 \bi
@@
 \bi
+\item rows: syzygies
 \item $J = \langle f_1,f_2,\dots,f_9\rangle$
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Syzygy Example}}
+\begin{frame}{\large{Syzygy Example: refine syzygy matrix}}
 \bi
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Syzygy Example}}
+\begin{frame}{\large{Syzygy example: derive interdependency}}
 \includegraphics[width=4.5in]{./syzygy_4.pdf}
@@
 \item Column $f_3$ contains ``1", means:\\
-$f_3 = ac f_2 + a f_5$
+$1\cdot f_3 + ac f_2 + a f_5 = 0\Leftrightarrow f_3 = acf_2+af_5$
 \item Generalization of this strategy refer to the paper
@@
 \item UNSAT core: \begin{align*}
-Core(I\land \neg p) =& G12+1+G1+G7+G1\cdot G7,\\
-&G13+1+G2+G12+G2\cdot G12, \\
-& Z+1+G1, G7;
+Core(I\land \neg p) =& G12+1+G1+\alert{G7}+G1\cdot \alert{G7},\\
+&\alert{G13}+1+G2+G12+G2\cdot G12, \\
+& Z+1+G1, \alert{G7};
 \end{align*}
@@
 \ei
+\pause
 \bi
@@
 \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{\large{Future work: Craig's interpolants}}
+\begin{figure}
+\centering
+\includegraphics[scale=0.3]{venn_craig.pdf}
+\end{figure}
+\vspace{-0.2in}
+\bi
+\item $ A \wedge B = \emptyset$, $A = (\overline{d})(\overline{c})(\overline{a}\vee
+d)$ and  $B = (a \vee b \vee c)(\overline{b})$
+\item $P = \overline{a}\wedge \overline{c}$ is an interpolant of $(A, B)$
+\item Find interpolant from resolution tree [K. Mcmillan '03]
+\item Algebraic geometry: projection in affine space
+\ei
+\end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\item {\it Word-level Traversal of Finite States Machines using Algebraic Geometry}. {\bf Xiaojun Sun}, Priyank Kalla, Florian Enescu. HLDVT 2016, Santa Cruz
+	\item Journal paper in preparation
 	\ei

### 2016-12-15 dissertation/slides/defense.tex
dissertation defense slides: final version

@@
 \end{figure}
+\hyperlink{pptpage2}{\beamergotobutton{Go back to example}}
 \end{column}
@@
 \end{columns}
-\hyperlink{pptpage2}{\beamergotobutton{Go back to example}}
+
 \end{frame}
@@
 \item Symbolic model checking (counterexample, IC3): SAT/BDDs in nature
-\item Word-level techniques (term rewriting, uninterpreted function): ?/no encoding
+\item Word-level techniques (term rewriting, uninterpreted function): no universal representation/no encoding
 \item Algebraic geometry methods
@@
 	\item $\B^k$: bit-vector
-	\item $\Z_{2^k}$, $\R_q$: approaches not compatible
+	\item $\Z_{2^k}$, $\R$: approaches not compatible
 	\ei
@@
 \item Functions $\B^k\to\B^k \Leftrightarrow \Func:\Fkk\to\Fkk$
+\pause
+\item State-space: finite set of $2^k$ points
+\item Treat them as solutions to polys in $\Fkk$
 \ei
@@
 \item $\mathbb{F}_{2^k} \equiv \mathbb{F}_{2}[x] \pmod {P(x)}$
-\item $P(x) \in \mathbb{F}_{2}[x]$, irreducible polynomial of degree
-  $k$
+\item $P(x) \in \mathbb{F}_{2}[x]$, irreducible polynomial of degree $k$
+\item Root $P(x)=0$: primitive element
+\bi
+\item E.g. $\C   = \R[x] \pmod{ x^2+1}$
+\ei
 \item Operations performed (mod $P(x)$) and coefficient reduced (mod 2)
-\bi
-\item E.g. $\C   = \R[x] \pmod{ x^2+1}$
-\ei
+
 \end{itemize}
@@
 \hyperlink{linkRATO}{\beamergotobutton{RATO}}
+\hyperlink{UNSAT}{\beamergotobutton{UNSAT}}
 \end{frame}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Gr\'obner basis with elimination term order}}
+\begin{frame}{\large{Gr\"obner basis with elimination term order}}
 \begin{itemize}
@@
  \KwIn{The circuit's characteristic polynomial ideal $J_{ckt}$,
-   initial state polynomial $\F(S)$, and LEX term order: bit-level
+   initial state polynomial $\Func(S)$, and LEX term order: bit-level
    variables $x, s, t$ $>$ PS word $S$ $>$ NS word $T$}
@@
 \end{figure}
-\hyperlink{Buch}{\beamergotobutton{Buchberger}}
+
 \end{frame}
@@
 \item Product criterion: $gcd(lt(f_i),lt(f_j)) = 1 \implies Spoly(f_i,f_j) \xrightarrow{J_{ckt}+J_0}_+ 0$
+\hyperlink{Buch}{\beamergotobutton{Buchberger}}
 \item Only one pair of poly with non-relatively-prime leading terms
@@
 \item $(s_0+s_1\alpha+\cdots+s_{k-1}\alpha^{k-1})^{2^n} = s_0^{2^n}+(s_1\alpha)^{2^n}+\cdots+(s_{k-1}\alpha^{k-1})^{2^n}$
-\item Build system of poly eqns by squaring:
+\item Build system of poly eqns by squaring poly:\\
+$S = s_0+s_1\alpha+\cdots+s_{k-1}\alpha^{2^{k-1}}$
 \begin{align*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{UNSAT problems: refutation within Buchberger's algorithm}}
+\begin{frame}[label = UNSAT]{\large{UNSAT problems: refutation within Buchberger's algorithm}}
 \begin{Theorem}[Weak Nullstellensatz]
@@
 $V_{\overline{\F}}(J)=\emptyset \iff 1\in J \iff reduced~GB(J) = \{1\}$
-\bi
-\item Recall Buchberger's algorithm: UNSAT $\to$ terminates with $1$
-\pause
-\item  Information lies in $Spoly$?
-\item Poly calculus stronger than resolution!
-\ei
-
 \begin{Definition}
@@
 \end{Definition}
+\bi
+\item Recall Buchberger's algorithm: UNSAT $\to$ terminates with $1$
+\hyperlink{Buch}{\beamergotobutton{Buchberger}}
+\pause
+\item  Information lies in $Spoly$?
+\item Poly calculus stronger than resolution!
+\ei
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frame}{\large{Reducing size: iterative refinement}}
-\bi
-\item Spoly pair selection strategy: $(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$
+\begin{frame}{\large{Reducing size: iterative refinement by reordering Spoly pairs}}
+\bi
+\item Spoly pair selection strategy: $(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$  (Animation 3)
 \item High likelihood in minimal core $\to$ Put ahead in Spoly queue $\to$ Faster approaching ``1" in GB-core $\to$ Smaller core

### 2016-12-17 dissertation/conclude.tex
dissertation: committee review version
dissertation:after defense_update 1

--- 2016-12-07 dissertation/conclude.tex
@@
 \label{ch:conclude}
-This dissertation first present a new approach to perform reachability
-analysis of FSMs at the word-level. It is facilitated by reproducing 
-the implicit state enumeration algorithm in computer algebra and algebraic
-geometry domain. The image function part is mapped to variety projection on 
+This dissertation presents a new approach to perform reachability
+analysis of FSMs at the word-level. It is facilitated by investigating the analog of 
+implicit state enumeration algorithm in computer algebra and algebraic
+geometry domains. The image function part of the computation is mapped to a variety projection on 
 next state variables. This projection is implemented by computing the 
-Gr\"obner basis of the elimination ideal with abstraction term orders.
-Moreover, the set operations in the state space is mapped to 
-the arithmetic of ideals, because algebraic geometry provides a 
-way to reason about the variety (solution) without actually solving the 
-system of polynomials. RATO is applied to improve the Gr\"obner bases 
-computation process, and a tool is developed to implement our 
+Gr\"obner basis of an elimination ideal with abstraction term orders.
+Moreover, the set operations in the state space are mapped to 
+the arithmetic of ideals, as algebraic geometry provides a 
+way to reason about the variety (solutions) by manipulating the ideals.
+A special term order (RATO) is utilized to improve the Gr\"obner bases 
+computation, and a tool is developed to implement our 
 word-level FSM traversal algorithm. Experiments are performed to 
-analyze the reachability for circuit benchmarks and gain success.
+analyze the reachability for ISCAS'89 and ITC'99 circuit benchmarks.
 
@@
 multipliers over $\Fkk$.  The core algorithm is based on word-level unrolling of 
-a Moore machine, which borrows ideas from our word-level FSM traversal 
-algorithm. As a result, bit-level transition relations are represented 
+a Moore machine, which applies concepts from the word-level FSM traversal 
+algorithm. As a result, transition relations are represented 
 as a polynomial with word-level inputs/outputs.
-We implement our algorithm with both \textsc{Singular} platform and C++ toolset 
-and perform experiments on Agnew's SMPO and RH-SMPO.
+We implement our algorithm with both the \textsc{Singular} platform and a custom C++ toolset 
+and perform experiments on two classes of circuits.
 Our approach is able to verify up to 162-bit sequential
@@
 
-At last we explore abstraction refinement techniques to boost sequential circuit verification.
-UNSAT core extraction, as an indispensable component of many refinement algorithms 
-which rely on UNSAT informations, is thus proposed in polynomial rings.
-We use weak Nullstellensatz as essential theory of UNSAT core extraction, 
-then develop heuristics depending on the structural information of the refutation.
+At last we explore abstraction UNSAT cores in algebraic geometry -- the foundation for refinement techniques
+to boost sequential circuit verification.
+% UNSAT core extraction, as an indispensable component of many refinement algorithms 
+% which rely on UNSAT informations, is thus proposed in polynomial rings.
+We use the Weak Nullstellensatz as the essential theory of UNSAT core extraction, 
+then develop heuristics to improve the core that exploit the structure of the refutation proof.
 An algorithm was implemented within the \textsc{Singular} computer algebra tool
@@
 Our approaches still have limitations: for word-level FSM traversal and 
-sequential GF multipliers verification, our method are more efficient
-on XOR-rich circuits, while most industrial designs are AND-OR gates dominant.
-For UNSAT core extractions, the abstraction refinement approach such as $k$-BMC 
+sequential GF multiplier verification, our methods are more efficient
+for XOR-rich circuits, while most industrial designs are AND-OR gate dominant.
+For UNSAT core extraction, the abstraction refinement approach such as $k$-BMC 
 has only limited application on certain model checking problems. To overcome these limitations,
-further explorations are needed for my research.
+the following further explorations are worthy of investigation.
 
 \section{Future Work}
-In this section, we point out some research problems that deserve further study.
-\subsection{Multivariate polynomial ideal based FSM Traversal}
-In Chapter \ref{ch:reacha}, we always use a single word-level variable $T$ to denote 
+In this section, we highlight some research problems that deserve further study.
+\subsection{Multivariate Polynomial Ideal based FSM Traversal}
+In Chapter \ref{ch:reacha}, we always use a single word-level variable $T$ to denote the
 next state. However, in some situations we need to keep recording the relations between $T$ and
@@
 While the elimination and set union/intersection are compatible with multiple NS variables,
-the set complement requires an extension to Theorem \ref{thm:quotient} on ideals with 
-multivariate polynomial generators. We make a conjecture as follows:
+the set complement requires an extension of Theorem \ref{thm:quotient} on ideals with 
+multivariate polynomial generators. We conjecture as follows:
 
 \begin{Conjecture}
-Assume we are given 2 ring variables $x,y$, and an ideal $J$ composed by 2 generators:
-$J = \langle f_y, v_x\rangle$. Here $f_y$ has the form $f_y = \mathcal F(x,y)$, where $\mathcal F$
-is a 2-argument function, e.g. $f_y = y^2+(1+x)y+x$. And $v_x$ is the vanishing polynomial for variable
-$x$: $v_x = x^{q_1} - x$. Now we have another generator serving as the universal set
-$J_0 = \langle v_x, v_y\rangle$, $v_y = y^{q_2} - y$.
+Assume we are given ring variables $x_i$, and an ideal $J$ composed by $s$ generators:
+$J = \langle f_1,\dots, f_s\rangle$.  Additionally $J_0$ is the vanishing polynomial for variables
+$x_i$: $ x_i^{q} - x_i$ where $q$ is the size of signal $x_i$ represents.
 We conjecture that
-$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J)$$
-And we believe this conjecture is scalable such that it is also valid for even more variables and generators like 
-$J = \langle f_z, v_x,v_y,v_u,v_w\rangle$, $f_z =\mathcal F (z,x,y,u,w)$.
+$$V(J') = V(J_0:J) = V(J_0)\setminus V(J) = \overline{V(J)}$$
+% And we believe this conjecture is scalable such that it is also valid for even more variables and generators like 
+% $J = \langle f_z, v_x,v_y,v_u,v_w\rangle$, $f_z =\mathcal F (z,x,y,u,w)$.
 \end{Conjecture}
+
+In univariate case, $\F[x]$ is the principle ideal domain thus Theorem \ref{thm:quotient} can be proved.
+However over multivariate case, the proof of this conjecture is not available.
 
@@
 
-\subsection{Use F-4 and ZDDs to Accelerate Polynomial Reduction}
-In Chapter \ref{ch:reacha} and Chapter \ref{ch:normal}, by using RATO we transform the GB 
-computation to a multivariate polynomial division. However, this division (reduction)
-is still with exponential complexity at worst case. In a situation where 
-a chain of OR gate exists, the size of polynomial will explode. For example, a chain of OR gates 
-which can be written as Boolean function 
+\subsection{Use of the $F_4$ Algorithm and ZDDs to Accelerate GB Reduction}
+In Chapters \ref{ch:reacha} and \ref{ch:normal}, by using RATO we transform the GB-based
+computation to that of a multivariate polynomial division. However, this division (reduction)
+still incurs exponential complexity in the worst case. In a situation where division is to be performed modulo
+a chain of OR gates, the size of remainder polynomial will explode. For example, a chain of OR gates 
+can be written as the Boolean function 
 \begin{equation}
@@
 \end{equation}
-equals to following polynomial function in $\F_2$:
+which equals to the following polynomial function in $\F_2$:
 $$f = abcd+abc+abd+ab+acd+ac+ad+a+bcd+bc+bd+b+cd+c+d$$
-Notice $f$ contains $2^4-1 = 15$ terms, which is exponential to number of variables.
-The polynomial size explosion is a major factor affect the efficiency of polynomial division.
+Notice, $f$ contains $2^4-1 = 15$ terms, which is exponential in the number of variables.
+The size explosion is a major factor affecting the efficiency of polynomial division.
 
 One way to further boost the efficiency is to adopt techniques from sparse linear algebra.
-Analysis on experiment results shows major time consumption is on multivariate polynomial division procedure.
-A matrix-based technique named as "F-4 style reduction" \cite{f4} can speed up the procedure
-dividing a low-degree polynomial with term-sparse polynomial ideal.
+Analysis of experimental results shows that multivariate polynomial division procedure
+consumes most of the verification time.
+A matrix-based technique named as "$F_4$ style reduction" \cite{f4} can speed up the procedure of
+dividing a low-degree polynomial with a term-sparse polynomial ideal.
 
 Another way is to utilize DDs, {\it e.g.} ZDDs. ZDDs can represent unate covers of Boolean formulas.
-We can represent $f$ using a ZDD, while the dotted edges denote the corresponding literal does not 
-exist in a minterm.
+We can represent $f$ using a ZDD, where every path in the ZDD from a root to a terminal corresponds
+to a monomial term.
 
@@
 \centerline{
-\includegraphics[width=\textwidth]{newfig/red_steps}
+\includegraphics[width=0.35\textwidth]{newfig/ZDD.eps}
 }
-\caption{The reduction process of chain of OR gates represent using ZDDs with order $d>c>b>a$}
+\caption{A ZDD representing remainder polynomial reducing by a chain of OR gates with order $d>c>b>a$}
 \label{fig:ZDD}
@@
 
-From Figure \ref{fig:ZDD} we find the size of the ZDD on the right-most side representing 
-Equation \ref{eqn:chainOR} is $2\times4 -1 = 7$, which is linear to the number of variables.
+Figure \ref{fig:ZDD} shows a ZDD representing 
+Equation \ref{eqn:chainOR} with size $2\times4 -1 = 7$, which is linear in the number of variables.
 The reduction process using ZDDs can be executed as in \cite{polybori:2009}. The graphical illustration 
-of reduction in ZDDs is also shown in Figure \ref{fig:ZDD}.
+of the remainder in ZDDs is also shown in Figure \ref{fig:ZDD}.
 
-\subsection{Interpolation extraction using GB Algorithm}
+\subsection{Craig Interpolants in Algebraic Geometry}
 The concept of Craig Interpolants (CI) and their existence comes from
@@
 (From \cite{mcmillan2003interpolation}) Let $A$ and $B$ be Boolean
-  formulae given as sets of clauses such that $A \wedge B$ is
+  formulas given as sets of clauses such that $A \wedge B$ is
   unsatisfiable (UNSAT). Then there exists a formula $P$ such that: 
-1) $A$ implies $P$ (or $A \subseteq P$); ~2) $P \wedge B$ is UNSAT; ~iii)
+1) $A$ implies $P$ (or $A \subseteq P$); ~2) $P \wedge B$ is UNSAT; ~3)
 $P$ refers to only the common variables of $A$ and $B$. The formula
@@
 
-{\bf Research problem 1:}
-{\it As we wish to perform MC with abstraction using algebraic
-  geometry, we introduce the algebraic geometry analog of CI
-  and explore algorithms to compute them. We conjecture that the
-  concept of CI should be related to elimination ideals, so our line
-  of investigation will focus on \Grobner basis computations with
+We introduce the algebraic geometry analog of CI.
+We conjecture that the
+  concept of CI should be related to elimination ideals, so future lines
+  of investigation should focus on \Grobner basis computations with
   elimination term orders for their computation.
-}
 
@@
 
-In the above definition $V_{\Fq}(J_A+J_B)=\emptyset$ implies that the
-ideal $J_A+J_B$ has an empty variety (UNSAT) problem.
+Existence of algebraic interpolant comes from \cite{craig-interpolate}.
+The question for exploring is that how do we compute it in algebraic geometry?
 
@@
 
-{\bf Research problem 2:}
-{\it How can this algebraic interpolant $f_p$ (or in the general case,
-  the set of polynomials $F_P$) be computed? As the interpolant $F_P$ 
-contains only variables $X_C$ that are common to the polynomial sets
-$F_A, F_B$, does this imply that the interpolants can be computed by
-means of \Grobner bases with an elimination term order $X_A>X_B>X_C$
-with $X_A, X_B$ eliminated from the problem? }
+% {\bf Research problem 2:}
+% {\it How can this algebraic interpolant $f_p$ (or in the general case,
+%   the set of polynomials $F_P$) be computed? As the interpolant $F_P$ 
+% contains only variables $X_C$ that are common to the polynomial sets
+% $F_A, F_B$, does this imply that the interpolants can be computed by
+% means of \Grobner bases with an elimination term order $X_A>X_B>X_C$
+% with $X_A, X_B$ eliminated from the problem? }
 
-We believe that algebraic interpolation is strongly related
+Algebraic interpolation is strongly related
 to the GB computation with the elimination order $X_A>X_B>X_C$, and
@@
 The aforementioned experiments in Example \ref{ex3} do not invalidate
-our conjectures, therefore we will attempt to prove them. If we are
-able to prove our conjectures, then we will further study the proof to
-derive efficient algorithms to compute the algebraic
-interpolants. Moreover, the above experiment shows that there can be
+our conjectures. Moreover, the above experiment shows that there can be
 multiple ways of computing the interpolants. What can these \Grobner
@@
 
-{\bf Research problem 3:}
-As there can be many interpolants for an ideal-pair $(J_A, J_B)$, we
-will further resolve these questions:
+As there can be many interpolants for an ideal-pair $(J_A, J_B)$, 
+the following questions should also be investigated in the future:
 \bi
-\item Find the minimal interpolant: i.e. find the interpolant $F_P$
+\item Find the minimal interpolant: {\it i.e.} find the interpolant $F_P$
   such that $V(J_P)$ is the smallest variety larger than $V(J_A)$ such
@@
 \includegraphics[width=4.5in]{newfig/project.eps}
-\caption{Algebraic interpolant: Projection of varieties on common variables? }
+\caption{Algebraic interpolant: Projection of varieties on common variables }
 \label{fig:projection}}
@@
 
-Once these problems are understood and algorithms for the computation
-of algebraic interpolants derived, then we will integrate them 
-with word-level reachability analysis for abstraction-based model
-checking.
+% 
+% Once these problems are understood and algorithms for the computation
+% of algebraic interpolants derived, then we will integrate them 
+% with word-level reachability analysis for abstraction-based model
+% checking.
 

### 2016-12-17 dissertation/slides/defense.tex
dissertation & slides: minor modifications before defense
dissertation:after defense_update 1

@@
 \item Recall Buchberger's algorithm: UNSAT $\to$ terminates with $1$
-\hyperlink{Buch}{\beamergotobutton{Buchberger}}
+\hyperlink{Buch}{\beamergotobutton{GB}}
 \pause
@@
 \item $f_{10}=f_1-acf_2-af_2-f_3-cf_2-f_2 = f_1+acf_2+af_2+f_3+cf_2+f_2$
-\item $1 = \F(f_1,f_2,f_3,f_4,f_5)\implies 1 \in \langle f_1,f_2,f_3,f_4,f_5\rangle \implies UNSAT$
+\item $1 = \Func(f_1,f_2,f_3,f_4,f_5)\implies 1 \in \langle f_1,f_2,f_3,f_4,f_5\rangle \implies UNSAT$
 \ei
@@
 \item Expand: \begin{align*}
-		1 &= \F(f_1,f_2,f_3,f_4,f_5)\\
+		1 &= \Func(f_1,f_2,f_3,f_4,f_5)\\
 		&= cf_2+f_5+\alert{f_{11}}\\
@@
 $1\cdot f_3 + ac f_2 + a f_5 = 0\Leftrightarrow f_3 = acf_2+af_5$
-\item Generalization of this strategy refer to the paper
+\item Generalization of this strategy refer to Ch.7 in dissertation
 \ei
@@
 \end{frame}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{frame}{\large{Auf Wiedersehen}}
+\begin{figure}
+\centering
+\includegraphics[scale=0.5]{thankyou.pdf}
+\end{figure}
+\end{frame}
 

### 2016-12-24 dissertation/submission/appendix1.tex
dissertation after defense update 2

--- 2016-12-09 dissertation/appendix.tex
@@
-\numberofappendices=1
-\appendix
-\chapter{}
-\fixchapterheading
-\section{Normal Basis Theory}
+\chapter{Normal Basis Theory}
 \label{append:NB}
 In Chapter \ref{ch:prelim_GF} we briefly introduced the concept of normal basis (NB), and 
-the benefits of using NB. In the section we describe more details about the normal basis theory by 
-characterizing NB from linear algebra perspective, constructing a general NB in an 
-arbitrary field and converting between NB and StdB. All theorems and lemmas refer to the 
-dissertation of S. Gao \cite{gao:phd_normal_basis}, and we deduce all the proofs for them.
-\subsection{Characterization of Normal Basis}
+the benefits of using NB. In this appendix chapter we describe more details about NB theory by 
+characterizing NB from a linear algebra perspective, constructing a general NB in an 
+arbitrary field, and converting between NB and StdB. All theorems and lemmas refer to the 
+dissertation of Gao \cite{gao:phd_normal_basis}, and we deduce all the proofs for them.
+\section{Characterization of Normal Basis}
 In order to depict the characterization of NB, we need the introduce some concepts from linear algebra domain.
@@
 A subspace $W\subset V$ is called {\bf $T$-invariant} when 
-$$Tu \in W, ~~ \forall \text{vector }u\in W$$
+$$Tu \in W, ~~ \forall \text{ vector }u\in W$$
 
@@
 \end{Definition}
-Field elements $x$ constitutes polynomials in ring $\F[x]$. Subsequently we can define the 
-nullspace of polynomial:
-\begin{Definition}[Nullspace of polynomial]
+
+We define the nullspace of a polynomial as:
+\begin{Definition}[Nullspace of a polynomial]
 \label{def:nullspace}
@@
 \end{Definition}
-Finally we get to the most important concept: $T$-Order. Meanwhile it derives the construction of a ring extended by a field requires
-a minimal polynomial (while we use stronger irreducible polynomial $P(x)$ in Chapter \ref{ch:prelim_GF}).
+Finally we get to the most important concept: $T$-Order. It derives the construction of a ring extended 
+by a field requiring a minimal polynomial.
 \begin{Definition}
@@
 
-That is, for an arbitrary element $\theta$ in $F_{q^n}$, find least positive integer $k$ such that
+Let map $T=\sigma$, for an arbitrary element $u=\theta$ in $F_{p^n}$, find least positive integer $k$ such that
 $\sigma^k\theta = \sum_{i=0}^{k-1} c_i\sigma^i\theta$, then the $\sigma$-Order of $\theta$ can be 
@@
 
-Using the concepts introduced above along with other basic concepts in linear algebra and finite field theory,
+Using the concepts introduced above, along with other basic concepts in linear algebra and finite field theory,
 we can derive following theorems.
@@
 
-Let $k$ be a polynomial whose degree is larger than $f(x)$, i.e. $deg(k(x))>deg(f(x))$, then
+Let $k$ be a polynomial whose degree is larger than that of $f(x)$, {\it i.e.} $deg(k(x))>deg(f(x))$, then
 $$f(x)|k(x)~~\text{iff}~k(T)=0$$
@@
 \end{Theorem}
-\subsection{Construction of General Normal Bases}
+\section{Construction of General Normal Bases}
 After proving the existence of normal bases, the upcoming question is that how to find such a normal basis/element.
@@
 \begin{Proposition}
-L\"uneburg's algorithm always generate a normal element over field $\F_{p^n}$.
+L\"uneburg's algorithm always generates a normal element over the field $\F_{p^n}$.
 \end{Proposition}
@@
 
-Lenstra's algorithm is an approximation algorithm in nature. For each iteration $Ord_\theta(x)$ monotonically increase,
+Lenstra's algorithm is an approximation algorithm in nature. For each iteration $Ord_\theta(x)$ monotonically increases,
 so it will finally reach termination condition $Ord_\theta(x) = x^n-1$.
-\subsection{Bases Conversion}
+\section{Bases Conversion and $\lambda$-Matrix Construction}
 % General \lambda matrix computation
+Given a normal basis (NB), we can easily convert it to a standard basis (StdB) by reducing $\alpha^t$
+with irreducible polynomial $p(\alpha)$ since 
+normal element is already given as $\beta = \alpha^t$. The reverse conversion requires 
+solving a system of equations. In Example \ref{ex:Rijndael}, the system of equations
+is as follows:
 
-\section{Optimal Normal Basis}
-\label{append:ONB}
-The number of non-zero entries in $\lambda$-Matrix or multiplication table is known as \emph{Complexity $(C_N)$}.
-To define optimal normal basis, it is necessary to find the lower bound of $C_N$.
-\begin{Theorem}
-If N is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
-\end{Theorem}
-\begin{Proof}
-Let basis $N = \{\beta, \beta^p, \beta^{p^2},\dots, \beta^{p^{n-1}}\}$, let us denote $\beta^{p^i}$ by $\beta_i$ for the sake of simplification.
-Then $\sum_{i=0}^{n-1} \beta_i = trace\ \beta$. 
-Denote $trace\ \beta$ by $b$, consider a $n\times n$ matrix $M^{(0)}$. Then
-$$b\beta_0 = \sum_{i=0}^{n-1} \beta \beta_i$$
+\begin{equation}
+\begin{cases}
+\alpha = \alpha \\
+\alpha^2 = \alpha^2 \\
+\alpha^4 = \alpha^4 \\
+\alpha^8 = \alpha^7 + \alpha^6+\alpha^4+\alpha^2+\alpha+1\\
+\alpha^{16} = \alpha^7+\alpha^6+\alpha^5+\alpha^3+\alpha+1\\
+\alpha^{32} = \alpha^7+\alpha^4+\alpha^3+\alpha^2+\alpha\\
+\alpha^{64} = \alpha^7+\alpha^6+\alpha^3+\alpha^2\\
+\alpha^{128} = \alpha^6+\alpha^5+\alpha^4+\alpha^3+1
+\end{cases}
+\end{equation}
 
-Therefore, the sum of all rows in $M^{(0)}$ is an n-tuple with $b$ as the first element and zeros elsewhere.
-So the first column always includes at least one nonzero element. 
-Since the sum of enries in each of other columns should equal to zero (modulo 2), they 
-need to include an even number of ``1"s.
-in order to maintain the linear independence among each row, they cannot be purely ``0"s. 
-Therefore, there are always at least two non-zero elements in each column.
+We treat StdB $\{1,\alpha,\dots,\alpha^7\}$ as unknowns and NB $\{\alpha,\alpha^2,\dots,\alpha^{128}\}$
+as parameters. Then we can obtain a solution to this system denoting mapping from StdB to NB:
 
-As a result, $C_N \geq 2(n-1)+1  = 2n-1$.
-\end{Proof}
+\begin{equation}
+\begin{cases}
+1 = \alpha^{128}+\alpha^{64}+\alpha^{32}+\alpha^{16}+\alpha^8+\alpha^4+\alpha^2+\alpha \\
+\alpha = \alpha \\
+\alpha^2 = \alpha^2 \\
+\alpha^3 = \alpha^{128} + \alpha^{32}+\alpha^{16}+\alpha^2\\
+\alpha^4 = \alpha^4\\
+\alpha^5 = \alpha^{128} + \alpha^{32}+\alpha^{8}+\alpha^4\\
+\alpha^6 = \alpha^{64} + \alpha^{32}+\alpha^{4}+\alpha\\
+\alpha^7 = \alpha^{128} + \alpha^{16}+\alpha^{4}+\alpha
+\end{cases}
+\end{equation}
 
-If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as \emph{Optimal Normal Basis}.
+Directly constructing $\lambda$-Matrix is difficult for general NBs. 
+In Section \ref{sec:nbdesign}, we introduced the concept of multiplication table (M-table) and explained how its 
+entries are transformed from the $0$-th $\lambda$-Matrix. 
+Recall the M-table is defined as
+\begin{equation}
+\label{eqn:appmt}
+\beta
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
+\end{bmatrix}
+= {\bf T}
+\begin{bmatrix}
+\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
+\end{bmatrix}
+\end{equation}
+Therefore we can compute M-table and then covert it to $\lambda$-Matrix.
+Using Equation \ref{eqn:appmt}, we derive entry $T_{i,j}$ in the M-table by equating it to
+the $j$-th bit of the NB representation of element $\beta\cdot\beta^{2^{i}}$.
+For example, the $i$-th row of the M-table for NB in Example \ref{ex:Rijndael} satisfies
+$$\beta\cdot\beta^{2^i} = \sum_{j=0}^{7} T_{i,j}\beta^{2^j}$$
+Thus, the M-table can be written as
+\begin{equation*}
+T = \left(\begin{array}{lccccccr}
+0 &1 &0 &0 &0 &0 &0 &1 \\
+0 &1 &0 &0 &1 &1 &0 &1 \\
+0 &0 &1 &0 &0 &1 &0 &1 \\
+0 &0 &0 &0 &0 &0 &0 &1 \\
+1 &1 &1 &1 &1 &1 &1 &1 \\
+0 &0 &0 &0 &1 &0 &0 &0 \\
+1 &1 &0 &1 &0 &1 &0 &0 \\
+1 &0 &0 &1 &1 &0 &1 &0
+\end{array}\right)
+\end{equation*}
+Use equation $M_{i,j}^{(0)} = T_{j-i,-i}$, we can obtain the
+$\lambda$-Matrix in Equation \ref{eqn:mat_R}.
 
-\begin{Example}
-In $\F_{2^4}$ constructed with $p(\alpha) = \alpha^4 + \alpha + 1$, two sets of normal bases can be found:
-$\beta = \alpha^3, N = \{ \alpha^3, \alpha^6, \alpha^{12}, \alpha^9\}$ and 
-$\beta = \alpha^7, N = \{\alpha^7, \alpha^{14}, \alpha^{13}, \alpha^{11}\}$. Multiplication table for $\beta = \alpha^3$:
-\begin{equation}
-T_1 = \left(
-\begin{array}{lccr}
-0 & 1 & 0 & 0\\
-0 & 0 & 0 & 1\\
-1 & 1 & 1 & 1\\
-0 & 0 & 1 & 0
-\end{array} \right)
-\end{equation}
-Complexity $C_N$ = 7, this is optimal. For $\beta = \alpha^7$:
-\begin{equation}
-T_2 = \left(
-\begin{array}{lccr}
-0 & 1 & 0 & 0\\
-1 & 1 & 0 & 1\\
-1 & 0 & 1 & 0\\
-1 & 0 & 1 & 1
-\end{array} \right)
-\end{equation}
-Complexity $C_N$ = 9, this is NOT optimal.
-\end{Example}
-
-There should be an efficient method to construct optimal normal basis. The following  subsection refers to \cite{rosingbook}.
-\subsection{Construction of Optimal Normal Basis}
-\subsection{Optimal Normal Basis Multiplier Design}
-% Customized Galois Field IEEE1363-2000
-% Multi table <=> \lambda matrix
-In Section \ref{sec:nbdesign}, we introduce the concept of multiplication table and explained how its 
-entries are transformed from the $0$-th $\lambda$-Matrix. Actually we can mathematically prove the relationship 
+Actually we can mathematically prove the relationship 
 between them:
 \begin{Theorem}
-Multiplication table $T$ is actually a conjugate of the $\lambda$-Matrix $M$. 
-We can use an equation $M_{i,j}^{(0)} = T_{j-i,-i}$ to demonstrate the relation between their entries.
+M-table $T$ is a conjugate of the $\lambda$-Matrix $M$. In other words,
+Equation $M_{i,j}^{(0)} = T_{j-i,-i}$ holds as the relation between their entries.
 \end{Theorem}

### 2016-12-24 dissertation/submission/appendix2.tex
dissertation after defense update 2

--- 2016-12-11 dissertation/appendix.tex
@@
-\numberofappendices=1
-\appendix
-\chapter{}
-\fixchapterheading
-\section{Normal Basis Theory}
-\label{append:NB}
-In Chapter \ref{ch:prelim_GF} we briefly introduced the concept of normal basis (NB), and 
-the benefits of using NB. In the section we describe more details about the normal basis theory by 
-characterizing NB from linear algebra perspective, constructing a general NB in an 
-arbitrary field and converting between NB and StdB. All theorems and lemmas refer to the 
-dissertation of S. Gao \cite{gao:phd_normal_basis}, and we deduce all the proofs for them.
-\subsection{Characterization of Normal Basis}
-In order to depict the characterization of NB, we need the introduce some concepts from linear algebra domain.
-\begin{Definition}[Frobenius Map]
-Define map $\sigma:x\to x^p,~x\in \F_{p^n}$. This map denoting the linear map of field extension $\F_{p^n}$
-over $\F_p$.
-\end{Definition}
-Additionally we can define a subspace based on linear map $T$:
-\begin{Definition}
-A subspace $W\subset V$ is called {\bf $T$-invariant} when 
-$$Tu \in W, ~~ \forall \text{vector }u\in W$$
-
-Subspace $Z(u,T) = \langle u,Tu,T^2u,\dots\rangle$ is called {\bf $T$-cyclic subspace} of $V$.
-If $Z(u,T) = V$ holds, then $u$ is called a {\bf cyclic vector} of $V$ for $T$.
-\end{Definition}
-Field elements $x$ constitutes polynomials in ring $\F[x]$. Subsequently we can define the 
-nullspace of polynomial:
-\begin{Definition}[Nullspace of polynomial]
-\label{def:nullspace}
-For any polynomial $g(x) \in \F[x]$, the null space of $g(T)$ consists of all vectors $u$ such that
-$g(T)u = 0$.
-\end{Definition}
-Finally we get to the most important concept: $T$-Order. Meanwhile it derives the construction of a ring extended by a field requires
-a minimal polynomial (while we use stronger irreducible polynomial $P(x)$ in Chapter \ref{ch:prelim_GF}).
-\begin{Definition}
-For any vector $u \in V$, the monic polynomial $g(x) \in \F[x]$ with smallest degree such that
-$g(T)u = 0$ is called the {\bf $T$-Order} of $u$ or {\bf minimal polynomial} of $u$.
-
-That is, for an arbitrary element $\theta$ in $F_{q^n}$, find least positive integer $k$ such that
-$\sigma^k\theta = \sum_{i=0}^{k-1} c_i\sigma^i\theta$, then the $\sigma$-Order of $\theta$ can be 
-written as $$Ord_{\theta,\sigma}(x) = x^k - \sum_{i=0}^{k-1} c_ix^i$$
-\end{Definition}
-
-Using the concepts introduced above along with other basic concepts in linear algebra and finite field theory,
-we can derive following theorems.
-
-\begin{Lemma}
-\label{lem:app1}
-Given $g(x) \in \F[x]$ and $W$ is its nullspace. Let $d(x) = gcd(f(x),g(x)), e(x) = f(x)/d(x)$. 
-Then $dim(W) = deg(d(x))$ and $W = \{e(T)u~|~u \in V\}$.
-\end{Lemma}
-\begin{Proof}
-Assume $f(x)$ is the minimal and characteristic polynomial for $T$. Then according to Definition \ref{def:nullspace}
-we obtain 
-$$W = \{u\in V~|~g(T)u = 0\}$$
-
-Let $k$ be a polynomial whose degree is larger than $f(x)$, {\it i.e.} $deg(k(x))>deg(f(x))$, then
-$$f(x)|k(x)~~\text{iff}~k(T)=0$$
-
-Since the construction of $f(x)$ relies on $T$, $f(T) = 0$, we get
-$$\forall u, f(T)u = 0$$
-
-Consider $W(u,T)$ subsequently relies on $g(T)$, we deduce 
-$$dim(W) = deg(d(x))$$
-
-According to the definition of $W$, we have
-\begin{align*}
-\forall u,~~& e(T)u \in W \Longleftrightarrow \\
-& g(T)e(T)u = g(T)\frac{f(T)}{d(T)} u = h(T)\cdot f(T) u = 0
-\end{align*}
-\end{Proof}
-
-The $T$-Order and corresponding nullspace also affect the factorization of polynomials:
-\begin{Lemma}[Factorization of $f(x)$]
-Factorization $f(x) = \prod_{i=1}^{r} f_{i}^{d_i} (x)$, where each $f_i(x)$ is prime to others. 
-Assume $V_i$ be nullspace of $f_{i}^{d_i} (x)$, then $V = V_1\oplus V_2 \oplus \dots\oplus V_r$.
-
-Furthermore, we can define a polynomial $\Psi_i(x) = f(x)/f_{i}^{d_i} (x)$, where 
-$$\forall u_j \in V_j, u_j \neq 0, \Psi_i(T)u_j \neq 0$$
-only if $i = j$.
-\end{Lemma}
-\begin{Proof}
-We can use Lemma \ref{lem:app1}:
-$$gcd(f(x),f_{i}^{d_i} (x)) = f_{i}^{d_i} (x),~~dim(V_i) = deg(f_{i}^{d_i} (x))$$
-
-Which implies 
-$$dim(V_i) = deg\left(\prod_i f_{i}^{d_i} (x)\right) = \prod_i dim(V_i) \implies V = \bigoplus_i V_i$$
-
-Assume $i\neq j$, $\Psi_i(x) = \frac{f(x)}{f_{i}^{d_i} (x)} = h(x)f_{j}^{d_j} (x)$. Then
-\begin{align*}
-&\forall u_j\in V_j, f_j(T)u_j = 0 \implies h(x)f_{j}^{d_j} (T)u_j = 0 \\
-& \implies \Psi_i(x)u_j = 0
-\end{align*}
-
-Conversely, if $i=j$, then
-\begin{align*}
-& \Psi_i(x) = \frac{f(x)}{f_{i}^{d_i} (x)} \perp f_i(x) \\
-& \implies \Psi_i(x)u_j \neq 0
-\end{align*}
-\end{Proof}
-
-For a given Frobenius map $\sigma$, corresponding minimal polynomial is restricted:
-\begin{Lemma}
-\label{lem:3}
-The minimal (and characteristic) polynomial for $\sigma$ is $x^n-1$.
-\end{Lemma}
-\begin{Proof}
-Consider Fermat's little theorem in $\F_{p^n}$. Let $\beta$ be an element in the field, then
-$$\sigma^n\beta = \beta^{p^n} = \beta \implies \sigma^n-I = 0$$
-where $I$ is the identity map. For characteristic polynomial, assume $\exists f(x) = \sum_i f_ix^i \in \F_p[x]$, 
-such that the degree of characteristic polynomial is lower than $n$:
-$$\sum_if_i\sigma^i = 0~~~\text{and}~~~deg(f(x))<n$$
-
-Then $\forall \beta\in\F_{p^n}$, 
-$$\left(\sum_if_i\sigma^i\right)\beta = \sum_if_i\beta^{p^i} = 0$$
-
-This equation denotes that $\beta$ is one of $p^n$ roots of polynomial $\Func(x) = \sum_if_i(x^{p^i})$.
-However, the maximum number of roots allowed equals to the degree of $\Func(x)$, which is 
-$p^{n-1}<p^n$, we find a contradiction. Thus we conclude that
-$$\textit{Both characteristic and minimal polynomial is }x^n-1$$
-\end{Proof}
-
-From lemmas above we can deduce the following corollary:
-\begin{Corollary}
-\label{corol:1}
-An element $\alpha \in \F_{p^n}$ is a {\bf normal element} if and only if $Ord_{\alpha,\sigma}(x) = x^n-1$.
-\end{Corollary}
-\begin{Proof}
-Normal bases require $\{\beta,\beta^2,\dots,\beta^{2^{n-1}}\}$ to be linearly independent,
-which is equivalent to the following expression:
-$$\forall f(x)\in \F_p[x],~deg(f(x))<n$$
-Furthermore, because there is no annihilators in the $T$-Order, this implies that
-$$Ord_{\alpha,\sigma}(x) = x^n-1$$
-\end{Proof}
-
-From corollary above we can deduce another form of criterion of normal element:
-\begin{Theorem}
-\label{thm:1}
-Given finite field $\F_{2^n}$, the field characteristic $p = 2$.
-Define $t = p^e$ such that $n = kp^e, gcd(k,p) = 1$, so $t=1$ if $n$ is an odd integer. Then 
-$x^n - 1$ can be factorized as 
-$$(\varphi_1(x)\varphi_2(x)\cdot\cdot\cdot\varphi_r(x))^t$$
-Additionally, define $\Phi_i(x) = (x^n - 1)/\varphi_i(x)$. We assert that
-
-An element $\alpha \in 	F_{p^n}$ is a normal element if and only if $\Phi_i(\sigma)\alpha \neq 0, i = 1,2,\dots,r$.
-\end{Theorem}
-\begin{Proof}
-Let us analyze the auxiliary polynomial $\Phi_i(x) = \frac{x^n - 1}{\varphi_i(x)}$ first.
-$$\Phi_i(\sigma)\alpha \neq 0 \Leftrightarrow \text{No factor in }x^n-1\text{ annihilates }\alpha$$
-On the other hand, according to Lemma \ref{lem:3}, it is a known 
-fact that \emph{any annihilator always divide} $x^n-1$. As a result, the only possible situation is
-$$Ord_{\alpha,\sigma}(x) = x^n-1$$
-Using Corollary \ref{corol:1} we deduce that $\alpha$ is a normal element.
-\end{Proof}
-
-The normal element identification can also be described from the nullspace perspective:
-\begin{Theorem}
-\label{thm:2}
-Let $W_i$ be the nullspace of $\varphi_{i}^{t} (x)$ and $\widetilde{W_i}$ the
-nullspace of $\varphi_{i}^{t-1} (x)$. Let $\overline{W_i}$ be any subspace of
-$W_i$ such that $W_i = \overline{W_i}\oplus \widetilde{W_i}$. Then
-$$\F_{p^n} = \displaystyle\sum_{i=1}^{r} \overline{W_i}\oplus \widetilde{W_i}$$
-is a direct sum where $dim(\overline{W_i}) = d_i$ and $dim(\widetilde{W_i}) = (t-1)d_i$.
-
-Using above setup we define an element $\alpha \in \F_{p^n}$ with 
-$\alpha = \sum_{i=1}^{r} (\overline{\alpha_i} + \widetilde{\alpha_i}), \overline{\alpha_i} \in \overline{W_i}, \widetilde{\alpha_i} \in \widetilde{W_i}$,
-as a normal element if and only if $\overline{\alpha_i} \neq 0,~ \forall i = 1,2,\dots,r$.
-\end{Theorem}
-
-Ultimately, since there always exist at least one element fulfilling requirements in Theorem \ref{thm:1}
-or Theorem \ref{thm:2}, we obtain following theorem:
-\begin{Theorem}[Normal Basis Theorem over Finite Fields]
-There always exists a normal basis of $\F_{p^n}$ over $\F_p$.
-\end{Theorem}
-\subsection{Construction of General Normal Bases}
-After proving the existence of normal bases, the upcoming question is that how to find such a normal basis/element.
-For general normal basis identification, there are two methods widely used: the L\"uneburg's algorithm
-and Lenstra's algorithm.\\
-\textbf{L\"uneburg's Algorithm:}
-\begin{enumerate}[{1)}]
-\item Randomly pick an element $\alpha$ from $\F_{p^n}$. For each $i = 0,1,\dots,n-1$, 
-compute $\sigma$-Order $f_i = Ord_{\alpha^i}(x)$. Then $x^n - 1 = lcm(f_0,f_1,...,f_{n-1})$.
-\item Apply factor refinement to set $\{f_0,\dots,f_{n-1}\}$ and obtain $f_i = \prod_{1\leq j\leq r} g_{j}^{e_{ij}}, i = 0,1,\dots,n-1$.
-We can write the result as an $i\times j$ matrix.
-\item For each $j$, find an index $i_j$ (denote as $i(j)$) such that $e_{ij}$ is maximum in the $j$-th column.
-\item Let $h_j = f_{i(j)}/g_{j}^{e_{i(j)j}}$, take $\beta_j = h_j(\sigma)\alpha^{i(j)}$. Then
-$$\beta = \displaystyle\sum_{j=1}^{r} \beta_j$$
-is a normal element.
-\end{enumerate}
-
-L\"uneburg's algorithm starts with a random element and ends with a normal element. The justification of 
-its soundness is as follows:
-\begin{Proposition}
-L\"uneburg's algorithm always generate a normal element over field $\F_{p^n}$.
-\end{Proposition}
-
-\begin{Proof}
-First, according to the definition of $T$-Order:
-$$f_i = Ord_{\alpha^i} \Leftrightarrow f_i(\sigma)\alpha^i = 0$$
-
-Use Lemma \ref{lem:3}, the minimal/characteristic polynomial for $\sigma$ is $x^n-1$ implies that
-any annihilator of $\alpha^i$ ({\it i.e.} $f_i(\sigma)$) divides $x^n-1$. Meanwhile, $\{\alpha^i~|~0<i<n-1\}$ forms a (standard) basis
-of field $\F_{p^n}$, this means $\alpha_i$ are linearly independent with each other ($i=0,1,\dots,n-1$).
-Since linear map $f(\sigma)$ guarantees that all elements in the field 
-have order $lcm(f_0,\dots,f_{n-1})$, {\it i.e.} $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
-By contradiction we can prove that there is no factor of $x^n-1$ that can be divided by the product 
-of elements in set $\{f_0,f_1,\dots,f_{n-1}\}$. This actually corresponds to the assertion in the first step of the algorithm:
-$$x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$$
-
-In the second step, after factorization of set $\{f_i\}$, we transform $f_i(\sigma)\alpha^i=0$ to
-$$h_j(\sigma)\cdot g_{j}^{e_{i(j)j}}(\sigma) \cdot \alpha^{i(j)} = 0$$
-
-Furthermore we have 
-$$g_{j}^{e_{i(j)j}}(\sigma)\beta_j = 0 \Leftrightarrow Ord_{\sigma,\beta_j}(x) = g_{j}^{e_{i(j)j}}(x)$$
-
-Consider the facts: 1) elements in set $\{g_j\}$ are relatively prime; 2) $g_{j}^{e_{i(j)j}}$ is the maximum factor; 
-3) $x^n-1 = lcm(f_0,f_1,\dots,f_{n-1})$, we deduce that 
-$$x^n-1 = \prod_j g_{j}^{e_{i(j)j}}(x) = \prod_jOrd_{\sigma,\beta_j}(x)$$
-
-As a result
-$$x^n-1 = Ord_{\sigma,\beta}(x) \implies \beta\text{ is a normal element.}$$
-\end{Proof}
-
-L\"uneburg's algorithm is an analytic method, which requires relatively high computational complexity 
-mainly because of the factorization in its second step. To overcome the high cost, an inductive method 
-is proposed, which is the Lenstra's algorithm. It allows for setup of heuristics to accelerate the procedure.
-Before introducing the details of the algorithm, we demonstrate two preliminary lemmas.
-\begin{Lemma}
-\label{lem:4}
-For an arbitrary element $\theta \in \F_{p^n}$ that $Ord_\theta(x) \neq x^n - 1$,
-let $g(x) = (x^n - 1)/Ord_\theta(x)$. There exists another element $\beta$ such that $g(\sigma)\beta = \theta$.
-\end{Lemma}
-\begin{Proof}
-Assume $\gamma$ is the desired normal element. From the definition of normal element we derive 
-$$\exists f(x)\in \F_{p^n}[x],~~f(\sigma)\gamma = \theta$$
-
-Then using the definition of $T$-Order, we derive
-$$Ord_{\sigma}\theta = 0 \implies (Ord_{\sigma}f(\sigma))\gamma = 0$$
-
-Since $\gamma$ is the normal element, $Ord_\gamma(x) = x^n-1\implies x^n-1~|~(Ord_\theta(x)f(x))$.
-Furthermore,
-$$g(x) = \frac{x^n-1}{Ord_\theta(x)} \implies x^n-1\Bigm|\left(\frac{x^n-1}{g(x)}\cdot f(x)\right)\implies g(x)~|~f(x)$$
-
-Let $f(x)=h(x)g(x)$. Then
-$$g(\sigma)(h(\sigma)\gamma) = \theta$$
-
-Therefore, $\exists \beta$ {\it i.e. }$g(\sigma)\beta = \theta$. Concretely, $\beta = h(\sigma)\gamma$.
-\end{Proof}
-
-\begin{Lemma}
-\label{lem:5}
-Define $\theta$ and $g(x)$ as in Lemma \ref{lem:4}. Assume there exists a solution 
-$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Respectively there exists a non-zero element $\eta$ such that
-$$g(\sigma)\eta = 0$$
-and
-$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
-\end{Lemma}
-\begin{Proof}
-Assume $\gamma$ is the desired normal element. Similarly,
-$$\exists \eta = Ord_\theta(\sigma)\gamma \neq 0,~~g(\sigma)\eta = 0$$
-
-Follow the setup in Lemma \ref{lem:4}, we have
-$$g(\sigma)\beta = \theta,~~\frac{x^n-1}{Ord_\theta(x)}\Bigm|_\sigma\cdot \beta = \theta$$
-
-Subsequently,
-$$Ord_\theta(x)\cdot \frac{x^n-1}{Ord_\theta(x)}\Bigm|_\sigma\beta = 0 = Ord_\beta(x)\beta$$
-which implies
-$$Ord_\theta(x)\bigm|Ord_\beta(x),~~deg(Ord_\theta(x)) \leq deg(Ord_\beta(x))$$
-Combining with the assumption in the lemma: $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$, we derive
-$$deg(Ord_\beta(x)) = deg(Ord_\theta(x))\implies Ord_\theta(x) =Ord_\beta(x)$$
-
-In the next part we need to prove $g(x)\perp Ord_\theta(x)$ by contradiction. Assume 
-$h(x) = gcd(g(x),Ord_\theta(x))\neq 1$, and
-$$g(\sigma)\beta = a(\sigma)h(\sigma)\beta = \theta,~~Ord_\theta(\sigma)\theta = a(\sigma)b(\sigma)h^2(\sigma)\beta$$
-However, $Ord_\theta(x) =Ord_\beta(x)$ indicates that $Ord_\beta(x)\beta = b(x)h(x)a(x)\beta = 0$. Thus,
-$Ord_\theta(x)=b(x)$. This is true if and only if $h(x)=1$. As a result, we assert that
-$$g(x) \perp Ord_\theta(x)$$
-
-Consider $g(\sigma)\eta = 0 \implies Ord_\eta(x)\bigm|g(x)$. It further implies that 
-$$Ord_\theta(x)\perp Ord_\eta(x) \implies Ord_{\theta+\eta}(x) = Ord_\theta(x)\cdot Ord_\eta(x)$$
-
-Since $\eta \neq 0$, we derive the result
-$$deg(Ord_{\theta+\eta}(x)) > deg(Ord_\theta(x))$$
-\end{Proof}
-
-Based on Lemmas \ref{lem:4} and \ref{lem:5}, the Lenstra's algorithm is described below:\\
-\textbf{Lenstra's Algorithm}
-\begin{enumerate}[{1)}]
-\item Take an arbitrary element $\theta \in \F_{p^n}$, determine $Ord_\theta(x)$.
-\item If $Ord_\theta(x) = x^n - 1$ then algorithm terminates and return $\theta$ as a normal element.
-\item Otherwise, compute $g(x) = (x^n - 1)/Ord_\theta(x)$, and solve $\beta$ from $g(\sigma)\beta = \theta$.
-\item Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to 2);
-otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a non-zero element $\eta$ such that $g(\sigma)\eta = 0$,
-replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, and go to 2).
-\end{enumerate}
-
-Lenstra's algorithm is an approximation algorithm in nature. For each iteration $Ord_\theta(x)$ monotonically increase,
-so it will finally reach termination condition $Ord_\theta(x) = x^n-1$.
-\subsection{Bases Conversion and $\lambda$-Matrix Construction}
-% General \lambda matrix computation
-Given a normal basis (NB), we can easily convert it to a standard basis (StdB) by reducing $\alpha^t$
-with irreducible polynomial $p(\alpha)$ since 
-normal element is already given as $\beta = \alpha^t$. The reverse conversion requires 
-solving a system of equations. In Example \ref{ex:Rijndael}, the system of equations
-is as follows:
-
-\begin{equation}
-\begin{cases}
-\alpha = \alpha \\
-\alpha^2 = \alpha^2 \\
-\alpha^4 = \alpha^4 \\
-\alpha^8 = \alpha^7 + \alpha^6+\alpha^4+\alpha^2+\alpha+1\\
-\alpha^{16} = \alpha^7+\alpha^6+\alpha^5+\alpha^3+\alpha+1\\
-\alpha^{32} = \alpha^7+\alpha^4+\alpha^3+\alpha^2+\alpha\\
-\alpha^{64} = \alpha^7+\alpha^6+\alpha^3+\alpha^2\\
-\alpha^{128} = \alpha^6+\alpha^5+\alpha^4+\alpha^3+1
-\end{cases}
-\end{equation}
-
-We treat StdB $\{1,\alpha,\dots,\alpha^7$ as unknowns and NB $\{\alpha,\alpha^2,\dots,\alpha^{128}\}$
-as parameters. Then we can obtain a solution to this system denoting mapping from StdB to NB:
-
-\begin{equation}
-\begin{cases}
-1 = \alpha^{128}+\alpha^{64}+\alpha^{32}+\alpha^{16}+\alpha^8+\alpha^4+\alpha^2+\alpha \\
-\alpha = \alpha \\
-\alpha^2 = \alpha^2 \\
-\alpha^3 = \alpha^{128} + \alpha^{32}+\alpha^{16}+\alpha^2\\
-\alpha^4 = \alpha^4\\
-\alpha^5 = \alpha^{128} + \alpha^{32}+\alpha^{8}+\alpha^4\\
-\alpha^6 = \alpha^{64} + \alpha^{32}+\alpha^{4}+\alpha\\
-\alpha^7 = \alpha^{128} + \alpha^{16}+\alpha^{4}+\alpha
-\end{cases}
-\end{equation}
-
-Directly constructing $\lambda$-Matrix is difficult for general NBs. 
-In Section \ref{sec:nbdesign}, we introduce the concept of multiplication table (M-table) and explained how its 
-entries are transformed from the $0$-th $\lambda$-Matrix. 
-Recall the M-table is defined as
-\begin{equation}
-\label{eqn:appmt}
-\beta
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
-\end{bmatrix}
-= {\bf T}
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
-\end{bmatrix}
-\end{equation}
-Therefore we can compute M-table and then covert it to $\lambda$-Matrix.
-Using Equation \ref{eqn:appmt}, we derive entry $T_{i,j}$ in the M-table equals to
-the $j$-th bit of the NB representation of element $\beta\cdot\beta^{2^{i}}$.
-For example, the $i$-th row of the M-table for NB in Example \ref{ex:Rijndael} satisfies
-$$\beta\cdot\beta^{2^i} = \sum_{j=0}^{7} T_{i,j}\beta^{2^j}$$
-Thus, the M-table can be written as
-\begin{equation*}
-T = \left(\begin{array}{lccccccr}
-0 &1 &0 &0 &0 &0 &0 &1 \\
-0 &1 &0 &0 &1 &1 &0 &1 \\
-0 &0 &1 &0 &0 &1 &0 &1 \\
-0 &0 &0 &0 &0 &0 &0 &1 \\
-1 &1 &1 &1 &1 &1 &1 &1 \\
-0 &0 &0 &0 &1 &0 &0 &0 \\
-1 &1 &0 &1 &0 &1 &0 &0 \\
-1 &0 &0 &1 &1 &0 &1 &0
-\end{array}\right)
-\end{equation*}
-Use equation $M_{i,j}^{(0)} = T_{j-i,-i}$, we can obtain the
-$\lambda$-Matrix in Equation \ref{eqn:mat_R}.
-
-Actually we can mathematically prove the relationship 
-between them:
-\begin{Theorem}
-M-table $T$ is actually a conjugate of the $\lambda$-Matrix $M$. 
-We can use an equation $M_{i,j}^{(0)} = T_{j-i,-i}$ to demonstrate the relation between their entries.
-\end{Theorem}
-\begin{Proof}
-Recall the definition of $\lambda$-Matrix over field $\F_{2^k}$:
-$$C = A\times B = \left(\sum_i a_i\beta^{2^i}\right)\left(\sum_jb_j\beta^{2^j}\right) = \sum_i\sum_ja_ib_j\beta^{2^i}\beta^{2^j}$$
-Then there always exist $\lambda_{ij}^{(l)}$, such that 
-$$\beta^{2^i}\beta^{2^j} = \sum_l\lambda_{ij}^{(l)}\beta^{2^l}$$
-We call it the cross-product term. The $l$-th bit of the product $C$ is
-$$c_l = \sum_i\sum_ja_ib_j\lambda_{ij}^{(l)}$$
-For the sake of simplification, we let $i=0$, thus 
-$$\beta\cdot\beta^{2^j} = \sum_l\lambda_{0j}^{(l)}\beta^{2^l}$$
-Since the multiplication table is defined as
-\begin{equation}
-\beta
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
-\end{bmatrix}
-= {\bf T}
-\begin{bmatrix}
-\beta \\ \beta^2 \\ \beta^{2^2} \\ \vdots \\ \beta^{2^{n-1}}
-\end{bmatrix}
-\end{equation}
-The $j$-th row of $T$ can be written as
-\begin{align}
-\beta\cdot\beta^{2^j} =& \sum_l\lambda_{0j}^{(l)}\beta^{2^l} & \text{(M-table definition)} \nonumber\\
-					=& \sum_l\lambda_{jl}^{(0)}\beta^{2^l} & \text{(cross-product term)} \label{eqn:mteq}
-\end{align}
-Notice that $\lambda$ corresponds to $\lambda$-Matrix entries. In Equation \ref{eqn:mteq} we assign $0$ to the 
-row index $i$, the proof can be extended to all row indices $i<n$ but fix $l$ to 0 because of the 
-conjugation generated by right-down cyclic-shift of $\lambda$-Matrix. Therefore we have 
-$$M_{i,j}^{(0)} = \lambda_{ij}^{(0)} = \lambda_{j,0}^{(i)} = T_{j-i,0-i}$$
-\end{Proof}
-
-\section{Optimal Normal Basis}
+\chapter{Optimal Normal Basis}
 \label{append:ONB}
@@
 to construct them.
-The following subsection refers to \cite{rosingbook}.
-\subsection{Construction of Optimal Normal Basis}
+The following section refers to \cite{rosingbook}.
+\section{Construction of Optimal Normal Basis}
 {\bf Type-I ONB} over $F_{2^n}$ satisfies following criteria:
@@
 
-In following part we derive a simple way to construct the $\lambda$-Matrix of type-I ONB 
+In following, we derive a simple way to construct the $\lambda$-Matrix of type-I ONB 
 from the criteria above. 
@@
 
-\subsection{Optimal Normal Basis Multiplier Design}
+\section{Optimal Normal Basis Multiplier Design}
 % Customized Galois Field IEEE1363-2000
@@
 
-For type-I ONB over $\F_{2^k}$, the irreducible polynomial $p(\alpha) = \alpha^k+\alpha^{k-1}+\cdots+\alpha+1
-= \sum_{i=0}^k \alpha^i$. For type-II ONB, the following iterative algorithm is required to generate desired 
+For type-I ONB over $\F_{2^k}$, the irreducible polynomial is 
+$$p(\alpha) = \alpha^k+\alpha^{k-1}+\cdots+\alpha+1
+= \sum_{i=0}^k \alpha^i$$
+
+For type-II ONB, the following iterative algorithm is required to generate desired 
 irreducible polynomial:

### 2016-12-24 dissertation/submission/conclude.tex
dissertation after defense update 2

--- 2016-12-17 dissertation/conclude.tex
@@
 $T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
-$S$ (i.e. $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. 
+$S$ ({\it i.e.} $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. 
 In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$
@@
 }
-\caption{A ZDD representing remainder polynomial reducing by a chain of OR gates with order $d>c>b>a$}
+\caption{A ZDD representing remainder polynomial reducing by a chain of OR gates with order $d>c>b>a$.}
 \label{fig:ZDD}
@@
 \includegraphics[width=4.5in]{newfig/project.eps}
-\caption{Algebraic interpolant: Projection of varieties on common variables }
+\caption{Algebraic interpolant: projection of varieties on common variables.}
 \label{fig:projection}}
@@
 
-\subsection{Technology Mapping for Word-level Functional Blocks}
+\subsection{Technology Mapping for Word-Level Functional Blocks}
 Technology mapping is an important problem in digital circuit synthesis.
-Designers are given a library of well-designed functional/macro blocks (including IP cores) and a raw netlist, 
-technology mapping's objective is to map as many as blocks to the raw netlist and 
+Designers are given a library of well-designed functional/macro blocks (including IP cores) and a raw netlist.
+Technology mapping's objective is to map as many as blocks to the raw netlist and 
 keep the functional equivalence. Contemporary techniques rely on bit-wise 
@@
 
-\begin{Problem}
 The objective of our approach is to map the macro blocks without boundary information.
@@
 
-Given a gate-level design $D$ and several word-level macro blocks $\{B_i\}$, we need to map macro
-blocks $B_i$ into design $D$, and write out the mapped design $D'$ which is equivalent to original
-design $D$. The objective is to generate a mapped design $D$ with as many of $B_i$ such that the area
+\begin{Problem}
+Given a gate-level design $N$ and several word-level macro blocks $\{B_i\}$, we need to map macro
+blocks $B_i$ into design $N$, and write out the mapped design $N'$ which is equivalent to original
+design $D$. The objective is to generate a mapped design $N'$ with as many of $B_i$ such that the area
 and timing is optimal. The procedure is also illustrated in Figure \ref{fig:macro}.
@@
 	\end{center}
-	\caption{The outline and flow of technology mapping of macro blocks}
+	\caption{The outline and flow of technology mapping of macro blocks.}
 	\label{fig:macro}
@@
 
-1) {\bf Template and state constraints:}  Ideal membership test technique requires a setup of ideal $J_{loop}$
+1) {\bf Template and state constraints:}  Ideal membership test requires a setup of ideal $J_{loop}$
 for loop invariants. 
-One heuristic can provide better coverage for loop invariant abstraction as well as
+One heuristic that can provide better coverage for loop invariant abstraction as well as a
 relatively small size is \emph{generic quadratic form} (GQF).
@@
 $$\Func = a_0x^2+a_1x+a_2xy+a_3y+a_4y^2+a_5$$
-It covers all possible terms with degree no more than 2. $a_0,a_1,\dots, a_5$ are usually real number parameters,
-some of their assignments can turn $\Func$ into desired invariant. This parameterized constrain
-polynomial covers all combinations of state variables can also be called a \emph{template}.
-Subsequently, finding a proper assignment is the essential part of in our proposed research.
+It covers all possible terms with degree no more than 2. Constants $a_0,a_1,\dots, a_5$ are 
+usually real number parameters,
+some of their assignments can turn $\Func$ into the desired invariant. This parameterized constraint
+polynomial covers all combinations of state variables, and is also called a \emph{template}.
+Subsequently, finding a proper assignment is the essential part of our proposed research.
 
-2) {\bf Initial state constraints:} For initial states, the constrains are explicit. A template is adopted and refined by Gr\"obner basis 
-generated by original constrains, by equaling the remainder to 0 we can get constrains on
+2) {\bf Initial state constraints:} For initial states, the constraints are explicit. 
+A template is adopted and refined by the Gr\"obner basis 
+generated by original constraints, by equaling the remainder to 0 we can get constraints on
 parameters from the template.
@@
 
-Its template is generic quadratic form of $\{s,i,j,j_0\}$, which is
+Its template is the generic quadratic form of $\{s,i,j,j_0\}$, which is
 \begin{align}
-\Func = & a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
-&a_5i^2+a_6i+a_7ij+a_8ij_0+a_9j^2\nonumber\\
-&a_{10}j+a_{11}jj_0+a_{12}j_0^2+a_{13}j_0+a_{14}\nonumber
+\Func &= a_0s^2+a_1s+a_2si+a_3sj+a_4sj_0\nonumber\\
+&+a_5i^2+a_6i+a_7ij+a_8ij_0+a_9j^2\nonumber\\
+&+a_{10}j+a_{11}jj_0+a_{12}j_0^2+a_{13}j_0+a_{14}\nonumber
 \end{align}
@@
 
-Constrains of initial state $s=0\land j=j_0$ can be interpreted to polynomials:
+Constraints of initial state $s=0\land j=j_0$ can be interpreted as polynomials:
 $$\{s, j-j_0\}$$
@@
 \begin{equation}
+\label{eqn:parasys}
 \left\{
@@
 \end{array}\right.
-\nonumber
 \end{equation}
@@
 Our proposed approach models the 2 states individually,
-{i.e.} performs polynomial reduction separately and obtains remainder $r_1$ and $r_2$. 
+{\it i.e.} performs polynomial reduction separately and obtains remainder $r_1$ and $r_2$. 
 Assume a constraint polynomial describing
 the state transition is $r_t$, then we require that when invariant of PS holds and transition $PS \to NS$ stands,
-invariant of NS should also be satisfied:
+the invariant of NS should also be satisfied:
 $$(r_1 = 0)\land (r_t = 0) \implies (r_2 = 0)$$
@@
 $$r_t = r_1 - \lambda r_2$$
-Theoretically $\lambda$ can be any polynomial in arbitrary rings. To make it practical solving the system, 
+Theoretically $\lambda$ can be any polynomial in arbitrary rings. To make it practical, in terms of solving the system, 
 we limit $\lambda$ to the polynomial ring with only real numbers.
 
-We take Equation \ref{eqn:loopinv} as the example (which refers to Example 10 in \cite{sankaranarayanan2004non}). 
+Take Equation \ref{eqn:loopinv} as the example (which refers to Example 10 in \cite{sankaranarayanan2004non}). 
 PS is initial state we just characterized 
@@
 We then perform reduction with the Gr\"obner basis generated from the transition relation modeling, 
-and record its remainder $r_2$.
-From equality $r_2 = \lambda r_1$, we get constraints for the parameters. Solve the system using 
+which is the Gr\"obner basis of polynomial ideal corresponding to the gates in the circuit partition.
+After recording the NS remainder $r_2$,
+by using equality $r_2 = \lambda r_1$ we get constraints for the parameters. We solve the system using 
 similar algorithm to solve binate covering problems
 which is also described in Section 4 as \emph{elimination by splitting} technique in \cite{sankaranarayanan2004non}. 
-One branching result is:
+One of the branching results is:
 \begin{align*}
@@
 {\bf Our proposed approach on technology mapping:}
-Our approach borrows inspiration of templates in \cite{sankaranarayanan2004non}. However, we make an improvement
-on the original approach:
-instead of using template polynomial to describe the system, we add some templates into the ideal
+Our approach is inspired by the concept of templates in \cite{sankaranarayanan2004non}. 
+However, we make an improvement on the original approach:
+instead of using a template polynomial to describe the system, we add some extra templates into the ideal
 of macro blocks which serve as technology mapping candidates. In this way we can cover all possibilities
@@
 1) {\bf System abstraction:}
-The polynomial we use to test ideal membership should include all information of a circuit partition,
+The polynomial we use to test ideal membership should include all the information of a circuit partition,
 which requires us to abstract information from the system and write it into a single polynomial.
@@
 $$Z + f(i_1,i_2,\dots,i_n)$$
-where $Z$ is the output. When there is only one output, $Z$ collapse to a bit-level variable. However in most
+where $Z$ is the output. When there is only one output, $Z$ collapses to a bit-level variable. However in most
 cases there are multiple outputs on the cut, indicating $Z$ as a word-level variable. Boolean function $\Func$ covers
@@
 	\end{center}
-	\caption{An example gate-level netlist to be mapped}
+	\caption{An example gate-level netlist of a target sub-circuit to be mapped. ``$\blacksquare$" denotes a pin, 
+	which is the boundary for the mapper.}
 	\label{fig:tobemapped}
@@
 
-Figure \ref{fig:tobemapped} shows an example circuit partition with $a_1,b_1,a_0,b_0$ as inputs and
+Figure \ref{fig:tobemapped} shows an example circuit partition with $s_0,s_1,s_2,s_3$ as inputs and
 $Z = \{z_0,z_1\}$ as outputs. If we use elements from Galois field $\F_{2^2}$ to represent word $Z$,
@@
 After imposing ATO of LEX with
-$$Other\ circuit\ variables > output\ word\ Z > all\ bit\ level\ inputs$$
+$$\textit{Other\ circuit\ variables} > \textit{output\ word\ } Z > \textit{all\ cut-set\ bit\ level\ inputs}$$
 on the ideal describing the partitioned circuit as well as ideal with vanishing polynomials,
-the reduced Gr\"obner basis has a single polynomial generator in the form of $Z + \Func(a_1,b_1,a_0,b_0)$.
+the reduced Gr\"obner basis has a single polynomial generator in the form of $Z + \Func(s_0,s_1,s_2,s_3)$.
 In this example:
@@
 \begin{itemize}
-\item Term ordering: $\{z_0,z_1,t,g_1,p_1,ci,t_1,t_2\}>Z>\{a_0,b_0,a_1,b_1$
-\item Gate description: $z_0+t+g_1+t\cdot g_1, t+t_1+t_2+t_1\cdot t_2, g_1+a_1\cdot b_1,
-			t_1+a_1\cdot ci, t_2+b_1\cdot ci, ci+a_0\cdot b_0, z_1+p_1+ci, p_1+a_1+b_1$
+\item Term ordering: $\{z_0,z_1,t,g_1,p_1,ci,t_1,t_2\}>Z>\{s_0,s_1,s_2,s_3\}$
+\item Gate description: $z_0+t+g_1+t\cdot g_1, t+t_1+t_2+t_1\cdot t_2, g_1+s_0\cdot s_1,
+			t_1+s_0\cdot ci, t_2+s_1\cdot ci, ci+s_2\cdot s_3, z_1+p_1+ci, p_1+s_0+s_1$
 \item Word definition: $Z+z_0+z_1\cdot \alpha$
 \item Vanishing polynomial ideal $(J_0)$: $z_0^2+z_0, z_1^2+z_1, t^2+t, g_1^2+g_1, t_1^2+t_1, t_2^2+t_2, p_1^2+p_1, ci^2+ci,
-			a_0^2+a_0, b_0^2+b_0, a_1^2+a_1, b_1^2+b_1, Z^4+Z$(since Z is a 2-bit word)
+			s_2^2+s_2, s_3^2+s_3, s_0^2+s_0, s_1^2+s_1, Z^4+Z$ (since Z is a 2-bit word)
 \end{itemize}
@@
 The result is a Gr\"obner basis with single polynomial generator. The polynomial has leading term $Z$:
-$$Z+a_0\cdot b_0\cdot a_1+a_0\cdot b_0\cdot b_1+\alpha\cdot a_0\cdot b_0+a_1\cdot b_1+\alpha\cdot a_1+\alpha\cdot b_1$$
+$$Z+s_2\cdot s_3\cdot s_0+s_2\cdot s_3\cdot s_1+\alpha\cdot s_2\cdot s_3+s_0\cdot s_1+\alpha\cdot s_0+\alpha\cdot s_1$$
 
@@
 	\end{center}
-	\caption{Standard 2-bit adder with input/output mapping}
+	\caption{Candidate cell: a standard 2-bit adder with one possible input/output mapping.}
 	\label{fig:template}
@@
 
-Figure \ref{fig:template} shows a 2-bit standard cell. It has 3 outputs mapped to 2 output pins,
-and 4 inputs mapped to 3 input pins (the rest pins are assigned to fixed 0/1 signal). Dashed connection
+Figure \ref{fig:template} shows a 2-bit standard cell. It has 3 outputs ($t_0,t_1,CO$) mapped to 2 output pins ($z_0,z_1$),
+and 4 inputs ($a_0,a_1,b_0,b_1$) mapped to 3 input pins ($s_0,s_1,s_2$, the excessive input {\it e.g.} $b_1$ is assigned to fixed 0/1 signal). Dashed connection
 lines show one possible mapping, to find out this kind of feasible mapping, we need to simulate
@@
 
-{\it Output template}: $z_0+c_{{t_0}{z_0}}\cdot t_0+c_{t_1z_0}\cdot t_1+c0_{CO}\cdot CO+c_{n_0}$
+{\it Output template}: $z_0+c_{{t_0}{z_0}}\cdot t_0+c_{t_1z_0}\cdot t_1+c_{z_0CO}\cdot CO+c_{n_0}$
 
-When $c_{n_0} = 0$, then $\{c_{t_0z_0},c_{t_1z_0},c0_{CO}\}$ equals to 1 denotes a valid mapping to corresponding output pin. 
-Conversely if $cn0 = 1$, the ``1" evaluation denotes mapping to negation of corresponding output pin.
+Above polynomial includes all possibilities where $z_0$ could be mapped to.
+When $c_{n_0} = 0$, then $\{c_{t_0z_0},c_{t_1z_0},c_{z_0CO}\}$ equals to 1 denotes a valid mapping to corresponding output pin. 
+Conversely if $c_{n_0} = 1$, the ``1" evaluation denotes mapping to negation of corresponding output pin.
 
-{\it Input template}: $a_0+p_{a_0}+c0_{A_0}\cdot s_0+c1_{A_0}\cdot s_1+c2_{A_0}\cdot s_2$
+{\it Input template}: $a_0+c_{s_0a_0}\cdot s_0+c_{s_1a_0}\cdot s_1+c_{s_2a_0}\cdot s_2+p_{a_0}$
 
-When $p_{a_0} = 0$, then $\{c0_{A_0},c1_{A_0},c2_{A_0}\}$ equal to 1 denotes a valid mapping to corresponding input
-pin. Conversely if every variable in set $\{c0_{A_0},c1_{A_0},c2_{A_0}\}$ equals to 0, it denotes mapping to fixed signal ``0" (when $p_{a_0} = 1$ then
+When $p_{a_0} = 0$, then $\{c_{s_0a_0},c_{s_1a_0},c_{s_2a_0}\}$ equal to 1 denotes a valid mapping to corresponding input
+pin. Conversely if every variable in set $\{c_{s_0a_0},c_{s_1a_0},c_{s_2a_0}\}$ equals to 0, 
+it denotes mapping to fixed signal ``0" (when $p_{a_0} = 1$ then
 mapping to fixed signal ``1").
@@
 \begin{itemize}
-\item First, choose a set of independent wires as input pins;
+\item First, choose a cut-set of independent wires in target circuit as input pins;
 
-\item Second, push forward input signal for a certain number of gates (depth), choose a cut set
-which fully dependent on these inputs as output pins;
+\item Second, propagate forward the cut for a certain number of gates (depth), choose a cut-set
+which fully dependents on these inputs as output pins. Thus we obtain a sub-circuit by partitioning;
 
-\item Third, abstract a description polynomial using GB based abstraction;
+\item Third, abstract a description polynomial $\Func$ of the target sub-circuit using GB based abstraction;
 
-\item Fourth, reduce the description polynomial with the GB we computed associated to macro block 
+\item Fourth, reduce the description polynomial $\Func$ with the GB we computed associated to macro block 
 and templates, obtain a remainder polynomial;
 
-\item Last but not least, abstract all coefficients from the remainder polynomial and set up a system of equations.
+\item Last, but not the least, abstract all coefficients from the remainder polynomial and set up 
+a system of equations similar to Equation \ref{eqn:parasys}.
 If we find a solution to this system, then the solution is a feasible mapping; no solution means we cannot make a

### 2016-12-24 dissertation/submission/normal.tex
dissertation after defense update 2

--- 2016-12-11 dissertation/normal.tex
@@
 \includegraphics[width=\textwidth]{newfig/Moore.eps}
-\caption{A typical Moore machine and its state transition graph}
+\caption{A typical Moore machine and its state transition graph.}
 \label{fig:Moore}}
@@
 \includegraphics[width=\textwidth]{newfig/convention.eps}
-\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking}
+\caption{Conventional verification techniques based on bit-level unrolling and equivalence checking.}
 \label{fig:convention}}
@@
 \includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
-\caption{Architecture of a combinational GF multiplier}
+\caption{Architecture of a combinational GF multiplier.}
 \label{fig:comb_RAB}}

### 2016-12-24 dissertation/submission/prelim_GF.tex
dissertation after defense update 2

--- 2016-12-11 dissertation/prelim_GF.tex
@@
 field $\F_2$.
-\begin{Example}
-Addition and multiplication operations over $\F_2$:
+
 \begin{table}[!h]
+\caption{Addition and multiplication operations over $\F_2$.}
 	\centering
@@
 Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.
-\end{Example}
 
@@
 \caption{Bit-vector, Exponential and Polynomial representation of
-elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$}\label{tab:gfelement} 
+elements in  $\mathbb{F}_{2^4} = \mathbb{F}_2[x] \pmod{x^4+x^3+1}$.}\label{tab:gfelement} 
 \begin{tabular}{|c|c|c||c|c|c|} 
@@
 \includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
-\caption{A typical SMSO structure of Massey-Omura multiplier}
+\caption{A typical SMSO structure of Massey-Omura multiplier.}
 \label{fig:MasseyOmura}}
@@
 Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, {\it i.e.} if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
-Index $w = 2i\pmod 5$}
+Index $w = 2i\pmod 5$.}
 \label{fig:SMPO}}
@@
 \includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
-\caption{5-bit RH-SMPO}
+\caption{A 5-bit RH-SMPO.}
 \label{fig:5bitRH}}
@@
 \includegraphics[width=3in]{newfig/multitable.pdf}
-\caption{A $5\times 5$ multiplication table}
+\caption{A $5\times 5$ multiplication table.}
 \label{fig:multitable}}

### 2016-12-24 dissertation/submission/prev.tex
dissertation after defense update 2

--- 2016-12-11 dissertation/prev.tex
@@
 }
-\caption{The unrolling of a sequential circuit}
+\caption{The unrolling of a sequential circuit.}
 \label{fig:unrolling}
@@
 }
-\caption{ROBDD representing Boolean function $\neg a \land b \land (c\lor d)$ with order $a>b>c>d$}
+\caption{ROBDD representing Boolean function $\neg a \land b \land (c\lor d)$ with order $a>b>c>d$.}
 \label{fig:BDD}
@@
 It implemented latch abstraction by removing "irrelevant" latches by analyzing the 
-UNSAT core from the $k$-BMC. Jain {\it et al.} \cite{HimanshuDAC2005} improved the abstraction refinement technique of \cite{clarke2000counterexample},
+UNSAT core from the $k$-BMC. Jain {\it et al.} \cite{jain2005word} improved the abstraction refinement technique of \cite{clarke2000counterexample},
 where they use CNF-SAT to perform the refinement instead of using BDDs. The new approach is applied to verify RTL Verilog
@@
 \cite{pruss:dac14} performs verification by deriving a canonical
-word-level polynomial representation $\F$ from the circuit $C$. Their
+word-level polynomial representation $\Func$ from the circuit $C$. Their
 approach views any arbitrary Boolean function (circuit) $f: \B^k
 \rightarrow \B^k$ as a polynomial function $f: \Fkk \rightarrow \Fkk$,
-and derives a canonical polynomial representation $\F$ over
+and derives a canonical polynomial representation $\Func$ over
 $\Fkk$. They show that this can be achieved by computing a reduced 

### 2016-12-24 dissertation/submission/recha.tex
dissertation after defense update 2

--- 2016-12-11 dissertation/recha.tex
@@
 \includegraphics[width=\textwidth]{newfig/MIPS.pdf}
-\caption{A 64-bit sequential multiplication hardware design}
+\caption{A 64-bit sequential multiplication hardware design.}
 \label{fig:MIPS}}
@@
 \includegraphics[width=4in]{newfig/Lagrange.pdf}
-\caption{Gate-level netlist for Lagrange's interpolation example}
+\caption{Gate-level netlist for Lagrange's interpolation example.}
 \label{fig:Lagrange}}
@@
 \label{tab:truthtable}
+\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
 \centering{\small
@@
 }
-\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
 \end{table}
@@
 % \vspace{-0.2in}
-\caption{FSM models of sequential circuits}
+\caption{FSM models of sequential circuits.}
 %\end{minipage}
@@
 \includegraphics[width=\textwidth]{newfig/proj_reacha.pdf}
-\caption{Projection of the variety of circuit description ideal}
+\caption{Projection of the variety of circuit description ideal.}
 \label{fig:proj_reacha}}
@@
 \includegraphics[width=\textwidth]{newfig/PI.pdf}
-\caption{PI partition of a sequential circuit}
+\caption{PI partition of a sequential circuit.}
 \label{fig:PI}}
@@
 \includegraphics[width=\textwidth]{newfig/flowchart.pdf}
-\caption{Execution process of word-level FSM traversal tool}
+\caption{Execution process of word-level FSM traversal tool.}
 \label{fig:flowchart}}
@@
 \includegraphics[width=\textwidth]{newfig/SING.pdf}
-\caption{Singular script for executing bit-to-word substitution and traversal loop}
+\caption{Singular script for executing bit-to-word substitution and traversal loop.}
 \label{fig:SING}}
@@
 \includegraphics[width=\textwidth]{newfig/Recha_result.pdf}
-\caption{The output given by our traversal tool}
+\caption{The output given by our traversal tool.}
 \label{fig:Recha_result}}
@@
  & \multirow{3}{0.9cm}{\centering \# States}
- & \multirow{3}{1.3cm}{\centering \# iterations}
+ & \multirow{3}{1.2cm}{\centering \# iterations}
  & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
- & \multirow{3}{1.8cm}{\centering Runtime of VIS (sec)} \\
+ & \multirow{3}{1.3cm}{\centering Runtime of VIS (sec)} \\
   & & & & & &\multicolumn{3}{c|}{}& \\

### 2016-12-24 dissertation/submission/unsat.tex
dissertation after defense update 2

--- 2016-12-11 dissertation/unsat.tex
@@
 \centering{
-\includegraphics[scale=0.3]{newfig/fig_SAT.eps}
-\caption{An example of Boolean satisfiability problem on circuits}
-\label{fig:SAT}}
+\includegraphics[scale=0.3]{newfig/fig_SAT.eps}}
+\caption{An example of Boolean satisfiability problem on circuits.}
+\label{fig:SAT}
 \end{figure}
@@
 
-\begin{table}[htb]
+\begin{table}[h]
+\caption{Mapping Boolean operators to functions over $\Fkk$}
 \centering
@@
 \end{tabular}
-\caption{Mapping Boolean operators to functions over $\Fkk$}
 \label{tab:booltof4}  
@@
 
-\begin{figure}[hbt]
+\begin{figure}[h]
 \centering{
-\includegraphics[width=5.5in]{newfig/refine.eps}
-\caption{Abstraction by reducing latches}
-\label{fig:refine}}
+\includegraphics[width=5.5in]{newfig/refine.eps}}
+\caption{Abstraction by reducing latches.}
+\label{fig:refine}
 \end{figure}
@@
 \end{Example}
-\begin{figure}[hbt]
+\begin{figure}[h]
 \centering{
 \includegraphics[width=\textwidth]{newfig/UNSAT.eps}
-\caption{DAG representing $Spoly$ computations and multivariate divisions}
-\label{fig:UNSAT}}
+\caption{DAG representing $Spoly$ computations and multivariate divisions.}}
+\label{fig:UNSAT}
 \end{figure}
@@
 
-\begin{figure}[hbt]
-\centering
-\includegraphics[width=4.5in]{newfig/refutation_tree.eps}
-\caption{Generating refutation trees to record UNSAT cores}
+\begin{figure}[h]
+\centering{
+\includegraphics[width=4.5in]{newfig/refutation_tree.eps}}
+\caption{Generating refutation trees to record UNSAT cores.}
 \label{fig:refute}
@@
 refutation tree as shown in Figure \ref{fig:refine}(a).  
-\begin{figure}[hbt]
-\centering
-\includegraphics[width=\textwidth]{newfig/core_refine.eps}
-\caption{Refutation trees of core refinement example}
+\begin{figure}[h]
+\centering{
+\includegraphics[width=\textwidth]{newfig/core_refine.eps}}
+\caption{Refutation trees of core refinement example.}
 \label{fig:refine}
@@
 \centering{
-\includegraphics[width=\textwidth]{newfig/s27.pdf}
-\caption{Gate-level Schematic of Example Circuit}
-\label{fig:s27}}
+\includegraphics[width=\textwidth]{newfig/s27.eps}
+\caption{Gate-level schematic of the example circuit.}}
+\label{fig:s27}
 \end{figure}
@@
 \centering{
-\includegraphics[width=4.5in]{newfig/s27_stg.pdf}
-\caption{State transition graph of example circuit}
-\label{fig:s27_stg}}
+\includegraphics[width=4.5in]{newfig/s27_stg.pdf}}
+\caption{State transition graph of example circuit.}
+\label{fig:s27_stg}
 \end{figure}
@@
 \begin{figure}[H]
-\centering{
-\includegraphics[width=4.5in]{newfig/s27_stg_refined.pdf}
-\caption{State transition graph of abstracted machine}
-\label{fig:s27_stg_refined}}
+\centerline{\includegraphics[width=4.5in]{newfig/s27_stg_refined.pdf}}
+\caption{State transition graph of abstracted machine.}
+\label{fig:s27_stg_refined}
 \end{figure}
@@
 \hline
-\multirow{3}{2.2cm}{\centering Benchmark} 
-& \multirow{3}{0.9cm}{\centering \# Polys} 
-& \multirow{3}{0.8cm}{\centering \# MUS} 
-& \multicolumn{3}{c|}{\multirow{2}{1.8cm}{\centering Size of core}}
- & \multirow{3}{1.6cm}{\centering \# GB-core iterations}
- & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
- & \multirow{3}{1.8cm}{\centering Runtime of PicoMUS (sec)} \\
+\multirow{3}{1.8cm}{\centering Benchmark} 
+& \multirow{3}{0.7cm}{\centering \# Polys} 
+& \multirow{3}{0.7cm}{\centering \# MUS} 
+& \multicolumn{3}{c|}{\multirow{2}{1.5cm}{\centering Size of core}}
+ & \multirow{3}{1.3cm}{\centering \# GB-core iterations}
+ & \multicolumn{3}{c|}{\multirow{2}{1.8cm}{\centering Runtime (sec)}}
+ & \multirow{3}{1.6cm}{\centering Runtime of PicoMUS (sec)} \\
   & & &\multicolumn{3}{c|}{}& &\multicolumn{3}{c|}{}& \\

### 2016-12-31 dissertation/submission/acknowledge.tex
dissertation milestone: submitting to thesis office

I was born in a small town in mid-inland-China. My grandpa, Shouli Sun, 
is the first man to educate me. He was so kind and patient, and his education 
was so successful that I could read 2,000 Chinese characters at the age of 3, 
when my grandpa passed away. With a good start, I became used to being the 
first place in a variety of exams throughout elementary school, middle school, 
and high school. Peking university seemed to be the end of my personal 
progress in education, until I learnt that there is a way to continue my 
graduate study abroad, more specifically, in the US. So I would like to 
thank Dr. Fu, Mr. Qiu and Mr. Yang for providing such information and 
suggestions as my roommates at that time.

The beginning of my graduate school experience was not 
so wonderful. I got lost among the labs of the ECE department since there are 
too many directions and categories of research. Finally, I found out what
I am really interested in after taking the lectures by my advisor, 
Prof. Priyank Kalla. He led the way towards designing a beautiful and powerful
algorithm and applying it to the ``mission impossible" -- verifying
a large circuit with 100\% fault coverage. I would like to thank him for
his valuable guidance, magnificent inspirations, and of course, his generous 
financial support that allow me to carry out my research. Meanwhile, I would 
like to thank four professors serving as my supervisory committee members, 
for their kind help in assisting my research and coursework, as well as the 
review of this dissertation.

At last, I want to say thank you to the most important people in my life.
My parents, Feng Sun and Qinghua Zhang, dedicate themselves to raising me
in a happy family for 26 years. They also keep motivating me to pursue higher 
education and more achievements in schools and universities. My dear wife,
Dr. Huizhong Li, has been my greatest support in the past 10 years. It is our 
precious time working together towards the same goal in the same university, and 
I will remember our life full of sweets and bitters in Salt Lake
City forever.

### 2016-12-31 dissertation/submission/appendix1.tex
dissertation milestone: submitting to thesis office

@@
 For general normal basis identification, there are two methods widely used: the L\"uneburg's algorithm
-and Lenstra's algorithm.\\
-\textbf{L\"uneburg's Algorithm:}
+and Lenstra's algorithm.
+
+\subsection{L\"uneburg's Algorithm}
+L\"uneburg's algorithm can be described with following steps:
+
 \begin{enumerate}[{1)}]
@@
 
+\subsection{Lenstra's Algorithm}
 L\"uneburg's algorithm is an analytic method, which requires relatively high computational complexity 
@@
 
-Based on Lemmas \ref{lem:4} and \ref{lem:5}, the Lenstra's algorithm is described below:\\
-\textbf{Lenstra's Algorithm}
+Based on Lemmas \ref{lem:4} and \ref{lem:5}, the Lenstra's algorithm is described below:
+
 \begin{enumerate}[{1)}]

### 2016-12-31 dissertation/submission/conclude.tex
dissertation milestone: submitting to thesis office

@@
 next state. However, in some situations we need to keep recording the relations between $T$ and
-inputs, {\it i.e.} the reached states contain multivariate polynomials in $T,S$ and PI $x$. 
-While the elimination and set union/intersection are compatible with multiple NS variables,
+the inputs, {\it e.g.} the reached states contain multivariate polynomials in $T$, as well as 
+PS variable $S$ and PI $x$. 
+While the elimination and set union/intersection are compatible with multivariate polynomials,
 the set complement requires an extension of Theorem \ref{thm:quotient} on ideals with 
@@
 \begin{Example}
-In Figure \ref{fig:fsm}
+In Figure \ref{fig:fsm}, we denote that
 $\{s_0, s_1\}$ are state/pseudo inputs, $\{t_0,t_1\}$ are state/pseudo outputs, and there is a primary input (1-bit) 
-$x$. We propose a new algorithm by modifying Algorithm \ref{alg:univa}:
+$x$. We propose a new algorithm (Algorithm \ref{alg:multi}) by modifying Algorithm \ref{alg:univa}.
 
@@
 
-The inputs of this algorithm includes transition polynomial (result of word-level abstraction) and initial 
-states description ideal, which contains 2 generators corresponds to constraints of PI and combinational input $S$. 
+The inputs of this algorithm includes the transition polynomial (result of word-level abstraction) and initial 
+states description ideal, which contains 2 generators corresponding to constraints of PI and combinational input $S$. 
 For example, $\langle S+1+\alpha, x^2+x\rangle$ means initial state$=\{11\}$,
@@
 
-Transition polynomial calculation uses the abstraction and bit-to-level conversion method from 
+Transition polynomial calculation uses the abstraction and bit-to-word conversion method from 
 Chapter \ref{ch:reacha}. After Constructing
-an elimination ideal, we impose RATO such that \emph{reverse\ topo\ order\ ckt\ variables }$> T > S > x$, the reduction
-remainder has the form $T+\mathcal F(s_0,s_1,x)$. From word definition $S+s_0+s_1\alpha$ we get
+an elimination ideal, we impose RATO such that \textit{reverse\ topo\ order\ ckt\ variables }$> T > S > x$, the reduction
+remainder has the form $T+\mathcal F(s_0,s_1,x)$. According to the bit-to-word correspondence $S+s_0+s_1\alpha$ we get
 $$s_0 = \alpha S^2+ (1+\alpha)S, s_1 = S^2+S$$
-Do substitution, the transition polynomial of example ckt is 
+After substitution, we obtain the transition polynomial of example circuit: 
 $$f_T = T+S^3\cdot x+\alpha S^3+(1+\alpha)S^2\cdot x+S^2+S\cdot x+(1+\alpha)x+1$$
 
-Assume we start from state $\{11\}$. In the first iteration, the reached state is $\{01\}$. Line 4 is to compose an
-ideal with 2 generators from $from^0$ and transition polynomial $f_T$, compute its Gr\"obner basis. Note this ideal
+Assume the initial state is $\{11\}$. In the first iteration, the reached state is $\{01\}$. Line 4 is to compose an
+ideal with 2 generators from $from^0$ and transition polynomial $f_t$, compute its Gr\"obner basis. Note this ideal
 has the form
@@
 \end {equation}
-$v_x$ is a polynomial containing only $q$-bit PI $x$, initially it should be vanishing polynomial $x^{q_1}-x$, 
+$v_x$ is a polynomial containing only $q_1$-bit PI $x$, initially it should be vanishing polynomial $x^{q_1}-x$, 
 with the program executing it may be factorized.
 
-Consider Buchberger's algorithm, all generators' leading terms are relatively prime, so it is a GB itself. 
-Then we need to reduce this GB, we will find out $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
-and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. So, at last we will get a polynomial
-$T + \mathcal F'(x)$ in reduce GB. We can include this polynomial and $v_x$, exclude the polynomial containing
-$S$ ({\it i.e.} $\mathcal H(S)$ in algorithm), to compose an ideal representing \emph{next states} $to^i$. 
-In iteration 1, result is $to^1 = \langle T+1, x^2+x\rangle$
+Considering the Buchberger's algorithm, all generators' leading terms are relatively prime, so $I_{tran}$ is a GB itself. 
+Furthermore, generator $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
+and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. In the end we obtain a polynomial
+$T + \mathcal F'(x)$ in the reduced GB. We can then include this polynomial and $v_x$ and exclude the polynomial containing
+$S$ ({\it i.e.} $\mathcal H(S)$ in Algorithm \ref{alg:multi}), to compose an ideal representing \emph{next states} $to^i$. 
+In iteration 1, the result is $to^1 = \langle T+1, x^2+x\rangle$
 
@@
 Line 6 is ideals' sum (intersection of their varieties), it is done by combining all generators
-from 2 ideals and compute GB. For first iteration result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.
+from 2 ideals and compute GB. In the first iteration, the result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.
 
 Line 7 is ideals' product (union of their varieties), it is done by multiplying all pairs of
-generators from both ideal. For first iteration result is $GB(\langle (T+1)(T+1+\alpha),
+generators from both ideal. For the first iteration it gives the result as $GB(\langle (T+1)(T+1+\alpha),
 (T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$ representing $\{01,11\}$.
 
-The traversal will run 3 iterations and terminate at 4th iteration. I list all intermediate 
+The traversal algorithm executes 3 iterations and terminate at 4th iteration. We list all intermediate 
 results below:
@@
 \end{itemize}
-The final reachable states are represented by multivariate polynomial ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
+The final reachable states are represented by a multivariate polynomial ideal $\langle T^3\cdot x+x, T^4+T, x^2+x\rangle$,
 which denotes $\{00,01,10,11\}$.
@@
 Equation \ref{eqn:chainOR} with size $2\times4 -1 = 7$, which is linear in the number of variables.
-The reduction process using ZDDs can be executed as in \cite{polybori:2009}. The graphical illustration 
+The reduction process using ZDDs can be executed as in \cite{PolyBoRi}. The graphical illustration 
 of the remainder in ZDDs is also shown in Figure \ref{fig:ZDD}.

### 2016-12-31 dissertation/submission/unsat.tex
dissertation milestone: submitting to thesis office

@@
 \item Update $F=F\cup r_2$;
-\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:UNSAT}(a) shows;
+\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:originUNSAT}(a) shows;
 \item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
@@
 terminate the Buchberger's algorithm;
-\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:UNSAT}(b). 
+\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:originUNSAT}(b). 
 From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
@@
 \end{Example}
+
 \begin{figure}[h]
 \centering{
-\includegraphics[width=\textwidth]{newfig/UNSAT.eps}
-\caption{DAG representing $Spoly$ computations and multivariate divisions.}}
-\label{fig:UNSAT}
+\includegraphics[width=\textwidth]{newfig/UNSAT.eps}}
+\caption{DAG representing $Spoly$ computations and multivariate divisions.}
+\label{fig:originUNSAT}
 \end{figure}
@@
 \centering{
-\includegraphics[width=\textwidth]{newfig/s27.eps}
-\caption{Gate-level schematic of the example circuit.}}
+\includegraphics[width=\textwidth]{newfig/s27.eps}}
+\caption{Gate-level schematic of the example circuit.}
 \label{fig:s27}

### 2017-01-24 dissertation/submission/intro.tex
dissertation after defense update 2
dissertation: after submission_1st revision

--- 2016-12-11 dissertation/intro.tex
@@
 \label{ch:intro}
-
+\vspace{-0.8cm}
 \section{Hardware Design and Verification Overview}
@@
 
-{
-\begin{figure}[h]
-\centerline{
-\includegraphics[width=0.7\textwidth]{newfig/designflow.eps}
-}
-\caption{Typical hardware design flow}
-\label{fig:designflow}
-\end{figure}
-}
-
 Another method developed besides simulation is \emph{formal verification}, it utilizes 
@@
 \section{Formal Verification: Property and Equivalence Checking}
-
 {\it Property checking} (or property verification) verifies
@@
 and checking layout verification between netlist and layout for fabrication.
+
+\begin{figure}[bp]
+\centerline{
+\includegraphics[width=0.7\textwidth]{newfig/designflow.eps}
+}
+\caption{Typical hardware design flow.}
+\label{fig:designflow}
+\end{figure}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[bp]
 \centering{
 \includegraphics[scale=0.3]{newfig/fig_SAT.eps}
-\caption{An example of equivalence checking on miter of circuit $A$ and $B$ using SAT}
+\caption{An example of equivalence checking on miter of circuit $A$ and $B$ using SAT.}
 \label{fig:miter}}

### 2017-01-24 dissertation/submission/normal.tex
dissertation: after submission_1st revision

@@
 
-\begin{figure}[H]
+\begin{figure}[bp]
 \centering{
@@
 \subsection{Implicit Unrolling based on Abstraction with ATO}
-\begin{figure}[H]
-\centering{
-\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
-\caption{Architecture of a combinational GF multiplier.}
-\label{fig:comb_RAB}}
-\end{figure}
-
 The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
@@
 
-For 2-input sequential NB multipliers, abstraction is utilized to implement following algorithm:
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/comb_RAB.pdf}
+\caption{Architecture of a combinational GF multiplier.}
+\label{fig:comb_RAB}}
+\end{figure}
+
 %\vspace{-0.1in}
-\IncMargin{1em}
+% \IncMargin{1em}
 \begin{algorithm}[hbt]
@@
 \end{algorithm}
-\DecMargin{1em}
+% \DecMargin{1em}
 
-Notations in the following figure as well as in next example comply with 
+For 2-input sequential NB multipliers, abstraction is utilized to implement Algorithm \ref{alg:modified}.
+Notations in Figure \ref{fig:sequential} as well as in next example comply with 
 the notations in this algorithm.
 
-\begin{figure}[H]
+\begin{figure}[bp]
 \centering{
@@
 $$R' = \Func(A,B,R)$$
-
 Using ATO, if we put $R$ ahead of $R',A,B$ in term ordering, $R$ is thus eliminated
@@
 $$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$
-
 \begin{eqnarray*}
@@
 with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
-in following table:
+in Table \ref{tab:slimgb}.
 
-\begin{table}[H]
+\begin{table}[bp]
 \begin{center}
@@
 $Spoly(f_w,f_g) \xrightarrow{J_{ckt}+J_0}_{+}r$.
-\item Construct matrices $M_0,\dots,M_{k-1}$, where $M_i$ is $M$ with the $i$-column replaced by vector 
-$[R~~R^2~~\cdots~~R^{2^{k-1}}]^T$, and $M$ is the reverse-circulant matrix generated by vector $(\beta,\beta^2,\dots,\beta^{2^{k-1}})$.
+\item Construct matrices $M_0,\dots,M_{k-1}$, where $M_i$ is $M$ with the $i$-column replaced by circulant vector 
+$[R~~R^2~~\cdots~~R^{2^{k-1}}]^T$, and $M$ is the reverse-circulant matrix generated by circulant vector $(\beta,\beta^2,\dots,\beta^{2^{k-1}})$.
 \item Symbolically compute the determinants $|M_i|$ to find set $F_R$, where $f_{R_i}:R_i+|M_i|$, for $0\leq i \leq k-1$.
@@
 
-\vspace{-0.1in}
-\begin{table}[htb]
+% \vspace{-0.1in}
+\begin{table}[bp]
 \centering
@@
 
-\begin{table}[htb]
+\begin{table}[bp]
 \centering
@@
 
-\begin{table}[hbt]
+\begin{table}[bp]
 \centering

### 2017-01-24 dissertation/submission/prelim_GF.tex
dissertation: after submission_1st revision

@@
 We are specifically interested in extension fields of type 
-$\Fkk$, where $k > 1$. These are extensions of the binary
-field $\F_2$.
-
-\begin{table}[!h]
-\caption{Addition and multiplication operations over $\F_2$.}
+$\Fkk$, where $k > 1$, because they are extensions of the binary
+field $\F_2$. Elements and operators in $\F_2$ can be mapped to 
+Boolean values and Boolean operators, as Table \ref{tab:1stTab} shows.
+Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
+because it is performed modulo $2$.
+Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.
+
+\begin{table}[bp]
+% \vspace{-0.3cm}
+\caption{Addition and multiplication operations over $\F_2$.\vspace{0.5cm}}
 	\centering
+	\label{tab:1stTab}
 	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
+% 	\indent\\
 	\hhline{~---~}
@@
 	\begin{tabular}{m{1cm}|l|ll|m{1cm}}
+% 	\indent\\
 	\hhline{~---~}
@@
 	\end{tabular}
+
 \end{table}
-
-Notice that addition over $\F_2$ is a Boolean {\sc XOR} operation, 
-because it is performed modulo $2$.
-Similarly, multiplication over $\F_2$ performs a Boolean {\sc AND} operation.
 
@@
 
-\begin{table}[h]
+\begin{table}[bp]
 \begin{center}
@@
 
-\begin{figure}[H]
-\centering{
-\includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
-\caption{A typical SMSO structure of Massey-Omura multiplier.}
-\label{fig:MasseyOmura}}
-\end{figure}
-
 Figure \ref{fig:MasseyOmura} shows the basic architecture of a Massey-Omura multiplier. The operands
@@
 
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/MasseyOmura.pdf}
+\caption{A typical SMSO structure of Massey-Omura multiplier.}
+\label{fig:MasseyOmura}}
+\end{figure}
+
 The gate-level logic boxes implement the following function:
@@
 was invented by Agnew {\it et al.} \cite{agnew1991implementation}.
-
-\begin{figure}[H]
-\centering{
-\includegraphics[width=\textwidth]{newfig/mySMPO.eps}
-\caption{5-bit Agnew's SMPO.
-Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, {\it i.e.} if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
-Index $w = 2i\pmod 5$.}
-\label{fig:SMPO}}
-\end{figure}
 
@@
 Then a typical design of 5-bit Agnew's SMPO is depicted in Figure \ref{fig:SMPO}.
+
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/mySMPO.eps}
+\caption{5-bit Agnew's SMPO.
+Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, {\it i.e.} if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
+Index $w = 2i\pmod 5$.}
+\label{fig:SMPO}}
+\end{figure}
 
@@
 
-\begin{figure}[H]
-\centering{
-\includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
-\caption{A 5-bit RH-SMPO.}
-\label{fig:5bitRH}}
-\end{figure}
-
 Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
@@
 \SetAlgoNoLine
-
+\LinesNumbered
  \KwIn{$A,B\in \Fkk$ given {\it w.r.t.} NB $N$}
@@
 So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
-$F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
+sequence $F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
 \end{Proof}
+
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
+\caption{A 5-bit RH-SMPO.}
+\label{fig:5bitRH}}
+\end{figure}
 
@@
 \end{equation}
-Consider indices $4+1=0\bmod 5,~4+2=1\bmod 5$, write down gates in $c$-layer and $d$-layer (besides $d_0$)
+Consider equality of indices $4+1=0\bmod 5,~4+2=1\bmod 5$, we can write down gates in $c$-layer and $d$-layer (besides $d_0$)
 $$c_1 = a_0+a_4,~c_2 = b_0+b_4,~d_1=d_{4,1}= c_1c_2 = (a_4+a_0)(b_4+b_0)$$

### 2017-01-24 dissertation/submission/prelim_basis.tex
dissertation after defense update 2
dissertation: after submission_1st revision

--- 2016-12-11 dissertation/prelim_basis.tex
@@
 \SetAlgoNoLine
-
+\LinesNumbered
  \KwIn{$f,f_{1},\dots,f_{s}$}
@@
 Thus, every monomial in $Spoly(f, g)$ is divisible by either $lt(f)$ 
-or $lt(g)$, so computing 
+or $lt(g)$, so computing the remainder of \textit{Spoly}: 
 $Spoly(f, g) \stackrel{f,g}{\textstyle\longrightarrow}_+r$ will give $r=0$.
@@
 {\it projecting} the variety $V(\langle f_1,\dots,f_s\rangle)$ from 
-$\Fq^d$ to $\Fq^{d-l}$. Figure \ref{fig:projection} is an example of projection in space of varieties
+$\Fq^d$ to $\Fq^{d-l}$. Figure \ref{fig:3Dprojection} is an example of projection in space of varieties
 from $\Fq^3$ to $\Fq^2$, corresponding to eliminating variable $x_1$ from 
@@
 
-\begin{figure}[H]
+\begin{figure}[bp]
 \centering{
 \includegraphics[width=\textwidth]{newfig/projection.pdf}
-\caption{An example of projection from $\Fq^3$ to $\Fq^2$}
-\label{fig:projection}}
+\caption{An example of projection from $\Fq^3$ to $\Fq^2$.}
+\label{fig:3Dprojection}}
 \end{figure}
@@
 \subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
-\begin{figure}[H]
-\centering{
-\includegraphics[width=\textwidth]{newfig/abs_setup.pdf}
-\caption{Galois field arithmetic circuit model for abstraction}
-\label{fig:abs_setup}}
-\end{figure}
-
 \begin{Problem}
@@
 
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/abs_setup.pdf}
+\caption{Galois field arithmetic circuit model for abstraction.}
+\label{fig:abs_setup}}
+\end{figure}
+
 The polynomial representation of $\Func$ over $\Fkk$ is:
-
 \begin{equation}
@@
 
-\begin{figure}[H]
+\begin{figure}[bp]
 \centerline{

### 2017-01-24 dissertation/submission/recha.tex
dissertation: after submission_1st revision

@@
 
-\IncMargin{1em}
+% \IncMargin{1em}
 \begin{algorithm}[hbt]
@@
 \LinesNumbered
-\Indm
+% \Indm
  \KwIn{Transition functions $\Delta$, initial state $S^0$}
-\Indp
+% \Indp
 
@@
 \end{algorithm}
-\DecMargin{1em}
+% \DecMargin{1em}
 
@@
 \begin{Example}
-
-\begin{figure}[H]
-\centering{
-\includegraphics[width=\textwidth]{newfig/MIPS.pdf}
-\caption{A 64-bit sequential multiplication hardware design.}
-\label{fig:MIPS}}
-\end{figure}
-
 Figure \ref{fig:MIPS} depicts a sequential multiplication hardware implementation within a 64-bit 
@@
 
-\IncMargin{1em}
+% \IncMargin{1em}
 \begin{algorithm}[H]
@@
 \LinesNumbered
-\Indm
+% \Indm
  \KwIn{Multiplicand $A,B$}
  \KwOut{Product $C$}
-\Indp
+% \Indp
 
@@
 \end{algorithm}
-\DecMargin{1em}
-
-\vspace{0.2in}
+% \DecMargin{1em}
+
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/MIPS.pdf}
+\caption{A 64-bit sequential multiplication hardware design.}
+\label{fig:MIPS}}
+\end{figure}
 
@@
 \begin{Example}[Lagrange's interpolation]
-
-\begin{figure}[H]
-\centering{
-\includegraphics[width=4in]{newfig/Lagrange.pdf}
-\caption{Gate-level netlist for Lagrange's interpolation example.}
-\label{fig:Lagrange}}
-\end{figure}
-
 Assume we are given gate-level netlist shown in Figure \ref{fig:Lagrange}. It can be written as 3 
@@
 
-For each element in $\F_{2^3}$, we write down the truth table as follows:
-\begin{table}[H]
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=4in]{newfig/Lagrange.pdf}
+\caption{Gate-level netlist for Lagrange's interpolation example.}
+\label{fig:Lagrange}}
+\end{figure}
+
+For each element in $\F_{2^3}$, we write down the truth table as Table \ref{tab:truthtable}.
+Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
+$Z = \Func(A)$. Recall Lagrange's interpolation formula:
+\begin{equation}
+\label{eqn:Lagrange}
+\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
+\end{equation}
+
+\begin{table}[bp]
+\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
 \label{tab:truthtable}
-\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
 \centering{\small
@@
 \end{table}
-Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
-$Z = \Func(A)$. Recall Lagrange's interpolation formula:
-\begin{equation}
-\label{eqn:Lagrange}
-\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
-\end{equation}
+
 The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ points with coordinates $(x_i,y_i)$,
@@
 $$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$
-\begin{figure}[H]
+
+In some cases, arithmetic computations are implemented as Moore machines where input operands
+are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
+We can simplify them to the model in Figure \ref{fig:seqmodel}(b).
+
+\begin{figure}[bp]
 \centering{
@@
 \end{figure}
-In some cases, arithmetic computations are implemented as Moore machines where input operands
-are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
-We can simplify them to the model in Figure \ref{fig:seqmodel}(b).
 
@@
 \LinesNumbered
-\Indm
+% \Indm
  \KwIn{The circuit's characteristic polynomial ideal $J_{ckt}$,
@@
    variables $x, s, t$ $>$ PS word $S$ $>$ NS word $T$}
-\Indp
+% \Indp
 
@@
 \LinesNumbered
-\Indm
+% \Indm
  \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\Func(S)$}
  \KwOut{Final reachable states represented by polynomial $\mathcal G(T)$}
-\Indp
+% \Indp
 
@@
 Therefore it is necessary to convert bit-level PIs to word-level PI variables.
+
+As in Figure \ref{fig:PI}, assume there exist $k$-bit datapath and $n$-bit PIs. In finite field, we need to carefully partition $n$ PIs
+such that states of each partition can be covered by a univariate polynomial respectively.
 
@@
 \end{figure}
-
-Assume we have $k$-bit datapath and $n$-bit PIs. In finite field, we need to carefully partition $n$ PIs
-such that states of each partition can be covered by a univariate polynomial respectively.
 
@@
 After the script is executed, the traversal finishes after 4 
-transition iterations, which denotes BFS depth equals to 4. The final reachable states 
+transition iterations, which denotes BFS depth equals to 4. According to the output in
+Figure \ref{fig:Recha_result}, the final reachable states 
 is a degree-9 polynomial in $T$, indicating final reachable states set contains 9 states.
@@
 
-\begin{table}[H]
+In Table \ref{tab:recha_result}, \# States denotes the final reachable states
+starting from given reset state, which given by our tool is the same with
+the return value of {\it compute\_reach} in VIS. Meanwhile, from observation
+of the experiment run-times, we find the reduction runtime increases
+as the number of gates grows. Also, iterative reachability convergence
+check's runtime reflects both the size of present state/next state
+words ($k$) and the number of final reached states, which corresponds
+to the degree of polynomial {\it reached} in Algorithm \ref{alg:univa}. 
+Although the efficiency of our initial implementation fails to compete
+with the BDD based FSM analyzer VIS, the experiment
+demonstrates the power of abstraction of algebraic geometry techniques
+for reachability analysis applications. 
+% Currently, we are investigating techniques that
+% can help us overcome the complexity of the GB computation with elimination
+% orders and speed-up our approach.
+
+\begin{table}[tbp]
 \centering
@@
 
-In Table \ref{tab:recha_result}, \# States denotes the final reachable states
-starting from given reset state, which given by our tool is the same with
-the return value of {\it compute\_reach} in VIS. Meanwhile, from observation
-of the experiment run-times, we find the reduction runtime increases
-as the number of gates grows. Also, iterative reachability convergence
-check's runtime reflects both the size of present state/next state
-words ($k$) and the number of final reached states, which corresponds
-to the degree of polynomial {\it reached} in Algorithm \ref{alg:univa}. 
-Although the efficiency of our initial implementation fails to compete
-with the BDD based FSM analyzer VIS, the experiment
-demonstrates the power of abstraction of algebraic geometry techniques
-for reachability analysis applications. 
-% Currently, we are investigating techniques that
-% can help us overcome the complexity of the GB computation with elimination
-% orders and speed-up our approach.
-
-%\vspace{-0.1in}
-
 \section{Concluding Remarks}

### 2017-01-24 dissertation/submission/unsat.tex
dissertation: after submission_1st revision

@@
 
-\vspace{-0.5in}
+\vspace{-0.7in}
 \begin{center}
@@
 mapping Boolean variables and connectors in CNF formulas to variables and operators in 
-Galois fields. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
+Galois fields as examples in Table \ref{tab:booltof4}. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
 algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
@@
 
-\begin{table}[h]
+\begin{table}[bp]
 \caption{Mapping Boolean operators to functions over $\Fkk$}
@@
 
-\IncMargin{1em}
-\begin{algorithm}[H]
+% \IncMargin{1em}
+\begin{algorithm}[htb]
 \SetAlgoNoLine
@@
 \end{algorithm}
-\DecMargin{1em}
-
-\begin{figure}[h]
-\centering{
-\includegraphics[width=5.5in]{newfig/refine.eps}}
-\caption{Abstraction by reducing latches.}
-\label{fig:refine}
-\end{figure}
-Assume that we are given a sequential circuit with $n$ latches as shown in Figure \ref{fig:refine}(a). 
+% \DecMargin{1em}
+
+Assume that we are given a sequential circuit with $n$ latches as shown in Figure \ref{fig:absrefine}(a). 
 This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
@@
 The $k$-BMC represents the initial states $I$, the transition relation $T$ and property $p$ as CNF formulas.
+
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=5.5in]{newfig/refine.eps}}
+\caption{Abstraction by reducing latches.}
+\label{fig:absrefine}
+\end{figure}
 
@@
 these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
-In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).
+In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:absrefine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).
 
@@
 the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
-when the size of datapath (number of latches in Figure \ref{fig:refine}) is very large. \textbf{Here we propose an altogether new 
+when the size of datapath (number of latches in Figure \ref{fig:absrefine}) is very large. \textbf{Here we propose an altogether new 
 method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
@@
 
-\begin{figure}[h]
+\begin{figure}[bp]
 \centering{
@@
 \SetAlgoNoLine
+\LinesNumbered
 	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
@@
 
-\begin{figure}[h]
+\begin{figure}[tbp]
 \centering{
@@
 refutation tree as shown in Figure \ref{fig:refine}(a).  
-\begin{figure}[h]
+\begin{figure}[bp]
 \centering{
@@
 \begin{Example}
-
-\begin{figure}[H]
+Figure \ref{fig:s27} shows a sequential circuit (``s27'' from the
+ISCAS benchmark set) with 3-bit state registers
+PS = $\{G7,G6,G5\}$ and NS = $\{G13,G11,G10\}$. 
+Its underlying FSM contains 8 states, as shown in Figure \ref{fig:s27_stg}
+
+\begin{figure}[tbp]
 \centering{
@@
 
-Figure \ref{fig:s27} shows a sequential circuit (``s27'' from the
-ISCAS benchmark set) with 3-bit state registers
-PS = $\{G7,G6,G5\}$ and NS = $\{G13,G11,G10\}$. 
-Its underlying FSM contains 8 states.
-
-\begin{figure}[H]
+\begin{figure}[tbp]
 \centering{
@@
 the number of state variables is reduced to 1, such that the new
-machine only contains 2 states.
-
-\begin{figure}[H]
+machine only contains 2 states, as the STG in Figure \ref{fig:s27_stg_refined}.
+
+\begin{figure}[tbp]
 \centerline{\includegraphics[width=4.5in]{newfig/s27_stg_refined.pdf}}
@@
 %\vspace{-0.1in}
-\begin{table}[htb]
+\begin{table}[tbp]
 \centering

### 2017-03-03 dissertation/submission/abstract.tex
dissertation after defense update 2
after submission: revision 2

--- 2016-12-11 dissertation/abstract.tex
@@
-\begin{center}{\bf ABSTRACT}\end{center}
+%\begin{center}{\bf ABSTRACT}\end{center}
 Formal verification of hardware designs has become an essential
@@
 size and complexity of the circuits causes the state explosion
-problem. Abstraction is key to tackle the scalability challenge.
+problem. Abstraction is the key to tackling the scalability challenges.
 
@@
 implementations, we demonstrate how to perform reachability
-analysis of finite-state machines purely at the word-level. Using this
+analysis of finite-state machines purely at the word level. Using this
 concept, we perform scalable verification of sequential arithmetic

### 2017-03-03 dissertation/submission/acknowledge.tex
after submission: revision 2

@@
 I was born in a small town in mid-inland-China. My grandpa, Shouli Sun, 
-is the first man to educate me. He was so kind and patient, and his education 
+was the first man to educate me. He was so kind and patient, and his education 
 was so successful that I could read 2,000 Chinese characters at the age of 3, 
-when my grandpa passed away. With a good start, I became used to being the 
+when my grandpa passed away. With a good start, I became used to taking the 
 first place in a variety of exams throughout elementary school, middle school, 
 and high school. Peking university seemed to be the end of my personal 
-progress in education, until I learnt that there is a way to continue my 
+progress in education, until I learnt that there is was way to continue my 
 graduate study abroad, more specifically, in the US. So I would like to 
@@
 his valuable guidance, magnificent inspirations, and of course, his generous 
-financial support that allow me to carry out my research. Meanwhile, I would 
+financial support, which allow me to carry out my research. Meanwhile, I would 
 like to thank four professors serving as my supervisory committee members, 

### 2017-03-03 dissertation/submission/appendix1.tex
after submission: revision 2

@@
 \section{Characterization of Normal Basis}
-In order to depict the characterization of NB, we need the introduce some concepts from linear algebra domain.
+In order to depict the characterization of NB, we need to introduce some concepts from the domain of linear algebra.
 \begin{Definition}[Frobenius Map]
@@
 \end{Definition}
-Finally we get to the most important concept: $T$-Order. It derives the construction of a ring extended 
+Finally we get to the most important concept: $T$-Order, which derives the construction of a ring extended 
 by a field requiring a minimal polynomial.
@@
 Using the concepts introduced above, along with other basic concepts in linear algebra and finite field theory,
-we can derive following theorems.
+we can derive the following theorems.
 
@@
 
-Let $k$ be a polynomial whose degree is larger than that of $f(x)$, {\it i.e.} $deg(k(x))>deg(f(x))$, then
+Let $k$ be a polynomial whose degree is larger than that of $f(x)$, i.e., $deg(k(x))>deg(f(x))$, then
 $$f(x)|k(x)~~\text{iff}~k(T)=0$$
@@
 This equation denotes that $\beta$ is one of $p^n$ roots of polynomial $\Func(x) = \sum_if_i(x^{p^i})$.
-However, the maximum number of roots allowed equals to the degree of $\Func(x)$, which is 
+However, the maximum number of roots allowed equals the degree of $\Func(x)$, which is 
 $p^{n-1}<p^n$, we find a contradiction. Thus we conclude that
@@
 
-From lemmas above we can deduce the following corollary:
+From the lemmas above we can deduce the following corollary:
 \begin{Corollary}
@@
 $$\forall f(x)\in \F_p[x],~deg(f(x))<n$$
-Furthermore, because there is no annihilators in the $T$-Order, this implies that
+Furthermore, because there are no annihilators in the $T$-Order, this implies that
 $$Ord_{\alpha,\sigma}(x) = x^n-1$$
@@
 Ultimately, since there always exist at least one element fulfilling requirements in Theorem \ref{thm:1}
-or Theorem \ref{thm:2}, we obtain following theorem:
+or Theorem \ref{thm:2}, we obtain the following theorem:
 \begin{Theorem}[Normal Basis Theorem over Finite Fields]
@@
 \section{Construction of General Normal Bases}
-After proving the existence of normal bases, the upcoming question is that how to find such a normal basis/element.
-For general normal basis identification, there are two methods widely used: the L\"uneburg's algorithm
+After proving the existence of normal bases, the upcoming question is how to find such a normal basis/element.
+For general normal basis identification, there are two methods widely used: L\"uneburg's algorithm
 and Lenstra's algorithm.
@@
 \subsection{L\"uneburg's Algorithm}
-L\"uneburg's algorithm can be described with following steps:
+L\"uneburg's algorithm can be described with the following steps:
 
@@
 Use Lemma \ref{lem:3}, the minimal/characteristic polynomial for $\sigma$ is $x^n-1$ implies that
-any annihilator of $\alpha^i$ ({\it i.e.} $f_i(\sigma)$) divides $x^n-1$. Meanwhile, $\{\alpha^i~|~0<i<n-1\}$ forms a (standard) basis
+any annihilator of $\alpha^i$ (i.e., $f_i(\sigma)$) divides $x^n-1$. Meanwhile, $\{\alpha^i~|~0<i<n-1\}$ forms a (standard) basis
 of field $\F_{p^n}$, this means $\alpha_i$ are linearly independent with each other ($i=0,1,\dots,n-1$).
 Since linear map $f(\sigma)$ guarantees that all elements in the field 
-have order $lcm(f_0,\dots,f_{n-1})$, {\it i.e.} $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
+have order $lcm(f_0,\dots,f_{n-1})$, i.e., $f(\sigma)\gamma = f(\sigma)(\sum_i \alpha^i) \implies f(\sigma) = lcm(f_0,\dots,f_{n-1})$.
 By contradiction we can prove that there is no factor of $x^n-1$ that can be divided by the product 
@@
 mainly because of the factorization in its second step. To overcome the high cost, an inductive method 
-is proposed, which is the Lenstra's algorithm. It allows for setup of heuristics to accelerate the procedure.
+is proposed, which is Lenstra's algorithm. It allows for setup of heuristics to accelerate the procedure.
 Before introducing the details of the algorithm, we demonstrate two preliminary lemmas.
@@
 Define $\theta$ and $g(x)$ as in Lemma \ref{lem:4}. Assume there exists a solution 
-$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Respectively there exists a non-zero element $\eta$ such that
+$\beta$ such that $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$. Respectively there exists a nonzero element $\eta$ such that
 $$g(\sigma)\eta = 0$$
@@
 
-Based on Lemmas \ref{lem:4} and \ref{lem:5}, the Lenstra's algorithm is described below:
+Based on Lemmas \ref{lem:4} and \ref{lem:5}, Lenstra's algorithm is described below:
 
@@
 \item Determine $Ord_\beta(x)$. If $deg(Ord_\beta(x)) > deg(Ord_\theta(x))$ then replace $\theta$ by $\beta$ and go to 2);
-otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a non-zero element $\eta$ such that $g(\sigma)\eta = 0$,
+otherwise if $deg(Ord_\beta(x)) \leq deg(Ord_\theta(x))$ then find a nonzero element $\eta$ such that $g(\sigma)\eta = 0$,
 replace $\theta$ by $\theta + \eta$ and determine the order of new $\theta$, and go to 2).
@@
 Directly constructing $\lambda$-Matrix is difficult for general NBs. 
-In Section \ref{sec:nbdesign}, we introduced the concept of multiplication table (M-table) and explained how its 
+In Section \ref{sec:nbdesign}, we introduced the concept of a multiplication table (M-table) and explained how its 
 entries are transformed from the $0$-th $\lambda$-Matrix. 
@@
 \end{equation*}
-Use equation $M_{i,j}^{(0)} = T_{j-i,-i}$, we can obtain the
+Using equation $M_{i,j}^{(0)} = T_{j-i,-i}$, we can obtain the
 $\lambda$-Matrix in Equation \ref{eqn:mat_R}.

### 2017-03-03 dissertation/submission/appendix2.tex
after submission: revision 2

@@
 \label{append:ONB}
-The number of non-zero entries in $\lambda$-Matrix or multiplication table (M-table) is known as \emph{Complexity $(C_N)$}.
+The number of nonzero entries in $\lambda$-Matrix or multiplication table (M-table) is known as \emph{Complexity $(C_N)$}.
 To define optimal normal basis, it is necessary to find the lower bound of $C_N$.
 \begin{Theorem}
-If $\mathcal{N}$ is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then non-zero entries in matrix $C_N\geq 2n-1$.
+If $\mathcal{N}$ is a normal basis over $\F_{p^n}$ with $\lambda$-Matrix $M^{(k)}$, then nonzero entries in matrix $C_N\geq 2n-1$.
 \end{Theorem}
@@
 So the first column always includes at least one nonzero element. 
-Since the sum of entries in each of other columns should equal to zero (modulo 2), they 
+Since the sum of entries in each of the other columns should equal zero (modulo 2), they 
 need to include an even number of ``1"s.
 Meanwhile, in order to maintain the linear independence among each row, they cannot be all ``0"s. 
-Therefore, there are always at least two non-zero elements in each column.
+Therefore, there are always at least two nonzero elements in each column.
 
@@
 
-ONBs are widely used in finite field circuit design, not only because its low complexity, 
+ONBs are widely used in finite field circuit design, not only because of its low complexity, 
 but also because of the convenience to construct the $\lambda$-Matrix and M-table for them. 
@@
 
-In following, we derive a simple way to construct the $\lambda$-Matrix of type-I ONB 
+In the following, we derive a simple way to construct the $\lambda$-Matrix of type-I ONB 
 from the criteria above. 
@@
 \end{equation}
-Suppose we only care about k=0. So simplified to following equations:
+Suppose we only care about $k=0$. So simplified to following equations:
 
@@
 
-Solution $(i,j)$ implies location of entries that equal to ``1" in $\lambda$-Matrix. 
+Solution $(i,j)$ implies location of entries that equal ``1" in $\lambda$-Matrix. 
 Let $\beta$ be an optimal normal element, 
@@
 we obtain $\lambda$-Matrix $M^{(0)}$. This construction method does not require information about 
-primitive polynomial nor normal element, so it is very convenient for circuit designers.
+primitive polynomial or normal element, so it is very convenient for circuit designers.
 Basis $\mathcal N_1$ in Example \ref{ex:app4} is a type-I ONB over $\F_{2^4}$.
@@
 structure of the circuit. However, our GB based approach requires specifying the exact normal element $\beta$,
-{\it i.e. } it is necessary to obtain $t$ as $\beta = \alpha^t$ before executing our technique.
-L\"uneburg's algorithm and Lenstra's algorithm do not guarantee the output normal element is optimal 
+i.e., it is necessary to obtain $t$ as $\beta = \alpha^t$ before executing our technique.
+L\"uneburg's algorithm and Lenstra's algorithm do not guarantee the output normal element is the optimal 
 normal element, and usually result in high computation cost. Actually if the type of ONB within the design 
 is known, instead of looking up the optimal normal element, we can construct a special irreducible $p(\alpha)$
-such that optimal normal element equals to the primitive element: $\beta=\alpha$. The concepts and algorithms 
+such that the optimal normal element is equal to the primitive element: $\beta=\alpha$. The concepts and algorithms 
 refer to IEEE standard 1363-2000 \cite{IEEE1363}.
@@
 
-For type-II ONB, the following iterative algorithm is required to generate desired 
+For type-II ONB, the following iterative algorithm is required to generate the desired 
 irreducible polynomial:

### 2017-03-03 dissertation/submission/conclude.tex
dissertation: after submission_1st revision
after submission: revision 2

@@
 
-Our approaches still have limitations: for word-level FSM traversal and 
+Our approaches still have limitations: For word-level FSM traversal and 
 sequential GF multiplier verification, our methods are more efficient
@@
 In this section, we highlight some research problems that deserve further study.
-\subsection{Multivariate Polynomial Ideal based FSM Traversal}
+\subsection{Multivariate Polynomial Ideal Based FSM Traversal}
 In Chapter \ref{ch:reacha}, we always use a single word-level variable $T$ to denote the
@@
 \end{algorithm}
-\DecMargin{1em}
+% \DecMargin{1em}
 
-The inputs of this algorithm includes the transition polynomial (result of word-level abstraction) and initial 
+The inputs of this algorithm include the transition polynomial (result of word-level abstraction) and initial 
 states description ideal, which contains 2 generators corresponding to constraints of PI and combinational input $S$. 
@@
 
-Considering the Buchberger's algorithm, all generators' leading terms are relatively prime, so $I_{tran}$ is a GB itself. 
+Considering Buchberger's algorithm, all generators' leading terms are relatively prime, so $I_{tran}$ is a GB itself. 
 Furthermore, generator $S + \mathcal G(x)$ could (possibly) be reduced by $v_x$,
-and $T+\mathcal F(S,x)$ will definitely reduced by $S + \mathcal G(x)$. In the end we obtain a polynomial
+and $T+\mathcal F(S,x)$ will definitely be reduced by $S + \mathcal G(x)$. In the end we obtain a polynomial
 $T + \mathcal F'(x)$ in the reduced GB. We can then include this polynomial and $v_x$ and exclude the polynomial containing
-$S$ ({\it i.e.} $\mathcal H(S)$ in Algorithm \ref{alg:multi}), to compose an ideal representing \emph{next states} $to^i$. 
+$S$ ({\it i.e.}, $\mathcal H(S)$ in Algorithm \ref{alg:multi}), to compose an ideal representing \emph{next states} $to^i$. 
 In iteration 1, the result is $to^1 = \langle T+1, x^2+x\rangle$
@@
 Line 5 is the ideal quotient of universal set and reached states. In the first iteration, 
-$reached$ is the initial state $\langle T+1+\alpha, x^2+x \rangle$. Result of ideal quotient
+The initial state $reached$ is $\langle T+1+\alpha, x^2+x \rangle$. The result of ideal quotient
 is $\langle T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle$ represents $\{00,01,10\}$.
 
-Line 6 is ideals' sum (intersection of their varieties), it is done by combining all generators
-from 2 ideals and compute GB. In the first iteration, the result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$ representing $\{01\}$.
+Line 6 is ideals' sum (intersection of their varieties) and is done by combining all generators
+from 2 ideals and computing GB. In the first iteration, the result is $GB(\langle T+1,T^3+(1+\alpha)T^2+\alpha T, x^2+x\rangle) = \langle T+1, x^2+x\rangle$,
+representing $\{01\}$.
 
-Line 7 is ideals' product (union of their varieties), it is done by multiplying all pairs of
-generators from both ideal. For the first iteration it gives the result as $GB(\langle (T+1)(T+1+\alpha),
-(T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$ representing $\{01,11\}$.
+Line 7 is ideals' product (union of their varieties) and is done by multiplying all pairs of
+generators from both ideals. For the first iteration it gives the result as $GB(\langle (T+1)(T+1+\alpha),
+(T+1)(x^2+x), (T+1+\alpha)(x^2+x), (x^4+x)\rangle) = \langle T^2+\alpha T+(1+\alpha), x^2+x\rangle$, representing $\{01,11\}$.
 
-The traversal algorithm executes 3 iterations and terminate at 4th iteration. We list all intermediate 
+The traversal algorithm executes 3 iterations and terminates at the 4th iteration. We list all intermediate 
 results below:
@@
 still incurs exponential complexity in the worst case. In a situation where division is to be performed modulo
-a chain of OR gates, the size of remainder polynomial will explode. For example, a chain of OR gates 
+a chain of OR gates, the size of the remainder polynomial will explode. For example, a chain of OR gates 
 can be written as the Boolean function 
@@
 \end{equation}
-which equals to the following polynomial function in $\F_2$:
+which equals the following polynomial function in $\F_2$:
 $$f = abcd+abc+abd+ab+acd+ac+ad+a+bcd+bc+bd+b+cd+c+d$$
@@
 
-Another way is to utilize DDs, {\it e.g.} ZDDs. ZDDs can represent unate covers of Boolean formulas.
+Another way is to utilize DDs, e.g., ZDDs. ZDDs can represent unate covers of Boolean formulas.
 We can represent $f$ using a ZDD, where every path in the ZDD from a root to a terminal corresponds
@@
 
-\begin{figure}[h]
+Figure \ref{fig:ZDD} shows a ZDD representing 
+Equation \ref{eqn:chainOR} with size $2\times4 -1 = 7$, which is linear in the number of variables.
+The reduction process using ZDDs can be executed as in \cite{PolyBoRi}. The graphical illustration 
+of the remainder in ZDDs is also shown in Figure \ref{fig:ZDD}.
+
+\begin{figure}[bp]
 \centerline{
@@
 
-Figure \ref{fig:ZDD} shows a ZDD representing 
-Equation \ref{eqn:chainOR} with size $2\times4 -1 = 7$, which is linear in the number of variables.
-The reduction process using ZDDs can be executed as in \cite{PolyBoRi}. The graphical illustration 
-of the remainder in ZDDs is also shown in Figure \ref{fig:ZDD}.
-
 \subsection{Craig Interpolants in Algebraic Geometry}
@@
 symbolic logic \cite{craig-interpolate}; later, algorithms were
-presented to find the CI for Boolean formulae
-\cite{pudlak:ci} \cite{mcmillan2003interpolation}. Assume that Boolean
-formulae are represented in Clause Normal Form (CNF) as 
+presented to find the CI for Boolean formulas
+\cite{pudlak:ci, mcmillan2003interpolation}. Assume that Boolean
+formulas are represented in Clause Normal Form (CNF) as 
 $f = C_1 \wedge C_2 \wedge \dots \wedge C_m$ where: 1) Each clause
@@
 
-CIs are used to derive abstractions to produce over-approximate image
+CIs are used to derive abstractions to produce overapproximate image
 operators in model checking \cite{mcmillan2003interpolation}. Since $A
@@
 Existence of algebraic interpolant comes from \cite{craig-interpolate}.
-The question for exploring is that how do we compute it in algebraic geometry?
+The question for exploring is, how do we compute it in algebraic geometry?
 
@@
 \bi
-\item Find the minimal interpolant: {\it i.e.} find the interpolant $F_P$
+\item Find the minimal interpolant, i.e., find the interpolant $F_P$
   such that $V(J_P)$ is the smallest variety larger than $V(J_A)$ such
@@
 It is possible to use the equivalence checking techniques proposed in this dissertation 
-as a alternative way to perform technology mapping, especially on word-level when 
+as an alternative way to perform technology mapping, especially at word level when 
 given blocks represent word-level functions.
@@
 design $D$. The objective is to generate a mapped design $N'$ with as many of $B_i$ such that the area
-and timing is optimal. The procedure is also illustrated in Figure \ref{fig:macro}.
+and timing are optimal. The procedure is also illustrated in Figure \ref{fig:macro}.
 
-\begin{figure}[h]
+\begin{figure}[bp]
 	\begin{center}
@@
 
-The following part describes the sketch of our proposed approach base on loop invariant constraints \cite{sankaranarayanan2004non}.
-First, a transition system is modeled by algebraic assertions; then \emph{ideal membership test} \cite{lv:phd}
+The following part describes the sketch of our proposed approach based on loop-invariant constraints \cite{sankaranarayanan2004non}.
+First, a transition system is modeled by algebraic assertions, then the \emph{ideal membership test} \cite{lv:phd}
 is applied on the set of assertions to help abstract the loop invariant. We introduce concept \emph{template}
@@
 
-1) {\bf Template and state constraints:}  Ideal membership test requires a setup of ideal $J_{loop}$
+1) {\bf Template and state constraints:}  The ideal membership test requires a setup of ideal $J_{loop}$
 for loop invariants. 
-One heuristic that can provide better coverage for loop invariant abstraction as well as a
+One heuristic that can provide better coverage for loop-invariant abstraction as well as a
 relatively small size is \emph{generic quadratic form} (GQF).
@@
 
-In the next step, we reduce the template with the Gr\"obner basis $G$: $\Func \xrightarrow{G}_+ r$, the remainder equals to
+In the next step, we reduce the template with the Gr\"obner basis $G$: $\Func \xrightarrow{G}_+ r$. The remainder equals to
 $$r = a_5i^2+a_6i+(a_7+a_8)ij_0+(a_9+a_{11}+a_{12})j_0^2+(a_{10}+a_{13})j_0+a_{14}$$
-Let it equal to 0, then each coefficient will generate a constraint. The solution to the system forms the candidate
+Let it equal 0, then each coefficient will generate a constraint. The solution to the system forms the candidate
 assignment to generate loop invariants.
@@
 Our proposed approach models the 2 states individually,
-{\it i.e.} performs polynomial reduction separately and obtains remainder $r_1$ and $r_2$. 
+i.e., performs polynomial reduction separately and obtains remainders $r_1$ and $r_2$. 
 Assume a constraint polynomial describing
@@
 Take Equation \ref{eqn:loopinv} as the example (which refers to Example 10 in \cite{sankaranarayanan2004non}). 
-PS is initial state we just characterized 
+PS is the initial state we just characterized 
 $$\Func = f(s,i,j,j_0)$$ 
@@
 by using equality $r_2 = \lambda r_1$ we get constraints for the parameters. We solve the system using 
-similar algorithm to solve binate covering problems
+similar algorithm to solve binate-covering problems,
 which is also described in Section 4 as \emph{elimination by splitting} technique in \cite{sankaranarayanan2004non}. 
-One of the branching results is:
+One of the branching results is
 \begin{align*}
@@
 \end{align*}
-The reduced remainder is: 
+The reduced remainder is
 $$r_2 = a_1s +(a_1-a_7)i+a_7ij+a_8ij_0 = a_1(s+ij-ij_0) = 0$$
-Thus the invariant of the program in Equation \ref{eqn:loopinv} is:
+Thus the invariant of the program in Equation \ref{eqn:loopinv} is
 $$s = i(j_0- j)$$
@@
 $$Z + f(i_1,i_2,\dots,i_n)$$
-where $Z$ is the output. When there is only one output, $Z$ collapses to a bit-level variable. However in most
+where $Z$ is the output. When there is only one output, $Z$ collapses to a bit-level variable. However, in most
 cases there are multiple outputs on the cut, indicating $Z$ as a word-level variable. Boolean function $\Func$ covers
@@
 
-\begin{figure}[h]
+\begin{figure}[tbp]
 	\begin{center}
@@
 $$\textit{Other\ circuit\ variables} > \textit{output\ word\ } Z > \textit{all\ cut-set\ bit\ level\ inputs}$$
-on the ideal describing the partitioned circuit as well as ideal with vanishing polynomials,
+on the ideal describing the partitioned circuit as well as the ideal with vanishing polynomials,
 the reduced Gr\"obner basis has a single polynomial generator in the form of $Z + \Func(s_0,s_1,s_2,s_3)$.
@@
 2) {\bf Templates on Boundary Information:}
-\begin{figure}[h]
+\begin{figure}[bp]
 	\begin{center}
@@
 lines show one possible mapping, to find out this kind of feasible mapping, we need to simulate
-all possible mappings, where the concept "template" can be used.
+all possible mappings where the concept "template" can be used.
 
@@
 
-Above polynomial includes all possibilities where $z_0$ could be mapped to.
-When $c_{n_0} = 0$, then $\{c_{t_0z_0},c_{t_1z_0},c_{z_0CO}\}$ equals to 1 denotes a valid mapping to corresponding output pin. 
+The above polynomial includes all possibilities where $z_0$ could be mapped to.
+When $c_{n_0} = 0$, then $\{c_{t_0z_0},c_{t_1z_0},c_{z_0CO}\}$ equals 1 denotes a valid mapping to a corresponding output pin. 
 Conversely if $c_{n_0} = 1$, the ``1" evaluation denotes mapping to negation of corresponding output pin.
@@
 \item Second, propagate forward the cut for a certain number of gates (depth), choose a cut-set
-which fully dependents on these inputs as output pins. Thus we obtain a sub-circuit by partitioning;
+which fully depends on these inputs as output pins. Thus we obtain a subcircuit by partitioning;
 
-\item Third, abstract a description polynomial $\Func$ of the target sub-circuit using GB based abstraction;
+\item Third, abstract a description polynomial $\Func$ of the target subcircuit using GB based abstraction;
 

### 2017-03-03 dissertation/submission/intro.tex
after submission: revision 2

@@
 \section{Hardware Design and Verification Overview}
-During the past decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
+During the past decades, the level of integration in modern VLSI systems is becoming higher and higher because of Moore's law.
 As a result, an entire system with billions of transistors can be built upon a single chip.
 The design process also evolves from manual design with little validation, to 
-a formal 3-step procedure which requires collaboration of teams with large number of 
+a formal 3-step procedure which requires collaboration of teams with large numbers of 
 engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
@@
 
-Nowadays the verification step is usually completed by a team that specializes on test, verification and validation of 
+Nowadays the verification step is usually completed by a team that specializes in the testing, verification and validation of 
 circuits. This step is also automated as an indispensable part of the CAD flow, when circuit synthesis is performed. 
 Figure \ref{fig:designflow} shows the typical synthesis flow, which covers procedures starting from the 
-register-transfer-level (RTL) description (using hardware design languages, {\it i.e.} HDL) to  the 
+register-transfer-level (RTL) description (using hardware design languages, i.e., HDL) to  the 
 physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
@@
 
-One way to perform verification is by {\it simulation}. It is the collage of all circuit validation 
+One way to perform verification is by simulation, which is the collage of all circuit validation 
 methods which apply stimulus on the inputs of circuit model and verify correctness of the outputs.
@@
 is a good example where simulation failed. Failure occurred with input assignments which were rarely used in most divisions. Because of the limitation of simulation,  
-test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollars recalling bill
-for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.
+test engineers from Intel failed to detect the bug,  which brought a roughly $\$475$ million-dollar recall bill
+for the company. Thus,  new methods that can guarantee the correctness of the design need to be explored.
 
-Another method developed besides simulation is \emph{formal verification}, it utilizes 
+Another method developed besides simulation is \emph{formal verification}, which utilizes 
 mathematical theory to reason about the correctness of hardware designs.
@@
 On the other hand,  it formalizes properties
-for the circuit model which are relevant only to specific input signals,  and prove the properties mathematically. 
-These descriptions are named as {\it specifications}.
+for the circuit model which are relevant only to specific input signals and prove the properties mathematically. 
+These descriptions are called {\it specifications}.
 Formal verification has two main forms: property checking and equivalence 
@@
 the specification as well as the circuit implementation
-is described as theorems in mathematical logic. Subsequently logic rules
+are described as theorems in mathematical logic. Subsequently, logic rules
 are employed to deduce new objective theorems. In practice, the tool can reduce
-a proof goal to simpler sub-goals for automatic verification.
+a proof goal to simpler subgoals for automatic verification.
 
@@
 properties are violated. If violation occurs, 
-a counter-example is then generated as a transition firing path that corresponds to the
+a counterexample is then generated as a transition firing path that corresponds to the
 false behavior in the design. 
@@
 equivalence checking techniques: graph-based,
-satisfiability-based (SAT-based) and induction-based.
+satisfiability-based (SAT-based), and induction-based.
 \begin{enumerate}[{1)}]
@@
 Many variants branch out from BDDs; some widely used variants include 
-ZDD \cite{minato1993zero}, BMD \cite{bmd}, FDD \cite{okfdd}, {\it etc.} The comparison algorithms can 
+ZDD \cite{minato1993zero}, BMD \cite{bmd}, FDD \cite{okfdd}, etc. The comparison algorithms can 
 determine whether the two graphs are isomorphic. The canonicity of the graph representation guarantees that
@@
 gives output ``1" only when the outputs of the circuits differ with 
-the same inputs, {\it e.g. }inputs $a,b,c$ shown in Figure \ref{fig:miter}. 
+the same inputs, e.g., inputs $a,b,c$ shown in Figure \ref{fig:miter}. 
 A SAT solver \cite{csat,mishchenko2006improvements} is then employed to simplify the problem 
@@
 miter output is ``1". If such an assignment exists, this solution acts as a 
-counter-example to equivalence; otherwise the circuits
+counterexample to equivalence; otherwise the circuits
 are functionally equivalent.
@@
 (SEC) problem. A miter model can also be built with two sequential circuits. Through the 
-miter model, SEC problem is transformed into a sequential backward justification problem.
+miter model, an SEC problem is transformed into a sequential backward justification problem.
 Equivalence of states and transitions between states can be proved using induction-based 
@@
 Word-level abstractions have a wide variety of applications in formal 
-verification. For example, it can work as automatic decision and canonical reduction engine in theorem proving;
+verification. For example, they can work as automatic decision and canonical reduction engines in theorem proving;
 for RTL composed of macro blocks, abstractions of these blocks also benefit RTL verification.
@@
 \item Model checking with abstractions \cite{kroening:model}, 
-where an over-approximation of RTL blocks is abstracted and used for property checking on a simplified model.
+where an overapproximation of RTL blocks is abstracted and used for property checking on a simplified model.
 \item Graph-based equivalence 
@@
 abstractions are used to analyze structural symmetries and similarities such that
-the Boolean formulas fed to SAT solver is simplified.
+the Boolean formulas fed to the SAT solver are simplified.
 \end{itemize}
@@
 Other equivalence checking techniques that employ abstractions 
-include {\it satisfiability modulo theory} (SMT) solvers \cite{boolector,bryant:tacas07}.
+include {\it satisfiability modulo theory} (SMT) solvers \cite{boolector,bryant:tacas07},
 as well as constraint programming (CP) techniques \cite{ms:research,tew:iccad08}.
@@
 datapaths. Moreover, 
-word-level abstractions facilitates the use of uninterpreted functions (UFs) \cite{UF3}, which 
+word-level abstractions facilitate the use of uninterpreted functions (UFs) \cite{UF3}, which 
 can be transformed into proposition formulas with word-level information and verified using 
@@
 errors are usually modeled as ``bad" states, and the
-circuits/functional components are modeled as finite state machines
+circuits/functional components are modeled as finite-state machines
 (FSMs). Once state reachability is analyzed, the existence of errors
@@
 model checking, optimized implementations of sequential circuits are
-synthesized. A subsequent problem then needs to be solved to prove
+synthesized. Then there is a subsequent problem that needs to be solved: how to prove
 that the sequential circuit implementations are equivalent to the
-original specification models, {\it i.e.} {\it Sequential Equivalence
-  Checking }(SEC). When the specification is given as an arithmetic function
+original specification models? This problem is called {\it Sequential Equivalence
+  Checking }(SEC) problem. When the specification is given as an arithmetic function
 which canonically represents the circuit, then the problem 
@@
 fewer steps.  
-\item Analyze over-approximations (or {\it abstractions})
+\item Analyze overapproximations (or {\it abstractions})
 of the state-space. Abstraction proves properties on the system by
-first simplifying it, and when the abstraction does not satisfy the
+first simplifying them, and when the abstraction does not satisfy the
 same properties as the original one, a process of refinement is
@@
 unsatisfiability (UNSAT cores) to refine the abstractions.
-\item The recent breakthrough method of \cite{bradley2011sat} where the set of
-over-approximations to forward reachable states are refined with
+\item The recent breakthrough method of \cite{bradley2011sat}, where the set of
+overapproximations to forward reachable states are refined with
 inductive constraints -- property directed reachability (PDR). 
@@
 the designs are expressed at the level of bit-vector words
-({\it e.g.} Matlab code, Verilog), and these word-level abstractions are
+(e.g., Matlab code, Verilog), and these word-level abstractions are
 rarely exploited in verification. The problem is further exacerbated
@@
 predicate abstraction \cite{jain2005word,mcmillan:cav06,mcmillan2010lazy}, 
-{\it using a purely word-level representation
+using a purely word-level representation
   of the state-space, the properties and their abstractions have not
   been fully explored as another dimension in improving sequential
-  verification.}
+  verification.
 
@@
 reachability analysis and abstraction} for sequential design verification techniques. 
-Our approaches operate at the word-level and are based
-largely on the concepts from {\it algebraic geometry}.
+Our approaches operate at the word level and are based
+largely on concepts from {\it algebraic geometry}.
 
@@
 relates to that of performing {\it FSM traversal} \cite{touati1990implicit}
-but  at the word-level. Analogously, in the case of MC, the problem
+but  at the word level. Analogously, in the case of MC, the problem
 is setup {\it w.r.t.} a FSM $\mathcal{M}$, a set of initial states $S_0$ and
 a set of property states $p$. Techniques are to be researched that
-verify that there exist no sequence of transitions from an initial
-state to a non-property state (``bad'' state). These problems have to
-be solved in the context of word-level verification -- {\it i.e.} data
-representation, abstraction using UNSAT cores and
-algorithm execution has to be carried out at word-level.
+verify that there exists no sequence of transitions from an initial
+state to a nonproperty state (``bad'' state). These problems have to
+be solved in the context of word-level verification -- i.e., data
+representation, abstraction using UNSAT cores, and
+algorithm execution have to be carried out at word level.
 
 \subsection{Word-level Reachability Analysis of FSMs}
-In this dissertation, we propose a method to perform reachability analysis at the word-level. 
+In this dissertation, we propose a method to perform reachability analysis at the word level. 
 	The given FSM is modeled as a system of polynomials over a finite field,
@@
 the FSMs -- the transition relations -- by means of a set of
-multi-variate polynomials with coefficients from the finite (Galois)
-field $\Fkk$ of $2^k$ elements, {\it i.e.} polynomials in the ring
+multivariate polynomials with coefficients from the finite (Galois)
+field $\Fkk$ of $2^k$ elements, i.e., polynomials in the ring
 $\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
@@
 $f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
-$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
+$f: \Fkk \rightarrow \Fkk$. It is well known that over the finite
 field ($\Fq$) of  $q$ elements, every function $f: \Fq
@@
 the theory and technology of {\it \Grobner bases}. GB-based
-algebraic reasoning is very powerful; in fact it is known to be
+algebraic reasoning is very powerful -- in fact it is known to be
 strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
@@
 Sequential Galois field (GF) arithmetic circuits can be modeled as a special type of 
-FSM, where the pre-loaded operands are mapped to initial states, and pseudo outputs
+FSM, where the preloaded operands are mapped to initial states, and pseudo outputs
 after $k$ clock-cycles are recognized as the final reached state after $k$ transitions.
@@
 The rest of the dissertation is organized as
-follows: Chapter \ref{ch:prev} reviews previous work, and analyzes their drawbacks with respect to 
+follows: Chapter \ref{ch:prev} reviews previous works and analyzes their drawbacks with respect to 
 the word-level sequential verification problem.
@@
 continuation of this work and concludes the dissertation. 
-An appendix provides theory and methodology on the characterization of finite field normal basis,
+An appendix provides theory and methodology on the characterization of finite-field normal basis,
 as well as the construction of optimal normal basis and application to normal basis multiplier design.

### 2017-03-03 dissertation/submission/normal.tex
after submission: revision 2

@@
-\chapter{Functional Verification of Sequential Normal Basis Multipliers}
+\chapter{Functional Verification of Sequential Normal Basis\protect\\ Multipliers}
 \label{ch:normal}
 With the size of sequential arithmetic circuits increasing rapidly, 
-conventional methods become inefficient on verifying the correctness of these circuits.
-In Chapter \ref{ch:reacha} we proposed a FSM traversal approach at word-level 
+conventional methods become inefficient for verifying the correctness of these circuits.
+In Chapter \ref{ch:reacha} we proposed a FSM traversal approach at word level 
 based on abstractions. Since the sequential arithmetic circuits can be modeled 
@@
 We show how this approach can be applied to verify sequential GF multipliers
-which we can apply our traversal algorithm to actually 
+which we can apply to our traversal algorithm to actually 
 verify its functional correctness.
@@
 
-\begin{figure}[bp]
-\centering{
-\includegraphics[width=\textwidth]{newfig/Moore.eps}
-\caption{A typical Moore machine and its state transition graph.}
-\label{fig:Moore}}
-\end{figure}
-
 In practice,
@@
 An example of such designs is the sequential GF arithmetic multipliers described in Chapter \ref{ch:prelim_GF}.
-However, it also brings difficulties in verifying the the circuit functions.
+However, it also brings difficulties in verifying the circuit functions.
 
-\begin{figure}[bp]
+\begin{figure}[tbp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/Moore.eps}
+\caption{A typical Moore machine and its state transition graph.}
+\label{fig:Moore}}
+\end{figure}
+
+Conventional methods to such a sequential circuit may consist of unrolling the circuit for 
+$k$ time-frames, and performing an equivalence checking between the unrolled machine and
+the specification function, as shown in Figure \ref{fig:convention}. However, the number of gates will grow fast when unrolling
+at bit level. Meanwhile the structural similarity-based equivalence checking techniques 
+will fail when the sequential circuit is highly customized and optimized from the naive specification 
+function. As a result, conventional techniques are grossly inefficient for large circuits.
+Therefore, a new method based on our proposed word-level FSM traversal technique is worthy of exploration.
+
+\begin{figure}[tbp]
 \centering{
@@
 
-Conventional methods to such a sequential circuit may consist of unrolling the circuit for 
-$k$ time-frames, and performing an equivalence checking between the unrolled machine and
-the specification function. However, the number of gates will grow fast when unrolling
-on bit-level. Meanwhile the structural similarity based equivalence checking techniques 
-will fail when the sequential circuit is highly customized and optimized from the naive specification 
-function. As a result, conventional techniques are grossly inefficient for large circuits.
-Therefore, a new method based on our proposed word-level FSM traversal technique is worthy of exploration.
-
-\section{Formal Verification of Normal Basis Multipliers using Gr\"obner Bases}
+\section{Formal Verification of Normal Basis Multipliers Using Gr\"obner Bases}
 The gate-level design of a NB multiplier can be generated using approaches introduced in 
@@
 an approach similar to that in Chapter \ref{ch:reacha}. 
-First we introduce the sketch of our approach using abstraction term order (ATO)
+First we introduce the sketch of our approach using abstraction term order (ATO),
 mentioned previously in Section \ref{sec:abstraction}, then refine our approach using 
-the concept of ``RATO", which is previously used in Section \ref{sec:improve}.
+the concept of ``RATO", which was previously used in Section \ref{sec:improve}.
 % The following part borrows contents
 % from my own conference paper \cite{myDATE}.
-\subsection{Implicit Unrolling based on Abstraction with ATO}
+\subsection{Implicit Unrolling Based on Abstraction with ATO}
 The concept of abstraction was discussed in Section \ref{sec:abstraction}. If we use an elimination term order with
@@
 $$R = \Func(A,B)$$
-To verify the functional correctness of a combinational NB multiplier ({\it e.g.} Mastrovito multiplier
+To verify the functional correctness of a combinational NB multiplier (e.g., Mastrovito multiplier
 or Montgomery multiplier), the function given by abstraction will be computed as
@@
 In other words, with the assistance of abstraction, we can execute implicit unrolling instead of 
-explicit unrolling and avoid bit-blasting problem.
+explicit unrolling and avoid a bit-blasting problem.
 
-\begin{figure}[bp]
+\begin{figure}[tbp]
 \centering{
@@
 For 2-input sequential NB multipliers, abstraction is utilized to implement Algorithm \ref{alg:modified}.
-Notations in Figure \ref{fig:sequential} as well as in next example comply with 
+Notations in Figure \ref{fig:sequential} as well as in the next example comply with 
 the notations in this algorithm.
-
-\begin{figure}[bp]
-\centering{
-\includegraphics[width=5in]{newfig/gf_seq_model.eps}
-\caption{A typical normal basis GF sequential circuit model. $A =
-  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
-  $A', B', R'$ denote next-state inputs.}
-\label{fig:sequential}}
-\end{figure}
 
@@
 
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=4.5in]{newfig/gf_seq_model.eps}
+\caption{A typical normal basis GF sequential circuit model. $A =
+  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
+  $A', B', R'$ denote next-state inputs.}
+\label{fig:sequential}}
+\end{figure}
+
 Each gate in the combinational logic is represented by a Boolean
@@
   of the circuit. The ideal of vanishing polynomials 
-$J_0 = \langle x_i^2-x_i,R_j^{2^k}-R_j\rangle,~\forall \text{ bit-level variables }x_i\text{ and word-level variables }R_j$, is also included. 
-Then the implicit FSM unrolling problem is setup for abstraction.
+$J_0 = \langle x_i^2-x_i,R_j^{2^k}-R_j\rangle,~\forall \text{ bit-level variables }x_i\text{, and word-level variables }R_j$, is also included. 
+Then the implicit FSM unrolling problem is set up for abstraction.
 
@@
 $>$ {\it word-level NS outputs} $>$ {\it word-level PS inputs}. This
-allows to eliminate all the bit-level variables 
+allows us to eliminate all the bit-level variables 
 %(corresponding to the combinational logic and the state variables),
@@
 logic actually implements function not only related to current operands $A$ and $B$,
-but also involved with PS variable ({\it i.e.} temporary product) $R$. Which can be 
+but also involved with PS variable (i.e., temporary product) $R$, which can be 
 obtained using the abstraction:
@@
 abstraction approach. Following ideal $J_{ckt}$ from line 5 in Algorithm \ref{alg:modified} is the ideal for 
-all gates in combinational logic block and definition of word-level variables.
+all gates in a combinational logic block and definition of word-level variables.
 $$J_{ckt} = \langle f_1,f_2,\dots,f_{19}\rangle$$
@@
 $$R$$
-The other two polynomials describes PS variable $A,B$ -- current multiplication operands, and we write them 
-in parameters $A_{init}, B_{init}$. Since this is the first iteration, they have the same form with $f_{16}$ and $f_{17}$:
+The other two polynomials describe PS variable $A,B$ -- current multiplication operands, and we write them 
+in parameters $A_{init}, B_{init}$. Since this is the first iteration, they have the same form as $f_{16}$ and $f_{17}$:
 \begin{align*}
@@
 
-\subsection{Overcome Computational Complexity using RATO}
+\subsection{Overcome Computational Complexity Using RATO}
 We implemented Algorithm \ref{alg:modified} in \textsc{Singular}, with similar setup and environments as
 with the experiments in Section \ref{sec:exp_reacha}. The result of verifying Agnew's SMPO is shown 
-in Table \ref{tab:slimgb}.
+in Table \ref{tab:slimgb}. It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 11 bits operands.
+Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} can also be applied to 
+accelerate the GB computation here. More specifically, we circumvent the GB computation 
+of Line 5 in Algorithm \ref{alg:modified} by turning the Buchberger's algorithm 
+into a single-step of multivariate polynomial division.
 
@@
 
-It indicates that our approach based on ATO cannot verify sequential GF multiplier with size larger than 11 bits operands.
-Similar to our improvements in Section \ref{sec:improve}, RATO \cite{TimDAC} can also be applied to 
-accelerate the GB computation here. More specifically, we circumvent the GB computation 
-of Line 5 in Algorithm \ref{alg:modified} by turning the Buchberger's algorithm 
-into a single-step of multivariate polynomial division.
-
 \begin{itemize}
@@
 \item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using a reverse topological order
-derived from the circuit structure, followed by word-level NS outputs ({\it e.g.} $R'$ in Figure \ref{fig:sequential}), 
+derived from the circuit structure, followed by word-level NS outputs (e.g., $R'$ in Figure \ref{fig:sequential}), 
 then word-level PS inputs ({\it e.g.} $A,B,R$ in Figure \ref{fig:sequential}).
@@
 \end{align}
-The variables in braces are at the same topological level, they are arranged in LEX order. 
+The variables in braces are at the same topological level and are arranged in LEX order. 
 We search among all generators of $J_{ckt}$ from Example \ref{ex:RHSMPO} using RATO, 
@@
 \end{Example}
-The above example indicates that RATO based abstraction on 5-bit RH-SMPO will result in a remainder containing both
+The above example indicates that RATO-based abstraction on 5-bit RH-SMPO will result in a remainder containing both
 bit-level variables and word-level variables. Moreover, the number of variables is still large such that
@@
 this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
-variables, {\it e.g.} $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
+variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
 approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.
@@
 % \end{align}
-to recursively eliminate $a_1$ from third row, $a_2$ from fourth row, etc. The final solution to this system of
+to recursively eliminate $a_1$ from the third row, $a_2$ from the fourth row, etc. The final solution to this system of
 equations is
@@
 furthermore, there will be one and
-only one term containing word-level output with degree equals to 1.
+only one term containing word-level output with degree equal to 1.
 \end{Proposition}
@@
 \begin{Proof}
-First proposition is easy to prove by contradiction: assume there exists an intermediate bit-level variable $v$ in the remainder,
+The first proposition is easy to prove by contradiction: Assume there exists an intermediate bit-level variable $v$ in the remainder,
 then this remainder must be divided further by a polynomial with leading term $v$. 
@@
 $R$ in $Spoly$. Meanwhile in other polynomials from $J_{ckt}+J_0$ there is no such term containing $R$, so this term 
-will appear in remainder $r$, with exponent equals to 1.
+will appear in remainder $r$, with exponent equal to 1.
 \end{Proof}
 
-By replacing all bit-level variables by corresponding word-level variable polynomials, we transform the remainder
-of $Spoly$ reduction to the form of $R'+R+\mathcal F'(A,B)$. Note $R$ is present state notion of output, which
-equals to initial value $R=0$ in first clock cycle, or value of $R'$ from last clock cycle. By substituting
+By replacing all bit-level variables with corresponding word-level variable polynomials, we transform the remainder
+of $Spoly$ reduction to the form of $R'+R+\mathcal F'(A,B)$. Note $R$ is the present state notion of output, which
+equals to initial value $R=0$ in first clock cycle, or value of $R'$ from the last clock cycle. By substituting
 $R$ with its corresponding value ($0$ or a polynomial only about $A$ and $B$), we get the desired polynomial function
@@
 In Example \ref{ex:BLVS} we show a Gaussian-elimination method to solve the system of polynomial
-equations. We describe another formal method to solve following system of equations
+equations. We describe another formal method to solve the following system of equations
 \begin{align}
@@
 \end{Definition}
-Moore determinant can be written as an explicit expression
+The Moore determinant can be written as an explicit expression
 \begin{equation}
@@
 \end{equation}
-We use an example to help understanding the notations in Equation \ref{eqn:Moore}:
+We use an example to help understand the notations in Equation \ref{eqn:Moore}:
 \begin{Example}
@@
 \end{align}
-First, let $i=1$,  we obtain $c_0\in \mathbb F_2$. When $c_0=0$, the product term equals to $\alpha_1$; when $c_0=1$ it equals to $(\alpha_1-\alpha_0)$.
+First, let $i=1$,  we obtain $c_0\in \mathbb F_2$. When $c_0=0$, the product term equals $\alpha_1$; when $c_0=1$ it equals $(\alpha_1-\alpha_0)$.
 Then let $i=2$, we obtain ${c_0,c_1}\in \mathbb F_2$, they can take value from $\{0,0\},\{0,1\},\{1,0\}$ and $\{1,1\}$.
@@
 is a (normal) basis, then all product terms are distinct and represents all elements in the field $\Fpk$. Since the product 
-of all elements of a field equals to 1, the Moore determinant $\M=1$.
+of all elements of a field equals 1, the Moore determinant $\M=1$.
 \end{Proof}
@@
 Based on the above concepts and improvements, the functional verification for sequential GF multiplier 
-on word-level with $k$-bit input operands $A,B$ and $k$-bit output $R$ is described as follows:
+on word level with $k$-bit input operands $A,B$ and $k$-bit output $R$ is described as follows:
 \begin{enumerate}[{1)}]
@@
 loop of our script file performs the same function as Algorithm \ref{alg:modified} describes, while the Gr\"obner basis
-computation in main loop is divided into 4 different functional parts:
+computation in the main loop is divided into 4 different functional parts:
 
-1) Pre-processing:
+1) Preprocessing:
 
-This step is executed only once before the main loop starts. The function of pre-processing is to compute following system
+This step is executed only once before the main loop starts. The function of preprocessing is to compute the following system
 of equations for bit-level inputs $a_0 \dots a_{k-1}$:
@@
 62 bits, we rewrite the degree using smaller integers (the feasibility of this rewriting can also be verified 
-in following steps). Our C++ tool does not have data size limit, but it accumulates exponent when computing 
+in the following steps). Our C++ tool does not have a data size limit, but it accumulates exponents when computing 
 the power of a variable. Thus if the exponent is as large as $2^{63}$, its runtime is very long. As a 
@@
 Since the bit-to-word substitution procedure
-only requires squaring of equations each time, the exponent of word $A$ can only be in the form $2^{i-1}$, {\it i.e.} 
+only requires squaring of equations each time, the exponent of word $A$ can only be in the form $2^{i-1}$, i.e., 
 $A^{2^0},A^{2^1},\dots,A^{2^{k-1}}$. To minimize the exponents presenting in both \textsc{Singular} and our tool, 
-we rewrite exponent $2^{i-1}$ to exponent $i$, {\it i.e.}
+we rewrite exponent $2^{i-1}$ to exponent $i$, i.e.,
 $(A^{2^0},A^{2^1},\dots,A^{2^{k-1}}) \to (A, A^2, \dots, A^k)$. In this way result is rewritten to be
@@
 This step requires limited substitution operations, so although we use the naive Gaussian elimination
-method (whose time complexity is $O(k^3)$), the time cost is trivial comparing to following steps. 
+method (whose time complexity is $O(k^3)$), the time cost is trivial comparing to the following steps. 
 % For 33 bits experiment,
@@
 
-For SMPO experiments, reduced $Spoly$ has following generic form (all coefficients are omitted):
+For SMPO experiments, reduced $Spoly$ has the following generic form (all coefficients are omitted):
 \begin{equation}
@@
 For our exponent rewriting approach to give a faithful result, it has to satisfy an important condition:
-there exists no cross-product terms simultaneously containing elements within the same set, {\it i.e. } sets
+there exist no cross-product terms simultaneously containing elements within the same set, i.e., sets
 $\{A,a_i\},\{B,b_j\}$ and $\{R,r_i\}$.
 We now show that $redSpoly$ actually satisfies this condition.
-This property of $redSpoly$ guarantees the word level variable can only exist in the form $A^{2^{i-1}}$,
-after substituting a bit-level variables with corresponding word-level variable.
+This property of $redSpoly$ guarantees the word-level variable can only exist in the form $A^{2^{i-1}}$,
+after substituting a bit-level variable with a corresponding word-level variable.
 
-3) Substitute bit-level variables in reduced $Spoly$, {\it i.e.} 
+3) Substitute bit-level variables in reduced $Spoly$, i.e., 
 $$\Func(a_i,b_j,r_l,A,B,R) \xrightarrow{a_i+\mathcal{G}_i(A),b_j+\mathcal{G}_j(B),r_l+\mathcal{G}_l(R)} \mathcal{H}(A,B,R)$$
 
-Use the result from pre-processing, get rid of $r_i, a_i$ and $b_i$ through substitution. This step yields following polynomial:
+Use the result from preprocessing, get rid of $r_i, a_i$ and $b_i$ through substitution. This step yields the following polynomial:
 \begin{equation}
@@
 
-4) Substitute PS word-level variable $R$ with inputs $A$ and $B$, {\it i.e.} 
+4) Substitute PS word-level variable $R$ with inputs $A$ and $B$, i.e.,
 $$\Func(A,B,R) \xrightarrow{R+\Func'(A,B)} \Func(A,B)$$
@@
 For PS variable $R$ we can use relation $R+\Func'(A,B)$, which is the last clock-cycle's output ($R'+\Func'(A,B)$) 
-with only leading term replaced in step ``$from^i\gets to^i$" in 
+with only the leading term replaced in step ``$from^i\gets to^i$" in 
 Algorithm \ref{alg:modified}. This step needs to be specially executed because of the exponent rewriting technique. 
@@
 
-{\it Evaluation of SAT/ABC/BDD based methods:} To verify circuit $S$
+{\it Evaluation of SAT/ABC/BDD-based methods:} To verify circuit $S$
 against the polynomial $\Func$, we unroll the SMPO over $k$ time-frames,
 and construct a miter against a combinational implementation of $\Func$.
-A (pre-verified) $\Fkk$ Mastrovito multiplier \cite{mastro:1989} is
+A (preverified) $\Fkk$ Mastrovito multiplier \cite{mastro:1989} is
 used as the {\it spec} model. This miter is checked for SAT using the
@@
 RH-SMPO against Mastrovito {\it   spec} are given in Table
-\ref{tbl:equiv} -- which shows that the techniques fail beyond 23 bit
+\ref{tbl:equiv} -- which shows that the techniques fail beyond 23-bit
 fields.
@@
 \centering
-\caption{\small Run-time for verification of bug-free RH-SMPO circuits
+\caption{Run-time for verification of bug-free RH-SMPO circuits
   for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
@@
 \centering
-\caption{\small Similarity between RH-SMPO and Agnew's SMPO}
+\caption{Similarity between RH-SMPO and Agnew's SMPO}
 \label{tbl:fraig}
@@
 
-\begin{table}[bp]
+\begin{table}[tbp]
 \centering
-\caption{\small Run-time (seconds) for verification of bug-free and
+\caption{Run-time (seconds) for verification of bug-free and
   buggy RH-SMPO with \textsc{Singular}} 
@@
 
-\begin{table}[bp]
+\begin{table}[tbp]
 \centering
-\caption{\small Run-time (seconds) for verification of
+\caption{Run-time (seconds) for verification of
   bug-free and buggy Agnew's SMPO with \textsc{Singular}}
@@
 
-\begin{table}[bp]
+\begin{table}[tbp]
 \centering
-\caption{\small Run-time (seconds) for verification of
+\caption{Run-time (seconds) for verification of
   RH-SMPO and Agnew's SMPO using C++ implementation}
@@
 This chapter has described a method to verify sequential Galois field
-multipliers over $\Fkk$ using computer algebra and algebraic geometry
-based approach. As sequential Galois field circuits perform the
+multipliers over $\Fkk$ using an approach based on computer algebra and algebraic geometry. 
+As sequential Galois field circuits perform the
 computations over $k$ clock-cycles, verification requires an efficient
-approach to unroll the computation, and represent it as a canonical
-word-level multi-variate polynomial. Using algebraic geometry, we show
-that the unrolling of the computation at word-level can be performed by
+approach to unroll the computation and represent it as a canonical
+word-level multivariate polynomial. Using algebraic geometry, we show
+that the unrolling of the computation at word level can be performed by
 \Grobner bases and elimination term orders. Subsequently, we show
@@
 elimination. Our approach is able to verify up to 162-bit sequential
-circuits, whereas contemporary techniques ({\it e.g.} BDDs) fail beyond 23-bit
+circuits, whereas contemporary techniques (e.g., BDDs) fail beyond 23-bit
 datapaths.

### 2017-03-03 dissertation/submission/prelim_GF.tex
after submission: revision 2

@@
 Additionally, we introduce a special type of sequential arithmetic hardware based on normal basis, as well
-as the normal basis theory behind the designing such hardware.
-
-The material is referred from \cite{galois_field:mceliece, ftheory:2006, ff:1997} for Galois field concepts, 
-\cite{mastro:1989, PT:1985, acar:1998, wu:2002, Knezevic:2008} for hardware design over Galois fields 
+as the normal basis theory behind designing such hardware.
+
+The material is taken from \cite{galois_field:mceliece, ftheory:2006, ff:1997} for Galois field concepts, 
+\cite{mastro:1989, PT:1985, acar:1998, wu:2002, Knezevic:2008} for hardware design over Galois fields, 
 and previous work by Lv \cite{lv:phd}.
-Normal basis theory in this section is referred from \cite{normal_book, gao:phd_normal_basis} and sequential
+Normal basis theory in this section is referred from \cite{normal_book, gao:phd_normal_basis}, and sequential
 normal basis arithmetic hardware designs come from \cite{mullinONB,MasseyOmura,agnew1991implementation, RHmulti}.
@@
 A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
-non-zero element in $\mathbb{F}$ has a multiplicative inverse; {\it i.e.} $\forall
+nonzero element in $\mathbb{F}$ has a multiplicative inverse; {\it i.e.} $\forall
 a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
@@
 A field is defined as a ring with one extra condition: the presence of a 
-multiplicative inverse for all non-zero elements.
+multiplicative inverse for all nonzero elements.
 Therefore, a field must be a ring while a ring is not necessarily a field.
@@
 field extension is to take a base field and construct a larger field which 
-contains the base field as well as satisfies additional properties. For example,
+contains the base field as well as satisfying additional properties. For example,
 the set of real numbers $\mathbb{R}$ forms a field; one extension of 
@@
 Finite fields find widespread applications in 
-many areas of electrical engineering and computer science such as error-
-correcting codes, elliptic curve cryptography, digital signal processing, 
+many areas of electrical engineering and computer science such as error-correcting
+codes, elliptic curve cryptography, digital signal processing, 
 testing of VLSI circuits, among others.
@@
 This section describes the relevant Galois field concepts
-\cite{galois_field:mceliece} \cite{ftheory:2006} \cite{ff:1997}
-and hardware arithmetic designs over such fields \cite{mastro:1989} \cite{PT:1985} 
-\cite{acar:1998} \cite{wu:2002} \cite{Knezevic:2008}.
+\cite{galois_field:mceliece,ftheory:2006,ff:1997}
+and hardware arithmetic designs over such fields \cite{mastro:1989,PT:1985,acar:1998,wu:2002,Knezevic:2008}.
 
@@
 Algebraic extensions of the binary field $\F_{2}$  
-are generally termed as {\it binary extension fields} $\Fkk$.
+are generally termed {\it binary extension fields} $\Fkk$.
 Where elements in $\F_2$ can only represent $1$ bit, elements in $\Fkk$ 
@@
 \begin{Definition}
-A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$, {\it i.e.} the set of all 
+A polynomial $P(x) \in \mathbb{F}_{2}\left[x\right]$, i.e., the set of all 
 polynomials in $x$ with coefficients in $\F_2$, is {\bf irreducible} 
-if $P(x)$ is non-constant with degree $k$ and cannot be 
+if $P(x)$ is nonconstant with degree $k$ and cannot be 
 factored into a product of polynomials of lower degree in $\mathbb{F}_2[x]$.
@@
 $\mathbb{F}_{2}$ if and only if it has no roots in $\mathbb{F}_{2}$,
-i.e if $\forall a \in \mathbb{F}_{2}$, $P(a)\neq 0$.
+i.e., if $\forall a \in \mathbb{F}_{2}$, $P(a)\neq 0$.
 For example, $x^2+x+1$ is an irreducible polynomial over $\mathbb{F}_{2}$
-because it has no solutions in $\mathbb{F}_{2}$, {\it i.e.} $(0)^2+(0)+1=1\neq0$ 
+because it has no solutions in $\mathbb{F}_{2}$, i.e., $(0)^2+(0)+1=1\neq0$ 
 and $(1)^2+(1)+1=1\neq0$ over $\F_2$.
@@
 $\mathbb{F}_{2^k} \equiv \mathbb{F}_2[x] \pmod{P(x)}$.
-Let $\alpha$ be a root of $P(x)$, {\it i.e.} $P(\alpha)=0$.
+Let $\alpha$ be a root of $P(x)$, i.e., $P(\alpha)=0$.
 Since $P(x)$ is irreducible over
@@
 To better understand this field extension, compare its similarities to another
-common-place
-field extension $\C$, the set of complex numbers. $\C$ is an extension of the field 
+commonplace field extension $\C$, the set of complex numbers. $\C$ is an extension of the field 
 of real numbers $\R$ with an additional element $i=\sqrt{-1}$, which is an imaginary
@@
 $P(x)=x^4+x^3+1 \in \mathbb{F}_2[x]$ is an irreducible polynomial of degree $k=4$. 
-Let $\alpha$ be the root of $P(x)$, {\it i.e.} $P(\alpha)=0$.
+Let $\alpha$ be the root of $P(x)$, i.e., $P(\alpha)=0$.
 
@@
 From an algebraic perspective, a field is a space, and field elements are points in the space. Those elements can be 
-represented with unique coordinates, which requires the pre-definition of a basis vector. In this
-section, we discuss a special basis called normal basis, as well as the advantages adopting it in 
+represented with unique coordinates, which requires the predefinition of a basis vector. In this
+section, we discuss a special basis called normal basis, as well as the advantages of adopting it in 
 GF operations, especially multiplication.
@@
 
-\subsection{Multiplication using Normal Basis}
+\subsection{Multiplication Using Normal Basis}
 Besides addition/subtraction, multiplication is also very common in arithmetic circuit design.
-The multiplication of GF elements in $\Fkk$ in StdB follows the rule of polynomial multiplication.
-However, it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
+The multiplication of GF elements in $\Fkk$ in StdB follows the rule of polynomial multiplication,
+but it will result in $O(k^2)$ bitwise operations. In other words, if we implement GF multiplication
 in a bit-level logic circuit, it will contain $O(k^2)$ gates. When the datapath size $k$ is large,
@@
 According to the definition, a normal basis is a vector where the next entry is the square of the former one.
-We note that the vector is cyclic, {\it i.e.} $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
+We note that the vector is cyclic, i.e., $\beta^{2^k} = \beta$ due to {\it Fermat's little theorem}.
 
@@
 
-This example shows that convenience to use NB when computing $2^k$ power of an element.
+This example shows the convenience of using NB when computing $2^k$ power of an element.
 Multiplication is more complicated than squaring; but when it is decomposed as bit-wise
@@
 Before discussing the details of the function $f_{mult}$, we can 
-square both sides of Equation \ref{eqn:multiMSB}, {\it i.e.} $C^2 = A^2\times B^2$.
+square both sides of Equation \ref{eqn:multiMSB}, {\it i.e.}, $C^2 = A^2\times B^2$.
 Obviously, using the property in Lemma \ref{lem:squareNB}, the original second most significant bit 
@@
 
-From above example, it is known that a universal structure that implements $f_{mult}$ can be reused
+From the above example, it is known that a universal structure that implements $f_{mult}$ can be reused
 $k$ times in NB multiplication over $\Fkk$. Compared to StdB, which requires a distinct design 
@@
 we obtain a series of conjugate matrices of $M$. 
-Which means instead of shifting operands $A$ and $B$, we can also shift the 
+This means that instead of shifting operands $A$ and $B$, we can also shift the 
 matrix.
@@
 $$c_l = A \times M^{(l)} \cdot B^T$$
-Meanwhile, the operator shifting rule in Equation \ref{eqn:shiftMSB} still holds. Then we have the relation 
+Meanwhile, the operator-shifting rule in Equation \ref{eqn:shiftMSB} still holds. Then we have the relation 
 $$c_{l-1} = A \cdot M^{(l-1)} \cdot B^T = shift(A) \cdot M^{(l)} \cdot shift(B)^T$$ which means
@@
 The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
-NB, {\it i.e.}
+NB, i.e.,
 \begin{equation}
@@
 nonzero entries in $\lambda$-Matrix.
-There usually exists multiple NBs in $\Fkk, k>3$. If we employ a random NB, there is no mathematical
+There usually exist multiple NBs in $\Fkk, k>3$. If we employ a random NB, there is no mathematical
 guarantee that $C_N$ has bound $o(k)$. However,
 Mullin {\it et al.} \cite{mullinONB} % This citation is valid
-proves that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
+prove that in certain GF $\mathbb F_{p^{k_{opt}}}$, there always exists at least one NB such that 
 its corresponding $\lambda$-Matrix has $C_N = 2n-1$ nonzero entries. A basis with this property is
-called optimal normal basis (ONB), details are introduced in Appendix \ref{append:ONB}.
-
-In practice, large size NB multipliers are usually designed in $\Fkk$ when ONB exists
+called optimal normal basis (ONB), details of which are introduced in Appendix \ref{append:ONB}.
+
+In practice, large NB multipliers are usually designed in $\Fkk$ when ONB exists
 to minimize the number of gates. So in the following part of this chapter and our experiments,
@@
 $\alpha^8+\alpha^7+\alpha^6+\alpha^4+\alpha^2+\alpha+1$. Coincidently, $\alpha$ is also a normal element,
-i.e. $\beta = \alpha$ can construct a NB $\{\alpha,\alpha^2,\alpha^4,\alpha^8,\alpha^{16},\alpha^{32},\alpha^{64},
+i.e., $\beta = \alpha$ can construct a NB $\{\alpha,\alpha^2,\alpha^4,\alpha^8,\alpha^{16},\alpha^{32},\alpha^{64},
 \alpha^{128}\}$.
@@
 
-The final remainder is $1$, {\it i.e.} the product equals to 1 in StdB.
+The final remainder is $1$, {\it i.e.} the product equals 1 in StdB.
 
@@
 As a result, the most naive NB multiplier design (or Massey-Omura multiplier \cite{MasseyOmura})
-contains 53 gates in total, which is a great saving in area cost comparing to StdB multipliers.
+contains 53 gates in total, which is a great saving in area cost compared to StdB multipliers.
 
@@
 \subsection{Sequential Multiplier with Parallel Outputs}
-The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
+The major benefit of NB multiplier originates from the sequential design. A straightforward design implementing 
 the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.
@@
 
-The behavior of Massey-Omura multiplier can be described as: pre-load operands $A,B$ and reset $C$ to 0, after 
+The behavior of Massey-Omura multiplier can be described as follows: preload operands $A,B$ and reset $C$ to 0, after 
 executing for $k$ clock cycles, the data stored in flip-flop array $C$ is the product $A\times B$.
 We note that there is only one output giving 1 bit of the product each clock cycle, which matches the 
-definition of serial output to communication channel. Therefore this type of design is named as 
+definition of serial output to communication channel. Therefore this type of design is called a 
 sequential multiplier with serial output (SMSO).
-The SMSO architecture need $C_N$ AND gates and $C_N-1$ XOR gates, which equals to $2k-1$ AND gates and $2k-2$ XOR gates
+The SMSO architecture needs $C_N$ AND gates and $C_N-1$ XOR gates, which equals to $2k-1$ AND gates and $2k-2$ XOR gates
 if it is designed using ONB. In fact, the number of gates can be reduced if the multiplication is 
@@
 clock cycle, make $k$ outputs and add them up using shift register after $k$ clock cycles, 
-it will generate the same result with SMSO. This kind of architecture is named as a 
+it will generate the same result with SMSO. This kind of architecture is called a 
 sequential multiplier with parallel outputs (SMPO). The basic SMPO, as a conjugate of Massey-Omura multiplier,
@@
 \caption{5-bit Agnew's SMPO.
-Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, {\it i.e.} if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
+Index $i$ satisfies $0<i<4$, indices $u,v$ are determined by column \# of nonzero entries in $i$-th row of $\lambda$-Matrix $M^{(0)}$, i.e., if entry $M_{ij}^{(0)}$ is a nonzero entry, $u$ or $v$ equals to $i+j \pmod 5$.
 Index $w = 2i\pmod 5$.}
@@
 It is written in the form of Equation \ref{eqn:SMPOterms}. In next clock cycles we can obtain 
-remaining bits of the product, which can be written in following general form polynomial:
+remaining bits of the product, which can be written in the following general form polynomial:
 \begin{align}
@@
 signal sent to the flip-flop in block $R_0$ denotes function:
-$$R_0^{(1)} = a_1b_0$$
-It equals to the first term of Equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
+$$R_0^{(1)} = a_1b_0.$$
+This equals the first term of Equation \ref{eqn:5bitSMPO_c0}. In the second clock cycle,
 this signal is sent to block $R_1$ through wire $r_0$, and this block also receives data 
@@
 $$R_1^{(2)} = R_0^{(1)} + (a_0+a_3)b_1 = a_1b_0+(a_0+a_3)b_1$$
-which forms first 2 terms of Equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
-on $R_2$ in third clock cycle, signal on $R_3$ in fourth clock cycle, finally we can get 
+which forms the first 2 terms of Equation \ref{eqn:5bitSMPO_c0}. Similarly, we track the signal 
+on $R_2$ in the third clock cycle, signal on $R_3$ in the fourth clock cycle, finally we can get 
 $$R_4^{(5)} = a_1b_0+(a_0+a_3)b_1 + (a_3+a_4)b_2 + (a_1+a_2)b_3 + (a_2+a_4)b_4$$
-which equals to $c_0$ in Equation \ref{eqn:5bitSMPO_c0}.
+which is equal to $c_0$ in Equation \ref{eqn:5bitSMPO_c0}.
 After the fifth clock cycle ends, this signal can be detected on wire $r_0$. It shows that 
@@
 
-If we track $R_1\to R_2\to R_3 \to \cdots \to R_0$, we can obtain $c_1$ respectively.
+If we track $R_1\to R_2\to R_3 \to \cdots \to R_0$, we can obtain $c_1$.
 Thus we conclude that Agnew's SMPO functions the same with Massey-Omura multiplier.
@@
 
-\subsection{Multiplier not based on the $\lambda$-Matrix}
+\subsection{Multiplier Not Based on the $\lambda$-Matrix}
 Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
@@
 After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
-some of them jump out of the box and are independent from $\lambda$-Matrix.
-One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
+some of which jump out of the box and are independent from $\lambda$-Matrix.
+One competitive multiplier design of this type was invented by Reyhani-Masoleh and Hasan 
 \cite{RHmulti}, which is therefore called RH-SMPO.
 
-Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO which is functionally equivalent to 5-bit Agnew's SMPO 
+Figure \ref{fig:5bitRH} is a 5-bit RH-SMPO, which is functionally equivalent to the 5-bit Agnew's SMPO 
 in Figure \ref{fig:SMPO}. A brief proof is as follows:
+
+\begin{figure}[bp]
+\centering{
+\includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
+\caption{A 5-bit RH-SMPO.}
+\label{fig:5bitRH}}
+\end{figure}
 
@@
 $$d_{i,v} = (a_i+a_{i+v})b_i$$
-With the auxiliary function, we can utilize following theorem (proof refers to \cite{RHmulti}):
+With the auxiliary function, we can utilize the following theorem (proof refers to \cite{RHmulti}):
 \begin{Theorem}
-Consider three elements $A,B$ and $R$ such that $R=A\times B$ over $\Fkk$. Then,
+Consider three elements $A,B$, and $R$ such that $R=A\times B$ over $\Fkk$. Then,
 $$R=(((F_{k-1}^2+F_{k-2})^2+F_{k-3})^2+\cdots+F_1)^2+F_0$$
+where auxiliary function $F_i$ is as given in Equation \ref{eqn:aux}.
+This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
+$R_i$ flip-flops.
 \end{Theorem}
-where $F$ is as given in Equation \ref{eqn:aux}.
-This form is called inductive sum of squares, and corresponds to the cyclic shifting on 
-$R_i$ flip-flops. Concretely, the multiplier behavior is an implementation of 
+Concretely, the multiplier behavior is an implementation of the
 following algorithm:
@@
 $$F_{k-l} = F_{k-1}(A^{2^{l-1}},B^{2^{l-1}}),~~1\le l\le k$$
-So using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
+So, using fixed $F_{k-1}$ and squaring $A^{2^i}$ every time inside the loop is equivalent to computing 
 sequence $F_{k-1},F_{k-2},\dots,F_0$ with fixed operands $A,B$.
 \end{Proof}
-
-\begin{figure}[bp]
-\centering{
-\includegraphics[width=\textwidth]{newfig/5bitRH.pdf}
-\caption{A 5-bit RH-SMPO.}
-\label{fig:5bitRH}}
-\end{figure}
 
@@
 
-Similarly, from row 2 of M-table we derive that $d_2$ has fanouts $e_3,e_4$:
+Similarly, from row 2 of the M-table we derive that $d_2$ has fanouts $e_3,e_4$:
 \begin{equation*}
@@
 
-Let us look back at Equation \ref{eqn:5bitRHaux}, we already dealt with the latter part.
+Let us look back at Equation \ref{eqn:5bitRHaux}; we already dealt with the latter part.
 The first term is always $d_0\beta$, which denotes $d_0$ should always be connected to $e_0(\beta)$.
@@
 Then the last mission is to implement the output $R_i$ layer. Assume $r_{i-1}$ is the output of 
-$R_{i-1}$ in last clock cycle, we can connect using relation 
+$R_{i-1}$ in the last clock cycle, we can connect using the relation 
 $$R_i = r_{i-1} + e_i$$
 In this example, according to the M-table in Figure \ref{fig:multitable}, columns $e_1,e_2$
-have only zeros in its intersection with row $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.
+have only zeros in their intersection with rows $d_1,d_2$. Thus gates for $e_1,e_2$ can be omitted.
 
@@
 
-The area cost of RH-SMPO is even smaller than Agnew's SMPO. XOR gates corresponds to all nonzero entries 
+The area cost of RH-SMPO is even smaller than Agnew's SMPO. XOR gates correspond to all nonzero entries 
 in M-table, which is with the same number of nonzero entries in $\lambda$-Matrix ($C_N$). The number of AND gates 

### 2017-03-03 dissertation/submission/prelim_basis.tex
after submission: revision 2

@@
 in subsequent chapters. The material of this chapter is
-mostly referred from the textbooks \cite{ideals:book,gb_book} and 
+mostly taken from the textbooks \cite{ideals:book,gb_book} and 
 previous work by Lv \cite{lv:phd} as well as Pruss \cite{pruss:tcad15}.
@@
 where $\alpha=({\alpha}_1,\cdots,{\alpha}_d)$ is a vector size $d$ of 
-integers $\ge 0$, {\it i.e.} $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.
+integers $\ge 0$, i.e., $\alpha \in \mathbb{Z}_{\ge 0}^{d}$.
 
@@
 A {\bf degree-reverse-lexicographic ordering} (DEGREVLEX) is the same as a
-DEGLEX ordering, however terms are LEXed in reverse.
+DEGLEX ordering, but terms are LEXed in reverse.
 
@@
 Let $\F$ be a field and let $f, g \in \F[x_1, x_2, \ldots, x_d]$ be polynomials
-over the field. {\bf Polynomial division} of $f$ by $g$ is computes following:
+over the field. {\bf Polynomial division} of $f$ by $g$ is computed following:
 \begin{equation}
@@
 where $r$ is the resulting polynomial of the division.
-If $\frac{lt(f)}{lt(g)}$ is non-zero, then $f$ is considered divisible by $g$, 
-{\it i.e.} $g \mid f$.
+If $\frac{lt(f)}{lt(g)}$ is nonzero, then $f$ is considered divisible by $g$, 
+{\it i.e.}, $g \mid f$.
 \end{Definition}
@@
 \end{equation}
-Since $\frac{lt(f)}{lt(g)}$ is non-zero $g|f$. The division, $f\xrightarrow{g} r$, 
+Since $\frac{lt(f)}{lt(g)}$ is nonzero $g|f$. The division, $f\xrightarrow{g} r$, 
 is computed as:
@@
     represents the reduced polynomial $r$ resulting from $f$ as reduced by a 
-    set of non-zero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
+    set of nonzero polynomials $F = \{f_1,\dots,f_s\}$.  The polynomial $r$ is considered {\bf reduced} if 
     $r = 0$  or no term in $r$ is divisible  by $lm(f_i), \forall f_i \in F$.
@@
 %%%%%%%%%%%%%%%%%%%%%  variety %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In computer-algebra based formal verification,
+In computer-algebra-based formal verification,
 it is often necessary to analyze the presence 
@@
 
-The above example shows the variety can be infinite, finite (non-empty set) 
+The above example shows the variety can be infinite, finite (nonempty set) 
 or empty. It is interesting to note that since we will be operating over
-finite fields $\F_{q}$, and any finite set of points is a variety. Likewise,
+finite fields $\F_{q}$, any finite set of points is a variety. Likewise,
 any variety over $\F_{q}$ is finite (or empty).
@@
 $\{f_1,\dots,f_{s}\}$ and $\{g_{1},\dots,g_{t}\}$ that may generate the same 
-ideal, {\it i.e.} 
-$\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
+ideal, i.e., $\langle f_{1},\dots,f_{s}\rangle=\langle g_{1},\dots,g_{t}\rangle$. Since 
 variety depends on the ideal, these sets of polynomials have the same 
@@
 
-Ideals and their varieties are a key part of computer-algebra based formal
+Ideals and their varieties are a key part of computer-algebra-based formal
 verification. A given hardware design can be transformed into a set
@@
 \end{eqnarray}
-If $I$ is the ideal generated by $f_1$ and $f_2$, {\it i.e.} 
+If $I$ is the ideal generated by $f_1$ and $f_2$, {\it i.e.}, 
 $I=\langle f_1, f_2 \rangle$, then we find $h \in I$ as follows:
@@
 
-\Grobner basis has an important property, and therefore can be used to perform ideal membership test.
+\Grobner basis has an important property, and therefore can be used to perform an ideal membership test.
 Formally speaking, a \Grobner basis gives a decision procedure to test 
@@
 \stackrel{f_1}{\textstyle\longrightarrow} x^2 - x  \neq 0$;
-{\it i.e.} dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
+{\it i.e.}, dividing $f$ by $f_2$ first and then by $f_1$ does not lead to a
 zero remainder. However,  if we compute the Gr\"{o}bner basis $G$ of
@@
 \rangle$, where $f_1 = yx - y, ~f_2 = y^2 - x$. 
-Assume a degree-lexicographic term ordering with $y > x$ is imposed.
+Assume a degree-lexicographic term ordering with $y > x$ imposed.
 
@@
 
-When computing a \Grobner basis, it's important to note that if $lt(f_i)$ 
+When computing a \Grobner basis, it is important to note that if $lt(f_i)$ 
 and $lt(f_j)$ have no common variables, the $Spoly$ reduction step in 
@@
 \begin{enumerate}
-\item Minimize every $g_i \in G$, i.e $g_i=g_i/lc(g_i)$
+\item Minimize every $g_i \in G$, i.e., $g_i=g_i/lc(g_i)$
 \item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
-{\it i.e.} remove every polynomial in $G$ whose leading term is divisible by the leading term 
+i.e., remove every polynomial in $G$ whose leading term is divisible by the leading term 
 of some other polynomial in $G$.
@@
 The high computational complexity of the Gr\"obner basis problem is an important issue
-because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
+because of its high cost in time and space. Concretely, for an arbitrary polynomial set, 
 the worst case computation time/space cost of its Gr\"obner basis is doubly exponential, as the total degree of 
@@
 $O(2(\frac{1}{2}d^2+d)^{2^{n-1}})$ where $d$ is the degree of the ideal and $n$ is the number of variables
-\cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
+\cite{dube1986complexity}. However, in situations such as applications of circuit verification, 
 the polynomial is well restricted rather than arbitrary.
-Gao {\it et al.} \cite{gao:qe-gf-gb} proves that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
+Gao {\it et al.} \cite{gao:qe-gf-gb} prove that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
 This provides the possibility to make use of Gr\"obner basis under restricted situations,
@@
 \end{Definition}
-In other words, any non-constant polynomial equation over 
+In other words, any nonconstant polynomial equation over 
 $\overline {\F}\left[x\right]$ always has at least one root 
@@
 be an ideal satisfying $V(J)=\emptyset$. 
-Then $J=\overline {\F}[x_1, x_2, \cdots, x_d]$, Or equivalently, 
+Then $J=\overline {\F}[x_1, x_2, \cdots, x_d]$, or equivalently, 
 \begin{equation}
@@
 
-\begin{Theorem}[Strong Nullstellensatz]
+\begin{Theorem}[Strong Nullstellensatz \cite{gb_book}]
 \label{thm:sns}
-\cite{gb_book} 
 Let $\overline{\F}$ be an algebraically closed field, and let $J$
@@
 %%%%%%%%%%%%%%%%weak nullstellensatz in finite field%%%%%%%%%%%%%
-\begin{Theorem}[Weak Nullstellensatz in $\Fq$]
+\begin{Theorem}[Weak Nullstellensatz in $\Fq$ \cite{null:1890}]
 \label{wnull:ff}
-\cite{null:1890}
 Given $f_1,f_2,\cdots,f_s \in \Fq[x_1,x_2,\cdots,x_d]$. 
@@
 Elimination of certain variables in a system of polynomials is a common 
-operation when some variables is not needed in modeling and analysis. 
-In this section, eliminating variables targets a tight-bound over-approximation.
+operation when some variables are not needed in modeling and analysis. 
+In this section, eliminating variables targets a tight-bound overapproximation.
 This is equivalent to the existential quantifier elimination in first-order logic,
-or variable smoothing in Boolean operations. We introduce a elimination method 
+or variable smoothing in Boolean operations. We introduce an elimination method 
 based on algebraic geometry concepts, and use it as the fundamental of 
@@
 
-Formally, we define concepts of projection and elimination ideal:
+Formally, we define concepts of the projection and elimination ideals:
 \begin{Definition}
@@
 The elimination ideal $J_l$ has eliminated all the variables 
-$x_1,\dots,x_l$, {\it i.e.} it only contains polynomials with variables in
+$x_1,\dots,x_l$, i.e., it only contains polynomials with variables in
 $x_{l+1},\dots,x_d$. 
@@
 that 
-$$\pi_l(V(J+J_0)) = V((J+J_0)_l)$$
-
+$$\pi_l(V(J+J_0)) = V((J+J_0)_l).$$
 This shows that in finite fields, projection of the variety $V_{d}(\langle f_1,\dots,f_s\rangle)$
@@
 
-Elimination theory provides the basis for following abstraction approach.
-\subsection{Abstraction using Nullstellensatz and Gr\"obner Basis}
+Elimination theory provides the basis for the following abstraction approach.
+\subsection{Abstraction Using Nullstellensatz and Gr\"obner Basis}
 \begin{Problem}
@@
 Since $f_{\Func}$ is ultimately derived from the circuit implementation, 
-it agrees with the solution to the system of polynomials $\{S\}=0$, {\it i.e.}:
+it agrees with the solution to the system of polynomials $\{S\}=0$, i.e.,:
 \begin{equation}
@@
 
-\begin{Theorem}[Abstraction Theorem]
-\cite{tim:phd}
+\begin{Theorem}[Abstraction Theorem \cite{tim:phd}]
 Using the notations from Problem Setup \ref{setup:abs} at the beginning of this subsection,

### 2017-03-03 dissertation/submission/prev.tex
dissertation: after submission_1st revision
after submission: revision 2

@@
 \label{ch:prev}
+\vspace{-0.8cm}
 \section{Sequential Equivalence Checking}
@@
 have been developed over decades and widely utilized in both academia and industry. 
-The specification of a sequential circuit can be modeled as a (golden model) state machine,
+The specification of a sequential circuit can be modeled as a (golden model) state machine;
 SEC is performed to compare the functionality between the circuit for test and the golden one.
@@
 Researchers proposed improvements by using Boolean functions to represent 
-a set of states/transitions \cite{coudert2003unified, coudert1990verification},  or by dividing the sequential circuit to 
-into a smaller subcircuit and re-model the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
-interfaces that focuses on only the designated initial states and remove redundancies in state space \cite{baumgartner2007scalable}.
+a set of states/transitions \cite{coudert2003unified, coudert1990verification},  or by dividing the sequential circuit
+into a smaller subcircuit and remodeling the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
+interfaces that focuses on only the designated initial states and removes redundancies in state space \cite{baumgartner2007scalable}.
 
@@
 is copied over several time-frames then connected together.
-The procedure is called {\it unrolling} of sequential circuit, as 
+The procedure is called {\it unrolling} of a sequential circuit, as 
 Figure \ref{fig:unrolling} shows.
 
-\begin{figure}[h]
+\begin{figure}[tbp]
 \centerline{
@@
 
-Unrolling provides a way to transform a sequential circuit to combinational 
+Unrolling provides a way to transform a sequential circuit into a combinational 
 circuit. Therefore,  methods which can be applied to combinational circuit 
@@
 \subsection{Canonical Decision Diagrams}
-The decision diagrams (DDs) are invented by inspirations of 
-using optimized data structures to accelerate formal verification.
+The decision diagrams (DDs) are optimized data structures which can significantly accelerate formal verification.
 The most fundamental DD is the Binary DD (BDD), which originates from the 
@@
 A BDD is usually represented as a binary tree.
-Its ordered and reduced form -- the Reduced Ordered Binary Decision Diagram (ROBBD)
+Its ordered and reduced form, the Reduced Ordered Binary Decision Diagram (ROBBD)
 \cite{BRYA86}, was the first significant contribution because of its canonicity.  
@@
 (DAG). Manipulation of Boolean functions can then be carried out as
-composition operations on their respective DAGs.
+composition operations on their respective DAGs. An example of ROBDD is shown as Figure \ref{fig:BDD}.
 
-Following BDDs,  variants of the Shannon's decomposition principle
+Following BDDs,  variants of Shannon's decomposition principle
 were explored to develop other functional decision diagrams such as
@@
 to represent the existence of variables in a cube, and result in lower 
-space complexity. It can be used to represent polynomials with integer coefficients.
+space complexity. They can be used to represent polynomials with integer coefficients.
 
-\begin{figure}[h]
+\begin{figure}[bp]
 \centerline{
@@
 
-Binary Moment Diagrams (BMDs) \cite{bmd}, and its derivatives K*BMDs
+Binary Moment Diagrams (BMDs) \cite{bmd}, and their derivatives K*BMDs
 \cite{kbmd} and *PHDDs \cite{phdd}, perform the decomposition of a {\it linear} function
@@
 \cite{BRYA86} and their word-level variants \cite{WLS}),
-non-canonical representations such as And-Invert-Graph (AIG) based reductions 
+noncanonical representations such as And-Invert-Graph-based (AIG-based) reductions 
 \cite{AIG:2002,alanmi:cec:iccad2006} are also very effective. 
@@
 as long as the miter of two circuits can be described using conjunctive normal form (CNF)
-formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, {\it etc.}
-If the circuits being compared are structurally highly similar, AIG and circuit-SAT 
-based approaches are known to be efficient.
+formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, etc.
+If the circuits being compared are structurally highly similar, AIG and circuit-SAT-based
+approaches are known to be efficient.
 However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
   contemporary techniques including quantifier-free bit-vector 
-  (QF-BV) theory based SMT-solvers \cite{Cryptol:fmcad09},
+  (QF-BV) theory-based SMT-solvers \cite{Cryptol:fmcad09},
   offer a practical solution.
 
-Recently integer polynomial based techniques \cite{ciesielski2014function,rolf:date16} are proposed  to verify the functional 
+Recently integer polynomial based techniques \cite{ciesielski2014function,rolf:date16} have been proposed  to verify the functional 
 correctness of integer arithmetic circuits. Their approach formulates the output signature as a polynomial function 
@@
 will be composed by only input variables. Then the polynomial is converted to a canonical representation, and compared
-with designated input signature. If they are equivalent, then the arithmetic circuit is successfully verified.
+with a designated input signature. If they are equivalent, then the arithmetic circuit is successfully verified.
 This approach incurs polynomial term explosion during the backward rewriting. The authors proposed a heuristic
 to levelize the arithmetic circuit, and substitute several gates' variables at the same time to minimize the risk. However, 
-the heuristic proved to be less effective when the inner symmetry of circuit structure is missing.
+the heuristic proved to be less effective when the inner symmetry of the circuit structure is missing.
 
@@
 Model checking is a way to verify certain safety and liveness properties 
-in sequential circuits. Symbolic model checking which avoids using explicit state encoding
+in sequential circuits. Symbolic model checking, which avoids using explicit state encoding,
 provides more flexibility to reduce the state space and enhance the 
@@
 
-The $k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
-on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
+The $k$-BMC with interpolation is a purely incremental model-checking approach, and the interpolation procedure relies
+on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker called IC3 is developed 
 \cite{bradley2011sat,bradley2011incremental}. IC3 works incrementally to find inductive subclauses
-of negations of reached states, meanwhile it is monolithic when computing over-approximations to sets of reachable
-states within $1,2,\dots,k$ steps. It is proved to be more efficient than interpolation based model checking
+of negations of reached states, meanwhile it is monolithic when computing overapproximations to sets of reachable
+states within $1,2,\dots,k$ steps. It is proved to be more efficient than interpolation-based model checking,
 although using similar mechanisms.
 
-Above techniques have limitations: they all rely on bit-level information from 
-the circuit, which prevents them from being applied to circuit with large datapaths.
+The above techniques have limitations: they all rely on bit-level information from 
+the circuit, which prevents them from being applied to circuits with large datapaths.
 Meanwhile, their implementation relies on SAT/BDDs, which is an extension of Boolean 
@@
 
-\section{Word-level Techniques applied to Sequential Circuit Synthesis and Validation}
+\section{Word-level Techniques Applied to Sequential Circuit Synthesis and Validation}
 To better verify word-level designs, word-level verification techniques have been 
@@
 \cite{devadas:algebraic_manipulation_iccd91}, or the ones based on
-term re-writing \cite{AST}, {\it etc.}
+term rewriting \cite{AST}, etc.
 
-Polynomial, integer and other non-linear representations have also
+Polynomial, integer, and other nonlinear representations have also
 been researched: Difference Decision Diagrams (DDDs) \cite{ddd-csl99,ddd-mt-98}, interval
@@
 
-\section{Verification using Algebraic Geometry}
+\section{Verification Using Algebraic Geometry}
 
@@
 \underline{a}bstraction \underline{t}erm \underline{o}rder (called
-RATO), that enables to compute the \Grobner basis of a smaller subset
-of polynomials. The authors show that their approach can prove
+RATO) that enables them to compute the \Grobner basis of a smaller subset
+of polynomials. The authors show that their approach can prove the
 correctness of up to 571-bit combinational GF multipliers.
@@
 IBM proposed a method to apply algebraic geometry techniques 
-on verifying error coding circuits \cite{BLUEVERI}.
+to verifying error coding circuits \cite{BLUEVERI}.
 Recent papers \cite{rolf:date16,rolf:FMCAD16} provide a way to utilize 
@@
 \cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
-GB-algorithms for CTL, LTL and 
+GB-algorithms for CTL, LTL, and 
 bounded MC over Boolean rings. However, these approaches are a
-straight-forward transformation of the problem to {\it bit-level}
+straightforward transformation of the problem to {\it bit-level}
 Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
 the concepts of word-level reachability, abstraction-refinement using
-interpolation or UNSAT cores, {\it etc.}, that we desire was not the focus of
+interpolation or UNSAT cores, etc., that we desire were not the focus of
 \cite{Avrunin:CAV,vardi-iasted07}.
@@
 From the investigation of previous work, techniques are to be researched that 
-can perform the FSM traversal on word-level to verify a property excluding spurious 
+can perform the FSM traversal at word level to verify a property excluding spurious 
 faults. Meanwhile, many abstraction refinement techniques utilize information from UNSAT cores.
 We propose to solve these problems in the context of word-level verification,
-with data representation, abstraction and algorithm execution all carried out 
-on word-level.
+with data representation, abstraction, and algorithm execution all carried out 
+at word level.
 
-In this dissertation, we propose a purely word-level reachability analysis approach which has never been done before.
-We achieve this by modeling the transition relations, states and the traversal algorithm at word-level.
+In this dissertation, we propose a purely word-level reachability analysis approach, which has never been done before.
+We achieve this by modeling the transition relations, states and the traversal algorithm at word level.
 We borrow inspirations from \cite{tim:phd,gao:qe-gf-gb} to perform state space abstraction.
-Moreover, we demonstrate applications of our proposed approach to sequential arithmetic verification which is 
-not done before, either.
+Moreover, we demonstrate applications of our proposed approach to sequential arithmetic verification, which has 
+not been done before, either.
 Finally, we show algebraic geometry analogs of UNSAT cores of polynomial ideals, and describe algorithms

### 2017-03-03 dissertation/submission/recha.tex
after submission: revision 2

@@
 formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
-tools, in order to improve various performance indicators such as latency, clock skew or power
+tools, in order to improve various performance indicators such as latency, clock skew, or power
 density, sequential optimization techniques such as retiming \cite{retiming}, scan logic \cite{scan},
-sequential power optimization \cite{mathur2009power} and clock-gating techniques \cite{clockgating} are applied.
+sequential power optimization \cite{mathur2009power}, and clock-gating techniques \cite{clockgating} are applied.
 % some major refactoring and attachments are made on original designs. 
 These modifications may introduce bugs, errors or malfunctions to the original logic and cause problems. 
-Based on traditional localized simulation or formal verification method ({\it e.g.} equivalence checking), 
+Based on traditional localized simulation or formal verification method (e.g., equivalence checking), 
 designers are reluctant to make aggressive optimization
-since the malfunctions are considered as ``faults" in the circuits.
+since the malfunctions are considered ``faults" in the circuits.
 However, if the circuit behavior is carefully investigated, it may become evident that 
-those ``faults" may never be activated during a restricted execution starting 
-from legal initial states and with legal inputs. Thus we will call those ``faults" as 
+those faults may never be activated during a restricted execution starting 
+from legal initial states and with legal inputs. Thus we will call those faults
 {\it spurious faults} (false negatives), since they will not affect the circuit's normal behavior.
@@
 take specific legal inputs, a set of reachable states can be derived. 
-As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
+As long as the faults can be modeled as bad states, we can judge whether they are 
 spurious faults by checking if they belong to the unreachable states. From the spurious fault validation 
-perspective, reachability analysis is a must when developing full set of sequential circuit verification
+perspective, reachability analysis is a must when developing a full set of sequential circuit verification
 techniques.
 
-There are quite a few methods to perform reachability checking on FSMs. One among them is 
+There are quite a few methods to perform reachability checking on FSMs. One of them is 
 state space traversal. Conventionally, the algorithm is based on bit-level techniques such as
-binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
+binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm at word level,
 which brings critical advantages. In this chapter the approach will be described and discussed in depth, 
@@
 \section{Motivation}
-Before introducing the details of our approach, there are a few questions to ask: what is 
-the benefit of executing FSM traversal at word-level? How can algebraic geometry make this happen?
+Before introducing the details of our approach, there are a few questions to ask: What is 
+the benefit of executing FSM traversal at word level? How can algebraic geometry make this happen?
 The answers can be found in this section, as a statement of the motivation of our research.
@@
 Sequential circuits are modeled as FSMs, which can be implemented as graphs. Thus a 
-graph-traversal based algorithm is created to analyze the reachable states \cite{coudert2003unified}.
+graph-traversal-based algorithm is created to analyze the reachable states \cite{coudert2003unified}.
 A traversal algorithm using the concept of implicit state enumeration is proposed \cite{cho1993redundancy}.
@@
 
-Above algorithm describes a breath-first-search (BFS) traversal in
-state space. The traversal algorithm is a simple variation of BFS algorithm where 
+The above algorithm describes a breath-first-search (BFS) traversal in
+state space. The traversal algorithm is a simple variation of a BFS algorithm where 
 states are nodes and transitions are arcs. Each state is uniquely encoded by a combination
@@
 component, the traversal algorithm analyzes the combinational logic and derives the transition 
-function for one-step reachability within current time-frame, and extends the result to complete 
-execution paths through unrolling. If each state encoding ({\it i.e.} exact values in the selected registers) is
+function for one-step reachability within the current time-frame, and extends the result to complete 
+execution paths through unrolling. If each state encoding (i.e., exact values in the selected registers) is
 explicitly analyzed and counted during the unrolling 
@@
 directly read in the execution; instead, they are implicitly represented using a conjunction of several 
-Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.
+Boolean formulas. Such techniques differ from explicit unrolling and are called {\bf implicit unrolling}.
 
@@
 Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
-the size of BDD explodes when the formulas become too long and too complicated. In \cite{cho1993redundancy},
+the size of a BDD explodes when the formulas become too long and too complicated. In \cite{cho1993redundancy},
 the authors make a compromise between accuracy and cost, and turn to approximate reachability. 
-In this research work, the aim is to explore a word-level technique which can make a accurate reachability
+In this research work, the aim is to explore a word-level technique which can make an accurate reachability
 analysis available.
@@
 The level of integration of modern digital circuit designs is very high. For example, processor A10 designed by Apple
-integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such a high density makes the silicon 
+integrates 3.3 billion transistors on a $125~\text{mm}^2$ chip \cite{AppleA10}. Such a high density makes the silicon 
 implementation of large datapaths possible. In recent decades, 64-bit or even larger datapaths frequently appear in 
-modern digital IC designs such as powerful central processing units (CPUs) and high bandwidth memory (HBM).
+modern digital IC designs such as powerful central processing units (CPUs) and high-bandwidth memory (HBM).
 Meanwhile, with the development of electronic design automation (EDA) tools, data flow 
 is described by the designer as word-level specifications. Therefore, it will be straightforward 
-and beneficial for users if formal verification tools can work on word-level. Moreover, adopting word-level 
+and beneficial for users if formal verification tools can work at word level. Moreover, adopting word-level 
 techniques will greatly reduce the state space and make verification more efficient.
 
-In order to throw light on the advantages using word-level techniques, we pick a typical 
-digital circuit component in modern 64-bit MIPS processor as an example.
+In order to throw light on the advantages of using word-level techniques, we pick a typical 
+digital circuit component in a modern 64-bit MIPS processor as an example.
 
@@
 Finally when the most significant bit (MSB) of preloaded multiplicand arrives at the MSB of product registers,
-the registers contains the result -- 128 bits product. The behavior can be described by the following algorithm:
+the registers contains the result -- a 128-bit product. The behavior can be described by the following algorithm:
 
@@
 
-\begin{figure}[bp]
+\begin{figure}[tbp]
 \centering{
@@
 % List the detailed bit-level property need to check P==P+B
-Traditionally, to verify the functional correctness of this multiplier, satisfiability (SAT) based or BDD based 
+Traditionally, to verify the functional correctness of this multiplier, SAT-based or BDD-based 
 model checking is applied on basic function units. For example, as a part of functional verification, 
@@
 \end{equation}
-
-The abstraction to word-level will reduce symbolic storage and execution cost; the complexity 
+the abstraction to word-level will reduce symbolic storage and execution cost; the complexity 
 to traverse the state space will be greatly reduced.
@@
 On the other hand, when implementation details of the datapath are not available, it is not convenient 
-any more for users of conventional model checker. The reason is that the user has to write all clauses 
-for the implementation, which contains cross-literals, {\it e.g.} $s_2$ may associate with $a_1, p_0$, {\it etc.} If the user 
+any more for users of conventional model checkers. The reason is that the user has to write all clauses 
+for the implementation, which contains cross-literals, e.g., $s_2$ may associate with $a_1, p_0$, etc. If the user 
 is not familiar with the implementation of this adder, those cross-literals will bring
-confusions. However, if word-level techniques allow specification like Equation \ref{eqn:wordspec},
+confusion. However, if word-level techniques allow specification like Equation \ref{eqn:wordspec},
 the verification tool will be very user-friendly and straightforward even if the implementation details
@@
 For each element in $\F_{2^3}$, we write down the truth table as Table \ref{tab:truthtable}.
-Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
+Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, i.e., 
 $Z = \Func(A)$. Recall Lagrange's interpolation formula:
@@
 
-\begin{table}[bp]
+\begin{table}[tbp]
 \caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
@@
 
-Based on the aforementioned discussions, the importance of performing FSM traversal at word-level 
+Based on the aforementioned discussions, the importance of performing FSM traversal at word level 
 can bring a dimension of abstraction in sequential circuit reachability analysis.
@@
 confirms the existence of word-level abstraction for bit-level circuits, the concept
-can be extended to encode the state space of sequential circuit by word-level polynomials in $\Fkk$.
+can be extended to encode the state space of a sequential circuit by word-level polynomials in $\Fkk$.
 % we prove the feasibility of word-level techniques on general circuits' verification.
@@
 
-\section{FSM Reachability using Algebraic Geometry}
+\section{FSM Reachability Using Algebraic Geometry}
 \label{sec:reach}
@@
 A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
-circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
+circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is called a \textit{Mealy machine};
 the formal definition is as follows:
@@
 The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
-its primary outputs only depend on the present states, {\it i.e.} the output function is defined as
+its primary outputs only depend on the present states, i.e., the output function is defined as
 $$\Lambda:\ S \to O$$
@@
   variables and $x$ represents the input variables. The {\it
-    transition relation of the FSM} is then represented as:  
+    transition relation of the FSM} is then represented as  
 \begin{equation} 
@@
 image of the initial states, under the transition function $\Delta$ is
-finally computed as:
+finally computed as
 \begin{align}
@@
 For the circuit in Figure \ref{fig:fsm}(b), we have the transition
-functions of the machine as:
-
+functions of the machine as
 \begin{align*}
@@
 using Gr\"obner bases over Galois fields. {\it
-This requires to perform quantifier elimination; which can be
+This requires performing quantifier elimination, which can be
 accomplished using the GB computation over $\Fkk$ using elimination
-ideals} \cite{gao:qe-gf-gb}. Finally, the set union, intersection and
+ideals} \cite{gao:qe-gf-gb}. Finally, the set union, intersection, and
 complement operations are also to be implemented in algebraic geometry.
@@
 Algorithm  \ref{alg:BFS}, we need  \textbf{union}, 
-\textbf{intersection} and \textbf{complement} of varieties over
+\textbf{intersection}, and \textbf{complement} of varieties over
 $\mathbb{F}_{2^k}$, for which we again use algebraic geometry concepts.
@@
 Now we make an assumption that $x \in V(g)$. Then $g(x) = 0 =
-d(x)g_1(x)$ which means either $h(x) = 0$ or $g_1 (x) = 0$.
+d(x)g_1(x)$, which means either $h(x) = 0$ or $g_1 (x) = 0$.
 
@@
 
-Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$
-{\it i.e.} the variety of vanishing ideal contains
+Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$,
+i.e., the variety of vanishing ideal contains
  all possible valuations of variables, so it constitutes the {\bf
@@
 $$f_T: T = t_0+t_1\alpha+t_2\alpha^2+\cdots+t_{k-1}\alpha^{k-1}$$
-\item Impose ATO for sequential circuits which is 
+\item Impose ATO for sequential circuits, which is 
 $$LEX: \text{ all bit-level variables in any order }>S>T$$
@@
 
-Since the formerly reached state ``$reached = T$'', its complement is
+Since the formerly reached state ``$reached = T$," its complement is
 computed using Corollary \ref{cor:complement} 
@@
 \end{align*}
- {\it i.e.} states $\{00,01,10\}$. We update the present states
+i.e., states $\{00,01,10\}$. We update the present states
 for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha.$$
@@
 $\{00,10\}$. The complement of $reached$ is $$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
-= \langle T + 1+\alpha\rangle$$ ({\it i.e.} states $\{11\}$). We compute the newly reached state 
+= \langle T + 1+\alpha\rangle$$ (i.e., states $\{11\}$). We compute the newly reached state 
 $$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle 1\rangle$$
@@
 
-\subsection{Significance of using Algebraic Geometry and Gr\"obner Bases}
+\subsection{Significance of Using Algebraic Geometry and Gr\"obner Bases}
 The essence of our approach is based on algebraic geometry and Gr\"obner basis concepts.
-One the one hand, we use GB with elimination term ordering as an analog of image function.
+On the one hand, we use GB with elimination term ordering as an analog of image function.
 As mentioned in Section \ref{sec:abstraction}, we construct the ideal $J+J_0$ to 
@@
 Given the present states, the next states are implicitly represented in the variety of 
-the elimination ideal obtained by quantify away the remaining variables. 
+the elimination ideal obtained by quantifying away the remaining variables. 
 This projects the variety on the NS output $T$ by eliminating all other variables.
@@
 
-On the other hand, we use the algebraic ideals to implement set operations. States are finite set of points,
+On the other hand, we use the algebraic ideals to implement set operations. States are finite sets of points,
 which can be mapped to a variety.
@@
 As a result, we create the analog of the FSM traversal algorithm in algebraic geometry which is 
-compatible with word-level variables ({\it e.g.} $S,T$). We show it is effective to perform the 
+compatible with word-level variables (e.g., $S,T$). We show it is effective to perform the 
 reachability analysis.
 
-\section{Improving our Approach}
+\section{Improving Our Approach}
 \label{sec:improve}
 
-Using elimination term ordering on computing GB for large set of polynomial is time-consuming, and usually 
+Using elimination term ordering on computing GB for a large set of polynomials is time-consuming, and usually 
 intractable. The reason is the exponential computational complexity \cite{gao:gf-gb-ms}:
@@
 $Spoly$ pairs, the other is dividing the $Spoly$ by the set of polynomials.
-Its actual cost is very sensitive to the term ordering: if there exists 
+Its actual cost is very sensitive to the term ordering: If there exists 
 a term order making most $Spoly$ computation unnecessary, then the cost to 
@@
 there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
-their leading monomials are not relatively prime, {\it i.e.} 
+their leading monomials are not relatively prime, i.e., 
 $$gcd(LM(f_w), LM(f_g)) \neq 1$$
@@
 We start from the outputs $t_0,t_1$, then intermediate bit-level signals $a,b,c,d$ because they are 
-the fanins of the corresponding gates which fanout $t_0,t_1$. Then we ends at the pseudo inputs $s_0,s_1$
+the fanins of the corresponding gates which fanout $t_0,t_1$. Then we ends at the pseudo inputs $s_0,s_1$,
 and primary input $x$.
@@
 containing only word-level PS variables. One possible method is 
-to rewrite bit-level variables in term of word-level variables, {\it i.e.}
+to rewrite bit-level variables in terms of word-level variables, i.e.,
 \begin{equation}
@@
 
-Solution to the system of equations can be obtained by Gaussian elimination, which could 
+The solution to the system of equations can be obtained by Gaussian elimination, which could 
 compute corresponding $\mathcal{G}(S)$ efficiently with time complexity $O(k^3)$.
@@
 
-In this approach we get word-level variable representation for each bit-level PS variables. 
+In this approach we get word-level variable representation for each bit-level PS variable. 
 By substitution, a new polynomial in word-level PS/NS variables could be obtained.
@@
 
-\begin{figure}[hbt]
+\begin{figure}[tbp]
 \centering{
@@
 Again, assume a partition $n_i~|~k$ and corresponding word-level variable is $P$. Then we can use polynomial
-$P^{2^{n_i}}-P$ to represent all signals at free-end PIs, according to following theorem about {\bf composite
+$P^{2^{n_i}}-P$ to represent all signals at free-end PIs, according to the following theorem about {\bf composite
 fields} \cite{ecc:software}:
@@
 $$P^4+P,~~\text{where}~~P=p_0+p_1\cdot\beta$$
-Using Theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
+Using Theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as an element from $\mathbb F_{2^4}$:
 $$P = p_0 + p_1\cdot\alpha^5$$
 \end{Example}
-Using this method we can efficiently partition large size PIs to small number of word-level PI variables.
+Using this method we can efficiently partition large-size PIs to a small number of word-level PI variables.
 One limitation of this approach is PIs cannot be partitioned when $k$ is prime.
@@
 RATO. This part is implemented using a scripting language such as \emph{Perl}. If 
-the given benchmark is a structural/behavior hybrid description, we perform pre-processing on it to get a synthesized netlist.
+the given benchmark is a structural/behavior hybrid description, we perform preprocessing on it to get a synthesized netlist.
 Secondly, the polynomial reduction is executed using our customized reduction engine, which is written in C++.
@@
 One of these languages is the Berkeley logic interchange format (BLIF) \cite{BLIF},
-it allows state behavior representation and logic component representation.
+which allows state behavior representation and logic component representation.
 
 \begin{Example}
-In this example, we use benchmark FSM ``lion9" from MCNC benchmark library.
-This benchmark circuit is given as state table based BLIF representation of a FSM.
+In this example, we use benchmark FSM ``lion9" from the MCNC benchmark library.
+This benchmark circuit is given as a state-table-based BLIF representation of a FSM.
 
@@
 After the script is executed, the traversal finishes after 4 
-transition iterations, which denotes BFS depth equals to 4. According to the output in
+transition iterations, which denotes BFS depth equal to 4. According to the output in
 Figure \ref{fig:Recha_result}, the final reachable states 
@@
 
-There are 2 bottlenecks which restricts the performance of our tool:
-one bottleneck is that the polynomial reduction engine is slow when the number of
+There are 2 bottlenecks, which restricts the performance of our tool:
+One bottleneck is that the polynomial reduction engine is slow when the number of
 gates (especially OR gates) is large; the other one is the high computational complexity
-of Gr\"obner basis engine in general. Therefore, we pick 10 FSM
+of a Gr\"obner basis engine in general. Therefore, we pick 10 FSM
 benchmarks of reasonable size for testing our tool. Among them ``b01,
@@
 where ITC/MCNC FSMs are given in unsynthesized {\it BLIF} format so we first turn them
-into gate-level netlists using AIG based synthesizer ABC. % \cite{brayton2010abc}.
+into gate-level netlists using AIG-based synthesizer ABC. % \cite{brayton2010abc}.
 Since the number of primary inputs $(m)$ is relatively small, in our experiments we partition
@@
 In Table \ref{tab:recha_result}, \# States denotes the final reachable states
-starting from given reset state, which given by our tool is the same with
+starting from given reset state, which, given by our tool, is the same with
 the return value of {\it compute\_reach} in VIS. Meanwhile, from observation
@@
 Although the efficiency of our initial implementation fails to compete
-with the BDD based FSM analyzer VIS, the experiment
+with the BDD-based FSM analyzer VIS, the experiment
 demonstrates the power of abstraction of algebraic geometry techniques
@@
 \caption{Results of running benchmarks using our tool. 
-\small{Parts I to III denote the time taken by polynomial divisions,
-  bit-level to word-level abstraction and iterative reachability
-  convergence checking part of our approach, respectively.}}
+Parts I to III denote the time taken by polynomial divisions,
+  bit-level to word-level abstraction, and iterative reachability
+  convergence checking part of our approach, respectively.}
 {\small 

### 2017-03-03 dissertation/submission/unsat.tex
after submission: revision 2

@@
 when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
-a set of constrains ({\it e.g.} clauses, formulas or polynomials) which is unsatisfiable, sometimes 
-it is worthy to explore the reasons for UNSAT. From the execution of the GB algorithm, 
+a set of constrains (e.g., clauses, formulas, or polynomials) which is unsatisfiable, sometimes 
+it is worthwhile to explore the reasons for UNSAT. From the execution of the GB algorithm, 
 an auxiliary structure can be obtained to help explore the reasons causing UNSAT. This chapter introduces the 
-details about the motivation, mechanism and implementation.
+details about the motivation, mechanism, and implementation.
 
@@
 The Boolean satisfiability (SAT) problem is a fundamental problem in computer science.
-In the following part we define the terminology related to SAT problem \cite{anaICCAD}.
+In the following part we define the terminology related to a SAT problem \cite{anaICCAD}.
 
@@
 A {\bf literal} $l$ is defined as a variable $v$, or its negation $\overline{v}$. A disjunction (OR relation)
-of literals forms a {\bf clause}, {\it i.e}. $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
+of literals forms a {\bf clause}, i.e., $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
 A Boolean formula can always be written as {\bf conjunctive normal form} (CNF),
@@
 
-Figure \ref{fig:SAT} shows a simple example of SAT problem on circuit verification:
+Figure \ref{fig:SAT} shows a simple example of a SAT problem on circuit verification:
 we need to verify whether subcircuit $A$ and subcircuit $B$ have the same function, so
@@
 In most cases, SAT problems are modeled as conjunctive normal form (CNF) formulas, and solved by procedures based on 
-Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of DPLL algorithm
-is recursive branching and backtrack searching. To improve the efficiency of DPLL-based
-SAT solver, efforts are made on minimizing number of branches, accelerating unit propagation
-and modifying backtrack algorithm. Recently state-of-art SAT solvers developed conflict-driven
-clause learning (CDCL) technique to prune the search space, which is effective to reduce
+the Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of the DPLL algorithm
+is recursive branching and backtrack searching. To improve the efficiency of a DPLL-based
+SAT solver, efforts are made to minimize the number of branches, accelerating unit propagation
+and modifying backtrack algorithm. Recently, state-of-art SAT solvers developed a conflict-driven
+clause learning (CDCL) technique to prune the search space, which is effective at reducing
 search time.
 
-When a SAT solver fails to give a SAT assignment, it will provide a {\it UNSAT proof} or {\it refutation proof}
+When a SAT solver fails to give a SAT assignment, it will provide an {\it UNSAT proof} or {\it refutation proof}
 to prove the problem is UNSAT. By analyzing clauses involved in the UNSAT proof, we can 
-generate a subformula which remains UNSAT. A n\"aive method is to collect all leaf 
+generate a subformula which remains UNSAT. A naive method is to collect all leaf 
 clauses in an UNSAT proof as an UNSAT core. In practice, a minimal UNSAT core is more valuable.
 There are mainly two kinds of methods to find minimal UNSAT cores.
-One is insertion-based method, which is achieved by adding clauses to the smallest subset 
-until the subset turns to be UNSAT. The other method is deletion-based method,
+One is the insertion-based method, which is achieved by adding clauses to the smallest subset 
+until the subset turns to be UNSAT. The other method is the deletion-based method,
 which is realized by deleting clauses from a larger subset until the subset turns to be SAT.
 Recently heuristics such as clause-set refinement \cite{modelrotation} and model rotation \cite{belov2011accelerating}
-are applied as improvements on the deletion/insertion based methods.
+have been applied as improvements on the deletion/insertion-based methods.
 These methods are also expanded to satisfiability modulo theories (SMT) \cite{cimatti2007simple}.
@@
 Using PC and its variations, researchers developed many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
-Besides, researchers borrow concepts from PC, combine them with traditional DPLL and clause learning techniques
+Besides, researchers borrow concepts from PC and combine them with traditional DPLL and clause learning techniques
 to build hybrid SAT solvers \cite{condratTACAS07,Zengler2010}.
@@
 The problem of finding small UNSAT cores has attracted interest for decades because of 
-its applications to various of verification and synthesis problems.
-In solving MaxSAT problems, a small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
+its applications to various verification and synthesis problems.
+In solving MaxSAT problems, a small UNSAT proof provides a lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
 It is applied to solving logic synthesis problems, such as Boolean function decomposition \cite{lee2008bi}. 
-Small explanation generation in general constraint programming problems also rely on small UNSAT cores \cite{cambazard2008reformulating}.
-
-UNSAT cores can find wide range of applications in circuit verification as well. Many abstraction refinement techniques require 
+Small explanation generation in general constraint programming problems also relies on small UNSAT cores \cite{cambazard2008reformulating}.
+
+UNSAT cores can find a wide range of applications in circuit verification as well. Many abstraction refinement techniques require 
 information mining from UNSAT proofs of intermediate abstractions. 
@@
 
-Bounded model checking (BMC) is a model checking technique which set upper bound to the length of all paths. 
-It can be solved by solving a SAT problem. Given a model $M$, property $p$ and bound $k$, 
-The $k$-BMC unrolls $M$ by $k$ clock-cycles and generates Boolean formula $F$ including violation check of $p$.
+Bounded model checking (BMC) is a model checking technique which set an upper bound to the length of all paths. 
+It can be solved by solving a SAT problem. Given a model $M$, property $p$, and bound $k$, 
+the $k$-BMC unrolls $M$ by $k$ clock-cycles and generates Boolean formula $F$ including violation check of $p$.
 Then $F$ is fed to a SAT solver, if the SAT solver returns SAT, then $p$ is violated in 
@@
 If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
-these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
+these variables will not affect the unsatisfiability of the original formula. Therefore, we can ignore them in the abstracted model.
 In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:absrefine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).
@@
 The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
-error trace, we can assert that property $p$ also holds on the original machine $M$. The reason of this assertion is that
-the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf over-approximation}
+error trace, we can assert that property $p$ also holds on the original machine $M$. The reason to this assertion is that
+the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf overapproximation}
 of $M$, such that 
@@
 
-If we find a violation on abstracted machine, then this
+If we find a violation on an abstracted machine, then this
 abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.
@@
 the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
-when the size of datapath (number of latches in Figure \ref{fig:absrefine}) is very large. \textbf{Here we propose an altogether new 
+when the size of the datapath (number of latches in Figure \ref{fig:absrefine}) is very large. Here we propose an altogether new 
 method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
-method according to the following observation:}
+method according to the following observation:
 
@@
 Given ideal $J\subseteq \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure $\overline{\F}$
-of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
+of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1."
 $${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow \text{reduced GB}(J) = \{1\}$$
@@
 will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
-Thus the number of polynomials in non-reduced GB is much smaller than usual.
+Thus the number of polynomials in nonreduced GB is much smaller than usual.
 Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
@@
 allow the use of word-level variables, which provides a strong potential for all applications which can be
-modeled as polynomials in finite fields extensions.
+modeled as polynomials in finite-field extensions.
 
@@
 %% disregarding minimality. 
-It is not hard to motivate that an UNSAT core should be identifiable
+It is not hard to figure out that an UNSAT core should be identifiable
 using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
@@
 
-A n\"aive way (and inefficient way) to identify {\it a minimal core}
+A naive way (and inefficient way) to identify {\it a minimal core}
 using the GB computation is as follows:
-select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
+Select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
 ({\it i.e.} if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
@@
 
-\subsection{An Example Motivating our Approach}
+\subsection{An Example Motivating Our Approach}
 
 Buchberger's algorithm picks pairs of polynomials from a given set, computes their $Spoly$, then reduces this $Spoly$
-with the given set of polynomials. If the remainder is non-zero,
+with the given set of polynomials. If the remainder is nonzero,
 it is added to the set of polynomials.
@@
 \item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
-$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
+$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1," we
 terminate the Buchberger's algorithm;
@@
 From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
-as these polynomials are the ``causes" of unsatisfiability of original set of polynomials.
+as these polynomials are the ``causes" of the unsatisfiability of the original set of polynomials.
 \end{enumerate}
@@
 
-We investigate if it is possible to identify a core by analyzing the
+We investigate whether it is possible to identify a core by analyzing the
 $Spoly(f_i,f_j)\xrightarrow{F}_+ g_{ij}$ reductions in Buchberger's
@@
 computation corresponding to Example \ref{ex:1}, the first 3 critical
-{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
+{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$, and
 $(f_2,f_5)$. It turns out that the $Spoly$-reductions corresponding to
@@
 node $i$ to node $j$ signifies that the polynomial at node $j$ results
-from polynomial at node $i$. For example, consider the computation
+from the polynomial at node $i$. For example, consider the computation
 $Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
@@
 To identify an $F_c \subset F$, we start from the refutation node
-``1'', and traverse the graph in reverse, all the way up to the
+``1'' and traverse the graph in reverse, all the way up to the
 leaves. Then, all the leaves in the transitive fanin of ``1''
@@
 We record the S-polynomial reduction
-$Spoly(f_i,f_j)\xrightarrow{F}_+{g_{ij}}$ that give a non-zero
+$Spoly(f_i,f_j)\xrightarrow{F}_+{g_{ij}}$, which gives a nonzero
 remainder when divided by the system of polynomials $F$ at that
@@
 $Spoly(f_i,f_j)$ and the current basis $F$; thus, it can be
-represented as:
+represented as
 \begin{equation}
@@
 $\{f_1,\ldots,f_m\}$ is the ``current'' system of polynomials
-$F$. For each non-zero $g_{ij}$, we will record the following data:
-
+$F$. For each nonzero $g_{ij}$, we will record the following data: 
 \begin{equation}
@@
 only twice in the division process. In both occasions, the quotient of
-the division is 1. From Figure \ref{fig:refute}, it follows that:
+the division is 1. From Figure \ref{fig:refute}, it follows that
 \begin{equation}
@@
 \end{equation}
-If the remainder $g_{ij}$ is non-zero, we denote it by
+If the remainder $g_{ij}$ is nonzero, we denote it by
 $f_{m+1}$ and add it to the current set of generators $G$. We
@@
 coefficients are associated with new elements (remainders) in the
-Gr\"obner basis, that are not a part of the UNSAT core. 
+Gr\"obner basis that are not a part of the UNSAT core. 
 %which cannot benefit the UNSAT core extraction. 
@@
 larger indexes get paired {\it later} in the order:
-
 $$(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$$
@@
 distances are used as divisors in later stages of polynomial
-reductions; which implies that they may generally have lower-degree
+reductions, which implies that they may generally have lower-degree
 leading terms. This is because we impose a degree-lexicographic term
@@
 The frequency of occurrence of a polynomial $f_i$ in a refutation tree
-corresponds to the number it appears in the refutation tree. 
+corresponds to the number of times it appears in the refutation tree. 
 \end{Definition}
@@
 
-We apply both heuristics: after the first iteration of the GB-core
+We apply both heuristics: After the first iteration of the GB-core
 algorithm, we analyze the refutation tree $T$ and sort the polynomials
@@
 so their relative ordering is unchanged. We reorder $f_2$ to be the
-polynomial with the  smallest index. We re-index the polynomial set 
+polynomial with the  smallest index. We reindex the polynomial set 
 $f_1'=f_2, f_2' = f_1, f_3' = f_3, f_4' = f_4$
@@
 f_4'\} = \{f_2,f_3,f_4\}$. Further iterations do not refine the core
--- {\it i.e.} a fix point is reached. 
+-- i.e., a fix point is reached. 
 \end{Example}
 
-\section{Refining the UNSAT Core using Syzygies}
+\section{Refining the UNSAT Core Using Syzygies}
 \label{sec:syz}
@@
 The UNSAT core obtained through our GB-core algorithm is by nature a
-refutation polynomial that equals to 1:  
+refutation polynomial equal to 1:  
 $$1 = \sum_{i=1}^s c_i\cdot f_i$$
@@
 $f_k \in F$ can be represented using a combination of the rest of the
-polynomials of the core, {\it e.g.}
+polynomials of the core, e.g.,
 $$f_k = \sum_{j\neq k} c_j'f_j.$$
@@
 redundant. One of the limitations of the GB-core algorithm and the
-re-labeling/refinement strategy is that they cannot easily identify
+relabeling/refinement strategy is that they cannot easily identify
 such polynomials $f_k$ in the generating set $F$ that can be composed
-of the other polynomials in the basis, {\it i.e.}
+of the other polynomials in the basis, i.e.,
 $f_k \in \langle F-\{f_k\} \rangle$. We present an approach targeted
@@
 implies that some polynomial combination of $\{f_1,\dots,f_s\}$
-vanishes: {\it i.e.} $c_1f_1 + c_2f_2 + \dots+c_sf_s=0$, for some
+vanishes: i.e., $c_1f_1 + c_2f_2 + \dots+c_sf_s=0$, for some
 $c_1,\dots,c_s$. These elements ($c_1,\dots,c_s$) form a syzygy on $f_1,\dots,f_s$.
@@
  Here $\{f_1,f_2,\dots,f_s\}$ is the given core.
- Take one column of the syzygy matrix ({\it e.g.} the set of polynomials in $j$-th column
+ Take one column of the syzygy matrix (e.g., the set of polynomials in $j$-th column
  $c_j^1, c_j^2, \dots, c_j^m$)  and compute its reduced Gr\"obner
@@
 discarded from the core. By repeating this procedure, some redundant
-polynomials can be identified and size of UNSAT core can be reduced
+polynomials can be identified and the size of UNSAT core can be reduced
 further.
@@
  S-polynomials that give 0 remainder. The coefficients can be
- represented as entries in matrix shown below. For example, the first
+ represented as entries in the matrix shown below. For example, the first
  row in the matrix corresponds to the syzygies generated by
@@
 reduction. In order to remove this extra column, we need to turn the
-non-zero entries in this column to 0 through standard matrix
+nonzero entries in this column to 0 through standard matrix
 manipulations.
@@
  By adding this row vector (Equation \ref{sr}) to the rows in
- Equation \ref{eqn:sm} corresponding to the non-zero entries in the
+ Equation \ref{eqn:sm} corresponding to the nonzero entries in the
  column for $f_{10}$, we obtain the syzygy matrix only for the
@@
 zero, so the number of rows of the syzygy matrix can be much larger than
-the size of original generating set. Full GB computation on each
+the size of the original generating set. Full GB computation on each
 column of the syzygy matrix can become prohibitive to apply
@@
 reduce the state space. Since those state variables contribute nothing to the violation of 
-property $p$, the abstracted model is a reasonable over-approximation for checking 
+property $p$, the abstracted model is a reasonable overapproximation for checking 
 $p$ without loss of accuracy. In the following example, we apply the UNSAT core extraction 
@@
 any constraints which can be written as polynomials in GF extensions,
-there is no such benchmark libraries clearly identifying a minimal core within 
-to test our tool.
+there is no benchmark library clearly identifying a minimal core within 
+which to test our tool.
 
@@
 library by modifying SAT benchmarks and translating from circuit benchmarks: 
-1) "aim-100" is a modified version  of the random 3-SAT
-benchmark "aim-50/100", modified by adding some redundant clauses; 2)
-The "subset" series are generated for random subset sum problems; 3)
-"cocktail" is similarly revised from a combination of factorization
-and a random 3-SAT benchmark; 4) and "phole4/5" are generated by
+1) ``aim-100" is a modified version  of the random 3-SAT
+benchmark ``aim-50/100", modified by adding some redundant clauses; 2)
+The ``subset" series are generated for random subset sum problems; 3)
+``cocktail" is similarly revised from a combination of factorization
+and a random 3-SAT benchmark; 4) and ``phole4/5" are generated by
 adding redundant clauses to pigeon hole benchmarks; 5) Moreover,
@@
 circuits \cite{agnew1991implementation,RHmulti}, compared against a golden model
-{\it spec}. Similarly, the "MasVMon" benchmarks are the equivalence
+{\it spec}. Similarly, the ``MasVMon" benchmarks are the equivalence
 checking  circuits corresponding to Mastrovito multipliers compared
@@
 \cite{condrat-tacas07}), and the GB-Core algorithm and the refinement
-approach is applied.   
+approach are applied.   
 %These benchmarks are in moderate size for our GB engine but not too trivial.
@@
 \caption{Results of running benchmarks using our tool. 
-\small{Asterisk($^*$) denotes that the benchmark was not translated from CNF.} 
-Our tool is composed by 3 parts: part I runs a single GB-core algorithm,
+Asterisk($^*$) denotes that the benchmark was not translated from CNF.
+Our tool is composed of 3 parts: part I runs a single GB-core algorithm,
 part II applies the iterative refinement heuristic to run the GB-core algorithm
@@
 number of calls to the GB-core engine to arrive at the reduced UNSAT
-core. The second last column shows the improvement in the minimal core
+core. The second-last column shows the improvement in the minimal core
 size by applying the syzygy heuristic on those cases which cannot be
@@
 computational algebraic geometry and solved using the Gr\"obner basis
-algorithm. We show that by recording the data produced by the
+algorithm. We show that by recording the data produced by
 Buchberger's algorithm -- the $Spoly(f_i,f_j)$ pairs, as well as the
@@
 by the SAT community. We hope that this technique will kindle some
-interest in this topic which is worthy of attention from the SAT
+interest in this topic, which is worthy of attention from the SAT
 community -- particularly when there seems to be a renewal of interest

### 2017-03-16 dissertation/submission/intro.tex
dissertation: after submission_final revision

@@
 that the sequential circuit implementations are equivalent to the
-original specification models? This problem is called {\it Sequential Equivalence
+original specification models. This problem is called {\it Sequential Equivalence
   Checking }(SEC) problem. When the specification is given as an arithmetic function

### 2017-03-16 dissertation/submission/normal.tex
dissertation: after submission_final revision

@@
 
-%\vspace{-0.1in}
-% \IncMargin{1em}
+For 2-input sequential NB multipliers, abstraction is utilized to implement Algorithm \ref{alg:modified}.
+Notations in Figure \ref{fig:sequential} as well as in the next example comply with 
+the notations in this algorithm.
+
+Given a sequential GF circuit in Figure \ref{fig:sequential} over $\Fkk=\F_2[x]$, irreducible polynomial $p(\alpha) = 0$
+defines the primitive element $\alpha$, $\beta = \alpha^t$ is the normal element.
+The circuit includes word-level variables $A, B, R$
+denoting the {\it present states (PS)} and $A', B', R'$ denoting the {\it next
+  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
+for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
+\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
+store the result, and $A, B\ (A', B')$ store input operands. For example,
+for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
+0$) are the initial values (operands) loaded into the registers,  and
+$R = \Func(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
+result after $k$-cycles. Our approach aims to find this polynomial
+representation for $R$.
+
+\begin{figure}[tbp]
+\centering{
+\includegraphics[width=4.5in]{newfig/gf_seq_model.eps}
+\caption{A typical normal basis GF sequential circuit model. $A =
+  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
+  $A', B', R'$ denote next-state inputs.}
+\label{fig:sequential}}
+\end{figure}
+
 \begin{algorithm}[hbt]
@@
 \end{algorithm}
-% \DecMargin{1em}
-
-For 2-input sequential NB multipliers, abstraction is utilized to implement Algorithm \ref{alg:modified}.
-Notations in Figure \ref{fig:sequential} as well as in the next example comply with 
-the notations in this algorithm.
-
-Given a sequential GF circuit in Figure \ref{fig:sequential} over $\Fkk=\F_2[x]$, irreducible polynomial $p(\alpha) = 0$
-defines the primitive element $\alpha$, $\beta = \alpha^t$ is the normal element.
-The circuit includes word-level variables $A, B, R$
-denoting the {\it present states (PS)} and $A', B', R'$ denoting the {\it next
-  states (NS)} of the machine; where $A = \sum_{i=0}^{k-1} a_i \beta^{2^i}$
-for the PS variables and $A' = \sum_{i=0}^{k-1} a_i'
-\beta^{2^i}$ for NS variables, and so on.  Variables $R\ (R')$ correspond to those that 
-store the result, and $A, B\ (A', B')$ store input operands. For example,
-for a GF multiplier, $A_{init}, B_{init}$ (and $R_{init} =
-0$) are the initial values (operands) loaded into the registers,  and
-$R = \Func(A_{init}, B_{init}) = A_{init} \times B_{init}$ is the final
-result after $k$-cycles. Our approach aims to find this polynomial
-representation for $R$.
-
-\begin{figure}[bp]
-\centering{
-\includegraphics[width=4.5in]{newfig/gf_seq_model.eps}
-\caption{A typical normal basis GF sequential circuit model. $A =
-  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
-  $A', B', R'$ denote next-state inputs.}
-\label{fig:sequential}}
-\end{figure}
 
@@
 this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
-variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
+variables, e.g., $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
 approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.

### 2017-03-16 dissertation/submission/prev.tex
dissertation: after submission_final revision

@@
 However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
-  contemporary techniques including quantifier-free bit-vector 
+  contemporary techniques, including quantifier-free bit-vector 
   (QF-BV) theory-based SMT-solvers \cite{Cryptol:fmcad09},

### 2017-03-16 dissertation/submission/unsat.tex
dissertation: after submission_final revision

@@
 The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
-error trace, we can assert that property $p$ also holds on the original machine $M$. The reason to this assertion is that
+error trace, we can assert that property $p$ also holds on the original machine $M$. The reason for this assertion is that
 the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf overapproximation}
@@
 Select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
-({\it i.e.} if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
+(i.e., if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
 from the core; otherwise retain $f_i$ in $F_c$. Select a different
@@
 
-\begin{figure}[tbp]
+\begin{figure}[bp]
 \centering{

### 2017-04-02 dissertation/submission/intro.tex
dissertation: SAGA ends. upload to proQuest

@@
 generalizes to functions with different input/output bit-vector sizes,
-{\it i.e.} functions of the type $f: \B^n \rightarrow \B^m$, modeled as
+i.e., functions of the type $f: \B^n \rightarrow \B^m$, modeled as
 polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
@@
 Therefore, the word-level FSM traversal algorithm can be applied to 
-verify the correctness of final reached state, {\it i.e.} the functional correctness of
+verify the correctness of final reached state, i.e., the functional correctness of
 a sequential arithmetic circuit.

### 2017-04-02 dissertation/submission/prelim_GF.tex
dissertation: SAGA ends. upload to proQuest

@@
 A {\bf field} $\mathbb{F}$ is a commutative ring with unity, where every
-nonzero element in $\mathbb{F}$ has a multiplicative inverse; {\it i.e.} $\forall
+nonzero element in $\mathbb{F}$ has a multiplicative inverse; i.e., $\forall
 a \in \mathbb{F} - \{0\}$, $\exists \hat{a} \in \mathbb{F}$ such that $ a \cdot
@@
 number of elements, $q$. The number of elements $q$ of the field is
-a power of a prime integer, {\it i.e.} $q = p^k$, where $p$ is a prime
+a power of a prime integer, i.e., $q = p^k$, where $p$ is a prime
 integer, and $k \geq 1$. Thus a Galois field can also be denoted as 
@@
 \begin{Theorem}
-$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, {\it i.e.} if $n$ divides $m$.
+$\F_{2^n}\subset\F_{2^m}$ iff $n \mid m$, {\it i.e.}, if $n$ divides $m$.
 \end{Theorem}
@@
 
-The final remainder is $1$, {\it i.e.} the product equals 1 in StdB.
+The final remainder is $1$, i.e., the product equals 1 in StdB.
 

### 2017-04-02 dissertation/submission/prelim_basis.tex
dissertation: after submission_final revision
dissertation: SAGA ends. upload to proQuest

@@
 A {\bf lexicographic ordering} (LEX) is a total-ordering $>$ such that 
-variables in the terms are lexicographically ordered, {\it i.e.} simply based on 
+variables in the terms are lexicographically ordered, i.e., simply based on 
 when the variables appear in the ordering. It is also a well-order, where the least element exists.
 Higher variable-degrees take 
-precedence over lower degrees for equivalent variables ({\it e.g.} $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
+precedence over lower degrees for equivalent variables (e.g., $a^3 > a^2$ due to $a \cdot a \cdot a > a \cdot a \cdot 1$).
 \begin{Definition}
@@
 \mathbb{F}[x_1,\dots, x_d]$ be a polynomial. We say that $f$ {\it vanishes} on $\mathbf{a}$ 
-if $f(\mathbf{a}) = 0$, {\it i.e.} $\mathbf{a}$ is in the variety of $f$.
+if $f(\mathbf{a}) = 0$, i.e., $\mathbf{a}$ is in the variety of $f$.
 
@@
 \begin{enumerate}
-\item Minimize every $g_i \in G$, i.e., $g_i=g_i/lc(g_i)$
+\item Minimize every $g_i \in G$, i.e., $g_i=g_i/lc(g_i)$.
 \item For $g_i, g_j \in G$ where $i\neq j$, remove $g_i$ from $G$ if $lt(g_i)\mid lt(g_j)$, 
@@
 in ${\overline {\F}}^d$. For this purpose, we only need to check if 
-the ideal is generated by the unit element, {\it i.e.} $1\in J$. 
+the ideal is generated by the unit element, i.e., $1\in J$. 
 This approach can be used to evaluate the feasibility of constraints in 
